### 2018-06-09 Misc/IRC18_backup.tex
added Misc folder

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference]{IEEEtran}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

\usepackage{graphicx}
\usepackage{amssymb,url,times}%,subfigure}% amsthm is the one!
\usepackage{caption,subcaption,hyperref}
\usepackage{color,comment}
\usepackage{curves,pgfgantt}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

\newtheorem{prop}{Proposition}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\abs}[1]{\ensuremath{\left| #1 \right|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\ip}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Figure \ref{fig:#1}}
\newcommand{\tr}[1]{\mbox{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mbox{tr}\ensuremath{\!\bracket{#1}}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathbf{d}}}
\newcommand{\pair}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\!\langle #1 \rangle\!\rangle}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\ad}{\ensuremath{\mathrm{ad}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\WriteBlue}[1]{{\color{blue}\protect #1}}
\graphicspath{{./Fig/}}

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Autonomous Quadrotor 3D Mapping and Exploration Using Exact Probabilities and Entropies}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations

%\author{Evan Kaufman, Taeyoung Lee, and Zhuming Ai%, and Ira S. Moskowitz%
%\thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 {\tt \{evankaufman,tylee\}@gwu.edu}}
%\thanks{Zhuming Ai, Information Management \& Decision Architectures, U.S. Naval Research Laboratory,  Washington, DC 20375}
%\thanks{This research has been supported by the U.S. Naval Research Laboratory Base Program ``Intelligent Microflyer'' and in part by NSF under the grants CMMI-1243000, CMMI-1335008, and CNS-1337722.}
%}

%\author{\IEEEauthorblockN{Evan Kaufman, Kuya Takami, Taeyoung Lee}
%\IEEEauthorblockA{Department of Mechanical and Aerospace Engineering\\The George Washington University\\
%Washington, DC 20052\\
%Email: \{evankaufman,kuya,tylee\}@gwu.edu}
%\and
%\IEEEauthorblockN{Zhuming Ai}
%\IEEEauthorblockA{Information Management \& Decision Architectures
%              \\ U.S. Naval Research Laboratory \\
%              Washington, DC 20375}
%}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
\author{\IEEEauthorblockN{Evan Kaufman\IEEEauthorrefmark{1},
Kuya Takami\IEEEauthorrefmark{1},
Zhuming Ai\IEEEauthorrefmark{2} and
Taeyoung Lee\IEEEauthorrefmark{1}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Mechanical and Aerospace Engineering\\
The George Washington University, Washington, DC 20052\\
Email: \{evankaufman, kuya, tylee\}@gwu.edu}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Information Management \& Decision Architectures
\\ U.S. Naval Research Laboratory, Washington, DC 20375}
}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
Robotic mapping and exploration has received significant attention recently because exploring uncertain environments can pose danger to humans, and robots can produce highly detailed maps. Commonly uncertain environments contain complex geometries and robots frequently face restrictive terrains, motivating a 3D representation of mapping from a vehicle capable of hover. This paper focusses on the popular occupancy grid mapping representation, with the goal of determining occupancy probabilities of evenly-spaced grid cells in 3D. With recent developments from the authors using the stochastic properties of depth sensors directly with a rigorous, probabilistic formulation, this paper extends this precise mapping technique to 3D for real-time implementation with realistic sensor capabilities. Furthermore, important aspects of the 3D map are projected onto 2D maps, where a predicted level of map uncertainty, known as Shannon's entropy, provides an exploration policy that governs robotic motion. Both mapping and exploration algorithms produce important information that impacts each other and are demonstrated simultaneously with numerical simulations and experiments in real-time.
\end{abstract}

\section{Introduction}

Robotic mapping is the process of generating maps, which represent the environment surrounding a robot. 
Effectively modeling this environment is crucial when building an understanding of an initially uncertain environment.
Robotic mapping serves an important role in tasks such as search-and-rescue, surveillance, and robotic cleaning.
Several grid-based map representations have been studied, e.g.,~\cite{WolSuk05,MeyBeiBur12,TanThoWolBus14}, because they provide abundant occupancy knowledge of the environment are computationally tractable.

Most grid-based mapping representations are designed for 2D, while others are extended to capture a full 3D environment. In particular, occupancy probabilities have been assigned to a uniform 3D grid in~\cite{Andert09,PirRutBisSch11}. Another highly-popular variation of occupancy grid mapping is the Octomap representation~\cite{WurHorBenStaBur10,HorWurBenStaBur13}, which provides powerful open-source software tools for representing free and occupied spaces nearby the robot, with the capability of extending the mapping limits and changing cell sizes. However, Octomaps produce cell occupancy probabilities that do not directly depend on the depth sensor stochastic properties, known as the forward sensor model. Ideally, the probability density provided by the forward sensor model should be used explicitly when determining occupancy probabilities of grid cells conditioned on the robot pose and measurements, known as the inverse sensor model. Instead, these approaches mentioned above determine occupancy probabilities using an approximate function to replace the true inverse sensor model. Additionally, this approach replaces probabilities with log-odds ratios, then combines past and current information with a potentially harmful Markov assumption.

Alternatively, another popular approach is to simulate maps, robot poses, and measurements and use a learning algorithm, to approximate the inverse sensor model~\cite{Thr01,ThrBurFox05}, which is implemented in 3D with~\cite{SouMaiGon12}. These approaches are undesirable in practice due to complexities associated with implementing a learning algorithm. For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples, or how many samples are required to obtain a reasonable approximation. Furthermore, it is challenging to apply any learning algorithm over the large dimensional space composed of maps, poses, and measurements.

The approximate nature of the inverse sensor model motivated the authors of this paper to design computationally-efficient algorithm to construct the exact inverse sensor model that follows a Bayesian framework~\cite{KauLeeAiMos16,KauTakAiLee17}. The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancies of the other cells blocked by it are irrelevant to the forward sensor model, and this model is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model. This method combines a priori probabilities of occupancy and multiple range measurements according to a Bayesian framework to obtain more accurate maps. As such, it contrasts from the existing framework based on log-odds ratios that impose an additional Markov assumption or particle-filtering methods such as~\cite{TanThoWolBus14} that require particle weighting and resampling to obtain a variation on a Bayesian solution. 
This paper proposes extending this algorithm to 3D with realistic sensor limitations. We demonstrate the efficacy of the proposed algorithm using two different depth sensors simultaneously to generate a 3D occupancy grid map with accurate probabilities in real-time with numerical simulations and experiments.

The second topic of this paper deals with autonomous exploration. A common approach to solve the autonomous exploration problem is known as frontier-based exploration, proposed for 2D environments in~\cite{Yam97,Yam98}. This is the process of executing actions that move the robot toward the closest boundary between visited and unvisited space, known as a frontier.
Then the robot takes measurements at this location such that the mapped territory expands, and thus the new frontiers are pushed back. This process is repeated until the map is well-known.
This frontier-based approach is extended to 3D with~\cite{ZhuDinLinWu15,SenWan16,KleDor13} using the Octomap representation, and similarly in~\cite{SawKriSri09} using a visibility metric.
Frontier-based exploration assumes that repeatedly moving toward the closest frontier and taking measurements are the best actions to gain new information about the map, though these systematic actions are not based on any consideration of the future uncertainty of the probabilistic map or optimality.

Alternatively, there have been exploration techniques that are based on a quantitative measure of uncertainty known as Shannon's entropy~\cite{StaGriBur05}, an uncertainty metric based on grid cell occupancy probabilities. However, the expected entropy of an action is not directly calculated. Instead, this approach assumes that expected entropy is equivalent to entropy based on expected measurement scans; a Rao-Blackwellized particle filter estimates the robot pose and map, where expected measurements are drawn from each particle when finding the expected information gains, then combined with the weighting on each particle. However, assuming an expected measurement is sufficient to predict expected information is subject to inaccuracies due to the nonlinear nature of object detection probability and entropy. For these reasons, the authors have proposed a direct calculation of expected information gain from occupancy probabilities and stochastic sensor models~\cite{KauAiLee16,KauTakAiLee17}. In this paper, this concept is extended to 3D, where grid cells above and below a predetermined exploration height are projected onto 2D maps for entropy and collision calculations. In this sense, the proposed approach is entropy-based in 3D similar to~\cite{JohStaPfaBur07} where uncertainties are projected onto a 2D map, except that the entropy measures are based on map probabilities and sensor properties directly.

The paper is organized as follows. The problem is formulated in Section II. Special considerations for mapping and exploration in 3D are established in Section III. A numerical example is demonstrated in Section IV. Experimental results are shown in Section V, followed by conclusions.

\section{Problem Formulation}

The key differences between the approaches presented in this paper and others is how map cell occupancy probability and measurement ray entropy are calculated directly from a priori cell probabilities and stochastic properties of the sensors involved. These are defined next.

\subsection{Probabilistic Occupancy Grid Mapping}
\label{subsec:POGM}
Let a map $m$ be an occupancy grid map decomposed into $n$ evenly-spaced cubic grid cells with fixed edge length $\alpha$. The $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,\ldots,n}$, that is defined as $\mathbf{m}_i=1$ when occupied, and $\mathbf{m}_i=0$ when free. The location and size of each grid cell is assumed known, where a smaller cell size better represents a space, but increases computation and memory. Therefore, a map $m$ is defined by $\{\mathbf{m}_1,\ldots, \mathbf{m}_{n}\}$ ($2^{n}$ possible maps).

Another random variable is defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience, which is the event that the $i$-th cell is free, i.e., $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$. The random variables composing $m$ are assumed mutually independent, 
\begin{align}
P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n})=\prod_{i=1}^{n}P(\mathbf{m}_i).
\end{align}

% {R ? R3×3 | RT R = I, det R = 1}
Consider a range sensor that provides scans of the surrounding environment in order to identify the closest occupied space. Let the measurement scan at the $t$-th time step be $Z_t$, where $z\in Z_t$ is a single measurement ray. Let the $t$-th pose be denoted by $X_t=\braces{x_t,R_t}$ such that $x_t\in\Re^3$ and $R_t\in\SO=\braces{R\in\Re^{3\times3}|R^\T R=I,\det{R}=1}$, where $X_t$ is assumed to be known. Given the sensor probabilistic properties $p(z|X_t,m)$, commonly referred to as the \emph{forward sensor model}, the goal is finding the \emph{inverse sensor model}, $P(m|z,X_{1:t},Z_{1:t-1})$. To obtain this probability for all cells,~\cite{KauLeeAiMos16,KauTakAiLee17} introduced a reduced map $r\subset m$ composed of $n_r$ cells along $z$, where cells are indexed by increasing distance from $x_t$, and $\mathbf{r}_{k+}$ refers to the event that the $k$-th cell is the closest occupied space. Then, the inverse sensor model for the $k$-th cell is
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})&=\eta\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1}),
\end{align}
where the unnormalized probability of the inverse sensor model is defined as
\begin{align}
\label{eqn:Unnormalized}
\tilde P(\mathbf{r}_{k}|z,X_{1:t},&Z_{1:t-1})\nonumber\\
&=\mathbf{P}_k^-
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]\nonumber\\
&\quad + \bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{k+},X_t)\mathbf{P}_k^-,
\end{align}
where a priori probability is $\mathbf{P}_k^-=P(\mathbf{r}_{k}|X_{1:t-1},Z_{1:t-1})$, its complement is $\bar{\mathbf{P}}_k^-=1-\mathbf{P}_k^-$, $P(\bar{\mathbf{r}}_{0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience, and $p(z|\mathbf{r}_{(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading. The normalizer $\eta$ is given by
\begin{align}
\label{eqn:allEta}
\eta
&=
\bigg[\sum_{i=1}^{n_{r}+1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\} p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]^{-1},
\end{align}
and it is independent of the cell index $k$. Since the terms of \refeqn{Unnormalized} are easily obtained and several are used repeatedly to obtain \refeqn{allEta}, the computational cost of \refeqn{RayISMAnswer} is trivial and can be applied in real-time. This process is repeated for all measurements composing scan $Z_t$.

\subsection{Predictive Entropy-Based Autonomous Exploration}

Since occupancy grid mapping provides a probabilistic representation of surrounding space, this mapping scheme holds probabilistic information about the uncertainty of the map. Shannon's entropy is commonly used as a measure of uncertainty, such as in~\cite{StaGriBur05}. Given the probabilities of each grid cell of map $m$, Shannon's entropy of an individual cell and the entire map are defined as
\begin{align}
\label{eqn:ShannonsEntropyCell}
H(P(\mathbf{m}_i))&=-P(\mathbf{m}_i)\log{P(\mathbf{m}_i})-P(\bar{\mathbf{m}}_i)\log{P(\bar{\mathbf{m}}_i}),
\\
\label{eqn:ShannonsEntropyMap}
H(P(m))&=\sum_{i=1}^{n_m}H(P(\mathbf{m}_i)),
\end{align}
respectively.
The entropy of the $i$-th grid cell is maximized when $P(\mathbf{m}_i)=0.5$ (greatest uncertainty) and minimized as $P(\mathbf{m}_i)$ approaches $0$ or $1$ (smallest uncertainty). %; thus Shannon's entropy is a measure of uncertainty.

Let an unvisited candidate future pose and its measurement scan be $X_c$ and $Z_c$, respectively. Since the robot has not visited $X_c$, any change to the probabilistic map must be predicted. Much like the probabilistic mapping, this is achieved ray-by-ray~\cite{KauAiLee16,KauTakAiLee17}. Since all grid cell probabilities are conditioned on the history of poses $X_{1:t}$ and measurement scans $Z_{1:t}$, these are removed from the remaining equations for simplicity. For candidate ray $z_c$, the expected entropy is
\begin{align}
\label{eqn:DiscExpEntropyRay}
&\text{E}[H(P(m|x_c,z_{c}))]=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(m|x_c,z_{c,k}))P(z_{c,k}|x_c)\bigg\},
\end{align}
where subscript $k$ refers to the event that the candidate \emph{measurement} captures the $k$-th cell along the ray, i.e., $z_{c,k}$ is the distance from $x_c$ to the $k$-th grid cell. The first term of the summation of \refeqn{DiscExpEntropyRay}, namely $H(P(m|x_c,z_{c,k}))$, is obtained with the inverse sensor model \refeqn{RayISMAnswer}--\refeqn{allEta} and entropy definitions \refeqn{ShannonsEntropyCell}, \refeqn{ShannonsEntropyMap}. The second term is found using \refeqn{allEta} with
\begin{align}
P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}}.
\end{align}

The goal of entropy-based autonomous exploration is to select the optimal pose $X^*_c$ that maximizes the expected information gain,
\begin{align}
\label{eqn:objectiveFunOptimized}
X_c^*&=\argmax_{X_c}{\ \mathcal I(X_c)},
\\
\label{eqn:objectiveFun}
\mathcal I(X_c)&=H(P(m|X_{1:t},Z_{1:t}))-\text{E}\left[H(P(m|X_c,Z_c))\right],
\end{align}
where \refeqn{DiscExpEntropyRay} is used repeatedly for several sample rays that compose scan $Z_c$ as shown in~\cite{KauAiLee16}.

Once an optimal pose $X^*_c$ is selected, a set of collision-free waypoints along the occupancy grid is easily obtained with Dijkstra's algorithm~\cite{Dij59}. Then, a constrained least squares optimization provides a trajectory as a tracking command for the robot controller.

\section{Mapping and Exploring in 3D}

\subsection{Mapping in 3D Space}

The inverse sensor model from~\cite{KauLeeAiMos16,KauTakAiLee17} is based on an arbitrary vector spanning 2D space. The distances from the robot to grid cells are obtained geometrically through ray casting. This method is easily extended to 3D space, and the inverse sensor model is applied identically.

However, implementation in 3D requires several careful considerations. First, the number of cell occupancy probabilities tracked with computer memory increases by a factor of the number of cells in the added dimension. In its current form, the proposed method assumes fixed map limits, so these should be carefully selected with grid cell resolution based on available memory. Second, measurement rays spanning 3D space typically involve more intersections with grid cells. The computational order of ray casting and the inverse sensor model are proportional to the number of grid cells along the ray $n_r$, so careful consideration must be placed on sensor limits and rays per scan that can be considered. Typically, computers small enough to be carried onboard flying robots lack the memory or processing capabilities to update massive maps quickly. Therefore, an onboard computer can simply stream the sensor data, and a more powerful computer can run the mapping.

Additionally, flying vehicles are frequently subject to fast movements, and the time stamps for pose estimates and depth sensor readings are not guaranteed to align. For example, the most recent pose estimate and measurement scan may have significantly different time stamps while the robot is rapidly rotating. Using these together violates the assumption that the robot pose is well-known for every measurement scan, i.e., $X_t$ and $Z_t$ are given. Therefore, pose estimates and sensor readings must be paired according to their time stamps. A useful tool that easily satisfies this goal is message filtering with the Robot Operating System (ROS), where stamped poses and measurement scans are synchronized using recently received data.

\subsection{Autonomous Exploration in 3D Space}

The goal of the proposed autonomous exploration strategy is moving to poses that maximize the information the robot expects to learn about the map, measured with entropy. To achieve this in real-time, careful consideration is placed on computational costs. In particular, the expected entropy computational order of a ray that intersects $n_r$ cells is $(n_r+1)^2$. Since most quadrotor flight is close to level, i.e., the third body-fixed axis (up/down) of the inertial and quadrotor body-fixed frames are nearly aligned, the proposed exploration uses 2D occupancy grid maps located at the desired exploration height. This approach decreases $n_r$ and simplifies the exploration problem with 3D space. These 2D maps project important properties of the 3D map, described next.

\subsubsection{Collision and Entropy Maps}
Occupancy grid map cell volume is frequently smaller than the size of the robot taking measurements, particularly with highly-detailed maps. This motivates a method to simply represent locations that risk collision, referred to as a \emph{collision map} $C$. Let $k_C\geq1$ be an integer representing the minimum number of 3D grid cells that form a cube that can completely encompass the robot, i.e., the collision map cell edge length is
\begin{align}
\label{eqn:alphaC}
\alpha_C=k_C\alpha.
\end{align}
Consider the set $m_{C,k}\subset m$ consisting of the 3D grid cells falling inside the limits of the $k$-th cell of collision map $C$, namely $\mathbf{C}_k$. Then the collision probability is
\begin{align}
P(\mathbf{C}_k)=\max{(P(\mathbf{m}_i)\ \forall \ i\in m_{C,k})}.
\end{align}
Then, the robot is only allowed to consider moving into $k$-th cell of the collision map if
\begin{align}
P(\mathbf{C}_k) \leq C_\text{thresh},
\end{align}
where $C_\text{thresh}>0$ is a maximum acceptable collision threshold, typically chosen slightly greater than $0$. In short, this conservative approach limits robotic motion to regions with a low risk of collision while neglecting objects far above or below the exploration plane.

In a similar method, \emph{entropy map} $E$ determines cells based on the entropy metric of \refeqn{ShannonsEntropyCell} over a different subset of cells. Let the $E$ be a 2D map composed of cells with edge length $\alpha$ located at the exploration height. Let the set $m_{E,k}\subset m$ be the cells directly above and below the $k$-th cell of $E$, neglecting any 3D cells located lower than the floor height. Selecting this set is advantageous because measurements are assumed not to penetrate occupied cells (edge length $\alpha$) and measurement rays are typically not closely-aligned with the third axis of the inertial frame (vertically long set). Furthermore, any cells located below the floor, which cannot be reached, should not impact the uncertainty of the space. Then, the occupancy probability of the $k$-th entropy map cell is selected as
\begin{align}
\label{eqn:ProbEntropyMap}
P(\mathbf{E}_k)=\argmax_{P(\mathbf{m}_i)}(H(P(\mathbf{m}_i))\ \forall \ i\in m_{E,k}).
\end{align}
After several measurements of the cells belonging to $m_{E,k}$, it is common for $H(P(\mathbf{E}_k))\approx0$, which implies that these cells are well-known. However, this does not imply that the captured space is \emph{entirely} free or occupied (e.g. a floor with free space above it, an object that does not breach the upper map boundary), so applying \refeqn{ProbEntropyMap} alone could arbitrarily set $P(\mathbf{E}_k)$ close to $0$ or $1$. This can have damaging effects on the expected information of cells beyond the $k$-th cell. Let $P_\text{min}>0$ be a lower bound such that $H(P_\text{min})=H(1-P_\text{min})\approx0$. If $H(P(\mathbf{E}_k))\leq H(P_\text{min})$, then this probability is corrected to
\begin{align}
P(\mathbf{E}_k)= 
\begin{cases}
    P_\text{min},			& \text{if } \text{mean}(P(\mathbf{m}_i)\ \forall \ i\in m_{E,k})\leq0.5\\
    1-P_\text{min},              & \text{otherwise}
\end{cases},
\end{align}
where the first case corresponds to mostly open space and the second case corresponds to mostly occupied space (e.g. walls and objects). In short, the candidates are only considered if they can be reached over collision map $C$ via Dijkstra's search, and the information gains of the reachable candidates are predicted using entropy map $E$.

\subsubsection{Collision and Entropy Combination Map}
The second proposed technique combines the advantages of both the collision and entropy maps into a 2D projected occupancy grid map, which can be used for both tasks. Let the \emph{collision and entropy combination map} $B$ be a 2D projected map located at the exploration height with grid cell edges the same as the collision map, namely $\alpha_C$ from \refeqn{alphaC}, using the same subset of grid cells from the collision map, namely $m_C$. Considering that the cells begin with uniform probability $P_\text{init}$ where $0<P_\text{init}<1$, define a threshold $P_\text{thresh}$ such that $P_\text{init}<P_\text{thresh}<1$. Then, the probability assigned to the $k$-th grid cell of $B$ is
\begin{align}
P(\mathbf{B}_k)= 
\begin{cases}
    P(\mathbf{C}_k),			& \text{if }P(\mathbf{C}_k)\geq P_\text{thresh}\\
    \min{(P(\mathbf{m}_i)\ \forall \ i\in m_{C,k})},              & \text{otherwise}
\end{cases}.
\end{align}

This approach has several advantages. First, measurements must enter the vicinity of the $k$-th cell for $P(\mathbf{B}_k)$ to change from $P_\text{init}$. Without prior measurements, capturing $P(\mathbf{B}_k)$ produces a large reward for the expected information gain in \refeqn{objectiveFun}. Second, the value of $P_\text{thresh}$ can be modified based on how conservative the collision risk assessment for a robot should be. Additionally, when $P(\mathbf{B}_k)<P_\text{thresh}$, this event tends to favor free cells such that expected information gains of cells beyond the $k$-th cell have greater impact on exploration. Furthermore, a single map is simpler, and since cell sizes satisfy $\alpha_C\geq\alpha$, it follows that $n_r$ can only be decreased with $\alpha_C$, thereby increasing computational speed.

\section{Numerical Results}

The 3D mapping and exploration algorithms are simulated using the first proposed approach with separate collision and entropy 2D projected maps. The Robot Operating System (ROS) provides an excellent framework for several programs to communicate easily. Gazebo serves as the simulator, which provides plugins to simulate an Asus Xtion color depth sensor and a Hokuyo LIDAR. It also provides position and attitude transforms from the world to the quadrotor body. Using fixed transforms between the quadrotor body and its onboard sensors, an original mapping node uses a variation of ROS message filters, known as transform message filters, to synchronize the sensor poses with the sensor reading time stamps. The node is designed to work for any number of sensors and uses the sensor properties directly when solving the inverse sensor model. In particular, the Xtion and Hokuyo are modeled using normalized Gaussian and uniform distributions for the sensor model, with a Gaussian probability of $0.9$ with $\sigma_\text{Xtion}=0.25$ and $\sigma_\text{Hokuyo}=0.1$, respectively, which are greater than their specified values to account for pose uncertainty and distortion with the Xtion. Since the Xtion tends to distort depths significantly beyond $4$m, this maximum reading considered for both the Xtion and Hokuyo, and the uniform distribution (probability $0.1$) covers the sensor ranges from $0.5$m to $4$m. A cell size of $\alpha=0.075$m is selected and the map limits are $-4$m to $4$m in the x-direction, $-6$m to $6$m in the y-direction, and $-0.15$m to $1.5$m in the z-direction. The initial probability of cells was selected at $P_\text{init}=0.1$.

The other original node is for autonomous exploration. For this section, separate 2D collision and entropy maps are used. For the collision map, the cell edge factor is $k_C=3$. Additionally, Dijkstra's algorithm produces a cost map composed of distance costs for each collision map cell, denoted $d_\text{cell}$, from the current robot position to the candidate pose locations over a collision-free path about the occupancy grid. A normalized variation on the \emph{bump function}~\cite{Joh06} with maximum distance $d_\text{max}>0$,% TODO: CHECK!
\begin{align}
\text{bump}(d_\text{cell})= 
\begin{cases}
    e^{1-\frac1{1-(d_\text{cell}/d_\text{max})^2}},			& \text{if }-d_\text{max}<d_\text{cell}<d_\text{max}\\
    0,              				& \text{otherwise}
\end{cases},
\end{align}
is multiplied to \refeqn{objectiveFun} when optimizing \refeqn{objectiveFunOptimized}. For this experiment, $d_\text{max}$ is chosen as the maximum value of $d_\text{cell}$ on the cost map. This serves to prioritize understanding the local surrounding space before the robot moves across the map to distant regions.

In the simulation, the robot explores the simulated environment for $5$min while taking measurements. The 3D occupancy grid map and true environment are shown in Fig. \ref{fig:sim3DMap}. The Gazebo simulated environment and two 2D projected maps for collision and entropy are shown in Fig. \ref{fig:sim2Dmaps}. The total map entropy as a function of time is shown in Fig. \ref{fig:simH}.

\begin{figure}[!t]
    	\begin{subfigure}[t]{0.24\columnwidth}
           	\centering
          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_1min.jpg}
        		\caption{$t=1$min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.24\columnwidth}
           	\centering
          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_2min.jpg}
        		\caption{$t=2$min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.24\columnwidth}
           	\centering
          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_3min.jpg}
        		\caption{$t=3$min}
   	\end{subfigure}
    	\begin{subfigure}[t]{0.24\columnwidth}
           	\centering
          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_4min.jpg}
        		\caption{$t=4$min}
    	\end{subfigure}
	\begin{subfigure}[t]{0.24\columnwidth}
           	\centering
          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_5min.jpg}
        		\caption{$t=5$min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.24\columnwidth}
           	\centering
          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_6min.jpg}
        		\caption{$t=6$min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.24\columnwidth}
           	\centering
          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_7min.jpg}
        		\caption{$t=7$min}
   	\end{subfigure}
    	\begin{subfigure}[t]{0.24\columnwidth}
           	\centering
          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_8min.jpg}
        		\caption{$t=8$min}
    	\end{subfigure}
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
\caption{Simulated 3D Occupancy Grid Map}
\label{fig:sim3DMap}
\end{figure}

\begin{figure}[!t]
	\centering
	\begin{subfigure}[t]{0.3\columnwidth}
           	\centering
          	\includegraphics[height=1.5\textwidth]{gazebo_view.png}
        		\caption{Environment}
    	\end{subfigure}
		\hspace*{0.05cm}
    	\begin{subfigure}[t]{0.3\columnwidth}
           	\centering
          	\includegraphics[trim = {3.6cm 17cm 52.5cm 4cm}, clip, height=1.5\textwidth]{gazebo_4min.jpg}
        		\caption{Collision Map}
    	\end{subfigure}
	\hspace*{0.1cm}
	\begin{subfigure}[t]{0.3\columnwidth}
           	\centering
          	\includegraphics[trim = {18cm 17cm 38.1cm 4cm}, clip, height=1.5\textwidth]{gazebo_4min.jpg}
        		\caption{Entropy Map}
    	\end{subfigure}
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
\caption{Simulated Environment and 2D Projected Maps at $4$min}
\label{fig:sim2Dmaps}
\end{figure}

\begin{figure}[!t]
	\centering
	\includegraphics[width=0.4\textwidth]{SimEntropy.pdf}
	\caption{Simulated Environment Map Entropy}
	\label{fig:simH}
\end{figure}

The two-map approach exhibits advantages and disadvantages. The exploration algorithm governs robotic motion to poses that were highly beneficial to the 3D occupancy grid of the local environment. The floor, which is not always visible, is properly estimated in most places. Conversely, this attention to detail has negative consequences that encourage the robot to repeatedly look at the same regions for different vantage points. Occasionally an entropy map cell, composed of all cells above the floor single-stacked vertically, is rarely captured, which yields large expected information gains, even though other regions might warrant observation more. This is because there is no explicit consideration for the \emph{number} of 3D cells with high uncertainty to projected onto the entropy map, only a single probability associated with the maximum entropy belonging to that set. In short, this algorithms successfully explores the space collision-free while producing a complete map, but over-attention to just a few uncertain cells has negative consequences on the exploration time.

\section{Experimental Results}

The experimental setup follows similar conditions with a few key differences. Most importantly, the second proposed exploration approach, namely a collision and entropy combination map, replaces the two-map approach. Additionally, instead of using a simulated environment such as Gazebo, where the robot state can be set and measured directly, a Vicon motion capture system provides pose data and nonlinear geometric flight controller~\cite{GooDaeLee13} provides the necessary forces and moments to follow the desired exploration trajectories, run onboard a Jetson TX2 computer. The Asus Xtion and Hokuyo Lidar are configured the same was as simulation, but their data is streamed from the Jetson TX2 via Wifi to a host computer running the mapping and exploration nodes with an Intel Core i7-6800K CPU (12$\times$3.40GHz).

The Laboratory for Autonomous Systems Research (LASR) at the U.S. Naval Research Laboratory (NRL) provided a large experimental space for testing. An experimental setup, involving Ram Board suspended from stanchions resembled a building floor plan, as well as miscellaneous objects like trash cans and tables. The map limits were $0.5$m to $10.3$m in the x-direction, $-8.3$m to $3.5$m in the y-direction, and $-0.15$m to $1.5$m in the z-direction. Pictures of the environment are shown in Fig. \ref{fig:exp3DEnvironment}, and the resulting 3D maps at various time steps are shown in Fig. \ref{fig:exp3DMap}. The 2D projected maps (x-direction upward) at corresponding time steps are shown in Figure \ref{fig:exp2DMap}, where trajectories are shown with curves to candidate poses, shown with arrows. The total map entropy as a function of time is shown in Figure \ref{fig:expH}.

% trim={<left> <lower> <right> <upper>}
\begin{figure}[!t]
\centering
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_north.jpg}
        		\caption{North Region}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_south.jpg}
        		\caption{South Region}
    	\end{subfigure}
	\caption{3D Environment}
	\label{fig:exp3DEnvironment}
\end{figure}

\begin{figure}[!t]
\centering
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_ogm3D_0min.jpg}
        		\caption{$0$min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_ogm3D_1min.jpg}
        		\caption{$1$min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_ogm3D_2min.jpg}
        		\caption{$2$min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_ogm3D_2min47sec.jpg}
        		\caption{$2$min $47$sec (end)}
    	\end{subfigure}
	\caption{3D Occupancy Grid Map and Point Clouds}
	\label{fig:exp3DMap}
\end{figure}

\begin{figure}[!t]
\centering
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_0min_2D.jpg}
        		\caption{$0$min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_1min_2D.jpg}
        		\caption{$1$min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_2min_2D.jpg}
        		\caption{$2$min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_2min47sec_2D.jpg}
        		\caption{$2$min $47$sec (end)}
    	\end{subfigure}
	\caption{2D Collision and Exploration Combination Map}
	\label{fig:exp2DMap}
\end{figure}

\begin{figure}[!t]
	\centering
	\includegraphics[width=0.4\textwidth]{ExpEntropy.pdf}
	\caption{Experimental Environment Map Entropy}
	\label{fig:expH}
\end{figure}

The projected map used for both collision and exploration demonstrated several benefits. The robot measured walls and objects with the 3D map, and nicely represented these as occupied space with the 2D projected map. Additionally, the robot explored the space much faster than with two separate maps, primarily because the exploration experiment avoided several

\begin{itemize}
	\item Important result: free space and walls correctly estimated, but features of the floor could be improved
\end{itemize}

% ATTENTION: left out patched segment size and polynomial order

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

\section{Conclusions and Future Work}

\begin{itemize}
	\item Paper demonstrated results for mapping and exploration can be successfully executed in 3D
	\item Note on differences with projected maps
	\item Future work: projected maps can serve multiple robots working cooperatively
	\item Future work: ray casting in 3D allows entropy calculation and exploration in 3D
\end{itemize}

% conference papers do not normally have an appendix

% use section* for acknowledgement
%\section*{Acknowledgment}
%
%
%The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliography{../../BibSources}
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}
%
%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%
%\end{thebibliography}

% that's all folks
\end{document}

### 2018-06-09 Misc/Other First Author/Attitude Observer/1_Introduction.tex
added Misc folder

%\begin{frame} %-----------------------------%
%\frametitle{}
%  \begin{itemize} \item \vspace*{0.1cm} \end{itemize} 
%\end{frame}   %-----------------------------%

\section*{Introduction}  
%\subsection*{Introduction}

\begin{frame} %-----------------------------%
\frametitle{Motivation}
	\begin{itemize} 
	\item Application of Attitude Control and Estimation
		\begin{itemize} 
		\item \Emph{Space systems (spacecraft)} 
		\vspace*{0.1cm} 
		\item Atmospheric flight vehicles (quadrotors)
		\vspace*{0.1cm} 
		\item Underwater vehicles, ground vehicles, robotic systems 
		\end{itemize}  
	\vspace*{0.3cm}
\pause	
	\item Attitude Dynamics of A Rigid Body  
		\begin{itemize} 
		\item Attitude: orientation of a body-fixed frame with respect to a reference frame.
		\vspace*{0.1cm} 
		\item Special Orthogonal Group:
	    	\begin{align*} %SO(3)%
	       	\SO = \{ R\in\Re^{3\times 3}\,|\, R^TR=I,~\mathrm{det}[R]=1\}
	       	\end{align*}		
		\end{itemize} 
%	\vspace*{0.3cm}
	\end{itemize} 
        	\begin{figure} %figure%
            \centerline{
            \includegraphics[width=2.4cm]{Mot_darwin1.jpg}\hspace*{0.3cm}
            \includegraphics[width=1.81cm]{sat.jpg} \hspace*{0.3cm}
            \includegraphics[width=1.8cm]{f22.jpg} \hspace*{0.3cm}
            \includegraphics[width=2.49cm]{orca2b.jpg} 
            }
          	\end{figure}	
\end{frame}   %-----------------------------%

\begin{frame} %-----------------------------%
\frametitle{Geometric Control and Estimation}
\begin{itemize} 
\item Geometric Control and Estimation on Nonlinear Manifold 
	\begin{itemize} 
	\item Questions of \Emph{global nature} require global analysis for satisfactory solutions.
	\vspace*{0.1cm} 	
	\item Control and estimation systems are designed from the perspective of \Emph{differential geometry}.
%	\vspace*{0.1cm} 
%	\item ``Coordinate-invariant'' and ``intrinsic formulations''.
	\end{itemize} 
	\vspace*{0.3cm}
%\pause	
\item Geometric Control and Estimation on $\SO$	
	\begin{itemize} 
	\item \tc{purple}{No singularities} in representing large angle rotational maneuvers.
	\vspace*{0.1cm} 
	\item \tc{purple}{No ambiguities} and unwinding associated with quaternions.
	\vspace*{0.1cm} 
	\item Remarkably \tc{purple}{compact form} of equations of motion for complex aerospace systems.
	\end{itemize} 	
\end{itemize} 
\end{frame}   %-----------------------------%

\begin{frame} %-----------------------------%
\frametitle{Topological Obstruction}
\begin{itemize} 
\item Topological Obstruction of $\SO$
	\begin{itemize} 
	\item Global asymptotic stability: \\ Lyapunov stability + \Emph{global attractivity}. 	
	\vspace*{0.05cm} 
	\item \tc{purple}{No \textit{continuous} controller} can globally asymptotically stabilize an equilibrium of an attitude control system	[Bhat,~2000].
	\end{itemize} 
	\vspace*{0.2cm}	
\item Continuous Attitude Control Law
	\begin{itemize}
	\item \textit{Almost} global asymptotic stability: the region of attraction excludes only a set of zero measure. 
	\vspace*{0.05cm} 
	\item \Emph{Convergence rate is slow} when the initial state is closed to undesired equilibria.
	\end{itemize} 	
\pause	
	\vspace*{0.2cm}	
	\item Literature
		\begin{itemize} 
		\item Attitude observer/estimator on $\SO$: [Mahony,~2008], [Vasconcelos,~2009], [Sanyal,~2012].
		\vspace*{0.05cm} 
		\item Hybrid attitude control system: [Mayhew,~2013], [Lee,~2013].
		\end{itemize}

\end{itemize}	
\end{frame}   %-----------------------------%

%\begin{frame} %-----------------------------%
%\frametitle{}
%\begin{itemize} \item \vspace*{0.1cm} \end{itemize} 
%\end{frame}   %-----------------------------%

### 2018-06-09 Conference/AAS 2015 DA/AAS15.tex
added Conference folder

% use paper, or submit
% use 11 pt (preferred), 12 pt, or 10 pt only

%\documentclass[letterpaper, preprint, paper,11pt]{AAS}	% for preprint proceedings
%\documentclass[letterpaper, paper,11pt]{AAS}		% for final proceedings (20-page limit)
%\documentclass[letterpaper, paper,12pt]{AAS}		% for final proceedings (20-page limit)
\documentclass[letterpaper, paper,10pt]{AAS}		% for final proceedings (20-page limit)
%\documentclass[letterpaper, submit]{AAS}			% to submit to JAS

\usepackage{bm}
\usepackage{amsmath}
\usepackage{subfigure}
%\usepackage[notref,notcite]{showkeys}  % use this to temporarily show labels
\usepackage[colorlinks=true, pdfstartview=FitV, linkcolor=black, citecolor= black, urlcolor= black]{hyperref}
\usepackage{overcite}
\usepackage{footnpag}	
\usepackage{amssymb,url}
\usepackage{graphicx,subfigure,tikz,graphicx}
\usepackage{color}
\usepackage{csquotes}		      	% make footnote symbols restart on each page
\usepackage{amssymb,amsmath,amsthm,times,graphicx,subfigure,tabularx,booktabs,colortbl,multirow,threeparttable}

\newcommand{\hilight}[1]{\colorbox{green}{#1}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\expm}{\ensuremath{\mathrm{expm}}}
\newcommand{\diag}{\mathop{\mathrm{diag}}\nolimits}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\ave}[1]{\mathrm{E}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand{\Ibd}{\ensuremath{\mathbf{1}_{bd}}}
\newcommand\circled[1]{%
  \tikz[baseline=(C.base)]\node[draw,circle,inner sep=0.5pt](C) {#1};\!
}

\PaperNumber{27-432}

\begin{document}

\title{Minimum Uncertainty JPDA Filter and Coalescence Avoidance Performance Evaluations}

\author{Evan Kaufman\thanks{Doctoral Student, Mechanical and Aerospace Engineering, George Washington University, 801 22nd St NW, Washington, DC 20052, Email: evankaufman@gwu.edu.},  
T. Alan Lovell\thanks{Research Aerospace Engineer, Air Force Research Laboratory, Space Vehicles Directorate, Kirtland AFB, NM.},
\ and Taeyoung Lee\thanks{Assistant Professor, Mechanical and Aerospace Engineering, George Washington University, 801 22nd St NW, Washington, DC 20052, Tel: 202-994-8710, Email: tylee@gwu.edu.}
\thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}
}

\maketitle{}

\begin{abstract}
Two variations of the joint probabilistic data association filter (JPDAF) are derived and simulated extensively in this paper.
First, an analytic solution for an optimal gain that minimizes posterior estimate uncertainty is derived, referred to as the minimum uncertainty JPDAF (MUJPDAF).
Second, the coalescence-avoiding optimal JPDAF (C-JPDAF) is derived, which removes coalescence by minimizing a weighted sum of the posterior uncertainty and a measure of similarity between estimated probability densities.
Both novel algorithms are tested in much further depth than any prior work to show how the algorithms perform in various scenarios.
In particular, the MUJPDAF more accurately tracks objects than the conventional JPDAF in all simulated cases.
When coalescence degrades the estimates at too great of a level, and the C-JPDAF is often superior at removing coalescence when its parameters are properly tuned.
\end{abstract}

\section{Introduction}

Data association has been extensively studied with various applications in estimation when measurements are not necessarily originating from a single object.
We consider applications wherein multiple objects of interest enter the field of view of a sensor; however, the measurement origins are unknown.
To properly track these objects, we require estimation techniques that associate objects with measurements.
When the objects are in very close proximity, the association uncertainties grow, and an effective algorithm is necessary to maintain close estimations of the objects~\cite{KauLovLee14}.

A variety of data association techniques are considered for close-proximity trajectories with varying performance.
Multiple hypothesis tracking (MHT) and probability hypothesis density (PHD) techniques involve a high computational load, especially when objects are in close proximity, making real-time implementation more difficult~\cite{MHT1,PHD1,PHD2}.
Many other techniques are based on the Kalman filter (KF) or the extended KF (EKF), which provide a framework for recursive data association techniques for easier real-time implementation.
There are two popular strategies among recursive approaches, namely \emph{hard decisions} and \emph{soft decisions}~\cite{JPDAF1}.
A hard decision is when estimates are fully associated with measurements according to some metric between the estimates and the measurements.
For example, a common heuristic approach is known as the nearest neighbor filter (NNF), which uses the Mahalanobis distance between each measurement and the predicted measurements of each potential object to determine likely associations.
The measurement with the smallest Mahalanobis distance is considered as the correct one~\cite{NN2}.
However, when estimates are in close proximity, there is a high likelihood of making incorrect associations, which may be detrimental to the estimated states, so this technique is avoided in the analysis of this paper.

Alternatively with a soft decision technique such as the joint probabilistic data association filter (JPDAF), measurements are shared among the estimates. With the JPDAF, the likelihoods of association between measurements and estimates serve as weighting factors inside the measurement update portion of the filtering~\cite{JPDAF1,TrackDataAssoc}.
The JPDAF is particularly advantageous in multi-object scenarios because when objects share measurements, the soft decisions prevent loss of information, which occurs with hard decision approaches.
%The a posteriori uncertainties of the estimates are changed because of an additional term representing the spread in the innovations; the Kalman gain serves as a close estimate to an optimal gain in the sense that in nearly minimizes this a posteriori gain as derived by the linearized equations of the JPDAF.
The JPDAF is based on the Kalman gain, which minimizes the posterior uncertainty of individual tracks.
However, the Kalman gain is not guaranteed to \emph{exactly} minimize the posterior uncertainties as they are defined for the JPDAF.
In fact, the JPDAF is applied to multi-target tracking applications widely with the incorrect assumption that the Kalman gain yields an optimal estimator.
%The unnecessarily large uncertainty of the suboptimal Kalman gain serves as motivation to derive an optimal gain that minimizes the a posteriori uncertainty of the linearized JPDAF equations.

In this paper, we show the derivation of the optimal estimation gain that minimizes the magnitude of the posterior covariance matrix that maintains the framework of the JPDAF.
The resulting minimum uncertainty JPDAF is referred to as the MUJPDAF, where this new algorithm is neither more complicated nor more computationally expensive than the conventional JPDAF.
We show that decreasing estimate uncertainty to its minimum value increases the confidence of associating measurements to estimates through various numerical examples.

Additionally, this paper deals with cases when neighboring objects share measurements, where the estimates are subject to coalescence.
Most techniques to avoid track coalescence involve pruning the tracks, or designing a set of rules that limit the possible outcomes to likely scenarios. These rules typically involve exact nearest neighbor probabilistic data association (ENNPDA), which applies a strong decision to the most likely associated estimate. Pruning algorithms incorporate the ENNPDA algorithm because  estimates do not share measurements, making it insensitive to track coalescence~\cite{Coal1}.
For example in~\cite{Fitzgerald}, the algorithm involves a set of rules to prune the tracks that share measurements.
The algorithm considers the possible combinations of estimate-to-measurement associations over a time window and applies a logic that assigns estimates to likely scenarios; only a small set of possible scenarios are considered and lower likelihood scenarios are deleted~\cite{Coal_d,Coal_e,Coal_c}.
However, track pruning assumes that low likelihood association scenarios are incorrect, which is not necessarily true.
Furthermore, removing the weighting scheme of the JPDAF causes the estimates to lose information.

Therefore, we propose a new coalescence-avoiding optimal JPDAF, referred to as the C-JPDAF, which is a systematic approach to remove coalescence by minimizing the weighted sum of the uncertainty and a similarity index between estimates.
The prior work from~\cite{KauLovLee14} covers the concept in simple cases, assuming that only two objects are considered with a single measurement at each time step, and that the state and measurement are very similar.
These assumptions are unrealistic in most scenarios.
Therefore, we generalize the C-JPDAF algorithm in this paper to handle (i) any number of objects and measurements, (ii) missed detections and measurements originating from extraneous clutter, and (iii) measurements with a much different interpretation than just the object states.

This paper serves two major purposes:
first, it provides a thorough derivation and relevant proofs for the MUJPDAF and C-JPDAF for a generalized system of estimates where measurement origins are uncertain.
Second, the algorithms undergo extensive numerical simulations with orbiting satellites that provide insight into the effects of various parameters on the algorithms.

% Cut from the abstract (summary of what is above)
%This paper analyzes variations on a popular approach known as the joint probabilistic data association filter (JPDAF).
%The JPDAF is advantageous because of its low computational requirements for real time implementation and its heavily tested successful data association with missed detections and measuring extraneous clutter.
%The JPDAF is based on the Kalman gain, which minimizes the posterior uncertainty of individual tracks.
%However, the Kalman gain is not guaranteed to \emph{exactly} minimize the posterior uncertainties of the JPDAF.
%Therefore, a minimum uncertainty JPDAF (MUJPDAF) is derived, which provides an analytic solution to a gain that minimizes the posterior uncertainty~\cite{KauLovLee15}.
%This algorithm shows significant improvements by decreasing estimate uncertainties and coalescence when compared with the JPDAF.
%
%Additionally, when measurements are shared among the estimates, the measurement updates tend to coalesce the estimates.
%A generalized coalescence-avoiding optimal JPDAF is proposed, where a similarity measure is included in the cost function to solve for a gain matrix inside the filter~\cite{KauLovLee15}.
%
%However, the MUJPDAF and C-JPDAF are only simulated in a few scenarios.
%This paper shows the performance of the JPDAF, MUJPDAF, and C-JPDAF in various cases with orbiting satellites where associating measurements to estimates is a challenge.

\section{Problem Formulation}

This section describes the dynamics of objects and the measurement model considered in this paper and a filtering problem is formulated.

\subsection{Dynamics of Objects and Measurements}

%
% Linearized equations of motion: assumed known
% Assume the Gaussian noise model and define unknown random variables $w$ (process noise) and $v$ (measurement noise)
% Initial conditions are assumed known, define $\hat x_{i,0}$, $P^+_{i,0}$

%Define the set of all objects $I=\{1,2,...,t\}$ such that the object index $i\in I$ where the total number of objects $t$ is assumed known.
Consider $n_t$ objects indexed by $i\in I=\{1,2,...,n_t\}$, where $n_t$ is assumed known.
The dynamics of the $i$-th object are modeled as a linear stochastic discrete-time system described by
\begin{align}
x_{i_{k}} & = F_{i_{k-1}} x_{i_{k-1}} + w_{i_{k}},\label{eqn:xkp}\\
z_{i_k} & = H_{i_k} x_{i_k} + v_{i_k},
\end{align}
where $x_{i_k}\in\Re^n$ is the state vector, $z_{i_k}\in\Re^m$ is the output vector, and  the matrices $F_{i_{k-1}}\in\Re^{n\times n}$ and $H_{i_k}\in\Re^{m\times n}$ describe system dynamics and outputs.
The process noise and the measurement noise are denoted by $w_{i_k}\in\Re^n$ and $v_{i_k}\in\Re^m$, respectively, where they are assumed to be zero-mean, mutually independent white, Gaussian noise sequences with known covariance matrices $Q_{i_k}\in\Re^{n\times n}$ and $R_{i_k}\in\Re^{m\times m}$, respectively, i.e.,
\begin{align}
w_{i_k} \sim \mathcal{N}[0,Q_{i_k}],\quad
v_{i_k} \sim \mathcal{N}[0,R_{i_k}],
\end{align}
where $\mathcal{N}[\mu,\sigma]$ denotes the Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.

Each initial state $\bar x_{i_0}\in\Re^n$ and covariance $P_{i_0}\in\Re^{n\times n}$ are also assumed to be mutually independent with noise. They are Gaussian with known mean values and covariance matrices:
\begin{align}
x_{i_0} \sim \mathcal{N}[\bar x_{i_0}, P_{i_0}].\label{eqn:xi0}
\end{align}

\subsection{Measurement Origination Model}
\label{DAP}

% Likelihoods of association are calculated from the innovations and innovation covariances only
% Two techniques are used to obtain the likelihoods of association, both of which may be applied to the JPDAF, MUJPDAF, and C-JPDAF
	% "Cheap JPDAF": inexpensive calculation, present the four equations and two references
	% True likelihood of association: infeasible for many real-time applications, calculated in the appendix

%Define the set $J=\{1,2,...,r\}$ such that the measurement index $j\in J$ where $r$ is a variable number of measurement returns such that $r\geq1$.
Suppose that there are $n_r$ measurements indexed by $j\in J=\{1,2,...,n_r\}$.
Let $\beta_{ij}$ be the probability that the $i$-th system is associated with measurement $j$.
Two techniques may be applied to solve for every $\beta_{ij}$, both of which require the innovation and its covariance (defined later) and information on detection probabilities, which are assumed known.
The first technique is applied in the standard JPDAF that calculates the true association probabilities from~\cite[Eq. 9.(45-46)]{TrackDataAssoc}.
However, the computation required with a standard JPDAF is inconsistent with the capabilities of the computational resources to account for an arbitrary number of objects to be tracked~\cite{Bar1990}.

A modified version of the JPDA, commonly known as the ``Cheap JPDA'', provides a simpler approximation of the likelihood $\beta_{ij}$.
The equations below are adopted from~\cite[Eq. 1.(1-2)]{Bar1990}.
Let $G_{ij}$ be a variable which takes a very similar form as the Gaussian probability density function
\begin{align}
G_{ij}=\frac1{\sqrt{\det{S_{i}}}}\exp{\{-\frac12e_{ij}^TS_{i}^{-1}e_{ij}\}},
\end{align}
where $e_{ij}$ is the innovation and $S_{i}$ is the a priori innovation covariance, defined later in \refeqn{eij} and \refeqn{S}, respectively.
This is used to calculate intermediate variables $\mathcal{S}_{ti}$ and $\mathcal{S}_{ri}$
\begin{align}
\mathcal{S}_{ti}=\sum\limits_{j\in J}G_{ij}, 
\quad 
\mathcal{S}_{ri}=\sum\limits_{i\in I}G_{ij}.
\end{align}
A close approximation of the probability that the $i$-th system is associated with the $j$-th measurement is
\begin{align}
\beta_{ij}=\frac{G_{ij}}{\mathcal{S}_{ti}+\mathcal{S}_{ri}-G_{ij}+B}, 
\end{align}
where $B$ is a correction factor to account for a non-unity probability of detection.

\section{Minimum Uncertainty Probabilistic Data Association Filter (MUJPDAF)}
\label{MUJPDAF}

In this section, we derive the MUJPDAF with a similar framework to that of the JPDAF.
Differences between the JPDAF and MUJPDAF occur in the measurement update where the gain is derived, which changes the posterior covariance equation.
As a result of the changes described in this section, the MUJPDAF serves to update state estimates from measurements in a soft decision approach such that the state estimate uncertainties are minimized; the conventional JPDAF
%does \emph{not} guarantee
fails
to minimize these uncertainties.

\subsection{Flow Update}

During the flow update, the a priori (denoted with superscript $-$) estimated state $\hat x_{i_k}^-$ is found with \refeqn{xkp} and its a priori covariance $P_{i_k}^-$ is calculated from the last posterior (denoted with superscript $+$) covariance $P_{i_{k-1}}^+$ for all $i$ as
\begin{gather}
%\hat x_{i_{k}}^- = f(\hat x_{i_{k-1}}^+),\\
P_{i_{k}}^- = F_{i_{k-1}} P_{i_{k-1}}^+ F_{i_{k-1}}^T + Q_{i_{k}}.
\end{gather}
%These equations provide predictions of the states and their uncertainties.
% Equations for a priori state estimates and covariances

\subsection{Measurement Update}

The measurements serve to update the state estimates. Because all equations in this section take place at the $k$-th time step, this subscript is removed from the equations. For a given measurement $z_j$, its measurement residual or innovation for the $i$-th object is given by
\begin{align}
e_{ij} = z_j - \hat z_i,\label{eqn:eij}
\end{align}
where $\hat z_i$ is the predicted measurement from $\hat x_{i}^-$ and a known measurement matrix $H_i$ such that
\footnote{Note that for nonlinear systems where $z_i = h_i(x_i)$, then $\hat z_i = h_i(\hat x_{i}^-)$ replaces \refeqn{zlin} and the linearized matrix $H_i=\deriv{h_i}{x_i}\bigg|_{x_i=\hat x_{i}^-}$ is used for the remaining equations.\label{fn:nonlinsys}}
\begin{align}
\label{eqn:zlin}
\hat z_i = H_i\hat x_{i}^-.
\end{align}
%where the true measurement is expected to be
%\begin{align}
%z_i=h(x_i)+v_i, \quad H_i=\frac{\partial h}{\partial x}\bigg|_{x=\hat x_{i}^- }.
%\end{align}
Given the uncertainty of the measurements represented with covariance $R_i$, the a priori innovation covariance $S_{i}$ is
\begin{align}
S_{i}=H_{i}P_{i}^{-}H_{i}^T+R_{i}.\label{eqn:S}
\end{align}

The weighted sum of all innovations associated with the object $i$ is defined as ${\bf\bar{e}}_{i}={\sum\limits_{j\in J} \beta_{ij}e_{ij}}$.
The resulting posterior estimate of the state is given by
%~\cite[Eq. 6-(25-26)]{TrackDataAssoc}
\begin{align}
\hat x^+_{i}=&\ \hat x^-_{i}+K_{i}{\bf\bar{e}}_{i}.\label{eqn:KalEst}
\end{align}
The conventional JPDAF defines $K_i$ from the gain of the Kalman filter, given by $K_i=P^-_iH_i^TS_i^{-1}$, based on an incorrect assumption that the Kalman gain yields an optimal estimator for the JPDAF.
Here we derive the optimal gain $K_i$ such that the posterior uncertainties are minimized with respect to the generalized JPDAF equations.

Let the probability that object $i$ is \emph{not} associated with any measurement be $\beta_{0,i}=1-{\sum\limits_{j} \beta_{ij}}$.
The resulting posterior covariance can be written as
%~\cite[Eq. D-32 and D-33]{TrackDataAssoc}
\begin{align}
\label{eqn:JPDAFCov}
P^+_{i}=&\ \beta_{0,i}P^-_{i}+(1-\beta_{0,i})P_i^c+\tilde P_i,
\end{align}
where $\tilde P\in\Re^{n\times n}$ is
\begin{align}
\tilde P_i=&\ K_i\parenth{{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf\bar{e}}_{i}}{\bf\bar{e}}_{i}^T}K_{i}^T,\label{eqn:tildeP}
\end{align}
and $P_i^c$ corresponds to the posterior covariance of object $i$ if only a single measurement serves to update this track~\cite{TrackDataAssoc}%, and it can be written as
%We can define $P_i^c$ without any assumptions on $K_i$ (i.e. $K_i$ is not necessarily the Kalman gain, so the well-known Kalman covariance update equations are not valid here) as~\cite[Eq. 4.2-12]{OptEst1}
\begin{align}
P^c_i=&\ (I-K_iH_i)P^-_i(I-K_iH_i)^T+K_iR_iK_i^T\nonumber
\\
=&\ P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i(H_iP^-_iH_i^T+R_i)K_i^T,
\label{eqn:GenCov}
\end{align}
(see~\cite{OptEst1}).
Substituting \refeqn{tildeP} and \refeqn{GenCov} into \refeqn{JPDAFCov},% and rearranging,
\begin{align}
P^+_{i}=&\ \beta_{0,i}P^-_{i}+(1-\beta_{0,i})[P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i(H_iP^-_iH_i^T+R_i)K_i^T]\nonumber
\\
&\ +K_i\left({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf\bar{e}}_{i}}{\bf\bar{e}}_{i}^T\right)K_{i}^T\nonumber
\\
=&\ P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i\left((1-\beta_{0,i})S_i+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf\bar{e}}_{i}}{\bf\bar{e}}_{i}^T\right)K_i^T\nonumber%asdf
\\
=&\ P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_iE_iK_i^T,\label{eqn:CovExpanded}
\end{align}
where the symmetric matrix $E_i=E_i^T\in\Re^{m\times m}$ is defined as
\begin{align}
E_i=&\ (1-\beta_{0,i})S_i+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf\bar{e}}_{i}}{\bf\bar{e}}_{i}^T.\label{eqn:E}
\end{align}
Note that $E_i$ is independent of $K_i$ and is invertible when $\beta_{0,i}\neq1$ since $S_i$ is positive-definite and
%may be calculated without knowledge of the gain matrix $K_i$, and is invertible because $P^-_i$ and $R_i$ are positive-definite and
$(\beta_{ij}e_{ij}e_{ij}^T-{\bf\bar{e}}_{i}{\bf\bar{e}}_{i}^T)$ is positive-semidefinite, as shown in the Appendix A. 
The case when $\beta_{0,i}=1$ corresponds with the $i$-th object \emph{not} being measured with full certainty, analyzed in Appendix B.

As with the Kalman filter, we choose the uncertainty cost function $J_{p_i}$ to be the trace of the posterior covariance
\begin{align}
\label{eqn:JpIntro}
J_{p_i}=\tr{P_i^+}.
\end{align}
We choose the gain that minimizes $J_{p_i}$, which requires that
\begin{align}
0&=\frac{\partial\tr{P^+_i}}{\partial K_i}\nonumber
\\
&=-2(1-\beta_{0,i})\frac{\partial\tr{K_iH_iP^-_i}}{\partial K_i}+\frac{\partial\tr{K_iE_iK_i^T}}{\partial K_i}\nonumber%asdf
\\
&=-2(1-\beta_{0,i})P^-_iH_i^T+2K_iE_i.\label{eqn:Uncertainty0}
\end{align}
Therefore the optimal gain is given by
\begin{align}
K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1},\label{eqn:GainK}
\end{align}
where this choice of $K_i$ yields a stationary point of $\tr{P_i^+}$, which is a minimum, shown in Appendix C.
%As with the Kalman filter, we choose the estimation gain to minimize the magnitude of the posterior covariance measured by $\tr{P_i^+}$.
%The first derivative necessary condition for optimality requires that
%\begin{align}
%0&=\frac{\partial\tr{P^+_i}}{\partial K_i},\nonumber
%\\
%&=-2(1-\beta_{0,i})\frac{\partial\tr{K_iH_iP^-_i}}{\partial K_i}+\frac{\partial\tr{K_iE_iK_i^T}}{\partial K_i},\nonumber
%\\
%&=-2(1-\beta_{0,i})P^-_iH_i^T+2K_iE_i.\label{eqn:Uncertainty0}
%\end{align}
%%Therefore the optimal gain is given by
%%\begin{align}
%%K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
%%\end{align}
%Solving \refeqn{Uncertainty0} to satisfy necessary condition for optimality,
%\begin{align}
%K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
%\end{align}
%The Hessian

Finally, \refeqn{GainK} is substituted into \refeqn{CovExpanded} to obtain
\begin{align}
\label{eqn:PNotSimplified}
P^+_{i}&=P^-_{i}-(1-\beta_{0,i})^2P^-_iH_i^TE_i^{-1}H_iP^-_i,
\end{align}
which may be simplified in two ways
\begin{align}
P^+_{i}&=P^-_{i}-K_iE_iK_i^T, \quad \mbox{or equivalently}\label{eqn:JPDAFPostCovSymmetricUpdate}
\\
P^+_{i}&=\left(I-(1-\beta_{0,i})K_iH_i\right)P^-_i.\label{eqn:JPDAFPostCovSimpleUpdate}
\end{align}
The proposed gain is compared with the Kalman gain explicitly in Appendix D.
In short, the differences between the MUJPDAF and the conventional JPDAF are shown in \refeqn{GainK} and \refeqn{JPDAFPostCovSymmetricUpdate} or \refeqn{JPDAFPostCovSimpleUpdate}. As the proposed optimal gain minimizes the posterior uncertainties, the corresponding MUJPDA performs better than the conventional JPDAF, as illustrated by numerical examples later.

% Abstract version:
%The measurements serve to update the state estimates. All equations in this section take place at the $k$ time step, so this is removed. For a given measurement $z_j$, its measurement innovation is
%\begin{align}
%e_{ij} = z_j - \hat z_i,\label{eqn:eij}
%\quad \hat z_i = H_i\hat x_{i}^-.
%\end{align}
%
%Let $\beta_{ij}$ be the probability that the $i$-th system is associated with measurement $j$ where $j\in J=\{1,2,...,n_r\}$~\cite{KauLovLee15}.
%The weighted sum of all innovations associated with the object $i$ is defined as ${\bf{e}}_{i}={\sum\limits_{j\in J} \beta_{ij}e_{ij}}$.
%The resulting posterior estimate of the state is given by
%%~\cite[Eq. 6-(25-26)]{TrackDataAssoc}
%\begin{align}
%\hat x^+_{i}=&\ \hat x^-_{i}+K_{i}{\bf{e}}_{i}.\label{eqn:KalEst}
%\end{align}
%The conventional JPDAF defines $K_i$ from the gain of the Kalman filter, given by $K_i=P^-_iH_i^TS_i^{-1}$ where $S_{i}=H_{i}P_{i}^{-}H_{i}^T+R_{i}$, based on an incorrect assumption that the Kalman gain is always optimal for data association with the JPDAF.
%
%Let the probability that object $i$ is not associated with any measurement be $\beta_{0,i}=1-{\sum\limits_{j} \beta_{ij}}$.
%The resulting posterior covariance can be written as~\cite{KauLovLee15}
%\begin{align}
%\label{eqn:JPDAFCov}
%P^+_{i}
%&=P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)
%+K_iE_iK_i^T
%\quad \mbox{where}
%\\
%E_i&=(1-\beta_{0,i})S_i+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T.
%\end{align}
%
%
%The gain $K_i$ that minimizes $J_P=\tr{P^+_i}$ is given by
%\begin{align}
%K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
%\end{align}
%
%Finally, \refeqn{GainK} is substituted into \refeqn{JPDAFCov} to obtain
%%\begin{align}
%%%\begin{split}
%%%P^+_{i}=&P^-_{i}-(1-\beta_{0,i})^2(P^-_iH_i^TE_i^{-1}H_iP^-_i
%%%\\
%%%&+P^-_iH_i^TE_i^{-1}H_iP^-_i
%%%-P^-_iH_i^TE_i^{-1}E_iE_i^{-1}H_iP^-_i)
%%%\end{split}
%%%\\
%%P^+_{i}=&P^-_{i}-(1-\beta_{0,i})^2P^-_iH_i^TE_i^{-1}H_iP^-_i,
%%\end{align}
%%which may be simplified in two ways
%\begin{align}
%P^+_{i}=&P^-_{i}-K_iE_iK_i^T, \quad \mbox{or equivalently}\label{eqn:JPDAFPostCovSymmetricUpdate}
%\\
%P^+_{i}=&\left(I-(1-\beta_{0,i})K_iH_i\right)P^-_i.\label{eqn:JPDAFPostCovSimpleUpdate}
%\end{align}
%The proposed gain is \emph{not} guaranteed to be equivalent to the Kalman gain because the Kalman gain does not include the uncertainty due to the spread in the innovations~\cite{KauLovLee15}.
%In short, the differences between the MUJPDAF and the conventional JPDAF are shown in \refeqn{GainK} and \refeqn{JPDAFPostCovSymmetricUpdate} or \refeqn{JPDAFPostCovSimpleUpdate}.

\section{Coalescence-Avoiding Optimal JPDAF (C-JPDAF)}

% Abstract version:
%Including a similarity index serves to remove this coalescence~\cite{KauLovLee14}.
%Let $\mathbf{J}\in\Re$ be a cost function% of coalescence avoiding filtering as
%\begin{align}
%\mathbf{J}=J_P+cJ_S,
%\end{align}
%where $J_P$ is an uncertainty cost defined by
%\begin{align}
%J_P=\sum\limits_{i}\tr{P^+_i},\label{eqn:JP}
%\end{align}
%and $J_S$ is an uncertainty cost defined by
%\begin{align}
%J_S=\sum\limits_{p,q\in I,p\neq q}\exp (-au_{pq}),
%\quad
%u_{pq}=(\hat z_{p}^+-\hat z^+_{q})^T{S^+_{pq}}^{-1}(\hat z^+_{p}-\hat z^+_{q}),
%\end{align}
%where $\hat z_i^+=H_i\hat x_i^+$, $S^+_i=H_iP^+_iH_i^T+R_i$, $p,q\in I$ such that $p\neq q$, and $S^+_{pq}=S^+_{p}+S^+_{q}$.
%
%Necessary conditions for optimality are given by
%\begin{align}
%\deriv{\mathbf{J}}{K_i} = \deriv{J_P}{K_i} + c \deriv{J_S}{K_i} =0.\label{eqn:NCO}
%\end{align}

Coalescence of the JPDAF (or variations thereof such as the MUJPDAF described in the last section) is the outcome of estimates becoming closer to each other as a result of sharing measurements. This occurs because the signs of the innovations typically cause updates that draw measurement estimates towards the observed measurements; the net effect tends to cause estimates to become more similar to each other (e.g. a large number of estimates in close proximity, estimates with large uncertainties).
The goal of the C-JPDAF is to minimize a cost that includes a similarity index in addition to uncertainty.
This concept is explained and very simple numerical examples are presented in~\cite{KauLovLee14}.
These are based on the assumptions that during every filter update, exactly two objects are present in the field of view of a sensor, and only a single measurement at each time step is available.
This measurement originates from one of the two objects with equal probability that it originated from either object.
These assumptions severely limit the scenarios where this algorithm can be useful.
Therefore in this paper, the derivation and numerical examples do not operate on such restrictive assumptions.
Additionally, the similarity measure is slightly changed to be in terms of measurements and measurement estimates (measurement space), so the technique may be applied to a wider variety of systems.
This approach to the problem further contrasts with~\cite{KauLovLee14}, which derives the similarity index in terms of the state variables (state space).

\subsection{Matusita's Similarity Measure of Many Estimates}

The Matusita's similarity measure is defined between two Gaussian distributions.
Let $y_i$ be a random variable corresponding to the $i$-th system of set $I$ such that $y_i \sim \mathcal{N}[\hat y_i,{\mathbf P}_i]$, and let indices $p,q\in I$ such that $p\neq q$.
From~\cite{KauLovLee14}, the portion of the Matusita's similarity measure corresponding to the spread in the estimate means is given by
\begin{align}
\zeta_{pq}=\exp \{-a(\hat y_{p}- \hat y_{q})^T({\mathbf P}_{p}
+{\mathbf P}_{q})^{-1}(\hat y_{p}-\hat y_{q})\},\label{eqn:Mat2Est}
\end{align}
where $a>0$ shapes the similarity measure.

However, the distributions of more than two estimates may exhibit a nonzero overlap.
The similarity measure is changed to suit \emph{any} number of estimates by summing all combinations of Matusita's similarity measures, illustrated in Figure \ref{fig:SimMeas}, expressed in a general sense as
\begin{align}
\label{eqn:zetaGeneral}
\zeta=\sum\limits_{p,q\in I,p\neq q}\exp \{-a(\hat y_{p}- \hat y_{q})^T({\mathbf P}_{p}
+{\mathbf P}_{q})^{-1}(\hat y_{p}-\hat y_{q})\},
\end{align}
which is the summation of every similarity measure from \refeqn{Mat2Est} between all different Gaussian distributions $y_i$.
\begin{figure}
\centerline{
		{\includegraphics[width=.7\columnwidth]{GaussianDistOverlap1D.png}}
	}
\caption{A similarity index between three Gaussian distributions is illustrated as summation of overlapping regions of two distributions (green) or three distributions (brown).
This concept is easily extended to multivariate distributions.
}\label{fig:SimMeas}
\end{figure}

\subsection{Optimal Coalescence Avoidance}

Soft decision algorithms such as the JPDAF and MUJPDAF, which only consider estimate uncertainties when determining estimator gains, yield coalescence among the estimates.
However, including a similarity index in an optimal fashion to determine each gain matrix serves to remove this coalescence.
Let $\mathbf{J}\in\Re$ be a cost function of coalescence avoiding filtering as
\begin{align}
\mathbf{J}=J_P+cJ_S,
\end{align}
where $J_P$ is an uncertainty cost that measures the confidence level of the estimates and $J_S$ is a similarity cost that measures a degree of coalescence. The constant $c$ is a positive weighting factor between $J_P$ and $J_S$. 
The estimator gain $K_i$ is selected to minimize the total cost $\mathbf{J}$.
In short, the estimator gain is chosen in the consideration of both uncertainty and coalescence, illustrated in Figure \ref{fig:CostTrends}.

\begin{figure}
\setlength{\unitlength}{0.05\columnwidth}
\centerline{\footnotesize\selectfont
\begin{picture}(8.5,7.3)(-0.5,-0.8)
\put(-0.5,3){\rotatebox{90}{Cost}}
\put(0,0){\includegraphics[width=0.4\columnwidth]{ECC14_fig1.pdf}}
\put(5.3,1.2){\shortstack[c]{Uncertainty\\ cost: $J_P$}}
\put(6.2,2.6){\shortstack[c]{Similarity\\ cost: $J_S$}}
\put(5.2,5.2){\shortstack[c]{Weighted sum:\\ $\mathbf{J}=J_P+cJ_S$}}
\put(4.98,0.78){\circle*{0.1}}
\put(2.48,1.57){\circle*{0.1}}
\put(4.8,-0.4){$K_{MUJPDAF}$}
\put(2.3,-0.4){$K_{C-JPDAF}$}
\put(3.1,-0.8){Estimator gain}
\end{picture}
}
\caption{Illustration of optimal coalescence avoidance: several costs are illustrated with respect to an element of an estimator gain $K$ (red: uncertainty cost $J_P$, blue: similarity cost $J_S$, black: weighted sum). The estimator gain $K_{MUJPDAF}$ is selected to minimize the uncertainty cost. The estimator gain $K_{C-JPDAF}$ is chosen such that the weighted sum of the uncertainty cost and the similarity cost is minimized. In short, at the expense of increased uncertainty at a single step, the proposed approach prevents coalescence. The increased uncertainty at each step is rewarded by reducing the overall estimation error drastically by maintaining correct data association during close proximity tracking scenarios or missions.}
\label{fig:CostTrends}
\end{figure}

\paragraph*{Uncertainty Cost}
The uncertainty cost $J_P$ is defined as the sum of the trace of each posterior covariance:
\begin{align}
J_P=\sum\limits_{i}\tr{P^+_i},\label{eqn:JP}
\end{align}
where $P^+_{i}$ is obtained by \refeqn{JPDAFCov}.
Therefore, $P_i$ does not depend on $K_j$ for $i\neq j$.
The derivative of $J_P$ is taken from \refeqn{Uncertainty0}
\begin{align}
\label{eqn:CostP}
\frac{\partial J_{P}}{\partial K_{i}}&=2K_iE_i-2(1-\beta_{0,i})P^-_iH_i^T.
\end{align}

\paragraph*{Similarity Cost}
Any shared measurement $z_j$ on measurement estimate $\hat z_i$ yields an innovation $e_{ij}$ according to \refeqn{eij}.
During the measurement update, this innovation typically has the net effect of drawing $\hat z_i$ towards $z_j$ if $\beta_{ij}$ is nonzero.
This occurs for all estimates and all measurements; coalescence results when measurement estimates are drawn towards common measurements.
Therefore, coalescence of the JPDAF or MUJPDAF occurs inside the measurement space (e.g. if a sensor provides ranges and bearings, these are in the measurement space as opposed to a Cartesian state vector inside the state space).
Contrary to~\cite{KauLovLee14} where the similarity index is in terms of state space variables, the coalescence avoidance presented in this paper takes place inside the measurement space to most directly remove the coalescence.
Let the posterior measurement estimate be $\hat z_i^+\in\Re^m$ and its innovation be $e_{i}^+\in\Re^m$ that corresponds with its true measurement $z_i\in\Re^m$ (although association is unknown) as
\footnote{See the footnote on page~\pageref{fn:nonlinsys} for nonlinear systems.}
\begin{align}
\hat z_i^+&=H_i\hat x_i^+,
\\
e_{i}^+&=z_i-\hat z_i^+=(H_ix_i+v_i)-H_i\hat x_{i}^+=H_i\tilde x_i^++v_i,
\end{align}
where $\tilde x_i^+=x_i-\hat x_{i}^+$ denotes the error between the true state and the posterior estimate. The posterior innovation covariance $S^+_i\in\Re^{m\times m}$ is given by
\begin{align}
S^+_i&=\mathrm{E}[e_i^+e_i^{+T}]=\mathrm{E}[(H_i\tilde x_i^++v_i)(H_i\tilde x_i^++v_i)^T]\nonumber
\\
&=H_i\mathrm{E}[\tilde x_i^+\tilde x_i^{+T}]H_i^T+\mathrm{E}[vv^T]\nonumber%asdf
\\
&=H_iP^+_iH_i^T+R_i.
\end{align}
From these, the similarity cost is defined according to the format of \refeqn{zetaGeneral} as
%which are used to define $u_{p q}$ between $\hat z^+_{p}$ and $\hat z^+_{q}$:
\begin{align}
\label{eqn:Js}
J_S&=\sum\limits_{p,q\in I,p\neq q}\exp (-au_{pq}),
\end{align}
where $u_{pq}\in\Re$ is
\begin{align}
u_{pq} & = (\hat z_{p}^+-\hat z^+_{q})^T{S^+_{pq}}^{-1}(\hat z^+_{p}-\hat z^+_{q})\label{eqn:U},
\end{align}
where $S^+_{pq}=S^+_{p}+S^+_{q}$. Expanding \refeqn{U} with respect to the $p$-th system,
\begin{align}
u_{pq}=&\ 
(H_pK_p{\bf\bar e}_p-\hat z^+_{q})^T
(H_pP^+_{p}H_p^T+R_p+S^+_q)^{-1}  
(H_pK_p{\bf\bar e}_p-\hat z^+_{q}).
\end{align}

Next, we find the derivative of the similarity cost as follows.
Let $\Ibd\in \Re^{n\times m}$ be defined such that its $b,d$-th element is one and other elements are zero.
The partial derivative of $u_{pq}$ with respect to the scalar value at the $b,d$-th element of $K_p$, denoted as $K_{p,bd}$, is
\begin{align}
\deriv{u_{pq}}{K_{p,bd}}
=&\ 
(H_p\Ibd{\bf\bar e}_p)^T
{S_{pq}}^{-1}
(H_pK_p{\bf\bar e}_p-\hat z^+_{q})+
(H_pK_p{\bf\bar e}_p-\hat z^+_{q})^T
\mathcal{S}
(H_pK_p{\bf\bar e}_p-\hat z^+_{q})\nonumber
\\
&\ +
(H_pK_p{\bf\bar e}_p-\hat z^+_{q})^T
{S_{pq}}^{-1}
(H_p\Ibd{\bf\bar e}_p)\nonumber
\\
=&\ \{
2(H_p\Ibd{\bf\bar e}_p)^T
{S_{pq}}^{-1} +(H_pK_p{\bf\bar e}_p-\hat z^+_{q})^T
\mathcal{S}
\}(H_pK_p{\bf\bar e}_p-\hat z^+_{q}),\label{eqn:derivu}
\\
\deriv{u_{pq}}{K_{r,bd}}=&\ 0\quad \mbox{if } r\not\in\{p,q\},\label{eqn:derivu0}
\end{align}
where
\begin{align}
\mathcal{S}=&\ \deriv{{S^+_{pq}}^{-1}}{K_{p,bd}}\nonumber
\\
=&\ -{S^+_{pq}}^{-1}
\deriv{\left(H_pP^+_{p}H_p^T+R_p+S^+_q\right)}{K_{p,bd}}
{S^+_{pq}}^{-1}\nonumber
\\
=&\ -{S^+_{pq}}^{-1}
H_p
[(1-\beta_{0,p})(\Ibd H_pP_p^-+P_p^-H_p^T\Ibd^T)-\Ibd E_pK_p^T -K_pE_p\Ibd^T]
H_p^T
{S^+_{pq}}^{-1},\label{eqn:SS}
\end{align}
because
\begin{align}
&\deriv{\left(H_pP^+_{p}H_p^T+R_p+S^+_q\right)}{K_{p,bd}}
=
H_p
\deriv{P^+_{p}}{K_{p,bd}}
H_p^T\nonumber
\\
&=
H_p
[(1-\beta_{0,p})(\Ibd H_pP_p^-+P_p^-H_p^T\Ibd^T)-\Ibd E_pK_p^T -K_pE_p\Ibd^T]
H_p^T.\label{eqn:SSmidpart}
\end{align}
Then, \refeqn{SS} and \refeqn{SSmidpart} are substituted into \refeqn{derivu} or \refeqn{derivu0} to solve
\begin{align}
\label{eqn:JSK}
\deriv{J_S}{K_{p,bd}}=\ -a\sum\limits_{q\in I,q\neq p}\exp (-au_{pq})\deriv{u_{pq}}{K_{p,bd}}.
\end{align}

Necessary conditions for optimality are given by
\begin{align}
\deriv{\mathbf{J}}{K_i} = \deriv{J_P}{K_i} + c \deriv{J_S}{K_i} =0,\label{eqn:NCO}
\end{align}
where $\deriv{J_P}{K_i}$ and $\deriv{J_S}{K_i}$ may be nonzero.
In short, \refeqn{CostP} and \refeqn{JSK} are substituted into \refeqn{NCO} and repeated for all $p$ to obtain the set of optimal gains $\mathcal{K}=\{K_1,K_2,...,K_{n_t}\}$.
Then, the C-JPDAF is completed by updating the state estimate and covariance by using \refeqn{KalEst} and \refeqn{JPDAFCov}, respectively.
As the optimality condition is a nonlinear function of estimator gains, there is no closed-form solution.
Hence, we require a nonlinear equation solver.
When $c=0$, we can show that the optimal gain reduces to that of the MUJPDAF, which is used as an initial guess of numerical solutions.
Therefore, if the number of estimates exhibiting significant similarity is $n_t^*\leq n_t$ and $\mathcal C$ is the number of times \refeqn{NCO} is solved depending on the numerical accuracy, then the required iterations is $n_t^*(n_t^*-1)\mathcal C$, where some variables may be reused (e.g. $S^+_{pq}\equiv S^+_{qp}$) and often the effect of \refeqn{derivu} is negligible.

\section{Numerical Examples}

We consider satellites in the lower earth orbit (LEO) and in the geosynchronous orbit (GEO) to compare the performances of the conventional JPDAF, the MUJPDAF, and the C-JPDAF, using equations of motion based off the two-body problem.

Let $\vec r=\begin{bmatrix}x_1 & x_2 & x_3\end{bmatrix}^T$ be the position vector of the $i$-th satellite represented with respect to the Earth-Centered Inertial (ECI) frame.
The equation of motion corresponding to the two-body problem is given by
\begin{align}
\ddot{\vec r}&=-\frac{\mu}{r^3}\vec r=-\mu(\vec r^T\vec r)^{-\frac32}\vec r,
\end{align}
where $\mu=398,600\ \frac{km^3}{sec^2}$ is the gravitational constant of the earth if position is estimated in kilometers and time is measured in seconds~\cite{Val01}.

Let $\delta$ denote an infinitesimal change of a state variable.
The linearized equations of state variable $\vec x=\begin{bmatrix}\vec r^T, & \dot{\vec r}^T\end{bmatrix}^T\in\Re^6$ are given by
\begin{align}
\begin{bmatrix}
\delta\dot{\vec r} \\ \delta\ddot{\vec r}
\end{bmatrix}
%\delta\dot{\vec x}
=
\begin{bmatrix}
0_{3\times3} & I_{3\times3} \\
\mu\left[3\vec r({\vec r}^T\vec r)^{-\frac52}{\vec r}^T-({\vec r}^T\vec r)^{-\frac32}I_{3\times3}\right] & 0_{3\times3}
\end{bmatrix}
\begin{bmatrix}
\delta\vec r \\ \delta\dot{\vec r}
\end{bmatrix}
%\delta\vec x
.
\end{align}
%Defining the above square matrix as $A_{lin}$, the linearized state transition matrix over the time step $dt$ is defined as
%\begin{align}
%F=\expm{(A_{lin}dt)}
%\end{align}
%where $\expm$ denotes the matrix exponential function.

The process uncertainty may be subject to a variety of perturbations beyond two-body forces, and the process covariance takes the form $Q=\diag[\sigma_r^2I_{3\times3}, \sigma_{\dot r}^2I_{3\times3}]$, which varies depending on the case considered.
%=\diag[10^{-6}I_{3\times3}, 10^{-8}I_{3\times3}]$.

For the cases where the satellites are orbiting in LEO, the satellites are measured with a single phased-array radar at location $x_s$ on the equator
\begin{align}
x_s=
\begin{bmatrix}
R_E\cos\frac{2\pi t}T & R_E\sin\frac{2\pi t}T & 0 & -\frac{2\pi tR_E}T\sin\frac{2\pi t}T  & \frac{2\pi tR_E}T\cos\frac{2\pi t}T & 0
\end{bmatrix}^T,
\end{align}
where $R_E=6378$ km is the radius of the earth, time of the day $t$ starts at $0$ seconds, and the time period of a day of the day is $T=86164.090518$ seconds.
Here, the measurement space consists of the range $\rho$ between the sensor and the satellite (km), the range rate $\dot\rho$ (km/sec), and topocentric right ascension $\alpha$ and declination $\delta$ (rad)~\cite{Val01}.
The standard deviations $\sigma_\rho$, $\sigma_\alpha$, and $\sigma_\delta$ are based on the Mahe Ground Station uncertainty data~\cite{VerSauSco04} and $\sigma_{\dot\rho}$ from paper examples~\cite{KurAriAriEfe10}.
The measurement uncertainty is $R=\diag[\sigma_\rho^2, \sigma_{\dot\rho}^2, \sigma_\alpha^2, \sigma_\delta^2]=\diag[0.15^2, 0.005^2, (\frac{0.01\pi}{180})^2, (\frac{0.01\pi}{180})^2]$.
For GEO, angles-only measurements are available with higher accuracy; namely the topocentric right ascension and declination are chosen as $\sigma_{\alpha,GEO}=\sigma_{\delta,GEO}=10^{-2}\sigma_\alpha$.
For all cases, there is a $10\%$ chance that a measurement is either deleted or an extraneous measurement is added at each time step.
These cases are outlined in Table \ref{tab:CaseDef}.

Further descriptions of the cases are available below, and the resulting accuracy is measured with RMS error, the uncertainty is summarized with $\tr{J_P}$, and the computation time is measured as well, tabulated in Tables \ref{tab:A1}--\ref{tab:B2}.

\begin{table}
\begin{center}
\caption{Case Definitions} \label{tab:CaseDef}
\begin{threeparttable}[h]
\begin{tabularx}{.59\textwidth}
{
>{$}c<{$} |
>{$}c<{$}
}
\toprule
\multirow{2}{*}{Symbol} & \multirow{2}{*}{Explanation}\\
\\
\midrule
\multirow{1}{*}{A} &  \multirow{1}{*}{Comparison of JPDAF and MUJPDAF}
\\
\multirow{1}{*}{B} &  \multirow{1}{*}{Comparison of JPDAF, MUJPDAF, and C-JPDAF}
\\
\multirow{1}{*}{1} &  \multirow{1}{*}{Orbits in LEO}
\\
\multirow{1}{*}{2} &  \multirow{1}{*}{Orbits in GEO}
\\
\multirow{1}{*}{a} &  \multirow{1}{*}{$\sigma^2_{r,a} = \epsilon\sigma^2_{r}$}
\\
\multirow{1}{*}{b} &  \multirow{1}{*}{$\sigma^2_{\dot r,b} = \epsilon\sigma^2_{\dot r}$}
\\
\multirow{1}{*}{c} &  \multirow{1}{*}{$\sigma^2_{r,c} = \epsilon\sigma^2_{r}$, $\sigma^2_{\dot r,c} = \epsilon\sigma^2_{\dot r}$}
\\
\multirow{1}{*}{d} &  \multirow{1}{*}{$\sigma^2_{\alpha,d} = \epsilon\sigma^2_{\alpha}$, $\sigma^2_{\delta,d} = \epsilon\sigma^2_{\delta}$}
\\
\multirow{1}{*}{e} &  \multirow{1}{*}{$\sigma^2_{\mathbf{range},e} = \epsilon\sigma^2_{\mathbf{range}}$ (case number 1 only)}
\\
\multirow{1}{*}{f} &  \multirow{1}{*}{$\sigma^2_{\mathbf{range-rate},f} = \epsilon\sigma^2_{\mathbf{range-rate}}$ (case number 1 only)}
\\
\bottomrule
\end{tabularx}
\end{threeparttable}
\end{center}
\end{table}

% Alan's note:
%The scenarios below (A and B) are those from the ACC paper submission; the results to be included in this paper are according to those outlined in Table 1.
%The brief summary of these results to come is that the minimum uncertainty JPDAF (MUJPDAF) outperforms the conventional JPDAF in every case (LEO and GEO) in terms of reducing uncertainty and improving state estimate accuracy, and the coalescence-avoiding optimal JPDAF (C-JPDAF) often outperforms the MUJPDAF and JPDAF in both LEO and GEO scenarios in terms of state estimate accuracy only, but this is not always guaranteed.

\paragraph*{Scenario A1: Comparison of the JPDAF and MUJPDAF with two LEO satellites in close proximity}

The first set of cases compares the JPDAF with the MUJPDAF for tracking two nearby satellites passing over the sensor field of view in LEO. The initial conditions of the first two orbits are chosen as
\begin{align}
x_{1,A1}=x_{ref,A1}+x_{pert,A1},& \quad x_{2,A1}=x_{ref,A1}-x_{pert,A1},\nonumber
\\
x_{ref,A1}=\begin{bmatrix}7000 & 0 & 1000 & 0 & 7.546 & 0\end{bmatrix}^T,& \quad x_{pert,A1}=\begin{bmatrix}
0.2 & 0 & 0 & 0 & -0.105 & 0
\end{bmatrix}^T.
\end{align}
In these cases, the process covariance is $Q=\diag[10^{-6}I_{3\times3}, 10^{-8}I_{3\times3}]$, the estimates have initial covariances matrix $P_{0,A1}=10^7\times Q$, time steps of $1$ second over $100$ seconds, and $\epsilon=10$.
These scenarios are chosen because the estimates exhibit large initial uncertainties, so the spread in the innovations affects the updates.
The gain of the MUJPDAF accounts for this portion of the state covariance matrix better than the gain for the conventional JPDAF in every single case.
Given identical parameters and measurements, the RMS errors of the position are lower in \emph{every} case when the MUJPDAF is applied compared with when the conventional JPDAF is applied.
In fact in two cases, namely A1b and A1c, the JPDAF is incapable of decreasing the estimate uncertainty as quickly as the MUJPDAF; the JPDAF initially suffers from such great coalescence that the estimates become associated with incorrect measurements consistently.
Furthermore, the computational time is roughly the same between the JPDAF and MUJPDAF, so the results suggest that the improvement in accuracy is at no increase in computation.
These results are tabulated in Table \ref{tab:A1}, and some cases are plotted in Figure \ref{fig:A1}.

\begin{center}
\begin{threeparttable}[h]
\caption{Cases of A1} \label{tab:A1}
\begin{tabularx}{.58\textwidth}
{
>{$}c<{$} |
*{2}{>{$}c<{$}} |
*{2}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multicolumn{2}{c}{\multirow{1}{*}{RMS Position Error (km)}} & \multicolumn{2}{c}{\multirow{1}{*}{Time (sec)}} \\
 & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{MUJPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{MUJPDAF}
\\
\midrule
A1  & 0.12026 & 0.10646 & 0.582294 & 0.560754 \\
A1a & 0.17910 & 0.17157 & 0.558011 & 0.561458 \\
A1b & \hilight{2.63710} & 0.14150 & 0.561605 & 0.558537 \\
A1c & \hilight{2.63520} & 0.19559 & 0.570204 & 0.578704 \\
A1d & 0.20204 & 0.18865 & 0.572659 & 0.578029 \\
A1e & 0.14541 & 0.13361 & 0.575572 & 0.573995 \\
A1f & 0.12175 & 0.10795 & 0.576113 & 0.559978 \\
\bottomrule
\end{tabularx}
{\small
\begin{tablenotes}
    \item \hilight{Track swapping occurs}
  \end{tablenotes}}
\end{threeparttable}
\end{center}

\begin{figure}
{
%\hspace*{0.05\columnwidth}
\centerline{
	\subfigure[\;Case A1: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1JPDAF.pdf}}
	\subfigure[\;Case A1: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1MUJPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case A1b: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1bJPDAF.pdf}}
	\subfigure[\;Case A1b: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1bMUJPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case A1d: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1dJPDAF.pdf}}
	\subfigure[\;Case A1d: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1dMUJPDAF.pdf}}
	}
}
\caption{The orbital radii of three cases are shown where the dashed plots are the true trajectories and the solid plots are the estimated trajectories.
In all cases the MUJPDAF is more accurate and experiences less coalescence than the JPDAF.
In particular, when the JPDAF is used in to Case A1b, the initial coalescence causes track swapping.
Additionally, Case A1d experiences the most inaccurate estimation without track swapping with very noisy angle measurements.
}\label{fig:A1}
\end{figure}

\paragraph*{Scenario A2: Comparison of the JPDAF and MUJPDAF with two near-GEO satellites in close proximity}

The second set of cases compares the JPDAF with the MUJPDAF for tracking two nearby satellites passing over the sensor field of view in GEO, where the orbital radius is $r_{GEO}=4.2164\times10^{4}\ km$. The initial conditions of the first two orbits are chosen as
\begin{align}
x_{1,A2}&=\begin{bmatrix}r_{GEO} & 0 & 0 & 0 & \sqrt{\frac{\mu}{r_{GEO}}}, & 0\end{bmatrix}^T,
\\
x_{2,A2}&=x_{1,A2}+\begin{bmatrix}
0 & 0 & -1 & 0 & 0 & 4\times10^{-4}
\end{bmatrix}^T,
\end{align}
or where $x_{1,A2}$ remains exactly in GEO while $x_{2,A2}$ starts slightly perturbed from a GEO orbit with a nonzero inclination such that the two satellites nearly collide as they pass by in close proximity.
Unlike the last cases, the A2 cases consider small initial covariance matrices $P_{0,A2}= Q$ where $Q=\diag[10^{-6}I_{3\times3}, 10^{-8}I_{3\times3}]$, time steps of $10$ seconds over $5000$ seconds, and $\epsilon=1.1$.
%These scenarios are chosen because the estimates exhibit large initial uncertainties, so the spread in the innovations affects the updates.
Even though the initial conditions are largely different from those of the A1 cases, the A2 cases display the same trends, where the computational costs among the algorithms are roughly the same but the tracking abilities of the MUJPDAF \emph{always} yield estimates closer to the truth than those of the JPDAF.
These results are tabulated in Table \ref{tab:A2}, and some cases are plotted in Figure \ref{fig:A2}.

\begin{center}
\begin{threeparttable}[h]
\caption{Case A2} \label{tab:A2}
\begin{tabularx}{.65\textwidth}
{
>{$}c<{$} |
*{2}{>{$}c<{$}} |
*{2}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multicolumn{2}{c}{\multirow{1}{*}{RMS Angular Error (rad)}} & \multicolumn{2}{c}{\multirow{1}{*}{Time (sec)}} \\
 & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{MUJPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{MUJPDAF}
\\
\midrule
A2  & 1.2140\times10^{-6} & 9.8784\times10^{-7} & 2.765200 & 2.732202 \\
A2a & 1.2141\times10^{-6} & 9.8780\times10^{-7} & 2.739866 & 2.743265 \\
A2b & \hilight{1.1728$\times10^{-5}$} & 1.0143\times10^{-6} & 2.733589 & 2.742162 \\
A2c & \hilight{1.1728$\times10^{-5}$} & 1.0142\times10^{-6} & 2.732890 & 2.751087 \\
A2d & \hilight{1.1700$\times10^{-5}$} & 1.0332\times10^{-6} & 2.819630 & 2.773270 \\
\bottomrule
\end{tabularx}
{\small
\begin{tablenotes}
    \item \hilight{Track swapping occurs}
  \end{tablenotes}}
\end{threeparttable}
\end{center}

\begin{figure}
{
%\hspace*{0.05\columnwidth}
\centerline{
	\subfigure[\;Case A2: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2JPDAF.pdf}}
	\subfigure[\;Case A2: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2MUJPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case A2a: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2aJPDAF.pdf}}
	\subfigure[\;Case A2a: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2aMUJPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case A2b: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2bJPDAF.pdf}}
	\subfigure[\;Case A2b: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2bMUJPDAF.pdf}}
	}
}
\caption{The orbital radii of three cases are shown where the dashed plots are the true trajectories and the solid plots are the estimated trajectories and $\alpha_{diff}$ is the difference between $\alpha_{1}$ and $\hat\alpha_{1,A2}$ or $\hat\alpha_{2,A2}$.
Cases A2 and A2a serve as examples where the coalescence during crossing periods with the JPDAF yields much more similar estimates than with the MUJPDAF.
In Case A2b, this coalescence is so severe that track swapping occurs with the JPDAF, but this is avoided with the MUJPDAF.
}\label{fig:A2}
\end{figure}

\paragraph*{Scenario B1: Comparison of the JPDAF, MUJPDAF, and C-JPDAF with three LEO satellites in close proximity}
The JPDAF, MUJPDAF, and C-JPDAF algorithms are compared for three satellites in LEO with initial conditions
\begin{align}
\begin{split}
x_{1,B1}&=x_{ref,B1}+x_{pert,B1}, \quad x_{2,B1}=x_{ref,B1}-x_{pert,B1}, \quad x_{3,B1}=x_{ref,B1},
\\
x_{ref,B1}&=\begin{bmatrix}7000 & 0 & 0 & 0 & 7.546 & 0\end{bmatrix}^T, \quad x_{pert,B1}=\begin{bmatrix}
0.025 & 0 & 0 & 0 & 0 & 0
\end{bmatrix}^T,
\end{split}
\end{align}
where $Q=\diag[10^{-6}I_{3\times3}, 10^{-8}I_{3\times3}]$, $P_{0,B1}=Q$, and $\epsilon=1.1$ for time steps of $10.0$ seconds over a $2000$ second time period.
This corresponds to a phased-array radar observing observing a slow separation of objects in a near-circular orbit; because the satellite states are in close proximity over a large time period, the JPDAF and MUJPDAF are subject to severe coalescence.
The C-JPDAF is designed to avoid this coalescence, where the parameters $a=5\times10^{-1}$ and $c=2.3\times10^{-4}$ are used in these scenarios.
The accuracy of the C-JPDAF is largely improved from that of the JPDAF and MUJPDAF; however, the computational cost of the C-JPDAF is roughly $60$ times that of either other algorithm.
The position RMS error and computation time are tabulated in Table \ref{tab:B1} and some example plots from these cases are shown in Figure \ref{fig:B1}.

\begin{center}
\begin{threeparttable}[h]
\caption{Case B1} \label{tab:B1}
\begin{tabularx}{.81\textwidth}
{
>{$}c<{$} |
*{3}{>{$}c<{$}} |
*{3}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multicolumn{3}{c}{\multirow{1}{*}{RMS Position Error (km)}} & \multicolumn{3}{c}{\multirow{1}{*}{Time (sec)}} \\
 & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{MUJPDAF} & \multirow{1}{*}{C-JPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{MUJPDAF} & \multirow{1}{*}{C-JPDAF}
\\
\midrule
B1   	& 0.14432 & 0.091133 & 0.057312 & 1.736272 & 1.670539 & 101.000529 \\
B1a 	& 0.14432 & 0.091134 & 0.057307 & 1.689807 & 1.726980 & 100.988662 \\
B1b 	& 0.14694 & 0.094506 & 0.055145 & 1.673762 & 1.678786 & 100.855376 \\
B1c 	& 0.14693 & 0.094507 & 0.055136 & 1.675486 & 1.669083 & 100.907850 \\
B1d 	& 0.14400 & 0.088933 & 0.058911 & 1.663424 & 1.660136 & 102.606102 \\
B1e 	& 0.14408 & 0.090823 & 0.059919 & 1.673660 & 1.666829 & 100.278073 \\
B1f 	& 0.14436 & 0.091212 & 0.064299 & 1.732541 & 1.721413 & 98.672870 \\
\bottomrule
\end{tabularx}
\end{threeparttable}
\end{center}

\begin{figure}
{
%\hspace*{0.05\columnwidth}
\centerline{
	\subfigure[\;Case B1: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1JPDAF.pdf}}
	\subfigure[\;Case B1: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1MUJPDAF.pdf}}
	\subfigure[\;Case B1: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1C-JPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case B1c: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1cJPDAF.pdf}}
	\subfigure[\;Case B1c: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1cMUJPDAF.pdf}}
	\subfigure[\;Case B1c: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1cC-JPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case B1f: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1fJPDAF.pdf}}
	\subfigure[\;Case B1f: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1fMUJPDAF.pdf}}
	\subfigure[\;Case B1f: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1fC-JPDAF.pdf}}
	}
}
\caption{The orbital radii of three cases are shown where the dashed plots are the true trajectories and the solid plots are the estimated trajectories.
In all cases the MUJPDAF is more accurate and experiences less coalescence than the JPDAF, but both algorithms suffer from coalescence.
The C-JPDAF systematically removes this coalescence from the updates at the cost of slightly larger covariance matrices and a large increase in required computational resources.
%Between Cases B1 and B1c, the process position uncertainty is increased, which last little effect on the estimates since this has little
The conditions that affect the C-JPDAF most are with Case B1f where the range-rate measurements have the greatest uncertainty.
}\label{fig:B1}
\end{figure}

Among the variations of the B1 cases, the range-rate measurement uncertainty (Case B1f) has the largest effect on tracking with the C-JPDAF.
In the B1 cases, the ranges and bearings of each track are close and changing, but the difference between the velocities of the three satellites remain largely constant after about $500$ seconds.
As long as the estimates are able to stay sufficiently far apart, as the C-JPDAF is able to accomplish, the component of the velocity measurement in the radial direction serves as a distinguishing portion of the full measurement, so that the correct association between measurement and estimate is more heavily weighted, and hence better tracking is achieved.
The quality of the range-rate measurement is degraded in Case B1f, and hence the C-JPDAF algorithm is not able to as accurately track each estimate when compared with the other B1 cases.

\paragraph*{Scenario B2: Comparison of the JPDAF, MUJPDAF, and C-JPDAF with two GEO satellites in close proximity}
The JPDAF, MUJPDAF, and C-JPDAF algorithms are compared for two satellites in GEO with initial conditions
\begin{align}
\begin{split}
x_{1,B2}&=f_{GEO}(0^\circ), \quad x_{2,B2}=f_{GEO}({10^{-4}}^\circ), \quad \mbox{where}
\\
f_{GEO}(\theta)&=\begin{bmatrix}r_{GEO}\cos\theta & r_{GEO}\sin\theta & 0 & \sqrt{\frac{\mu}{r_{GEO}}}\cos(90^\circ+\theta) & \sqrt{\frac{\mu}{r_{GEO}}}\sin(90^\circ+\theta) & 0\end{bmatrix}^T,
\end{split}
\end{align}
where $Q=\diag[10^{-8}I_{3\times3}, 10^{-10}I_{3\times3}]$, $P_{0,B}=Q$, and $\epsilon=1.1$ for time steps of $10.0$ seconds over a $2000$ second time period.
This corresponds to a scenario where two satellite orbit steadily nearby in GEO.
For the C-JPDAF, the parameters $a=1.5\times10^{2}$ and $c=1\times10^{10}$ are used in these scenarios.
In theory, topocentric right ascension measurements should be the same distance apart throughout the cases and the declination should always be zero.
Much like the B1 cases, the accuracy of the C-JPDAF yields a large improvement from that of the JPDAF and MUJPDAF; however, the increase in computational cost of the C-JPDAF is roughly $30$ times that of either other algorithm, which is about half of the increase found with the B1 cases.
This is because the similarity measure of two estimates is less complex with the B2 cases than with three estimates in the B1 cases.
The angular RMS error and computation time are tabulated in Table \ref{tab:B2} and some example plots from these cases are shown in Figure \ref{fig:B2}.

\begin{center}
\begin{threeparttable}[h]
\caption{Case B2} \label{tab:B2}
\begin{tabularx}{.95\textwidth}
{
>{$}c<{$} |
*{3}{>{$}c<{$}} |
*{3}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multicolumn{3}{c}{\multirow{1}{*}{RMS Angular Error (rad)}} & \multicolumn{3}{c}{\multirow{1}{*}{Time (sec)}} \\
 & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{MUJPDAF} & \multirow{1}{*}{C-JPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{MUJPDAF} & \multirow{1}{*}{C-JPDAF}
\\
\midrule
B2   	& 1.1112\times10^{-6} & 9.4780\times10^{-7} & 5.6708\times10^{-7} & 1.149378 & 1.119439 & 34.318254 \\
B2a 	& 1.1112\times10^{-6} & 9.4779\times10^{-7} & 5.6347\times10^{-7} & 1.154256 & 1.150072 & 34.758609 \\
B2b 	& 1.1306\times10^{-6} & 9.8287\times10^{-7} & 5.7387\times10^{-7} & 1.126598 & 1.132463 & 35.327683 \\
B2c 	& 1.1306\times10^{-6} & 9.8286\times10^{-7} & 5.6435\times10^{-7} & 1.118436 & 1.131690 & 34.817179 \\
B2d 	& 1.1115\times10^{-6} & 9.5439\times10^{-7} & 5.6136\times10^{-7} & 1.113892 & 1.116399 & 35.213445 \\
\bottomrule
\end{tabularx}
\end{threeparttable}
\end{center}

\begin{figure}
{
%\hspace*{0.05\columnwidth}
\centerline{
	\subfigure[\;Case B2: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2JPDAF.pdf}}
	\subfigure[\;Case B2: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2MUJPDAF.pdf}}
	\subfigure[\;Case B2: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2C-JPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case B2b: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2bJPDAF.pdf}}
	\subfigure[\;Case B2b: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2bMUJPDAF.pdf}}
	\subfigure[\;Case B2b: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2bC-JPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case B2d: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2dJPDAF.pdf}}
	\subfigure[\;Case B2d: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2dMUJPDAF.pdf}}
	\subfigure[\;Case B2d: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2dC-JPDAF.pdf}}
	}
}
\caption{In the various cases, the C-JPDAF is always able to produce more distinguishable estimates.
The perturbations among the cases yield slightly different results, but none are drastically different.
}\label{fig:B2}
\end{figure}

The most important result is that the C-JPDAF consistently avoids track-swapping, whereas the JPDAF and MUJPDAF do not.
However unlike the B1 cases, the B2 cases do not have a measurement that is consistently different among the estimates except for topocentric right ascension.
However, this component of the measurement is consistently very small.
Therefore, the variations from Case B2 all yield very close results.

An interesting observation from the data is the nonzero estimated declination, which should be zero because the GEO orbit lies on the equatorial plane; any shift from zero is due to process and measurement uncertainties.
Rather than observing independent errors among the two estimates, the declination errors are correlated and coalescing together.
The correlation of the sets of measurement estimates is easily explained by the soft decisions applied by all of the compared algorithms; measurement updates are largely shared among the estimates, so uncertainties affect the two estimates in a similar way.
Coalescence occurs when measurement updates are shared, shown in all cases in this paper, but the declination errors are less similar with the C-JPDAF.

\section{Conclusion}

The conventional JPDAF is a well-known algorithm with various applications in data association.
The soft decision approach makes the algorithm robust to missed detections and other clutter.
This paper enhances the JPDAF in two different ways.
First, the conventional JPDAF employs a Kalman gain, which does not minimize the trace of the state covariance matrix because the JPDAF yields an important spread in innovations term.
A minimum uncertainty JPDAF (MUJPDAF) is derived with a new gain that minimizes the full posterior uncertainty, including the spread in innovations state covariance matrix term.
This gain is computationally comparable to the JPDAF, and shows superior performance in estimate accuracy when the spread in the innovations is nonzero.

Second, both the JPDAF and the MUJPDAF involve measurement sharing, which tends to coalesce estimates in close proximity.
From~\cite{KauLovLee14}, the C-JPDAF is generalized for any number of estimates and measurements and the similarity cost is found inside the measurement space where coalescence occurs.
When using the C-JPDAF, there must be a trade-off between uncertainty increase and coalescence removal and powerful computational resources are also required.

The results in this paper show how the JPDAF, MUJPDAF, and C-JPDAF perform in LEO and GEO scenarios with various changes in parameters.
Numerous simulations show that the MUJPDAF operates at a similar computational cost to the JPDAF with a consistent improvement in accuracy.
In some scenarios where coalescence of the estimates is expected to be severe, the C-JPDAF can better track objects at the cost of increased computation.

Future work includes determining a systematic method of choosing the parameters for the C-JPDAF and deriving a suboptimal solution for the C-JPDAF that does not involve the large computation of the optimal solution.

\begin{appendix}
\label{append}

\subsection{Appendix A: Invertibility of the new $E_i$} \label{NewPartOfEIsPSD}
Consider that object $i$ is not measured (probability $\beta_{0,i}$), then this portion of the innovation update $e_0=0_{m\times1}$ because no innovation exists where measurements are not available.
Then for all innovations in the set $J^*=\{0\}\cup J$, the Total Probability Theorem holds true for object $i$:
\begin{align}
\sum\limits_{j^*\in J^*}\beta_{ij^*}=1.
\end{align}
From~\cite[Eq. 1.4.16-(1-10)]{ShaRonThi2001}, the covariance spread in the means term of the mixture of $e$ is defined as% (note that ${\bf e}_{i}$ remains unchanged)
\begin{align}
\tilde P_e=\sum\limits_{j^*}\beta_{ij^*}(e_{ij^*}-{\bf\bar e}_{i})(e_{ij^*}-{\bf\bar e}_{i})^T,
\end{align}
which is positive semi-definite. Manipulating this equation,
\begin{align}
\tilde P_e=&\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}e_{ij^*}^T
-\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}{\bf\bar e}_{i}^T
-{\bf\bar e}_{i}\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}^T
+{\bf\bar e}_{i}{\bf\bar e}_{i}^T\nonumber%asdf
\\
=&\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}e_{ij^*}^T-{\bf\bar e}_{i}{\bf\bar e}_{i}^T\nonumber
\\
=&\sum\limits_{j}\beta_{ij}e_{ij}e_{ij}^T-{\bf\bar e}_{i}{\bf\bar e}_{i}^T\geq0,
\label{eqn:CovSpreadPSD}
\end{align}
because $e_{i0}\equiv0$. Therefore $E_i$ is invertible $\forall\ 0\leq\beta_{0,i}<1$.

\subsection{Appendix B: The singularity of $E$ is not an infinite limit}
Consider the singularity when $E_i=0$ when no association is made to a track, or equivalently when $\beta^*=(1-\beta_{0,i})=\sum\limits_{j}\beta_j=0$.
It is clear from \refeqn{KalEst} and \refeqn{CovExpanded} that the updates are independent of $K_i$ exactly at this singularity.
We show that $K_i$ is \emph{not} growing without bound nearby this singularity, shown by evaluating the limit at this singularity to ensure boundedness.
Note that if $\sum\limits_{j}\beta_j\rightarrow0\Rightarrow \beta_j\rightarrow0\ \forall\ j$ because of the bounds $\sum\limits_{j}\beta_j\geq\beta_j\geq0\ \forall \ j$.
For clarity, let $\tilde\beta_p=\sum\limits_{j,j\neq p}\beta_j/\sum\limits_{j}\beta_j$ and drop the subscript $i$ because this analysis focuses on a single track.

First, we factor $\beta^*$ from $\sum\limits_{j} \beta_{j}e_{j}e_{j}^T$,
\begin{align}
\sum\limits_{j} \beta_{j}e_{j}e_{j}^T&=(\beta_1e_1e_1^T+\beta_2e_2e_2^T+\cdots)\nonumber
%\\
%&=\beta^*
%(e_1e_1^T-\tilde\beta_1e_1e_1^T+
%e_2e_2^T-\tilde\beta_2e_2e_2^T+\cdots),\nonumber
\\
&=\beta^*(\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T).
\end{align}
Similarly, we factor $\beta^*$ from ${{\bf\bar{e}}}{\bf\bar{e}}^T$,
\begin{align}
{{\bf\bar{e}}}{\bf\bar{e}}^T&=\sum\limits_{j} \beta_{j}e_{j}\sum\limits_{j} \beta_{j}e_{j}^T\nonumber%asdf
\\
%&=
%(\beta_1e_1\sum\limits_{j} \beta_{j}e_{j}^T+\beta_2e_2\sum\limits_{j} \beta_{j}e_{j}^T+\cdots),\nonumber
%\\
%&=\beta^*(e_1\sum\limits_{j} \beta_{j}e_{j}^T-\tilde\beta_1e_1\sum\limits_{j} \beta_{j}e_{j}^T\nonumber
%\\
%&\ \ \ +e_2\sum\limits_{j} \beta_{j}e_{j}^T-\tilde\beta_2e_2\sum\limits_{j} \beta_{j}e_{j}^T+\cdots),\nonumber
%\\
&=\beta^*(\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T).
\end{align}
Now that the summations inside $E$ are evaluated, their effects on $K$ are analyzed,
\begin{align}
\lim_{\beta^* \to 0}K&=\lim_{\beta^* \to 0}\beta^*P^-H^T
(\beta^*[S+\sum\limits_{j}e_{j}e_{j}^T
-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T-\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T+\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T])^{-1}\nonumber
\\
&=P^-H^T\lim_{\beta^* \to 0}
(S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T-\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T+\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T)^{-1},
\label{eqn:OptGainNoSingularity}
\end{align}
which not an infinite limit because
\begin{align}
\lim_{\beta^* \to 0}\tilde\beta_p&=\lim_{\beta^* \to 0}\frac{...+\beta_{p-1}+0+\beta_{p+1}+...}{...+\beta_{p-1}+\beta_p+\beta_{p+1}+...}\ \forall\ p\in J\nonumber%asdf
\\
&=1-\lim_{\beta^* \to 0}\frac{\beta_p}{\beta^*},
\end{align}
and therefore the following inequalities hold:
\begin{align}
&0\leq\beta_p\leq\beta^*\leq1\Rightarrow0\leq\lim_{\beta^* \to 0}\frac{\beta_p}{\beta^*}\leq1\Rightarrow0\leq\lim_{\beta^* \to 0}\tilde\beta_p\leq1,
\end{align}
and therefore \refeqn{OptGainNoSingularity} may be solved
\begin{align}
\lim_{\beta^* \to 0}K
%&=
%P^-H^T
%\left(S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}{B_j}e_{j}e_{j}^T\right)^{-1}
%\\
&=P^-H^T
(S+\sum\limits_{j}(1-B_j)e_{j}e_{j}^T)^{-1},
\label{eqn:OptGainNoSingularitySolved}
\end{align}
for $0\leq B_j\leq1\ \forall\ j$ and it can be easily seen that the sum of the terms inside the parentheses are invertible.
% and we know that $S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}{B_j}e_{j}e_{j}^T$ is invertible because $S$ is positive definite and the remaining terms are a simplified form of \refeqn{CovSpreadPSD}.
From a computational standpoint, for every $\beta^*_i\approx0$, the effect of the measurement update should be negligible, so we may choose $K_i=0_{n\times m}$ near this singularity as it will have little effect on the filter performance; this typically occurs when the object is not measured.
Choosing this $K_i$ results in no difference between the a priori and a posteriori update results.

\subsection{Appendix C: The optimal gain yields an absolute minimum to the uncertainty cost}

From \refeqn{CovExpanded} and \refeqn{JpIntro} the uncertainty cost for each estimate is given by
\begin{align}
\label{eqn:JpAppendixC}
J_p=\tr{P^--(1-\beta_{0})(KHP^-+P^-H^TK^T)+KEK^T}.
\end{align}
It can be easily seen that $J_p$ is the trace of a quadratic, and its second derivative is only a function of
\begin{align}
\tr{KEK^T}=\sum\limits_{i=1}^nk_i^TEk_i
\end{align}
where $k_i$ corresponds to the $i$-th row of $K$.
Taking derivatives with respect to $k_i$,
\begin{align}
\deriv{(k_i^TEk_i)}{k_i}&=2Ek_i,
\\
\frac{\partial^2(k_i^TEk_i)}{\partial k_i^2}&=2E,
\end{align}
where every Hessian is positive definite.
Therefore the only stationary point when $K=(1-\beta_{0})P^-H^TE^{-1}$ is a minimum.
%%%%% Direct method:
%From \refeqn{CovExpanded} and \refeqn{JpIntro} the uncertainty cost for each estimate is given by
%\begin{align}
%\label{eqn:JpAppendixC}
%J_p(K)=\tr{P^--(1-\beta_{0})(KHP^-+P^-H^TK^T)+KEK^T}
%\end{align}
%where the uncertainty cost associated with the MUJPDAF gain $K_{MUJPDAF}=(1-\beta_{0})P^-H^TE^{-1}$ can be easily found using \refeqn{GainK} and \refeqn{PNotSimplified} as
%\begin{align}
%\label{eqn:JpOpt}
%J_{p}(K_{MUJPDAF})=\tr{P^--(1-\beta_{0})^2P^-H^TE^{-1}HP^-
%}.
%\end{align}
%Let $\epsilon\in\Re^{m\times n}$ be an arbitrary matrix such that at least one of its elements is nonzero.
%Consider $\tilde K\in\Re^{n\times m}$ is any gain matrix that is \emph{not} $K_{MUJPDAF}$, i.e. $\tilde K=K_{MUJPDAF}+\epsilon^T=(1-\beta_{0})(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)E^{-1}$.
%Then the posterior covariance from \refeqn{CovExpanded} is
%\begin{align}
%\label{eqn:CovSubOpt}
%P^+(\tilde K)=&\ P^--(1-\beta_{0})^2
%\left(
%(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)E^{-1}HP^-
%+P^-H^TE^{-1}(HP^-+\frac1{(1-\beta_{0})}E\epsilon)
%\right)\nonumber
%\\
%&+(1-\beta_{0})^2\left(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE\right)E^{-1}EE^{-1}\left(HP^-+\frac1{(1-\beta_{0})}E\epsilon\right)\nonumber
%\\
%=&\ P^-+(1-\beta_{0})^2
%\left(
%P^-H^TE^{-1}HP^--P^-H^TE^{-1}HP^--P^-H^TE^{-1}HP^-
%\right)\nonumber
%\\
%&+(1-\beta_{0})(\epsilon^THP^-+P^-H^T\epsilon-\epsilon^THP^--P^-H^T\epsilon)+\epsilon^TE\epsilon\nonumber
%\\
%=&\ P^--(1-\beta_{0})^2P^-H^TE^{-1}HP^-+\epsilon^TE\epsilon
%\end{align}
%
%If $J_{p}(\tilde K)-J_{p}(K_{MUJPDAF})>0\ \forall\ \epsilon$, then $K_{MUJPDAF}$ is the \emph{only} gain matrix that yields an absolute minimum of the cost function $J_{p}$.
%We solve for this inequality with \refeqn{JpOpt} and \refeqn{CovSubOpt},
%\begin{align}
%J_{p}(\tilde K)-J_{p}(K_{MUJPDAF})=&\ \tr{P^--(1-\beta_{0})^2P^-H^TE^{-1}HP^-+\epsilon^TE\epsilon}\nonumber
%\\
%&-\tr{P^--(1-\beta_{0})^2P^-H^TE^{-1}HP^-}\nonumber
%\\
%=&\ \tr{\epsilon^TE\epsilon}\nonumber
%\\
%=&\ \sum\limits_{i=1}^n\epsilon_i^TE\epsilon_i>0
%\end{align}
%because $E$ is a positive-definite symmetric matrix (see Appendices A and B) and where $\epsilon_i\in\Re^{m\times1}$ corresponds to the $i$-th column of $\epsilon$, where at least a single column contains at least a single nonzero element.
%%\begin{align}
%%J_{p}(\tilde K)&=J_{p}((1-\beta_{0})P^-H^TE^{-1}+\epsilon^T)\nonumber
%%\\
%%&=J_{p}((1-\beta_{0})(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)E^{-1}),\nonumber
%%\\
%%&=\tr{P^-}\nonumber
%%\\
%%&\ \ -(1-\beta_{0})^2\tr{(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)HP^-+P^-H^T(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)^T}\nonumber
%%\\
%%&\ \ +(1-\beta_{0})^2\tr{(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)E(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)^T}\nonumber
%%\end{align}

\subsection{Appendix D: The optimal gain and Kalman gain are not equivalent}
The Kalman filter gain $K_{KF}=P^-H^TS^{-1}$ from~\cite{TrackDataAssoc} and the MUJPDAF gain $K_{MUJPDAF}=(1-\beta_{0})P^-H^TE^{-1}$ from \refeqn{GainK} are not always equivalent with a proof by contradiction.
Consider that these gains are equivalent and multiplying both gains by $E$ from \refeqn{E} (object subscript $i$ is omitted for clarity),
\begin{align}
&K_{KF}E=K_{MUJPDAF}E=(1-\beta_{0})P^-H^T\nonumber
\\
&=P^-H^TS^{-1}\{(1-\beta_{0})S+{\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf\bar{e}}}{\bf\bar{e}}^T\}\nonumber
\\
&=(1-\beta_{0})P^-H^T+K_{KF}({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf\bar{e}}}{\bf\bar{e}}^T).\label{eqn:KalMUareSame}
\end{align}
Therefore if \refeqn{KalMUareSame} is true, then $K_{KF}({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf\bar{e}}}{\bf\bar{e}}^T)=0$, which
%\begin{align}
%K_{KF}({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T)=0.\label{eqn:ContradictionResult}
%\end{align}
%%which implies that the spread in the innovations $\tilde P\equiv0$, not assumed in the JPDAF.
%The result of \refeqn{ContradictionResult} 
may only occur if $K_{KF}=0$, the probabilities that each measurement $j$ belongs to each object $i$ are exactly $0$ or $1$, or $K_{KF}\perp({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf\bar{e}}}{\bf\bar{e}}^T)$.
None of these conditions are satisfied in general, so it follows that the Kalman gain is different in general from the optimal gain proposed in this paper.

\end{appendix}

% Old table:
%\begin{table}
%\caption{JPDAF, MUJPDAF, and C-JPDAF Comparisons}\label{tab:RMS}
%\centering
%\makebox[0pt][c]{\parbox{1\textwidth}{%
%    \begin{minipage}[b]{0.5\hsize}\centering
%\begin{tabular}{cccc}\hline
%Case & Method & RMS Err. (km) & $\bar{J_P}$\\\hline
%\vspace*{-0.02\columnwidth}\\
%A1 & JPDAF &  0.12024 & 0.51254\\
%A1 & MUJPDAF & 0.10641 & 0.49055\\\hline
%\vspace*{-0.02\columnwidth}\\
%A2 & JPDAF & 2.6353 & 5.7157\\
%A2 & MUJPDAF & 0.19554 & 4.8456\\\hline
%\end{tabular}
%	\end{minipage}
%\hfill
%    \begin{minipage}[b]{0.5\hsize}\centering
%\begin{tabular}{ccc}\hline
%Method & RMS Err. (km) & Time (sec)\\\hline
%\vspace*{-0.02\columnwidth}\\
%JPDAF & 0.14904 &  1.729\\
%MUJPDAF & 0.09935 & 1.725\\
%C-JPDAF & 0.046411 & 104.605\\\hline
%\end{tabular}
%    \end{minipage} 
%}}
%\end{table}

% Old figure:
%\begin{figure}
%{
%%\hspace*{0.05\columnwidth}
%\centerline{
%	\subfigure[\;Case A1 Orbital Radii]
%		{\hspace*{0.\columnwidth}\includegraphics[width=0.25\columnwidth]{CaseA1_rad.pdf}}
%	\subfigure[\;Case A1 Uncertainty]
%		{\hspace*{0.\columnwidth}\includegraphics[width=0.25\columnwidth]{CaseA1_Jp.pdf}}
%	\subfigure[\;Case A2 Orbital Radii]
%		{\hspace*{0.\columnwidth}\includegraphics[width=0.25\columnwidth]{CaseA2_rad.pdf}}
%	\subfigure[\;Case A2 Uncertainty]
%		{\hspace*{0.\columnwidth}\includegraphics[width=0.25\columnwidth]{CaseA2_Jp.pdf}}
%	}
%}
%\caption{The orbital radii and uncertainty costs are compared among the algorithms (JPDAF is dashed, MUJPDAF is solid).
%These examples show that the posterior uncertainty of the MUJPDAF decreases faster than that of the JPDAF.
%The MUJPDAF shares measurement updates less, thereby decreasing coalescence, shown in both Cases A1 and A2.
%In Case A2 with the JPDAF, the coalescence causes consistent incorrect associations among the objects, known as track swapping.
%}\label{fig:CompareJPDAFandOJPDAF}
%\end{figure}

\bibliographystyle{AAS_publication}   % Number the references.
\bibliography{BibMaster}   % Use references.bib to resolve the labels.

\end{document}

%Equations are centered with the equation number flush to the right. In the text, these equations should be referenced by name as Eq.~\eqref{eq:ab} or Equation~\eqref{eq:ab} (\emph{e.g}., not eq. 1, (1), or \emph{Equation 1}). To improve readability, scalar variable names such as $a$ and $b^{2}$ are usually italicized when appearing in text and equations.\footnote{A section on mathematical notation is provided in the sequel.}

%\subsubsection{Abbreviations.} 
%When abbreviations for units of measure are used, lower case without periods is preferred in most instances; \emph{e.g}. ft, yd, sec, ft/sec, \emph{etc}., but in. for inch.
%
%
%
%
%\begin{figure}[htb]
%	\centering\includegraphics[width=3.5in]{Figures/test}
%	\caption{Illustration Caption Goes Here}
%	\label{fig:xxx}
%\end{figure}
%
%\subsubsection{Figures.}   
%Illustrations are referenced by name and without formatting embellishments, such as Figure~\ref{fig:xxx}, Figure 2, \emph{etc}., or, Figures 3 and 4 (\emph{e.g}., not figure (1), Fig. 1, \underline{Figure 1}, \emph{Figure 1}, \emph{etc}.). Each illustration should have a caption unless it is a mere sketch. Single-phrase captions are usually in title case; they are bold 10-point serif font and centered below the figure as shown in Figure~\ref{fig:xxx}. An explanatory caption of several sentences is permissible. Ideally, every illustration should be legibly sized -- usually about one-half or one-quarter page -- and appear in the text just before it is called out or mentioned. Alternatively, it is also permissible to place all figures together at the end of the text as a separate appendix; however, these two conventions should not be mixed. All figures and callouts should remain clearly legible after reduction. All illustrations appear as black and white in the final printing, although colors are retained in the electronic (CD-ROM) version.

%\subsubsection{Graphic Formats.} 
%The highest quality formats are Encapsulated PostScript (EPS) and PDF vector-graphic formats. These formats are recommended for all illustrations, unless they create document files that are excessively large. Specifically, you should change the graphic format or compress the image resolution whenever an illustrated page takes more than two seconds to render onscreen, or, whenever the total manuscript file size starts to approach 5 Mb. Photographs, illustrations that use heavy toner or ink (such as bar graphs), and figures without text callouts, may be suitably displayed with picture formats such as BMP, GIF, JPEG, PNG, TIFF, \emph{etc}. Line drawings, plots, and callouts on illustrations, should not use picture formats that do not provide sharp reproduction. All graphical content must be embedded when creating a PDF document, especially any fonts used within the illustration. Note that the Windows Metafile Format (WMF) is sometimes problematic and should be avoided.

%\subsubsection{References and Citations.} 
%The citation of bibliographical endnote references is indicated in the text by superscripted Arabic numerals, preferably at the end of a sentence.\cite{doe2005, style1959}   If this citation causes confusion in mathematics, or if a superscript is inappropriate for other reasons, this may be alternately expressed as (Reference~\citenum{doe2005}) or (see References~\citenum{doe2005} and \citenum{style1959}), (\emph{e.g}., not [1], Ref. (1), \emph{etc}.). While there is no singly prescribed format for every bibliographical endnote, references should be consistent in form. Citations should be sufficient to allow the reader to precisely find the information being cited, and should include specific pages, editions, and printing numbers where necessary. URL citations are discouraged, especially when an archival source for the same information is available. If a URL citation is required, it should appear completely and as a footnote instead of a bibliographical reference.\footnote{\url{http://www.univelt.com/FAQ.html\#SUBMISSION}}  The citation of private communication is especially discouraged, but if required it should be cited as a footnote and include the date, professional affiliation, and location of the person cited.\footnote{Gangster, Maurice (1999), personal correspondence of March 21st. Sr. Consultant, Space Cowboy Associates, Inc., Colorado Springs, CO.}

%\begin{table}[htbp]
%	\fontsize{10}{10}\selectfont
%    \caption{A Caption Goes Here}
%   \label{tab:label}
%        \centering 
%   \begin{tabular}{c | r | r } % Column formatting, 
%      \hline 
%      Animal    & Description & Price (\$)\\
%      \hline 
%      Gnat      & per gram & 13.65 \\
%                & each     &  0.01 \\
%      Gnu       & stuffed  & 92.50 \\
%      Emu       & stuffed  & 33.33 \\
%      Armadillo & frozen   &  8.99 \\
%      \hline
%   \end{tabular}
%\end{table}

%\emph{Tables.} 
%Tables are referred to by name in the text as Table~\ref{tab:label}, or, Tables 2 and 3 (\emph{e.g}., not table 1, Tbl. 1, or \emph{Table 1}). The title is centered above the table, as shown in Table 1. The font size inside tables should be no larger than the body text, but may be adjusted down to 9-point if necessary (10-point serif font is considered nominal). Note that table units are in parentheses. Only the minimum number of table lines needed for clarity is desired. Ideally, every table should appear within the text just before it is called out, but, it is also permissible to place all tables together at the end of the text as a separate appendix. If so, these two conventions should not be mixed.
%
%Equations, figures, and tables must be sequentially numbered with no repeated numbers or gaps. Each figure and table shall be called out in the text; gratuitous figures and tables that are not called out should be eliminated. Intermediate equations may be numbered without being called out.

%Space-based surveillance can be highly effective at estimating foreign spacecraft or objects. Angles-only orbit determination is a visual sensing technique for relative orbit determination passively; i.e. light from a deputy satellite is captured by the chief, unlike a radar system wherein waves must be emitted from an external source.

### 2018-06-09 Conference/AAS 2015 DA/ACC 2015/ACC15.tex
added Conference folder

\documentclass[letterpaper, 10pt, conference]{ieeeconf}
\IEEEoverridecommandlockouts \overrideIEEEmargins
\usepackage{amsmath,amssymb,url}
\usepackage{graphicx,subfigure,tikz,graphicx}
\usepackage{color}
\usepackage{csquotes}

\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\expm}{\ensuremath{\mathrm{expm}}}
\newcommand{\diag}{\mathop{\mathrm{diag}}\nolimits}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\ave}[1]{\mathrm{E}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand\circled[1]{%
  \tikz[baseline=(C.base)]\node[draw,circle,inner sep=0.5pt](C) {#1};\!
}

\title{\LARGE \bf
%Systematic Reduction of Uncertainty and Coalescence from JPDA Filtering
%Minimum Uncertainty JPDA Filtering and Advancements on Avoiding Coalescence
Minimum Uncertainty JPDA Filter and Coalescence Avoidance}
\author{Evan Kaufman, Thomas Alan Lovell, and Taeyoung Lee$^*$
 \thanks{Evan Kaufman and Taeyoung Lee are with Department of Aerospace Engineering, George Washington University, Washington, DC. Email: {\tt\footnotesize \{evankaufman, tylee\}@gwu.edu }}
\thanks{Thomas Alan Lovell is a Research Aerospace Engineer at the Space Vehicles Directorate, Air Force Research Lab (AFRL), Albuquerque, NM. Email: {\tt\footnotesize afrl.rvsv@kirtland.af.mil }}
\thanks{$^*$This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}}

%\newcommand{\EditTL}[1]{{\color{blue}\protect #1}}
%\renewcommand{\EditTL}[1]{{\protect #1}}

\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{remark}{Remark}

%\usepackage{ulem}
\begin{document}
\allowdisplaybreaks

\maketitle \thispagestyle{empty} \pagestyle{empty}

\begin{abstract}
This paper deals with an estimation problem where a known number of objects in close proximity are observed but the measurement origins are uncertain.
It is well known that the joint probabilistic data association filter (JPDAF) is effective in handling uncertain measurement originations with clutter, but the algorithm suffers it two ways addressed in this paper: first, the JPDAF directly adopts the gain of Kalman filters that does not minimize the posterior uncertainty of the JPDAF.
In this paper, an analytic solution for an optimal gain is derived that minimizes this uncertainty, and the corresponding filter is referred to as the minimum uncertainty JPDAF (MUJPDAF).
Second, the JPDAF is well-known to cause coalescence of estimates when objects are in close proximity.
This paper introduces a new variation of JPDAF, referred to as the coalescence-avoiding optimal JPDAF (C-JPDAF), which removes coalescence by minimizing a weighted sum of the posterior uncertainty and a measure of similarity between estimated probability densities.
This approach takes a simpler structure than other coalescence avoiding approaches based on pruning, while exhibiting excellent filtering performance for objects in close proximity with close or crossing tracks. Both algorithms are illustrated by numerical examples of satellites in close proximity orbiting around the Earth.
\end{abstract}

%_______________________________________________%

\section{Introduction}

Data association has been extensively studied with various applications in estimation when measurements are not necessarily originating from a single object.
We consider applications wherein multiple objects of interest enter the field of view of a sensor; however, the measurement origins are unknown.
To properly track these objects, we require estimation techniques that associate objects with measurements.
When the objects are in very close proximity, the association uncertainties grow, and an effective algorithm is necessary to maintain close estimations of the objects~\cite{KauLovLee14}.

A variety of data association techniques are considered for close-proximity trajectories with varying performance.
Multiple hypothesis tracking (MHT) and probability hypothesis density (PHD) techniques involve a high computational load, especially when objects are in close proximity, making real-time implementation more difficult~\cite{MHT1},~\cite{PHD1},~\cite{PHD2}.
Many other techniques are based on the Kalman filter (KF) or extended KF (EKF), which provide a framework for recursive data association techniques for easier real-time implementation.
There are two popular strategies among recursive approaches, namely \emph{hard decisions} and \emph{soft decisions}~\cite{JPDAF1}.
A hard decision is when estimates are fully associated with measurements according to some metric between the estimates and the measurements.
For example, a common heuristic approach is known as the nearest neighbor filter (NNF), which uses the Mahalanobis distance between each measurement and the predicted measurements of each potential object that may be associated with that measurement.
The measurement with the smallest Mahalanobis distance is considered as the correct one~\cite{NN2}.
However, when estimates are in close proximity, there is a high likelihood of making incorrect associations, which may be detrimental to the estimated states, so this technique is avoided in the analysis of this paper.

Alternatively with a soft decision technique such as the JPDAF, measurements are shared among the estimates. With the JPDAF, the likelihoods of association between measurements and estimates serve as weighting factors inside the measurement update portion of the filtering~\cite{JPDAF1},~\cite{TrackDataAssoc}.
The JPDAF is particularly advantageous in multi-object scenarios because when objects share measurements, the soft decisions prevent loss of information, which occurs with hard decision approaches.
%The a posteriori uncertainties of the estimates are changed because of an additional term representing the spread in the innovations; the Kalman gain serves as a close estimate to an optimal gain in the sense that in nearly minimizes this a posteriori gain as derived by the linearized equations of the JPDAF.
The JPDAF is based on the Kalman gain, which minimizes the posterior uncertainty of individual tracks.
However, the Kalman gain is not guaranteed to \emph{exactly} minimize the posterior uncertainties of the JPDAF.
%The unnecessarily large uncertainty of the suboptimal Kalman gain serves as motivation to derive an optimal gain that minimizes the a posteriori uncertainty of the linearized JPDAF equations.
In this paper, we derive the optimal estimation gain that minimizes the magnitude of the posterior covariance matrix, and the resulting minimum uncertainty JPDAF is referred to as the MUJPDAF.

The second part of this paper deals with the cases when neighboring objects share measurements, where the estimates are subject to coalescence.
Most techniques to avoid track coalescence involve pruning the tracks, or designing a set of rules that limit the possible outcomes to likely scenarios. These rules typically involve exact nearest neighbor probabilistic data association (ENNPDA), which applies a strong decision to the most likely associated estimate. Pruning algorithms incorporate the ENNPDA algorithm because  estimates do not share measurements, making it insensitive to track coalescence~\cite{Coal1}.
For example in~\cite{Fitzgerald}, the algorithm involves a set of rules to prune the tracks that share measurements.
The algorithm considers the possible combinations of estimate-to-measurement associations over a time window and applies a logic that assigns estimates to likely scenarios; a set of possible scenarios are considered and lower likelihood scenarios are deleted~\cite{Coal_d},~\cite{Coal_e},~\cite{Coal_c}.
However, track pruning assumes that low likelihood association scenarios are incorrect, which is not necessarily true.
Furthermore, removing the weighting scheme of the JPDAF causes the estimates to lose information.

Therefore, we propose a new coalescence-avoiding optimal JPDAF referred to as the C-JPDAF, which is a systematic approach to remove coalescence by minimizing the weighted sum of the uncertainty and a similarity index between estimates~\cite{KauLovLee14}.
The prior research covers the concept in simple cases, such that only two estimates are considered with a single measurement at each time step, making the algorithm infeasible in many realistic applications.
Therefore, we generalize the algorithm to handle any number of objects and measurements, as well as missed detections and measurements originating from extraneous clutter.
%Therefore we mature the algorithm to have the same assumptions as the JPDAF.
%Most importantly, there may be a variable number of measurements for a known number of estimates such that missed detections and false alarms are no longer an issue.

The paper is organized as follows: the data association filtering problem is formally presented in Section \ref{ProbDef}.
The posterior uncertainty according to the JPDAF equations is minimized to derive the MUJPDAF, which is compared with the conventional JPDAF equations in Section \ref{MUJPDAF}.
An optimal gain that minimizes a weighted cost of uncertainty and similarity in a general sense is derived in Section \ref{C-JPDAF}.
Numerical examples show how each of these algorithms can be advantageous and how they compare to each other in Section \ref{NumRes}, followed by conclusions.

\section{Data Association Filtering Problem Formulation}
\label{ProbDef}

 This section describes the dynamics of objects and the measurement model considered in this paper and a filtering problem is formulated.

\subsection{Dynamics of Objects and Measurements}

%
% Linearized equations of motion: assumed known
% Assume the Gaussian noise model and define unknown random variables $w$ (process noise) and $v$ (measurement noise)
% Initial conditions are assumed known, define $\hat x_{i,0}$, $P^+_{i,0}$

%Define the set of all objects $I=\{1,2,...,t\}$ such that the object index $i\in I$ where the total number of objects $t$ is assumed known.
Consider $n_t$ objects indexed by $i\in I=\{1,2,...,n_t\}$, where $n_t$ is assumed known.
The dynamics of the $i$-th object are modeled as a linear stochastic discrete-time system described by
\begin{align}
x_{i_{k}} & = F_{i_{k-1}} x_{i_{k-1}} + w_{i_{k}},\label{eqn:xkp}\\
z_{i_k} & = H_{i_k} x_{i_k} + v_{i_k},
\end{align}
where $x_{i_k}\in\Re^n$ is the state vector, $z_{i_k}\in\Re^m$ is the output vector, and  the matrices $F_{i_{k-1}}\in\Re^{n\times n}$ and $H_{i_k}\in\Re^{m\times n}$ describe system dynamics and outputs.
The process noise and the measurement noise are denoted by $w_{i_k}\in\Re^n$ and $v_{i_k}\in\Re^m$, respectively, where they are assumed to be zero-mean, mutually independent white, Gaussian noise sequences with known covariance matrices $Q_{i_k}\in\Re^{n\times n}$ and $R_{i_k}\in\Re^{m\times m}$, respectively, i.e.,
\begin{align}
w_{i_k} \sim \mathcal{N}[0,Q_{i_k}],\quad
v_{i_k} \sim \mathcal{N}[0,R_{i_k}],
\end{align}
where $\mathcal{N}[\mu,\sigma]$ denotes the Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.

The initial states are also assumed to be mutually independent with noise. They are Gaussian with known mean values and covariance matrices:
\begin{align}
x_{i_0} \sim \mathcal{N}[\bar x_{i_0}, P_{i_0}],\label{eqn:xi0}
\end{align}
where $\bar x_{i_0}\in\Re^n$ and $P_{i_0}\in\Re^{n\times n}$.

%To demonstrate the main idea of the proposed data association filter concisely, it is assumed that there are two objects, i.e., $i\in\{1,2\}$. But, the results of this paper are readily generalized to arbitrary number of systems provided the number of objects is known.

\subsection{Measurement Origination Model}
\label{DAP}

% Likelihoods of association are calculated from the innovations and innovation covariances only
% Two techniques are used to obtain the likelihoods of association, both of which may be applied to the JPDAF, MUJPDAF, and C-JPDAF
	% "Cheap JPDAF": inexpensive calculation, present the four equations and two references
	% True likelihood of association: infeasible for many real-time applications, calculated in the appendix

%Define the set $J=\{1,2,...,r\}$ such that the measurement index $j\in J$ where $r$ is a variable number of measurement returns such that $r\geq1$.
Suppose that there are $n_r$ measurements indexed by $j\in J=\{1,2,...,n_r\}$.
Let $\beta_{ij}$ be the probability that the $i$-th system is associated with measurement $j$.
Two techniques may be applied to solve for every $\beta_{ij}$, both of which require the innovation and its covariance (defined later) and information on detection probabilities, which are assumed known.
The first technique is applied in the standard JPDAF that calculates the true association probabilities from~\cite[Eq. 9.(45-46)]{TrackDataAssoc}.
However, the computation required with a standard JPDAF is inconsistent with the capabilities of the computational resources to account for an arbitrary number of objects to be tracked~\cite{Bar1990}.

A modified version of the JPDA, commonly known as the ``Cheap JPDA'', provides a simpler approximation of the likelihood $\beta_{ij}$.
The equations below are adopted from~\cite[Eq. 1.(1-2)]{Bar1990}.
Let $G_{ij}$ be a variable which takes a very similar form as the Gaussian probability density function
\begin{align}
G_{ij}=\frac1{\sqrt{\det{S_{i}}}}\exp{\{-\frac12e_{ij}^TS_{i}^{-1}e_{ij}\}},
\end{align}
where $e_{ij}$ is the innovation and $S_{i}$ is the a priori innovation covariance, defined later in \refeqn{eij} and \refeqn{S}, respectively.
This is used to calculate intermediate variables $\mathcal{S}_{ti}$ and $\mathcal{S}_{ri}$
\begin{align}
\mathcal{S}_{ti}=\sum\limits_{j\in J}G_{ij}, 
\quad 
\mathcal{S}_{ri}=\sum\limits_{i\in I}G_{ij}.
\end{align}
The probability that the $i$-th system is associated with the $j$-th measurement is
\begin{align}
\beta_{ij}=\frac{G_{ij}}{\mathcal{S}_{ti}+\mathcal{S}_{ri}-G_{ij}+B}, 
\end{align}
where $B$ is a correction factor to account for a non-unity probability of detection.

\section{Minimum Uncertainty Probabilistic Data Association Filter (MUJPDAF)}
\label{MUJPDAF}

In this section we derive the MUJPDAF that minimizes the posterior estimate uncertainties.
Differences between the JPDAF and MUJPDAF occur in the measurement update where the gain is derived, which changes the posterior covariance equation.

\subsection{Flow Update}

During the flow update, the a priori (denoted with superscript $-$) estimated state $\hat x_{i_k}^-$ is found with \refeqn{xkp} and its a priori covariance $P_{i_k}^-$ is calculated from the last posterior (denoted with superscript $+$) covariance $P_{i_{k-1}}^+$ for all $i$ as
\begin{gather}
%\hat x_{i_{k}}^- = f(\hat x_{i_{k-1}}^+),\\
P_{i_{k}}^- = F_{i_{k-1}} P_{i_{k-1}}^+ F_{i_{k-1}}^T + Q_{i_{k}}.
\end{gather}
These equations are applied between measurements to provide predictions of the states and their uncertainties.
% Equations for a priori state estimates and covariances

\subsection{Measurement Update}

The measurements serve to update the state estimates. Because all equations in this section take place at the $k$ time step, this subscript is removed from the equations. For a given measurement $z_j$, its measurement residual or innovation for the $i$-th object is given by
\begin{align}
e_{ij} = z_j - \hat z_i,\label{eqn:eij}
\end{align}
where $\hat z_i$ is the predicted measurement from $\hat x_{i}^-$ and a known measurement matrix $H_i$ such that
\begin{align}
\hat z_i = H_i\hat x_{i}^-.
\end{align}
%where the true measurement is expected to be
%\begin{align}
%z_i=h(x_i)+v_i, \quad H_i=\frac{\partial h}{\partial x}\bigg|_{x=\hat x_{i}^- }.
%\end{align}
Given the uncertainty of the measurements represented with covariance $R_i$, the a priori innovation covariance $S_{i}$ is
\begin{align}
S_{i}=H_{i}P_{i}^{-}H_{i}^T+R_{i}.\label{eqn:S}
\end{align}

The weighted sum of all innovations associated with the object $i$ is defined as ${\bf{e}}_{i}={\sum\limits_{j\in J} \beta_{ij}e_{ij}}$.
The resulting posterior estimate of the state is given by
%~\cite[Eq. 6-(25-26)]{TrackDataAssoc}
\begin{align}
\hat x^+_{i}=&\ \hat x^-_{i}+K_{i}{\bf{e}}_{i}.\label{eqn:KalEst}
\end{align}
The conventional JPDAF defines $K_i$ from the gain of the Kalman filter, given by $K_i=P^-_iH_i^TS_i^{-1}$, based on an incorrect assumption that the Kalman gain is always optimal for data association with the JPDAF.
%%____________________________________BEGIN NEW SECTION __________________________________%
%
%From \refeqn{eij} and \refeqn{KalEst}, $\hat x^+_{i}$ may be rewritten as
%\begin{align}
%\hat x_i^+&=\hat x^-+K_i\sum\limits_{j\in J}\beta_{ij}(z_j-H_i\hat x_i^-),\nonumber
%\\
%&=\left(I-(1-\beta_{0,i})K_iH_i\right)\hat x_i^-+K_i\sum\limits_{j\in J}\beta_{ij}z_j.\label{eqn:xhatpDEF}
%\end{align}
%Because $z_i=H_ix_i+v$, the posterior state error is
%\begin{align}
%\tilde x_i^+&=\hat x_i^+-x_i,\label{eqn:xperr}
%\\
%&=\left(I-(1-\beta_{0,i})K_iH_i\right)\hat x^-+K_i\sum\limits_{j\in J}\beta_{ij}z_j-
%\end{align}
%
%The definition of the posterior covariance matrix $P^+_i$ is
%\begin{align}
%P^+_i=\mathrm{E}[\tilde x^+\tilde x^{+T}],
%\end{align}
%
%In this paper, we find $K_i$ to minimize the posterior uncertainty of the state estimate.
%The optimal gain $K_i$ is derived below in the sense that the posterior uncertainties are minimized with respect to the generalized equations of the JPDAF.
%%Of course, the estimator cannot be optimal if the system is nonlinear, because the linearized equations to not fully represent the dynamics of a nonlinear system.
%
%Let the probability that object $i$ is not associated with any measurement be $\beta_{0,i}=1-{\sum\limits_{j} \beta_{ij}}$.
%The resulting posterior covariance can be written as
%%~\cite[Eq. D-32 and D-33]{TrackDataAssoc}
%\begin{align}
%\label{eqn:JPDAFCov}
%P^+_{i}=&\ \beta_{0,i}P^-_{i}+(1-\beta_{0,i})P_i^c+\tilde P,
%\end{align}
%where $\tilde P\in\Re^{n\times n}$ is
%\begin{align}
%\tilde P=&\ K_i\left({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_{i}^T,\label{eqn:tildeP}
%\end{align}
%and $P_i^c$ corresponds to the posterior covariance of object $i$ if only a single measurement serves to update this track~\cite{TrackDataAssoc}, and it can be written as
%%We can define $P_i^c$ without any assumptions on $K_i$ (i.e. $K_i$ is not necessarily the Kalman gain, so the well-known Kalman covariance update equations are not valid here) as~\cite[Eq. 4.2-12]{OptEst1}
%\begin{align}
%P^c_i=&\ (I-K_iH_i)P^-_i(I-K_iH_i)^T+K_iR_iK_i^T,\nonumber
%\\
%=&\ P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\nonumber
%\\
%&\ +K_i(H_iP^-_iH_i^T+R_i)K_i^T,
%\label{eqn:GenCov}
%\end{align}
%(see~\cite{OptEst1}).
%Substituting \refeqn{tildeP} and \refeqn{GenCov} into \refeqn{JPDAFCov} and rearranging,
%\begin{align}
%P^+_{i}=&\ \beta_{0,i}P^-_{i}+(1-\beta_{0,i})[P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\nonumber
%\\
%&\ +K_i(H_iP^-_iH_i^T+R_i)K_i^T]\nonumber
%\\
%&\ +K_i\left({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_{i}^T,\nonumber
%\\
%=&\ P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\nonumber
%\\
%&\ +K_i\left((1-\beta_{0,i})S_i+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_i^T,\nonumber
%\\
%=&\ P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\nonumber
%\\
%&\ +K_iE_iK_i^T,\label{eqn:CovExpanded}
%\end{align}
%where the symmetric matrix $E_i=E_i^T\in\Re^{m\times m}$, defined as
%\begin{align}
%E_i=&\ (1-\beta_{0,i})S_i+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T.
%\end{align}
%Note that $E_i$ is independent of $K_i$ and is invertible when $\beta_{0,i}\neq1$ since $S_i$ is positive-definite and
%%may be calculated without knowledge of the gain matrix $K_i$, and is invertible because $P^-_i$ and $R_i$ are positive-definite and
%$(\beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}{\bf{e}}_{i}^T)$ is positive-semidefinite, as shown in the Appendix A. 
%The case when $\beta_{0,i}=1$ corresponds with the $i$-th object \emph{not} being measured with full certainty.
%This case is analyzed in Appendix B.
%
%As with the Kalman filter, we choose the estimation gain to minimize the magnitude of the posterior covariance measured by $\tr{P_i^+}$.
%It requires that
%\begin{align}
%0&=\frac{\partial\tr{P^+_i}}{\partial K_i},\nonumber
%\\
%&=-2(1-\beta_{0,i})\frac{\partial\tr{K_iH_iP^-_i}}{\partial K_i}+\frac{\partial\tr{K_iE_iK_i^T}}{\partial K_i},\nonumber
%\\
%&=-2(1-\beta_{0,i})P^-_iH_i^T+2K_iE_i.\label{eqn:Uncertainty0}
%\end{align}
%Therefore the optimal gain is given by
%\begin{align}
%K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
%\end{align}
%
%Finally, \refeqn{GainK} is substituted into \refeqn{CovExpanded} to obtain
%\begin{align}
%%\begin{split}
%%P^+_{i}=&P^-_{i}-(1-\beta_{0,i})^2(P^-_iH_i^TE_i^{-1}H_iP^-_i
%%\\
%%&+P^-_iH_i^TE_i^{-1}H_iP^-_i
%%-P^-_iH_i^TE_i^{-1}E_iE_i^{-1}H_iP^-_i)
%%\end{split}
%%\\
%P^+_{i}=&P^-_{i}-(1-\beta_{0,i})^2P^-_iH_i^TE_i^{-1}H_iP^-_i,
%\end{align}
%which may be simplified in two ways
%\begin{align}
%P^+_{i}=&P^-_{i}-K_iE_iK_i^T, \quad \mbox{or equivalently}\label{eqn:JPDAFPostCovSymmetricUpdate}
%\\
%P^+_{i}=&\left(I-(1-\beta_{0,i})K_iH_i\right)P^-_i.\label{eqn:JPDAFPostCovSimpleUpdate}
%\end{align}
%The proposed gain is compared with the Kalman gain explicitly in Appendix C.
%In short, the differences between the MUJPDAF and the conventional JPDAF are shown in \refeqn{GainK} and \refeqn{JPDAFPostCovSymmetricUpdate} or \refeqn{JPDAFPostCovSimpleUpdate}.
The optimal gain $K_i$ is derived below in the sense that the posterior uncertainties are minimized with respect to the generalized equations of the JPDAF.
%Of course, the estimator cannot be optimal if the system is nonlinear, because the linearized equations to not fully represent the dynamics of a nonlinear system.

Let the probability that object $i$ is not associated with any measurement be $\beta_{0,i}=1-{\sum\limits_{j} \beta_{ij}}$.
The resulting posterior covariance can be written as
%~\cite[Eq. D-32 and D-33]{TrackDataAssoc}
\begin{align}
\label{eqn:JPDAFCov}
P^+_{i}=&\ \beta_{0,i}P^-_{i}+(1-\beta_{0,i})P_i^c+\tilde P,
\end{align}
where $\tilde P\in\Re^{n\times n}$ is
\begin{align}
\tilde P=&\ K_i\left({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_{i}^T,\label{eqn:tildeP}
\end{align}
and $P_i^c$ corresponds to the posterior covariance of object $i$ if only a single measurement serves to update this track~\cite{TrackDataAssoc}, and it can be written as
%We can define $P_i^c$ without any assumptions on $K_i$ (i.e. $K_i$ is not necessarily the Kalman gain, so the well-known Kalman covariance update equations are not valid here) as~\cite[Eq. 4.2-12]{OptEst1}
\begin{align}
P^c_i=&\ (I-K_iH_i)P^-_i(I-K_iH_i)^T+K_iR_iK_i^T,\nonumber
\\
=&\ P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\nonumber
\\
&\ +K_i(H_iP^-_iH_i^T+R_i)K_i^T,
\label{eqn:GenCov}
\end{align}
(see~\cite{OptEst1}).
Substituting \refeqn{tildeP} and \refeqn{GenCov} into \refeqn{JPDAFCov} and rearranging,
\begin{align}
P^+_{i}=&\ \beta_{0,i}P^-_{i}+(1-\beta_{0,i})[P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\nonumber
\\
&\ +K_i(H_iP^-_iH_i^T+R_i)K_i^T]\nonumber
\\
&\ +K_i\left({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_{i}^T,\nonumber
\\
=&\ P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\nonumber
\\
&\ +K_i\left((1-\beta_{0,i})S_i+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_i^T,\nonumber
\\
=&\ P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\nonumber
\\
&\ +K_iE_iK_i^T,\label{eqn:CovExpanded}
\end{align}
where the symmetric matrix $E_i=E_i^T\in\Re^{m\times m}$, defined as
\begin{align}
E_i=&\ (1-\beta_{0,i})S_i+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T.
\end{align}
Note that $E_i$ is independent of $K_i$ and is invertible when $\beta_{0,i}\neq1$ since $S_i$ is positive-definite and
%may be calculated without knowledge of the gain matrix $K_i$, and is invertible because $P^-_i$ and $R_i$ are positive-definite and
$(\beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}{\bf{e}}_{i}^T)$ is positive-semidefinite, as shown in the Appendix A. 
The case when $\beta_{0,i}=1$ corresponds with the $i$-th object \emph{not} being measured with full certainty.
This case is analyzed in Appendix B.

As with the Kalman filter, we choose the estimation gain to minimize the magnitude of the posterior covariance measured by $\tr{P_i^+}$.
It requires that
%The only case where $E_i$ is not invertible is when $\beta_{0,i}=1$, or when there is full confidence that a particular object is not observed by any measurement.
%This is evaluated in the appendix to show that this case does not yield an infinite limit of the gain matrix derived below.
%
%The goal of the choosing $K_i$ is to minimize the uncertainty of the object state estimate.
%With a Kalman filter, this goal is accomplished by minimizing $\tr{P^+_i}$ with respect to $K_i$ for a single track and measurement; the same approach may be applied to the JPDAF, except the a posteriori covariance equation has changed, so $K_i$ is subject to change as well:
\begin{align}
0&=\frac{\partial\tr{P^+_i}}{\partial K_i},\nonumber
\\
&=-2(1-\beta_{0,i})\frac{\partial\tr{K_iH_iP^-_i}}{\partial K_i}+\frac{\partial\tr{K_iE_iK_i^T}}{\partial K_i},\nonumber
\\
&=-2(1-\beta_{0,i})P^-_iH_i^T+2K_iE_i.\label{eqn:Uncertainty0}
\end{align}
Therefore the optimal gain is given by
\begin{align}
K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
\end{align}

Finally, \refeqn{GainK} is substituted into \refeqn{CovExpanded} to obtain
\begin{align}
%\begin{split}
%P^+_{i}=&P^-_{i}-(1-\beta_{0,i})^2(P^-_iH_i^TE_i^{-1}H_iP^-_i
%\\
%&+P^-_iH_i^TE_i^{-1}H_iP^-_i
%-P^-_iH_i^TE_i^{-1}E_iE_i^{-1}H_iP^-_i)
%\end{split}
%\\
P^+_{i}=&P^-_{i}-(1-\beta_{0,i})^2P^-_iH_i^TE_i^{-1}H_iP^-_i,
\end{align}
which may be simplified in two ways
\begin{align}
P^+_{i}=&P^-_{i}-K_iE_iK_i^T, \quad \mbox{or equivalently}\label{eqn:JPDAFPostCovSymmetricUpdate}
\\
P^+_{i}=&\left(I-(1-\beta_{0,i})K_iH_i\right)P^-_i.\label{eqn:JPDAFPostCovSimpleUpdate}
\end{align}
The proposed gain is compared with the Kalman gain explicitly in Appendix C.
In short, the differences between the MUJPDAF and the conventional JPDAF are shown in \refeqn{GainK} and \refeqn{JPDAFPostCovSymmetricUpdate} or \refeqn{JPDAFPostCovSimpleUpdate}.

\section{Coalescence-Avoiding Optimal Joint Probabilistic Data Association Filter (C-JPDAF)}
\label{C-JPDAF}

% The main idea: coalescence occurs when the similarity overlap between estimates is large, so we seek to minimize a cost that includes a similarity index
% Explain how the assumptions of old algorithm make realistic scenarios with measurement scans infeasible

Coalescence of the JPDAF (or variations thereof) is the outcome of estimates becoming closer to each other as a result of sharing measurements. This occurs because the signs of the innovations typically cause updates that draw measurement estimates towards the observed measurements; the net effect tends to cause estimates to become more similar to each other (e.g. a large number of estimates in close proximity, estimates with large uncertainties).
The goal of the C-JPDAF is to minimize a cost that includes a similarity index in addition to uncertainty.
This concept and simple numerical examples are presented in~\cite{KauLovLee14}.
These are based on the assumptions that exactly two estimates are present with a single measurement at each time step with equal likelihood of detection between the estimates; this paper eliminates such an assumption.
Additionally, the similarity measure is slightly changed to be in terms of measurements and measurement estimates, derived below, so the technique may be applied to a wider variety of systems.

\subsection{Matusita's Similarity Measure of Many Estimates}

The Matusita's similarity measure is defined between two Gaussian distributions.
Let $y_i$ be a random variable corresponding to the $i$-th system of set $I$ such that $y_i \sim \mathcal{N}[\hat y_i,{\mathbf P}_i]$, and let indices $p,q\in I$ such that $p\neq q$.
From~\cite{KauLovLee14}, the portion of the Matusita's similarity measure corresponding to the spread in the estimate means is given by
\begin{align}
\zeta_{pq}=\exp \{-a(y_{p}- y_{q})^T({\mathbf P}_{p}
+{\mathbf P}_{q})^{-1}(y_{p}-y_{q})\},\label{eqn:Mat2Est}
\end{align}
where $a>0$ shapes the similarity measure.

However, the distributions of more than two estimates may exhibit a nonzero overlap.
The similarity measure is changed to suit \emph{any} number of estimates by summing all combinations of Matusita's similarity measures, illustrated in Figure \ref{fig:SimMeas}, expressed in a general sense as
\begin{align}
\zeta=\sum\limits_{p,q\in I,p\neq q}\exp \{-a(y_{p}- y_{q})^T({\mathbf P}_{p}
+{\mathbf P}_{q})^{-1}(y_{p}-y_{q})\},
\end{align}
which is the summation of every similarity measure from \refeqn{Mat2Est} between all different Gaussian distributions $y_i$.
\begin{figure}
\centerline{
		{\includegraphics[width=\columnwidth]{GaussianDistOverlap1D.png}}
	}
\caption{The similarity index between three Gaussian distributions is a function the summation measures of overlapping regions of two distributions (green) or three distributions (brown). This concept is easily extended to multivariate distributions.
}\label{fig:SimMeas}
\end{figure}

\subsection{Optimal Coalescence Avoidance}

% Explain motivation: minimizing just the uncertainty causes coalescence, so minimizing a similarity index serves to remove coalescence
% Present cost function and the definition of u (from $J_s=\exp(-au)$)
% Derive the remaining terms in a similar fashion to the last paper except with summations where necessary
% Explain which similarities need not be used (when a priori similarity is below some level, the gain of the C-JPDAF is equivalent to that of the MUJPDAF)

Simply minimizing uncertainties of the estimates is well-known to cause coalescence with the JPDAF, but including a similarity index serves to remove this coalescence.
Let $\mathbf{J}\in\Re$ be a cost function of coalescence avoiding filtering as
\begin{align}
\mathbf{J}=J_P+cJ_S,
\end{align}
where $J_P$ is an uncertainty cost that measures the confidence level of the estimates and $J_S$ is a similarity cost that measures a degree of coalescence. The constant $c$ is a positive weighting factor between $J_P$ and $J_S$. 
The estimator gain $K_i$ is selected to minimize the total cost $\mathbf{J}$.
In short, the estimator gain is chosen in the consideration of both uncertainty and coalescence, illustrated in Figure \ref{fig:CostTrends}.

\begin{figure}
\setlength{\unitlength}{0.1\columnwidth}
\centerline{\footnotesize\selectfont
\begin{picture}(8.5,7.3)(-0.5,-0.8)
\put(-0.5,3){\rotatebox{90}{Cost}}
\put(0,0){\includegraphics[width=0.8\columnwidth]{ECC14_fig1.pdf}}
\put(6.2,0.2){\shortstack[c]{Uncertainty\\ cost: $J_P$}}
\put(6.2,2.8){\shortstack[c]{Similarity\\ cost: $J_S$}}
\put(5.7,5.2){\shortstack[c]{Weighted sum:\\ $\mathbf{J}=J_P+cJ_S$}}
\put(4.98,0.78){\circle*{0.1}}
\put(2.48,1.57){\circle*{0.1}}
\put(5.0,-0.3){$K_{MUJPDAF}$}
\put(2.5,-0.3){$K_{C-JPDAF}$}
\put(3.1,-0.8){Estimator gain}
\end{picture}
}
\caption{Illustration of optimal coalescence avoidance: several costs are illustrated with respect to an element of an estimator gain $K$ (red: uncertainty cost $J_P$, blue: similarity cost $J_S$, black: weighted sum). The estimator gain $K_{MUJPDAF}$ is selected to minimize the uncertainty cost. The estimator gain $K_{C-JPDAF}$ is chosen such that the weighted sum of the uncertainty cost and the similarity cost is minimized. In short, at the expense of increased uncertainty at a single step, the proposed approach prevents coalescence. The increased uncertainty at each step is rewarded by reducing the overall estimation error drastically by maintaining correct data association during close proximity missions.}
\label{fig:CostTrends}
\end{figure}

\paragraph*{Uncertainty Cost}
The uncertainty cost $J_P$ is defined as the sum of the trace of each posterior covariance:
\begin{align}
J_P=\sum\limits_{i}\tr{P^+_i},\label{eqn:JP}
\end{align}
where $P^+_{i}$ is obtained by \refeqn{JPDAFCov}.
Therefore, $P_i$ does not depend on $K_j$ for $i\neq j$.
The derivative of $J_P$ is taken from \refeqn{Uncertainty0}
\begin{align}
\label{eqn:CostP}
\frac{\partial J_{P}}{\partial K_{i}}&=2K_iE_i-2(1-\beta_{0,i})P^-_iH_i^T.
\end{align}

\paragraph*{Similarity Cost}
Any shared $z_j$ on estimate $i$ yields an innovation $e_{ij}$.
During the measurement update, this innovation typically has the net effect of drawing the measurement estimate towards $z_j$ if $\beta_{ij}$ is nonzero.
This occurs for all estimates and all measurements; coalescence results when measurement estimates are drawn towards common measurements.
Therefore, coalescence of the JPDAF or MUJPDAF occurs inside the measurement space (e.g. if a sensor provides ranges and bearings, these are in the measurement space as opposed to a Cartesian state vector inside the state space).
Contrary to~\cite{KauLovLee14} where the similarity index is in terms of state space variables, the coalescence avoidance presented in this paper takes place inside the measurement space to most directly remove the coalescence.
Let the posterior measurement estimate be $\hat z_i^+\in\Re^m$ and its innovation be $e_{i}^+\in\Re^m$ that corresponds with its true measurement $z_i\in\Re^m$ (although association is unknown) as
\begin{align}
\hat z_i^+&=H_i\hat x_i^+,
\\
e_{i}^+&=z_i-\hat z_i^+=(H_ix_i+v_i)-H_i\hat x_{i}^+=H_i\tilde x_i^++v_i,
\end{align}
where $\tilde x_i^+=x_i-\hat x_{i}^+$ denotes the error between the true state and the posterior estimate. The posterior innovation covariance $S^+_i\in\Re^{m\times m}$ is given by
\begin{align}
S^+_i&=\mathrm{E}[e_i^+e_i^{+T}]=\mathrm{E}[(H_i\tilde x_i^++v_i)(H_i\tilde x_i^++v_i)^T],\nonumber
\\
&=H_i\mathrm{E}[\tilde x_i^+\tilde x_i^{+T}]H_i^T+\mathrm{E}[vv^T],\nonumber
\\
&=H_iP^+_iH_i^T+R_i.
\end{align}
From these, the similarity cost is defined as
%which are used to define $u_{p q}$ between $\hat z^+_{p}$ and $\hat z^+_{q}$:
\begin{align}
J_S&=\sum\limits_{p,q\in I,p\neq q}\exp (-au_{pq}),
\end{align}
where $u_{pq}\in\Re$ is
\begin{align}
u_{pq} & = (\hat z_{p}^+-\hat z^+_{q})^T{S^+_{pq}}^{-1}(\hat z^+_{p}-\hat z^+_{q})\label{eqn:U},
\end{align}
where $S^+_{pq}=S^+_{p}+S^+_{q}$.

Expanding \refeqn{U} with respect to the $p$-th system,
\begin{align}
u_{pq}=&\ 
(H_pK_p{\bf e}_p-\hat z^+_{q})^T
(H_pP^+_{p}H_p^T+R_p+S^+_q)^{-1}\nonumber
\\
&\ 
(H_pK_p{\bf e}_p-\hat z^+_{q}).
\end{align}
\newcommand{\Ibd}{\ensuremath{\mathbf{1}_{bd}}}

Next, we find the derivative of the similarity cost as follows.
Let $\Ibd\in \Re^{n\times m}$ be defined such that its $b,d$-th element is one and other elements are zero.
The partial derivative of $u_{pq}$ with respect to the scalar value at the $b,d$-th element of $K_p$, denoted as $K_{p,bd}$, is
\begin{align}
\deriv{u_{pq}}{K_{p,bd}}
=&\ 
(H_p\Ibd{\bf e}_p)^T
{S_{pq}}^{-1}
(H_pK_p{\bf e}_p-\hat z^+_{q})\nonumber
\\
&\ +
(H_pK_p{\bf e}_p-\hat z^+_{q})^T
\mathcal{S}
(H_pK_p{\bf e}_p-\hat z^+_{q})\nonumber
\\
&\ +
(H_pK_p{\bf e}_p-\hat z^+_{q})^T
{S_{pq}}^{-1}
(H_p\Ibd{\bf e}_p),\nonumber
\\
=&\ \{
2(H_p\Ibd{\bf e}_p)^T
{S_{pq}}^{-1}\nonumber
\\
&\ +(H_pK_p{\bf e}_p-\hat z^+_{q})^T
\mathcal{S}
\}
(H_pK_p{\bf e}_p-\hat z^+_{q}),\label{eqn:derivu}
\\
\deriv{u_{pq}}{K_{r,bd}}=&\ 0\quad \mbox{if } r\not\in\{p,q\},\label{eqn:derivu0}
\end{align}
where
\begin{align}
\mathcal{S}=&\ \deriv{{S^+_{pq}}^{-1}}{K_{p,bd}},\nonumber
\\
=&\ {S^+_{pq}}^{-1}
\deriv{\left(H_pP^+_{p}H_p^T+R_p+S^+_q\right)}{K_{p,bd}}
{S^+_{pq}}^{-1},\nonumber
\\
=&\ {S^+_{pq}}^{-1}
H_p
[(1-\beta_{0,p})(\Ibd H_pP_p^-+P_p^-H_p^T\Ibd^T)\nonumber
\\
&\ -\Ibd E_pK_p^T -K_pE_p\Ibd^T]
H_p^T
{S^+_{pq}}^{-1},\label{eqn:SS}
\end{align}
because
\begin{align}
&\deriv{\left(H_pP^+_{p}H_p^T+R_p+S^+_q\right)}{K_{p,bd}}
=
H_p
\deriv{P^+_{p}}{K_{p,bd}}
H_p^T\nonumber
\\
&=
H_p
[(1-\beta_{0,p})(\Ibd H_pP_p^-+P_p^-H_p^T\Ibd^T)\nonumber
\\
&\ \ \ -\Ibd E_pK_p^T -K_pE_p\Ibd^T]
H_p^T.\label{eqn:SSmidpart}
\end{align}
Then, \refeqn{SS} and \refeqn{SSmidpart} are substituted into \refeqn{derivu} or \refeqn{derivu0} to solve
\begin{align}
\label{eqn:JSK}
\deriv{J_S}{K_{p,bd}}=\ -a\sum\limits_{q\in I,q\neq p}\exp (-au_{pq})\deriv{u_{pq}}{K_{p,bd}}.
\end{align}

Necessary conditions for optimality are given by
\begin{align}
\deriv{\mathbf{J}}{K_i} = \deriv{J_P}{K_i} + c \deriv{J_S}{K_i} =0,\label{eqn:NCO}
\end{align}
where $\deriv{J_P}{K_i}$ and $\deriv{J_S}{K_i}$ may be nonzero.
In short, \refeqn{CostP} and \refeqn{JSK} are substituted into \refeqn{NCO} and repeated for all $p$ to obtain the set of optimal gains $\mathcal{K}=\{K_1,K_2,...,K_{n_t}\}$.
Then, the C-JPDAF is completed by updating the state estimate and covariance by using \refeqn{KalEst} and \refeqn{JPDAFCov}, respectively.
As the optimality condition is a nonlinear function of estimator gains, there is no closed-form solution.
Hence, we require a nonlinear equation solver.
When $c=0$, we can show that the optimal gain reduces to that of the MUJPDAF, which is used as an initial guess of numerical solutions.
Therefore, if the number of estimates exhibiting significant similarity is $n_t^*\leq n_t$ and $\mathcal C$ is the number of times \refeqn{NCO} is solved depending on the numerical accuracy, the required iterations is $n_t^*(n_t^*-1)\mathcal C$, where some variables may be reused (e.g. $S^+_{pq}\equiv S^+_{qp}$) and often the effect of \refeqn{derivu} is negligible.

\section{Numerical Examples}
\label{NumRes}

% Two examples with identical realistic (from Vallado book) sensors
% Provide conditions for both scenarios

We consider satellites in the lower earth orbit (LEO) to compare the performances of the conventional JPDAF, the MUJPDAF, and the C-JPDAF.
%The state vector $x=\begin{bmatrix}x_1 & x_2 & x_3 & \dot x_1 & \dot x_2 & \dot x_3\end{bmatrix}^T$ is defined in terms of Cartesian coordinates such that $(x_1,x_2,x_3)$ align with the first three axes of the earth-centered inertial (ECI) frame, where positions are in km and velocities are in km/sec.
%Two-body motion is considered so that for $\vec r=\begin{bmatrix}x_1 & x_2 & x_3\end{bmatrix}^T$, so the nonlinear equations of motion are
Let $\vec r=\begin{bmatrix}x_1 & x_2 & x_3\end{bmatrix}^T$ be the position vector of the $i$-th satellite represented with respect to the Earth-Centered Inertial (ECI) frame.
The equation of motion corresponding to the two-body problem is given by
\begin{align}
\ddot{\vec r}&=-\frac{\mu}{r^3}\vec r=-\mu(\vec r^T\vec r)^{-\frac32}\vec r,
\end{align}
where $\mu=398,600\ \frac{km^3}{sec^2}$ is the gravitational constant of the earth if position is estimated in kilometers and time is measured in seconds~\cite{Val01}.

Let $\delta$ denote an infinitesimal change of a state variable.
The linearized equations of state variable $\vec x=\begin{bmatrix}\vec r & \dot{\vec r}\end{bmatrix}^T\in\Re^6$ are given by
\begin{align}
\begin{bmatrix}
\delta\dot{\vec r} \\ \delta\ddot{\vec r}
\end{bmatrix}
%\delta\dot{\vec x}
=
\begin{bmatrix}
0_{3\times3} & I_{3\times3} \\
\mu\left[3\vec r({\vec r}^T\vec r)^{-\frac52}{\vec r}^T-({\vec r}^T\vec r)^{-\frac32}I_{3\times3}\right] & 0_{3\times3}
\end{bmatrix}
\begin{bmatrix}
\delta\vec r \\ \delta\dot{\vec r}
\end{bmatrix}
%\delta\vec x
.
\end{align}
Defining the above square matrix as $A_{lin}$, the linearized state transition matrix over the time step $dt$ is defined as
\begin{align}
F=\expm{(A_{lin}dt)}
\end{align}
where $\expm$ denotes the matrix exponential function.
%The discretized linearized equations of motion from step $k-1$ to step $k$ are
%\begin{align}
%x_{k}=F_{k-1}x_{k-1}.
%\end{align} 
The process uncertainty may be subject to a variety of perturbations beyond two-body forces, and the process covariance is chosen as $Q=\diag[10^{-6}I_{3\times3}, 10^{-8}I_{3\times3}]$.

The satellites are measured with a single phased-array radar; the measurement space consists of the range $\rho$ between the sensor and the satellite (km), the range rate $\dot\rho$ (km/sec), and topocentric right ascension $\alpha$ and declination $\delta$ (rad) as defined in~\cite{Val01}.
The standard deviations $\sigma_\rho$, $\sigma_\alpha$, and $\sigma_\delta$ are based on the Mahe Ground Station uncertainty data from~\cite{VerSauSco04} and $\sigma_{\dot\rho}$ is taken from~\cite{KurAriAriEfe10}. The measurement uncertainty is $R=\diag[\sigma_\rho^2, \sigma_{\dot\rho}^2, \sigma_\alpha^2, \sigma_\delta^2]=\diag[0.15^2, 0.005^2, (\frac{0.01\pi}{180})^2, (\frac{0.01\pi}{180})^2]$. There is also a $10\%$ chance that a measurement is either deleted or an extraneous measurement is added at each time step.

The conventional JPDAF, MUJPDAF, and C-JPDAF have all been tested with a wide range of objects inside the sensor field of view.
However, this paper focuses on the uncertainty and coalescence of objects in close proximity, so only two or three objects are considered at a time.

%If these objects are not in close proximity with the tracks of interest in the following section, these objects are omitted from the results of this paper because they have an insignificant effect on the tracks of interest.
\paragraph*{Scenario A: Satellites in close proximity pass over a sensor field of view}
The first example compares the JPDAF with the MUJPDAF for tracking two nearby satellites passing over the sensor field of view. The initial conditions of the first two orbits are chosen as
\begin{align}
&x_{1,A}=x_{ref,A}+x_{pert,A}, \quad x_{2,A}=x_{ref,A}-x_{pert,A},\nonumber
\\
&x_{ref,A}=\begin{bmatrix}7000 & 0 & 1000 & 0 & 7.546 & 0\end{bmatrix}^T,\nonumber
\\
&x_{pert,A}=\begin{bmatrix}
0.2 & 0 & 0 & 0 & -0.105 & 0
\end{bmatrix}^T.
\end{align}
%and the initial conditions of the remaining six orbits are defined as
%\begin{align}
%x_b=x_{ref}+\begin{bmatrix}
%3\times10^{-4}\mathcal{R}_{1,3}, & 10^{-6}\mathcal{R}_{1,3}
%\end{bmatrix}\ \forall \ 2\leq b \leq 8,
%\end{align}
%where $\mathcal{R}_{1,3}$ is a $1\times3$ vector of Gaussian random numbers with mean $0$ and standard deviation $1$ defined by the MATLAB \emph{randn} command.
Two cases are considered with different initial covariances matrices: $P_{0,A1}=10^7\times Q$ for Case A1 and $P_{0,A2}=10^8\times Q$ for Case A2.
These scenarios are chosen because the estimates exhibit large initial uncertainties, so the spread in the innovations term $\tilde P$ from \refeqn{tildeP} affects the updates.
The gain of the MUJPDAF accounts well for $\tilde P$ whereas the gain for the conventional JPDAF does not.
Both data association algorithms are subject to the same set of measurements.
The RMS errors of the position and $J_P$ are summarized in Table \ref{tab:RMSptA}.
The results are plotted in Figure \ref{fig:CompareJPDAFandOJPDAF}.

\begin{table}
\caption{JPDAF and MUJPDAF Comparison: Scenario A}\label{tab:RMSptA}
\centerline{
\begin{tabular}{cccc}\hline
Case & Method & Radial RMS Error (km) & Mean $J_P$\\\hline
\vspace*{-0.02\columnwidth}\\
A1 & JPDAF &  0.12024 & 0.51254\\
A1 & MUJPDAF & 0.10641 & 0.49055\\\hline
\vspace*{-0.02\columnwidth}\\
A2 & JPDAF & 2.6353 & 5.7157\\
A2 & MUJPDAF & 0.19554 & 4.8456\\\hline
\end{tabular}}
\end{table}

\begin{figure}
{
\hspace*{0.05\columnwidth}
\centerline{
	\subfigure[\;Case A1 Orbital Radii]
		{\hspace*{-0.1\columnwidth}\includegraphics[width=0.5\columnwidth]{CaseA1_rad.pdf}}
	\subfigure[\;Case A1 Uncertainty Cost]
		{\hspace*{-0.0\columnwidth}\includegraphics[width=0.5\columnwidth]{CaseA1_Jp.pdf}}
	}
}
{
\hspace*{0.05\columnwidth}
\centerline{
	\subfigure[\;Case A2 Orbital Radii]
		{\hspace*{-0.1\columnwidth}\includegraphics[width=0.5\columnwidth]{CaseA2_rad.pdf}}
	\subfigure[\;Case A2 Uncertainty Cost]
		{\hspace*{-0.0\columnwidth}\includegraphics[width=0.5\columnwidth]{CaseA2_Jp.pdf}}
	}
}
\caption{The orbital radii and uncertainty costs are compared among the algorithms (JPDAF is dashed, MUJPDAF is solid).
These examples show that the posterior uncertainty of the MUJPDAF decreases faster than that of the JPDAF.
The MUJPDAF shares measurement updates less, thereby decreasing coalescence, shown in both Cases A1 and A2.
In Case A2 with the JPDAF, the coalescence causes consistent incorrect associations among the objects, known as track swapping.
}\label{fig:CompareJPDAFandOJPDAF}
\end{figure}

\paragraph*{Scenario B: Three satellites in close proximity have identical initial velocities}
The JPDAF, MUJPDAF, and C-JPDAF algorithms are compared for three satellites with initial conditions
\begin{align}
\begin{split}
&x_{1,B}=x_{ref,B}+x_{pert,B}, \quad x_{2,B}=x_{ref,B}-x_{pert,B},
\\
&x_{3,B}=x_{ref,B},
\\
&x_{ref,B}=\begin{bmatrix}7000 & 0 & 0 & 0 & 7.546 & 0\end{bmatrix}^T,
\\
&x_{pert,B}=\begin{bmatrix}
0.025 & 0 & 0 & 0 & 0 & 0
\end{bmatrix}^T,
\end{split}
\end{align}
where $P_{0,B}=Q$.
For the C-JPDAF, the parameters $a=5\times10^{-1}$ and $c=2.3\times10^{-4}$ are used.
The computation time is compared when the algorithms run on a conventional desktop computer.
Given the same measurement set, the results are shown in Table \ref{tab:RMSptB} and Figure \ref{fig:CompareJPDAFandOJPDAFandCJPDAF}.

	% Show coalescence with the JPDAF and MUJPDAF, but avoidance with the C-JPDAF
	% Compare the covariance costs and the computation time

\begin{table}
\caption{JPDAF, MUJPDAF, and C-JPDAF Comparison: Scenario B}\label{tab:RMSptB}
\centerline{
\begin{tabular}{ccc}\hline
Method & Radial RMS Error (km) & Computation time (sec)\\\hline
\vspace*{-0.02\columnwidth}\\
JPDAF & 0.14904 &  1.729\\
MUJPDAF & 0.09935 & 1.725\\
C-JPDAF & 0.046411 & 104.605\\\hline
\end{tabular}
}
\vspace*{-0.08\columnwidth}
\end{table}

\begin{figure}
\centerline{
	\subfigure[\;JPDAF]
		{\includegraphics[width=1\columnwidth]{JPDAF_2000sectest.pdf}}
	}
\centerline{
	\subfigure[\;MUJPDAF]
		{\includegraphics[width=1\columnwidth]{OJPDAF_2000sectest.pdf}}
	}
\centerline{
	\subfigure[\;C-JPDAF]
		{\includegraphics[width=1\columnwidth]{CJPDAF_2000sectest.pdf}}
	}
\caption{The true orbital radii (dashed) and their estimates (solid) for three tracks (blue, black, and red) are compared among all algorithms in Scenario B.
The JPDAF experiences the most coalescence such that the the estimates are not well distinguishable.
The estimates of the MUJPDAF have less uncertainty and therefore less similarity and coalescence.
The C-JPDAF greatly reduces the coalescence, even when measurement are heavily shared, at the cost of increased uncertainty.
}\label{fig:CompareJPDAFandOJPDAFandCJPDAF}
\end{figure}

\section{Conclusions}
\label{Conclusion}

The conventional JPDAF is a well-known algorithm with various applications in data association.
The soft decision approach makes the algorithm robust to missed detections and other clutter.
This paper enhances the JPDAF in two different ways.
First, the JPDAF suffers from a suboptimal Kalman gain.
A minimum uncertainty JPDAF (MUJPDAF) is derived with a new gain that minimizes the full posterior uncertainty.
This gain is computationally comparable to the JPDAF, and shows superior performance in decreasing the estimate uncertainties when the spread in the innovations is nonzero.

Second, both the JPDAF and the MUJPDAF involve measurement sharing, which tends to coalesce estimates in close proximity.
The work from~\cite{KauLovLee14} of the C-JPDAF involves a similarity cost when choosing the estimator gains.
In this paper, the research is matured; the C-JPDAF is generalized for any number of estimates and measurements and the similarity cost is found inside the measurement space where coalescence occurs.
When using the C-JPDAF, there must be a trade-off between uncertainty increase and coalescence removal and powerful computational resources are also required.

%There are several benefits to choosing non-Kalman gains. The Kalman gain is suboptimal with respect to minimizing the linearized a posteriori covariance of the JPDAF. This issue is resolved with using the MUJPDAF, which is computationally comparable and is shown to consistently exhibit less RMS position error and state estimate uncertainty from simulations. The only drawback is that there exists a singularity, which is shown to be easily avoidable and has negligible effect on the filter performance.
%
%However, both the JPDAF and MUJPDAF suffer from coalescence when estimates are in close proximity. The C-JPDAF is derived in a more general sense (same assumptions as the JPDAF and MUJPDAF), which minimizes a weighted cost of state estimate uncertainty and a similarity index of the cluster of any number of estimates. By choosing this index inside the measurement space, the coalescence is more directly removed than prior work in~\cite{KauLovLee14}. To remove coalescence this way, there must be a trade-off between uncertainty increase and coalescence removal and powerful computational resources are also required.

Future work may take two directions: first, the computational requirements of the C-JPDAF make the algorithm infeasible in many scenarios; hence a suboptimal analytic solution that still removes coalescence would make the algorithm suitable for more applications. Second, all three algorithms may undergo further Monte Carlo simulations to show their performance in a variety of scenarios and applications.
These simulations may also serve as direct methods to find the weighting parameters of the C-JPDAF.

% Future work
	% Finding a suboptimal analytic solution to the C-JPDAF to lower the computational requirements
	% Running Monte Carlo simulations on both algorithms to show their performance in a variety of scenarios and applications
	% These simulations may serve as direct methods to find the weighting parameters of the C-JPDAF

\begin{appendix}
\label{append}

\subsection{Invertibility of the new $E_i$} \label{NewPartOfEIsPSD}
Consider that object $i$ is not measured (probability $\beta_{0,i}$), then this portion of the innovation update $e_0=0_{m\times1}$ because no innovation exists where measurements are not available.
Then for all innovations in the set $J^*=\{0\}\cup J$, the Total Probability Theorem holds true for object $i$:
\begin{align}
\sum\limits_{j^*\in J^*}\beta_{ij^*}=1.
\end{align}
From~\cite[Eq. 1.4.16-(1-10)]{ShaRonThi2001}, the covariance spread in the means term of the mixture of $e$ is defined as (note that ${\bf e}_{i}$ remains unchanged)
\begin{align}
\tilde P_e=\sum\limits_{j^*}\beta_{ij^*}(e_{ij^*}-{\bf e}_{i})(e_{ij^*}-{\bf e}_{i})^T,
\end{align}
which is positive semi-definite. Manipulating this equation,
\begin{align}
\tilde P_e=&\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}e_{ij^*}^T
-\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}{\bf e}_{i}^T\nonumber
\\
&
-{\bf e}_{i}\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}^T
+{\bf e}_{i}{\bf e}_{i}^T\nonumber
\\
=&\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}e_{ij^*}^T-{\bf e}_{i}{\bf e}_{i}^T\nonumber
\\
=&\sum\limits_{j}\beta_{ij}e_{ij}e_{ij}^T-{\bf e}_{i}{\bf e}_{i}^T\geq0,
\label{eqn:CovSpreadPSD}
\end{align}
because $e_{i0}\equiv0$. Therefore $E_i$ is invertible $\forall\ 0\leq\beta_{0,i}<1$.

\subsection{The singularity of $E$ is not an infinite limit}
Consider the singularity when $E_i=0$ when no association is made to a track, or equivalently when $\beta^*=(1-\beta_{0,i})=\sum\limits_{j}\beta_j=0$.
It is clear from \refeqn{KalEst} and \refeqn{CovExpanded} that the updates are independent of $K_i$ exactly at this singularity.
We show that $K_i$ is \emph{not} growing without bound nearby this singularity, shown by evaluating the limit at this singularity to ensure boundedness.
Note that if $\sum\limits_{j}\beta_j\rightarrow0\Rightarrow \beta_j\rightarrow0\ \forall\ j$ because of the bounds $\sum\limits_{j}\beta_j\geq\beta_j\geq0\ \forall \ j$.
For clarity, let $\tilde\beta_p=\sum\limits_{j,j\neq p}\beta_j/\sum\limits_{j}\beta_j$ and drop the subscript $i$ because this analysis focuses on a single track.

First, we factor $\beta^*$ from $\sum\limits_{j} \beta_{j}e_{j}e_{j}^T$,
\begin{align}
\sum\limits_{j} \beta_{j}e_{j}e_{j}^T&=(\beta_1e_1e_1^T+\beta_2e_2e_2^T+...),\nonumber
\\
&=\beta^*
(e_1e_1^T-\tilde\beta_1e_1e_1^T+
e_2e_2^T-\tilde\beta_2e_2e_2^T+...),\nonumber
\\
&=\beta^*(\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T).
\end{align}
Similarly, we factor $\beta^*$ from ${{\bf{e}}}{\bf{e}}^T$,
\begin{align}
{{\bf{e}}}{\bf{e}}^T&=\sum\limits_{j} \beta_{j}e_{j}\sum\limits_{j} \beta_{j}e_{j}^T,\nonumber
\\
&=
(\beta_1e_1\sum\limits_{j} \beta_{j}e_{j}^T+\beta_2e_2\sum\limits_{j} \beta_{j}e_{j}^T+...),\nonumber
\\
&=\beta^*(e_1\sum\limits_{j} \beta_{j}e_{j}^T-\tilde\beta_1e_1\sum\limits_{j} \beta_{j}e_{j}^T\nonumber
\\
&\ \ \ +e_2\sum\limits_{j} \beta_{j}e_{j}^T-\tilde\beta_2e_2\sum\limits_{j} \beta_{j}e_{j}^T+...),\nonumber
\\
&=\beta^*(\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T).
\end{align}
Now that the summations inside $E$ are evaluated, their effects on $K$ are analyzed,
\begin{align}
\lim_{\beta^* \to 0}K&=\lim_{\beta^* \to 0}\beta^*P^-H^T
(\beta^*[S+\sum\limits_{j}e_{j}e_{j}^T\nonumber
\\
&\ \ \ -\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T
-\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T\nonumber
\\
&\ \ \ +\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T])^{-1},\nonumber
\\
&=P^-H^T\lim_{\beta^* \to 0}
(S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T\nonumber
\\
&\ \ \ -\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T+\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T)^{-1},
\label{eqn:OptGainNoSingularity}
\end{align}
which not an infinite limit because
\begin{align}
\lim_{\beta^* \to 0}\tilde\beta_p&=\lim_{\beta^* \to 0}\frac{...+\beta_{p-1}+0+\beta_{p+1}+...}{...+\beta_{p-1}+\beta_p+\beta_{p+1}+...}\ \forall\ p\in J,\nonumber
\\
&=1-\lim_{\beta^* \to 0}\frac{\beta_p}{\beta^*},
\end{align}
and therefore the following inequalities hold:
\begin{align}
&0\leq\beta_p\leq\beta^*\leq1\Rightarrow0\leq\lim_{\beta^* \to 0}\frac{\beta_p}{\beta^*}\leq1\Rightarrow0\leq\lim_{\beta^* \to 0}\tilde\beta_p\leq1,
\end{align}
and therefore \refeqn{OptGainNoSingularity} may be solved
\begin{align}
\lim_{\beta^* \to 0}K
%&=
%P^-H^T
%\left(S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}{B_j}e_{j}e_{j}^T\right)^{-1}
%\\
&=P^-H^T
(S+\sum\limits_{j}(1-B_j)e_{j}e_{j}^T)^{-1},
\label{eqn:OptGainNoSingularitySolved}
\end{align}
for $0\leq B_j\leq1\ \forall\ j$ and it can be easily seen that the sum of the terms inside the parentheses are invertible.
% and we know that $S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}{B_j}e_{j}e_{j}^T$ is invertible because $S$ is positive definite and the remaining terms are a simplified form of \refeqn{CovSpreadPSD}.
From a computational standpoint, for every $\beta^*_i\approx0$, the effect of the measurement update should be negligible, so we may choose $K_i=0_{n\times m}$ near this singularity as it will have little effect on the filter performance; this typically occurs when the object is not measured.
 Choosing this $K_i$ results in no difference between the a priori and a posteriori update results.

\subsection{The optimal gain and Kalman gain are not guaranteed to be equivalent}
The Kalman gain $K_{Kalman}$ and the MUJPDAF gain $K_{MUJPDAF}$ are not always equivalent with a proof by contradiction. Consider that these gains are equivalent (object subscript $i$ is omitted for clarity):
\begin{align}
&K_{Kalman}=K_{MUJPDAF},\nonumber
\\
%&P^-H^TS^{-1}=(1-\beta_{0})P^-H^TE^{-1},
%\\
&P^-H^TS^{-1}E=(1-\beta_{0})P^-H^T,\nonumber
\\
%&=P^-H^TS^{-1}\left((1-\beta_{0})S+{\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T\right)
%\\
&=(1-\beta_{0})P^-H^T+P^-H^TS^{-1}({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T).\label{eqn:KalMUareSame}
\end{align}
Therefore if \refeqn{KalMUareSame} is true, then
\begin{align}
K_{Kalman}({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T)=0,\label{eqn:ContradictionResult}
\end{align}
which implies that the spread in the innovations $\tilde P\equiv0$, not assumed in the JPDAF.
The result of \refeqn{ContradictionResult} may only occur if $K_{Kalman}=0$, the probabilities that each measurement $j$ belongs to each object $i$ are exactly $0$ or $1$, or $K_{Kalman}\perp({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T)$ . None of these conditions are guaranteed, so it follows that $K_{MUJPDAF}=K_{Kalman}$ is not guaranteed.

\end{appendix}

\bibliography{BibMaster}
\bibliographystyle{IEEEtran}

\end{document}

%\section{Numerical Examples}
%\label{NumRes}
%
%We consider estimation of two satellites in a low-Earth orbit with crossing orbital trajectories. It is assumed that the range toward satellites are measured with unknown origination. 
%
%The dynamics of each satellite is modeled as a two-body problem between each satellite and Earth, i.e., the equations of motion for each orbit only involve force caused by the gravity between the Earth and the satellite and neglect all other forces. For simplicity, both orbits occupy the same plane, so polar coordinates are used to define our state variable $x=\begin{bmatrix} r & \dot r & \theta & \dot \theta\end{bmatrix}^{T}$. The resulting equations of motion of each satellite are given by
%\begin{gather*}
%\ddot r - r\dot\theta^2 = -\frac{\mu}{r^2},\\
%r\ddot\theta + 2\dot r\dot\theta =0,
%\end{gather*}
%where $\mu=3.986\times10^5\, \mathrm{km^3/s^2}$ is the gravitational parameter of the Earth. These are linearized to obtain the system matrix of the discrete-time linear equations of motion considered in this paper:
%\begin{align}
%A_{i_k}=
%\exp\left(\begin{bmatrix}
%0 & 1 & 0 & 0 \\
%\frac{2\mu}{r_{i_k}^3} +\dot\theta^2_{i_k} & 0 & 0 & 2r_{i_k}\dot \theta_{i_k} \\
%0 & 0 & 0 & 1 \\
%2r_{i_k}^{-2}\dot r_{i_k}\dot\theta_{i_k} & -2r^{-1}\dot\theta_{i_k} & 0 &  -2r_{i_k}^{-1}\dot r_{i_k}\\
%\end{bmatrix}\Delta t\right).
%\end{align}
%%where  and $H=\begin{bmatrix} 1 & 0 & 0 & 0\end{bmatrix}$.
%
%
%The initial estimates of the states are designed to occur shortly before the two trajectories cross paths near $r_{ref}=7000km$ from the center of the earth; for the time to complete a circular orbit at this radius, $t_{ref}=2\pi \sqrt{\frac{r_{ref}^3}{\mu}}$, we can solve for a reference circular orbit $X_{ref}=\begin{bmatrix} r_{ref} & 0 & 0 & \frac{2\pi}{t_{ref}}\end{bmatrix}^T$, then perturb this orbit slightly by $X_{pert}=\begin{bmatrix} 0.40 & 0 & 0 & -1.5\times10^{-5}\end{bmatrix}^T$. We can easily choose the initial conditions to be $X_{1_0}= X_{ref}+X_{pert}$ and $X_{2_0}= X_{ref}-X_{pert}$. The initial estimates are assumed to be relatively accurate: there is no need to simulate large uncertainty when the paths are initially sufficiently far apart because this paper is concerned with the coalescence when the objects are close together, so the state and covariance will have been determined by simulated measurements while the estimates were still spread significantly apart; hence for both estimates, the initial covariance was chosen to be $P_0=diag(\begin{bmatrix}
%2.5\times10^{-4} & 2.5\times10^{-6} & 3.9\times10^{-8} & 3.9\times10^{-8} \end{bmatrix})$.
%The process noise was chosen to be equivalent for both estimates such that $Q=diag(\begin{bmatrix}
%2.5\times10^{-5} & 2.5\times10^{-7} & 3.9\times10^{-9}& 3.9\times10^{-9}\end{bmatrix})$ and $R=\begin{bmatrix}
%2.25\times10^{-2}
%\end{bmatrix}$ applied at each time step.
%
%%\EditTL{Put the values of $P_0$, $Q$ and $R$ here. Describe how the measurements are switched between two satellites. }
%
%
%Three filtering approaches are applied, namely extended Kalman filter (EKF), JPDAF, and C-JPDAF. For the EKF, it is assumed that the correct measurement origination is know with certainty, i.e., all of the measurements are associated to the correct object. This serves as an ideal reference case to be compared with JPDAF and C-JPDAF. For the C-JPDAF, the weighting parameters of the cost are chosen as $a=0.02$ and $c=5\times10^{-5}$. We propagate the nonlinear equations of motion numerically using Matlab, and we generate a measurement at every $0.1$ seconds with an equal likelihood that its origination is from either object such that all three algorithms are subject to identical noisy measurements. The estimates are propagated with their nonlinear equations of motion; the data association and filtering algorithms use the linearized form.
%
%\begin{figure}
%\centerline{
%	\subfigure[EKF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{EKF.png}\hspace*{0.07\textwidth}}
%	}
%\centerline{
%	\subfigure[JPDAF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{JPDAF.png}\hspace*{0.07\textwidth}}
%	}
%\centerline{
%	\subfigure[C-JPDAF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{CAOJPDAF.png}\hspace*{0.05\textwidth}}
%	}
%\caption{Object 1 is represented in blue and object 2 is represented in red. A solid line represents an estimate and a dotted line represents the truth. The JPDAF is subject to coalescence of the neighboring track, causing the estimations to switch trajectories. The C-JPDAF is subject to much less coalescence and hence does not exhibit this problem.}
%\label{AllAlgorithms}
%\end{figure}
%
%The corresponding results for each of the EKF, JPDAF, and C-JPDAF algorithms are illustrated in Figure \ref{AllAlgorithms}. 
%First, EKF exhibits good estimation performance as there is no error or uncertainty in data association. The performance of JPDA is similar to EKF when two satellites are far apart from each other. However, when $50\leq t\leq 70$ seconds, two satellites become closer, and the estimated states are coalesced. When $t\geq 70$ seconds, the estimated states are incorrectly crossed. This illustrates undesirable performance of JPDAF especially for crossing objects. 
%
%The proposed C-JPDAF does not exhibit the incorrect, crossed estimation of JPDA, and the estimated states follow each of the correct trajectories throughout the crossing orbital trajectories. The resulting RMS errors for the orbital radius of each method are summarized at Table \ref{tab:RMS}. This illustrates that the performance of the proposed C-JPDAF is comparable to EKF where there is no error in data association. 
%
%\begin{table}
%\caption{Estimation results}\label{tab:RMS}
%\centerline{
%\begin{tabular}{ccc}\hline
%Method & Radial RMS Error (km) & Computation time (sec) \\\hline
%EKF & 1.438e-05 &6.118\\
%JPDAF & 7.1242e-05 &6.130\\
%C-JPDAF & 2.3145e-05  & 19.864 \\\hline
%\end{tabular}}
%\end{table}
%
%The major drawback of the C-JPDAF is the added computation time. However, the results of all three simulated algorithms (EKF, JPDAF, and C-JPDAF) are nearly identical when the estimates are not near the crossing region. Hence, the C-JPDAF can be turned on only during these close proximity regions, then turned off when unnecessary to improve the computational efficiency. 
%
%%_______________________________________________%
%
%\section{Conclusions and Future Work}
%\label{ConclusionFutureWork}
%We explore a new approach to reduce coalescence of close proximity multi-object tracking. The estimator gains are chosen such that a weighted cost with respect to posterior uncertainty and similarity is minimized. The proximity similarity portion of the Matusita's measure is chosen to evaluate the similarity. Necessary conditions for optimality are derived, and they are solved numerically to obtain optimal estimator gains. Numerical examples illustrate that it exhibits excellent estimation and data association properties for crossing tracks in close proximity. Compared with other techniques to avoid coalescence, the proposed approach has simpler structures. Future works include generalizing C-JPDAF for arbitrary number of objects with clutter, and improving the computational efficiencies.

### 2018-06-09 Conference/AAS 2015 DA/ACC 2015/ACC15_Final.tex
added Conference folder

--- 2018-06-09 Conference/AAS 2015 DA/ACC 2015/ACC15.tex
@@
 Minimum Uncertainty JPDA Filter and Coalescence Avoidance}
-\author{Evan Kaufman, Thomas Alan Lovell, and Taeyoung Lee$^*$
+\author{Evan Kaufman and Taeyoung Lee$^*$% Thomas Alan Lovell,
  \thanks{Evan Kaufman and Taeyoung Lee are with Department of Aerospace Engineering, George Washington University, Washington, DC. Email: {\tt\footnotesize \{evankaufman, tylee\}@gwu.edu }}
-\thanks{Thomas Alan Lovell is a Research Aerospace Engineer at the Space Vehicles Directorate, Air Force Research Lab (AFRL), Albuquerque, NM. Email: {\tt\footnotesize afrl.rvsv@kirtland.af.mil }}
+%\thanks{Thomas Alan Lovell is a Research Aerospace Engineer at the Space Vehicles Directorate, Air Force Research Lab (AFRL), Albuquerque, NM. Email: {\tt\footnotesize afrl.rvsv@kirtland.af.mil }}
 \thanks{$^*$This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}}
@@
 To properly track these objects, we require estimation techniques that associate objects with measurements.
-When the objects are in very close proximity, the association uncertainties grow, and an effective algorithm is necessary to maintain close estimations of the objects~\cite{KauLovLee14}.
+When the objects are in very close proximity, the association uncertainties grow, and an effective algorithm is necessary to maintain accurate estimations of the objects~\cite{KauLovLee14}.
 
@@
 
-Alternatively with a soft decision technique such as the JPDAF, measurements are shared among the estimates.
+Alternatively with a soft decision technique such as the joint probabilistic data association filter (JPDAF), measurements are shared among the estimates.
 With the JPDAF, the likelihoods of association between measurements and estimates serve as weighting factors inside the measurement update portion of the filtering~\cite{JPDAF1},~\cite{TrackDataAssoc}.
@@
 The JPDAF is based on the Kalman gain, which minimizes the posterior uncertainty of individual tracks.
-However, the Kalman gain is not guaranteed to \emph{exactly} minimize the posterior uncertainties of the JPDAF.
+However, the Kalman gain is not guaranteed to \emph{exactly} minimize the posterior uncertainties for data association of the JPDAF.
 %The unnecessarily large uncertainty of the suboptimal Kalman gain serves as motivation to derive an optimal gain that minimizes the a posteriori uncertainty of the linearized JPDAF equations.
+
 In this paper, we derive the optimal estimation gain that minimizes the magnitude of the posterior covariance matrix, and the resulting minimum uncertainty JPDAF is referred to as the MUJPDAF.
@@
 Therefore, we propose a new coalescence-avoiding optimal JPDAF referred to as the C-JPDAF, which is a systematic approach to remove coalescence by minimizing the weighted sum of the uncertainty and a similarity index between estimates~\cite{KauLovLee14}.
-The prior research covers the concept in simple cases, such that only two estimates are considered with a single measurement at each time step, making the algorithm infeasible in many realistic applications.
+The prior work covers the concept in simple cases, such that only two estimates are considered with a single measurement at each time step, making the algorithm infeasible in many realistic applications.
 Therefore, we generalize the algorithm to handle any number of objects and measurements, as well as missed detections and measurements originating from extraneous clutter.
@@
 
-During the flow update, the a priori (denoted with superscript $-$) estimated state $\hat x_{i_k}^-$ is found with \refeqn{xkp} and its a priori covariance $P_{i_k}^-$ is calculated from the last posterior (denoted with superscript $+$) covariance $P_{i_{k-1}}^+$ for all $i$ as
+During the flow update, the a priori (denoted with the superscript $-$) estimate $\hat x_{i_k}^-$ is found with \refeqn{xkp} and its a priori covariance $P_{i_k}^-$ is calculated from the last posterior (denoted with the superscript $+$) covariance $P_{i_{k-1}}^+$ for all $i$ as
 \begin{gather}
@@
 %In short, the differences between the MUJPDAF and the conventional JPDAF are shown in \refeqn{GainK} and \refeqn{JPDAFPostCovSymmetricUpdate} or \refeqn{JPDAFPostCovSimpleUpdate}.
-The optimal gain $K_i$ is derived below in the sense that the posterior uncertainties are minimized with respect to the generalized equations of the JPDAF.
+Here we derive the optimal gain $K_i$ such that the posterior uncertainties are minimized with respect to the generalized JPDAF equations.
 %Of course, the estimator cannot be optimal if the system is nonlinear, because the linearized equations to not fully represent the dynamics of a nonlinear system.
@@
 \begin{align}
-\tilde P=&\ K_i\left({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_{i}^T,\label{eqn:tildeP}
-\end{align}
-and $P_i^c$ corresponds to the posterior covariance of object $i$ if only a single measurement serves to update this track~\cite{TrackDataAssoc}, and it can be written as
+\tilde P=&\ K_i\parenth{{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T}K_{i}^T,\label{eqn:tildeP}
+\end{align}
+and $P_i^c$ corresponds to the posterior covariance of object $i$ if only a single measurement serves to update this track~\cite{TrackDataAssoc}%, and it can be written as
 %We can define $P_i^c$ without any assumptions on $K_i$ (i.e.
@@
 (see~\cite{OptEst1}).
-Substituting \refeqn{tildeP} and \refeqn{GenCov} into \refeqn{JPDAFCov} and rearranging,
+Substituting \refeqn{tildeP} and \refeqn{GenCov} into \refeqn{JPDAFCov},% and rearranging,
 \begin{align}
@@
 \begin{align}
-E_i=&\ (1-\beta_{0,i})S_i+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T.
+E_i=&\ (1-\beta_{0,i})S_i+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T.\label{eqn:E}
 \end{align}
@@
 $(\beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}{\bf{e}}_{i}^T)$ is positive-semidefinite, as shown in the Appendix A. 
-The case when $\beta_{0,i}=1$ corresponds with the $i$-th object \emph{not} being measured with full certainty.
-This case is analyzed in Appendix B.
+The case when $\beta_{0,i}=1$ corresponds with the $i$-th object \emph{not} being measured with full certainty, analyzed in Appendix B.
 
@@
 In short, the differences between the MUJPDAF and the conventional JPDAF are shown in \refeqn{GainK} and \refeqn{JPDAFPostCovSymmetricUpdate} or \refeqn{JPDAFPostCovSimpleUpdate}.
+As the proposed optimal gain minimizes the posteriori uncertainties, the corresponding JPDA performs better, as illustrated by numerical examples later.
 
@@
 \begin{align}
-\zeta_{pq}=\exp \{-a(y_{p}- y_{q})^T({\mathbf P}_{p}
-+{\mathbf P}_{q})^{-1}(y_{p}-y_{q})\},\label{eqn:Mat2Est}
+\zeta_{pq}=\exp \{-a(\hat y_{p}- \hat y_{q})^T({\mathbf P}_{p}
++{\mathbf P}_{q})^{-1}(\hat y_{p}-\hat y_{q})\},\label{eqn:Mat2Est}
 \end{align}
@@
 \begin{align}
-\zeta=\sum\limits_{p,q\in I,p\neq q}\exp \{-a(y_{p}- y_{q})^T({\mathbf P}_{p}
-+{\mathbf P}_{q})^{-1}(y_{p}-y_{q})\},
+\zeta=\sum\limits_{p,q\in I,p\neq q}\exp \{-a(\hat y_{p}- \hat y_{q})^T({\mathbf P}_{p}
++{\mathbf P}_{q})^{-1}(\hat y_{p}-\hat y_{q})\},
 \end{align}
@@
 \end{align}
-where $S^+_{pq}=S^+_{p}+S^+_{q}$.
-
-Expanding \refeqn{U} with respect to the $p$-th system,
+where $S^+_{pq}=S^+_{p}+S^+_{q}$. Expanding \refeqn{U} with respect to the $p$-th system,
 \begin{align}
@@
 \\
-&\ 
+&\times  
 (H_pK_p{\bf e}_p-\hat z^+_{q}).
@@
 2(H_p\Ibd{\bf e}_p)^T
-{S_{pq}}^{-1}\nonumber
-\\
-&\ +(H_pK_p{\bf e}_p-\hat z^+_{q})^T
+{S_{pq}}^{-1} +(H_pK_p{\bf e}_p-\hat z^+_{q})^T
 \mathcal{S}
-\}
-(H_pK_p{\bf e}_p-\hat z^+_{q}),\label{eqn:derivu}
+\}\nonumber\\
+&\times (H_pK_p{\bf e}_p-\hat z^+_{q}),\label{eqn:derivu}
 \\
@@
 \begin{align}
-&x_{1,A}=x_{ref,A}+x_{pert,A}, \quad x_{2,A}=x_{ref,A}-x_{pert,A},\nonumber
-\\
-&x_{ref,A}=\begin{bmatrix}7000 & 0 & 1000 & 0 & 7.546 & 0\end{bmatrix}^T,\nonumber
-\\
-&x_{pert,A}=\begin{bmatrix}
+x_{1,A}&=x_{ref,A}+x_{pert,A}, \quad x_{2,A}=x_{ref,A}-x_{pert,A},\nonumber
+\\
+x_{ref,A}&=\begin{bmatrix}7000 & 0 & 1000 & 0 & 7.546 & 0\end{bmatrix}^T,\nonumber
+\\
+x_{pert,A}&=\begin{bmatrix}
 0.2 & 0 & 0 & 0 & -0.105 & 0
@@
 The results are plotted in Figure \ref{fig:CompareJPDAFandOJPDAF}.
+The filtering and data association of the MUJPDAF shows a faster reduction in the estimate uncertainty, thereby more accurately associating measurements to the satellites.
+Obtaining correct measurement associations increases the accuracy of the estimates and decreases the sharing of the measurements, leading to coalescence.
+In Case A2, the coalescence of the JPDAF is so great that the estimates are consistently incorrectly associated with measurements, known as track swapping; by employing the MUJPDAF, we avoid this issue.
 
@@
 {
-\hspace*{0.05\columnwidth}
 \centerline{
-	\subfigure[\;Case A1 Orbital Radii]
-		{\hspace*{-0.1\columnwidth}\includegraphics[width=0.5\columnwidth]{CaseA1_rad.pdf}}
-	\subfigure[\;Case A1 Uncertainty Cost]
-		{\hspace*{-0.0\columnwidth}\includegraphics[width=0.5\columnwidth]{CaseA1_Jp.pdf}}
+	\subfigure[\;Case A1: orbital radius (JPDAF), orbital radius (MUJPDAF), and uncertainty costs for both]
+		{\includegraphics[width=1\columnwidth]{CaseA1_3by1.pdf}}
 	}
@@
 {
-\hspace*{0.05\columnwidth}
 \centerline{
-	\subfigure[\;Case A2 Orbital Radii]
-		{\hspace*{-0.1\columnwidth}\includegraphics[width=0.5\columnwidth]{CaseA2_rad.pdf}}
-	\subfigure[\;Case A2 Uncertainty Cost]
-		{\hspace*{-0.0\columnwidth}\includegraphics[width=0.5\columnwidth]{CaseA2_Jp.pdf}}
+	\subfigure[\;Case A2: orbital radius (JPDAF), orbital radius (MUJPDAF), and uncertainty costs for both]
+		{\includegraphics[width=1\columnwidth]{CaseA2_3by1.pdf}}
 	}
 }
-\caption{The orbital radii and uncertainty costs are compared among the algorithms (JPDAF is dashed, MUJPDAF is solid).
+\caption{The orbital radii and uncertainty costs are compared among the algorithms (JPDAF: dashed, MUJPDAF: solid, true: solid and thick).
 These examples show that the posterior uncertainty of the MUJPDAF decreases faster than that of the JPDAF.
@@
 \end{figure}
+%\begin{figure}
+%{
+%\hspace*{0.05\columnwidth}
+%\centerline{
+%	\subfigure[\;Case A1 Orbital Radii]
+%		{\hspace*{-0.1\columnwidth}\includegraphics[width=0.5\columnwidth]{CaseA1_rad.pdf}}
+%	\subfigure[\;Case A1 Uncertainty Cost]
+%		{\hspace*{-0.0\columnwidth}\includegraphics[width=0.5\columnwidth]{CaseA1_Jp.pdf}}
+%	}
+%}
+%{
+%\hspace*{0.05\columnwidth}
+%\centerline{
+%	\subfigure[\;Case A2 Orbital Radii]
+%		{\hspace*{-0.1\columnwidth}\includegraphics[width=0.5\columnwidth]{CaseA2_rad.pdf}}
+%	\subfigure[\;Case A2 Uncertainty Cost]
+%		{\hspace*{-0.0\columnwidth}\includegraphics[width=0.5\columnwidth]{CaseA2_Jp.pdf}}
+%	}
+%}
+%\caption{The orbital radii and uncertainty costs are compared among the algorithms (JPDAF is dashed, MUJPDAF is solid).
+%These examples show that the posterior uncertainty of the MUJPDAF decreases faster than that of the JPDAF.
+%The MUJPDAF shares measurement updates less, thereby decreasing coalescence, shown in both Cases A1 and A2.
+%In Case A2 with the JPDAF, the coalescence causes consistent incorrect associations among the objects, known as track swapping.
+%}\label{fig:CompareJPDAFandOJPDAF}
+%\end{figure}
 
@@
 \begin{align}
-\begin{split}
-&x_{1,B}=x_{ref,B}+x_{pert,B}, \quad x_{2,B}=x_{ref,B}-x_{pert,B},
-\\
-&x_{3,B}=x_{ref,B},
-\\
-&x_{ref,B}=\begin{bmatrix}7000 & 0 & 0 & 0 & 7.546 & 0\end{bmatrix}^T,
-\\
-&x_{pert,B}=\begin{bmatrix}
+x_{1,B}&=x_{ref,B}+x_{pert,B}, \quad x_{2,B}=x_{ref,B}-x_{pert,B},
+\nonumber\\
+x_{3,B}&=x_{ref,B},
+\nonumber\\
+x_{ref,B}&=\begin{bmatrix}7000 & 0 & 0 & 0 & 7.546 & 0\end{bmatrix}^T,
+\nonumber\\
+x_{pert,B}&=\begin{bmatrix}
 0.025 & 0 & 0 & 0 & 0 & 0
 \end{bmatrix}^T,
-\end{split}
 \end{align}
@@
 Given the same measurement set, the results are shown in Table \ref{tab:RMSptB} and Figure \ref{fig:CompareJPDAFandOJPDAFandCJPDAF}.
+Both the JPDAF and the MUJPDAF experience coalescence due to measurement sharing; however, the coalescence is less severe with the MUJPDAF, which is consistent with the conclusions of Scenario A.
+Measurements are shared with the C-JPDAF as well, but the filter coalescence minimization prevents measurement sharing from causing estimates to coalesce.
 
@@
 
-%There are several benefits to choosing non-Kalman gains.
-The Kalman gain is suboptimal with respect to minimizing the linearized a posteriori covariance of the JPDAF.
-This issue is resolved with using the MUJPDAF, which is computationally comparable and is shown to consistently exhibit less RMS position error and state estimate uncertainty from simulations.
-The only drawback is that there exists a singularity, which is shown to be easily avoidable and has negligible effect on the filter performance.
-%
-%However, both the JPDAF and MUJPDAF suffer from coalescence when estimates are in close proximity.
-The C-JPDAF is derived in a more general sense (same assumptions as the JPDAF and MUJPDAF), which minimizes a weighted cost of state estimate uncertainty and a similarity index of the cluster of any number of estimates.
-By choosing this index inside the measurement space, the coalescence is more directly removed than prior work in~\cite{KauLovLee14}.
-To remove coalescence this way, there must be a trade-off between uncertainty increase and coalescence removal and powerful computational resources are also required.
-
-Future work may take two directions: first, the computational requirements of the C-JPDAF make the algorithm infeasible in many scenarios; hence a suboptimal analytic solution that still removes coalescence would make the algorithm suitable for more applications.
+%Future work may take two directions: first, the computational requirements of the C-JPDAF make the algorithm infeasible in many scenarios; hence a suboptimal analytic solution that still removes coalescence would make the algorithm suitable for more applications.
 Second, all three algorithms may undergo further Monte Carlo simulations to show their performance in a variety of scenarios and applications.
-These simulations may also serve as direct methods to find the weighting parameters of the C-JPDAF.
-
-% Future work
-	% Finding a suboptimal analytic solution to the C-JPDAF to lower the computational requirements
-	% Running Monte Carlo simulations on both algorithms to show their performance in a variety of scenarios and applications
-	% These simulations may serve as direct methods to find the weighting parameters of the C-JPDAF
+%These simulations may also serve as direct methods to find the weighting parameters of the C-JPDAF.
 
@@
 \end{align}
-From~\cite[Eq. 1.4.16-(1-10)]{ShaRonThi2001}, the covariance spread in the means term of the mixture of $e$ is defined as (note that ${\bf e}_{i}$ remains unchanged)
+From~\cite[Eq. 1.4.16-(1-10)]{ShaRonThi2001}, the covariance spread in the means term of the mixture of $e$ is defined as% (note that ${\bf e}_{i}$ remains unchanged)
 \begin{align}
@@
 \begin{align}
-\sum\limits_{j} \beta_{j}e_{j}e_{j}^T&=(\beta_1e_1e_1^T+\beta_2e_2e_2^T+...),\nonumber
-\\
-&=\beta^*
-(e_1e_1^T-\tilde\beta_1e_1e_1^T+
-e_2e_2^T-\tilde\beta_2e_2e_2^T+...),\nonumber
+\sum\limits_{j} \beta_{j}e_{j}e_{j}^T&=(\beta_1e_1e_1^T+\beta_2e_2e_2^T+\cdots)\nonumber
+%\\
+%&=\beta^*
+%(e_1e_1^T-\tilde\beta_1e_1e_1^T+
+%e_2e_2^T-\tilde\beta_2e_2e_2^T+\cdots),\nonumber
 \\
@@
 \\
-&=
-(\beta_1e_1\sum\limits_{j} \beta_{j}e_{j}^T+\beta_2e_2\sum\limits_{j} \beta_{j}e_{j}^T+...),\nonumber
-\\
-&=\beta^*(e_1\sum\limits_{j} \beta_{j}e_{j}^T-\tilde\beta_1e_1\sum\limits_{j} \beta_{j}e_{j}^T\nonumber
-\\
-&\ \ \ +e_2\sum\limits_{j} \beta_{j}e_{j}^T-\tilde\beta_2e_2\sum\limits_{j} \beta_{j}e_{j}^T+...),\nonumber
-\\
+%&=
+%(\beta_1e_1\sum\limits_{j} \beta_{j}e_{j}^T+\beta_2e_2\sum\limits_{j} \beta_{j}e_{j}^T+\cdots),\nonumber
+%\\
+%&=\beta^*(e_1\sum\limits_{j} \beta_{j}e_{j}^T-\tilde\beta_1e_1\sum\limits_{j} \beta_{j}e_{j}^T\nonumber
+%\\
+%&\ \ \ +e_2\sum\limits_{j} \beta_{j}e_{j}^T-\tilde\beta_2e_2\sum\limits_{j} \beta_{j}e_{j}^T+\cdots),\nonumber
+%\\
 &=\beta^*(\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T).
@@
 \lim_{\beta^* \to 0}K&=\lim_{\beta^* \to 0}\beta^*P^-H^T
-(\beta^*[S+\sum\limits_{j}e_{j}e_{j}^T\nonumber
-\\
-&\ \ \ -\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T
--\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T\nonumber
-\\
-&\ \ \ +\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T])^{-1},\nonumber
+(\beta^*[S+\sum\limits_{j}e_{j}e_{j}^T
+-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T\nonumber\\
+&\ \ -\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T+\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T])^{-1}\nonumber
 \\
@@
 \\
-&\ \ \ -\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T+\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T)^{-1},
+&\ \ -\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T+\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T)^{-1},
 \label{eqn:OptGainNoSingularity}
@@
 
-\subsection{The optimal gain and Kalman gain are not guaranteed to be equivalent}
-The Kalman gain $K_{Kalman}$ and the MUJPDAF gain $K_{MUJPDAF}$ are not always equivalent with a proof by contradiction.
-Consider that these gains are equivalent (object subscript $i$ is omitted for clarity):
-\begin{align}
-&K_{Kalman}=K_{MUJPDAF},\nonumber
-\\
-%&P^-H^TS^{-1}=(1-\beta_{0})P^-H^TE^{-1},
-%\\
-&P^-H^TS^{-1}E=(1-\beta_{0})P^-H^T,\nonumber
-\\
-%&=P^-H^TS^{-1}\left((1-\beta_{0})S+{\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T\right)
-%\\
-&=(1-\beta_{0})P^-H^T+P^-H^TS^{-1}({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T).\label{eqn:KalMUareSame}
-\end{align}
-Therefore if \refeqn{KalMUareSame} is true, then
-\begin{align}
-K_{Kalman}({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T)=0,\label{eqn:ContradictionResult}
-\end{align}
-which implies that the spread in the innovations $\tilde P\equiv0$, not assumed in the JPDAF.
-The result of \refeqn{ContradictionResult} may only occur if $K_{Kalman}=0$, the probabilities that each measurement $j$ belongs to each object $i$ are exactly $0$ or $1$, or $K_{Kalman}\perp({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T)$ .
-None of these conditions are guaranteed, so it follows that $K_{MUJPDAF}=K_{Kalman}$ is not guaranteed.
+\subsection{The optimal gain and Kalman gain are not equivalent}
+The Kalman filter gain $K_{KF}=P^-H^TS^{-1}$ from~\cite{TrackDataAssoc} and the MUJPDAF gain $K_{MUJPDAF}=(1-\beta_{0})P^-H^TE^{-1}$ from \refeqn{GainK} are not always equivalent with a proof by contradiction.
+Consider that these gains are equivalent and multiplying by $E$ from \refeqn{E} (object subscript $i$ is omitted for clarity):
+\begin{align}
+&K_{KF}E=K_{MUJPDAF}E=(1-\beta_{0})P^-H^T\nonumber
+\\
+&=P^-H^TS^{-1}\{(1-\beta_{0})S+{\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T\}\nonumber
+\\
+&=(1-\beta_{0})P^-H^T+K_{KF}({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T).\label{eqn:KalMUareSame}
+\end{align}
+Therefore if \refeqn{KalMUareSame} is true, then $K_{KF}({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T)=0$, which
+%\begin{align}
+%K_{KF}({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T)=0.\label{eqn:ContradictionResult}
+%\end{align}
+%%which implies that the spread in the innovations $\tilde P\equiv0$, not assumed in the JPDAF.
+%The result of \refeqn{ContradictionResult} 
+may only occur if $K_{KF}=0$, the probabilities that each measurement $j$ belongs to each object $i$ are exactly $0$ or $1$, or $K_{KF}\perp({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T)$.
+None of these conditions are satisfied in general, so it follows that the Kalman gain is different from the optimal gain proposed in this paper.
 

### 2018-06-09 Conference/AAS 2015 DA/ACC 2015/Generalized CJPDAF Derivation.tex
added Conference folder

\documentclass[10pt]{article}
\usepackage[letterpaper,text={6.5in,8.7in},centering]{geometry}
\usepackage{epic,eepic}
\usepackage{amssymb,amsmath,times,subfigure,graphicx,theorem}
%\usepackage{warmread}
%\usepackage[all,import]{xy}
%\usepackage{eepic}
\usepackage{subfigure}

\usepackage{amsmath}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Figure \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}

\renewcommand{\baselinestretch}{1.2}
\date{}

\renewcommand{\thesubsection}{\arabic{subsection}. }
\renewcommand{\thesubsubsection}{\arabic{subsection}.\arabic{subsubsection} }

\theoremstyle{plain}\theorembodyfont{\normalfont}
\newtheorem{prob}{Question}[section]
%\renewcommand{\theprob}{\arabic{section}.\arabic{prob}}
\renewcommand{\theprob}{\arabic{prob}}

\newenvironment{subprob}%
{\renewcommand{\theenumi}{\alph{enumi}}\renewcommand{\labelenumi}{(\theenumi)}\begin{enumerate}}%
{\end{enumerate}}%

\begin{document}

\section{Optimal Joint Probabilistic Data Assocation Filter (O-JPDAF)}

%\section*{Coalescence Avoiding Optimal Joint Probabilistic Data Association Filter (C-JPDAF)}
%
%The difference between the JPDAF and the C-JPDAF is that the JPDAF is optimal in the sence that estimate uncertainty is minimized. However, the C-JPDAF minimizes a weighted sum of uncertainty and similarity to decrease track coalescence, from which the JPDAF suffers when estimates are in close proximity over long periods of time.

% Paper copy & paste starts below:

\subsection{Flow Update}
Similar to Kalman filters, the JPDAF is composed of a flow update and a measurement update. During the flow update (or prediction step), the estimated state $\hat x_{i_k}$ and its covariance $P_{i_k}$ for each track $i=\{1,2,...,t\}$ of $t$ tracks at time step $k$ is updated as follows:
\begin{gather}
\hat x_{i_{k}} = f_{i_{k-1}}( \hat x_{i_{k-1}}),\\
P_{i_{k}} = F_{i_{k-1}} P_{i_{k-1}} F_{i_{k-1}}^T + Q_{i_{k-1}}, \quad F_{i_{k-1}}=\frac{\partial f}{\partial x}\bigg|_{x=\hat x_{i_{k}} }
\end{gather}
where the uncertainty of the process is represented with covariance $Q_{i_{k-1}}$.
These equations are applied between measurements to provide predictions of the states and their uncertainties. % paper change

\subsection{Measurement Update}

The measurement serves to update the state estimate. Because all equations in this section take place at the $k$ time step, this subscript is removed from the equations. Let  the a priori value of a variable and its a posteriori value be denoted by the superscript $-$ and $+$, respectively. Let subscript $j$ indicate a measurement index such that the set $j=\{1,2,...,r\}$ for the number of returns $r$ (or measurements). For a given measurement $z_j$, its measurement residual or innovation for the $i$-th object is given by
\begin{align*}
e_{ij} = z_j - \hat z_i
\end{align*}
where $\hat z_i$ is the predicted measurement of the linearized equation
\begin{align}
\hat z_i = H_i \hat x_{i}^-
\end{align}
such that the measurement is a function of state
\begin{align}
z=h(x), \quad H_i=\frac{\partial h}{\partial x}\bigg|_{x=\hat x_{i_{k}}^- }.
\end{align}•
Given the uncertainty of the measurements represented with covariance $R_i$, the innovation covariance $S_{i}$ is
\begin{align}
S_{i}=H_{i}P_{i}^{-}H_{i}^T+R_{i}.
\end{align}

% [Eq. 1.4.16-(1-10)]

\subsubsection{Likelihood of Association}
The computation required with a standard JPDA is inconsistent with the capabilities of the computational resources to account for an arbitrary number of objects to be tracked~\cite[Eq. 1.(1-2)]{Bar1990}. A modified version of the JPDA, commonly known as the "Cheap JPDA", provides a simpler approximation of the likelihood $\beta_{ij}$ that a return $j$ (or measurement) belongs to a particular track $i$ (or object). We define $G_{ij}$, which takes a very similar form as the Gaussian probability density function
\begin{align}
G_{ij}=\frac1{\sqrt{\det{S_{ij}}}}\exp{\{-\frac12e_{ij}^TS_{ij}^{-1}e_{ij}\}}
\end{align}
which is used to calculate the probability of association $\beta_{ij}$ as well as intermediate variables $\mathcal{S}_{ti}$ and $\mathcal{S}_{ri}$
\begin{align}
\beta_{ij}=\frac{G_{ij}}{\mathcal{S}_{ti}+\mathcal{S}_{ri}-G_{ij}+B}, 
\quad 
\mathcal{S}_{ti}=\sum\limits_{j}G_{ij}, 
\quad 
\mathcal{S}_{ri}=\sum\limits_{i}G_{ij}
\end{align}
where $B$ is a correction factor to account for the nonunity probability of detection as well as clutter and false alarms. Unless there are reasons that either phenomena may occur, we assume $B=0$.

\subsubsection{State and Covariance Update Equations}
The weighted sum of all innovations associated with the object $i$ is defined as ${\bf{e}}_{i}={\sum\limits_{j} \beta_{ij}e_{ij}}$.
The resulting a posteriori estimate of the state is given by~\cite[Eq. 6-(25-26)]{TrackDataAssoc}
\begin{align}
\hat x^+_{i}=&\hat x^-_{i}+K_{i}{\bf{e}}_{i}\label{eqn:KalEst}
\end{align}
where $K_i$ is a gain chosen to minimize the a posteriori uncertainty of the state estimate; $K_i$ is derived below optimally in the sence that the a posteriori uncertainties are minimized with respect to the generalized linearized equations of a JPDAF. Of course, the estimator cannot be optimal if the system is nonlinear, because the linearized equations to not fully represent the dynamics of a nonlinear system.

We define the likelihood that object $i$ is not detected is $\beta_{0,i}=1-{\sum\limits_{j} \beta_{ij}}$. The resulting a posteriori covariance can be written as [Tracking and Data Association, pg. 326, equations D-32 and D-33]
\begin{align}
\label{eqn:JPDAFCov}
P^+_{i}=& \beta_{0,i}P^-_{i}+(1-\beta_{0,i})P_i^c+\tilde P
\\
\tilde P=&K_i\left({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_{i}^T\label{eqn:tildeP}
\end{align}
where $P_i^c$ is defined as the a posteriori covariance of object $i$ if only a single measurement serves to update this track.
%From [Tracking and Data Association, pg. 326, equation D-27], which is referenced in numerous publications, $P_i^c$ is calculated
%\begin{align}
%\label{eqn:PcKalman1}
%P_i^c=(I-K_{i,Kalman}H_i)P_i^-
%\end{align}
%or equivalently
%\begin{align}
%\label{eqn:PcKalman2}
%P_i^c=P_i^--K_{i,Kalman}SK_{i,Kalman}^T
%\end{align}
%where Kalman gain $K_{i,Kalman}=P_iH_i^TS_i^{-1}$ is used. In fact, (\ref{eqn:PcKalman1}) and (\ref{eqn:PcKalman2}) are derived from equations that require $K_{i,Kalman}$ be used. Below we show that $K_{i,Kalman}$ cannot guarantee that $\tr{P^+_{i}}$ is minimized:
%\begin{align}
%P^+_{i}=& \beta_{0,i}P^-_{i}+(1-\beta_{0,i})\left(P_i^--K_{i,Kalman}SK_{i,Kalman}^T\right)+K_{i,Kalman}\left({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_{i,Kalman}^T\\
%=&P^-_{i}-K_{i,Kalman}\left((1-\beta_{0,i})S-({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T)\right)K_{i,Kalman}^T\\
%=&P^-_{i}-K_{i,Kalman}S^*K_{i,Kalman}^T, \quad S^*=(1-\beta_{0,i})S-({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T).
%\end{align}
%We define the cost function $J^*=\tr{K_{i,Kalman}}$ to test for optimality with respect to $K_i$ assuming that $K_i=K_{i,Kalman}$ to show if this is true:
%\begin{align}
%0&=\frac{\partial\tr{P^+_i}}{\partial K_i}
%\\
%0&=-2(1-\beta_{0,i})\frac{\partial\tr{K_iH_iP^-_i}}{\partial K_i}+\frac{\partial\tr{K_iS^*K_i^T}}{\partial K_i}
%\\
%0&=-2(1-\beta_{0,i})P^-_iH_i^T+2K_iE_i
%\\
%K_iE_i&=(1-\beta_{0,i})P^-_iH_i^T
%\\
%K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1}\label{eqn:GainK}
%\end{align}
We can define $P_i^c$ without any assumptions on $K_i$ (i.e. $K_i$ is not necessarily the Kalman gain, so the well-known Kalman covariance update equations are not valid here) [Applied Optimal Estimation, pg. 108, equation 4.2-12]
\begin{align}
\begin{split}
P^c_i&=(I-K_iH_i)P^-_i(I-K_iH_i)^T+K_iR_iK_i^T
\\
&=P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i(H_iP^-_iH_i^T+R_i)K_i^T.
\label{eqn:GenCov}
\end{split}
\end{align}
Substituting \refeqn{tildeP} and \refeqn{GenCov} into \refeqn{JPDAFCov},
\begin{align}
\begin{split}
P^+_{i}=&\beta_{0,i}P^-_{i}+(1-\beta_{0,i})\left(P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i(H_iP^-_iH_i^T+R_i)K_i^T\right)\\&+K_i\left({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_{i}^T
\end{split}
\\
\begin{split}
P^+_{i}=&P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\\&+K_i\left((1-\beta_{0,i})(H_iP^-_iH_i^T+R_i)+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_i^T
\end{split}
\\
\begin{split}
P^+_{i}=&P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_iE_iK_i^T,\\
&E_i=(1-\beta_{0,i})(H_iP^-_iH_i^T+R_i)+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\label{eqn:CovExpanded}
\end{split}
\end{align}
where $E_i=E_i^T$ and may be calculated without knowledge of the gain matrix $K_i$, and is invertible because $P^-_i$ and $R_i$ are positive-definite and $(\beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}{\bf{e}}_{i}^T)$ are positive-semidefinite, shown below.

Consider that object $i$ is not measured (likelihood $\beta_{0,i}$), then this portion of the innovation update $e_0=0_{m\times1}$ because no innovation exists where measurments are not available. Then for all innovations in the set $j^*=\{0,1, 2,...,r\}$, the Total Probability Theorem holds true for object $i$:
\begin{align}
\sum\limits_{j^*}\beta_{ij^*}=1.
\end{align}
From~\cite[Eq. 1.4.16-(1-10)]{ShaRonThi2001}, the covariance spread in the means term of the mixture of $e$ is defined as (note that ${\bf e}_{i}$ remains unchanged)
\begin{align}
\tilde P_e=\sum\limits_{j^*}\beta_{ij^*}(e_{ij^*}-{\bf e}_{i})(e_{ij^*}-{\bf e}_{i})^T
\end{align}
which is positive semi-definite. Manipulating this equation,
\begin{align}
\tilde P_e&=\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}e_{ij^*}^T
-\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}{\bf e}_{i}^T
-{\bf e}_{i}\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}^T
+{\bf e}_{i}{\bf e}_{i}^T
\\
\tilde P_e&=\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}e_{ij^*}^T-{\bf e}_{i}{\bf e}_{i}^T
\\
\tilde P_e&=\sum\limits_{j}\beta_{ij}e_{ij}e_{ij}^T-{\bf e}_{i}{\bf e}_{i}^T\geq0
\label{eqn:CovSpreadPSD}
\end{align}
because $e_{ij^*}\equiv0$. Therefore $E_i$ is invertible $\forall\ \beta_{0,i}<1$.

The only other case where $E_i$ is not invertible is when $\beta_{0,i}=1$, or when there is full confidence that a particular object is not observed by any measurement. This case is evaluated after the gain matrix is derived.

The goal of the choosing $K_i$ is to minimize the uncertainty of the object state. With a Kalman filter, this goal is accomplished by minimizing $\tr{P^+_i}$ with respect to $K_i$ such that the total number of tracks $t=1$; the same approach may be applied to the JPDAF, except the a posteriori covariance equation has changed, so $K_i$ changes as well ($t=1$ is not guaranteed):
\begin{align}
0&=\frac{\partial\tr{P^+_i}}{\partial K_i}
\\
0&=-2(1-\beta_{0,i})\frac{\partial\tr{K_iH_iP^-_i}}{\partial K_i}+\frac{\partial\tr{K_iE_iK_i^T}}{\partial K_i}
\\
0&=-2(1-\beta_{0,i})P^-_iH_i^T+2K_iE_i\label{eqn:Uncertainty0}
\\
K_iE_i&=(1-\beta_{0,i})P^-_iH_i^T
\\
K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1}\label{eqn:GainK}
\end{align}
such that \refeqn{GainK} may be substituted into \refeqn{CovExpanded}
\begin{align}
\begin{split}
P^+_{i}=&P^-_{i}-(1-\beta_{0,i})^2(P^-_iH_i^TE_i^{-1}H_iP^-_i+P^-_iH_i^TE_i^{-1}H_iP^-_i)
\\
&+(1-\beta_{0,i})^2P^-_iH_i^TE_i^{-1}E_iE_i^{-1}H_iP^-_i
\end{split}
\\
P^+_{i}=&P^-_{i}-(1-\beta_{0,i})^2P^-_iH_i^TE_i^{-1}H_iP^-_i
\end{align}
which may be simplified in two ways
\begin{align}
P^+_{i}=&P^-_{i}-K_iE_iK_i^T,\label{eqn:JPDAFPostCovSymmetricUpdate}
\\
P^+_{i}=&\left(I-(1-\beta_{0,i})K_iH_i\right)P^-_i.\label{eqn:JPDAFPostCovSimpleUpdate}
\end{align}

Consider the singularity when $E_i=0$ if no association is made to a track, or where $(1-\beta_{0,i})=\sum\limits_{j}\beta_j=0$.
While it is clear from \refeqn{KalEst} and \refeqn{CovExpanded} that the updates are independent of $K_i$ at this singularity, we show that $K_i$ is not growing without bound nearby the singularity.
Hence, we evaluate the limit at this singularity to ensure it is not infinite.
Note that if $\sum\limits_{j}\beta_j\rightarrow0\Rightarrow \beta_j\rightarrow0\ \forall j$ because of the bounds $\sum\limits_{j}\beta_j\geq\beta_j\geq0\ \forall j$. For clarity, let $\beta^*=\sum\limits_{j}\beta_j$ and let $\tilde\beta_p=\sum\limits_{j,j\neq p}\beta_j/\sum\limits_{j}\beta_j$ and we drop the subscript $i$ because this analysis focuses on a single track. First we evaluate the summations inside $E$, then consider the effect on $K$:
\begin{align}
\begin{split}
&\sum\limits_{j} \beta_{j}e_{j}e_{j}^T=(\beta_1e_1e_1^T+\beta_2e_2e_2^T+...)
\\
&=(\beta_1+\beta_2+...)
(e_1e_1^T-\tilde\beta_1e_1e_1^T+
e_2e_2^T-\tilde\beta_2e_2e_2^T+...)
\\
&=\beta^*(\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T)
\end{split}
\\
\begin{split}
&{{\bf{e}}}{\bf{e}}^T=\sum\limits_{j} \beta_{j}e_{j}\sum\limits_{j} \beta_{j}e_{j}^T=
(\beta_1e_1\sum\limits_{j} \beta_{j}e_{j}^T+\beta_2e_2\sum\limits_{j} \beta_{j}e_{j}^T+...)
\\
&=(\beta_1+\beta_2+...)(e_1\sum\limits_{j} \beta_{j}e_{j}^T-\tilde\beta_1e_1\sum\limits_{j} \beta_{j}e_{j}^T+
e_2\sum\limits_{j} \beta_{j}e_{j}^T-\tilde\beta_2e_2\sum\limits_{j} \beta_{j}e_{j}^T+...)
\\
&=\beta^*(\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T)
\end{split}
\\
\begin{split}
&\lim_{\beta^* \to 0}K=\lim_{\beta^* \to 0}\beta^*P^-H^T
\left(\beta^*[S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T-\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T+\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T]\right)^{-1}
\\
&=P^-H^T\lim_{\beta^* \to 0}
\left(S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T-\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T+\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T\right)^{-1}
\label{eqn:OptGainNoSingularity}
\end{split}
\end{align}
which not an infinite limit because of the following bounded conditions:
\begin{align}
\begin{split}
&\lim_{\beta^* \to 0}\tilde\beta_p=\lim_{\beta^* \to 0}\frac{\beta_1+\beta_2+...+\beta_{p-1}+0+\beta_{p+1}+...}{\beta_1+\beta_2+...+\beta_{p-1}+\beta_p+\beta_{p+1}+...}\ \forall p\in j
\\
&=1-\lim_{\beta^* \to 0}\frac{\beta_p}{\beta^*},\quad 0\leq\beta_p\leq\beta^*\leq1\Rightarrow0\leq\lim_{\beta^* \to 0}\frac{\beta_p}{\beta^*}\leq1\Rightarrow0\leq\lim_{\beta^* \to 0}\tilde\beta_p\leq1.
\end{split}
\end{align}
and therefore \refeqn{OptGainNoSingularity} may be solved
\begin{align}
\lim_{\beta^* \to 0}K&=P^-H^T
\left(S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}{B_j}e_{j}e_{j}^T\right)^{-1}
\\
&=P^-H^T
\left(S+\sum\limits_{j}(1-B_j)e_{j}e_{j}^T\right)^{-1}
\label{eqn:OptGainNoSingularitySolved}
\end{align}
for $0\leq B_j\leq1\ \forall j$ and it can be easily seen that the sum of the terms inside the parentheses are invertible.
% and we know that $S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}{B_j}e_{j}e_{j}^T$ is invertible because $S$ is positive definite and the remaining terms are a simplified form of \refeqn{CovSpreadPSD}.
From a computational standpoint, for every $\beta^*_i\approx0$, the effect of the measurement update should be negligible, so we may choose $K_i=0_{n\times m}$ near this singularity as it will have little effect on the filter performance. Choosing $K_i$ this way results in no difference between the a priori and a posteriori update results (no measurement update where no measurement is observed). 
%\begin{align}
%\begin{split}
%\lim_{\beta^* \to 0}K&=\lim_{\beta^* \to 0}P^-H^T
%\left(S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T-\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T+\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T\right)^{-1}
%\\
%&asfd
%\end{split}
%\end{align}

We show that the Kalman gain $K_{Kalman}$ and the optimal JPDAF gain $K_{O-JPDAF}$ are not always the same with a proof by contradiction. Consider that these gains are equivalent (object subscript $i$ is omitted for clarity):
\begin{align}
&K_{Kalman}=K_{O-JPDAF}
\\
&P^-H^TS^{-1}=(1-\beta_{0})P^-H^TE^{-1}
\\
&P^-H^TS^{-1}E=(1-\beta_{0})P^-H^T
\\
&P^-H^TS^{-1}\left((1-\beta_{0})S+{\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T\right)=(1-\beta_{0})P^-H^T
\\
&(1-\beta_{0})P^-H^T+P^-H^TS^{-1}\left({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T\right)=(1-\beta_{0})P^-H^T
\\
&K_{Kalman}\left({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T\right)=0.\label{eqn:ContradictionResult}
\end{align}

The final statement implies that the spread in the innovations $\tilde P\equiv0$, which is not assumed in the JPDAF. The result of \refeqn{ContradictionResult} may only occur if $K_{Kalman}=0$, the probabilities that each measurement $j$ belongs to each object $i$ is exactly $0$ or $1$, or $K_{Kalman}\perp\left({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T\right)$ . None of these conditions are guaranteed, so it follows that $K_{O-JPDAF}=K_{Kalman}$ is not guaranteed.

While many of the equations of the O-JPDAF and JPDAF are identical, the changes from the JPDAF are shown in \refeqn{GainK} and \refeqn{JPDAFPostCovSymmetricUpdate} or \refeqn{JPDAFPostCovSimpleUpdate}. Because these changes yield an update to the states that minimize a posteriori uncertainty specifically suited for the JPDAF, we refer to this technique as the optimal JPDAF, or O-JPDAF for short.

\subsection{Numerical Example}
We consider an example where four satellites in LEO are in the field of view of a ground sensor. The sensor measures range (km), range rate (km/s), and two angles (rad) according to local right-ascension and declination convensions with a $10\%$ chance that an object is not measured or an additional false alarm comes into the field of view. The conventional JPDAF and the O-JPDAF algorithms are compared with identical initial condions and measurements.

\begin{figure}
\centerline{
	\subfigure[\;JPDAF Radii]
		{\includegraphics[width=0.45\columnwidth]{JPDAF_Radii_Tracks1234}}
	\subfigure[\;JPDAF Uncertainty]
		{\includegraphics[width=0.45\columnwidth]{JPDAF_UncertaintyCosts_Tracks1234}}
	}
\centerline{
	\subfigure[\;O-JPDAF Radii]
		{\includegraphics[width=0.45\columnwidth]{O-JPDAF_Radii_Tracks1234}}
	\subfigure[\;O-JPDAF Uncertainty]
		{\includegraphics[width=0.45\columnwidth]{O-JPDAF_UncertaintyCosts_Tracks1234}}
	}
\centerline{
	\subfigure[\;Comparison of Radii of the First Two Tracks]
		{\includegraphics[width=0.45\columnwidth]{ComparisonRadii_Tracks12}}
	\subfigure[\;Comparison of Uncertainties of the First Two Tracks]
		{\includegraphics[width=0.45\columnwidth]{ComparisonUncertaintyCosts_Tracks12}}
	}
\caption{The orbital radii and costs are compared among the algorithms to show that their is a small improvement in the mean of the estimates with the O-JPDAF and large improvement in the uncertainty in the estimates. Addtitionally, both algorithms handled an uncertain number of measurements and false alarms in these simulations.
}\label{fig:hoverRegions}
\end{figure}

%\\
%P^+_{i}=&P^-_{i}+K_i\{(1-\beta_{0,i})S_i+({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e_{ij}}}}{\bf{e_{ij}}}^T)\}K_{i}^T
%\\
%P^+_{i}=&P^-_{i}+K_iS^*K_i^T, \quad S^*=(1-\beta_{0,i})S_i+({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e_{ij}}}}{\bf{e_{ij}}}^T)

%
%
\section{Coalescence Avoiding Optimal Joint Probabilistic Data Association Filter (C-JPDAF)}% Or another name that you choose
\label{C-JPDAF}

The O-JPDAF described in the previous section is effective in handling clutter or missed detection of multiple objects. However, it is prone to coalescing estimates, especially with objects in close-proximity. Here, we present a coalescence-avoiding optimal JPDAF, hereafter C-JPDAF for short, to avoid coalescence of the estimates. The main idea is choosing the estimator gain such that a weighted sum of the size of uncertainty and a measure of similarity is minimized.

A similarity measure is introduced first, and an optimization problem is formulated. Throughout this section, the subscript $k$ denoting the time step is omitted for simplicity.

\subsection{Matusita's Similarity Measure}

Coalescence of neighboring estimates only occurs when these estimates are very similar. We evaluate the Matusita's measure of similarity because it takes the form of a differentiable analytic index. For two Gaussian probability density functions, $f_1(x)$ and $f_2(x)$, Matusita's measure is defined as
\begin{align*}
\zeta = \int \sqrt{f_1(x)f_2(x)}dx,
\end{align*}
which is between 0 and 1, and it measures the similarity between two densities ~\cite{Coal_k}.

From~\cite{Coal_k}, using the assumption that the a posteriori states $\hat x_\mathcal{A}^+$ and $\hat x_\mathcal{B}^+$ are Gaussian, the Matusita's measure between two objects can be written as
\begin{align}
\zeta=\zeta_1\zeta_2,
\end{align}
where
\begin{align}
\label{Q}
\zeta_1=&\frac{2^\frac{n}{2}|P_\mathcal{A}^+|^{\frac{1}{4}}|P_\mathcal{B}^+|^{\frac{1}{4}}}{|P_\mathcal{A}^++P_\mathcal{B}^+|^{\frac{1}{2}}},
\end{align}
\begin{align}
\label{R}
\zeta_2=&\exp{\left\{-\frac{1}{4}(\hat x^+_\mathcal{A}-\hat x^+_\mathcal{B})^T(P_\mathcal{A}^++P_\mathcal{B}^+)^{-1}(\hat x^+_\mathcal{A}-\hat x^+_\mathcal{B})\right\}}.
\end{align}
Here $\zeta_1$ measures the differences in the covariances and $\zeta_2$ represents the spread in the means.

%Coalescence is well known to occur when estimates share measurements over long periods of time; estimates with largely different state covariances do not tend to coalesce
%; consider two satellites $1$ and $2$ such that $S_1>>S_2$. Inside the measurement space, if a measurement falls nearby $\hat z_1$, then estimate with low measurement uncertainty...
	% Introduce the definition of M-measure
	% Introduce the definition of Bhattacharyya distance

\subsection{Optimal Coalescence Avoidance}

Define a cost function of coalescence avoiding filtering as

\begin{align}
J=J_P+cJ_S
\end{align}
where $J_P$ is an uncertainty cost that measures the confidence level of estimation and $J_S$ is a similarity cost that measures degree of coalescence. The constant $c$ is a positive weighting factor between $J_P$ and $J_S$.

In the proposed C-JPDAF, the a posteriori estimate of the state is given in the form of O-JPDA given by \refeqn{GainK}, but the estimator gain $K_i$ is selected to minimize the total cost $J$. In short, the estimator gain is chosen in the consideration of both uncertainty and coalescence. These are illustrated at Figure \ref{fig:CostTrends}.

\begin{figure}
\setlength{\unitlength}{0.1\columnwidth}
\centerline{\footnotesize\selectfont
\begin{picture}(8.5,7.3)(-0.5,-0.8)
\put(-0.5,3){\rotatebox{90}{Cost}}
\put(0,0){\includegraphics[width=0.8\columnwidth]{ECC14_fig1.pdf}}
\put(6.2,0.2){\shortstack[c]{Uncertainty\\ cost: $J_P$}}
\put(6.2,2.8){\shortstack[c]{Similarity\\ cost: $J_S$}}
\put(5.7,5.2){\shortstack[c]{Weighted sum:\\ $J=J_P+cJ_S$}}
\put(4.98,0.78){\circle*{0.1}}
\put(2.48,1.57){\circle*{0.1}}
\put(5.0,-0.3){$K_{O-JPDAF}$}
\put(2.5,-0.3){$K_{C-JPDAF}$}
\put(3.1,-0.8){Estimator gain}
\end{picture}
}
\caption{Illustration of optimal coalescence avoidance: several costs are illustrated with respect to an element of an estimator gain $K$ (red: uncertainty cost $J_P$, blue: similarity cost $J_S$, black: weighted sum). The estimator gain $K_{O-JPDAF}$ of O-JPDAF is selected to minimize the uncertainty cost. In the proposed approach, the estimator gain $K_{C-JPDAF}$ is chosen such that the weighted sum of the uncertainty cost and the similarity cost is minimized. In short, at the expense of increased uncertainty at a single step, the proposed approach prevents coalescence. The increased uncertainty at each step is rewarded by reducing the overall estimation error drastically by maintaining correct data association throughout crossing tracks in close proximity.}
\label{fig:CostTrends}
\end{figure}

More explicitly, the uncertainty cost $J_P$ is defined as the sum of the trace of each a posteriori covariance:
\begin{align}
J_P=\sum\limits_{i}^{t}\tr{P^+_i},\label{eqn:JP}
\end{align}
where $P^+_{i}$ is obtained by \refeqn{JPDAFCov}. The similarity cost is chosen as the second part of the Matusita's measure that represents the weighted-spread of the mean values:
\begin{align}
J_S&=\sum\limits_{\mathcal{A}}^{t}\sum\limits_{\mathcal{B}}^{t}\exp (-au_{\mathcal{A}\mathcal{B}})\bigg|_{\mathcal{A}\neq\mathcal{B}}\\
u_{\mathcal{A}\mathcal{B}} & = \begin{cases} (\hat x_{\mathcal{A}}^+-\hat x^+_{\mathcal{B}})^T(P^+_{\mathcal{A}}+P^+_{\mathcal{B}})^{-1}(\hat x^+_{\mathcal{A}}-\hat x^+_{\mathcal{B}})\label{eqn:u}, & \mbox{if } \zeta_2^->\zeta_{2,min} \\ 0 & \mbox{otherwise}  \end{cases}
\end{align}
%%%% \norm{\hat x_{\mathcal{A}}^+-\hat x^+_{\mathcal{B}}}
% OR
%\begin{align}
%J_S&=\sum\limits_{\mathcal{A}}^{t}\sum\limits_{\mathcal{B}}^{t}\exp (-au_{\mathcal{A}\mathcal{B}})\bigg|_{\mathcal{A}\neq\mathcal{B}}\\
%u_{\mathcal{A}\mathcal{B}} & = (\hat x_{\mathcal{A}}^+-\hat x^+_{\mathcal{B}})^T(P^+_{\mathcal{A}}+P^+_{\mathcal{B}})^{-1}(\hat x^+_{\mathcal{A}}-\hat x^+_{\mathcal{B}})
%\end{align}
where $\zeta_2^-$ is calculated using $\hat x^-_{\mathcal A}$ and $\hat x^-_{\mathcal B}$ as the means and $P^-_\mathcal{A}$ and $P^-_\mathcal{B}$ as the covariances, $\zeta_{2,min}$ defines a region of separation, and positive constant $a$ serves to amplify or attenuate the function as well as alter its shape.

Necessary conditions for optimality are given by
\begin{align}
\deriv{J}{K_i} = \deriv{J_P}{K_i} + c \deriv{J_S}{K_i} =0,\label{eqn:NCO}
\end{align}
where $\deriv{J_S}{K_i}$ may be nonzero. These are solved for the set of all gains $\mathcal{K}=\{K_1,K_2,...,K_t\}$.

Next, we find the derivative of each cost. The derivative of $J_P$ is taken from \refeqn{Uncertainty0}
\begin{align}
\label{eqn:CostP}
\frac{\partial J_{P}}{\partial K_{i}}&=2K_iE_i-2(1-\beta_{0,i})P^-_iH_i^T
\end{align}
Note that $J_{P_\mathcal{X}}$ does not depend on $K_\mathcal{Y}$ for $\mathcal{X}\neq \mathcal{Y}$.

%
%To appropriately define the similarity cost, we only consider the $R$ term from Equation \ref{R} because coalescence is well known to occur when estimates are within close proximity. For simplicity, we define ${P^+_{12}}^{-1}$ as
%\begin{align}
%{P^+_{12}}^{-1}=(P_1^++P_2^+)^{-1}
%\end{align}
%and define define $u$ as
We expand the nonzero portion of \refeqn{u} so it may be differentiated with respect to $K_\mathcal{A}$,
\begin{align}
\begin{split}
u=&(\hat x^+_\mathcal{A}-\hat x^+_\mathcal{B})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^+_\mathcal{A}-\hat x^+_\mathcal{B})
\\
=&\hat x^{+T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^+_\mathcal{A}
-2\hat x^{+T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^+_\mathcal{A}
+\hat x^{+T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^+_\mathcal{B}
\\
=&(\hat x^-_\mathcal{A}+K_\mathcal{A}{\bf e}_\mathcal{A})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^-_\mathcal{A}+K_\mathcal{A}{\bf e}_\mathcal{A})%
-2(\hat x^-_\mathcal{B}+K_\mathcal{B}{\bf e}_\mathcal{B})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^-_\mathcal{A}+K_\mathcal{A}{\bf e}_\mathcal{A})
\\
&+(\hat x^-_\mathcal{B}+K_\mathcal{B}{\bf e}_\mathcal{B})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^-_\mathcal{B}+K_\mathcal{B}{\bf e}_\mathcal{B})
\\
=&\hat x^{-T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^-_\mathcal{A}
+2\hat x^{-T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
+{\bf e}_\mathcal{A}^TK_\mathcal{A}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
\\
&+\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^-_\mathcal{B}
+2\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B}
+{\bf e}_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B}
\\
&-2\{\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^-_\mathcal{A}
+\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
+\hat x^{-T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B}
+{\bf e}_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}\}
\\
=&\hat x_\mathcal{A}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{A}^-
-2\hat x_\mathcal{A}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{B}^-
\\
&+2(\hat x_\mathcal{A}^{-}-\hat x_\mathcal{B}^{-})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
+\hat x_\mathcal{B}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{B}^-
\\
&+2(\hat x_\mathcal{B}^{-}-\hat x_\mathcal{A}^{-})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B}
+{\bf e}_\mathcal{A}^TK_\mathcal{A}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
\\
&-2{\bf e}_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
+{\bf e}_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B},
\label{eqn:uexpanded}
\end{split}
\end{align}
where ${P^+_{\mathcal{A}\mathcal{B}}}^{-1}=(P_\mathcal{A}^++P_\mathcal{B}^+)^{-1}$.

%which is the most reduced form to differentiate with respect to $K_1$ and $K_2$.

\newcommand{\Iij}{\ensuremath{\mathbf{1}_{ij}}}

Let $\Iij\in \Re^{n\times m}$ be defined such that its $i,j$-th element is one and other elements are zero. The derivative of $(P_{\mathcal{A}\mathcal{B}}^+)^{-1}$ with respect to the $i,j$-th element of $K_\mathcal{A}$ is defined as $\mathcal{P}_{\mathcal{A}\mathcal{B}_{ij}}\in\Re^{n\times n}$:
%such that $ij1(i,j)=1$ and $0$ elsewhere, we define $P^*_{12}$ as
\begin{align}
\begin{split}
\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}&=\frac{\partial {P^+_{\mathcal{A}\mathcal{B}}}^{-1}}{\partial K_{\mathcal{A},ij}}
\\
&=-{P_{\mathcal{A}\mathcal{B}}^+}^{-1}\frac{\partial}{\partial K_{\mathcal{A},ij}}
\left(
P^-_\mathcal{A}-(1-\beta_{0,\mathcal{A}})(K_\mathcal{A}H_\mathcal{A}P^-_\mathcal{A}+P^-_\mathcal{A}H_\mathcal{A}^TK_\mathcal{A}^T)+K_\mathcal{A}E_\mathcal{A}K_\mathcal{A}^T
\right)
{P_{\mathcal{A}\mathcal{B}}^+}^{-1}
\\
&={P_{\mathcal{A}\mathcal{B}}^+}^{-1}[(1-\beta_{0,\mathcal{A}})(\Iij H_\mathcal{A}P_\mathcal{A}^-+P_\mathcal{A}^-H_\mathcal{A}^T\Iij^T)-\Iij E_\mathcal{A}K_\mathcal{A}^T -K_\mathcal{A}E_\mathcal{A}\Iij^T]{P_{\mathcal{A}\mathcal{B}}^+}^{-1}.
\end{split}
\end{align}
Using this, the derivative of $u_{\mathcal{A}\mathcal{B}}$ with respect to the $i,j$-th element of $K_\mathcal{A}$ can be written as
\begin{align}
\begin{split}
\frac{\partial u_{\mathcal{A}\mathcal{B}}}{\partial K_{\mathcal{A},ij}}
=&\frac{\partial}{\partial K_{\mathcal{A},ij}}\{\hat x_\mathcal{A}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{A}^-
-2\hat x_\mathcal{A}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{B}^-
\\
&+2(\hat x_\mathcal{A}^{-}-\hat x_\mathcal{B}^{-})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
+\hat x_\mathcal{B}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{B}^-
\\
&+2(\hat x_\mathcal{B}^{-}-\hat x_\mathcal{A}^{-})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B}
+{\bf e}_\mathcal{A}^TK_\mathcal{A}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
\\
&-2{\bf e}_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
+{\bf e}_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B}\}
\\
=& \hat x_\mathcal{A}^{-T}\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}\hat x_\mathcal{A}^-
-2\hat x_\mathcal{A}^{-T}\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}\hat x_\mathcal{B}^-
\\
&+2(\hat x_\mathcal{A}^{-}-\hat x_\mathcal{B}^{-})^T
[\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}K_\mathcal{A}+{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\Iij]
{\bf e}_\mathcal{A}
\\
&+\hat x_\mathcal{B}^{-T}\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}\hat x_\mathcal{B}^-
+2(\hat x_\mathcal{B}^{-}-\hat x_\mathcal{A}^{-})^T\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}K_\mathcal{B}{\bf e}_\mathcal{B}
\\
&+{\bf e}_\mathcal{A}^T[\Iij^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}
+K_\mathcal{A}^T\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}K_\mathcal{A}
\\
&+K_\mathcal{A}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\Iij]{\bf e}_\mathcal{A}
-2{\bf e}_\mathcal{B}^TK_\mathcal{B}^T[\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}K_\mathcal{A}
\\
&+{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\Iij]{\bf e}_\mathcal{A}
+{\bf e}_\mathcal{B}^TK_\mathcal{B}^T\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}K_\mathcal{B}{\bf e}_\mathcal{B}.
\end{split}
\end{align}
From this, the derivation of the similarity cost $J_S$ with respect to the $i,j$-th element of $K_\mathcal{A}$ is given by
\begin{align}
\frac{\partial J_{S}}{\partial K_{\mathcal{A},ij}}=-a\exp(-au_{\mathcal{A}\mathcal{B}})\frac{\partial u_{\mathcal{A}\mathcal{B}}}{\partial K_{\mathcal{A},ij}}.\label{eqn:JSK}
\end{align}

In summary, \refeqn{CostP} and \refeqn{JSK} are substituted into \refeqn{NCO}, and it is solved to obtain the set of optimal gains $\mathcal{K}$. Then, the algorithm of C-JPDAF is completed by updating the state estimate and covariance by using \refeqn{KalEst} and \refeqn{JPDAFCov}, respectively.

As the optimality condition is a nonlinear function of estimator gains, there is no closed-form solution. Hence, we require a nonlinear equation solver. Note that ${P_\mathcal{A}^+}^{-1}$, ${P_\mathcal{B}^+}^{-1}$, and ${P_{\mathcal{A}\mathcal{B}}^+}^{-1}$ are symmetric inverses and must only be calculated once for each $(\mathcal{A},\mathcal{B})$ where the similarity measure is nontrivial. When $c=0$, we can show that the optimal gain reduces to that of the O-JPDAF, which is used as an initial guess of numerical solutions.

%\subsection{Suboptimal Solution}
%Because optimal solution of \ref{eqn:JSK} does not have a closed-form solution, a close initial guess to the actual solution can decrease the computational requirements of the problem. Therefore for objects $1$ and $2$, we can minimize the two equations:
%\begin{align}
%\frac{\partial J}{\partial K_1}
%&=\frac{\partial J_{P}}{\partial K_1}+c\frac{\partial J_{S}}{\partial K_1}
%\approx2\beta_1({K_1S_1-P_1^-H_1^T})-ac\exp(-au^-)\frac{\partial u^*}{\partial K_1}
%\label{eqn:subopt1}\\
%\frac{\partial J}{\partial K_2}
%&=\frac{\partial J_{P}}{\partial K_2}+c\frac{\partial J_{S}}{\partial K_2}
%\approx2\beta_2({K_2S_2-P_2^-H_2^T})-ac\exp(-au^-)\frac{\partial u^*}{\partial K_2}
%\label{eqn:subopt2}
%\end{align}
%where
%\begin{align}
%u^-&= (\hat x^-_1-\hat x^-_2)^T(P^-_{1}+P^-_{2})^{-1}(\hat x^-_1-\hat x^-_2)\label{eqn:uminus}\\
%u^*&= (\hat x^+_1-\hat x^+_2)^T(P^-_{1}+P^-_{2})^{-1}(\hat x^+_1-\hat x^+_2)\label{eqn:uminus}
%\end{align}
%so that we define a new variable $\Omega=ac\exp(-au^-)$, which may be treated as constant in \ref{eqn:subopt1} and \ref{eqn:subopt2}. We modify \ref{eqn:uexpanded} to solve for $u^*$,
%\begin{align}
%\label{eqn:ustarexpanded}
%u^*=&\{\hat x_1^{-T}{P^-_{12}}^{-1}\hat x_1^-
%-2\hat x_1^{-T}{P^-_{12}}^{-1}\hat x_2^-\nonumber\\
%&+2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T{P^-_{12}}^{-1}K_1e_1
%+\hat x_2^{-T}{P^-_{12}}^{-1}\hat x_2^-\nonumber\\
%&+2\beta_2(\hat x_2^{-}-\hat x_1^{-})^T{P^-_{12}}^{-1}K_2e_2
%+\beta_1^2e_1^TK_1^T{P^-_{12}}^{-1}K_1e_1\nonumber\\
%&-2\beta_1\beta_2e_2^TK_2^T{P^-_{12}}^{-1}K_1e_1
%+\beta_2^2e_2^TK_2^T{P^-_{12}}^{-1}K_2e_2\}
%\end{align}
%where ${P^-_{12}}^{-1}=(P_1^-+P_2^-)^{-1}$, which is fixed with respect to $K_1$ and $K_2$. We take the derivative of \ref{eqn:ustarexpanded} with respect to $K_1$ and $K_2$
%\begin{align}
%\begin{split}
%\frac{\partial u^*}{\partial K_1}=&\frac{\partial}{\partial K_1}\left(2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T{P^-_{12}}^{-1}K_1e_1
%+\beta_1^2e_1^TK_1^T{P^-_{12}}^{-1}K_1e_1
%-2\beta_1\beta_2e_2^TK_2^T{P^-_{12}}^{-1}K_1e_1\right)\\
%=&2\beta_1{P^-_{12}}^{-1}(\hat x_1^{-}-\hat x_2^{-})e_1^T
%+2\beta_1^2{P^-_{12}}^{-1}K_1e_1e_1^T
%-2\beta_1\beta_2{P^-_{12}}^{-1}K_2e_2e_1^T.
%\label{eqn:ustarderivexpanded}
%\end{split}
%\end{align}
%Let $\lambda$ be an eigenvalue of ${P^-_{12}}^{-1}$ and $q$ be the associated eigenvector, i.e. ${P^-_{12}}^{-1}q=\lambda q\Rightarrow q^T{P^-_{12}}^{-1}=\lambda q^T$ because ${P^-_{12}}^{-1}$ is symmetric.
%Then the derviative of \ref{eqn:ustarexpanded} with resepect to $K_1$ and $K_2$ and setting this to zero yields
%\begin{align}
%0&=2\beta_1({K_1S_1-P_1^-H_1^T})-\Omega\frac{\partial u^*}{\partial K_1}\\
%0&=2\beta_1({K_1S_1-P_1^-H_1^T})-2\beta_1\Omega\left(
%{P^-_{12}}^{-1}(\hat x_1^{-}-\hat x_2^{-})e_1^T
%+\beta_1{P^-_{12}}^{-1}K_1e_1e_1^T
%-\beta_2{P^-_{12}}^{-1}K_2e_2e_1^T\right)\\
%0&=K_1S_1-P_1^-H_1^T
%-\Omega{P^-_{12}}^{-1}(\hat x_1^{-}-\hat x_2^{-})e_1^T
%-\Omega\beta_1{P^-_{12}}^{-1}K_1e_1e_1^T
%+\Omega\beta_2{P^-_{12}}^{-1}K_2e_2e_1^T\\
%0&=q^TK_1S_1-q^TP_1^-H_1^T
%-\Omega q^T{P^-_{12}}^{-1}(\hat x_1^{-}-\hat x_2^{-})e_1^T
%-\Omega\beta_1q^T{P^-_{12}}^{-1}K_1e_1e_1^T
%+\Omega\beta_2q^T{P^-_{12}}^{-1}K_2e_2e_1^T\\
%0&=q^TK_1S_1-q^TP_1^-H_1^T
%-\Omega q^T\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T
%-\Omega\beta_1q^T\lambda K_1e_1e_1^T
%+\Omega\beta_2q^T\lambda K_2e_2e_1^T\\
%0&=q^T\left\{K_1S_1-P_1^-H_1^T
%-\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T
%-\Omega\beta_1\lambda K_1e_1e_1^T
%+\Omega\beta_2\lambda K_2e_2e_1^T\right\},
%\quad q\neq0\\
%0&=K_1(S_1-\Omega\beta_1\lambda e_1e_1^T)
%-P_1^-H_1^T
%-\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T
%+\Omega\beta_2\lambda K_2e_2e_1^T\\
%K_1&=\left(P_1^-H_1^T
%+\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T
%-\Omega\beta_2\lambda K_2e_2e_1^T\right)
%(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}
%\label{eqn:K1coupled}
%\end{align}
%and similarly
%\begin{align}
%K_2&=\left(P_2^-H_2^T
%+\Omega\lambda(\hat x_2^{-}-\hat x_1^{-})e_2^T
%-\Omega\beta_1\lambda K_1e_1e_2^T\right)
%(S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}
%\label{eqn:K2coupled}.
%\end{align}
%The equations \ref{eqn:K1coupled} and \ref{eqn:K2coupled} may be substituted into each other to seclude $K_1$ and $K_2$
%\begin{align}
%\begin{split}
%K_1&=(P_1^-H_1^T
%+\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T
%-\Omega\beta_2\lambda[P_2^-H_2^T
%+\Omega\lambda(\hat x_2^{-}-\hat x_1^{-})e_2^T\\
%&-\Omega\beta_1\lambda K_1e_1e_2^T]
%(S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}e_2e_1^T)
%(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}
%\end{split}
%\end{align}
%which may be rearranged to
%\begin{align}
%\begin{split}
%&K_1\left(I-\Omega^2\lambda^2\beta_1\beta_2e_1e_2^T(S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}e_2e_1^T
%(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}\right)=(P_1^-H_1^T
%+\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T\\
%&-\Omega\beta_2\lambda[P_2^-H_2^T+\Omega\lambda(\hat x_2^{-}-\hat x_1^{-})e_2^T](S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}e_2e_1^T)
%(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}.
%\end{split}
%\end{align}
%Solving for $K_1$,
%\begin{align}
%\begin{split}
%K_1=&(P_1^-H_1^T
%+\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T
%-\Omega\beta_2\lambda[P_2^-H_2^T+\Omega\lambda(\hat x_2^{-}-\hat x_1^{-})e_2^T]
%(S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}
%e_2e_1^T)
%(S_1-\\
%&\Omega\beta_1\lambda e_1e_1^T)^{-1}\left(I-\Omega^2\lambda^2\beta_1\beta_2e_1e_2^T(S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}e_2e_1^T
%(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}\right)^{-1}
%\end{split}
%\end{align}
%and similarly
%\begin{align}
%\begin{split}
%K_2=&(P_2^-H_2^T
%+\Omega\lambda(\hat x_2^{-}-\hat x_1^{-})e_2^T
%-\Omega\beta_1\lambda[P_1^-H_1^T+\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T]
%(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}
%e_1e_2^T)
%(S_2-\\
%&\Omega\beta_2\lambda e_2e_2^T)^{-1}\left(I-\Omega^2\lambda^2\beta_2\beta_1e_2e_1^T(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}e_1e_2^T
%(S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}\right)^{-1}.
%\end{split}
%\end{align}

\bibliography{BibMaster}
\bibliographystyle{IEEEtran}

\end{document}

%\begin{align}
%\begin{split}
%u=&(\hat x^+_\mathcal{A}-\hat x^+_\mathcal{B})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^+_\mathcal{A}-\hat x^+_\mathcal{B})
%\end{split}
%\\
%\begin{split}
%=&\hat x^{+T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^+_\mathcal{A}
%-2\hat x^{+T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^+_\mathcal{A}
%+\hat x^{+T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^+_\mathcal{B}
%\end{split}
%\\
%\begin{split}
%=&(\hat x^-_\mathcal{A}+K_\mathcal{A}{\bf e}_\mathcal{A})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^-_\mathcal{A}+K_\mathcal{A}{\bf e}_\mathcal{A})%
%-2(\hat x^-_\mathcal{B}+\beta_\mathcal{B}K_\mathcal{B}e_\mathcal{B})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^-_\mathcal{A}+K_\mathcal{A}{\bf e}_\mathcal{A})
%\\
%&+(\hat x^-_\mathcal{B}+\beta_\mathcal{B}K_\mathcal{B}e_\mathcal{B})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^-_\mathcal{B}+\beta_\mathcal{B}K_\mathcal{B}e_\mathcal{B})
%\end{split}
%\\
%\begin{split}
%=&\hat x^{-T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^-_\mathcal{A}+2\hat x^{-T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\beta_\mathcal{A}K_\mathcal{A}e_\mathcal{A}
%+e_\mathcal{A}^TK_\mathcal{A}^T\beta_\mathcal{A}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
%\\
%&+\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^-_\mathcal{B}+2\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\beta_\mathcal{B}K_\mathcal{B}e_\mathcal{B}+e_\mathcal{B}^TK_\mathcal{B}^T\beta_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\beta_\mathcal{B}K_\mathcal{B}e_\mathcal{B}
%\\
%&-2\{\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^-_\mathcal{A}+\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\beta_\mathcal{A}K_\mathcal{A}e_\mathcal{A}+\hat x^{-T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\beta_\mathcal{B}K_\mathcal{B}e_\mathcal{B}+e_\mathcal{B}^TK_\mathcal{B}^T\beta_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\beta_\mathcal{A}K_\mathcal{A}e_\mathcal{A}\}
%\end{split}
%\\
%\begin{split}
%=&\{\hat x_\mathcal{A}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{A}^-
%-2\hat x_\mathcal{A}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{B}^-
%\\
%&+2\beta_\mathcal{B}(\hat x_\mathcal{A}^{-}-\hat x_\mathcal{B}^{-})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}e_\mathcal{A}
%+\hat x_\mathcal{B}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{B}^-
%\\
%&+2\beta_\mathcal{B}(\hat x_\mathcal{B}^{-}-\hat x_\mathcal{A}^{-})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}e_\mathcal{B}
%+\beta_\mathcal{A}^2e_\mathcal{A}^TK_\mathcal{A}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}e_\mathcal{A}
%\\
%&-2\beta_\mathcal{A}\beta_\mathcal{B}e_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}e_\mathcal{A}
%+\beta_\mathcal{B}^2e_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}e_\mathcal{B}\},
%\label{eqn:uexpanded}
%\end{split}
%\end{align}

### 2018-06-09 Conference/AAS 2015 DA/Presentation/AAS15_DataAssociation.tex
added Conference folder

\documentclass[hyperref={pdftex,pdfpagemode=none,pdfstartview=FitH},10pt]{beamer}
\usepackage[english]{babel}
\usepackage{times}
%\usepackage{mathptmx,helvet}
\usepackage{amsmath,amssymb}
\usepackage{subfigure,graphicx,tabularx,booktabs,colortbl}
\usepackage{amssymb,amsmath,amsthm,times,graphicx,subfigure,tabularx,booktabs,colortbl,multirow,threeparttable,graphics}
\usepackage{movie15}
\usepackage[labelformat=empty]{caption}
\usepackage{color}
\newcommand{\hilight}[1]{\colorbox{green}{#1}}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mbox{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mbox{tr}\ensuremath{\bracket{#1}}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO}(3)}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\SE}{\ensuremath{\mathrm{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\S}{\ensuremath{\mathbb{S}}}
\newcommand{\abs}{\mathop{\mathrm{abs}}\nolimits}

\renewcommand{\r}{\mathbf{r}}
\renewcommand{\u}{\mathbf{u}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\x}{\mathbf{x}}

\definecolor{RoyalBlue}{rgb}{0.25,0.41,0.88}
\def\Emph{\textcolor{RoyalBlue}}

\definecolor{mygray}{gray}{0.9}
\definecolor{tmp}{rgb}{0.804,0.941,1.0}
\setbeamercolor{numerical}{fg=black,bg=tmp}
\setbeamercolor{exact}{fg=black,bg=red}

\mode<presentation> {
  \usetheme{Warsaw}
  \usefonttheme{serif}
  \setbeamercovered{transparent}
}

\setbeamertemplate{footline}%{split theme}
{%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm plus1fill,rightskip=.3cm]{title in head/foot}%
    \usebeamerfont{title in
    head/foot}\insertframenumber/\inserttotalframenumber
  \end{beamercolorbox}}%
  \vskip0pt%
} \setbeamercolor{box}{fg=black,bg=yellow}

\title[Minimum Uncertainty JPDA and Coalescence-Avoiding Optimal JPDA Filtering]
{Minimum Uncertainty JPDA Filter and Coalescence
\\
Avoidance Performance Evaluations
}

\author{
Evan Kaufman\\
{\footnotesize\selectfont Department of Mechanical and Aerospace Engineering,\\
The George Washington University, Washington DC}
$ $\\
\vspace*{0.015\columnwidth}
T. Alan Lovell\\
{\footnotesize\selectfont Air Force Research Laboratory,\\ Kirtland AFB, Albuquerque NM}
$ $\\
\vspace*{0.015\columnwidth}
Taeyoung Lee\\
{\footnotesize\selectfont Department of Mechanical and Aerospace Engineering,\\
The George Washington University, Washington DC}
}

%\institute[GWU]{\footnotesize\selectfont
%$^*$ Air Force Research Laboratory, Kirtland AFB, Albuquerque NM\\$ $\\
%Department of Mechanical and Aerospace Engineering\\
%George Washington University, Washington DC\\$ $\\}

\date{\vspace*{-0.8cm}\\AAS/AIAA Spaceflight Mechanics Meeting 2015\vspace*{0.1cm}\\
\footnotesize\selectfont (Supported by AFRL Summer Faculty and Space Scholars Programs,\\ NSF Grants CMMI-\#1029551, \#1335008, and CNS-\#1337722)}

%	\graphicspath{{/Users/tylee@seas.gwu.edu/Documents/Seminar/ISSFD14/Figs/}}
%\graphicspath{{/Users/evankaufman/Google Drive/FDCL/Evan/Publications/AAS 2015 OD EKF/Presentation/Figs}}
\graphicspath{{Figs/}}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section*{}
\subsection*{I. Introduction}

\begin{frame}
\frametitle{Introduction}

\begin{itemize}
\item Satellite orbit estimation
	\begin{itemize}
	\item As the number of satellites orbiting around the earth increases, accurate orbital estimation becomes important
	\item Sensors may provide many measurements, which may originate from any of the objects in the sensor field of view
	\item When we receive these measurements, we must associate measurements with objects consistently and correctly to perform accurate estimation of the satellites
	\end{itemize}
\end{itemize}

\vfill
\centerline{
\includegraphics[height=2.7cm]{PhasedArrayRadar}\hspace*{0.2cm}
\includegraphics[height=2.7cm]{TelescopeTrackingSats}\hspace*{0.2cm}
\includegraphics[height=2.7cm]{SatDots}}
\centerline{\scriptsize\selectfont
\hspace*{0.6cm}Phased Array Radar\hspace*{1.7cm} Satellite Tracking Telescope\hspace*{1.0cm} Space Object Illustration}
\end{frame}

\begin{frame}
\frametitle{Introduction}

\begin{itemize}
\item Data association
	\begin{itemize}
	\item Several techniques have been applied to associate measurements with objects
	\item This research focuses on recursive data association techniques that obtain new satellite estimations based on prior estimations and current measurements, not measurements over a long time period--more computationally expensive
	\item \Emph{Hard decision} data association techniques like the nearest neighbor filter \Emph{completely associate} measurements to object estimates, even if there is a nonzero probability that this association is wrong, which can be detrimental to the estimation
	\item \Emph{Soft decision} data association techniques update the satellite tracks with measurements in a \Emph{weighted fashion} based on the probability of association, but often leads to an increase in estimate uncertainties and coalescence of the estimates
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Introduction}

\begin{itemize}
\item The Joint Probabilistic Data Association Filter (JPDAF)
	\begin{itemize}
	\item Well-known recursive data association technique
	\item The probabilities of association between measurements and objects serve as weighting factors during the measurement update portion of the filter
	\item Designed for linear systems, but easily extended to nonlinear systems if the linearized dynamics are accurate% in the time period between measurement updates
	\end{itemize}
\item The JPDAF suffers from two issues addressed in this paper:
\begin{enumerate}
\item The Kalman gain, which normally minimizes the uncertainty of estimates of individual tracks, is applied but does \Emph{not} minimize the uncertainty of the satellite states with the JPDAF
\item When measurements are shared among multiple objects, the object \Emph{estimates} tend to come together, known as \Emph{coalescence}
\end{enumerate}
\item To solve these problems, we derive the minimum uncertainty JPDAF (MUJPDAF) and the coalescence-avoiding optimal JPDAF (C-JPDAF) and perform numerous simulations to show the benefits and drawbacks of each approach
\end{itemize}

\end{frame}

\section*{}
\subsection*{II. MUJPDAF Problem Formulation}

\begin{frame}
\frametitle{Flow Update}

\begin{itemize}
\item The $i$-th linear stochastic discrete-time system is described by
\begin{align}
x_{i_{k}} & = F_{i_{k-1}} x_{i_{k-1}} + w_{i_{k}},\label{eqn:xkp}\\
z_{i_k} & = H_ix_{i_k} + v_{i_k}
\end{align}
\item Gaussian distributions for the $x_{i_0} \sim \mathcal{N}[\bar x_{i_0}, P_{i_0}]$
\item Zero-mean, mutually independent, white Gaussian noise sequences for the process noise $w_{i_k} \sim \mathcal{N}[0,Q_{i_k}]$ and measurement noise $v_{i_k} \sim \mathcal{N}[0,R_{i_k}]$
\item A priori state covariance
\begin{align}
P_{i_{k}}^- = F_{i_{k-1}} P_{i_{k-1}}^+ F_{i_{k-1}}^T + Q_{i_{k}}
\end{align}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Measurement Update}

\begin{itemize}
\item The measurement prediction of the $i$-th estimate $\hat z_i = H_i\hat x_{i}^-$ (linear) or $\hat z_i = h_i(\hat x_{i}^-)$ where $H_i=\deriv{h_i}{x_i}\bigg|_{x_i=\hat x_{i}^-}$ (nonlinear) is compared with the $j$-th measurement to obtain the innovations
\begin{align}
e_{ij} = z_j - \hat z_i\ \ \forall\ i,j
\end{align}
such that the $i$-th covariance matrix is
\begin{align}
S_{i}=H_{i}P_{i}^{-}H_{i}^T+R_{i}=\mathrm{E}[e_{i}e_i^{T}]
\end{align}
\item Measurement origination model based on the "Cheap JPDAF" to obtain the probability that the $i$-th estimate is associated with the $j$-th measurement
\begin{align}
G_{ij}&=\frac1{\sqrt{\det{S_{i}}}}\exp{\{-\frac12e_{ij}^TS_{i}^{-1}e_{ij}\}},
\quad
\mathcal{S}_{ti}=\sum\limits_{j\in J}G_{ij}, 
\quad 
\mathcal{S}_{ri}=\sum\limits_{i\in I}G_{ij},
\\
\beta_{ij}&=\frac{G_{ij}}{\mathcal{S}_{ti}+\mathcal{S}_{ri}-G_{ij}+B}
\end{align}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Measurement Update}

\begin{itemize}
\item A posteriori state is
\begin{align}
\hat x^+_{i}=&\ \hat x^-_{i}+K_{i}{\bf{e}}_{i}
\end{align}
where the gain matrix $K_i$ is chosen to minimize the trace of the covariance matrix $P^+_{i}$ where
\begin{align}
P^c_i=&\ (I-K_iH_i)P^-_i(I-K_iH_i)^T+K_iR_iK_i^T,
\\
\tilde P_i=&\ K_i\parenth{{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T}K_{i}^T,
\\
P^+_{i}=&\ \beta_{0,i}P^-_{i}+(1-\beta_{0,i})P_i^c+\tilde P_i \label{eqn:PostCovBasic}
\end{align}
\item However, the Kalman gain $K_{i,Kalman}=P^-_iH_i^TS_i^{-1}$ is \Emph{not} guaranteed to minimize $\tr{P^+_{i}}$ unless $\tilde P_i=0$, even though it is \Emph{widely used with the JPDAF}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Measurement Update}

\begin{itemize}
\item We can rewrite \refeqn{PostCovBasic} as 
\begin{align}
P^+_i&=P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)
+K_iE_iK_i^T,\label{eqn:CovExpanded}
\\
E_i=&\ (1-\beta_{0,i})S_i+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T
\end{align}
\item Then, we choose the gain $K_i$ that minimizes $\tr{P^+_{i}}$ as
\begin{align}
K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1},\label{eqn:GainK}
\end{align}
such that
\begin{align}
P^+_{i}=&P^-_{i}-K_iE_iK_i^T, \quad \mbox{or equivalently}\label{eqn:JPDAFPostCovSymmetricUpdate}
\\
P^+_{i}=&\left(I-(1-\beta_{0,i})K_iH_i\right)P^-_i\label{eqn:JPDAFPostCovSimpleUpdate}
\end{align}
\end{itemize}

\end{frame}

\section*{}
\subsection*{II. MUJPDAF Numerical Examples}

\begin{frame}
\frametitle{Numerical Problem Formulation}

\begin{itemize}
\item Let $\vec r=\begin{bmatrix}x_1 & x_2 & x_3\end{bmatrix}^T$ be the position vector of the $i$-th satellite represented with respect to the Earth-Centered Inertial (ECI) frame such that $\vec x=\begin{bmatrix}\vec r^T & \dot{\vec r}^T\end{bmatrix}^T\in\Re^6$
\item The nonlinear equation of motion is
\begin{align}
\ddot{\vec r}&=-\frac{\mu}{r^3}\vec r=-\mu(\vec r^T\vec r)^{-\frac32}\vec r
\end{align}
\item The linearized equation is given by
\begin{align}
\begin{bmatrix}
\delta\dot{\vec r} \\ \delta\ddot{\vec r}
\end{bmatrix}
%\delta\dot{\vec x}
=
\begin{bmatrix}
0_{3\times3} & I_{3\times3} \\
\mu\left[3\vec r({\vec r}^T\vec r)^{-\frac52}{\vec r}^T-({\vec r}^T\vec r)^{-\frac32}I_{3\times3}\right] & 0_{3\times3}
\end{bmatrix}
\begin{bmatrix}
\delta\vec r \\ \delta\dot{\vec r}
\end{bmatrix}
\end{align}
\item The ground-based sensor is situated on the equator such that
\begin{align}
x_s=R_E
\begin{bmatrix}
\cos\frac{2\pi t}T & \sin\frac{2\pi t}T & 0 & -\frac{2\pi t}T\sin\frac{2\pi t}T  & \frac{2\pi t}T\cos\frac{2\pi t}T & 0
\end{bmatrix}^T
\end{align}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Scenario A1: Comparison of the JPDAF and MUJPDAF with two LEO satellites in close proximity}

\begin{itemize}
\item Two satellites pass nearby with large initial uncertainties
\item Initial conditions:
\begin{align}
x_{1,A1}&=x_{ref,A1}+x_{pert,A1}, \quad x_{2,A1}=x_{ref,A1}-x_{pert,A1},\nonumber
\\
x_{ref,A1}&=\begin{bmatrix}7000 & 0 & 1000 & 0 & 7.546 & 0\end{bmatrix}^T,
\\
x_{pert,A1}&=\begin{bmatrix}
0.2 & 0 & 0 & 0 & -0.105 & 0
\end{bmatrix}^T\nonumber
\end{align}
\item Range, range-rate, topocentric right-ascension and declination measurements are used
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Scenario A1: Comparison of the JPDAF and MUJPDAF with two LEO satellites in close proximity}

\begin{itemize}
\item A1b: more uncertainty in position
\item A1d: more uncertainty in angular measurements
\end{itemize}

\begin{figure}
\centerline{
	\subfigure[\;Case A1: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.3\columnwidth]{A1JPDAF.pdf}}
	\subfigure[\;Case A1b: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.3\columnwidth]{A1bJPDAF.pdf}}
	\subfigure[\;Case A1d: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.3\columnwidth]{A1dJPDAF.pdf}}
}
\centerline{
	\subfigure[\;Case A1: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.3\columnwidth]{A1MUJPDAF.pdf}}
	\subfigure[\;Case A1b: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.3\columnwidth]{A1bMUJPDAF.pdf}}
	\subfigure[\;Case A1d: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.3\columnwidth]{A1dMUJPDAF.pdf}}
}
\caption{}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Scenario A2: Comparison of the JPDAF and MUJPDAF with two near-GEO satellites in close proximity}

\begin{itemize}
\item One satellite is on a geostationary orbit while the other has the same state and a small $z$-component
\item Initial conditions:
\begin{align}
x_{1,A2}&=\begin{bmatrix}r_{GEO} & 0 & 0 & 0 & \sqrt{\frac{\mu}{r_{GEO}}}, & 0\end{bmatrix}^T,\nonumber
\\
x_{2,A2}&=x_{1,A2}+\begin{bmatrix}
0 & 0 & -1 & 0 & 0 & 4\times10^{-4}
\end{bmatrix}^T\nonumber
\end{align}
\item Topocentric right-ascension and declination measurements are used
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Scenario A2: Comparison of the JPDAF and MUJPDAF with two near-GEO satellites in close proximity}

\begin{itemize}
\item A2a: more uncertainty in position
\item A2b: more uncertainty in velocity
\end{itemize}

\begin{figure}
\centerline{
	\subfigure[\;Case A2: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.3\columnwidth]{A2JPDAF.pdf}}
	\subfigure[\;Case A2a: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.3\columnwidth]{A2aJPDAF.pdf}}
	\subfigure[\;Case A2b: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.3\columnwidth]{A2bJPDAF.pdf}}
}
\centerline{
	\subfigure[\;Case A2: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.3\columnwidth]{A2MUJPDAF.pdf}}
	\subfigure[\;Case A2a: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.3\columnwidth]{A2aMUJPDAF.pdf}}
	\subfigure[\;Case A2b: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.3\columnwidth]{A2bMUJPDAF.pdf}}
}
\caption{}
\end{figure}

\end{frame}

\section*{}
\subsection*{III. C-JPDAF Problem Formulation}

\begin{frame}
\frametitle{Similarity Measure}

\begin{itemize}
\item Coalescence is well-known to occur when measurements are shared, which happens when the similarity among estimates is high
\item Illustration of similarity in a single dimension:
\begin{figure}
\centerline{
		{\includegraphics[width=.7\columnwidth]{GaussianDistOverlap1D.png}}
	}
\end{figure}
\item Define the similarity measure between an arbitrary number of Gaussian distributions:
\begin{align}
\zeta=\sum\limits_{p,q\in I,p\neq q}\exp \{-a(\hat y_{p}- \hat y_{q})^T({\mathbf P}_{p}
+{\mathbf P}_{q})^{-1}(\hat y_{p}-\hat y_{q})\}
\end{align}
where $\hat y_i$ has an uncertainty described by the covariance matrix ${\mathbf P}_{i}$
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Coalescence-Avoiding Gain Matrix}

\begin{itemize}
\item Main idea: choose the gain matrix $K_i$ to minimize a weighted sum of uncertainty and similarity
\item Cost function $\mathbf{J}\in\Re$ to minimize:
\begin{align}
\mathbf{J}=J_P+cJ_S,\quad J_P=\sum\limits_{i}\tr{P^+_i}, \quad J_S&=\sum\limits_{p,q\in I,p\neq q}\exp (-au_{pq}),
\end{align}
for $a,c>0$.
\item Nonlinear equations do not yield a closed-form solution
\end{itemize}
\begin{figure}
\setlength{\unitlength}{0.05\columnwidth}
\centerline{\footnotesize\selectfont
\begin{picture}(8.5,7.3)(-0.5,-0.8)
\put(-0.5,3){\rotatebox{90}{Cost}}
\put(0,0){\includegraphics[width=0.4\columnwidth]{ECC14_fig1.pdf}}
\put(5.0,1.1){\shortstack[c]{Uncertainty\\ cost: $J_P$}}
\put(5.7,2.3){\shortstack[c]{Similarity\\ cost: $J_S$}}
\put(2.2,4.0){\shortstack[c]{Weighted sum:\\ $\mathbf{J}=J_P+cJ_S$}}
\put(4.98,0.78){\circle*{0.1}}
\put(2.48,1.57){\circle*{0.1}}
\put(4.8,-0.4){$K_{MUJPDAF}$}
\put(2.3,-0.4){$K_{C-JPDAF}$}
\end{picture}
}
\end{figure}

\end{frame}

\section*{}
\subsection*{IV. C-JPDAF Numerical Examples}

\begin{frame}
\frametitle{Scenario B1: Comparison of the JPDAF, MUJPDAF, and C-JPDAF with three LEO satellites in close proximity}

\begin{itemize}
\item Three satellites have nearly the same initial state except for a small perturbation in the $x$-direction
\item Initial conditions:
\begin{align}
\begin{split}
x_{1,B1}&=x_{ref,B1}+x_{pert,B1}, \quad x_{2,B1}=x_{ref,B1}-x_{pert,B1}, \quad x_{3,B1}=x_{ref,B1},
\\
x_{ref,B1}&=\begin{bmatrix}7000 & 0 & 0 & 0 & 7.546 & 0\end{bmatrix}^T,\nonumber
\\
x_{pert,B1}&=\begin{bmatrix}
0.025 & 0 & 0 & 0 & 0 & 0
\end{bmatrix}^T\nonumber
\end{split}
\end{align}
\item Range, range-rate, topocentric right-ascension and declination measurements are used
\item B1c: position and velocity are more uncertain
\item B1f: range-rate measurement is more uncertain
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Scenario B1: JPDAF, MUJPDAF, and C-JPDAF in LEO}

%\begin{itemize}
%\item B1b: more uncertainty in measurement range rate
%\item B1d: more uncertainty in angular measurements
%\end{itemize}
\begin{figure}
%\vspace*{-0.3cm}
\centerline{
	\subfigure[\;B1: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B1JPDAF.pdf}}
	\subfigure[\;B1: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B1MUJPDAF.pdf}}
	\subfigure[\;B1: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B1C-JPDAF.pdf}}
	}
%\vspace*{-0.2cm}
\centerline{
	\subfigure[\;B1c: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B1cJPDAF.pdf}}
	\subfigure[\;B1c: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B1cMUJPDAF.pdf}}
	\subfigure[\;B1c: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B1cC-JPDAF.pdf}}
	}
%\vspace*{-0.2cm}
\centerline{
	\subfigure[\;B1f: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B1fJPDAF.pdf}}
	\subfigure[\;B1f: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B1fMUJPDAF.pdf}}
	\subfigure[\;B1f: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B1fC-JPDAF.pdf}}
	}
\caption{}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Scenario B2: Comparison of the JPDAF, MUJPDAF, and C-JPDAF with two GEO satellites in close proximity}

\begin{itemize}
\item Two satellites in a geostationary orbit are offset by a very small phase shift
\item Initial conditions:
\begin{align}
\begin{split}
f_{GEO,r}(\theta)&=\begin{bmatrix}r_{GEO}\cos\theta & r_{GEO}\sin\theta & 0\end{bmatrix}^T,
\\
\quad f_{GEO,\dot r}(\theta)&=\begin{bmatrix}\sqrt{\frac{\mu}{r_{GEO}}}\cos(90^\circ+\theta) & \sqrt{\frac{\mu}{r_{GEO}}}\sin(90^\circ+\theta) & 0\end{bmatrix}^T
\\
f_{GEO}(\theta)&=\begin{bmatrix}f_{GEO,r} & f_{GEO,\dot r}\end{bmatrix}^T,
\\
x_{1,B2}&=f_{GEO}(0^\circ), \quad x_{2,B2}=f_{GEO}({10^{-4}}^\circ)
\end{split}\nonumber
\end{align}
\item Topocentric right-ascension and declination measurements are used
\item B2b: velocity is more uncertain
\item B2d: angle measurements are more uncertain
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Scenario B2: JPDAF, MUJPDAF, and C-JPDAF in GEO}

\begin{figure}
%\vspace*{-0.3cm}
\centerline{
	\subfigure[\;B2: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B2JPDAF.pdf}}
	\subfigure[\;B2: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B2MUJPDAF.pdf}}
	\subfigure[\;B2: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B2C-JPDAF.pdf}}
	}
\centerline{
	\subfigure[\;B2b: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B2bJPDAF.pdf}}
	\subfigure[\;B2b: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B2bMUJPDAF.pdf}}
	\subfigure[\;B2b: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B2bC-JPDAF.pdf}}
	}
\centerline{
	\subfigure[\;B2d: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B2dJPDAF.pdf}}
	\subfigure[\;B2d: MUJPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B2dMUJPDAF.pdf}}
	\subfigure[\;B2d: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.23\columnwidth]{B2dC-JPDAF.pdf}}
	}
\caption{}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Computation Time}

\begin{center}
\resizebox{0.66\columnwidth}{!}{
\begin{threeparttable}[h]
\caption{B1 Scenarios} \label{tab:B1}
\begin{tabularx}{1.16\columnwidth}
{
>{$}c<{$} |
*{3}{>{$}c<{$}} |
*{3}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multicolumn{3}{c}{\multirow{1}{*}{RMS Position Error (km)}} & \multicolumn{3}{c}{\multirow{1}{*}{Time (sec)}} \\
 & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{MUJPDAF} & \multirow{1}{*}{C-JPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{MUJPDAF} & \multirow{1}{*}{C-JPDAF}
\\
\midrule
B1   	& 0.14432 & 0.091133 & 0.057312 & 1.736272 & 1.670539 & 101.000529 \\
B1a 	& 0.14432 & 0.091134 & 0.057307 & 1.689807 & 1.726980 & 100.988662 \\
B1b 	& 0.14694 & 0.094506 & 0.055145 & 1.673762 & 1.678786 & 100.855376 \\
B1c 	& 0.14693 & 0.094507 & 0.055136 & 1.675486 & 1.669083 & 100.907850 \\
B1d 	& 0.14400 & 0.088933 & 0.058911 & 1.663424 & 1.660136 & 102.606102 \\
B1e 	& 0.14408 & 0.090823 & 0.059919 & 1.673660 & 1.666829 & 100.278073 \\
B1f 	& 0.14436 & 0.091212 & 0.064299 & 1.732541 & 1.721413 & 98.672870 \\
\bottomrule
\end{tabularx}
\end{threeparttable}
}
\end{center}

\begin{center}
\resizebox{0.8\columnwidth}{!}{
\begin{threeparttable}[h]
\caption{B2 Scenarios} \label{tab:B2}
\begin{tabularx}{1.35\columnwidth}
{
>{$}c<{$} |
*{3}{>{$}c<{$}} |
*{3}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multicolumn{3}{c}{\multirow{1}{*}{RMS Angular Error (rad)}} & \multicolumn{3}{c}{\multirow{1}{*}{Time (sec)}} \\
 & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{MUJPDAF} & \multirow{1}{*}{C-JPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{MUJPDAF} & \multirow{1}{*}{C-JPDAF}
\\
\midrule
B2   	& 1.1112\times10^{-6} & 9.4780\times10^{-7} & 5.6708\times10^{-7} & 1.149378 & 1.119439 & 34.318254 \\
B2a 	& 1.1112\times10^{-6} & 9.4779\times10^{-7} & 5.6347\times10^{-7} & 1.154256 & 1.150072 & 34.758609 \\
B2b 	& 1.1306\times10^{-6} & 9.8287\times10^{-7} & 5.7387\times10^{-7} & 1.126598 & 1.132463 & 35.327683 \\
B2c 	& 1.1306\times10^{-6} & 9.8286\times10^{-7} & 5.6435\times10^{-7} & 1.118436 & 1.131690 & 34.817179 \\
B2d 	& 1.1115\times10^{-6} & 9.5439\times10^{-7} & 5.6136\times10^{-7} & 1.113892 & 1.116399 & 35.213445 \\
\bottomrule
\end{tabularx}
\end{threeparttable}
}
\end{center}

\end{frame}

\section*{}
\subsection*{IV. Conclusions}

\begin{frame}
\frametitle{Conclusions}

\begin{itemize}
\item The Kalman gain is not guaranteed to minimize the state uncertainty generated by the JPDAF
\begin{itemize}
\item This motivates the derivation of the minimum uncertainty JPDAF, designed to account for all parts of the state covariance by changing the gain matrix (not Kalman), derived as a research contribution
\item Numerous simulations show that the MUJPDAF experiences smaller RMS errors and covariances in \Emph{every} simulated scenario
\item The computation is nearly identical between the algorithms
\end{itemize}
\item Both the JPDAF and MUJPDAF experience coalescence
\begin{itemize}
\item Research contribution: the C-JPDAF, which systematically removes coalescence from the soft decision techniques
\item Similar to the derivation of the MUJPDAF, we choose a gain that minimizes a weighted sum of uncertainty and similarity
\item In many cases, this leads to improved accuracy of the estimation in exchange for a greater computational load
\end{itemize}
\item Future work
\begin{itemize}
\item Determine systematic way of choosing weighting parameters for the C-JPDAF
\item Solve a sub-optimal solution for the C-JPDAF to decrease its computational load to a practical level
\end{itemize}
\end{itemize}

\end{frame}

\end{document}

### 2018-06-09 Conference/AAS 2015 OD EKF/AAS15.tex
added Conference folder

% use paper, or submit
% use 11 pt (preferred), 12 pt, or 10 pt only

%\documentclass[letterpaper, preprint, paper,11pt]{AAS}	% for preprint proceedings
%\documentclass[letterpaper, paper,11pt]{AAS}		% for final proceedings (20-page limit)
%\documentclass[letterpaper, paper,12pt]{AAS}		% for final proceedings (20-page limit)
\documentclass[letterpaper, paper,10pt]{AAS}		% for final proceedings (20-page limit)
%\documentclass[letterpaper, submit]{AAS}			% to submit to JAS

\usepackage{bm}
\usepackage{amssymb,amsmath,amsthm,times,graphicx,subfigure,tabularx,booktabs,colortbl,multirow,threeparttable}
\usepackage{subfigure}
%\usepackage[notref,notcite]{showkeys}  % use this to temporarily show labels
\usepackage[colorlinks=true, pdfstartview=FitV, linkcolor=black, citecolor= black, urlcolor= black]{hyperref}
\usepackage{overcite}
\usepackage{footnpag}			      	% make footnote symbols restart on each page

\usepackage{color}
\newcommand{\hilight}[1]{\colorbox{green}{#1}}

\newtheorem{prop}{Proposition}

\newcommand{\abs}{\mathop{\mathrm{abs}}\nolimits}
\newcommand{\diag}{\mathop{\mathrm{diag}}\nolimits}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\ave}[1]{\mathrm{E}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}

\newcommand{\x}{\ensuremath{\mathbf{x}}}
\renewcommand{\r}{\mathbf{r}}
\renewcommand{\u}{\mathbf{u}}
\newcommand{\y}{\mathbf{y}}

\newcommand{\bfi}{\bfseries\itshape\selectfont}

%\renewcommand{\baselinestretch}{1.2}

\PaperNumber{23-451}

\begin{document}

\title{Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements}

\author{Evan Kaufman\thanks{Doctoral Student, Mechanical and Aerospace Engineering, George Washington University, 801 22nd St NW, Washington, DC 20052, Email: evankaufman@gwu.edu.},  
T. Alan Lovell\thanks{Research Aerospace Engineer, Air Force Research Laboratory, Space Vehicles Directorate, Kirtland AFB, NM.},
\ and Taeyoung Lee\thanks{Assistant Professor, Mechanical and Aerospace Engineering, George Washington University, 801 22nd St NW, Washington, DC 20052, Tel: 202-994-8710, Email: tylee@gwu.edu.}
\thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}
}

\maketitle{}

\begin{abstract}
%An angles-only extended Kalman filter (EKF), based off of two-body motion, is developed to estimate the position of a Deputy satellite relative to a chief satellite.
A new nonlinear observability measure is proposed for relative orbit determination when lines-of-sight between satellites are measured only. It corresponds to a generalization of the observability Gramian in linear dynamic systems to the nonlinear relative orbit dynamics represented by the two-body problems. An extended Kalman filter (EKF) is adapted to this problem and is evaluated with various gravitational harmonics and initial orbital determination (IOD) predictions. Extensive results illustrate correspondence between the proposed observability measure with filtering errors. An extensive numerical analysis in realistic scenarios includes satellite propagation of the two-body problem the $J_2$ perturbation effects.
\end{abstract}

\section{Introduction}

Space-based surveillance or relative navigation is desirable for many spacecraft missions, such as formation control and rendezvous. Spacecraft maneuvers based only on on-board measurements reduce the total operating cost significantly, and it improves safety against communication interruptions with ground stations.  Relative navigation between spacecraft in close-proximity essentially corresponds to space-based orbit determination.  In particular, vision-based navigation and estimation of relative orbits have received attention recently, since optical sensors have the desirable properties of low cost and minimal maintenance, while providing accurate line-of-sight measurements, or angles from a chief to a deputy.

Relative navigation based on angles-only measurements has been investigated in~\cite{WofGelITAES09,Tom11,PatLovPASFMM12}.  The problem is to determine the relative orbit between a chief spacecraft and a deputy spacecraft by using the line-of-sight between the two objects, assuming that the orbit of the chief is prescribed exactly.  Reference \cite{WofGelITAES09} shows that the relative orbit is unobservable from angles-only measurements when linear relative orbital dynamics are assumed, unless there are thrusting maneuvers. Reference \cite{Tom11} investigates observability by using a relative orbit model linearized in terms of spherical coordinates.  Reference~\cite{PatLovPASFMM12} introduces the concept of partial observability to determine a basis vector representing a family of relative orbits, and an initial orbit determination technique is developed for this method.

All of these results are based on linear relative orbital dynamics. It is straightforward to see that the relative orbit is not observable with angles-only measurements through its linearized dynamics. In other words, there are an infinite number of relative orbits that yield the identical line-of-sight measurements, and the orbital distance between a deputy and a chief cannot be determined by angles-only measurements. As such, it is required to study the nonlinear relative orbital dynamics to determine observability with angles-only measurements. Linear observability analysis is performed numerically for a particular case in~\cite{YimCraPASMM04}.

Recently, observability criteria have been derived for the nonlinear relative orbital dynamics represented by the solutions of the two-body problem without linearization~\cite{LovLeePISSFD14}.  Assuming that a chief is on a circular orbit with a prescribed orbital radius, nonlinear equations of motion for the relative orbital motion of a deputy with respect to the chief are presented.  A differential geometric method, based on the Lie derivatives of the line-of-sight from the chief to the deputy, is applied to obtain sufficient conditions for observability.  It is shown that under certain geometric conditions on the relative configuration between the chief and the deputy, the nonlinear relative motion is observable from angles-only measurements.

However, this result does not provide any information of the degree of observability. The main contribution of this paper is proposing a new  measure of observability for relative orbit determination with angles-only measurements. For a given initial condition of relative orbits, the proposed observability measure determines how much the corresponding relative orbits are easy or difficult to estimate, thereby providing a quantitative measure for relative orbit determination.

Furthermore, this paper serves as a feasibility study of an extended Kalman filter (EKF) for orbit determination using angles-only measurements in realistic scenarios. The filter is subject to measurement uncertainty and perturbation forces not considered with the two-body problem, on which the filter is based.
%The EKF is simulated with various time steps to show how the filter performs on a wide range of systems with different sample rates.
Then, the performance of the EKF is compared with the proposed observability measure to show the correspondence.

\section{Nonlinear Relative Orbital Dynamics}\label{sec:ND}

Consider two satellites orbiting around the Earth, where each satellite is modeled as a rigid body. Suppose that a \textit{chief} satellite is on a circular orbit with a pre-determined orbital radius of $a\in\Re$.

Define a local-vertical, local-horizontal (LVLH) frame as follows. Its origin is located at the chief satellite. The $x$-axis is along the radial direction from the Earth to the chief, and the $y$-axis is along the velocity vector of the chief. The $z$-axis is normal to the orbital plane, and it is parallel to the angular momentum vector of the chief. The LVLH frame is rotating with the angular velocity of $\mathbf{\omega}=[0,0,n]^T\in\Re^3$, where $n=\sqrt{\frac{\mu}{a^3}}$ is the mean motion of the chief satellite, and $\mu$ denotes the gravitational parameters of the Earth. Note that the velocity of the chief is given by $\mathbf{v}_{chief}=[0,na,0]^T\in\Re^3$. Let the relative position of a \textit{deputy} satellite with respect to the chief satellite be given by $\mathbf{r}=[x,y,z]^T\in\Re^3$ in the LVLH frame.

\subsection{Nonlinear Equations of Motion}

Nonlinear equations of motion for the relative motion of the deputy with respect to the chief can be derived as follows based on Lagrangian mechanics~\cite{LovLeePISSFD14}.

\paragraph{Lagrangian}

Considering that the LVLH frame is rotating with the angular velocity of $\mathbf{\omega}=[0,0,n]^T$, the inertial velocity of the deputy satellite is given by
\begin{align*}
\mathbf{v}_{deputy}
=
\mathbf{v}_{chief} + \dot{\r} + \mathbf{\omega}\times \mathbf{r} = 
\begin{bmatrix}
0 \\ na \\ 0
\end{bmatrix}
+
\begin{bmatrix}
\dot x \\ \dot y \\ \dot z
\end{bmatrix}
+
\begin{bmatrix}
-n y \\ nx \\ 0
\end{bmatrix}
=
\begin{bmatrix}
\dot x - ny \\ \dot y + nx + na \\ \dot z
\end{bmatrix}.
\end{align*}
Therefore, the (normalized) kinetic energy of the deputy satellite is 
\begin{align*}
T = \frac{1}{2}\|\mathbf{v}_{deputy}\|^2 = \frac{1}{2} \braces{(\dot x -ny)^2+(\dot y + nx +na)^2 + \dot z^2}.
\end{align*}
The location of the Earth from the chief is given by $[-a,0,0]^T$ in the LVLH frame. Therefore, the position vector of the deputy from the center of the Earth is given by 
$\r_a=[x+a,y,z]^T\in\Re^3$. The gravitational potential energy is 
\begin{align*}
U = -\frac{\mu}{\sqrt{(x+a)^2 + y^2 + z^2}} = -\frac{\mu}{\|\r_a\|}.
\end{align*}
From the above equations, the Lagrangian of the deputy satellite is expressed in terms of $(x,y,z)$ as 
\begin{align}
L & = T-U = \frac{1}{2} \braces{(\dot x -ny)^2+(\dot y + nx +na)^2 + \dot z^2}
+\frac{\mu}{\sqrt{(x+a)^2 + y^2 + z^2}}.
\end{align}

\paragraph{Euler-Lagrange Equations}

Using the Euler-Lagrange equations, given by
\begin{align*}
\frac{d}{dt}\deriv{L}{\dot q}-\deriv{L}{q}=0,
\end{align*}
for $q\in\{x,y,z\}$, we obtain the nonlinear equations of motion for the relative orbit as follows.
\begin{align}
\ddot x - 2n\dot y -n^2 x&=n^2 a - \frac{\mu (x+a)}{((x+a)^2 + y^2 + z^2)^{3/2}},\label{eqn:ddotx}\\
\ddot y + 2n\dot x -n^2 y &=   - \frac{\mu y}{((x+a)^2 + y^2 + z^2)^{3/2}},\label{eqn:ddoty}\\
\ddot z &= - \frac{\mu z}{((x+a)^2 + y^2 + z^2)^{3/2}}.\label{eqn:ddotz}
\end{align}

These can be written as the standard form of the state equation,
\begin{align}
\dot{\mathbf{x}} = f (\mathbf{x}), \label{eqn:xxdot}
\end{align}
where the state vector is $\mathbf{x}=[x,y,z,\dot x, \dot y,\dot z]^T\in\Re^N$ with $N=6$, and
\begin{align}
f(\mathbf{x}) 
%= \begin{bmatrix}
%\dot x\\
%\dot y\\
%\dot z\\
%2n\dot y +n^2 (x+a) - \dfrac{\mu (x+a)}{((x+a)^2 + y^2 + z^2)^{3/2}}\\
%-2n\dot x +n^2 y  - \dfrac{\mu y}{((x+a)^2 + y^2 + z^2)^{3/2}}\\
%- \dfrac{\mu z}{((x+a)^2 + y^2 + z^2)^{3/2}}
%\end{bmatrix}
=\begin{bmatrix}
\dot \r \\
-2\omega\times \dot\r - \omega\times(\omega\times \r_a) -\dfrac{\mu \r_a}{\|\r_a\|^3}
\end{bmatrix}.\label{eqn:f}
\end{align}
%where $\r=[x,y,z]^T\in\Re^3$, $\r_a = [x+a,y,z]^T\in\Re^3$, and $\dot\r =[\dot x,\dot y,\dot x]^T\in\Re^3$.

\subsection{Line-of-sight Measurement}

We assume that the line-of-sight from the chief to the deputy is measured by onboard sensing, such as optical sensors. The measurement is represented by the unit-vector of the relative position vector, i.e.,
\begin{align}
\y = h(\mathrm{x}) =\frac{\r}{\|\r\|},\label{eqn:y}
\end{align}
where $\y\in\Re^M$ with $M=3$, and it satisfies $\|\y\|=1$.

\section{Nonlinear Observability Criteria}\label{sec:OC}

Based on the nonlinear dynamic model presented at the previous section, here we analyze the observability of the relative orbit with line-of-sight measurements.
This section serves as a summary of~\cite{LovLeePISSFD14}.

\subsection{Observability Criteria for Nonlinear Systems}

Observability of nonlinear systems has been studied in~\cite{HerKreITAC77}, and it is summarized as follows. For a given nonlinear dynamic system described with \refeqn{xxdot} and \refeqn{y}, a pair of points $\x_0$ and $\x_1$ are called \textit{indistinguishable} if the outputs of the corresponding solutions starting from each of $\x_0$ and $\x_1$ are identical for a certain time period. The systems is \textit{locally weakly observable} at $\x_0$, if there exists an open neighborhood $V$ of $\x_0$ such that for every open neighborhood $U$ of $\x_0$ contained in $V$, the only indistinguishable point to $\x_0$ is the point $\x_0$ itself.

The Lie-derivative of the output $h(\x)$ along $f(\x)$ is defined as follows:
\begin{align*}
L_f h(\x) = \deriv{h(\x)}{\x}f(\x)\in\Re^{M\times 1},
\end{align*}
which corresponds to the directional derivative of $h(\x)$ along $f(\x)$. For a non-negative integer $i$, the $i$-th order Lie-derivative is defined by induction as $L_f^i h = L_f (L_f^{i-1} h)$ with $L_f^0 h = h$. Define an observability matrix $\mathcal{O}\in\Re^{NM\times N}$ as
\begin{align*}
\mathcal{O}(\x_0) = \deriv{}{\x} \begin{bmatrix} L_f^0 h(\x) \\ L_f^1 h(\x)\\ \vdots \\L_f^{N-1} h(\x)\end{bmatrix}\bigg|_{\x=\x_0}.
\end{align*}
It has been shown that the system is locally weakly observable at $\x_0$ if the rank of the observability matrix $\mathcal{O}(\x_0) = N$. When applied to linear dynamics, this yields the well-known observability rank condition for linear systems. Note that when there is more than a single measurement, i.e., $M>1$, the observability rank condition can be satisfied without a need for computing the higher-order Lie derivatives up to the $N-1$-th order.

\subsection{Observability Criteria for Relative Orbital Dynamics}

\paragraph{Observability Matrix}
We apply the above observability criteria for the nonlinear relative orbital dynamics. Using the fact that $\x=[\r^T,\dot \r^T]$, the observability matrix of the relative orbital dynamics can be written as
\begin{align}
\mathcal{O} =
\begin{bmatrix}
\deriv{\y}{\r} & \deriv{\y}{\dot\r}\\
\deriv{\dot\y}{\r} & \deriv{\dot\y}{\dot\r}\\
\deriv{\ddot\y}{\r} & \deriv{\ddot\y}{\dot\r}
\end{bmatrix}
\triangleq
\begin{bmatrix}
\mathcal{O}_{00} & 0_{3\times 3}\\
\mathcal{O}_{10} & \mathcal{O}_{11}\\
\mathcal{O}_{20} & \mathcal{O}_{21}
\end{bmatrix}.\label{eqn:OO}
\end{align}
Here we consider the observability matrix obtained by up to the second order Lie derivatives of the measurement due to complexity. But, this still provides sufficient conditions for observability.

After straightforward but tedious algebraic manipulations using the following identity repeatedly,
\begin{align*}
\delta \parenth{\frac{1}{\|\mathbf{r}\|^i}} 
& = -i\frac{ \mathbf{r}^T \delta \mathbf{r}}{\|\mathbf{r}\|^{i+2}},
\end{align*}
for any positive integer $i$, we can show that each of the sub-matrices $\mathcal{O}_{ij}$ of the observability matrix $\mathcal{O}$ is given by
\begin{align}
\mathcal{O}_{00} & = \deriv{\y}{\r} = \frac{1}{\|\r\|}(I - \y\y^T),\label{eqn:O_00}\\
\mathcal{O}_{10} & = \deriv{\dot\y}{\r} = \deriv{}{\r} \parenth{\deriv{\y}{\r}\dot \r}= -\frac{1}{\|\r\|^2}\braces{
{\dot\r\y^T +\y^T\dot\r I +\y\dot\r^T}
- 3\y\y^T\dot\r \y^T},\\
\mathcal{O}_{11} & = \deriv{\dot\y}{\dot\r} = \deriv{}{\dot\r} \parenth{\deriv{\y}{\r}\dot \r}= \deriv{\y}{\r} = \mathcal{O}_{00},\label{eqn:O11}\\
\mathcal{O}_{20} & = \deriv{\ddot\y}{\r} = -\dfrac{2\dot\r \dot\r^T  +\dot\r^T\dot\r I}{\|\r\|^3}
+3\dfrac{2(\r^T\dot\r) \dot\r\r^T +(\dot\r^T\dot\r)\r\r^T
+(\r^T\dot\r)^2 I
+2(\r^T\dot\r)\r  \dot\r^T
}{\|\r\|^5}
- 15 \dfrac{(\r^T\dot\r)^2\r  \r^T}{\|\r\|^7}\nonumber \\
&\quad -\dfrac{\ddot{\r}\r^T +\r^T\ddot{\r} I +\r \ddot{\r}^T}{\|\r\|^3}
+ 3 \dfrac{(\r^T\ddot{\r})\r \r^T}{\|\r\|^5}
+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}},\\
\mathcal{O}_{21} & = \deriv{\ddot\y}{\dot\r} = \deriv{}{\dot\r}\parenth{\deriv{}{\r}\parenth{\deriv{\y}{\r}\dot\r}\dot\r + \deriv{\y}{\r}\ddot\r} = 2 \deriv{}{\r} \parenth{\deriv{\y}{\r}\dot\r} + \deriv{\y}{\r} \deriv{\ddot\r}{\dot\r}
\nonumber\\
& \quad = 2\mathcal{O}_{10} 
- 2 \mathcal{O}_{00}[\omega]_\times,\label{eqn:O_21}
\end{align}
where $[\omega]_\times\in\Re^{3\times 3}$ is defined as
\begin{align*}
[\omega]_\times = \begin{bmatrix} 0 & -n & 0 \\ n & 0 & 0 \\ 0 & 0 & 0 \end{bmatrix}.
\end{align*}
These expressions were fully verified by the Matlab symbolic computation toolbox.

After some algebraic manipulations, we can show that the sub-matrices satisfy the following identities. 
\begin{align}
\mathcal{O}_{10}\r & = 
%-\frac{1}{\|\r\|}\braces{{\dot\r\y^T +\y^T\dot\r I +\y\dot\r^T}- 3\y\y^T\dot\r \y^T}\y\\
%&=-\frac{1}{\|\r\|}\braces{\dot\r +(\y^T\dot\r)\y +\y(\dot\r^T\y)- 3\y(\y^T\dot\r) }\\
%&=-\frac{1}{\|\r\|}\braces{\dot\r -\y(\y^T\dot\r) }=
-\mathcal{O}_{00}\dot\r,\label{eqn:O10r}\\
\mathcal{O}_{10}\dot\r & = -\frac{1}{\|\r\|^2}\braces{
{2\dot\r(\y^T\dot \r)  +\y(\dot\r^T\dot \r)}
- 3\y(\y^T\dot\r)^2},\label{eqn:O10dotr}\\
\mathcal{O}_{20}\r %& = 
%-\dfrac{2\dot\r \dot\r^T\y  +\dot\r^T\dot\r \y}{\|\r\|^2}
%+3\dfrac{2(\y^T\dot\r) \dot\r\y^T\y +(\dot\r^T\dot\r)\y\y^T\y+(\y^T\dot\r)^2 y +2(\y^T\dot\r)\y \dot\r^T\y}{\|\r\|^2}
%- 15 \dfrac{(\y^T\dot\r)^2\y  \y^T\y}{\|\r\|^2}\nonumber \\
%&\quad -\dfrac{\ddot{\r}\y^Ty +\y^T\ddot{\r} \y +\y \ddot{\r}^T\y}{\|\r\|^1}
%+ 3 \dfrac{(\y^T\ddot{\r})\y \y^T\y}{\|\r\|}
%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
%&=
%-\dfrac{2\dot\r (\dot\r^T\y)  +(\dot\r^T\dot\r) \y}{\|\r\|^2}
%+3\dfrac{2(\y^T\dot\r) \dot\r +(\dot\r^T\dot\r)\y+(\y^T\dot\r)^2 \y +2(\y^T\dot\r)^2\y }{\|\r\|^2}
%- 15 \dfrac{(\y^T\dot\r)^2\y}{\|\r\|^2}\nonumber \\
%&\quad -\dfrac{\ddot{\r} +(\y^T\ddot{\r}) \y +\y (\ddot{\r}^T\y)}{\|\r\|^1}
%+ 3 \dfrac{(\y^T\ddot{\r})\y }{\|\r\|}
%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
%&=
%-\dfrac{2\dot\r (\dot\r^T\y)  +(\dot\r^T\dot\r) \y}{\|\r\|^2}
%+3\dfrac{2(\y^T\dot\r) \dot\r +(\dot\r^T\dot\r)\y+(\y^T\dot\r)^2 \y +2(\y^T\dot\r)^2\y }{\|\r\|^2}
%- 15 \dfrac{(\y^T\dot\r)^2\y}{\|\r\|^2}\nonumber \\
%&\quad -\dfrac{\ddot{\r} +(\y^T\ddot{\r}) \y +\y (\ddot{\r}^T\y)}{\|\r\|^1}
%+ 3 \dfrac{(\y^T\ddot{\r})\y }{\|\r\|}
%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
%&=
%\dfrac{4\dot\r (\dot\r^T\y)  -2(\dot\r^T\dot\r)\y-6(\y^T\dot\r)^2 \y }{\|\r\|^2}
% -\mathcal{O}_{00}\ddot \r
%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
& = -2\mathcal{O}_{10}\dot\r +\mathcal{O}_{00}\braces{-\ddot \r +\deriv{\ddot\r}{\r}\r},\label{eqn:O20r}\\
\mathcal{O}_{21}\r & = -2\mathcal{O}_{00} (\dot\r+\omega\times\r),\label{eqn:O21r}\\
\mathcal{O}_{21}\dot\r & = 2\mathcal{O}_{10} \dot\r
- 2 \mathcal{O}_{00}[\omega]_\times\dot\r,\label{eqn:O21rdot}
\end{align}
which are useful to derive the observability criteria.

\paragraph{Observability Rank Condition}

Now we present sufficient conditions that the observability matrix $\mathcal{O}$ has full rank.

\begin{prop}
Define three vectors $\mathbf{v}_{rel}$, $\mathbf{a}_1,\mathbf{a}_2\in\Re^3$ as 
\begin{align}
\mathbf{v}_{rel} & = \dot\r +\omega\times\r,\label{eqn:vrel}\\
\mathbf{a}_1 & = \ddot\r-\deriv{\ddot\r}{\r}\r = -2\omega\times \dot\r - [\omega]_\times^2 a \mathbf{e}_1 -\dfrac{\mu a }{\|\r_a\|^3}\mathbf{e}_1- \dfrac{3\mu\r_a^T\r}{\|\r_a\|^5}\r_a,\label{eqn:a1}\\
\mathbf{a}_2 & = \ddot\r-\deriv{\ddot\r}{\r}\r -\deriv{\ddot\r}{\dot\r}\dot\r = \mathbf{a}_1+2\omega\times\dot\r= - [\omega]_\times^2 a \mathbf{e}_1 -\dfrac{\mu a }{\|\r_a\|^3}\mathbf{e}_1- \dfrac{3\mu\r_a^T\r}{\|\r_a\|^5}\r_a.\label{eqn:a2}
\end{align}
The nonlinear relative orbital dynamics are locally weakly observable at $\x=[\r^T,\dot\r^T]$ if
%\begin{list}{}{\setlength{\leftmargin}{1cm}\setlength{\itemsep}{0mm}\setlength{\parsep}{0mm}\setlength{\topsep}{0mm}\setlength{\parskip}{0mm}\setlength{\labelwidth}{2cm}}
%\item[(i)] when $\r\times\dot\r =0$, $\r\times\mathbf{a}_1\neq 0$ and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_1)\neq 0$,
%\item[(ii)] when $\r\times\dot\r \neq 0$, $\r\times\mathbf{a}_2\neq 0$ and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_2)\neq 0$.
%\end{list}
\begin{alignat}{2}
(i)&\; &\text{when $\r\times\dot\r =0$, $\r\times\mathbf{v}_{ref}\neq0$, $\r\times\mathbf{a}_1\neq 0$,  and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_1)\neq 0$},\label{eqn:cond1}\\
(ii)& &\text{when $\r\times\dot\r \neq 0$, $\r\times\mathbf{v}_{ref}\neq0$, $\r\times\mathbf{a}_2\neq 0$ and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_2)\neq 0$.}\label{eqn:cond2}
\end{alignat}
\end{prop}

\begin{proof}
We show that if the above conditions are satisfied, then the six columns of the observability matrix are linearly independent. Suppose that for a constant vector $\mathbf{c}=[\mathbf{c}_1^T,\mathbf{c}_2^T]\in\Re^6$, where $\mathbf{c}_1,\mathbf{c}_2\in\Re^3$, we have $\mathcal{O}\mathbf{c}=0$, i.e., 
\begin{align}
\mathcal{O} \mathbf{c} = \begin{bmatrix}
\mathcal{O}_{00} & 0_{3\times 3}\\
\mathcal{O}_{10} & \mathcal{O}_{11}\\
\mathcal{O}_{20} & \mathcal{O}_{21}
\end{bmatrix}
\begin{bmatrix} \mathbf{c}_1 \\ \mathbf{c_2} \end{bmatrix}
=
\begin{bmatrix}
\mathcal{O}_{00}\mathbf{c}_1\\
\mathcal{O}_{10}\mathbf{c}_1+ \mathcal{O}_{11}\mathbf{c}_2\\
\mathcal{O}_{20}\mathbf{c}_1+ \mathcal{O}_{21}\mathbf{c}_2
\end{bmatrix}
=
\begin{bmatrix}
0_{3\times 1}\\
0_{3\times 1}\\
0_{3\times 1}
\end{bmatrix}.\label{eqn:Oc}
\end{align}

We wish to show that there is no non-zero vector $\mathbf{c}$ satisfying \refeqn{Oc}. At the first three rows of \refeqn{Oc}, we have
\begin{align*}
\mathcal{O}_{00} \mathbf{c}_1 = \frac{1}{\|\r\|}(I-\y\y^T) \mathbf{c}_1 = 0.
\end{align*}
The matrix $\mathcal{O}_{00}$ has a one-dimensional null space spanned by $\y$. Therefore, without loss of generality, we can choose $\mathbf{c}_1=0$ or $\mathbf{c}_1=\r$. Suppose $\mathbf{c}_1=\r$ for the subsequent development.

For the chosen value of $\mathbf{c}_1=\r$, we find the next three rows of \refeqn{Oc} as
\begin{align}
\mathcal{O}_{10} \mathbf{c}_1 +\mathcal{O}_{11} \mathbf{c}_2 & = \mathcal{O}_{10} \r +\mathcal{O}_{11} \mathbf{c}_2  = \mathcal{O}_{00} (-\dot\r + \mathbf{c}_2)\nonumber\\
& = \frac{1}{\|\r\|}(I-\y\y^T)(-\dot\r + \mathbf{c}_2)=0,\label{eqn:row2}
\end{align}
where we have used \refeqn{O11} and \refeqn{O10r}. Next, we consider two cases of \refeqn{row2}, namely (i) when $\r\times\dot\r=0$, and (ii) when $\r\times\dot\r\neq 0$.

\paragraph{Case (i): $\r\times\dot \r=0$} In this case, $\dot \r$ can be written as $\dot \r = \alpha\r$ for some constant $\alpha$ as $\r$ is parallel to $\dot\r$. Then, \refeqn{row2} reduces to 
\begin{align*}
\frac{1}{\|\r\|}(I-\y\y^T)\mathbf{c}_2=0,
\end{align*}
which implies that $\mathbf{c}_2= c\r$ for an arbitrary constant $c$. For the given choice of $\mathbf{c}=[\r^T,c\r^T]^T$, the last three rows of \refeqn{Oc} are given by
\begin{align*}
\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
& = \mathcal{O}_{20} \r + c\mathcal{O}_{21} \r.
\end{align*}
Using \refeqn{O20r} and \refeqn{O21r}, this can be rewritten as
\begin{align*}
\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
& =
-2\mathcal{O}_{10}\dot\r +\mathcal{O}_{00}\braces{-\ddot \r +\deriv{\ddot\r}{\r}\r}
-2c\mathcal{O}_{00} (\dot\r+\omega\times\r).
\end{align*}
But, from \refeqn{O10dotr}, we can show that $2\mathcal{O}_{10}\dot\r=0$ when $\r$ is parallel to $\dot \r$. Using \refeqn{vrel} and \refeqn{a1}, this further reduces to
\begin{align}
\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
& =
-\mathcal{O}_{00}(\mathbf{a}_1 +2c \mathbf{v}_{rel})\nonumber\\
& = - \frac{1}{\|\r\|}(I-\y\y^T)(\mathbf{a}_1 +2c \mathbf{v}_{rel})=0.\label{eqn:row3i}
\end{align}
The matrix $(I-\y\y^T)$ represents the orthogonal projection of a vector into the plane normal to $\y$. From the second condition of \refeqn{cond1}, namely $\r\times\mathbf{a}_1\neq 0$, we have $(I-\y\y^T)\mathbf{a}_1\neq 0$, which implies that the constant $c$ cannot be simply chosen as $c=0$. Therefore, the only possible case to satisfy the above equation is when $\mathbf{a}_1 +2c \mathbf{v}_{rel}$ is parallel to $\r$ for some values of $c$. However, that is not feasible since the third condition of \refeqn{cond1}, namely $\r^T(\mathbf{v}_{ref}\times\mathbf{a}_1)\neq0$,  implies that the three vectors $\r$, $\mathbf{a}_1$, and $\mathbf{v}_{rel}$ do not belong to a common plane, i.e., there is no constant $c$ such that $\mathbf{a}_1 +2c \mathbf{v}_{rel}$ is parallel to $\r$.

Therefore, there is no $\mathbf{c}\in\Re^6$ satisfying \refeqn{Oc} if $\mathbf{c}_1=\r$, under the given condition \refeqn{cond1}. This implies that $\mathbf{c}_1=0$. Substituting this back to \refeqn{Oc}, we have $\mathcal{O}_{11}\mathbf{c}_2=\mathcal{O}_{00}\mathbf{c}_2=0$, which follows that $\mathbf{c}_2=c\r$ for some constant $c$. However, when $\mathbf{c}_2=c\r$, the last three rows of \refeqn{Oc} are given by
\begin{align}
\mathcal{O}_{21}\mathbf{c}_2 & = 2c(\mathcal{O}_{10}-\mathcal{O}_{00}[\omega]_\times) \r \nonumber
\\
& = -2c\mathcal{O}_{00}\mathbf{v}_{rel}=0,\label{eqn:row3i1}
\end{align}
where \refeqn{O10r} is used. But, from the first condition of \refeqn{cond1}, we have $\mathcal{O}_{00}\mathbf{v}_{rel}\neq 0$, and therefore $c=0$, i.e., $\mathbf{c}_2=c\r =0$.

In short, under the given condition \refeqn{cond1}, the equation \refeqn{Oc} implies that $\mathbf{c}=0$. Therefore, the six columns of the observability matrix $\mathcal{O}$ are linearly independent.

\paragraph{Case (ii): $\r\times\dot \r\neq0$} Next, we consider the second case of \refeqn{row2}. It implies that $-\dot\r + \mathbf{c}_2$ is parallel to $\r$, or equivalently, $\mathbf{c}_2 = \dot\r + c\r$ for an arbitrary constant $c$. For the given choice of $\mathbf{c}=[\r^T,\dot\r+ c\r^T]^T$, the last three rows of \refeqn{Oc} are given by
\begin{align*}
\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
& = (\mathcal{O}_{20} \r+\mathcal{O}_{21}\dot\r) + c\mathcal{O}_{21} \r.
\end{align*}
From \refeqn{O20r}, \refeqn{O21r}, \refeqn{O21rdot}, this can be rewritten as
\begin{align*}
\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
& = -\mathcal{O}_{00}(\mathbf{a}_2+2c\mathrm{v}_{ref})\\
& = -\frac{1}{\|\r\|}(I-\y\y^T)(\mathbf{a}_2+2c\mathbf{v}_{ref})=0.
\end{align*}
By following the same argument given after \refeqn{row3i}, under \refeqn{cond2}, there is no $c$ satisfying the above equation.

This implies that $\mathbf{c}_1=0$. Then, by the same argument given at \refeqn{row3i1}, we have $\mathbf{c}_2=0$. In short, under the given condition \refeqn{cond2}, the equation \refeqn{Oc} implies that $\mathbf{c}=0$. Therefore, the six columns of the observability matrix $\mathcal{O}$ are linearly independent. 
\end{proof}

\paragraph{Remarks} We consider the cases where the given sufficient conditions \refeqn{cond1} and \refeqn{cond2} are violated. 
For both cases, we have $\r\times\mathbf{v}_{rel}\neq 0$. At \refeqn{vrel}, $\mathbf{v}_{rel}$ corresponds to the relative velocity observed in the inertial frame. Therefore, the given sufficient conditions for observability are violated when the relative velocity vector is parallel to the relative position vector in the inertial frame.

The second condition of each of \refeqn{cond1} and \refeqn{cond2} is satisfied in general, as the expressions for $\mathbf{a}_1$ and $\mathbf{a}_2$ are relatively arbitrary at \refeqn{a1} and \refeqn{a2}.

The third condition of \refeqn{cond1} implies that three vectors $\r$, $\mathbf{v}_{rel}$ and $\mathbf{a}_1$ do not belong to the same plane. The third condition of \refeqn{cond2} has a similar structure as well. This condition can be easily violated if the relative motion is planar, i.e., when $z(t)\equiv 0$ for all $t$.

However, \refeqn{cond1} and \refeqn{cond2} are sufficient conditions for observability, and the fact that any of these condition is not satisfied does not necessarily mean that the considered point is not observable. In such case, the third or higher order Lie derivatives should be checked to determine observability. The main contribution of this section of the paper is showing that under certain geometric conditions, the nonlinear relative orbital dynamics are indeed observable via angles-only measurements.

\section{Observability Measure of Nonlinear Systems}

The measure of observability has been generalized to nonlinear dynamic systems in order to obtain balanced realizations.
An energy-like observability function is introduced~\cite{SchSCL93,Sch94,NewKriPICDC98} to measure the degree of contribution of an initial state to the output, and it is applied to a pendulum system. But, this approach is only applicable to asymptotically stable equilibrium states. An empirical observability Gramian is introduced~\cite{LalMarPIWC99,LalMarIJRNC02}, which is essentially a covariance matrix of the output computed by a number of sample trajectories.
It has been applied to a balanced realization of chemical processes~\cite{HahEdgJPC03,HahEdgCCE02}. However, this approach is based on the assumption that the output converges to a steady-state value, and it is not clear how  sample trajectories are selected.

A similar observability Gramian is introduced~\cite{KreIdePICDC09}, which is basically the Gramian for the sensitivity of the output with respect to the initial condition:
\begin{align}
\mathcal{W} (\x_0,t_0,t_f) = \int_{t_0}^{t_f} \parenth{\deriv{\y(\tau)}{\x_0}}^T\deriv{\y(\tau)}{\x_0}\,d\tau,\label{eqn:Wo_NL}
\end{align}
where $\y(t)$ denotes the output for the initial condition given by $\x(t_0)=\x_0$, and $\deriv{\y}{\x_0}\in\Re^{p\times n}$ is defined such that its $i,j$-th element is $\deriv{\y_i}{\x_{0_j}}$.

In general, the Gramian of a set of time-dependent functions represents the degree of linear independence of those functions over a given period of time. Therefore, the observability Gramian defined at \refeqn{Wo_NL} measures how much the sensitivities of the output with respect to the initial condition are linearly independent of each other. This represents the degree of observability, since for example, if $\deriv{\y}{\x_{0_i}}$ is linearly dependent to $\deriv{\y}{\x_{0_j}}$, then it is difficult to distinguish the effects on $\delta \x_{0_i}$ to the output $\y$ from the effects of $\delta \x_{0_j}$ on $\y$. Therefore, the condition number of $\mathcal{W} (\x_0,t_0,t_f)$ can be used as a measure of observability.

%Furthermore, we can easily show that \refeqn{Wo_NL} reduces to the observability Gramian for linear dynamic systems
%\begin{align}
%\mathcal{W}(t_0,t_f) = \int_{t_0}^{t_f}\Phi(\tau,t_0)^T C^T(\tau)  C(\tau) \Phi(\tau,t_0) d\tau,\label{eqn:Wo_Lin}
%\end{align}
%where $\Phi(t,\tau)$ is the state transition matrix, since the output is given by $\y(t) = \Phi(t,t_0)\x_0$.

A numerical approach is proposed as follows. The $i,j$--th element of $\mathcal{W}_o$ can be approximated by
\begin{align}
[\mathcal{W}]_{ij} = \frac{1}{4\epsilon^2}\sum_{k=0}^N (\y^{i+}_k-\y^{i-}_k)(\y^{j+}_k-\y^{j-}_k) \Delta t_k,
\end{align}
where $\y^{i\pm}_k$ denotes the value of $\y^i$ at $t=t_k$ with the initial condition of $\x(t_0)=\x_0\pm \epsilon \mathbf{e}_i $ for a positive constant $\epsilon$, and the standard basis $\mathbf{e}_i$ of $\Re^n$.  The above expression can be computed numerically.

\section{Kalman Filter}

%Preliminary results between the filter performance of an extended Kalman filter and the nonlinear observability Gramian are shown.

We consider the relative orbital dynamics between two satellites, where the observability of the deputy relative to the chief satellite is on a prescribed circular orbit when only $\mathbf{y}$ is assumed to be measured.
Then, an extended Kalman filter is developed for numerical analyses, where the satellites are either subject to two-body forces or with gravitational $J_2$ perturbations included.

The initial conditions for an estimate of the state vector $\hat \x_0^+$ and its initial covariance matrix $P^+_0$ are assumed to be known from an initial orbit determination (IOD) technique.

\paragraph{Flow Update}
The flow update provides an a priori state estimate $\hat \x^-_k$ with covariance matrix $P^-_k$ based on the equations of motion.
The linearized state transition matrix from \refeqn{xxdot} and \refeqn{f} from time step $k-1$ to time step $k$ is written as
\begin{align}
F_{k-1}=\exp\left(
\begin{bmatrix}
0_{3\times3} & I_{3\times3} \\
\left(-[\omega]_\times^2-\frac{\mu I_{3\times3}}{\norm{\r_a}^3}+\frac{3\mu \r_a \r_a^T}{\norm{\r_a}^5}\right) & -2[\omega]_\times
%0 & 0 & 0 & 1 & 0 & 0\\
%0 & 0 & 0 & 0 & 1 & 0\\
%0 & 0 & 0 & 0 & 0 & 1\\
%3*n^2 & 0 & 0 & 0 & 2*n & 0\\
%0 & 0 & 0 & -2*n & 0 & 0\\
%0 & 0 & -n^2 & 0 & 0 & 0
\end{bmatrix}
(t_k-t_{k-1})\right),
\end{align}
such that
\begin{align}
\hat \x_{k}^-=F_{k-1}\hat \x_{k-1}^+.
\end{align}
The process covariance matrix $Q_k$ is assumed known.
The a priori covariance matrix is defined as
\begin{align}
P^-_k=F_{k-1}P^+_{k-1}F_{k-1}^T+Q_k.
\end{align}

\paragraph{Measurement Update}
The effect of the measurement $\y_k$ with covariance matrix $R_k$ is included to determine the posterior state estimate $\hat \x^+_k$ and $P^+_k$.
The linearized measurement matrix $H_k$ is derived from \refeqn{y},
\begin{align}
H_k=
\begin{bmatrix}
\left(\frac1{\norm{\hat \r_{k}^-}}I_{3\times3}-\frac{\hat \r_{k}^-{\hat \r_{k}^{-T}}}{\norm{\hat \r_{k}^-}^3}\right) & 0_{3\times3}
\end{bmatrix},
\end{align}
where $\hat \r_{k}^-$ is the a priori estimate of $\r$ given by the first three elements of $\hat \x_{k}^-$.
Based on \refeqn{y}, the measurement prediction is given by
\begin{align}
\hat{\y}_k=h(\hat \r_{k}^-)=\frac{\hat \r_{k}^-}{\norm{\hat \r_{k}^-}}.
\end{align}
The posterior state estimate and its covariance matrix are given by
\begin{align}
\x^+_k&=\x^-_k+K_k(\y_k-\hat{\y}_k),
\\
P^+_k&=(I_{6\times6}-K_kH_k)P^-_k,
\end{align}
such that $K_k$ is chosen to minimize $\tr{P^+_k}$ with
\begin{align}
K_k=P^-_kH_k^T(H_kP^-_kH_k+R_k)^{-1}.
\end{align}

%The initial conditions of the deputy satellite are chosen as
%\begin{align*}
%\mathbf{r}_0=[r_{x_0},\, 0,\, 0]^T\,\mathrm{km},\quad \mathbf{v}_0=[0,\, -2r_{x_0}n,\, 0]^T\,\mathrm{km/s},
%\end{align*}
%for a varying $r_{x_0}$. These yield a series of elliptic orbits around the chief when applied to the linearized relative orbital dynamics. The state estimation error is represented by a time averaged normalized magnitude error, $e_{mag} = \frac{\|\mathbf{x}-\hat{\mathbf{x}}\|}{\|\mathbf{x}\|}$, where $\mathbf{x}=[\mathbf{r}^T,\dot{\mathbf{r}}^T]^T\in\Re^6$ denotes the true state and $\hat{\mathbf{x}}$ denotes the estimated state. 
%
%The normalized magnitude error $e_{mag}$ and the condition number of the proposed observability measure $\mathrm{cond}\braces{\mathcal{W}}$ are computed among various values for $r_{x_0}$, tabulated below.
%
%\hspace*{0.1cm}\centerline{
%\begin{tabular}{ccc}\hline
%$r_{x_0}$ & $e_{mag}$ & $\mathrm{cond}\braces{\mathcal{W}}$\\\hline
%\vspace*{-0.02\columnwidth}\\
%$1$ & $11.5962$ &  $10^{16.6653}$\\
%$2$ & $9.6374$ & $10^{16.8323}$\\
%$5$ & $4.0041$ & $10^{16.3535}$\\
%$10$ & $2.6318$ &  $10^{15.8194}$\\
%$20$ & $0.9250$ &  $10^{15.2832}$\\
%$50$ & $0.5315$ &  $10^{14.4052}$\\
%$100$ & $0.1675$ &  $10^{12.9884}$\\
%\hline
%\end{tabular}
%}
%
%These preliminary results illustrate that there exists a certain correspondence between the filter error and the proposed observability measure, and therefore, it can be used as a quantitative measure for strength of observability.

\section{Numerical Results}

We choose numerous cases of relative motion between chief and deputy satellite orbits.
With each case, we evaluate the observability measure and process angles-only measurements with the EKF, using both the two-body equations of motion and higher fidelity propagation with $J_2$ effects as the truth models.
%In particular, the $J_2$ gravitational harmonics will be included in the satellite propagation, so the EKF must handle simulated process uncertainty.
To begin each simulation, an initial orbit determination (IOD) estimate is determined by various angles-only observations.
These IOD estimates serve as initial estimates for the EKF.
Then, we evaluate the performance of the EKF for these scenarios, specifically focusing on the observability measure and issues arising from initial conditions.

Two IOD methods are developed in~\cite{NewLovPra14,PraLovNew14}, referred to as \emph{IOD1} and \emph{IOD2}, and employed to provide initial conditions for $13$ cases, outlined in Table \ref{tab:IODerr}.
%Two IOD methods are referred to as \emph{IOD1} and \emph{IOD2}, and employed to provide initial conditions for $13$ cases, outlined in Table \ref{tab:IODerr}.
These cases yield relative orbit trajectories illustrated in Figures \ref{fig:0CasesTraj}, \ref{fig:1CasesTraj}, \ref{fig:2CasesTraj}, \ref{fig:3CasesTraj}, and \ref{fig:4CasesTraj}.
The most important distinction between the various cases is that in Case $\mathbf{1}$ the two-body solution and the linearized dynamics of the Hill-Clohessy-Wiltshire (HCW) solution are very close and the trajectory remains largely planar with a circular shape.

\begin{figure}[h]
\centerline{
	\subfigure[Case 0A]{
		\includegraphics[width=0.3\textwidth]{OM_IOD_0A_traj.pdf}}
	\hfill
	\subfigure[Case 0B]{
		\includegraphics[width=0.3\textwidth]{OM_IOD_0B_traj.pdf}}
	\hfill
	\subfigure[Case 0C]{
		\includegraphics[width=0.3\textwidth]{OM_IOD_0C_traj.pdf}}
}
\caption{The trajectories of the Cases 0A, 0B, and 0C (solid red: HCW, dashed blue: two-body solution) show that the linear and nonlinear solutions are close.
The resulting trajectories are not cyclic as the trajectories change in the $y$-direction with time.}
\label{fig:0CasesTraj}
\end{figure}

\begin{figure}[h]
\label{fig:CaseTraj}
\centerline{
	\subfigure[Case 1]{
		\includegraphics[width=0.3\textwidth]{OM_IOD_1_traj.pdf}}
}
\caption{The trajectory of the Case 1 (solid red: HCW, dashed blue: two-body solution) shows that the linear and nonlinear solutions are very close, where both solutions are largely planar and cyclic.}
\label{fig:1CasesTraj}
\end{figure}

\begin{figure}[h]
\centerline{
	\subfigure[Case 2A]{
		\includegraphics[width=0.3\textwidth]{OM_IOD_2A_traj.pdf}}
	\hfill
	\subfigure[Case 2B]{
		\includegraphics[width=0.3\textwidth]{OM_IOD_2B_traj.pdf}}
	\hfill
	\subfigure[Case 2C]{
		\includegraphics[width=0.3\textwidth]{OM_IOD_2C_traj.pdf}}
}
\caption{The trajectories of the Cases 2A, 2B, and 2C (solid red: HCW, dashed blue: two-body solution) show differences between the linear and nonlinear solutions that vary about the $y$-axis, increasing the observabilities.}
\label{fig:2CasesTraj}
\end{figure}

\begin{figure}[h]
\centerline{
	\subfigure[Case 3A]{
		\includegraphics[width=0.3\textwidth]{OM_IOD_3A_traj.pdf}}
	\hfill
	\subfigure[Case 3B]{
		\includegraphics[width=0.3\textwidth]{OM_IOD_3B_traj.pdf}}
	\hfill
	\subfigure[Case 3C]{
		\includegraphics[width=0.3\textwidth]{OM_IOD_3C_traj.pdf}}
}
\caption{The trajectories of the Cases 3A, 3B, and 3C (solid red: HCW, dashed blue: two-body solution) exhibit large observabilities due to nonlinear solutions that move far away from the linearized solutions.}
\label{fig:3CasesTraj}
\end{figure}

\begin{figure}[h]
\centerline{
	\subfigure[Case 4A]{
		\includegraphics[width=0.3\textwidth]{OM_IOD_4A_traj.pdf}}
	\hfill
	\subfigure[Case 4B]{
		\includegraphics[width=0.3\textwidth]{OM_IOD_4B_traj.pdf}}
	\hfill
	\subfigure[Case 4C]{
		\includegraphics[width=0.3\textwidth]{OM_IOD_4C_traj.pdf}}
}
\caption{The trajectories of the Cases 4A, 4B, and 4C (solid red: HCW, dashed blue: two-body solution) experience the largest observabilities due to nonlinear solutions that move farther away from the linearized solutions than any of the trajectories from other cases considered in this paper.}
\label{fig:4CasesTraj}
\end{figure}

The initial error functions between the true orbit and and IOD predictions are defined as
\begin{align}
e_{mag,0}=\abs\left(1-{\frac{\norm{\r^+_0}}{\norm{\r_0}}}\right), \quad e_{dir,0}=\cos^{-1}\left(\frac{\r_0}{\norm{\r_0}}\cdot \frac{\r^+_0}{\norm{\r^+_0}}\right).
\end{align}

\begin{center}
\begin{threeparttable}[h]
\caption{Case Observabilities and Initial Errors}
\begin{tabularx}{0.72\textwidth}
{
>{$}c<{$}
*{1}{>{$}c<{$}} |
*{2}{>{$}c<{$}} |
*{2}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{\multirow{1}{*}{IOD1}} & \multicolumn{2}{c}{\multirow{1}{*}{IOD2}} \\
& &  e_{mag,0} & e_{dir,0} & e_{mag,0} & e_{dir,0} \\\midrule
0A & 10^{11.1486} & 0.0003 & 1.4901\times10^{-8} & 0.0012  & 2.1073\times10^{-8} \\
0B & 10^{10.9387} & 0.0233 & 0 & 0.0088 & 0 \\
0C & 10^{10.9851} & 0.0152 & 2.1073\times10^{-8} & 0.0139 & 1.4901\times10^{-8}  \\
\midrule
\mathbf{1} & 10^{16.6871} & 0.0001 & 0 & 0 & 0  \\
\midrule
2A & 10^{10.6492} & 0.0550 & 1.8150\times10^{-4} & 0.0238 & 1.8150\times10^{-4}  \\
2B & 10^{10.4163} & 0.9571 & 0.0734                     & 0.9571 & 0.0734  \\
2C & 10^{10.2155} & 0.0219 & 0.0044 & 0.0065 & 0.0044 \\
\midrule
3A & 10^{8.3490} & 0.0191 & 0.0031 & 0.1830 & 0.0031  \\
3B & 10^{8.5372} & 0.0612 & 0.0533 & 0.2315 & 0.0533  \\
3C & 10^{8.6586} & 6.8962 & 0.7348 & 5.4921 & 0.7348  \\
\midrule
4A & 10^{7.9062} & 8.7347 & 0.4587 & 0.1575 & 0.4587  \\
4B & 10^{7.9903} & 0.9276 & 0.1819 & 0.0621 & 0.1819  \\
4C & 10^{8.0645} & 0.1161 & 0.0554 & 0.0354 & 0.0554  \\
\bottomrule
\end{tabularx}
{\small
\begin{tablenotes}
    \item Note: $0$ indicates numbers small enough to be below Matlab numerical accuracy.
  \end{tablenotes}}
\label{tab:IODerr}
\end{threeparttable}
\end{center}

The Kalman filter is used in each case. The initial state covariance matrix is chosen as
\begin{align*}
P_0^+=\diag[50^2,\ 50^2,\ 50^2,\ (50n)^2,\ (50n)^2,\ (50n)^2],
\end{align*}
and the process and measurement covariance matrices are chosen as
\begin{align*}
Q_k=\diag[10^{-7},\ 10^{-7},\ 10^{-7},\ 10^{-9},\ 10^{-9},\ 10^{-9}], \quad R_k=\tan^2{2^\circ}I_{3\times3}\ \forall\ k.
\end{align*}
%\begin{align*}
%Q_k=10^q\diag[10^{-8},\ 10^{-8},\ 10^{-8},\ 10^{-10},\ 10^{-10},\ 10^{-10}], \quad R_k=\tan^2{2^\circ}I_{3\times3}\ \forall\ k
%\end{align*}
%where $q\in[-1,\ 0,\ 1,\ 2]$ is varied in the simulations, thereby changing the impact of measurement updates inside the Kalman filter.

Each case is simulated for every IOD estimate. The running time is $59540$ seconds (slightly more than $\frac23$ of a day) with time steps of $10$ seconds between measurements. The following RMS errors are collected from each trial:
\begin{align}
\bar e_{mag}=\left(\frac1N\sum_{k=1}^N\left\{1-{\frac{\norm{\r^+_k}}{\norm{\r_k}}}\right\}^2\right)^{\frac12}, \quad \bar e_{dir}=\left(\frac1N\sum_{k=1}^N
\left\{\cos^{-1}\left(\frac{\r_k}{\norm{\r_k}}\cdot \frac{\r^+_k}{\norm{\r^+_k}}\right)\right\}^{2}
\right)^{\frac12},
\end{align}
for $N$ time steps.
These metrics are tabulated in Tables \ref{tab:ResultsTBP} and \ref{tab:ResultsJ2}.
Over the simulated time period, figures were generated for these analyses that  serve to compare the true and estimated state magnitudes and the maximum eigenvalues of the state covariance matrices as a measure of uncertainty for the state estimates.
Because of paper constraints, a sample of these plots are shown that highlight various aspects of the observability and filter performance.

%The figure below illustrates preliminary results for angles-only relative orbit estimation with $J_2$ perturbation in the measurements.
%Note that the EKF performs well in this particular case, as evidenced by the close agreement between the true and estimated trajectory (Figure 1a) and the decrease in the maximum eigenvalue of the covariance matrix over time (Figure 1b).
%For the various cases to be chosen, we will evaluate the performance of the EKF (using metrics such as those illustrated in Figure 1) and seek a consistent correlation between the filter performance and the new observability measure proposed above.

%\begin{center}
%\begin{threeparttable}[h]
%\caption{Error variables for EKF with varying $Q_k$ Considering Only Two-Body Problem Forces}
%\begin{tabularx}{\textwidth}
%{
%>{$}c<{$}
%*{1}{>{$}c<{$}} |
%*{4}{>{$}c<{$}} |
%*{4}{>{$}c<{$}}
%}
%% & \bar e_{mag} & \bar e_{dir}
%\toprule
%\multirow{2}{*}{Case} & \multirow{2}{*}{IOD} & \multicolumn{4}{c}{$\bar e_{mag}$} & \multicolumn{4}{c}{$\bar e_{dir}$} \\
%& & q=-1 & q=0 & q=1 & q=2 & q=-1 & q=0 & q=1 & q=2 \\\midrule
%0A & IOD 1 & 0.0895 &  0.0611  &  0.3766  &  0.2861  &  0.0430  &  0.0135  &  0.0144  &  0.0183 \\
%0A & IOD 2 & 0.1793 &  0.2811  &  0.5145  &  0.7716  &  0.0203  &  0.0313  &  0.0691  &  0.1046 \\
%0B & IOD 1 & 0.3241 &  0.0531  &  0.5186  &  0.0620  &  0.1695  &  0.0143  &  0.0810  &  0.0172 \\
%0B & IOD 2 & 0.2330 &  0.2115  &  0.3092  &  0.2733  &  0.1007  &  0.0879  &  0.0804  &  0.1048 \\
%0C & IOD 1 & 0.0973 &   0.0748 &   0.2723 &   0.0607 &   0.0292 &   0.0068  &  0.1362 &   0.0443 \\
%0C & IOD 2 & 0.1473 &   0.1017 &   0.1707 &   0.3545  &  0.0702  &  0.0517  &  0.0614  &  0.1224 \\
%\midrule
%\mathbf{1} & IOD 1 & \hilight{8.922} &   \hilight{2.964}   & \hilight{4.078}  &  \hilight{3.571} &   0.0071  &  0.0089  &  0.0109  &  0.0418 \\
%\mathbf{1} & IOD 2 & \hilight{13.635} &  \hilight{14.121} &  \hilight{42.486} &  \hilight{42.573} &   0.0113 &   0.0107  &  0.0445  &  0.0115 \\
%\midrule
%2A & IOD 1 & 0.2045 &   0.0629 &   0.1592  &  0.1329  &  0.0041 &   0.0041  &  0.0043  &  0.0039 \\
%2A & IOD 2 & 0.1013 &   0.1797  &  0.1348  &  0.2247  &  0.0050 &   0.0048 &   0.0065 &   0.0061 \\
%2B & IOD 1 & \hilight{0.9886} &   \hilight{0.9895}   & \hilight{0.9843}  &  \hilight{0.9851}  &  0.0114 &   0.0116  &  0.0114  &  0.0116 \\
%2B & IOD 2 & \hilight{0.9557} &   \hilight{0.9685} &   \hilight{0.9278}  &  \hilight{0.9431} &   0.0105  &  0.0108 &   0.0113 &   0.0153 \\
%2C & IOD 1 & 0.0219 &   0.0184  &  0.0258  &  0.0238 &   0.0026 &   0.0031  &  0.0044 &   0.0038 \\
%2C & IOD 2 & 0.0303 &   0.0287  &  0.0360  &  0.0470  &  0.0044  &  0.0054  &  0.0059 &   0.0062 \\
%\midrule
%3A & IOD 1 &  0.2377 &   0.1684  &  0.0486 &   0.0695 &   0.0094 &   0.0070 &   0.0040 &  0.0039 \\
%3A & IOD 2 &  0.0854  &  0.0874 &   0.0603  &  0.0544  &  0.0035 &   0.0033  &  0.0053  &  0.0042 \\
%3B & IOD 1 &  0.0724 &   0.0260 &   0.0680  &  0.0289  &  0.0035 &   0.0029  &  0.0033  &  0.0027 \\
%3B & IOD 2 &  0.0385  &  0.0432  &  0.1459  &  0.0284  &  0.0038  &  0.0037  &  0.0070  &  0.0047 \\
%3C & IOD 1 &  \hilight{2.2433}  &  \hilight{1.5231} &   \hilight{2.0611} &   \hilight{1.5110}  &  0.1803  &  0.1172  &  0.1595  &  0.1103 \\
%3C & IOD 2 &  \hilight{4.0635} &   \hilight{1.7717}  & \hilight{10.567}  &  \hilight{4.0277} &   0.4475 &   0.1506 &   0.9011 &   0.4028 \\
%\midrule
%4A & IOD 1 &  \hilight{6.0259} &   \hilight{0.0464}  &  \hilight{6.0706}  &  \hilight{0.0325}  &  1.3353  &  0.0087   & 1.3182  &  0.0065 \\
%4A & IOD 2 &  6.3646 &   0.0173  & 11.8882 &   0.0256 &   1.3423  &  0.0064  &  1.5319 &   0.0066 \\
%4B & IOD 1 &  \hilight{0.2922}  &  \hilight{0.0258}  &  \hilight{0.2723}  &  \hilight{0.0182}   & 0.0929 &   0.0032  &  0.0789  &  0.0036 \\
%4B & IOD 2 &  0.1815  &  0.0217  &  0.3155 &   0.0771  &  0.0268   & 0.0034  &  0.1061 &   0.0044 \\
%4C & IOD 1 &  0.0247  &  \hilight{0.0421}  &  2.3699  &  0.0182  &  0.0032  &  0.0032  &  0.8486 &   0.0028 \\
%4C & IOD 2 &  0.0200 &   0.0128  &  0.0278  &  0.0299 &   0.0040  &  0.0029  &  0.0034  &  0.0033 \\
%\bottomrule
%\end{tabularx}
%{\small
%\begin{tablenotes}
%    \item \hilight{Filter diverges}
%  \end{tablenotes}}
%\label{tab:ResultsTBP}
%\end{threeparttable}
%\end{center}
%
%\begin{center}
%\begin{threeparttable}[h]
%\caption{Error variables for EKF with varying $Q_k$ with $J_2$ Perturbations Included}
%\begin{tabularx}{\textwidth}
%{
%>{$}c<{$}
%*{1}{>{$}c<{$}} |
%*{4}{>{$}c<{$}} |
%*{4}{>{$}c<{$}}
%}
%\toprule
%\multirow{2}{*}{Case} & \multirow{2}{*}{IOD} & \multicolumn{4}{c}{$\bar e_{mag}$} & \multicolumn{4}{c}{$\bar e_{dir}$} \\
%& & q=-1 & q=0 & q=1 & q=2 & q=-1 & q=0 & q=1 & q=2 \\\midrule
%0A & IOD 1 & 0.0553  &  0.1105  &  0.1918  &  0.2298  &  0.0059  &  0.0062  &  0.0069  &  0.0069 \\
%0A & IOD 2 & 0.2177  &  0.3227  &  0.3140  &  0.5013  &  0.0082  &  0.0074  &  0.0129  &  0.0085 \\
%0B & IOD 1 & 0.0473  &  0.1246  &  0.7262  &  0.0655  &  0.0077  &  0.0088  &  0.0505  &  0.0066\\
%0B & IOD 2 & 0.8872  &  0.4186  &  0.1571  &  0.1561  &  0.0390  &  0.0168  &  0.0086  &  0.0293 \\
%0C & IOD 1 & 0.0652  &  0.0809  &  0.0900  &  0.1113  &  0.0082  &  0.0079  &  0.0061  &  0.0106 \\
%0C & IOD 2 & 0.0367  &  0.0204  &  0.1002  &  0.7253  &  0.0145  &  0.0084  &  0.0154  &  0.0579 \\
%\midrule
%\mathbf{1} & IOD 1 & \hilight{9.1098}  &  \hilight{2.900}  &  \hilight{4.059}  &  \hilight{3.567}  &  0.0071  &  0.0090  &  0.0109  &   0.0418 \\
%\mathbf{1} & IOD 2 & \hilight{13.594} & \hilight{14.033} & \hilight{42.381} & \hilight{42.979}  & 0.0113   & 0.0107   & 0.0445   &   0.0115 \\
%\midrule
%2A & IOD 1 & 0.2301  &  0.0423  &  0.1978  &  0.1740  &  0.0048  &  0.0049  &  0.0047  &  0.0042 \\
%2A & IOD 2 & 0.1058  &  0.2164  &  0.1638  &  0.2644  &  0.0051  &  0.0049  &  0.0065  &  0.0061 \\
%2B & IOD 1 & \hilight{0.9895}  &  \hilight{0.9902}  &  \hilight{0.9853}  &  \hilight{0.9859}  &  0.0115  &  0.0117  &  0.0115  &  0.0117 \\
%2B & IOD 2 & \hilight{0.9585}  &  \hilight{0.9701}  &  \hilight{0.9307}  &  \hilight{0.9455}  &  0.0106  &  0.0109  &  0.0114  &  0.0153 \\
%2C & IOD 1 & 0.0481  &  0.0474  &  0.0563  &  0.0415  &  0.0041  &  0.0042  &  0.0047  &  0.0042 \\
%2C & IOD 2 & 0.0535  &  0.0397  &  0.0536  &  0.0705  &  0.0044  &  0.0056  &  0.0059  &  0.0062 \\
%\midrule
%3A & IOD 1 &  0.2462  &  0.1827  &  0.0516  &  0.0699 &   0.0097 &   0.0077  &  0.0041 &   0.0041 \\
%3A & IOD 2 &  0.1038  &  0.0986  &  0.0643  &  0.0692 &   0.0036  &  0.0034  &  0.0053  &  0.0041 \\
%3B & IOD 1 &  0.0770  &  0.0310  &  0.0739  &  0.0458 &   0.0038 &   0.0034  &  0.0035  &  0.0031 \\
%3B & IOD 2 &  0.0461  &  0.0585   & 0.1485 &   0.0391  &  0.0039  &  0.0040  &  0.0071  &  0.0047 \\
%3C & IOD 1 &  \hilight{2.239}  &  \hilight{1.522}  &  \hilight{2.063}  &  \hilight{1.514}  &  0.1817  &  0.1179  &  0.1614  &  0.1116 \\
%3C & IOD 2 &  \hilight{4.188}  &  \hilight{1.821}  & \hilight{10.136}  &  \hilight{4.510} &   0.4873 &   0.1592  &  0.9195 &   0.5532 \\
%\midrule
%4A & IOD 1 &  \hilight{6.013}  &  \hilight{0.0454}  &  \hilight{6.359}  &  \hilight{0.031}  &  1.3131  &  0.0086  &  1.4235 &   0.0066 \\
%4A & IOD 2 &  6.3272  &  0.0197  & 12.0062 &  0.0248  &  1.3116  &  0.0065  &  1.5310  &  0.0066 \\
%4B & IOD 1 &  \hilight{0.2936}  &  \hilight{0.0272}  &  \hilight{0.2773}  &  \hilight{0.0217}  &  0.0935  &  0.0032  &  0.0814  &  0.0038 \\
%4B & IOD 2 &  0.1853  &  0.0210  &  0.5244  &  0.0785  &  0.0293  &  0.0036  &  0.2305  &  0.0045 \\
%4C & IOD 1 &  0.0260  &  \hilight{0.0410}  &  2.0105  &  0.0164  &  0.0034  &  0.0031  &  0.9144  &  0.0030 \\
%4C & IOD 2 &  0.0210  &  0.0145  &  0.0303  &  0.0349  &  0.0042  &  0.0030  &  0.0036  &  0.0034 \\
%\bottomrule
%\end{tabularx}
%{\small
%\begin{tablenotes}
%    \item \hilight{Filter diverges}
%  \end{tablenotes}}
%\label{tab:ResultsJ2}
%\end{threeparttable}
%\end{center}

\begin{center}
\begin{threeparttable}[h]
\caption{Error variables for EKF Considering Only Two-Body Problem Forces}
\begin{tabularx}{.58\textwidth}
{
>{$}c<{$}
*{1}{>{$}c<{$}} |
*{2}{>{$}c<{$}} |
*{2}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{IOD1} & \multicolumn{2}{c}{IOD2} \\
& & \bar e_{mag} & \bar e_{dir} & \bar e_{mag} & \bar e_{dir} \\\midrule
0A & 10^{11.1486} &  0.3766  &  0.0144  &  0.5145  &  0.0691 \\
0B & 10^{10.9387} &  0.5186  &  0.0810  &  0.3092  &  0.0804\\
0C & 10^{10.9851} &  0.2723  &  0.1362 &  0.1707 &  0.0614\\
\midrule
\mathbf{1} & 10^{16.6871} & \cellcolor{green} 4.078  & \cellcolor{green}  0.0109 & \cellcolor{green} 42.486 & \cellcolor{green}  0.0445 \\
\midrule
2A & 10^{10.6492} &  0.1592  &   0.0043  &  0.1348  &   0.0065   \\
2B & 10^{10.4163} & \cellcolor{green} 0.9843   & \cellcolor{green}  0.0114  & \cellcolor{green}  0.9278 & \cellcolor{green}   0.0113 \\
2C & 10^{10.2155} & 0.0258 &  0.0044 & 0.0360 &  0.0059 \\
\midrule
3A & 10^{8.3490} &  0.0486 & 0.0040   &  0.0603  &  0.0053 \\
3B & 10^{8.5372} &  0.0680  &  0.0033 &  0.1459 &  0.0070 \\
3C & 10^{8.6586} & \cellcolor{green}  2.0611 & \cellcolor{green}  0.1595   & \cellcolor{green}  10.567  & \cellcolor{green}   0.9011 \\
\midrule
4A & 10^{7.9062} & \cellcolor{green}  6.0706  & \cellcolor{green} 1.3182  &  11.8882   &  1.5319  \\
4B & 10^{7.9903} & \cellcolor{green}  0.2723  & \cellcolor{green}  0.0789   &  0.3155  &  0.1061 \\
4C & 10^{8.0645} &  2.3699 &  0.8486   &  0.0278 &  0.0034 \\
\bottomrule
\end{tabularx}
{\small
\begin{tablenotes}
    \item \hilight{Filter diverges}
  \end{tablenotes}}
\label{tab:ResultsTBP}
\end{threeparttable}
\end{center}

\begin{center}
\begin{threeparttable}[h]
\caption{Error variables for EKF Considering $J_2$ Perturbations}
\begin{tabularx}{.58\textwidth}
{
>{$}c<{$}
*{1}{>{$}c<{$}} |
*{2}{>{$}c<{$}} |
*{2}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{IOD1} & \multicolumn{2}{c}{IOD2} \\
& & \bar e_{mag} & \bar e_{dir} & \bar e_{mag} & \bar e_{dir} \\\midrule
0A & 10^{11.1486} &  0.1918  &  0.0069  &  0.3140  &  0.0129 \\
0B & 10^{10.9387} &  0.7262  &  0.0505  &  0.1571  &  0.0086\\
0C & 10^{10.9851} &  0.0900  &  0.0061 &  0.1002  &  0.0154\\
\midrule
\mathbf{1} & 10^{16.6871} & \cellcolor{green} 4.059  & \cellcolor{green}  0.0109 & \cellcolor{green} 42.381 & \cellcolor{green}  0.0445 \\
\midrule
2A & 10^{10.6492} &  0.1978  &  0.0047  &  0.1638  &   0.0065   \\
2B & 10^{10.4163} & \cellcolor{green}  0.9853  & \cellcolor{green}  0.9307 & \cellcolor{green}  0.9278 & \cellcolor{green}   0.0114 \\
2C & 10^{10.2155} & 0.0563 &  0.0047 & 0.0536 &  0.0059 \\
\midrule
3A & 10^{8.3490} &  0.0516 & 0.0041   &  0.0643  &  0.0053 \\
3B & 10^{8.5372} &  0.0739  &  0.0035 &  0.1485 &  0.0071 \\
3C & 10^{8.6586} & \cellcolor{green}  2.063 & \cellcolor{green} 0.1614    & \cellcolor{green}  10.136  & \cellcolor{green}   0.9195 \\
\midrule
4A & 10^{7.9062} & \cellcolor{green}  6.359  & \cellcolor{green} 1.4235  &  12.0062   &  1.5310  \\
4B & 10^{7.9903} & \cellcolor{green}  0.2773  & \cellcolor{green}  0.0814   &  0.5244  &  0.2305 \\
4C & 10^{8.0645} &  2.0105 &  0.9144   &  0.0303 &  0.0036 \\
\bottomrule
\end{tabularx}
{\small
\begin{tablenotes}
    \item \hilight{Filter diverges}
  \end{tablenotes}}
\label{tab:ResultsJ2}
\end{threeparttable}
\end{center}

These data from the simulations show several important aspects and challenges with relative orbit determination involving angles-only measurements.
First, an important observation from the data is that the IOD guess is critical to the success of the Kalman filter.
Cases 2B, 3C, 4A, and 4B experience filter divergence when the IOD estimate is largely inaccurate, shown in Tables \ref{tab:IODerr}, \ref{tab:ResultsTBP}, and \ref{tab:ResultsJ2}, even when cases with similar observabilities do not experience filter divergence.
The Kalman filter, which is based on linear dynamic systems, is modeled as a Markov chain and hence is memoryless.
Therefore when nothing more than an inaccurate IOD prediction is poorly assumed to contain accurate information about the system, the recursive formulation of the Kalman filter is unable to converge.
In these diverging cases, a greater level of observability or an algorithm that relies on a larger history of measurements may improve convergence.

In contrast, every variation of Case $\mathbf{1}$ has a closer IOD than the IODs of other cases, but experiences the most severe divergence.
Case $\mathbf{1}$ has a greater condition number of $\mathcal W$, which corresponds with low observability according to the proposed observability measure.
A series of angles-only measurements may correspond with a large range of possible relative orbits, even with a low variation in those measurements.
Therefore, small levels of noise in the sensor or uncertainties in the orbit prevent the Kalman filter from converging to an accurate estimate, even with an accurate IOD.
The state magnitudes of both variations of Case $\mathbf{1}$ are illustrated against various converging cases in Figure \ref{fig:ExampesEachCase} to show how important this observability is to the filter performance.

\begin{figure}[h]
\centerline{
	\subfigure[Case 0C, IOD1]{
		\includegraphics[width=0.35\textwidth]{Case0Cq1IOD1TBPnormX.pdf}}
	\hfill
	\subfigure[Case 0C, IOD2]{
	\includegraphics[width=0.35\textwidth]{Case0Cq1IOD2TBPnormX.pdf}}
}
\centerline{
	\subfigure[Case $\mathbf{1}$, IOD1]{
		\includegraphics[width=0.35\textwidth]{Case1q1IOD1TBPnormX.pdf}}
	\hfill
	\subfigure[Case $\mathbf{1}$, IOD2]{
	\includegraphics[width=0.35\textwidth]{Case1q1IOD2TBPnormX.pdf}}
}
\centerline{
	\subfigure[Case 2A, IOD1]{
		\includegraphics[width=0.35\textwidth]{Case2Aq1IOD1TBPnormX.pdf}}
	\hfill
	\subfigure[Case 2A, IOD2]{
	\includegraphics[width=0.35\textwidth]{Case2Aq1IOD2TBPnormX.pdf}}
}
\centerline{
	\subfigure[Case 3A, IOD1]{
		\includegraphics[width=0.35\textwidth]{Case3Aq1IOD1TBPnormX.pdf}}
	\hfill
	\subfigure[Case 3A, IOD2]{
	\includegraphics[width=0.35\textwidth]{Case3Aq1IOD2TBPnormX.pdf}}
}
\centerline{
	\subfigure[Case 4C, IOD1]{
		\includegraphics[width=0.35\textwidth]{Case4Cq1IOD1TBPnormX.pdf}}
	\hfill
	\subfigure[Case 4C, IOD2]{
	\includegraphics[width=0.35\textwidth]{Case4Cq1IOD2TBPnormX.pdf}}
}
\caption{The magnitudes of the relative states (solid red: true, blue dashed: estimated) are shown for various two-body propagated cases with close IOD estimates where the filter only diverges when the observability is low.}\label{fig:ExampesEachCase}
\end{figure}

Another observation from the data is that the $J_2$ perturbation, which is well-known to have a noticeable impact perturbing orbits from the two-body model for satellites around the earth, yields an insignificant change when compared to the data when only two-body forces are considered in the process model.
In fact, every case when the filter diverged occurred with both the two-body problem and $J_2$ simulations.
Examples of these results are tabulated in Tables \ref{tab:ResultsTBP} and \ref{tab:ResultsJ2} and illustrated in Figure \ref{fig:TBPvsJ2}.

\begin{figure}[h]
\centerline{
	\subfigure[State magnitude with the two-body solution]{
		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1TBPnormX.pdf}}
	\hfill
	\subfigure[State magnitude with the $J_2$ solution]{
		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1HighFidelitynormX.pdf}}
}
\centerline{
	\subfigure[Uncertainty measure with the two-body solution]{
		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1TBPeigP.pdf}}
	\hfill
	\subfigure[Uncertainty measure with the $J_2$ solution]{
		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1eigP.pdf}}
}
\caption{The magnitudes and maximum eigenvalues of the state covariance matrices of the relative states (solid red: true, blue dashed: estimated) in Case 3B with IOD1 serve as an example where the $J_2$ propagated simulations give very close results to those of the two-body solution.
}\label{fig:TBPvsJ2}
\end{figure}

\section{Conclusions}

It is well known that the linearized relative orbital dynamics are not observable with angles-only
measurements.
This paper formulates a measure of nonlinear observability, and investigates a less conservative observability criteria by using higher-order Lie derivatives.
Numerous relative orbit determination cases are illustrated numerically by an extended Kalman filter, which shows that the degree of observability by the proposed measure and an accurate IOD prediction are useful for predicting filter convergence.

\bibliographystyle{AAS_publication}   % Number the references.
\bibliography{references}   % Use references.bib to resolve the labels.

\end{document}

### 2018-06-09 Conference/AAS 2015 OD EKF/AAS15_.tex
added Conference folder

--- 2018-06-09 Conference/AAS 2015 OD EKF/AAS15.tex
@@
 
-%\documentclass[letterpaper, preprint, paper,11pt]{AAS}	% for preprint proceedings
+\documentclass[letterpaper, preprint, paper,11pt]{AAS}	% for preprint proceedings
 %\documentclass[letterpaper, paper,11pt]{AAS}		% for final proceedings (20-page limit)
 %\documentclass[letterpaper, paper,12pt]{AAS}		% for final proceedings (20-page limit)
-\documentclass[letterpaper, paper,10pt]{AAS}		% for final proceedings (20-page limit)
+%\documentclass[letterpaper, paper,10pt]{AAS}		% for final proceedings (20-page limit)
 %\documentclass[letterpaper, submit]{AAS}			% to submit to JAS
@@
 \usepackage{amssymb,amsmath,amsthm,times,graphicx,subfigure,tabularx,booktabs,colortbl,multirow,threeparttable}
+
 \usepackage{subfigure}
@@
 
-\usepackage{color}
-\newcommand{\hilight}[1]{\colorbox{green}{#1}}
-
-\newtheorem{prop}{Proposition}
-
-\newcommand{\abs}{\mathop{\mathrm{abs}}\nolimits}
-\newcommand{\diag}{\mathop{\mathrm{diag}}\nolimits}
 \newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
@@
 
-\PaperNumber{23-451}
+\PaperNumber{XX-XXX}
 
@@
 \ and Taeyoung Lee\thanks{Assistant Professor, Mechanical and Aerospace Engineering, George Washington University, 801 22nd St NW, Washington, DC 20052, Tel: 202-994-8710, Email: tylee@gwu.edu.}
-\thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}
 }
@@
 It corresponds to a generalization of the observability Gramian in linear dynamic systems to the nonlinear relative orbit dynamics represented by the two-body problems.
-An extended Kalman filter (EKF) is adapted to this problem and is evaluated with various gravitational harmonics and initial orbital determination (IOD) predictions.
-Extensive results illustrate correspondence between the proposed observability measure with filtering errors.
+An extended Kalman filter (EKF) is adapted to this problem and is evaluated with various gravitational harmonics, initial orbital determination (IOD) predictions, and time step sizes.
+Preliminary results illustrate correspondence between the proposed observability measure with filtering errors.
 An extensive numerical analysis in realistic scenarios includes satellite propagation of the two-body problem the $J_2$ perturbation effects.
@@
 Relative navigation between spacecraft in close-proximity essentially corresponds to space-based orbit determination.
-In particular, vision-based navigation and estimation of relative orbits have received attention recently, since optical sensors have the desirable properties of low cost and minimal maintenance, while providing accurate line-of-sight measurements, or angles from a chief to a deputy.
+In particular, vision-based navigation and estimation of relative orbit have received attention recently, since optical sensors have desirable properties of low cost and minimal maintenance, while providing accurate line-of-sight measurements, or angles from a chief to a deputy.
 
@@
 However, this result does not provide any information of the degree of observability.
-The main contribution of this paper is proposing a new  measure of observability for relative orbit determination with angles-only measurements.
+The main contribution of this paper is proposing a new  measure of observability for relative orbit determination with angles-only measurement.
 For a given initial condition of relative orbits, the proposed observability measure determines how much the corresponding relative orbits are easy or difficult to estimate, thereby providing a quantitative measure for relative orbit determination.
@@
 
-\section{Nonlinear Relative Orbital Dynamics}\label{sec:ND}
+% General background (very similar to the first two paragraphs of the last abstract)
+%Relative navigation is desirable for many spacecraft missions, such as formation control and rendezvous, since spacecraft maneuvers based only on on-board measurements can reduce the total operating cost significantly, and it improves safety against communication interruptions with ground stations.
+For these close proximity missions, relative navigation essentially reduces to space-based orbit determination.
+Vision-based navigation and estimation problems have received attention, since optical sensors have desirable properties of having low cost and minimal maintenance, while providing accurate line-of-sight measurements.
 
-Consider two satellites orbiting around the Earth, where each satellite is modeled as a rigid body.
-Suppose that a \textit{chief} satellite is on a circular orbit with a pre-determined orbital radius of $a\in\Re$.
+%Relative navigation based on angles-only measurements has been the subject of much recent investigation~\cite{WofGel09,Tom11,PatLovAllRusSin12}.
+The problem is to determine the relative orbit between a chief spacecraft and a deputy spacecraft by using the line-of-sight measurements between the two objects, assuming that the orbit of the chief is known.
 
-Define a local-vertical, local-horizontal (LVLH) frame as follows.
-Its origin is located at the chief satellite.
-The $x$-axis is along the radial direction from the Earth to the chief, and the $y$-axis is along the velocity vector of the chief.
-The $z$-axis is normal to the orbital plane, and it is parallel to the angular momentum vector of the chief.
-The LVLH frame is rotating with the angular velocity of $\mathbf{\omega}=[0,0,n]^T\in\Re^3$, where $n=\sqrt{\frac{\mu}{a^3}}$ is the mean motion of the chief satellite, and $\mu$ denotes the gravitational parameters of the Earth.
-Note that the velocity of the chief is given by $\mathbf{v}_{chief}=[0,na,0]^T\in\Re^3$.
-Let the relative position of a \textit{deputy} satellite with respect to the chief satellite be given by $\mathbf{r}=[x,y,z]^T\in\Re^3$ in the LVLH frame.
+%A common problem in relative navigation with angles-only measurements is range ambiguity.
+Without range information, the deputy states becomes less observable.
+We propose a generalized observability Gramian to nonlinear systems.
 
-\subsection{Nonlinear Equations of Motion}
+%Reference 1 shows that the relative orbit is unobservable from angles-only measurements when linear relative orbital dynamics are assumed, unless there are thrusting maneuvers.
+%Reference 2 investigates observability by using a relative orbit model linearized about spherical coordinates.
+%Reference 3 introduces the concept of partial observability to determine a basis vector representing a family of relative orbits, and an initial orbit determination technique for this method.
+%However, all of these results are based on linear relative orbital dynamics.
 
-Nonlinear equations of motion for the relative motion of the deputy with respect to the chief can be derived as follows based on Lagrangian mechanics~\cite{LovLeePISSFD14}.
+% Motivation: the performance of relative OD EKF is largely untested for various scenarios, including higher-fidelity gravitational models and time between measurements, for various initial predictions from IOD techniques
 
-\paragraph{Lagrangian}
-
-Considering that the LVLH frame is rotating with the angular velocity of $\mathbf{\omega}=[0,0,n]^T$, the inertial velocity of the deputy satellite is given by
-\begin{align*}
-\mathbf{v}_{deputy}
-=
-\mathbf{v}_{chief} + \dot{\r} + \mathbf{\omega}\times \mathbf{r} = 
-\begin{bmatrix}
-0 \\ na \\ 0
-\end{bmatrix}
-+
-\begin{bmatrix}
-\dot x \\ \dot y \\ \dot z
-\end{bmatrix}
-+
-\begin{bmatrix}
--n y \\ nx \\ 0
-\end{bmatrix}
-=
-\begin{bmatrix}
-\dot x - ny \\ \dot y + nx + na \\ \dot z
-\end{bmatrix}.
-\end{align*}
-Therefore, the (normalized) kinetic energy of the deputy satellite is 
-\begin{align*}
-T = \frac{1}{2}\|\mathbf{v}_{deputy}\|^2 = \frac{1}{2} \braces{(\dot x -ny)^2+(\dot y + nx +na)^2 + \dot z^2}.
-\end{align*}
-The location of the Earth from the chief is given by $[-a,0,0]^T$ in the LVLH frame. Therefore, the position vector of the deputy from the center of the Earth is given by 
-$\r_a=[x+a,y,z]^T\in\Re^3$. The gravitational potential energy is 
-\begin{align*}
-U = -\frac{\mu}{\sqrt{(x+a)^2 + y^2 + z^2}} = -\frac{\mu}{\|\r_a\|}.
-\end{align*}
-From the above equations, the Lagrangian of the deputy satellite is expressed in terms of $(x,y,z)$ as 
-\begin{align}
-L & = T-U = \frac{1}{2} \braces{(\dot x -ny)^2+(\dot y + nx +na)^2 + \dot z^2}
-+\frac{\mu}{\sqrt{(x+a)^2 + y^2 + z^2}}.
-\end{align}
-
-\paragraph{Euler-Lagrange Equations}
-
-Using the Euler-Lagrange equations, given by
-\begin{align*}
-\frac{d}{dt}\deriv{L}{\dot q}-\deriv{L}{q}=0,
-\end{align*}
-for $q\in\{x,y,z\}$, we obtain the nonlinear equations of motion for the relative orbit as follows.
-\begin{align}
-\ddot x - 2n\dot y -n^2 x&=n^2 a - \frac{\mu (x+a)}{((x+a)^2 + y^2 + z^2)^{3/2}},\label{eqn:ddotx}\\
-\ddot y + 2n\dot x -n^2 y &=   - \frac{\mu y}{((x+a)^2 + y^2 + z^2)^{3/2}},\label{eqn:ddoty}\\
-\ddot z &= - \frac{\mu z}{((x+a)^2 + y^2 + z^2)^{3/2}}.\label{eqn:ddotz}
-\end{align}
-
-These can be written as the standard form of the state equation,
-\begin{align}
-\dot{\mathbf{x}} = f (\mathbf{x}), \label{eqn:xxdot}
-\end{align}
-where the state vector is $\mathbf{x}=[x,y,z,\dot x, \dot y,\dot z]^T\in\Re^N$ with $N=6$, and
-\begin{align}
-f(\mathbf{x}) 
-%= \begin{bmatrix}
-%\dot x\\
-%\dot y\\
-%\dot z\\
-%2n\dot y +n^2 (x+a) - \dfrac{\mu (x+a)}{((x+a)^2 + y^2 + z^2)^{3/2}}\\
-%-2n\dot x +n^2 y  - \dfrac{\mu y}{((x+a)^2 + y^2 + z^2)^{3/2}}\\
-%- \dfrac{\mu z}{((x+a)^2 + y^2 + z^2)^{3/2}}
-%\end{bmatrix}
-=\begin{bmatrix}
-\dot \r \\
--2\omega\times \dot\r - \omega\times(\omega\times \r_a) -\dfrac{\mu \r_a}{\|\r_a\|^3}
-\end{bmatrix}.\label{eqn:f}
-\end{align}
-%where $\r=[x,y,z]^T\in\Re^3$, $\r_a = [x+a,y,z]^T\in\Re^3$, and $\dot\r =[\dot x,\dot y,\dot x]^T\in\Re^3$.
-
-\subsection{Line-of-sight Measurement}
-
-We assume that the line-of-sight from the chief to the deputy is measured by onboard sensing, such as optical sensors.
-The measurement is represented by the unit-vector of the relative position vector, i.e.,
-\begin{align}
-\y = h(\mathrm{x}) =\frac{\r}{\|\r\|},\label{eqn:y}
-\end{align}
-where $\y\in\Re^M$ with $M=3$, and it satisfies $\|\y\|=1$.
-
-\section{Nonlinear Observability Criteria}\label{sec:OC}
-
-Based on the nonlinear dynamic model presented at the previous section, here we analyze the observability of the relative orbit with line-of-sight measurements.
-This section serves as a summary of~\cite{LovLeePISSFD14}.
-
-\subsection{Observability Criteria for Nonlinear Systems}
-
-Observability of nonlinear systems has been studied in~\cite{HerKreITAC77}, and it is summarized as follows.
-For a given nonlinear dynamic system described with \refeqn{xxdot} and \refeqn{y}, a pair of points $\x_0$ and $\x_1$ are called \textit{indistinguishable} if the outputs of the corresponding solutions starting from each of $\x_0$ and $\x_1$ are identical for a certain time period.
-The systems is \textit{locally weakly observable} at $\x_0$, if there exists an open neighborhood $V$ of $\x_0$ such that for every open neighborhood $U$ of $\x_0$ contained in $V$, the only indistinguishable point to $\x_0$ is the point $\x_0$ itself.
-
-The Lie-derivative of the output $h(\x)$ along $f(\x)$ is defined as follows:
-\begin{align*}
-L_f h(\x) = \deriv{h(\x)}{\x}f(\x)\in\Re^{M\times 1},
-\end{align*}
-which corresponds to the directional derivative of $h(\x)$ along $f(\x)$.
-For a non-negative integer $i$, the $i$-th order Lie-derivative is defined by induction as $L_f^i h = L_f (L_f^{i-1} h)$ with $L_f^0 h = h$.
-Define an observability matrix $\mathcal{O}\in\Re^{NM\times N}$ as
-\begin{align*}
-\mathcal{O}(\x_0) = \deriv{}{\x} \begin{bmatrix} L_f^0 h(\x) \\ L_f^1 h(\x)\\ \vdots \\L_f^{N-1} h(\x)\end{bmatrix}\bigg|_{\x=\x_0}.
-\end{align*}
-It has been shown that the system is locally weakly observable at $\x_0$ if the rank of the observability matrix $\mathcal{O}(\x_0) = N$.
-When applied to linear dynamics, this yields the well-known observability rank condition for linear systems.
-Note that when there is more than a single measurement, i.e., $M>1$, the observability rank condition can be satisfied without a need for computing the higher-order Lie derivatives up to the $N-1$-th order.
-
-\subsection{Observability Criteria for Relative Orbital Dynamics}
-
-\paragraph{Observability Matrix}
-We apply the above observability criteria for the nonlinear relative orbital dynamics.
-Using the fact that $\x=[\r^T,\dot \r^T]$, the observability matrix of the relative orbital dynamics can be written as
-\begin{align}
-\mathcal{O} =
-\begin{bmatrix}
-\deriv{\y}{\r} & \deriv{\y}{\dot\r}\\
-\deriv{\dot\y}{\r} & \deriv{\dot\y}{\dot\r}\\
-\deriv{\ddot\y}{\r} & \deriv{\ddot\y}{\dot\r}
-\end{bmatrix}
-\triangleq
-\begin{bmatrix}
-\mathcal{O}_{00} & 0_{3\times 3}\\
-\mathcal{O}_{10} & \mathcal{O}_{11}\\
-\mathcal{O}_{20} & \mathcal{O}_{21}
-\end{bmatrix}.\label{eqn:OO}
-\end{align}
-Here we consider the observability matrix obtained by up to the second order Lie derivatives of the measurement due to complexity. But, this still provides sufficient conditions for observability.
-
-After straightforward but tedious algebraic manipulations using the following identity repeatedly,
-\begin{align*}
-\delta \parenth{\frac{1}{\|\mathbf{r}\|^i}} 
-& = -i\frac{ \mathbf{r}^T \delta \mathbf{r}}{\|\mathbf{r}\|^{i+2}},
-\end{align*}
-for any positive integer $i$, we can show that each of the sub-matrices $\mathcal{O}_{ij}$ of the observability matrix $\mathcal{O}$ is given by
-\begin{align}
-\mathcal{O}_{00} & = \deriv{\y}{\r} = \frac{1}{\|\r\|}(I - \y\y^T),\label{eqn:O_00}\\
-\mathcal{O}_{10} & = \deriv{\dot\y}{\r} = \deriv{}{\r} \parenth{\deriv{\y}{\r}\dot \r}= -\frac{1}{\|\r\|^2}\braces{
-{\dot\r\y^T +\y^T\dot\r I +\y\dot\r^T}
-- 3\y\y^T\dot\r \y^T},\\
-\mathcal{O}_{11} & = \deriv{\dot\y}{\dot\r} = \deriv{}{\dot\r} \parenth{\deriv{\y}{\r}\dot \r}= \deriv{\y}{\r} = \mathcal{O}_{00},\label{eqn:O11}\\
-\mathcal{O}_{20} & = \deriv{\ddot\y}{\r} = -\dfrac{2\dot\r \dot\r^T  +\dot\r^T\dot\r I}{\|\r\|^3}
-+3\dfrac{2(\r^T\dot\r) \dot\r\r^T +(\dot\r^T\dot\r)\r\r^T
-+(\r^T\dot\r)^2 I
-+2(\r^T\dot\r)\r  \dot\r^T
-}{\|\r\|^5}
-- 15 \dfrac{(\r^T\dot\r)^2\r  \r^T}{\|\r\|^7}\nonumber \\
-&\quad -\dfrac{\ddot{\r}\r^T +\r^T\ddot{\r} I +\r \ddot{\r}^T}{\|\r\|^3}
-+ 3 \dfrac{(\r^T\ddot{\r})\r \r^T}{\|\r\|^5}
-+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
-\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}},\\
-\mathcal{O}_{21} & = \deriv{\ddot\y}{\dot\r} = \deriv{}{\dot\r}\parenth{\deriv{}{\r}\parenth{\deriv{\y}{\r}\dot\r}\dot\r + \deriv{\y}{\r}\ddot\r} = 2 \deriv{}{\r} \parenth{\deriv{\y}{\r}\dot\r} + \deriv{\y}{\r} \deriv{\ddot\r}{\dot\r}
-\nonumber\\
-& \quad = 2\mathcal{O}_{10} 
-- 2 \mathcal{O}_{00}[\omega]_\times,\label{eqn:O_21}
-\end{align}
-where $[\omega]_\times\in\Re^{3\times 3}$ is defined as
-\begin{align*}
-[\omega]_\times = \begin{bmatrix} 0 & -n & 0 \\ n & 0 & 0 \\ 0 & 0 & 0 \end{bmatrix}.
-\end{align*}
-These expressions were fully verified by the Matlab symbolic computation toolbox.
-
-After some algebraic manipulations, we can show that the sub-matrices satisfy the following identities. 
-\begin{align}
-\mathcal{O}_{10}\r & = 
-%-\frac{1}{\|\r\|}\braces{{\dot\r\y^T +\y^T\dot\r I +\y\dot\r^T}- 3\y\y^T\dot\r \y^T}\y\\
-%&=-\frac{1}{\|\r\|}\braces{\dot\r +(\y^T\dot\r)\y +\y(\dot\r^T\y)- 3\y(\y^T\dot\r) }\\
-%&=-\frac{1}{\|\r\|}\braces{\dot\r -\y(\y^T\dot\r) }=
--\mathcal{O}_{00}\dot\r,\label{eqn:O10r}\\
-\mathcal{O}_{10}\dot\r & = -\frac{1}{\|\r\|^2}\braces{
-{2\dot\r(\y^T\dot \r)  +\y(\dot\r^T\dot \r)}
-- 3\y(\y^T\dot\r)^2},\label{eqn:O10dotr}\\
-\mathcal{O}_{20}\r %& = 
-%-\dfrac{2\dot\r \dot\r^T\y  +\dot\r^T\dot\r \y}{\|\r\|^2}
-%+3\dfrac{2(\y^T\dot\r) \dot\r\y^T\y +(\dot\r^T\dot\r)\y\y^T\y+(\y^T\dot\r)^2 y +2(\y^T\dot\r)\y \dot\r^T\y}{\|\r\|^2}
-%- 15 \dfrac{(\y^T\dot\r)^2\y  \y^T\y}{\|\r\|^2}\nonumber \\
-%&\quad -\dfrac{\ddot{\r}\y^Ty +\y^T\ddot{\r} \y +\y \ddot{\r}^T\y}{\|\r\|^1}
-%+ 3 \dfrac{(\y^T\ddot{\r})\y \y^T\y}{\|\r\|}
-%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
-%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
-%&=
-%-\dfrac{2\dot\r (\dot\r^T\y)  +(\dot\r^T\dot\r) \y}{\|\r\|^2}
-%+3\dfrac{2(\y^T\dot\r) \dot\r +(\dot\r^T\dot\r)\y+(\y^T\dot\r)^2 \y +2(\y^T\dot\r)^2\y }{\|\r\|^2}
-%- 15 \dfrac{(\y^T\dot\r)^2\y}{\|\r\|^2}\nonumber \\
-%&\quad -\dfrac{\ddot{\r} +(\y^T\ddot{\r}) \y +\y (\ddot{\r}^T\y)}{\|\r\|^1}
-%+ 3 \dfrac{(\y^T\ddot{\r})\y }{\|\r\|}
-%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
-%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
-%&=
-%-\dfrac{2\dot\r (\dot\r^T\y)  +(\dot\r^T\dot\r) \y}{\|\r\|^2}
-%+3\dfrac{2(\y^T\dot\r) \dot\r +(\dot\r^T\dot\r)\y+(\y^T\dot\r)^2 \y +2(\y^T\dot\r)^2\y }{\|\r\|^2}
-%- 15 \dfrac{(\y^T\dot\r)^2\y}{\|\r\|^2}\nonumber \\
-%&\quad -\dfrac{\ddot{\r} +(\y^T\ddot{\r}) \y +\y (\ddot{\r}^T\y)}{\|\r\|^1}
-%+ 3 \dfrac{(\y^T\ddot{\r})\y }{\|\r\|}
-%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
-%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
-%&=
-%\dfrac{4\dot\r (\dot\r^T\y)  -2(\dot\r^T\dot\r)\y-6(\y^T\dot\r)^2 \y }{\|\r\|^2}
-% -\mathcal{O}_{00}\ddot \r
-%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
-%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
-& = -2\mathcal{O}_{10}\dot\r +\mathcal{O}_{00}\braces{-\ddot \r +\deriv{\ddot\r}{\r}\r},\label{eqn:O20r}\\
-\mathcal{O}_{21}\r & = -2\mathcal{O}_{00} (\dot\r+\omega\times\r),\label{eqn:O21r}\\
-\mathcal{O}_{21}\dot\r & = 2\mathcal{O}_{10} \dot\r
-- 2 \mathcal{O}_{00}[\omega]_\times\dot\r,\label{eqn:O21rdot}
-\end{align}
-which are useful to derive the observability criteria.
-
-\paragraph{Observability Rank Condition}
-
-Now we present sufficient conditions that the observability matrix $\mathcal{O}$ has full rank.
-
-\begin{prop}
-Define three vectors $\mathbf{v}_{rel}$, $\mathbf{a}_1,\mathbf{a}_2\in\Re^3$ as 
-\begin{align}
-\mathbf{v}_{rel} & = \dot\r +\omega\times\r,\label{eqn:vrel}\\
-\mathbf{a}_1 & = \ddot\r-\deriv{\ddot\r}{\r}\r = -2\omega\times \dot\r - [\omega]_\times^2 a \mathbf{e}_1 -\dfrac{\mu a }{\|\r_a\|^3}\mathbf{e}_1- \dfrac{3\mu\r_a^T\r}{\|\r_a\|^5}\r_a,\label{eqn:a1}\\
-\mathbf{a}_2 & = \ddot\r-\deriv{\ddot\r}{\r}\r -\deriv{\ddot\r}{\dot\r}\dot\r = \mathbf{a}_1+2\omega\times\dot\r= - [\omega]_\times^2 a \mathbf{e}_1 -\dfrac{\mu a }{\|\r_a\|^3}\mathbf{e}_1- \dfrac{3\mu\r_a^T\r}{\|\r_a\|^5}\r_a.\label{eqn:a2}
-\end{align}
-The nonlinear relative orbital dynamics are locally weakly observable at $\x=[\r^T,\dot\r^T]$ if
-%\begin{list}{}{\setlength{\leftmargin}{1cm}\setlength{\itemsep}{0mm}\setlength{\parsep}{0mm}\setlength{\topsep}{0mm}\setlength{\parskip}{0mm}\setlength{\labelwidth}{2cm}}
-%\item[(i)] when $\r\times\dot\r =0$, $\r\times\mathbf{a}_1\neq 0$ and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_1)\neq 0$,
-%\item[(ii)] when $\r\times\dot\r \neq 0$, $\r\times\mathbf{a}_2\neq 0$ and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_2)\neq 0$.
-%\end{list}
-\begin{alignat}{2}
-(i)&\; &\text{when $\r\times\dot\r =0$, $\r\times\mathbf{v}_{ref}\neq0$, $\r\times\mathbf{a}_1\neq 0$,  and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_1)\neq 0$},\label{eqn:cond1}\\
-(ii)& &\text{when $\r\times\dot\r \neq 0$, $\r\times\mathbf{v}_{ref}\neq0$, $\r\times\mathbf{a}_2\neq 0$ and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_2)\neq 0$.}\label{eqn:cond2}
-\end{alignat}
-\end{prop}
-
-\begin{proof}
-We show that if the above conditions are satisfied, then the six columns of the observability matrix are linearly independent.
-Suppose that for a constant vector $\mathbf{c}=[\mathbf{c}_1^T,\mathbf{c}_2^T]\in\Re^6$, where $\mathbf{c}_1,\mathbf{c}_2\in\Re^3$, we have $\mathcal{O}\mathbf{c}=0$, i.e.,
-\begin{align}
-\mathcal{O} \mathbf{c} = \begin{bmatrix}
-\mathcal{O}_{00} & 0_{3\times 3}\\
-\mathcal{O}_{10} & \mathcal{O}_{11}\\
-\mathcal{O}_{20} & \mathcal{O}_{21}
-\end{bmatrix}
-\begin{bmatrix} \mathbf{c}_1 \\ \mathbf{c_2} \end{bmatrix}
-=
-\begin{bmatrix}
-\mathcal{O}_{00}\mathbf{c}_1\\
-\mathcal{O}_{10}\mathbf{c}_1+ \mathcal{O}_{11}\mathbf{c}_2\\
-\mathcal{O}_{20}\mathbf{c}_1+ \mathcal{O}_{21}\mathbf{c}_2
-\end{bmatrix}
-=
-\begin{bmatrix}
-0_{3\times 1}\\
-0_{3\times 1}\\
-0_{3\times 1}
-\end{bmatrix}.\label{eqn:Oc}
-\end{align}
-
-We wish to show that there is no non-zero vector $\mathbf{c}$ satisfying \refeqn{Oc}. At the first three rows of \refeqn{Oc}, we have
-\begin{align*}
-\mathcal{O}_{00} \mathbf{c}_1 = \frac{1}{\|\r\|}(I-\y\y^T) \mathbf{c}_1 = 0.
-\end{align*}
-The matrix $\mathcal{O}_{00}$ has a one-dimensional null space spanned by $\y$.
-Therefore, without loss of generality, we can choose $\mathbf{c}_1=0$ or $\mathbf{c}_1=\r$.
-Suppose $\mathbf{c}_1=\r$ for the subsequent development.
-
-For the chosen value of $\mathbf{c}_1=\r$, we find the next three rows of \refeqn{Oc} as
-\begin{align}
-\mathcal{O}_{10} \mathbf{c}_1 +\mathcal{O}_{11} \mathbf{c}_2 & = \mathcal{O}_{10} \r +\mathcal{O}_{11} \mathbf{c}_2  = \mathcal{O}_{00} (-\dot\r + \mathbf{c}_2)\nonumber\\
-& = \frac{1}{\|\r\|}(I-\y\y^T)(-\dot\r + \mathbf{c}_2)=0,\label{eqn:row2}
-\end{align}
-where we have used \refeqn{O11} and \refeqn{O10r}. Next, we consider two cases of \refeqn{row2}, namely (i) when $\r\times\dot\r=0$, and (ii) when $\r\times\dot\r\neq 0$.
-
-\paragraph{Case (i): $\r\times\dot \r=0$} In this case, $\dot \r$ can be written as $\dot \r = \alpha\r$ for some constant $\alpha$ as $\r$ is parallel to $\dot\r$. Then, \refeqn{row2} reduces to 
-\begin{align*}
-\frac{1}{\|\r\|}(I-\y\y^T)\mathbf{c}_2=0,
-\end{align*}
-which implies that $\mathbf{c}_2= c\r$ for an arbitrary constant $c$. For the given choice of $\mathbf{c}=[\r^T,c\r^T]^T$, the last three rows of \refeqn{Oc} are given by
-\begin{align*}
-\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
-& = \mathcal{O}_{20} \r + c\mathcal{O}_{21} \r.
-\end{align*}
-Using \refeqn{O20r} and \refeqn{O21r}, this can be rewritten as
-\begin{align*}
-\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
-& =
--2\mathcal{O}_{10}\dot\r +\mathcal{O}_{00}\braces{-\ddot \r +\deriv{\ddot\r}{\r}\r}
--2c\mathcal{O}_{00} (\dot\r+\omega\times\r).
-\end{align*}
-But, from \refeqn{O10dotr}, we can show that $2\mathcal{O}_{10}\dot\r=0$ when $\r$ is parallel to $\dot \r$. Using \refeqn{vrel} and \refeqn{a1}, this further reduces to
-\begin{align}
-\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
-& =
--\mathcal{O}_{00}(\mathbf{a}_1 +2c \mathbf{v}_{rel})\nonumber\\
-& = - \frac{1}{\|\r\|}(I-\y\y^T)(\mathbf{a}_1 +2c \mathbf{v}_{rel})=0.\label{eqn:row3i}
-\end{align}
-The matrix $(I-\y\y^T)$ represents the orthogonal projection of a vector into the plane normal to $\y$.
-From the second condition of \refeqn{cond1}, namely $\r\times\mathbf{a}_1\neq 0$, we have $(I-\y\y^T)\mathbf{a}_1\neq 0$, which implies that the constant $c$ cannot be simply chosen as $c=0$.
-Therefore, the only possible case to satisfy the above equation is when $\mathbf{a}_1 +2c \mathbf{v}_{rel}$ is parallel to $\r$ for some values of $c$.
-However, that is not feasible since the third condition of \refeqn{cond1}, namely $\r^T(\mathbf{v}_{ref}\times\mathbf{a}_1)\neq0$,  implies that the three vectors $\r$, $\mathbf{a}_1$, and $\mathbf{v}_{rel}$ do not belong to a common plane, i.e., there is no constant $c$ such that $\mathbf{a}_1 +2c \mathbf{v}_{rel}$ is parallel to $\r$.
-
-Therefore, there is no $\mathbf{c}\in\Re^6$ satisfying \refeqn{Oc} if $\mathbf{c}_1=\r$, under the given condition \refeqn{cond1}.
-This implies that $\mathbf{c}_1=0$.
-Substituting this back to \refeqn{Oc}, we have $\mathcal{O}_{11}\mathbf{c}_2=\mathcal{O}_{00}\mathbf{c}_2=0$, which follows that $\mathbf{c}_2=c\r$ for some constant $c$.
-However, when $\mathbf{c}_2=c\r$, the last three rows of \refeqn{Oc} are given by
-\begin{align}
-\mathcal{O}_{21}\mathbf{c}_2 & = 2c(\mathcal{O}_{10}-\mathcal{O}_{00}[\omega]_\times) \r \nonumber
-\\
-& = -2c\mathcal{O}_{00}\mathbf{v}_{rel}=0,\label{eqn:row3i1}
-\end{align}
-where \refeqn{O10r} is used. But, from the first condition of \refeqn{cond1}, we have $\mathcal{O}_{00}\mathbf{v}_{rel}\neq 0$, and therefore $c=0$, i.e., $\mathbf{c}_2=c\r =0$.
-
-In short, under the given condition \refeqn{cond1}, the equation \refeqn{Oc} implies that $\mathbf{c}=0$. Therefore, the six columns of the observability matrix $\mathcal{O}$ are linearly independent.
-
-\paragraph{Case (ii): $\r\times\dot \r\neq0$} Next, we consider the second case of \refeqn{row2}.
-It implies that $-\dot\r + \mathbf{c}_2$ is parallel to $\r$, or equivalently, $\mathbf{c}_2 = \dot\r + c\r$ for an arbitrary constant $c$.
-For the given choice of $\mathbf{c}=[\r^T,\dot\r+ c\r^T]^T$, the last three rows of \refeqn{Oc} are given by
-\begin{align*}
-\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
-& = (\mathcal{O}_{20} \r+\mathcal{O}_{21}\dot\r) + c\mathcal{O}_{21} \r.
-\end{align*}
-From \refeqn{O20r}, \refeqn{O21r}, \refeqn{O21rdot}, this can be rewritten as
-\begin{align*}
-\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
-& = -\mathcal{O}_{00}(\mathbf{a}_2+2c\mathrm{v}_{ref})\\
-& = -\frac{1}{\|\r\|}(I-\y\y^T)(\mathbf{a}_2+2c\mathbf{v}_{ref})=0.
-\end{align*}
-By following the same argument given after \refeqn{row3i}, under \refeqn{cond2}, there is no $c$ satisfying the above equation.
-
-This implies that $\mathbf{c}_1=0$.
-Then, by the same argument given at \refeqn{row3i1}, we have $\mathbf{c}_2=0$.
-In short, under the given condition \refeqn{cond2}, the equation \refeqn{Oc} implies that $\mathbf{c}=0$.
-Therefore, the six columns of the observability matrix $\mathcal{O}$ are linearly independent.
-\end{proof}
-
-\paragraph{Remarks} We consider the cases where the given sufficient conditions \refeqn{cond1} and \refeqn{cond2} are violated. 
-For both cases, we have $\r\times\mathbf{v}_{rel}\neq 0$.
-At \refeqn{vrel}, $\mathbf{v}_{rel}$ corresponds to the relative velocity observed in the inertial frame.
-Therefore, the given sufficient conditions for observability are violated when the relative velocity vector is parallel to the relative position vector in the inertial frame.
-
-The second condition of each of \refeqn{cond1} and \refeqn{cond2} is satisfied in general, as the expressions for $\mathbf{a}_1$ and $\mathbf{a}_2$ are relatively arbitrary at \refeqn{a1} and \refeqn{a2}.
-
-The third condition of \refeqn{cond1} implies that three vectors $\r$, $\mathbf{v}_{rel}$ and $\mathbf{a}_1$ do not belong to the same plane.
-The third condition of \refeqn{cond2} has a similar structure as well.
-This condition can be easily violated if the relative motion is planar, i.e., when $z(t)\equiv 0$ for all $t$.
-
-However, \refeqn{cond1} and \refeqn{cond2} are sufficient conditions for observability, and the fact that any of these condition is not satisfied does not necessarily mean that the considered point is not observable.
-In such case, the third or higher order Lie derivatives should be checked to determine observability.
-The main contribution of this section of the paper is showing that under certain geometric conditions, the nonlinear relative orbital dynamics are indeed observable via angles-only measurements.
+% Assumptions:
+% 	The chief orbit initial conditions (initially circular) assumed to be well-known.
+%	Measurements: a unit vector representing the direction between the chief and deputy is subject to Gaussian white noise
+%		the measurements are always available (nothing blocking line-of-sight)
+%		the measurements always originate from the deputy satellite
+%	Initial conditions: an initial state estimate and uncertainty are assumed well-known
 
@@
 
-The measure of observability has been generalized to nonlinear dynamic systems in order to obtain balanced realizations.
+The measure of observability has been generalized to nonlinear dynamic systems, in order to obtain balanced realizations.
 An energy-like observability function is introduced~\cite{SchSCL93,Sch94,NewKriPICDC98} to measure the degree of contribution of an initial state to the output, and it is applied to a pendulum system.
@@
 
-\section{Kalman Filter}
+\section{Nonlinear Observability Measure for Relative Orbits}
 
@@
 
-We consider the relative orbital dynamics between two satellites, where the observability of the deputy relative to the chief satellite is on a prescribed circular orbit when only $\mathbf{y}$ is assumed to be measured.
-Then, an extended Kalman filter is developed for numerical analyses, where the satellites are either subject to two-body forces or with gravitational $J_2$ perturbations included.
+We consider the relative orbital dynamics between two satellites, referred to as a chief and a deputy, where the chief satellite is on a prescribed circular orbit.
+The dynamics of the deputy are defined by the two-body problem.
+The relative position of the deputy from the chief is denoted by $\mathbf{r}\in\Re^3$, and the line-of-sight from the chief to the deputy, namely $\mathbf{y}=\frac{\mathbf{r}}{\|\mathbf{r}\|}\in\Re^3$ is assumed to be measured.
+An extended Kalman filter is developed for numerical analysis.
 
-The initial conditions for an estimate of the state vector $\hat \x_0^+$ and its initial covariance matrix $P^+_0$ are assumed to be known from an initial orbit determination (IOD) technique.
-
-\paragraph{Flow Update}
-The flow update provides an a priori state estimate $\hat \x^-_k$ with covariance matrix $P^-_k$ based on the equations of motion.
-The linearized state transition matrix from \refeqn{xxdot} and \refeqn{f} from time step $k-1$ to time step $k$ is written as
-\begin{align}
-F_{k-1}=\exp\left(
-\begin{bmatrix}
-0_{3\times3} & I_{3\times3} \\
-\left(-[\omega]_\times^2-\frac{\mu I_{3\times3}}{\norm{\r_a}^3}+\frac{3\mu \r_a \r_a^T}{\norm{\r_a}^5}\right) & -2[\omega]_\times
-%0 & 0 & 0 & 1 & 0 & 0\\
-%0 & 0 & 0 & 0 & 1 & 0\\
-%0 & 0 & 0 & 0 & 0 & 1\\
-%3*n^2 & 0 & 0 & 0 & 2*n & 0\\
-%0 & 0 & 0 & -2*n & 0 & 0\\
-%0 & 0 & -n^2 & 0 & 0 & 0
-\end{bmatrix}
-(t_k-t_{k-1})\right),
-\end{align}
-such that
-\begin{align}
-\hat \x_{k}^-=F_{k-1}\hat \x_{k-1}^+.
-\end{align}
-The process covariance matrix $Q_k$ is assumed known.
-The a priori covariance matrix is defined as
-\begin{align}
-P^-_k=F_{k-1}P^+_{k-1}F_{k-1}^T+Q_k.
-\end{align}
-
-\paragraph{Measurement Update}
-The effect of the measurement $\y_k$ with covariance matrix $R_k$ is included to determine the posterior state estimate $\hat \x^+_k$ and $P^+_k$.
-The linearized measurement matrix $H_k$ is derived from \refeqn{y},
-\begin{align}
-H_k=
-\begin{bmatrix}
-\left(\frac1{\norm{\hat \r_{k}^-}}I_{3\times3}-\frac{\hat \r_{k}^-{\hat \r_{k}^{-T}}}{\norm{\hat \r_{k}^-}^3}\right) & 0_{3\times3}
-\end{bmatrix},
-\end{align}
-where $\hat \r_{k}^-$ is the a priori estimate of $\r$ given by the first three elements of $\hat \x_{k}^-$.
-Based on \refeqn{y}, the measurement prediction is given by
-\begin{align}
-\hat{\y}_k=h(\hat \r_{k}^-)=\frac{\hat \r_{k}^-}{\norm{\hat \r_{k}^-}}.
-\end{align}
-The posterior state estimate and its covariance matrix are given by
-\begin{align}
-\x^+_k&=\x^-_k+K_k(\y_k-\hat{\y}_k),
-\\
-P^+_k&=(I_{6\times6}-K_kH_k)P^-_k,
-\end{align}
-such that $K_k$ is chosen to minimize $\tr{P^+_k}$ with
-\begin{align}
-K_k=P^-_kH_k^T(H_kP^-_kH_k+R_k)^{-1}.
-\end{align}
-
-%The initial conditions of the deputy satellite are chosen as
-%\begin{align*}
-%\mathbf{r}_0=[r_{x_0},\, 0,\, 0]^T\,\mathrm{km},\quad \mathbf{v}_0=[0,\, -2r_{x_0}n,\, 0]^T\,\mathrm{km/s},
-%\end{align*}
-%for a varying $r_{x_0}$.
+The initial conditions of the deputy satellite are chosen as
+\begin{align*}
+\mathbf{r}_0=[r_{x_0},\, 0,\, 0]^T\,\mathrm{km},\quad \mathbf{v}_0=[0,\, -2r_{x_0}n,\, 0]^T\,\mathrm{km/s},
+\end{align*}
+for a varying $r_{x_0}$.
 These yield a series of elliptic orbits around the chief when applied to the linearized relative orbital dynamics.
 The state estimation error is represented by a time averaged normalized magnitude error, $e_{mag} = \frac{\|\mathbf{x}-\hat{\mathbf{x}}\|}{\|\mathbf{x}\|}$, where $\mathbf{x}=[\mathbf{r}^T,\dot{\mathbf{r}}^T]^T\in\Re^6$ denotes the true state and $\hat{\mathbf{x}}$ denotes the estimated state.
+
+The normalized magnitude error $e_{mag}$ and the condition number of the proposed observability measure $\mathrm{cond}\braces{\mathcal{W}}$ are computed among various values for $r_{x_0}$, tabulated below.
+
+\hspace*{0.1cm}\centerline{
+\begin{tabular}{ccc}\hline
+$r_{x_0}$ & $e_{mag}$ & $\mathrm{cond}\braces{\mathcal{W}}$\\\hline
+\vspace*{-0.02\columnwidth}\\
+$1$ & $11.5962$ &  $10^{16.6653}$\\
+$2$ & $9.6374$ & $10^{16.8323}$\\
+$5$ & $4.0041$ & $10^{16.3535}$\\
+$10$ & $2.6318$ &  $10^{15.8194}$\\
+$20$ & $0.9250$ &  $10^{15.2832}$\\
+$50$ & $0.5315$ &  $10^{14.4052}$\\
+$100$ & $0.1675$ &  $10^{12.9884}$\\
+\hline
+\end{tabular}
+}
+
+These preliminary results illustrate that there exists a certain correspondence between the filter error and the proposed observability measure, and therefore, it can be used as a quantitative measure for strength of observability.
+
+\section{Expected Results and Significance}
+
+%It is expected that the given results are generalized to realistic scenarios that include orbital perturbations.
+We will choose numerous cases of relative motion between chief and deputy satellite orbits, and with each case, we plan to evaluate the observability measure and process angles-only measurements and evaluate the performance of an EKF based on the perturbed two-body equations of motion.
 %
-%The normalized magnitude error $e_{mag}$ and the condition number of the proposed observability measure $\mathrm{cond}\braces{\mathcal{W}}$ are computed among various values for $r_{x_0}$, tabulated below.
-%
-%\hspace*{0.1cm}\centerline{
-%\begin{tabular}{ccc}\hline
-%$r_{x_0}$ & $e_{mag}$ & $\mathrm{cond}\braces{\mathcal{W}}$\\\hline
-%\vspace*{-0.02\columnwidth}\\
-%$1$ & $11.5962$ &  $10^{16.6653}$\\
-%$2$ & $9.6374$ & $10^{16.8323}$\\
-%$5$ & $4.0041$ & $10^{16.3535}$\\
-%$10$ & $2.6318$ &  $10^{15.8194}$\\
-%$20$ & $0.9250$ &  $10^{15.2832}$\\
-%$50$ & $0.5315$ &  $10^{14.4052}$\\
-%$100$ & $0.1675$ &  $10^{12.9884}$\\
-%\hline
-%\end{tabular}
-%}
-%
-%These preliminary results illustrate that there exists a certain correspondence between the filter error and the proposed observability measure, and therefore, it can be used as a quantitative measure for strength of observability.
+%In particular, the $J_2$ gravitational harmonics will be included in the satellite propagation, so the EKF must handle simulated process uncertainty.
+Then, we will consider various time steps between measurements.
+Finally, we will consider various initial orbit determination (IOD) predictions to show when and how the filter converges.
+The following figure illustrates preliminary results for angles-only relative orbit estimation with $J_2$ perturbation.
 
-\section{Numerical Results}
-
-We choose numerous cases of relative motion between chief and deputy satellite orbits.
-With each case, we evaluate the observability measure and process angles-only measurements with the EKF, using both the two-body equations of motion and higher fidelity propagation with $J_2$ effects as the truth models.
-%In particular, the $J_2$ gravitational harmonics will be included in the satellite propagation, so the EKF must handle simulated process uncertainty.
-To begin each simulation, an initial orbit determination (IOD) estimate is determined by various angles-only observations.
-These IOD estimates serve as initial estimates for the EKF.
-Then, we evaluate the performance of the EKF for these scenarios, specifically focusing on the observability measure and issues arising from initial conditions.
-
-Two IOD methods are developed in~\cite{NewLovPra14,PraLovNew14}, referred to as \emph{IOD1} and \emph{IOD2}, and employed to provide initial conditions for $13$ cases, outlined in Table \ref{tab:IODerr}.
-%Two IOD methods are referred to as \emph{IOD1} and \emph{IOD2}, and employed to provide initial conditions for $13$ cases, outlined in Table \ref{tab:IODerr}.
-These cases yield relative orbit trajectories illustrated in Figures \ref{fig:0CasesTraj}, \ref{fig:1CasesTraj}, \ref{fig:2CasesTraj}, \ref{fig:3CasesTraj}, and \ref{fig:4CasesTraj}.
-The most important distinction between the various cases is that in Case $\mathbf{1}$ the two-body solution and the linearized dynamics of the Hill-Clohessy-Wiltshire (HCW) solution are very close and the trajectory remains largely planar with a circular shape.
-
-\begin{figure}[h]
-\centerline{
-	\subfigure[Case 0A]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_0A_traj.pdf}}
-	\hfill
-	\subfigure[Case 0B]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_0B_traj.pdf}}
-	\hfill
-	\subfigure[Case 0C]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_0C_traj.pdf}}
-}
-\caption{The trajectories of the Cases 0A, 0B, and 0C (solid red: HCW, dashed blue: two-body solution) show that the linear and nonlinear solutions are close.
-The resulting trajectories are not cyclic as the trajectories change in the $y$-direction with time.}
-\label{fig:0CasesTraj}
-\end{figure}
-
-\begin{figure}[h]
-\label{fig:CaseTraj}
-\centerline{
-	\subfigure[Case 1]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_1_traj.pdf}}
-}
-\caption{The trajectory of the Case 1 (solid red: HCW, dashed blue: two-body solution) shows that the linear and nonlinear solutions are very close, where both solutions are largely planar and cyclic.}
-\label{fig:1CasesTraj}
-\end{figure}
-
-\begin{figure}[h]
-\centerline{
-	\subfigure[Case 2A]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_2A_traj.pdf}}
-	\hfill
-	\subfigure[Case 2B]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_2B_traj.pdf}}
-	\hfill
-	\subfigure[Case 2C]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_2C_traj.pdf}}
-}
-\caption{The trajectories of the Cases 2A, 2B, and 2C (solid red: HCW, dashed blue: two-body solution) show differences between the linear and nonlinear solutions that vary about the $y$-axis, increasing the observabilities.}
-\label{fig:2CasesTraj}
-\end{figure}
-
-\begin{figure}[h]
-\centerline{
-	\subfigure[Case 3A]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_3A_traj.pdf}}
-	\hfill
-	\subfigure[Case 3B]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_3B_traj.pdf}}
-	\hfill
-	\subfigure[Case 3C]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_3C_traj.pdf}}
-}
-\caption{The trajectories of the Cases 3A, 3B, and 3C (solid red: HCW, dashed blue: two-body solution) exhibit large observabilities due to nonlinear solutions that move far away from the linearized solutions.}
-\label{fig:3CasesTraj}
-\end{figure}
-
-\begin{figure}[h]
-\centerline{
-	\subfigure[Case 4A]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_4A_traj.pdf}}
-	\hfill
-	\subfigure[Case 4B]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_4B_traj.pdf}}
-	\hfill
-	\subfigure[Case 4C]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_4C_traj.pdf}}
-}
-\caption{The trajectories of the Cases 4A, 4B, and 4C (solid red: HCW, dashed blue: two-body solution) experience the largest observabilities due to nonlinear solutions that move farther away from the linearized solutions than any of the trajectories from other cases considered in this paper.}
-\label{fig:4CasesTraj}
-\end{figure}
-
-The initial error functions between the true orbit and and IOD predictions are defined as
-\begin{align}
-e_{mag,0}=\abs\left(1-{\frac{\norm{\r^+_0}}{\norm{\r_0}}}\right), \quad e_{dir,0}=\cos^{-1}\left(\frac{\r_0}{\norm{\r_0}}\cdot \frac{\r^+_0}{\norm{\r^+_0}}\right).
-\end{align}
+For the final paper, we will choose numerous cases of relative motion between chief and deputy satellite orbits, and with each case, we plan to both evaluate the observability measure and process angles-only measurements with an EKF, using the perturbed two-body equations of motion as a truth model.
+In particular, the $J_2$ gravitational harmonics will be included in the satellite propagation, so the EKF must handle simulated process uncertainty.
+The figure below illustrates preliminary results for angles-only relative orbit estimation with $J_2$ perturbation in the measurements.
+Note that the EKF performs well in this particular case, as evidenced by the close agreement between the true and estimated trajectory (Figure 1a) and the decrease in the maximum eigenvalue of the covariance matrix over time (Figure 1b).
+For the various cases to be chosen, we will evaluate the performance of the EKF (using metrics such as those illustrated in Figure 1) and seek a consistent correlation between the filter performance and the new observability measure proposed above.
 
@@
 \begin{threeparttable}[h]
-\caption{Case Observabilities and Initial Errors}
-\begin{tabularx}{0.72\textwidth}
-{
->{$}c<{$}
-*{1}{>{$}c<{$}} |
-*{2}{>{$}c<{$}} |
-*{2}{>{$}c<{$}}
-}
-\toprule
-\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{\multirow{1}{*}{IOD1}} & \multicolumn{2}{c}{\multirow{1}{*}{IOD2}} \\
-& &  e_{mag,0} & e_{dir,0} & e_{mag,0} & e_{dir,0} \\\midrule
-0A & 10^{11.1486} & 0.0003 & 1.4901\times10^{-8} & 0.0012  & 2.1073\times10^{-8} \\
-0B & 10^{10.9387} & 0.0233 & 0 & 0.0088 & 0 \\
-0C & 10^{10.9851} & 0.0152 & 2.1073\times10^{-8} & 0.0139 & 1.4901\times10^{-8}  \\
-\midrule
-\mathbf{1} & 10^{16.6871} & 0.0001 & 0 & 0 & 0  \\
-\midrule
-2A & 10^{10.6492} & 0.0550 & 1.8150\times10^{-4} & 0.0238 & 1.8150\times10^{-4}  \\
-2B & 10^{10.4163} & 0.9571 & 0.0734                     & 0.9571 & 0.0734  \\
-2C & 10^{10.2155} & 0.0219 & 0.0044 & 0.0065 & 0.0044 \\
-\midrule
-3A & 10^{8.3490} & 0.0191 & 0.0031 & 0.1830 & 0.0031  \\
-3B & 10^{8.5372} & 0.0612 & 0.0533 & 0.2315 & 0.0533  \\
-3C & 10^{8.6586} & 6.8962 & 0.7348 & 5.4921 & 0.7348  \\
-\midrule
-4A & 10^{7.9062} & 8.7347 & 0.4587 & 0.1575 & 0.4587  \\
-4B & 10^{7.9903} & 0.9276 & 0.1819 & 0.0621 & 0.1819  \\
-4C & 10^{8.0645} & 0.1161 & 0.0554 & 0.0354 & 0.0554  \\
+\caption{Error variables for EKF with varying $Q_k$ Considering Only Two-Body Problem Forces}
+\begin{tabularx}{0.7\textwidth}{>{\centering $}X<{$}*{2}|*{5}{>{$}c<{$}}}\toprule
+\multirow{2}{*}{Case} & \multirow{2}{*}{IOD} & \multicolumn{4}{c}{$\bar e_{mag}$} \\
+& & q=-1 & q=0 & q=1 & q=2 \\\midrule
+0A & IOD 1  & 0.1129  & 0.1405  & 0.2859  & 0.4726 \\
+0A & IOD 2  & 0.0400  & 0.1470  & 0.2559  & 0.5269 \\
+0B & IOD 1  & 0.0153  & 0.0553  & 0.2734  & 0.7942 \\
+0B & IOD 2  & 0.0330  & 0.0531  & 0.2286  & 0.9262 \\
+0C & IOD 1 & 0.0441   & 0.0286  & 0.1191  & 0.6596 \\
+0C & IOD 2 & 0.0292   & 0.0290  & 0.2174  & 0.8249 \\\midrule
 \bottomrule
@@
 \begin{tablenotes}
-    \item Note: $0$ indicates numbers small enough to be below Matlab numerical accuracy.
+    \item[$(\cdot )$] Filter diverges. 
   \end{tablenotes}}
-\label{tab:IODerr}
-\end{threeparttable}
-\end{center}
-
-The Kalman filter is used in each case. The initial state covariance matrix is chosen as
-\begin{align*}
-P_0^+=\diag[50^2,\ 50^2,\ 50^2,\ (50n)^2,\ (50n)^2,\ (50n)^2],
-\end{align*}
-and the process and measurement covariance matrices are chosen as
-\begin{align*}
-Q_k=\diag[10^{-7},\ 10^{-7},\ 10^{-7},\ 10^{-9},\ 10^{-9},\ 10^{-9}], \quad R_k=\tan^2{2^\circ}I_{3\times3}\ \forall\ k.
-\end{align*}
-%\begin{align*}
-%Q_k=10^q\diag[10^{-8},\ 10^{-8},\ 10^{-8},\ 10^{-10},\ 10^{-10},\ 10^{-10}], \quad R_k=\tan^2{2^\circ}I_{3\times3}\ \forall\ k
-%\end{align*}
-%where $q\in[-1,\ 0,\ 1,\ 2]$ is varied in the simulations, thereby changing the impact of measurement updates inside the Kalman filter.
-
-Each case is simulated for every IOD estimate.
-The running time is $59540$ seconds (slightly more than $\frac23$ of a day) with time steps of $10$ seconds between measurements.
-The following RMS errors are collected from each trial:
-\begin{align}
-\bar e_{mag}=\left(\frac1N\sum_{k=1}^N\left\{1-{\frac{\norm{\r^+_k}}{\norm{\r_k}}}\right\}^2\right)^{\frac12}, \quad \bar e_{dir}=\left(\frac1N\sum_{k=1}^N
-\left\{\cos^{-1}\left(\frac{\r_k}{\norm{\r_k}}\cdot \frac{\r^+_k}{\norm{\r^+_k}}\right)\right\}^{2}
-\right)^{\frac12},
-\end{align}
-for $N$ time steps.
-These metrics are tabulated in Tables \ref{tab:ResultsTBP} and \ref{tab:ResultsJ2}.
-Over the simulated time period, figures were generated for these analyses that  serve to compare the true and estimated state magnitudes and the maximum eigenvalues of the state covariance matrices as a measure of uncertainty for the state estimates.
-Because of paper constraints, a sample of these plots are shown that highlight various aspects of the observability and filter performance.
-
-%The figure below illustrates preliminary results for angles-only relative orbit estimation with $J_2$ perturbation in the measurements.
-%Note that the EKF performs well in this particular case, as evidenced by the close agreement between the true and estimated trajectory (Figure 1a) and the decrease in the maximum eigenvalue of the covariance matrix over time (Figure 1b).
-%For the various cases to be chosen, we will evaluate the performance of the EKF (using metrics such as those illustrated in Figure 1) and seek a consistent correlation between the filter performance and the new observability measure proposed above.
-
-%\begin{center}
-%\begin{threeparttable}[h]
-%\caption{Error variables for EKF with varying $Q_k$ Considering Only Two-Body Problem Forces}
-%\begin{tabularx}{\textwidth}
-%{
-%>{$}c<{$}
-%*{1}{>{$}c<{$}} |
-%*{4}{>{$}c<{$}} |
-%*{4}{>{$}c<{$}}
-%}
-%% & \bar e_{mag} & \bar e_{dir}
-%\toprule
-%\multirow{2}{*}{Case} & \multirow{2}{*}{IOD} & \multicolumn{4}{c}{$\bar e_{mag}$} & \multicolumn{4}{c}{$\bar e_{dir}$} \\
-%& & q=-1 & q=0 & q=1 & q=2 & q=-1 & q=0 & q=1 & q=2 \\\midrule
-%0A & IOD 1 & 0.0895 &  0.0611  &  0.3766  &  0.2861  &  0.0430  &  0.0135  &  0.0144  &  0.0183 \\
-%0A & IOD 2 & 0.1793 &  0.2811  &  0.5145  &  0.7716  &  0.0203  &  0.0313  &  0.0691  &  0.1046 \\
-%0B & IOD 1 & 0.3241 &  0.0531  &  0.5186  &  0.0620  &  0.1695  &  0.0143  &  0.0810  &  0.0172 \\
-%0B & IOD 2 & 0.2330 &  0.2115  &  0.3092  &  0.2733  &  0.1007  &  0.0879  &  0.0804  &  0.1048 \\
-%0C & IOD 1 & 0.0973 &   0.0748 &   0.2723 &   0.0607 &   0.0292 &   0.0068  &  0.1362 &   0.0443 \\
-%0C & IOD 2 & 0.1473 &   0.1017 &   0.1707 &   0.3545  &  0.0702  &  0.0517  &  0.0614  &  0.1224 \\
-%\midrule
-%\mathbf{1} & IOD 1 & \hilight{8.922} &   \hilight{2.964}   & \hilight{4.078}  &  \hilight{3.571} &   0.0071  &  0.0089  &  0.0109  &  0.0418 \\
-%\mathbf{1} & IOD 2 & \hilight{13.635} &  \hilight{14.121} &  \hilight{42.486} &  \hilight{42.573} &   0.0113 &   0.0107  &  0.0445  &  0.0115 \\
-%\midrule
-%2A & IOD 1 & 0.2045 &   0.0629 &   0.1592  &  0.1329  &  0.0041 &   0.0041  &  0.0043  &  0.0039 \\
-%2A & IOD 2 & 0.1013 &   0.1797  &  0.1348  &  0.2247  &  0.0050 &   0.0048 &   0.0065 &   0.0061 \\
-%2B & IOD 1 & \hilight{0.9886} &   \hilight{0.9895}   & \hilight{0.9843}  &  \hilight{0.9851}  &  0.0114 &   0.0116  &  0.0114  &  0.0116 \\
-%2B & IOD 2 & \hilight{0.9557} &   \hilight{0.9685} &   \hilight{0.9278}  &  \hilight{0.9431} &   0.0105  &  0.0108 &   0.0113 &   0.0153 \\
-%2C & IOD 1 & 0.0219 &   0.0184  &  0.0258  &  0.0238 &   0.0026 &   0.0031  &  0.0044 &   0.0038 \\
-%2C & IOD 2 & 0.0303 &   0.0287  &  0.0360  &  0.0470  &  0.0044  &  0.0054  &  0.0059 &   0.0062 \\
-%\midrule
-%3A & IOD 1 &  0.2377 &   0.1684  &  0.0486 &   0.0695 &   0.0094 &   0.0070 &   0.0040 &  0.0039 \\
-%3A & IOD 2 &  0.0854  &  0.0874 &   0.0603  &  0.0544  &  0.0035 &   0.0033  &  0.0053  &  0.0042 \\
-%3B & IOD 1 &  0.0724 &   0.0260 &   0.0680  &  0.0289  &  0.0035 &   0.0029  &  0.0033  &  0.0027 \\
-%3B & IOD 2 &  0.0385  &  0.0432  &  0.1459  &  0.0284  &  0.0038  &  0.0037  &  0.0070  &  0.0047 \\
-%3C & IOD 1 &  \hilight{2.2433}  &  \hilight{1.5231} &   \hilight{2.0611} &   \hilight{1.5110}  &  0.1803  &  0.1172  &  0.1595  &  0.1103 \\
-%3C & IOD 2 &  \hilight{4.0635} &   \hilight{1.7717}  & \hilight{10.567}  &  \hilight{4.0277} &   0.4475 &   0.1506 &   0.9011 &   0.4028 \\
-%\midrule
-%4A & IOD 1 &  \hilight{6.0259} &   \hilight{0.0464}  &  \hilight{6.0706}  &  \hilight{0.0325}  &  1.3353  &  0.0087   & 1.3182  &  0.0065 \\
-%4A & IOD 2 &  6.3646 &   0.0173  & 11.8882 &   0.0256 &   1.3423  &  0.0064  &  1.5319 &   0.0066 \\
-%4B & IOD 1 &  \hilight{0.2922}  &  \hilight{0.0258}  &  \hilight{0.2723}  &  \hilight{0.0182}   & 0.0929 &   0.0032  &  0.0789  &  0.0036 \\
-%4B & IOD 2 &  0.1815  &  0.0217  &  0.3155 &   0.0771  &  0.0268   & 0.0034  &  0.1061 &   0.0044 \\
-%4C & IOD 1 &  0.0247  &  \hilight{0.0421}  &  2.3699  &  0.0182  &  0.0032  &  0.0032  &  0.8486 &   0.0028 \\
-%4C & IOD 2 &  0.0200 &   0.0128  &  0.0278  &  0.0299 &   0.0040  &  0.0029  &  0.0034  &  0.0033 \\
-%\bottomrule
-%\end{tabularx}
-%{\small
-%\begin{tablenotes}
-%    \item \hilight{Filter diverges}
-%  \end{tablenotes}}
-%\label{tab:ResultsTBP}
-%\end{threeparttable}
-%\end{center}
-%
-%\begin{center}
-%\begin{threeparttable}[h]
-%\caption{Error variables for EKF with varying $Q_k$ with $J_2$ Perturbations Included}
-%\begin{tabularx}{\textwidth}
-%{
-%>{$}c<{$}
-%*{1}{>{$}c<{$}} |
-%*{4}{>{$}c<{$}} |
-%*{4}{>{$}c<{$}}
-%}
-%\toprule
-%\multirow{2}{*}{Case} & \multirow{2}{*}{IOD} & \multicolumn{4}{c}{$\bar e_{mag}$} & \multicolumn{4}{c}{$\bar e_{dir}$} \\
-%& & q=-1 & q=0 & q=1 & q=2 & q=-1 & q=0 & q=1 & q=2 \\\midrule
-%0A & IOD 1 & 0.0553  &  0.1105  &  0.1918  &  0.2298  &  0.0059  &  0.0062  &  0.0069  &  0.0069 \\
-%0A & IOD 2 & 0.2177  &  0.3227  &  0.3140  &  0.5013  &  0.0082  &  0.0074  &  0.0129  &  0.0085 \\
-%0B & IOD 1 & 0.0473  &  0.1246  &  0.7262  &  0.0655  &  0.0077  &  0.0088  &  0.0505  &  0.0066\\
-%0B & IOD 2 & 0.8872  &  0.4186  &  0.1571  &  0.1561  &  0.0390  &  0.0168  &  0.0086  &  0.0293 \\
-%0C & IOD 1 & 0.0652  &  0.0809  &  0.0900  &  0.1113  &  0.0082  &  0.0079  &  0.0061  &  0.0106 \\
-%0C & IOD 2 & 0.0367  &  0.0204  &  0.1002  &  0.7253  &  0.0145  &  0.0084  &  0.0154  &  0.0579 \\
-%\midrule
-%\mathbf{1} & IOD 1 & \hilight{9.1098}  &  \hilight{2.900}  &  \hilight{4.059}  &  \hilight{3.567}  &  0.0071  &  0.0090  &  0.0109  &   0.0418 \\
-%\mathbf{1} & IOD 2 & \hilight{13.594} & \hilight{14.033} & \hilight{42.381} & \hilight{42.979}  & 0.0113   & 0.0107   & 0.0445   &   0.0115 \\
-%\midrule
-%2A & IOD 1 & 0.2301  &  0.0423  &  0.1978  &  0.1740  &  0.0048  &  0.0049  &  0.0047  &  0.0042 \\
-%2A & IOD 2 & 0.1058  &  0.2164  &  0.1638  &  0.2644  &  0.0051  &  0.0049  &  0.0065  &  0.0061 \\
-%2B & IOD 1 & \hilight{0.9895}  &  \hilight{0.9902}  &  \hilight{0.9853}  &  \hilight{0.9859}  &  0.0115  &  0.0117  &  0.0115  &  0.0117 \\
-%2B & IOD 2 & \hilight{0.9585}  &  \hilight{0.9701}  &  \hilight{0.9307}  &  \hilight{0.9455}  &  0.0106  &  0.0109  &  0.0114  &  0.0153 \\
-%2C & IOD 1 & 0.0481  &  0.0474  &  0.0563  &  0.0415  &  0.0041  &  0.0042  &  0.0047  &  0.0042 \\
-%2C & IOD 2 & 0.0535  &  0.0397  &  0.0536  &  0.0705  &  0.0044  &  0.0056  &  0.0059  &  0.0062 \\
-%\midrule
-%3A & IOD 1 &  0.2462  &  0.1827  &  0.0516  &  0.0699 &   0.0097 &   0.0077  &  0.0041 &   0.0041 \\
-%3A & IOD 2 &  0.1038  &  0.0986  &  0.0643  &  0.0692 &   0.0036  &  0.0034  &  0.0053  &  0.0041 \\
-%3B & IOD 1 &  0.0770  &  0.0310  &  0.0739  &  0.0458 &   0.0038 &   0.0034  &  0.0035  &  0.0031 \\
-%3B & IOD 2 &  0.0461  &  0.0585   & 0.1485 &   0.0391  &  0.0039  &  0.0040  &  0.0071  &  0.0047 \\
-%3C & IOD 1 &  \hilight{2.239}  &  \hilight{1.522}  &  \hilight{2.063}  &  \hilight{1.514}  &  0.1817  &  0.1179  &  0.1614  &  0.1116 \\
-%3C & IOD 2 &  \hilight{4.188}  &  \hilight{1.821}  & \hilight{10.136}  &  \hilight{4.510} &   0.4873 &   0.1592  &  0.9195 &   0.5532 \\
-%\midrule
-%4A & IOD 1 &  \hilight{6.013}  &  \hilight{0.0454}  &  \hilight{6.359}  &  \hilight{0.031}  &  1.3131  &  0.0086  &  1.4235 &   0.0066 \\
-%4A & IOD 2 &  6.3272  &  0.0197  & 12.0062 &  0.0248  &  1.3116  &  0.0065  &  1.5310  &  0.0066 \\
-%4B & IOD 1 &  \hilight{0.2936}  &  \hilight{0.0272}  &  \hilight{0.2773}  &  \hilight{0.0217}  &  0.0935  &  0.0032  &  0.0814  &  0.0038 \\
-%4B & IOD 2 &  0.1853  &  0.0210  &  0.5244  &  0.0785  &  0.0293  &  0.0036  &  0.2305  &  0.0045 \\
-%4C & IOD 1 &  0.0260  &  \hilight{0.0410}  &  2.0105  &  0.0164  &  0.0034  &  0.0031  &  0.9144  &  0.0030 \\
-%4C & IOD 2 &  0.0210  &  0.0145  &  0.0303  &  0.0349  &  0.0042  &  0.0030  &  0.0036  &  0.0034 \\
-%\bottomrule
-%\end{tabularx}
-%{\small
-%\begin{tablenotes}
-%    \item \hilight{Filter diverges}
-%  \end{tablenotes}}
-%\label{tab:ResultsJ2}
-%\end{threeparttable}
-%\end{center}
-
-\begin{center}
-\begin{threeparttable}[h]
-\caption{Error variables for EKF Considering Only Two-Body Problem Forces}
-\begin{tabularx}{.58\textwidth}
-{
->{$}c<{$}
-*{1}{>{$}c<{$}} |
-*{2}{>{$}c<{$}} |
-*{2}{>{$}c<{$}}
-}
-\toprule
-\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{IOD1} & \multicolumn{2}{c}{IOD2} \\
-& & \bar e_{mag} & \bar e_{dir} & \bar e_{mag} & \bar e_{dir} \\\midrule
-0A & 10^{11.1486} &  0.3766  &  0.0144  &  0.5145  &  0.0691 \\
-0B & 10^{10.9387} &  0.5186  &  0.0810  &  0.3092  &  0.0804\\
-0C & 10^{10.9851} &  0.2723  &  0.1362 &  0.1707 &  0.0614\\
-\midrule
-\mathbf{1} & 10^{16.6871} & \cellcolor{green} 4.078  & \cellcolor{green}  0.0109 & \cellcolor{green} 42.486 & \cellcolor{green}  0.0445 \\
-\midrule
-2A & 10^{10.6492} &  0.1592  &   0.0043  &  0.1348  &   0.0065   \\
-2B & 10^{10.4163} & \cellcolor{green} 0.9843   & \cellcolor{green}  0.0114  & \cellcolor{green}  0.9278 & \cellcolor{green}   0.0113 \\
-2C & 10^{10.2155} & 0.0258 &  0.0044 & 0.0360 &  0.0059 \\
-\midrule
-3A & 10^{8.3490} &  0.0486 & 0.0040   &  0.0603  &  0.0053 \\
-3B & 10^{8.5372} &  0.0680  &  0.0033 &  0.1459 &  0.0070 \\
-3C & 10^{8.6586} & \cellcolor{green}  2.0611 & \cellcolor{green}  0.1595   & \cellcolor{green}  10.567  & \cellcolor{green}   0.9011 \\
-\midrule
-4A & 10^{7.9062} & \cellcolor{green}  6.0706  & \cellcolor{green} 1.3182  &  11.8882   &  1.5319  \\
-4B & 10^{7.9903} & \cellcolor{green}  0.2723  & \cellcolor{green}  0.0789   &  0.3155  &  0.1061 \\
-4C & 10^{8.0645} &  2.3699 &  0.8486   &  0.0278 &  0.0034 \\
-\bottomrule
-\end{tabularx}
-{\small
-\begin{tablenotes}
-    \item \hilight{Filter diverges}
-  \end{tablenotes}}
-\label{tab:ResultsTBP}
 \end{threeparttable}
@@
 \begin{threeparttable}[h]
-\caption{Error variables for EKF Considering $J_2$ Perturbations}
-\begin{tabularx}{.58\textwidth}
-{
->{$}c<{$}
-*{1}{>{$}c<{$}} |
-*{2}{>{$}c<{$}} |
-*{2}{>{$}c<{$}}
-}
-\toprule
-\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{IOD1} & \multicolumn{2}{c}{IOD2} \\
-& & \bar e_{mag} & \bar e_{dir} & \bar e_{mag} & \bar e_{dir} \\\midrule
-0A & 10^{11.1486} &  0.1918  &  0.0069  &  0.3140  &  0.0129 \\
-0B & 10^{10.9387} &  0.7262  &  0.0505  &  0.1571  &  0.0086\\
-0C & 10^{10.9851} &  0.0900  &  0.0061 &  0.1002  &  0.0154\\
-\midrule
-\mathbf{1} & 10^{16.6871} & \cellcolor{green} 4.059  & \cellcolor{green}  0.0109 & \cellcolor{green} 42.381 & \cellcolor{green}  0.0445 \\
-\midrule
-2A & 10^{10.6492} &  0.1978  &  0.0047  &  0.1638  &   0.0065   \\
-2B & 10^{10.4163} & \cellcolor{green}  0.9853  & \cellcolor{green}  0.9307 & \cellcolor{green}  0.9278 & \cellcolor{green}   0.0114 \\
-2C & 10^{10.2155} & 0.0563 &  0.0047 & 0.0536 &  0.0059 \\
-\midrule
-3A & 10^{8.3490} &  0.0516 & 0.0041   &  0.0643  &  0.0053 \\
-3B & 10^{8.5372} &  0.0739  &  0.0035 &  0.1485 &  0.0071 \\
-3C & 10^{8.6586} & \cellcolor{green}  2.063 & \cellcolor{green} 0.1614    & \cellcolor{green}  10.136  & \cellcolor{green}   0.9195 \\
-\midrule
-4A & 10^{7.9062} & \cellcolor{green}  6.359  & \cellcolor{green} 1.4235  &  12.0062   &  1.5310  \\
-4B & 10^{7.9903} & \cellcolor{green}  0.2773  & \cellcolor{green}  0.0814   &  0.5244  &  0.2305 \\
-4C & 10^{8.0645} &  2.0105 &  0.9144   &  0.0303 &  0.0036 \\
+\caption{Error variables for EKF with varying $Q_k$ with $J_2$ Perturbations Included}
+\begin{tabularx}{0.7\textwidth}{>{\centering $}X<{$}*{2}{>{$}c<{$}}|*{5}{>{$}c<{$}}}\toprule
+\multirow{2}{*}{Case} & \multirow{2}{*}{IOD} & \multirow{2}{*}{$e_{mag}(0)$} & \multicolumn{4}{c}{$\bar e_{mag}$} \\
+& & & q=-1 & q=0 & q=1 & q=2 \\\midrule
+0A & IOD 1 & 0.9997  & 0.1129   & 0.1405  & 0.2859  & 0.4726 \\
+0A & IOD 2 & 1.0012  & 0.0400   & 0.1470  & 0.2559  & 0.5269 \\
+0B & IOD 1 & 0.9767  & 0.0153   & 0.0553  & 0.2734  & 0.7942 \\
+0B & IOD 2 & 1.0088  & 0.0330   & 0.0531  & 0.2286  & 0.9262 \\
+0C & IOD 1 & 1.0152  & 0.0441   & 0.0286  & 0.1191  & 0.6596 \\
+0C & IOD 2 & 1.0139  & 0.0292   & 0.0290  & 0.2174  & 0.8249 \\\midrule
+\mathbf{1} & IOD 1 & 0.9999  & (4.9960)   & (20.6584)  & (52.8485)  & (159.8720) \\
+\mathbf{1} & IOD 2 & 1.0000  & (12.0855)   & (16.9126)  & (54.0381)  & (168.9124) \\\midrule
+2A & IOD 1 & 1.0550  & 0.1925   & 0.1262  & 0.1677  & 0.6115 \\
+2A & IOD 2 & 1.0238  & 0.1444   & 0.0283  & 0.1975  & 0.5445 \\
+2B & IOD 1 & \mathbf{0.0429}  & (0.9810)   & (0.9494)  & (0.9120)  & (0.7438) \\
+2B & IOD 2 & \mathbf{0.0429}  & (0.9813)   & (0.9506)  & (0.9069)  & (0.7565) \\
+2C & IOD 1 & 0.9781  & 0.0313   & 0.0205  & 0.0284  & 0.0792 \\
+2C & IOD 2 & 0.9935  & 0.0221   & 0.0230  & 0.0257  & 0.0842 \\\midrule
+3A & IOD 1 & 1.0191  & 0.0841   & 0.0180  & 0.1445  & 0.3300 \\
+3A & IOD 2 & 0.8170  & 0.0864   & 0.0416  & 0.1936  & 0.2913 \\
+3B & IOD 1 & 1.0612  & (0.9761)   & (0.9785)  & 0.0283  & 0.0985 \\
+3B & IOD 2 & 0.7685  & 0.0128   & 0.0275  & 0.0206  & 0.0864 \\
+3C & IOD 1 & 7.8962  & (1.5927)   & (2.2080)  & 0.8156  & 0.7897 \\
+3C & IOD 2 & 6.4921  & 0.9330   & 0.7311  & 0.6269  & 0.6311 \\\midrule
+4A & IOD 1 & \mathbf{9.7347}  & (6.2650)   & (6.1141)  & (8.8398)  & (10.0667) \\
+4A & IOD 2 & 0.8425  & 0.0212   & 0.0506  & 0.0228  & 0.0268 \\
+4B & IOD 1 & 1.9276  & (0.2523)   & 0.2016  & 0.1417  & 0.1669 \\
+4B & IOD 2 & 0.9379  & 0.0068   & 0.0147  & 0.0483  & 0.0157 \\
+4C & IOD 1 & 1.1161  & 0.0347   & 0.0713  & 0.0187  & 0.0352 \\
+4C & IOD 2 & 0.9646  & 0.0185   & 0.0191  & 0.0094  & (0.9824) \\
 \bottomrule
@@
 \begin{tablenotes}
-    \item \hilight{Filter diverges}
+    \item[$(\cdot )$] Filter diverges. 
   \end{tablenotes}}
-\label{tab:ResultsJ2}
 \end{threeparttable}
@@
 
-These data from the simulations show several important aspects and challenges with relative orbit determination involving angles-only measurements.
-First, an important observation from the data is that the IOD guess is critical to the success of the Kalman filter.
-Cases 2B, 3C, 4A, and 4B experience filter divergence when the IOD estimate is largely inaccurate, shown in Tables \ref{tab:IODerr}, \ref{tab:ResultsTBP}, and \ref{tab:ResultsJ2}, even when cases with similar observabilities do not experience filter divergence.
-The Kalman filter, which is based on linear dynamic systems, is modeled as a Markov chain and hence is memoryless.
-Therefore when nothing more than an inaccurate IOD prediction is poorly assumed to contain accurate information about the system, the recursive formulation of the Kalman filter is unable to converge.
-In these diverging cases, a greater level of observability or an algorithm that relies on a larger history of measurements may improve convergence.
-
-In contrast, every variation of Case $\mathbf{1}$ has a closer IOD than the IODs of other cases, but experiences the most severe divergence.
-Case $\mathbf{1}$ has a greater condition number of $\mathcal W$, which corresponds with low observability according to the proposed observability measure.
-A series of angles-only measurements may correspond with a large range of possible relative orbits, even with a low variation in those measurements.
-Therefore, small levels of noise in the sensor or uncertainties in the orbit prevent the Kalman filter from converging to an accurate estimate, even with an accurate IOD.
-The state magnitudes of both variations of Case $\mathbf{1}$ are illustrated against various converging cases in Figure \ref{fig:ExampesEachCase} to show how important this observability is to the filter performance.
-
+%An example is shown in Figure \ref{ODEKF} where the IOD prediction is not exactly the true state vector and $J_2$ gravitational forces affect the satellite propagation.
+The time step is fixed at $1$ second.
 \begin{figure}[h]
 \centerline{
-	\subfigure[Case 0C, IOD1]{
-		\includegraphics[width=0.35\textwidth]{Case0Cq1IOD1TBPnormX.pdf}}
+	\subfigure[Magnitude Error (true:red, estimated:blue)]{
+		\includegraphics[width=0.48\textwidth]{normX.pdf}}
 	\hfill
-	\subfigure[Case 0C, IOD2]{
-	\includegraphics[width=0.35\textwidth]{Case0Cq1IOD2TBPnormX.pdf}}
+	\subfigure[Uncertainty in estimation]{
+	\includegraphics[width=0.48\textwidth]{eigP.pdf}}
 }
-\centerline{
-	\subfigure[Case $\mathbf{1}$, IOD1]{
-		\includegraphics[width=0.35\textwidth]{Case1q1IOD1TBPnormX.pdf}}
-	\hfill
-	\subfigure[Case $\mathbf{1}$, IOD2]{
-	\includegraphics[width=0.35\textwidth]{Case1q1IOD2TBPnormX.pdf}}
-}
-\centerline{
-	\subfigure[Case 2A, IOD1]{
-		\includegraphics[width=0.35\textwidth]{Case2Aq1IOD1TBPnormX.pdf}}
-	\hfill
-	\subfigure[Case 2A, IOD2]{
-	\includegraphics[width=0.35\textwidth]{Case2Aq1IOD2TBPnormX.pdf}}
-}
-\centerline{
-	\subfigure[Case 3A, IOD1]{
-		\includegraphics[width=0.35\textwidth]{Case3Aq1IOD1TBPnormX.pdf}}
-	\hfill
-	\subfigure[Case 3A, IOD2]{
-	\includegraphics[width=0.35\textwidth]{Case3Aq1IOD2TBPnormX.pdf}}
-}
-\centerline{
-	\subfigure[Case 4C, IOD1]{
-		\includegraphics[width=0.35\textwidth]{Case4Cq1IOD1TBPnormX.pdf}}
-	\hfill
-	\subfigure[Case 4C, IOD2]{
-	\includegraphics[width=0.35\textwidth]{Case4Cq1IOD2TBPnormX.pdf}}
-}
-\caption{The magnitudes of the relative states (solid red: true, blue dashed: estimated) are shown for various two-body propagated cases with close IOD estimates where the filter only diverges when the observability is low.}\label{fig:ExampesEachCase}
+\caption{Relative orbits determination with $J_2$ perturbation}\label{ODEKF}
 \end{figure}
-
-Another observation from the data is that the $J_2$ perturbation, which is well-known to have a noticeable impact perturbing orbits from the two-body model for satellites around the earth, yields an insignificant change when compared to the data when only two-body forces are considered in the process model.
-In fact, every case when the filter diverged occurred with both the two-body problem and $J_2$ simulations.
-Examples of these results are tabulated in Tables \ref{tab:ResultsTBP} and \ref{tab:ResultsJ2} and illustrated in Figure \ref{fig:TBPvsJ2}.
-
-\begin{figure}[h]
-\centerline{
-	\subfigure[State magnitude with the two-body solution]{
-		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1TBPnormX.pdf}}
-	\hfill
-	\subfigure[State magnitude with the $J_2$ solution]{
-		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1HighFidelitynormX.pdf}}
-}
-\centerline{
-	\subfigure[Uncertainty measure with the two-body solution]{
-		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1TBPeigP.pdf}}
-	\hfill
-	\subfigure[Uncertainty measure with the $J_2$ solution]{
-		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1eigP.pdf}}
-}
-\caption{The magnitudes and maximum eigenvalues of the state covariance matrices of the relative states (solid red: true, blue dashed: estimated) in Case 3B with IOD1 serve as an example where the $J_2$ propagated simulations give very close results to those of the two-body solution.
-}\label{fig:TBPvsJ2}
-\end{figure}
-
-\section{Conclusions}
-
-It is well known that the linearized relative orbital dynamics are not observable with angles-only
-measurements.
-This paper formulates a measure of nonlinear observability, and investigates a less conservative observability criteria by using higher-order Lie derivatives.
-Numerous relative orbit determination cases are illustrated numerically by an extended Kalman filter, which shows that the degree of observability by the proposed measure and an accurate IOD prediction are useful for predicting filter convergence.
 

### 2018-06-09 Conference/AAS 2015 OD EKF/EKF_IOD.tex
added Conference folder

\documentclass[10pt]{article}
\usepackage[letterpaper,text={6.5in,8.6in},centering]{geometry}
\usepackage{amssymb,amsmath,amsthm,times,graphicx,subfigure,tabularx,booktabs,colortbl,multirow,threeparttable}
\usepackage[small]{caption}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\ave}[1]{\mathrm{E}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}

\renewcommand{\r}{\mathbf{r}}
\renewcommand{\u}{\mathbf{u}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\e}{\mathbf{e}}

\newcommand{\bfi}{\bfseries\itshape\selectfont}

\renewcommand{\baselinestretch}{1.2}

\date{}

\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{remark}{Remark}

\renewcommand{\thesubsection}{\arabic{subsection}. }
\renewcommand{\thesubsubsection}{\arabic{subsection}.\arabic{subsubsection} }

%\graphicspath{{/Users/tylee@seas.gwu.edu/Documents/Research/Estimation/Simulation/GloUncPro/DistSph/Build/Products/Debug/Diff/}}

\begin{document}
\thispagestyle{empty}

\noindent{\large \bf Air Force Summer Faculty Fellowship Program:}\vspace*{0.1cm}\\
\noindent{\Large \bf Nonlinear Observability for Relative Orbits with Line-of-Sight Measurements}\vspace*{0.1cm}\\
\noindent{T. Alan Lovell and Taeyoung Lee}\\
\noindent{July 25, 2013}\\

\subsection{Numerical Examples}

\paragraph{Initial Condition}

Five cases of the IOD data are simulated.

\paragraph{Extended Kalman Filter}

The covariance of the process noise is parameterized as follows:
\begin{align*}
Q_k = 10^{q}\times  \mathrm{diag}[10^{-8}I_{3\times 3}, \,10^{-10}I_{3\times 3}],
\end{align*}
where the integer parameter $q$ is varied as $q=-1,\,0,\,1,\,2$.

The resulting estimation errors are summarized at the table at the next page. Overall, the filter performed relatively well: 21 cases converged out of 26 cases for a proper level of process noise.

The cases where filter diverted can be categorized as follows:
\begin{itemize}
\item \textbf{IOD} When the initial estimate is poor, more explicitly when $e_{mag}(0)>9$ or $e_{mag}<0.05$, the filter diverged. This corresponds to the following cases:\\
Case 2B, Case 4A (IOD1)
\item \textbf{HCW} When the nonlinear trajectory is too close to the linearized HCW model, the filter diverged. This corresponds to the following case:
Case 1
\item \textbf{Process noise} Filter diverged when the process noise is too low or too high. This corresponds to the following cases:
Case 3B (IOD1), Case 3C (IOD1), Case 4B (IOD1), Case 4C (IOD2)
\end{itemize}

The numerical results when $q=0$ are illustrated at Figures 1-13. From these figures, filter divergence can be determined as follows:
\begin{itemize}
\item \textbf{Too low $\lambda_{\max}{P}$ } The maximum eigenvalue of $P$ drops instantaneously and remain at a very low number. It represents that the filter is falsely too confident about a wrong estimate. This corresponds to the following cases:\\
Case 1, Case 2B, Case 3B (IOD1)
\item \textbf{Too high error residual} The residual error is too large. This corresponds to the following cases:
Case 3C (IOD1), Case 4A (IOD1)
\end{itemize}

\begin{center}
\begin{threeparttable}[h]
\caption{Error variables for EKF with varying $Q_k$}
\begin{tabularx}{0.7\textwidth}{>{\centering $}X<{$}*{2}{>{$}c<{$}}|*{5}{>{$}c<{$}}}\toprule
\multirow{2}{*}{Case} & \multirow{2}{*}{IOD} & \multirow{2}{*}{$e_{mag}(0)$} & \multicolumn{4}{c}{$\bar e_{mag}$} \\
& & & q=-1 & q=0 & q=1 & q=2 \\\midrule
0A & IOD 1 & 0.9997  & 0.1129   & 0.1405  & 0.2859  & 0.4726 \\
0A & IOD 2 & 1.0012  & 0.0400   & 0.1470  & 0.2559  & 0.5269 \\
0B & IOD 1 & 0.9767  & 0.0153   & 0.0553  & 0.2734  & 0.7942 \\
0B & IOD 2 & 1.0088  & 0.0330   & 0.0531  & 0.2286  & 0.9262 \\
0C & IOD 1 & 1.0152  & 0.0441   & 0.0286  & 0.1191  & 0.6596 \\
0C & IOD 2 & 1.0139  & 0.0292   & 0.0290  & 0.2174  & 0.8249 \\\midrule
\mathbf{1} & IOD 1 & 0.9999  & (4.9960)   & (20.6584)  & (52.8485)  & (159.8720) \\
\mathbf{1} & IOD 2 & 1.0000  & (12.0855)   & (16.9126)  & (54.0381)  & (168.9124) \\\midrule
2A & IOD 1 & 1.0550  & 0.1925   & 0.1262  & 0.1677  & 0.6115 \\
2A & IOD 2 & 1.0238  & 0.1444   & 0.0283  & 0.1975  & 0.5445 \\
2B & IOD 1 & \mathbf{0.0429}  & (0.9810)   & (0.9494)  & (0.9120)  & (0.7438) \\
2B & IOD 2 & \mathbf{0.0429}  & (0.9813)   & (0.9506)  & (0.9069)  & (0.7565) \\
2C & IOD 1 & 0.9781  & 0.0313   & 0.0205  & 0.0284  & 0.0792 \\
2C & IOD 2 & 0.9935  & 0.0221   & 0.0230  & 0.0257  & 0.0842 \\\midrule
3A & IOD 1 & 1.0191  & 0.0841   & 0.0180  & 0.1445  & 0.3300 \\
3A & IOD 2 & 0.8170  & 0.0864   & 0.0416  & 0.1936  & 0.2913 \\
3B & IOD 1 & 1.0612  & (0.9761)   & (0.9785)  & 0.0283  & 0.0985 \\
3B & IOD 2 & 0.7685  & 0.0128   & 0.0275  & 0.0206  & 0.0864 \\
3C & IOD 1 & 7.8962  & (1.5927)   & (2.2080)  & 0.8156  & 0.7897 \\
3C & IOD 2 & 6.4921  & 0.9330   & 0.7311  & 0.6269  & 0.6311 \\\midrule
4A & IOD 1 & \mathbf{9.7347}  & (6.2650)   & (6.1141)  & (8.8398)  & (10.0667) \\
4A & IOD 2 & 0.8425  & 0.0212   & 0.0506  & 0.0228  & 0.0268 \\
4B & IOD 1 & 1.9276  & (0.2523)   & 0.2016  & 0.1417  & 0.1669 \\
4B & IOD 2 & 0.9379  & 0.0068   & 0.0147  & 0.0483  & 0.0157 \\
4C & IOD 1 & 1.1161  & 0.0347   & 0.0713  & 0.0187  & 0.0352 \\
4C & IOD 2 & 0.9646  & 0.0185   & 0.0191  & 0.0094  & (0.9824) \\
\bottomrule
\end{tabularx}
{\small
\begin{tablenotes}
    \item[$(\cdot )$] Filter diverges. 
  \end{tablenotes}}
\end{threeparttable}
\end{center}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0A_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0A_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0A_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0A_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 0A (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0B_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0B_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0B_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0B_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 0B (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0C_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0C_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0C_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0C_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 0C (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_1_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_1_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_1_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_1_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 1 (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2A_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2A_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2A_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2A_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 2A (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2B_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2B_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2B_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2B_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 2B (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2C_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2C_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2C_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2C_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 2C (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3A_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3A_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3A_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3A_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 3A (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3B_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3B_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3B_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3B_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 3B (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3C_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3C_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3C_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3C_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 3C (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4A_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4A_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4A_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4A_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 4A (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4B_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4B_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4B_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4B_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 4B (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4C_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4C_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4C_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4C_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 4C (IOD1: blue, IOD2: black)}
\end{figure}

\end{document}

### 2018-06-09 Conference/AAS 2015 OD EKF/Final Report 13.tex
added Conference folder

\documentclass[10pt]{article}
\usepackage[letterpaper,text={6.5in,8.6in},centering]{geometry}
\usepackage{amssymb,amsmath,amsthm,times,graphicx,subfigure,tabularx,booktabs,colortbl}
\usepackage[small]{caption}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\ave}[1]{\mathrm{E}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}

\newcommand{\x}{\ensuremath{\mathbf{x}}}
\renewcommand{\r}{\mathbf{r}}
\renewcommand{\u}{\mathbf{u}}
\newcommand{\y}{\mathbf{y}}

\newcommand{\bfi}{\bfseries\itshape\selectfont}

\renewcommand{\baselinestretch}{1.2}

\date{}

\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{remark}{Remark}

\renewcommand{\thesubsection}{\arabic{subsection}. }
\renewcommand{\thesubsubsection}{\arabic{subsection}.\arabic{subsubsection} }

%\graphicspath{{/Users/tylee@seas.gwu.edu/Documents/Research/Estimation/Simulation/GloUncPro/DistSph/Build/Products/Debug/Diff/}}

\begin{document}
\thispagestyle{empty}

\noindent{\large \bf Air Force Summer Faculty Fellowship Program Final Report:}\vspace*{0.1cm}\\
\noindent{\Large \bf Nonlinear Observability and Estimation\\for Relative Orbits with Line-of-Sight Measurements}\vspace*{0.1cm}\\
\noindent{Taeyoung Lee}\\
\noindent{July 29, 2013}\\

%\subsubsection*{Abstract}

\tableofcontents

\subsection{Relative Orbital Dynamics}\label{sec:PD}

Suppose that a chief satellite on a circular orbit with an orbital radius of $a$. Define a local-vertical, local-horizontal (LVLH) whose origin is located at the chief satellite. In this frame, the $x$ axis is along the radial direction, the $y$-axis is along the velocity vector of the chief, and the $z$ axis is normal to the orbital plane. The LVLH frame is rotating with the angular velocity of $\mathbf{\omega}=[0,0,n]^T$, where $n=\sqrt{\frac{\mu}{a^3}}$ is the mean motion of the chief satellite. Let the relative position of a deputy satellite with respect to the chief satellite be given by $\mathbf{r}=[x,y,z]^T\in\Re^3$ in the LVLH frame.

\subsubsection{Nonlinear Equations of Motion}

Nonlinear equations of motion for the relative motion can be derived as follows based on Lagrangian mechanics.

\paragraph{Lagrangian}

Considering that the LVLH frame is rotating with the angular velocity of $\mathbf{\omega}=[0,0,n]^T$, the inertial velocity of the deputy satellite is given by
\begin{align*}
\mathbf{v}_{deputy}
=
\mathbf{v}_{chief} + \dot{\r} + \mathbf{\omega}\times \mathbf{r} = 
\begin{bmatrix}
0 \\ na \\ 0
\end{bmatrix}
+
\begin{bmatrix}
\dot x \\ \dot y \\ \dot z
\end{bmatrix}
+
\begin{bmatrix}
-n y \\ nx \\ 0
\end{bmatrix}
=
\begin{bmatrix}
\dot x - ny \\ \dot y + nx + na \\ \dot z
\end{bmatrix}.
\end{align*}
Therefore, the (normalized) kinetic energy of the deputy satellite is 
\begin{align*}
T = \frac{1}{2} \braces{(\dot x -ny)^2+(\dot y + nx +na)^2 + \dot z^2}.
\end{align*}
The gravitational potential energy is 
\begin{align*}
U = -\frac{\mu}{\sqrt{(x+a)^2 + y^2 + z^2}} = -\frac{\mu}{\|\r_a\|},
\end{align*}
where $\r_a=[x+a,y,z]^T\in\Re^3$, which represents the location of the deputy from the central body (e.g, Earth). Therefore, the Lagrangian of the deputy satellite is expressed in terms of $(x,y,z)$ as 
\begin{align}
L & = T-U = \frac{1}{2} \braces{(\dot x -ny)^2+(\dot y + nx +na)^2 + \dot z^2}
+\frac{\mu}{\sqrt{(x+a)^2 + y^2 + z^2}}.
\end{align}

\paragraph{Euler-Lagrange Equations}

%The derivatives of the Lagrangian are given by
%\allowdisplaybreaks
%\begin{align*}
%\frac{d}{dt}\deriv{L}{\dot x} & = \ddot x - n\dot y\\
%\frac{d}{dt}\deriv{L}{\dot y} & = \ddot y + n\dot x \\
%\frac{d}{dt}\deriv{L}{\dot z} & = \ddot z,\\
%\deriv{L}{x} &= n\dot y + n^2x +n^2 a - \frac{\mu (x+a)}{((x+a)^2 + y^2 + z^2)^{3/2}}\\
%\deriv{L}{y} &= -n\dot x + n^2 y  - \frac{\mu y}{((x+a)^2 + y^2 + z^2)^{3/2}}\\
%\deriv{L}{z} &=  - \frac{\mu z}{((x+a)^2 + y^2 + z^2)^{3/2}}.
%\end{align*}
Substituting these into the Euler-Lagrange equations, we obtain the nonlinear equations of motion for the relative orbit as follows:
\begin{align}
\ddot x - 2n\dot y -n^2 x&=n^2 a - \frac{\mu (x+a)}{((x+a)^2 + y^2 + z^2)^{3/2}},\label{eqn:ddotx}\\
\ddot y + 2n\dot x -n^2 y &=   - \frac{\mu y}{((x+a)^2 + y^2 + z^2)^{3/2}},\label{eqn:ddoty}\\
\ddot z &= - \frac{\mu z}{((x+a)^2 + y^2 + z^2)^{3/2}}.\label{eqn:ddotz}
\end{align}

These can be written as 
\begin{align}
\dot{\mathbf{x}} = f (\mathbf{x}), \label{eqn:xxdot}
\end{align}
where the state vector is $\mathbf{x}=[x,y,z,\dot x, \dot y,\dot z]^T\in\Re^6$, and
\begin{align}
f(\mathbf{x}) = \begin{bmatrix}
\dot x\\
\dot y\\
\dot z\\
2n\dot y +n^2 (x+a) - \dfrac{\mu (x+a)}{((x+a)^2 + y^2 + z^2)^{3/2}}\\
-2n\dot x +n^2 y  - \dfrac{\mu y}{((x+a)^2 + y^2 + z^2)^{3/2}}\\
- \dfrac{\mu z}{((x+a)^2 + y^2 + z^2)^{3/2}}
\end{bmatrix}
=\begin{bmatrix}
\dot \r \\
-2\omega\times \dot\r - \omega\times(\omega\times \r_a) -\dfrac{\mu \r_a}{\|\r_a\|^3}
\end{bmatrix},\label{eqn:f}
\end{align}
where $\r=[x,y,z]^T\in\Re^3$, $\r_a = [x+a,y,z]^T\in\Re^3$, and $\dot\r =[\dot x,\dot y,\dot x]^T\in\Re^3$.

%\subsubsection{Equilibrium and Jacobi-Constant}
%
%\paragraph{Equilibrium}
%
%Define an effective (negative) potential function as follows:
%\begin{align*}
%U(x,y,z) = \frac{1}{2}n^2 ((x+a)^2+y^2) + \frac{\mu}{\sqrt{(x+a)^2 + y^2 + z^2}}.
%\end{align*}
%Then, the nonlinear equation of motion can be written as
%\begin{align}
%\ddot x - 2n\dot y = \deriv{U}{x},\\
%\ddot y + 2n\dot x = \deriv{U}{y},\\
%\ddot z = \deriv{U}{z}.\label{eqn:zddotU}
%\end{align}
%
%Then, the equilibrium of the nonlinear equations of motion corresponds to the critical point of the effective potential, where the derivatives of $U$ are zero. From \refeqn{zddotU}, $z=0$ when $\deriv{U}{z}=0$. This implies that there are only planar equilibrium. When $z=0$, the effective potential can be written as
%\begin{align*}
%U(r_a) = \frac{1}{2}n^2 r_a^2 + \frac{\mu}{r_a},
%\end{align*}
%where $r_a = \sqrt{(x+a)^2 + y^2}$. The critical points of $U(r_a)$ is given by
%\begin{align*}
%\deriv{U(r_a)}{r_a} = n^2 r_a - \frac{\mu}{r_a^2} = 0\quad\Rightarrow\quad r_a^3 = \frac{\mu}{n^2} = a^3.
%\end{align*}
%Therefore, there are infinitely many equilibria given by
%\begin{align}
%r_a = \sqrt{(x+a)^2 + y^2} = a,
%\end{align}
%which corresponds to the cases where the deputy satellite is on the same circular orbit of the chief satellite. 
%
%\paragraph{Jacobi Constant}
%
%Define a Jacobi constant:
%\begin{align}
%C & = 2U-(\dot x^2 + \dot y^2 + \dot z^2)\nonumber\\
%& = 
%n^2 ((x+a)^2+y^2) + \frac{2\mu}{\sqrt{(x+a)^2 + y^2 + z^2}}
%-(\dot x^2 + \dot y^2 + \dot z^2).
%\end{align}
%We can show that the Jacobi constant is fixed along the solution of the nonlinear equations of motion as follows. The time-derivative of $C$ is given by
%\begin{align*}
%\frac{1}{2}\frac{dC}{dt} & =  \deriv{U}{x}\dot x + \deriv{U}{y}\dot y + \deriv{U}{z}\dot z-\dot x \ddot x - \dot y \ddot y - \dot z \ddot z\\
%& = -\dot x (2n\dot y) + \dot y (2n\dot x) = 0.
%\end{align*}
%
%
%
%For a given initial condition $\r(0),\dot \r(0)$, we compute the value of the Jacobi constant $C$. Then, the feasible region of the solution of the nonlinear  equations of motion is described by
%\begin{align}
%v^2 = 2U - C = n^2 ((x+a)^2+y^2) + \frac{2\mu}{\sqrt{(x+a)^2 + y^2 + z^2}}-C \geq 0.
%\end{align}
%
%\paragraph{Feasible Region of Planar Maneuvers}
%
%For a planar maneuver, this condition can be rewritten as
%\begin{align}
%n^2 r_a^2 + \frac{2\mu}{r_a}\geq C,\text{\quad or equivalently,\quad} \parenth{\frac{r_a}{a}}^2 + 2\parenth{\frac{r_a}{a}}^{-1} \geq \frac{a}{\mu}C
%\end{align}
%The curve for $\parenth{\frac{r_a}{a}}^2 + 2\parenth{\frac{r_a}{a}}^{-1}$ is illustrated at \reffig{jacobi_cri}.
%
%
%\begin{figure}
%\centerline{
%\subfigure[The curve of $\parenth{\frac{r_a}{a}}^2 + 2\parenth{\frac{r_a}{a}}^{-1}$]{
%	\includegraphics[width=0.38\textwidth]{jacobi_cri.pdf}\label{fig:jacobi_cri}}
%	\hspace*{0.1\textwidth}
%\subfigure[The forbidden region of the deputy is the inside of the circular strip $C=\frac{4\mu}{a}$.]{
%	\includegraphics[width=0.41\textwidth]{jacobi_for.pdf}\label{fig:feafor}}}
%\caption{Relative orbit analysis using the Jacobi constant}
%\end{figure}
%
%
%The left hand side of the above expression is minimized when $r_a=a$, and the corresponding minimum value is given by
%\begin{align}
%1^2 + 2\times 1^{-1} = 3.
%\end{align}
%Therefore, if $C\leq 3\frac{\mu}{a}$, then the above inequality is always satisfied. In other words, if $C\leq 3\frac{\mu}{a}$, or if the Jacobi constant $C$ is less than three times of the orbital energy of the chief, there is no forbidden region for the deputy satellite, and the deputy can coast to anywhere.
%
%If $C> 3\frac{\mu}{a}$, the feasible region is described by
%\begin{align*}
%r_a \leq \alpha_1 <a \text{\quad or\quad} r_a \geq \alpha_2> a,
%\end{align*}
%for some positive constants $\alpha_1,\alpha_2$. This states that the deputy satellite cannot pass through a circular strip that contains the chief. In other words, the deputy stays inside of a circular orbit smaller than the circular orbit of the chief, or the deputy stays outside of a circular orbit larger than the circular orbit of the chief. 
%
%For example, when $C=4\frac{\mu}{a}$, we have $\alpha_1=0.5392a$ and $\alpha_2=1.6751a$. The corresponding boundary between the feasible region and the forbidden region of the deputy are illustrated at \reffig{feafor}, where the forbidden region corresponds to a donut-shaped area between two circles, i.e., the deputy can stay either inside of the smaller circle, or the outside of the larger circle. A  sample trajectory of the deputy inside of the smaller circle is also illustrated by blue curve.

%\subsubsection{Linearized Equations of Motion}
%
%To implement an extended Kalman filter, the nonlinear equation of motion \refeqn{xxdot} should be linearized. The variation of $f(\mathbf{x})$ can be written as
%\begin{align*}
%\delta f(\mathbf{x}) = 
%\begin{bmatrix}
%\delta\dot\r\\
%-2\omega\times\delta \dot\r - \omega\times(\omega\times\delta \r) -\dfrac{\mu \delta\r}{\|\r_a\|^3} + 3\dfrac{\mu \r_a\r_a^T}{\|\r_a\|^5}\delta\r
%\end{bmatrix}.
%\end{align*}
%Using this, the system matrix of the linearized equations of motion is given by
%\begin{align}
%A(\mathbf{x}) & = \deriv{f(\mathbf{x})}{\mathbf{x}}= \begin{bmatrix}
%0_{3\times 3} & I_{3\times 3} \\
%-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5} & -2[\omega]_\times 
%\end{bmatrix},
%\end{align}
%where $[\omega]_\times\in\Re^{3\times 3}$ is defined as
%\begin{align*}
%[\omega]_\times = \begin{bmatrix} 0 & -n & 0 \\ n & 0 & 0 \\ 0 & 0 & 0 \end{bmatrix}.
%\end{align*}
%The system matrix $A$ can be rewritten in terms of the coordinates as
%\begin{align}
%A(\mathbf{x}) = \begin{bmatrix}
%0 & 0 & 0 & 1 & 0 & 0\\
%0 & 0 & 0 & 0 & 1 & 0\\
%0 & 0 & 0 & 0 & 0 & 1\\
%n^2 -\frac{\mu}{r_a^3} + \frac{3\mu(x+a)^2}{r_a^5} &
%\frac{3\mu(x+a)y}{r_a^5} &
%\frac{3\mu(x+a)z}{r_a^5} &
%0 & 2n & 0 \\
%\frac{3\mu y(x+a)}{r_a^5} &
%n^2 -\frac{\mu}{r_a^3} + \frac{3\mu y^2 }{r_a^5} &
%\frac{3\mu yz}{r_a^5} &
%-2n & 0 & 0 \\
%\frac{3\mu z(x+a) }{r_a^5} &
%\frac{3\mu zy }{r_a^5} &
%-\frac{\mu}{r_a^3} + \frac{3\mu z^2 }{r_a^5} &
%0 & 0 & 0
%\end{bmatrix},\label{eqn:Ax}
%\end{align}
%where $r_a  = \sqrt{(x+a)^2+y^2+z^2}$.

%\paragraph{HCW Equations}
%
%When $\|\mathbf{r}\|\ll 1$, we have $r_a=a$, and $\frac{\mu}{r_a} = n^2$. Substituting these into \refeqn{Ax}, we obtain the Hills-Clohessy-Wiltshire (HCW) equations:
%\begin{align}
%A(0_{6\times 1}) = \begin{bmatrix}
%0 & 0 & 0 & 1 & 0 & 0\\
%0 & 0 & 0 & 0 & 1 & 0\\
%0 & 0 & 0 & 0 & 0 & 1\\
%3n^2 &
%0 &
%0 &
%0 & 2n & 0 \\
%0 &
%0 &
%0 &
%-2n & 0 & 0 \\
%0 &
%0 &
%-n^2 &
%0 & 0 & 0
%\end{bmatrix}.\label{eqn:A}
%\end{align}
%
%
%\clearpage\newpage
\subsection{Observability}

We assume that the line-of-sight from the chief to the deputy is measured by a vision-based sensor. In this section, using the Lie-derivatives of the measurement, we determine the observability for the nonlinear two-body model and the observability for the linear HCW model.

\subsubsection{Measurement Equation}

The measurement equation for the line-of-sight is given by
\begin{align}
\mathbf{y}= h(\mathbf{x}) =\frac{\r}{\|\r\|}
= \frac{1}{\sqrt{x^2+y^2+z^2}}\begin{bmatrix} x\\y\\z\end{bmatrix}.
\end{align}
Note that it has the unit-length, i.e. $\y^T\y=1$.

\subsubsection{Observability Test for the Nonlinear Two-Body Model}\label{sec:OT}

\paragraph{Lie derivatives}

To check observability, we find Lie derivatives of the measurement equation. The following identity is repeatedly used:
\begin{align*}
\delta \parenth{\frac{1}{\|\mathbf{r}\|^n}} 
& = -n\frac{ \mathbf{r}^T \delta \mathbf{r}}{\|\mathbf{r}\|^{n+2}},
\end{align*}
for any integer $n$.

The derivative of the measurement equation is given by
\begin{align}
\deriv{h(\mathbf{x})}{\mathbf{x}} = 
\begin{bmatrix}
\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}
&
0_{3\times 3}
\end{bmatrix}=
\frac{1}{\|\r\|}
\begin{bmatrix}
I - \y\y^T & 0_{3\times 3}
\end{bmatrix}.\label{eqn:dh}
\end{align}
The first-order Lie derivative of the output is 
\begin{align*}
\dot\y = L_f h = \deriv{h(\mathbf{x})}{\mathbf{x}} f(\mathbf{x}) = 
\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}\dot{\mathbf{r}}.
\end{align*}
The variation of the first-order Lie derivative can be written as
\begin{align}
\delta (L_f h) & = 
\parenth{-\frac{\r^T \delta \r}{\|\r\|^3}\dot\r
-\frac{\delta\r \r^T +\r\delta\r^T}{\|\r\|^3}\dot \r
+ 3 \frac{\r\r^T \r^T\delta\r}{\|\r\|^5}\dot\r }
+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}\delta\dot{\mathbf{r}} \nonumber\\
& = 
\begin{bmatrix}
-\dfrac{\dot\r\r^T +\r^T\dot\r I +\r\dot\r^T}{\|\r\|^3}
+ 3 \dfrac{\r\r^T\dot\r \r^T}{\|\r\|^5} & 
\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
\end{bmatrix}
\delta\mathbf{x}\nonumber\\
& = 
\begin{bmatrix}
-\frac{1}{\|\r\|^2}\braces{
{\dot\r\y^T +\y^T\dot\r I +\y\dot\r^T}
- 3\y\y^T\dot\r \y^T} & 
\frac{1}{\|\r\|}\braces{I-\y\y^T}
\end{bmatrix}
\delta\mathbf{x}.\label{eqn:dLfh}
\end{align}

From \refeqn{f}, the second-order Lie derivative of the output is given by
\begin{align*}
\ddot\y =L_f^2 h = 
-\dfrac{2\dot\r\r^T\dot\r +\r\dot\r^T\dot\r}{\|\r\|^3}
+ 3 \dfrac{\r\r^T\dot\r \r^T\dot\r}{\|\r\|^5}
+ \parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
\parenth{-2\omega\times \dot\r - \omega\times(\omega\times \r_a) -\dfrac{\mu \r_a}{\|\r_a\|^3}
}.
\end{align*}
The variation of the second-order Lie derivative can be written as
\begin{align}
 \delta(L_f^2 h) %& = 
%-\dfrac{2\dot\r \delta\r^T\dot\r +\delta\r\dot\r^T\dot\r}{\|\r\|^3}
%+3\dfrac{2\dot\r\r^T\dot\r +\r\dot\r^T\dot\r}{\|\r\|^5}\r^T\delta \r
%+ 3 \dfrac{\delta\r\r^T\dot\r \r^T\dot\r
%+\r\delta\r^T\dot\r \r^T\dot\r
%+\r\r^T\dot\r \delta\r^T\dot\r}{\|\r\|^5}
%- 15 \dfrac{\r\r^T\dot\r \r^T\dot\r}{\|\r\|^7}\r^T\delta\r\nonumber\\
%&\quad 
%-\dfrac{2\delta\dot\r\r^T\dot\r +2\dot\r\r^T\delta\dot\r +2\r\dot\r^T\delta\dot\r}{\|\r\|^3}
%+ 3 \dfrac{\r\r^T\delta\dot\r \r^T\dot\r+\r\r^T\dot\r \r^T\delta\dot\r}{\|\r\|^5}\nonumber\\
%&\quad + \braces{-\dfrac{\ddot{\r}\r^T +\r^T\ddot{\r} I +\r \ddot{\r}^T}{\|\r\|^3}
%+ 3 \dfrac{\r\r^T\ddot{\r} \r^T}{\|\r\|^5}}\delta \r \nonumber\\
%&\quad + \parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
%\braces{\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\delta\r  -2[\omega]_\times\delta\dot\r }
%\nonumber\\
%%
& = \braces{
-\dfrac{2\dot\r \dot\r^T  +\dot\r^T\dot\r I}{\|\r\|^3}
+3\dfrac{2\dot\r\r^T\dot\r\r^T +\r\dot\r^T\dot\r\r^T
+\r^T\dot\r \r^T\dot\r I
+2\r \r^T\dot\r \dot\r^T
}{\|\r\|^5}
- 15 \dfrac{\r\r^T\dot\r \r^T\dot\r\r^T}{\|\r\|^7}
}\delta\r\nonumber\\
&\quad + \braces{-\dfrac{\ddot{\r}\r^T +\r^T\ddot{\r} I +\r \ddot{\r}^T}{\|\r\|^3}
+ 3 \dfrac{\r\r^T\ddot{\r} \r^T}{\|\r\|^5}
+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}
}\delta \r \nonumber\\
&\quad +
\braces{-\dfrac{2\r^T\dot\r I +2\dot\r\r^T +2\r\dot\r^T}{\|\r\|^3}
+ 6 \dfrac{\r\r^T\dot\r \r^T}{\|\r\|^5}
- 2\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}[\omega]_\times
}\delta\dot\r.\label{eqn:dLf2h}
\end{align}

\paragraph{Nonlinear Observability Matrix}

Using \refeqn{dh}, \refeqn{dLfh}, and \refeqn{dLf2h}, the first three blocks of rows of the observability matrix can be written as follows:
\begin{align}
\mathcal{O} = 
\begin{bmatrix}
\mathcal{O}_{00} & 0_{3\times 3}\\
\mathcal{O}_{10} & \mathcal{O}_{11}\\
\mathcal{O}_{20} & \mathcal{O}_{21}
\end{bmatrix}
=\begin{bmatrix}
\deriv{\y}{\r} & \deriv{\y}{\dot\r}\\
\deriv{\dot\y}{\r} & \deriv{\dot\y}{\dot\r}\\
\deriv{\ddot\y}{\r} & \deriv{\ddot\y}{\dot\r}
\end{bmatrix}
,\label{eqn:OO}
\end{align}
where
\begin{align}
\mathcal{O}_{00} & = \deriv{\y}{\r} = \frac{1}{\|\r\|}(I - \y\y^T),\label{eqn:O_00}\\
\mathcal{O}_{10} & = \deriv{\dot\y}{\r} = \deriv{}{\r} \parenth{\deriv{\y}{\r}\dot \r}= -\frac{1}{\|\r\|^2}\braces{
{\dot\r\y^T +\y^T\dot\r I +\y\dot\r^T}
- 3\y\y^T\dot\r \y^T},\\
\mathcal{O}_{11} & = \deriv{\dot\y}{\dot\r} = \deriv{}{\dot\r} \parenth{\deriv{\y}{\r}\dot \r}= \deriv{\y}{\r} = \mathcal{O}_{00},\label{eqn:O11}\\
\mathcal{O}_{20} & = \deriv{\ddot\y}{\r} = -\dfrac{2\dot\r \dot\r^T  +\dot\r^T\dot\r I}{\|\r\|^3}
+3\dfrac{2(\r^T\dot\r) \dot\r\r^T +(\dot\r^T\dot\r)\r\r^T
+(\r^T\dot\r)^2 I
+2(\r^T\dot\r)\r  \dot\r^T
}{\|\r\|^5}
- 15 \dfrac{(\r^T\dot\r)^2\r  \r^T}{\|\r\|^7}\nonumber \\
&\quad -\dfrac{\ddot{\r}\r^T +\r^T\ddot{\r} I +\r \ddot{\r}^T}{\|\r\|^3}
+ 3 \dfrac{(\r^T\ddot{\r})\r \r^T}{\|\r\|^5}
+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}},\\
\mathcal{O}_{21} & = \deriv{\ddot\y}{\dot\r} = \deriv{}{\dot\r}\parenth{\deriv{}{\r}\parenth{\deriv{\y}{\r}\dot\r}\dot\r + \deriv{\y}{\r}\ddot\r} = 2 \deriv{}{\r} \parenth{\deriv{\y}{\r}\dot\r} + \deriv{\y}{\r} \deriv{\ddot\r}{\dot\r}
\nonumber\\
& \quad = 2\mathcal{O}_{10} 
- 2 \mathcal{O}_{00}[\omega]_\times.\label{eqn:O_21}
\end{align}
(\textit{These expressions for the observability matrix has been verified by Matlab Symbolic Toolbox.})

\paragraph{Observability Rank Condition}

Now we find sufficient conditions that the observability matrix $\mathcal{O}$ has full rank. After some algebraic analysis, we can show that the relative orbit with direction-only measurements is observable if the following three conditions are satisfied:
\begin{itemize}
\item[(i)] $\r\times\dot\r \neq 0$ or $\r\times\mathbf{o}_1 \neq 0$,
\item[(ii)] $\r\times\mathbf{o}_2 \neq 0$.
\item[(iii)] $(\mathcal{O}_{20}\r + \mathcal{O}_{21}\dot\r)\times \mathcal{O}_{21}\r\neq 0$, or equivalently, $\mathrm{Proj}[\mathbf{o}_2] \times \mathrm{Proj}[\mathbf{o}_3]\neq 0$,
\end{itemize}
where
\begin{align}
\mathbf{o}_1 & = \ddot\r-\deriv{\ddot\r}{\r}\r = -2\omega\times \dot\r - [\omega]_\times^2 a \mathbf{e}_1 -\dfrac{\mu a }{\|\r_a\|^3}\mathbf{e}_1- \dfrac{3\mu\r_a^T\r}{\|\r_a\|^5}\r_a,\\
\mathbf{o}_2 & = \dot\r + [\omega]_\times \r,\\
\mathbf{o}_3 & = \ddot\r-\deriv{\ddot\r}{\r}\r -\deriv{\ddot\r}{\dot\r}\dot\r = \mathbf{o}_1+2\omega\times\dot\r= - [\omega]_\times^2 a \mathbf{e}_1 -\dfrac{\mu a }{\|\r_a\|^3}\mathbf{e}_1- \dfrac{3\mu\r_a^T\r}{\|\r_a\|^5}\r_a.
\end{align}
%
Note that the condition (iii) implies the conditions (ii), as if the condition (ii) is violated, then the condition (iii) is also violated (if $\mathbf{o}_2$ is parallel to $\r$, then the projection of $\mathbf{o}_2$ to the plane normal to $\r$ is zero).

We consider several cases where these conditions are violated. The condition (i) is violated when $\r$ is parallel to either $\dot\r$ or $\mathbf{o}_1$. The vector $\mathbf{o}_2$ represents the the difference between the inertial velocity of the chief and the inertial velocity of the deputy. The condition (ii) is not satisfied if the relative velocity vector in the inertial frame is either zero or parallel to the relative position. The condition (iii) is always violated if the relative motion is co-planar with the orbital plane of the chief, i.e., $z(t)\equiv 0$. This is because if there is no out-of-plane relative motion, then the vectors $\r$, $\mathbf{o}_2$ and $\mathbf{o}_3$ always lie in the $x-y$ plane, and the projections of $\mathbf{o}_2$ and $\mathbf{o}_3$ to the plane normal to $\r$ is always parallel with each other.

However, these are only sufficient conditions for observability derived from the second-order Lie derivatives of the measurement equation. If any of these conditions is not satisfied, we may need to check the higher-order Lie derivatives of the measurement equation to determine observability.

\subsection{Observability Measure}

The observability rank conditions derived in the previous section provide sufficient conditions for observability. But, it does not state how strong or weak the observability is. In this section, we present two quantitative measures of observability. The first one is an analytic expression based on the second order approximated solution of \refeqn{xxdot}, and the second one is a generalization of observability gramian to nonlinear systems.

\subsubsection{Analytic Observability Measure}

The goal of this analysis is to define an observability measure that can be derived analytically. It is based on a series solution obtained by Volterra series. For simplicity, we consider the class of initial conditions that can be represented by
\begin{align*}
\r_0=[x_0,\,0,\,0]^T,\quad \dot\r=[0,\, \dot y_0,\, 0]^T.
\end{align*}
And, we only consider 2-dimensional relative orbits confined to the orbital plane of the chief.

\paragraph{Analytic Solution based on Volterra Series}
The solution based on the linearized HCW equations are given by
\begin{align}
x_{CW}(t;x_0,\dot y_0)&=(4-3\cos nt )x_0+\frac{2}{n}(1-\cos nt) \dot y_0,\\
y_{CW}(t;x_0,\dot y_0)&=6(\sin nt -nt) x_0+\frac{1}{n}(4\sin nt -3nt)  \dot y_0,
\end{align}
where $x_{CW}(t;x_0,\dot y_0)$ denotes the value of $x(t)$ when the initial condition is specified by $x(0)=x_0$, $\dot y(0)=\dot y_0$ for the HCW dyanmics.

A second-order solution is obtained by using Volterra series as
\begin{align}
x(t;x_0,\dot y_0)&=x_{CW}(t)    +\frac{3}{2R_0}(7-10\cos nt +3\cos2nt +12nt \sin nt -12n^2t^2)x_0^2 \nonumber\\
&\quad +\frac{1}{2n^2 R_0}(6-10\cos nt +4\cos2nt +12nt \sin nt -9n^2t^2)\dot y_0^2 \nonumber\\
&\quad +\frac{3}{nR_0}(4-6\cos nt +2\cos2nt +7nt \sin nt -6n^2t^2)x_0\dot y_0,\\
y(t;x_0,\dot y_0)&=y_{CW}(t)    +\frac{3}{4R_0}(40\sin nt +3\sin 2nt -22nt-24nt\cos nt )x_0^2\nonumber\\
&\quad    +\frac{1}{n^2R_0}(10\sin nt +\sin 2nt -6nt-6nt\cos nt )\dot y_0^2\nonumber\\
&\quad    +\frac{3}{nR_0}(12\sin nt +\sin 2nt -7nt-7nt\cos nt )x_0\dot y_0.
\end{align}

\paragraph{Measurement}

The line of sight measurement can be written as
\begin{align}
\y(t;x_0,\dot y_0) =\begin{bmatrix} y_1(t;x_0,\dot y_0) \\ y_2(t;x_0,\dot y_0) \end{bmatrix} = \frac{1}{\sqrt{x^2(t;x_0,\dot y_0) + y^2(t;x_0,\dot y_0)}}\begin{bmatrix} x(t;x_0,\dot y_0)\\y(t;x_0,\dot y_0)\end{bmatrix}.
\end{align}

\paragraph{Observability Measure}

We first introduce a very simple form of observability measure. Consider the sensitivity of each element of the measurement with respect to the magnitude of the initial condition. More explicitly,
\begin{align*}
z_i (t; x_0,\dot y_0) = \deriv{}{\alpha}\bigg|_{\alpha=1} y_i(t; \alpha x_0,\alpha \dot y_0) =\deriv{y_i(t;x_0,\dot y_0)}{x_0}x_0 + \deriv{y_i(t;x_0,\dot y_0)}{\dot y_0}\dot y_0,
\end{align*}
which is essentially the directional derivative of $y_i$, and it represents how much $y_i$ is sensitive to the variation of the magnitude of the initial condition.

The state will be more observable, if $z_1$ and $z_2$ are more linearly independent. A simplest way to check linear independence of a set of functions is formulating their Wronskian:
\begin{align*}
W(t;x_0,\dot y_0) = \begin{bmatrix} z_1(t;x_0,\dot y_0) & z_2(t;x_0,\dot y_0)\\
\dot z_1(t;x_0,\dot y_0) & \dot z_2(t;x_0,\dot y_0)
\end{bmatrix}.
\end{align*}
If $\mathrm{det}(W(t';x_0,\dot y_0))\neq 0$ for some time $t=t'$, then $z_1$ and $z_2$ are linearly independent at any time interval containing $t'$. If it is zero over some interval of time, then they are linearly dependent.

We choose its absolute value, namely $|\mathrm{det}(W(t';x_0,\dot y_0))|$ as a measure of observability. It is straightforward to show that $W_{CW}(t;x_0,\dot y_0)=0$ for the linearized HCW dynamics.

For the second order solution, the determinant is computed via Matlab symbolic computational tool, and it is evaluated for several values of $t$. For example, when $t=\frac{2\pi}{n}$, we have
\begin{align}
\mathrm{det}(W(\frac{2\pi}{n};x_0,\dot y_0)) = \frac{C_0 C_1^2 C_2}{C_3^3},
\end{align}
where
{\small
\begin{verbatim}
C_0 = -9*pi^2*R^2*n^3,
C_1 = 23*n^3*x0^3 + 36*pi^2*y_dot0^3 + 8*n*x0*y_dot0^2 + 28*n^2*x0^2*y_dot0 + 288*pi^2*n^3*x0^3 
      + 432*pi^2*n^2*x0^2*y_dot0 + 216*pi^2*n*x0*y_dot0^2,
C_2 = - R^2*n^3*x0*y_dot0 + 432*pi^2*R*n^4*x0^3 + 648*pi^2*R*n^3*x0^2*y_dot0 
      + 324*pi^2*R*n^2*x0*y_dot0^2 + 54*pi^2*R*n*y_dot0^3 + 2484*pi^2*n^4*x0^4 
      + 5094*pi^2*n^3*x0^3*y_dot0 + 3798*pi^2*n^2*x0^2*y_dot0^2 + 1224*pi^2*n*x0*y_dot0^3 
      + 144*pi^2*y_dot0^4,
C_3 = 576*pi^2*y_dot0^4 + 324*pi^4*y_dot0^4 + 4761*pi^2*n^4*x0^4 + 5184*pi^4*n^4*x0^4 
      + R^2*n^4*x0^2 + 1512*pi^2*R*n^4*x0^3 + 11592*pi^2*n^3*x0^3*y_dot0 
      + 10368*pi^4*n^3*x0^3*y_dot0 + 144*pi^2*R^2*n^4*x0^2 + 36*pi^2*R^2*n^2*y_dot0^2 
      + 10368*pi^2*n^2*x0^2*y_dot0^2 + 7776*pi^4*n^2*x0^2*y_dot0^2 + 288*pi^2*R*n*y_dot0^3 
      + 4032*pi^2*n*x0*y_dot0^3 + 2592*pi^4*n*x0*y_dot0^3 + 1548*pi^2*R*n^2*x0*y_dot0^2 
      + 2700*pi^2*R*n^3*x0^2*y_dot0 + 144*pi^2*R^2*n^3*x0*y_dot0.
\end{verbatim}}

We introduced a simple form of observability measure. But, the above expression is still too complicated to be analyzed to understand the physical nature of observability.

Instead, we assume that the chief is on a circular orbit with the altitude of $500\,\mathrm{km}$, and we vary the semi-major axis and the eccentricity of the deputy. The observability measure is computed for several values of $t$, and they are illustrated at the following figures. Note that the range of $e_{deputy}$ and $a_{deputy}$ variations is quite limited, as the second order solution is valid for small values of $e_{deputy}$ and $a_{deputy}$ variations.

It turned out that the observability measure increases as $e_{deputy}$ increases. For a fixed value of $e_{deputy}$, the semi-major axis $a_{deputy}$ should be chosen slightly less than $a_{deputy}$ to maximize the observability measure, i.e., the deputy is below the chief and it is orbiting faster.

\clearpage\newpage

\begin{figure}[h!]
\centerline{
	\subfigure[$|\mathrm{det}W(t;x_0,\dot y_0))|$]{
		\includegraphics[width=0.48\textwidth]{July23/detW_ae_1.png}}
	\hfill
	\subfigure[$\log_{10}|\mathrm{det}W(t;x_0,\dot y_0))|$]{
	\includegraphics[width=0.48\textwidth]{July23/detW_ae_log_1.png}}
}
\caption{Observability Measure: $t=0.5T$}
\end{figure}
\begin{figure}[h!]
\centerline{
	\subfigure[$|\mathrm{det}W(t;x_0,\dot y_0))|$]{
		\includegraphics[width=0.48\textwidth]{July23/detW_ae_2.png}}
	\hfill
	\subfigure[$\log_{10}|\mathrm{det}W(t;x_0,\dot y_0))|$]{
	\includegraphics[width=0.48\textwidth]{July23/detW_ae_log_2.png}}
}
\caption{Observability Measure: $t=T$}
\end{figure}

\subsubsection{Observability Gramian}\label{sec:OM}

\paragraph{Linear Systems} The above observability rank condition provides sufficient conditions for observability, and it does not state how difficult to observe each mode of the state vector. For linear dynamic systems, observability Gramian is commonly used as a measure of observability:
\begin{align}
\mathcal{W}(t_0,t_f) = \int_{t_0}^{t_f}\Phi(\tau,t_0)^T C^T(\tau)  C(\tau) \Phi(\tau,t_0) d\tau,\label{eqn:Wo_Lin}
\end{align}
where $\Phi(t,\tau)$ is the state transition matrix. Since the initial state can be computed by the inverse of $W_o(t_0,t_f)$ as 
\begin{align*}
x(t_0) = \mathcal{W}^{-1}(t_0,t_f)\int_{t_0}^{t_f}\Phi(\tau,t_0)^T C^T(\tau)  y(\tau) d\tau,
\end{align*}
the condition number of $W_o(t_0,t_f)$ is often used as a quantitive measure of observability: if the condition number is high, computing the inverse of $W_o$ becomes more sensitive to noise, and it is difficult to determine $x(t_0)$.

\paragraph{Nonlinear Systems} The measure of observability has been generalized to nonlinear dynamic systems, in order to obtain balanced realizations. In~\cite{SchSCL93,Sch94}, an energy-like observability function is introduced to measure the degree of contribution of an initial state to the output, and it is applied to a pendulum system~\cite{NewKriPICDC98}. But, this approach is only applicable to asymptotically stable equilibrium states. An empirical observability Gramian is introduced in~\cite{LalMarPIWC99,LalMarIJRNC02}, which is essentially a covariance matrix of the output computed by a number of sample trajectories. It has been applied to balanced realization of chemical processes in~\cite{HahEdgJPC03,HahEdgCCE02}. However, this approach is based on the assumption that the output converges to a steady-state value, and it is not clear how  sample trajectories are selected.

A similar observability Gramian is introduced in~\cite{KreIdePICDC09}, which is basically the Gramian for the sensitivity of the output with respect to the initial condition:
\begin{align}
\mathcal{W} (\x_0,t_0,t_f) = \int_{t_0}^{t_f} \parenth{\deriv{\y(\tau)}{\x_0}}^T\deriv{\y(\tau)}{\x_0}\,d\tau,\label{eqn:Wo_NL}
\end{align}
where $\y(t)$ denotes the output for the initial condition given by $\x(t_0)=\x_0$, and $\deriv{\y}{\x_0}\in\Re^{p\times n}$ is defined such that its $i,j$-th element is $\deriv{\y_i}{\x_{0_j}}$.

In general, the Gramian of a set of time-dependent functions represents the degree of linear independence of those functions over a given period of time. Therefore, the observability Gramian defined at \refeqn{Wo_NL} measures how much the sensitivities of the output with respect to the initial condition are linearly independent with each other. This represents the degree of observability, since for example, if $\deriv{\y}{\x_{0_i}}$ is linearly dependent to $\deriv{\y}{\x_{0_j}}$, then it is difficult to distinguish the effects of $\delta \x_{0_i}$ to the output $\y$ from the effects of $\delta \x_{0_j}$ to $\y$. Therefore, the condition number of $W_o (\x_0,t_0,t_f)$ can be used as a measure of observability. Futhermore, we can easily show that \refeqn{Wo_NL} reduces to \refeqn{Wo_Lin} for linear dynamic systems since the output is given by $\y(t) = \Phi(t,t_0)\x_0$.

A numerical approach is proposed as follows. The $i,j$--th element of $\mathcal{W}_o$ can be approximated by
\begin{align}
[\mathcal{W}]_{ij} = \frac{1}{4\epsilon^2}\sum_{k=0}^N (\y^{i+}_k-\y^{i-}_k)(\y^{j+}_k-\y^{j-}_k) \Delta t_k,
\end{align}
where $\y^{i\pm}_k$ denotes the value of $\y^i$ at $t=t_k$ with the initial condition of $\x(t_0)=\x_0\pm \epsilon \mathbf{e}_i $ for a positive constant $\epsilon$, and the standard basis $\mathbf{e}_i$ of $\Re^n$.  The above expression can be computed numerically.

\subsection{Extended Kalman Filter}

Extended Kalman filter is developed for the relative orbital dynamics. The observability measures are computed for several cases.

\paragraph{Initial Condition}

The initial condition of a deputy satellite is chosen as
\begin{align*}
\mathbf{r}_0=[r_{x_0},\, 0,\, 0]^T\,\mathrm{km},\quad \mathbf{v}_0=[0,\, -2r_{x_0}n,\, 0]^T\,\mathrm{km/s},
\end{align*}
which yields an elliptic orbit around the chief for the HCW equations.

\subsubsection{Observability Measure}

The initial range $r_{x_0}$ is varied as $1,\,2,\,5,\,10,\,20,\,50,\,100\,\mathrm{km}$, and the observability matrix $\mathcal{O}$ and the nonlinear observability Gramian $\mathcal{W}$ are computed for 20 orbits of the chief. The logarithmic mean of the minimum singular value and the condition number of $\mathcal{O}$ and $\mathcal{W}$ are summarized as follows.

\definecolor{Gray}{gray}{0.85}

\begin{center}
\begin{tabularx}{0.56\textwidth}{*{4}{>{$}c<{$}}{>{\cellcolor{Gray}$}c<{$}}}\toprule
r_{x_0} & \text{min} \braces{\sigma(\mathcal{O})} & \mathrm{cond}\braces{\mathcal{O}} &
\text{min} \braces{\sigma(\mathcal{W})} & \mathrm{cond}\braces{\mathcal{W}}\\\midrule
    1 &   10^{-16.1780} & 10^{16.0019} & 10^{-2.9347} & 10^{16.6653}\\
    2 &   10^{-16.4667} & 10^{15.9894} & 10^{-3.7041} & 10^{16.8323}\\
    5 &   10^{-16.8806} & 10^{16.0050} & 10^{-4.0220} & 10^{16.3535}\\
   10 &   10^{-17.1783} & 10^{16.0009} & 10^{-4.0906} & 10^{15.8194}\\
   20 &   10^{-17.4863} & 10^{16.0066} & 10^{-4.1552} & 10^{15.2832}\\
   50 &   10^{-17.8846} & 10^{16.0034} & 10^{-4.0462} & 10^{14.4052}\\
  100 &   10^{-18.2198} & 10^{16.0106} & 10^{-2.7198} & 10^{12.9884}\\\bottomrule
\end{tabularx}
\end{center}

%    1.0000  -16.1780   16.0019   -2.9347   16.6653
%    2.0000  -16.4667   15.9894   -3.7041   16.8323
%    5.0000  -16.8806   16.0050   -4.0220   16.3535
%   10.0000  -17.1783   16.0009   -4.0906   15.8194
%   20.0000  -17.4863   16.0066   -4.1552   15.2832
%   50.0000  -17.8846   16.0034   -4.0462   14.4052
%  100.0000  -18.2198   16.0106   -2.7198   12.9884

The minimum singular value of $\mathcal{O}$ decreases for increasing $r_{x_0}$. However, since they are close to, or smaller than the machine epsilon ($10^{-16}$), it is difficult to draw meaningful results from them. There is no significant change in the condition number of $\mathcal{O}$.

There is no consistent variation of the minimum singular value of $\mathcal{W}$. But, it is interesting to observe that the condition number of $\mathcal{W}$ becomes smaller as $r_{x_0}$ increases at the last column. This is consistent with the proposition that it is relatively easier to estimate the state as the effects of nonlinear dynamics become stronger.

\subsubsection{Filtering Results}

\paragraph{Initial Estimate}

It is assumed that the initial estimate is a constant multiple of the true state:
\begin{align*}
\hat \r_0=\alpha\r_0,\quad \hat{\mathbf{v}}_0=\alpha{\mathbf{v}}_0,
\end{align*}
where the scaling factor $\alpha$ is varied as $\alpha=1.5,\,2$, and $3$.

\paragraph{EKF}

It is simulated for 10 orbits of the chief. To compare the filtering performance, we define the following error variables, for the direction and the magnitudes. For varying $r_{x_0}$, the error variables are summarized as follows:
\begin{table}[h]
\begin{center}
\caption{Estimation errors for varying $r_{x_0}$ when $\alpha=1.5$}
\begin{tabularx}{0.65\textwidth}{>{\centering $}X<{$}*{4}{>{$}c<{$}}}\toprule
r_{x_0} & e_{dir} \;(\mathrm{rad}) & \cellcolor{Gray}{e_{mag}} \;(\mathrm{unitless}) 
& e_{mag,r} \;(\mathrm{km}) & e_{mag,v} \;(\mathrm{km/s})\\\midrule
    1 &   0.0031  & \cellcolor{Gray}{11.5962} &  18.3371  &  0.0203\\
    2 &   0.0028  & \cellcolor{Gray}{9.6374} &  30.4988  &  0.0338\\
    5 &   0.0030  & \cellcolor{Gray}{4.0041} &  31.7018  &  0.0351\\
   10 &   0.0027  & \cellcolor{Gray}{2.6318} &  41.8535  &  0.0461\\
   20 &   0.0027  & \cellcolor{Gray}{0.9250} &  29.7000  &  0.0325\\
   50 &   0.0023  & \cellcolor{Gray}{0.5315} &  42.3871  &  0.0470\\
  100 &   0.0021  & \cellcolor{Gray}{0.1675} &  26.4643  &  0.0298\\\bottomrule
\end{tabularx}
\end{center}
\end{table}

\noindent There is no meaningful difference in $e_{dir}$, as the direction is directly measured. However, the normalized magnitude error $e_{mag}$ decreases as $r_{x_0}$ increases. This is consistent with the condition number of the nonlinear observability gramian, and it illustrates that the proposed observability measure can be used as a quantitative measure for strength of observability. Filtering results for the case of $r_{x_0}=100\,\mathrm{km}$ is illustrated at Figure 3.

\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{June24/EKF_PE_r0_100_e_s.pdf}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{June24/EKF_PE_r0_100_xx_norm_s.pdf}}
}
\caption{Extended Kalman Filter $r_{x_0}=100\,\mathrm{km}$ (red:true, blue:$\alpha=1.5$, green:$\alpha=2$, black:$\alpha=3$)}
\end{figure}

\bibliography{/Users/tylee@seas.gwu.edu/Documents/BibMaster}
\bibliographystyle{IEEEtran}

\end{document}

### 2018-06-09 Conference/AAS 2015 OD EKF/ISSFD14/ISSFD14.tex
added Conference folder

--- 2018-06-09 Conference/AAS 2015 OD EKF/AAS15.tex
@@
-% use paper, or submit
-% use 11 pt (preferred), 12 pt, or 10 pt only
+\documentclass[12pt,letterpaper]{ISSFD_v01}
+\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
+\usepackage{amssymb,amsmath,amsthm,times,graphicx,subfigure,tabularx,booktabs,colortbl}
 
-%\documentclass[letterpaper, preprint, paper,11pt]{AAS}	% for preprint proceedings
-%\documentclass[letterpaper, paper,11pt]{AAS}		% for final proceedings (20-page limit)
-%\documentclass[letterpaper, paper,12pt]{AAS}		% for final proceedings (20-page limit)
-\documentclass[letterpaper, paper,10pt]{AAS}		% for final proceedings (20-page limit)
-%\documentclass[letterpaper, submit]{AAS}			% to submit to JAS
+\usepackage{times} % use times font
+\usepackage{mathptmx} %use same font for math mode
+\DeclareSymbolFont{largesymbols}{OMX}{cmex}{m}{n} %undo the definition of large symbols from the mathptmx package
+\usepackage{authblk} % use authblk package for author formatting
+\renewcommand\Authfont{\bfseries} % make authors bold
+\renewcommand\Affilfont{\normalfont\itshape} % make affiliation italic
+\setlength{\affilsep}{0em} % no spacing between authors and affiliations
+\usepackage[pdffitwindow=false,pdfstartview={FitH},colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}
+\usepackage{microtype} % makes detailed spacing look nicer.  This is only optional -- not required.
 
-\usepackage{bm}
-\usepackage{amssymb,amsmath,amsthm,times,graphicx,subfigure,tabularx,booktabs,colortbl,multirow,threeparttable}
-\usepackage{subfigure}
-%\usepackage[notref,notcite]{showkeys}  % use this to temporarily show labels
-\usepackage[colorlinks=true, pdfstartview=FitV, linkcolor=black, citecolor= black, urlcolor= black]{hyperref}
-\usepackage{overcite}
-\usepackage{footnpag}			      	% make footnote symbols restart on each page
-
-\usepackage{color}
-\newcommand{\hilight}[1]{\colorbox{green}{#1}}
-
-\newtheorem{prop}{Proposition}
-
-\newcommand{\abs}{\mathop{\mathrm{abs}}\nolimits}
-\newcommand{\diag}{\mathop{\mathrm{diag}}\nolimits}
+\usepackage{xspace}
+\newcommand{\twentyfourth}{24\textsuperscript{th}\xspace}
+\newcommand{\figText}{Fig.\ }
+\newcommand{\tabText}{Tab.\ }
+\newcommand{\eqText}{Eq.\ }
 \newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
@@
 
-\newcommand{\x}{\ensuremath{\mathbf{x}}}
 \renewcommand{\r}{\mathbf{r}}
@@
 \newcommand{\y}{\mathbf{y}}
+\newcommand{\x}{\mathbf{x}}
 
 \newcommand{\bfi}{\bfseries\itshape\selectfont}
+\newtheorem{prop}{Proposition}
 
-%\renewcommand{\baselinestretch}{1.2}
+\title{NONLINEAR OBSERVABILITY FOR RELATIVE SATELLITE ORBITS
+WITH ANGLES-ONLY MEASUREMENTS}
 
-\PaperNumber{23-451}
+\author[(1)]{T. Alan Lovell}
+\author[(2)]{Taeyoung Lee}
+\affil[(1)]{Air Force Research Laboratory, 3550 Aberdeen Ave SE, Kirtland AFB, NM 87117 USA,\newline Tel: +1-505-853-4132, Email: splash73@yahoo.com}
+\affil[(2)]{Department of Mechanical and Aerospace Engineering, The George Washington University,\newline
+801 22nd St NW, Washington DC 20052, Tel: +1-202-994-8710, Email:
+tylee@gwu.edu}
+
+\date{}  % don't display a date
 
 \begin{document}
-
-\title{Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements}
-
-\author{Evan Kaufman\thanks{Doctoral Student, Mechanical and Aerospace Engineering, George Washington University, 801 22nd St NW, Washington, DC 20052, Email: evankaufman@gwu.edu.},  
-T. Alan Lovell\thanks{Research Aerospace Engineer, Air Force Research Laboratory, Space Vehicles Directorate, Kirtland AFB, NM.},
-\ and Taeyoung Lee\thanks{Assistant Professor, Mechanical and Aerospace Engineering, George Washington University, 801 22nd St NW, Washington, DC 20052, Tel: 202-994-8710, Email: tylee@gwu.edu.}
-\thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}
-}
-
-\maketitle{}
+\maketitle
 
 \begin{abstract}
-%An angles-only extended Kalman filter (EKF), based off of two-body motion, is developed to estimate the position of a Deputy satellite relative to a chief satellite.
-A new nonlinear observability measure is proposed for relative orbit determination when lines-of-sight between satellites are measured only.
-It corresponds to a generalization of the observability Gramian in linear dynamic systems to the nonlinear relative orbit dynamics represented by the two-body problems.
-An extended Kalman filter (EKF) is adapted to this problem and is evaluated with various gravitational harmonics and initial orbital determination (IOD) predictions.
-Extensive results illustrate correspondence between the proposed observability measure with filtering errors.
-An extensive numerical analysis in realistic scenarios includes satellite propagation of the two-body problem the $J_2$ perturbation effects.
+In this paper, nonlinear observability criteria are presented for the relative orbital dynamics represented by the solutions of the two-body problem.
+It is assumed that a chief is on a circular orbit with a prescribed orbital radius, and it measures line-of-sight toward a deputy.
+A differential geometric method, based on the Lie derivatives is used to derive sufficient conditions for observability of the orbital properties of deputy.
+It is shown that under certain geometric conditions on the relative configuration between the chief and the deputy, the nonlinear relative motion is observable from angles-only measurements.
+An extended Kalman filter is also developed to numerically illustrate the observability of nonlinear relative orbits with angles-only measurements.
 \end{abstract}
 
+\begin{keywords}
+Relative orbit, Observability, Line-of-sight measurement, Extended Kalman filter.
+\end{keywords}
+
 \section{Introduction}
-
 Space-based surveillance or relative navigation is desirable for many spacecraft missions, such as formation control and rendezvous.
@@
 Relative navigation between spacecraft in close-proximity essentially corresponds to space-based orbit determination.
-In particular, vision-based navigation and estimation of relative orbits have received attention recently, since optical sensors have the desirable properties of low cost and minimal maintenance, while providing accurate line-of-sight measurements, or angles from a chief to a deputy.
+In particular, vision-based navigation and estimation of relative orbit have received attention recently, since optical sensors have desirable properties of low cost and minimal maintenance, while providing accurate line-of-sight measurements.%, or angles from a chief to a deputy.
 
@@
 All of these results are based on linear relative orbital dynamics.
-It is straightforward to see that the relative orbit is not observable with angles-only measurements through its linearized dynamics.
-In other words, there are an infinite number of relative orbits that yield the identical line-of-sight measurements, and the orbital distance between a deputy and a chief cannot be determined by angles-only measurements.
-As such, it is required to study the nonlinear relative orbital dynamics to determine observability with angles-only measurements.
+It is straightforward to see that the relative orbit is not observable with angles-only measurements through its linearized dynamics, due the homogeneity property of linear dynamics implying that any solution of linear systems is directly proportional to its initial conditions.
+In other words, there are infinite number of relative orbits that yield the identical line-of-sight measurements, and the orbital distance between a deputy and a chief cannot be determined by angles-only measurements.
+As such, it is required to study the nonlinear relative orbital dynamics to determine observability with angles-only measurement.
 Linear observability analysis is performed numerically for a particular case in~\cite{YimCraPASMM04}.
+However, there have been no comprehensive analytic in nonlinear observability of relative orbits.
 
-Recently, observability criteria have been derived for the nonlinear relative orbital dynamics represented by the solutions of the two-body problem without linearization~\cite{LovLeePISSFD14}.
+In this paper, the authors derive observability criteria for the nonlinear relative orbital dynamics represented by the solutions of the two-body problem without linearization.
 Assuming that a chief is on a circular orbit with a prescribed orbital radius, nonlinear equations of motion for the relative orbital motion of a deputy with respect to the chief are presented.
@@
 It is shown that under certain geometric conditions on the relative configuration between the chief and the deputy, the nonlinear relative motion is observable from angles-only measurements.
+We also develop an extended Kalman filter to illustrate the observability properties numerically.
 
-However, this result does not provide any information of the degree of observability.
-The main contribution of this paper is proposing a new  measure of observability for relative orbit determination with angles-only measurements.
-For a given initial condition of relative orbits, the proposed observability measure determines how much the corresponding relative orbits are easy or difficult to estimate, thereby providing a quantitative measure for relative orbit determination.
+The main contribution of this paper is analytically confirming that the relative orbit is observable via line-of-sight measurements if its nonlinear dynamic characteristics are properly incorporated.
+To the authors' best knowledge, the sufficient conditions for nonlinear observability presented in this paper have not studied in prior publications.
 
-Furthermore, this paper serves as a feasibility study of an extended Kalman filter (EKF) for orbit determination using angles-only measurements in realistic scenarios.
-The filter is subject to measurement uncertainty and perturbation forces not considered with the two-body problem, on which the filter is based.
-%The EKF is simulated with various time steps to show how the filter performs on a wide range of systems with different sample rates.
-Then, the performance of the EKF is compared with the proposed observability measure to show the correspondence.
+This paper is organized as follows.
+The nonlinear dynamics of relative orbit is presented at Section 2, and observability criteria are developed at Section 3.
+These are followed by numerical examples of extended Kalman filtering and conclusions.
 
@@
 
-Consider two satellites orbiting around the Earth, where each satellite is modeled as a rigid body.
+Consider two satellites orbiting around the Earth, and each satellite is modeled as a rigid body.
 Suppose that a \textit{chief} satellite is on a circular orbit with a pre-determined orbital radius of $a\in\Re$.
@@
 
-Nonlinear equations of motion for the relative motion of the deputy with respect to the chief can be derived as follows based on Lagrangian mechanics~\cite{LovLeePISSFD14}.
+Nonlinear equations of motion for the relative motion of the deputy with respect to the chief can be derived as follows based on Lagrangian mechanics.
 
@@
 
-We assume that the line-of-sight from the chief to the deputy is measured by onboard sensing, such as optical sensors.
+We assume that the line-of-sight from the chief to the deputy is measured by an onboard sensor, such as optical sensors.
 The measurement is represented by the unit-vector of the relative position vector, i.e.,
@@
 
-Based on the nonlinear dynamic model presented at the previous section, here we analyze the observability of the relative orbit with line-of-sight measurements.
-This section serves as a summary of~\cite{LovLeePISSFD14}.
+Based on the nonlinear dynamic model presented at the previous section, here we present analyze the observability of the relative orbit with line-of-sight measurements.
 
@@
 Observability of nonlinear systems has been studied in~\cite{HerKreITAC77}, and it is summarized as follows.
-For a given nonlinear dynamic system described with \refeqn{xxdot} and \refeqn{y}, a pair of points $\x_0$ and $\x_1$ are called \textit{indistinguishable} if the outputs of the corresponding solutions starting from each of $\x_0$ and $\x_1$ are identical for a certain time period.
+For a given nonlinear dynamic system \refeqn{xxdot} and \refeqn{y}, a pair of points $\x_0$ and $\x_1$ are called \textit{indistinguishable} if the outputs of the corresponding solutions starting from each of $\x_0$ and $\x_1$ are identical for a certain time period.
 The systems is \textit{locally weakly observable} at $\x_0$, if there exists an open neighborhood $V$ of $\x_0$ such that for every open neighborhood $U$ of $\x_0$ contained in $V$, the only indistinguishable point to $\x_0$ is the point $\x_0$ itself.
 
-The Lie-derivative of the output $h(\x)$ along $f(\x)$ is defined as follows:
+Let the Lie-derivative of the output $h(\x)$ along $f(\x)$ as follows:
 \begin{align*}
@@
 When applied to linear dynamics, this yields the well-known observability rank condition for linear systems.
-Note that when there is more than a single measurement, i.e., $M>1$, the observability rank condition can be satisfied without a need for computing the higher-order Lie derivatives up to the $N-1$-th order.
+Note that when there are more than a single measurement, i.e., $M>1$, the observability rank condition can be satisfied without need for computing the higher-order Lie derivatives up to the $N-1$-th order.
 
@@
 
-After straightforward but tedious algebraic manipulations using the following identity repeatedly,
+After straightforward but tedious algebraic manipulations using the following identify repeatedly,
 \begin{align*}
@@
 \end{align}
-The nonlinear relative orbital dynamics are locally weakly observable at $\x=[\r^T,\dot\r^T]$ if
+The nonlinear relative orbital dynamics is locally weakly observable at $\x=[\r^T,\dot\r^T]$ if
 %\begin{list}{}{\setlength{\leftmargin}{1cm}\setlength{\itemsep}{0mm}\setlength{\parsep}{0mm}\setlength{\topsep}{0mm}\setlength{\parskip}{0mm}\setlength{\labelwidth}{2cm}}
@@
 \end{align*}
-The matrix $\mathcal{O}_{00}$ has a one-dimensional null space spanned by $\y$.
+The matrix $\mathcal{O}_{00}$ has one-dimensional null space spanned by $\y$.
 Therefore, without loss of generality, we can choose $\mathbf{c}_1=0$ or $\mathbf{c}_1=\r$.
@@
 The matrix $(I-\y\y^T)$ represents the orthogonal projection of a vector into the plane normal to $\y$.
-From the second condition of \refeqn{cond1}, namely $\r\times\mathbf{a}_1\neq 0$, we have $(I-\y\y^T)\mathbf{a}_1\neq 0$, which implies that the constant $c$ cannot be simply chosen as $c=0$.
+From the second condition of \refeqn{cond1}, namely $\r\times\mathbf{a}_1\neq 0$, we have $(I-\y\y^T)\mathbf{a}_1\neq 0$, which implies that the constant $c$ cannot be simply chose as $c=0$.
 Therefore, the only possible case to satisfy the above equation is when $\mathbf{a}_1 +2c \mathbf{v}_{rel}$ is parallel to $\r$ for some values of $c$.
@@
 \begin{align}
-\mathcal{O}_{21}\mathbf{c}_2 & = 2c(\mathcal{O}_{10}-\mathcal{O}_{00}[\omega]_\times) \r \nonumber
-\\
+\mathcal{O}_{21}\mathbf{c}_2 & = 2c(\mathcal{O}_{10}-\mathcal{O}_{00}[\omega]_\times) \r \\
 & = -2c\mathcal{O}_{00}\mathbf{v}_{rel}=0,\label{eqn:row3i1}
@@
 At \refeqn{vrel}, $\mathbf{v}_{rel}$ corresponds to the relative velocity observed in the inertial frame.
-Therefore, the given sufficient conditions for observability are violated when the relative velocity vector is parallel to the relative position vector in the inertial frame.
+Therefore, the given sufficient conditions for observability is violated when the relative velocity vector is parallel to the relative position vector in the inertial frame.
 
@@
 The third condition of \refeqn{cond1} implies that three vectors $\r$, $\mathbf{v}_{rel}$ and $\mathbf{a}_1$ do not belong to the same plane.
-The third condition of \refeqn{cond2} has a similar structure as well.
+The third condition of \refeqn{cond2} has similar structure as well.
 This condition can be easily violated if the relative motion is planar, i.e., when $z(t)\equiv 0$ for all $t$.
@@
 In such case, the third or higher order Lie derivatives should be checked to determine observability.
-The main contribution of this section of the paper is showing that under certain geometric conditions, the nonlinear relative orbital dynamics are indeed observable via angles-only measurements.
+The main contribution of this paper is showing that under certain geometric conditions, the nonlinear relative orbital dynamics are indeed observable via angles-only measurements.
 
-\section{Observability Measure of Nonlinear Systems}
+\section{Numerical Examples}
 
-The measure of observability has been generalized to nonlinear dynamic systems in order to obtain balanced realizations.
-An energy-like observability function is introduced~\cite{SchSCL93,Sch94,NewKriPICDC98} to measure the degree of contribution of an initial state to the output, and it is applied to a pendulum system.
-But, this approach is only applicable to asymptotically stable equilibrium states.
-An empirical observability Gramian is introduced~\cite{LalMarPIWC99,LalMarIJRNC02}, which is essentially a covariance matrix of the output computed by a number of sample trajectories.
-It has been applied to a balanced realization of chemical processes~\cite{HahEdgJPC03,HahEdgCCE02}.
-However, this approach is based on the assumption that the output converges to a steady-state value, and it is not clear how  sample trajectories are selected.
+We illustrate the observability of nonlinear relative orbit by several numerical examples.
+A chief is assumed to be on a circular equatorial orbit with an orbital altitude of $500\,\mathrm{km}$.
+The orbit of a deputy is specified by orbital elements as follows.
+The eccentricity of the deputy is fixed at $e_{deputy}=0.2$, and the semi-major axis of the deputy is chosen such that its orbital period is identical to the chief, i.e., $a_{deputy} = (\frac{T_{chief}\sqrt{\mu}}{2\pi})^{2/3}$.
+Both of longitude of the ascending node and argument of periapsis are chosen to be zero.
+The inclination of the deputy is varied as $i_{deputy}=0^\circ,10^\circ,\ldots 50^\circ$.
+The corresponding orbital trajectories for the case of $i_{deputy}=30^\circ$ is illustrated at Fig.
+\ref{fig:OT}.
 
-A similar observability Gramian is introduced~\cite{KreIdePICDC09}, which is basically the Gramian for the sensitivity of the output with respect to the initial condition:
-\begin{align}
-\mathcal{W} (\x_0,t_0,t_f) = \int_{t_0}^{t_f} \parenth{\deriv{\y(\tau)}{\x_0}}^T\deriv{\y(\tau)}{\x_0}\,d\tau,\label{eqn:Wo_NL}
-\end{align}
-where $\y(t)$ denotes the output for the initial condition given by $\x(t_0)=\x_0$, and $\deriv{\y}{\x_0}\in\Re^{p\times n}$ is defined such that its $i,j$-th element is $\deriv{\y_i}{\x_{0_j}}$.
-
-In general, the Gramian of a set of time-dependent functions represents the degree of linear independence of those functions over a given period of time.
-Therefore, the observability Gramian defined at \refeqn{Wo_NL} measures how much the sensitivities of the output with respect to the initial condition are linearly independent of each other.
-This represents the degree of observability, since for example, if $\deriv{\y}{\x_{0_i}}$ is linearly dependent to $\deriv{\y}{\x_{0_j}}$, then it is difficult to distinguish the effects on $\delta \x_{0_i}$ to the output $\y$ from the effects of $\delta \x_{0_j}$ on $\y$.
-Therefore, the condition number of $\mathcal{W} (\x_0,t_0,t_f)$ can be used as a measure of observability.
-
-%Furthermore, we can easily show that \refeqn{Wo_NL} reduces to the observability Gramian for linear dynamic systems
-%\begin{align}
-%\mathcal{W}(t_0,t_f) = \int_{t_0}^{t_f}\Phi(\tau,t_0)^T C^T(\tau)  C(\tau) \Phi(\tau,t_0) d\tau,\label{eqn:Wo_Lin}
-%\end{align}
-%where $\Phi(t,\tau)$ is the state transition matrix, since the output is given by $\y(t) = \Phi(t,t_0)\x_0$.
-
-A numerical approach is proposed as follows. The $i,j$--th element of $\mathcal{W}_o$ can be approximated by
-\begin{align}
-[\mathcal{W}]_{ij} = \frac{1}{4\epsilon^2}\sum_{k=0}^N (\y^{i+}_k-\y^{i-}_k)(\y^{j+}_k-\y^{j-}_k) \Delta t_k,
-\end{align}
-where $\y^{i\pm}_k$ denotes the value of $\y^i$ at $t=t_k$ with the initial condition of $\x(t_0)=\x_0\pm \epsilon \mathbf{e}_i $ for a positive constant $\epsilon$, and the standard basis $\mathbf{e}_i$ of $\Re^n$.
-The above expression can be computed numerically.
-
-\section{Kalman Filter}
-
-%Preliminary results between the filter performance of an extended Kalman filter and the nonlinear observability Gramian are shown.
-
-We consider the relative orbital dynamics between two satellites, where the observability of the deputy relative to the chief satellite is on a prescribed circular orbit when only $\mathbf{y}$ is assumed to be measured.
-Then, an extended Kalman filter is developed for numerical analyses, where the satellites are either subject to two-body forces or with gravitational $J_2$ perturbations included.
-
-The initial conditions for an estimate of the state vector $\hat \x_0^+$ and its initial covariance matrix $P^+_0$ are assumed to be known from an initial orbit determination (IOD) technique.
-
-\paragraph{Flow Update}
-The flow update provides an a priori state estimate $\hat \x^-_k$ with covariance matrix $P^-_k$ based on the equations of motion.
-The linearized state transition matrix from \refeqn{xxdot} and \refeqn{f} from time step $k-1$ to time step $k$ is written as
-\begin{align}
-F_{k-1}=\exp\left(
-\begin{bmatrix}
-0_{3\times3} & I_{3\times3} \\
-\left(-[\omega]_\times^2-\frac{\mu I_{3\times3}}{\norm{\r_a}^3}+\frac{3\mu \r_a \r_a^T}{\norm{\r_a}^5}\right) & -2[\omega]_\times
-%0 & 0 & 0 & 1 & 0 & 0\\
-%0 & 0 & 0 & 0 & 1 & 0\\
-%0 & 0 & 0 & 0 & 0 & 1\\
-%3*n^2 & 0 & 0 & 0 & 2*n & 0\\
-%0 & 0 & 0 & -2*n & 0 & 0\\
-%0 & 0 & -n^2 & 0 & 0 & 0
-\end{bmatrix}
-(t_k-t_{k-1})\right),
-\end{align}
-such that
-\begin{align}
-\hat \x_{k}^-=F_{k-1}\hat \x_{k-1}^+.
-\end{align}
-The process covariance matrix $Q_k$ is assumed known.
-The a priori covariance matrix is defined as
-\begin{align}
-P^-_k=F_{k-1}P^+_{k-1}F_{k-1}^T+Q_k.
-\end{align}
-
-\paragraph{Measurement Update}
-The effect of the measurement $\y_k$ with covariance matrix $R_k$ is included to determine the posterior state estimate $\hat \x^+_k$ and $P^+_k$.
-The linearized measurement matrix $H_k$ is derived from \refeqn{y},
-\begin{align}
-H_k=
-\begin{bmatrix}
-\left(\frac1{\norm{\hat \r_{k}^-}}I_{3\times3}-\frac{\hat \r_{k}^-{\hat \r_{k}^{-T}}}{\norm{\hat \r_{k}^-}^3}\right) & 0_{3\times3}
-\end{bmatrix},
-\end{align}
-where $\hat \r_{k}^-$ is the a priori estimate of $\r$ given by the first three elements of $\hat \x_{k}^-$.
-Based on \refeqn{y}, the measurement prediction is given by
-\begin{align}
-\hat{\y}_k=h(\hat \r_{k}^-)=\frac{\hat \r_{k}^-}{\norm{\hat \r_{k}^-}}.
-\end{align}
-The posterior state estimate and its covariance matrix are given by
-\begin{align}
-\x^+_k&=\x^-_k+K_k(\y_k-\hat{\y}_k),
-\\
-P^+_k&=(I_{6\times6}-K_kH_k)P^-_k,
-\end{align}
-such that $K_k$ is chosen to minimize $\tr{P^+_k}$ with
-\begin{align}
-K_k=P^-_kH_k^T(H_kP^-_kH_k+R_k)^{-1}.
-\end{align}
-
-%The initial conditions of the deputy satellite are chosen as
-%\begin{align*}
-%\mathbf{r}_0=[r_{x_0},\, 0,\, 0]^T\,\mathrm{km},\quad \mathbf{v}_0=[0,\, -2r_{x_0}n,\, 0]^T\,\mathrm{km/s},
-%\end{align*}
-%for a varying $r_{x_0}$.
-These yield a series of elliptic orbits around the chief when applied to the linearized relative orbital dynamics.
-The state estimation error is represented by a time averaged normalized magnitude error, $e_{mag} = \frac{\|\mathbf{x}-\hat{\mathbf{x}}\|}{\|\mathbf{x}\|}$, where $\mathbf{x}=[\mathbf{r}^T,\dot{\mathbf{r}}^T]^T\in\Re^6$ denotes the true state and $\hat{\mathbf{x}}$ denotes the estimated state.
-%
-%The normalized magnitude error $e_{mag}$ and the condition number of the proposed observability measure $\mathrm{cond}\braces{\mathcal{W}}$ are computed among various values for $r_{x_0}$, tabulated below.
-%
-%\hspace*{0.1cm}\centerline{
-%\begin{tabular}{ccc}\hline
-%$r_{x_0}$ & $e_{mag}$ & $\mathrm{cond}\braces{\mathcal{W}}$\\\hline
-%\vspace*{-0.02\columnwidth}\\
-%$1$ & $11.5962$ &  $10^{16.6653}$\\
-%$2$ & $9.6374$ & $10^{16.8323}$\\
-%$5$ & $4.0041$ & $10^{16.3535}$\\
-%$10$ & $2.6318$ &  $10^{15.8194}$\\
-%$20$ & $0.9250$ &  $10^{15.2832}$\\
-%$50$ & $0.5315$ &  $10^{14.4052}$\\
-%$100$ & $0.1675$ &  $10^{12.9884}$\\
-%\hline
-%\end{tabular}
-%}
-%
-%These preliminary results illustrate that there exists a certain correspondence between the filter error and the proposed observability measure, and therefore, it can be used as a quantitative measure for strength of observability.
-
-\section{Numerical Results}
-
-We choose numerous cases of relative motion between chief and deputy satellite orbits.
-With each case, we evaluate the observability measure and process angles-only measurements with the EKF, using both the two-body equations of motion and higher fidelity propagation with $J_2$ effects as the truth models.
-%In particular, the $J_2$ gravitational harmonics will be included in the satellite propagation, so the EKF must handle simulated process uncertainty.
-To begin each simulation, an initial orbit determination (IOD) estimate is determined by various angles-only observations.
-These IOD estimates serve as initial estimates for the EKF.
-Then, we evaluate the performance of the EKF for these scenarios, specifically focusing on the observability measure and issues arising from initial conditions.
-
-Two IOD methods are developed in~\cite{NewLovPra14,PraLovNew14}, referred to as \emph{IOD1} and \emph{IOD2}, and employed to provide initial conditions for $13$ cases, outlined in Table \ref{tab:IODerr}.
-%Two IOD methods are referred to as \emph{IOD1} and \emph{IOD2}, and employed to provide initial conditions for $13$ cases, outlined in Table \ref{tab:IODerr}.
-These cases yield relative orbit trajectories illustrated in Figures \ref{fig:0CasesTraj}, \ref{fig:1CasesTraj}, \ref{fig:2CasesTraj}, \ref{fig:3CasesTraj}, and \ref{fig:4CasesTraj}.
-The most important distinction between the various cases is that in Case $\mathbf{1}$ the two-body solution and the linearized dynamics of the Hill-Clohessy-Wiltshire (HCW) solution are very close and the trajectory remains largely planar with a circular shape.
-
-\begin{figure}[h]
+\begin{figure}
 \centerline{
-	\subfigure[Case 0A]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_0A_traj.pdf}}
-	\hfill
-	\subfigure[Case 0B]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_0B_traj.pdf}}
-	\hfill
-	\subfigure[Case 0C]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_0C_traj.pdf}}
+\subfigure[Trajectory in the inertial frame (red:chief, blue:deputy)]{
+	\includegraphics[width=0.55\textwidth]{OT_itraj_PE_i_30}}
+\hspace*{0.05\textwidth}
+\subfigure[Relative trajectory in the LVLH frame]{
+	\includegraphics[width=0.35\textwidth]{OT_traj_PE_i_30}}
 }
-\caption{The trajectories of the Cases 0A, 0B, and 0C (solid red: HCW, dashed blue: two-body solution) show that the linear and nonlinear solutions are close.
-The resulting trajectories are not cyclic as the trajectories change in the $y$-direction with time.}
-\label{fig:0CasesTraj}
+\caption{Orbital trajectories when $i_{deputy}=30^\circ$}\label{fig:OT}
 \end{figure}
 
-\begin{figure}[h]
-\label{fig:CaseTraj}
-\centerline{
-	\subfigure[Case 1]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_1_traj.pdf}}
-}
-\caption{The trajectory of the Case 1 (solid red: HCW, dashed blue: two-body solution) shows that the linear and nonlinear solutions are very close, where both solutions are largely planar and cyclic.}
-\label{fig:1CasesTraj}
-\end{figure}
+\subsection{Observability Test}
+For the varying inclinations of the deputy, the minimum singular values and the condition numbers of the observability matrix are summarized as follows.
 
-\begin{figure}[h]
-\centerline{
-	\subfigure[Case 2A]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_2A_traj.pdf}}
-	\hfill
-	\subfigure[Case 2B]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_2B_traj.pdf}}
-	\hfill
-	\subfigure[Case 2C]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_2C_traj.pdf}}
-}
-\caption{The trajectories of the Cases 2A, 2B, and 2C (solid red: HCW, dashed blue: two-body solution) show differences between the linear and nonlinear solutions that vary about the $y$-axis, increasing the observabilities.}
-\label{fig:2CasesTraj}
-\end{figure}
+\begin{table}[h]
+\caption{Singular values and conditions number of observability matrix}\label{tab:SVD}
+\begin{center}
+\begin{tabularx}{0.38\textwidth}{*{3}{>{$}c<{$}}}\toprule
+i_{deputy} & \text{min} \braces{\sigma(\mathcal{O})} & \mathrm{cond}\braces{\mathcal{O}} \\\midrule
+0^\circ & 10^{-19.3041} & 10^{15.9893}  \\
+10^\circ & 10^{-10.3766} & 10^{7.0369}  \\
+20^\circ & 10^{-10.2584} & 10^{6.8603}  \\
+30^\circ & 10^{-10.2290} & 10^{6.7634}  \\
+40^\circ & 10^{-10.2176} & 10^{6.6873}  \\
+50^\circ & 10^{-10.2155} & 10^{6.6273}  \\\bottomrule
+\end{tabularx}
+\end{center}
+\end{table}
 
-\begin{figure}[h]
-\centerline{
-	\subfigure[Case 3A]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_3A_traj.pdf}}
-	\hfill
-	\subfigure[Case 3B]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_3B_traj.pdf}}
-	\hfill
-	\subfigure[Case 3C]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_3C_traj.pdf}}
-}
-\caption{The trajectories of the Cases 3A, 3B, and 3C (solid red: HCW, dashed blue: two-body solution) exhibit large observabilities due to nonlinear solutions that move far away from the linearized solutions.}
-\label{fig:3CasesTraj}
-\end{figure}
+When $i_{deputy}=0^\circ$, the relative motion becomes planar.
+As discussed in the previous section, in this case, the sufficient conditions for observability are violated.
+Therefore, the minimum singular value of the observability matrix $\mathcal{O}$ is close to zero.
+As $i_{deputy}$ is increased, out-of-plane orbital motions are more excited, and as a result, the condition number of $\mathcal{O}$ is increased.
+This implies that the relative motion is easier to estimate when the $i_{deputy}$ becomes large.
 
-\begin{figure}[h]
-\centerline{
-	\subfigure[Case 4A]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_4A_traj.pdf}}
-	\hfill
-	\subfigure[Case 4B]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_4B_traj.pdf}}
-	\hfill
-	\subfigure[Case 4C]{
-		\includegraphics[width=0.3\textwidth]{OM_IOD_4C_traj.pdf}}
-}
-\caption{The trajectories of the Cases 4A, 4B, and 4C (solid red: HCW, dashed blue: two-body solution) experience the largest observabilities due to nonlinear solutions that move farther away from the linearized solutions than any of the trajectories from other cases considered in this paper.}
-\label{fig:4CasesTraj}
-\end{figure}
+\subsection{Extended Kalman Filter}
 
-The initial error functions between the true orbit and and IOD predictions are defined as
-\begin{align}
-e_{mag,0}=\abs\left(1-{\frac{\norm{\r^+_0}}{\norm{\r_0}}}\right), \quad e_{dir,0}=\cos^{-1}\left(\frac{\r_0}{\norm{\r_0}}\cdot \frac{\r^+_0}{\norm{\r^+_0}}\right).
-\end{align}
+To illustrate observability of nonlinear relative orbital dynamics, we develop an extended Kalman filter.
+The estimate of the initial state is chosen as two times of the true state, i.e., $\hat \x_0=2\x_0$.
+This implies that there is a large initial error in the magnitude of the state, which is difficult to estimate accurately using angles-only measurements.
+The initial covariance of the state is chosen as $P_0=\mathrm{diag}[50^2I_{3\times 3}, (50n)^2I_{3\times 3}]$.
+The covariance matrices for the process noise and the measurement noise are $Q_k=\mathrm{diag}[10^{-8}I_{3\times 3}, 10^{-10}I_{3\times 3}]$, and $R_k=1.306^2 I_{3\times 3}\,\mathrm{deg}^2$.
+It is assumed that the line-of-sight is measured at every $\Delta t = 0.47$ seconds.
+It is simulated for ten orbits of the chief around the Earth.
 
+To compare the convergence property of each case, the following measures for estimation errors are introduced.
+\begin{align*}
+e_{dir}  & = \sqrt{\frac{1}{N}\sum_{k=0}^N\parenth{\cos^{-1}\parenth{\frac{\mathbf{x}(t_k)^T \hat{\mathbf{x}} (t_k)}
+{\|\mathbf{x}(t_k)\| \|\hat{\mathbf{x}} (t_k)\|}}}^2},\\
+e_{mag}  & = \sqrt{\frac{1}{N}\sum_{k=0}^N\parenth{
+\frac{\|\mathbf{x}(t_k)\|-\|\hat{\mathbf{x}} (t_k)\|}{\|\mathbf{x}(t_k)\|}}^2},\\
+e_{mag,r}  & = \sqrt{\frac{1}{N}\sum_{k=0}^N\parenth{
+\frac{\|\mathbf{r}(t_k)\|-\|\hat{\mathbf{r}} (t_k)\|}{\|\r(t_k)\|}}^2},\\
+e_{mag,v}  & = \sqrt{\frac{1}{N}\sum_{k=0}^N\parenth{
+\frac{\|\mathbf{v}(t_k)\|-\|\hat{\mathbf{v}} (t_k)\|}{\|\mathbf{v}(t_k)}}^2},
+\end{align*}
+where $N$ is the total number of time steps.
+The error variable $e_{dir}$ represents the mean squared value of the angle between the true state and the estimated state, and the variables $e_{mag},e_{mag,r},e_{mag,v}$ correspond to the normalized magnitude error for the state vector, position vector, and velocity vector, respectively.
+
+These are summarized at Table \ref{tab:EE}.
+For all cases, the direction errors are fairly small, as the line-of-sight is directly measured.
+Overall, the magnitude error decreases as the inclination of the deputy increases.
+This is consistent with the condition number of the observability matrix summarized at Table \ref{tab:SVD}.
+
+\begin{table}[h]
+\caption{Estimation Errors}\label{tab:EE}
 \begin{center}
-\begin{threeparttable}[h]
-\caption{Case Observabilities and Initial Errors}
-\begin{tabularx}{0.72\textwidth}
-{
->{$}c<{$}
-*{1}{>{$}c<{$}} |
-*{2}{>{$}c<{$}} |
-*{2}{>{$}c<{$}}
-}
-\toprule
-\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{\multirow{1}{*}{IOD1}} & \multicolumn{2}{c}{\multirow{1}{*}{IOD2}} \\
-& &  e_{mag,0} & e_{dir,0} & e_{mag,0} & e_{dir,0} \\\midrule
-0A & 10^{11.1486} & 0.0003 & 1.4901\times10^{-8} & 0.0012  & 2.1073\times10^{-8} \\
-0B & 10^{10.9387} & 0.0233 & 0 & 0.0088 & 0 \\
-0C & 10^{10.9851} & 0.0152 & 2.1073\times10^{-8} & 0.0139 & 1.4901\times10^{-8}  \\
-\midrule
-\mathbf{1} & 10^{16.6871} & 0.0001 & 0 & 0 & 0  \\
-\midrule
-2A & 10^{10.6492} & 0.0550 & 1.8150\times10^{-4} & 0.0238 & 1.8150\times10^{-4}  \\
-2B & 10^{10.4163} & 0.9571 & 0.0734                     & 0.9571 & 0.0734  \\
-2C & 10^{10.2155} & 0.0219 & 0.0044 & 0.0065 & 0.0044 \\
-\midrule
-3A & 10^{8.3490} & 0.0191 & 0.0031 & 0.1830 & 0.0031  \\
-3B & 10^{8.5372} & 0.0612 & 0.0533 & 0.2315 & 0.0533  \\
-3C & 10^{8.6586} & 6.8962 & 0.7348 & 5.4921 & 0.7348  \\
-\midrule
-4A & 10^{7.9062} & 8.7347 & 0.4587 & 0.1575 & 0.4587  \\
-4B & 10^{7.9903} & 0.9276 & 0.1819 & 0.0621 & 0.1819  \\
-4C & 10^{8.0645} & 0.1161 & 0.0554 & 0.0354 & 0.0554  \\
-\bottomrule
+\begin{tabularx}{0.75\textwidth}{>{\centering $}X<{$}*{4}{>{$}c<{$}}}\toprule
+i_{deputy} & e_{dir} \;(\mathrm{rad}) & {e_{mag}} \;(\mathrm{unitless}) 
+& e_{mag,r} \;(\mathrm{unitless}) & e_{mag,v} \;(\mathrm{unitless})\\\midrule
+0 & 0.0211  & {0.3361} & 0.3361 & 0.3428\\
+10 & 0.0172  & {0.2204} & 0.2204 & 0.2385\\
+20 & 0.0187  & {0.1681} & 0.1681 & 0.1976\\
+30 & 0.0248  & {0.1596} & 0.1596 & 0.1922\\
+40 & 0.0421  & {0.1703} & 0.1703 & 0.1995\\
+50 & 0.0599  & {0.1841} & 0.1841 & 0.2092\\\bottomrule
 \end{tabularx}
-{\small
-\begin{tablenotes}
-    \item Note: $0$ indicates numbers small enough to be below Matlab numerical accuracy.
-  \end{tablenotes}}
-\label{tab:IODerr}
-\end{threeparttable}
 \end{center}
+\end{table}
 
-The Kalman filter is used in each case. The initial state covariance matrix is chosen as
-\begin{align*}
-P_0^+=\diag[50^2,\ 50^2,\ 50^2,\ (50n)^2,\ (50n)^2,\ (50n)^2],
-\end{align*}
-and the process and measurement covariance matrices are chosen as
-\begin{align*}
-Q_k=\diag[10^{-7},\ 10^{-7},\ 10^{-7},\ 10^{-9},\ 10^{-9},\ 10^{-9}], \quad R_k=\tan^2{2^\circ}I_{3\times3}\ \forall\ k.
-\end{align*}
-%\begin{align*}
-%Q_k=10^q\diag[10^{-8},\ 10^{-8},\ 10^{-8},\ 10^{-10},\ 10^{-10},\ 10^{-10}], \quad R_k=\tan^2{2^\circ}I_{3\times3}\ \forall\ k
-%\end{align*}
-%where $q\in[-1,\ 0,\ 1,\ 2]$ is varied in the simulations, thereby changing the impact of measurement updates inside the Kalman filter.
-
-Each case is simulated for every IOD estimate.
-The running time is $59540$ seconds (slightly more than $\frac23$ of a day) with time steps of $10$ seconds between measurements.
-The following RMS errors are collected from each trial:
-\begin{align}
-\bar e_{mag}=\left(\frac1N\sum_{k=1}^N\left\{1-{\frac{\norm{\r^+_k}}{\norm{\r_k}}}\right\}^2\right)^{\frac12}, \quad \bar e_{dir}=\left(\frac1N\sum_{k=1}^N
-\left\{\cos^{-1}\left(\frac{\r_k}{\norm{\r_k}}\cdot \frac{\r^+_k}{\norm{\r^+_k}}\right)\right\}^{2}
-\right)^{\frac12},
-\end{align}
-for $N$ time steps.
-These metrics are tabulated in Tables \ref{tab:ResultsTBP} and \ref{tab:ResultsJ2}.
-Over the simulated time period, figures were generated for these analyses that  serve to compare the true and estimated state magnitudes and the maximum eigenvalues of the state covariance matrices as a measure of uncertainty for the state estimates.
-Because of paper constraints, a sample of these plots are shown that highlight various aspects of the observability and filter performance.
-
-%The figure below illustrates preliminary results for angles-only relative orbit estimation with $J_2$ perturbation in the measurements.
-%Note that the EKF performs well in this particular case, as evidenced by the close agreement between the true and estimated trajectory (Figure 1a) and the decrease in the maximum eigenvalue of the covariance matrix over time (Figure 1b).
-%For the various cases to be chosen, we will evaluate the performance of the EKF (using metrics such as those illustrated in Figure 1) and seek a consistent correlation between the filter performance and the new observability measure proposed above.
-
-%\begin{center}
-%\begin{threeparttable}[h]
-%\caption{Error variables for EKF with varying $Q_k$ Considering Only Two-Body Problem Forces}
-%\begin{tabularx}{\textwidth}
-%{
-%>{$}c<{$}
-%*{1}{>{$}c<{$}} |
-%*{4}{>{$}c<{$}} |
-%*{4}{>{$}c<{$}}
-%}
-%% & \bar e_{mag} & \bar e_{dir}
-%\toprule
-%\multirow{2}{*}{Case} & \multirow{2}{*}{IOD} & \multicolumn{4}{c}{$\bar e_{mag}$} & \multicolumn{4}{c}{$\bar e_{dir}$} \\
-%& & q=-1 & q=0 & q=1 & q=2 & q=-1 & q=0 & q=1 & q=2 \\\midrule
-%0A & IOD 1 & 0.0895 &  0.0611  &  0.3766  &  0.2861  &  0.0430  &  0.0135  &  0.0144  &  0.0183 \\
-%0A & IOD 2 & 0.1793 &  0.2811  &  0.5145  &  0.7716  &  0.0203  &  0.0313  &  0.0691  &  0.1046 \\
-%0B & IOD 1 & 0.3241 &  0.0531  &  0.5186  &  0.0620  &  0.1695  &  0.0143  &  0.0810  &  0.0172 \\
-%0B & IOD 2 & 0.2330 &  0.2115  &  0.3092  &  0.2733  &  0.1007  &  0.0879  &  0.0804  &  0.1048 \\
-%0C & IOD 1 & 0.0973 &   0.0748 &   0.2723 &   0.0607 &   0.0292 &   0.0068  &  0.1362 &   0.0443 \\
-%0C & IOD 2 & 0.1473 &   0.1017 &   0.1707 &   0.3545  &  0.0702  &  0.0517  &  0.0614  &  0.1224 \\
-%\midrule
-%\mathbf{1} & IOD 1 & \hilight{8.922} &   \hilight{2.964}   & \hilight{4.078}  &  \hilight{3.571} &   0.0071  &  0.0089  &  0.0109  &  0.0418 \\
-%\mathbf{1} & IOD 2 & \hilight{13.635} &  \hilight{14.121} &  \hilight{42.486} &  \hilight{42.573} &   0.0113 &   0.0107  &  0.0445  &  0.0115 \\
-%\midrule
-%2A & IOD 1 & 0.2045 &   0.0629 &   0.1592  &  0.1329  &  0.0041 &   0.0041  &  0.0043  &  0.0039 \\
-%2A & IOD 2 & 0.1013 &   0.1797  &  0.1348  &  0.2247  &  0.0050 &   0.0048 &   0.0065 &   0.0061 \\
-%2B & IOD 1 & \hilight{0.9886} &   \hilight{0.9895}   & \hilight{0.9843}  &  \hilight{0.9851}  &  0.0114 &   0.0116  &  0.0114  &  0.0116 \\
-%2B & IOD 2 & \hilight{0.9557} &   \hilight{0.9685} &   \hilight{0.9278}  &  \hilight{0.9431} &   0.0105  &  0.0108 &   0.0113 &   0.0153 \\
-%2C & IOD 1 & 0.0219 &   0.0184  &  0.0258  &  0.0238 &   0.0026 &   0.0031  &  0.0044 &   0.0038 \\
-%2C & IOD 2 & 0.0303 &   0.0287  &  0.0360  &  0.0470  &  0.0044  &  0.0054  &  0.0059 &   0.0062 \\
-%\midrule
-%3A & IOD 1 &  0.2377 &   0.1684  &  0.0486 &   0.0695 &   0.0094 &   0.0070 &   0.0040 &  0.0039 \\
-%3A & IOD 2 &  0.0854  &  0.0874 &   0.0603  &  0.0544  &  0.0035 &   0.0033  &  0.0053  &  0.0042 \\
-%3B & IOD 1 &  0.0724 &   0.0260 &   0.0680  &  0.0289  &  0.0035 &   0.0029  &  0.0033  &  0.0027 \\
-%3B & IOD 2 &  0.0385  &  0.0432  &  0.1459  &  0.0284  &  0.0038  &  0.0037  &  0.0070  &  0.0047 \\
-%3C & IOD 1 &  \hilight{2.2433}  &  \hilight{1.5231} &   \hilight{2.0611} &   \hilight{1.5110}  &  0.1803  &  0.1172  &  0.1595  &  0.1103 \\
-%3C & IOD 2 &  \hilight{4.0635} &   \hilight{1.7717}  & \hilight{10.567}  &  \hilight{4.0277} &   0.4475 &   0.1506 &   0.9011 &   0.4028 \\
-%\midrule
-%4A & IOD 1 &  \hilight{6.0259} &   \hilight{0.0464}  &  \hilight{6.0706}  &  \hilight{0.0325}  &  1.3353  &  0.0087   & 1.3182  &  0.0065 \\
-%4A & IOD 2 &  6.3646 &   0.0173  & 11.8882 &   0.0256 &   1.3423  &  0.0064  &  1.5319 &   0.0066 \\
-%4B & IOD 1 &  \hilight{0.2922}  &  \hilight{0.0258}  &  \hilight{0.2723}  &  \hilight{0.0182}   & 0.0929 &   0.0032  &  0.0789  &  0.0036 \\
-%4B & IOD 2 &  0.1815  &  0.0217  &  0.3155 &   0.0771  &  0.0268   & 0.0034  &  0.1061 &   0.0044 \\
-%4C & IOD 1 &  0.0247  &  \hilight{0.0421}  &  2.3699  &  0.0182  &  0.0032  &  0.0032  &  0.8486 &   0.0028 \\
-%4C & IOD 2 &  0.0200 &   0.0128  &  0.0278  &  0.0299 &   0.0040  &  0.0029  &  0.0034  &  0.0033 \\
-%\bottomrule
-%\end{tabularx}
-%{\small
-%\begin{tablenotes}
-%    \item \hilight{Filter diverges}
-%  \end{tablenotes}}
-%\label{tab:ResultsTBP}
-%\end{threeparttable}
-%\end{center}
-%
-%\begin{center}
-%\begin{threeparttable}[h]
-%\caption{Error variables for EKF with varying $Q_k$ with $J_2$ Perturbations Included}
-%\begin{tabularx}{\textwidth}
-%{
-%>{$}c<{$}
-%*{1}{>{$}c<{$}} |
-%*{4}{>{$}c<{$}} |
-%*{4}{>{$}c<{$}}
-%}
-%\toprule
-%\multirow{2}{*}{Case} & \multirow{2}{*}{IOD} & \multicolumn{4}{c}{$\bar e_{mag}$} & \multicolumn{4}{c}{$\bar e_{dir}$} \\
-%& & q=-1 & q=0 & q=1 & q=2 & q=-1 & q=0 & q=1 & q=2 \\\midrule
-%0A & IOD 1 & 0.0553  &  0.1105  &  0.1918  &  0.2298  &  0.0059  &  0.0062  &  0.0069  &  0.0069 \\
-%0A & IOD 2 & 0.2177  &  0.3227  &  0.3140  &  0.5013  &  0.0082  &  0.0074  &  0.0129  &  0.0085 \\
-%0B & IOD 1 & 0.0473  &  0.1246  &  0.7262  &  0.0655  &  0.0077  &  0.0088  &  0.0505  &  0.0066\\
-%0B & IOD 2 & 0.8872  &  0.4186  &  0.1571  &  0.1561  &  0.0390  &  0.0168  &  0.0086  &  0.0293 \\
-%0C & IOD 1 & 0.0652  &  0.0809  &  0.0900  &  0.1113  &  0.0082  &  0.0079  &  0.0061  &  0.0106 \\
-%0C & IOD 2 & 0.0367  &  0.0204  &  0.1002  &  0.7253  &  0.0145  &  0.0084  &  0.0154  &  0.0579 \\
-%\midrule
-%\mathbf{1} & IOD 1 & \hilight{9.1098}  &  \hilight{2.900}  &  \hilight{4.059}  &  \hilight{3.567}  &  0.0071  &  0.0090  &  0.0109  &   0.0418 \\
-%\mathbf{1} & IOD 2 & \hilight{13.594} & \hilight{14.033} & \hilight{42.381} & \hilight{42.979}  & 0.0113   & 0.0107   & 0.0445   &   0.0115 \\
-%\midrule
-%2A & IOD 1 & 0.2301  &  0.0423  &  0.1978  &  0.1740  &  0.0048  &  0.0049  &  0.0047  &  0.0042 \\
-%2A & IOD 2 & 0.1058  &  0.2164  &  0.1638  &  0.2644  &  0.0051  &  0.0049  &  0.0065  &  0.0061 \\
-%2B & IOD 1 & \hilight{0.9895}  &  \hilight{0.9902}  &  \hilight{0.9853}  &  \hilight{0.9859}  &  0.0115  &  0.0117  &  0.0115  &  0.0117 \\
-%2B & IOD 2 & \hilight{0.9585}  &  \hilight{0.9701}  &  \hilight{0.9307}  &  \hilight{0.9455}  &  0.0106  &  0.0109  &  0.0114  &  0.0153 \\
-%2C & IOD 1 & 0.0481  &  0.0474  &  0.0563  &  0.0415  &  0.0041  &  0.0042  &  0.0047  &  0.0042 \\
-%2C & IOD 2 & 0.0535  &  0.0397  &  0.0536  &  0.0705  &  0.0044  &  0.0056  &  0.0059  &  0.0062 \\
-%\midrule
-%3A & IOD 1 &  0.2462  &  0.1827  &  0.0516  &  0.0699 &   0.0097 &   0.0077  &  0.0041 &   0.0041 \\
-%3A & IOD 2 &  0.1038  &  0.0986  &  0.0643  &  0.0692 &   0.0036  &  0.0034  &  0.0053  &  0.0041 \\
-%3B & IOD 1 &  0.0770  &  0.0310  &  0.0739  &  0.0458 &   0.0038 &   0.0034  &  0.0035  &  0.0031 \\
-%3B & IOD 2 &  0.0461  &  0.0585   & 0.1485 &   0.0391  &  0.0039  &  0.0040  &  0.0071  &  0.0047 \\
-%3C & IOD 1 &  \hilight{2.239}  &  \hilight{1.522}  &  \hilight{2.063}  &  \hilight{1.514}  &  0.1817  &  0.1179  &  0.1614  &  0.1116 \\
-%3C & IOD 2 &  \hilight{4.188}  &  \hilight{1.821}  & \hilight{10.136}  &  \hilight{4.510} &   0.4873 &   0.1592  &  0.9195 &   0.5532 \\
-%\midrule
-%4A & IOD 1 &  \hilight{6.013}  &  \hilight{0.0454}  &  \hilight{6.359}  &  \hilight{0.031}  &  1.3131  &  0.0086  &  1.4235 &   0.0066 \\
-%4A & IOD 2 &  6.3272  &  0.0197  & 12.0062 &  0.0248  &  1.3116  &  0.0065  &  1.5310  &  0.0066 \\
-%4B & IOD 1 &  \hilight{0.2936}  &  \hilight{0.0272}  &  \hilight{0.2773}  &  \hilight{0.0217}  &  0.0935  &  0.0032  &  0.0814  &  0.0038 \\
-%4B & IOD 2 &  0.1853  &  0.0210  &  0.5244  &  0.0785  &  0.0293  &  0.0036  &  0.2305  &  0.0045 \\
-%4C & IOD 1 &  0.0260  &  \hilight{0.0410}  &  2.0105  &  0.0164  &  0.0034  &  0.0031  &  0.9144  &  0.0030 \\
-%4C & IOD 2 &  0.0210  &  0.0145  &  0.0303  &  0.0349  &  0.0042  &  0.0030  &  0.0036  &  0.0034 \\
-%\bottomrule
-%\end{tabularx}
-%{\small
-%\begin{tablenotes}
-%    \item \hilight{Filter diverges}
-%  \end{tablenotes}}
-%\label{tab:ResultsJ2}
-%\end{threeparttable}
-%\end{center}
-
-\begin{center}
-\begin{threeparttable}[h]
-\caption{Error variables for EKF Considering Only Two-Body Problem Forces}
-\begin{tabularx}{.58\textwidth}
-{
->{$}c<{$}
-*{1}{>{$}c<{$}} |
-*{2}{>{$}c<{$}} |
-*{2}{>{$}c<{$}}
-}
-\toprule
-\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{IOD1} & \multicolumn{2}{c}{IOD2} \\
-& & \bar e_{mag} & \bar e_{dir} & \bar e_{mag} & \bar e_{dir} \\\midrule
-0A & 10^{11.1486} &  0.3766  &  0.0144  &  0.5145  &  0.0691 \\
-0B & 10^{10.9387} &  0.5186  &  0.0810  &  0.3092  &  0.0804\\
-0C & 10^{10.9851} &  0.2723  &  0.1362 &  0.1707 &  0.0614\\
-\midrule
-\mathbf{1} & 10^{16.6871} & \cellcolor{green} 4.078  & \cellcolor{green}  0.0109 & \cellcolor{green} 42.486 & \cellcolor{green}  0.0445 \\
-\midrule
-2A & 10^{10.6492} &  0.1592  &   0.0043  &  0.1348  &   0.0065   \\
-2B & 10^{10.4163} & \cellcolor{green} 0.9843   & \cellcolor{green}  0.0114  & \cellcolor{green}  0.9278 & \cellcolor{green}   0.0113 \\
-2C & 10^{10.2155} & 0.0258 &  0.0044 & 0.0360 &  0.0059 \\
-\midrule
-3A & 10^{8.3490} &  0.0486 & 0.0040   &  0.0603  &  0.0053 \\
-3B & 10^{8.5372} &  0.0680  &  0.0033 &  0.1459 &  0.0070 \\
-3C & 10^{8.6586} & \cellcolor{green}  2.0611 & \cellcolor{green}  0.1595   & \cellcolor{green}  10.567  & \cellcolor{green}   0.9011 \\
-\midrule
-4A & 10^{7.9062} & \cellcolor{green}  6.0706  & \cellcolor{green} 1.3182  &  11.8882   &  1.5319  \\
-4B & 10^{7.9903} & \cellcolor{green}  0.2723  & \cellcolor{green}  0.0789   &  0.3155  &  0.1061 \\
-4C & 10^{8.0645} &  2.3699 &  0.8486   &  0.0278 &  0.0034 \\
-\bottomrule
-\end{tabularx}
-{\small
-\begin{tablenotes}
-    \item \hilight{Filter diverges}
-  \end{tablenotes}}
-\label{tab:ResultsTBP}
-\end{threeparttable}
-\end{center}
-
-\begin{center}
-\begin{threeparttable}[h]
-\caption{Error variables for EKF Considering $J_2$ Perturbations}
-\begin{tabularx}{.58\textwidth}
-{
->{$}c<{$}
-*{1}{>{$}c<{$}} |
-*{2}{>{$}c<{$}} |
-*{2}{>{$}c<{$}}
-}
-\toprule
-\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{IOD1} & \multicolumn{2}{c}{IOD2} \\
-& & \bar e_{mag} & \bar e_{dir} & \bar e_{mag} & \bar e_{dir} \\\midrule
-0A & 10^{11.1486} &  0.1918  &  0.0069  &  0.3140  &  0.0129 \\
-0B & 10^{10.9387} &  0.7262  &  0.0505  &  0.1571  &  0.0086\\
-0C & 10^{10.9851} &  0.0900  &  0.0061 &  0.1002  &  0.0154\\
-\midrule
-\mathbf{1} & 10^{16.6871} & \cellcolor{green} 4.059  & \cellcolor{green}  0.0109 & \cellcolor{green} 42.381 & \cellcolor{green}  0.0445 \\
-\midrule
-2A & 10^{10.6492} &  0.1978  &  0.0047  &  0.1638  &   0.0065   \\
-2B & 10^{10.4163} & \cellcolor{green}  0.9853  & \cellcolor{green}  0.9307 & \cellcolor{green}  0.9278 & \cellcolor{green}   0.0114 \\
-2C & 10^{10.2155} & 0.0563 &  0.0047 & 0.0536 &  0.0059 \\
-\midrule
-3A & 10^{8.3490} &  0.0516 & 0.0041   &  0.0643  &  0.0053 \\
-3B & 10^{8.5372} &  0.0739  &  0.0035 &  0.1485 &  0.0071 \\
-3C & 10^{8.6586} & \cellcolor{green}  2.063 & \cellcolor{green} 0.1614    & \cellcolor{green}  10.136  & \cellcolor{green}   0.9195 \\
-\midrule
-4A & 10^{7.9062} & \cellcolor{green}  6.359  & \cellcolor{green} 1.4235  &  12.0062   &  1.5310  \\
-4B & 10^{7.9903} & \cellcolor{green}  0.2773  & \cellcolor{green}  0.0814   &  0.5244  &  0.2305 \\
-4C & 10^{8.0645} &  2.0105 &  0.9144   &  0.0303 &  0.0036 \\
-\bottomrule
-\end{tabularx}
-{\small
-\begin{tablenotes}
-    \item \hilight{Filter diverges}
-  \end{tablenotes}}
-\label{tab:ResultsJ2}
-\end{threeparttable}
-\end{center}
-
-These data from the simulations show several important aspects and challenges with relative orbit determination involving angles-only measurements.
-First, an important observation from the data is that the IOD guess is critical to the success of the Kalman filter.
-Cases 2B, 3C, 4A, and 4B experience filter divergence when the IOD estimate is largely inaccurate, shown in Tables \ref{tab:IODerr}, \ref{tab:ResultsTBP}, and \ref{tab:ResultsJ2}, even when cases with similar observabilities do not experience filter divergence.
-The Kalman filter, which is based on linear dynamic systems, is modeled as a Markov chain and hence is memoryless.
-Therefore when nothing more than an inaccurate IOD prediction is poorly assumed to contain accurate information about the system, the recursive formulation of the Kalman filter is unable to converge.
-In these diverging cases, a greater level of observability or an algorithm that relies on a larger history of measurements may improve convergence.
-
-In contrast, every variation of Case $\mathbf{1}$ has a closer IOD than the IODs of other cases, but experiences the most severe divergence.
-Case $\mathbf{1}$ has a greater condition number of $\mathcal W$, which corresponds with low observability according to the proposed observability measure.
-A series of angles-only measurements may correspond with a large range of possible relative orbits, even with a low variation in those measurements.
-Therefore, small levels of noise in the sensor or uncertainties in the orbit prevent the Kalman filter from converging to an accurate estimate, even with an accurate IOD.
-The state magnitudes of both variations of Case $\mathbf{1}$ are illustrated against various converging cases in Figure \ref{fig:ExampesEachCase} to show how important this observability is to the filter performance.
-
-\begin{figure}[h]
-\centerline{
-	\subfigure[Case 0C, IOD1]{
-		\includegraphics[width=0.35\textwidth]{Case0Cq1IOD1TBPnormX.pdf}}
-	\hfill
-	\subfigure[Case 0C, IOD2]{
-	\includegraphics[width=0.35\textwidth]{Case0Cq1IOD2TBPnormX.pdf}}
-}
-\centerline{
-	\subfigure[Case $\mathbf{1}$, IOD1]{
-		\includegraphics[width=0.35\textwidth]{Case1q1IOD1TBPnormX.pdf}}
-	\hfill
-	\subfigure[Case $\mathbf{1}$, IOD2]{
-	\includegraphics[width=0.35\textwidth]{Case1q1IOD2TBPnormX.pdf}}
-}
-\centerline{
-	\subfigure[Case 2A, IOD1]{
-		\includegraphics[width=0.35\textwidth]{Case2Aq1IOD1TBPnormX.pdf}}
-	\hfill
-	\subfigure[Case 2A, IOD2]{
-	\includegraphics[width=0.35\textwidth]{Case2Aq1IOD2TBPnormX.pdf}}
-}
-\centerline{
-	\subfigure[Case 3A, IOD1]{
-		\includegraphics[width=0.35\textwidth]{Case3Aq1IOD1TBPnormX.pdf}}
-	\hfill
-	\subfigure[Case 3A, IOD2]{
-	\includegraphics[width=0.35\textwidth]{Case3Aq1IOD2TBPnormX.pdf}}
-}
-\centerline{
-	\subfigure[Case 4C, IOD1]{
-		\includegraphics[width=0.35\textwidth]{Case4Cq1IOD1TBPnormX.pdf}}
-	\hfill
-	\subfigure[Case 4C, IOD2]{
-	\includegraphics[width=0.35\textwidth]{Case4Cq1IOD2TBPnormX.pdf}}
-}
-\caption{The magnitudes of the relative states (solid red: true, blue dashed: estimated) are shown for various two-body propagated cases with close IOD estimates where the filter only diverges when the observability is low.}\label{fig:ExampesEachCase}
-\end{figure}
-
-Another observation from the data is that the $J_2$ perturbation, which is well-known to have a noticeable impact perturbing orbits from the two-body model for satellites around the earth, yields an insignificant change when compared to the data when only two-body forces are considered in the process model.
-In fact, every case when the filter diverged occurred with both the two-body problem and $J_2$ simulations.
-Examples of these results are tabulated in Tables \ref{tab:ResultsTBP} and \ref{tab:ResultsJ2} and illustrated in Figure \ref{fig:TBPvsJ2}.
-
-\begin{figure}[h]
-\centerline{
-	\subfigure[State magnitude with the two-body solution]{
-		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1TBPnormX.pdf}}
-	\hfill
-	\subfigure[State magnitude with the $J_2$ solution]{
-		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1HighFidelitynormX.pdf}}
-}
-\centerline{
-	\subfigure[Uncertainty measure with the two-body solution]{
-		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1TBPeigP.pdf}}
-	\hfill
-	\subfigure[Uncertainty measure with the $J_2$ solution]{
-		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1eigP.pdf}}
-}
-\caption{The magnitudes and maximum eigenvalues of the state covariance matrices of the relative states (solid red: true, blue dashed: estimated) in Case 3B with IOD1 serve as an example where the $J_2$ propagated simulations give very close results to those of the two-body solution.
-}\label{fig:TBPvsJ2}
-\end{figure}
+Estimation results are also illustrated at Figures \ref{fig:EKF_i_0}--\ref{fig:EKF_i_50}.
+At Figure \ref{fig:EKF_i_50}, the magnitude of the estimated state becomes quite close to the true value after four orbits.
+This confirms the observability of nonlinear relative orbital dynamics with angles-only measurements.
 
@@
 
-It is well known that the linearized relative orbital dynamics are not observable with angles-only
-measurements.
-This paper formulates a measure of nonlinear observability, and investigates a less conservative observability criteria by using higher-order Lie derivatives.
-Numerous relative orbit determination cases are illustrated numerically by an extended Kalman filter, which shows that the degree of observability by the proposed measure and an accurate IOD prediction are useful for predicting filter convergence.
+It is well known that the linearized relative orbital dynamics are not observable with angles-only measurements.
+This paper shows that the nonlinear relative orbital dynamics are observable under certain geometric conditions.
+Sufficient conditions for observability are derived based on the Lie derivatives of the measurements, and they are illustrated numerically by an extended Kalman filter.
+Future directions include formulating measure of nonlinear observability, and investigating less conservative observability criteria by using higher-order Lie derivatives.
 
-\bibliographystyle{AAS_publication}   % Number the references.
-\bibliography{references}   % Use references.bib to resolve the labels.
+\bibliographystyle{ISSFD_v01}
+\bibliography{references}
+
+\clearpage\newpage
+\begin{figure}
+\centerline{
+\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
+	\includegraphics[width=0.44\textwidth]{PE_i_0_e}}
+\hfill
+\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
+	\includegraphics[width=0.44\textwidth]{PE_i_0_xx_norm}}
+}
+\caption{Extended Kalman Filter $i_{deputy}=0^\circ$}\label{fig:EKF_i_0}
+\end{figure}
+
+\begin{figure}
+\centerline{
+\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
+	\includegraphics[width=0.44\textwidth]{PE_i_10_e}}
+\hfill
+\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
+	\includegraphics[width=0.44\textwidth]{PE_i_10_xx_norm}}
+}
+\caption{Extended Kalman Filter $i_{deputy}=10^\circ$}
+\end{figure}
+
+\begin{figure}
+\centerline{
+\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
+	\includegraphics[width=0.44\textwidth]{PE_i_20_e}}
+\hfill
+\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
+	\includegraphics[width=0.44\textwidth]{PE_i_20_xx_norm}}
+}
+\caption{Extended Kalman Filter $i_{deputy}=20^\circ$}
+\end{figure}
+
+\begin{figure}
+\centerline{
+\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
+	\includegraphics[width=0.44\textwidth]{PE_i_30_e}}
+\hfill
+\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
+	\includegraphics[width=0.44\textwidth]{PE_i_30_xx_norm}}
+}
+\caption{Extended Kalman Filter $i_{deputy}=30^\circ$}
+\end{figure}
+
+\begin{figure}
+\centerline{
+\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
+	\includegraphics[width=0.44\textwidth]{PE_i_40_e}}
+\hfill
+\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
+	\includegraphics[width=0.44\textwidth]{PE_i_40_xx_norm}}
+}
+\caption{Extended Kalman Filter $i_{deputy}=40^\circ$}
+\end{figure}
+
+\begin{figure}
+\centerline{
+\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
+	\includegraphics[width=0.44\textwidth]{PE_i_50_e}}
+\hfill
+\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
+	\includegraphics[width=0.44\textwidth]{PE_i_50_xx_norm}}
+}
+\caption{Extended Kalman Filter $i_{deputy}=50^\circ$}\label{fig:EKF_i_50}
+\end{figure}
 

### 2018-06-09 Conference/AAS 2015 OD EKF/ISSFD14/LaTeX_paper_template_ISSFD24.tex
added Conference folder

\documentclass[12pt,letterpaper]{ISSFD_v01}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{graphicx}

\usepackage{times} % use times font
\usepackage{mathptmx} %use same font for math mode
\DeclareSymbolFont{largesymbols}{OMX}{cmex}{m}{n} %undo the definition of large symbols from the mathptmx package
\usepackage{authblk} % use authblk package for author formatting
\renewcommand\Authfont{\bfseries} % make authors bold
\renewcommand\Affilfont{\normalfont\itshape} % make affiliation italic
\setlength{\affilsep}{0em} % no spacing between authors and affiliations
\usepackage[pdffitwindow=false,pdfstartview={FitH},colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}
\usepackage{microtype} % makes detailed spacing look nicer.  This is only optional -- not required.

\usepackage{xspace}
\newcommand{\twentyfourth}{24\textsuperscript{th}\xspace}
\newcommand{\figText}{Fig.\ }
\newcommand{\tabText}{Tab.\ }
\newcommand{\eqText}{Eq.\ }

\title{NONLINEAR OBSERVABILITY FOR RELATIVE SATELLITE ORBITS
WITH ANGLES-ONLY MEASUREMENTS}

\author[(1)]{T. Alan Lovell}
\author[(2)]{Taeyoung Lee}
\affil[(1)]{Air Force Research Laboratory, 3550 Aberdeen Ave SE, Kirtland AFB, NM 87117 USA,\newline Tel: +1-505-853-4132, Email: splash73@yahoo.com}
\affil[(2)]{Department of Mechanical and Aerospace Engineering, The George Washington University,\newline
801 22nd St NW, Washington DC 20052, Tel: +1-202-994-8710, Email:
tylee@gwu.edu}

\date{}  % don't display a date

\begin{document}
\maketitle

\begin{abstract}
In this paper, nonlinear observability criteria are presented for the relative orbital dynamics represented by the solutions of the two-body problem. It is assumed that a chief is on a circular orbit with a prescribed orbital radius, and it measures line-of-sight toward a deputy. A differential geometric method, based on the Lie derivatives is used to derive sufficient conditions for observability of the orbital properties of deputy. It is shown that under certain geometric conditions on the relative configuration between the chief and the deputy, the nonlinear relative motion is observable from angles-only measurements. An extended Kalman filter is also developed to numerically illustrate the observability of nonlinear relative orbits with angles-only measurements.
\end{abstract}

\begin{keywords}
Relative orbit, Observability, Line-of-sight measurement, Extended Kalman filter.
\end{keywords}

\section{Introduction}
Space-based surveillance or relative navigation is desirable for many spacecraft missions, such as formation control and rendezvous. Spacecraft maneuvers based only on on-board measurements reduce the total operating cost significantly, and it improves safety against communication interruptions with ground stations.  Relative navigation between spacecraft in close-proximity essentially correspond to space-based orbit determination.  In particular, vision-based navigation and estimation of relative orbit have received attention recently, since optical sensors have desirable properties of low cost and minimal maintenance, while providing accurate line-of-sight measurements.

Relative navigation based on angles-only measurements has been investigated in.  The problem is to determine the relative orbit between a chief spacecraft and a deputy spacecraft by using the line-of-sight between the two objects, as-suming that the orbit of the chief is known.  Reference 1 shows that the relative orbit is unobserv-able from angles-only measurements when linear relative orbital dynamics are assumed, unless there are thrusting maneuvers.  Reference 2 investigates observability by using a relative orbit model linearized about spherical coordinates.  Reference 3 introduces the concept of partial ob-servability to determine a basis vector representing a family of relative orbits, and an initial orbit determination technique for this method.  However, all of these results are based on linear relative orbital dynamics.

In this paper, the authors derive observability criteria for the nonlinear relative orbital dynam-ics represented by the solutions of the two-body problem.  Assuming that a chief is on a circular orbit with a prescribed orbital radius, nonlinear equations of motion for the relative orbital motion of a deputy with respect to the chief are derived.  A differential geometric method, based on the Lie derivatives of the line-of-sight from the chief to the deputy4, is applied to obtain sufficient conditions for observability.  It is shown that under certain geometric conditions on the relative configuration between the chief and the deputy, the nonlinear relative motion is observable from angles-only measurements.  We also develop an extended Kalman filter to illustrate the observa-bility properties numerically.

\section{Problem Formulation}

\section{Nonlinear Observability Criteria}

\section{Extended Kalman Filtering}

\section{Conclusions}

\bibliographystyle{ISSFD_v01}
\bibliography{references}

\end{document}

\subsection{Language}
All papers shall be presented and written in English.

\subsection{Page Layout and Paper Size}
The paper shall be written in standard letter paper format (8.5x11 inches, 21.59x27.94 cm) with 1-inch (25.4 mm) margins on all sides (left, right, top and bottom), 
as illustrated in \figText \ref{fig:page_layout}.

\begin{figure}[ht]
 \centering
 \includegraphics{./page_layout.pdf}
 \caption{Page layout}
 \label{fig:page_layout}
\end{figure}

\section{Title and Author Names}
The article should have a centered title of up to 15 words using 12 pt Times New Roman font, in boldface capital 
letters. The author names, which follow the title, should also be centered and written in boldface 12 pt Times New Roman. The author affiliations, addresses and emails shall follow the author names and be
written in 12 pt Times New Roman, formatted in italic style, and centered. These specifications are 
exemplified by the title and author section of this template.

\section{Text Specification}
The text shall be written according to the format specifications presented in \tabText \ref{formatTable}.

\begin{table}[ht]
\caption{Text Formatting}\label{formatTable}
\centering
\begin{tabular}{|l|l|}
\hline
Line spacing & Single \\ \hline
Alignment & Justified \\ \hline
Font type & Times New Roman \\ \hline
Font size & 12 pt \\ \hline
Font style & Regular text \\ \hline
\end{tabular}
\end{table}

\section{Figures and Tables}
All figures and tables should have a caption and be placed as close as possible to their first reference in the text. 
Captions of tables or figures should be formatted in bold 12 pt Times New Roman font, and be centered below the figures 
and above the tables. See \figText \ref{fig:page_layout} and \tabText \ref{formatTable}, above, as examples. Refer to figures in the text with the abbreviation ``\figText 1'', except at the beginning of a sentence, where ``Figure 1'' should be used.

Figures should have good quality and may be grayscale or color. Lettering in figures should be large enough to 
be clearly legible.

\section{Equations}
Each equation should be on a separate line with one blank line before and after. Equations should be numbered 
in increasing order with the numbering flush right to the page margin. Refer to equations in the text with the abbreviation 
 ``\eqText 1'', except at the beginning of a sentence, where ``Equation 1'' should be used. Equation \ref{eq1} is an 
example.
\begin{equation}
 U = \sum_{n=0}^{\infty} \rho_n \sum_{m=0}^{n} A_{nm}(u)D_{nm}(s,t) \label{eq1}
\end{equation}

\nocite{Radice2006}

\end{document}

### 2018-06-09 Conference/AAS 2015 OD EKF/Presentation/AAS15_Observability_RelativeOD_EKF.tex
added Conference folder

\documentclass[hyperref={pdftex,pdfpagemode=none,pdfstartview=FitH},10pt]{beamer}
\usepackage[english]{babel}
\usepackage{times}
%\usepackage{mathptmx,helvet}
\usepackage{amsmath,amssymb}
\usepackage{subfigure,graphicx,tabularx,booktabs,colortbl}
\usepackage{amssymb,amsmath,amsthm,times,graphicx,subfigure,tabularx,booktabs,colortbl,multirow,threeparttable,graphics}
\usepackage{movie15}

\usepackage{color}
\newcommand{\hilight}[1]{\colorbox{green}{#1}}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mbox{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mbox{tr}\ensuremath{\bracket{#1}}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO}(3)}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\SE}{\ensuremath{\mathrm{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\S}{\ensuremath{\mathbb{S}}}
\newcommand{\abs}{\mathop{\mathrm{abs}}\nolimits}

\renewcommand{\r}{\mathbf{r}}
\renewcommand{\u}{\mathbf{u}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\x}{\mathbf{x}}

\definecolor{RoyalBlue}{rgb}{0.25,0.41,0.88}
\def\Emph{\textcolor{RoyalBlue}}

\definecolor{mygray}{gray}{0.9}
\definecolor{tmp}{rgb}{0.804,0.941,1.0}
\setbeamercolor{numerical}{fg=black,bg=tmp}
\setbeamercolor{exact}{fg=black,bg=red}

\mode<presentation> {
  \usetheme{Warsaw}
  \usefonttheme{serif}
  \setbeamercovered{transparent}
}

\setbeamertemplate{footline}%{split theme}
{%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm plus1fill,rightskip=.3cm]{title in head/foot}%
    \usebeamerfont{title in
    head/foot}\insertframenumber/\inserttotalframenumber
  \end{beamercolorbox}}%
  \vskip0pt%
} \setbeamercolor{box}{fg=black,bg=yellow}

\title[Nonlinear Observability Measure for Relative Orbit Determination]
{Nonlinear Observability Measure for Relative Orbit
\\
Determination with Angles-Only Measurements}

\author{
Evan Kaufman\\
{\footnotesize\selectfont Department of Mechanical and Aerospace Engineering,\\
The George Washington University, Washington DC}
$ $\\
\vspace*{0.015\columnwidth}
T. Alan Lovell\\
{\footnotesize\selectfont Air Force Research Laboratory,\\ Kirtland AFB, Albuquerque NM}
$ $\\
\vspace*{0.015\columnwidth}
Taeyoung Lee\\
{\footnotesize\selectfont Department of Mechanical and Aerospace Engineering,\\
The George Washington University, Washington DC}
}

%\institute[GWU]{\footnotesize\selectfont
%$^*$ Air Force Research Laboratory, Kirtland AFB, Albuquerque NM\\$ $\\
%Department of Mechanical and Aerospace Engineering\\
%George Washington University, Washington DC\\$ $\\}

\date{\vspace*{-0.8cm}\\AAS/AIAA Spaceflight Mechanics Meeting 2015\vspace*{0.1cm}\\
\footnotesize\selectfont (Supported by AFRL Summer Faculty and Space Scholars Programs,\\ NSF Grants CMMI-\#1029551, \#1335008, and CNS-\#1337722)}

%	\graphicspath{{/Users/tylee@seas.gwu.edu/Documents/Seminar/ISSFD14/Figs/}}
%\graphicspath{{/Users/evankaufman/Google Drive/FDCL/Evan/Publications/AAS 2015 OD EKF/Presentation/Figs}}
\graphicspath{{Figs/}}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section*{}
\subsection*{I. Introduction}

\begin{frame}
\frametitle{Introduction}
\begin{itemize}
\item Space Relative Navigation
	\begin{itemize}
	\item Space-based orbit determination
	\item Determine relative orbital properties between satellites by on-board measurements
	\item Critical for accurate formation control and close-proximity missions\\
	\end{itemize}
\end{itemize}

\vfill
\centerline{
\includegraphics[height=2.7cm]{XSS11.jpg}\hspace*{0.2cm}
\includegraphics[height=2.7cm]{DLRPrisma.jpg}\hspace*{0.2cm}
\includegraphics[height=2.7cm]{OrbitalExpress.jpg}}
\centerline{\scriptsize\selectfont
AFRL XSS-11\hspace*{1.5cm} DLR PRISMA\hspace*{1.5cm} DARPA Orbital Express}
\end{frame}

\begin{frame}
\frametitle{Introduction}

\begin{itemize}
\item Space-Based Space Surveillance (SBSS)
	\begin{itemize}
	\item Detection, collection, identification and tracking of man-made space objects from deep space to low-earth orbits
%	\item Ground-based optical systems are limited to clear weather, night-time observations
	\item SBSS provides all-weather, day and night, near real-time space situation awareness data
	\item Ability to search for lost/unknown space objects in deep space
	\end{itemize}
	\vspace*{0.3cm}\pause
\item Relative Orbit Determination with \Emph{Angles-Only Measurements}
	\begin{itemize}
	\item Lines-of-sight between satellites are measured by optical sensors
	\item Optical sensors have desirable properties of being low-cost and minimal maintenance
	\item Range is not available:\\ \Emph{Unobservable for linearized relative orbital dynamics} [Woffinden 09]
	\item Observability for nonlinear relative orbital dynamics has been unknown
	\end{itemize}
\end{itemize}

\vspace*{0.3cm}\pause
    \centerline{
    \begin{beamercolorbox}[wd=10.68cm,sep=0.2cm,center,shade=on]{numerical}
        \small\bfseries\selectfont Goal: Investigative Observability for \textit{Nonlinear} Relative Orbital Dynamics
    \end{beamercolorbox}}

\end{frame}

\section*{}
\subsection*{II. Problem Formulation}

\begin{frame}
\frametitle{Relative Orbit Dynamics}

\begin{columns}
\begin{column}{6.5cm}
\footnotesize\selectfont
\only<1>{
\setlength{\unitlength}{1cm}
\begin{picture}(7,7)(0,0)
\put(0,0){\includegraphics[width=7cm]{PF0.pdf}}
\put(3.2,1.3){$a$}
\put(4.8,2.8){Chief}
\end{picture}
}
\only<2>{
\setlength{\unitlength}{1cm}
\begin{picture}(7,7)(0,0)
\put(0,0){\includegraphics[width=7cm]{PF1.pdf}}
\put(3.2,1.3){$a$}
\put(6.15,3.55){$x$}
\put(4.5,4.5){$y$}
\put(4.8,2.8){Chief}
\end{picture}
}
\only<3>{
\setlength{\unitlength}{1cm}
\begin{picture}(7,7)(0,0)
\put(0,0){\includegraphics[width=7cm]{PF2.pdf}}
\put(3.2,1.3){$a$}
\put(6.15,3.55){$x$}
\put(4.5,4.5){$y$}
\put(5.7,4.3){$\mathbf{r}$}
\put(4.8,2.8){Chief}
\put(5.3,6.1){Deputy}
\end{picture}
}
\end{column}
%
\begin{column}{6cm}
\begin{itemize}
\item Chief Satellite
	\begin{itemize}
	\item On a circular orbit
	\item Orbital radius $a$ is known
	\end{itemize}
\vspace*{0.3cm}\pause
\item Local-vertical, local-horizontal (LVLH) frame
	\begin{itemize}
	\item $x$--axis: radial
	\item $y$--axis: tangent to the path
	\item $z$--axis: normal to the orbital plane
	\item Rotating about the $z$--axis with $\mathbf{\omega}=\sqrt{\frac{\mu}{a^3}}$.
	\end{itemize}
\vspace*{0.3cm}\pause
\item Deputy Satellite
	\begin{itemize}
	\item On an unknown, arbitrary orbit\\
	\item Relative position $\mathbf{r}\in\Re^3$

	\end{itemize}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Relative Orbit Dynamics}

\begin{itemize}
\item Nonlinear Equations of Motion (Two-Body Gravity Only)
	\begin{itemize}
	\item State vector: $\mathbf{x}=[\r,\;\dot\r]\in\Re^6$
	\item State equation: $\dot \x = f(\x)$, where
\begin{align*}
f(\mathbf{x}) 
=\begin{bmatrix}
\dot \r \\
-2\omega\times \dot\r - \omega\times(\omega\times \r_a) -\dfrac{\mu \r_a}{\|\r_a\|^3}
\end{bmatrix}.
\end{align*}
($\r_a=\r+[a,0,0]^T$: position vector of deputy from the center of Earth)
	\end{itemize}
\vspace*{0.3cm}\pause
\item Hill-Clohessey-Wiltshire (HCW) Equations
	\begin{itemize}
	\item Linearization of above ODEs about $\x=0$: $\delta\dot\x = A\delta \x$, where
\begin{align*}
A=\deriv{f(\x)}{\x}\bigg|_{\x=0}
=\begin{bmatrix}
0 & 0 & 0 & 1 & 0 & 0\\
0 & 0 & 0 & 0 & 1 & 0\\
0 & 0 & 0 & 0 & 0 & 1\\
3n^2 & 0 & 0 & 0 & 2n & 0\\
0 & 0 & 0 & -2n & 0 & 0\\
0 & 0 & -n^2 & 0 & 0 & 0
\end{bmatrix}.
\end{align*}
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Line-of-sight Measurements}

\begin{columns}
\begin{column}{6.5cm}
\footnotesize\selectfont
\setlength{\unitlength}{1cm}
\begin{picture}(7,7)(0,0)
\put(0,0){\includegraphics[width=7cm]{PF3.pdf}}
\put(3.2,1.3){$a$}
\put(6.15,3.55){$x$}
\put(4.5,4.5){$y$}
\put(5.6,4.0){$\y$}
\put(4.8,2.8){Chief}
\put(5.3,6.1){Deputy}
\end{picture}
\end{column}
%
\begin{column}{6cm}
\begin{itemize}
\item Measurements
	\begin{itemize}
	\item Line-of-sight from chief to deputy is measured
	\item Measurement:
	\begin{align*}
	\y = \frac{\r}{\|\r\|}\in\Re^3
	\end{align*}
	(unit-length: $\|\y\|$=1)
	\end{itemize}
\vspace*{0.3cm}\pause
\item Relative Orbit Determination
	\begin{itemize}
	\item Wish to determine the relative orbit $\x=[\r,\dot\r]$ completely from $\y(t)$ only
	\end{itemize}	
\end{itemize}	
\end{column}
\end{columns}
\end{frame}

\section*{}
\subsection*{III. Observability}

\begin{frame}
\frametitle{Observability of HCW Equations}

\begin{itemize}
\item Solution of the Linear HCW Equation
	\begin{itemize}
	\item Let $\x(t)$ be the solution with an initial condition $\x(t_0)=\x_0$.
	\begin{align*}
	\x(t) = \exp(A(t-t_0)) \x_0
	\end{align*}
	\item \Emph{Homogeneity property}: let $\x'(t)$ be the solution with a \textit{scaled} initial condition $\x'(t_0)=\alpha\x_0$ for any $\alpha\in\Re$, then 
	\begin{align*}
	\x'(t) = \exp(A(t-t_0)) (\alpha\x_0) = \alpha \x(t).
	\end{align*}
	\end{itemize}
\vspace*{0.3cm}\pause
\item Observability of the HCW Equation with Angles-Only Measurements
	\begin{itemize}
	\item The two distinct relative orbits $\x(t)=[\r(t),\dot\r(t)]$ and $\x'(t)=[\r'(t),\dot\r'(t)]$ yield the \Emph{identical} line-of-sight measurements:
	\begin{align*}
	\y(t)=\frac{\r(t)}{\|\r(t)\|},\quad \y'(t)=\frac{\r'(t)}{\|\r'(t)\|}=\frac{\alpha\r(t)}{\|\alpha\r(t)\|}\equiv\y(t).
	\end{align*}
	\end{itemize}
\end{itemize}

\vspace*{0.2cm}\pause
    \centerline{
    \begin{beamercolorbox}[wd=10.68cm,sep=0.2cm,center,shade=on]{numerical}
        \small\bfseries\selectfont Linearized relative orbit is NOT observable with angles-only measurements!
    \end{beamercolorbox}}

\end{frame}

\begin{frame}
\frametitle{Observability of Nonlinear Systems}

\begin{itemize}
\item Observability Concepts 
	\begin{itemize}
	\item Nonlinear system: 
	\begin{align*}\dot \x =f(\x)\in\Re^N,\qquad \y=h(\x)\in\Re^M\end{align*}
	\item A pair of states $\x_0$ and $\x_1$ is \Emph{\textit{indistinguishable}} if the outputs of the corresponding solutions starting from $\x_0$ and $\x_1$ are identical. 
	\item The system is \Emph{\textit{locally weakly observable at $\x_0$}}, if there exists an open neighborhood $U$ of $\x_0$ such that the only indistinguishable point to $\x_0$ in $U$ is the point $\x_0$ itself. 
	\end{itemize}
\vspace*{0.3cm}\pause
\item Observability Criteria  {\small\selectfont [Hermann, Krener 1977]}
	\begin{itemize}
	\item Lie-derivative: $L_f h(\x) = \deriv{h(\x)}{\x}f(\x)=\dot h(\x)\in\Re^{M}$.
	\item Observability Matrix
	\begin{align*}
\mathcal{O}(\x_0) = \deriv{}{\x} \begin{bmatrix} h(\x) \\ L_f^1 h(\x)\\ \vdots \\L_f^{N-1} h(\x)\end{bmatrix}\bigg|_{\x=\x_0}\in\Re^{NM\times N}.
\end{align*}
	\item The system is \Emph{\textit{locally weakly observable at $\x_0$}} if \Emph{$\mathrm{rank}[\mathcal{O}(\x_0)]= N$}.
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Observability of Nonlinear Relative Orbits}

\begin{itemize}
\item Observability Matrix {\footnotesize\selectfont (up to the second order Lie derivative)}
\begin{align}
\mathcal{O} =
\begin{bmatrix}
\deriv{\y}{\r} & \deriv{\y}{\dot\r}\\
\deriv{\dot\y}{\r} & \deriv{\dot\y}{\dot\r}\\
\deriv{\ddot\y}{\r} & \deriv{\ddot\y}{\dot\r}
\end{bmatrix}
\triangleq
\begin{bmatrix}
\mathcal{O}_{00} & 0_{3\times 3}\\
\mathcal{O}_{10} & \mathcal{O}_{11}\\
\mathcal{O}_{20} & \mathcal{O}_{21}
\end{bmatrix}\in\Re^{9\times 6},
\end{align}
\only<1>{\scriptsize\selectfont where
\begin{align*}
\mathcal{O}_{00} & = \deriv{\y}{\r} = \frac{1}{\|\r\|}(I - \y\y^T),\\
\mathcal{O}_{10} & = \deriv{\dot\y}{\r} = \deriv{}{\r} \parenth{\deriv{\y}{\r}\dot \r}= -\frac{1}{\|\r\|^2}\braces{
{\dot\r\y^T +\y^T\dot\r I +\y\dot\r^T}
- 3\y\y^T\dot\r \y^T},\\
\mathcal{O}_{11} & = \deriv{\dot\y}{\dot\r} = \deriv{}{\dot\r} \parenth{\deriv{\y}{\r}\dot \r}= \deriv{\y}{\r} = \mathcal{O}_{00},\\
\mathcal{O}_{20} & = \deriv{\ddot\y}{\r} = -\dfrac{2\dot\r \dot\r^T  +\dot\r^T\dot\r I}{\|\r\|^3}
+3\dfrac{2(\r^T\dot\r) \dot\r\r^T +(\dot\r^T\dot\r)\r\r^T
+(\r^T\dot\r)^2 I
+2(\r^T\dot\r)\r  \dot\r^T
}{\|\r\|^5}
- 15 \dfrac{(\r^T\dot\r)^2\r  \r^T}{\|\r\|^7}\nonumber \\
&\quad -\dfrac{\ddot{\r}\r^T +\r^T\ddot{\r} I +\r \ddot{\r}^T}{\|\r\|^3}
+ 3 \dfrac{(\r^T\ddot{\r})\r \r^T}{\|\r\|^5}
+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}},\\
\mathcal{O}_{21} & = \deriv{\ddot\y}{\dot\r} = \deriv{}{\dot\r}\parenth{\deriv{}{\r}\parenth{\deriv{\y}{\r}\dot\r}\dot\r + \deriv{\y}{\r}\ddot\r} = 2 \deriv{}{\r} \parenth{\deriv{\y}{\r}\dot\r} + \deriv{\y}{\r} \deriv{\ddot\r}{\dot\r}
= 2\mathcal{O}_{10} 
- 2 \mathcal{O}_{00}[\omega]_\times,
\vspace*{-0.7cm}\end{align*}
(These expressions have been verified by \Emph{Matlab Symbolic Computation Toolbox})}
%
\only<2>{\scriptsize\selectfont They satisfy the following identities:
\begin{align*}
\mathcal{O}_{10}\r & = 
-\mathcal{O}_{00}\dot\r,\label{eqn:O10r}\\
\mathcal{O}_{10}\dot\r & = -\frac{1}{\|\r\|^2}\braces{
{2\dot\r(\y^T\dot \r)  +\y(\dot\r^T\dot \r)}
- 3\y(\y^T\dot\r)^2},\label{eqn:O10dotr}\\
\mathcal{O}_{20}\r %& = 
& = -2\mathcal{O}_{10}\dot\r +\mathcal{O}_{00}\braces{-\ddot \r +\deriv{\ddot\r}{\r}\r},\label{eqn:O20r}\\
\mathcal{O}_{21}\r & = -2\mathcal{O}_{00} (\dot\r+\omega\times\r),\label{eqn:O21r}\\
\mathcal{O}_{21}\dot\r & = 2\mathcal{O}_{10} \dot\r
- 2 \mathcal{O}_{00}[\omega]_\times\dot\r,\label{eqn:O21rdot}
\end{align*}
which are useful to derive the observability criteria.}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Observability of Nonlinear Relative Orbits}

\begin{theorem}[Sufficient Conditions for Observability]
Define three vectors $\mathbf{v}_{rel}$, $\mathbf{a}_1,\mathbf{a}_2\in\Re^3$ as 
\begin{align}
\mathbf{v}_{rel} & = \dot\r +\omega\times\r,\label{eqn:vrel}\\
\mathbf{a}_1 & = \ddot\r-\deriv{\ddot\r}{\r}\r = -2\omega\times \dot\r - [\omega]_\times^2 a \mathbf{e}_1 -\dfrac{\mu a }{\|\r_a\|^3}\mathbf{e}_1- \dfrac{3\mu\r_a^T\r}{\|\r_a\|^5}\r_a,\label{eqn:a1}\\
\mathbf{a}_2 & = \ddot\r-\deriv{\ddot\r}{\r}\r -\deriv{\ddot\r}{\dot\r}\dot\r = \mathbf{a}_1+2\omega\times\dot\r.\label{eqn:a2}
\end{align}
The nonlinear relative orbital dynamics is locally weakly observable at $\x=[\r^T,\dot\r^T]$ if
%\begin{list}{}{\setlength{\leftmargin}{1cm}\setlength{\itemsep}{0mm}\setlength{\parsep}{0mm}\setlength{\topsep}{0mm}\setlength{\parskip}{0mm}\setlength{\labelwidth}{2cm}}
%\item[(i)] when $\r\times\dot\r =0$, $\r\times\mathbf{a}_1\neq 0$ and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_1)\neq 0$,
%\item[(ii)] when $\r\times\dot\r \neq 0$, $\r\times\mathbf{a}_2\neq 0$ and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_2)\neq 0$.
%\end{list}
\begin{align}
(i)\; &\text{when $\r\times\dot\r =0$, $\r\times\mathbf{v}_{rel}\neq0$, $\r\times\mathbf{a}_1\neq 0$,  and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_1)\neq 0$},\label{eqn:cond1}\\
(ii)\; &\text{when $\r\times\dot\r \neq 0$, $\r\times\mathbf{v}_{rel}\neq0$, $\r\times\mathbf{a}_2\neq 0$ and $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_2)\neq 0$.}\label{eqn:cond2}
\end{align}
\end{theorem}

\vspace*{0.2cm}\pause
    \centerline{
    \begin{beamercolorbox}[wd=11.5cm,sep=0.2cm,center,shade=on]{numerical}
        \small\bfseries\selectfont Nonlinear relative orbits are OBSERVABLE under certain geometric conditions!
    \end{beamercolorbox}}

\end{frame}

\begin{frame}
\frametitle{Observability of Nonlinear Relative Orbits}
\setcounter{subfigure}{0}

\begin{itemize}
\item Sufficient condition: $\r\times\mathbf{v}_{rel}\neq0$
	\begin{itemize}
	\item $\mathbf{v}_{rel} = \dot\r +\omega\times\r$: relative velocity in the inertial frame
	\item violated if deputy is flying directly toward or away from chief
	\end{itemize}
	\vspace*{0.3cm}\pause
\item Sufficient condition: $\r\times\mathbf{a}_1\neq 0$ (or $\r\times\mathbf{a}_2\neq 0$)
	\begin{itemize}
	\item $\mathbf{a}_1 = \ddot\r-\deriv{\ddot\r}{\r}\r$: difference between acceleration and its variation along $\r$
	\item satisfied in general as $\mathbf{a}_1$ (or $\mathbf{a}_2$) is pretty arbitrary
	\end{itemize}
	\vspace*{0.3cm}\pause
\item Sufficient condition: $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_1)\neq 0$ (or $\r^T(\mathbf{v}_{ref}\times \mathbf{a}_2)\neq 0$)
	\begin{itemize}
	\item The (signed) volume of the \textit{parallelepiped} defined by $\r$, $\mathbf{v}_{rel}$, and $\mathbf{a}_1$ (or $\mathbf{a}_2$)

	\centerline{\setlength{\unitlength}{1cm}\footnotesize\selectfont
	\begin{picture}(3,2.2)(0,0)
	\only<3>{
	\put(0,0){\includegraphics[width=3cm]{pp.jpg}}}
	\put(1.,0.25){$\r$}
	\put(0.5,0.6){$\mathbf{v}_{rel}$}
	\put(0.2,0.9){$\mathbf{a}$}
	\end{picture}}
	\item violated for planar relative orbits
	\end{itemize}
%\vspace*{0.3cm}\pause
\item How observable is a relative orbit? Goal: define a measure of observability.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Measure of Observability}

\begin{itemize}
\item Main idea: the linearized dynamics of a relative orbit are unobservable, but the nonlinear properties of these systems may increase the observability by some amount that we wish to measure.
\item Approach: define an observability Gramian, which measures the sensitivity of the output with respect to the relative orbit initial condition over the estimation time period:
\begin{align}
\mathcal{W} (\x_0,t_0,t_f) = \int_{t_0}^{t_f} \parenth{\deriv{\y(\tau)}{\x_0}}^T\deriv{\y(\tau)}{\x_0}\,d\tau,\label{eqn:Wo_NL}
\end{align}
which is numerically calculated element-by-element as
\begin{align}
[\mathcal{W}]_{ij} = \frac{1}{4\epsilon^2}\sum_{k=0}^N (\y^{i+}_k-\y^{i-}_k)(\y^{j+}_k-\y^{j-}_k) \Delta t_k.
\end{align}
\item The smaller $\mathrm{cond}\braces{\mathcal{W}}$ is, the more the nonlinear system becomes observable.
\end{itemize}

\end{frame}

\section*{}
\subsection*{IV. Numerical Example}

\begin{frame}
\frametitle{Numerical Examples}

\begin{itemize}
\item Orbital Properties
	\begin{itemize}
	\item Chief: a circular orbit with an orbital radius of $7100\,\mathrm{km}$ such that $x_{chief,0}=[5020,\ 5020,\ 0,\ -1.812,\ 1.812,\ 7.041]^T$
	\item Deputy orbits are chosen at various locations, with relative orbits plotted:
	\end{itemize}	
\end{itemize}
\begin{figure}[h]
\centerline{
	\subfigure[Case 0C]{
		\includegraphics[width=0.19\textwidth]{OM_IOD_0C_traj.pdf}}
	\hfill
	\subfigure[Case $\mathbf{1}$]{
		\includegraphics[width=0.19\textwidth]{OM_IOD_1_traj.pdf}}
	\hfill
	\subfigure[Case 2A]{
		\includegraphics[width=0.19\textwidth]{OM_IOD_2A_traj.pdf}}
	\hfill
	\subfigure[Case 3A]{
		\includegraphics[width=0.19\textwidth]{OM_IOD_3A_traj.pdf}}
	\hfill
	\subfigure[Case 4C]{
		\includegraphics[width=0.19\textwidth]{OM_IOD_4C_traj.pdf}}
}
\caption{An example from each set of relative trajectories (solid red: HCW, dashed blue: two-body solution). Better tracking was shown with those trajectories that changed with time (not entirely cyclic) or varied far from the linearized solution.}
\label{fig:0CasesTraj}
\end{figure}

%\begin{figure}[h]
%\label{fig:CaseTraj}
%\centerline{
%	\subfigure[Case 1]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_1_traj.pdf}}
%}
%\caption{The trajectory of the Case 1 (solid red: HCW, dashed blue: two-body solution) shows that the linear and nonlinear solutions are very close, where both solutions are largely planar and cyclic.}
%\label{fig:1CasesTraj}
%\end{figure}
%
%
%\begin{figure}[h]
%\centerline{
%	\subfigure[Case 2A]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_2A_traj.pdf}}
%	\hfill
%	\subfigure[Case 2B]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_2B_traj.pdf}}
%	\hfill
%	\subfigure[Case 2C]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_2C_traj.pdf}}
%}
%\caption{The trajectories of the Cases 2A, 2B, and 2C (solid red: HCW, dashed blue: two-body solution) show differences between the linear and nonlinear solutions that vary about the $y$-axis, increasing the observabilities.}
%\label{fig:2CasesTraj}
%\end{figure}
%
%\begin{figure}[h]
%\centerline{
%	\subfigure[Case 3A]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_3A_traj.pdf}}
%	\hfill
%	\subfigure[Case 3B]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_3B_traj.pdf}}
%	\hfill
%	\subfigure[Case 3C]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_3C_traj.pdf}}
%}
%\caption{The trajectories of the Cases 3A, 3B, and 3C (solid red: HCW, dashed blue: two-body solution) exhibit large observabilities due to nonlinear solutions that move far away from the linearized solutions.}
%\label{fig:3CasesTraj}
%\end{figure}
%
%\begin{figure}[h]
%\centerline{
%	\subfigure[Case 4A]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_4A_traj.pdf}}
%	\hfill
%	\subfigure[Case 4B]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_4B_traj.pdf}}
%	\hfill
%	\subfigure[Case 4C]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_4C_traj.pdf}}
%}
%\caption{The trajectories of the Cases 4A, 4B, and 4C (solid red: HCW, dashed blue: two-body solution) experience the largest observabilities due to nonlinear solutions that move farther away from the linearized solutions than any of the trajectories from other cases considered in this paper.}
%\label{fig:4CasesTraj}
%\end{figure}

\end{frame}

\begin{frame}
\frametitle{Numerical Examples}
\framesubtitle{Observability Measure for the Various Cases}

\begin{itemize}
\item An initial orbit determination (IOD) approximation serves as the initial value for the Kalman filter.
\end{itemize}

\begin{center}
\resizebox{.5\columnwidth}{!}{
\begin{threeparttable}[h]
\caption{Case Observabilities and Initial Errors}
\begin{tabularx}{\textwidth}
{
>{$}c<{$}
*{1}{>{$}c<{$}} |
*{2}{>{$}c<{$}} |
*{2}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{\multirow{1}{*}{IOD1}} & \multicolumn{2}{c}{\multirow{1}{*}{IOD2}} \\
& &  e_{mag,0} & e_{dir,0} & e_{mag,0} & e_{dir,0} \\\midrule
0A & 10^{11.1486} & 0.0003 & 1.4901\times10^{-8} & 0.0012  & 2.1073\times10^{-8} \\
0B & 10^{10.9387} & 0.0233 & 0 & 0.0088 & 0 \\
0C & 10^{10.9851} & 0.0152 & 2.1073\times10^{-8} & 0.0139 & 1.4901\times10^{-8}  \\
\midrule
\mathbf{1} & 10^{16.6871} & 0.0001 & 0 & 0 & 0  \\
\midrule
2A & 10^{10.6492} & 0.0550 & 1.8150\times10^{-4} & 0.0238 & 1.8150\times10^{-4}  \\
2B & 10^{10.4163} & 0.9571 & 0.0734                     & 0.9571 & 0.0734  \\
2C & 10^{10.2155} & 0.0219 & 0.0044 & 0.0065 & 0.0044 \\
\midrule
3A & 10^{8.3490} & 0.0191 & 0.0031 & 0.1830 & 0.0031  \\
3B & 10^{8.5372} & 0.0612 & 0.0533 & 0.2315 & 0.0533  \\
3C & 10^{8.6586} & 6.8962 & 0.7348 & 5.4921 & 0.7348  \\
\midrule
4A & 10^{7.9062} & 8.7347 & 0.4587 & 0.1575 & 0.4587  \\
4B & 10^{7.9903} & 0.9276 & 0.1819 & 0.0621 & 0.1819  \\
4C & 10^{8.0645} & 0.1161 & 0.0554 & 0.0354 & 0.0554  \\
\bottomrule
\end{tabularx}
{\small
\begin{tablenotes}
    \item Note: $0$ indicates numbers small enough to be below Matlab numerical accuracy.
  \end{tablenotes}}
\label{tab:IODerr}
\end{threeparttable}
}
\end{center}
\vspace*{-0.1cm}
where $e_{mag,0}=\abs\left(1-{\frac{\norm{\r^+_0}}{\norm{\r_0}}}\right)$
and
$e_{dir,0}=\cos^{-1}\left(\frac{\r_0}{\norm{\r_0}}\cdot \frac{\r^+_0}{\norm{\r^+_0}}\right)$.

\end{frame}

\begin{frame}
\frametitle{Numerical Examples}
\framesubtitle{Kalman filter metrics with the two-body solution and $J_2$ effects}

\begin{center}
\begin{minipage}{0.45\columnwidth}
\resizebox{\columnwidth}{!}{
\begin{threeparttable}[h]
\caption{Error variables for EKF Considering Only Two-Body Problem Forces}
\begin{tabularx}{1.8\columnwidth}
{
>{$}c<{$}
*{1}{>{$}c<{$}} |
*{2}{>{$}c<{$}} |
*{2}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{IOD1} & \multicolumn{2}{c}{IOD2} \\
& & \bar e_{mag} & \bar e_{dir} & \bar e_{mag} & \bar e_{dir} \\\midrule
0A & 10^{11.1486} &  0.1918  &  0.0069  &  0.3140  &  0.0129 \\
0B & 10^{10.9387} &  0.7262  &  0.0505  &  0.1571  &  0.0086\\
0C & 10^{10.9851} &  0.0900  &  0.0061 &  0.1002  &  0.0154\\
\midrule
\mathbf{1} & 10^{16.6871} & \cellcolor{green} 4.059  & \cellcolor{green}  0.0109 & \cellcolor{green} 42.381 & \cellcolor{green}  0.0445 \\
\midrule
2A & 10^{10.6492} &  0.1978  &  0.0047  &  0.1638  &   0.0065   \\
2B & 10^{10.4163} & \cellcolor{green}  0.9853  & \cellcolor{green}  0.9307 & \cellcolor{green}  0.9278 & \cellcolor{green}   0.0114 \\
2C & 10^{10.2155} & 0.0563 &  0.0047 & 0.0536 &  0.0059 \\
\midrule
3A & 10^{8.3490} &  0.0516 & 0.0041   &  0.0643  &  0.0053 \\
3B & 10^{8.5372} &  0.0739  &  0.0035 &  0.1485 &  0.0071 \\
3C & 10^{8.6586} & \cellcolor{green}  2.063 & \cellcolor{green} 0.1614    & \cellcolor{green}  10.136  & \cellcolor{green}   0.9195 \\
\midrule
4A & 10^{7.9062} & \cellcolor{green}  6.359  & \cellcolor{green} 1.4235  &  12.0062   &  1.5310  \\
4B & 10^{7.9903} & \cellcolor{green}  0.2773  & \cellcolor{green}  0.0814   &  0.5244  &  0.2305 \\
4C & 10^{8.0645} &  2.0105 &  0.9144   &  0.0303 &  0.0036 \\
\bottomrule
\end{tabularx}
{\small
\begin{tablenotes}
    \item \hilight{Filter diverges}
  \end{tablenotes}}
\label{tab:ResultsTBP}
\end{threeparttable}
}
\end{minipage}
\begin{minipage}{0.45\columnwidth}
\resizebox{\columnwidth}{!}{
\begin{threeparttable}[h]
\caption{Error variables for EKF Considering Only Two-Body Problem Forces}
\begin{tabularx}{1.8\columnwidth}
{
>{$}c<{$}
*{1}{>{$}c<{$}} |
*{2}{>{$}c<{$}} |
*{2}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{IOD1} & \multicolumn{2}{c}{IOD2} \\
& & \bar e_{mag} & \bar e_{dir} & \bar e_{mag} & \bar e_{dir} \\\midrule
0A & 10^{11.1486} &  0.3766  &  0.0144  &  0.5145  &  0.0691 \\
0B & 10^{10.9387} &  0.5186  &  0.0810  &  0.3092  &  0.0804\\
0C & 10^{10.9851} &  0.2723  &  0.1362 &  0.1707 &  0.0614\\
\midrule
\mathbf{1} & 10^{16.6871} & \cellcolor{green} 4.078  & \cellcolor{green}  0.0109 & \cellcolor{green} 42.486 & \cellcolor{green}  0.0445 \\
\midrule
2A & 10^{10.6492} &  0.1592  &   0.0043  &  0.1348  &   0.0065   \\
2B & 10^{10.4163} & \cellcolor{green} 0.9843   & \cellcolor{green}  0.0114  & \cellcolor{green}  0.9278 & \cellcolor{green}   0.0113 \\
2C & 10^{10.2155} & 0.0258 &  0.0044 & 0.0360 &  0.0059 \\
\midrule
3A & 10^{8.3490} &  0.0486 & 0.0040   &  0.0603  &  0.0053 \\
3B & 10^{8.5372} &  0.0680  &  0.0033 &  0.1459 &  0.0070 \\
3C & 10^{8.6586} & \cellcolor{green}  2.0611 & \cellcolor{green}  0.1595   & \cellcolor{green}  10.567  & \cellcolor{green}   0.9011 \\
\midrule
4A & 10^{7.9062} & \cellcolor{green}  6.0706  & \cellcolor{green} 1.3182  &  11.8882   &  1.5319  \\
4B & 10^{7.9903} & \cellcolor{green}  0.2723  & \cellcolor{green}  0.0789   &  0.3155  &  0.1061 \\
4C & 10^{8.0645} &  2.3699 &  0.8486   &  0.0278 &  0.0034 \\
\bottomrule
\end{tabularx}
{\small
\begin{tablenotes}
    \item \hilight{Filter diverges}
  \end{tablenotes}}
\label{tab:ResultsTBP}
\end{threeparttable}
}
\end{minipage}
\end{center}
\resizebox{.9\columnwidth}{!}{
where $\bar e_{mag}=\left(\frac1N\sum_{k=1}^N\left\{1-{\frac{\norm{\r^+_k}}{\norm{\r_k}}}\right\}^2\right)^{\frac12}$ and $\bar e_{dir}=\left(\frac1N\sum_{k=1}^N
\left\{\cos^{-1}\left(\frac{\r_k}{\norm{\r_k}}\cdot \frac{\r^+_k}{\norm{\r^+_k}}\right)\right\}^{2}
\right)^{\frac12}$.
}

\end{frame}

%\begin{frame}
%\setcounter{subfigure}{0}
%\frametitle{Numerical Examples}
%\framesubtitle{Deputy inclination $i=20^\circ$}
%
%\begin{figure}
%\only<1>{
%\centerline{
%	\subfigure[Trajectory in the inertial frame]{
%		\includegraphics[width=0.5\columnwidth]{ISSFD14_i_20_OT_iner.pdf}}
%	\hspace*{0.05\textwidth}
%	\subfigure[Relative trajectory in the LVLH frame]{
%		\includegraphics[width=0.5\columnwidth]{ISSFD14_i_20_OT_lvlh.pdf}}
%}}
%\only<2>{
%\centerline{
%	\subfigure[Min. singular value of $\mathcal{O}$]{
%		\includegraphics[width=0.5\columnwidth]{ISSFD14_i_20_OT_sig.pdf}}
%	\hspace*{0.05\textwidth}
%	\subfigure[Condition number of $\mathcal{O}$]{
%		\includegraphics[width=0.5\columnwidth]{ISSFD14_i_20_OT_cond.pdf}}
%}}
%\only<3>{
%\centerline{
%	\subfigure[Estimation errors]{
%		\includegraphics[width=0.5\columnwidth]{PE_i_20_e.pdf}}
%	\hspace*{0.05\textwidth}
%	\subfigure[True $\|\x\|$ (red) and Estimated $\|\hat\x\|$ (blue)]{
%		\includegraphics[width=0.5\columnwidth]{PE_i_20_xx_norm.pdf}}
%}}
%\end{figure}
%\end{frame}

\begin{frame}
\setcounter{subfigure}{0}
\frametitle{Numerical Examples}
\framesubtitle{True and estimated solution with the Kalman fitler}

\begin{figure}[h]
\centerline{
	\subfigure[Case 0C, IOD1]{
		\includegraphics[width=0.2\textwidth]{Case0Cq1IOD1TBPnormX.pdf}}
	\hfill
	\subfigure[Case $\mathbf{1}$, IOD1]{
		\includegraphics[width=0.2\textwidth]{Case1q1IOD1TBPnormX.pdf}}
	\hfill
	\subfigure[Case 2A, IOD1]{
		\includegraphics[width=0.2\textwidth]{Case2Aq1IOD1TBPnormX.pdf}}
	\hfill
	\subfigure[Case 3A, IOD1]{
		\includegraphics[width=0.2\textwidth]{Case3Aq1IOD1TBPnormX.pdf}}
	\hfill
	\subfigure[Case 4C, IOD1]{
		\includegraphics[width=0.2\textwidth]{Case4Cq1IOD1TBPnormX.pdf}}
}
\centerline{
	\subfigure[Case 0C, IOD2]{
		\includegraphics[width=0.2\textwidth]{Case0Cq1IOD2TBPnormX.pdf}}
	\hfill
	\subfigure[Case $\mathbf{1}$, IOD2]{
		\includegraphics[width=0.2\textwidth]{Case1q1IOD2TBPnormX.pdf}}
	\hfill
	\subfigure[Case 2A, IOD2]{
		\includegraphics[width=0.2\textwidth]{Case2Aq1IOD2TBPnormX.pdf}}
	\hfill
	\subfigure[Case 3A, IOD2]{
		\includegraphics[width=0.2\textwidth]{Case3Aq1IOD2TBPnormX.pdf}}
	\hfill
	\subfigure[Case 4C, IOD2]{
		\includegraphics[width=0.2\textwidth]{Case4Cq1IOD2TBPnormX.pdf}}
}
\caption{The magnitudes of the relative states (solid red: true, blue dashed: estimated) are shown for various two-body propagated cases with close IOD estimates where the filter only diverges when the observability is low.}\label{fig:ExampesEachCase}
\end{figure}

\end{frame}

%\begin{frame}
%\setcounter{subfigure}{0}
%\frametitle{Numerical Examples}
%\framesubtitle{Deputy inclination $i=40^\circ$}
%
%\begin{figure}
%\only<1>{
%\centerline{
%	\subfigure[Trajectory in the inertial frame]{
%		\includegraphics[width=0.5\columnwidth]{ISSFD14_i_40_OT_iner.pdf}}
%	\hspace*{0.05\textwidth}
%	\subfigure[Relative trajectory in the LVLH frame]{
%		\includegraphics[width=0.5\columnwidth]{ISSFD14_i_40_OT_lvlh.pdf}}
%}}
%\only<2>{
%\centerline{
%	\subfigure[Min. singular value of $\mathcal{O}$]{
%		\includegraphics[width=0.5\columnwidth]{ISSFD14_i_40_OT_sig.pdf}}
%	\hspace*{0.05\textwidth}
%	\subfigure[Condition number of $\mathcal{O}$]{
%		\includegraphics[width=0.5\columnwidth]{ISSFD14_i_40_OT_cond.pdf}}
%}}
%\only<3>{
%\centerline{
%	\subfigure[Estimation errors]{
%		\includegraphics[width=0.5\columnwidth]{PE_i_40_e.pdf}}
%	\hspace*{0.05\textwidth}
%	\subfigure[True $\|\x\|$ (red) and Estimated $\|\hat\x\|$ (blue)]{
%		\includegraphics[width=0.5\columnwidth]{PE_i_40_xx_norm.pdf}}
%}}
%\end{figure}
%\end{frame}

\begin{frame}
\setcounter{subfigure}{0}
\frametitle{Numerical Examples}
\framesubtitle{Plotted $J_2$ Perturbation Effects}

%\begin{itemize}
%\item Results show $J_2$ perturbations have little effect.
%\end{itemize}

\begin{figure}[h]
\centerline{
	\subfigure[State magnitude with the two-body solution]{
		\includegraphics[width=0.2\textwidth]{Case3Bq1IOD1TBPnormX.pdf}}
\hspace*{0.07\textwidth}
	\subfigure[State magnitude with the $J_2$ solution]{
		\includegraphics[width=0.2\textwidth]{Case3Bq1IOD1HighFidelitynormX.pdf}}
}
\centerline{
	\subfigure[Uncertainty measure with the two-body solution]{
		\includegraphics[width=0.2\textwidth]{Case3Bq1IOD1TBPeigP.pdf}}
\hspace*{0.07\textwidth}
	\subfigure[Uncertainty measure with the $J_2$ solution]{
		\includegraphics[width=0.2\textwidth]{Case3Bq1IOD1eigP.pdf}}
}
\caption{The magnitudes and maximum eigenvalues of the state covariance matrices of the relative states (solid red: true, blue dashed: estimated) in Case 3B with IOD1 serve as an example where the $J_2$ propagated simulations give very close results to those of the two-body solution.
}\label{fig:TBPvsJ2}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Numerical Examples}
\framesubtitle{Summary}

\begin{itemize}
\item Success or failure of the Kalman filter using angles-only measurements for relative orbit estimation depends on 2 criteria:
\begin{enumerate}
\item If the condition number of $\mathcal{W}$ is smaller, then the observability of the nonlinear system is greater.
\item If the IOD or an initial estimate is far from the truth, then the Kalman filter may not converge on the true solution.
\end{enumerate}
\item The $J_2$ perturbations had an insignificant effect on the results of these simulations.
\end{itemize}

\end{frame}

\section*{}
\subsection*{V. Conclusions}

\begin{frame}
\frametitle{Conclusions}

\begin{itemize}
\item When linearized dynamics are considered for angles-only relative orbit estimation, the system is unobservable.
\item When certain criteria are met, the unobservable linearized systems become observable if the nonlinear properties of those systems are considered.
\item An observability measure is derived to quantify the degree to which relative orbits become observable.
\item A Kalman filter serves to show how cases with greater observability (shown by the proposed measure) are more likely to converge on the true solution provided an accurate IOD estimate.
\item Cases with greater observability tend to converge more accurately to the true solution.
\item When $J_2$ perturbations are included into the process model, the Kalman filter performs similarly.
\end{itemize}

\end{frame}
\end{document}

### 2018-06-09 Conference/ACC 2016/ACC16.1 Old Version 2/ACC16.1 (1)[Conflict].tex
added Conference folder

\documentclass[letterpaper, 10pt, conference]{ieeeconf}
\IEEEoverridecommandlockouts \overrideIEEEmargins
\usepackage{amsmath,amssymb,url,times,subfigure,graphicx,theorem}
\usepackage{graphicx,subfigure,hyperref}
\usepackage{color,comment}
\usepackage{csquotes}
%\usepackage{refcheck}

\usepackage{epic,eepic}

\usepackage[]{algorithm2e}
%\usepackage{algorithmicx}
%\usepackage{algpseudocode}

%\usepackage{algorithm}
%\usepackage{algpseudocode}
%\usepackage{pifont}

%\usepackage{amsmath}
%\usepackage{algorithm}
%\usepackage[noend]{algpseudocode}
%
%\makeatletter
%\def\BState{\State\hskip-\ALG@thistlm}
%\makeatother

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\abs}[1]{\ensuremath{\left| #1 \right|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\ip}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mbox{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mbox{tr}\ensuremath{\!\bracket{#1}}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathbf{d}}}
\newcommand{\pair}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\!\langle #1 \rangle\!\rangle}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\ad}{\ensuremath{\mathrm{ad}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}

\title{\LARGE \bf
Bayesian Occupancy Grid Mapping via Exact Inverse Sensor Model}

\author{Evan Kaufman, Taeyoung Lee%
\thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 {\tt \{evankaufman,tylee\}@gwu.edu}}
\thanks{This research has been supported in part by NRL under the grant *********, and NSF under the grants CMMI-1243000, CMMI-1335008, and CNS-1337722.}
}

\newcommand{\EditTL}[1]{{\color{red}\protect #1}}
%\renewcommand{\EditTL}[1]{{\protect #1}}

\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{cor}{Corollary}
\newtheorem{remark}{Remark}

\begin{document}
\allowdisplaybreaks

\maketitle \thispagestyle{empty} \pagestyle{empty}

\begin{abstract}
Occupancy grid maps are spatial representations of environments, where the space of interest is decomposed into a number of cells that are considered either occupied or free. This paper is focused on occupancy grid mapping, which is to estimate the probability of occupancy for each cell based on range measurements from a known location. For a given probabilistic model of a range sensor, we propose a computationally efficient method to obtain an exact inverse sensor model, and it is utilized to construct a probabilisitic mapping algorithm according to the Bayesian framework. Compared with the existing occupancy grid mapping techniques that rely on approximate inverse sensor models, the proposed approach yields substantially more accurate maps for the same set of measurements. These are illustrated by numerical examples and experiments. 
\end{abstract}

\section{Introduction}

Mapping in the context of robotics is to generate a representation of the environment around the robot.
This problem is crucial to understanding the surrounding regions when a robot enters an uncertain environment
Various map representations have been applied to the mapping problem such as occupancy grids~\cite{WolSuk05}, octomaps~\cite{WurHorBenStaBur10}, or with distinguishable features in a landmark representation~\cite{MonThrKolWeg02}.
To build maps in real time, a computationally-efficient algorithm is required.% \EditTL{rewrite this paragraph for mapping, without introducing SLAM here}

This paper focusses on the popular occupancy grid representation, where the map space is decomposed into evenly-spaced grid cells that are considered as either occupied or free. Obtaining the probability of occupancy for a grid cell based on range measurements and the configuration of the mobile robot, known as the \emph{inverse sensor model}, is of fundamental importance with occupancy grid mapping. However, the exact solution to the inverse sensor model has not been utilized in occupancy grid mapping, because the computational requirements appear too cumbersome to compute the exact model in real time~\cite{ThrBurFox05}.

Instead, several techniques have been proposed to approximate the exact inverse sensor model. %Due to the computational limitations, two common approaches are applied to obtain a function close to the inverse sensor model.
In the early occupancy grid mappings based on sonar sensors, \emph{ad hoc} techniques are employed to obtain an inverse sensor model, where the bearings of depth measurements are considered uncertain because the directions of bouncing sound waves are highly inaccurate~\cite{MorElf85,Elf89}.
%These occupancy grid mapping schemes employ \emph{ad hoc} techniques to obtain the inverse sensor model.
These techniques are based on approximate functions to represent the inverse sensor model with questionable accuracy~\cite{ChoLynHutKanBurKavThr05} and applied to more modern sensors in~\cite{And09},~\cite{PirRutBisSch11}.

The other popular approach to obtain an approximate inverse sensor model is to simulate maps, robot poses, and measurements and use a learning algorithm, such as neural network to approximate the inverse sensor model~\cite{Thr01,ThrBurFox05}. These approaches are undesirable in practice due to complexities associated to implementing a learning algorithm. For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples or how many samples are required to obtain a reasonable approximation. Furthermore, it is challenging to apply any learning algorithm over the large dimensional space composed of maps, poses, and measurements.

%and they only yield probabilistic approximations without strong mathematical justification.

%These approaches simply treat occlusions with discrete probabilities of detection, regardless of whether one cell may completely, partially, or not block another cell in the sensor field of view.

%Furthermore, both approaches follow a log-odds ratio Bayesian framework, imposing an additional Markov assumption.

%These techniques are not also suitable for modern range sensor with the highly accurate bearing~\cite{Thr03}. Based on recent improvements in sensor technology, the occupancy grid mapping problem has been restructured using the fact that depth measurement bearings are known. \EditTL{do we need this paragraph?}

This paper proposes a computationally efficient algorithm to construct the exact inverse sensor model. More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, this algorithm yields a posteriori probability of occupancy of all of the cells within the area covered by the range sensor from the range measurements. The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancy of the other cells blocked by it is irrelevant to the forward sensor model, and it is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model. Furthermore, the proposed approach integrates a priori probability of occupancy and multiple range measurements according to the Bayesian framework to obtain more accurate maps. As such, it is contrast to the existing framework based on log-odds ratios that impose an additional Markov assumption.

%these are constructed

In short, the main contribution of this paper is proposing the exact inverse sensor model and utilizing it into Bayesian occupancy grid mapping. Compared with the current mapping algorithms based on an approximate inverse sensor model, this approach yields more accurate occupancy probabilities. This constructs substantially more accurate occupancy grid maps with less uncertainty for the same set of measurements, and these are directly illustrated by numerical examples and preliminary experimental results. While this paper is focused on the mapping problem in the two-dimensional environments, the presented results can be certainly utilized in SLAM or stochastic motion planning in three dimensions.

%occupancy probabilities that are more accurate than prior work, which may be applied for more accurate real-time map generation in the SLAM problem and more effective trajectory planning strategies based on map uncertainty.

The paper is organized as follows.
The problem is formulated in Section \ref{sec:ProbForm}, and the exact solution to the occupancy grid mapping problem is derived in Section \ref{sec:ISM}.
A numerical result shows the improved accuracy in Section \ref{sec:NumRes}, and an experimental result with the Kinect depth sensor shows the improvement in map certainty in Section \ref{sec:ExpRes}, followed by conclusions.

\section{Problem Formulation}
\label{sec:ProbForm}

\subsection{Range Sensor}

Consider a range sensor that provides scans of surrounding maps to identify the closest occupied space. The location and the direction of the sensor, referred to as \emph{pose} at time $t$ is denoted by $X_t$. The range sensor returns a measurement \emph{scan} $Z_t$, which is composed of $n_z$ measurement \emph{rays} $z_{t,1},z_{t,2},...,z_{t,n_z}$. The $l$-th measurement ray for $l\in\braces{1,2,...,n_z}$ contains the depth (range) and bearing (direction) of a small region inside the sensor field of view (FOV) that corresponds to a circular sector.

The probability density function, namely $p(z_{t,l}|m,X_t)$ with respect to depth of the $l$-th measurement ray conditioned on the map $m$ and the pose $X_t$ is commonly referred to as a \emph{forward sensor model}, which characterizes the corresponding depth sensor, such as the maximum range or accuracy. The forward sensor model satisfies  ($i$) the ranges of all depth measurements are positive and finite, and ($ii$) measurement rays cannot pass through occupied regions.

Throughout this paper, it is assumed that the forward sensor model of the selected sensor is given. This can be determined empirically or analytically. For example, the \emph{beam model for range finders} satisfying  the above criteria is described in~\cite{ThrBurFox05}.
%, which contains Gaussian, exponential, uniform, and point pass portions~\cite{ThrBurFox05}.
%At time $t$ and given the pose $X_t$ and map $m$, the beam model with respect to the $l$-th measurement ray $z_{t,l}$ is the probability density defined as
%\begin{align}
%p(z_{t,l})=\begin{cases}
%\frac1{z_{max}-z_{min}}P_{hit}+P_{short}+P_{rand}\ \text{if},\ 
%\\
%P_{max},
%\end{cases}
%\end{align}
%where $z_{min}\leq z_{t,l}\leq z_{max}$ and $\hat z_{t,l}$ is the distance between the robot and the closest occupied location along the direction $z_{t,l}$.
%The probabilities are defined such that $P_{hit}$ is the probability of the depth originating from this occupied location (Gaussian), $P_{exp}$ is the probability that an unexpected occlusion blocks the first occupied location of the map (exponential), $P_{rand}$ is the probability of a phantom reading (uniform), and $P_{max}$ is the probability that the sensor returns its maximum value (point mass).
%These probabilities are normalized such that
%\begin{align}
%P_{hit}+P_{short}+P_{rand}+P_{max}=1.
%\end{align}

%We assume that this forward sensor model depends on the closest occupied location, and is independent of the occupancy beyond this point.

\subsection{Occupancy Grid Mapping}

Let a map $m$ be decomposed into $n_m$ evenly-spaced grid cells, where the $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,...,n_m}$, that is defined as $\mathbf{m}_i=1$ when it is occupied, and $\mathbf{m}_i=0$ when it is free. Another binary random variable is defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience.
Therefore, the probability that the $i$-th cell is occupied is $P(\mathbf{m}_i)$, and the probability that it is free is $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$. The location and size of each grid cell are assumed to be known.
%, but its occupancy is uncertain; the occupancy is treated as a binary random variable such that $P(\mathbf{m}_i)$ is the probability that $\mathbf{m}_i$ is occupied and $P(\bar{\mathbf{m}}_i)$ is the probability that $\mathbf{m}_i$ is free, such that $P(\mathbf{m}_i)+P(\bar{\mathbf{m}}_i)=1$.
The occupancies of different cells are assumed independent such that
\begin{align}
P(m)=P(\mathbf{m}_1,\mathbf{m}_2,...,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i).
\end{align}

%Occupancy grid mapping produces a probabilistic map; none of the $n_m$ grid cell occupancies are completely certain.
Note that there are $2^{n_{m}}$ possible maps. The occupancy grid mapping problem is to determine these probabilities based on robot poses, measurement scans, and the forward sensor model. More explicitly, let $X_{1:t}$ denotes the history of poses from the initial time to the current time, $\{X_1,X_2,\ldots, X_t\}$, and let $Z_{1:t}$ be the measurement history. Therefore, the goal is to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, commonly referred to as the \emph{inverse sensor model}, for all grid cells.

	% Def of occupancy grid
	% Cell occupancies are assumed independent
	% Probabilistic maps
	% Notations
	% Mapping problem

\subsection{Mapping via Log-Odds Ratio}

One of the popular frameworks of updating binary random variables with static state is with a binary Bayesian filter using the log-odds ratio formulation.
The main idea is that instead of multiplying terms from prior and current time steps, the properties of logarithms allow these terms to be simply added, while avoiding truncation issues associated with probabilities close to $0$ or $1$~\cite{ThrBurFox05}.
The log-odds ratio is also popular because learning techniques with forward models~\cite{Thr01},~\cite{Thr03} or ad hoc techniques~\cite{MorElf85},~\cite{Elf89} are simpler to derive when they need not consider $P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})$ as part of the inverse sensor model, but rather only $P(\mathbf{m}_i)$, which is typically fixed and uniform for all cells.

However, the log-odds ratio formulation makes an assumption that is not consistent with the occupancy grid mapping problem. Early on, the derivation assumes that if $X_t$ and $\mathbf{m}_i$ are known, then $Z_t$ is independent of $X_{1:t-1}$ and $Z_{1:t-1}$, i.e.,
%\begin{align}
%\label{eqn:FirstBayesRule}
%P(\mathbf{m}_i|X_{1:t},Z_{1:t})
%&
%=\frac{P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})}{P(Z_t|X_{1:t},Z_{1:t-1})}.
%\end{align}
\begin{align}
\label{eqn:AssumptionEarly}
P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})\approx P(Z_t|\mathbf{m}_i,X_t).
\end{align}
The above equation can be justified if both sides are conditioned by the complete map instead of $\mathbf{m}_i$, since the probability distribution of the sensor is completely determined by its pose and the map. However, since it is conditioned only by the $i$-th cell, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may have certain information about the occupancy of the other cells, which should be considered in calculating the probability distribution of the measurement $Z_t$.  In short, this log-odds framework is based on the assumption that neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$.

%However, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may suggest that there exist occupied regions that occlude some or all of $Z_t$ from detecting $\mathbf{m}_i$ from the current pose $X_t$.
%Since this assumption neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$, this framework is subject to poor results.
%Hence, the log-odds ratio formulation is completely avoided in further analysis.

	% approximated inverse sensor model
	% markov assumption
	% shortcoming, reason this is avoided

\section{Exact Inverse Sensor Model}
\label{sec:ISM}

Due to the computational complexity, approximate inverse sensor models have been used for occupancy grid mapping with simplifying assumptions. In this section, we propose an algorithm to compute the exact inverse sensor model efficiently. 
%Much like the forward sensor model that reasons measurements from a known map, he inverse sensor model reasons that map from known measurements~\cite{ThrBurFox05}.
While the inverse sensor model varies slightly in meaning in numerous publications, two versions are derived in this paper, both based on a Bayesian framework.
The first type is referred to as the \emph{ray inverse sensor model} $P(\mathbf{m}_i|X_{1:t},Z_{1:t-1},z_{t,l})$, which only considers $l$-th ray at time $t$.
The second type is referred to as the \emph{scan inverse sensor model} $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, which combines the information from \emph{all} measurement rays \emph{simultaneously} at the $t$-th time step.

These are contrary to inverse sensor models commonly used with the log-odds ratio formulations, and the proposed algorithm integrates all of past poses and measurement scans with the current ones to obtain more accurate estimate of the map.

\subsection{Ray Inverse Sensor Model}

Suppose the probability of the map conditioned on the past poses and measurements, namely $P(m|X_{1:t-1},Z_{1:t-1})$ are is. Here we construct a posteriori probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$, based on the current pose $X_t$, a measurement ray $z_{t,l}$, and the given forward sensor model $p(z_{t_l}|m,X_t)$.

\paragraph{Bayesian Framework}
Assuming that the probability density of the sensor measurement is fixed over a small measurement region $\Delta z$, Bayes' rule yields
%First, Bayes' rule is applied for a particular map $m$ and arbitrarily small measurement region $\Delta z$, yielding
\begin{align}
\label{eqn:BayesRuleRayISM}
P(&m|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\frac{p(z_{t,l}|m,X_{1:t},Z_{1:t-1})\Delta zP(m|X_{1:t-1},Z_{1:t-1})}{P(z_{t,l}|X_{1:t},Z_{1:t-1})},
\end{align}
where we have used the fact that $X_t$ carries no information about $m$ without $Z_t$.
If the current pose $X_t$ and map $m$ are known, then the measurement ray $z_{t,l}$ is independent of past poses $X_{1:t-1}$ and past measurements $Z_{1:t-1}$.
%Note that this is contrary to approaches that consider $X_t$ and $\mathbf{m}_i$ (not the full map $m$) as enough information, ignoring the occupancy of other cells in this term.
Then, \refeqn{BayesRuleRayISM} can be rewritten as
\begin{align}
P(m|z_{t,l},&X_{1:t},Z_{1:t-1})\nonumber
\\
&
=\eta_{t,l}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}),
\end{align}
where the normalizing constant $\eta_{t,l}=\frac{\Delta z}{P(z_{t,l}|X_{1:t},Z_{1:t-1})}\in\Re$ absorbs all terms independent of the map $m$.

%Assuming $z_{t,l}$ measures an occupied cell, then $z_{t,l}$ may be a result of the occupancy of $\mathbf{m}_i$, or $z_{t,l}$ may measure a different occupied cell.

Next, we compute the occupancy probability of each cell. Let $\mathcal{M}_i$ be the set of maps where the $i$-th cell is occupied, i.e., $\mathcal{M}_i =\{m\in\{0,1\}^{2_{n_m}}\,|\ \mathbf{m}_i=1\}$. To compute the probability of occupancy of the $i$-th cell, all possible combinations of map in $\mathcal{M}_i$ should be considered, i.e., 
%All possible combinations of map $m$ must be considered in which $\mathbf{m}_i$ is occupied, denoted as $m:m(i)=\mathbf{m}_i$,
\begin{align}
\label{eqn:InvSenModWithProbDens}
&P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\eta_{t,l}\sum_{m\in\mathcal{M}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
\end{align}
Furthermore, to determine the normalizing constant $\eta_{t,l}$, its its complement $P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in the similar manner. Since there are $2^{n_{m}-1}$ maps in $\mathcal{M}_i$, it appears that these equations require additions $2^{n_m}$ times for the $i$-th grid cell, and it should be repeated for other cells.
%However, this is not the case when the bearing of $z_{t,l}$ is known. (\EditTL{Not sure why bearing matters})

\paragraph{Computationally Efficient Approach}
Define a reduced map $r$ where the $k$-th cell is denoted $\mathbf{r}_k$ such that $k\in\braces{1,2,...,n_r}$ with $n_r\leq n_m$.
This reduced map is chosen such that each cell of $r$ corresponds to a grid cell of $m$ that $z_{t,l}$ intersects, ordered by increasing distance, which is easily determined with geometry.
Then let the term $\mathbf{r}_{k+}$ mean that the $k$-th cell is occupied, cells with lower index are free, and cells with greater index may or may not be occupied; put differently, the $k$-th cell $\mathbf{r}_k$ is the closest occupied cell to current pose $X_t$ along reduced map $r$.
This means that the forward sensor model $p(z_{t,l}|\mathbf{r}_{k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell, and \emph{no other cells}.

\begin{prop}
Let the $k$-th unnormalized probability of the inverse sensor model be defined as
\begin{align}
\label{eqn:Unnormalized}
\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=\frac1{\eta_{t,l}}P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\&=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
%\\
%\label{eqn:UnnormalizedCompl}
%\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\frac1{\eta_{t,l}}P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})
,
\end{align}
where the normalizer $\eta_{t,l}$ is defined as
\begin{align}
\label{eqn:allEta}
\eta_{t,l}
&=
\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1}.
\end{align}
Then, the ray inverse sensor model is computed simply as
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}).
\end{align}
\begin{proof}
See the Appendix 1.
\end{proof}
\end{prop}

%
%%respectively.
%These have the follow solutions that may be obtained in a computationally-efficient manner:
%\begin{align}
%\label{eqn:tildePresult}
%\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
%\nonumber
%\\
%&\quad
%+
%\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{k+},X_t)P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1}),
%\\
%\label{eqn:tildePcomplementresult}
%\tilde P(\bar{\mathbf{r}}_k|z_{t,l},&X_{1:t},Z_{1:t-1})
%=P(\bar{\mathbf{r}}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
%\nonumber
%\\
%&\quad
%+
%\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}).
%\end{align}

%
%
%Because $\mathbf{r}_k$ is binary, the following holds:
%\begin{align}
%\label{eqn:allEta}
%\eta_{t,l}&=\frac1{\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})+\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})}\nonumber
%\\
%&=
%\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
%\end{align}
%where $\eta_{t,l}$ is independent of the index $k$ of the reduced map.
%Rearranging \refeqn{Unnormalized} and \refeqn{UnnormalizedCompl}, the ray inverse sensor model is 
%\begin{align}
%\label{eqn:RayISMAnswer}
%P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
%\\
%P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
%\end{align}
%respectively, where $\eta_{t,l}$ comes from \refeqn{allEta} and the unnormalized probabilities come from \refeqn{tildePresult} and \refeqn{tildePcomplementresult}.

Compared with the ray inverse sensor model described with \refeqn{InvSenModWithProbDens} considering $n_r$ possible cells, where the summation should be repeated $2^{n_r}$ times \emph{per cell of the reduced map}, the proposed expressions  \refeqn{allEta} and \refeqn{RayISMAnswer} are \textit{substantially} simpler.
In fact, if the terms inside the summations of \refeqn{Unnormalized} are obtained recursively, only $n_r$ repetitions are required for \emph{all cells of the reduced map}, yielding an algorithm that is up to $\mathbf{2^{n_r}}$ \textbf{times faster}.

\subsection{Scan Inverse Sensor Model}

The goal of occupancy grid mapping is to obtain the probabilistic map based on all measurements and poses, including the latest measurement scan.
At the $t$-th time step, the mesauremenet scan is the set of all rays $Z_t=\{z_{t,1},z_{t,2},...,z_{t,n_z}\}$, which serve to update the map occupancy probabilities in a unified way.
\paragraph{Bayesian Framework} Assuming the measurement rays are mutually independent, Bayes' rule yields
\begin{align}
\label{eqn:ManyMeasBayesRule}
P&(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\
%&
%=\frac{P(z_{t,1:n_z}|\mathbf{m}_i,X_{1:t},Z_{1:t-1})P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})}{P(z_{t,1:n_z}|X_{1:t},Z_{1:t-1})}
%\nonumber
%\\
&
=
\frac{\prod_{l=1}^{n_z}\braces{P(z_{t,l}|\mathbf{m}_i,X_{1:t},Z_{1:t-1})}P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})}{P(z_{t,1:n_z}|X_{1:t},Z_{1:t-1})}.
\end{align}
Using Bayes' rule again, it can be rearranged as
\begin{align}
P(\mathbf{m}_i|X_{1:t},Z_{1:t})&=
%\frac{P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})^{1-n_z}}{P(z_{t,1:n_z}|X_{1:t},Z_{1:t-1})}
%\nonumber\\&\quad\times
%\prod_{l=1}^{n_z}\left\{\frac{P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})P(z_{t,l}|X_{1:t},Z_{1:t-1})P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})}\right\}
%\nonumber
%\\
%&=
\zeta_i P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})^{1-n_z}\nonumber\\&\quad\times
\prod_{l=1}^{n_z}\left\{P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\right\},
\label{eqn:ThirdBayesRule}
\end{align}
considering that $X_t$ carries no information about $\mathbf{m}_i$ without $Z_t$, and where the normalizer $\zeta_i = \frac{\prod_{l=1}^{n_z} P(z_{t,l}|X_{1:t},Z_{1:t-1})}{P(z_{t,1:n_z}|X_{1:t},Z_{1:t-1})}$.
%\refeqn{ThirdBayesRule} is simplified to 
%\begin{align}
%P(\mathbf{m}_i|X_{1:t},Z_{1:t})&=
%\zeta_i P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})^{1-n_z}\nonumber\\&\quad\times\prod_{l=1}^{n_z}\left\{P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\right\}.
%\label{eqn:FourthBayesRule}
%\end{align}

% and we assume $Z_t$ is independent of $X_{1:t-1}$ and $Z_{1:t-1}$.
% and $P(\mathbf{m}_i|X_{1:t},z_{t,l},Z_{1:t-1})=P(\mathbf{m}_i|X_t,z_{t,l})$.

%Similarly, the complement is
%\begin{align}
%\label{eqn:CompBayesProbDef}
%P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})&=
%\zeta_i P(\bar{\mathbf{m}}_i|X_{1:t-1},Z_{1:t-1})^{1-n_z}\prod_{l=1}^{n_z}\left\{P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})\right\}
%\nonumber
%\\
%&=\zeta_i (1-P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1}))^{1-n_z}\prod_{l=1}^{n_z}\left\{(1-P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1}))\right\}.
%\end{align}
%Because $\mathbf{m}_i$ is a binary random variable,
%\begin{align}
%\label{eqn:BinNorm}
%P(\mathbf{m}_i|X_{1:t},Z_{1:t})+P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})=1,
%\end{align}
%so \refeqn{FourthBayesRule} and \refeqn{CompBayesProbDef} can be substituted into \refeqn{BinNorm} to solve for $\zeta_i$,
%\begin{align}
%\label{eqn:xi}
%\zeta_i&=
%\bigg[
%P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})^{1-n_z}\prod_{l=1}^{n_z}\left\{P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\right\}
%\nonumber\\&\qquad
%+
%(1-P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1}))^{1-n_z}\prod_{l=1}^{n_z}\left\{(1-P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1}))\right\}
%\bigg]^{-1}.
%\end{align}

% Part 2...

Consider two cases with the ray inverse sensor model $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})$: either measurement ray $z_{t,l}$ intersects with the cell, in which case we have $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})=P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=\eta_{t,l}\tilde P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$, or measurement ray $l$ misses $\mathbf{m}_i$, in which case we have $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})=P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})$ because $X_t$ holds no information about $\mathbf{m}_i$ without $z_{t,l}$.
The $i$-th grid cell $\mathbf{m}_i$ is measured by $n_{i,l}$ rays belonging to the set $\mathcal L_i$, where $0\leq n_{i,l}\leq n_z$.
Then, \refeqn{ThirdBayesRule} is simplified to
\begin{align}
P(\mathbf{m}_i|X_{1:t},Z_{1:t})
%&=
%\zeta_i P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})^{1-n_z}\prod_{l=1}^{n_z}\left\{P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\right\}
%\nonumber\\&
%=\zeta_i P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})^{1-n_{i,l}}\prod_{\mathcal L_i}\left\{\eta_{t,l}\tilde P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\right\}
%\nonumber\\
&=\xi_i P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})\nonumber\\&\quad\times
\prod_{\mathcal L_i}
\hat P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
,
\label{eqn:ISM_Fusion}
\end{align}
where the normalizer $\xi_i$ is composed of the product of normalizers $\zeta_i$ and $\eta_i$ and the term $\hat P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ is developed for convenience because it requires less computation than $\tilde P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$,
\begin{align}
\hat P(&\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
\nonumber\\&=\frac{\tilde P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})}
\nonumber\\&=
\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
\nonumber\\&\quad
+
\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}p(z_{t,l}|\mathbf{r}_{k+},X_t).
\end{align}
Similarly, we have
\begin{align}
P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})
&=\xi_i P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t-1})
\nonumber\\&\quad\times\prod_{\mathcal L_i}
\hat P(\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
,
\label{eqn:ISM_Bar_Fusion}
\end{align}
where $\hat P(\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ is defined as
\begin{align}
\hat P(&\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\
&=\frac{\tilde P(\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t-1})}
\nonumber\\
&=\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
\nonumber
\\
&\quad
+
\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\qquad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}).
\end{align}
Then, because the occupancy of $\mathbf{m}_i$ is binary, the following holds,
\begin{align}
P(\mathbf{m}_i|X_{1:t},Z_{1:t})+P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})=1
\end{align}
so $\xi_i$ is obtained using \refeqn{ISM_Fusion} and \refeqn{ISM_Bar_Fusion},
\begin{align}
\xi_i&=
\bigg[
P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})
\prod_{\mathcal L_i}
\hat P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
\nonumber\\&\quad
+
P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t-1})
\prod_{\mathcal L_i}
\hat P(\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
\bigg]^{-1},
\end{align}
which is substituted into \refeqn{ISM_Fusion} to obtain the complete scan inverse sensor model $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$.

There are two key aspects of $\xi_i$ that make the variable valuable to the occupancy grid mapping problem.
The first is that $\xi_i$ holds important information about probability terms that are otherwise difficult or impossible to calculate accurately.
The second is that since $\xi_i$ is the same for both $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$ and $P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})$, which are based off of products, then there is no need to associate unnormalized probabilities with specific measurement rays; put differently, if individual measurement ray normalizers are \emph{not} equal, this difference has no affect (such as weighting) when computing $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, thereby decreasing the necessary memory for the algorithm.

\subsection{Summary of Algorithm}

While the derivation of the complete inverse sensor model is complicated, the algorithm is quite straightforward.
Temporary variables,
\begin{align*}
a_k&=\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{t,l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{t,l,i},X_t)P(\mathbf{r}_{t,l,i}|X_{1:t-1},Z_{1:t-1}),
\\
b_k&=\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{t,l,j}|X_{1:t-1},Z_{1:t-1}),
\nonumber\\
c_k&=\sum_{i=k+1}^{n_{r,l}}\left\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_i,X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}),
\end{align*}
serve to simplify the algorithm such that repeated calculations are removed, and updated occupancy probabilities are obtained recursively.
The pseudocode is summarized in Table \ref{tab:Alg_ISM_2D}.

%\begin{table}
%\begin{algorithm}[H]
% \KwData{Robot pose, all measurements from a single scan, prior occupancy probabilities}
% \KwResult{Probability of occupancy conditioned on the robot state and the measurements}
%\For{$l = 1,2,...,n_z$}{
%	Obtain temporary map $r_{t,l}$ of $n_r$ ordered cells (close to far) from $m$ inside the FOV\\
%	Define/redefine temporary variables: 
%	$P(\mathbf{r}_{t,l,0}|X_{1:t-1},Z_{1:t-1})=0$, $P(\bar{\mathbf{r}}_{t,l,0}|X_{1:t-1},Z_{1:t-1})=1$\;
%	\For{$k = 1,2,...,n_r$}{
%		Obtain $p(z_l|x,\mathbf{r}_k)$ from the beam model\;
%		\eIf{$k=1$}{
%			$a_1=0$, $b_1=1$\;
%		}
%		{
%		$a_k=a_{k-1}+b_{k-1}p(z_{t,l}|\mathbf{r}_{t,l,k-1},X_t)P(\mathbf{r}_{t,l,k-1}|X_{1:t-1},Z_{1:t-1})$\;
%		$b_k=b_{k-1}P(\bar{\mathbf{r}}_{t,l,k-1}|X_{1:t-1},Z_{1:t-1})$\;
%		}
%	}
%	$c_{n_r}=0$\;
%	\For{$k = n_r-1,n_r-2,...,1$}{
%		$c_k=c_{k+1}+b_{k}p(z_{t,l}|\mathbf{r}_{t,l,k+1},X_t)P(\mathbf{r}_{t,l,k+1}|X_{1:t-1},Z_{1:t-1})$\;
%	}
%	\For{$k = 1,2,...,n_r$}{
%		$\hat P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=a_k+b_kp(z_{t,l}|\mathbf{r}_{t,l,k},X_t)$\;
%		$\hat P(\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=a_k+c_k$\;
%	}
%}
%\For{i =\normalfont all cells inside the FOV of the $n_z$ measurement rays}{
%	Obtain the set $\mathcal L_i$ of $l_i$ measurement rays that may pass through this cell\\
%	\eIf{$l_i>0$}{
%		$d=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})\prod_{\mathcal L_i}\hat P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$\;
%		$e=P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t-1})\prod_{\mathcal L_i}\hat P(\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$\;
%		$P(\mathbf{m}_i|X_{1:t},Z_{1:t})=\frac{d}{d+e}$\;
%	}{
%	$P(\mathbf{m}_i|X_{1:t},Z_{1:t})=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})$\;
%	}
%}
%\Return $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$ for all $i\in n_m$
%\end{algorithm}
%\caption{Fast Algorithm of the Exact Solution of the Inverse Sensor Model for $l$ Simultaneous Measurement Rays}
%\label{tab:Alg_ISM_2D}
%\end{table}

	% Mathematical computation order of complexity
	% Temporary variables: limited number to avoid repeated calculations

\section{Numerical Example}
\label{sec:NumRes}

	% Approximation/Kinect OGM algorithm for comparison
	% Forward/backward figure 8

\section{Preliminary Experimental Results}
\label{sec:ExpRes}
	% Kinect result

\section{Conclusions}

\bibliography{BibSources}
\bibliographystyle{IEEEtran}

\end{document}

\end{document}

### 2018-06-09 Conference/ACC 2016/ACC16.1 Old Version 2/Old Files and Conflicts/ACC16.1[Conflict].tex
added Conference folder

--- 2018-06-09 Conference/ACC 2016/ACC16.1 Old Version 2/ACC16.1 (1)[Conflict].tex
@@
 \thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 {\tt \{evankaufman,tylee\}@gwu.edu}}
-\thanks{This research has been supported in part by NRL under the grant *********, and NSF under the grants CMMI-1243000, CMMI-1335008, and CNS-1337722.}
+\thanks{\textsuperscript{\footnotesize\ensuremath{*}}This research has been supported in part by NSF under the grants CMMI-1243000, CMMI-1335008, and CNS-1337722.}
 }
@@
 \newcommand{\EditTL}[1]{{\color{red}\protect #1}}
-%\renewcommand{\EditTL}[1]{{\protect #1}}
+\renewcommand{\EditTL}[1]{{\protect #1}}
 
@@
 \begin{abstract}
-Occupancy grid maps are spatial representations of environments, where the space of interest is decomposed into a number of cells that are considered either occupied or free.
+Occupancy grid maps are spatial representations of environments, where the space of interest is decomposed into a number of cells.
 This paper is focused on occupancy grid mapping, which is to estimate the probability of occupancy for each cell based on range measurements from a known location.
-For a given probabilistic model of a range sensor, we propose a computationally efficient method to obtain an exact inverse sensor model, and it is utilized to construct a probabilisitic mapping algorithm according to the Bayesian framework.
-Compared with the existing occupancy grid mapping techniques that rely on approximate inverse sensor models, the proposed approach yields substantially more accurate maps for the same set of measurements.
+For a given probabilistic model of a range sensor, we propose a computationally efficient method to obtain an exact inverse sensor model, and it is utilized to construct a probabilisitic mapping algorithm according to a Bayesian framework.
+Compared with the existing occupancy grid mapping techniques that rely on simplifying assumptions, the proposed approach yields substantially more accurate maps for the same set of measurements.
 These are illustrated by numerical examples and experiments.
@@
 
-Mapping in the context of robotics is to generate a representation of the environment around the robot.
-This problem is crucial to understanding the surrounding regions when a robot enters an uncertain environment
-Various map representations have been applied to the mapping problem such as occupancy grids~\cite{WolSuk05}, octomaps~\cite{WurHorBenStaBur10}, or with distinguishable features in a landmark representation~\cite{MonThrKolWeg02}.
-To build maps in real time, a computationally-efficient algorithm is required.% \EditTL{rewrite this paragraph for mapping, without introducing SLAM here}
+Simultaneous localization and mapping (SLAM) is crucial for exploring uncertain environments autonomously.
+Various map representations have been applied to the SLAM problem such as occupancy grids~\cite{WolSuk05}, octomaps~\cite{WurHorBenStaBur10}, or with distinguishable features in a landmark representation~\cite{MonThrKolWeg02}.
+The SLAM problem is useful for building maps with autonomous robots in real time, so a computationally-efficient algorithm is required.
 
 This paper focusses on the popular occupancy grid representation, where the map space is decomposed into evenly-spaced grid cells that are considered as either occupied or free.
-Obtaining the probability of occupancy for a grid cell based on range measurements and the configuration of the mobile robot, known as the \emph{inverse sensor model}, is of fundamental importance with occupancy grid mapping.
-However, the exact solution to the inverse sensor model has not been utilized in occupancy grid mapping, because the computational requirements appear too cumbersome to compute the exact model in real time~\cite{ThrBurFox05}.
+Obtaining the probability of occupancy for a grid cell based on measurements and poses, known as the \emph{inverse sensor model}, is of fundamental importance with occupancy grid mapping.
+However, the exact solution to the inverse sensor model is rarely even considered because the computational requirements appear too cumbersome for the SLAM problem to handle in real time~\cite{ThrBurFox05}.
 
-Instead, several techniques have been proposed to approximate the exact inverse sensor model.
-%Due to the computational limitations, two common approaches are applied to obtain a function close to the inverse sensor model.
-In the early occupancy grid mappings based on sonar sensors, \emph{ad hoc} techniques are employed to obtain an inverse sensor model, where the bearings of depth measurements are considered uncertain because the directions of bouncing sound waves are highly inaccurate~\cite{MorElf85,Elf89}.
-%These occupancy grid mapping schemes employ \emph{ad hoc} techniques to obtain the inverse sensor model.
-These techniques are based on approximate functions to represent the inverse sensor model with questionable accuracy~\cite{ChoLynHutKanBurKavThr05} and applied to more modern sensors in~\cite{And09},~\cite{PirRutBisSch11}.
+Due to the computational limitations, two common approaches are applied to obtain a function close to the inverse sensor model.
+In the early occupancy grid mapping work, which is based on sonar sensors available at the time, the bearings of depth measurements are considered uncertain because the directions of bouncing sound waves are highly inaccurate~\cite{MorElf85},~\cite{Elf89}.
+These occupancy grid mapping schemes employ \emph{ad hoc} techniques to obtain the inverse sensor model.
+These techniques are based on approximate functions to represent the inverse sensor model with questionable accuracy, which is described further detail in~\cite{ChoLynHutKanBurKavThr05} and applied to more modern sensors in~\cite{And09},~\cite{PirRutBisSch11}.
+The other popular approach to obtain the inverse sensor model is to simulate maps, robot poses, and measurements and use a learning algorithm to obtain the inverse sensor model~\cite{Thr01},~\cite{ThrBurFox05}.
+These approaches are undesirable because of their complexity and they only yield probabilistic approximations without strong mathematical justification.
+These approaches simply treat occlusions with discrete probabilities of detection, regardless of whether one cell may completely, partially, or not block another cell in the sensor field of view.
+Furthermore, both approaches follow a log-odds ratio Bayesian framework, imposing an additional Markov assumption.
 
-The other popular approach to obtain an approximate inverse sensor model is to simulate maps, robot poses, and measurements and use a learning algorithm, such as neural network to approximate the inverse sensor model~\cite{Thr01,ThrBurFox05}.
-These approaches are undesirable in practice due to complexities associated to implementing a learning algorithm.
-For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples or how many samples are required to obtain a reasonable approximation.
-Furthermore, it is challenging to apply any learning algorithm over the large dimensional space composed of maps, poses, and measurements.
-
-%and they only yield probabilistic approximations without strong mathematical justification.
-
-%These approaches simply treat occlusions with discrete probabilities of detection, regardless of whether one cell may completely, partially, or not block another cell in the sensor field of view.
-
-%Furthermore, both approaches follow a log-odds ratio Bayesian framework, imposing an additional Markov assumption.
-
-%These techniques are not also suitable for modern range sensor with the highly accurate bearing~\cite{Thr03}.
-Based on recent improvements in sensor technology, the occupancy grid mapping problem has been restructured using the fact that depth measurement bearings are known.
-\EditTL{do we need this paragraph?}
-
-This paper proposes a computationally efficient algorithm to construct the exact inverse sensor model.
-More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, this algorithm yields a posteriori probability of occupancy of all of the cells within the area covered by the range sensor from the range measurements.
-The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancy of the other cells blocked by it is irrelevant to the forward sensor model, and it is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model.
-Furthermore, the proposed approach integrates a priori probability of occupancy and multiple range measurements according to the Bayesian framework to obtain more accurate maps.
-As such, it is contrast to the existing framework based on log-odds ratios that impose an additional Markov assumption.
-
-%these are constructed
-
-In short, the main contribution of this paper is proposing the exact inverse sensor model and utilizing it into Bayesian occupancy grid mapping.
-Compared with the current mapping algorithms based on an approximate inverse sensor model, this approach yields more accurate occupancy probabilities.
-This constructs substantially more accurate occupancy grid maps with less uncertainty for the same set of measurements, and these are directly illustrated by numerical examples and preliminary experimental results.
-While this paper is focused on the mapping problem in the two-dimensional environments, the presented results can be certainly utilized in SLAM or stochastic motion planning in three dimensions.
-
-%occupancy probabilities that are more accurate than prior work, which may be applied for more accurate real-time map generation in the SLAM problem and more effective trajectory planning strategies based on map uncertainty.
+These techniques are not very applicable to modern range finders due to the highly accurate bearing of these sensors~\cite{Thr03}.
+Based on recent improvements in technology, the occupancy grid mapping problem is restructured using the assumption that depth measurement bearings are known.
+Using this and applying various probabilistic properties, a computationally-efficient solution to the inverse sensor model is derived.
+This approach yields occupancy probabilities that are more accurate than prior work, which may be applied for more accurate real-time map generation in the SLAM problem and more effective trajectory planning strategies based on map uncertainty.
 
@@
 
-Consider a range sensor that provides scans of surrounding maps to identify the closest occupied space.
-The location and the direction of the sensor, referred to as \emph{pose} at time $t$ is denoted by $X_t$.
-The range sensor returns a measurement \emph{scan} $Z_t$, which is composed of $n_z$ measurement \emph{rays} $z_{t,1},z_{t,2},...,z_{t,n_z}$.
-The $l$-th measurement ray for $l\in\braces{1,2,...,n_z}$ contains the depth (range) and bearing (direction) of a small region inside the sensor field of view (FOV) that corresponds to a circular sector.
+A range sensor provides scans of surrounding maps to identify the closest occupied space.
+At time $t$, the range sensor returns a measurement \emph{scan} $Z_t$, which is composed of $n_z$ measurement \emph{rays} $z_{t,1},z_{t,2},...,z_{t,n_z}$. 
+The $l$-th measurement ray contains the depth (range) and bearing (direction) of a small region inside the sensor field of view (FOV) where $l\in\braces{1,2,...,n_z}$.
 
-The probability density function, namely $p(z_{t,l}|m,X_t)$ with respect to depth of the $l$-th measurement ray conditioned on the map $m$ and the pose $X_t$ is commonly referred to as a \emph{forward sensor model}, which characterizes the corresponding depth sensor, such as the maximum range or accuracy.
-The forward sensor model satisfies  ($i$) the ranges of all depth measurements are positive and finite, and ($ii$) measurement rays cannot pass through occupied regions.
-
-Throughout this paper, it is assumed that the forward sensor model of the selected sensor is given.
-This can be determined empirically or analytically.
-For example, the \emph{beam model for range finders} satisfying  the above criteria is described in~\cite{ThrBurFox05}.
-%, which contains Gaussian, exponential, uniform, and point pass portions~\cite{ThrBurFox05}.
+The probability density function with respect to depth of the $l$-th measurement ray conditioned on the map $m$ and robot pose $p(z_{t,l}|m,X_t)$ is commonly referred to as a \emph{forward sensor model}, which varies among depth sensors.
+The following analysis of this paper requires the forward sensor model satisfies  ($i$) the ranges of all depth measurements are positive and finite, and ($ii$) measurement rays cannot pass through occupied regions.
+The \emph{beam model for range finders} with known bearings satisfies these criteria, which contains Gaussian, exponential, uniform, and point pass portions~\cite{ThrBurFox05}.
 %At time $t$ and given the pose $X_t$ and map $m$, the beam model with respect to the $l$-th measurement ray $z_{t,l}$ is the probability density defined as
@@
 
-\subsection{Occupancy Grid Mapping}
+\subsection{Occupancy Grid}
 
-Let a map $m$ be decomposed into $n_m$ evenly-spaced grid cells, where the $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,...,n_m}$, that is defined as $\mathbf{m}_i=1$ when it is occupied, and $\mathbf{m}_i=0$ when it is free.
-Another binary random variable is defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience.
-Therefore, the probability that the $i$-th cell is occupied is $P(\mathbf{m}_i)$, and the probability that it is free is $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$.
-The location and size of each grid cell are assumed to be known.
-%, but its occupancy is uncertain; the occupancy is treated as a binary random variable such that $P(\mathbf{m}_i)$ is the probability that $\mathbf{m}_i$ is occupied and $P(\bar{\mathbf{m}}_i)$ is the probability that $\mathbf{m}_i$ is free, such that $P(\mathbf{m}_i)+P(\bar{\mathbf{m}}_i)=1$.
-The occupancies of different cells are assumed independent such that
+Let map $m$ be decomposed into $n_m$ evenly-spaced grid cells, where the $i$-th grid cell is referred to as $\mathbf{m}_i$ such that $i\in\braces{1,2,...,n_m}$.
+The location and size of each grid cell is known, but its occupancy is uncertain; the occupancy is treated as a binary random variable such that $P(\mathbf{m}_i)$ is the probability that $\mathbf{m}_i$ is occupied and $P(\bar{\mathbf{m}}_i)$ is the probability that $\mathbf{m}_i$ is free, such that $P(\mathbf{m}_i)+P(\bar{\mathbf{m}}_i)=1$.
+The occupancies of different cells are assumed independent, i.e.,
 \begin{align}
@@
 
-%Occupancy grid mapping produces a probabilistic map; none of the $n_m$ grid cell occupancies are completely certain.
-Note that there are $2^{n_{m}}$ possible maps.
-The occupancy grid mapping problem is to determine these probabilities based on robot poses, measurement scans, and the forward sensor model.
-More explicitly, let $X_{1:t}$ denotes the history of poses from the initial time to the current time, $\{X_1,X_2,\ldots, X_t\}$, and let $Z_{1:t}$ be the measurement history.
+Occupancy grid mapping produces a probabilistic map; none of the $n_m$ grid cell occupancies are completely certain.
+The mapping problem is to determine these probabilities based on robot poses and measurement scans.
+At time step $t$, the occupancy probability of the $i$-th grid cell should reflect all poses $X_{1:t}$ and measurement scans $Z_{1:t}$.
 Therefore, the goal is to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, commonly referred to as the \emph{inverse sensor model}, for all grid cells.
@@
 
-One of the popular frameworks of updating binary random variables with static state is with a binary Bayesian filter using the log-odds ratio formulation.
+One popular framework of updating binary random variables with static state is with a binary Bayesian filter using the log-odds ratio formulation.
 The main idea is that instead of multiplying terms from prior and current time steps, the properties of logarithms allow these terms to be simply added, while avoiding truncation issues associated with probabilities close to $0$ or $1$~\cite{ThrBurFox05}.
 The log-odds ratio is also popular because learning techniques with forward models~\cite{Thr01},~\cite{Thr03} or ad hoc techniques~\cite{MorElf85},~\cite{Elf89} are simpler to derive when they need not consider $P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})$ as part of the inverse sensor model, but rather only $P(\mathbf{m}_i)$, which is typically fixed and uniform for all cells.
-
 However, the log-odds ratio formulation makes an assumption that is not consistent with the occupancy grid mapping problem.
-Early on, the derivation assumes that if $X_t$ and $\mathbf{m}_i$ are known, then $Z_t$ is independent of $X_{1:t-1}$ and $Z_{1:t-1}$, i.e.,
+Early on, the derivation assumes that if $X_t$ and $\mathbf{m}_i$ are known, then $Z_t$ is independent of $X_{1:t-1}$ and $Z_{1:t-1}$, thereby enforcing a harmful Markov assumption:
 %\begin{align}
@@
 \end{align}
-The above equation can be justified if both sides are conditioned by the complete map instead of $\mathbf{m}_i$, since the probability distribution of the sensor is completely determined by its pose and the map.
-However, since it is conditioned only by the $i$-th cell, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may have certain information about the occupancy of the other cells, which should be considered in calculating the probability distribution of the measurement $Z_t$.
-In short, this log-odds framework is based on the assumption that neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$.
-
-%However, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may suggest that there exist occupied regions that occlude some or all of $Z_t$ from detecting $\mathbf{m}_i$ from the current pose $X_t$.
-%Since this assumption neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$, this framework is subject to poor results.
-%Hence, the log-odds ratio formulation is completely avoided in further analysis.
+However, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may suggest that there exist occupied regions that occlude some or all of $Z_t$ from detecting $\mathbf{m}_i$ from the current pose $X_t$.
+Since this assumption neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$, this framework is subject to poor results.
+Hence, the log-odds ratio formulation is completely avoided in further analysis.
 
@@
 
-Due to the computational complexity, approximate inverse sensor models have been used for occupancy grid mapping with simplifying assumptions.
-In this section, we propose an algorithm to compute the exact inverse sensor model efficiently.
-%Much like the forward sensor model that reasons measurements from a known map, he inverse sensor model reasons that map from known measurements~\cite{ThrBurFox05}.
+Much like the forward sensor model that reasons measurements from a known map, he inverse sensor model reasons that map from known measurements~\cite{ThrBurFox05}.
 While the inverse sensor model varies slightly in meaning in numerous publications, two versions are derived in this paper, both based on a Bayesian framework.
@@
 The second type is referred to as the \emph{scan inverse sensor model} $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, which combines the information from \emph{all} measurement rays \emph{simultaneously} at the $t$-th time step.
-
-These are contrary to inverse sensor models commonly used with the log-odds ratio formulations, and the proposed algorithm integrates all of past poses and measurement scans with the current ones to obtain more accurate estimate of the map.
+These are contrary to inverse sensor models commonly used with the log-odds ratio formulations, which are independent of past poses and measurement scans.
 
@@
 
-Suppose the probability of the map conditioned on the past poses and measurements, namely $P(m|X_{1:t-1},Z_{1:t-1})$ are is.
-Here we construct a posteriori probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$, based on the current pose $X_t$, a measurement ray $z_{t,l}$, and the given forward sensor model $p(z_{t_l}|m,X_t)$.
-
 \paragraph{Bayesian Framework}
-Assuming that the probability density of the sensor measurement is fixed over a small measurement region $\Delta z$, Bayes' rule yields
-%First, Bayes' rule is applied for a particular map $m$ and arbitrarily small measurement region $\Delta z$, yielding
+First, Bayes' rule is applied for a particular map $m$ and arbitrarily small measurement region $\Delta z$, yielding
 \begin{align}
@@
 \end{align}
-where we have used the fact that $X_t$ carries no information about $m$ without $Z_t$.
+because $X_t$ carries no information about $m$ without $Z_t$.
 If the current pose $X_t$ and map $m$ are known, then the measurement ray $z_{t,l}$ is independent of past poses $X_{1:t-1}$ and past measurements $Z_{1:t-1}$.
@@
 \end{align}
-where the normalizing constant $\eta_{t,l}=\frac{\Delta z}{P(z_{t,l}|X_{1:t},Z_{1:t-1})}\in\Re$ absorbs all terms independent of the map $m$.
+where normalizing constant $\eta_{t,l}=\frac{\Delta z}{P(z_{t,l}|X_{1:t},Z_{1:t-1})}$ absorbs all terms independent of the map $m$.
 
-%Assuming $z_{t,l}$ measures an occupied cell, then $z_{t,l}$ may be a result of the occupancy of $\mathbf{m}_i$, or $z_{t,l}$ may measure a different occupied cell.
-
-Next, we compute the occupancy probability of each cell.
-Let $\mathcal{M}_i$ be the set of maps where the $i$-th cell is occupied, i.e., $\mathcal{M}_i =\{m\in\{0,1\}^{2_{n_m}}\,|\ \mathbf{m}_i=1\}$.
-To compute the probability of occupancy of the $i$-th cell, all possible combinations of map in $\mathcal{M}_i$ should be considered, i.e.,
-%All possible combinations of map $m$ must be considered in which $\mathbf{m}_i$ is occupied, denoted as $m:m(i)=\mathbf{m}_i$,
+Assuming $z_{t,l}$ measures an occupied cell, then $z_{t,l}$ may be a result of the occupancy of $\mathbf{m}_i$, or $z_{t,l}$ may measure a different occupied cell.
+All possible combinations of map $m$ must be considered in which $\mathbf{m}_i$ is occupied, denoted as $m:m(i)=\mathbf{m}_i$,
 \begin{align}
@@
 \\
-&=\eta_{t,l}\sum_{m\in\mathcal{M}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
+&=\eta_{t,l}\sum_{m:m(i)=\mathbf{m}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
 \end{align}
-Furthermore, to determine the normalizing constant $\eta_{t,l}$, its its complement $P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in the similar manner.
-Since there are $2^{n_{m}-1}$ maps in $\mathcal{M}_i$, it appears that these equations require additions $2^{n_m}$ times for the $i$-th grid cell, and it should be repeated for other cells.
-%However, this is not the case when the bearing of $z_{t,l}$ is known. (\EditTL{Not sure why bearing matters})
+It appears that $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ and its complement $P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in a cumbersome manor, with an order of complexity of $2^{n_m}$ for \emph{each} grid cell.
+However, this is not the case when the bearing of $z_{t,l}$ is known.
 
@@
 \begin{prop}
-Let the $k$-th unnormalized probability of the inverse sensor model be defined as
+Let the $k$-th unnormalized probability and its complement of the inverse sensor model be defined as
 \begin{align}
 \label{eqn:Unnormalized}
-\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=\frac1{\eta_{t,l}}P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\&=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
+\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\frac1{\eta_{t,l}}P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
+\\
+\label{eqn:UnnormalizedCompl}
+\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\frac1{\eta_{t,l}}P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
+\end{align}
+respectively.
+These have the follow solutions that may be obtained in a computationally-efficient manner:
+\begin{align}
+\label{eqn:tildePresult}
+\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
 &\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
 &\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
-%\\
-%\label{eqn:UnnormalizedCompl}
-%\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\frac1{\eta_{t,l}}P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})
-,
+\nonumber
+\\
+&\quad
++
+\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
+&\quad\times p(z_{t,l}|\mathbf{r}_{k+},X_t)P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1}),
+\\
+\label{eqn:tildePcomplementresult}
+\tilde P(\bar{\mathbf{r}}_k|z_{t,l},&X_{1:t},Z_{1:t-1})
+=P(\bar{\mathbf{r}}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
+&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
+&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
+\nonumber
+\\
+&\quad
++
+\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
+&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}).
 \end{align}
-where the normalizer $\eta_{t,l}$ is defined as
-\begin{align}
-\label{eqn:allEta}
-\eta_{t,l}
-&=
-\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1}.
-\end{align}
-Then, the ray inverse sensor model is computed simply as
-\begin{align}
-\label{eqn:RayISMAnswer}
-P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}).
-\end{align}
+\end{prop}
 \begin{proof}
@@
 \end{proof}
-\end{prop}
 
-%
-%%respectively.
-%These have the follow solutions that may be obtained in a computationally-efficient manner:
-%\begin{align}
-%\label{eqn:tildePresult}
-%\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
-%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
-%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
-%\nonumber
-%\\
-%&\quad
-%+
-%\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
-%&\quad\times p(z_{t,l}|\mathbf{r}_{k+},X_t)P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1}),
-%\\
-%\label{eqn:tildePcomplementresult}
-%\tilde P(\bar{\mathbf{r}}_k|z_{t,l},&X_{1:t},Z_{1:t-1})
-%=P(\bar{\mathbf{r}}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
-%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
-%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
-%\nonumber
-%\\
-%&\quad
-%+
-%\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
-%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}).
-%\end{align}
+Because $\mathbf{r}_k$ is binary, the following holds:
+\begin{align}
+\label{eqn:allEta}
+\eta_{t,l}&=\frac1{\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})+\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})}\nonumber
+\\
+&=
+\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
+\end{align}
+where $\eta_{t,l}$ is independent of the index $k$ of the reduced map.
+Rearranging \refeqn{Unnormalized} and \refeqn{UnnormalizedCompl}, the ray inverse sensor model is 
+\begin{align}
+\label{eqn:RayISMAnswer}
+P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
+\\
+P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
+\end{align}
+respectively, where $\eta_{t,l}$ comes from \refeqn{allEta} and the unnormalized probabilities come from \refeqn{tildePresult} and \refeqn{tildePcomplementresult}.
 
-%
-%
-%Because $\mathbf{r}_k$ is binary, the following holds:
-%\begin{align}
-%\label{eqn:allEta}
-%\eta_{t,l}&=\frac1{\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})+\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})}\nonumber
-%\\
-%&=
-%\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
-%\end{align}
-%where $\eta_{t,l}$ is independent of the index $k$ of the reduced map.
-%Rearranging \refeqn{Unnormalized} and \refeqn{UnnormalizedCompl}, the ray inverse sensor model is 
-%\begin{align}
-%\label{eqn:RayISMAnswer}
-%P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
-%\\
-%P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
-%\end{align}
-%respectively, where $\eta_{t,l}$ comes from \refeqn{allEta} and the unnormalized probabilities come from \refeqn{tildePresult} and \refeqn{tildePcomplementresult}.
-
-Compared with the ray inverse sensor model described with \refeqn{InvSenModWithProbDens} considering $n_r$ possible cells, where the summation should be repeated $2^{n_r}$ times \emph{per cell of the reduced map}, the proposed expressions  \refeqn{allEta} and \refeqn{RayISMAnswer} are \textit{substantially} simpler.
-In fact, if the terms inside the summations of \refeqn{Unnormalized} are obtained recursively, only $n_r$ repetitions are required for \emph{all cells of the reduced map}, yielding an algorithm that is up to $\mathbf{2^{n_r}}$ \textbf{times faster}.
+Compared with the ray inverse sensor model described with \refeqn{InvSenModWithProbDens} considering $n_r$ possible cells, where the summation should be repeated $2^{n_r}$ times \emph{per cell of the reduced map}, the proposed expressions  \refeqn{allEta}, \refeqn{tildePresult}, and \refeqn{RayISMAnswer} are \textit{substantially} simpler.
+In fact, if the terms inside the summations of \refeqn{tildePresult} are obtained recursively, only $n_r$ repetitions are required for \emph{all cells of the reduced map}, yielding an algorithm that is up to $\mathbf{2^{n_r}}$ \textbf{times faster}.
 
@@
 
-	% Approximation/Kinect OGM algorithm for comparison
-	% Forward/backward figure 8
+ It is common practice to approximate\refeqn{InvSenModWithProbDens} with a simplified function, frequently following the structure of the approximate inverse sensor model proposed in~\cite{And09}.
+The main idea of this approach is that the probability of a cell being occupied ($i$) near a measurement is high (measurement likely hits this cell), ($ii$) between the robot and the measurement is low (measurement passes through this cell), and ($iii$) beyond the measurement is unchanged (the robot cannot measure through a wall/object).
+The function is based on intuition, not thorough mathematical derivation.
+In short, this approach simplifies the inverse sensor model at the cost of accuracy.
+
+The purpose of this section is to compare the exact solution to the inverse sensor model from this paper with an approximate algorithm.
+The Kinect depth sensor is used as an example, with specifications from~\cite{KhoElb12}.
+With these parameters and the identical set of measurements, the exact solution is compared with a benchmark algorithm for a Kinect sensor taken from~\cite{PirRutBisSch11}, summarized as follows.
+The probability that the $i$-th grid cell $\mathbf{m}_i$ conditioned on the measurement ray $z_{t,l}$ (the $l$-th ray at the $t$-th time step) at the pose $X_t$ is the continuous function
+\begin{align}
+\label{eqn:ISM_Approx_1}
+&P(\mathbf{m}_i|z_{t,l},X_t)\nonumber\\&=\begin{cases}
+0.3+(\frac{k}{\sigma\sqrt{2\pi}}+0.2)e^{-\frac12\left(\frac{\hat z_{t,l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ z_{t,l}\leq \hat z_{t,l,i},
+\\
+0.5+\frac{k}{\sigma\sqrt{2\pi}}e^{-\frac12\left(\frac{\hat z_{t,l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ z_{t,l}>\hat z_{t,l,i},
+\end{cases}
+\end{align}
+which is based on the expected distance to the cell $\hat z_{t,l,i}$ with parameters $k=\sigma=0.6$.
+Then, these probabilities are combined in a weighted fashion such that all measurements rays of scan $Z_t$ simultaneously update the same grid cell in a log-odds format,
+\begin{align}
+\label{eqn:ISM_Approx_2}
+&\log\left(\frac{P(\mathbf{m}_i|Z_{t},X_t)}{1-P(\mathbf{m}_i|Z_{t},X_t)}\right)
+\nonumber\\&=
+\frac1{\sum_{z_{t,l}\in\mathbf{m}_i}\hat z_{t,l,i}}\sum_{z_{t,l}\in\mathbf{m}_i}\log\left(\frac{P(\mathbf{m}_i|z_{t,l},X_t)}{1-P(\mathbf{m}_i|z_{t,l},X_t)}\hat z_{t,l,i}\right).
+\end{align}
+
+In this numerical example, the robot follows a lissajous curve, then turns around and completes the same lissajous curve in the reverse direction.
+The same set of measurements are used with both occupancy grid mapping algorithms to obtain the map of the room.
+The occupancy probabilities are illustrated in Figure \ref{fig:NumResOccProbs}. The uncertainty of the cells, measured with map entropy defined as
+\begin{align}
+&H(P(m|X_{1:t},Z_{1:t}))=\nonumber\\&-\sum_{i=1}^n\bigg\{P(\mathbf{m}_i|X_{1:t},Z_{1:t})\log P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\&\quad+(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\log(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\bigg\},
+\end{align}
+is depicted in Figure \ref{fig:NumResOccH}.
+
+\begin{figure}[h]
+\label{fig:NumResOccProbs}
+\centerline{
+	\subfigure[Exact Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_1.pdf}
+	}
+	\subfigure[Approximate Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_1.pdf}
+	}
+}
+\centerline{
+	\subfigure[Exact Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_2.pdf}
+	}
+	\subfigure[Approximate Model, First Measurement]
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_2.pdf}
+	}
+}
+\centerline{
+	\subfigure[Exact Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_3.pdf}
+	}
+	\subfigure[Approximate Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_3.pdf}
+	}
+}
+\centerline{
+	\subfigure[Exact Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_4.pdf}
+	}
+	\subfigure[Approximate Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_4.pdf}
+	}
+}
+\centerline{
+	\subfigure[Exact Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_5.pdf}
+	}
+	\subfigure[Approximate Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_5.pdf}
+	}
+}
+\caption{The exact inverse sensor model approach proposed in this paper is compared with with an approximate algorithm designed for the Kinect depth sensor.
+The goal of the occupancy grid mapping algorithms is to update the grid cells from uncertain (grey, $0.5$ occupancy probability) to either occupied (black, $1$ occupancy probability) or free (white, $0$ occupancy probability).
+Measured cells have occupancy probabilities closer to $0$ or $1$ as more information becomes available, and unmeasured cells remain uncertain.
+The exact inverse sensor model approach generates a faster and more accurate result.}
+\end{figure}
+
+\begin{figure}[h]
+\label{fig:NumResOccH}
+\centerline{
+	\subfigure[Entropy Change (blue: exact model, red: approximation)]
+		{\includegraphics[width=0.8\columnwidth]{EntropyEvolution.png}
+	}
+}
+\centerline{
+	\subfigure[Final Exact Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_H.pdf}
+	}
+	\subfigure[Final Approximate Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_H.pdf}
+	}
+}
+\caption{Entropy serves as a measure of uncertainty of the occupancy grid, where more blue regions are more certain, and red regions are more uncertain.
+The uncertainty is always less when the exact solution derived in this paper is applied.}
+\end{figure}
+
+The occupancy grids generated by the exact inverse sensor model are yield clearer maps with less uncertainty.
+The entropy of the map from the first measurement to the last is less with the exact model, where sudden decreases correspond to sharp turns in the robot motion.
+During these turns, the field of view changes greatly, so the information gain from the measurements is large.
 
@@
 
+A preliminary experiment shows how the exact inverse sensor model is successfully implemented with the Kinect sensor.
+Again, the parameters for the Kinect are obtained from~\cite{KhoElb12}, and the OpenKinect library is used on a Mac computer.
+The Kinect sensor faces walls underneath a work desk, where the results are depicted in Figure \ref{fig:ExpRes}
+
+\begin{figure}[h]
+\label{fig:ExpRes}
+\centerline{
+	\subfigure[Test Setup]
+		{\includegraphics[height=0.3\columnwidth]{TestSetup.png}
+	}
+	\subfigure[Kinect View]
+		{\includegraphics[height=0.3\columnwidth]{KinectView.png}
+	}
+}
+\centerline{
+	\subfigure[Exact Model Probabilities]
+		{\includegraphics[height=0.4\columnwidth]{KinectDeskISM.png}
+	}
+	\hspace*{0.1\columnwidth}
+	\subfigure[Approximate Model Probabilities]
+		{\includegraphics[height=0.4\columnwidth]{KinectDeskApprox.png}
+	}
+}
+\centerline{
+	\subfigure[Exact Model Entropies]
+		{\includegraphics[height=0.4\columnwidth]{KinectDeskISM_H.png}
+	}
+	\hspace*{0.1\columnwidth}
+	\subfigure[Exact Model Entropies]
+		{\includegraphics[height=0.4\columnwidth]{KinectDeskApprox_H.png}
+	}
+}
+\caption{The Kinect captures depth measurements of the underside of a metal desk.
+The measurement scan serves to update two occupancy grid mapping schemes.
+The proposed approach, based on the the exact solution of the inverse sensor model, provides an update with substantially less uncertainty.}
+\end{figure}
+
+%The following challenges were encountered:
+%\begin{itemize}
+%\item there were some issues with saving depth measurements that are now resolved,
+%\item direct distances are not measured (instead from the plane of the sensor to the depth image pixel), and
+%\item the conversion between a $16$-bit integer to a distance appears to be nonlinear.
+%\end{itemize}
+%In short, we used the Kinect camera with a few challenges.
+%
+%\subsection{Experimental Results}
+%
+%We compare the occupancy grid mapping approach based on the exact inverse sensor model proposed for this research project with the conventional approach based on \refeqn{ISM_Approx_1} and \refeqn{ISM_Approx_2}.
+%The same measurement scan data of the lower part of a desk (Figures 1a and 1b) is used for both algorithms.
+
+The proposed approach uses all the information from the measurement scans accurately.
+After the first scan, the free space has a substantially lower probability of occupancy.
+Furthermore, the edges of the objects captured by the Kinect are well defined with the proposed approach with an edge thickness of roughly one or two cells, but two or three cells thick with the conventional approach; of course, the Kinect cannot measure through the metal desks measured, so increases in occupancy probability beyond this edge are not necessary accurate.
+Additionally, the decrease in entropy is greater with the proposed approach at $-3487$ compared with just $-557$ with the conventional approach.
+
 \section{Conclusions}
+
+Occupancy grid mapping produces a probabilistic map of occupied space, but each probability is approximated in practice due to the perceived complications of the exact solution.
+However, this paper provides an exact solution to this complicated probability problem based using cell occlusion and assuming that the depth measurements have accurate bearing.
+The numerical and experimental results show that maps using this exact approach are substantially improved without computational barriers.
 

### 2018-06-09 Conference/ACC 2016/ACC16.1 Old Version/ACC16.1.tex
added Conference folder

--- 2018-06-09 Conference/ACC 2016/ACC16.1 Old Version 2/ACC16.1 (1)[Conflict].tex
@@
 \title{\LARGE \bf
-Bayesian Occupancy Grid Mapping via Exact Inverse Sensor Model}
+Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model}
 
@@
 
-Mapping in the context of robotics is to generate a representation of the environment around the robot.
-This problem is crucial to understanding the surrounding regions when a robot enters an uncertain environment
+Mapping in the context of robotics is to generate a map, which represents the environment in close proximity of a robot.
+This problem is crucial to understanding the surrounding regions when a robot enters an uncertain environment.
 Various map representations have been applied to the mapping problem such as occupancy grids~\cite{WolSuk05}, octomaps~\cite{WurHorBenStaBur10}, or with distinguishable features in a landmark representation~\cite{MonThrKolWeg02}.
@@
 This paper focusses on the popular occupancy grid representation, where the map space is decomposed into evenly-spaced grid cells that are considered as either occupied or free.
-Obtaining the probability of occupancy for a grid cell based on range measurements and the configuration of the mobile robot, known as the \emph{inverse sensor model}, is of fundamental importance with occupancy grid mapping.
+The \emph{inverse sensor model} is the probability of occupancy for a grid cell based on range measurements and the configuration of the mobile robot, which is of fundamental importance with occupancy grid mapping.
 However, the exact solution to the inverse sensor model has not been utilized in occupancy grid mapping, because the computational requirements appear too cumbersome to compute the exact model in real time~\cite{ThrBurFox05}.
@@
 %Due to the computational limitations, two common approaches are applied to obtain a function close to the inverse sensor model.
-In the early occupancy grid mappings based on sonar sensors, \emph{ad hoc} techniques are employed to obtain an inverse sensor model, where the bearings of depth measurements are considered uncertain because the directions of bouncing sound waves are highly inaccurate~\cite{MorElf85,Elf89}.
+In the early occupancy grid mapping applications based on sonar sensors, \emph{ad hoc} techniques are employed to obtain an inverse sensor model, where the bearings of depth measurements are considered uncertain because the directions of bouncing sound waves are highly inaccurate~\cite{MorElf85,Elf89}.
 %These occupancy grid mapping schemes employ \emph{ad hoc} techniques to obtain the inverse sensor model.
@@
 
-The other popular approach to obtain an approximate inverse sensor model is to simulate maps, robot poses, and measurements and use a learning algorithm, such as neural network to approximate the inverse sensor model~\cite{Thr01,ThrBurFox05}.
-These approaches are undesirable in practice due to complexities associated to implementing a learning algorithm.
+The other popular approach to obtain an approximate inverse sensor model is to simulate maps, robot poses, and measurements and use a learning algorithm, such as a neural network, to approximate the inverse sensor model~\cite{Thr01,ThrBurFox05}.
+These approaches are undesirable in practice due to complexities associated with implementing a learning algorithm.
 For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples or how many samples are required to obtain a reasonable approximation.
@@
 This paper proposes a computationally efficient algorithm to construct the exact inverse sensor model.
-More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, this algorithm yields a posteriori probability of occupancy of all of the cells within the area covered by the range sensor from the range measurements.
-The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancy of the other cells blocked by it is irrelevant to the forward sensor model, and it is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model.
+More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, this algorithm yields a posterior probability of occupancy of all of the cells within the area covered by the range sensor from the range measurements.
+The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancy of the other cells blocked by it is irrelevant to the forward sensor model, and this model is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model.
 Furthermore, the proposed approach integrates a priori probability of occupancy and multiple range measurements according to the Bayesian framework to obtain more accurate maps.
-As such, it is contrast to the existing framework based on log-odds ratios that impose an additional Markov assumption.
+As such, the proposed approach contrasts from the existing framework based on log-odds ratios that impose an additional Markov assumption.
 
@@
 Therefore, the probability that the $i$-th cell is occupied is $P(\mathbf{m}_i)$, and the probability that it is free is $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$.
-The location and size of each grid cell are assumed to be known.
+The location and size of each grid cell is assumed to be known.
 %, but its occupancy is uncertain; the occupancy is treated as a binary random variable such that $P(\mathbf{m}_i)$ is the probability that $\mathbf{m}_i$ is occupied and $P(\bar{\mathbf{m}}_i)$ is the probability that $\mathbf{m}_i$ is free, such that $P(\mathbf{m}_i)+P(\bar{\mathbf{m}}_i)=1$.
@@
 The above equation can be justified if both sides are conditioned by the complete map instead of $\mathbf{m}_i$, since the probability distribution of the sensor is completely determined by its pose and the map.
-However, since it is conditioned only by the $i$-th cell, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may have certain information about the occupancy of the other cells, which should be considered in calculating the probability distribution of the measurement $Z_t$.
+However, since it is conditioned only by the $i$-th cell, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may have certain information about the occupancy of the other cells, which should be considered in calculating the probability distribution of the measurement scan $Z_t$.
 In short, this log-odds framework is based on the assumption that neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$.
@@
 
-These are contrary to inverse sensor models commonly used with the log-odds ratio formulations, and the proposed algorithm integrates all of past poses and measurement scans with the current ones to obtain more accurate estimate of the map.
+%These are contrary to inverse sensor models commonly used with the log-odds ratio formulations, and the proposed algorithm integrates all of past poses and measurement scans with the current ones to obtain more accurate estimate of the map.
 
@@
 
-Suppose the probability of the map conditioned on the past poses and measurements, namely $P(m|X_{1:t-1},Z_{1:t-1})$ are is.
+Suppose the probability of the map conditioned on the past poses and measurements, namely $P(m|X_{1:t-1},Z_{1:t-1})$, is known.
 Here we construct a posteriori probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$, based on the current pose $X_t$, a measurement ray $z_{t,l}$, and the given forward sensor model $p(z_{t_l}|m,X_t)$.
@@
 \end{align}
-Furthermore, to determine the normalizing constant $\eta_{t,l}$, its its complement $P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in the similar manner.
+Furthermore, to determine the normalizing constant $\eta_{t,l}$, the complement $P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in the similar manner.
 Since there are $2^{n_{m}-1}$ maps in $\mathcal{M}_i$, it appears that these equations require additions $2^{n_m}$ times for the $i$-th grid cell, and it should be repeated for other cells.
@@
 This reduced map is chosen such that each cell of $r$ corresponds to a grid cell of $m$ that $z_{t,l}$ intersects, ordered by increasing distance, which is easily determined with geometry.
-Then let the term $\mathbf{r}_{k+}$ mean that the $k$-th cell is occupied, cells with lower index are free, and cells with greater index may or may not be occupied; put differently, the $k$-th cell $\mathbf{r}_k$ is the closest occupied cell to current pose $X_t$ along reduced map $r$.
-This means that the forward sensor model $p(z_{t,l}|\mathbf{r}_{k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell, and \emph{no other cells}.
+Then let the term $\mathbf{r}_{l,k+}$ mean that the $k$-th cell of the $l$-th ray is occupied, cells with lower index are free, and cells with greater index may or may not be occupied; put differently, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along reduced map $r$.
+This means that the forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell, and \emph{no other cells}.
 
 \begin{prop}
-Let the $k$-th unnormalized probability of the inverse sensor model be defined as
+For the $l$-th measurement ray, let the $k$-th unnormalized probability of the inverse sensor model be defined as
 \begin{align}
 \label{eqn:Unnormalized}
-\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=\frac1{\eta_{t,l}}P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\&=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
-&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
-&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
+\tilde P&(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})%=\frac1{\eta_{t,l}}P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})
+\nonumber\\&=P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\
+&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
+&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})
 %\\
@@
 %\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\frac1{\eta_{t,l}}P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})
-,
+.
+\end{align}
+Then, the ray inverse sensor model is computed simply as
+\begin{align}
+\label{eqn:RayISMAnswer}
+P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}),
 \end{align}
@@
 &=
-\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1}.
-\end{align}
-Then, the ray inverse sensor model is computed simply as
-\begin{align}
-\label{eqn:RayISMAnswer}
-P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}).
+\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1}.
 \end{align}
@@
 %\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
-%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
+%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
 %&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
@@
 %+
-%\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
+%\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
 %&\quad\times p(z_{t,l}|\mathbf{r}_{k+},X_t)P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1}),
@@
 %=P(\bar{\mathbf{r}}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
-%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
+%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
 %&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
@@
 %+
-%\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
+%\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
 %&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}).
@@
 %&=
-%\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
+%\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
 %\end{align}
@@
 Compared with the ray inverse sensor model described with \refeqn{InvSenModWithProbDens} considering $n_r$ possible cells, where the summation should be repeated $2^{n_r}$ times \emph{per cell of the reduced map}, the proposed expressions  \refeqn{allEta} and \refeqn{RayISMAnswer} are \textit{substantially} simpler.
-In fact, if the terms inside the summations of \refeqn{Unnormalized} are obtained recursively, only $n_r$ repetitions are required for \emph{all cells of the reduced map}, yielding an algorithm that is up to $\mathbf{2^{n_r}}$ \textbf{times faster}.
+In fact, if the terms inside the summations of \refeqn{Unnormalized} and \refeqn{allEta} are obtained recursively, only $n_r$ repetitions are required for \emph{all cells of the reduced map}, yielding an algorithm that is up to $\mathbf{2^{n_r}}$ \textbf{times faster}.
 
@@
 
-Consider two cases with the ray inverse sensor model $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})$: either measurement ray $z_{t,l}$ intersects with the cell, in which case we have $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})=P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=\eta_{t,l}\tilde P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$, or measurement ray $l$ misses $\mathbf{m}_i$, in which case we have $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})=P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})$ because $X_t$ holds no information about $\mathbf{m}_i$ without $z_{t,l}$.
+Consider two cases with the ray inverse sensor model $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})$: either measurement ray $z_{t,l}$ intersects with the cell, in which case we have $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})=P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$, or measurement ray $l$ misses $\mathbf{m}_i$, in which case we have $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})=P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})$ because $X_t$ holds no information about $\mathbf{m}_i$ without $z_{t,l}$.
 The $i$-th grid cell $\mathbf{m}_i$ is measured by $n_{i,l}$ rays belonging to the set $\mathcal L_i$, where $0\leq n_{i,l}\leq n_z$.
@@
 %\nonumber\\&
-%=\zeta_i P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})^{1-n_{i,l}}\prod_{\mathcal L_i}\left\{\eta_{t,l}\tilde P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\right\}
+%=\zeta_i P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})^{1-n_{i,l}}\prod_{\mathcal L_i}\left\{\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\right\}
 %\nonumber\\
@@
 \prod_{\mathcal L_i}
-\hat P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
+\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
 ,
@@
 \end{align}
-where the normalizer $\xi_i$ is composed of the product of normalizers $\zeta_i$ and $\eta_i$ and the term $\hat P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ is developed for convenience because it requires less computation than $\tilde P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$,
+where the normalizer $\xi_i$ is composed of the product of normalizers $\zeta_i$ and $\eta_i$ and the term $\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ is developed for convenience because it requires less computation than $\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$,
 \begin{align}
-\hat P(&\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
-\nonumber\\&=\frac{\tilde P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})}
+\hat P(&\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
+\nonumber\\&=\frac{\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})}
 \nonumber\\&=
-\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
+\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})
 \nonumber\\&\quad
 +
-\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}p(z_{t,l}|\mathbf{r}_{k+},X_t).
+\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}p(z_{t,l}|\mathbf{r}_{l,k+},X_t).
 \end{align}
-Similarly, we have
+Similarly, its complement is
 \begin{align}
@@
 \nonumber\\&\quad\times\prod_{\mathcal L_i}
-\hat P(\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
+\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
 ,
@@
 \end{align}
-where $\hat P(\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ is defined as
+where $\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ is defined as
 \begin{align}
-\hat P(&\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\
-&=\frac{\tilde P(\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t-1})}
+\hat P(&\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\
+&=\frac{\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t-1})}
 \nonumber\\
-&=\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
+&=\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})
 \nonumber
@@
 +
-\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\qquad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}).
+\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\qquad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}).
 \end{align}
@@
 \prod_{\mathcal L_i}
-\hat P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
+\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
 \nonumber\\&\quad
@@
 \prod_{\mathcal L_i}
-\hat P(\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
+\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
 \bigg]^{-1},
@@
 \begin{align*}
-a_k&=\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{t,l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{t,l,i},X_t)P(\mathbf{r}_{t,l,i}|X_{1:t-1},Z_{1:t-1}),
+a_k&=\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
 \\
-b_k&=\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{t,l,j}|X_{1:t-1},Z_{1:t-1}),
+b_k&=\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1}),
 \nonumber\\
-c_k&=\sum_{i=k+1}^{n_{r,l}}\left\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_i,X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}),
+c_k&=\sum_{i=k+1}^{n_{r,l}}\left\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
 \end{align*}
@@
 
-%\begin{table}
-%\begin{algorithm}[H]
-% \KwData{Robot pose, all measurements from a single scan, prior occupancy probabilities}
-% \KwResult{Probability of occupancy conditioned on the robot state and the measurements}
-%\For{$l = 1,2,...,n_z$}{
-%	Obtain temporary map $r_{t,l}$ of $n_r$ ordered cells (close to far) from $m$ inside the FOV\\
-%	Define/redefine temporary variables: 
-%	$P(\mathbf{r}_{t,l,0}|X_{1:t-1},Z_{1:t-1})=0$, $P(\bar{\mathbf{r}}_{t,l,0}|X_{1:t-1},Z_{1:t-1})=1$\;
-%	\For{$k = 1,2,...,n_r$}{
-%		Obtain $p(z_l|x,\mathbf{r}_k)$ from the beam model\;
-%		\eIf{$k=1$}{
-%			$a_1=0$, $b_1=1$\;
-%		}
-%		{
-%		$a_k=a_{k-1}+b_{k-1}p(z_{t,l}|\mathbf{r}_{t,l,k-1},X_t)P(\mathbf{r}_{t,l,k-1}|X_{1:t-1},Z_{1:t-1})$\;
-%		$b_k=b_{k-1}P(\bar{\mathbf{r}}_{t,l,k-1}|X_{1:t-1},Z_{1:t-1})$\;
-%		}
-%	}
-%	$c_{n_r}=0$\;
-%	\For{$k = n_r-1,n_r-2,...,1$}{
-%		$c_k=c_{k+1}+b_{k}p(z_{t,l}|\mathbf{r}_{t,l,k+1},X_t)P(\mathbf{r}_{t,l,k+1}|X_{1:t-1},Z_{1:t-1})$\;
-%	}
-%	\For{$k = 1,2,...,n_r$}{
-%		$\hat P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=a_k+b_kp(z_{t,l}|\mathbf{r}_{t,l,k},X_t)$\;
-%		$\hat P(\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=a_k+c_k$\;
-%	}
-%}
-%\For{i =\normalfont all cells inside the FOV of the $n_z$ measurement rays}{
-%	Obtain the set $\mathcal L_i$ of $l_i$ measurement rays that may pass through this cell\\
-%	\eIf{$l_i>0$}{
-%		$d=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})\prod_{\mathcal L_i}\hat P(\mathbf{r}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$\;
-%		$e=P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t-1})\prod_{\mathcal L_i}\hat P(\bar{\mathbf{r}}_{t,l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$\;
-%		$P(\mathbf{m}_i|X_{1:t},Z_{1:t})=\frac{d}{d+e}$\;
-%	}{
-%	$P(\mathbf{m}_i|X_{1:t},Z_{1:t})=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})$\;
-%	}
-%}
-%\Return $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$ for all $i\in n_m$
-%\end{algorithm}
-%\caption{Fast Algorithm of the Exact Solution of the Inverse Sensor Model for $l$ Simultaneous Measurement Rays}
-%\label{tab:Alg_ISM_2D}
-%\end{table}
+\begin{table}
+\begin{tabular}{ l }
+  for $l = 1,2,...,n_z$\\
+   \ \ Obtain temporary map $r_{t,l}$;\\
+   \ \ Define $P(\mathbf{r}_{l,0}|X_{1:t-1},Z_{1:t-1})=0$, $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=1$;\\
+   \ \ for $k = 1,2,...,n_r$\\
+   \ \ \ \ Obtain $p(z_{t,l}|X_t,\mathbf{r}_{l,k})$ from a forward sensor model;\\
+   \ \ \ \ if $k=1$\\
+   \ \ \ \ \ \ $a_1=0$, $b_1=1$;\\
+   \ \ \ \ else\\
+   \ \ \ \ \ \ $a_k=a_{k-1}+b_{k-1}p(z_{t,l}|\mathbf{r}_{l,k-1},X_t)P(\mathbf{r}_{l,k-1}|X_{1:t-1},Z_{1:t-1})$;\\
+   \ \ \ \ \ \ $b_k=b_{k-1}P(\bar{\mathbf{r}}_{l,k-1}|X_{1:t-1},Z_{1:t-1})$;\\
+   \ \ \ \ end if\\
+   \ \ end for\\
+   \ \ $c_{n_r}=0$;\\
+   \ \ for $k = n_r-1,n_r-2,...,1$\\
+   \ \ \ \ $c_k=c_{k+1}+b_{k}p(z_{t,l}|\mathbf{r}_{l,k+1},X_t)P(\mathbf{r}_{l,k+1}|X_{1:t-1},Z_{1:t-1})$;\\
+   \ \ end for\\
+   \ \ for $k = 1,2,...,n_r$\\
+   \ \ \ \ $\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=a_k+b_kp(z_{t,l}|\mathbf{r}_{l,k},X_t)$;\\
+   \ \ \ \ $\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=a_k+c_k$;\\
+   \ \ end for\\
+   end for\\
+   for $i$ is the set of all cells inside the FOV\\
+   \ \ Obtain the set $\mathcal L_i$ of $l_i$ measurement rays that may pass through this cell\\
+   \ \ if $l_i>0$\\
+   \ \ \ \ $d=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})\prod_{\mathcal L_i}\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$;\\
+   \ \ \ \ $e=P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t-1})\prod_{\mathcal L_i}\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$;\\
+   \ \ else\\
+   \ \ \ \ $P(\mathbf{m}_i|X_{1:t},Z_{1:t})=\frac{d}{d+e}$;\\
+   \ \ end if\\
+   end for\\
+Return: $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$\\
+\end{tabular}
+\caption{Fast Algorithm of the Exact Solution of the Inverse Sensor Model for $l$ Simultaneous Measurement Rays}
+\label{tab:Alg_ISM_2D}
+\end{table}
 
@@
 
-	% Approximation/Kinect OGM algorithm for comparison
-	% Forward/backward figure 8
+ It is common practice to approximate \refeqn{InvSenModWithProbDens} with a simplified function, frequently following the structure of the approximate inverse sensor model proposed in~\cite{And09}.
+The main idea of this approach is that the probability of a cell being occupied ($i$) near a measurement is high (measurement likely hits this cell), ($ii$) between the robot and the measurement is low (measurement passes through this cell), and ($iii$) beyond the measurement is unchanged (the robot cannot measure through a wall/object).
+The function is based on intuition, not thorough mathematical derivation.
+In short, this approach simplifies the inverse sensor model at the cost of accuracy.
+
+The purpose of this section is to compare the exact solution to the inverse sensor model from this paper with an approximate algorithm.
+The Kinect depth sensor is used as an example, with specifications from~\cite{KhoElb12}.
+With these parameters and the identical set of measurements, the exact solution is compared with a benchmark algorithm for a Kinect sensor taken from~\cite{PirRutBisSch11}, summarized as follows.
+The probability that the $i$-th grid cell $\mathbf{m}_i$ is occupied conditioned on the measurement ray $z_{t,l}$ (the $l$-th ray at the $t$-th time step) at the pose $X_t$ is the continuous function
+\begin{align}
+\label{eqn:ISM_Approx_1}
+&P(\mathbf{m}_i|z_{t,l},X_t)\nonumber\\&=\begin{cases}
+0.3+(\frac{k}{\sigma\sqrt{2\pi}}+0.2)e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ z_{t,l}\leq \hat z_{l,i},
+\\
+0.5+\frac{k}{\sigma\sqrt{2\pi}}e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ z_{t,l}>\hat z_{l,i},
+\end{cases}
+\end{align}
+which is based on the expected distance to the cell $\hat z_{l,i}$ with parameters $k=\sigma=0.6$.
+Then, these probabilities are combined in a weighted fashion such that all measurements rays of scan $Z_t$ simultaneously update the same grid cell in a log-odds format,
+\begin{align}
+\label{eqn:ISM_Approx_2}
+&\log\left(\frac{P(\mathbf{m}_i|Z_{t},X_t)}{1-P(\mathbf{m}_i|Z_{t},X_t)}\right)
+\nonumber\\&=
+\frac1{\sum_{z_{t,l}\in\mathbf{m}_i}\hat z_{l,i}}\sum_{z_{t,l}\in\mathbf{m}_i}\log\left(\frac{P(\mathbf{m}_i|z_{t,l},X_t)}{1-P(\mathbf{m}_i|z_{t,l},X_t)}\hat z_{l,i}\right).
+\end{align}
+
+\paragraph{Computation Time}
+The proposed ray inverse sensor model is designed to maintain the accuracy of the exact solution with substantial computational improvements.
+A $1$-dimensional example is simulated on a Mac desktop computer with $11$ grid cells, $10$ of which are visible to the robot, where the same depth measurement is used for all occupancy grid algorithms.
+The computational time of the exact solution is computed in the conventional manner according to \refeqn{InvSenModWithProbDens} in $2.3314$ seconds to obtain the same solution of \refeqn{RayISMAnswer} from \refeqn{Unnormalized} and \refeqn{allEta} in just $0.0027$ seconds, which is $\mathbf{874.77}$ times faster.
+The ray inverse sensor model from \refeqn{ISM_Approx_1} completes in $0.0011$ seconds, but this algorithm does not follow a forward sensor model, so the accuracy is inaccurate.
+
+\paragraph{Trajectory Map Building}
+In this numerical example, the robot follows a lissajous curve, then turns around and completes the same lissajous curve in the reverse direction.
+The same set of measurements are used with both occupancy grid mapping algorithms to obtain the map of the room.
+The occupancy probabilities are illustrated in Figure \ref{fig:NumResOccProbs}. The uncertainties of the cells, measured with map entropy defined as
+\begin{align}
+&H(P(m|X_{1:t},Z_{1:t}))=\nonumber\\&-\sum_{i=1}^n\bigg\{P(\mathbf{m}_i|X_{1:t},Z_{1:t})\log P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\&\quad+(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\log(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\bigg\},
+\end{align}
+are depicted in Figure \ref{fig:NumResOccH}.
+
+\begin{figure}[h]
+\label{fig:NumResOccProbs}
+\centerline{
+	\subfigure[Exact Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_1.pdf}
+	}
+	\subfigure[Approximate Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_1.pdf}
+	}
+}
+\centerline{
+	\subfigure[Exact Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_2.pdf}
+	}
+	\subfigure[Approximate Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_2.pdf}
+	}
+}
+\centerline{
+	\subfigure[Exact Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_3.pdf}
+	}
+	\subfigure[Approximate Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_3.pdf}
+	}
+}
+\centerline{
+	\subfigure[Exact Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_4.pdf}
+	}
+	\subfigure[Approximate Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_4.pdf}
+	}
+}
+\centerline{
+	\subfigure[Exact Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_5.pdf}
+	}
+	\subfigure[Approximate Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_5.pdf}
+	}
+}
+\caption{The exact inverse sensor model approach proposed in this paper is compared with with an approximate algorithm designed for the Kinect depth sensor.
+The goal of the occupancy grid mapping algorithms is to update the grid cells from uncertain (grey, $0.5$ occupancy probability) to either occupied (black, $1$ occupancy probability) or free (white, $0$ occupancy probability).
+Measured cells have occupancy probabilities closer to $0$ or $1$ as more information becomes available, and unmeasured cells remain near $0.5$, or uncertain.
+The exact inverse sensor model approach generates a faster and more accurate result.}
+\end{figure}
+
+\begin{figure}[h]
+\label{fig:NumResOccH}
+\centerline{
+	\subfigure[Entropy (black: exact model, green: approximation)]
+		{\includegraphics[width=0.8\columnwidth]{EntropyEvolution.png}
+	}
+}
+\centerline{
+	\subfigure[Final Exact Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_H.pdf}
+	}
+	\subfigure[Final Approximate Model]
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_H.pdf}
+	}
+}
+\caption{Entropy serves as a measure of uncertainty of the occupancy grid, where more blue regions are more certain, and red regions are more uncertain.
+The uncertainty is always less when the exact solution is applied.}
+\end{figure}
+
+The occupancy grids generated by the exact inverse sensor model yield clearer maps with less uncertainty.
+The entropy of the map from the first measurement to the last is less with the exact model, where sudden decreases correspond to sharp turns in the robot motion.
+During these turns, the field of view changes greatly, so the information gain from the measurements is large, and therefore the decrease in map uncertainty is also large.
 
@@
 
+A preliminary experiment shows how the exact inverse sensor model is successfully implemented with the Kinect sensor.
+Again, the parameters for the Kinect are obtained from~\cite{KhoElb12}, and the OpenKinect library is used on a Mac computer.
+The Kinect sensor faces walls underneath a work desk, where the results are depicted in Figure \ref{fig:ExpRes}
+
+\begin{figure}[h]
+\label{fig:ExpRes}
+\centerline{
+	\subfigure[Test Setup]
+		{\includegraphics[height=0.3\columnwidth]{TestSetup.png}
+	}
+	\subfigure[Kinect View]
+		{\includegraphics[height=0.3\columnwidth]{KinectView.png}
+	}
+}
+\centerline{
+	\subfigure[Exact Model Probabilities]
+		{\includegraphics[height=0.4\columnwidth]{KinectDeskISM.png}
+	}
+	\hspace*{0.1\columnwidth}
+	\subfigure[Approximate Model Probabilities]
+		{\includegraphics[height=0.4\columnwidth]{KinectDeskApprox.png}
+	}
+}
+\centerline{
+	\subfigure[Exact Model Entropies]
+		{\includegraphics[height=0.4\columnwidth]{KinectDeskISM_H.png}
+	}
+	\hspace*{0.1\columnwidth}
+	\subfigure[Exact Model Entropies]
+		{\includegraphics[height=0.4\columnwidth]{KinectDeskApprox_H.png}
+	}
+}
+\caption{The Kinect captures depth measurements of the underside of a metal desk.
+The measurement scan serves to update two occupancy grid mapping schemes.
+The proposed approach, based on the the exact solution of the inverse sensor model, provides an update with substantially less uncertainty than the approximate model update.}
+\end{figure}
+
+The proposed approach uses all the information from the measurement scans accurately.
+After the first scan, the free space has a substantially lower probability of occupancy.
+Furthermore, the edges of the objects captured by the Kinect are well defined with the proposed approach with an edge thickness of roughly one or two cells, but two or three cells thick with the conventional approach; of course, the Kinect cannot measure through the metal desks, so increases in occupancy probability beyond these edges are not necessary accurate.
+Additionally, the decrease in entropy is greater with the proposed approach at $-3487$ compared with just $-557$ with the conventional approach.
+
 \section{Conclusions}
+
+Occupancy grid mapping produces a probabilistic map of occupied space, but each probability is approximated in practice due to the perceived complications of the exact solution.
+However, this paper provides an exact solution to this complicated probability problem using cell occlusion in the forward sensor model.
+This yields a substantially simpler inverse sensor model, which avoids a potentially harmful Markov assumption.
+The numerical and experimental results show that maps using this exact approach are substantially improved without computational barriers.
+
+\begin{appendix}
+\label{append}
+
+\paragraph{Unnormalized Reduced Map Inverse Sensor Model}
+Consider the $l$-th measurement ray at time $t$.
+The $k$-th cell of the reduced map $r$ is denoted $\mathbf{r}_{l,k}$, where $k$ corresponds to cells from map $m$, and is ordered by increasing distance from pose $X_t$.
+Define $P(\bar{\mathbf{r}}_0)=1$ and $p(z_{t,l}|X_t,\mathbf{r}_0)=0$ for convenience. Then, the probability of occupancy of the first three cells are
+\begin{align}
+\label{eqn:tildePk1}
+&\tilde P(\mathbf{r}_{l,1}|X_t,z_{t,l})\nonumber\\&=\left\{\prod_{j=0}^0P(\bar{\mathbf{r}}_{l,j})\right\}p(z_{t,l}|X_t,\mathbf{r}_{l,1})P(\mathbf{r}_{l,1})
+,
+\\
+&\tilde P(\mathbf{r}_{l,2}|X_t,z_{t,l})\nonumber\\&=
+P(\mathbf{r}_{l,2})\left\{\prod_{j=0}^0P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,1})p(z_{t,l}|X_t,\mathbf{r}_{l,1})\nonumber\\&\quad+\left\{\prod_{j=0}^1P(\bar{\mathbf{r}}_{l,j})\right\}p(z_{t,l}|X_t,\mathbf{r}_{l,2})P(\mathbf{r}_{l,2})
+,
+\\
+&\tilde P(\mathbf{r}_{l,3}|X_t,z_{t,l})\nonumber\\&=
+P(\mathbf{r}_{l,3})\sum_{i=1}^2
+\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,i})p(z_{t,l}|\mathbf{r}_{l,i},X_t)\nonumber\\
+&+\left\{\prod_{j=0}^2P(\bar{\mathbf{r}}_{l,j})\right\}p(z_{t,l}|X_t,\mathbf{r}_{l,3})P(\mathbf{r}_{l,3})
+,
+\end{align}
+respectively. This is generalized to the arbitrary $k$-th cell as
+\begin{align}
+\label{eqn:GenSol}
+&\tilde P(\mathbf{r}_k|X,z)
+\nonumber\\&=P(\mathbf{r}_{l,k})\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i})
+\nonumber\\&\quad+
+\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j})\right\}p(z_{t,l}|\mathbf{r}_{l,k},X_t)P(\mathbf{r}_{l,k}),
+\end{align}
+where special cases when the lower bound of a summation is greater than the upper bound results in a summation value of $0$.
+
+\emph{Proof by Induction of the Reduced Inverse Sensor Model \refeqn{GenSol}}
+Consider the base case when $k=1$ such that the $k$-th cell would block all others.
+This corresponds to $\tilde P(\mathbf{r}_{l,1}|X_t,z_{t,l})=p(z_{t,l}|X_t,\mathbf{r}_{l,1})P(\mathbf{r}_{l,1})$, which is equivalent to \refeqn{tildePk1}.
+Next, the inductive step is solved for the $k+1$ index,
+\begin{align}
+\label{eqn:GenSol_InductiveStep}
+\tilde P(\mathbf{r}_{l,k+1}|X_t,z_{t,l})
+%&=
+%P(\mathbf{r}_{k+1})\sum_{r:r({k+1})=\mathbf{r}_{k+1}}p(z|X,r)P(r)\nonumber
+%\\
+%&=P(\mathbf{r}_{k+1})\{
+%P(\mathbf{r}_1)\sum_{\substack{r:r(1)=\mathbf{r}_1\\r({k+1})=\mathbf{r}_{k+1}}}p(z|X,r)P(r|\mathbf{r}_1,\mathbf{r}_{k+1})
+%\nonumber
+%\\
+%&\quad
+%+P(\bar{\mathbf{r}}_1)\sum_{\substack{r:r(1)=\bar{\mathbf{r}}_1\\r({k+1})=\mathbf{r}_{k+1}}}p(z|X,r)P(r|\bar{\mathbf{r}}_1,\mathbf{r}_{k+1})\}
+%\nonumber
+%\\
+&=P(\mathbf{r}_{l,k+1})\bigg\{
+P(\bar{\mathbf{r}}_{l,0})P(\mathbf{r}_{l,1})p(z_{t,l}|X_t,\mathbf{r}_{l,1})+\nonumber\\&\quad P(\bar{\mathbf{r}}_{l,0})P(\bar{\mathbf{r}}_{l,1})P(\mathbf{r}_{l,2})p(z_{t,l}|X_t,\mathbf{r}_{l,2})+...\nonumber
+\\
+&\quad +\left(
+P(\bar{\mathbf{r}}_{l,0})P(\bar{\mathbf{r}}_{l,1})...P(\bar{\mathbf{r}}_{l,k-1})
+\right)\nonumber
+\\
+&\quad\times P(\mathbf{r}_{l,k})p(z_{t,l}|X_t,\mathbf{r}_{l,k})\bigg\}\nonumber
+\\
+&\quad
++\left(
+P(\bar{\mathbf{r}}_{l,0})P(\bar{\mathbf{r}}_{l,1})...P(\bar{\mathbf{r}}_{l,k})
+\right)\nonumber
+\\
+&\quad\times p(z_{t,l}|X_t,\mathbf{r}_{l,k+1})P(\mathbf{r}_{l,k+1})\nonumber
+\\
+&=P(\mathbf{r}_{l,k+1})\sum_{i=1}^{k}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}\nonumber
+\\
+&\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i})\nonumber\\&\quad
++
+\left\{\prod_{j=0}^{k}P(\bar{\mathbf{r}}_{l,j})\right\}\nonumber
+\\
+&\quad\times p(z_{t,l}|X_t,\mathbf{r}_{l,k+1})P(\mathbf{r}_{l,k+1})
+.
+\end{align}
+
+\paragraph{Complement of the Unnormalized Reduced Map Inverse Sensor Model}
+An analytic expression for the complement of the unnormalized reduced map inverse sensor model is vital for obtaining the normalizer of the $l$-th ray at time $t$, namely $\eta_{t,l}$.
+
+The unnormalized complement of the $k$-th cell inverse sensor model, namely $\tilde P(\bar{\mathbf{r}}_k|X,z)$, is required to solve for $\eta$ based on \refeqn{allEta}.
+Consider the complement of the occupancy of the first and second cells, i.e.,
+\begin{align}
+\label{eqn:InvSenModknot1}
+\tilde P(\bar{\mathbf{r}}_{l,1}|X_t,z_{t,l})
+&=P(\bar{\mathbf{r}}_{l,1})
+\bigg\{P(\mathbf{r}_{l,2})
+p(z_{t,l}|X_t,\mathbf{r}_{l,2})\nonumber\\&\quad
++P(\bar{\mathbf{r}}_{l,2})P(\mathbf{r}_{l,3})p(z_{t,l}|X_t,\mathbf{r}_{l,3})+...\nonumber\\&\quad+\prod_{j=2}^{n_r-1}P(\bar{\mathbf{r}}_{l,j})P(\mathbf{r}_{l,n_r})p(z_{t,l}|X_t,\mathbf{r}_{l,n_r})
+\bigg\}\nonumber
+\\
+&=\sum_{i=2}^{n_r}
+\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})
+\right\}P(\mathbf{r}_{i})p(z_{t,l}|X_t,\mathbf{r}_{l,i}).
+\\
+\label{eqn:InvSenModknot2}
+\tilde P(\bar{\mathbf{r}}_{l,2}|X_t,z_{t,l})
+%&=\sum_{r:r(2)=\bar{\mathbf{r}}_2}p(z|X,r)P(r)\nonumber
+%\\
+%&=P(\bar{\mathbf{r}}_2)\sum_{r:r(2)=\bar{\mathbf{r}}_2}p(z|X,r)P(r|\bar{\mathbf{r}}_2)\nonumber
+%\\
+%&=P(\bar{\mathbf{r}}_2)\left\{P(\mathbf{r}_1)
+%\sum_{\substack{r:r(1)=\mathbf{r}_1\\r(2)=\bar{\mathbf{r}}_2}}p(z|X,r)P(r|\mathbf{r}_1,\bar{\mathbf{r}}_2)
+%+P(\bar{\mathbf{r}}_1)\sum_{\substack{r:r(1)=\bar{\mathbf{r}}_1\\r(2)=\bar{\mathbf{r}}_2}}p(z|X,r)P(r|\bar{\mathbf{r}}_1,\bar{\mathbf{r}}_2)
+%\right\}\nonumber
+%\\
+%&=P(\bar{\mathbf{r}}_2)
+%\{
+%P(\mathbf{r}_1)p(z|X,\mathbf{r}_1)
+%+P(\bar{\mathbf{r}}_1)
+%[
+%P(\mathbf{r}_3)\sum_{\substack{r:r(1)=\bar{\mathbf{r}}_1\\r(2)=\bar{\mathbf{r}}_2\\r(3)=\mathbf{r}_3}}p(z|X,r)P(r|\bar{\mathbf{r}}_1,\bar{\mathbf{r}}_2,\mathbf{r}_3)\nonumber
+%\\
+%&\quad
+%+
+%P(\bar{\mathbf{r}}_3)\sum_{\substack{r:r(1)=\bar{\mathbf{r}}_1\\r(2)=\bar{\mathbf{r}}_2\\r(3)=\bar{\mathbf{r}}_3}}p(z|X,r)P(r|\bar{\mathbf{r}}_1,\bar{\mathbf{r}}_2,\bar{\mathbf{r}}_3)
+%]
+%\}\nonumber
+%\\
+&=P(\bar{\mathbf{r}}_{l,2})
+\bigg\{
+P(\mathbf{r}_{l,1})p(z_{t,l}|X_t,\mathbf{r}_{l,1})\nonumber\\&\quad
++
+P(\bar{\mathbf{r}}_{l,1})P(\mathbf{r}_{l,3})p(z_{t,l}|X_t,\mathbf{r}_{l,3})\nonumber\\&\quad
++P(\bar{\mathbf{r}}_{l,1})P(\bar{\mathbf{r}}_{l,3})P(\mathbf{r}_{l,4})\nonumber\\&\qquad\times p(z_{t,l}|X_t,\mathbf{r}_{l,4})+...\nonumber
+\\
+&\quad+\prod_{j=1,j\neq2}^{n_r-1}P(\bar{\mathbf{r}}_{l,j})P(\mathbf{r}_{l,n_r})p(z_{t,l}|X_t,\mathbf{r}_{l,n_r})
+\bigg\}\nonumber
+\\
+&=
+P(\bar{\mathbf{r}}_{l,2})\sum_{i=1}^{1}
+\left\{
+\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}\nonumber\\&\quad\times P(\mathbf{r}_{l,i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})\nonumber\\&\quad+\sum_{i=3}^{n_r}
+\left\{
+\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}\nonumber\\&\quad\times P(\mathbf{r}_{l,i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})
+.
+\end{align}
+
+In general, the $k$-th unnormalized inverse sensor model complement is
+\begin{align}
+\label{eqn:ComplGenSol}
+\tilde P&(\bar{\mathbf{r}}_{l,k}|X,z)\nonumber\\&=P(\bar{\mathbf{r}}_{l,k})\sum_{i=1}^{k-1}
+\left\{
+\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})\nonumber\\&\quad
++\sum_{i=k+1}^{n_r}
+\left\{
+\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})
+,
+\end{align}
+where special cases when the lower bound of a summation is greater than the upper bound results in a summation value of $0$.
+
+\paragraph{Proof by Induction of the Complement of the Reduced Inverse Sensor Model \refeqn{ComplGenSol}}
+The base case for the occupancy probability is obtain directly from substituting $k=1$ into \refeqn{ComplGenSol} to obtain $\tilde P(\bar{\mathbf{r}}_1|X,z)=\sum_{i=2}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})$.
+The inductive step is solved for the $k+1$ index,
+\begin{align}
+\tilde P&(\bar{\mathbf{r}}_{l,k+1}|X_t,z_{t,l})\nonumber\\
+%&=P(\bar{\mathbf{r}}_{k+1})
+%\{
+%[
+%P(\mathbf{r}_1)p(z|X,\mathbf{r}_1)
+%+
+%P(\bar{\mathbf{r}}_1)P(\mathbf{r}_2)p(z|X,\mathbf{r}_2)
+%+P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)P(\mathbf{r}_3)p(z|X,\mathbf{r}_3)+...\nonumber
+%\\
+%&\quad
+%+P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)...P(\bar{\mathbf{r}}_{k-1})P(\mathbf{r}_k)p(z_{t,l}|\mathbf{r}_k,X_t)
+%]\nonumber
+%\\
+%&\quad
+%+
+%P(\bar{\mathbf{r}}_{k+1})^{-1}[
+%P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)...P(\bar{\mathbf{r}}_{k+1})P(\mathbf{r}_{k+2})p(z|X,\mathbf{r}_{k+2})+...
+%\nonumber
+%\\
+%&\quad
+%+P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)...P(\bar{\mathbf{r}}_{n_r-1})P(\mathbf{r}_{n_r})p(z|X,\mathbf{r}_{n_r})
+%]
+%\}\nonumber
+%\\
+&=
+P(\bar{\mathbf{r}}_{l,k+1})\nonumber
+\\
+&\quad
+\times\bigg[
+\left\{\prod_{j=0}^{0}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,1})p(z_{t,l}|X_t,\mathbf{r}_{l,1})\nonumber
+\\
+&\quad
++
+\left\{\prod_{j=0}^{1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,2})p(z_{t,l}|X_t,\mathbf{r}_{l,2})\nonumber
+\\
+&\quad
++\left\{\prod_{j=0}^{2}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,3})p(z_{t,l}|X_t,\mathbf{r}_{l,3})+...\nonumber
+\\
+&\quad
++\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,k})p(z_{t,l}|\mathbf{r}_{l,k},X_t)
+\bigg]\nonumber
+\\
+&\quad
++
+\bigg[
+\left\{\prod_{j=0}^{k+1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,k+2})p(z_{t,l}|X_t,\mathbf{r}_{l,k+2})+...\nonumber
+\\
+&\quad
++\left\{\prod_{j=0}^{n_r-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,n_r})p(z_{t,l}|X_t,\mathbf{r}_{l,n_r})
+\bigg]
+\nonumber
+\\
+&=P(\bar{\mathbf{r}}_{l,k+1})\sum_{i=1}^{k}
+\left\{
+\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})\nonumber
+\\
+&\quad
++\sum_{i=k+2}^{n_r}
+\left\{
+\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})
+.
+\end{align}
+
+\paragraph{Normalizer}
+Because the occupancy of the $k$-th cell is binary, the normalizer is defined as
+\begin{align}
+\eta_{t,l}&=\bigg[\tilde P(\bar{\mathbf{r}}_{l,k}|X,z)+\tilde P(\bar{\mathbf{r}}_{l,k}|X,z)\bigg]^{-1}\nonumber\\
+&=\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
+\end{align}
+which is independent of the reduced cell index $k$.
+
+\end{appendix}
 

### 2018-06-09 Conference/ACC 2016/ACC16.1 Old Version/Old Files and Conflicts/ACC16.1.tex
added Conference folder

--- 2018-06-09 Conference/ACC 2016/ACC16.1 Old Version 2/ACC16.1 (1)[Conflict].tex
@@
 \end{align}
-Then, the ray inverse sensor model is computed simply as
+
+%respectively.
+These have the follow solutions that may be obtained in a computationally-efficient manner:
 \begin{align}
-\label{eqn:RayISMAnswer}
-P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}).
+\label{eqn:tildePresult}
+\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
+&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
+&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
+\nonumber
+\\
+&\quad
++
+\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
+&\quad\times p(z_{t,l}|\mathbf{r}_{k+},X_t)P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1}),
+\\
+\label{eqn:tildePcomplementresult}
+\tilde P(\bar{\mathbf{r}}_k|z_{t,l},&X_{1:t},Z_{1:t-1})
+=P(\bar{\mathbf{r}}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
+&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
+&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
+\nonumber
+\\
+&\quad
++
+\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
+&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}).
 \end{align}
+\end{prop}
 \begin{proof}

### 2018-06-09 Conference/ACC 2016/Final Abridged/ACC16.1 IRA.tex
added Conference folder

\documentclass[letterpaper, 10pt, conference]{ieeeconf}
\IEEEoverridecommandlockouts \overrideIEEEmargins
\usepackage{amsmath,amssymb,url,times,subfigure,graphicx,theorem}
\usepackage{graphicx,subfigure,hyperref}
\usepackage{color,comment}
\usepackage{csquotes}
%\usepackage{refcheck}

\usepackage{epic,eepic}

\usepackage[]{algorithm2e}
%\usepackage{algorithmicx}
%\usepackage{algpseudocode}

%\usepackage{algorithm}
%\usepackage{algpseudocode}
%\usepackage{pifont}

%\usepackage{amsmath}
%\usepackage{algorithm}
%\usepackage[noend]{algpseudocode}
%
%\makeatletter
%\def\BState{\State\hskip-\ALG@thistlm}
%\makeatother

%\definecolor{red}{rgb}{0,0,0}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\abs}[1]{\ensuremath{\left| #1 \right|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\ip}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mbox{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mbox{tr}\ensuremath{\!\bracket{#1}}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathbf{d}}}
\newcommand{\pair}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\!\langle #1 \rangle\!\rangle}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\ad}{\ensuremath{\mathrm{ad}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}

\title{\LARGE \bf
Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model}

\author{Evan Kaufman, Taeyoung Lee, Zhuming Ai, and Ira S. Moskowitz%
\thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 {\tt \{evankaufman,tylee\}@gwu.edu}}
\thanks{Zhuming Ai, Ira S. Moskowitz, Information Management \& Decision Architectures Branch, Code, 5580,  Naval Research Laboratory,  Washington, DC 20375}
\thanks{This research has been supported by the U.S. Naval Research Laboratory Base Program Work Unit ``Intelligent Microflyer'' and in part by NSF under the grants CMMI-1243000, CMMI-1335008, and CNS-1337722.}
}

\newcommand{\EditEK}[1]{{\color{blue}\protect #1}}
%\renewcommand{\EditTL}[1]{{\protect #1}}

\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{cor}{Corollary}
\newtheorem{remark}{Remark}

\begin{document}
\allowdisplaybreaks

\maketitle \thispagestyle{empty} \pagestyle{empty}

\begin{abstract}
Occupancy grid maps are spatial representations of environments where the space of interest is decomposed into a number of cells that are considered either occupied or free. This paper focuses on occupancy grid mapping which   estimates the probability of occupancy for each cell based upon range measurements from a known location. For a given probabilistic model of a range sensor, we propose a computationally efficient method to obtain an exact inverse sensor model. We use this to   construct a probabilistic mapping algorithm according to a Bayesian framework. Compared with the existing occupancy grid mapping techniques that rely on approximate, heuristic inverse sensor models, the our approach yields substantially more accurate maps for the same set of measurements. This is illustrated by numerical examples and experiments. 
\end{abstract}

\section{Introduction}

In the context of robotics, mapping is done to    to generate maps which represent the environment in close proximity of a robot. This problem is crucial to understanding the surrounding regions when a robot enters an uncertain environment.
Various map representations include occupancy grids~\cite{WolSuk05}, octomaps~\cite{WurHorBenStaBur10}, or feature-based maps%in a landmark representation
~\cite{MonThrKolWeg02}.
%To build maps in real time, a computationally-efficient algorithm is required.% \EditTL{rewrite this paragraph for mapping, without introducing SLAM here}

This paper focuses on the popular occupancy grid representation, where the map space is decomposed into evenly-spaced grid cells, each cell considered as  occupied or free.
The \emph{inverse sensor model} is a probabilistic model  of occupancy for a grid cell based on range measurements and the configuration of the mobile robot (which is of fundamental importance with occupancy grid mapping). However, the exact solution to the inverse sensor model has not been utilized in occupancy grid mapping, because the computational requirements are too cumbersome to compute the exact model in real time~\cite{ThrBurFox05}.

Instead, several techniques have been proposed to approximate the exact inverse sensor model. %Due to the computational limitations, two common approaches are applied to obtain a function close to the inverse sensor model.
In the early occupancy grid mapping applications based on sonar sensors, \emph{ad hoc} techniques to obtain an inverse sensor model were developed based off highly inaccurate  reflected sound waves~\cite{MorElf85,Elf89}.
%These occupancy grid mapping schemes employ \emph{ad hoc} techniques to obtain the inverse sensor model.
These techniques are based on approximate functions to represent the inverse sensor model with questionable accuracy~\cite{ChoLynHutKanBurKavThr05} and applied to more modern sensors in~\cite{And09},~\cite{PirRutBisSch11}.

The other popular approach to obtain an approximate inverse sensor model is to simulate maps, robot poses, and measurements, in conjunction with a learning algorithm to approximate the inverse sensor model~\cite{ThrBurFox05,Thr01}. These approaches are undesirable in practice due to the complexities associated to implementing a learning algorithm. For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples, or how many samples are required to obtain a reasonable approximation. Furthermore, it is challenging to apply any learning algorithm over the large dimensional space composed of maps, poses, and measurements.

%and they only yield probabilistic approximations without strong mathematical justification.

%These approaches simply treat occlusions with discrete probabilities of detection, regardless of whether one cell may completely, partially, or not block another cell in the sensor field of view.

%Furthermore, both approaches follow a log-odds ratio Bayesian framework, imposing an additional Markov assumption.

%These techniques are not also suitable for modern range sensor with the highly accurate bearing~\cite{Thr03}. Based on recent improvements in sensor technology, the occupancy grid mapping problem has been restructured using the fact that depth measurement bearings are known. \EditTL{do we need this paragraph?}

This paper proposes a computationally efficient algorithm to construct the exact inverse sensor model. More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, our algorithm yields the a posteriori probability of occupancy of all the cells within the area covered by the range sensor via the range measurements. The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancy of the other cells blocked by it is irrelevant to the forward sensor model, and this approach is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model. Furthermore, our proposed approach integrates the a priori probability of occupancy and multiple range measurements according to a Bayesian framework to obtain more accurate maps. As such, it contrasts from the existing framework based on log-odds ratios that impose an additional Markov assumption.

%these are constructed

In short, the main contribution of this paper is in proposing an exact inverse sensor model and constructing a Bayesian occupancy grid mapping based upon  it. Compared with the current mapping algorithms based on   approximate inverse sensor models, this approach yields more accurate occupancy probabilities. This constructs substantially more accurate occupancy grid maps with less uncertainty for the same set of measurements, and these are directly illustrated by numerical examples and preliminary experimental results. While this paper is focused on the mapping problem in two-dimensional environments, the presented results can be certainly utilized in SLAM or stochastic motion planning in three dimensions.

%occupancy probabilities that are more accurate than prior work, which may be applied for more accurate real-time map generation in the SLAM problem and more effective trajectory planning strategies based on map uncertainty.

% Abridged:
%The paper is organized as follows.
%The problem is formulated in Section \ref{sec:ProbForm}, and the exact solution to the occupancy grid mapping problem is derived in Section \ref{sec:ISM}.
%A numerical result shows the improved accuracy in Section \ref{sec:NumRes}, and an experimental result with the Kinect depth sensor shows the improvement in map certainty in Section \ref{sec:ExpRes}, followed by conclusions.

\section{Problem Formulation}
\label{sec:ProbForm}

\subsection{Range Sensor}

Consider a range sensor that provides scans of the surrounding environment in order to identify the closest occupied space. The location and the direction of the sensor, referred to as the \emph{pose} at time $t$, is denoted by $X_t$. The range sensor returns a measurement \emph{scan} $Z_t$, composed of $n_z$ measurement \emph{rays} $z_{t,1},z_{t,2},\ldots,z_{t,n_z}$. The $l$-th measurement ray for $l\in\braces{1,2,\ldots,n_z}$ contains the depth (range) and bearing (direction) inside the sensor field of view (FOV) that corresponds to a circular sector. Given $X_t$, we assume the origin and directions of each ray are known deterministically.

The probability density function, namely $p(z_{t,l}|m,X_t)$ with respect to the depth of the $l$-th measurement ray conditioned on the map $m$ and the pose $X_t$ is commonly referred to as the \emph{forward sensor model}, which characterizes the corresponding depth sensor, such as the maximum range or accuracy. The forward sensor model satisfies  (i) the ranges of all depth measurements are positive and finite, and (ii) measurement rays cannot pass through occupied regions.

Throughout this paper, we assume that the forward sensor model of the selected sensor is given. This can be determined empirically or analytically. For example, the \emph{beam model for range finders} satisfying  the above criteria is described in~\cite{ThrBurFox05}.
%, which contains Gaussian, exponential, uniform, and point pass portions~\cite{ThrBurFox05}.
%At time $t$ and given the pose $X_t$ and map $m$, the beam model with respect to the $l$-th measurement ray $z_{t,l}$ is the probability density defined as
%\begin{align}
%p(z_{t,l})=\begin{cases}
%\frac1{z_{max}-z_{min}}P_{hit}+P_{short}+P_{rand}\ \text{if},\ 
%\\
%P_{max},
%\end{cases}
%\end{align}
%where $z_{min}\leq z_{t,l}\leq z_{max}$ and $\hat z_{t,l}$ is the distance between the robot and the closest occupied location along the direction $z_{t,l}$.
%The probabilities are defined such that $P_{hit}$ is the probability of the depth originating from this occupied location (Gaussian), $P_{exp}$ is the probability that an unexpected occlusion blocks the first occupied location of the map (exponential), $P_{rand}$ is the probability of a phantom reading (uniform), and $P_{max}$ is the probability that the sensor returns its maximum value (point mass).
%These probabilities are normalized such that
%\begin{align}
%P_{hit}+P_{short}+P_{rand}+P_{max}=1.
%\end{align}

%We assume that this forward sensor model depends on the closest occupied location, and is independent of the occupancy beyond this point.

\subsection{Occupancy Grid Mapping}

Let a map $m$ be decomposed into $n_m$ evenly-spaced grid cells, where the $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,\ldots,n_m}$, that is defined as $\mathbf{m}_i=1$ when occupied, and $\mathbf{m}_i=0$ when free. The location and size of each grid cell is assumed known. Therefore, a map $m$ is defined by $\{\mathbf{m}_1,\ldots, \mathbf{m}_{n_m}\}$ ($2^{n_{m}}$ possible maps).

Another random variable is defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience. The probability that the $i$-th cell is occupied is $P(\mathbf{m}_i)$, and the probability that it is free is $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$. The random variables $\mathbf{m}_i$ are mutually independent, 
\begin{align}
P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i).
\end{align}

%Occupancy grid mapping produces a probabilistic map; none of the $n_m$ grid cell occupancies are completely certain.
Occupancy grid mapping  is used to determine these probabilities based on robot poses and measurement scans. More explicitly, let $X_{1:t}$ denote the history of poses from the initial time to the current time, $\{X_1,X_2,\ldots, X_t\}$, and let $Z_{1:t}$ be the measurement history. The goal is to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, commonly referred to as the \emph{inverse sensor model} for given forward sensor model $p(z|m,X)$ and the initial estimate of the map $P(m)$ with $X_{1:t}$ and $Z_{1:t}$.

	% Def of occupancy grid
	% Cell occupancies are assumed independent
	% Probabilistic maps
	% Notations
	% Mapping problem

\subsection{Mapping via Log-Odds Ratio}

A popular frameworks for updating binary random variables with static state, is to use a binary Bayesian filter using the log-odds ratio formulation.
The main idea is that instead of multiplying terms from prior and current time steps, the properties of logarithms allow these terms to be simply added, while avoiding truncation issues associated with probabilities close to $0$ or $1$~\cite{ThrBurFox05}.
The log-odds ratio is also popular because learning techniques with forward models~\cite{Thr01},~\cite{Thr03}, or ad hoc techniques~\cite{MorElf85},~\cite{Elf89} are simpler to derive when they need not consider $P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})$ as part of the inverse sensor model, but rather only $P(\mathbf{m}_i)$, which is typically fixed and uniform for all cells.

However, the log-odds ratio formulation makes an assumption that is not consistent with the occupancy grid mapping problem. It is assumed that if $X_t$ and $\mathbf{m}_i$ are known, then $Z_t$ is independent of $X_{1:t-1}$ and $Z_{1:t-1}$, i.e.,
%\begin{align}
%\label{eqn:FirstBayesRule}
%P(\mathbf{m}_i|X_{1:t},Z_{1:t})
%&
%=\frac{P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})}{P(Z_t|X_{1:t},Z_{1:t-1})}.
%\end{align}
\begin{align}
\label{eqn:AssumptionEarly}
P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})\approx P(Z_t|\mathbf{m}_i,X_t).
\end{align}

However, forward models are based on the entire map $m$. Since \refeqn{AssumptionEarly} conditioned only by the $i$-th cell, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may contain important information about the occupancy of the other cells, which should be considered in calculating the probability distribution of the measurement scan $Z_t$.  %In short, this log-odds ratio framework is based on the assumption that neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$.

%However, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may suggest that there exist occupied regions that occlude some or all of $Z_t$ from detecting $\mathbf{m}_i$ from the current pose $X_t$.
%Since this assumption neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$, this framework is subject to poor results.
%Hence, the log-odds ratio formulation is completely avoided in further analysis.

	% approximated inverse sensor model
	% markov assumption
	% shortcoming, reason this is avoided

\section{Exact Inverse Sensor Model}
\label{sec:ISM}

Due to computational complexity, approximate inverse sensor models have been used for occupancy grid mapping with simplifying assumptions. In this section, we propose an efficient algorithm to compute the exact inverse sensor model. 
%Much like the forward sensor model that reasons measurements from a known map, he inverse sensor model reasons that map from known measurements~\cite{ThrBurFox05}.
%While the inverse sensor model varies slightly in meaning in numerous publications, two versions are derived in this paper, both based on a Bayesian framework.
We present two versions according to a Bayesian framework. The first type is referred to as the \emph{ray inverse sensor model} $P(\mathbf{m}_i|X_{1:t},Z_{1:t-1},z_{t,l})$, which only considers $l$-th ray at time $t$, i.e., one-dimensional inverse sensor model. The second type is referred to as the \emph{scan inverse sensor model} $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, which combines the information from \emph{all} measurement rays \emph{simultaneously} at the $t$-th time step for two-dimensional cases. These are developed in a sequential form to be repeated at each time step.

%These are contrary to inverse sensor models commonly used with the log-odds ratio formulations, and the proposed algorithm integrates all of past poses and measurement scans with the current ones to obtain more accurate estimate of the map.

\subsection{Ray Inverse Sensor Model}

Suppose the probability of the map conditioned on the past poses and measurements, namely $P(m|X_{1:t-1},Z_{1:t-1})$, is known. Here we construct a posteriori probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$, based on the current pose $X_t$, the measurement from the $l$-th ray $z_{t,l}$, and the given forward sensor model $p(z_{t_l}|m,X_t)$.

\paragraph{Bayesian Framework}
%Assuming that the probability density of thxe sensor measurement is fixed over a small measurement region $\Delta z$, 
Bayes' rule yields
%First, Bayes' rule is applied for a particular map $m$ and arbitrarily small measurement region $\Delta z$, yielding
\begin{align}
\label{eqn:BayesRuleRayISM}
P(&m|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\frac{p(z_{t,l}|m,X_{1:t},Z_{1:t-1})P(m|X_{1:t-1},Z_{1:t-1})}{p(z_{t,l}|X_{1:t},Z_{1:t-1})},
\end{align}
considering that $X_t$ carries no information about $m$ without $Z_t$.
If the current pose $X_t$ and map $m$ are known, then the measurement ray $z_{t,l}$ is independent of past poses $X_{1:t-1}$ and past measurements $Z_{1:t-1}$ to obtain
%Note that this is contrary to approaches that consider $X_t$ and $\mathbf{m}_i$ (not the full map $m$) as enough information, ignoring the occupancy of other cells in this term.
%Then, \refeqn{BayesRuleRayISM} can be rewritten as
\begin{align}
P(m|z_{t,l},&X_{1:t},Z_{1:t-1})\nonumber
\\
&
=\eta_{t,l}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}),
\end{align}
where the normalizing constant $\eta_{t,l}\in\Re$ absorbs all terms independent of the map $m$.

%Assuming $z_{t,l}$ measures an occupied cell, then $z_{t,l}$ may be a result of the occupancy of $\mathbf{m}_i$, or $z_{t,l}$ may measure a different occupied cell.

Next, we compute the occupancy probability of each cell. Let $\mathcal{M}_i$ be the set of maps where the $i$-th cell is occupied, i.e., $\mathcal{M}_i =\{m\in\{0,1\}^{{n_m}}\,|\ \mathbf{m}_i=1\}$. To compute the probability of occupancy of the $i$-th cell, all possible combinations of map in $\mathcal{M}_i$ should be considered, i.e., 
%All possible combinations of map $m$ must be considered in which $\mathbf{m}_i$ is occupied, denoted as $m:m(i)=\mathbf{m}_i$,
\begin{align}
\label{eqn:InvSenModWithProbDens}
&P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\eta_{t,l}\sum_{m\in\mathcal{M}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
\end{align}
Furthermore, to determine the normalizing constant $\eta_{t,l}$, the complement $P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in a similar manner. Since there are $2^{n_{m}-1}$ maps in $\mathcal{M}_i$, these equations require $2^{n_m}$ terms to compute the summation for the $i$-th grid cell, and it should be repeated for other cells. This is the main reason why the current results are based on certain approximations of the above expression.

\paragraph{Computationally Efficient Approach}

We propose a computational algorithm to evaluate \refeqn{InvSenModWithProbDens} efficiently. 
Since the cells outside of the sensor FOV are not affected, we focus on a reduced map $r_l$ in the FOV of the $l$-th ray. This reduced map is chosen such that each cell of $r_l$ corresponds to a grid cell of $m$ that the $l$-th ray intersects, ordered by increasing distance, which is easily determined from geometry. Let $\mathbf{r}_{l,k}$ be the binary random variable representing the occupancy of the $k$-th cell of the $l$-th ray. The number of cells in the reduced map is denoted by $n_{r,l}\leq n_m$.

Next, let $\mathbf{r}_{l,k+}$ correspond to the set of maps that the $k$-th cell of the $l$-th ray is occupied, cells with lower index are free, and cells with greater index may or may not be occupied. In other words, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray. Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell. Based on this, we present the following computational algorithm to compute the inverse sensor model.

\begin{prop}
For the $l$-th measurement ray, the a posteriori probability of the occupancy of the $k$-th cell, namely the ray inverse sensor model, is given by
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}),
\end{align}
where the unnormalized probability of the inverse sensor model is defined as
\begin{align}
\label{eqn:Unnormalized}
& \tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\
&=P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\
&\quad\times 
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]\nonumber\\
&\quad + \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
&\quad\times p(z_{t,l}|\mathbf{r}_{l,k+},X_t)P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1}),
\end{align}
where it is considered $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading for convenience. The normalizer $\eta_{t,l}$ is given by
\begin{align}
\label{eqn:allEta}
\eta_{t,l}
&=
\bigg[\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
\end{align}
and it is independent of the cell index $k$.
\end{prop}
\begin{proof}
See Appendix A.
\end{proof}

%
%%respectively.
%These have the follow solutions that may be obtained in a computationally-efficient manner:
%\begin{align}
%\label{eqn:tildePresult}
%\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
%\nonumber
%\\
%&\quad
%+
%\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{k+},X_t)P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1}),
%\\
%\label{eqn:tildePcomplementresult}
%\tilde P(\bar{\mathbf{r}}_k|z_{t,l},&X_{1:t},Z_{1:t-1})
%=P(\bar{\mathbf{r}}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
%\nonumber
%\\
%&\quad
%+
%\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}).
%\end{align}

%
%
%Because $\mathbf{r}_k$ is binary, the following holds:
%\begin{align}
%\label{eqn:allEta}
%\eta_{t,l}&=\frac1{\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})+\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})}\nonumber
%\\
%&=
%\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
%\end{align}
%where $\eta_{t,l}$ is independent of the index $k$ of the reduced map.
%Rearranging \refeqn{Unnormalized} and \refeqn{UnnormalizedCompl}, the ray inverse sensor model is 
%\begin{align}
%\label{eqn:RayISMAnswer}
%P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
%\\
%P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
%\end{align}
%respectively, where $\eta_{t,l}$ comes from \refeqn{allEta} and the unnormalized probabilities come from \refeqn{tildePresult} and \refeqn{tildePcomplementresult}.

Note that the a priori estimate, $P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ and its compliment $P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})=1-P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ are available at the $t$-th step. Then, \refeqn{RayISMAnswer} yields a sequential occupancy grid mapping that can be applied whenever new measurements are available.

Compared with \refeqn{InvSenModWithProbDens}, where the terms of the summation should be repeated $2^{n_{r,l}}$ times \emph{per each cell of the reduced map}, the proposed expressions \refeqn{RayISMAnswer} and \refeqn{allEta} are \textit{substantially} simpler as they require $k$ summation terms for the $k$-th cell. In fact, if the terms inside the summations of \refeqn{Unnormalized} and \refeqn{allEta} are obtained recursively, only $n_{r,l}+1$ repetitions are required for \emph{all cells of the reduced map}, yielding an algorithm that is nearly $\mathbf{2^{n_{r,l}}}$ \textbf{times faster}.

\subsection{Scan Inverse Sensor Model}

%The goal of occupancy grid mapping is to obtain the probabilistic map based on all measurements and poses, including the latest measurement scan.

%At the $t$-th time step, the measurement scan is the set of all rays $Z_t=\{z_{t,1},z_{t,2},\ldots,z_{t,n_z}\}$.

%However, when there are a large number of rays, it is certainly possible that a single cell is covered by multiple rays.

%One may apply the results of Proposition 1 repeatedly for each ray to obtain a two-dimensional inverse sensor model,
%The scan inverse sensor model is generalized naturally by applying the conditions recursively,
The above one-dimensional ray inverse sensor model may be generalized into a two-dimensional scan inverse sensor model by updating the map sequentially for each ray, according to
%The most natural way to generalize the ray inverse sensor model into a scan inverse sensor model is applying the conditions in a recursive way,
\begin{align}
P(&\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\&%=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1},z_{t,2},\ldots,z_{t,n_z})\nonumber\\&
=P(((((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z}).
\end{align}
Alternatively, measurements from multiple rays can be integrated synergistically, based on the assumption that the measurements of rays are mutually independent, i.e., $p(z_{t,1},z_{t,2},\ldots,z_{t,n_z}|\mathbf{m}_i,X_{1:t},Z_{1:t-1})=\prod_{l=1}^{n_z}p(z_{t,l}|\mathbf{m}_i,X_{1:t},Z_{1:t-1})$.
Let $\mathcal L_i\subset\{1,\ldots, n_z\}$ be the set of rays intersecting the $i$-th cell, $\mathbf{m}_i$. Applying Bayes' rule repeatedly, we obtain
\begin{align}
%\label{eqn:ManyMeasBayesRule}
%P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\
%&
%=\frac{P(z_{t,1:n_z}|\mathbf{m}_i,X_{1:t},Z_{1:t-1})P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})}{P(z_{t,1:n_z}|X_{1:t},Z_{1:t-1})}
%\nonumber
%\\
&P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\
&=\tilde\zeta_i\braces{\prod_{l\in\mathcal{L}_i}p(z_{t,l}|\mathbf{m}_i,X_{1:t},Z_{1:t-1})}P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})\nonumber\\
%
%\end{align}
%Using Bayes' rule again, it can be rearranged as
%\begin{align}
%P(\mathbf{m}_i|X_{1:t},Z_{1:t})&=
%\frac{P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})^{1-n_z}}{P(z_{t,1:n_z}|X_{1:t},Z_{1:t-1})}
%\nonumber\\&\quad\times
%\prod_{l=1}^{n_z}\left\{\frac{P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})P(z_{t,l}|X_{1:t},Z_{1:t-1})P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})}\right\}
%\nonumber
%\\
&=
\zeta_i P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})%\nonumber\\&\quad\times
\prod_{l\in\mathcal{L}_i}\frac{P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})},
\label{eqn:ThirdBayesRule}
\end{align}
where $\tilde\zeta_i,\zeta_i\in\Re$ correspond to normalizing constants.% independent of $\mathbf{m}_i$.

%considering that $X_t$ carries no information about $\mathbf{m}_i$ without $Z_t$, and where the normalizer $\zeta_i = \frac{\prod_{l=1}^{n_z} P(z_{t,l}|X_{1:t},Z_{1:t-1})}{P(z_{t,1:n_z}|X_{1:t},Z_{1:t-1})}$.
%\refeqn{ThirdBayesRule} is simplified to 
%\begin{align}
%P(\mathbf{m}_i|X_{1:t},Z_{1:t})&=
%\zeta_i P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})^{1-n_z}\nonumber\\&\quad\times\prod_{l=1}^{n_z}\left\{P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\right\}.
%\label{eqn:FourthBayesRule}
%\end{align}

% and we assume $Z_t$ is independent of $X_{1:t-1}$ and $Z_{1:t-1}$.
% and $P(\mathbf{m}_i|X_{1:t},z_{t,l},Z_{1:t-1})=P(\mathbf{m}_i|X_t,z_{t,l})$.

%Similarly, the complement is
%\begin{align}
%\label{eqn:CompBayesProbDef}
%P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})&=
%\zeta_i P(\bar{\mathbf{m}}_i|X_{1:t-1},Z_{1:t-1})^{1-n_z}\prod_{l=1}^{n_z}\left\{P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})\right\}
%\nonumber
%\\
%&=\zeta_i (1-P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1}))^{1-n_z}\prod_{l=1}^{n_z}\left\{(1-P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1}))\right\}.
%\end{align}
%Because $\mathbf{m}_i$ is a binary random variable,
%\begin{align}
%\label{eqn:BinNorm}
%P(\mathbf{m}_i|X_{1:t},Z_{1:t})+P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})=1,
%\end{align}
%so \refeqn{FourthBayesRule} and \refeqn{CompBayesProbDef} can be substituted into \refeqn{BinNorm} to solve for $\zeta_i$,
%\begin{align}
%\label{eqn:xi}
%\zeta_i&=
%\bigg[
%P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})^{1-n_z}\prod_{l=1}^{n_z}\left\{P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\right\}
%\nonumber\\&\qquad
%+
%(1-P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1}))^{1-n_z}\prod_{l=1}^{n_z}\left\{(1-P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1}))\right\}
%\bigg]^{-1}.
%\end{align}

% Part 2...

%Consider two cases with the ray inverse sensor model $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})$.

Suppose the $l$-th measurement ray intersects with the cell $\mathbf{m}_i$, where $\mathbf{r}_{l,k}$ corresponds to the cell $\mathbf{m}_i$ in the $l$-th reduced map, so $\mathbf{r}_{l,k}=\mathbf{m}_i$.  Then, we have $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})=P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$. 
%
%Or, when the measurement ray $l$ misses $\mathbf{m}_i$, we have $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})=P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})$ because $X_t$ holds no information about $\mathbf{m}_i$ without $z_{t,l}$.
%
Using this, \refeqn{ThirdBayesRule} is rewritten as
\begin{align}
P(\mathbf{m}_i|X_{1:t},Z_{1:t})
%&=
%\zeta_i P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})^{1-n_z}\prod_{l=1}^{n_z}\left\{P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\right\}
%\nonumber\\&
%=\zeta_i P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})^{1-n_{i,l}}\prod_{\mathcal L_i}\left\{\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\right\}
%\nonumber\\
&=\xi_i P(\mathbf{m}_i|{X_{1:t-1}},Z_{1:t-1})\nonumber\\&\quad\times
\prod_{l\in\mathcal L_i}
\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
,
\label{eqn:ISM_Fusion}
\end{align}
where the normalizer $\xi_i$ is composed of the product of normalizers $\zeta_i$, and $\eta_i$. Here, we introduce a new term $\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ for computational efficiency as
\begin{align}
&\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
\triangleq \frac{\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})}
\nonumber\\&=
\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})
\nonumber\\&\quad
+
\bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,k+},X_t),
\end{align}
where we have used \refeqn{Unnormalized}. Similarly, its complement is
\begin{align}
P(\bar{\mathbf{m}}_i|{{X_{1:t}}},Z_{1:t})
&=\xi_i P(\bar{\mathbf{m}}_i|{X_{1:t-1}},Z_{1:t-1})
\nonumber\\&\quad\times\prod_{l\in\mathcal L_i}
\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
,
\label{eqn:ISM_Bar_Fusion}
\end{align}
where $\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ is defined as
\begin{align}
&\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
\triangleq\frac{\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t-1})}
\nonumber\\
&=\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})
\nonumber
\\
&\quad
+
\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0,j\neq k}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
\end{align}
which is obtained from \refeqn{tildePbar}. %, where the division of $P(\bar{\mathbf{m}}_i|X_{1:t-1},Z_{1:t-1})=P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})$ is absorbed into the second summation.
Since 
\begin{align*}
P(\mathbf{m}_i|X_{1:t},Z_{1:t})+P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})=1,
\end{align*}
the normalizer $\xi_i$ is obtained using \refeqn{ISM_Fusion} and \refeqn{ISM_Bar_Fusion} as,
\begin{align}
&\xi_i=
\bigg[
P(\mathbf{m}_i|{X_{1:t-1}},Z_{1:t-1})
\prod_{\mathcal L_i}
\hat P(\mathbf{r}_{l,k}|z_{t,l},{X_{1:t}},Z_{1:t-1})
\nonumber\\&\quad
+
P(\bar{\mathbf{m}}_i|{X_{1:t-1}},Z_{1:t-1})
\prod_{\mathcal L_i}
\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
\bigg]^{-1},\label{eqn:xi}
\end{align}
which is substituted into \refeqn{ISM_Fusion} to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$.%  the complete scan inverse sensor model

%{\color{red} There are two key aspects of $\xi_i$ that make the variable valuable to the occupancy grid mapping problem. The first is that $\xi_i$ holds important information about probability terms that are otherwise difficult or impossible to calculate accurately.
%The second is that since $\xi_i$ is the same for both $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$ and $P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})$, which are based off of products, then there is no need to associate unnormalized probabilities with specific measurement rays; put differently, if individual measurement ray normalizers are \emph{not} equal, this difference has no affect (such as weighting) when computing $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, thereby decreasing the necessary memory for the algorithm.}

\begin{table}
\begin{tabular}{ l }
  for $l = 1,2,\ldots,n_z$\\
   \ \ Obtain reduced map $r_{t,l}$\\
   \ \ Define $P(\mathbf{r}_{l,0}|X_{1:t-1},Z_{1:t-1})=0$, $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=1$, \\  \ \ \ \ \ \ \ \ \ \   $P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$;\\
   \ \ for $k = 1,2,\ldots,n_r$\\
   \ \ \ \ Obtain $p(z_{t,l}|X_t,\mathbf{r}_{l,k})$ from a forward sensor model\\
   \ \ \ \ if $k=1$\\
   \ \ \ \ \ \ $a_1=0$, $b_1=1$;\\
   \ \ \ \ else\\
   \ \ \ \ \ \ $a_k=a_{k-1}+b_{k-1}p(z_{t,l}|\mathbf{r}_{l,k-1},X_t)P(\mathbf{r}_{l,k-1}|X_{1:t-1},Z_{1:t-1})$;\\
   \ \ \ \ \ \ $b_k=b_{k-1}P(\bar{\mathbf{r}}_{l,k-1}|X_{1:t-1},Z_{1:t-1})$;\\
   \ \ \ \ end if\\
   \ \ end for\\
   \ \ $c_{n_{r,l}+1}=0$;\\
   \ \ for $k = n_{r,l},n_{r,l}-1,\ldots,1$\\
   \ \ \ \ $c_k=\frac{P(\mathbf{r}_{l,k+1}|X_{1:t-1},Z_{1:t-1})}{P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})}c_{k+1}$\\
   \ \ \ \ \ \ \ \ \ \ \ \ $+b_{k}p(z_{t,l}|\mathbf{r}_{l,k+1},X_t)P(\mathbf{r}_{l,k+1}|X_{1:t-1},Z_{1:t-1})$;\\
   \ \ end for\\
   \ \ for $k = 1,2,\ldots,n_r$\\
   \ \ \ \ $\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=a_k+b_kp(z_{t,l}|\mathbf{r}_{l,k},X_t)$;\\
   \ \ \ \ $\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=a_k+c_k$;\\
   \ \ end for\\
   end for\\
   for all $i$ map cells inside the FOV\\
   \ \ Obtain the set $\mathcal L_i$ of $l_i$ measurement rays intersecting this cell\\
   \ \ if $l_i>0$\\
   \ \ \ \ $d=P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})\prod_{\mathcal L_i}\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$;\\
   \ \ \ \ $e=P(\bar{\mathbf{m}}_i|X_{1:t-1},Z_{1:t-1})\prod_{\mathcal L_i}\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$;\\
   \ \ \ \ $P(\mathbf{m}_i|X_{1:t},Z_{1:t})=\frac{d}{d+e}$;\\
   \ \ else\\
   \ \ \ \ $P(\mathbf{m}_i|X_{1:t},Z_{1:t})=P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1});$\\
   \ \ end if\\
   \ \ Return: $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$\\
   end for\\

\end{tabular}
\caption{Fast Algorithm of the Exact Solution of the Inverse Sensor Model for $l$ Simultaneous Measurement Rays}
\label{tab:Alg_ISM_2D}
\end{table}

\subsection{Summary of Algorithm}

In short, the proposed scan inverse sensor model is computed by \refeqn{ISM_Fusion}-\refeqn{xi}. These are summarized in Table \ref{tab:Alg_ISM_2D} as a computationally efficient, recursive algorithm that avoids several repeated calculations of the same quantity. This algorithm utilizes the following temporary variables to develop the algorithm in an efficient recursive form, %defined as
\begin{align*}
a_k&=\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
\\
b_k&=\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1}),
\nonumber\\
c_k&=
\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0,j\neq k}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}).%,\sum_{i=k+1}^{n_{r,l}}\bigg\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}).
\end{align*}

In contrast to the current approximate inverse sensor models, the proposed algorithm evaluates the exact inverse sensor model \refeqn{InvSenModWithProbDens} efficiently without relying on assumptions like \refeqn{AssumptionEarly}, while integrating multiple measurements synergistically. Therefore, it yields substantially more accurate maps for the same set of measurements. These are illustrated by numerical examples and experiments as follows.

%While the derivation of the complete inverse sensor model is complicated, the algorithm is quite straightforward.
%Temporary variables,
%\begin{align*}
%a_k&=\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
%\\
%b_k&=\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1}),
%\nonumber\\
%c_k&=\sum_{i=k+1}^{n_{r,l}}\left\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
%\end{align*}
%serve to simplify the algorithm such that repeated calculations are removed, and updated occupancy probabilities are obtained recursively.
%The pseudocode is summarized in Table \ref{tab:Alg_ISM_2D}.

	% Mathematical computation order of complexity
	% Temporary variables: limited number to avoid repeated calculations

\begin{figure}
\centerline{
	\subfigure[Map at $t=0$ (exact)]
		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_1.png}
	}
	\subfigure[Map at $t=0$ (approx.)]
		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_1.png}
	}
}
\centerline{
	\subfigure[Map at $t=13$ (exact)]
		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_2.png}
	}
	\subfigure[Map at $t=13$ (approx.)]
		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_2.png}
	}
}
\centerline{
	\subfigure[Map at $t=25$ (exact)]
		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_3.png}
	}
	\subfigure[Map at $t=25$ (approx.)]
		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_3.png}
	}
}
\centerline{
	\subfigure[Map at $t=38$ (exact)]
		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_4.png}
	}
	\subfigure[Map at $t=38$ (approx.)]
		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_4.png}
	}
}
\centerline{
	\subfigure[Map at $t=50$ (exact)]
		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_5.png}
	}
	\subfigure[Map at $t=50$ (approx.)]
		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_5.png}
	}
}
\caption{A robot (green crossed circle, green curve of its previous $13$ seconds) measures a room with a Kinect depth sensor. The exact inverse sensor model is compared with an approximate algorithm designed for the Kinect depth sensor. The algorithms update the grid cells from uncertain (grey, $0.5$ occupancy probability) to either occupied (black, $1$ occupancy probability) or free (white, $0$ occupancy probability). Measured cells have occupancy probabilities closer to $0$ or $1$ as more information becomes available, and unmeasured cells remain near $0.5$, or uncertain.
%The exact inverse sensor model approach generates a more accurate result.
}
\label{fig:NumResOccProbs}
\end{figure}

\section{Numerical Example}
\label{sec:NumRes}

% It is common practice to approximate\refeqn{InvSenModWithProbDens} with a simplified function, frequently following the structure of the approximate inverse sensor model proposed in~\cite{And09}. The main idea of this approach is that the probability of a cell being occupied ($i$) near a measurement is high (measurement likely hits this cell), ($ii$) between the robot and the measurement is low (measurement passes through this cell), and ($iii$) beyond the measurement is unchanged (the robot cannot measure through a wall/object). The function is based on intuition, not thorough mathematical derivation. In short, this approach simplifies the inverse sensor model at the cost of accuracy.

%The Kinect depth sensor is used as an example, with specifications from~\cite{KhoElb12}. With these parameters and the identical set of measurements, the exact solution is compared with a benchmark algorithm for a Kinect sensor taken from summarized as follows.

The purpose of this section is to compare the proposed exact solution to the inverse sensor model with an approximate algorithm developed for a Microsoft Kinect sensor~\cite{PirRutBisSch11,KhoElb12}, summarized as follows. The probability that the $i$-th grid cell $\mathbf{m}_i$ is occupied conditioned on the measurement ray $z_{t,l}$ at the pose $X_t$ is the continuous function
\begin{align*}
%\label{eqn:ISM_Approx_1}
&P(\mathbf{m}_i|z_{t,l},X_t)\nonumber\\&=\begin{cases}
0.3+(\frac{k}{\sigma\sqrt{2\pi}}+0.2)e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ z_{t,l}\leq \hat z_{l,i},
\\
0.5+\frac{k}{\sigma\sqrt{2\pi}}e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ z_{t,l}>\hat z_{l,i},
\end{cases}
\end{align*}
which is based on the expected distance to the cell $\hat z_{l,i}$ with parameters $k=\sigma=0.6$. This follows the structure of the approximate inverse sensor model proposed in~\cite{And09}. The main idea of this approach is that the probability of a cell being occupied (i) near a measurement is high (measurement likely hits this cell), (ii) between the robot and the measurement is low (measurement passes through this cell), and (iii) beyond the measurement is unchanged (the robot cannot measure through a wall/object).

Then, these probabilities are combined in a weighted fashion such that all measurements rays of scan $Z_t$ simultaneously update the same grid cell in a log-odds format,
\begin{align*}
%\label{eqn:ISM_Approx_2}
&\log\left(\frac{P(\mathbf{m}_i|Z_{t},X_t)}{1-P(\mathbf{m}_i|Z_{t},X_t)}\right)
\nonumber\\&=
\frac1{\sum_{z_{t,l}\in\mathbf{m}_i}\hat z_{l,i}}\sum_{z_{t,l}\in\mathbf{m}_i}\log\left(\frac{P(\mathbf{m}_i|z_{t,l},X_t)}{1-P(\mathbf{m}_i|z_{t,l},X_t)}\hat z_{l,i}\right).
\end{align*}

\subsection{Computation Time}
The proposed ray inverse sensor model is designed to maintain the accuracy of the exact solution with substantial computational improvements.
A $1$-dimensional example is simulated on a Mac desktop computer with $11$ grid cells, $10$ of which are visible to the robot, where the same depth measurement is used for all occupancy grid algorithms.
The computational time of the exact solution is computed in the conventional manner according to \refeqn{InvSenModWithProbDens} in $2.3314$ seconds to obtain the same solution of \refeqn{RayISMAnswer} from \refeqn{Unnormalized} and \refeqn{allEta} in just $0.0027$ seconds, which is $874.77$ times faster.
The approximate ray inverse sensor model completes in $0.0011$ seconds, but this algorithm does not follow a forward sensor model, so the probabilities of the map may be inaccurate.

\subsection{Trajectory Map Building}
In this numerical example, we consider a robot in a two-dimensional environment composed of ten wall edges, and the robot follows a figure-eight curve, then turns around and completes the same curve in the reverse direction.
The map is composed of $100\times100$ square grid cells with edge length of $1$ cm.
A forward sensor model $p(z_t|m,X_t)$~\cite{ThrBurFox05} is constructed according to the specifications of the Kinect sensor~\cite{KhoElb12}, and a pseudo-random measurement scan is sampled at each time step via the inverse transform sampling~\cite{DevBK86}. The same set of measurements updated each second are used with both occupancy grid mapping algorithms to construct the map.

The resulting maps are illustrated in Figure \ref{fig:NumResOccProbs} for both algorithms, where it is shown that the proposed algorithm yields a substantially more accurate and clear map with less uncertainty. To quantify the degree of map uncertainty, we define the entropy of the map as 
\begin{align*}
&H(P(m|X_{1:t},Z_{1:t}))=\nonumber\\
&-\sum_{i=1}^n\big\{P(\mathbf{m}_i|X_{1:t},Z_{1:t})\log P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\
&+(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\log(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\big\},
\end{align*}
which is maximized when the probability of occupancy is $0.5$ for all cells (more uncertain), and it is minimized when they are either $0$ or $1$ (less uncertain).

The change of the map entropy over time, and the entropy of the completed maps, for both methods are depicted in Figure \ref{fig:NumResOccH}. The subfigure (a) illustrates that the proposed exact inverse sensor model exhibits rapid decreases of entropies, and smaller entropies always. The resulting terminal map obtained from the proposed approach, shown in the subfigure (b), has less uncertainty than (c) constructed by the approximate model. In short, the proposed approach is more efficient at extracting information about the environment from the same set of set measurements.

%The occupancy grids generated by the exact inverse sensor model yield clearer maps with less uncertainty. The entropy of the map from the first measurement to the last is less with the exact model, where sudden decreases correspond to sharp turns in the robot motion. During these turns, the field of view changes greatly, so the information gain from the measurements is large, and therefore the decrease in map uncertainty is also large.

\begin{figure}
\centerline{
	\subfigure[Entropy change]% (black: exact model, green: approx.)]
		{\includegraphics[height=0.3\columnwidth]{EntropyEvolution.png}
	}
	\hspace*{-.075\columnwidth}
	\subfigure[$t=50$ (exact)]
		{\hspace*{0.3cm}\includegraphics[height=0.3\columnwidth]{Compare_ISM_H.png}\hspace*{0.3cm}
	}
	\hspace*{-.075\columnwidth}
	\subfigure[$t=50$  (approx.)]
		{\hspace*{0.3cm}\includegraphics[height=0.3\columnwidth]{Compare_Approx_H.png}\hspace*{0.3cm}
	}
}
\caption{Entropy serves as a measure of uncertainty of the occupancy grid, where more blue regions are more certain, and red regions are more uncertain. In (a), the entropy with the exact solution (black) decreases faster than the entropy with the approximate solution (green) over $20$ seconds. The uncertainty is always less when the exact solution is applied, where the final map ($50$ seconds) with the approximate solution has more uncertainty.}\label{fig:NumResOccH}
\end{figure}

\section{Preliminary Experimental Results}
\label{sec:ExpRes}
	% Kinect result

%A preliminary experiment shows how the exact inverse sensor model is successfully implemented with the Kinect sensor.

The proposed algorithm is also implemented to range measurements obtained from a Kinect sensor. The map is composed of $200\times200$ square grid cells with edge length of $1$ cm. The parameters for the Kinect sensor are obtained from~\cite{KhoElb12}, and the OpenKinect library is used on a Mac computer to extract range measurements. The Kinect sensor faces walls underneath a work desk, as illustrated in Figures \ref{fig:ExpRes}.(a) and (b). The resulting occupancy grid map and entropy map are obtained from both the proposed and approximate algorithms after a single scan~\cite{PirRutBisSch11,KhoElb12}.

%\begin{figure}
%\centerline{
%	\subfigure[Test Setup]
%		{\includegraphics[trim={1.5cm 0 1.5cm 0},clip,height=0.3\columnwidth]{TestSetup.jpg}
%	}
%	\subfigure[Kinect View]
%		{\includegraphics[height=0.3\columnwidth]{KinectView.jpg}
%	}
%}
%\centerline{
%	\subfigure[Occupancy map (exact)]
%		{\includegraphics[trim={0 0 0 3cm},clip,height=0.34\columnwidth]{KinectDeskISM.png}
%	} 
%	\hfill
%	\subfigure[Occupancy map (approx.)]
%		{\includegraphics[trim={0 0 0 3cm},clip,height=0.34\columnwidth]{KinectDeskApprox.png}
%	}
%}
%\centerline{
%	\subfigure[Entropy map (exact)]
%		{\includegraphics[trim={0 0 0 3cm},clip,height=0.34\columnwidth]{KinectDeskISM_H.png}
%	}
%	\hfill
%	\subfigure[Entropy map (approx.)]
%		{\includegraphics[trim={0 0 0 3cm},clip,height=0.34\columnwidth]{KinectDeskApprox_H.png}
%	}
%}
%\caption{The Kinect captures depth measurements of the underside of a metal desk. The measurement scan serves to update two occupancy grid mapping schemes. The proposed approach, based on the the exact solution of the inverse sensor model, provides an update with substantially less uncertainty than the approximate model update.}\label{fig:ExpRes}
%\end{figure}
\begin{figure}
\centerline{
	\subfigure[Test Setup]
		{\includegraphics[height=0.3\columnwidth]{TestSetup.jpg}
	}
	\subfigure[Exact Occupancy]
		{\includegraphics[height=0.3\columnwidth]{KinectDeskISM.png}
	} 
	\subfigure[Exact Entropy]
		{\includegraphics[height=0.3\columnwidth]{KinectDeskISM_H.png}
	}
}
\centerline{
	\subfigure[Kinect View]
		{\includegraphics[height=0.3\columnwidth]{KinectView.jpg}
	}
	\subfigure[Approx. Occupancy]
		{\includegraphics[height=0.3\columnwidth]{KinectDeskApprox.png}
	}
	\subfigure[Approx. Entropy]
		{\includegraphics[height=0.3\columnwidth]{KinectDeskApprox_H.png}
	}
}
\caption{The Kinect captures depth measurements of the underside of a metal desk. The measurement scan serves to update two occupancy grid mapping schemes. The proposed approach, based on the the exact solution of the inverse sensor model, provides an update with substantially less uncertainty than the approximate model update.}\label{fig:ExpRes}
\end{figure}

The proposed approach uses all the information from the measurement scans accurately. After the first scan, the free space has a substantially lower probability of occupancy. Furthermore, the edges of the objects captured by the Kinect are well defined with the proposed approach with an edge thickness of roughly one or two cells, but two or three cells thick with the conventional approach. Additionally, the decrease in entropy is greater with the proposed approach at $-3487$, compared with $-557$ with the approximate model.

\section{Conclusions}

Occupancy grid mapping produces a probabilistic map of occupied space, but each probability is approximated in practice due to the perceived complications of the exact solution.
This paper provides an exact solution to this complicated probability problem using cell occlusion in the forward sensor model. This yields a substantially simpler inverse sensor model, which avoids a potentially harmful Markov assumption that commonly appears in the log-odd ratio formations. The main contribution of this paper is that all of the information available in the measurements and the a priori estimate is extracted and integrated synergistically to generate a more accurate a posteriori map. The numerical and experimental results show that maps using this exact approach are substantially improved. %Future study includes developing the presented experimental results further for mobile robotic mapping, and utilizing the proposed exact inverse sensor model into SLAM and autonomous explorations.

\begin{appendix}
\label{append}

\subsection{Proof of Proposition 1}

\paragraph{Unnormalized Reduced Map Inverse Sensor Model}
For the reduced map of the $l$-th ray, the unnormalized probability corresponding to terms without $\eta_{t,l}$ in \refeqn{InvSenModWithProbDens} is %given by
\begin{align}
\tilde P&(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\
& = \sum_{r\in\mathcal{M}_{k}} p(z_{t,l}|r,X_t) P(r|X_{1:t-1},Z_{1:t-1}).\label{eqn:tildePlk}
\end{align}
Recall $\mathcal{M}_k$ corresponds to the set of maps where the $k$-th cell is occupied. Define a subset $\mathcal{N}_{i,k}\subset \mathcal{M}_k$ for $1\leq i\leq k$ be the set of maps where the $i$-th cell is the first occupied cell. More explicitly, 
\begin{align*}
\mathcal{N}_{i,k} & = \{ r\in\mathcal{M}_k| \mathbf{r}_{l,i+}\}\\
&= \{ r\in\mathcal{M}_k| \mathbf{r}_{l,1}=0,\ldots, \mathbf{r}_{l,i-1}=0, \mathbf{r}_{l,i}=1,
\mathbf{r}_{l,k}=1\}.
\end{align*}
Then, $\mathcal{M}_k$ can be written as $\mathcal{M}_k =\bigcup_{i=1}^{k} \mathcal{N}_{i,k}$. Using this, the summation over $\mathcal{M}_k$ at \refeqn{tildePlk} can be decomposed of the summation over each $\mathcal{N}_{i,k}$ to obtain
\begin{align*}
\tilde P&(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\
& = \sum_{i=1}^k\braces{\sum_{r\in\mathcal{N}_{i,k}} p(z_{t,l}|r,X_t) P(r|X_{1:t-1},Z_{1:t-1})}.
\end{align*}
This is motivated by the fact that the forward sensor model $p(z_{t,l}|r,X_t)$ is identical for all maps in $\mathcal{N}_{i,k}$, such that $p(z_{t,l}|r,X_t)=p(z_{t,l}|\mathbf{r}_{l,i+},X_t)$ and it is moved left of the summation to obtain
\begin{align}
\tilde P&(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\
& = \sum_{i=1}^k \braces{p(z_{t,l}|\mathbf{r}_{l,i+},X_t) \sum_{r\in\mathcal{N}_{i,k}} P(r|X_{1:t-1},Z_{1:t-1})}.\label{eqn:tildePlk0}
\end{align}
The last term of the above expression corresponds to the a priori probability of $\mathcal{N}_{i,k}$. When $i<k$, it is given by
\begin{align}
\sum_{r\in\mathcal{N}_{i,k}}  &P(r|X_{1:t-1},Z_{1:t-1}) = \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
 \times &P({\mathbf{r}}_{l,i}|X_{1:t-1},Z_{1:t-1})P({\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1}),\label{eqn:PNik}
\end{align}
and when $i=k$, 
\begin{align}
\sum_{r\in\mathcal{N}_{k,k}} & P(r|X_{1:t-1},Z_{1:t-1})= \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
&\quad \times P({\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1}).\label{eqn:PNkk}
\end{align}
Substituting \refeqn{PNik} and \refeqn{PNkk} into \refeqn{tildePlk0}, we obtain \refeqn{Unnormalized}.

%%Consider the $l$-th measurement ray at time $t$. The $k$-th cell of the reduced map $r$ is denoted $\mathbf{r}_{l,k}$, where $k$ corresponds to cells from map $m$, and is ordered by increasing distance from pose $X_t$.
%%Define $P(\bar{\mathbf{r}}_0)=1$ and $p(z_{t,l}|X_t,\mathbf{r}_0)=0$ for convenience. Then, the probability of occupancy of the first three cells are
%%\begin{align}
%%\label{eqn:tildePk1}
%%&\tilde P(\mathbf{r}_{l,1}|X_t,z_{t,l})\nonumber\\&=\left\{\prod_{j=0}^0P(\bar{\mathbf{r}}_{l,j})\right\}p(z_{t,l}|X_t,\mathbf{r}_{l,1})P(\mathbf{r}_{l,1})
%%,
%%\\
%%&\tilde P(\mathbf{r}_{l,2}|X_t,z_{t,l})\nonumber\\&=
%%P(\mathbf{r}_{l,2})\left\{\prod_{j=0}^0P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,1})p(z_{t,l}|X_t,\mathbf{r}_{l,1})\nonumber\\&\quad+\left\{\prod_{j=0}^1P(\bar{\mathbf{r}}_{l,j})\right\}p(z_{t,l}|X_t,\mathbf{r}_{l,2})P(\mathbf{r}_{l,2})
%%,
%%\\
%%&\tilde P(\mathbf{r}_{l,3}|X_t,z_{t,l})\nonumber\\&=
%%P(\mathbf{r}_{l,3})\sum_{i=1}^2
%%\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,i})p(z_{t,l}|\mathbf{r}_{l,i},X_t)\nonumber\\
%%&+\left\{\prod_{j=0}^2P(\bar{\mathbf{r}}_{l,j})\right\}p(z_{t,l}|X_t,\mathbf{r}_{l,3})P(\mathbf{r}_{l,3})
%%,
%%\end{align}
%%respectively. This is generalized to the arbitrary $k$-th cell as
%%\begin{align}
%%\label{eqn:GenSol}
%%&\tilde P(\mathbf{r}_k|X,z)
%%\nonumber\\&=P(\mathbf{r}_{l,k})\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i})
%%\nonumber\\&\quad+
%%\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j})\right\}p(z_{t,l}|\mathbf{r}_{l,k},X_t)P(\mathbf{r}_{l,k}),
%%\end{align}
%%where special cases when the lower bound of a summation is greater than the upper bound results in a summation value of $0$.
%%
%%\emph{Proof by Induction of the Reduced Inverse Sensor Model \refeqn{GenSol}}
%%Consider the base case when $k=1$ such that the $k$-th cell would block all others. This corresponds to $\tilde P(\mathbf{r}_{l,1}|X_t,z_{t,l})=p(z_{t,l}|X_t,\mathbf{r}_{l,1})P(\mathbf{r}_{l,1})$, which is equivalent to \refeqn{tildePk1}.
%%Next, the inductive step is solved for the $k+1$ index,
%%\begin{align}
%%\label{eqn:GenSol_InductiveStep}
%%\tilde P(\mathbf{r}_{l,k+1}|X_t,z_{t,l})
%%%&=
%%%P(\mathbf{r}_{k+1})\sum_{r:r({k+1})=\mathbf{r}_{k+1}}p(z|X,r)P(r)\nonumber
%%%\\
%%%&=P(\mathbf{r}_{k+1})\{
%%%P(\mathbf{r}_1)\sum_{\substack{r:r(1)=\mathbf{r}_1\\r({k+1})=\mathbf{r}_{k+1}}}p(z|X,r)P(r|\mathbf{r}_1,\mathbf{r}_{k+1})
%%%\nonumber
%%%\\
%%%&\quad
%%%+P(\bar{\mathbf{r}}_1)\sum_{\substack{r:r(1)=\bar{\mathbf{r}}_1\\r({k+1})=\mathbf{r}_{k+1}}}p(z|X,r)P(r|\bar{\mathbf{r}}_1,\mathbf{r}_{k+1})\}
%%%\nonumber
%%%\\
%%&=P(\mathbf{r}_{l,k+1})\bigg\{
%%P(\bar{\mathbf{r}}_{l,0})P(\mathbf{r}_{l,1})p(z_{t,l}|X_t,\mathbf{r}_{l,1})+\nonumber\\&\quad P(\bar{\mathbf{r}}_{l,0})P(\bar{\mathbf{r}}_{l,1})P(\mathbf{r}_{l,2})p(z_{t,l}|X_t,\mathbf{r}_{l,2})+\ldots\nonumber
%%\\
%%&\quad +\left(
%%P(\bar{\mathbf{r}}_{l,0})P(\bar{\mathbf{r}}_{l,1})\ldotsP(\bar{\mathbf{r}}_{l,k-1})
%%\right)\nonumber
%%\\
%%&\quad\times P(\mathbf{r}_{l,k})p(z_{t,l}|X_t,\mathbf{r}_{l,k})\bigg\}\nonumber
%%\\
%%&\quad
%%+\left(
%%P(\bar{\mathbf{r}}_{l,0})P(\bar{\mathbf{r}}_{l,1})\ldotsP(\bar{\mathbf{r}}_{l,k})
%%\right)\nonumber
%%\\
%%&\quad\times p(z_{t,l}|X_t,\mathbf{r}_{l,k+1})P(\mathbf{r}_{l,k+1})\nonumber
%%\\
%%&=P(\mathbf{r}_{l,k+1})\sum_{i=1}^{k}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}\nonumber
%%\\
%%&\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i})\nonumber\\&\quad
%%+
%%\left\{\prod_{j=0}^{k}P(\bar{\mathbf{r}}_{l,j})\right\}\nonumber
%%\\
%%&\quad\times p(z_{t,l}|X_t,\mathbf{r}_{l,k+1})P(\mathbf{r}_{l,k+1})
%%.
%%\end{align}

\paragraph{Complement of the Unnormalized Reduced Map Inverse Sensor Model}
An analytic expression for the complement of the unnormalized inverse sensor model is also required  for obtaining the normalizer of the $l$-th ray at time $t$, namely $\eta_{t,l}$. Let $\bar{\mathcal{M}}_k$ be the set of maps where the $k$-th cell is unoccupied, i.e., $\bar{\mathcal{M}}_k = \{ r\in\{0,1\}^{n_{r,l}}\,|\, \mathbf{r}_{l,k}=0\}$. Similar to \refeqn{tildePlk},
\begin{align}
\tilde P&(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\
& = \sum_{r\in\bar{\mathcal{M}}_{k}} p(z_{t,l}|r,X_t) P(r|X_{1:t-1},Z_{1:t-1}).\label{eqn:tildePbarlk}
\end{align}

Let $\bar{\mathcal{N}}_{i,k}\subset \bar{\mathcal{M}}_k $ for $1\leq i\leq n_{r,l}$ and $i\neq k$ be the set of maps where the $i$-th cell is the first occupied cell. More explicitly, 
\begin{align*}
\bar{\mathcal{N}}&_{i,k}\nonumber\\&=\{r\in\bar{\mathcal{M}}_k\,|\, \mathbf{r}_{l,1}=0,\ldots,\mathbf{r}_{l,i-1}=0,
\mathbf{r}_{l,i}=1,\mathbf{r}_{l,k}=0\}.
\end{align*}
Then, we have $\bar{\mathcal{M}}_k=\bigcup_{\substack{i=1\\i\neq k}}^{n_{r,l}} \bar{\mathcal{N}}_{i,k}$. Note that the forward sensor model $p(z_{t,l}|r,X_t)$ is identical for any maps in $\bar{\mathcal{N}}_{i,k}$, such that $p(z_{t,l}|r,X_t)=p(z_{t,l}|\mathbf{r}_{l,i+},X_t)$. Similar to \refeqn{tildePlk0},
\begin{align}
\tilde P&(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\
& = \sum_{\substack{i=1\\i\neq k}}^{n_{r,l}} \braces{p(z_{t,l}|\mathbf{r}_{l,i+},X_t) \sum_{r\in\bar{\mathcal{N}}_{i,k}} P(r|X_{1:t-1},Z_{1:t-1})},\label{eqn:tildePbarlk0}
\end{align}
where the last term corresponds to the a priori probability of $\bar{\mathcal{N}}_{i,k}$. When $i<k$, it is given by
\begin{align}
&\sum_{r\in\bar{\mathcal{N}}_{i,k}} P(r|X_{1:t-1},Z_{1:t-1}) = \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
&\quad \times P({\mathbf{r}}_{l,i}|X_{1:t-1},Z_{1:t-1})P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1}),\label{eqn:PbarNik1}
\end{align}
and when $k<i$,
\begin{align}
\sum_{r\in\bar{\mathcal{N}}_{i,k}} & P(r|X_{1:t-1},Z_{1:t-1}) = \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
&\quad \times P({\mathbf{r}}_{l,i}|X_{1:t-1},Z_{1:t-1}).\label{eqn:PbarNik2}
\end{align}
Substituting \refeqn{PbarNik1} and \refeqn{PbarNik2} into \refeqn{tildePbarlk0}, we obtain
\begin{align}
\tilde P &(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})
=P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\
&\quad\times \bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]
\nonumber
\\
&\quad
+
\bigg[\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg],\label{eqn:tildePbar}
\end{align}
where $p(z_{t,l}|\mathbf{r}_{(n_r+1)+},X_t)$ corresponds to the forward sensor model of maximum reading and $P(\mathbf{r}_{l,n_{r,l}+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience for the empty map case.

% ADD: where it is considered $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$

%%The unnormalized complement of the $k$-th cell inverse sensor model, namely $\tilde P(\bar{\mathbf{r}}_k|X,z)$, is required to solve for $\eta$ based on \refeqn{allEta}.
%%Consider the complement of the occupancy of the first and second cells, i.e.,
%%\begin{align}
%%\label{eqn:InvSenModknot1}
%%\tilde P(\bar{\mathbf{r}}_{l,1}|X_t,z_{t,l})
%%&=P(\bar{\mathbf{r}}_{l,1})
%%\bigg\{P(\mathbf{r}_{l,2})
%%p(z_{t,l}|X_t,\mathbf{r}_{l,2})\nonumber\\&\quad
%%+P(\bar{\mathbf{r}}_{l,2})P(\mathbf{r}_{l,3})p(z_{t,l}|X_t,\mathbf{r}_{l,3})+\ldots\nonumber\\&\quad+\prod_{j=2}^{n_r-1}P(\bar{\mathbf{r}}_{l,j})P(\mathbf{r}_{l,n_r})p(z_{t,l}|X_t,\mathbf{r}_{l,n_r})
%%\bigg\}\nonumber
%%\\
%%&=\sum_{i=2}^{n_r}
%%\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})
%%\right\}P(\mathbf{r}_{i})p(z_{t,l}|X_t,\mathbf{r}_{l,i}).
%%\\
%%\label{eqn:InvSenModknot2}
%%\tilde P(\bar{\mathbf{r}}_{l,2}|X_t,z_{t,l})
%%%&=\sum_{r:r(2)=\bar{\mathbf{r}}_2}p(z|X,r)P(r)\nonumber
%%%\\
%%%&=P(\bar{\mathbf{r}}_2)\sum_{r:r(2)=\bar{\mathbf{r}}_2}p(z|X,r)P(r|\bar{\mathbf{r}}_2)\nonumber
%%%\\
%%%&=P(\bar{\mathbf{r}}_2)\left\{P(\mathbf{r}_1)
%%%\sum_{\substack{r:r(1)=\mathbf{r}_1\\r(2)=\bar{\mathbf{r}}_2}}p(z|X,r)P(r|\mathbf{r}_1,\bar{\mathbf{r}}_2)
%%%+P(\bar{\mathbf{r}}_1)\sum_{\substack{r:r(1)=\bar{\mathbf{r}}_1\\r(2)=\bar{\mathbf{r}}_2}}p(z|X,r)P(r|\bar{\mathbf{r}}_1,\bar{\mathbf{r}}_2)
%%%\right\}\nonumber
%%%\\
%%%&=P(\bar{\mathbf{r}}_2)
%%%\{
%%%P(\mathbf{r}_1)p(z|X,\mathbf{r}_1)
%%%+P(\bar{\mathbf{r}}_1)
%%%[
%%%P(\mathbf{r}_3)\sum_{\substack{r:r(1)=\bar{\mathbf{r}}_1\\r(2)=\bar{\mathbf{r}}_2\\r(3)=\mathbf{r}_3}}p(z|X,r)P(r|\bar{\mathbf{r}}_1,\bar{\mathbf{r}}_2,\mathbf{r}_3)\nonumber
%%%\\
%%%&\quad
%%%+
%%%P(\bar{\mathbf{r}}_3)\sum_{\substack{r:r(1)=\bar{\mathbf{r}}_1\\r(2)=\bar{\mathbf{r}}_2\\r(3)=\bar{\mathbf{r}}_3}}p(z|X,r)P(r|\bar{\mathbf{r}}_1,\bar{\mathbf{r}}_2,\bar{\mathbf{r}}_3)
%%%]
%%%\}\nonumber
%%%\\
%%&=P(\bar{\mathbf{r}}_{l,2})
%%\bigg\{
%%P(\mathbf{r}_{l,1})p(z_{t,l}|X_t,\mathbf{r}_{l,1})\nonumber\\&\quad
%%+
%%P(\bar{\mathbf{r}}_{l,1})P(\mathbf{r}_{l,3})p(z_{t,l}|X_t,\mathbf{r}_{l,3})\nonumber\\&\quad
%%+P(\bar{\mathbf{r}}_{l,1})P(\bar{\mathbf{r}}_{l,3})P(\mathbf{r}_{l,4})\nonumber\\&\qquad\times p(z_{t,l}|X_t,\mathbf{r}_{l,4})+\ldots\nonumber
%%\\
%%&\quad+\prod_{j=1,j\neq2}^{n_r-1}P(\bar{\mathbf{r}}_{l,j})P(\mathbf{r}_{l,n_r})p(z_{t,l}|X_t,\mathbf{r}_{l,n_r})
%%\bigg\}\nonumber
%%\\
%%&=
%%P(\bar{\mathbf{r}}_{l,2})\sum_{i=1}^{1}
%%\left\{
%%\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}\nonumber\\&\quad\times P(\mathbf{r}_{l,i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})\nonumber\\&\quad+\sum_{i=3}^{n_r}
%%\left\{
%%\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}\nonumber\\&\quad\times P(\mathbf{r}_{l,i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})
%%.
%%\end{align}
%%
%%
%%
%%In general, the $k$-th unnormalized inverse sensor model complement is
%%\begin{align}
%%\label{eqn:ComplGenSol}
%%\tilde P&(\bar{\mathbf{r}}_{l,k}|X,z)\nonumber\\&=P(\bar{\mathbf{r}}_{l,k})\sum_{i=1}^{k-1}
%%\left\{
%%\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})\nonumber\\&\quad
%%+\sum_{i=k+1}^{n_r}
%%\left\{
%%\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})
%%,
%%\end{align}
%%where special cases when the lower bound of a summation is greater than the upper bound results in a summation value of $0$.
%%
%%\paragraph{Proof by Induction of the Complement of the Reduced Inverse Sensor Model \refeqn{ComplGenSol}}
%%The base case for the occupancy probability is obtain directly from substituting $k=1$ into \refeqn{ComplGenSol} to obtain $\tilde P(\bar{\mathbf{r}}_1|X,z)=\sum_{i=2}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})$.
%%The inductive step is solved for the $k+1$ index,
%%\begin{align}
%%\tilde P&(\bar{\mathbf{r}}_{l,k+1}|X_t,z_{t,l})\nonumber\\
%%%&=P(\bar{\mathbf{r}}_{k+1})
%%%\{
%%%[
%%%P(\mathbf{r}_1)p(z|X,\mathbf{r}_1)
%%%+
%%%P(\bar{\mathbf{r}}_1)P(\mathbf{r}_2)p(z|X,\mathbf{r}_2)
%%%+P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)P(\mathbf{r}_3)p(z|X,\mathbf{r}_3)+\ldots\nonumber
%%%\\
%%%&\quad
%%%+P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)\ldotsP(\bar{\mathbf{r}}_{k-1})P(\mathbf{r}_k)p(z_{t,l}|\mathbf{r}_k,X_t)
%%%]\nonumber
%%%\\
%%%&\quad
%%%+
%%%P(\bar{\mathbf{r}}_{k+1})^{-1}[
%%%P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)\ldotsP(\bar{\mathbf{r}}_{k+1})P(\mathbf{r}_{k+2})p(z|X,\mathbf{r}_{k+2\ldots...
%%%\nonumber
%%%\\
%%%&\quad
%%%+P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)\ldotsP(\bar{\mathbf{r}}_{n_r-1})P(\mathbf{r}_{n_r})p(z|X,\mathbf{r}_{n_r})
%%%]
%%%\}\nonumber
%%%\\
%%&=
%%P(\bar{\mathbf{r}}_{l,k+1})\nonumber
%%\\
%%&\quad
%%\times\bigg[
%%\left\{\prod_{j=0}^{0}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,1})p(z_{t,l}|X_t,\mathbf{r}_{l,1})\nonumber
%%\\
%%&\quad
%%+
%%\left\{\prod_{j=0}^{1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,2})p(z_{t,l}|X_t,\mathbf{r}_{l,2})\nonumber
%%\\
%%&\quad
%%+\left\{\prod_{j=0}^{2}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,3})p(z_{t,l}|X_t,\mathbf{r}_{l,3})+\ldots\nonumber
%%\\
%%&\quad
%%+\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,k})p(z_{t,l}|\mathbf{r}_{l,k},X_t)
%%\bigg]\nonumber
%%\\
%%&\quad
%%+
%%\bigg[
%%\left\{\prod_{j=0}^{k+1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,k+2})p(z_{t,l}|X_t,\mathbf{r}_{l,k+2})+\ldots\nonumber
%%\\
%%&\quad
%%+\left\{\prod_{j=0}^{n_r-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,n_r})p(z_{t,l}|X_t,\mathbf{r}_{l,n_r})
%%\bigg]
%%\nonumber
%%\\
%%&=P(\bar{\mathbf{r}}_{l,k+1})\sum_{i=1}^{k}
%%\left\{
%%\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})\nonumber
%%\\
%%&\quad
%%+\sum_{i=k+2}^{n_r}
%%\left\{
%%\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,i})p(z_{t,l}|X_t,\mathbf{r}_{l,i})
%%.
%%\end{align}

\paragraph{Normalizer}
We have 
\begin{align*}
P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})+
P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=1.
\end{align*}
Since they share the same normalizer $\eta_{t,l}$, this implies 
\begin{align*}
&\eta_{t,l}=\frac1{\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})+
\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}.
\end{align*}
Substituting \refeqn{Unnormalized} and \refeqn{tildePbar}, and rearranging, we obtain \refeqn{allEta}.
%
%\begin{align}
%\eta_{t,l}
%&=\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
%\end{align}
%which is independent of the reduced cell index $k$.

\end{appendix}

\bibliography{BibSources}
\bibliographystyle{IEEEtran}

\end{document}

\end{document}

### 2018-06-09 Conference/ACC 2016/Final Abridged/ACC16.1.tex
added Conference folder

--- 2018-06-09 Conference/ACC 2016/Final Abridged/ACC16.1 IRA.tex
@@
 \thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 {\tt \{evankaufman,tylee\}@gwu.edu}}
-\thanks{Zhuming Ai, Ira S. Moskowitz, Information Management \& Decision Architectures Branch, Code, 5580,  Naval Research Laboratory,  Washington, DC 20375}
+\thanks{Zhuming Ai, Ira S. Moskowitz, Information Management \& Decision Architectures, U.S. Naval Research Laboratory,  Washington, DC 20375}
 \thanks{This research has been supported by the U.S.
@@
 \begin{abstract}
-Occupancy grid maps are spatial representations of environments where the space of interest is decomposed into a number of cells that are considered either occupied or free.
-This paper focuses on occupancy grid mapping which   estimates the probability of occupancy for each cell based upon range measurements from a known location.
-For a given probabilistic model of a range sensor, we propose a computationally efficient method to obtain an exact inverse sensor model.
-We use this to   construct a probabilistic mapping algorithm according to a Bayesian framework.
-Compared with the existing occupancy grid mapping techniques that rely on approximate, heuristic inverse sensor models, the our approach yields substantially more accurate maps for the same set of measurements.
-This is illustrated by numerical examples and experiments.
+Occupancy grid maps are spatial representations of environments, where the space of interest is decomposed into a number of cells that are considered either occupied or free.
+This paper focuses on occupancy grid mapping, which is to estimate the probability of occupancy for each cell based on range measurements from a known location.
+For a given probabilistic model of a range sensor, we propose a computationally efficient method to obtain an exact inverse sensor model, and it is utilized to construct a probabilistic mapping algorithm according to the Bayesian framework.
+Compared with the existing occupancy grid mapping techniques that rely on approximate, heuristic inverse sensor models, the proposed approach yields substantially more accurate maps for the same set of measurements.
+These are illustrated by numerical examples and experiments.
 \end{abstract}
@@
 
-In the context of robotics, mapping is done to    to generate maps which represent the environment in close proximity of a robot.
+Robotic mapping is the process of generating maps, which represent the environment in close proximity of a robot. 
+%Mapping in the context of robotics is to generate maps, which represent the environment in close proximity of a robot. 
 This problem is crucial to understanding the surrounding regions when a robot enters an uncertain environment.
@@
 
-This paper focuses on the popular occupancy grid representation, where the map space is decomposed into evenly-spaced grid cells, each cell considered as  occupied or free.
-The \emph{inverse sensor model} is a probabilistic model  of occupancy for a grid cell based on range measurements and the configuration of the mobile robot (which is of fundamental importance with occupancy grid mapping).
+This paper focusses on the popular occupancy grid representation, where the map space is decomposed into evenly-spaced grid cells, considered as either occupied or free.
+The \emph{inverse sensor model} is the probability of occupancy for a grid cell based on range measurements and the configuration of the mobile robot, which is of fundamental importance with occupancy grid mapping.
 However, the exact solution to the inverse sensor model has not been utilized in occupancy grid mapping, because the computational requirements are too cumbersome to compute the exact model in real time~\cite{ThrBurFox05}.
@@
 %Due to the computational limitations, two common approaches are applied to obtain a function close to the inverse sensor model.
-In the early occupancy grid mapping applications based on sonar sensors, \emph{ad hoc} techniques to obtain an inverse sensor model were developed based off highly inaccurate  reflected sound waves~\cite{MorElf85,Elf89}.
+In the early occupancy grid mapping applications based on sonar sensors, \emph{ad hoc} techniques to obtain an inverse sensor model were developed based off highly-innacurate bouncing sound waves~\cite{MorElf85,Elf89}.
 %These occupancy grid mapping schemes employ \emph{ad hoc} techniques to obtain the inverse sensor model.
@@
 
-The other popular approach to obtain an approximate inverse sensor model is to simulate maps, robot poses, and measurements, in conjunction with a learning algorithm to approximate the inverse sensor model~\cite{ThrBurFox05,Thr01}.
-These approaches are undesirable in practice due to the complexities associated to implementing a learning algorithm.
+The other popular approach to obtain an approximate inverse sensor model is to simulate maps, robot poses, and measurements and use a learning algorithm to approximate the inverse sensor model~\cite{ThrBurFox05,Thr01}.
+These approaches are undesirable in practice due to complexities associated to implementing a learning algorithm.
 For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples, or how many samples are required to obtain a reasonable approximation.
@@
 This paper proposes a computationally efficient algorithm to construct the exact inverse sensor model.
-More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, our algorithm yields the a posteriori probability of occupancy of all the cells within the area covered by the range sensor via the range measurements.
-The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancy of the other cells blocked by it is irrelevant to the forward sensor model, and this approach is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model.
-Furthermore, our proposed approach integrates the a priori probability of occupancy and multiple range measurements according to a Bayesian framework to obtain more accurate maps.
+More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, this algorithm yields the a posteriori probability of occupancy of all the cells within the area covered by the range sensor from the range measurements.
+The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancy of the other cells blocked by it is irrelevant to the forward sensor model, and this model is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model.
+Furthermore, the proposed approach integrates a priori probability of occupancy and multiple range measurements according to the Bayesian framework to obtain more accurate maps.
 As such, it contrasts from the existing framework based on log-odds ratios that impose an additional Markov assumption.
@@
 
-In short, the main contribution of this paper is in proposing an exact inverse sensor model and constructing a Bayesian occupancy grid mapping based upon  it.
-Compared with the current mapping algorithms based on   approximate inverse sensor models, this approach yields more accurate occupancy probabilities.
+In short, the main contribution of this paper is proposing the exact inverse sensor model and constructing a Bayesian occupancy grid mapping based on that.
+Compared with the current mapping algorithms based on an approximate inverse sensor model, this approach yields more accurate occupancy probabilities.
 This constructs substantially more accurate occupancy grid maps with less uncertainty for the same set of measurements, and these are directly illustrated by numerical examples and preliminary experimental results.
@@
 %Occupancy grid mapping produces a probabilistic map; none of the $n_m$ grid cell occupancies are completely certain.
-Occupancy grid mapping  is used to determine these probabilities based on robot poses and measurement scans.
+Occupancy grid mapping  is to determine these probabilities based on robot poses and measurement scans.
 More explicitly, let $X_{1:t}$ denote the history of poses from the initial time to the current time, $\{X_1,X_2,\ldots, X_t\}$, and let $Z_{1:t}$ be the measurement history.
-The goal is to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, commonly referred to as the \emph{inverse sensor model} for given forward sensor model $p(z|m,X)$ and the initial estimate of the map $P(m)$ with $X_{1:t}$ and $Z_{1:t}$.
+The goal is to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, commonly referred to as the \emph{inverse sensor model} for given forward sensor model $p(z_{t,l}|m,X_t)$ and the initial estimate of the map $P(m)$ with $X_{1:t}$ and $Z_{1:t}$.
 
@@
 
-A popular frameworks for updating binary random variables with static state, is to use a binary Bayesian filter using the log-odds ratio formulation.
+One of the popular frameworks of updating binary random variables with static state is with a binary Bayesian filter using the log-odds ratio formulation.
 The main idea is that instead of multiplying terms from prior and current time steps, the properties of logarithms allow these terms to be simply added, while avoiding truncation issues associated with probabilities close to $0$ or $1$~\cite{ThrBurFox05}.
-The log-odds ratio is also popular because learning techniques with forward models~\cite{Thr01},~\cite{Thr03}, or ad hoc techniques~\cite{MorElf85},~\cite{Elf89} are simpler to derive when they need not consider $P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})$ as part of the inverse sensor model, but rather only $P(\mathbf{m}_i)$, which is typically fixed and uniform for all cells.
+The log-odds ratio is also popular because learning techniques with forward models~\cite{Thr01},~\cite{Thr03} or ad hoc techniques~\cite{MorElf85},~\cite{Elf89} are simpler to derive when they need not consider $P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})$ as part of the inverse sensor model, but rather only $P(\mathbf{m}_i)$, which is typically fixed and uniform for all cells.
 
@@
 Due to computational complexity, approximate inverse sensor models have been used for occupancy grid mapping with simplifying assumptions.
-In this section, we propose an efficient algorithm to compute the exact inverse sensor model.
+In this section, we propose an algorithm to compute the exact inverse sensor model efficiently.
 %Much like the forward sensor model that reasons measurements from a known map, he inverse sensor model reasons that map from known measurements~\cite{ThrBurFox05}.
 %While the inverse sensor model varies slightly in meaning in numerous publications, two versions are derived in this paper, both based on a Bayesian framework.
-We present two versions according to a Bayesian framework.
+We present two versions according to Bayesian framework.
 The first type is referred to as the \emph{ray inverse sensor model} $P(\mathbf{m}_i|X_{1:t},Z_{1:t-1},z_{t,l})$, which only considers $l$-th ray at time $t$, i.e., one-dimensional inverse sensor model.
@@
 \end{align}
-Furthermore, to determine the normalizing constant $\eta_{t,l}$, the complement $P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in a similar manner.
+Furthermore, to determine the normalizing constant $\eta_{t,l}$, the complement $P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in the similar manner.
 Since there are $2^{n_{m}-1}$ maps in $\mathcal{M}_i$, these equations require $2^{n_m}$ terms to compute the summation for the $i$-th grid cell, and it should be repeated for other cells.
@@
 \begin{prop}
-For the $l$-th measurement ray, the a posteriori probability of the occupancy of the $k$-th cell, namely the ray inverse sensor model, is given by
+For the $l$-th measurement ray, the a posteriori probability of the occupancy of the $k$-th cell, namely the ray inverse sensor model is given by
 \begin{align}
@@
 \end{align}
-where it is considered $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading for convenience.
+where $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})\\=1$ is chosen for convenience and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading.
 The normalizer $\eta_{t,l}$ is given by
@@
 Compared with \refeqn{InvSenModWithProbDens}, where the terms of the summation should be repeated $2^{n_{r,l}}$ times \emph{per each cell of the reduced map}, the proposed expressions \refeqn{RayISMAnswer} and \refeqn{allEta} are \textit{substantially} simpler as they require $k$ summation terms for the $k$-th cell.
-In fact, if the terms inside the summations of \refeqn{Unnormalized} and \refeqn{allEta} are obtained recursively, only $n_{r,l}+1$ repetitions are required for \emph{all cells of the reduced map}, yielding an algorithm that is nearly $\mathbf{2^{n_{r,l}}}$ \textbf{times faster}.
+In fact, if \refeqn{Unnormalized} and \refeqn{allEta} are obtained recursively, requiring $O(n_{r,l}+1)$ rather than $O(n_{r,l}\times2^{n_{r,l}})$ time as previously thought for \emph{all cells of the reduced map}, this yields an algorithm that is $\mathbf{n_{r,l}\times2^{n_{r,l}}/(n_{r,l}+1)}$ \textbf{times faster}.
 
@@
 \begin{align}
-P(&\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\&%=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1},z_{t,2},\ldots,z_{t,n_z})\nonumber\\&
-=P(((((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z}).
+P&(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\&%=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1},z_{t,2},\ldots,z_{t,n_z})\nonumber\\&
+=P((\dots(((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z}).
 \end{align}

### 2018-06-09 Conference/ACC 2016/Final Unabridged/ACC16.1.tex
added Conference folder

--- 2018-06-09 Conference/ACC 2016/Final Abridged/ACC16.1.tex
@@
 For a given probabilistic model of a range sensor, we propose a computationally efficient method to obtain an exact inverse sensor model, and it is utilized to construct a probabilistic mapping algorithm according to the Bayesian framework.
-Compared with the existing occupancy grid mapping techniques that rely on approximate, heuristic inverse sensor models, the proposed approach yields substantially more accurate maps for the same set of measurements.
+Compared with the existing occupancy grid mapping techniques that rely on approximate inverse sensor models, the proposed approach yields substantially more accurate maps for the same set of measurements.
 These are illustrated by numerical examples and experiments.
@@
 
-Robotic mapping is the process of generating maps, which represent the environment in close proximity of a robot. 
-%Mapping in the context of robotics is to generate maps, which represent the environment in close proximity of a robot. 
+Mapping in the context of robotics is to generate maps, which represents the environment in close proximity of a robot.
 This problem is crucial to understanding the surrounding regions when a robot enters an uncertain environment.
-Various map representations include occupancy grids~\cite{WolSuk05}, octomaps~\cite{WurHorBenStaBur10}, or feature-based maps%in a landmark representation
+Various map representations have been applied to the mapping problem such as occupancy grids~\cite{WolSuk05}, octomaps~\cite{WurHorBenStaBur10}, or feature-based maps%in a landmark representation
 ~\cite{MonThrKolWeg02}.
@@
 
-This paper focusses on the popular occupancy grid representation, where the map space is decomposed into evenly-spaced grid cells, considered as either occupied or free.
+This paper focusses on the popular occupancy grid representation, where the map space is decomposed into evenly-spaced grid cells that are considered as either occupied or free.
 The \emph{inverse sensor model} is the probability of occupancy for a grid cell based on range measurements and the configuration of the mobile robot, which is of fundamental importance with occupancy grid mapping.
@@
 %Due to the computational limitations, two common approaches are applied to obtain a function close to the inverse sensor model.
-In the early occupancy grid mapping applications based on sonar sensors, \emph{ad hoc} techniques to obtain an inverse sensor model were developed based off highly-innacurate bouncing sound waves~\cite{MorElf85,Elf89}.
+In the early occupancy grid mapping applications based on sonar sensors, \emph{ad hoc} techniques are employed to obtain an inverse sensor model, where the bearings of depth measurements are considered uncertain because the directions of bouncing sound waves are highly inaccurate~\cite{MorElf85,Elf89}.
 %These occupancy grid mapping schemes employ \emph{ad hoc} techniques to obtain the inverse sensor model.
@@
 
-The other popular approach to obtain an approximate inverse sensor model is to simulate maps, robot poses, and measurements and use a learning algorithm to approximate the inverse sensor model~\cite{ThrBurFox05,Thr01}.
+The other popular approach to obtain an approximate inverse sensor model is to simulate maps, robot poses, and measurements and use a learning algorithm, to approximate the inverse sensor model~\cite{Thr01,ThrBurFox05}.
 These approaches are undesirable in practice due to complexities associated to implementing a learning algorithm.
@@
 
-% Abridged:
-%The paper is organized as follows.
-%The problem is formulated in Section \ref{sec:ProbForm}, and the exact solution to the occupancy grid mapping problem is derived in Section \ref{sec:ISM}.
-%A numerical result shows the improved accuracy in Section \ref{sec:NumRes}, and an experimental result with the Kinect depth sensor shows the improvement in map certainty in Section \ref{sec:ExpRes}, followed by conclusions.
+The paper is organized as follows.
+The problem is formulated in Section \ref{sec:ProbForm}, and the exact solution to the occupancy grid mapping problem is derived in Section \ref{sec:ISM}.
+A numerical result shows the improved accuracy in Section \ref{sec:NumRes}, and an experimental result with the Kinect depth sensor shows the improvement in map certainty in Section \ref{sec:ExpRes}, followed by conclusions.
 
@@
 
-Consider a range sensor that provides scans of the surrounding environment in order to identify the closest occupied space.
-The location and the direction of the sensor, referred to as the \emph{pose} at time $t$, is denoted by $X_t$.
-The range sensor returns a measurement \emph{scan} $Z_t$, composed of $n_z$ measurement \emph{rays} $z_{t,1},z_{t,2},\ldots,z_{t,n_z}$.
-The $l$-th measurement ray for $l\in\braces{1,2,\ldots,n_z}$ contains the depth (range) and bearing (direction) inside the sensor field of view (FOV) that corresponds to a circular sector.
-Given $X_t$, we assume the origin and directions of each ray are known deterministically.
+Consider a range sensor that provides scans of the surrounding environment to identify the closest occupied space.
+The location and the direction of the sensor, referred to as the \emph{pose} at time $t$ is denoted by $X_t$.
+The range sensor returns a measurement \emph{scan} $Z_t$, which is composed of $n_z$ measurement \emph{rays} $z_{t,1},z_{t,2},...,z_{t,n_z}$.
+The $l$-th measurement ray for $l\in\braces{1,2,...,n_z}$ contains the depth (range) and bearing (direction) of a small region inside the sensor field of view (FOV) that corresponds to a circular sector.
+For a given $X_t$, it is assumed that the origin and the direction of each ray is known deterministically.
 
@@
 
-Throughout this paper, we assume that the forward sensor model of the selected sensor is given.
+Throughout this paper, it is assumed that the forward sensor model of the selected sensor is given.
 This can be determined empirically or analytically.
@@
 
-Let a map $m$ be decomposed into $n_m$ evenly-spaced grid cells, where the $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,\ldots,n_m}$, that is defined as $\mathbf{m}_i=1$ when occupied, and $\mathbf{m}_i=0$ when free.
-The location and size of each grid cell is assumed known.
-Therefore, a map $m$ is defined by $\{\mathbf{m}_1,\ldots, \mathbf{m}_{n_m}\}$ ($2^{n_{m}}$ possible maps).
+Let a map $m$ be decomposed into $n_m$ evenly-spaced grid cells, where the $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,...,n_m}$, that is defined as $\mathbf{m}_i=1$ when it is occupied, and $\mathbf{m}_i=0$ when it is free.
+The location and size of each grid cell is assumed to be known.
+Therefore, a map $m$ is defined by $\{\mathbf{m}_1,\ldots, \mathbf{m}_{n_m}\}$, and there are $2^{n_{m}}$ possible maps.
 
@@
 \begin{align}
-P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i).
+P(m)=P(\mathbf{m}_1,\mathbf{m}_2,...,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i).
 \end{align}
@@
 More explicitly, let $X_{1:t}$ denote the history of poses from the initial time to the current time, $\{X_1,X_2,\ldots, X_t\}$, and let $Z_{1:t}$ be the measurement history.
-The goal is to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, commonly referred to as the \emph{inverse sensor model} for given forward sensor model $p(z_{t,l}|m,X_t)$ and the initial estimate of the map $P(m)$ with $X_{1:t}$ and $Z_{1:t}$.
+The goal is to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, commonly referred to as the \emph{inverse sensor model} for given forward sensor model $p(z|m,X)$ and the initial estimate of the map $P(m)$ with $X_{1:t}$ and $Z_{1:t}$.
 
@@
 \end{align}
-
-However, forward models are based on the entire map $m$.
-Since \refeqn{AssumptionEarly} conditioned only by the $i$-th cell, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may contain important information about the occupancy of the other cells, which should be considered in calculating the probability distribution of the measurement scan $Z_t$.
-%In short, this log-odds ratio framework is based on the assumption that neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$.
+The above equation can be justified if both sides are conditioned by the complete map $m$ instead of just the cell $\mathbf{m}_i$, since the probability distribution of the sensor is completely determined by its pose and the map through the forward sensor model.
+However, since it is conditioned only by the $i$-th cell, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may have certain information about the occupancy of the other cells, which should be considered in calculating the probability distribution of the measurement scan $Z_t$.
+In short, this log-odds ratio framework is based on the assumption that neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$.
 
@@
 
-Due to computational complexity, approximate inverse sensor models have been used for occupancy grid mapping with simplifying assumptions.
+Due to the computational complexity, approximate inverse sensor models have been used for occupancy grid mapping with simplifying assumptions.
 In this section, we propose an algorithm to compute the exact inverse sensor model efficiently.
@@
 \end{align}
-considering that $X_t$ carries no information about $m$ without $Z_t$.
+where we have used the fact that $X_t$ carries no information about $m$ without $Z_t$.
 If the current pose $X_t$ and map $m$ are known, then the measurement ray $z_{t,l}$ is independent of past poses $X_{1:t-1}$ and past measurements $Z_{1:t-1}$ to obtain
@@
 
-We propose a computational algorithm to evaluate \refeqn{InvSenModWithProbDens} efficiently. 
+In this paper, we propose a computational algorithm to evaluate \refeqn{InvSenModWithProbDens} efficiently. 
 Since the cells outside of the sensor FOV are not affected, we focus on a reduced map $r_l$ in the FOV of the $l$-th ray.
@@
 \end{align}
-where $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})\\=1$ is chosen for convenience and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading.
+where it is considered $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading for convenience.
 The normalizer $\eta_{t,l}$ is given by
@@
 Compared with \refeqn{InvSenModWithProbDens}, where the terms of the summation should be repeated $2^{n_{r,l}}$ times \emph{per each cell of the reduced map}, the proposed expressions \refeqn{RayISMAnswer} and \refeqn{allEta} are \textit{substantially} simpler as they require $k$ summation terms for the $k$-th cell.
-In fact, if \refeqn{Unnormalized} and \refeqn{allEta} are obtained recursively, requiring $O(n_{r,l}+1)$ rather than $O(n_{r,l}\times2^{n_{r,l}})$ time as previously thought for \emph{all cells of the reduced map}, this yields an algorithm that is $\mathbf{n_{r,l}\times2^{n_{r,l}}/(n_{r,l}+1)}$ \textbf{times faster}.
+In fact, if the terms inside the summations of \refeqn{Unnormalized} and \refeqn{allEta} are obtained recursively, only $n_{r,l}+1$ repetitions are required for \emph{all cells of the reduced map}, yielding an algorithm that is nearly $\mathbf{2^{n_{r,l}}}$ \textbf{times faster}.
 
@@
 
-%At the $t$-th time step, the measurement scan is the set of all rays $Z_t=\{z_{t,1},z_{t,2},\ldots,z_{t,n_z}\}$.
-
+At the $t$-th time step, the measurement scan is the set of all rays $Z_t=\{z_{t,1},z_{t,2},...,z_{t,n_z}\}$.
+One may apply the results of Proposition 1 repeatedly for each ray to obtain a two-dimensional inverse sensor model because
+\begin{align}
+P(\mathbf{m}_i|X_{1:t},Z_{1:t})&%=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1},z_{t,2},...,z_{t,n_z})\nonumber\\&
+=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1}|z_{t,2}|...|z_{t,n_z}).
+\end{align}
 %However, when there are a large number of rays, it is certainly possible that a single cell is covered by multiple rays. 
-
-%One may apply the results of Proposition 1 repeatedly for each ray to obtain a two-dimensional inverse sensor model,
-%The scan inverse sensor model is generalized naturally by applying the conditions recursively,
-The above one-dimensional ray inverse sensor model may be generalized into a two-dimensional scan inverse sensor model by updating the map sequentially for each ray, according to
-%The most natural way to generalize the ray inverse sensor model into a scan inverse sensor model is applying the conditions in a recursive way,
-\begin{align}
-P&(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\&%=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1},z_{t,2},\ldots,z_{t,n_z})\nonumber\\&
-=P((\dots(((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z}).
-\end{align}
-Alternatively, measurements from multiple rays can be integrated synergistically, based on the assumption that the measurements of rays are mutually independent, i.e., $p(z_{t,1},z_{t,2},\ldots,z_{t,n_z}|\mathbf{m}_i,X_{1:t},Z_{1:t-1})=\prod_{l=1}^{n_z}p(z_{t,l}|\mathbf{m}_i,X_{1:t},Z_{1:t-1})$.
-Let $\mathcal L_i\subset\{1,\ldots, n_z\}$ be the set of rays intersecting the $i$-th cell, $\mathbf{m}_i$. Applying Bayes' rule repeatedly, we obtain
+Alternatively, measurements from multiple rays can be integrated synergistically to obtain a more accurate estimate.
+
+%\paragraph{Bayesian Framework}
+
+%Assuming the measurement rays are mutually independent, Bayes' rule yields
+
+Here, we construct such two-dimensional inverse sensor model, referred to as the scan inverse sensor model, based on the assumption that the measurements of rays are mutually independent.
+Let $\mathcal L_i\subset\{1,\ldots, n_z\}$ be the set of rays that pass through the $i$-th cell, $\mathbf{m}_i$. Applying Bayes' rule repeatedly, we obtain
 \begin{align}
@@
 \end{align}
-where $\tilde\zeta_i,\zeta_i\in\Re$ correspond to normalizing constants.% independent of $\mathbf{m}_i$.
+where $\tilde\zeta_i,\zeta_i\in\Re$ correspond to normalizing constants independent of $\mathbf{m}_i$.
 
@@
 
-Suppose the $l$-th measurement ray intersects with the cell $\mathbf{m}_i$, where $\mathbf{r}_{l,k}$ corresponds to the cell $\mathbf{m}_i$ in the $l$-th reduced map, so $\mathbf{r}_{l,k}=\mathbf{m}_i$.
+Suppose the $l$-th measurement ray intersects with the cell $\mathbf{m}_i$, and it is the $k$-th cell of the corresponding reduced map, i.e., $\mathbf{r}_{l,k}$ corresponds to the cell $\mathbf{m}_i$ in the $l$-th reduced map, and $\mathbf{r}_{l,k}=\mathbf{m}_i$ since they represent the same cell.
 Then, we have $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})=P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$.
@@
 &\quad
-+
-\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0,j\neq k}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
++P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})^{-1}
+\nonumber\\&\quad\times 
+\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
 \end{align}
@@
 \end{align}
-which is substituted into \refeqn{ISM_Fusion} to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$.%  the complete scan inverse sensor model
+which is substituted into \refeqn{ISM_Fusion} to obtain the complete scan inverse sensor model $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$.
 
@@
 \begin{tabular}{ l }
-  for $l = 1,2,\ldots,n_z$\\
+  for $l = 1,2,...,n_z$\\
    \ \ Obtain reduced map $r_{t,l}$\\
    \ \ Define $P(\mathbf{r}_{l,0}|X_{1:t-1},Z_{1:t-1})=0$, $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=1$, \\  \ \ \ \ \ \ \ \ \ \   $P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$;\\
-   \ \ for $k = 1,2,\ldots,n_r$\\
+   \ \ for $k = 1,2,...,n_r$\\
    \ \ \ \ Obtain $p(z_{t,l}|X_t,\mathbf{r}_{l,k})$ from a forward sensor model\\
@@
    \ \ $c_{n_{r,l}+1}=0$;\\
-   \ \ for $k = n_{r,l},n_{r,l}-1,\ldots,1$\\
+   \ \ for $k = n_{r,l},n_{r,l}-1,...,1$\\
    \ \ \ \ $c_k=\frac{P(\mathbf{r}_{l,k+1}|X_{1:t-1},Z_{1:t-1})}{P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})}c_{k+1}$\\
@@
    \ \ end for\\
-   \ \ for $k = 1,2,\ldots,n_r$\\
+   \ \ for $k = 1,2,...,n_r$\\
    \ \ \ \ $\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=a_k+b_kp(z_{t,l}|\mathbf{r}_{l,k},X_t)$;\\
@@
 In short, the proposed scan inverse sensor model is computed by \refeqn{ISM_Fusion}-\refeqn{xi}.
-These are summarized in Table \ref{tab:Alg_ISM_2D} as a computationally efficient, recursive algorithm that avoids several repeated calculations of the same quantity.
-This algorithm utilizes the following temporary variables to develop the algorithm in an efficient recursive form, %defined as
+These are summarized in Table \ref{tab:Alg_ISM_2D} as a computationally efficient, recursive algorithm that avoids repeated calculations of the same quantity.
+This algorithm utilizes the following temporary variables to develop the algorithm in an efficient recursive form, defined as
 \begin{align*}
@@
 \nonumber\\
-c_k&=
-\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0,j\neq k}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}).%,\sum_{i=k+1}^{n_{r,l}}\bigg\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}).
+c_k&=P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})^{-1}
+\nonumber\\&\quad\times 
+\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}).%,\sum_{i=k+1}^{n_{r,l}}\bigg\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}).
 \end{align*}
@@
 The purpose of this section is to compare the proposed exact solution to the inverse sensor model with an approximate algorithm developed for a Microsoft Kinect sensor~\cite{PirRutBisSch11,KhoElb12}, summarized as follows.
-The probability that the $i$-th grid cell $\mathbf{m}_i$ is occupied conditioned on the measurement ray $z_{t,l}$ at the pose $X_t$ is the continuous function
+The probability that the $i$-th grid cell $\mathbf{m}_i$ is occupied conditioned on the measurement ray $z_{t,l}$ (the $l$-th ray at the $t$-th time step) at the pose $X_t$ is the continuous function
 \begin{align*}
@@
 \centerline{
-	\subfigure[Entropy change]% (black: exact model, green: approx.)]
-		{\includegraphics[height=0.3\columnwidth]{EntropyEvolution.png}
-	}
-	\hspace*{-.075\columnwidth}
-	\subfigure[$t=50$ (exact)]
-		{\hspace*{0.3cm}\includegraphics[height=0.3\columnwidth]{Compare_ISM_H.png}\hspace*{0.3cm}
-	}
-	\hspace*{-.075\columnwidth}
-	\subfigure[$t=50$  (approx.)]
-		{\hspace*{0.3cm}\includegraphics[height=0.3\columnwidth]{Compare_Approx_H.png}\hspace*{0.3cm}
+	\subfigure[Entropy (black: exact model, green: approx.)]
+		{\includegraphics[width=0.7\columnwidth]{EntropyEvolution.png}
+	}
+}
+\centerline{
+	\subfigure[Entropy map at $t=50$ (exact)]
+		{\hspace*{0.3cm}\includegraphics[height=0.4\columnwidth]{Compare_ISM_H.png}\hspace*{0.3cm}
+	}
+	\subfigure[Entropy map at $t=50$  (approx.)]
+		{\hspace*{0.3cm}\includegraphics[height=0.4\columnwidth]{Compare_Approx_H.png}\hspace*{0.3cm}
 	}
@@
 \caption{Entropy serves as a measure of uncertainty of the occupancy grid, where more blue regions are more certain, and red regions are more uncertain.
-In (a), the entropy with the exact solution (black) decreases faster than the entropy with the approximate solution (green) over $20$ seconds.
-The uncertainty is always less when the exact solution is applied, where the final map ($50$ seconds) with the approximate solution has more uncertainty.}\label{fig:NumResOccH}
+The uncertainty is always less when the exact solution is applied.}\label{fig:NumResOccH}
 \end{figure}
@@
 The Kinect sensor faces walls underneath a work desk, as illustrated in Figures \ref{fig:ExpRes}.(a) and (b).
-The resulting occupancy grid map and entropy map are obtained from both the proposed and approximate algorithms after a single scan~\cite{PirRutBisSch11,KhoElb12}.
-
-%\begin{figure}
-%\centerline{
-%	\subfigure[Test Setup]
-%		{\includegraphics[trim={1.5cm 0 1.5cm 0},clip,height=0.3\columnwidth]{TestSetup.jpg}
-%	}
-%	\subfigure[Kinect View]
-%		{\includegraphics[height=0.3\columnwidth]{KinectView.jpg}
-%	}
-%}
-%\centerline{
-%	\subfigure[Occupancy map (exact)]
-%		{\includegraphics[trim={0 0 0 3cm},clip,height=0.34\columnwidth]{KinectDeskISM.png}
-%	} 
-%	\hfill
-%	\subfigure[Occupancy map (approx.)]
-%		{\includegraphics[trim={0 0 0 3cm},clip,height=0.34\columnwidth]{KinectDeskApprox.png}
-%	}
-%}
-%\centerline{
-%	\subfigure[Entropy map (exact)]
-%		{\includegraphics[trim={0 0 0 3cm},clip,height=0.34\columnwidth]{KinectDeskISM_H.png}
-%	}
-%	\hfill
-%	\subfigure[Entropy map (approx.)]
-%		{\includegraphics[trim={0 0 0 3cm},clip,height=0.34\columnwidth]{KinectDeskApprox_H.png}
-%	}
-%}
-%\caption{The Kinect captures depth measurements of the underside of a metal desk.
-The measurement scan serves to update two occupancy grid mapping schemes.
-The proposed approach, based on the the exact solution of the inverse sensor model, provides an update with substantially less uncertainty than the approximate model update.}\label{fig:ExpRes}
-%\end{figure}
+The resulting occupancy grid map and entropy map, obtained from both the proposed algorithm and an approximate model after a single scan~\cite{PirRutBisSch11,KhoElb12}, are also presented.
+
 \begin{figure}
@@
 	\subfigure[Test Setup]
-		{\includegraphics[height=0.3\columnwidth]{TestSetup.jpg}
-	}
-	\subfigure[Exact Occupancy]
-		{\includegraphics[height=0.3\columnwidth]{KinectDeskISM.png}
-	} 
-	\subfigure[Exact Entropy]
-		{\includegraphics[height=0.3\columnwidth]{KinectDeskISM_H.png}
-	}
-}
-\centerline{
+		{\includegraphics[trim={1.5cm 0 1.5cm 0},clip,height=0.3\columnwidth]{TestSetup.jpg}
+	}
 	\subfigure[Kinect View]
@@
 	}
-	\subfigure[Approx. Occupancy]
-		{\includegraphics[height=0.3\columnwidth]{KinectDeskApprox.png}
-	}
-	\subfigure[Approx. Entropy]
-		{\includegraphics[height=0.3\columnwidth]{KinectDeskApprox_H.png}
+}
+\centerline{
+	\subfigure[Occupancy map (exact)]
+		{\includegraphics[trim={0 0 0 3cm},clip,height=0.34\columnwidth]{KinectDeskISM.png}
+	} 
+	\hfill
+	\subfigure[Occupancy map (approx.)]
+		{\includegraphics[trim={0 0 0 3cm},clip,height=0.34\columnwidth]{KinectDeskApprox.png}
+	}
+}
+\centerline{
+	\subfigure[Entropy map (exact)]
+		{\includegraphics[trim={0 0 0 3cm},clip,height=0.34\columnwidth]{KinectDeskISM_H.png}
+	}
+	\hfill
+	\subfigure[Entropy map (approx.)]
+		{\includegraphics[trim={0 0 0 3cm},clip,height=0.34\columnwidth]{KinectDeskApprox_H.png}
 	}
@@
 Furthermore, the edges of the objects captured by the Kinect are well defined with the proposed approach with an edge thickness of roughly one or two cells, but two or three cells thick with the conventional approach.
-Additionally, the decrease in entropy is greater with the proposed approach at $-3487$, compared with $-557$ with the approximate model.
+Additionally, the decrease in entropy is greater with the proposed approach at $-3487$, compared with $-557$ with the approximate inverse sensor model.
 
@@
 The numerical and experimental results show that maps using this exact approach are substantially improved.
-%Future study includes developing the presented experimental results further for mobile robotic mapping, and utilizing the proposed exact inverse sensor model into SLAM and autonomous explorations.
+Future study includes developing the presented experimental results further for mapping with a mobile robot, and utilizing the proposed exact inverse sensor model into SLAM and autonomous explorations.
 
@@
 \paragraph{Unnormalized Reduced Map Inverse Sensor Model}
-For the reduced map of the $l$-th ray, the unnormalized probability corresponding to terms without $\eta_{t,l}$ in \refeqn{InvSenModWithProbDens} is %given by
+For the reduced map of the $l$-th ray, the unnormalized probability that corresponds to the terms without $\eta_{t,l}$ in \refeqn{InvSenModWithProbDens} is given by
 \begin{align}
@@
 %%&=P(\mathbf{r}_{l,k+1})\bigg\{
-%%P(\bar{\mathbf{r}}_{l,0})P(\mathbf{r}_{l,1})p(z_{t,l}|X_t,\mathbf{r}_{l,1})+\nonumber\\&\quad P(\bar{\mathbf{r}}_{l,0})P(\bar{\mathbf{r}}_{l,1})P(\mathbf{r}_{l,2})p(z_{t,l}|X_t,\mathbf{r}_{l,2})+\ldots\nonumber
+%%P(\bar{\mathbf{r}}_{l,0})P(\mathbf{r}_{l,1})p(z_{t,l}|X_t,\mathbf{r}_{l,1})+\nonumber\\&\quad P(\bar{\mathbf{r}}_{l,0})P(\bar{\mathbf{r}}_{l,1})P(\mathbf{r}_{l,2})p(z_{t,l}|X_t,\mathbf{r}_{l,2})+...\nonumber
 %%\\
 %%&\quad +\left(
-%%P(\bar{\mathbf{r}}_{l,0})P(\bar{\mathbf{r}}_{l,1})\ldotsP(\bar{\mathbf{r}}_{l,k-1})
+%%P(\bar{\mathbf{r}}_{l,0})P(\bar{\mathbf{r}}_{l,1})...P(\bar{\mathbf{r}}_{l,k-1})
 %%\right)\nonumber
@@
 %%+\left(
-%%P(\bar{\mathbf{r}}_{l,0})P(\bar{\mathbf{r}}_{l,1})\ldotsP(\bar{\mathbf{r}}_{l,k})
+%%P(\bar{\mathbf{r}}_{l,0})P(\bar{\mathbf{r}}_{l,1})...P(\bar{\mathbf{r}}_{l,k})
 %%\right)\nonumber
@@
 \begin{align}
-&\sum_{r\in\bar{\mathcal{N}}_{i,k}} P(r|X_{1:t-1},Z_{1:t-1}) = \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
+\sum_{r\in\bar{\mathcal{N}}_{i,k}} & P(r|X_{1:t-1},Z_{1:t-1}) = \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
 &\quad \times P({\mathbf{r}}_{l,i}|X_{1:t-1},Z_{1:t-1})P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1}),\label{eqn:PbarNik1}
@@
 %%p(z_{t,l}|X_t,\mathbf{r}_{l,2})\nonumber\\&\quad
-%%+P(\bar{\mathbf{r}}_{l,2})P(\mathbf{r}_{l,3})p(z_{t,l}|X_t,\mathbf{r}_{l,3})+\ldots\nonumber\\&\quad+\prod_{j=2}^{n_r-1}P(\bar{\mathbf{r}}_{l,j})P(\mathbf{r}_{l,n_r})p(z_{t,l}|X_t,\mathbf{r}_{l,n_r})
+%%+P(\bar{\mathbf{r}}_{l,2})P(\mathbf{r}_{l,3})p(z_{t,l}|X_t,\mathbf{r}_{l,3})+...\nonumber\\&\quad+\prod_{j=2}^{n_r-1}P(\bar{\mathbf{r}}_{l,j})P(\mathbf{r}_{l,n_r})p(z_{t,l}|X_t,\mathbf{r}_{l,n_r})
 %%\bigg\}\nonumber
@@
 %%P(\bar{\mathbf{r}}_{l,1})P(\mathbf{r}_{l,3})p(z_{t,l}|X_t,\mathbf{r}_{l,3})\nonumber\\&\quad
-%%+P(\bar{\mathbf{r}}_{l,1})P(\bar{\mathbf{r}}_{l,3})P(\mathbf{r}_{l,4})\nonumber\\&\qquad\times p(z_{t,l}|X_t,\mathbf{r}_{l,4})+\ldots\nonumber
+%%+P(\bar{\mathbf{r}}_{l,1})P(\bar{\mathbf{r}}_{l,3})P(\mathbf{r}_{l,4})\nonumber\\&\qquad\times p(z_{t,l}|X_t,\mathbf{r}_{l,4})+...\nonumber
 %%\\
@@
 %%%P(\bar{\mathbf{r}}_1)P(\mathbf{r}_2)p(z|X,\mathbf{r}_2)
-%%%+P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)P(\mathbf{r}_3)p(z|X,\mathbf{r}_3)+\ldots\nonumber
+%%%+P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)P(\mathbf{r}_3)p(z|X,\mathbf{r}_3)+...\nonumber
 %%%\\
 %%%&\quad
-%%%+P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)\ldotsP(\bar{\mathbf{r}}_{k-1})P(\mathbf{r}_k)p(z_{t,l}|\mathbf{r}_k,X_t)
+%%%+P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)...P(\bar{\mathbf{r}}_{k-1})P(\mathbf{r}_k)p(z_{t,l}|\mathbf{r}_k,X_t)
 %%%]\nonumber
@@
 %%%P(\bar{\mathbf{r}}_{k+1})^{-1}[
-%%%P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)\ldotsP(\bar{\mathbf{r}}_{k+1})P(\mathbf{r}_{k+2})p(z|X,\mathbf{r}_{k+2\ldots...
+%%%P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)...P(\bar{\mathbf{r}}_{k+1})P(\mathbf{r}_{k+2})p(z|X,\mathbf{r}_{k+2})+...
 %%%\nonumber
@@
 %%%&\quad
-%%%+P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)\ldotsP(\bar{\mathbf{r}}_{n_r-1})P(\mathbf{r}_{n_r})p(z|X,\mathbf{r}_{n_r})
+%%%+P(\bar{\mathbf{r}}_1)P(\bar{\mathbf{r}}_2)...P(\bar{\mathbf{r}}_{n_r-1})P(\mathbf{r}_{n_r})p(z|X,\mathbf{r}_{n_r})
 %%%]
@@
 %%&\quad
-%%+\left\{\prod_{j=0}^{2}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,3})p(z_{t,l}|X_t,\mathbf{r}_{l,3})+\ldots\nonumber
+%%+\left\{\prod_{j=0}^{2}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,3})p(z_{t,l}|X_t,\mathbf{r}_{l,3})+...\nonumber
 %%\\
@@
 %%\bigg[
-%%\left\{\prod_{j=0}^{k+1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,k+2})p(z_{t,l}|X_t,\mathbf{r}_{l,k+2})+\ldots\nonumber
+%%\left\{\prod_{j=0}^{k+1}P(\bar{\mathbf{r}}_{l,j})\right\}P(\mathbf{r}_{l,k+2})p(z_{t,l}|X_t,\mathbf{r}_{l,k+2})+...\nonumber
 %%\\
@@
 \begin{align*}
-&\eta_{t,l}=\frac1{\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})+
-\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}.
+&\eta_{t,l}\nonumber\\&=[\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})+
+\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})]^{-1}.
 \end{align*}

### 2018-06-09 Conference/ACC 2016/First Submition/ACC16.1.tex
added Conference folder

--- 2018-06-09 Conference/ACC 2016/Final Unabridged/ACC16.1.tex
@@
 
-\newcommand{\EditEK}[1]{{\color{blue}\protect #1}}
+\newcommand{\EditTL}[1]{{\color{red}\protect #1}}
 %\renewcommand{\EditTL}[1]{{\protect #1}}
@@
 \end{align}
-where it is considered $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading for convenience.
-The normalizer $\eta_{t,l}$ is given by
+where it is considered $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=1$ for convenience. The normalizer $\eta_{t,l}$ is given by
 \begin{align}
@@
 &=
-\bigg[\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
+\bigg[\sum_{i=1}^{n_{r,l}}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
 \end{align}
@@
 Compared with \refeqn{InvSenModWithProbDens}, where the terms of the summation should be repeated $2^{n_{r,l}}$ times \emph{per each cell of the reduced map}, the proposed expressions \refeqn{RayISMAnswer} and \refeqn{allEta} are \textit{substantially} simpler as they require $k$ summation terms for the $k$-th cell.
-In fact, if the terms inside the summations of \refeqn{Unnormalized} and \refeqn{allEta} are obtained recursively, only $n_{r,l}+1$ repetitions are required for \emph{all cells of the reduced map}, yielding an algorithm that is nearly $\mathbf{2^{n_{r,l}}}$ \textbf{times faster}.
+In fact, if the terms inside the summations of \refeqn{Unnormalized} and \refeqn{allEta} are obtained recursively, only $n_{r,l}$ repetitions are required for \emph{all cells of the reduced map}, yielding an algorithm that is up to $\mathbf{2^{n_{r,l}}}$ \textbf{times faster}.
 
@@
 At the $t$-th time step, the measurement scan is the set of all rays $Z_t=\{z_{t,1},z_{t,2},...,z_{t,n_z}\}$.
-One may apply the results of Proposition 1 repeatedly for each ray to obtain a two-dimensional inverse sensor model because
-\begin{align}
-P(\mathbf{m}_i|X_{1:t},Z_{1:t})&%=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1},z_{t,2},...,z_{t,n_z})\nonumber\\&
-=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1}|z_{t,2}|...|z_{t,n_z}).
-\end{align}
-%However, when there are a large number of rays, it is certainly possible that a single cell is covered by multiple rays. 
-Alternatively, measurements from multiple rays can be integrated synergistically to obtain a more accurate estimate.
+One may apply the results of Proposition 1 repeatedly for each ray to obtain a two-dimensional inverse sensor model.
+However, when there are a large number of rays, it is certainly possible that a single cell is covered by multiple rays.
+For those cells, measurements from multiple rays should be integrated synergistically to obtain a more accurate estimate.
+Here, we construct such two-dimensional inverse sensor model, referred to as the scan inverse sensor model, based on the assumption that the measurements of rays are mutually independent.
 
@@
 
-Here, we construct such two-dimensional inverse sensor model, referred to as the scan inverse sensor model, based on the assumption that the measurements of rays are mutually independent.
 Let $\mathcal L_i\subset\{1,\ldots, n_z\}$ be the set of rays that pass through the $i$-th cell, $\mathbf{m}_i$. Applying Bayes' rule repeatedly, we obtain
@@
 &\quad
-+P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})^{-1}
-\nonumber\\&\quad\times 
-\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
-\end{align}
-which is obtained from \refeqn{tildePbar}. %, where the division of $P(\bar{\mathbf{m}}_i|X_{1:t-1},Z_{1:t-1})=P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})$ is absorbed into the second summation.
-Since 
++
+\sum_{i=k+1}^{n_r}\bigg\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\qquad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
+\end{align}
+which is obtained from \refeqn{tildePbar}, where the division of $P(\bar{\mathbf{m}}_i|X_{1:t-1},Z_{1:t-1})=P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})$ is absorbed into the second summation. Since 
 \begin{align*}
@@
    \ \ Obtain reduced map $r_{t,l}$\\
-   \ \ Define $P(\mathbf{r}_{l,0}|X_{1:t-1},Z_{1:t-1})=0$, $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=1$, \\  \ \ \ \ \ \ \ \ \ \   $P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$;\\
+   \ \ Define $P(\mathbf{r}_{l,0}|X_{1:t-1},Z_{1:t-1})=0$, $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=1$;\\
    \ \ for $k = 1,2,...,n_r$\\
@@
    \ \ end for\\
-   \ \ $c_{n_{r,l}+1}=0$;\\
-   \ \ for $k = n_{r,l},n_{r,l}-1,...,1$\\
-   \ \ \ \ $c_k=\frac{P(\mathbf{r}_{l,k+1}|X_{1:t-1},Z_{1:t-1})}{P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})}c_{k+1}$\\
-   \ \ \ \ \ \ \ \ \ \ \ \ $+b_{k}p(z_{t,l}|\mathbf{r}_{l,k+1},X_t)P(\mathbf{r}_{l,k+1}|X_{1:t-1},Z_{1:t-1})$;\\
+   \ \ $c_{n_r}=0$;\\
+   \ \ for $k = n_r-1,n_r-2,...,1$\\
+   \ \ \ \ $c_k=c_{k+1}+b_{k}p(z_{t,l}|\mathbf{r}_{l,k+1},X_t)P(\mathbf{r}_{l,k+1}|X_{1:t-1},Z_{1:t-1})$;\\
    \ \ end for\\
@@
    \ \ end if\\
-   \ \ Return: $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$\\
    end for\\
-
+Return: $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$\\
 \end{tabular}
@@
 \nonumber\\
-c_k&=P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})^{-1}
-\nonumber\\&\quad\times 
-\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}).%,\sum_{i=k+1}^{n_{r,l}}\bigg\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}).
+c_k&=\sum_{i=k+1}^{n_{r,l}}\bigg\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}).
 \end{align*}
@@
 	\subfigure[Map at $t=0$ (exact)]
-		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_1.png}
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_1.pdf}
 	}
 	\subfigure[Map at $t=0$ (approx.)]
-		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_1.png}
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_1.pdf}
 	}
@@
 	\subfigure[Map at $t=13$ (exact)]
-		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_2.png}
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_2.pdf}
 	}
 	\subfigure[Map at $t=13$ (approx.)]
-		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_2.png}
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_2.pdf}
 	}
@@
 	\subfigure[Map at $t=25$ (exact)]
-		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_3.png}
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_3.pdf}
 	}
 	\subfigure[Map at $t=25$ (approx.)]
-		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_3.png}
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_3.pdf}
 	}
@@
 	\subfigure[Map at $t=38$ (exact)]
-		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_4.png}
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_4.pdf}
 	}
 	\subfigure[Map at $t=38$ (approx.)]
-		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_4.png}
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_4.pdf}
 	}
@@
 	\subfigure[Map at $t=50$ (exact)]
-		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_5.png}
+		{\includegraphics[height=0.4\columnwidth]{Compare_ISM_5.pdf}
 	}
 	\subfigure[Map at $t=50$ (approx.)]
-		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_5.png}
+		{\includegraphics[height=0.4\columnwidth]{Compare_Approx_5.pdf}
 	}
@@
 The subfigure (a) illustrates that the proposed exact inverse sensor model exhibits rapid decreases of entropies, and smaller entropies always.
-The resulting terminal map obtained from the proposed approach, shown in the subfigure (b), has less uncertainty than (c) constructed by the approximate model.
+The resulting terminal map obtained from the proposed approach, shown in the subfigure (b), has less uncertainties than (c) constructed by the approximate model.
 In short, the proposed approach is more efficient at extracting information about the environment from the same set of set measurements.
@@
 \centerline{
-	\subfigure[Entropy map at $t=50$ (exact)]
-		{\hspace*{0.3cm}\includegraphics[height=0.4\columnwidth]{Compare_ISM_H.png}\hspace*{0.3cm}
-	}
-	\subfigure[Entropy map at $t=50$  (approx.)]
-		{\hspace*{0.3cm}\includegraphics[height=0.4\columnwidth]{Compare_Approx_H.png}\hspace*{0.3cm}
+	\subfigure[Entropy map at $t=20$ (exact)]
+		{\hspace*{0.3cm}\includegraphics[height=0.4\columnwidth]{Compare_ISM_H.pdf}\hspace*{0.3cm}
+	}
+	\subfigure[Entropy map at $t=20$  (approx.)]
+		{\hspace*{0.3cm}\includegraphics[height=0.4\columnwidth]{Compare_Approx_H.pdf}\hspace*{0.3cm}
 	}
@@
 +
-\bigg[\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
-&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg],\label{eqn:tildePbar}
-\end{align}
-where $p(z_{t,l}|\mathbf{r}_{(n_r+1)+},X_t)$ corresponds to the forward sensor model of maximum reading and $P(\mathbf{r}_{l,n_{r,l}+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience for the empty map case.
-
-% ADD: where it is considered $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$
+\bigg[\sum_{i=k+1}^{n_r}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
+&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg].\label{eqn:tildePbar}
+\end{align}
 

### 2018-06-09 Conference/ACC 2016/Old Presentation Files/ACC16.1_Presentation.tex
added Conference folder

\documentclass[11pt,professionalfonts,hyperref={pdftex,pdfpagemode=none,pdfstartview=FitH}]{beamer}
%\usepackage{times}
%\usefonttheme{serif}
%\usepackage{helvet}
%\usepackage{amsmath,amssymb}
\usepackage{graphicx,multirow}
\usepackage[scriptsize]{subfigure}
%\usepackage{movie15}

%\usepackage{warmread}
%\usepackage[all,import]{xy}

%\renewcommand\mathfamilydefault{\rmdefault}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand{\refprop}[1]{Proposition \ref{prop:#1}}

\definecolor{mygray}{gray}{0.9}

\mode<presentation> {
  \usetheme{Warsaw}
  \usefonttheme{serif}
  \setbeamercovered{transparent}
}

\newcommand{\mypaper}{}

\setbeamertemplate{footline}%{split theme}
{%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=.7\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm]{title in head/foot}
%    \usebeamerfont{title in head/foot}\mypaper\hfill \insertframenumber/\inserttotalframenumber
    \usebeamerfont{title in head/foot}\hfill \insertframenumber/\inserttotalframenumber
  \end{beamercolorbox}}%
  \vskip0pt%
} \setbeamercolor{box}{fg=black,bg=yellow}

\title[Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model ]{\large Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model}

\author{\vspace*{-0.3cm}}

\institute{\footnotesize
{\normalsize Evan Kaufman, Taeyoung Lee, \\Zhuming Ai, and Ira S. Moskowitz}\vspace*{0.2cm}\\
  Mechanical and Aerospace Engineering\\ George Washington University}

\date{}

\definecolor{tmp}{rgb}{0.804,0.941,1.0}
\setbeamercolor{numerical}{fg=black,bg=tmp}
\setbeamercolor{exact}{fg=black,bg=red}

\newtheorem{prop}{Proposition}

\renewcommand{\emph}[1]{\textit{\textbf{\color{blue}{#1}}}}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section*{}
\subsection*{Introduction}

\begin{frame}
\frametitle{Introduction}
\begin{itemize}
    \item Robotic Mapping
    \begin{itemize}
    	\item Goal: generate a map representing surrounding regions
	\item Crucial for simulataneous localization and mapping (SLAM) and autonomous exploration of uncertain environments
    	\item Popular mapping representations include occupancy grids, octomaps, and feature-based maps
    \end{itemize}
\item Occupancy Grid Mapping
\begin{itemize}
	\item The environment may be decomposed into evenly spaced grid cells that are either \emph{occupied} or \emph{free}
	\item Probabilistic map: the goal is to obtain the  probability of each grid cell being occupied
\end{itemize}
\end{itemize}

\only<1->{
\begin{figure}
\centerline{
    \includegraphics[height=2.1cm]{ogm_ex1.jpg}\hspace*{0.1cm}
\hspace*{0.5cm}
    \includegraphics[height=2.1cm]{ogm_ex2.png}\hspace*{0.1cm}
\hspace*{0.5cm}
    \includegraphics[height=2.1cm]{ogm_ex3.jpeg}\hspace*{0.1cm}
}
\end{figure}}

\end{frame}

\begin{frame}
\frametitle{Introduction}
\framesubtitle{Problem Definition}

\begin{itemize}
	\item The Map and the Robot
	\begin{itemize}
	\item Map $m$ is composed of $n_m$ grid cells
	\item The $i$-th grid cell $\mathbf{m}_i$ is a \emph{binary} random variable, independent of other grid cells: $P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i)$
	\item The robot \emph{pose} $X_t$ contains the \emph{position} and \emph{attitude} of the robot at the $t$-th timestep, and the history of poses $X_{1:t}$ is known	
	\end{itemize}
\vspace*{0.3cm}\pause
	\item Depth Measurements
	\begin{itemize}
	\item The measurement origins and directions are known \emph{deterministically}
	\item A measurement \emph{scan} $Z_t=\braces{z_{t,1},z_{t,2},\ldots,z_{t,n_z}}$ contains $n_z$ measurement \emph{rays} (depth readings) at the $t$-th time step, and the history of measurement scans $Z_{1:t}$ is known
	\item The \emph{forward sensor model} is the probability density distribution $p(z_{t,l}|m,X_{t})$  known from the sensor properties
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Introduction}
\framesubtitle{Problem Definition}

\begin{itemize}
	\item Bayesian Framework
	\begin{itemize}
	\item Markov Assumption
	\item Benefit of Log-Odds
	\item Assumption with Log-Odds
	\end{itemize}
\vspace*{0.3cm}\pause
	\item Inverse Sensor Model
	\begin{itemize}
	\item Equation
	\item Computationally intractable equation
	\item Motivation for another solution...
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Introduction}
\framesubtitle{Problem Definition}

\begin{itemize}
	\item Heuristic Approach
	\begin{itemize}
	\item Idea
	\item Figure
	\item Mathematically innacurate
	\end{itemize}
\vspace*{0.3cm}\pause
	\item Learning
	\begin{itemize}
	\item Idea
	\item Complicated, not clear how parameters are chosen
	\end{itemize}
\vspace*{0.3cm}\pause
	\item Goal: design a simple and accurate method for occupancy grid mapping that avoids log-odds ratio assumptions
\end{itemize}

\end{frame}

\section*{}
\subsection*{Problem Formulation}

\begin{frame}
\frametitle{Single Measurement Ray}

\begin{itemize}
    \item Main idea: make use of occupancy grid mapping \emph{assumptions} and extract \emph{paterns} from probabilistic properties to find a \emph{computationally-efficient} solution
	\begin{itemize}
		\item Since the origin and direction of each measurement ray is known deterministically, the set of grid cells that the ray intersects is known through geometry
		\item A depth reading follows the forward sensor model, which \emph{only depends} on the first occupied grid cell along the measurement ray
		\item (figure on forward sensor model)
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Single Measurement Ray}

\begin{itemize}
    \item Patterns...
	\begin{itemize}
		\item 
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Full Measurement Scan}

\begin{itemize}
	\item 
\end{itemize}

\end{frame}

\section*{}
\subsection*{Results}

\begin{frame}
\frametitle{Numerical Example}

\begin{itemize}
\item 1D Case
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Numerical Example}

\begin{itemize}
\item 2D Case
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Experiment}

\begin{itemize}
\item Kinect
\end{itemize}

\end{frame}

\section*{}
\subsection*{Conclusions}

\begin{frame}
\frametitle{Conclusions}
\begin{itemize}
    \item 
    \begin{itemize}
        \item 
    \end{itemize}
\end{itemize}
\end{frame}

\end{document}

### 2018-06-09 Conference/ACC 2016/Presentation/ACC16.1_Presentation.tex
added Conference folder

\documentclass[11pt,professionalfonts,hyperref={pdftex,pdfpagemode=none,pdfstartview=FitH}]{beamer}
%\usepackage{times}
%\usefonttheme{serif}
%\usepackage{helvet}
%\usepackage{amsmath,amssymb}
\usepackage{graphicx,multirow}
\usepackage[scriptsize]{subfigure}
%\usepackage{movie15}

%\usepackage{warmread}
%\usepackage[all,import]{xy}

%\renewcommand\mathfamilydefault{\rmdefault}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand{\refprop}[1]{Proposition \ref{prop:#1}}

\definecolor{mygray}{gray}{0.9}

\mode<presentation> {
  \usetheme{Warsaw}
  \usefonttheme{serif}
  \setbeamercovered{transparent}
}

\newcommand{\mypaper}{}

\setbeamertemplate{footline}%{split theme}
{%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=.7\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm]{title in head/foot}
    \usebeamerfont{title in head/foot}\mypaper\hfill \insertframenumber/\inserttotalframenumber
    \usebeamerfont{title in head/foot}\hfill \insertframenumber/\inserttotalframenumber
  \end{beamercolorbox}}%
  \vskip0pt%
} \setbeamercolor{box}{fg=black,bg=yellow}

\title[Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model ]{\large Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model}

\author{\vspace*{-0.3cm}}

\institute{\footnotesize
{\normalsize Evan Kaufman, Taeyoung Lee, \\Zhuming Ai, and Ira S. Moskowitz}\vspace*{0.2cm}\\
  Mechanical and Aerospace Engineering\\ George Washington University}

\date{}

\definecolor{tmp}{rgb}{0.804,0.941,1.0}
\setbeamercolor{numerical}{fg=black,bg=tmp}
\setbeamercolor{exact}{fg=black,bg=red}

\newtheorem{prop}{Proposition}

\renewcommand{\emph}[1]{\textit{\textbf{\color{blue}{#1}}}}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section*{}
\subsection*{Introduction}

\begin{frame}
\frametitle{Introduction}
\begin{itemize}
    \item Robotic Mapping
    \begin{itemize}
    	\item Goal: generate a map representing surrounding regions
	\item Crucial for simultaneous localization and mapping (SLAM) and autonomous exploration of uncertain environments
    	\item Popular mapping representations include \emph{occupancy grids}, octomaps, and feature-based maps
    \end{itemize}
\item Occupancy Grid Mapping
\begin{itemize}
	\item The environment may be decomposed into evenly spaced grid cells that are either \emph{occupied} or \emph{free}
	\item Probabilistic map: the goal is to obtain the \emph{probability} of each grid cell being occupied
\end{itemize}
\end{itemize}

\only<1->{
\begin{figure}
\centerline{
    \includegraphics[height=2.1cm]{ogm_ex3.jpeg}\hspace*{0.1cm}
\hspace*{0.5cm}
    \includegraphics[height=2.1cm]{ogm_ex1.jpg}\hspace*{0.1cm}
\hspace*{0.5cm}
    \includegraphics[height=2.1cm]{ogm_ex2.png}\hspace*{0.1cm}
}
\end{figure}}

\end{frame}

\begin{frame}
\frametitle{Introduction}
\framesubtitle{Problem Definition}

\begin{itemize}
	\item The Map and the Robot
	\begin{itemize}
	\item Map $m$ is composed of $n_m$ grid cells with known location and size
	\item The $i$-th grid cell $\mathbf{m}_i$ is a \emph{static binary} random variable, independent of other grid cells: $P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i)$
	\item \emph{Pose} $X_t$ is known, containing robot \emph{position} and \emph{attitude}
	\end{itemize}
\vspace*{0.0cm}\pause
\end{itemize}
\begin{minipage}[t]{7.0cm}
\begin{itemize}
	\item Depth Measurements
	\begin{itemize}
	\item Each measurement origin and direction is known \emph{deterministically}
	\item A measurement \emph{scan} $Z_t=\braces{z_{t,1},z_{t,2},\ldots,z_{t,n_z}}$ contains $n_z$ measurement \emph{rays} (depths)% at the $t$-th time step, and the history of measurement scans $Z_{1:t}$ is known
\item The \emph{forward sensor model} is known from the sensor properties
\end{itemize}
\end{itemize}
\end{minipage}
\begin{minipage}[t]{3.0cm}
%The \emph{forward sensor model} is the probability density distribution $p(z_{t,l}|m,X_{t})$  known from the sensor properties
\hspace*{0.25cm}
\begin{figure}[!htbp]
\vspace*{-0.25cm}
%\vspace*{0.25cm}
\centerline{
    \includegraphics[width=3.5cm]{BeamModel.png}\hspace*{0.1cm}
    }
\vspace*{0.25cm}
\centerline{
    \includegraphics[width=2.5cm]{1D_True_Grid.png}\hspace*{0.1cm}
%    \hspace*{0.75cm}
    }
{Beam Model for Range Finders}
\end{figure}

\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Introduction}
\framesubtitle{Problem Definition}

\begin{itemize}
	\item Bayesian Framework
	\begin{itemize}
	\item Markov Assumption: latest a priori cell occupancy probabilities capture the information from all prior observations
	\item Log-Odds Ratio: popular representation due to simple additive update structure and probability truncation avoidance, but requires the \emph{assumption}
	\begin{align*}
		P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})\approx P(Z_t|\mathbf{m}_i,X_t)
	\end{align*}
	\end{itemize}
\vspace*{0.0cm}\pause
	\item Inverse Sensor Model
	\begin{align*}
&P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\eta_{t,l}\sum_{m\in\mathcal{M}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
\end{align*}
	\begin{itemize}
	\item Given $n$ grid cells: $\mathcal O(2^n)$ is \emph{computationally intractable}, motivating a different solution
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Introduction}
\framesubtitle{Problem Definition}

\begin{minipage}[t]{7.0cm}
\begin{itemize}
	\item Approximate a function for the inverse sensor model based on intuition
	\begin{itemize}
		\item Simple to implement, but mathematically inaccurate
	\end{itemize}
\end{itemize}
\end{minipage}
\begin{minipage}[t]{3.0cm}
\begin{figure}[!htbp]
\centerline{
    \vspace*{-0.5cm}
%	\hspace*{0.25cm}
    \includegraphics[width=4.0cm]{Approx_ISM_shortened.png}\hspace*{0.1cm}
    }
\centerline{
    \includegraphics[width=3.5cm]{1D_True_Grid.png}
%    \hspace*{0.75cm}
    }
		\end{figure}
\end{minipage}
\vspace*{-0.5cm}
\begin{itemize}
	\item Find a solution through learning
	\begin{itemize}
		\item Simulate maps, poses, and measurements and use learning to obtain an inverse sensor model
		\item Complicated, unclear how these parameters are chosen
	\end{itemize}
	\item Goal: design a \emph{simple} and \emph{accurate} occupancy grid mapping method avoiding log-odds ratio assumptions
\end{itemize}
%
%	\begin{minipage}[width=1cm]
%		Approximate a function for the inverse sensor model based on intuition
%	\end{minipage}
%	\begin{minipage}[width=1cm]
%		Approximate a function for the inverse sensor model based on intuition2
%	\end{minipage}
%\begin{itemize}
%	\item Heuristic Approach
%	\begin{itemize}
%	\item asdf
%\begin{figure}
%\centerline{
%    \includegraphics[width=3.5cm]{ISM_Approx_Single_Ray.png}\hspace*{0.1cm}
%    }
%\centerline{
%    \includegraphics[width=3.5cm]{1D_True_Grid.png}\hspace*{0.1cm}
%%    \hspace*{0.75cm}
%    }
%\end{figure}
%	\item Simple to implement, but mathematically inaccurate
%	\end{itemize}
%\vspace*{0.3cm}\pause
%\begin{itemize}
%	\item Learning
%	\begin{itemize}
%	\item Idea
%	\item Complicated, not clear how parameters are chosen
%	\end{itemize}
%\vspace*{0.3cm}\pause
%	\item Goal: design a simple and accurate method for occupancy grid mapping that avoids log-odds ratio assumptions
%\end{itemize}

\end{frame}

\section*{}
\subsection*{Probabilistic Mapping}

\begin{frame}
\frametitle{Single Measurement Ray}

\begin{itemize}
    \item Main idea: make use of occupancy grid mapping \emph{assumptions} and extract \emph{patterns} from probabilistic properties to find a \emph{computationally-efficient} solution
	\begin{itemize}
		\item Since the origin and direction of each measurement ray is known deterministically, the set of grid cells that the ray intersects is \emph{known through geometry}
		\item A depth reading follows the forward sensor model, which \emph{only depends} on the first occupied grid cell along the measurement ray
	\end{itemize}
\end{itemize}
\setcounter{subfigure}{0}
\begin{figure}[!htbp]
\vspace*{-0.25cm}
\centerline{
	\subfigure[True Grid]{
	\includegraphics[width=4.0cm]{simple_true_grid.png}\hspace*{-0.5cm}}
	\subfigure[Probabilistic Map]{
	\includegraphics[width=4.0cm]{simple_sweep.png}}
 }
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Single Measurement Ray}

\begin{itemize}
    \item Reduced Map
	\begin{itemize}
		\item Consider a reduced map of the $l$-th measurement ray, namely $r_l=\braces{\mathbf{r}_{l,1},\mathbf{r}_{l,2},\ldots,\mathbf{r}_{l,n_{r,l}}}$, consisting of grid cells \emph{intersected} by the measurement ray within the sensor limits, \emph{indexed by increasing distance}
		\item Since reduced map $r_l$ is composed of binary random variables, it has \emph{$2^{n_{r,l}}$} possible solutions
		\item If $\mathbf{r}_{l,1}$ is occupied (a priori probability $P(\mathbf{r}_{l,1}|X_{1:t-1},Z_{1:t-1})$), then each occupancy $\mathbf{r}_{l,2:n_{r,l}}$ does \emph{not} affect forward sensor model $P(z_{t,l}|\mathbf{r}_{l,1},X_t)$
		\item Similarly, if $\mathbf{r}_{l,2}$ is occupied (a priori probability $P(\bar{\mathbf{r}}_{l,1}|X_{1:t-1},Z_{1:t-1})P(\mathbf{r}_{l,2}|X_{1:t-1},Z_{1:t-1})$), then forward sensor model $P(z_{t,l}|\mathbf{r}_{l,2},X_t)$ is independent of grid cells of higher index
		\item This concept is extended for all grid cells in $r_l$
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Single Measurement Ray}

\begin{itemize}
	\item Define $P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ and its complement $P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=1-P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=\eta_{t,l}\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$
	\item The $k$-th unnormalized probability is written generally as
	\begin{align*}
		& \tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\
		&=P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\times 
		\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
		&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]\nonumber\\
		&\quad + \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
		&\quad\times p(z_{t,l}|\mathbf{r}_{l,k+},X_t)P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})
	\end{align*}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Single Measurement Ray}

\begin{itemize}
	\item The unnormalized complement $\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ takes a similar form
	\item Because $P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})+P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=1$, the normalizer is
	\begin{align*}
		\eta_{t,l}&=
		\bigg[\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
		&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1}
	\end{align*}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Single Measurement Ray}

\begin{itemize}
	\item Computational Efficiency
	\begin{itemize}
		\item Several of the terms required for $\eta_{t,l}$ and $\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ for each $k=1,2,\ldots,n_{r,l}$ are repeated and need not be computed more than once
		\item \emph{Computational order $\mathcal O(n_{r,l}+1)$} for all grid cells in reduced map $r$: each grid cell once and the empty map case
		\item Analytic solution is valid for any forward sensor model dependent on the closest occupied space
	\end{itemize}
\pause
	\item Comparison to other approaches
	\begin{itemize}
		\item It is commonly believed that the computational order is $\mathcal O(2^{n_{r,l}})$; the proposed method is \emph{$2^{n_{r,l}}\left(\frac{n_{r,l}}{n_{r,l}+1}\right)$ times faster} for the \emph{same solution}
		\item Proposed method completely avoids inaccuracies associated with heuristic solutions, learned methods, and log-odds ratio update assumptions
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Full Measurement Scan}

\begin{itemize}
	\item Two Methods
	\begin{itemize}
		\item \emph{Ray-By-Ray Inverse Sensor Model}: each measurement ray is evaluated individually for updating the map
		\item \emph{Synergistic Scan Inverse Sensor Model}: all measurement rays are evaluated simultaneously
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Ray-By-Ray Inverse Sensor Model}

\begin{itemize}
	\item Main idea: given a scan of measurement rays, update the map based on each measurement ray \emph{individually} and \emph{sequentially}
	\item Conditional probability:
	\begin{align*}
		P(\mathbf{m}_i|X_{1:t}&,Z_{1:t})\nonumber\\&
		=P((\dots(((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z})
	\end{align*}
	\item Benefit: the measurement rays maintain some dependency as they capture the same map
	\item Drawback: the best order for evaluating rays is not necessarily chosen
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Synergistic Scan Inverse Sensor Model}

\begin{itemize}
	\item Main idea: consider every measurement ray inside a scan \emph{simultaneously} to update all grid cells inside the FOV
	\item Probability derived from a Bayesian approach:
	\begin{align*}
		P(\mathbf{m}_i|X_{1:t},Z_{1:t})
		&=\xi_i P(\mathbf{m}_i|{X_{1:t-1}},Z_{1:t-1})\nonumber\\&\quad\times
		\prod_{l\in\mathcal L_i}
		\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}),
		\\
		\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})&
		\triangleq \frac{\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})}
	\end{align*}
	\item Benefit: synergistic update provides a method where all rays are considered simultaneously
	\item Drawback: required assumption that the measurement rays are independent, though they capture the same map
\end{itemize}

\end{frame}

\section*{}
\subsection*{Results}

\begin{frame}
\frametitle{Numerical Example}

\begin{itemize}
\item The proposed occupancy grid mapping algorithm and a heuristic solution are compared in a simulated scenario where a robot moves around a closed room
\begin{figure}
\vspace*{0.2cm}
\centerline{
\subfigure[Exact Proposed Occupancy Grid Mapping]{
	\includegraphics[width=1.5cm]{Compare_ISM_1.png}\hspace*{0.1cm}
	\includegraphics[width=1.5cm]{Compare_ISM_2.png}\hspace*{0.1cm}
	\includegraphics[width=1.5cm]{Compare_ISM_3.png}\hspace*{0.1cm}
	\includegraphics[width=1.5cm]{Compare_ISM_4.png}\hspace*{0.1cm}
	\includegraphics[width=1.5cm]{Compare_ISM_5.png}\hspace*{0.1cm}
	\includegraphics[width=1.5cm]{Compare_ISM_H.png}
	}}
\vspace*{0.2cm}
\centerline{
\subfigure[Heuristic Occupancy Grid Mapping]{
	\includegraphics[width=1.5cm]{Compare_Approx_1.png}\hspace*{0.1cm}
	\includegraphics[width=1.5cm]{Compare_Approx_2.png}\hspace*{0.1cm}
	\includegraphics[width=1.5cm]{Compare_Approx_3.png}\hspace*{0.1cm}
	\includegraphics[width=1.5cm]{Compare_Approx_4.png}\hspace*{0.1cm}
	\includegraphics[width=1.5cm]{Compare_Approx_5.png}\hspace*{0.1cm}
	\includegraphics[width=1.5cm]{Compare_Approx_H.png}
	}}
\end{figure}
\item The robot trajectory and the same set of measurements are used with both approaches
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Experimental Example}

\begin{itemize}
\item Kinect is placed under a desk and receives a single measurement scan
\setcounter{subfigure}{0}
\begin{figure}[!htbp]
\centerline{
\vspace*{0.5cm}
		\subfigure[Test Setup]{
\includegraphics[height=1.5cm]{TestSetup.png}\hspace*{0.5cm}}
		\subfigure[Camera View]{
\includegraphics[height=1.5cm]{KinectView.png}}}
\end{figure}
\vspace{-0.5cm}\pause
\item Exact and approximate inverse sensor models are compared
\end{itemize}
\setcounter{subfigure}{0}
\begin{figure}[!htbp]
	\centerline{
		\subfigure[Exact \newline \hspace*{0.4cm} Occupancy]{\includegraphics[height=2.1cm]{KinectDeskISM.png}}\hspace*{0.3cm}
		\subfigure[Exact \newline \hspace*{0.4cm} Uncertainty]{\includegraphics[height=2.1cm]{KinectDeskISM_H.png}}\hspace*{0.3cm}
		\subfigure[Approx. \newline \hspace*{0.4cm} Occupancy]{\includegraphics[height=2.1cm]{KinectDeskApprox.png}}\hspace*{0.3cm}
		\subfigure[Approx. \newline \hspace*{0.4cm} Uncertainty]{\includegraphics[height=2.1cm]{KinectDeskApprox_H.png}}
		}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Numerical Example}

\begin{itemize}
\item Similar concepts are extended to autonomous exploration
\item The poses are chosen to maximize map information gain
\setcounter{subfigure}{0}
\begin{figure}
\vspace*{0.2cm}
\centerline{
	\subfigure[$t=1$ sec]{\includegraphics[width=3.0cm]{RDShowcaseMotionPlanning1sec.png}}\hspace*{0.2cm}
	\subfigure[$t=100$ sec]{\includegraphics[width=3.0cm]{RDShowcaseMotionPlanning100sec.png}}\hspace*{0.2cm}
	\subfigure[$t=200$ sec]{\includegraphics[width=3.0cm]{RDShowcaseMotionPlanning200sec.png}}
	}
\vspace*{0.0cm}
\centerline{
	\subfigure[$t=300$ sec]{\includegraphics[width=3.0cm]{RDShowcaseMotionPlanning300sec.png}}\hspace*{0.2cm}
	\subfigure[$t=400$ sec]{\includegraphics[width=3.0cm]{RDShowcaseMotionPlanning400sec.png}}\hspace*{0.2cm}
	\subfigure[$t=500$ sec]{\includegraphics[width=3.0cm]{RDShowcaseMotionPlanning500sec.png}}
	}
\end{figure}
\end{itemize}

\end{frame}

\section*{}
\subsection*{Conclusions and Future Work}

\begin{frame}
\frametitle{Conclusions}
\begin{itemize}
        	\item Proposed an occupancy grid mapping technique that uses the \emph{exact probabilistic solution}
	\item Computational cost is reduced \emph{substantially} for \emph{real-time implementation} using probabilisitic properties and exploiting mathematical patterns
	\item Assumptions required for log-odds ratio Bayesian updating are avoided
	\item The proposed technique is compared with a heuristic solution to show the improvement in algorithm performance 
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Current and Future Work}
\begin{itemize}
    \item Current Reseach
    \begin{itemize}
        	\item Using the normalizer from this research, a method to \emph{predict map information gain} governs an \emph{autonomous exploration} algorithm
	\item Both the mapping and autonomous exploration algorithms are successfully tested with C++ ROS nodes
	\item We are completing ground vehicle tests where a robot autonomously explores a room
    \end{itemize}
    \item Future Work
    \begin{itemize}
        	\item Apply occupancy grid mapping and autonomous exploration in a \emph{3D} setting with a \emph{flying robot}
	\item Extend problem to \emph{multiple vehicles} for exploring uncertain spaces
    \end{itemize}
\end{itemize}
\end{frame}

\end{document}

### 2018-06-09 Conference/AIAA 2019/AIAA19.tex
added Conference folder

\documentclass[conf]{new-aiaa}
%\documentclass[journal]{new-aiaa} for journal papers
\usepackage[utf8]{inputenc}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage{longtable,tabularx}
\setlength\LTleft{0pt}

\title{Entropy-Based 3D Autonomous Exploration on Mars}

\author{Evan Kaufman\footnote{Graduate Research Assistant, Department of Mechanical and Aerospace Engineering, The George Washington University, evankaufman@gwu.edu.} and Taeyoung Lee\footnote{Associate Professor, Department of Mechanical and Aerospace Engineering, The George Washington University, tylee@gwu.edu.}}

\begin{document}

\maketitle

\begin{abstract}
In this paper, we present a novel approach to explore the 3D complex environments near the surface of Mars. We apply recent developments in probabilistic 3D occupancy grid mapping to develop an exploration strategy to maximize expected map information over a complex 3D terrain. The exploration technique predicts a measure of map uncertainty using Shannon's entropy for large 3D measurement scans to determine its future pose. Then, Dijkstra's algorithm is applied in 3D with the occupancy grid to provide collision-free waypoints to safe motion planning. These contributions are demonstrated with a simulation using a height map and image from the Viking Orbiter Visual Imaging Subsystem (VIS).
%These instructions give you guidelines for preparing papers for AIAA Technical Papers using \LaTeX{}. Define all symbols used in the abstract. Do not cite references in the abstract. The footnote on the first page should list the Job Title and AIAA Member Grade for each author, if known Authors do not have to be AIAA members.
\end{abstract}

%\section{Nomenclature}
%
%{\renewcommand\arraystretch{1.0}
%\noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
%$A$  & amplitude of oscillation \\
%$a$ &    cylinder diameter \\
%$C_p$& pressure coefficient \\
%$Cx$ & force coefficient in the \textit{x} direction \\
%$Cy$ & force coefficient in the \textit{y} direction \\
%c   & chord \\
%d$t$ & time step \\
%$Fx$ & $X$ component of the resultant pressure force acting on the vehicle \\
%$Fy$ & $Y$ component of the resultant pressure force acting on the vehicle \\
%$f, g$   & generic functions \\
%$h$  & height \\
%$i$  & time index during navigation \\
%$j$  & waypoint index \\
%$K$  & trailing-edge (TE) nondimensional angular deflection rate
%\end{longtable*}}

\section{Introduction}

Mars has been captured

%\lettrine{T}{his} document is a \LaTeX{} template for preparation of papers for AIAA Technical Conferences. If you are reading a hard-copy or .pdf version of this document, download the electronic file, new-aiaa.cls, and use it to prepare your manuscript.
%
%Authors using \url{https://www.overleaf.com} may simply open the AIAA template from the Overleaf gallery to work online; no local installation of any files is required. Authors using a local \LaTeX{} installation will need to open the template in Overleaf and use the ``Download as zip'' option from the project menu to download a local copy of the template files. To create your formatted manuscript, type your own text over sections of the template, or cut and paste from another document and then use the available markup styles. Note that special formatting such as subscripts, superscripts, and italics may be lost when you copy your text into the template from a Word Processing program such as Microsoft Word. See Sec. V for more detailed formatting guidelines.

\section{Procedure for Paper Submission}

All manuscripts are to be submitted electronically to the ScholarOne Abstracts site created for each conference. The manuscript upload will be enabled several weeks after acceptance notices have been sent. Presenting authors of accepted papers will receive an email with instructions when manuscript submission is open. It is important that presenting authors keep their email addresses up-to- date so they do not miss this notice.

Before completing manuscript submission, submitters must also select the copyright statement that will appear on the paper, and complete other acknowledgments. It is also necessary to click both the ``Accept'' and ``Save'' buttons to complete a submission. All completed manuscript submission will be confirmed by email. Completed submissions will also have a status of ``Accepted'' at the top of your manuscript submission page.

All files must be in pdf format. Please be sure that all security settings are removed from the pdf file before uploading to ensure proper processing of your manuscript file.

\section{General Guidelines}

The following section outlines general (nonformatting) guidelines to follow, drawn from the original AIAA Manuscript Preparation Kit. These guidelines are applicable to all authors (except as noted), and include information on the policies and practices relevant to the publication of your manuscript.

\subsection{Publication by AIAA}
Your manuscript cannot be published by AIAA if:
\begin{enumerate}
\item It has been published previously or

\item The work contains copyright-infringing material or

\item An appropriate copyright statement has not yet been selected.
\end{enumerate}

\subsection{Paper Review and Visa Considerations}

It is the responsibility of the author to obtain any required government or company reviews for their papers in advance of publication. Start early to determine if the reviews are required; this process can take several weeks.

If you plan to attend an AIAA Forum, technical conference or professional development course held in the United States and you require a visa for travel, it is incumbent upon you to apply for a visa with the U.S.~embassy (consular division) or consulate with ample time for processing.  To avoid bureaucratic problems, AIAA strongly suggests that you submit your formal application to U.S.~ authorities a minimum of 120 days in advance of the date of anticipated travel.

Prospective conference and course attendees requiring a visa to travel to the United States should first contact

AIAA to request an official letter of invitation. This letter and a copy of the conference call for papers should be presented along with the required documentation to the U.S. consular officials as part of the formal application process.  AIAA cannot directly intervene with the U.S. Department of State, consular offices, or embassies on behalf of individuals applying for visas. A letter of invitation can be requested at \url{https://www.aiaa.org/Secondary.aspx?id=6258} or you may contact AIAA Customer Service at \url{custserv@aiaa.org} for more information.

\subsection{Control ID Number vs Paper Number}

Your paper was assigned a control ID number at the time you submitted your abstract. It is critical that you reference the tracking number and conference name when contacting AIAA regarding your submission. The control ID number is \emph{not} the final AIAA paper number. The paper number, which appears in the format AIAA-201X-XXXX, will be used to refer to your paper in the program and in any publication format. It will not be assigned until shortly before the conference. \textbf{Do not include a paper number anywhere on your paper, as this number will be stamped automatically in the top right corner of your paper at the time of processing.}

\subsection{Copyright}

Before AIAA can print or publish any paper, the copyright information must be completed on our website. Failure to complete the form correctly could result in your paper not being published. The following fields must be completed:

\begin{enumerate}
\item Clearance Statement
\item Non-Infringement Statement
\item Publication Status Statement
\item One Copyright Assignment Statement (Select either A, B, C, or D)
\end{enumerate}

Be sure to read the copyright statements carefully. AIAA requires a copyright transfer from the author(s) to AIAA or a license to publish and distribute your material; government authors can assert that the work is in the public domain. If you are not sure which copyright statement to use, contact your legal department; AIAA cannot help you determine which statement to use. A hard copy of the form is found in the Author Kit for your reference. As you will be completing this form online, you do not need to fill out the hard-copy form. Do not include a copyright statement anywhere on your paper, and do not upload a copyright form with your paper. The correct statement will be stamped automatically at the time of processing.

\subsection{Submission Deadlines}

Manuscripts will be accepted for upload to the system from the receipt of the email invitation until the deadline set for the conference. You will be notified of the specific manuscript submission deadline in your acceptance letter, and the deadline will also be listed on the conference web page at AIAA. Please do not upload a draft version of your manuscript with the intent to upload a final version later. \textbf{Please review your manuscript very carefully before completing your submission to ensure that your paper is complete and final in all respects. Once the manuscript deadline has passed you will be locked out of the manuscript site so it is critical that you upload a final, carefully proofed document.}

On-line conference proceedings will be made accessible to attendees who have registered for the ``full conference'' when the conference opens. No updates to the proceedings will be made during the conference. Revision access will be provided to presenting authors starting the first day of the conference through 2000hrs/8pm Eastern Time 7 business days after the last day of the conference. Revisions submitted during this period are limited to minor changes (e.g.~typos, etc.). Changes to content are not permitted. The proceedings will then be updated shortly after that date.

To ensure conference quality, session chairs will enforce a ``no paper, no podium'' rule. This policy is intended to eliminate no-shows, to improve the quality of the conference for all participants, and to ensure that the published proceedings accurately represent the presentations made at a conference.

\subsection{Oversize Papers and Color Illustrations}
As of 2006, all papers presented at AIAA events are published in an electronic-only (PDF) format. Papers published will be reproduced in color automatically. There is no ``oversize paper'' or color illustration charge associated with published AIAA papers.

\section{Detailed Formatting Instructions}

The styles and formats for the AIAA Papers Template have been incorporated into the structure of this document. If you are using \LaTeX{}, please use this template to prepare your manuscript. A Microsoft Word template is also available from AIAA's website if you prefer to use Microsoft Word 2001 or later. Regardless of which program you use to prepare your manuscript, please use the formatting instructions contained in this document as a guide.

If you are using the AIAA Papers \LaTeX{} Template file to prepare your manuscript, you can simply type your own text over sections of this document, or cut and paste from another document and use the available markup styles. If you choose to cut and paste, select the text from your original document and choose Edit>Copy. (Do not select your title and author information, since the document spacing may be affected. It is a simple task to reenter your title and author information in the template.) Open the \LaTeX{} template file. Place your cursor in the text area of the template and select Edit>Paste. Please note that special formatting (e.g., subscripts, superscripts, italics) may be lost when you copy your text into the template. Use italics for emphasis; do not underline. Use the compiled \LaTeX{} pdf to see the most accurate representation of how your final paper will appear.

\subsection{Document Text}
The default font for AIAA papers is Times New Roman, 10-point size.  The first line of every paragraph should be indented, and all lines should be single-spaced. Default margins are 1'' on all sides. In the electronic version of this template, all margins and other formatting is preset. There should be no additional lines between paragraphs.

\begin{quoting}
Extended quotes, such as this example, are to be used when material being cited is longer than a few sentences, or the standard quotation format is not practical. In this \LaTeX template, the appropriate command environment is \verb|\begin{quoting}...\end{quoting}|. Extended quotes are to be in Times New Roman, 9-point font, indented 0.4'' and full justified.
\end{quoting}

\emph{NOTE:} If you are using the electronic \LaTeX{} template to format your manuscript, the required spacing and formatting will be applied automatically.

\subsection{Headings}
The title of your paper should be typed in bold, 18-point type, with capital and lower-case letters, and centered at the top of the page. The names of the authors, business or academic affiliation, city, and state/province should follow on separate lines below the title. The names of authors with the same affiliation can be listed on the same line above their collective affiliation information. Author names are centered, and affiliations are centered and in italic type. The affiliation line for each author is to include that author’s city, state, and zip/postal code (or city, province, zip/postal code and country, as appropriate). The first footnote (lower left-hand side) is to contain the job title and department name, street address/mail stop, and AIAA member grade for each author. Author email addresses may be included also.

Major headings (``sections'' in the \LaTeX{} template commands) are bold 11-point font, centered, and numbered with Roman numerals.

Subheadings (``subsections'' in the \LaTeX{} template commands) are bold, flush left, and numbered with capital letters.

Sub-Subheadings (``subsubsections'' in the \LaTeX{} template commands) are italic, flush left, and numbered (1. 2. 3. etc.)

\subsection{Abstract}
The abstract should appear at the beginning of your paper. It should be one paragraph long (not an introduction) and complete in itself (no reference numbers). It should indicate subjects dealt with in the paper and state the objectives of the investigation. Newly observed facts and conclusions of the experiment or argument discussed in the paper must be stated in summary form; readers should not have to read the paper to understand the abstract. The abstract should be bold, indented 3 picas (1/2'') on each side, and separated from the rest of the document by two blank lines.

\subsection{Nomenclature}
Papers with many symbols may benefit from a nomenclature list that defines all symbols with units, inserted between the abstract and the introduction. If one is used, it must contain all the symbology used in the manuscript, and the definitions should not be repeated in the text. In all cases, identify the symbols used if they are not widely recognized in the profession. Define acronyms in the text, not in the nomenclature.

\subsection{Footnotes and References}
Footnotes, where they appear, should be placed above the 1'' margin at the bottom of the page. To insert footnotes into the template, use the Insert>Footnote feature from the main menu as necessary. Footnotes are formatted automatically in the template, but if another medium is used, should appear in superscript as symbols in the sequence, *, $\dag$, $\ddag$, \S, \P, **, $\dag\dag$, $\ddag\ddag$, \S\S, etc.

List and number all references at the end of the paper. Corresponding bracketed numbers are used to cite references in the text \cite{vatistas1986reverse}, including citations that are an integral part of the sentence (e.g., ``It is shown in \cite{dornheim1996planetary} that\ldots '') or follow a mathematical expression: ``$A^{2} + B = C$ (Ref.~\cite{terster1997nasa}).'' For multiple citations, separate reference numbers with commas \cite{peyret2012computational,oates1997aerothermodynamics}, or use a dash to show a range \cite{volpe1994techniques,thompsonspacecraft,chi1993fluid}. Reference citations in the text should be in numerical order.

In the reference list, give all authors' names; do not use ``et al.'' unless there are six authors or more. Papers that have not been published should be cited as ``unpublished''; papers that have been submitted or accepted for publication should be cited as ``submitted for publication.'' Private communications and personal website should appear as footnotes rather than in the reference list.

References should be cited according to the standard publication reference style (for examples, see the ``References'' section of this template). Never edit titles in references to conform to AIAA style of spellings, abbreviations, etc. Names and locations of publishers should be listed; month and year should be included for reports and papers. For papers published in translation journals, please give the English citation first, followed by the original foreign language citation.

\subsection{Images, Figures and Tables}
All artwork, captions, figures, graphs, and tables will be reproduced exactly as submitted. Be sure to position any figures, tables, graphs, or pictures as you want them printed. AIAA will not be responsible for incorporating your figures, tables, etc. (Company logos and identification numbers are not permitted on your illustrations.)

Do not insert your tables and figures in text boxes. Figures should have no background, borders, or outlines. In the \LaTeX{} template, use the ``caption'' command to type caption text. Captions are bold with a single tab (no hyphen or other character) between the figure number and figure description.

% \begin{table}
% \caption{\label{tab:table1} Transitions selected for thermometry}
% \centering
% \begin{tabular}{lcccccc}
% \hline
% & Transition& & \multicolumn{2}{c}{}\\\cline{2-2}
% Line& $\nu''$& & $J'' $& Frequency, cm$^{-1}$& $FJ$, cm$^{-1}$& $G\nu $, cm$^{-1}$\\\hline
% a& 0& P$_{12}$& 2.5& 44069.416& 73.58& 948.66\\
% b& 1& R$_{2}$& 2.5& 42229.348& 73.41& 2824.76\\
% c& 2& R$_{21}$& 805& 40562.179& 71.37& 4672.68\\
% d& 0& R$_{2}$& 23.5& 42516.527& 1045.85& 948.76\\
% \hline
% \end{tabular}
% \end{table}

\begin{figure}[hbt!]
\centering
\includegraphics[width=.5\textwidth]{graph}
\caption{Magnetization as a function of applied fields.}
\end{figure}

Place figure captions below all figures; place table titles above the tables. If your figure has multiple parts, include the labels ``a),'' ``b),'' etc. below and to the left of each part, above the figure caption. Please verify that the figures and tables you mention in the text actually exist. \emph{Please do not include captions as part of the figures, and do not put captions in separate text boxes linked to the figures.} When citing a figure in the text, use the abbreviation ``Fig.'' except at the beginning of a sentence. Do not abbreviate ``Table.'' Number each different type of illustration (i.e., figures, tables, images) sequentially with relation to other illustrations of the same type.

Figure axis labels are often a source of confusion. Use words rather than symbols. As in the example to the right, write the quantity ``Magnetization'' rather than just ``M.'' Do not enclose units in parenthesis, but rather separate them from the preceding text by commas. Do not label axes only with units. As in Fig.~1, for example, write ``Magnetization, \si[per-mode=symbol]{\ampere\per\meter}'' or ``Magnetization, \si[inter-unit-product=\ensuremath{{}\cdot{}}]{\ampere\per\meter},'' not just ``A/m.'' Do not label axes with a ratio of quantities and units. For example, write ``Temperature, K,'' not ``Temperature/K.''

Multipliers can be especially confusing. Write ``Magnetization, \si[per-mode=symbol]{\kilo\ampere\per\meter}'' or ``Magnetization, \SI[per-mode=symbol]{e3}{\ampere\per\meter}.'' Do not write ``Magnetization (A/m) x 1000'' because the reader would not then know whether the top axis label in Fig.~1 meant 16000 A/m or 0.016 A/m. Figure labels must be legible, approximately 8--12 point type.

\subsection{Equations, Numbers, Symbols, and Abbreviations}
Equations are numbered consecutively, with equation numbers in parentheses flush right, as in Eq.~\eqref{sample:equation}. Insert a blank line on either side of the equation. To insert an equation into the \LaTeX{} document, use the \verb|\begin{equation}...\end{equation}| command environment.

A sample equation is included here, formatted using the preceding instructions. To make your equation more compact, you can use the solidus (/), the exp function, or appropriate exponents. Use parentheses to avoid ambiguities in denominators.

\begin{equation}
\label{sample:equation}
\int^{r_2}_0 F(r,\varphi){\rm d}r\,{\rm d}\varphi = [\sigma r_2/(2\mu_0)]\int^{\infty}_0\exp(-\lambda|z_j-z_i|)\lambda^{-1}J_1 (\lambda r_2)J_0 (\lambda r_i\,\lambda {\rm d}\lambda)
\end{equation}

Be sure that the symbols in your equation are defined before the equation appears, or immediately following. Italicize symbols ($T$ might refer to temperature, but T is the unit tesla). Refer to ``Eq.~(1),'' not ``(1)'' or ``equation (1)'' except at the beginning of a sentence: ``Equation (1) is\ldots'' Equations can be labeled other than ``Eq.'' should they represent inequalities, matrices, or boundary conditions. If what is represented is really more than one equation, the abbreviation ``Eqs.'' can be used.

Define abbreviations and acronyms the first time they are used in the text, even after they have already been defined in the abstract. Very common abbreviations such as AIAA, SI, ac, and dc do not have to be defined. Abbreviations that incorporate periods should not have spaces: write ``P.R.,'' not ``P.~R.'' Delete periods between initials if the abbreviation has three or more initials; e.g., U.N.~but ESA. Do not use abbreviations in the title unless they are unavoidable (for instance, ``AIAA'' in the title of this article).

\subsection{General Grammar and Preferred Usage}
Use only one space after periods or colons. Hyphenate complex modifiers: ``zero-field-cooled magnetization.'' Avoid dangling participles, such as, ``Using Eq.~(1), the potential was calculated.'' [It is not clear who or what used Eq.~(1).] Write instead ``The potential was calculated using Eq.~(1),'' or ``Using Eq.~(1), we calculated the potential.''

Insert a zero before decimal points: ``0.25,'' not ``.25.'' Use ``\si{\centi\meter\squared}'' not ``cc.'' Indicate sample dimensions as ``$\SI{0.1}{\centi\meter} \times \SI{0.2}{\centi\meter}$,'' not ``$0.1 \times \SI{0.2}{\centi\meter\squared}$.'' The preferred abbreviation for ``seconds'' is ``s,'' not ``sec.'' Do not mix complete spellings and abbreviations of units: use ``\si[per-mode=symbol]{\weber\per\meter\squared}'' or ``webers per square meter,'' not ``webers/m$^2$.'' When expressing a range of values, write ``7 to 9'' or ``7--9,'' not ``7$\sim$9.''

A parenthetical statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within parenthesis.) In American English, periods and commas are placed within quotation marks, like ``this period.'' Other punctuation is ``outside''! Avoid contractions; for example, write ``do not'' instead of ``don’t.'' The serial comma is preferred: ``A, B, and C'' instead of ``A, B and C.''

If you wish, you may write in the first person singular or plural and use the active voice (``I observed that\ldots'' or ``We observed that\ldots'' instead of ``It was observed that\ldots''). Remember to check spelling. If your native language is not English, please ask a native English-speaking colleague to proofread your paper.

Be aware of the different meanings of the homophones ``affect'' (usually a verb) and ``effect'' (usually a noun), ``complement'' and ``compliment,'' ``discreet'' and ``discrete,'' ``principal'' (e.g., ``principal investigator'') and ``principle'' (e.g., ``principle of measurement''). Do not confuse ``imply'' and ``infer.''

The word ``data'' is plural, not singular (i.e., ``data are,'' not ``data is''). The subscript for the permeability of vacuum $\mu_0$ is zero, not a lowercase letter ``o.'' The term for residual magnetization is ``remanence''; the adjective is ``remanent''; do not write ``remnance'' or ``remnant.'' The word ``micrometer'' is preferred over ``micron'' when spelling out this unit of measure. A graph within a graph is an ``inset,'' not an ``insert.'' The word ``alternatively'' is preferred to the word ``alternately'' (unless you really mean something that alternates). Use the word ``whereas'' instead of ``while'' (unless you are referring to simultaneous events). Do not use the word ``essentially'' to mean ``approximately'' or ``effectively.'' Do not use the word ``issue'' as a euphemism for ``problem.'' When compositions are not specified, separate chemical symbols by en-dashes; for example, ``NiMn'' indicates the intermetallic compound \ce{Ni_{0.5}Mn_{0.5}} whereas ``Ni--Mn'' indicates an alloy of some composition \ce{Ni_{x}Mn_{1-x}}.

Be aware of the different meanings of the homophones ``affect'' (usually a verb) and ``effect'' (usually a noun), ``complement'' and ``compliment,'' ``discreet'' and ``discrete,'' ``principal'' (e.g., ``principal investigator'') and ``principle'' (e.g., ``principle of measurement''). Do not confuse ``imply'' and ``infer.''

Prefixes such as ``non,'' ``sub,'' ``micro,'' ``multi,'' and ``"ultra'' are not independent words; they should be joined to the words they modify, usually without a hyphen. There is no period after the ``et'' in the abbreviation ``et al.'' The abbreviation ``i.e.,'' means ``that is,'' and the abbreviation ``e.g.,'' means ``for example'' (these abbreviations are not italicized). If desired, more detailed style and formatting instructions can be found in the AIAA style guide, AIAA Style (available from AIAA upon request).

\section{Conclusion}
A conclusion section is not required, though it is preferred. Although a conclusion may review the main points of the paper, do not replicate the abstract as the conclusion. A conclusion might elaborate on the importance of the work or suggest applications and extensions. \textit{Note that the conclusion section is the last section of the paper that should be numbered. The appendix (if present), acknowledgment, and references should be listed without numbers.}

\section*{Appendix}

An Appendix, if needed, should appear before the acknowledgments.

\section*{Acknowledgments}
An Acknowledgments section, if used, \textbf{immediately precedes} the References. Sponsorship and financial support acknowledgments should be included here. The preferred spelling of the word ``acknowledgment'' in American English is without the ``e'' after the ``g.'' Avoid expressions such as ``One of us (S.B.A.) would like to thank\ldots'' Instead, write ``F.~A.~Author thanks\ldots'' Sponsor and financial support acknowledgments are also to be listed in the ``acknowledgments'' section.

\bibliography{sample}

\end{document}

### 2018-06-09 Conference/ASC 2014/IEEE_Aerospace_Conference_2014_Abstract.tex
added Conference folder

\documentclass[10pt,a4paper]{article}
\usepackage[letterpaper,text={6.5in,9.5in},centering]{geometry}
\usepackage{epic,eepic}
\usepackage{amssymb,amsmath,times,subfigure,graphicx,theorem}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\title{Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers}
\author{Evan Kaufman\thanks{evankaufman@gwu.edu}}
\author{Kiren Caldwell\thanks{kirencaldwell@gmail.com}}
\author{Taeyoung Lee\thanks{tylee@gwu.edu}}
\affil{Department of Mechanical and Aerospace Engineering, The George Washington University}

\renewcommand\Authands{ and }

%\usepackage{warmread}
%\usepackage[all,import]{xy}
%\usepackage{eepic}
\usepackage{subfigure}

\usepackage{amsmath}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Figure \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}

\renewcommand{\baselinestretch}{1.2}
\date{September 3, 2013}

\renewcommand{\thesubsection}{\arabic{subsection}. }
\renewcommand{\thesubsubsection}{\arabic{subsection}.\arabic{subsubsection} }

\theoremstyle{plain}\theorembodyfont{\normalfont}
\newtheorem{prob}{Question}[section]
%\renewcommand{\theprob}{\arabic{section}.\arabic{prob}}
\renewcommand{\theprob}{\arabic{prob}}

\newenvironment{subprob}%
{\renewcommand{\theenumi}{\alph{enumi}}\renewcommand{\labelenumi}{(\theenumi)}\begin{enumerate}}%
{\end{enumerate}}%

\begin{document}
\maketitle
We designed an experimental platform for testing nonlinear control on a fully-actuated system. We have designed a nonlinear controller, developed in a geometric framework such that the nonlinearities of the attitude configuration manifold are explicitly considered in controller design, which is tested extensively with numerical simulations. The research presented in this paper provides a technique to experimentally validate this control on a dynamic system with six decoupled degrees of freedom--three position and three attitude--with an unmanned aerial vehicle. Quadrotor UAVs have been tested for geometric control; however, quadrotors are unable to hover when the propeller thrust is not parallel to the gravitational force. Hence, we propose a hexrotor UAV testbed that is fully-actuated to maneuver in the special Euclidean group $\SE$. The system design is completed and the actuators are successfully tested.

The UAV testbed for fully-actuated geometric control uses variable pitch propeller actuators, capable of quick reversal of thrust, so that the UAV may produce large angle rotations. The UAV frame holds two of these propellers on each of three orthogonal planes. Thorough calibration tests show a wide range of thrust values from the motor and propeller combination, yielding nearly identical thrust magnitudes in both positive and negative directions. The expected flight of this UAV will take place at George Washington University's (GWU) Motion Capture (MoCa) laboratory, which uses line-of-sight measurements to capture markers rigidly attached to the UAVs. We expect preliminary flight results of this UAV in the coming months.

The key results of the testing show how the variable pitch propellers are able to accurately and consistently provide up to 3.5 Newtons of force in each direction; these results yield large admissible hovering and maneuvering regions of attitude for the UAV. Hence, we provide evidence that the system is not constrained to a single plane for hovering; maneuvers involving more complex and nontrivial trajectories are possible. Furthermore, these results serve to improve aspects of numerical modeling of the control, namely saturation regions and dynamic response times. These results confirmed the feasibility of the experimental testbed.

Geometric control has direct applications with spacecraft formation control. Controllers using inertial measurement unit (IMU) data to determine relative attitude with local coordinates yield a complicated controller structure and measurement error accumulation of the sensors. A technique developed at GWU involves group of spacecraft using line-of-sight measurements in a geometric framework to control relative attitude directly with no accumulation of sensor error. Conventional testbeds to validate this control scheme such as spherical air bearings have severe restrictions in tumbling motions; however, the UAV testbed presented in this paper is capable of hover in a wide range of directions to perform ground tests of large angle rotational maneuvers of spacecraft attitude dynamics. This hardware platform is also readily extended to other single- or multi-agent systems for cooperative controls. Experimental results serve as evidence for the merit of geometric control in space systems.

\end{document}

### 2018-06-09 Conference/ASC 2014/Presentation/ACC13.1.tex
added Conference folder

\documentclass[11pt,professionalfonts,hyperref={pdftex,pdfpagemode=none,pdfstartview=FitH}]{beamer}
%\usepackage{times}
%\usefonttheme{serif}
%\usepackage{helvet}
%\usepackage{amsmath,amssymb}
\usepackage{graphicx,multirow}
\usepackage[scriptsize]{subfigure}
%\usepackage{movie15}

%\usepackage{warmread}
%\usepackage[all,import]{xy}

%\renewcommand\mathfamilydefault{\rmdefault}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand{\refprop}[1]{Proposition \ref{prop:#1}}

\definecolor{mygray}{gray}{0.9}

\mode<presentation> {
  \usetheme{Warsaw}
  \usefonttheme{serif}
  \setbeamercovered{transparent}
}

\newcommand{\mypaper}{}

\setbeamertemplate{footline}%{split theme}
{%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm]{title in head/foot}
%    \usebeamerfont{title in head/foot}\mypaper\hfill \insertframenumber/\inserttotalframenumber
    \usebeamerfont{title in head/foot}\hfill \insertframenumber/\inserttotalframenumber
  \end{beamercolorbox}}%
  \vskip0pt%
} \setbeamercolor{box}{fg=black,bg=yellow}

\title[Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers]{\large Design and Development of a Free-Floating\\Hexrotor UAV for 6-DOF Maneuvers}

\author{\vspace*{-0.3cm}}

\institute{\footnotesize
{\normalsize Taeyoung Lee}\vspace*{0.2cm}\\
  Mechanical and Aerospace Engineering\\ George Washington University}

\date{}

\definecolor{tmp}{rgb}{0.804,0.941,1.0}
\setbeamercolor{numerical}{fg=black,bg=tmp}
\setbeamercolor{exact}{fg=black,bg=red}

\newtheorem{prop}{Proposition}

\renewcommand{\emph}[1]{\textit{\textbf{\color{blue}{#1}}}}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section*{}
\subsection*{Introduction}

\begin{frame}
\frametitle{Introduction}
\begin{itemize}
    \item Attitude Dynamics of a Rigid Body
    \begin{itemize}
    	\item Examples: robotics, flight dynamics, and space systems
    	\item Attitude: orientation of a body-fixed frame with respect to a reference frame
		\item Special Orthogonal Group
        \begin{align*}
        \SO = \{ R\in\Re^{3\times 3}\,|\, R^TR=I,\quad\mathrm{det}[R]=1\}.
        \end{align*}
        \item Compact nonlinear configuration manifold
        \item Unique dynamic properties that cannot be observed in $\Re^n$
    \end{itemize}
\end{itemize}

\only<1->{
\begin{figure}
\centerline{
    \includegraphics[width=2.1cm]{f22.jpg}\hspace*{0.1cm}
    \includegraphics[width=2.1cm]{sat.jpg}\hspace*{0.1cm}
    \includegraphics[width=2.1cm]{ast.jpg}\hspace*{0.1cm}
    \includegraphics[width=2.1cm]{rov.pdf}
}
\end{figure}}

\end{frame}

\begin{frame}
\frametitle{Introduction}
\framesubtitle{Attitude Control Systems: Classification Based on Representation}

\begin{itemize}
	\item Euler angles
	\begin{itemize}
	\item Any minimal representation of attitude has singularities
	\item Switching between several types to cover $\SO$
	\item Complex expressions for trigonometric functions
	\end{itemize}
\vspace*{0.3cm}\pause
	\item Quaternions in $\Sph^3=\{q\in\Re^4\,|\, \|q\|=1\}$
	\begin{itemize}
	\item No singularities
	\item Ambiguity: $\Sph^3$ double-covers $\SO$
	\pause
	\item Unwinding: a rigid body rotates unnecessarily through a large angle, even if the initial attitude error is small %{\footnotesize [Bhat, Bernstein 00]}
	\item Sensitive to measurement noise %{\footnotesize [Mayhew et. al. 11]}
	\pause
	\item Attitude in $\SO$ should be \emph{carefully} lifted to $\Sph^3$
	\end{itemize}
\vspace*{0.3cm}\pause
	\item Rotation matrices in $\SO$	
	\begin{itemize}
	\item No singularity, and no ambiguity
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Introduction}
\framesubtitle{Attitude Control Systems: Classification Based on Stability Properties}

\begin{itemize}
	\item Topological Obstruction
	\begin{itemize}
	\item No \emph{smooth} control system to \emph{globally} asymptotically stabilize a desired attitude
	\item Region of attraction is homeomorphic to $\Re^n$
	\end{itemize}
\vspace*{0.3cm}\pause
	\item \emph{Almost} Global Asymptotic Stabilization
	\begin{itemize}
	\item Smooth control system
	\item Desired attitude is asymptotically stable
	\item Region of attraction excludes a zero-measure set
	\end{itemize}
\vspace*{0.3cm}\pause
	\item \emph{Hybrid} Global Asymptotic Stabilization
	\begin{itemize}
	\item Hysteresis-based switching control system
	\item Asymptotic stability from hybrid invariance principle
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Introduction}
\framesubtitle{Contributions}

\begin{itemize}
\item \emph{Exponential} Stability
	\begin{itemize}
	\item Prior Work: asymptotic stability
		\begin{itemize}
		\item Stabilization: LaSalle's principle
		\item Tracking: an exogenous system to reformulate it into a hybrid autonomous system
		\end{itemize} 
	\item Lyapunov analysis for time-varying systems on $\SO$
	\end{itemize}
\vspace*{0.3cm}\pause
\item Simple Hybrid Control System for \emph{Global} Attractiveness
	\begin{itemize}
	\item Prior Work: error function based on stretched rotations, etc
	\item New intuitive form of attitude error function is introduced for simpler, explicit stability conditions
	\end{itemize}
\vspace*{0.3cm}\pause
\item \emph{Robustness}
	\begin{itemize}
	\item Prior Work: integral term is used heuristically without stability analysis
	\item New integral term is introduced to guarantee stability with fixed disturbances
	\end{itemize}
\end{itemize}
\end{frame}

\section*{}
\subsection*{Problem Formulation}

\begin{frame}
\frametitle{Problem Formulation}

\begin{itemize}
    \item Attitude Dynamics of a Rigid Body
    \begin{itemize}
    \item Equations of motion
    \begin{gather*}
J\dot \Omega + \Omega\times J\Omega = u +\Delta,\\
\dot R = R\hat\Omega = \hat\omega R,\quad (\omega=R\Omega).
\end{gather*}
	\item Fixed unknown disturbance: $\Delta\in\Re^3$
    \end{itemize}
\vspace*{0.4cm}\pause    
    \item Attitude Tracking Problem
    \begin{itemize}
    \item Desired attitude trajectory: $R_d(t):[t_0,t_f]\rightarrow\SO$
    \begin{gather*}
    \dot R_d = R_d \hat\Omega_d = \hat\omega_d R_d,\quad (\omega_d=R_d\Omega_d).
    \end{gather*}
    \item Goal: Design control input $u$ such that $(R,\omega)=(R_d,\omega_d)$ is globally exponentially stable
    \end{itemize}
\end{itemize}
\end{frame}

\section*{}
\subsection*{Control Systems Design}

\begin{frame}
\frametitle{Control Systems Design}

\begin{itemize}
\item Four Types of Attitude Control Systems
\end{itemize}
\vspace*{-0.8cm}
\begin{figure}
\setlength{\unitlength}{0.1\textwidth}\footnotesize
\begin{picture}(10,6)(-5.3,-2.8)
\only<1->{
%
\put(-6,0.9){\shortstack[c]{Without\\ disturbance\\ $\Delta=0$}}
\put(-3.3,2.5){\shortstack[c]{Smooth control}}
\put(-4.3,0.3){\framebox(4,2)[c]{\shortstack[c]{Smooth Attitude Tracking\\for \textit{Almost Global}\\ \textit{Exponential} Stability}}}}
%
\only<2->{
\put(-0.3,1.3){\vector(1,0){0.8}}
\put(1.6,2.5){\shortstack[c]{Hybrid control}}
\put(0.5,0.3){\dashbox{0.08}(4,2)[c]{\shortstack[c]{Hybrid Attitude Tracking\\for \textit{Global}\\ \textit{Exponential} Stability}}}}
%
\only<3->{
\put(-2.3,0.3){\vector(0,-1){0.6}}
\put(2.5,0.3){\vector(0,-1){0.6}}
\put(-6,-1.7){\shortstack[c]{With\\ disturbance\\ $\Delta\neq0$}}
{\linethickness{1.0pt}
\put(-4.3,-2.3){\framebox(4,2)[c]{\shortstack[c]{Smooth Attitude Tracking\\for \textit{\bf Robust} \textit{Almost Global}\\ \textit{Exponential} Stability}}}}
{\linethickness{1.0pt}
\put(0.5,-2.3){\dashbox{0.08}(4,2)[c]{\shortstack[c]{Hybrid Attitude Tracking\\for \textit{{\bf Robust} Global}\\ \textit{Exponential} Stability}}}}}
\end{picture}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Smooth Attitude Tracking}

\begin{itemize}
\item Attitude Error Function: {\small $\Psi(R,R_d):\SO^2\rightarrow \Re_+$}
\begin{itemize}
\item Choose characteristic direction $b_1,b_2\in\Sph^2$ fixed to the body\\
(ex: direction of an antenna of satellite)
\item \textit{Current} direction of $b_i$ in the inertial frame: $r_i = R b_i\in\Sph^2$
\item \textit{Desired} direction of $b_i$ in the inertial frame: $r_{i_d} = R_d b_i\in\Sph^2$
\item Direction error function: $\Psi_i=\frac{1}{2}\|r_i-r_{i_d}\|^2=1- r_i \cdot r_{i_d}$
\item Attitude error function:
\begin{gather*}
\Psi = k_1 \Psi_1 + k_2 \Psi_2,\quad (k_1\neq k_2).
\end{gather*}
\end{itemize}
\pause\vspace*{0.1cm}
\item Attitude Error Vector: {\small $e_r(R,R_d):\SO^2\rightarrow \Re^3$}
	\begin{itemize}
	\item Derivatives of the error function
	\begin{align*}
	\frac{d}{d\epsilon}\bigg|_{\epsilon=0} \Psi(R\exp(\epsilon\hat\eta),R_d) = e_r\cdot \eta.%(R^T r_{i_d}\times b_i)\cdot \eta \equiv e_{r_i}\cdot\eta.
	\end{align*}
%	\item Attitude error vector: $e_r = k_1 e_{r_1} + k_2 e_{r_2}$.
	\end{itemize}
	\vspace*{0.1cm}
\pause	
\item Angular Velocity Error Vector: $e_\Omega=\Omega- R^T\omega_d$.	
\end{itemize}
\end{frame}

%\begin{frame}
%\frametitle{Smooth Attitude Tracking}
%
%\begin{itemize}
%\item Angular Velocity Error Vector: $e_\Omega=\Omega- R^T\omega_d$.
%\end{itemize}
%
%\begin{prop}
%\begin{itemize}\setlength{\itemsep}{0pt}
%\item[(i)] $\Psi$ is positive definite about $R=R_d$.
%\item[(ii)] the left-trivialized derivative of $\Psi$ is $\T^*_I \L_R\, (\D_R\Psi(R,R_d))= e_r$.
%\item[(iii)] $\dot\Psi= e_r\cdot e_\Omega$, $\dot e_r\leq (k_1+k_2)\|e_\Omega\|$.
%\item[(iv)] there exist $h_1,h_2>0$ such that
%\begin{align*}
%h_1\|e_r\|^2 \leq \Psi(R,R_d) \leq h_2 \|e_r\|^2.
%\end{align*}
%\end{itemize}
%\end{prop}
%\end{frame}

\begin{frame}
\frametitle{Smooth Attitude Tracking}

\begin{itemize}
\item Smooth Attitude Tracking Control on $\SO$

\begin{prop}\label{prop:AGASSO}
For $k_1,k_2,k_\omega >0$ with $k_1\neq k_2$ a control input is chosen as
\begin{align*}
u & = -e_r -k_\Omega e_\Omega  +(R^T\omega_d)^\wedge JR^T\omega_d
+ JR^T\dot\omega_d.\label{eqn:uSO}
\end{align*}
Then, the following properties hold:
\begin{enumerate}
\item There are four equilibrium attitudes, given by $R_d$ and $180^\circ$ rotation of $R_d$ about each of $r_{1_d}$, $r_{2_d}$, and $r_{1_d}\times r_{2_d}$.
\item The desired attitude is almost globally asymptotically stable and almost semi-globally exponentially stable.
\item The three undesired equilibria are unstable.
\end{enumerate}
\end{prop}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Numerical Example}

\begin{itemize}
\item Simulation Parameters
	\begin{itemize}
	\item Inertia Matrix: $J=\mathrm{diag}[3,\,2,\,1]\,\mathrm{kgm^2}$.
	\vspace*{0.3cm}
	\item Desired Attitude Trajectory
	\begin{gather*}
	R_d(t) =\exp(\psi(t)\hat e_3)\exp(\theta(t)\hat e_2)\exp(\phi(t)\hat e_1),\\
	\phi(t)=\sin 0.5t,\quad \theta(t)=0.1(-1+t),\quad \psi(t)=1-\cos t.
	\end{gather*}
	\item Controller Parameters
	\begin{gather*}
	k_1=10,\quad k_2=11,\quad k_\Omega=4.42
	\end{gather*}
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Numerical Example}

\only<1>{
\begin{itemize}
\item Initial Conditions: {\small $R(0)=I$, $\Omega(0)=0_{3\times 1}$ }
\begin{figure}
\centerline{
	\subfigure[Attitude Error $\|R-R_d\|$]{\includegraphics[width=0.47\textwidth]{ACC13_1_neR.pdf}}
	\hfill
	\subfigure[Angular Velocity Error $e_\Omega$]{\includegraphics[width=0.47\textwidth]{ACC13_1_eW.pdf}}
	}
\end{figure}
\end{itemize}}

\only<2>{
\begin{itemize}
\item Initial Conditions: {\small $R(0) = \exp(0.9999\pi (r_{1_d}\times r_{2_d})^\wedge) R_d(0)$, $\Omega(0)=R(0)^T \omega_d(0)$ (close to one of the undesired equilibria)}
\begin{figure}\setcounter{subfigure}{0}
\centerline{
	\subfigure[Attitude Error $\|R-R_d\|$]{\includegraphics[width=0.47\textwidth]{ACC13_3_neR.pdf}}
	\hfill
	\subfigure[Angular Velocity Error $e_\Omega$]{\includegraphics[width=0.47\textwidth]{ACC13_3_eW.pdf}}
	}
\end{figure}
\end{itemize}}

\end{frame}

\begin{frame}
\frametitle{Hybrid Attitude Tracking}

\only<1>{
\begin{itemize}
\item Smooth Attitude Tracking
\begin{figure}\setcounter{subfigure}{0}
\centerline{
	\subfigure[Error Function $\Psi_i=1-r_i\cdot r_{i_d}$]{\includegraphics[width=0.47\textwidth]{ACC13_H_P0.pdf}}
	\hfill
	\subfigure[Error Vector $e_{r_i}$]{\includegraphics[width=0.475\textwidth]{ACC13_H_er0.pdf}}
	}
\caption{Direction error function $\Psi_i$ and error vector $e_{r_1}$ as a function of angle between $r_i$ and $r_{i_d}$}
\end{figure}
\end{itemize}}

\only<2>{
\begin{itemize}
\item Hybrid Attitude Tracking
\begin{itemize}
\item Introduce \textbf{\textit{\color{red}{expelling}}} functions: $\Psi_{E_i}=\alpha+\beta r_i\cdot (r_{1_d}\times r_{2_d})$
\end{itemize}
\begin{figure}\setcounter{subfigure}{0}
\centerline{
	\subfigure[Error Function $\Psi_i,\Psi_{E_i}$]{\includegraphics[width=0.47\textwidth]{ACC13_H_P1.pdf}}
	\hfill
	\subfigure[Error Vector $e_{r_i},e_{E_i}$]{\includegraphics[width=0.475\textwidth]{ACC13_H_er1.pdf}}
	}
\caption{Direction error function $\Psi_i$ and error vector $e_{r_1}$ as a function of angle between $r_i$ and $r_{i_d}$}
\end{figure}
\end{itemize}
}
\end{frame}

\begin{frame}
\frametitle{Hybrid Attitude Tracking}

\only<1>{
\begin{itemize}
\item Switching Algorithm
\begin{itemize}
\item Choose the error function with the minimum value
\item Sensitive to noise, chattering
\end{itemize}
\begin{figure}\setcounter{subfigure}{0}
\centerline{
	\subfigure[Error Function $\Psi_i,\Psi_{E_i}$]{\includegraphics[width=0.47\textwidth]{ACC13_H_P2.pdf}}
	\hfill
	\subfigure[Error Vector $e_{r_i},e_{E_i}$]{\includegraphics[width=0.475\textwidth]{ACC13_H_er2.pdf}}
	}
\caption{Direction error function $\Psi_i$ and error vector $e_{r_1}$ as a function of angle between $r_i$ and $r_{i_d}$}
\end{figure}
\end{itemize}}

\only<2>{
\begin{itemize}
\item Hysteresis-Based Switching Algorithm
\begin{itemize}
\item Switch if the difference to the min. value is greater than $\delta$
\item Robust to noise, eliminate chattering
\end{itemize}
\begin{figure}\setcounter{subfigure}{0}
\centerline{
	\subfigure[Error Function $\Psi_i,\Psi_{E_i}$]{\includegraphics[width=0.47\textwidth]{ACC13_H_P3.pdf}}
	\hfill
	\subfigure[Error Vector $e_{r_i},e_{E_i}$]{\includegraphics[width=0.475\textwidth]{ACC13_H_er3.pdf}}
	}
\caption{Direction error function $\Psi_i$ and error vector $e_{r_1}$ as a function of angle between $r_i$ and $r_{i_d}$}
\end{figure}
\end{itemize}}

\end{frame}

\begin{frame}
\frametitle{Hybrid Attitude Tracking}

\begin{itemize}
\item Global Exponential Attitude Tracking on $\SO$

\begin{prop}\label{prop:GES}
Consider the hybrid control system described above. For given constants $k_1,k_2,\alpha,\beta$ satisfying $k_1,k_2>0$, $k_1\neq k_2$, $1<\alpha<2$ and $|\beta|< \alpha-1$, choose the hysteresis gap $\delta$ such that
\begin{align}
0<\delta < \min\{k_1,k_2\}\min\{2-\alpha, \alpha-|\beta|-1\}.\label{eqn:deltaSO}
\end{align}
Then, the desired equilibrium $(R_d,\omega_d)$ is globally exponentially stable.
\end{prop}

\begin{itemize}
\item Sufficient condition for stability is \emph{explicitly} given by (1)
\item Stronger global \emph{exponential} stability is guaranteed
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Numerical Example}

\only<1>{
\begin{itemize}
\item Smooth Attitude Control
\begin{itemize}
\item Initial Conditions: {\small $R(0) = \exp(0.9999\pi (r_{1_d}\times r_{2_d})^\wedge) R_d(0)$, $\Omega(0)=R(0)^T \omega_d(0)$ (close to one of the undesired equilibria)}
\end{itemize}
\begin{figure}\setcounter{subfigure}{0}
\centerline{
	\subfigure[Attitude Error $\|R-R_d\|$]{\includegraphics[width=0.47\textwidth]{ACC13_3_neR.pdf}}
	\hfill
	\subfigure[Angular Velocity Error $e_\Omega$]{\includegraphics[width=0.47\textwidth]{ACC13_3_eW.pdf}}
	}
\end{figure}
\end{itemize}}

\only<2>{
\begin{itemize}
\item Hybrid Attitude Control
\begin{itemize}
\item Initial Conditions: {\small $R(0) = \exp(0.9999\pi (r_{1_d}\times r_{2_d})^\wedge) R_d(0)$, $\Omega(0)=R(0)^T \omega_d(0)$ (close to one of the undesired equilibria)}
\end{itemize}
\begin{figure}\setcounter{subfigure}{0}
\centerline{
	\subfigure[Attitude Error $\|R-R_d\|$]{\includegraphics[width=0.47\textwidth]{ACC13_4_neR.pdf}}
	\hfill
	\subfigure[Angular Velocity Error $e_\Omega$]{\includegraphics[width=0.47\textwidth]{ACC13_4_eW.pdf}}
	}
\end{figure}
\end{itemize}}

\end{frame}

\begin{frame}
\frametitle{Control Systems Design}

\begin{itemize}
\item Four Types of Attitude Control Systems
\end{itemize}
\vspace*{-0.8cm}
\begin{figure}
\setlength{\unitlength}{0.1\textwidth}\footnotesize
\begin{picture}(10,6)(-5.3,-2.8)
\only<1->{
%
\put(-6,0.9){\shortstack[c]{Without\\ disturbance\\ $\Delta=0$}}
\put(-3.3,2.5){\shortstack[c]{Smooth control}}
\put(-4.3,0.3){\framebox(4,2)[c]{\shortstack[c]{Smooth Attitude Tracking\\for \textit{Almost Global}\\ \textit{Exponential} Stability}}}}
%
\only<1->{
\put(-0.3,1.3){\vector(1,0){0.8}}
\put(1.6,2.5){\shortstack[c]{Hybrid control}}
\put(0.5,0.3){\dashbox{0.08}(4,2)[c]{\shortstack[c]{Hybrid Attitude Tracking\\for \textit{Global}\\ \textit{Exponential} Stability}}}}
%
\only<2->{
\put(-2.3,0.3){\vector(0,-1){0.6}}
\put(2.5,0.3){\vector(0,-1){0.6}}
\put(-6,-1.7){\shortstack[c]{With\\ disturbance\\ $\Delta\neq0$}}
{\linethickness{1.0pt}
\put(-4.3,-2.3){\framebox(4,2)[c]{\shortstack[c]{Smooth Attitude Tracking\\for \textit{\bf Robust} \textit{Almost Global}\\ \textit{Exponential} Stability}}}}
{\linethickness{1.0pt}
\put(0.5,-2.3){\dashbox{0.08}(4,2)[c]{\shortstack[c]{Hybrid Attitude Tracking\\for \textit{{\bf Robust} Global}\\ \textit{Exponential} Stability}}}}}
\end{picture}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Robust Attitude Tracking}

\begin{itemize}
\item Generalized Integral Term
\begin{itemize}
\item Definition

\begin{center}\hspace*{0.01cm}
\begin{beamercolorbox}[wd=8cm,dp=0.15cm,ht=1.1cm,sep=0.05cm,center,shade=on]{numerical}
$e_I(t) = \displaystyle\int_{0}^t ce_r(\tau)+e_\Omega(\tau) \,d\tau.$
\end{beamercolorbox}
\end{center}

%
%
%\begin{align}
%e_I(t) = \int_{0}^t ce_r(\tau)+e_\Omega(\tau) \,d\tau,\label{eqn:eI}
%\end{align}
%for some positive constant $c$.
\item The first part is equivalent to the \textit{classical} integral term
\item The second term increases the proportional term effectively\\
(Note: $\dot e_r \neq e_\Omega$) 
\pause\vspace*{0.3cm}
\item Control inputs are augmented with the proposed, generalized integral term
\item Stabilities are guaranteed by rigorous Lyapunov analysis for time-varying systems
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Robust Attitude Tracking}

\begin{itemize}
\item Smooth Attitude Tracking
\begin{prop}\label{prop:RAGAS}
Control input is chosen as
{\small
\begin{align*}
u & = -e_r -k_\Omega e_\Omega -k_I e_I+(R^T\omega_d)^\wedge JR^T\omega_d
+ JR^T\dot\omega_d.\label{eqn:uI}
\end{align*}}
Then, the desired equilibrium is almost globally asymptotically stable, and locally exponentially stable with respect to $e_r,e_\Omega$. The integral term $e_I$ is globally uniformly bounded.
\end{prop}
\pause\vspace*{0.1cm}
\item Hybrid Attitude Tracking
\begin{prop}
The desired equilibrium $(R_d,\omega_d)$ is globally exponentially stable with respect to $e_r$ and $e_\Omega$, and the integral term $e_{I}$ is globally uniformly bounded.
\end{prop}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Numerical Example}

\only<1>{
\begin{itemize}
\item Hybrid Attitude Control
\begin{itemize}
\item Initial Conditions: {\small $R(0) = \exp(0.9999\pi (r_{1_d}\times r_{2_d})^\wedge) R_d(0)$, $\Omega(0)=R(0)^T \omega_d(0)$ (close to one of the undesired equilibria)}
\item Fixed Disturbance: $\Delta=[-1,2,1]^T$.
\end{itemize}
\begin{figure}\setcounter{subfigure}{0}
\centerline{
	\subfigure[Attitude Error $\|R-R_d\|$]{\includegraphics[width=0.47\textwidth]{ACC13_7_neR.pdf}}
	\hfill
	\subfigure[Angular Velocity Error $e_\Omega$]{\includegraphics[width=0.47\textwidth]{ACC13_7_eW.pdf}}
	}
\end{figure}
\end{itemize}}

\only<2>{
\begin{itemize}
\item Robust Hybrid Attitude Control
\begin{itemize}
\item Initial Conditions: {\small $R(0) = \exp(0.9999\pi (r_{1_d}\times r_{2_d})^\wedge) R_d(0)$, $\Omega(0)=R(0)^T \omega_d(0)$ (close to one of the undesired equilibria)}
\item Fixed Disturbance: $\Delta=[-1,2,1]^T$.
\end{itemize}
\begin{figure}\setcounter{subfigure}{0}
\centerline{
	\subfigure[Attitude Error $\|R-R_d\|$]{\includegraphics[width=0.47\textwidth]{ACC13_8_neR.pdf}}
	\hfill
	\subfigure[Angular Velocity Error $e_\Omega$]{\includegraphics[width=0.47\textwidth]{ACC13_8_eW.pdf}}
	}
\end{figure}
\end{itemize}}
\end{frame}

\section*{}
\subsection*{Conclusions}

\begin{frame}
\frametitle{Conclusions}
\begin{itemize}
    \item Robust Global Exponential Attitude Tracking
    \begin{itemize}
        \item Lyapunov stability analysis for time-varying systems
        \item Hybrid attitude control with simpler controller design
        \item Robust control with new generalized integral control term
    \end{itemize}
\vspace*{0.4cm}\pause
    \item Geometric Controls on $\SO$
    \begin{itemize}
        \item \textit{Coordinate-free} nonlinear controls on a nonlinear manifold.
        \item Completely avoid \textit{singularities, ambiguities, and complexities}
        \item No need to introduce exogenous system to lift $\SO$ to $\Sph^3$
    \end{itemize}
\end{itemize}
\end{frame}

\end{document}

### 2018-06-09 Conference/ECC 2014/ECC14.tex
added Conference folder

\documentclass[letterpaper, 10pt, conference]{ieeeconf}
\IEEEoverridecommandlockouts \overrideIEEEmargins
\usepackage{amsmath,amssymb,url}
\usepackage{graphicx,subfigure,tikz}
\usepackage{color}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\ave}[1]{\mathrm{E}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand\circled[1]{%
  \tikz[baseline=(C.base)]\node[draw,circle,inner sep=0.5pt](C) {#1};\!
}

\title{\LARGE \bf
%Recursive Kalman-Mahanobis JPDAF Minimization to avoid Coalescence of Estimates in Close Proximity
Coalescence Avoiding Optimal Joint Probabilistic Data Association Filter}
\author{Evan Kaufman and Taeyoung Lee
 \thanks{Evan Kaufman and Taeyoung Lee are with Department of Aerospace Engineering, George Washington University, Washington, DC. Email: {\tt\footnotesize \{evankaufman, tylee\}@gwu.edu }}
}

\newcommand{\EditTL}[1]{{\color{red}\protect #1}}
\renewcommand{\EditTL}[1]{{\protect #1}}

\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{remark}{Remark}

\begin{document}
\allowdisplaybreaks

\maketitle \thispagestyle{empty} \pagestyle{empty}

\begin{abstract}
This paper deals with an estimation problem where a known number of targets in close proximity are observed but the measurement originations are uncertain. It is well known that joint probabilistic data association filters (JPDAF) are effective in handing uncertain measurement originations with clutters, but they are prone to estimation coalescence, particularly when close neighboring targets share measurements. This paper proposes a Coalescence Avoiding Optimal JPDAF (C-JPDAF) that minimizes the weighted sum of the posterior uncertainty and a measure of similarity between estimated probability densities. The proposed approach has simpler structures than other coalescence avoiding approaches based on pruning, while exhibiting excellent filtering performance for targets in close proximity with crossing tracks. These are illustrated by a numerical example of two satellites on crossing orbits around the Earth.
%The algorithm is derived and explained, involving a higher computational cost that the JPDAF, but not high enough to prevent the possibility real-time implementation. A numerical example demonstrates how the C-JPDAF is able to maintain the correct trajectories of slowly crossing estimates whereas the JPDAF is not. The results show that the Kalman gain is an effective initial guess for the C-JPDAF, and we discuss the circumstances that warrant the algorithm's use.
\end{abstract}

%_______________________________________________%

\section{Introduction}

Data association is a heavily studied field with many applications in estimation when measurements are not necessarily originating from a single target.
We consider applications wherein multiple targets of interest enter the field of view of a sensor; however, the measurement origin is unknown.
To properly track these targets, we require estimation techniques to associate targets with measurements.
When the targets are in very close proximity, the association uncertainty grows, and an effective algorithm is necessary to maintain close estimations of the targets.

A variety of data association techniques are considered for the close-proximity data association with varying performance.
Some approaches such as multiple hypothesis tracking (MHT) and probability hypothesis density (PHD) techniques involve a high computational load, especially when targets are in close proximity, making real-time implementation more difficult~\cite{MHT1},~\cite{PHD1}, and~\cite{PHD2}.
Many other techniques may involve the Kalman filter (KF) or extended KF (EKF), which are well studied and provide a framework for recursive data association techniques that are much easier to implement in real-time.
A common heuristic approach is known as the nearest neighbor filter (NNF), which is a simple algorithm that only updates estimates for measurements with the closest square Mahalnobis distance, updating other targets only with their predictions~\cite{NN2}.
This is considered a \emph{hard decision} in~\cite{JPDAF1}, or that every measurement update receives full weighting; if the decision is incorrect, these measurement updates can result in detrimental effects to the associations of the estimates, so this technique is avoided in the analysis of this paper.
Alternatively, some approaches calculate the minimum mean square error (MMSE) of the estimates and their associated uncertainty.
The probabilistic data association (PDA) technique is effective in tracking a single target because it updates with a \emph{soft decision}; in particular, this approach updates provides a weighting from a likelihood of measurement association.

The joint PDA (JPDA) is an extension of the PDA to incorporate multiple targets~\cite{JPDAF1},~\cite{TrackDataAssoc}.
The JPDA filter (JPDAF) is particularly advantageous in multi-target scenarios because when targets share measurements, the soft decisions prevent loss of information, which occurs with hard decisions.
However, when neighboring targets share measurements, the estimates are subject to coalescence.

Most techniques to avoid track coalescence involve pruning the tracks, or designing a set of rules that limit the possible outcomes to likely scenarios. These rules typically involve the exact nearest neighbor probabilistic data association (ENNPDA), which applies a strong decision to the most likely associated estimate. Pruning algorithms incorporate the ENNPDA algorithm because  estimates do not share measurements, making it insensitive to track coalescence~\cite{Coal1}. For example in~\cite{Fitzgerald}, the algorithm involves a set of rules to prune the tracks that share measurements. The algorithm considers the possible tracks and applies a logic that assigns estimates to likely scenarios; a set of possible tracks are considered and lower likelihood track scenarios are deleted. Updates to tracks considered likely are given a strong decision. This technique is extended in~\cite{Coal_d},~\cite{Coal_e}, and~\cite{Coal_c} because of its successful resistance to coalescence. However, track pruning assumes that low likelihood tracks are incorrect, which is not necessarily true. Removing the weighting scheme of the JPDA causes the estimates to lose information, which this paper seeks to avoid.

We propose a novel approach to removing coalescence by minimizing a similarity index of the estimates.
Instead of choosing a different means of weighting the measurement updates, we design an optimal recursive estimation algorithm that chooses gains to compensate for the effects of coalescence.
The algorithm derived in this paper can be applied in real-time without the heuristic nature, loss of information, or the risk of hard decisions from the current coalescence avoidance approaches.

The paper is organized as follows: the problem is formally presented in Section \ref{ProbDef}.
The derivation and explanation of a modified JPDAF under the assumptions of this paper is described in Section \ref{JPDAF}.
The optimal gain that minimizes a weighted cost of uncertainty and similarity is derived in Section \ref{C-JPDAF}.
We consider a satellite tracking example of this technique in Section \ref{NumRes}.
Finally, we discuss results and applications for the C-JPDAF in Section \ref{ConclusionFutureWork}.

% Dr. Lee Edits:
\section{Data Association Filtering Problem Formulation}
\label{ProbDef}

This section describes the dynamics of targets and the measurement model considered in this paper, and a filtering problem is formulated.

% Equations of motion
\subsection{Dynamics of Targets and Measurements}

The dynamics of each target is modeled as a linear stochastic discrete-time system. The equations of motion for the $i$-th system are described by
\begin{align}
x_{i_{k+1}} & = A_{i_k} x_{i_k} + w_{i_k},\label{eqn:xkp}\\
z_{i_k} & = H_{i_k} x_{i_k} + v_{i_k},
\end{align}
where $x_{i_k}\in\Re^n$ is the state vector, $z_{i_k}\in\Re^m$ is the output vector, and  the matrices $A_{i_k}\in\Re^{n\times n}$ and $H_{i_k}\in\Re^{m\times n}$ describe system dynamics and outputs. The process noise and the measurement noise are denoted by $w_{i_k}\in\Re^n$ and $v_{i_k}\in\Re^m$, respectively, and it is assumed that they are zero-mean, mutually independent white, Gaussian noise sequences with known covariance matrices $Q_{i_k}\in\Re^{n\times n}$ and $R_{i_k}\in\Re^{m\times m}$, respectively, i.e.,
\begin{align}
w_{i_k} \sim \mathcal{N}[0,Q_{i_k}],\quad
v_{i_k} \sim \mathcal{N}[0,R_{i_k}].
\end{align}

The initial states are also assumed to be mutually independent and independent to noise. They are Gaussian with known mean values and covariance matrices:
\begin{align}
x_{i_0} \sim \mathcal{N}[\bar x_{i_0}, P_{i_0}],\label{eqn:xi0}
\end{align}
where $\bar x_{i_0}\in\Re^n$ and $P_{i_0}\in\Re^{n\times n}$.

To demonstrate the main idea of the proposed data association filter concisely, it is assumed that there are two targets, i.e., $i\in\{1,2\}$. But, the results of this paper are readily generalized to arbitrary number of systems provided the number of targets is known.

%%%%

\subsection{Measurement Origination Model}
\label{DAP}

%Consider the data association problem when two estimates are in very close proximity.
%For this scenario, several assumptions are made.

The following assumptions are made on the measurement origination, considering the case where two target are in close proximity. First, it is assumed that there is only a single measurement at each time in the region of close proximity. Secondly, we assume that this measurement originates from either one of the targets and it is not from other sources such as clutter. Also, the case when neither target is measured is not considered. Like the JPDA, all targets are assumed to fall inside the validation gate. The measurement origination between targets is random, unknown, and equally likely.

More explicitly, let $\Theta_i$ be the event that the measurement originated from the $i$-th target, for $i\in\{1,2\}$. According to the above assumptions, we have
\begin{align}
P(\Theta_1\cup \Theta_2)&=1,\quad
P(\Theta_1\cap \Theta_2)=0,\\
P(\Theta_1)&=P(\Theta_2)=0.5.\label{eqn:MeaOrig}
\end{align}

%\EditTL{
%The only source of knowledge to associate data at time $k$ is the measurement $z_k$.
%Using the estimate prediction $\hat x^-$, we define its prediction
%\begin{align}
%z_{i_{k|k-1}}=H_{i_k}\hat x_{i_k}^-
%\end{align}
%which is used to determine the innovation
%\begin{align}
%e_{i_k}=z_k-z_{i_{k|k-1}}
%\end{align}
%and its covariance
%\begin{align}
%S_{i_k}=H_{i_k}P_{i_k}^{-}H_{i_k}^T+R_{i_k}
%\end{align}
%which is used to determine the weighting of the measurement update.
%
%We can modify JPDAF weighting and update from~\cite{JPDAF1} and~\cite{TrackDataAssoc}.
%The assumptions in this section are applied and significantly simplify the weighting calculation. For the $i$-th target, the weighting $\beta_i$ is reduced to (see proof in the appendix)
%\begin{gather}
%\beta_{i_k}=\frac{f_{i_k}}{\displaystyle\sum\limits_{y=1}^2 f_{y_k}} \\
%f_{y_k}=\mathcal{N}[z_{k};\hat z_{i_{k|k-1}},S_{i_k}].
%\end{gather}
%where $\mathcal{N}[x;\mu,\Sigma]$ is the Gaussian probability density function of mean $\mu$ and covariance matrix $\Sigma$ evaluated at $x$.}

\subsection{Problem Formulation}

A filtering problem is formulated as recursively updating the estimated state of each target and its covariance, namely $(\hat x_{i_k}, P_{i_k})$ based on the above assumptions given by \refeqn{xkp}--\refeqn{MeaOrig}. In particular, we wish to avoid track coalescence when two targets are in close proximity.

%The filtering problem involves updating each state using this weighting. We wish to update the estimates of both systems on parallel filters; however, when these weightings are close, multiple estimates share a single measurement, causing the estimates to coalesce, which we wish to avoid.

% Association Problem

	% Assumptions on measurement
		% only one measurement is available at each time 
		% the case of no measurement is not included
		% measurement origination is random and unknown
		% these assumptions represent two targets in close proximity, such as crossing targets

	% Data association event
		% there are only two events for data association
		% data association probability (beta) is defined as...
		% (Proper justification is needed, as this is one of the key assumptions)

	% Filtering problem
		% We wish to estimate states of both system

%_______________________________________________%

\section{Joint Probabilistic Data Association Filter}
\label{JPDAF}

Joint probabilistic data association filter is a suboptimal Bayesian filter for multiple targets where the probabilities of association is computed for the latest set of measurements only~\cite{TrackDataAssoc}. This section summarizes the approaches of JPDAF.

\subsection{Flow Update}
Similar to Kalman filters, the JPDAF is composed of a flow update and a measurement update. During the flow update (or prediction step), the estimated state $\hat x_{i_k}$ and its covariance $P_{i_k}$ of each target are updated as in the standard filter as follows:
\begin{gather}
\hat x_{i_{k}} = A_{i_{k-1}} \hat x_{i_{k-1}},\\
P_{i_{k}} = A_{i_{k-1}} P_{i_{k-1}} A_{i_{k-1}}^T + Q_{i_{k-1}}.
\end{gather}
These are initialized by \refeqn{xi0}, and they are repeated between measurements.

\subsection{Measurement Update}

Next, we consider the measurement update step. Let  the priori value of a variable and its posteriori value be denoted by the superscript $-$ and $+$, respectively. For a given measurement $z_{k}$, its measurement residual or innovation for the $i$-th target is given by
\begin{align*}
e_{i_k} = z_k - z_{i_{k|k-1}},
\end{align*}
where $z_{i_{k|k-1}}$ is the predicted measurement given by
\begin{align}
z_{i_{k|k-1}} = H_{i_k} \hat x_{i_k}^-.
\end{align}
The innovation covariance $S_{i_k}$ is
\begin{align}
S_{i_k}=H_{i_k}P_{i_k}^{-}H_{i_k}^T+R_{i_k}.
\end{align}

The data association probability is defined as follows. Let $\beta_i$ be the probability of the $i$-th data association event conditioned by the given measurement:
\begin{align*}
\beta_{i_k} = P[\Theta_i|z_k].
\end{align*}
For the given assumptions summarized in Section \ref{ProbDef}, it can be obtained as
\begin{align}
\beta_{i_k}=\frac{f_{i_k}}{\sum\limits_{y=1}^2 f_{i_k}},\label{eqn:betaik}
\end{align}
where $f_{i_k}$ is defined as
\begin{align}
f_{i_k}=\mathcal{N}[z_{k};\hat z_{i_{k|k-1}},S_{i_k}].
\end{align}
Here $\mathcal{N}[x;\mu,\Sigma]$ denotes the Gaussian probability density function of mean $\mu$ and covariance matrix $\Sigma$ evaluated at $x$ (see Appendix for the derivation of \refeqn{betaik}).

According to the total probability theorem with respect to the association events, the resulting posterior estimate of the state is given by
\begin{align}
\hat x^+_{i_k}=&\hat x^-_{i_k}+\beta_{i_k}K_{i_k}e_{i_k}\label{KalEst},
\end{align}
where the estimator gain is equal to the Kalman gain, i.e., $K_i = P_{i_k}^{-} H_{i_k}^T S_{i_k}^{-1}$, as there is no measurement origination uncertainty when conditioned on the event $\Theta_i$.

The resulting posteriori covariance is given by
\begin{align}
\label{KalCov}
%P^+_{i_k}=&P^-_{i_k}-\beta_{i_k}K_{i_k}S_{i_k}K_{i_k}^T.
P^+_{i_k}=&P^-_{i_k}-\beta_i[K_iH_iP_i^-+P_i^-H_i^TK_i^T-K_iS_{i_k}K_i^T].
\end{align}
(see Appendix for derivation of (\ref{KalCov})).

%%\EditTL{A typical JPDAF considers scans of many measurements and the possibility of clutter. Applying the assumptions from Section \ref{ProbDef} we can greatly reduce the measurement update terms to much more manipulatable forms (see appendix for proofs):
%%\begin{align}
%%\label{KalEst}
%%\hat x^+_{i_k}=&\hat x^-_{i_k}+\beta_{i_k}K_{i_k}e_{i_k}\\
%%\label{KalCov}
%%P^+_{i_k}=&P^-_{i_k}-\beta_{i_k}K_{i_k}S_{i_k}K_{i_k}^T.
%%\end{align}
%%These equations are derived under the key assumption that the states of the systems are assumed to be Gaussian according to the most recent state estimate and covariance.
%%The gain $K_{i_k}$ is equivalent to a standard Kalman gain
%%\begin{align}
%%\label{KalGain}
%%K_{i_k}=P_{i_k}^-H_{i_k}^TS_{i_k}^{-1}
%%\end{align}
%%which serves to minimize $\tr{P^+_{i_k}}$ because here there is no measurement uncertainty when conditioned on event $\Theta_{i_k}$~\cite{TrackDataAssoc}.
%%However, Equation \ref{KalCov} only holds true for $K_{i_k}$ as defined in Equation \ref{KalGain}.
%%The general form of $P_{i_k}^+$ is defined as
%%\begin{equation}
%%\label{genPp}
%%P_{i_k}^{+}=(I-K_{i_k}H_{i_k})P_{i_k}^{-}(I-K_{i_k}H_{i_k})^T+K_{i_k}R_{i_k}K_{i_k}^T
%%\end{equation}
%%which may be manipulated with Equation \ref{KalGain} to form Equation \ref{KalCov} (without the $\beta_{i_k}$ term)~\cite{OptEst1}.}

	% The key assumption of JPDA: eqn (6-16) of the book
	% JPDA: see pages 164-166
		% Derive how the mean value of the state is updated 
		% Show that the gain becomes the same as the standard filter that minimizes tr[P]
		% Derive how the covariance is updated

%_______________________________________________%

\section{Coalescence Avoiding Optimal Joint Probabilistic Data Association Filter}% Or another name that you choose
\label{C-JPDAF}

The joint probabilistic data association filter described in the previous section is effective in handling clutter or missed detection of multiple targets. However, it is prone to coalescing estimates, especially with targets in close-proximity. Here, we present a coalescence-avoiding optimal JPDAF, hereafter C-JPDAF for short, to avoid data coalescence. The main idea is choosing the estimator gain such that a weighted sum of the size of uncertainty and a measure of similarity is minimized.

A similarity measure is introduced first, and an optimization problem is formulated. Throughout this section, the subscript $k$ is omitted for simplicity.

%The goal of the C-JPDAF is to minimize a weighted sum of the uncertainty and the similarity.
%The filter is designed to slightly change each gain $K_{i_k}$ from that of the JPDAF when the estimates are in very close proximity to resist coalescence.

%This section describes the changes necessary to form the C-JPDAF from the JPDAF. We consider the changes to the measurement update at time $k$, so these subscripts will not be included in this section.

% Derivation of a new filter

	% Short description of the main idea: minimize the weighted sum of the uncertainty and the similarity

\subsection{Matusita's Similarity Measure}

For two probability density functions, $p_1(x)$ and $p_2(x)$, Matusita's measure is defined as
\begin{align*}
\zeta = \int \sqrt{f_1(x)f_2(x) dx},
\end{align*}
which is between 0 and 1, and it measures the similarity between two densities~\cite{Coal_k}.

Using the fact that the posteriori states $\hat x_1^+$ and $\hat x_2^+$ are Gaussian, Matusita's measure between two targets can be written as
\begin{align}
\zeta=\zeta_1\zeta_2,
\end{align}
where
\begin{align}
\label{Q}
\zeta_1=&\frac{2^\frac{n}{2}|P_1^+|^{\frac{1}{4}}|P_2^+|^{\frac{1}{4}}}{|P_1^++P_2^+|^{\frac{1}{2}}},\\
\label{R}
\zeta_2=&\exp{\left\{-\frac{1}{4}(\hat x^+_1-\hat x^+_2)^T(P_1^++P_2^+)^{-1}(\hat x^+_1-\hat x^+_2)\right\}}.
\end{align}
Here $\zeta_1$ measures the differences in the covariance matrices and $\zeta_2$ represents the spread in the means.
	% Introduce the definition of M-measure
	% Introduce the definition of Bhattacharyya distance

\subsection{Optimal Coalescence Avoidance}

Define a cost function of coalescence avoiding filtering as
%The choice of optimal cost involves two parts: posterior uncertainty minimization, much like the Kalman filter, and similarity minimization, based on the Matusita's measure. The total cost function is defined as
\begin{align}
J=J_P+cJ_S
\end{align}
where $J_P$ is an uncertainty cost that measures the confidence level of estimation and $J_S$ is a similarity cost that measure degree of coalescence. The constant $c$ is a positive weighting factor between $J_P$ and $J_S$.

In the proposed C-JPDAF, the posterior estimate of the state is given in the form of JPDA given by (\ref{KalEst}), but the estimator gain $K_i$ is selected to minimize the total cost $J$. In short, the estimator gain is chosen in the consideration of both uncertainty and coalescence. These are illustrated at Figure \ref{CostTrends}.

%The variation of these costs with respect to a controller gain is illustrated at Figure \ref{CostTrends}.

%, and $c$ is a positive constant for relative weighting.

%The cost function is shown by varying the first element of $K_1$ in Figure \ref{CostTrends}, although $c$ is chosen much larger for this figure than in practice to depict the effect of including a similarity cost rather than just uncertainty cost.

\begin{figure}
\setlength{\unitlength}{0.1\columnwidth}
\centerline{\footnotesize\selectfont
\begin{picture}(8.5,7.3)(-0.5,-0.8)
\put(-0.5,3){\rotatebox{90}{Cost}}
\put(0,0){\includegraphics[width=0.8\columnwidth]{ECC14_fig1.pdf}}
\put(6.2,0.2){\shortstack[c]{Uncertainty\\ cost: $J_P$}}
\put(6.2,2.8){\shortstack[c]{Similarity\\ cost: $J_S$}}
\put(5.7,5.2){\shortstack[c]{Weighted sum:\\ $J=J_P+cJ_S$}}
\put(4.98,0.78){\circle*{0.1}}
\put(2.48,1.57){\circle*{0.1}}
\put(5.0,-0.3){$K_{JPDA}$}
\put(2.5,-0.3){$K$}
\put(3.1,-0.8){Estimator gain}
\end{picture}
}
\caption{Illustration of optimal coalescence avoidance: several costs are illustrated with respect to the estimator gain $K$ (red: uncertainty cost $J_P$, blue: similarity cost $J_S$, black: weighted sum). The estimator gain $K_{JPDA}$ of JPDA is selected to minimize the uncertainty cost. In the proposed approach, the estimator gain $K$ is chosen such that the weighted sum of the uncertainty cost and the similarity cost is minimized. In short, at the expense of increased uncertainty at a single step, the proposed approach prevents target coalescence. The increased uncertainty at each step is rewarded by reducing the overall estimation error drastically by maintaining correct data association throughout crossing tracks in close proximity.}
%The cost functions of target 1 are examined under the variation of a single element of $K_1$. The resulting cost function (black) is minimized slightly to the right of the uncertainty cost function (blue). Note that the total cost when the Kalman gain is applied does not lie on the total cost curve because other elements of this gain that are not optimal. In Kalman filtering applications, including the JPDAF, only the uncertainty cost function is minimized.}
\label{CostTrends}
\end{figure}

More explicitly, the uncertainty cost $J_P$ is defined as the sum of the trace of each posteriori covariance:
\begin{align}
J_P=\tr{P^+_1}+\tr{P^+_2},\label{eqn:JP}
\end{align}
where $P^+_{i_k}$ is obtained by (\ref{KalCov}). The similarity cost is chosen as the second part of the Matusita's measure that represent the weighted-spread of the mean values:
\begin{align}
J_S&=\exp (-au),\\
u & = (\hat x^+_1-\hat x^+_2)^T(P^+_{1}+P^+_{2})^{-1}(\hat x^+_1-\hat x^+_2).\label{eqn:u}
\end{align}
for positive constant $a$, which serves to amplify or attenuate the function as well as alter its shape.

Necessary conditions for optimality are given by
\begin{align}
\deriv{J}{K_i} = \deriv{J_P}{K_i} + c \deriv{J_S}{K_i} =0,\label{eqn:NCO}
\end{align}
for $i=\{1,2\}$. These are solved for $K_1$ and $K_2$ to obtain optimal estimator gains.

%%\EditTL{Note that $P^+_i$ must be valid for all $K_i$. Hence, from the general form of Equation \ref{genPp}, we seclude a $P_i^-$ term to take the form of Equation \ref{KalCov} to obtain
%%\begin{align}
%%\label{newPp}
%%P^+_i=&P^-_i-\beta_i[K_iH_iP_i^-+P_i^-H_i^TK_i^T\\
%%&-K_i(H_iP_i^-H_i^T+R_i)K_i^T].
%%\end{align}}

%For the remaining equations of this section, the only derivatives taken are with respect to $K_1$ for the similarity between targets $1$ and $2$ for simplicity; taking derivatives with respect other $K_i$ values are identical except for subscripts.

Next, we derive the derivative of each cost. For simplicity, derivatives with respect to the first gain $K_1$ are presented, and derivatives with respect to $K_2$ can be obtain by symmetry. By substituting (\ref{KalCov}) into \refeqn{JP} and using the property $\frac{\partial}{\partial X}\tr{XYX^T}=2XY$ for any matrix $X$ and any symmetric matrix $Y$, we obtain
\begin{align}
\label{CostP}
\frac{\partial J_{P,1}}{\partial K_{1}}&=2\beta_1({K_1S_1-P_1^-H_1^T}).
\end{align}
%
%To appropriately define the similarity cost, we only consider the $R$ term from Equation \ref{R} because coalescence is well known to occur when estimates are within close proximity. For simplicity, we define ${P^+_{12}}^{-1}$ as
%\begin{align}
%{P^+_{12}}^{-1}=(P_1^++P_2^+)^{-1}
%\end{align}
%and define define $u$ as
Substituting (\ref{KalEst}) and (\ref{KalCov}) into \refeqn{u}, and rearranging, 
\begin{align}
u%=&(\hat x^+_1-\hat x^+_2)^T{P^+_{12}}^{-1}(\hat x^+_1-\hat x^+_2)\\
=&\{\hat x_1^{-T}{P^+_{12}}^{-1}\hat x_1^-
-2\hat x_1^{-T}{P^+_{12}}^{-1}\hat x_2^-\nonumber\\
&+2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T{P^+_{12}}^{-1}K_1e_1
+\hat x_2^{-T}{P^+_{12}}^{-1}\hat x_2^-\nonumber\\
&+2\beta_2(\hat x_2^{-}-\hat x_1^{-})^T{P^+_{12}}^{-1}K_2e_2
+\beta_1^2e_1^TK_1^T{P^+_{12}}^{-1}K_1e_1\nonumber\\
&-2\beta_1\beta_2e_2^TK_2^T{P^+_{12}}^{-1}K_1e_1+\beta_2^2e_2^TK_2^T{P^+_{12}}^{-1}K_2e_2\},
\end{align}
where ${P^+_{12}}^{-1}=(P_1^++P_2^+)^{-1}$.

%which is the most reduced form to differentiate with respect to $K_1$ and $K_2$.

\newcommand{\Iij}{\ensuremath{\mathbf{1}_{ij}}}

Let $\Iij\in \Re^{n\times m}$ be a $n\times m$ matrix defined such that its $i,j$-th element is one and other elements are zero. The derivative of $(P_{12}^+)^{-1}$ with respect to the $i,j$-th element of $K_1$ is defined as $\mathcal{P}_{ij}\in\Re$:
%such that $ij1(i,j)=1$ and $0$ elsewhere, we define $P^*_{12}$ as
\begin{align}
\mathcal{P}_{ij}&=\frac{\partial {P^+_{12}}^{-1}}{\partial K_{1,ij}}\nonumber\\
&=\beta_1{P_1^+}^{-1}[\Iij H_1P_1^-+P_1^-H_1^T\Iij^T\nonumber\\
&\quad -\Iij S_1K_1^T -K_1S_1\Iij^T]{P_1^+}^{-1}.
\end{align}
Using this, the derivative of $u$ with respect to the $i,j$-th element of $K_1$ can be written as
\begin{align}
\frac{\partial u}{\partial K_{1,ij}}%=&\frac{\partial}{\partial K_{1,ij}}\left\{\hat x^+_1-\hat x^+_2)^T{P^+_{12}}^{-1}(\hat x^+_1-\hat x^+_2\right\}\\
=& \{\hat x_1^{-T}\mathcal{P}_{ij}\hat x_1^-
-2\hat x_1^{-T}\mathcal{P}_{ij}\hat x_2^-\nonumber\\
&+2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T[\mathcal{P}_{ij}K_1+\mathcal{P}_{ij}^{-1}\Iij]e_1\nonumber\\
&+\hat x_2^{-T}\mathcal{P}_{ij}\hat x_2^-+2\beta_2(\hat x_2^{-}-\hat x_1^{-})^T\mathcal{P}_{ij}K_2e_2\nonumber\\
&+\beta_1^2e_1^T[\Iij^T{\mathcal{P}_{ij}}^{-1}K_1+K_1^T\mathcal{P}_{ij}K_1\nonumber\\
&+K_1^T{\mathcal{P}_{ij}}^{-1}\Iij]e_1-2\beta_1\beta_2e_2^TK_2^T[\mathcal{P}_{ij}K_1\nonumber\\
&+{\mathcal{P}_{ij}}^{-1}\Iij]e_1+\beta_2^2e_2^TK_2^T\mathcal{P}_{ij}K_2e_2\}.
\end{align}
From this, the derivation of the similarity cost $J_S$ with respect to the $i,j$-th element of $K_1$ is given by
\begin{align}
\frac{\partial J_{S}}{\partial K_{1,ij}}=-a\exp(-au)\frac{\partial u}{\partial K_{1,ij}}.\label{eqn:JSK}
\end{align}

In summary, (\ref{CostP}) and \refeqn{JSK} are substituted into \refeqn{NCO}, and it is solved to obtain the optimal gain $K_1$ and $K_2$. Then, the algorithm of C-JPDAF is completed by updating the state estimate and covariance by using (\ref{KalEst}) and (\ref{KalCov}), respectively.

%: equation \ref{KalEst} 
%is used for $\hat x^+_i$ because this equation does not require a Kalman gain; however we must use Equation \ref{newPp} to obtain $P^+_i$ because Equation \ref{KalCov} is only valid for a Kalman gain.

%
%
%for $i=1:n$ and $j=1:m$ to obtain $\frac{\partial u}{\partial K_{1}}$.
%
%The weighted cost is minimized when $K_1$ and $K_2$ are chosen such that
%\begin{align}
%\frac{\partial J}{\partial K}=\frac{\partial J_{P,1}}{\partial K_{1}}+\frac{\partial J_{P,2}}{\partial K_{2}}
%+c\left(\frac{\partial J_{S,1}}{\partial K_{1}}+\frac{\partial J_{S,2}}{\partial K_{2}}\right)=0.
%\end{align}

As the optimality condition is a nonlinear function of estimator gains, there is no closed-form solution. Hence, we require a nonlinear equation solver. Note that ${P_1^+}^{-1}$, ${P_2^+}^{-1}$, and $(P^+_1+P^+_2)^{-1}$ are symmetric inverses and must only be calculated once each iteration of the solver. When $c=0$, we can show that the optimal gain reduces to the Kalman gain, which is used as an initial guess of numerical solutions.

%Furthermore, this weighted cost may be readily extended to an arbitrary known number of targets provided the Matusita's measure is included between each estimate.

%The algorithm is completed by updating the state estimate and covariance. Equation \ref{KalEst} may be used for $\hat x^+_i$ because this equation does not require a Kalman gain; however we must use Equation \ref{newPp} to obtain $P^+_i$ because Equation \ref{KalCov} is only valid for a Kalman gain.

	% Choice of optimal cost
		% the part of B-distance representing the difference of mean is used
		% the role of the weighting factor of c

	% Necessary conditions for optimality
		% Derivation of \deriv{J}{K}

	% Filter equations
		% Show why we can use the same expression for the mean value and the covariance of JPDA described in the previous section

%	\subsection{Suboptimal Coalescence Avoidance}

	% Linear approximation of the gain to obtain explicit solution

\section{Numerical Examples}
\label{NumRes}
	% Describe two body problem of orbital mechanics
	% Describe the measured values and the initial conditions

	% Results of the perfect association with 100% confidence (imaginary case for comparison)
	% Results of JPDA (incorrect association after crossing)
	% Results of the proposed optimal filtering approach
	% Results of the proposed suboptimal filtering approach with linear approximation

	% Discussion that the proposed approach is between JPDA and perfect association in terms of filtering performance
	% Show how much additional computational load is required compared with JPDA

%We present a numerical example of  where some parameters are varied to evaluate the performance of the C-JPDAF with the JPDAF.

We consider estimation of two satellites in a low-Earth orbit with crossing orbital trajectories. It is assumed that the range toward satellites are measured with unknown origination.

The dynamics of each satellite is modeled as a two-body problem between each satellite and Earth, i.e., the of equations of motion for each orbit only involve force caused by the gravity between the Earth and the satellite and neglect all other forces. For simplicity, both orbits occupy the same plane, so cylindrical coordinates are used to define our state variable $x=\begin{bmatrix} r & \dot r & \theta & \dot \theta\end{bmatrix}^{T}$. The resulting equations of motion of each satellite are given by
\begin{gather*}
\ddot r - r\dot\theta^2 = -\frac{\mu}{r^2},\\
r\ddot\theta + 2\dot r\dot\theta =0,
\end{gather*}
where $\mu=3.986\times10^5\, \mathrm{km^3/s^2}$ is the gravitational parameter of the Earth. These are linearized to obtain the system matrix of the discrete-time linear equations of motion considered in this paper:
\begin{align}
A_{i_k}=
\exp\left(\begin{bmatrix}
0 & 1 & 0 & 0 \\
\frac{2\mu}{r_{i_k}^3} +\dot\theta^2_{i_k} & 0 & 0 & 2r_{i_k}\dot \theta_{i_k} \\
0 & 0 & 0 & 1 \\
2r_{i_k}^{-2}\dot r_{i_k}\dot\theta_{i_k} & -2r^{-1}\dot\theta_{i_k} & 0 &  -2r_{i_k}^{-1}\dot r_{i_k}\\
\end{bmatrix}\Delta t\right).
\end{align}
%where  and $H=\begin{bmatrix} 1 & 0 & 0 & 0\end{bmatrix}$.

The initial estimate of the states are designed to occur shortly before the two trajectories cross paths near $r_{ref}=7000km$ from the center of the earth; for the time to complete a circular orbit at this radius, $t_{ref}=2\pi \sqrt{\frac{r_{ref}^3}{\mu}}$, we can solve for a reference circular orbit $X_{ref}=\begin{bmatrix} r_{ref} & 0 & 0 & \frac{2\pi}{t_{ref}}\end{bmatrix}^T$, then perturb this orbit slightly by $X_{pert}=\begin{bmatrix} 0.40 & 0 & 0 & -1.5\times10^{-5}\end{bmatrix}^T$. We can easily choose the initial conditions to be $X_{1_0}= X_{ref}+X_{pert}$ and $X_{2_0}= X_{ref}-X_{pert}$. The initial estimates are assumed to be relatively accurate: there is no need to simulate large uncertainty when the paths are initially sufficiently far apart because this paper is concerned with the coalescence when the targets are close together, so the state and covariance will have been determined by simulated measurements while the estimates were still spread significantly apart; hence for both estimates, the initial covariance was chosen to be $P_0=diag(\begin{bmatrix}
2.5\times10^{-4} & 2.5\times10^{-6} & 3.9\times10^{-8} & 3.9\times10^{-8} \end{bmatrix})$.
The process noise was chosen to be equivalent for both estimate such that $Q=diag(\begin{bmatrix}
2.5\times10^{-5} & 2.5\times10^{-7} & 3.9\times10^{-9}& 3.9\times10^{-9}\end{bmatrix})$ and $R=\begin{bmatrix}
2.25\times10^{-2}
\end{bmatrix}$ applied at each time step.

Three filtering approaches are applied, namely extended Kalman filter (EKF), JPDAF, and C-JPDAF, such that a single measurement was evaluated at each time step, and its origination was random and equally likely from both targets. For EKF, it is assumed that the correct measurement origination is know with certainty, i.e., all of the measurements are associated to correct target. This serves as an ideal reference case to be compared with JPDAF and C-JPDAF. For the C-JPDAF, the weighting parameters of the cost are chosen as $a=0.02$ and $c=5\times10^{-5}$. We propagate the nonlinear equations of motion numerically using Matlab, and we generate measurement at every $0.1$ seconds with an equal likelihood that its origination is from either target.

\begin{figure}
\centerline{
	\subfigure[EKF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{EKF.png}\hspace*{0.07\textwidth}}
	}
\centerline{
	\subfigure[JPDAF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{JPDAF.png}\hspace*{0.07\textwidth}}
	}
\centerline{
	\subfigure[C-JPDAF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{CAOJPDAF.png}\hspace*{0.05\textwidth}}
	}
\caption{Target 1 is represented in blue and target 2 is represented in red. A solid line represents an estimate and a dotted line represents the truth.The JPDAF is subject to coalescence of the neighboring track, causing the estimations to switch trajectories. The C-JPDAF is subject to much less coalescence and hence does not exhibit this problem.}
\label{AllAlgorithms}
\end{figure}

The corresponding results for each of  EKF, JPDAF, and C-JPDAF algorithms are illustrated in Figure \ref{AllAlgorithms}. 
First, EKF exhibits good estimation performance as there is no error or uncertainty in data association. The performance of JPDA is similar to EKF when two satellites are far apart from each other. However, when $50\leq t\leq 70$ seconds, two satellites become closer, and the estimated states are coalesced. When $t\geq 70$ seconds, the estimated states are incorrectly crossed. This illustrates undesirable performance of JPDAF especially for crossing targets.

The proposed C-JPDAF does not exhibit the incorrect, crossed estimation of JPDA, and the estimated states follows each of the correct trajectories throughout the crossing orbital trajectories. The resulting RMS errors for the orbital radius of each method are summarized at Table \ref{tab:RMS}. This illustrates that the performance of the proposed C-JPDAF is comparable to EKF where there is no error in data association.

\begin{table}
\caption{Estimation results}\label{tab:RMS}
\centerline{
\begin{tabular}{ccc}\hline
Method & Radial RMS Error (km) & Computation time (sec) \\\hline
EKF & 1.438e-05 &6.118\\
JPDAF & 7.1242e-05 &6.130\\
C-JPDAF & 2.3145e-05  & 19.864 \\\hline
\end{tabular}}
\end{table}

The major drawback of the C-JPDAF is the added computation time. However, the results of all three simulated algorithms (EKF, JPDAF, and C-JPDAF) are nearly identical when the estimates are not near the crossing region. Hence, the C-JPDAF can be turned on only during these close proximity regions, then turned off when unnecessary to improve the computational efficiency.

%First, JPDAF is unable to maintain the correct trajectories of the estimates because the tracks coalesce so much that %correctly labeling the estimates becomes very difficult; when measurements from either target are sufficiently spread after the crossing, the JPDAF decision of trajectories that the estimates follow becomes ambiguous as the estimates become ambiguous.

%Secondly, within $20 sec$ of crossing ($201$ measurements), the number of incorrect associated (weighted less than $50\%$ likelihood of association) for the JPDAF was $92$ whereas for the C-JPDAF this number was $70$. Comparing the performance of all three algorithms, the radial RMS error in the same window was calculated for the EKF as $1.4425\times10^{-05} km$, for the JPDAF $2.3014\times10^{-05} km$, and for the C-JPDAF $2.3858\times10^{-05} km$. This metric demonstrates the uncertainty of association degrades the estimates as the EKF performs significantly better than algorithms where measurement origination is unknown. The JPDAF is performs slightly better than the C-JPDAF with this metric, but this is a small drawback when compared with the ability to maintain correct tracks after crossing. The largest drawback of the C-JPDAF is the computational cost; running the simulations on the same computer, the JPDAF only required $7.518sec$ to run the program whereas the C-JPDAF required $25.785sec$ to run.

%_______________________________________________%

\section{Conclusions and Future Work}
\label{ConclusionFutureWork}
We explore a new approach to reduce coalescence of multiple target tracking. The estimator gains are chosen such that a weighted cost with respect to posterior uncertainty and similarity is minimized. The proximity similarity portion of the Matusita's measure is chosen to evaluate the similarity. Necessary conditions for optimality are derived, and they are solved numerically to obtain optimal estimator gains. Numerical examples illustrate that it exhibits excellent estimation and data association properties for crossing tracks in close proximity. Compared with other techniques to avoid coalescence, the proposed approach has simpler structures. Future works include generalizing C-JPDAF for arbitrary number of targets with clutters, and improving the computational efficiencies.

%We take the derivative of the similarity cost for both estimates, yielding a result with no closed form solution, but involving expressions with low computational cost that can be solved with a nonlinear equation solver. In practice, a small similarity cost relative to the uncertainty cost yields best results; hence, the algorithm requires less time if the initial guesses of the Kalman gains are chosen.

%We validate the theory with numerical simulations. The C-JPDAF shares measurements in the same way as the JPDAF, but the tracks tend to maintain separation; the C-JPDAF does not prevent the estimates from crossing as the targets in the trajectory are designed to do. The major drawback of the C-JPDAF is the added computation. However, the results of all three simulated algorithms (EKF, JPDAF, and C-JPDAF) are nearly identical when the estimates are not near the crossing region (even if on the wrong trajectory). Hence, the C-JPDAF can be turned on only during these close proximity regions, then turned off when unnecessary to be replaced by a less computationally expensive algorithm.

%The C-JPDAF can be extended to a variety of multi-target tracking applications when targets share measurements due to close proximity. One example, very close to that presented in Section \ref{NumRes}, is keeping logs of satellites using phased-array radar systems. In cases where satellites have similar measurement predictions over long periods of time, satellite tracking becomes much more difficult. Implementing a simple JPDAF can serve to coalesce the estimates, which can cause the estimates of the satellites to switch trajectories as shown in this paper. Therefore, including the C-JPDAF under certain proximity conditions can expand our ability to track satellites. The C-JPDAF has tracking applications for space, aerial, and underwater vehicles in close proximity and identification of individual agents in multi-agent systems.

%Much like the Kalman filter, the C-JPDAF can be effective for systems with well-known Gaussian distributions of state, sensor noise, and process. This paper has assumed that measurements come from a single target of interest at a single time and not clutter. Future work may involve expanding the capabilities of the algorithm. We can consider the measurements as a scan, or so that multiple measurements, including those from external sources of clutter, may belong to multiple targets at any time step as a JPDAF does without the assumptions of this paper. A general form of this algorithm could be derived for these conditions. Furthermore we can explore other techniques to measure similarity. By only using the proximity portion Matusita's measure, we are assuming the covariances of the estimates are close, which is not always accurate in all applications, and another future direction of this work.

%%_______________________________________________%
%
%\section{Data Association Problem Formulation}
%\label{ProbDef}
%
%% Equations of motion
%\subsection{Equations of Motion}
%
%Consider a set of linear stochastic discrete-time systems. The equations of motion for the $i$-th system are described by
%\begin{align}
%x_{i_{k+1}} & = A_{i_k} x_{i_k} + w_{i_k},\\
%z_{i_k} & = H_{i_k} x_{i_k} + v_{i_k},
%\end{align}
%where $x_{i_k}\in\Re^n$ is the state vector, $z_{i_k}\in\Re^m$ is the output vector, and  the matrices $A_{i_k}\in\Re^{n\times n}$ and $H_{i_k}\in\Re^{m\times n}$ describe system dynamics and outputs. The process noise and the measurement noise are denoted by $w_{i_k}\in\Re^n$ and $v_{i_k}\in\Re^m$, respectively, and it is assumed that they are zero-mean, mutually independent white, Gaussian noise sequences with known covariance matrices $Q_{i_k}\in\Re^{n\times n}$ and $R_{i_k}\in\Re^{m\times m}$, respectively:
%\begin{align}
%w_{i_k} \sim \mathcal{N}(0,Q_{i_k}),\quad
%v_{i_k} \sim \mathcal{N}(0,R_{i_k}).
%\end{align}
%
%
%The initial states are also assumed to be mutually independent and independent to noise. They are Gaussian with known mean values and covariance matrices:
%\begin{align}
%x_{i_k} \sim \mathcal{N}(\bar x_{i_k}, P_{i_k}),
%\end{align}
%where $\bar x_{i_k}\in\Re^n$ and $P_{i_k}\in\Re^{n\times n}$. 
%
%To demonstrate the main idea of the proposed data association filter concisely, it is assumed that there are two dynamic systems, i.e., $i\in\{1,2\}$. But, the results of this paper are readily generalized to arbitrary number of systems provided the number of systems is known.
%
%%%%%
%
%\subsection{Data Association Problem}
%\label{DAP}
%
%Consider the data association problem when two estimates are in very close proximity.
%For this scenario, several assumptions are made.
%First, we assume that we are only observing a single measurement from this close proximity region, i.e., we consider a single measurement at a time, not a scan of multiple measurements.
%Secondly, we assume that this measurement originates from either one of the targets and not from other sources of clutter; the case when neither target is measured is not considered.
%Like the JPDA, all targets are assumed to fall inside the validation gate.
%Furthermore, the measurement origination between targets is random, unknown, and equally likely. In summary, for the event $\Theta_1$ as measuring target $1$ and the event $\Theta_2$ as measuring target $2$,
%\begin{align}
%P(\Theta_1\cup \Theta_2)&=1,\quad
%P(\Theta_1\cap \Theta_2)=0,\\
%P(\Theta_1)&=P(\Theta_2)=0.5.
%\end{align}
%
%The only source of knowledge to associate data at time $k$ is the measurement $z_k$.
%Using the estimate prediction $\hat x^-$, we define its prediction
%\begin{align}
%z_{i_{k|k-1}}=H_{i_k}\hat x_{i_k}^-
%\end{align}
%which is used to determine the innovation
%\begin{align}
%e_{i_k}=z_k-z_{i_{k|k-1}}
%\end{align}
%and its covariance
%\begin{align}
%S_{i_k}=H_{i_k}P_{i_k}^{-}H_{i_k}^T+R_{i_k}
%\end{align}
%which is used to determine the weighting of the measurement update.
%
%We can modify JPDAF weighting and update from~\cite{JPDAF1} and~\cite{TrackDataAssoc}.
%The assumptions in this section are applied and significantly simplify the weighting calculation. For the $i$-th target, the weighting $\beta_i$ is reduced to (see proof in the appendix)
%\begin{gather}
%\beta_{i_k}=\frac{f_{i_k}}{\displaystyle\sum\limits_{y=1}^2 f_{y_k}} \\
%f_{y_k}=\mathcal{N}[z_{k};\hat z_{i_{k|k-1}},S_{i_k}].
%\end{gather}
%where $\mathcal{N}[x;\mu,\Sigma]$ is the Gaussian probability density function of mean $\mu$ and covariance matrix $\Sigma$ evaluated at $x$.
%
%The filtering problem involves updating each state using this weighting. We wish to update the estimates of both systems on parallel filters; however, when these weightings are close, multiple estimates share a single measurement, causing the estimates to coalesce, which we wish to avoid.
%
%% Association Problem
%
%	% Assumptions on measurement
%		% only one measurement is available at each time 
%		% the case of no measurement is not included
%		% measurement origination is random and unknown
%		% these assumptions represent two targets in close proximity, such as crossing targets
%
%	% Data association event
%		% there are only two events for data association
%		% data association probability (beta) is defined as...
%		% (Proper justification is needed, as this is one of the key assumptions)
%		
%	% Filtering problem
%		% We wish to estimate states of both system
%
%%_______________________________________________%
%
%\section{Joint Probabilistic Data Association Filter}
%\label{JPDAF}
%
%A typical JPDAF considers scans of many measurements and the possibility of clutter. Applying the assumptions from Section \ref{ProbDef} we can greatly reduce the measurement update terms to much more manipulatable forms (see appendix for proofs):
%\begin{align}
%\label{KalEst}
%\hat x^+_{i_k}=&\hat x^-_{i_k}+\beta_{i_k}K_{i_k}e_{i_k}\\
%\label{KalCov}
%P^+_{i_k}=&P^-_{i_k}-\beta_{i_k}K_{i_k}S_{i_k}K_{i_k}^T.
%\end{align}
%These equations are derived under the key assumption that the states of the systems are assumed to be Gaussian according to the most recent state estimate and covariance.
%The gain $K_{i_k}$ is equivalent to a standard Kalman gain
%\begin{align}
%\label{KalGain}
%K_{i_k}=P_{i_k}^-H_{i_k}^TS_{i_k}^{-1}
%\end{align}
%which serves to minimize $\tr{P^+_{i_k}}$ because here there is no measurement uncertainty when conditioned on event $\Theta_{i_k}$~\cite{TrackDataAssoc}.
%However, Equation \ref{KalCov} only holds true for $K_{i_k}$ as defined in Equation \ref{KalGain}.
%The general form of $P_{i_k}^+$ is defined as
%\begin{equation}
%\label{genPp}
%P_{i_k}^{+}=(I-K_{i_k}H_{i_k})P_{i_k}^{-}(I-K_{i_k}H_{i_k})^T+K_{i_k}R_{i_k}K_{i_k}^T
%\end{equation}
%which may be manipulated with Equation \ref{KalGain} to form Equation \ref{KalCov} (without the $\beta_{i_k}$ term)~\cite{OptEst1}.
%
%	% The key assumption of JPDA: eqn (6-16) of the book
%	% JPDA: see pages 164-166
%		% Derive how the mean value of the state is updated 
%		% Show that the gain becomes the same as the standard filter that minimizes tr[P]
%		% Derive how the covariance is updated
%
%%_______________________________________________%
%
%\section{Coalescence Avoiding Optimal Joint Probabilistic Data Association Filter}% Or another name that you choose
%\label{CAOJPDAF}
%
%This section describes the changes necessary to form the CAOJPDAF from the JPDAF.
%We consider the changes to the measurement update at time $k$, so these subscripts will not be included in this section.
%The goal of the CAOJPDAF is to minimize a weighted sum of the uncertainty and the similarity.
%The filter is designed to slightly change each gain $K_{i_k}$ from that of the JPDAF when the estimates are in very close proximity to resist coalescence.
%
%% Derivation of a new filter
%
%	% Short description of the main idea: minimize the weighted sum of the uncertainty and the similarity
%	
%\subsection{Matusita's Similarity Measure}
%
%Two similarity measures are considered to determine an appropriate cost function. First, the Matusita's measure is considered because it ranges between 0 and 1. For the $m$-variate Gaussian distributions, the Matusita's measure can be decomposed into $Q$ and $R$
%\begin{align}
%\zeta=QR
%\end{align}
%such that
%\begin{align}
%\label{Q}
%Q=&\frac{2^\frac{m}{2}|P_1^+|^{\frac{1}{4}}|P_2^+|^{\frac{1}{4}}}{|P_1^++P_2^+|^{\frac{1}{2}}}\\
%\label{R}
%R=&\exp{\left\{-\frac{1}{4}(\hat x^+_1-\hat x^+_2)^T(P_1^++P_2^+)^{-1}(\hat x^+_1-\hat x^+_2)\right\}}
%\end{align}
%where $Q$ measures the differences in the covariance matrices and $R$ measures the spread in the means~\cite{Coal_k}.
%	% Introduce the definition of M-measure
%	% Introduce the definition of Bhattacharyya distance
%	
%\subsection{Optimal Coalescence Avoidance}
%
%The choice of optimal cost involves two parts: posterior uncertainty minimization, much like the Kalman filter, and similarity minimization, based on the Matusita's measure. The total cost function is defined as
%\begin{align}
%J=J_P+cJ_S
%\end{align}
%where $J_P$ is the uncertainty cost, $J_S$ is the similarity cost, and $c$ is a positive constant for relative weighting. The cost function is shown by varying the first element of $K_1$ in Figure \ref{CostTrends}, although $c$ is chosen much larger for this figure than in practice to depict the effect of including a similarity cost rather than just uncertainty cost.
%
%\begin{figure}
%\centerline{
%\includegraphics[width=9cm]{CostTrends.png}
%	}
%\caption{The cost functions of target 1 are examined under the variation of a single element of $K_1$. The resulting cost function (black) is minimized slightly to the right of the uncertainty cost function (blue). Note that the total cost when the Kalman gain is applied does not lie on the total cost curve because other elements of this gain that are not optimal. In Kalman filtering applications, including the JPDAF, only the uncertainty cost function is minimized.}
%\label{CostTrends}
%\end{figure}
%
%The uncertainty cost $J_P$ is defined as
%\begin{align}
%J_P=\tr{P^+_1}+\tr{P^+_2}.
%\end{align}
%Note that $P^+_i$ must be valid for all $K_i$. Hence, from the general form of Equation \ref{genPp}, we seclude a $P_i^-$ term to take the form of Equation \ref{KalCov} to obtain
%\begin{align}
%\label{newPp}
%P^+_i=&P^-_i-\beta_i[K_iH_iP_i^-+P_i^-H_i^TK_i^T\\
%&-K_i(H_iP_i^-H_i^T+R_i)K_i^T].
%\end{align}
%
%For the remaining equations of this section, the only derivatives taken are with respect to $K_1$ for the similarity between targets $1$ and $2$ for simplicity; taking derivatives with respect other $K_i$ values are identical except for subscripts.
%
%First,  using the property $\frac{\partial}{\partial X}\tr{XYX^T}=2XY$ for symmetric $Y$, we take the derivative
%\begin{align}
%\label{CostP}
%\begin{split}
%\frac{\partial J_{P,1}}{\partial K_{1}}&=2\beta_1\{{K_1(H_1P_1^-H_1^T+R_1)-P_1^-H_1^T}\}.
%\end{split}
%\end{align}
%
%To appropriately define the similarity cost, we only consider the $R$ term from Equation \ref{R} because coalescence is well known to occur when estimates are within close proximity. For simplicity, we define ${P^+_{12}}^{-1}$ as
%\begin{align}
%{P^+_{12}}^{-1}=(P_1^++P_2^+)^{-1}
%\end{align}
%and define define $u$ as
%\begin{align}
%\begin{split}
%u=&(\hat x^+_1-\hat x^+_2)^T{P^+_{12}}^{-1}(\hat x^+_1-\hat x^+_2)\\
%=&\{\hat x_1^{-T}{P^+_{12}}^{-1}\hat x_1^-
%-2\hat x_1^{-T}{P^+_{12}}^{-1}\hat x_2^-\\
%&+2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T{P^+_{12}}^{-1}K_1e_1
%+\hat x_2^{-T}{P^+_{12}}^{-1}\hat x_2^-\\
%&+2\beta_2(\hat x_2^{-}-\hat x_1^{-})^T{P^+_{12}}^{-1}K_2e_2
%+\beta_1^2e_1^TK_1^T{P^+_{12}}^{-1}K_1e_1\\
%&-2\beta_1\beta_2e_2^TK_2^T{P^+_{12}}^{-1}K_1e_1+\beta_2^2e_2^TK_2^T{P^+_{12}}^{-1}K_2e_2\}
%\end{split}
%\end{align}
%which is the most reduced form to differentiate with respect to $K_1$ and $K_2$.
%We generalize the similarity cost further to
%\begin{align}
%J_S=&\exp{\left\{-au\right\}}
%\end{align}
%for positive constant $a$, which serves to amplify or attenuate the function as well as alter its shape.
%Defining $ij1\in \Re^{n\times m}$ such that $ij1(i,j)=1$ and $0$ elsewhere, we define $P^*_{12}$ as
%\begin{align}
%\begin{split}
%P^*_{12,1}=&\frac{\partial {P^+_{12}}^{-1}}{\partial K_{1,ij}}\\
%=&\beta_1{P_1^+}^{-1}[ij1H_1P_1^-+P_1^-H_1^Tij1^T\\
%&-ij1(H_1P_1H_1^T+R_1)\\
%&-K_1(H_1P_1^-H_1^T+R_1)ij1^T]{P_1^+}^{-1}
%\end{split}
%\end{align}
%and using the chain rule, we take the derivative
%\begin{align}
%\frac{\partial J_{S,1}}{\partial K_{1,ij}}=-a\exp{\left\{-au\right\}}\frac{\partial u}{\partial K_{1,ij}}
%\end{align}
%where
%\begin{align}
%\begin{split}
%\frac{\partial u}{\partial K_{1,ij}}=&\frac{\partial}{\partial K_{1,ij}}\left\{\hat x^+_1-\hat x^+_2)^T{P^+_{12}}^{-1}(\hat x^+_1-\hat x^+_2\right\}\\
%=&\{\hat x_1^{-T}P^*_{12,1}\hat x_1^-
%-2\hat x_1^{-T}P^*_{12,1}\hat x_2^-\\
%&+2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T[P^*_{12,1}K_1+{P^+_{12}}^{-1}ij1]e_1\\
%&+\hat x_2^{-T}P^*_{12,1}\hat x_2^-+2\beta_2(\hat x_2^{-}-\hat x_1^{-})^TP^*_{12,1}K_2e_2\\
%&+\beta_1^2e_1^T[ij1^T{P^+_{12}}^{-1}K_1+K_1^TP^*_{12,1}K_1\\
%&+K_1^T{P^+_{12}}^{-1}ij1]e_1-2\beta_1\beta_2e_2^TK_2^T[P^*_{12,1}K_1\\
%&+{P^+_{12}}^{-1}ij1]e_1+\beta_2^2e_2^TK_2^TP^*_{12,1}K_2e_2\}
%\end{split}
%\end{align}
%for $i=1:n$ and $j=1:m$ to obtain $\frac{\partial u}{\partial K_{1}}$.
%
%The weighted cost is minimized when $K_1$ and $K_2$ are chosen such that
%\begin{align}
%\frac{\partial J}{\partial K}=\frac{\partial J_{P,1}}{\partial K_{1}}+\frac{\partial J_{P,2}}{\partial K_{2}}
%+c\left(\frac{\partial J_{S,1}}{\partial K_{1}}+\frac{\partial J_{S,2}}{\partial K_{2}}\right)=0.
%\end{align}
%The complexities of $(P^+_1+P^+_2)^{-1}$ prevents a closed-form solution for this equation; hence, we require a nonlinear equation solver.
%The equations vary in computational cost depending on the level of precision of the nonlinear equation solver.
%Note that ${P_1^+}^{-1}$, ${P_2^+}^{-1}$, and $(P^+_1+P^+_2)^{-1}$ are symmetric inverses and must only be calculated once each iteration of the solver.
%Furthermore, this weighted cost may be readily extended to an arbitrary known number of targets provided the Matusita's measure is included between each estimate.
%
%The algorithm is completed by updating the state estimate and covariance. Equation \ref{KalEst} may be used for $\hat x^+_i$ because this equation does not require a Kalman gain; however we must use Equation \ref{newPp} to obtain $P^+_i$ because Equation \ref{KalCov} is only valid for a Kalman gain.
%
%	% Choice of optimal cost
%		% the part of B-distance representing the difference of mean is used
%		% the role of the weighting factor of c
%		
%	% Necessary conditions for optimality
%		% Derivation of \deriv{J}{K}
%		
%	% Filter equations
%		% Show why we can use the same expression for the mean value and the covariance of JPDA described in the previous section
%		
%%	\subsection{Suboptimal Coalescence Avoidance}
%	
%	% Linear approximation of the gain to obtain explicit solution
%	
%\section{Numerical Examples}
%\label{NumRes}
%	% Describe two body problem of orbital mechanics
%	% Describe the measured values and the initial conditions
%
%	% Results of the perfect association with 100% confidence (imaginary case for comparison)
%	% Results of JPDA (incorrect association after crossing)
%	% Results of the proposed optimal filtering approach
%	% Results of the proposed suboptimal filtering approach with linear approximation
%
%	% Discussion that the proposed approach is between JPDA and perfect association in terms of filtering performance
%	% Show how much additional computational load is required compared with JPDA
%
%We present a satellite tracking example where some parameters are varied to evaluate the performance of the CAOJPDAF with the JPDAF. We consider the scenario when we observe two satellites in LEO with crossing trajectories wherein we receive range measurements of unknown origination. This problem is simplified to a two-body problem between each satellite and Earth, i.e., the of equations of motion for each orbit only involve force caused by the gravity between the Earth and the satellite and neglect all other forces. For simplicity, both orbits occupy the same plane, so cylindrical coordinates are used to define our state variable $x=\begin{bmatrix} r & \dot r & \theta & \dot \theta\end{bmatrix}^{T}$. We linearize the equations of motion and bring to the appropriate form of this paper
%\begin{align}
%A_{i_k}=
%exp\left(\begin{bmatrix}
%0 & 1 & 0 & 0 \\
%\frac{2\mu}{r_{i_k}^3} +\dot\theta^2_{i_k} & 0 & 0 & 2r_{i_k}\dot \theta_{i_k} \\
%0 & 0 & 0 & 1 \\
%2r_{i_k}^{-2}\dot r_{i_k}\dot\theta_{i_k} & -2r^{-1}\dot\theta_{i_k} & 0 &  -2r_{i_k}^{-1}\dot r_{i_k}\\
%\end{bmatrix}dt\right)
%\end{align}
%where $\mu=GM_{Earth}=3.986\times10^5 km^3/s^2$ and $H=\begin{bmatrix} 1 & 0 & 0 & 0\end{bmatrix}$.
%We assume we have very nice priori knowledge from other measurement devices; there was no need to simulate large uncertainty when the paths are initially sufficiently far apart because this paper is concerned with the coalescence when the targets are close together, so the state and covariance will have been determined by simulated measurements while the estimates were still spread significantly apart.
%The initial estimate of the states are designed to occur shortly before the two trajectories cross paths near $r_{ref}=7000km$ from the center of the earth; for the time to complete a circular orbit at this radius, $t_{ref}=2\pi \sqrt{\frac{r_{ref}^3}{\mu}}$, we can solve for a reference circular orbit $X_{ref}=\begin{bmatrix} r_{ref} & 0 & 0 & \frac{2\pi}{t_{ref}}\end{bmatrix}^T$, then perturb this orbit slightly by $X_{pert}=\begin{bmatrix} 0.40 & 0 & 0 & -1.5\times10^{-5}\end{bmatrix}^T$. We can easily choose the initial conditions to be $X_{1_0}= X_{ref}+X_{pert}$ and $X_{2_0}= X_{ref}-X_{pert}$.
%
%We provide identical noisy range measurements for the EKF, the JPDAF, and the CAOJPDAF algorithms in Figure \ref{AllAlgorithms}. The meaning of the EKF in this example is that there exists some way to perfectly associate all measurements with the correct targets. This scenario is, of course, unrealistic but serves as a nice comparison for data association algorithms.
%For the CAOJPDAF, we choose $a=0.02$ and $c=5\times10^{-5}$. We propagate the nonlinear equations of motion using \emph{ode45} in Matlab, and produce a measurement update every $0.1sec$ with an equal likelihood that its origination is from either target.
%
%\begin{figure}
%\centerline{
%	\subfigure[EKF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{EKF.png}\hspace*{0.07\textwidth}}
%	}
%\centerline{
%	\subfigure[JPDAF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{JPDAF.png}\hspace*{0.07\textwidth}}
%	}
%\centerline{
%	\subfigure[CAOJPDAF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{CAOJPDAF.png}\hspace*{0.05\textwidth}}
%	}
%\caption{Target 1 is represented in blue and target 2 is represented in red. A solid line represents an estimate and a dotted line represents the truth.The JPDAF is subject to coalescence of the neighboring track, causing the estimations to switch trajectories. The CAOJPDAF is subject to much less coalescence and hence does not exhibit this problem.}
%\label{AllAlgorithms}
%\end{figure}
%
%The results to these numerical simulations demonstrate very important aspects of the algorithms. Most importantly, the JPDAF is unable to maintain the correct trajectories of the estimates because the tracks coalesce so much that correctly labeling the estimates becomes very difficult; when measurements from either target are sufficiently spread after the crossing, the JPDAF decision of trajectories that the estimates follow becomes ambiguous as the estimates become ambiguous. Secondly, within $20 sec$ of crossing ($201$ measurements), the number of incorrect associated (weighted less than $50\%$ likelihood of association) for the JPDAF was $92$ whereas for the CAOJPDAF this number was $70$. Comparing the performance of all three algorithms, the radial RMS error in the same window was calculated for the EKF as $1.4425\times10^{-05} km$, for the JPDAF $2.3014\times10^{-05} km$, and for the CAOJPDAF $2.3858\times10^{-05} km$. This metric demonstrates the uncertainty of association degrades the estimates as the EKF performs significantly better than algorithms where measurement origination is unknown. The JPDAF is performs slightly better than the CAOJPDAF with this metric, but this is a small drawback when compared with the ability to maintain correct tracks after crossing. The largest drawback of the CAOJPDAF is the computational cost; running the simulations on the same computer, the JPDAF only required $7.518sec$ to run the program whereas the CAOJPDAF required $25.785sec$ to run.
%
%%_______________________________________________%
%
%\section{Conclusions and Future Work}
%\label{ConclusionFutureWork}
%We explore a new approach to removing coalescence by choosing measurement update gains that minimize weighted cost with respect to posterior uncertainty in addition to similarity. The proximity similarity portion of the Matusita's measure is chosen to evaluate the similarity because of its range, scalar result, and association with coalescence. We take the derivative of the similarity cost for both estimates, yielding a result with no closed form solution, but involving expressions with low computational cost that can be solved with a nonlinear equation solver. In practice, a small similarity cost relative to the uncertainty cost yields best results; hence, the algorithm requires less time if the initial guesses of the Kalman gains are chosen.
%
%We validate the theory with numerical simulations. The CAOJPDAF shares measurements in the same way as the JPDAF, but the tracks tend to maintain separation; the CAOJPDAF does not prevent the estimates from crossing as the targets in the trajectory are designed to do. The major drawback of the CAOJPDAF is the added computation. However, the results of all three simulated algorithms (EKF, JPDAF, and CAOJPDAF) are nearly identical when the estimates are not near the crossing region (even if on the wrong trajectory). Hence, the CAOJPDAF can be turned on only during these close proximity regions, then turned off when unnecessary to be replaced by a less computationally expensive algorithm.
%
%The CAOJPDAF can be extended to a variety of multi-target tracking applications when targets share measurements due to close proximity. One example, very close to that presented in Section \ref{NumRes}, is keeping logs of satellites using phased-array radar systems. In cases where satellites have similar measurement predictions over long periods of time, satellite tracking becomes much more difficult. Implementing a simple JPDAF can serve to coalesce the estimates, which can cause the estimates of the satellites to switch trajectories as shown in this paper. Therefore, including the CAOJPDAF under certain proximity conditions can expand our ability to track satellites. The CAOJPDAF has tracking applications for space, aerial, and underwater vehicles in close proximity and identification of individual agents in multi-agent systems.
%
%Much like the Kalman filter, the CAOJPDAF can be effective for systems with well-known Gaussian distributions of state, sensor noise, and process. This paper has assumed that measurements come from a single target of interest at a single time and not clutter. Future work may involve expanding the capabilities of the algorithm. We can consider the measurements as a scan, or so that multiple measurements, including those from external sources of clutter, may belong to multiple targets at any time step as a JPDAF does without the assumptions of this paper. A general form of this algorithm could be derived for these conditions. Furthermore we can explore other techniques to measure similarity. By only using the proximity portion Matusita's measure, we are assuming the covariances of the estimates are close, which is not always accurate in all applications, and another future direction of this work.

\begin{appendix}
\label{append}

We incorporate all of the assumptions from Section \ref{DAP} to simplify the weighting term for the estimates.
We consider time step $k$ for the following proofs so subscripts omitted in this appendix.
We define ${\bf{\Theta}}\in[\Theta_1, \Theta_2]$, as all possible outcomes, namely that target $1$ is measured or target $2$ is measured from $m_k=1$ total measurements. Because all events ${\bf{\Theta}}$ fall inside an arbitrary volume $V$, or are subject to being measured, we define the measurement association indicator $\tau=1$, meaning that measurement $k$ is associated with exactly one of the two targets. Hence, the number of unassociated measurements $\phi({\bf{\Theta}})=0$ because the event that neither target is measured $\Theta_0\notin{\bf{\Theta}}$. We define the binary target detection indicator $\delta_i(\Theta_i)$, which indicates the association of any measurement with target $i$; hence, $\delta_i(\Theta_i)=1$ if measurement $k$ is associated with target $i$ and $\delta_i(\Theta_i)=0$ otherwise.
We also consider equal likelihood of detection $P(\Theta_i)$ for each measurement; we can generalize the number of estimates to $y$ such that $P(\Theta_i)=\frac{1}{y}$ for equal likelihood of event $\Theta_i$.
From~\cite[Section 9.3]{TrackDataAssoc}, we determine the weighting for $T$ possible targets to be
\begin{align}
\beta_i\equiv P(\Theta_i|z)=&\frac{1}{c}\frac{\phi!}{V}\displaystyle\prod_{j=1}^{m_k} f_{i_j}^{\tau_j}\displaystyle\prod_{i=1}^{T} P(\Theta_i)^{\delta_i}(1-P(\Theta_i))^{1-\delta_i}\\
=&\frac{1}{cV}f_{i}\frac{1}{y}(1-\frac1y)^{T-1}
\end{align}
for normalizing constant $c$ and Gaussian pdf $f_{i}$. Note that $T-1$ targets are not associated and only one target is associated. Hence, $c$ is defined as
\begin{align}
c=&\frac{1}{V}\frac{1}{y}(1-\frac1y)^{T-1}\displaystyle\sum\limits_{l=1}^y f_{l}.
\end{align}
Therefore, we solve for the simplified weighting
\begin{align}
\beta_i=\frac{f_i}{\displaystyle\sum\limits_{l=1}^y f_{l}}
\end{align}
which is much simpler after assumptions.

The second proof shows how the measurement updates may take a more manipulatable form under the assumptions. From~\cite[Section 6.4]{TrackDataAssoc},
\begin{align}
\bar e=&\displaystyle\sum\limits_{i=1}^{m_k}\beta_ie_i=\beta_ie_i\\
\hat x^+_i=&\hat x^-_i+K_i\bar e=\hat x^-_i+\beta_iK_ie_i.
\end{align}
 , so this portion of the covariance Let $\tilde{P}$ be spread in the innovations term on each target. Because only one innovation is calculated per target, there is no spread in this term; hence $\tilde{P}=0$. Note that his term would be nonzero in the JDPAF without the single measurement assumption. Let $P_{i}^C$ be the posterior covariance of target $i$ conditioned on event $\Theta_i$. For any $K_i$ from ~\cite[Section 4.2]{OptEst1}:
\begin{align}
\begin{split}
P_{i}^C=&(I-K_{i}H_{i})P_{i}^{-}(I-K_{i}H_{i})^T+K_{i}R_{i}K_{i}^T\\
=&P^--K_{i}H_{i}P_{i}^{-}-P_{i}^{-}H_{i}^TK_{i}^T\\
&+K_{i}(H_{i}P_{i}^{-}H_{i}^T+R_i)K_{i}^T.
\end{split}
\end{align}
We define $\beta_{0_i}$ as the probability that target $i$ is not measured; with only one measurement considered each time step; hence, $\beta_{0_i}=1-\beta_i$. Again from~\cite[Section 6.4]{TrackDataAssoc},
\begin{align}
\begin{split}
P^+_i=&\beta_{0_i}P^-_i+(1-\beta_{0_i})P_i^C+\tilde{P}\\
=&(1-\beta_i)P^-+\beta_i[P^--K_{i}H_{i}P_{i}^{-}\\
&-P_{i}^{-}H_{i}^TK_{i}^T+K_{i}(H_{i}P_{i}^{-}H_{i}^T+R_i)K_{i}^T]\\
=&P^-_i-\beta_i[K_{i}H_{i}P_{i}^{-}+P_{i}^{-}H_{i}^TK_{i}^T\\
&-K_{i}(H_{i}P_{i}^{-}H_{i}^T+R_i)K_{i}^T].
\end{split}
\end{align}

\end{appendix}

\bibliography{BibMaster}
\bibliographystyle{IEEEtran}
\end{document}

%Many techniques to avoid track coalescence involve the concept that when the exact nearest neighbor probabilistic data association (ENNPDA) is applied, which associates a measurement with a strong decision to the most likely estimate, the estimates are insensitive to track coalescence~\cite{Coal1}.
%By pruning the tracks, or designing a set of rules to determine high likelihood scenarios, an algorithm choice or weighing scheme is applied to the measurement updates.
%For example in~\cite{Fitzgerald}, the algorithm involves a set of rules based on the data received with respect to the estimates.
%A special logic is introduced that determines how weighting of the updates is defined or if a new track should be considered.
%In some cases, it is possible for an incorrectly associated estimate measurement may receive full weighting in the update.
%Other criteria or extensions of this approach are detailed in~\cite{Coal_d},~\cite{Coal_e}, and~\cite{Coal_c}, which can be very effective, but information may be lost when measurements are omitted from updating an estimates with small probabilities of association because the pruning rule prevented these association possibilities.

%A considerable amount of work has been done to solve the data association problem. Approaches vary in their performance, including the portion of correct associations, the ability to maintain the correct tracks, and the computational cost. In this section, we examine some of the current techniques with respect to tracking two objects in close proximity and propose a new approach to solve this problem.
%
%The multiple hypothesis tracking (MHT) approach considers multiple hypotheses for how the data may be associated for multiple tracks~\cite{MHT1}. As new information becomes available, hypotheses are evaluated, in some cases removed, and predicted again. However in complicated problems, the number hypothesis may explode. One approach to remove this issue from the close proximity problem is known as track pruning, where incompatible tracks are highlighted and only the most likely tracks are considered. This approach is complex and involves a high computational cost, so implementing this approach in real-time is very difficult. Furthermore according to~\cite{PHD1}, MHT is subject to poor results with objects in close proximity.
%
%The probability hypothesis density (PHD) process is a propagation of the first moment of a joint distribution~\cite{PHD2}. A typical PHD considers a sub-area of interest where peaks from the integral of the PHD are considered objects. In many cases, the PHD filter is hindered by computational complexity, growing exponentially with the number of objects and further increasing when the possibility that objects may enter or leave the sub-area considered. An approach described in~\cite{PHD1} maintains a tracker for each object of interest. Each tracker accounts for a priori estimates of state and dynamics to predict the probability densities for the next time step. The process is complex and involves separating the sub-spaces into cells wherein likelihoods of each object are used to define peaks and dummy peaks, where a single track or multiple tracks exist, respectively. This approach can be more computationally expensive than other recursive approaches as the algorithm is among the most complex and the resolution (defined by size of the cells where peaks lie) will require modification as objects become close.
%
%The nearest neighbor filter (NNF) is a heuristic approach that calculates the magnitude between the measurement and its prediction~\cite{JPDAF1}; the state with smallest square Mahalanobis distance of the innovation and its covariance receives a filtered update with the measurement while all other states of interest are updated with only their predictions~\cite{NN2}; this is referred to as a \emph{hard decision}~\cite{JPDAF1}. This approach adds a simple step to the EKF at the cost of accuracy as the probability of association is not explicitly calculated.
%
%The JPDAF is an extension of the probabilistic data association filter (PDAF) described in~\cite{JPDAF1} and~\cite{TrackDataAssoc}, which calculates the probabilities that measurements belong to an individual estimate. The JPDAF is designed to track multiple objects of interest. The estimates are updated with a \emph{soft decision}, or by a weighting factor based from the probability of association. However, when more than one object of interest share measurements, the objects are subject to a measurement updates that draw the estimates together, or coalesce the estimates.
%
%There are a few approaches to remove coalescence from the JPDAF. Much like the MHT, some approaches iinvolve track pruning and are detailed in~\cite{Coal1},~\cite{Coal2},~\cite{Coal_d}, and~\cite{Coal_e}, but this is computationally complex and can suffer when implementing the approach recursively. Other techniques such as those described in~\cite{Coal_c},~\cite{Coal_g},~\cite{Coal_i}, and~\cite{Coal_j} involve assigning hard decisions. The criteria and performance of these techniques vary, but each eliminate coalescence by never sharing a measurement, which can yield fully weighted incorrect associations that may cause detrimental effects such as estimates switching trajectories of crossing tracks. To decrease this problem, an arbitrary scaling factor is applied in~\cite{Coal_f} wherein the more likely association is given more weight as the less likely association weightings are decreased, yielding a hybrid technique between the NNF and the JPDAF. This approach may yield improved results with a heuristic weighting scheme but results vary with the application and expected data.
%
%Therefore, we propose a systematic approach that minimizes a weighted cost of uncertainty and similarity. We design the KMMJPDAF to minimize the posterior uncertainty much like the Kalman filter in addition to evaluating the Mahalanobis distance involved with updating the states. The paper is organized as follows: the problem is formally presented in Section \ref{ProbDef}. The optimal gain derivation and algorithm description is presented in Section \ref{KMM}. We discuss how to implement the algorithm and how to appropriately weight the minimized cost function in section \ref{AlgImpTuning}. We consider a satellite tracking example of this technique in Section \ref{NumRes}. Finally, we discuss results and applications for the KMMJPDAF in Section \ref{ConclusionFutureWork}.

%\section{Problem Definition}
%\label{ProbDef}
%The objects of interest are assumed to have nonlinear dynamics and each have Gaussian distributed estimates. We evaluate objects at time $k$ and linearize the system dynamics with respect to time $k-1$. We define the JPDAF algorithm such that we update the state as measurements are received, i.e. we only consider a single measurement rather than a scan of measurements. Furthermore, this research is concerned with the two estimates in very close proximity; hence we assume that the measurement originations are from a single object of interest and not from other sources of clutter.
%
%\subsection{Time Update} We denote an estimate of state with a $\hat{.}$, and the estimates before and after the measurement update with a $.^-$ and $.^+$, respectively. We define $x\in \Re^{n\times1}$ as the state vector and $P\in\Re^{n\times n}$ as its covariance. The time update provides the predicted state and covariance
%\begin{gather}
%\hat{x}_{k}^{-}=A_{k-1}\hat{x}_{k-1} \\
%P_{k}^{-}=A_{k-1}P_{k-1}A_{k-1}^T+Q_k.
%\end{gather}
%The covariance of process noise is captured in $Q_k$ and no control input is considered.
%
%\subsection{Measurement Update} The measurement update requires an innovation
%\begin{equation}
%\label{innovation}
%e_{k}=z_{k}-\hat z_{k|k-1}
%\end{equation}
%where $z_{k}\in\Re^{m\times1}$ is the measurement received and measurement prediction $\hat z_{k|k-1}=H_{k}\hat{x}_{k}^{-}$; $H_{k}$ maps $\hat{x}_{k}^{-}$ from the state space to the measurement space.% Check this line terminology
%We simplify the equations for probability of association from~\cite{JPDAF1} and~\cite{TrackDataAssoc} to include only a single measurement that always belongs to either one of the two objects considered; hence the process of association is limited two mutually exclusive scenarios: we measure object 1 and not object 2 or we measure object 2 and not object 1. After some manipulation we can show for object $t$, the relative probability $\beta_t$ is simplified to
%\begin{gather}
%\beta_t=\frac{f_t}{\displaystyle\sum\limits_{\tau=1}^2 f_{\tau}} \\
%f_t=\mathcal{N}[z_{k};\hat z_{k|k-1},S].
%\end{gather}
%The innovation covariance $S$ is defined as
%\begin{equation}
%S_k=H_{k}P_{k}^{-}H_{k}^T+R_k
%\end{equation}
%where $R_k$ is the covariance of the measurement. The gain to minimize the posterior uncertainty with cost function $J_P=\tr{P_{k+1}^{+}}$ is equivalent to the Kalman gain
%\begin{equation}
%\label{KalmanGain}
%K_{k}=P_{k}^{-}H_{k}^TS_k^{-1}
%\end{equation}
%which is used to calculate the updated state and its covariance for each estimate
%\begin{gather}
%\label{xhatp}
%\hat{x}_{k}^{+}=\hat{x}_{k}^{-}+\beta K_{k}e_{k} \\
%\label{oldPp}
%P_{k}^{+}=P^-_{k}-\beta K_{k}S_kK_{k}^{T}
%\end{gather}
%for the JPDAF, but not for the KMMJPDAF. For the KMMJPDAF we derive a new gain which is not compatible with Equation \ref{oldPp}, which is a manipulation involving Equation \ref{KalmanGain} of the original form (ignoring the $\beta$ term)
%\begin{equation}
%\label{genPp}
%P_{k}^{+}=(I-K_{k}H_{k})P_{k}^{-}(I-K_{k}H_{k})^T+K_{k}R_kK_{k}^T
%\end{equation}
%from ~\cite{OptEst1}. Hence in order to properly update the posterior covariance when we include coalescence into the minimization, we manipulate Equation \ref{genPp} to involve relative weighting $\beta$ of the measurement update. Note that time step subscripts will be assumed at $k$ from here forward for simplicity. Obtaining a gain accounting for posterior uncertainty and coalescence takes the generalized form
%\begin{equation}
%\label{newPp}
%P^+=P^--\beta[KHP^-+P^-H^TK^T-K(HP^-H^T+R)K^T].
%\end{equation}
%Note that if we insert the Kalman gain from Equation \ref{KalmanGain} into Equation \ref{newPp}, we can manipulate the result to Equation \ref{oldPp}.

%\subsection{Uncertainty Cost} The derivation of $J_P$ is a fairly straightforward adaptation of the Kalman filter from~\cite{OptEst1} with the included JPDAF weighting. Evaluating the trace of Equation \ref{newPp} for object 1,
%\begin{align}
%\begin{split}
%J_{P,1}=&\tr{P_1^+} \\
%=& \tr{P_1^-}+\beta_1(\tr{K_1(H_1P_1^-H_1^T+R_1)K_1^T}\\
%&-2\tr{K_1H_1P_1^-}).
%\end{split}
%\end{align}
%We can take its derivative $\frac{\partial J_{P,1}}{\partial K_{1}}$ using the property $\frac{\partial}{\partial X}\tr{XYX^T}=2XY$ for symmetric $Y$ to obtain
%\begin{align}
%\label{CostP}
%\begin{split}
%\frac{\partial J_{P,1}}{\partial K_{1}}&=2\beta_1\{{K_1(H_1P_1^-H_1^T+R_1)-P_1^-H_1^T}\}.
%\end{split}
%\end{align}
%
% \subsection{Similarity Cost} We choose the Mahalanobis distance to measure similarity because it yields a scalar result and taking its derivative with respect to measurement update gains is possible. We use $P_1^-$ as the covariace estimation of object 1 for the Mahalanobis distance with respect to the estimated mean of object $2$ can take the form
%\begin{align}
%\label{M1}
%M_1=u_1^{\frac{1}{2}}
%\end{align}
%where
%\begin{align} \label{u_1}
%\begin{split}
%u_1=&(\hat x_1^+-\hat x_2^+)^{T}{P_1^-}^{-1}(\hat x_1^+-\hat x_2^+)\\
%=&\{\hat x_1^{-T}{P_1^-}^{-1}\hat x_1^-
%-2\hat x_1^{-T}{P_1^-}^{-1}\hat x_2^-\\
%&+2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T{P_1^-}^{-1}K_1e_1
%+\hat x_2^{-T}{P_1^-}^{-1}\hat x_2^-\\
%&+2\beta_2(\hat x_2^{-}-\hat x_1^{-})^T{P_1^-}^{-1}K_2e_2
%+\beta_1^2e_1^TK_1^T{P_1^-}^{-1}K_1e_1\\
%&-2\beta_1\beta_2e_2^TK_2^T{P_1^-}^{-1}K_1e_1+\beta_2^2e_2^TK_2^T{P_1^-}^{-1}K_2e_2\}
%\end{split}
%\end{align}
%which is the most reduced form to differentiate with respect to $K_1$ and $K_2$. However, the Mahalanobis distance is an increasing measure as similarity decreases. Therefore we wish to minimize $M_1^{-2a}$ for $a>0$. Furthermore as the Mahalanobis distance is a scalar quantity, we can use the property $\frac{\partial g(u_1)}{\partial K_1}=\frac{\partial g(u_1)}{\partial u_1}\frac{\partial u_1}{\partial K_1}$ such that $J_{C,1}=M_1^{-2a}=g(u_1)=u_1^{-a}$. We take the derivative of the similarity measure using $u_1=f(K_1,K_2)$ from Equations \ref{M1} and \ref{u_1} as well as the properties $\frac{\partial}{\partial X} \tr{AXB}=A^TB^T$ and $\frac{\partial}{\partial X}\tr{AXBX^TC}=A^TC^TXB^T+CAXB$ to obtain, after some manipulation,
%\begin{align}
%\begin{split}
%\label{CostS}
%\frac{\partial J_{S,1}}{\partial K_1}=&-au_1^{-(a+1)}\frac{\partial u}{\partial K_1}\\
%=&2a\beta_1u_1^{-(a+1)}\{
%{P_1^-}^{-1}(\hat x_2^--\hat x_1^-)\\
%&-\beta_1{P_1^-}^{-1}K_1e_1
%+\beta_2{P_1^-}^{-1}K_2e_2
%\}e_1^T.
%\end{split}
%\end{align}
%
%We absorb the constants from Equations \ref{CostP} and \ref{CostS} into $c$ from Equation \ref{CostDef} and use a nonlinear equation solver for the optimal gains $K_1$ and $K_2$ in the joint equations
%\begin{align}
%\label{Opt1}
%\begin{split}
%\frac{\partial J_{1}}{\partial K_{1}}=&K_1(H_1P_1^-H_1^T+R_1)-P_1^-H_1^T\\
%&+cu_1^{-(a+1)}\{
%{P_1^-}^{-1}(\hat x_2^--\hat x_1^-)\\
%&-\beta_1{P_1^-}^{-1}K_1e_1
%+\beta_2{P_1^-}^{-1}K_2e_2
%\}e_1^T\\
%=&0_{n\times m}
%\end{split}
%\end{align}
%and similarly
%\begin{align}
%\label{Opt2}
%\begin{split}
%\frac{\partial J_{2}}{\partial K_{2}}=&K_2(H_2P_2^-H_2^T+R_2)-P_2^-H_2^T\\
%&+cu_2^{-(a+1)}\{
%{P_2^-}^{-1}(\hat x_1^--\hat x_2^-)\\
%&-\beta_2{P_2^-}^{-1}K_2e_2
%+\beta_1{P_2^-}^{-1}K_1e_1
%\}e_2^T\\
%=&0_{n\times m}.
%\end{split}
%\end{align}
%These equations vary in computational cost depending on the level of precision of the nonlinear equation solver. Note that ${P_1^-}^{-1}$ and ${P_2^-}^{-1}$ are symmetric inverses and must only be calculated once each.

%\section{Algorithm Implementation and Tuning}
%\label{AlgImpTuning}
%This algorithm requires considerably more computational cost than the JPDAF alone; hence, we only implement this algorithm when the likelihood of coalescence is high. For example, if the association of a measurement belongs primarily to a single estimate, the KMMJPDAF and JPDAF yield similar results so we choose JPDAF because of its closed form solution and hence lower computational cost. Therefore depending on the application, we determine thresholds such that beyond some spacing between the a priori estimates we default to the JPDAF algorithm for speed.
%
%We must also examine the shape of the cost function. Consider Figure \ref{CostTrends}, which contains a variation of a single element of $K_1$ from an example in Section \ref{NumRes} and plots both cost functions of object 1. The uncertainty cost minimizes at a single point. However, the similarity cost contains a maximum. If we incorrectly choose $K_1$ and $K_2$ to maximize $J_S$, this would serve to increase coalescence because similarity would be increasing at a maximum rate due only to a measurement update.
%
%\begin{figure}
%\centerline{
%\includegraphics[width=9cm]{CostTrends.png}
%	}
%\caption{The cost functions of object 1 ($J_{P,1}$, $cJ_{C,1}$, and $J_{1}$) are examined under the variation of a single element of $K_1$. The resulting cost function (black) is minimized slightly to the left of the uncertainty cost function (blue). In Kalman filtering applications, including the JPDAF, only the uncertainty cost function is minimized.}
%\label{CostTrends}
%\end{figure}
%
%There are a few approaches to avoid this detrimental scenario. First, we can choose $a$ and $c$ properly. Decreasing $a$ serves to amplify and flatten out $J_S$. Once a desired cost function shape as achieved, we can choose $c$ to bring $J_S$ to an appropriate scaling. In practice, the JPDAF is a very effective algorithm for estimation of multiple objects, which only relies on $J_P$, so $J_S$ is chosen to be very small in comparison only to remove coalescence from the estimates (Figure \ref{CostTrends} shows $cJ_{S,1}$ much larger in magnitude than in practice for illustration). Choice of these parameters leads to trade-offs: if $a$ is decreased along with $c$, the similarity cost varies less, decreasing the beneficial effects of the KMMJPDAF over the JPDAF. However, when $a$ is increased (and $c$ is increased to compensate), we may incorrectly solve for the similarity maximizing point.
%
%We explore two techniques to find an effective minimization. First, we choose the initial guess of $K_1$ and $K_2$ to be that of the EKF, which is already located near the minimized point of uncertainty. Secondly, we only implement the KMMJPDAF when the a priori estimates are sufficiently spaced apart. This way we can maintain the desirable shape of the cost function and implement the KMMJPDAF for the majority of the region where coalescence poses an issue. Therefore, we have both minimum and maximum estimate spacing thresholds for which KMMJPDAF is appropriate.

%In fact, the results are very similar with two noteworthy differences. The first is that within $20 sec$ of crossing ($201$ measurements), the number of incorrect associated (weighted less than $50\%$ likelihood of association) for the JPDAF was $64$ whereas for the KMMJPDAF the number was $58$, but this is a minor difference in the metric. The second difference is that the JPDAF completed in $4.212sec$ while the KMMJPDAF ran in $6.8952sec$ on the same computer. For real-time implementation, either algorithm is capable of receiving these simple range measurements as described above, but when measurements are more complicated or are received at a faster rate, there can be cases where the performance of the JPDAF computational time makes the KMMJPDAF an inferior algorithm.
%
%The noise of the measurements are increased in second example (Figure \ref{Case2}), and this causes the JPDAF to permanently lose the correct association after the crossing. The optimization of the KMMJPDAF compensates for the coalescence of the tracks, and hence correct track maintenance is achieved properly with this algorithm. The result of this example does not imply that the KMMJPDAF is impervious the the effects of coalescence; rather including similarity in the minimization serves to remove some of the coalescence from measurement updates. For the case presented here, choosing the KMMJPDAF rather than the JPDAF increases the computational time from $4.1964sec$ to $6.8018sec$ to maintain the estimates with the correct trajectories.
%
%We now use the last example to demonstrate the importance of choosing the parameters correctly for the effective implementation of the KMMJPDAF. In Figure \ref{changeParameters}.(a), we increase $a$ such that $a=1$, which serves to attenuate the similarity cost function. As a result, the gains are chosen to minimize uncertainty with greater weight much like the JPDAF. This results in increasing coalescence, yielding a poorer estimation of the states. In this case, the estimates are drawn very close together and diverge incorrectly.
%
%We consider the same example but this time increase $c$ such that $c=1\times10^{-7}$, shown in Figure \ref{changeParameters}.(b). Increasing $c$ increases the similarity cost function proportionally according to Equation \ref{CostDef}. Contrary to the last example, the gains are chosen to decrease similarity of the estimates so much that the measurement updates serve to separate the estimates when the estimates become close. We experience the opposite effect with increasing $c$ than increasing $a$, but choosing either parameter poorly can lead to estimates switching trajectories.

### 2018-06-09 Conference/ECC 2014/ECC14_Lee.tex
added Conference folder

--- 2018-06-09 Conference/ECC 2014/ECC14.tex
@@
 \usepackage{amsmath,amssymb,url}
-\usepackage{graphicx,subfigure,tikz}
+\usepackage{graphicx,subfigure,tikz,graphicx}
 \usepackage{color}
@@
 %Recursive Kalman-Mahanobis JPDAF Minimization to avoid Coalescence of Estimates in Close Proximity
-Coalescence Avoiding Optimal Joint Probabilistic Data Association Filter}
-\author{Evan Kaufman and Taeyoung Lee
+Optimal Joint Probabilistic Data Association Filter\\ Avoiding Coalescence in Close Proximity }
+\author{Evan Kaufman and Taeyoung Lee$^*$
  \thanks{Evan Kaufman and Taeyoung Lee are with Department of Aerospace Engineering, George Washington University, Washington, DC. Email: {\tt\footnotesize \{evankaufman, tylee\}@gwu.edu }}
-}
+\thanks{$^*$This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}}
 
 \newcommand{\EditTL}[1]{{\color{red}\protect #1}}
-\renewcommand{\EditTL}[1]{{\protect #1}}
+%\renewcommand{\EditTL}[1]{{\protect #1}}
 
@@
 This paper deals with an estimation problem where a known number of targets in close proximity are observed but the measurement originations are uncertain.
-It is well known that joint probabilistic data association filters (JPDAF) are effective in handing uncertain measurement originations with clutters, but they are prone to estimation coalescence, particularly when close neighboring targets share measurements.
+It is well known that joint probabilistic data association filters (JPDAF) are effective in handing uncertain measurement originations with clutter, but they are prone to estimation coalescence, particularly when close neighboring targets share measurements.
 This paper proposes a Coalescence Avoiding Optimal JPDAF (C-JPDAF) that minimizes the weighted sum of the posterior uncertainty and a measure of similarity between estimated probability densities.
@@
 
-A variety of data association techniques are considered for the close-proximity data association with varying performance.
+A variety of data association techniques are considered for close-proximity trajectories with varying performance.
 Some approaches such as multiple hypothesis tracking (MHT) and probability hypothesis density (PHD) techniques involve a high computational load, especially when targets are in close proximity, making real-time implementation more difficult~\cite{MHT1},~\cite{PHD1}, and~\cite{PHD2}.
@@
 
-% Dr. Lee Edits:
+% Old introduction:
+
+%\section{Introduction}
+%
+%Data association has been widely studied with various applications in estimation of a single or multiple targets with measurements that are not necessarily originating from correct targets in order.
+To properly track multiple targets with measurement origin uncertainties, measured data should be properly associated to targets with certain probabilities.
+%When the targets are in very close proximity, the association uncertainty grows, and an effective algorithm is required to maintain close estimations of the targets.
+Otherwise, estimators tend to coalesce neighboring tracks.
+%
+%%This paper considers estimation wherein multiple targets of interest enter the field of view of a sensor; however, the measurement origin is unknown. 
+%
+%A variety of data association techniques with varying performance have been studied.
+A common heuristic approach is known as the nearest neighbor filter (NNF), where measurements with the closest distance are only applied~\cite{NN2}.
+This is considered a \emph{hard decision} in~\cite{JPDAF1}, as every selected measurement update receives full weighting.
+Therefore, if the decision is incorrect, these measurement updates can result in detrimental effects to the associations of the estimates.
+Alternatively, there are approaches that calculate the minimum mean square error (MMSE) of the estimates and their associated uncertainty.
+The probabilistic data association (PDA) technique is effective in tracking a single target because it updates with a \emph{soft decision}; in particular, this approach updates provides a weighting from a likelihood of measurement association.
+Other approaches include multiple hypothesis tracking (MHT) and probability hypothesis density (PHD) techniques~\cite{MHT1,PHD1,PHD2}.
+%
+%%Many other techniques may involve the Kalman filter (KF) or extended KF (EKF), which are well studied and provide a framework for recursive data association techniques that are much easier to implement in real-time.
+%
+%
+%The joint PDA (JPDA) is an extension of the PDA to incorporate multiple targets~\cite{JPDAF1},~\cite{TrackDataAssoc}.
+%The JPDA filter (JPDAF) is particularly advantageous in multi-target scenarios because when targets share measurements, the soft decisions prevent loss of information that occurs with hard decisions.
+As such, it is well known that JPDAF are effective in handling data clutter or missed detection.
+However, when the targets are in very close proximity, the association uncertainty grows significantly, thereby degrading the filtering performance.
+Furthermore, the estimated probability distributions tend to be coalesced especially if the paths of neighboring targets cross with each other.
+%
+%%neighboring targets share measurements, the estimates are subject to coalescence.
+%
+%Many techniques to avoid track coalescence involve the concept that when the exact nearest neighbor probabilistic data association (ENNPDA) is applied, which associates a measurement with a strong decision to the most likely estimate, the estimates are insensitive to track coalescence~\cite{Coal1}.
+By pruning the tracks, or designing a set of rules to determine high likelihood scenarios, an algorithm choice or weighing scheme is applied to the measurement updates.
+%For example in~\cite{Fitzgerald}, the algorithm involves a set of rules based on the data received with respect to the estimates.
+A special logic is introduced that determines how weighting of the updates is defined or if a new track should be considered.
+In some cases, it is possible for an incorrectly associated estimate measurement may receive full weighting in the update.
+Other criteria or extensions of this approach are detailed in~\cite{Coal_d},~\cite{Coal_e}, and~\cite{Coal_c}, which can be very effective, but information may be lost when measurements are omitted from updating an estimates with small probabilities of association because the pruning rule prevented these association possibilities.
+%
+%This paper is focused on optimal joint probabilistic data association filter that avoids coalescence in close proximity.
+We propose a novel approach to remove coalescence by minimizing a similarity index of the estimates.
+More explicitly, the gain of the proposed estimator is chosen such that a weighted sum of the size of uncertainties and a similarity measure between estimated states is minimized at each update.
+This yields an optimal recursive estimation algorithm that compensates for the effects of coalescence systematically.
+The proposed approach has simpler structures compared with pruning-based approaches, and it is computationally efficient.
+As it is still based on probabilistic data association, it avoids the heuristic nature, loss of information, or the risk of other coalescence avoidance approaches based on hard-decision.
+%
+%The paper is organized as follows: the dynamics of target and the measurement models are described in Section \ref{ProbDef}.
+The proposed Coalescence Avoiding Optimal JPDAF (C-JPDAF) are presented at Sections \ref{JPDAF}, and the optimal gain that minimizes a weighted cost of uncertainty and similarity is derived in Section \ref{C-JPDAF}.
+%These are illustrated by a numerical example for estimating two satellites in crossing orbits in Section \ref{NumRes}, followed by conclusions in Section \ref{ConclusionFutureWork}.
+
+%_______________________________________________%
+
 \section{Data Association Filtering Problem Formulation}
@@
 
-More explicitly, let $\Theta_i$ be the event that the measurement originated from the $i$-th target, for $i\in\{1,2\}$. According to the above assumptions, we have
+More explicitly, let $\Theta_i$ be the event that the measurement originates from the $i$-th target, for $i\in\{1,2\}$. According to the above assumptions, we have
 \begin{align}
@@
 
-Joint probabilistic data association filter is a suboptimal Bayesian filter for multiple targets where the probabilities of association is computed for the latest set of measurements only~\cite{TrackDataAssoc}.
+Joint probabilistic data association filter is a suboptimal Bayesian filter for multiple targets where the probabilities of association are computed for the latest set of measurements only~\cite{TrackDataAssoc}.
 This section summarizes the approaches of JPDAF.
@@
 
-The data association probability is defined as follows. Let $\beta_i$ be the probability of the $i$-th data association event conditioned by the given measurement:
+The date association probability is defined as follows. Let $\beta_i$ be the probability of the $i$-th data association event conditioned by the given measurement:
 \begin{align*}
@@
 However, it is prone to coalescing estimates, especially with targets in close-proximity.
-Here, we present a coalescence-avoiding optimal JPDAF, hereafter C-JPDAF for short, to avoid data coalescence.
+Here, we present a coalescence-avoiding optimal JPDAF, hereafter C-JPDAF for short, to avoid coalescence of the estimates.
 The main idea is choosing the estimator gain such that a weighted sum of the size of uncertainty and a measure of similarity is minimized.
@@
 
-For two probability density functions, $p_1(x)$ and $p_2(x)$, Matusita's measure is defined as
+For two probability density functions, $f_1(x)$ and $f_2(x)$, Matusita's measure is defined as
 \begin{align*}
@@
 In the proposed approach, the estimator gain $K$ is chosen such that the weighted sum of the uncertainty cost and the similarity cost is minimized.
-In short, at the expense of increased uncertainty at a single step, the proposed approach prevents target coalescence.
+In short, at the expense of increased uncertainty at a single step, the proposed approach prevents coalescence.
 The increased uncertainty at each step is rewarded by reducing the overall estimation error drastically by maintaining correct data association throughout crossing tracks in close proximity.}
@@
 Next, we derive the derivative of each cost.
-For simplicity, derivatives with respect to the first gain $K_1$ are presented, and derivatives with respect to $K_2$ can be obtain by symmetry.
+For simplicity, derivatives with respect to the first gain $K_1$ are presented, and derivatives with respect to $K_2$ can be obtained by symmetry.
 By substituting (\ref{KalCov}) into \refeqn{JP} and using the property $\frac{\partial}{\partial X}\tr{XYX^T}=2XY$ for any matrix $X$ and any symmetric matrix $Y$, we obtain
@@
 -2\hat x_1^{-T}\mathcal{P}_{ij}\hat x_2^-\nonumber\\
-&+2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T[\mathcal{P}_{ij}K_1+\mathcal{P}_{ij}^{-1}\Iij]e_1\nonumber\\
+&+2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T[\mathcal{P}_{ij}K_1+{P^+_{12}}^{-1}\Iij]e_1\nonumber\\
 &+\hat x_2^{-T}\mathcal{P}_{ij}\hat x_2^-+2\beta_2(\hat x_2^{-}-\hat x_1^{-})^T\mathcal{P}_{ij}K_2e_2\nonumber\\
-&+\beta_1^2e_1^T[\Iij^T{\mathcal{P}_{ij}}^{-1}K_1+K_1^T\mathcal{P}_{ij}K_1\nonumber\\
-&+K_1^T{\mathcal{P}_{ij}}^{-1}\Iij]e_1-2\beta_1\beta_2e_2^TK_2^T[\mathcal{P}_{ij}K_1\nonumber\\
-&+{\mathcal{P}_{ij}}^{-1}\Iij]e_1+\beta_2^2e_2^TK_2^T\mathcal{P}_{ij}K_2e_2\}.
+&+\beta_1^2e_1^T[\Iij^T{P^+_{12}}^{-1}K_1+K_1^T\mathcal{P}_{ij}K_1\nonumber\\
+&+K_1^T{P^+_{12}}^{-1}\Iij]e_1-2\beta_1\beta_2e_2^TK_2^T[\mathcal{P}_{ij}K_1\nonumber\\
+&+{P^+_{12}}^{-1}\Iij]e_1+\beta_2^2e_2^TK_2^T\mathcal{P}_{ij}K_2e_2\}.
 \end{align}
@@
 
-Three filtering approaches are applied, namely extended Kalman filter (EKF), JPDAF, and C-JPDAF, such that a single measurement was evaluated at each time step, and its origination was random and equally likely from both targets.
-For EKF, it is assumed that the correct measurement origination is know with certainty, i.e., all of the measurements are associated to correct target.
+%\EditTL{Put the values of $P_0$, $Q$ and $R$ here. Describe how the measurements are switched between two satellites. }
+
+Three filtering approaches are applied, namely extended Kalman filter (EKF), JPDAF, and C-JPDAF.
+For the EKF, it is assumed that the correct measurement origination is know with certainty, i.e., all of the measurements are associated to the correct target.
 This serves as an ideal reference case to be compared with JPDAF and C-JPDAF.
 For the C-JPDAF, the weighting parameters of the cost are chosen as $a=0.02$ and $c=5\times10^{-5}$.
-We propagate the nonlinear equations of motion numerically using Matlab, and we generate measurement at every $0.1$ seconds with an equal likelihood that its origination is from either target.
+We propagate the nonlinear equations of motion numerically using Matlab, and we generate a measurement at every $0.1$ seconds with an equal likelihood that its origination is from either target such that identical noisy measurements are used in all three algorithms.
 
@@
 \caption{Target 1 is represented in blue and target 2 is represented in red.
-A solid line represents an estimate and a dotted line represents the truth.The JPDAF is subject to coalescence of the neighboring track, causing the estimations to switch trajectories.
+A solid line represents an estimate and a dotted line represents the truth.
+The JPDAF is subject to coalescence of the neighboring track, causing the estimations to switch trajectories.
 The C-JPDAF is subject to much less coalescence and hence does not exhibit this problem.}
@@
 \label{ConclusionFutureWork}
-We explore a new approach to reduce coalescence of multiple target tracking.
+We explore a new approach to reduce coalescence of close proximity multi-target tracking.
 The estimator gains are chosen such that a weighted cost with respect to posterior uncertainty and similarity is minimized.
@@
 Compared with other techniques to avoid coalescence, the proposed approach has simpler structures.
-Future works include generalizing C-JPDAF for arbitrary number of targets with clutters, and improving the computational efficiencies.
+Future works include generalizing C-JPDAF for arbitrary number of targets with clutter, and improving the computational efficiencies.
 
@@
 
-%%_______________________________________________%
-%
-%\section{Data Association Problem Formulation}
-%\label{ProbDef}
-%
-%% Equations of motion
-%\subsection{Equations of Motion}
-%
-%Consider a set of linear stochastic discrete-time systems. The equations of motion for the $i$-th system are described by
-%\begin{align}
-%x_{i_{k+1}} & = A_{i_k} x_{i_k} + w_{i_k},\\
-%z_{i_k} & = H_{i_k} x_{i_k} + v_{i_k},
-%\end{align}
-%where $x_{i_k}\in\Re^n$ is the state vector, $z_{i_k}\in\Re^m$ is the output vector, and  the matrices $A_{i_k}\in\Re^{n\times n}$ and $H_{i_k}\in\Re^{m\times n}$ describe system dynamics and outputs.
-The process noise and the measurement noise are denoted by $w_{i_k}\in\Re^n$ and $v_{i_k}\in\Re^m$, respectively, and it is assumed that they are zero-mean, mutually independent white, Gaussian noise sequences with known covariance matrices $Q_{i_k}\in\Re^{n\times n}$ and $R_{i_k}\in\Re^{m\times m}$, respectively:
-%\begin{align}
-%w_{i_k} \sim \mathcal{N}(0,Q_{i_k}),\quad
-%v_{i_k} \sim \mathcal{N}(0,R_{i_k}).
-%\end{align}
-%
-%
-%The initial states are also assumed to be mutually independent and independent to noise. They are Gaussian with known mean values and covariance matrices:
-%\begin{align}
-%x_{i_k} \sim \mathcal{N}(\bar x_{i_k}, P_{i_k}),
-%\end{align}
-%where $\bar x_{i_k}\in\Re^n$ and $P_{i_k}\in\Re^{n\times n}$. 
-%
-%To demonstrate the main idea of the proposed data association filter concisely, it is assumed that there are two dynamic systems, i.e., $i\in\{1,2\}$.
-But, the results of this paper are readily generalized to arbitrary number of systems provided the number of systems is known.
-%
-%%%%%
-%
-%\subsection{Data Association Problem}
-%\label{DAP}
-%
-%Consider the data association problem when two estimates are in very close proximity.
-%For this scenario, several assumptions are made.
-%First, we assume that we are only observing a single measurement from this close proximity region, i.e., we consider a single measurement at a time, not a scan of multiple measurements.
-%Secondly, we assume that this measurement originates from either one of the targets and not from other sources of clutter; the case when neither target is measured is not considered.
-%Like the JPDA, all targets are assumed to fall inside the validation gate.
-%Furthermore, the measurement origination between targets is random, unknown, and equally likely.
-In summary, for the event $\Theta_1$ as measuring target $1$ and the event $\Theta_2$ as measuring target $2$,
-%\begin{align}
-%P(\Theta_1\cup \Theta_2)&=1,\quad
-%P(\Theta_1\cap \Theta_2)=0,\\
-%P(\Theta_1)&=P(\Theta_2)=0.5.
-%\end{align}
-%
-%The only source of knowledge to associate data at time $k$ is the measurement $z_k$.
-%Using the estimate prediction $\hat x^-$, we define its prediction
-%\begin{align}
-%z_{i_{k|k-1}}=H_{i_k}\hat x_{i_k}^-
-%\end{align}
-%which is used to determine the innovation
-%\begin{align}
-%e_{i_k}=z_k-z_{i_{k|k-1}}
-%\end{align}
-%and its covariance
-%\begin{align}
-%S_{i_k}=H_{i_k}P_{i_k}^{-}H_{i_k}^T+R_{i_k}
-%\end{align}
-%which is used to determine the weighting of the measurement update.
-%
-%We can modify JPDAF weighting and update from~\cite{JPDAF1} and~\cite{TrackDataAssoc}.
-%The assumptions in this section are applied and significantly simplify the weighting calculation. For the $i$-th target, the weighting $\beta_i$ is reduced to (see proof in the appendix)
-%\begin{gather}
-%\beta_{i_k}=\frac{f_{i_k}}{\displaystyle\sum\limits_{y=1}^2 f_{y_k}} \\
-%f_{y_k}=\mathcal{N}[z_{k};\hat z_{i_{k|k-1}},S_{i_k}].
-%\end{gather}
-%where $\mathcal{N}[x;\mu,\Sigma]$ is the Gaussian probability density function of mean $\mu$ and covariance matrix $\Sigma$ evaluated at $x$.
-%
-%The filtering problem involves updating each state using this weighting.
-We wish to update the estimates of both systems on parallel filters; however, when these weightings are close, multiple estimates share a single measurement, causing the estimates to coalesce, which we wish to avoid.
-%
-%% Association Problem
-%
-%	% Assumptions on measurement
-%		% only one measurement is available at each time 
-%		% the case of no measurement is not included
-%		% measurement origination is random and unknown
-%		% these assumptions represent two targets in close proximity, such as crossing targets
-%
-%	% Data association event
-%		% there are only two events for data association
-%		% data association probability (beta) is defined as...
-%		% (Proper justification is needed, as this is one of the key assumptions)
-%		
-%	% Filtering problem
-%		% We wish to estimate states of both system
-%
-%%_______________________________________________%
-%
-%\section{Joint Probabilistic Data Association Filter}
-%\label{JPDAF}
-%
-%A typical JPDAF considers scans of many measurements and the possibility of clutter.
-Applying the assumptions from Section \ref{ProbDef} we can greatly reduce the measurement update terms to much more manipulatable forms (see appendix for proofs):
-%\begin{align}
-%\label{KalEst}
-%\hat x^+_{i_k}=&\hat x^-_{i_k}+\beta_{i_k}K_{i_k}e_{i_k}\\
-%\label{KalCov}
-%P^+_{i_k}=&P^-_{i_k}-\beta_{i_k}K_{i_k}S_{i_k}K_{i_k}^T.
-%\end{align}
-%These equations are derived under the key assumption that the states of the systems are assumed to be Gaussian according to the most recent state estimate and covariance.
-%The gain $K_{i_k}$ is equivalent to a standard Kalman gain
-%\begin{align}
-%\label{KalGain}
-%K_{i_k}=P_{i_k}^-H_{i_k}^TS_{i_k}^{-1}
-%\end{align}
-%which serves to minimize $\tr{P^+_{i_k}}$ because here there is no measurement uncertainty when conditioned on event $\Theta_{i_k}$~\cite{TrackDataAssoc}.
-%However, Equation \ref{KalCov} only holds true for $K_{i_k}$ as defined in Equation \ref{KalGain}.
-%The general form of $P_{i_k}^+$ is defined as
-%\begin{equation}
-%\label{genPp}
-%P_{i_k}^{+}=(I-K_{i_k}H_{i_k})P_{i_k}^{-}(I-K_{i_k}H_{i_k})^T+K_{i_k}R_{i_k}K_{i_k}^T
-%\end{equation}
-%which may be manipulated with Equation \ref{KalGain} to form Equation \ref{KalCov} (without the $\beta_{i_k}$ term)~\cite{OptEst1}.
-%
-%	% The key assumption of JPDA: eqn (6-16) of the book
-%	% JPDA: see pages 164-166
-%		% Derive how the mean value of the state is updated 
-%		% Show that the gain becomes the same as the standard filter that minimizes tr[P]
-%		% Derive how the covariance is updated
-%
-%%_______________________________________________%
-%
-%\section{Coalescence Avoiding Optimal Joint Probabilistic Data Association Filter}% Or another name that you choose
-%\label{CAOJPDAF}
-%
-%This section describes the changes necessary to form the CAOJPDAF from the JPDAF.
-%We consider the changes to the measurement update at time $k$, so these subscripts will not be included in this section.
-%The goal of the CAOJPDAF is to minimize a weighted sum of the uncertainty and the similarity.
-%The filter is designed to slightly change each gain $K_{i_k}$ from that of the JPDAF when the estimates are in very close proximity to resist coalescence.
-%
-%% Derivation of a new filter
-%
-%	% Short description of the main idea: minimize the weighted sum of the uncertainty and the similarity
-%	
-%\subsection{Matusita's Similarity Measure}
-%
-%Two similarity measures are considered to determine an appropriate cost function.
-First, the Matusita's measure is considered because it ranges between 0 and 1.
-For the $m$-variate Gaussian distributions, the Matusita's measure can be decomposed into $Q$ and $R$
-%\begin{align}
-%\zeta=QR
-%\end{align}
-%such that
-%\begin{align}
-%\label{Q}
-%Q=&\frac{2^\frac{m}{2}|P_1^+|^{\frac{1}{4}}|P_2^+|^{\frac{1}{4}}}{|P_1^++P_2^+|^{\frac{1}{2}}}\\
-%\label{R}
-%R=&\exp{\left\{-\frac{1}{4}(\hat x^+_1-\hat x^+_2)^T(P_1^++P_2^+)^{-1}(\hat x^+_1-\hat x^+_2)\right\}}
-%\end{align}
-%where $Q$ measures the differences in the covariance matrices and $R$ measures the spread in the means~\cite{Coal_k}.
-%	% Introduce the definition of M-measure
-%	% Introduce the definition of Bhattacharyya distance
-%	
-%\subsection{Optimal Coalescence Avoidance}
-%
-%The choice of optimal cost involves two parts: posterior uncertainty minimization, much like the Kalman filter, and similarity minimization, based on the Matusita's measure.
-The total cost function is defined as
-%\begin{align}
-%J=J_P+cJ_S
-%\end{align}
-%where $J_P$ is the uncertainty cost, $J_S$ is the similarity cost, and $c$ is a positive constant for relative weighting.
-The cost function is shown by varying the first element of $K_1$ in Figure \ref{CostTrends}, although $c$ is chosen much larger for this figure than in practice to depict the effect of including a similarity cost rather than just uncertainty cost.
-%
-%\begin{figure}
-%\centerline{
-%\includegraphics[width=9cm]{CostTrends.png}
-%	}
-%\caption{The cost functions of target 1 are examined under the variation of a single element of $K_1$.
-The resulting cost function (black) is minimized slightly to the right of the uncertainty cost function (blue).
-Note that the total cost when the Kalman gain is applied does not lie on the total cost curve because other elements of this gain that are not optimal.
-In Kalman filtering applications, including the JPDAF, only the uncertainty cost function is minimized.}
-%\label{CostTrends}
-%\end{figure}
-%
-%The uncertainty cost $J_P$ is defined as
-%\begin{align}
-%J_P=\tr{P^+_1}+\tr{P^+_2}.
-%\end{align}
-%Note that $P^+_i$ must be valid for all $K_i$. Hence, from the general form of Equation \ref{genPp}, we seclude a $P_i^-$ term to take the form of Equation \ref{KalCov} to obtain
-%\begin{align}
-%\label{newPp}
-%P^+_i=&P^-_i-\beta_i[K_iH_iP_i^-+P_i^-H_i^TK_i^T\\
-%&-K_i(H_iP_i^-H_i^T+R_i)K_i^T].
-%\end{align}
-%
-%For the remaining equations of this section, the only derivatives taken are with respect to $K_1$ for the similarity between targets $1$ and $2$ for simplicity; taking derivatives with respect other $K_i$ values are identical except for subscripts.
-%
-%First,  using the property $\frac{\partial}{\partial X}\tr{XYX^T}=2XY$ for symmetric $Y$, we take the derivative
-%\begin{align}
-%\label{CostP}
-%\begin{split}
-%\frac{\partial J_{P,1}}{\partial K_{1}}&=2\beta_1\{{K_1(H_1P_1^-H_1^T+R_1)-P_1^-H_1^T}\}.
-%\end{split}
-%\end{align}
-%
-%To appropriately define the similarity cost, we only consider the $R$ term from Equation \ref{R} because coalescence is well known to occur when estimates are within close proximity.
-For simplicity, we define ${P^+_{12}}^{-1}$ as
-%\begin{align}
-%{P^+_{12}}^{-1}=(P_1^++P_2^+)^{-1}
-%\end{align}
-%and define define $u$ as
-%\begin{align}
-%\begin{split}
-%u=&(\hat x^+_1-\hat x^+_2)^T{P^+_{12}}^{-1}(\hat x^+_1-\hat x^+_2)\\
-%=&\{\hat x_1^{-T}{P^+_{12}}^{-1}\hat x_1^-
-%-2\hat x_1^{-T}{P^+_{12}}^{-1}\hat x_2^-\\
-%&+2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T{P^+_{12}}^{-1}K_1e_1
-%+\hat x_2^{-T}{P^+_{12}}^{-1}\hat x_2^-\\
-%&+2\beta_2(\hat x_2^{-}-\hat x_1^{-})^T{P^+_{12}}^{-1}K_2e_2
-%+\beta_1^2e_1^TK_1^T{P^+_{12}}^{-1}K_1e_1\\
-%&-2\beta_1\beta_2e_2^TK_2^T{P^+_{12}}^{-1}K_1e_1+\beta_2^2e_2^TK_2^T{P^+_{12}}^{-1}K_2e_2\}
-%\end{split}
-%\end{align}
-%which is the most reduced form to differentiate with respect to $K_1$ and $K_2$.
-%We generalize the similarity cost further to
-%\begin{align}
-%J_S=&\exp{\left\{-au\right\}}
-%\end{align}
-%for positive constant $a$, which serves to amplify or attenuate the function as well as alter its shape.
-%Defining $ij1\in \Re^{n\times m}$ such that $ij1(i,j)=1$ and $0$ elsewhere, we define $P^*_{12}$ as
-%\begin{align}
-%\begin{split}
-%P^*_{12,1}=&\frac{\partial {P^+_{12}}^{-1}}{\partial K_{1,ij}}\\
-%=&\beta_1{P_1^+}^{-1}[ij1H_1P_1^-+P_1^-H_1^Tij1^T\\
-%&-ij1(H_1P_1H_1^T+R_1)\\
-%&-K_1(H_1P_1^-H_1^T+R_1)ij1^T]{P_1^+}^{-1}
-%\end{split}
-%\end{align}
-%and using the chain rule, we take the derivative
-%\begin{align}
-%\frac{\partial J_{S,1}}{\partial K_{1,ij}}=-a\exp{\left\{-au\right\}}\frac{\partial u}{\partial K_{1,ij}}
-%\end{align}
-%where
-%\begin{align}
-%\begin{split}
-%\frac{\partial u}{\partial K_{1,ij}}=&\frac{\partial}{\partial K_{1,ij}}\left\{\hat x^+_1-\hat x^+_2)^T{P^+_{12}}^{-1}(\hat x^+_1-\hat x^+_2\right\}\\
-%=&\{\hat x_1^{-T}P^*_{12,1}\hat x_1^-
-%-2\hat x_1^{-T}P^*_{12,1}\hat x_2^-\\
-%&+2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T[P^*_{12,1}K_1+{P^+_{12}}^{-1}ij1]e_1\\
-%&+\hat x_2^{-T}P^*_{12,1}\hat x_2^-+2\beta_2(\hat x_2^{-}-\hat x_1^{-})^TP^*_{12,1}K_2e_2\\
-%&+\beta_1^2e_1^T[ij1^T{P^+_{12}}^{-1}K_1+K_1^TP^*_{12,1}K_1\\
-%&+K_1^T{P^+_{12}}^{-1}ij1]e_1-2\beta_1\beta_2e_2^TK_2^T[P^*_{12,1}K_1\\
-%&+{P^+_{12}}^{-1}ij1]e_1+\beta_2^2e_2^TK_2^TP^*_{12,1}K_2e_2\}
-%\end{split}
-%\end{align}
-%for $i=1:n$ and $j=1:m$ to obtain $\frac{\partial u}{\partial K_{1}}$.
-%
-%The weighted cost is minimized when $K_1$ and $K_2$ are chosen such that
-%\begin{align}
-%\frac{\partial J}{\partial K}=\frac{\partial J_{P,1}}{\partial K_{1}}+\frac{\partial J_{P,2}}{\partial K_{2}}
-%+c\left(\frac{\partial J_{S,1}}{\partial K_{1}}+\frac{\partial J_{S,2}}{\partial K_{2}}\right)=0.
-%\end{align}
-%The complexities of $(P^+_1+P^+_2)^{-1}$ prevents a closed-form solution for this equation; hence, we require a nonlinear equation solver.
-%The equations vary in computational cost depending on the level of precision of the nonlinear equation solver.
-%Note that ${P_1^+}^{-1}$, ${P_2^+}^{-1}$, and $(P^+_1+P^+_2)^{-1}$ are symmetric inverses and must only be calculated once each iteration of the solver.
-%Furthermore, this weighted cost may be readily extended to an arbitrary known number of targets provided the Matusita's measure is included between each estimate.
-%
-%The algorithm is completed by updating the state estimate and covariance.
-Equation \ref{KalEst} may be used for $\hat x^+_i$ because this equation does not require a Kalman gain; however we must use Equation \ref{newPp} to obtain $P^+_i$ because Equation \ref{KalCov} is only valid for a Kalman gain.
-%
-%	% Choice of optimal cost
-%		% the part of B-distance representing the difference of mean is used
-%		% the role of the weighting factor of c
-%		
-%	% Necessary conditions for optimality
-%		% Derivation of \deriv{J}{K}
-%		
-%	% Filter equations
-%		% Show why we can use the same expression for the mean value and the covariance of JPDA described in the previous section
-%		
-%%	\subsection{Suboptimal Coalescence Avoidance}
-%	
-%	% Linear approximation of the gain to obtain explicit solution
-%	
-%\section{Numerical Examples}
-%\label{NumRes}
-%	% Describe two body problem of orbital mechanics
-%	% Describe the measured values and the initial conditions
-%
-%	% Results of the perfect association with 100% confidence (imaginary case for comparison)
-%	% Results of JPDA (incorrect association after crossing)
-%	% Results of the proposed optimal filtering approach
-%	% Results of the proposed suboptimal filtering approach with linear approximation
-%
-%	% Discussion that the proposed approach is between JPDA and perfect association in terms of filtering performance
-%	% Show how much additional computational load is required compared with JPDA
-%
-%We present a satellite tracking example where some parameters are varied to evaluate the performance of the CAOJPDAF with the JPDAF.
-We consider the scenario when we observe two satellites in LEO with crossing trajectories wherein we receive range measurements of unknown origination.
-This problem is simplified to a two-body problem between each satellite and Earth, i.e., the of equations of motion for each orbit only involve force caused by the gravity between the Earth and the satellite and neglect all other forces.
-For simplicity, both orbits occupy the same plane, so cylindrical coordinates are used to define our state variable $x=\begin{bmatrix} r & \dot r & \theta & \dot \theta\end{bmatrix}^{T}$.
-We linearize the equations of motion and bring to the appropriate form of this paper
-%\begin{align}
-%A_{i_k}=
-%exp\left(\begin{bmatrix}
-%0 & 1 & 0 & 0 \\
-%\frac{2\mu}{r_{i_k}^3} +\dot\theta^2_{i_k} & 0 & 0 & 2r_{i_k}\dot \theta_{i_k} \\
-%0 & 0 & 0 & 1 \\
-%2r_{i_k}^{-2}\dot r_{i_k}\dot\theta_{i_k} & -2r^{-1}\dot\theta_{i_k} & 0 &  -2r_{i_k}^{-1}\dot r_{i_k}\\
-%\end{bmatrix}dt\right)
-%\end{align}
-%where $\mu=GM_{Earth}=3.986\times10^5 km^3/s^2$ and $H=\begin{bmatrix} 1 & 0 & 0 & 0\end{bmatrix}$.
-%We assume we have very nice priori knowledge from other measurement devices; there was no need to simulate large uncertainty when the paths are initially sufficiently far apart because this paper is concerned with the coalescence when the targets are close together, so the state and covariance will have been determined by simulated measurements while the estimates were still spread significantly apart.
-%The initial estimate of the states are designed to occur shortly before the two trajectories cross paths near $r_{ref}=7000km$ from the center of the earth; for the time to complete a circular orbit at this radius, $t_{ref}=2\pi \sqrt{\frac{r_{ref}^3}{\mu}}$, we can solve for a reference circular orbit $X_{ref}=\begin{bmatrix} r_{ref} & 0 & 0 & \frac{2\pi}{t_{ref}}\end{bmatrix}^T$, then perturb this orbit slightly by $X_{pert}=\begin{bmatrix} 0.40 & 0 & 0 & -1.5\times10^{-5}\end{bmatrix}^T$.
-We can easily choose the initial conditions to be $X_{1_0}= X_{ref}+X_{pert}$ and $X_{2_0}= X_{ref}-X_{pert}$.
-%
-%We provide identical noisy range measurements for the EKF, the JPDAF, and the CAOJPDAF algorithms in Figure \ref{AllAlgorithms}.
-The meaning of the EKF in this example is that there exists some way to perfectly associate all measurements with the correct targets.
-This scenario is, of course, unrealistic but serves as a nice comparison for data association algorithms.
-%For the CAOJPDAF, we choose $a=0.02$ and $c=5\times10^{-5}$.
-We propagate the nonlinear equations of motion using \emph{ode45} in Matlab, and produce a measurement update every $0.1sec$ with an equal likelihood that its origination is from either target.
-%
-%\begin{figure}
-%\centerline{
-%	\subfigure[EKF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{EKF.png}\hspace*{0.07\textwidth}}
-%	}
-%\centerline{
-%	\subfigure[JPDAF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{JPDAF.png}\hspace*{0.07\textwidth}}
-%	}
-%\centerline{
-%	\subfigure[CAOJPDAF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{CAOJPDAF.png}\hspace*{0.05\textwidth}}
-%	}
-%\caption{Target 1 is represented in blue and target 2 is represented in red.
-A solid line represents an estimate and a dotted line represents the truth.The JPDAF is subject to coalescence of the neighboring track, causing the estimations to switch trajectories.
-The CAOJPDAF is subject to much less coalescence and hence does not exhibit this problem.}
-%\label{AllAlgorithms}
-%\end{figure}
-%
-%The results to these numerical simulations demonstrate very important aspects of the algorithms.
-Most importantly, the JPDAF is unable to maintain the correct trajectories of the estimates because the tracks coalesce so much that correctly labeling the estimates becomes very difficult; when measurements from either target are sufficiently spread after the crossing, the JPDAF decision of trajectories that the estimates follow becomes ambiguous as the estimates become ambiguous.
-Secondly, within $20 sec$ of crossing ($201$ measurements), the number of incorrect associated (weighted less than $50\%$ likelihood of association) for the JPDAF was $92$ whereas for the CAOJPDAF this number was $70$.
-Comparing the performance of all three algorithms, the radial RMS error in the same window was calculated for the EKF as $1.4425\times10^{-05} km$, for the JPDAF $2.3014\times10^{-05} km$, and for the CAOJPDAF $2.3858\times10^{-05} km$.
-This metric demonstrates the uncertainty of association degrades the estimates as the EKF performs significantly better than algorithms where measurement origination is unknown.
-The JPDAF is performs slightly better than the CAOJPDAF with this metric, but this is a small drawback when compared with the ability to maintain correct tracks after crossing.
-The largest drawback of the CAOJPDAF is the computational cost; running the simulations on the same computer, the JPDAF only required $7.518sec$ to run the program whereas the CAOJPDAF required $25.785sec$ to run.
-%
-%%_______________________________________________%
-%
-%\section{Conclusions and Future Work}
-%\label{ConclusionFutureWork}
-%We explore a new approach to removing coalescence by choosing measurement update gains that minimize weighted cost with respect to posterior uncertainty in addition to similarity.
-The proximity similarity portion of the Matusita's measure is chosen to evaluate the similarity because of its range, scalar result, and association with coalescence.
-We take the derivative of the similarity cost for both estimates, yielding a result with no closed form solution, but involving expressions with low computational cost that can be solved with a nonlinear equation solver.
-In practice, a small similarity cost relative to the uncertainty cost yields best results; hence, the algorithm requires less time if the initial guesses of the Kalman gains are chosen.
-%
-%We validate the theory with numerical simulations.
-The CAOJPDAF shares measurements in the same way as the JPDAF, but the tracks tend to maintain separation; the CAOJPDAF does not prevent the estimates from crossing as the targets in the trajectory are designed to do.
-The major drawback of the CAOJPDAF is the added computation.
-However, the results of all three simulated algorithms (EKF, JPDAF, and CAOJPDAF) are nearly identical when the estimates are not near the crossing region (even if on the wrong trajectory).
-Hence, the CAOJPDAF can be turned on only during these close proximity regions, then turned off when unnecessary to be replaced by a less computationally expensive algorithm.
-%
-%The CAOJPDAF can be extended to a variety of multi-target tracking applications when targets share measurements due to close proximity.
-One example, very close to that presented in Section \ref{NumRes}, is keeping logs of satellites using phased-array radar systems.
-In cases where satellites have similar measurement predictions over long periods of time, satellite tracking becomes much more difficult.
-Implementing a simple JPDAF can serve to coalesce the estimates, which can cause the estimates of the satellites to switch trajectories as shown in this paper.
-Therefore, including the CAOJPDAF under certain proximity conditions can expand our ability to track satellites.
-The CAOJPDAF has tracking applications for space, aerial, and underwater vehicles in close proximity and identification of individual agents in multi-agent systems.
-%
-%Much like the Kalman filter, the CAOJPDAF can be effective for systems with well-known Gaussian distributions of state, sensor noise, and process.
-This paper has assumed that measurements come from a single target of interest at a single time and not clutter.
-Future work may involve expanding the capabilities of the algorithm.
-We can consider the measurements as a scan, or so that multiple measurements, including those from external sources of clutter, may belong to multiple targets at any time step as a JPDAF does without the assumptions of this paper.
-A general form of this algorithm could be derived for these conditions.
-Furthermore we can explore other techniques to measure similarity.
-By only using the proximity portion Matusita's measure, we are assuming the covariances of the estimates are close, which is not always accurate in all applications, and another future direction of this work.
-
 \begin{appendix}
@@
 
+\subsection{Proofs}
+
+\paragraph{Derivation of the Association Probability}
+The association probability at time step $k$ (this subscript will be omitted from this appendix) is given as follows from~\cite[Section 9.3]{TrackDataAssoc}
+\begin{align}
+\label{beta_all_terms}
+\beta_i\equiv P(\Theta_i|z)=&\frac{1}{c}\frac{\phi!}{V}\displaystyle\prod_{j=1}^{m_k} f_{i_j}^{\tau_j}\displaystyle\prod_{i=1}^{T} P(\Theta_i)^{\delta_i}(1-P(\Theta_i))^{1-\delta_i}
+\end{align}
+where $c$ denotes a normalizing constant, $\phi$ denotes the number of unassociated measurements, $\delta$ denotes the binary target detection indicator, $V$ denotes an arbitrary volume that all targets fall inside, $j$ denotes an individual meausurement in a single update, $m_k$ denotes the total number of measurements that may be associated with a target, $f_{i}$ is the Gaussian pdf of target $i$, $\tau$ denotes the measurement association indicator, $T$ denotes the total number of targets, and $P(\Theta_i)$ denotes the probability of detecting target $i$.
 We incorporate all of the assumptions from Section \ref{DAP} to simplify the weighting term for the estimates.
-We consider time step $k$ for the following proofs so subscripts omitted in this appendix.
-We define ${\bf{\Theta}}\in[\Theta_1, \Theta_2]$, as all possible outcomes, namely that target $1$ is measured or target $2$ is measured from $m_k=1$ total measurements.
-Because all events ${\bf{\Theta}}$ fall inside an arbitrary volume $V$, or are subject to being measured, we define the measurement association indicator $\tau=1$, meaning that measurement $k$ is associated with exactly one of the two targets.
+Let ${\bf{\Theta}}\in[\Theta_1, \Theta_2]$, as both possible outcomes, namely that target $1$ is measured or target $2$ is measured.
+Each estimate may receive at most a single measurement, so $m_k=1$.
+Because all events ${\bf{\Theta}}$ fall inside an arbitrary volume $V$, or are subject to being measured, the measurement association indicator $\tau=1$, meaning that measurement $k$ is associated with exactly one of the two targets.
 Hence, the number of unassociated measurements $\phi({\bf{\Theta}})=0$ because the event that neither target is measured $\Theta_0\notin{\bf{\Theta}}$.
-We define the binary target detection indicator $\delta_i(\Theta_i)$, which indicates the association of any measurement with target $i$; hence, $\delta_i(\Theta_i)=1$ if measurement $k$ is associated with target $i$ and $\delta_i(\Theta_i)=0$ otherwise.
-We also consider equal likelihood of detection $P(\Theta_i)$ for each measurement; we can generalize the number of estimates to $y$ such that $P(\Theta_i)=\frac{1}{y}$ for equal likelihood of event $\Theta_i$.
-From~\cite[Section 9.3]{TrackDataAssoc}, we determine the weighting for $T$ possible targets to be
-\begin{align}
-\beta_i\equiv P(\Theta_i|z)=&\frac{1}{c}\frac{\phi!}{V}\displaystyle\prod_{j=1}^{m_k} f_{i_j}^{\tau_j}\displaystyle\prod_{i=1}^{T} P(\Theta_i)^{\delta_i}(1-P(\Theta_i))^{1-\delta_i}\\
-=&\frac{1}{cV}f_{i}\frac{1}{y}(1-\frac1y)^{T-1}
-\end{align}
-for normalizing constant $c$ and Gaussian pdf $f_{i}$. Note that $T-1$ targets are not associated and only one target is associated. Hence, $c$ is defined as
+The binary target detection indicator is defined as $\delta_i(\Theta_i)=1$ if measurement $k$ is associated with target $i$ and $\delta_i(\Theta_i)=0$ otherwise.
+The detection probabilities are generalized to $y$ estimates such that $P(\Theta_i)=\frac{1}{y}$ for equal likelihood of event $\Theta_i$.
+Therefore, Equation \ref{beta_all_terms} reduces to
+\begin{align}
+\beta_i=&\frac{1}{cV}f_{i}\frac{1}{y}(1-\frac1y)^{T-1}.
+\end{align}
+Solving for $c$,
 \begin{align}
@@
 \end{align}
-Therefore, we solve for the simplified weighting
+Hence, the simplified weighting considering the assumptions of this paper is
 \begin{align}
@@
 \end{align}
-which is much simpler after assumptions.
+which is equivalent to Equation \ref{eqn:betaik} for two estimates.
+
+\paragraph{Derivation of Posteriori Covariance}
 
@@
 \end{align}
- , so this portion of the covariance Let $\tilde{P}$ be spread in the innovations term on each target.
-Because only one innovation is calculated per target, there is no spread in this term; hence $\tilde{P}=0$.
-Note that his term would be nonzero in the JDPAF without the single measurement assumption.
-Let $P_{i}^C$ be the posterior covariance of target $i$ conditioned on event $\Theta_i$.
-For any $K_i$ from ~\cite[Section 4.2]{OptEst1}:
+Because only one innovation is calculated per target, there is no spread in the innovations term on each target, so this portion of the covariance $\tilde{P}=0$, which would normally be included in the JDPAF without these assumptions.
+The posterior covariance of target $i$ conditioned on event $\Theta_i$ is defined for any $K_i$ from ~\cite[Section 4.2]{OptEst1}:
 \begin{align}
@@
 \end{align}
-We define $\beta_{0_i}$ as the probability that target $i$ is not measured; with only one measurement considered each time step; hence, $\beta_{0_i}=1-\beta_i$.
+Let $\beta_{0_i}$ be the probability that target $i$ is not measured; with only one measurement considered each time step; hence, $\beta_{0_i}=1-\beta_i$.
 Again from~\cite[Section 6.4]{TrackDataAssoc},
@@
 \bibliographystyle{IEEEtran}
+
 \end{document}
-
-%Many techniques to avoid track coalescence involve the concept that when the exact nearest neighbor probabilistic data association (ENNPDA) is applied, which associates a measurement with a strong decision to the most likely estimate, the estimates are insensitive to track coalescence~\cite{Coal1}.
-%By pruning the tracks, or designing a set of rules to determine high likelihood scenarios, an algorithm choice or weighing scheme is applied to the measurement updates.
-%For example in~\cite{Fitzgerald}, the algorithm involves a set of rules based on the data received with respect to the estimates.
-%A special logic is introduced that determines how weighting of the updates is defined or if a new track should be considered.
-%In some cases, it is possible for an incorrectly associated estimate measurement may receive full weighting in the update.
-%Other criteria or extensions of this approach are detailed in~\cite{Coal_d},~\cite{Coal_e}, and~\cite{Coal_c}, which can be very effective, but information may be lost when measurements are omitted from updating an estimates with small probabilities of association because the pruning rule prevented these association possibilities.
 

### 2018-06-09 Conference/ECC 2014/ECC14_OLD.tex
added Conference folder

\documentclass[letterpaper, 10pt, conference]{ieeeconf}
\IEEEoverridecommandlockouts \overrideIEEEmargins
\usepackage{amsmath,amssymb,url}
\usepackage{graphicx,subfigure,tikz}
\usepackage{color}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\ave}[1]{\mathrm{E}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand\circled[1]{%
  \tikz[baseline=(C.base)]\node[draw,circle,inner sep=0.5pt](C) {#1};\!
}

\title{\LARGE \bf
%Recursive Kalman-Mahanobis JPDAF Minimization to avoid Coalescence of Estimates in Close Proximity
Optimal Joint Probabilistic Data Association Filter Avoiding Track Coalescence}
\author{Evan Kaufman and Taeyoung Lee
 \thanks{Evan Kaufman and Taeyoung Lee are with Department of Aerospace Engineering, George Washington University, Washington, DC. Email: {\tt\footnotesize \{evankaufman, tylee\}@gwu.edu }}
}

\newcommand{\EditTL}[1]{{\color{red}\protect #1}}
\renewcommand{\EditTL}[1]{{\protect #1}}

\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{remark}{Remark}

\begin{document}
\allowdisplaybreaks

\maketitle \thispagestyle{empty} \pagestyle{empty}

\begin{abstract}
We consider the tracking problem of two objects in close proximity where the measurement originations are unknown. The joint probabilistic data association filter (JPDAF) is useful for real-time applications because the measurement updates can be recursively and effectively applied based on the likelihood of association. However, when close neighboring tracks share measurements, the JPDAF tends to causes estimates to coalesce. Therefore we propose a Kalman-Mahalanobis minimizing JPDAF (KMMJPDAF) that chooses the gains of the measurement update to minimize both the posterior uncertainty and a measure of similarity based from the Mahalanobis distance. The algorithm is derived and explained, involving a higher computational cost that the JPDAF, but not high enough to prevent the possiblity real-time implementation. We discuss how to tune the parameters and when the algorithm is best used. Numerical simulations are used to evaluate the performance of the algorithms and conclusions are drawn on these results.
\end{abstract}

%_______________________________________________%

\section{Introduction}
Data association is a heavily studied field with many applications involving estimation involving measurements that are not necessarily originating from a single object.
We consider applications wherein multiple objects of interest enter the field of view of a sensor; however, the measurement origin is unknown.
To properly track these objects, we require estimation techniques to associate objects with measurements.
When the objects are in very close proximity, the association uncertainty grows, and an effective algorithm is necessary to maintain close estimations of the objects.

A variety of data association techniques with varying performance.
Some approaches such as multiple hypothesis tracking (MHT) and probability hypothesis density (PHD) techniques involve a high computational load, especially when objects are in close proximity, making real-time implementation more difficult~\cite{MHT1},~\cite{PHD1}~\cite{PHD2}.
Many other techniques may involve the Kalman filter (KF) or extended KF (EKF), which are well studied and provide a framework for recursive data association techniques that is much easier to implement in real-time.
A common heuristic approach is known as the nearest neighbor filter (NNF), which is a simple algorithm that only updates estimates for measurements with the closest square Mahalnobis distance, updating other objects only with their predictions~\cite{NN2}.
This is considered a \emph{hard decision} in~\cite{JPDAF1}, or that every measurement update receives full weighting; if the decision is incorrect, these measurement updates can result in detrimental effects to the associations of the estimates, so this technique is avoided in the analysis of this paper.
Alternatively, some approaches calculate the minimum mean square error (MMSE) of the estmates and their associated uncertainty.
The probabilistic data association (PDA) technique is effective in tracking a single object because it updates with a \emph{soft decision}; in particular, this approach updates provides a weighting from a likelihood of measurement association.

The joint PDA (JPDA) is an extension of the PDA to incorporate multiple objects~\cite{JPDAF1},~\cite{TrackDataAssoc}.
The JPDA filter (JPDAF) is particularly advantageous in multi-object scenarios because when objects share measurements, the soft decisions prevent loss of information, which occurs with hard decisions.
However, when neighboring objects share measurements, the estimates are subject to coalescence.

Some approaches to avoid track coalescence involve the concept that when the exact nearest neighbor probabilistic data association (ENNPDA) is applied, which associates a measurement with a strong decision to the most likely estimate, the estimates are insensitive to track coalescence [Blom 2000].
By pruning the tracks, or designing a set of rules to determine high likelihood scenarios, a weighing scheme is applied to the measurement updates.
For example in [Fitzgerald], the algorithm involves a set of rules based on the data received with respect to the estimates.
A special logic is introduced that determines how weighting of the updates is defined or if a new track should be considered.
In some cases, an incorrectly associated estimate measurement may receive full weighting in the update.
Other criteria for choosing hard and soft decisions to avoid track coalescence are detailed in~\cite{Coal_d} and~\cite{Coal_e}, which may be effective, but are rather heuristic by nature and risk assigning an incorrect hard decision.
Some other approaches involve coupling joint states. Explain more and validate last citation...

We propose a novel approach to removing coalescence by minimizing a similarity index of the estimates.
Instead of choosing a different means of weighting the measurement updates or coupling their distributions, we design an optimal recursive estimation algorithm that chooses gains to compensate for the effects of coalescence.
The algorithm derived in this paper can be applied in real-time without the heuristic nature, coupling, or the risk of hard decisions from the current coalescence avoidance approaches.

The paper is organized as follows: the problem is formally presented in Section \ref{ProbDef}.
The optimal gain derivation and algorithm description is presented in Section \ref{KMM}.
We discuss how to implement the algorithm and how to appropriately weight the minimized cost function in section \ref{AlgImpTuning}.
We consider a satellite tracking example of this technique in Section \ref{NumRes}.
Finally, we discuss results and applications for the KMMJPDAF in Section \ref{ConclusionFutureWork}.

%A considerable amount of work has been done to solve the data association problem. Approaches vary in their performance, including the portion of correct associations, the ability to maintain the correct tracks, and the computational cost. In this section, we examine some of the current techniques with respect to tracking two objects in close proximity and propose a new approach to solve this problem.
%
%The multiple hypothesis tracking (MHT) approach considers multiple hypotheses for how the data may be associated for multiple tracks~\cite{MHT1}. As new information becomes available, hypotheses are evaluated, in some cases removed, and predicted again. However in complicated problems, the number hypothesis may explode. One approach to remove this issue from the close proximity problem is known as track pruning, where incompatible tracks are highlighted and only the most likely tracks are considered. This approach is complex and involves a high computational cost, so implementing this approach in real-time is very difficult. Furthermore according to~\cite{PHD1}, MHT is subject to poor results with objects in close proximity.
%
%The probability hypothesis density (PHD) process is a propagation of the first moment of a joint distribution~\cite{PHD2}. A typical PHD considers a sub-area of interest where peaks from the integral of the PHD are considered objects. In many cases, the PHD filter is hindered by computational complexity, growing exponentially with the number of objects and further increasing when the possibility that objects may enter or leave the sub-area considered. An approach described in~\cite{PHD1} maintains a tracker for each object of interest. Each tracker accounts for a priori estimates of state and dynamics to predict the probability densities for the next time step. The process is complex and involves separating the sub-spaces into cells wherein likelihoods of each object are used to define peaks and dummy peaks, where a single track or multiple tracks exist, respectively. This approach can be more computationally expensive than other recursive approaches as the algorithm is among the most complex and the resolution (defined by size of the cells where peaks lie) will require modification as objects become close.
%
%The nearest neighbor filter (NNF) is a heuristic approach that calculates the magnitude between the measurement and its prediction~\cite{JPDAF1}; the state with smallest square Mahalanobis distance of the innovation and its covariance receives a filtered update with the measurement while all other states of interest are updated with only their predictions~\cite{NN2}; this is referred to as a \emph{hard decision}~\cite{JPDAF1}. This approach adds a simple step to the EKF at the cost of accuracy as the probability of association is not explicitly calculated.
%
%The JPDAF is an extension of the probabilistic data association filter (PDAF) described in~\cite{JPDAF1} and~\cite{TrackDataAssoc}, which calculates the probabilities that measurements belong to an individual estimate. The JPDAF is designed to track multiple objects of interest. The estimates are updated with a \emph{soft decision}, or by a weighting factor based from the probability of association. However, when more than one object of interest share measurements, the objects are subject to a measurement updates that draw the estimates together, or coalesce the estimates.
%
%There are a few approaches to remove coalescence from the JPDAF. Much like the MHT, some approaches iinvolve track pruning and are detailed in~\cite{Coal1},~\cite{Coal2},~\cite{Coal_d}, and~\cite{Coal_e}, but this is computationally complex and can suffer when implementing the approach recursively. Other techniques such as those described in~\cite{Coal_c},~\cite{Coal_g},~\cite{Coal_i}, and~\cite{Coal_j} involve assigning hard decisions. The criteria and performance of these techniques vary, but each eliminate coalescence by never sharing a measurement, which can yield fully weighted incorrect associations that may cause detrimental effects such as estimates switching trajectories of crossing tracks. To decrease this problem, an arbitrary scaling factor is applied in~\cite{Coal_f} wherein the more likely association is given more weight as the less likely association weightings are decreased, yielding a hybrid technique between the NNF and the JPDAF. This approach may yield improved results with a heuristic weighting scheme but results vary with the application and expected data.
%
%Therefore, we propose a systematic approach that minimizes a weighted cost of uncertainty and similarity. We design the KMMJPDAF to minimize the posterior uncertainty much like the Kalman filter in addition to evaluating the Mahalanobis distance involved with updating the states. The paper is organized as follows: the problem is formally presented in Section \ref{ProbDef}. The optimal gain derivation and algorithm description is presented in Section \ref{KMM}. We discuss how to implement the algorithm and how to appropriately weight the minimized cost function in section \ref{AlgImpTuning}. We consider a satellite tracking example of this technique in Section \ref{NumRes}. Finally, we discuss results and applications for the KMMJPDAF in Section \ref{ConclusionFutureWork}.

%_______________________________________________%

\section{Problem Definition}
\label{ProbDef}
The objects of interest are assumed to have nonlinear dynamics and each have Gaussian distributed estimates. We evaluate objects at time $k$ and linearize the system dynamics with respect to time $k-1$. We define the JPDAF algorithm such that we update the state as measurements are received, i.e. we only consider a single measurement rather than a scan of measurements. Furthermore, this research is concerned with the two estimates in very close proximity; hence we assume that the measurement originations are from a single object of interest and not from other sources of clutter.

\subsection{Time Update} We denote an estimate of state with a $\hat{.}$, and the estimates before and after the measurement update with a $.^-$ and $.^+$, respectively. We define $x\in \Re^{n\times1}$ as the state vector and $P\in\Re^{n\times n}$ as its covariance. The time update provides the predicted state and covariance
\begin{gather}
\hat{x}_{k}^{-}=A_{k-1}\hat{x}_{k-1} \\
P_{k}^{-}=A_{k-1}P_{k-1}A_{k-1}^T+Q_k.
\end{gather}
The covariance of process noise is captured in $Q_k$ and no control input is considered.

\subsection{Measurement Update} The measurement update requires an innovation
\begin{equation}
\label{innovation}
e_{k}=z_{k}-\hat z_{k|k-1}
\end{equation}
where $z_{k}\in\Re^{m\times1}$ is the measurement received and measurement prediction $\hat z_{k|k-1}=H_{k}\hat{x}_{k}^{-}$; $H_{k}$ maps $\hat{x}_{k}^{-}$ from the state space to the measurement space.% Check this line terminology
We simplify the equations for probability of association from~\cite{JPDAF1} and~\cite{TrackDataAssoc} to include only a single measurement that always belongs to either one of the two objects considered; hence the process of association is limited two mutually exclusive scenarios: we measure object 1 and not object 2 or we measure object 2 and not object 1. After some manipulation we can show for object $t$, the relative probability $\beta_t$ is simplified to
\begin{gather}
\beta_t=\frac{f_t}{\displaystyle\sum\limits_{\tau=1}^2 f_{\tau}} \\
f_t=\mathcal{N}[z_{k};\hat z_{k|k-1},S].
\end{gather}
The innovation covariance $S$ is defined as
\begin{equation}
S_k=H_{k}P_{k}^{-}H_{k}^T+R_k
\end{equation}
where $R_k$ is the covariance of the measurement. The gain to minimize the posterior uncertainty with cost function $J_P=\tr{P_{k+1}^{+}}$ is equivalent to the Kalman gain
\begin{equation}
\label{KalmanGain}
K_{k}=P_{k}^{-}H_{k}^TS_k^{-1}
\end{equation}
which is used to calculate the updated state and its covariance for each estimate
\begin{gather}
\label{xhatp}
\hat{x}_{k}^{+}=\hat{x}_{k}^{-}+\beta K_{k}e_{k} \\
\label{oldPp}
P_{k}^{+}=P^-_{k}-\beta K_{k}S_kK_{k}^{T}
\end{gather}
for the JPDAF, but not for the KMMJPDAF. For the KMMJPDAF we derive a new gain which is not compatible with Equation \ref{oldPp}, which is a manipulation involving Equation \ref{KalmanGain} of the original form (ignoring the $\beta$ term)
\begin{equation}
\label{genPp}
P_{k}^{+}=(I-K_{k}H_{k})P_{k}^{-}(I-K_{k}H_{k})^T+K_{k}R_kK_{k}^T
\end{equation}
from ~\cite{OptEst1}. Hence in order to properly update the posterior covariance when we include coalescence into the minimization, we manipulate Equation \ref{genPp} to involve relative weighting $\beta$ of the measurement update. Note that time step subscripts will be assumed at $k$ from here forward for simplicity. Obtaining a gain accounting for posterior uncertainty and coalescence takes the generalized form
\begin{equation}
\label{newPp}
P^+=P^--\beta[KHP^-+P^-H^TK^T-K(HP^-H^T+R)K^T].
\end{equation}
Note that if we insert the Kalman gain from Equation \ref{KalmanGain} into Equation \ref{newPp}, we can manipulate the result to Equation \ref{oldPp}.

%_______________________________________________%

\section{Kalman-Mahalanobis Optimization}
\label{KMM}
We derive the gain of the measurement update for object $t$ to minimize a weighted sum of the uncertainty and a similarity measure
\begin{equation} \label{CostDef}
J=J_P+cJ_S
\end{equation}
where $J_P$ is the cost associated with the posterior state covariance, $J_S$ is the cost associated with the similarity of the two distributions, and $c$ is a positive scalar constant for relative weighting. The following equations in this section will focus on object 1 but the derivations with respect to object 2 are indentical except with subscripts switched.

\subsection{Uncertainty Cost} The derivation of $J_P$ is a fairly straightforward adaptation of the Kalman filter from~\cite{OptEst1} with the included JPDAF weighting. Evaluating the trace of Equation \ref{newPp} for object 1,
\begin{align}
\begin{split}
J_{P,1}=&\tr{P_1^+} \\
=& \tr{P_1^-}+\beta_1(\tr{K_1(H_1P_1^-H_1^T+R_1)K_1^T}\\
&-2\tr{K_1H_1P_1^-}).
\end{split}
\end{align}
We can take its derivative $\frac{\partial J_{P,1}}{\partial K_{1}}$ using the property $\frac{\partial}{\partial X}\tr{XYX^T}=2XY$ for symmetric $Y$ to obtain
\begin{align}
\label{CostP}
\begin{split}
\frac{\partial J_{P,1}}{\partial K_{1}}&=2\beta_1\{{K_1(H_1P_1^-H_1^T+R_1)-P_1^-H_1^T}\}.
\end{split}
\end{align}

 \subsection{Similarity Cost} We choose the Mahalanobis distance to measure similarity because it yields a scalar result and taking its derivative with respect to measurement update gains is possible. We use $P_1^-$ as the covariace estimation of object 1 for the Mahalanobis distance with respect to the estimated mean of object $2$ can take the form
\begin{align}
\label{M1}
M_1=u_1^{\frac{1}{2}}
\end{align}
where
\begin{align} \label{u_1}
\begin{split}
u_1=&(\hat x_1^+-\hat x_2^+)^{T}{P_1^-}^{-1}(\hat x_1^+-\hat x_2^+)\\
=&\{\hat x_1^{-T}{P_1^-}^{-1}\hat x_1^-
-2\hat x_1^{-T}{P_1^-}^{-1}\hat x_2^-\\
&+2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T{P_1^-}^{-1}K_1e_1
+\hat x_2^{-T}{P_1^-}^{-1}\hat x_2^-\\
&+2\beta_2(\hat x_2^{-}-\hat x_1^{-})^T{P_1^-}^{-1}K_2e_2
+\beta_1^2e_1^TK_1^T{P_1^-}^{-1}K_1e_1\\
&-2\beta_1\beta_2e_2^TK_2^T{P_1^-}^{-1}K_1e_1+\beta_2^2e_2^TK_2^T{P_1^-}^{-1}K_2e_2\}
\end{split}
\end{align}
which is the most reduced form to differentiate with respect to $K_1$ and $K_2$. However, the Mahalanobis distance is an increasing measure as similarity decreases. Therefore we wish to minimize $M_1^{-2a}$ for $a>0$. Furthermore as the Mahalanobis distance is a scalar quantity, we can use the property $\frac{\partial g(u_1)}{\partial K_1}=\frac{\partial g(u_1)}{\partial u_1}\frac{\partial u_1}{\partial K_1}$ such that $J_{C,1}=M_1^{-2a}=g(u_1)=u_1^{-a}$. We take the derivative of the similarity measure using $u_1=f(K_1,K_2)$ from Equations \ref{M1} and \ref{u_1} as well as the properties $\frac{\partial}{\partial X} \tr{AXB}=A^TB^T$ and $\frac{\partial}{\partial X}\tr{AXBX^TC}=A^TC^TXB^T+CAXB$ to obtain, after some manipulation,
\begin{align}
\begin{split}
\label{CostS}
\frac{\partial J_{S,1}}{\partial K_1}=&-au_1^{-(a+1)}\frac{\partial u}{\partial K_1}\\
=&2a\beta_1u_1^{-(a+1)}\{
{P_1^-}^{-1}(\hat x_2^--\hat x_1^-)\\
&-\beta_1{P_1^-}^{-1}K_1e_1
+\beta_2{P_1^-}^{-1}K_2e_2
\}e_1^T.
\end{split}
\end{align}

We absorb the constants from Equations \ref{CostP} and \ref{CostS} into $c$ from Equation \ref{CostDef} and use a nonlinear equation solver for the optimal gains $K_1$ and $K_2$ in the joint equations
\begin{align}
\label{Opt1}
\begin{split}
\frac{\partial J_{1}}{\partial K_{1}}=&K_1(H_1P_1^-H_1^T+R_1)-P_1^-H_1^T\\
&+cu_1^{-(a+1)}\{
{P_1^-}^{-1}(\hat x_2^--\hat x_1^-)\\
&-\beta_1{P_1^-}^{-1}K_1e_1
+\beta_2{P_1^-}^{-1}K_2e_2
\}e_1^T\\
=&0_{n\times m}
\end{split}
\end{align}
and similarly
\begin{align}
\label{Opt2}
\begin{split}
\frac{\partial J_{2}}{\partial K_{2}}=&K_2(H_2P_2^-H_2^T+R_2)-P_2^-H_2^T\\
&+cu_2^{-(a+1)}\{
{P_2^-}^{-1}(\hat x_1^--\hat x_2^-)\\
&-\beta_2{P_2^-}^{-1}K_2e_2
+\beta_1{P_2^-}^{-1}K_1e_1
\}e_2^T\\
=&0_{n\times m}.
\end{split}
\end{align}
These equations vary in computational cost depending on the level of precision of the nonlinear equation solver. Note that ${P_1^-}^{-1}$ and ${P_2^-}^{-1}$ are symmetric inverses and must only be calculated once each.

%_______________________________________________%

\section{Algorithm Implementation and Tuning}
\label{AlgImpTuning}
This algorithm requires considerably more computational cost than the JPDAF alone; hence, we only implement this algorithm when the likelihood of coalescence is high. For example, if the association of a measurement belongs primarily to a single estimate, the KMMJPDAF and JPDAF yield similar results so we choose JPDAF because of its closed form solution and hence lower computational cost. Therefore depending on the application, we determine thresholds such that beyond some spacing between the a priori estimates we default to the JPDAF algorithm for speed.

We must also examine the shape of the cost function. Consider Figure \ref{CostTrends}, which contains a variation of a single element of $K_1$ from an example in Section \ref{NumRes} and plots both cost functions of object 1. The uncertainty cost minimizes at a single point. However, the similarity cost contains a maximum. If we incorrectly choose $K_1$ and $K_2$ to maximize $J_S$, this would serve to increase coalescence because similarity would be increasing at a maximum rate due only to a measurement update.

\begin{figure}
\centerline{
\includegraphics[width=9cm]{CostTrends.png}
	}
\caption{The cost functions of object 1 ($J_{P,1}$, $cJ_{C,1}$, and $J_{1}$) are examined under the variation of a single element of $K_1$. The resulting cost function (black) is minimized slightly to the left of the uncertainty cost function (blue). In Kalman filtering applications, including the JPDAF, only the uncertainty cost function is minimized.}
\label{CostTrends}
\end{figure}

There are a few approaches to avoid this detrimental scenario. First, we can choose $a$ and $c$ properly. Decreasing $a$ serves to amplify and flatten out $J_S$. Once a desired cost function shape as achieved, we can choose $c$ to bring $J_S$ to an appropriate scaling. In practice, the JPDAF is a very effective algorithm for estimation of multiple objects, which only relies on $J_P$, so $J_S$ is chosen to be very small in comparison only to remove coalescence from the estimates (Figure \ref{CostTrends} shows $cJ_{S,1}$ much larger in magnitude than in practice for illustration). Choice of these parameters leads to trade-offs: if $a$ is decreased along with $c$, the similarity cost varies less, decreasing the beneficial effects of the KMMJPDAF over the JPDAF. However, when $a$ is increased (and $c$ is increased to compensate), we may incorrectly solve for the similarity maximizing point.

We explore two techniques to find an effective minimization. First, we choose the initial guess of $K_1$ and $K_2$ to be that of the EKF, which is already located near the minimized point of uncertainty. Secondly, we only implement the KMMJPDAF when the a priori estimates are sufficiently spaced apart. This way we can maintain the desirable shape of the cost function and implement the KMMJPDAF for the majority of the region where coalescence poses an issue. Therefore, we have both minimum and maximum estimate spacing thresholds for which KMMJPDAF is appropriate.

%_______________________________________________%

\section{Numerical Results}
\label{NumRes}
We present a satellite tracking example where some parameters are varied to evaluate the performance of the KMMJPDAF with the JPDAF. We consider the scenario when we observe two satellites in LEO with crossing trajectories wherein we receive range measurements of unknown origination. For simplicity, both orbits occupy the same plane, so cylindrical coordinates are used to define our state variable $x=\begin{bmatrix} r & \dot r & \theta & \dot \theta\end{bmatrix}^{T}$. We only consider gravity, so we linearize the equations of motion and bring to the appropriate form
\begin{align}
A_{k-1}=
exp\left(\begin{bmatrix}
0 & 1 & 0 & 0 \\
\frac{2\mu}{r^3} +\dot{\theta}^2 & 0 & 0 & 2r\dot{\theta} \\
0 & 0 & 0 & 1 \\
2r^{-2}\dot{r}\dot{\theta} & -2r^{-1}\dot{\theta} & 0 &  -2r^{-1}\dot{r}\\
\end{bmatrix}dt\right)
\end{align}
where $\mu=GM_{Earth}=3.986\times10^5 km^3/s^2$ and $H=\begin{bmatrix} 1 & 0 & 0 & 0\end{bmatrix}$. We assume identical dynamics of the satellites (although this is not a requirement). We assume we have a priori knowlege from other measurement devices that provide an estimate of the state and covariance shortly before the two trajectories cross paths near $7000km$ from the center of the earth. We provide identical noisy range measurements to both the JPDAF and the KMMJPDAF algorithms. For the KMMJPDAF, we choose $a=0.01$ and $c=3\times10^{-8}$. We propagate the nonlinear equations of motion using \emph{ode45} in Matlab, and produce a measurement update every $0.1sec$ with an equal likelihood that its origination is from either object. For the results below, we refer to the KMMJPDAF algorithm as the KMMJPDAF when $0.05km\leq\norm{\hat z_1-\hat z_2}\leq0.4km$ and the JPDAF elsewhere due to the observations discussed in Section \ref{AlgImpTuning}.

\begin{figure}
\centerline{
	\subfigure[JPDAF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{Case1JPDAF.png}\hspace*{0.07\textwidth}}
	}
\centerline{
	\subfigure[KMMJPDAF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{Case1KMMJPDAF.png}\hspace*{0.05\textwidth}}
	}
\caption{The estimation of object 1 (solid blue) is overlaid with the true trajectory of object 1 (dashed blue) and the estimation of object 2 (solid red) is overlaid with the true trajectory of object 2 (dashed red). When the JPDAF is able to properly track the trajectories, the KMMJPDAF yields similar results.}	
\label{Case1}
\end{figure}

The first results (Figure \ref{Case1}) shows a result wherein both algorithms are able to maintain the correct tracks. In fact, the results are very similar with two noteworthy differences. The first is that within $10 sec$ of crossing ($101$ measurements), the number of incorrect associated (weighted less than $50\%$ likelihood of association) for the JPDAF was $64$ whereas for the KMMJPDAF the number was $58$, but this is a minor difference in the metric. The second difference is that the JPDAF completed in $4.212sec$ while the KMMJPDAF ran in $6.8952sec$ on the same computer. For real-time implementation, either algorithm is capable of receiving these simple range measurements as described above, but when measurements are more complicated or are received at a faster rate, there can be cases where the performance of the JPDAF computational time makes the KMMJPDAF an inferior algorithm.

\begin{figure}
\centerline{
	\subfigure[JPDAF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{Case2JPDAF.png}\hspace*{0.07\textwidth}}
	}
\centerline{
	\subfigure[KMMJPDAF]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{Case2KMMJPDAF.png}\hspace*{0.05\textwidth}}
	}
\caption{The JPDAF is subject to coalescence of the neighboring track, causing the estimations to switch trajectories. The KMMJPDAF is subject to less coalescence and hence does not exhibit this problem. Refer to Figure \ref{Case1} for the legend.}
\label{Case2}
\end{figure}

The noise of the measurements are increased in second example (Figure \ref{Case2}), and this causes the JPDAF to permanently lose the correct association after the crossing. The optimization of the KMMJPDAF compensates for the coalescence of the tracks, and hence correct track maintenance is achieved properly with this algorithm. The result of this example does not imply that the KMMJPDAF is impervious the the effects of coalescence; rather including similarity in the minimization serves to remove some of the coalescence from measurement updates. For the case presented here, choosing the KMMJPDAF rather than the JPDAF increases the computational time from $4.1964sec$ to $6.8018sec$ to maintain the estimates with the correct trajectories.

\begin{figure}
\centerline{
	\subfigure[Parameter change: $a=1$]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{a1.png}\hspace*{0.07\textwidth}}
	}
\centerline{
	\subfigure[Parameter change: $c=1\times10^{-7}$]{\hspace*{0.07\textwidth}\includegraphics[width=9cm]{c1e7.png}\hspace*{0.05\textwidth}}
	}
\caption{These results show detrimental effects of choosing $a$ too great, attenuating the similarity cost function, yielding numerical results wherein coalescence causes poor estimation or choosing $c$ too great, increasing the similarity cost so much that the estimates diverge before they can ever cross. Refer to Figure \ref{Case1} for the legend.}
\label{changeParameters}
\end{figure}

We now use the last example to demonstrate the importance of choosing the parameters correctly for the effective implementation of the KMMJPDAF. In Figure \ref{changeParameters}.(a), we increase $a$ such that $a=1$, which serves to attenuate the similarity cost function. As a result, the gains are chosen to minimize uncertainty with greater weight much like the JPDAF. This results in increasing coalescence, yielding a poorer estimation of the states. In this case, the estimates are drawn very close together and diverge incorrectly.

We consider the same example but this time increase $c$ such that $c=1\times10^{-7}$, shown in Figure \ref{changeParameters}.(b). Increasing $c$ increases the similarity cost function proportionally according to Equation \ref{CostDef}. Contrary to the last example, the gains are chosen to decrease similarity of the estimates so much that the measurement updates serve to separate the estimates when the estimates become close. We experience the opposite effect with increasing $c$ than increasing $a$, but choosing either parameter poorly can lead to estimates switching trajectories.

%_______________________________________________%

\section{Conclusions and Future Work}
\label{ConclusionFutureWork}
We explore a new approach to removing coalescence by choosing measurement update gains that minimize with respect to similarity in addition to posterior uncertainty. The Mahalanobis distance is chosen as the measure of similarity because of its simple form and scalar result. We take the derivative of the similarity cost for both estimates, yielding a result with no closed form solution, but involving expressions with low computational cost that can be solved with a nonlinear equation solver. We discuss the importance of initial prediction for gains as well as thresholds for which the KMMJPDAF is appropriate.

We validate the theory with numerical simulations of a satellite tracking problem. We observe that if the JPDAF can properly handle tracking the objects of interest, this algorithm should be used due to its low computational cost. However, we show an example where the coalescence of the JPDAF causes the estimates to switch to incorrect trajectories. The KMMJPDAF is more robust to this issue and the estimates maintain their correct trajectories in the example. These results are only valid if the parameters of the cost function are chosen well; decreasing or increasing the similarity cost has the effect of coalescing or repelling the estimates, respectively, which can cause the estimates to lose the correct trajectories.

The KMMJPDAF can be extended to a variety of multi-object tracking applications when objects share measurements due to close proximity. One example, very close to that presented in Section \ref{NumRes}, is keeping logs of satellites using phased-array radar systems. In cases where satellites have similar measurement predictions over long periods of time, satellite tracking becomes much more difficult. Implementing a simple JPDAF can serve to coalesce the estimates, which can cause the estimates of the satellites to switch trajectories as shown in this paper. Therefore, including the KMMJPDAF under certain proximitiy conditions can expand our ability to track satellites. The KMMJPDAF has tracking applications for space, aerial, and underwater vehicles in close proximity and identification of individual agents in multi-agent systems.

Much like the Kalman filter, the KMMJPDAF can be effective for systems with well-known Gaussian distributions of state, sensor noise, and process. This paper has assumed that measurements come from a single object of interest at a single time and not clutter. Future work may involve expanding the capabilities of the algorithm. We can consider the measurements as a scan, or so that multiple measurements, including those from external sources of clutter, may belong to multiple objects at any timestep as a JPDAF does without the assumptions of this paper. A general form of this algorithm could be derived for these conditions.

Furthermore we can explore other techniques to measure similarity. The Mahanobis distance assumes the covariances of the estimates are close, which is not always accurate in all applications. Furthermore, we approximate the covariance in the Mahanobis distance with the a priori covariance, which changes with the measurement update. Hence, there are a few directions to improve the measure of posterior similarity.

\bibliography{BibMaster}
\bibliographystyle{IEEEtran}

\end{document}

### 2018-06-09 Conference/ECC 2014/ECC14__Final.tex
added Conference folder

--- 2018-06-09 Conference/ECC 2014/ECC14_Lee.tex
@@
 Optimal Joint Probabilistic Data Association Filter\\ Avoiding Coalescence in Close Proximity }
-\author{Evan Kaufman and Taeyoung Lee$^*$
+\author{Evan Kaufman, Thomas Alan Lovell, and Taeyoung Lee$^*$
  \thanks{Evan Kaufman and Taeyoung Lee are with Department of Aerospace Engineering, George Washington University, Washington, DC. Email: {\tt\footnotesize \{evankaufman, tylee\}@gwu.edu }}
+\thanks{Thomas Alan Lovell is a Research Aerospace Engineer at the Space Vehicles Directorate, Air Force Research Lab (AFRL), Albuquerque, NM. Email: {\tt\footnotesize afrl.rvsv@kirtland.af.mil }}
 \thanks{$^*$This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}}
@@
 \begin{abstract}
-This paper deals with an estimation problem where a known number of targets in close proximity are observed but the measurement originations are uncertain.
-It is well known that joint probabilistic data association filters (JPDAF) are effective in handing uncertain measurement originations with clutter, but they are prone to estimation coalescence, particularly when close neighboring targets share measurements.
+This paper deals with an estimation problem where a known number of objects in close proximity are observed but the measurement originations are uncertain.
+It is well known that joint probabilistic data association filters (JPDAF) are effective in handing uncertain measurement originations with clutter, but they are prone to estimation coalescence, particularly when close neighboring objects share measurements.
 This paper proposes a Coalescence Avoiding Optimal JPDAF (C-JPDAF) that minimizes the weighted sum of the posterior uncertainty and a measure of similarity between estimated probability densities.
-The proposed approach has simpler structures than other coalescence avoiding approaches based on pruning, while exhibiting excellent filtering performance for targets in close proximity with crossing tracks.
+The proposed approach has simpler structures than other coalescence avoiding approaches based on pruning, while exhibiting excellent filtering performance for objects in close proximity with crossing tracks.
 These are illustrated by a numerical example of two satellites on crossing orbits around the Earth.
@@
 
-Data association is a heavily studied field with many applications in estimation when measurements are not necessarily originating from a single target.
-We consider applications wherein multiple targets of interest enter the field of view of a sensor; however, the measurement origin is unknown.
-To properly track these targets, we require estimation techniques to associate targets with measurements.
-When the targets are in very close proximity, the association uncertainty grows, and an effective algorithm is necessary to maintain close estimations of the targets.
+Data association is a heavily studied field with many applications in estimation when measurements are not necessarily originating from a single object.
+We consider applications wherein multiple objects of interest enter the field of view of a sensor; however, the measurement origin is unknown.
+To properly track these objects, we require estimation techniques to associate objects with measurements.
+When the objects are in very close proximity, the association uncertainty grows, and an effective algorithm is necessary to maintain close estimations of the objects.
 
 A variety of data association techniques are considered for close-proximity trajectories with varying performance.
-Some approaches such as multiple hypothesis tracking (MHT) and probability hypothesis density (PHD) techniques involve a high computational load, especially when targets are in close proximity, making real-time implementation more difficult~\cite{MHT1},~\cite{PHD1}, and~\cite{PHD2}.
+Some approaches such as multiple hypothesis tracking (MHT) and probability hypothesis density (PHD) techniques involve a high computational load, especially when objects are in close proximity, making real-time implementation more difficult~\cite{MHT1},~\cite{PHD1}, and~\cite{PHD2}.
 Many other techniques may involve the Kalman filter (KF) or extended KF (EKF), which are well studied and provide a framework for recursive data association techniques that are much easier to implement in real-time.
-A common heuristic approach is known as the nearest neighbor filter (NNF), which is a simple algorithm that only updates estimates for measurements with the closest square Mahalnobis distance, updating other targets only with their predictions~\cite{NN2}.
+A common heuristic approach is known as the nearest neighbor filter (NNF), which is a simple algorithm that only updates estimates for measurements with the closest square Mahalnobis distance, updating other objects only with their predictions~\cite{NN2}.
 This is considered a \emph{hard decision} in~\cite{JPDAF1}, or that every measurement update receives full weighting; if the decision is incorrect, these measurement updates can result in detrimental effects to the associations of the estimates, so this technique is avoided in the analysis of this paper.
 Alternatively, some approaches calculate the minimum mean square error (MMSE) of the estimates and their associated uncertainty.
-The probabilistic data association (PDA) technique is effective in tracking a single target because it updates with a \emph{soft decision}; in particular, this approach updates provides a weighting from a likelihood of measurement association.
+The probabilistic data association (PDA) technique is effective in tracking a single object because it updates with a \emph{soft decision}; in particular, this approach updates provides a weighting from a likelihood of measurement association.
 
-The joint PDA (JPDA) is an extension of the PDA to incorporate multiple targets~\cite{JPDAF1},~\cite{TrackDataAssoc}.
-The JPDA filter (JPDAF) is particularly advantageous in multi-target scenarios because when targets share measurements, the soft decisions prevent loss of information, which occurs with hard decisions.
-However, when neighboring targets share measurements, the estimates are subject to coalescence.
+The joint PDA (JPDA) is an extension of the PDA to incorporate multiple objects~\cite{JPDAF1},~\cite{TrackDataAssoc}.
+The JPDA filter (JPDAF) is particularly advantageous in multi-object scenarios because when objects share measurements, the soft decisions prevent loss of information, which occurs with hard decisions.
+However, when neighboring objects share measurements, the estimates are subject to coalescence.
 
@@
 
-% Old introduction:
-
-%\section{Introduction}
-%
-%Data association has been widely studied with various applications in estimation of a single or multiple targets with measurements that are not necessarily originating from correct targets in order.
-To properly track multiple targets with measurement origin uncertainties, measured data should be properly associated to targets with certain probabilities.
-%When the targets are in very close proximity, the association uncertainty grows, and an effective algorithm is required to maintain close estimations of the targets.
-Otherwise, estimators tend to coalesce neighboring tracks.
-%
-%%This paper considers estimation wherein multiple targets of interest enter the field of view of a sensor; however, the measurement origin is unknown. 
-%
-%A variety of data association techniques with varying performance have been studied.
-A common heuristic approach is known as the nearest neighbor filter (NNF), where measurements with the closest distance are only applied~\cite{NN2}.
-This is considered a \emph{hard decision} in~\cite{JPDAF1}, as every selected measurement update receives full weighting.
-Therefore, if the decision is incorrect, these measurement updates can result in detrimental effects to the associations of the estimates.
-Alternatively, there are approaches that calculate the minimum mean square error (MMSE) of the estimates and their associated uncertainty.
-The probabilistic data association (PDA) technique is effective in tracking a single target because it updates with a \emph{soft decision}; in particular, this approach updates provides a weighting from a likelihood of measurement association.
-Other approaches include multiple hypothesis tracking (MHT) and probability hypothesis density (PHD) techniques~\cite{MHT1,PHD1,PHD2}.
-%
-%%Many other techniques may involve the Kalman filter (KF) or extended KF (EKF), which are well studied and provide a framework for recursive data association techniques that are much easier to implement in real-time.
-%
-%
-%The joint PDA (JPDA) is an extension of the PDA to incorporate multiple targets~\cite{JPDAF1},~\cite{TrackDataAssoc}.
-%The JPDA filter (JPDAF) is particularly advantageous in multi-target scenarios because when targets share measurements, the soft decisions prevent loss of information that occurs with hard decisions.
-As such, it is well known that JPDAF are effective in handling data clutter or missed detection.
-However, when the targets are in very close proximity, the association uncertainty grows significantly, thereby degrading the filtering performance.
-Furthermore, the estimated probability distributions tend to be coalesced especially if the paths of neighboring targets cross with each other.
-%
-%%neighboring targets share measurements, the estimates are subject to coalescence.
-%
-%Many techniques to avoid track coalescence involve the concept that when the exact nearest neighbor probabilistic data association (ENNPDA) is applied, which associates a measurement with a strong decision to the most likely estimate, the estimates are insensitive to track coalescence~\cite{Coal1}.
-By pruning the tracks, or designing a set of rules to determine high likelihood scenarios, an algorithm choice or weighing scheme is applied to the measurement updates.
-%For example in~\cite{Fitzgerald}, the algorithm involves a set of rules based on the data received with respect to the estimates.
-A special logic is introduced that determines how weighting of the updates is defined or if a new track should be considered.
-In some cases, it is possible for an incorrectly associated estimate measurement may receive full weighting in the update.
-Other criteria or extensions of this approach are detailed in~\cite{Coal_d},~\cite{Coal_e}, and~\cite{Coal_c}, which can be very effective, but information may be lost when measurements are omitted from updating an estimates with small probabilities of association because the pruning rule prevented these association possibilities.
-%
-%This paper is focused on optimal joint probabilistic data association filter that avoids coalescence in close proximity.
-We propose a novel approach to remove coalescence by minimizing a similarity index of the estimates.
-More explicitly, the gain of the proposed estimator is chosen such that a weighted sum of the size of uncertainties and a similarity measure between estimated states is minimized at each update.
-This yields an optimal recursive estimation algorithm that compensates for the effects of coalescence systematically.
-The proposed approach has simpler structures compared with pruning-based approaches, and it is computationally efficient.
-As it is still based on probabilistic data association, it avoids the heuristic nature, loss of information, or the risk of other coalescence avoidance approaches based on hard-decision.
-%
-%The paper is organized as follows: the dynamics of target and the measurement models are described in Section \ref{ProbDef}.
-The proposed Coalescence Avoiding Optimal JPDAF (C-JPDAF) are presented at Sections \ref{JPDAF}, and the optimal gain that minimizes a weighted cost of uncertainty and similarity is derived in Section \ref{C-JPDAF}.
-%These are illustrated by a numerical example for estimating two satellites in crossing orbits in Section \ref{NumRes}, followed by conclusions in Section \ref{ConclusionFutureWork}.
-
-%_______________________________________________%
-
 \section{Data Association Filtering Problem Formulation}
@@
 
-This section describes the dynamics of targets and the measurement model considered in this paper, and a filtering problem is formulated.
+This section describes the dynamics of objects and the measurement model considered in this paper, and a filtering problem is formulated.
 
 % Equations of motion
-\subsection{Dynamics of Targets and Measurements}
+\subsection{Dynamics of Objects and Measurements}
 
-The dynamics of each target is modeled as a linear stochastic discrete-time system. The equations of motion for the $i$-th system are described by
+The dynamics of each object is modeled as a linear stochastic discrete-time system belonging to the set $S$ such that
+\begin{align}
+S=\left\{i=1,2,...n\right\}
+\end{align}
+where $n$ is the number of objects.
+The equations of motion for the $i$-th system are described by
 \begin{align}
@@
 
-To demonstrate the main idea of the proposed data association filter concisely, it is assumed that there are two targets, i.e., $i\in\{1,2\}$.
-But, the results of this paper are readily generalized to arbitrary number of systems provided the number of targets is known.
+To demonstrate the main idea of the proposed data association filter concisely, it is assumed that there are two objects, i.e., $i\in\{1,2\}$.
+But, the results of this paper are readily generalized to arbitrary number of systems provided the number of objects is known.
 
@@
 
-The following assumptions are made on the measurement origination, considering the case where two target are in close proximity.
+The following assumptions are made on the measurement origination, considering the case where two object are in close proximity.
 First, it is assumed that there is only a single measurement at each time in the region of close proximity.
-Secondly, we assume that this measurement originates from either one of the targets and it is not from other sources such as clutter.
-Also, the case when neither target is measured is not considered.
-Like the JPDA, all targets are assumed to fall inside the validation gate.
-The measurement origination between targets is random, unknown, and equally likely.
+Secondly, we assume that this measurement originates from either one of the objects and it is not from other sources such as clutter.
+Also, the case when neither object is measured is not considered.
+Like the JPDA, all objects are assumed to fall inside the validation gate.
+The measurement origination between objects is random, unknown, and equally likely.
 
-More explicitly, let $\Theta_i$ be the event that the measurement originates from the $i$-th target, for $i\in\{1,2\}$. According to the above assumptions, we have
+More explicitly, let $\Theta_i$ be the event that the measurement originates from the $i$-th object, for $i\in\{1,2\}$. According to the above assumptions, we have
 \begin{align}
@@
 P(\Theta_1\cap \Theta_2)=0,\\
-P(\Theta_1)&=P(\Theta_2)=0.5.\label{eqn:MeaOrig}
+P(\Theta_1)&=P(\Theta_2)=0.5\label{eqn:MeaOrig}
 \end{align}
-
-%\EditTL{
-%The only source of knowledge to associate data at time $k$ is the measurement $z_k$.
-%Using the estimate prediction $\hat x^-$, we define its prediction
-%\begin{align}
-%z_{i_{k|k-1}}=H_{i_k}\hat x_{i_k}^-
-%\end{align}
-%which is used to determine the innovation
-%\begin{align}
-%e_{i_k}=z_k-z_{i_{k|k-1}}
-%\end{align}
-%and its covariance
-%\begin{align}
-%S_{i_k}=H_{i_k}P_{i_k}^{-}H_{i_k}^T+R_{i_k}
-%\end{align}
-%which is used to determine the weighting of the measurement update.
-%
-%We can modify JPDAF weighting and update from~\cite{JPDAF1} and~\cite{TrackDataAssoc}.
-%The assumptions in this section are applied and significantly simplify the weighting calculation. For the $i$-th target, the weighting $\beta_i$ is reduced to (see proof in the appendix)
-%\begin{gather}
-%\beta_{i_k}=\frac{f_{i_k}}{\displaystyle\sum\limits_{y=1}^2 f_{y_k}} \\
-%f_{y_k}=\mathcal{N}[z_{k};\hat z_{i_{k|k-1}},S_{i_k}].
-%\end{gather}
-%where $\mathcal{N}[x;\mu,\Sigma]$ is the Gaussian probability density function of mean $\mu$ and covariance matrix $\Sigma$ evaluated at $x$.}
+where the equal probability of detection is due to their close proximity.
 
@@
 
-A filtering problem is formulated as recursively updating the estimated state of each target and its covariance, namely $(\hat x_{i_k}, P_{i_k})$ based on the above assumptions given by \refeqn{xkp}--\refeqn{MeaOrig}.
-In particular, we wish to avoid track coalescence when two targets are in close proximity.
-
-%The filtering problem involves updating each state using this weighting.
-We wish to update the estimates of both systems on parallel filters; however, when these weightings are close, multiple estimates share a single measurement, causing the estimates to coalesce, which we wish to avoid.
-
-% Association Problem
-
-	% Assumptions on measurement
-		% only one measurement is available at each time 
-		% the case of no measurement is not included
-		% measurement origination is random and unknown
-		% these assumptions represent two targets in close proximity, such as crossing targets
-
-	% Data association event
-		% there are only two events for data association
-		% data association probability (beta) is defined as...
-		% (Proper justification is needed, as this is one of the key assumptions)
-
-	% Filtering problem
-		% We wish to estimate states of both system
-
-%_______________________________________________%
+A filtering problem is formulated as recursively updating the estimated state of each object and its covariance, namely $(\hat x_{i_k}, P_{i_k})$ based on the above assumptions given by \refeqn{xkp}--\refeqn{MeaOrig}.
+In particular, we wish to avoid track coalescence when two objects are in close proximity.
 
@@
 
-Joint probabilistic data association filter is a suboptimal Bayesian filter for multiple targets where the probabilities of association are computed for the latest set of measurements only~\cite{TrackDataAssoc}.
+Joint probabilistic data association filter is a suboptimal Bayesian filter for multiple objects where the probabilities of association are computed for the latest set of measurements only~\cite{TrackDataAssoc}.
 This section summarizes the approaches of JPDAF.
@@
 Similar to Kalman filters, the JPDAF is composed of a flow update and a measurement update.
-During the flow update (or prediction step), the estimated state $\hat x_{i_k}$ and its covariance $P_{i_k}$ of each target are updated as in the standard filter as follows:
+During the flow update (or prediction step), the estimated state $\hat x_{i_k}$ and its covariance $P_{i_k}$ of each object are updated as in the standard filter as follows:
 \begin{gather}
@@
 Next, we consider the measurement update step.
-Let  the priori value of a variable and its posteriori value be denoted by the superscript $-$ and $+$, respectively.
-For a given measurement $z_{k}$, its measurement residual or innovation for the $i$-th target is given by
+Let  the a priori value of a variable and its a posteriori value be denoted by the superscript $-$ and $+$, respectively.
+For a given measurement $z_{k}$, its measurement residual or innovation for the $i$-th object is given by
 \begin{align*}
@@
 
-The resulting posteriori covariance is given by
+The resulting a posteriori covariance is given by
 \begin{align}
@@
 
-The joint probabilistic data association filter described in the previous section is effective in handling clutter or missed detection of multiple targets.
-However, it is prone to coalescing estimates, especially with targets in close-proximity.
+The joint probabilistic data association filter described in the previous section is effective in handling clutter or missed detection of multiple objects.
+However, it is prone to coalescing estimates, especially with objects in close-proximity.
 Here, we present a coalescence-avoiding optimal JPDAF, hereafter C-JPDAF for short, to avoid coalescence of the estimates.
@@
 
-For two probability density functions, $f_1(x)$ and $f_2(x)$, Matusita's measure is defined as
+For two Gaussian probability density functions, $f_1(x)$ and $f_2(x)$, Matusita's measure is defined as
 \begin{align*}
-\zeta = \int \sqrt{f_1(x)f_2(x) dx},
+\zeta = \int \sqrt{f_1(x)f_2(x)}dx,
 \end{align*}
@@
 
-Using the fact that the posteriori states $\hat x_1^+$ and $\hat x_2^+$ are Gaussian, Matusita's measure between two targets can be written as
+From~\cite{Coal_k}, using the assumption that the a posteriori states $\hat x_1^+$ and $\hat x_2^+$ are Gaussian, the Matusita's measure between two objects can be written as
 \begin{align}
@@
 \label{Q}
-\zeta_1=&\frac{2^\frac{n}{2}|P_1^+|^{\frac{1}{4}}|P_2^+|^{\frac{1}{4}}}{|P_1^++P_2^+|^{\frac{1}{2}}},\\
+\zeta_1=&\frac{2^\frac{n}{2}|P_1^+|^{\frac{1}{4}}|P_2^+|^{\frac{1}{4}}}{|P_1^++P_2^+|^{\frac{1}{2}}},
+\end{align}
+\begin{align}
 \label{R}
@@
 The increased uncertainty at each step is rewarded by reducing the overall estimation error drastically by maintaining correct data association throughout crossing tracks in close proximity.}
-%The cost functions of target 1 are examined under the variation of a single element of $K_1$.
+%The cost functions of object 1 are examined under the variation of a single element of $K_1$.
 The resulting cost function (black) is minimized slightly to the right of the uncertainty cost function (blue).
@@
 
-More explicitly, the uncertainty cost $J_P$ is defined as the sum of the trace of each posteriori covariance:
+More explicitly, the uncertainty cost $J_P$ is defined as the sum of the trace of each a posteriori covariance:
 \begin{align}
@@
 
-%For the remaining equations of this section, the only derivatives taken are with respect to $K_1$ for the similarity between targets $1$ and $2$ for simplicity; taking derivatives with respect other $K_i$ values are identical except for subscripts.
+%For the remaining equations of this section, the only derivatives taken are with respect to $K_1$ for the similarity between objects $1$ and $2$ for simplicity; taking derivatives with respect other $K_i$ values are identical except for subscripts.
 
 Next, we derive the derivative of each cost.
-For simplicity, derivatives with respect to the first gain $K_1$ are presented, and derivatives with respect to $K_2$ can be obtained by symmetry.
+For simplicity, derivatives with respect to the first gain $K_1$ are presented, and derivatives with respect to $K_2$ can be obtained in the same way.
 By substituting (\ref{KalCov}) into \refeqn{JP} and using the property $\frac{\partial}{\partial X}\tr{XYX^T}=2XY$ for any matrix $X$ and any symmetric matrix $Y$, we obtain
@@
 Let $\Iij\in \Re^{n\times m}$ be a $n\times m$ matrix defined such that its $i,j$-th element is one and other elements are zero.
-The derivative of $(P_{12}^+)^{-1}$ with respect to the $i,j$-th element of $K_1$ is defined as $\mathcal{P}_{ij}\in\Re$:
+The derivative of $(P_{12}^+)^{-1}$ with respect to the $i,j$-th element of $K_1$ is defined as $\mathcal{P}_{ij}\in\Re^{m\times n}$:
 %such that $ij1(i,j)=1$ and $0$ elsewhere, we define $P^*_{12}$ as
@@
 
-%Furthermore, this weighted cost may be readily extended to an arbitrary known number of targets provided the Matusita's measure is included between each estimate.
+%Furthermore, this weighted cost may be readily extended to an arbitrary known number of objects provided the Matusita's measure is included between each estimate.
 
@@
 
-The dynamics of each satellite is modeled as a two-body problem between each satellite and Earth, i.e., the of equations of motion for each orbit only involve force caused by the gravity between the Earth and the satellite and neglect all other forces.
-For simplicity, both orbits occupy the same plane, so cylindrical coordinates are used to define our state variable $x=\begin{bmatrix} r & \dot r & \theta & \dot \theta\end{bmatrix}^{T}$.
+The dynamics of each satellite is modeled as a two-body problem between each satellite and Earth, i.e., the equations of motion for each orbit only involve force caused by the gravity between the Earth and the satellite and neglect all other forces.
+For simplicity, both orbits occupy the same plane, so polar coordinates are used to define our state variable $x=\begin{bmatrix} r & \dot r & \theta & \dot \theta\end{bmatrix}^{T}$.
 The resulting equations of motion of each satellite are given by
@@
 
-The initial estimate of the states are designed to occur shortly before the two trajectories cross paths near $r_{ref}=7000km$ from the center of the earth; for the time to complete a circular orbit at this radius, $t_{ref}=2\pi \sqrt{\frac{r_{ref}^3}{\mu}}$, we can solve for a reference circular orbit $X_{ref}=\begin{bmatrix} r_{ref} & 0 & 0 & \frac{2\pi}{t_{ref}}\end{bmatrix}^T$, then perturb this orbit slightly by $X_{pert}=\begin{bmatrix} 0.40 & 0 & 0 & -1.5\times10^{-5}\end{bmatrix}^T$.
+The initial estimates of the states are designed to occur shortly before the two trajectories cross paths near $r_{ref}=7000km$ from the center of the earth; for the time to complete a circular orbit at this radius, $t_{ref}=2\pi \sqrt{\frac{r_{ref}^3}{\mu}}$, we can solve for a reference circular orbit $X_{ref}=\begin{bmatrix} r_{ref} & 0 & 0 & \frac{2\pi}{t_{ref}}\end{bmatrix}^T$, then perturb this orbit slightly by $X_{pert}=\begin{bmatrix} 0.40 & 0 & 0 & -1.5\times10^{-5}\end{bmatrix}^T$.
 We can easily choose the initial conditions to be $X_{1_0}= X_{ref}+X_{pert}$ and $X_{2_0}= X_{ref}-X_{pert}$.
-The initial estimates are assumed to be relatively accurate: there is no need to simulate large uncertainty when the paths are initially sufficiently far apart because this paper is concerned with the coalescence when the targets are close together, so the state and covariance will have been determined by simulated measurements while the estimates were still spread significantly apart; hence for both estimates, the initial covariance was chosen to be $P_0=diag(\begin{bmatrix}
+The initial estimates are assumed to be relatively accurate: there is no need to simulate large uncertainty when the paths are initially sufficiently far apart because this paper is concerned with the coalescence when the objects are close together, so the state and covariance will have been determined by simulated measurements while the estimates were still spread significantly apart; hence for both estimates, the initial covariance was chosen to be $P_0=diag(\begin{bmatrix}
 2.5\times10^{-4} & 2.5\times10^{-6} & 3.9\times10^{-8} & 3.9\times10^{-8} \end{bmatrix})$.
-The process noise was chosen to be equivalent for both estimate such that $Q=diag(\begin{bmatrix}
+The process noise was chosen to be equivalent for both estimates such that $Q=diag(\begin{bmatrix}
 2.5\times10^{-5} & 2.5\times10^{-7} & 3.9\times10^{-9}& 3.9\times10^{-9}\end{bmatrix})$ and $R=\begin{bmatrix}
@@
 Three filtering approaches are applied, namely extended Kalman filter (EKF), JPDAF, and C-JPDAF.
-For the EKF, it is assumed that the correct measurement origination is know with certainty, i.e., all of the measurements are associated to the correct target.
+For the EKF, it is assumed that the correct measurement origination is know with certainty, i.e., all of the measurements are associated to the correct object.
 This serves as an ideal reference case to be compared with JPDAF and C-JPDAF.
 For the C-JPDAF, the weighting parameters of the cost are chosen as $a=0.02$ and $c=5\times10^{-5}$.
-We propagate the nonlinear equations of motion numerically using Matlab, and we generate a measurement at every $0.1$ seconds with an equal likelihood that its origination is from either target such that identical noisy measurements are used in all three algorithms.
+We propagate the nonlinear equations of motion numerically using Matlab, and we generate a measurement at every $0.1$ seconds with an equal likelihood that its origination is from either object such that all three algorithms are subject to identical noisy measurements.
+The estimates are propagated with their nonlinear equations of motion; the data association and filtering algorithms use the linearized form.
 
@@
 	}
-\caption{Target 1 is represented in blue and target 2 is represented in red.
+\caption{Object 1 is represented in blue and object 2 is represented in red.
 A solid line represents an estimate and a dotted line represents the truth.
@@
 
-The corresponding results for each of  EKF, JPDAF, and C-JPDAF algorithms are illustrated in Figure \ref{AllAlgorithms}. 
+The corresponding results for each of the EKF, JPDAF, and C-JPDAF algorithms are illustrated in Figure \ref{AllAlgorithms}. 
 First, EKF exhibits good estimation performance as there is no error or uncertainty in data association.
@@
 When $t\geq 70$ seconds, the estimated states are incorrectly crossed.
-This illustrates undesirable performance of JPDAF especially for crossing targets.
+This illustrates undesirable performance of JPDAF especially for crossing objects.
 
-The proposed C-JPDAF does not exhibit the incorrect, crossed estimation of JPDA, and the estimated states follows each of the correct trajectories throughout the crossing orbital trajectories.
+The proposed C-JPDAF does not exhibit the incorrect, crossed estimation of JPDA, and the estimated states follow each of the correct trajectories throughout the crossing orbital trajectories.
 The resulting RMS errors for the orbital radius of each method are summarized at Table \ref{tab:RMS}.
@@
 
-%First, JPDAF is unable to maintain the correct trajectories of the estimates because the tracks coalesce so much that %correctly labeling the estimates becomes very difficult; when measurements from either target are sufficiently spread after the crossing, the JPDAF decision of trajectories that the estimates follow becomes ambiguous as the estimates become ambiguous.
+%First, JPDAF is unable to maintain the correct trajectories of the estimates because the tracks coalesce so much that %correctly labeling the estimates becomes very difficult; when measurements from either object are sufficiently spread after the crossing, the JPDAF decision of trajectories that the estimates follow becomes ambiguous as the estimates become ambiguous.
 
@@
 \label{ConclusionFutureWork}
-We explore a new approach to reduce coalescence of close proximity multi-target tracking.
+We explore a new approach to reduce coalescence of close proximity multi-object tracking.
 The estimator gains are chosen such that a weighted cost with respect to posterior uncertainty and similarity is minimized.
@@
 Compared with other techniques to avoid coalescence, the proposed approach has simpler structures.
-Future works include generalizing C-JPDAF for arbitrary number of targets with clutter, and improving the computational efficiencies.
+Future works include generalizing C-JPDAF for arbitrary number of objects with clutter, and improving the computational efficiencies.
 
@@
 %We validate the theory with numerical simulations.
-The C-JPDAF shares measurements in the same way as the JPDAF, but the tracks tend to maintain separation; the C-JPDAF does not prevent the estimates from crossing as the targets in the trajectory are designed to do.
+The C-JPDAF shares measurements in the same way as the JPDAF, but the tracks tend to maintain separation; the C-JPDAF does not prevent the estimates from crossing as the objects in the trajectory are designed to do.
 The major drawback of the C-JPDAF is the added computation.
@@
 
-%The C-JPDAF can be extended to a variety of multi-target tracking applications when targets share measurements due to close proximity.
+%The C-JPDAF can be extended to a variety of multi-object tracking applications when objects share measurements due to close proximity.
 One example, very close to that presented in Section \ref{NumRes}, is keeping logs of satellites using phased-array radar systems.
@@
 %Much like the Kalman filter, the C-JPDAF can be effective for systems with well-known Gaussian distributions of state, sensor noise, and process.
-This paper has assumed that measurements come from a single target of interest at a single time and not clutter.
+This paper has assumed that measurements come from a single object of interest at a single time and not clutter.
 Future work may involve expanding the capabilities of the algorithm.
-We can consider the measurements as a scan, or so that multiple measurements, including those from external sources of clutter, may belong to multiple targets at any time step as a JPDAF does without the assumptions of this paper.
+We can consider the measurements as a scan, or so that multiple measurements, including those from external sources of clutter, may belong to multiple objects at any time step as a JPDAF does without the assumptions of this paper.
 A general form of this algorithm could be derived for these conditions.
@@
 
-\subsection{Proofs}
-
 \paragraph{Derivation of the Association Probability}
@@
 \end{align}
-where $c$ denotes a normalizing constant, $\phi$ denotes the number of unassociated measurements, $\delta$ denotes the binary target detection indicator, $V$ denotes an arbitrary volume that all targets fall inside, $j$ denotes an individual meausurement in a single update, $m_k$ denotes the total number of measurements that may be associated with a target, $f_{i}$ is the Gaussian pdf of target $i$, $\tau$ denotes the measurement association indicator, $T$ denotes the total number of targets, and $P(\Theta_i)$ denotes the probability of detecting target $i$.
+where $c$ denotes a normalizing constant, $\phi$ denotes the number of unassociated measurements, $\delta$ denotes the binary object detection indicator, $V$ denotes an arbitrary volume that all objects fall inside, $j$ denotes an individual meausurement in a single update, $m_k$ denotes the total number of measurements that may be associated with a object, $f_{i}$ is the Gaussian pdf of object $i$, $\tau$ denotes the measurement association indicator, $T$ denotes the total number of objects, and $P(\Theta_i)$ denotes the probability of detecting object $i$.
 We incorporate all of the assumptions from Section \ref{DAP} to simplify the weighting term for the estimates.
-Let ${\bf{\Theta}}\in[\Theta_1, \Theta_2]$, as both possible outcomes, namely that target $1$ is measured or target $2$ is measured.
+Let ${\bf{\Theta}}\in[\Theta_1, \Theta_2]$, as both possible outcomes, namely that object $1$ is measured or object $2$ is measured.
 Each estimate may receive at most a single measurement, so $m_k=1$.
-Because all events ${\bf{\Theta}}$ fall inside an arbitrary volume $V$, or are subject to being measured, the measurement association indicator $\tau=1$, meaning that measurement $k$ is associated with exactly one of the two targets.
-Hence, the number of unassociated measurements $\phi({\bf{\Theta}})=0$ because the event that neither target is measured $\Theta_0\notin{\bf{\Theta}}$.
-The binary target detection indicator is defined as $\delta_i(\Theta_i)=1$ if measurement $k$ is associated with target $i$ and $\delta_i(\Theta_i)=0$ otherwise.
+Because all events ${\bf{\Theta}}$ fall inside an arbitrary volume $V$, or are subject to being measured, the measurement association indicator $\tau=1$, meaning that measurement $k$ is associated with exactly one of the two objects.
+Hence, the number of unassociated measurements $\phi({\bf{\Theta}})=0$ because the event that neither object is measured $\Theta_0\notin{\bf{\Theta}}$.
+The binary object detection indicator is defined as $\delta_i(\Theta_i)=1$ if measurement $k$ is associated with object $i$ and $\delta_i(\Theta_i)=0$ otherwise.
 The detection probabilities are generalized to $y$ estimates such that $P(\Theta_i)=\frac{1}{y}$ for equal likelihood of event $\Theta_i$.
@@
 
-\paragraph{Derivation of Posteriori Covariance}
+\paragraph{Derivation of a Posteriori Covariance}
 
-The second proof shows how the measurement updates may take a more manipulatable form under the assumptions. From~\cite[Section 6.4]{TrackDataAssoc},
+The second proof shows how the measurement updates may take a more manipulable form under the assumptions. From~\cite[Section 6.4]{TrackDataAssoc},
 \begin{align}
@@
 \end{align}
-Because only one innovation is calculated per target, there is no spread in the innovations term on each target, so this portion of the covariance $\tilde{P}=0$, which would normally be included in the JDPAF without these assumptions.
-The posterior covariance of target $i$ conditioned on event $\Theta_i$ is defined for any $K_i$ from ~\cite[Section 4.2]{OptEst1}:
+Because only one innovation is calculated per object, there is no spread in the innovations term on each object, so this portion of the covariance $\tilde{P}=0$, which would normally be included in the JDPAF without these assumptions.
+The posterior covariance of object $i$ conditioned on event $\Theta_i$ is defined for any $K_i$ from ~\cite[Section 4.2]{OptEst1}:
 \begin{align}
@@
 \end{align}
-Let $\beta_{0_i}$ be the probability that target $i$ is not measured; with only one measurement considered each time step; hence, $\beta_{0_i}=1-\beta_i$.
+Let $\beta_{0_i}$ be the probability that object $i$ is not measured; with only one measurement considered each time step; hence, $\beta_{0_i}=1-\beta_i$.
 Again from~\cite[Section 6.4]{TrackDataAssoc},

### 2018-06-09 Conference/ECC 2014/last_bit.tex
added Conference folder

--- 2018-06-09 Conference/ECC 2014/ECC14.tex
@@
 
-Data association is a heavily studied field with many applications in estimation when measurements are not necessarily originating from a single target.
-We consider applications wherein multiple targets of interest enter the field of view of a sensor; however, the measurement origin is unknown.
-To properly track these targets, we require estimation techniques to associate targets with measurements.
-When the targets are in very close proximity, the association uncertainty grows, and an effective algorithm is necessary to maintain close estimations of the targets.
-
-A variety of data association techniques are considered for the close-proximity data association with varying performance.
-Some approaches such as multiple hypothesis tracking (MHT) and probability hypothesis density (PHD) techniques involve a high computational load, especially when targets are in close proximity, making real-time implementation more difficult~\cite{MHT1},~\cite{PHD1}, and~\cite{PHD2}.
-Many other techniques may involve the Kalman filter (KF) or extended KF (EKF), which are well studied and provide a framework for recursive data association techniques that are much easier to implement in real-time.
-A common heuristic approach is known as the nearest neighbor filter (NNF), which is a simple algorithm that only updates estimates for measurements with the closest square Mahalnobis distance, updating other targets only with their predictions~\cite{NN2}.
-This is considered a \emph{hard decision} in~\cite{JPDAF1}, or that every measurement update receives full weighting; if the decision is incorrect, these measurement updates can result in detrimental effects to the associations of the estimates, so this technique is avoided in the analysis of this paper.
-Alternatively, some approaches calculate the minimum mean square error (MMSE) of the estimates and their associated uncertainty.
+Data association has been widely studied with various applications in estimation of a single or multiple targets with measurements that are not necessarily originating from correct targets in order.
+To properly track multiple targets with measurement origin uncertainties, measured data should be properly associated to targets with certain probabilities.
+%When the targets are in very close proximity, the association uncertainty grows, and an effective algorithm is required to maintain close estimations of the targets.
+Otherwise, estimators tend to coalesce neighboring tracks.
+
+%This paper considers estimation wherein multiple targets of interest enter the field of view of a sensor; however, the measurement origin is unknown.
+
+A variety of data association techniques with varying performance have been studied.
+A common heuristic approach is known as the nearest neighbor filter (NNF), where measurements with the closest distance are only applied~\cite{NN2}.
+This is considered a \emph{hard decision} in~\cite{JPDAF1}, as every selected measurement update receives full weighting.
+Therefore, if the decision is incorrect, these measurement updates can result in detrimental effects to the associations of the estimates.
+Alternatively, there are approaches that calculate the minimum mean square error (MMSE) of the estimates and their associated uncertainty.
 The probabilistic data association (PDA) technique is effective in tracking a single target because it updates with a \emph{soft decision}; in particular, this approach updates provides a weighting from a likelihood of measurement association.
+Other approaches include multiple hypothesis tracking (MHT) and probability hypothesis density (PHD) techniques~\cite{MHT1,PHD1,PHD2}.
+
+%Many other techniques may involve the Kalman filter (KF) or extended KF (EKF), which are well studied and provide a framework for recursive data association techniques that are much easier to implement in real-time.
 
 The joint PDA (JPDA) is an extension of the PDA to incorporate multiple targets~\cite{JPDAF1},~\cite{TrackDataAssoc}.
-The JPDA filter (JPDAF) is particularly advantageous in multi-target scenarios because when targets share measurements, the soft decisions prevent loss of information, which occurs with hard decisions.
-However, when neighboring targets share measurements, the estimates are subject to coalescence.
-
-Most techniques to avoid track coalescence involve pruning the tracks, or designing a set of rules that limit the possible outcomes to likely scenarios.
-These rules typically involve the exact nearest neighbor probabilistic data association (ENNPDA), which applies a strong decision to the most likely associated estimate.
-Pruning algorithms incorporate the ENNPDA algorithm because  estimates do not share measurements, making it insensitive to track coalescence~\cite{Coal1}.
-For example in~\cite{Fitzgerald}, the algorithm involves a set of rules to prune the tracks that share measurements.
-The algorithm considers the possible tracks and applies a logic that assigns estimates to likely scenarios; a set of possible tracks are considered and lower likelihood track scenarios are deleted.
-Updates to tracks considered likely are given a strong decision.
-This technique is extended in~\cite{Coal_d},~\cite{Coal_e}, and~\cite{Coal_c} because of its successful resistance to coalescence.
-However, track pruning assumes that low likelihood tracks are incorrect, which is not necessarily true.
-Removing the weighting scheme of the JPDA causes the estimates to lose information, which this paper seeks to avoid.
-
-We propose a novel approach to removing coalescence by minimizing a similarity index of the estimates.
-Instead of choosing a different means of weighting the measurement updates, we design an optimal recursive estimation algorithm that chooses gains to compensate for the effects of coalescence.
-The algorithm derived in this paper can be applied in real-time without the heuristic nature, loss of information, or the risk of hard decisions from the current coalescence avoidance approaches.
-
-The paper is organized as follows: the problem is formally presented in Section \ref{ProbDef}.
-The derivation and explanation of a modified JPDAF under the assumptions of this paper is described in Section \ref{JPDAF}.
-The optimal gain that minimizes a weighted cost of uncertainty and similarity is derived in Section \ref{C-JPDAF}.
-We consider a satellite tracking example of this technique in Section \ref{NumRes}.
-Finally, we discuss results and applications for the C-JPDAF in Section \ref{ConclusionFutureWork}.
-
-% Dr. Lee Edits:
+The JPDA filter (JPDAF) is particularly advantageous in multi-target scenarios because when targets share measurements, the soft decisions prevent loss of information that occurs with hard decisions.
+As such, it is well known that JPDAF are effective in handling data clutter or missed detection.
+However, when the targets are in very close proximity, the association uncertainty grows significantly, thereby degrading the filtering performance.
+Furthermore, the estimated probability distributions tend to be coalesced especially if the paths of neighboring targets cross with each other.
+
+%neighboring targets share measurements, the estimates are subject to coalescence.
+
+Many techniques to avoid track coalescence involve the concept that when the exact nearest neighbor probabilistic data association (ENNPDA) is applied, which associates a measurement with a strong decision to the most likely estimate, the estimates are insensitive to track coalescence~\cite{Coal1}.
+By pruning the tracks, or designing a set of rules to determine high likelihood scenarios, an algorithm choice or weighing scheme is applied to the measurement updates.
+For example in~\cite{Fitzgerald}, the algorithm involves a set of rules based on the data received with respect to the estimates.
+A special logic is introduced that determines how weighting of the updates is defined or if a new track should be considered.
+In some cases, it is possible for an incorrectly associated estimate measurement may receive full weighting in the update.
+Other criteria or extensions of this approach are detailed in~\cite{Coal_d},~\cite{Coal_e}, and~\cite{Coal_c}, which can be very effective, but information may be lost when measurements are omitted from updating an estimates with small probabilities of association because the pruning rule prevented these association possibilities.
+
+This paper is focused on optimal joint probabilistic data association filter that avoids coalescence in close proximity.
+We propose a novel approach to remove coalescence by minimizing a similarity index of the estimates.
+More explicitly, the gain of the proposed estimator is chosen such that a weighted sum of the size of uncertainties and a similarity measure between estimated states is minimized at each update.
+This yields an optimal recursive estimation algorithm that compensates for the effects of coalescence systematically.
+The proposed approach has simpler structures compared with pruning-based approaches, and it is computationally efficient.
+As it is still based on probabilistic data association, it avoids the heuristic nature, loss of information, or the risk of other coalescence avoidance approaches based on hard-decision.
+
+The paper is organized as follows: the dynamics of target and the measurement models are described in Section \ref{ProbDef}.
+The proposed Coalescence Avoiding Optimal JPDAF (C-JPDAF) are presented at Sections \ref{JPDAF}, and the optimal gain that minimizes a weighted cost of uncertainty and similarity is derived in Section \ref{C-JPDAF}.
+These are illustrated by a numerical example for estimating two satellites in crossing orbits in Section \ref{NumRes}, followed by conclusions in Section \ref{ConclusionFutureWork}.
+
+%_______________________________________________%
+
 \section{Data Association Filtering Problem Formulation}
@@
 
-More explicitly, let $\Theta_i$ be the event that the measurement originated from the $i$-th target, for $i\in\{1,2\}$. According to the above assumptions, we have
+More explicitly, let $\Theta_i$ be the event that the measurement is originated from the $i$-th target, for $i\in\{1,2\}$. According to the above assumptions, we have
 \begin{align}
@@
 
-A filtering problem is formulated as recursively updating the estimated state of each target and its covariance, namely $(\hat x_{i_k}, P_{i_k})$ based on the above assumptions given by \refeqn{xkp}--\refeqn{MeaOrig}.
+A filtering problem is formulated as recursively updating the estimated state of each target and its covariance, namely $(\hat x_{i_i}, P_{i_i})$ based on the above assumptions given by \refeqn{xkp}--\refeqn{MeaOrig}.
 In particular, we wish to avoid track coalescence when two targets are in close proximity.
@@
 \subsection{Flow Update}
-Similar to Kalman filters, the JPDAF is composed of a flow update and a measurement update.
-During the flow update (or prediction step), the estimated state $\hat x_{i_k}$ and its covariance $P_{i_k}$ of each target are updated as in the standard filter as follows:
+Similar to Kalman filters, JPDAF is composed of flow update and measurement update.
+At flow update (or prediction step), the estimated state $\hat x_{i_k}$ and its covariance $P_{i_k}$ of each target are updated as in the standard filter as follows:
 \begin{gather}
@@
 \end{align}
-where the estimator gain is equal to the Kalman gain, i.e., $K_i = P_{i_k}^{-} H_{i_k}^T S_{i_k}^{-1}$, as there is no measurement origination uncertainty when conditioned on the event $\Theta_i$.
+where the estimator gain is equal to the Kalman gain, i.e., $K_i = P_{i_k}^{-} H_{i_k^T} S_{i_k}^{-1}$, as there is no measurement origination uncertainty when conditioned on the event $\Theta_i$.
 
@@
 The joint probabilistic data association filter described in the previous section is effective in handling clutter or missed detection of multiple targets.
-However, it is prone to coalescing estimates, especially with targets in close-proximity.
+However, it is prone to coalesce estimates, especially in targets in close-proximity.
 Here, we present a coalescence-avoiding optimal JPDAF, hereafter C-JPDAF for short, to avoid data coalescence.
@@
 We can easily choose the initial conditions to be $X_{1_0}= X_{ref}+X_{pert}$ and $X_{2_0}= X_{ref}-X_{pert}$.
-The initial estimates are assumed to be relatively accurate: there is no need to simulate large uncertainty when the paths are initially sufficiently far apart because this paper is concerned with the coalescence when the targets are close together, so the state and covariance will have been determined by simulated measurements while the estimates were still spread significantly apart; hence for both estimates, the initial covariance was chosen to be $P_0=diag(\begin{bmatrix}
-2.5\times10^{-4} & 2.5\times10^{-6} & 3.9\times10^{-8} & 3.9\times10^{-8} \end{bmatrix})$.
-The process noise was chosen to be equivalent for both estimate such that $Q=diag(\begin{bmatrix}
-2.5\times10^{-5} & 2.5\times10^{-7} & 3.9\times10^{-9}& 3.9\times10^{-9}\end{bmatrix})$ and $R=\begin{bmatrix}
-2.25\times10^{-2}
-\end{bmatrix}$ applied at each time step.
-
-Three filtering approaches are applied, namely extended Kalman filter (EKF), JPDAF, and C-JPDAF, such that a single measurement was evaluated at each time step, and its origination was random and equally likely from both targets.
+The initial estimates are assumed to be relatively accurate: there is no need to simulate large uncertainty when the paths are initially sufficiently far apart because this paper is concerned with the coalescence when the targets are close together, so the state and covariance will have been determined by simulated measurements while the estimates were still spread significantly apart.
+
+\EditTL{Put the values of $P_0$, $Q$ and $R$ here. Describe how the measurements are switched between two satellites. }
+
+Three filtering approaches are applied, namely extended Kalman filter (EKF), JPDAF, and C-JPDAF.
 For EKF, it is assumed that the correct measurement origination is know with certainty, i.e., all of the measurements are associated to correct target.
@@
 \begin{tabular}{ccc}\hline
-Method & Radial RMS Error (km) & Computation time (sec) \\\hline
-EKF & 1.438e-05 &6.118\\
-JPDAF & 7.1242e-05 &6.130\\
-C-JPDAF & 2.3145e-05  & 19.864 \\\hline
+Method & RMS Error & Computation time \\\hline
+EKF & &\\
+JPDAF & &\\
+C-JPDAF & & \\\hline
 \end{tabular}}
@@
 
+\subsection{Proofs}
+
+\paragraph{Derivation of the Association Probability}
+
 We incorporate all of the assumptions from Section \ref{DAP} to simplify the weighting term for the estimates.
-We consider time step $k$ for the following proofs so subscripts omitted in this appendix.
-We define ${\bf{\Theta}}\in[\Theta_1, \Theta_2]$, as all possible outcomes, namely that target $1$ is measured or target $2$ is measured from $m_k=1$ total measurements.
-Because all events ${\bf{\Theta}}$ fall inside an arbitrary volume $V$, or are subject to being measured, we define the measurement association indicator $\tau=1$, meaning that measurement $k$ is associated with exactly one of the two targets.
-Hence, the number of unassociated measurements $\phi({\bf{\Theta}})=0$ because the event that neither target is measured $\Theta_0\notin{\bf{\Theta}}$.
-We define the binary target detection indicator $\delta_i(\Theta_i)$, which indicates the association of any measurement with target $i$; hence, $\delta_i(\Theta_i)=1$ if measurement $k$ is associated with target $i$ and $\delta_i(\Theta_i)=0$ otherwise.
-We also consider equal likelihood of detection $P(\Theta_i)$ for each measurement; we can generalize the number of estimates to $y$ such that $P(\Theta_i)=\frac{1}{y}$ for equal likelihood of event $\Theta_i$.
-From~\cite[Section 9.3]{TrackDataAssoc}, we determine the weighting for $T$ possible targets to be
-\begin{align}
-\beta_i\equiv P(\Theta_i|z)=&\frac{1}{c}\frac{\phi!}{V}\displaystyle\prod_{j=1}^{m_k} f_{i_j}^{\tau_j}\displaystyle\prod_{i=1}^{T} P(\Theta_i)^{\delta_i}(1-P(\Theta_i))^{1-\delta_i}\\
-=&\frac{1}{cV}f_{i}\frac{1}{y}(1-\frac1y)^{T-1}
-\end{align}
-for normalizing constant $c$ and Gaussian pdf $f_{i}$. Note that $T-1$ targets are not associated and only one target is associated. Hence, $c$ is defined as
-\begin{align}
-c=&\frac{1}{V}\frac{1}{y}(1-\frac1y)^{T-1}\displaystyle\sum\limits_{l=1}^y f_{l}.
+ we consider time step $k$ (subscripts omitted here) that the during the \emph{single} event $\Theta_i$ target $i$ is measured out of $m_k=1$ total measurements inside arbitrary volume $V$.
+Because all targets fall inside $V$, the measurement association indicator $\tau=1$ while the number of unassociated measurements $\phi(\Theta_i)=0$; at least one measurement is associated, so the binary target detection indicator $\delta_t(\Theta_i)=1$.
+We also consider equal likelihood of detection $P_D$ for each measurement; we generalize the number of estimates to $y$ such that $P_D=\frac{1}{y}$.
+From~\cite[Section ?.?]{TrackDataAssoc}, we determine the weighting to be
+\begin{align}
+\beta_i\equiv P(\Theta_i|z)=&\frac{1}{c}\frac{\phi!}{V}\displaystyle\prod_{j=1}^{m_k} f_{i_j}^{\tau_j}\displaystyle\prod_{i=1}^{2} (P_{D_i})^{\delta_t}(1-P_{D_i})^{1-\delta_t}\\
+=&\frac{1}{cV}f_{i}
+\end{align}
+for normalizing constant $c$ and Gaussian pdf $f_{i}$. Hence, $c$ is defined as
+\begin{align}
+c=&\frac{1}{V}\displaystyle\sum\limits_{l=1}^y f_{l}.
 \end{align}
@@
 
-The second proof shows how the measurement updates may take a more manipulatable form under the assumptions. From~\cite[Section 6.4]{TrackDataAssoc},
+\paragraph{Derivation of Posteriori Covariance}
+
+The second proof shows how the measurement updates may take a more manipulatable form under the assumptions. From~\cite{TrackDataAssoc},
 \begin{align}
@@
 \end{align}
- , so this portion of the covariance Let $\tilde{P}$ be spread in the innovations term on each target.
-Because only one innovation is calculated per target, there is no spread in this term; hence $\tilde{P}=0$.
-Note that his term would be nonzero in the JDPAF without the single measurement assumption.
-Let $P_{i}^C$ be the posterior covariance of target $i$ conditioned on event $\Theta_i$.
-For any $K_i$ from ~\cite[Section 4.2]{OptEst1}:
-\begin{align}
-\begin{split}
-P_{i}^C=&(I-K_{i}H_{i})P_{i}^{-}(I-K_{i}H_{i})^T+K_{i}R_{i}K_{i}^T\\
-=&P^--K_{i}H_{i}P_{i}^{-}-P_{i}^{-}H_{i}^TK_{i}^T\\
-&+K_{i}(H_{i}P_{i}^{-}H_{i}^T+R_i)K_{i}^T.
-\end{split}
-\end{align}
-We define $\beta_{0_i}$ as the probability that target $i$ is not measured; with only one measurement considered each time step; hence, $\beta_{0_i}=1-\beta_i$.
-Again from~\cite[Section 6.4]{TrackDataAssoc},
-\begin{align}
-\begin{split}
-P^+_i=&\beta_{0_i}P^-_i+(1-\beta_{0_i})P_i^C+\tilde{P}\\
-=&(1-\beta_i)P^-+\beta_i[P^--K_{i}H_{i}P_{i}^{-}\\
-&-P_{i}^{-}H_{i}^TK_{i}^T+K_{i}(H_{i}P_{i}^{-}H_{i}^T+R_i)K_{i}^T]\\
-=&P^-_i-\beta_i[K_{i}H_{i}P_{i}^{-}+P_{i}^{-}H_{i}^TK_{i}^T\\
-&-K_{i}(H_{i}P_{i}^{-}H_{i}^T+R_i)K_{i}^T].
-\end{split}
+Because only 1 innovation is calculated per target, there is no spread in the innovations term on each target, which is included in the JDPAF without these assumptions.
+Furthermore we define $\beta_{0_i}$ as the probability that target $i$ is not measured; with only one measurement considered each time step, $\beta_{0_i}=1-\beta_i$.
+Again from~\cite{TrackDataAssoc},
+\begin{align}
+P^+_i=&\beta_{0_i}P^-_i+(1-\beta_{0_i})(P^-_i-K_iS_iK_i^T)\\
+P^+_i=&(1-\beta_i+\beta_i)P^-_i-\beta_iK_iS_iK_i^T\\
+P^+_i=&P^-_i-\beta_iK_iS_iK_i^T
 \end{align}

### 2018-06-09 Conference/ICRA 2015/Abstract/ICRA15.tex
added Conference folder

\documentclass[letterpaper, 10pt, conference]{ieeeconf}
\IEEEoverridecommandlockouts \overrideIEEEmargins
\usepackage{amsmath,amssymb,url}
\usepackage{graphicx,subfigure,hyperref}
\usepackage{color}
%\usepackage{refcheck}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\abs}[1]{\ensuremath{\left| #1 \right|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\ip}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mbox{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mbox{tr}\ensuremath{\!\bracket{#1}}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathbf{d}}}
\newcommand{\pair}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\!\langle #1 \rangle\!\rangle}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\ad}{\ensuremath{\mathrm{ad}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}

\title{\LARGE \bf
Geometric Adaptive Control for Aerial Transportation of a Rigid Body}
%Geometric control and differential flatness of a quadrotor {UAV} with a cable-suspended load on {$\SE\times\Sph^2$}}

\author{Evan Kaufman, Taeyoung Lee%\authorrefmark{1}%, Koushil Sreenath, and Vijay Kumar%\authorrefmark{2}%
\thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington, DC 20052 {\tt tylee@gwu.edu}}%
%\thanks{Koushil Sreenath, Vijay Kumar, GRASP Lab, University of Pennsylvania, Philadelphia, PA 190104 {\tt \{koushils,kumar\}@seas.upenn.edu}}%
%\thanks{Vijay Kumar, GRASP Lab, University of Pennsylvania, Philadelphia, PA 190104 {\tt kumar@seas.upenn.edu}}%
\thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}
%\thanks{\textsuperscript{\footnotesize\ensuremath{\dagger}}This research has been supported in part by .}
}

\newcommand{\EditTL}[1]{{\color{red}\protect #1}}
\renewcommand{\EditTL}[1]{{\protect #1}}

\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{remark}{Remark}

\begin{document}
\allowdisplaybreaks

\maketitle \thispagestyle{empty} \pagestyle{empty}

\begin{abstract}
This paper is focused on tracking control for a rigid body payload, that is connected to an arbitrary number of quadrotor unmanned aerial vehicles via rigid links. A geometric adaptive controller is constructed such that the payload asymptotically follows a given desired trajectory for its position and attitude in the presence of uncertainties. The coupled dynamics between the rigid body payload, links, and quadrotors are explicitly incorporated into control system design and stability analysis. These are developed directly on the nonlinear configuration manifold in a coordinate-free fashion to avoid singularities and complexities that are associated with local parameterizations. %The desirable features of the proposed control system are illustrated by a numerical example.
\end{abstract}

\section{Introduction}

%Quadrotor unmanned aerial vehicles have desirable features of simple mechanical structures, hovering capabilities, and high thrusts. It has been widely studied for various applications, such as aerial photography and commercials~\cite{GuiAck13}, and complex maneuvers are also demonstrated~\cite{LeeLeoPICDC10,GilHuaPICRA10,MelMicIJRR12}. In particular, the high thrust-to-weight ratio of quadrotors or multi-rotor aerial vehicles is useful for aerial load transportation.

%Aerial transport of payloads by towed cables is common in emergency response, industrial operations, and military missions. Examples of aerial transportation range from search and rescue missions where individuals are lifted from dangerous situations, to the delivery of heavy equipment to the top of a tall building that is difficult to reach by other means.

%Transportation of a cable-suspended load has been studied traditionally for human-operated helicopters~\cite{CicKanJAHS95,BerPICRA09}. Recently,

By utilizing the high thrust-to-weight ratio, quadrotor unmanned aerial vehicles have been envisaged for aerial load transportation. Most of the existing results for the control of quadrotors to transport a cable-suspended payload are based on the assumption that the dynamics of the payload is decoupled from the dynamics of quadrotors. For example, the effects of the payload are considered as arbitrary external forces and torques exerted to quadrotors. As such, these results may not be suitable for agile load transportation where the motion of cable and payload should be actively suppressed.

Recently, the full dynamic model for an arbitrary number of quadrotors transporting a payload has been developed, and based on that, geometric tracking controllers are constructed in an intrinsic fashion. In particular, autonomous transportation of a point mass connected to quadrotors via rigid links is developed in~\cite{LeeSrePICDC13}. It has been generalized into a more realistic dynamic model that considers the deformation of cables in~\cite{GooLeePACC14}, and also the attitude dynamics of a payload, that is considered as a rigid body instead of a point mass, is incorporated in~\cite{LeePICDC14}. However, these results are based on the assumption that the exact properties of the quadrotors and the payload are available, and that there are no external disturbances, thereby making it challenging to implement those results in  actual hardware systems.

The objective of this paper is to construct a control system for an arbitrary number of quadrotors connected to a rigid body payload via rigid links with explicit consideration on uncertainties. A coordinate-free form of the equations of motion that have been developed in~\cite{LeePICDC14} is extended to include the effects of unknown forces and moments acting on each of the quadrotors, the cables, and the payload. A geometric nonlinear adaptive control system is designed such that both the position and the attitude of the payload asymptotically follow their desired trajectories, while maintaining a certain formation of quadrotors relative to the payload.

The unique property is that the coupled dynamics of the payload, the cables, and quadrotors are explicitly incorporated in control system design for agile load transportation, where the motions of the payload relative to the quadrotors are excited nontrivially. Another distinct feature is that the equations of motion and the control systems are developed directly on the nonlinear configuration manifold intrinsically. Therefore, singularities of local parameterization are completely avoided.

As such, the proposed control system is particularly useful for rapid and safe payload transportation in complex terrain, where the position and attitude of the payload should be controlled concurrently. Most of the existing control systems of aerial load transportation suffer from limited agility as they are based on reactive assumptions that ignore the inherent complexities in the dynamics of aerial load transportation. The proposed control system explicitly integrates the comprehensive dynamic characteristics to achieve extreme maneuverability in aerial load transportation. Nonlinear adaptive tracking controls of a cable-suspended rigid body with uncertainties have not been studied as mathematically rigorously as presented in this paper.

\begin{figure}
\centerline{
		\includegraphics[width=0.8\columnwidth]{Quad_N_Rgbd_Adaptive_3d0}
}
%\centerline{
%	\subfigure[Top view]{
%	\setlength{\unitlength}{0.1\columnwidth}\scriptsize
%	\begin{picture}(9.5,3.2)(0,0)
%	\put(0,0){\includegraphics[width=0.95\columnwidth]{Quad_N_Rgbd_Adaptive_3d2}}
%	\put(8.4,3.2){$t=0$}
%	\put(6.0,3.2){$t=3.3$}
%	\put(2.2,-0.2){$t=6.6$}
%	\put(0.2,2.3){$t=10$}
%	\put(2.2,3.2){$t=13.3$}
%	\put(5.8,-0.2){$t=16.6$}
%	\put(7.9,0.4){$t=20$}
%	\end{picture}}	
%}
%\centerline{
%	\subfigure[Side view]{
%		\includegraphics[width=0.95\columnwidth]{Quad_N_Rgbd_Adaptive_3d1}}
%}
\caption{Snapshots of controlled maneuver (red:desired trajectory, blue:actual trajectory). A short animation illustrating this maneuver is available at {\href{http://youtu.be/nOWErfdzZLU}{http://youtu.be/nOWErfdzZLU}}.}\label{fig:SS}
\end{figure}

\vfill
\bibliography{/Users/tylee/Documents/BibMaster}
\bibliographystyle{IEEEtran}

\end{document}

%This paper is organized as follows. A dynamic model is presented and the problem is formulated at Section \ref{sec:DM}. Control systems are constructed at Sections \ref{sec:SDM} and \ref{sec:FDM}, which are followed by a numerical example at Section V.

%
%
%These properties of quadrotors are also desirable for load carrying and transportation. %Areal transportation of a cable-suspended load has been studied traditionally for helicopters~\cite{CicKanJAHS95,BerPICRA09}. 
%Small-size single or multiple autonomous vehicles are considered for load transportation and deployment~\cite{PalCruIRAM12,MicFinAR11,MazKonJIRS10}. Nonlinear tracking control systems are developed for a single quadrotor UAV with a cable-suspended load in~\cite{SreMicPICRA13} and a companion paper~\cite{SreLeePICDC13}.
%
%Load transportation with multiple quadrotors is useful when the load is heavy compared with the maximum thrust of a single quadrotor, or when additional redundancy is required for safety. But, this is challenging since \textit{dynamically coupled} quadrotors should cooperate safely to transport load. This is in contrast to the existing results on formation control of decoupled multi-agent systems. 
%
%In this paper, we consider an arbitrary number of quadrotors that are connected to a point mass via rigid links. The equations of motion are derived from the variational principle, and control systems are developed such that the point mass asymptotically follows a given smooth desired trajectory. Two formation flight modes are introduced to control the formation of quadrotors with respect to the point mass.  In the existing control systems for a load-carrying quadrotor, such as~\cite{PalCruIRAM12}, a quadrotor is designed to follow pre-computed minimum swing trajectories while rejecting the force and moment exerted by the load that are considered as disturbances. The control systems proposed in this paper explicitly consider the coupling effects between the load dynamics and the dynamics of multiple quadrotors for safe transportation along complex trajectories. 
%
%Another distinct feature is that the equations of motion and the control systems are developed directly on the nonlinear configuration manifold in a coordinate-free fashion. This yields remarkably compact expressions for the dynamic model and controllers, compared with local coordinates that often require symbolic computational tools due to complexity of multibody systems. Furthermore, singularities of local parameterization are completely avoided to generate agile maneuvers in a uniform way.
%
%If the links are assumed to be rigid, the proposed control system is also applied to the cases where selected quadrotors are below the load, to obtain so-called \textit{flying inverted pendulum}. Linear control systems have been developed and implemented to stabilize few selected nominal trajectories in~\cite{HehDAnPICRA11}. The proposed control system is based on the full nonlinear dynamics, and it guarantees exponential stability of flying inverted pendulum for arbitrary desired trajectories.
%
%Compared with the companion paper~\cite{SreLeePICDC13} that is focused on tracking, differential flatness, and experiments of a single quadrotor with a suspended load, this paper proposes a cooperative framework of an arbitrary number of quadrotors.
%%The paper is organized as follows. We develop a globally defined dynamic in Section \ref{sec:DM}. Tracking control results are presented in Section \ref{sec:CSD}, followed by numerical examples. %.  Several numerical results are presented in Section IV.% Proofs are relegated to the Appendix.

\section{Problem Formulation}\label{sec:DM}

Consider $n$ quadrotor UAVs that are connected to a payload, that is modeled as a rigid body, via massless links (see Figure \ref{fig:DM}). Throughout this paper, the variables related to the payload is denoted by the subscript $0$, and the variables for the $i$-th quadrotor are denoted by the subscript $i$, which is assumed to be an element of $\mathcal{I}=\{1,\cdots\, n\}$ if not specified. We choose an inertial reference frame $\{\vec e_1,\vec e_2,\vec e_3\}$ and body-fixed frames $\{\vec b_{j_1},\vec b_{j_2},\vec b_{j_3}\}$ for $0\leq j\leq n$ as follows. For the inertial frame, the third axis $\vec e_3$ points downward along the gravity and the other axes are chosen to form an orthonormal frame. %The origin of the $j$-th body-fixed frame is located at the center of mass of the payload for $j=0$ and at the mass center the quadrotor for $1\leq j\leq n$. The third body-fixed axis $\vec b_{i_3}$ is normal to the plane defined by the centers of rotors, and it points downward.

\begin{figure}
\centerline{\setlength{\unitlength}{0.1\columnwidth}\footnotesize
\begin{picture}(10,6.5)(0.0,0)
\put(1.25,0){\includegraphics[width=0.75\columnwidth]{Quad_N_Rgbd_Adaptive_3d.pdf}}
\put(2.35,1.45){$\vec e_1$}
\put(1.9,0.3){$\vec e_2$}
\put(1.4,-0.1){$\vec e_3$}
\put(3.1,1.0){$x_0\in\Re^3$}
\put(6.7,0.5){\shortstack[c]{$m_0,J_0$\\$R_0\in\SO$}}
\put(5.3,5.3){\shortstack[c]{$m_i,J_i$\\$R_i\in\SO$}}
\put(7.8,4.2){$q_i\in\Sph^2$}
\put(6.8,1.8){$\rho_i$}
\put(7.65,3.3){$l_i$}
\end{picture}}
\caption{Dynamics model: $n$ quadrotors are connect to a rigid body $m_0$ via massless links $l_i$. The configuration manifold is $\Re^3\times\SO\times(\Sph^2\times\SO)^n$.}\label{fig:DM}
\vspace*{-0.1cm}
\end{figure}

The location of the mass center of the payload is denoted by $x_0\in\Re^3$, and its attitude is given by $R_0\in\SO$, where the special orthogonal group is defined by $\SO=\{R\in\Re^{3\times 3}\,|\, R^TR=I,\,\mathrm{det}[R]=1\}$. Let $\rho_i\in\Re^3$ be the point on the payload where the $i$-th link is attached, and it is represented with respect to the zeroth body-fixed frame. The other end of the link is attached to the mass center of the $i$-th quadrotor. The direction of the link from the mass center of the $i$-th quadrotor toward the payload is defined by the unit-vector $q_i\in\Sph^2$, where $\Sph^2=\{q\in\Re^3\,|\,\|q\|=1\}$, and the length of the $i$-th link is denoted by $l_i\in\Re$.

Let $x_i\in\Re^3$ be the location of the mass center of the $i$-th quadrotor with respect to the inertial frame. As the link is assumed to be rigid, we have $x_i=x_0+R_0\rho_i-l_i q_i$. The attitude of the $i$-th quadrotor is defined by $R_i\in\SO$, which represents the linear transformation of the representation of a vector from the $i$-th body-fixed frame to the inertial frame.

In summary, the configuration of the presented system is described by the position $x_0$ and the attitude $R_0$ of the payload, the direction $q_i$ of the links, and the attitudes $R_i$ of the quadrotors. The corresponding configuration manifold of this system is $\Re^3\times\SO\times(\Sph^2\times \SO)^n$.

The mass and the inertia matrix of the payload are denoted by $m_0\in\Re$ and $J_0\in\Re^{3\times 3}$, respectively. The dynamic model of each quadrotor is identical to~\cite{LeeLeoPICDC10}. The mass and the inertia matrix of the $i$-th quadrotor are denoted by $m_i\in\Re$ and $J_i\in\Re^{3\times 3}$, respectively. The $i$-th quadrotor can generates a thrust $-f_iR_ie_3\in\Re^3$ with respect to the inertial frame, where $f_i\in\Re$ is the total thrust magnitude and $e_3=[0,0,1]^T\in\Re^3$. It also generates a moment $M_i\in\Re^3$ with respect to its body-fixed frame. The control input of this system corresponds to $\{f_i,M_i\}_{1\leq i\leq n}$.

In this paper, we include external disturbances as follows. Let $\Delta_{x_0},\Delta_{R_0}\in\Re^3$ be arbitrary, but fixed disturbance force and moment acting on the mass center of the payload, respectively. The disturbance force and moment exerted on the $i$-th quadrotor are denoted by $\Delta_{x_i},\Delta_{R_i}\in\Re^3$, respectively. The disturbance forces are represented with respect to the inertial frame, and the disturbance moments are represented with respect to the corresponding body-fixed frame.

Throughout this paper, the 2-norm of a matrix $A$ is denoted by $\|A\|$, and its maximum eigenvalue and minimum eigenvalues are denoted by $\lambda_{M}[A]$ and $\lambda_{m}[A]$, respectively. The standard dot product is denoted by $x \cdot y = x^Ty$ for any $x,y\in\Re^3$.

\subsection{Equations of Motion}

The kinematic equations for the payload, quadrotors, and links are given by
\begin{gather}
\dot q_{i} = \omega_i\times q_i=\hat\omega_i q_i,\label{eqn:dotqi}\\
\dot R_0  = R_0\hat\Omega_0,\quad \dot R_i  = R_i\hat\Omega_i,\label{eqn:dotRi}
\end{gather}
where $\omega_i\in\Re^3$ is the angular velocity of the $i$-th link, satisfying $q_i\cdot\omega_i=0$, and $\Omega_0$ and $\Omega_i\in\Re^3$ are the angular velocities of the payload and the $i$-th quadrotor expressed with respect to its body-fixed frame, respectively. The \textit{hat map} $\hat\cdot:\Re^3\rightarrow\so$ is defined by the condition that $\hat x y=x\times y$ for all $x,y\in\Re^3$, and the inverse of the hat map is denoted by the \textit{vee map} $\vee:\so\rightarrow\Re^3$, where $\so$ denotes the set of $3\times 3$ skew-symmetric matrices, i.e., $\so=\{S\in\Re^{3\times 3}\,|\, S^T=-S\}$, and it corresponds to the Lie algebra of $\SO$. %Several properties of the hat map are summarized as follows~\cite{ShuJAS93}:
%\begin{gather}
%    \hat x y = x\times y = - y\times x = - \hat y x,\\
%%    \hat x^T \hat x = (x^T x) I - x x^T,\\
%%    \hat x \hat y \hat x=-(y^Tx)\hat x,\\
%%    -\frac{1}{2}\tr{\hat x \hat y} = x^T y,\\
%%    \widehat{x\times y} = \hat x \hat y -\hat y \hat x = yx^T-xy^T,\\
%    %\tr{\hat x A}=
%    \tr{A\hat x }=\frac{1}{2}\tr{\hat x (A-A^T)}=-x^T (A-A^T)^\vee,\label{eqn:hat1}\\
%%    \widehat{Ax} = \hat x \parenth{\frac{1}{2}\tr{A}I-A} + \parenth{\frac{1}{2}\tr{A}I-A}^T \hat x,\\
%    \hat x  A+A^T\hat x=(\braces{\tr{A}I_{3\times 3}-A}x)^{\wedge},\label{eqn:xAAx}\\
%R\hat x R^T = (Rx)^\wedge,\label{eqn:RxR}
%\end{gather}
%for any $x,y\in\Re^3$, $A\in\Re^{3\times 3}$, and $R\in\SO$.

We derive equations of motion according to Lagrangian mechanics. The velocity of the $i$-th quadrotor is given by $\dot x_i = \dot x_0+\dot R_0\rho_i - l_i\dot q_i$. The kinetic energy of the system is composed of the translational kinetic energy and the rotational kinetic energy of the payload and quadrotors:
\begin{align}
\mathcal{T} & = \frac{1}{2}m_0 \|\dot x_0\|^2 + \frac{1}{2}\Omega_0\cdot J_0\Omega_0\nonumber\\
&\quad +\sum_{i=1}^n \frac{1}{2} m_i \|\dot x_0+\dot R_0\rho_i - l_i\dot q_i\|^2+ \frac{1}{2}\Omega_i\cdot J_i\Omega_i.\label{eqn:TT}
\end{align}
The gravitational potential energy is given by
\begin{align}
\mathcal{U} = - m_0g e_3 \cdot x_0 - \sum_{i=1}^n m_ige_3\cdot (x_0+R_0\rho_i-l_iq_i),\label{eqn:UU}
\end{align}
where the unit-vector $e_3$ points downward along the gravitational acceleration as shown at \reffig{DM}. The resulting Lagrangian of the system is $\mathcal{L}=\mathcal{T}-\mathcal{U}$.

The corresponding Euler-Lagrange equations have been developed according to Hamilton's principle in~\cite{LeePICDC14},  Here, it is generalized to include the effects of disturbances via the Lagrange-d'Alembert principle. Let the action integral be $\mathfrak{G}=\int_{t_0}^{t_f} \mathcal{L}\, dt$. Next, let the total control thrust at the $i$-th quadrotor with respect to the inertial frame be denoted by $u_i = -f_iR_ie_3\in\Re^3$ and the total control moment at the $i$-th quadrotor is defined as $M_i\in\Re^3$. There exist the disturbances $\Delta_{x_0},\Delta_{R_0}$ for the payload, and the disturbances $\Delta_{x_i},\Delta_{R_i}$ for the $i$-th quadrotor. The corresponding virtual work can be written as
\begin{align*}
\delta\mathcal{W} & = \int_{t_0}^{t_f} \sum_{i=1}^n (u_i+\Delta_{x_i})\cdot\braces{\delta x_0 + R_0\hat\eta_0\rho_i -l_i\xi_i\times q_i}\\
&\quad + \sum_{i=1}^n(M_i+\Delta_{R_i}) \cdot\eta_i+\Delta_{x_0}\cdot\delta x_0 + \Delta_{R_0}\cdot \eta_0\,dt.
\end{align*}
The Lagrange-d'Alembert principle states that $\delta\mathfrak{G}=-\delta\mathcal{W}$ for any variation of trajectories with fixed end points. This yields the following equations of motion (see~\cite{LeeACC151ext} for detailed derivations),
\begin{gather}
M_q (\ddot x_0-ge_3) - \sum_{i=1}^n m_iq_iq_i^T R_0\hat\rho_i\dot\Omega_0 = \Delta_{x_0}\nonumber\\
 +\sum_{i=1}^n u_i^\parallel+\Delta_{x_i}^\parallel-m_il_i\|\omega_i\|^2q_i- m_iq_iq_i^T R_0\hat\Omega_0^2\rho_i,
\label{eqn:ddotx0}\\
%
  (J_0-\sum_{i=1}^n m_i\hat\rho_i R_0^T q_iq_i^T R_0\hat\rho_i)\dot\Omega_0 \nonumber\\
 + \sum_{i=1}^n m_i\hat\rho_i R_0^Tq_iq_i^T(\ddot x_0-ge_3) + \hat\Omega_0   J_0\Omega_0
=\Delta_{R_0}\nonumber \\
+\sum_{i=1}^n \hat\rho_i R_0^T (u_i^\parallel+\Delta_{x_i}^\parallel-m_il_i\|\omega_i\|^2 q_i-m_iq_iq_i^TR_0\hat\Omega_0^2\rho_i),\label{eqn:dotW0}\\
%
\dot\omega_i =\frac{1}{l_i}\hat q_i(\ddot x_0-ge_3 - R_0\hat\rho_i\dot\Omega_0+
R_0\hat\Omega_0^2\rho_i)\nonumber\\
 -\frac{1}{m_il_i}\hat q_i(u_i^\perp+\Delta_{x_i}^\perp),\label{eqn:dotwi}\\
% \dot\omega_i = \frac{1}{l_i}\hat q_i (\ddot y-ge_3) -\frac{1}{m_il_i} \hat q_i u_i^\perp,\label{eqn:dotwi}\\
J_i\dot\Omega_i + \Omega_i\times J_i\Omega_i = M_i+\Delta_{R_i},\label{eqn:dotWi}
\end{gather}
where $M_q=m_y I + \sum_{i=1}^n m_i q_i q_i^T\in\Re^{3\times 3}$, which is symmetric, positive-definite for any $q_i$.

Recall the vector $u_i\in\Re^3$ represents the control force at the $i$-th quadrotor, i.e., $u_i=-f_i R_i e_3$.  The vectors $u_i^\parallel$ and $u_i^\perp\in\Re^3$ denote the orthogonal projection of $u_i$ along $q_i$, and the orthogonal projection of $u_i$ to the plane  normal to $q_i$, respectively, i.e.,
\begin{align}
u_i^\parallel & = q_iq_i^T u_i,\\
u_i^\perp &= -\hat q_i^2 u_i =(I-q_iq_i^T) u_i.
\end{align}
Therefore, $u_i = u_i^\parallel + u_i^\perp$. Throughout this paper, the subscripts $\parallel$ and $\perp$ of a vector denote the component of the vector that is parallel to $q_i$ and the other component of the vector that is perpendicular to $q_i$. Similarly, the disturbance force at the $i$-th quadrotor is decomposed as
\begin{align}
\Delta_{x_i}^\parallel & = q_iq_i^T \Delta_{x_i},\label{eqn:delxiparallel}\\
\Delta_{x_i}^\perp & = -\hat q_i^2 \Delta_{x_i} = (I-q_iq_i^T) \Delta_{x_i}.\label{eqn:delxiperp}
\end{align}

\subsection{Tracking Problem}

Define a fixed matrix $\mathcal{P} \in \Re^{6\times 3n}$ as 
\begin{align}
\mathcal{P} = \begin{bmatrix} I_{3\times 3} & \cdots & I_{3\times 3} \\ \hat\rho_1 & \cdots & \hat\rho_n \end{bmatrix}. \label{eqn:P}
\end{align}
Recall that $\rho_i$ describe the point on the payload where the $i$-th link is attached. 
Assume the links are attached to the payload such that
\begin{align}
\mathrm{rank}[\mathcal{P}] \geq 6.\label{eqn:A1}
\end{align}
This is to guarantee that there exist enough degrees of freedom in control inputs for both the translational motion and the rotational maneuver of the payload. The assumption \refeqn{A1} requires that the number of quadrotor is at least three, i.e., $n\geq 3$.

It is also assumed that the bounds of the disturbance forces and moments are available, i.e., for a known positive constant $B_{\delta}\in\Re$, we have
\begin{gather}
\max\{\|\Delta_{x_{0}}\|,\,
\|\Delta_{R_{0}}\|,\,
\|[\Delta_{x_{1}}^T,\cdots\Delta_{x_n}^T]^T\|,\,\|\Delta_{R_i}\|\} < B_{\delta}.\label{eqn:Bdelta}
\end{gather}

Suppose that the desired trajectories for the position and the attitude of the payload are given as smooth functions of time, namely $x_{0_d}(t)\in\Re^3$ and $R_{0_d}(t)\in\SO$. From the attitude kinematics equation, we have
\begin{align*}
\dot R_{0_d}(t) = R_{0_d}(t) \hat\Omega_{0_d}(t),
\end{align*}
where $\Omega_{0_d}(t)\in\Re^3$ corresponds to the desired angular velocity of the payload. It is assumed that the velocity and the acceleration of the desired trajectories are bounded by known constants.

We wish to design a control input of each quadrotor $\{f_i,M_i\}_{1\leq i\leq n}$ such that  the tracking errors asymptotically converge to zero along the solution of the controlled dynamics.

\section{Control System Design For Simplified Dynamic Model}\label{sec:SDM}

In this section, we consider a simplified dynamic model where the attitude dynamics of each quadrotor is ignored, and we design a control input by assuming that the thrust at each quadrotor, namely $u_i$ can be arbitrarily chosen. It corresponds to the case where every quadrotor is replaced by a fully actuated aerial vehicle that can generates a thrust along any direction arbitrarily. The effects of the attitude dynamics of quadrotors will be incorporated in the next section.

In the simplified dynamic model given by \refeqn{ddotx0}-\refeqn{dotwi}, the dynamics of the payload are affected by the parallel components $u_i^\parallel$ of the thrusts, and the dynamics of the links are directly affected by the normal components $u_i^\perp$ of the thrusts. This structure motivates the following control system design procedure: first, the parallel components $u_i^\parallel$ are chosen such that the payload follows the desired position and attitude trajectory while yielding the desired direction of each link, namely $q_{i_d}\in\Sph^2$; next, the normal components $u_i^\perp$ are designed such that the actual direction of the links $q_i$ follows the desired direction $q_{i_d}$.

\subsection{Design of Parallel Components}

Let $a_i\in\Re^3$ be the acceleration of the point on the payload where the $i$-th link is attached, that is measured relative to the gravitational acceleration:
\begin{align}
a_i = \ddot x_0 - ge_3 + R_0\hat\Omega_0^2\rho_i -R_0\hat\rho_i\dot\Omega_0.\label{eqn:ai0}
\end{align}
The parallel component of the control input is chosen as
\begin{align}
u_i^\parallel & = \mu_i + m_il_i\|\omega_i\|^2q_i + m_i q_iq_i^T a_i,\label{eqn:uip}
\end{align}
where $\mu_i\in\Re^3$ is a virtual control input that is designed later, with a constraint that $\mu_i$ is parallel to $q_i$. Note that the expression of $u_i^\parallel$ is guaranteed to be parallel to $q_i$ due to the projection operator $q_iq_i^T$ at the last term of the right-hand side of the above expression.

The motivation for the proposed parallel components becomes clear if \refeqn{uip} is substituted into \refeqn{ddotx0}-\refeqn{dotW0} and rearranged to obtain
\begin{gather}
m_0 (\ddot x_0 -g e_3) = \Delta_{x_0}+\sum_{i=1}^n (\mu_i+\Delta_{x_i}^\parallel),\label{eqn:ddotx0s}\\
J_0\dot\Omega_0 +\hat\Omega_0 J_0\Omega_0 = \Delta_{R_0}+\sum_{i=1}^n \hat\rho_i R_0^T (\mu_i+\Delta_{x_i}^\parallel).\label{eqn:dotW0s}
\end{gather}
Therefore, considering a free-body diagram of the payload, the virtual control input $\mu_i$ corresponds to the force exerted to the payload by the $i$-link, or the tension of the $i$-th link in the absence of disturbances.

%When there is no control force from each quadrotor, i.e., $u_i^\parallel=0$, the tension of the $i$-th link is composed of the projected relative inertial force at the point where the $i$-th link is attached to the payload and the centrifugal force due to the rotation of the link. Substituting \refeqn{ddotx0s} and \refeqn{dotW0s} back into \refeqn{ai0}, we obtain
%\begin{align}
%a_i & = \frac{1}{m_0}(\Delta_{x_0}+\sum_{j=1}^n(\mu_j+\Delta_{x_j}^\parallel))+R_0\hat\Omega_0^2\rho_i\nonumber\\
%&\quad + R_0\hat\rho_iJ_0^{-1}(\hat\Omega_0 J_0\Omega_0-\Delta_{R_0} - \sum_{j=1}^n \hat\rho_j R_0^T(\mu_j+\Delta_{x_j}^\parallel)).\label{eqn:ai}
%%\\
%%u_i^\parallel & = \mu_i + m_il_i\|\omega_i\|^2q_i +\frac{m_i}{m_0}q_iq_i^T\sum_{j=1}^n\mu_j \nonumber\\
%%&\quad+m_iq_iq_i^T R_0 \{\hat\Omega_0^2 \rho_i+\hat\rho_iJ_0^{-1}(\hat\Omega_0 J_0\Omega_0 - \sum_{j=1}^n \hat\rho_j R_0^T\mu_j)\},\label{eqn:uip}
%\end{align}
%In short, the parallel component of the control input $u_i^\parallel$ is determined by the virtual control input $\mu_i$.

Next, we determine the virtual control input $\mu_i$. As in~\cite{GooLeePECC13}, define position, attitude, and angular velocity tracking error vectors $e_{x_0},e_{R_0},e_{\Omega_0}\in\Re^3$ for the payload as
\begin{align*}
e_{x_0} & = x_0 -x_{0_d},\\
e_{R_0} & = \frac{1}{2} (R_{0_d}^T R_0-R_0^T R_{0_d})^\vee,\\
e_{\Omega_0} & = \Omega_0 - R_0^T R_{0_d}\Omega_{0_d}.
\end{align*}
The desired resultant control force $F_d\in\Re^3$ and moment $M_d\in\Re^3$ acting on the payload are given as
\begin{align}
F_d & = m_0 (-k_{x_0}e_{x_0} - k_{\dot x_0} \dot e_{x_0} + \ddot x_{0_d} - g e_3)\nonumber\\
&\quad -\bar\Delta_{x_0} - \sum_{i=1}^n\bar\Delta_{x_i}^\parallel,\label{eqn:Fd}\\
M_d & = -k_{R_0} e_{R_0} - k_{\Omega_0} e_{\Omega_0}   +(R_0^TR_{0_d}\Omega_{0_d})^\wedge J_0 R_0^TR_{0_d} \Omega_{0_d}\nonumber\\
& \quad + J_0 R_0^T R_{0_d} \dot\Omega_{0_d}-\bar\Delta_{R_0}-\sum_{i=1}^n\hat\rho_iR_0 \bar\Delta_{x_i}^\parallel,\label{eqn:Md}
\end{align}
for positive constants $k_{x_0},k_{\dot x_0},k_{R_0},k_{\Omega_0}\in\Re$. Here, the estimates of $\Delta_{x_0},\Delta_{x_i},\Delta_{R_0}$ are denoted by $\bar\Delta_{x_0},\bar\Delta_{x_i},\bar\Delta_{R_0}\in\Re^3$, and the orthogonal projection of $\bar\Delta_{x_i}$ along $q_i$ is denoted by $\bar\Delta_{x_i}^\parallel = q_iq_i^T \bar\Delta_{x_i}\in\Re^3$. Adaptive control laws to update the estimates of disturbances are introduced later at Section \ref{sec:AL}.

These are the ideal resultant force and moment to achieve the control objectives. One may try to choose the virtual control input $\mu_i$ by making the expressions in the right-hand sides of \refeqn{ddotx0s} and \refeqn{dotW0s}, namely $\sum_{i}\mu_i$ and $\sum_i\hat\rho_iR_0^T\mu_i$, become identical to $F_d$ and $M_d$, respectively. But, this is not valid in general, as each $\mu_i$ is constrained to be parallel to $q_i$. Instead, we choose the desired value of $\mu_i$, without any constraint, such that 
\begin{align}
\sum_{i=1}^n \mu_{i_d} = F_d,\quad \sum_{i=1}^n \hat\rho_i R_0^T \mu_{i_d} = M_d,\label{eqn:muid0}
\end{align}
or equivalently, using the matrix $\mathcal{P}$ defined at \refeqn{P},
\begin{align*}
\mathcal{P}
\begin{bmatrix}
R_0^T\mu_{1_d} \\ \vdots \\ R_0^T\mu_{n_d}
\end{bmatrix}
=
\begin{bmatrix}
R_0^T F_d \\ M_d
\end{bmatrix}.
\end{align*}
From the assumption stated at \refeqn{A1}, there exists at least one solution to the above matrix equation for any $F_d, M_d$. Here, we find the minimum-norm solution given by
\begin{align}
\begin{bmatrix} \mu_{1_d}\\\vdots\\\mu_{n_d} \end{bmatrix} 
= \mathrm{diag}[R_0,\cdots R_0]\;\mathcal{P}^T (\mathcal{P}\mathcal{P}^T)^{-1}\begin{bmatrix} R_0^T F_d\\M_d\end{bmatrix}.\label{eqn:muid}
\end{align}
The virtual control input $\mu_i$ is selected as the projection of its desired value $\mu_{i_d}$ along $q_i$,
\begin{align}
\mu_i = (\mu_{i_d}\cdot q_i) q_i=q_iq_i^T\mu_{i_d},\label{eqn:mui}
\end{align}
and the desired direction of each link, namely $q_{i_d}\in\Sph^2$ is defined as
\begin{align}
q_{i_d} = -\frac{\mu_{i_d}}{\|\mu_{i_d}\|}.\label{eqn:qid}
\end{align}
It is straightforward to verify that when $q_{i}=q_{i_d}$, the resultant force and moment acting on the payload become identical to their desired values.

%Here, the extra degrees of freedom in control inputs are used to minimize the magnitude of the desired tension at \refeqn{muid}, but they can be applied to other tasks, such as controlling the relative configuration of links~\cite{LeeSrePICDC13}. This is referred to future investigation.

\subsection{Design of Normal Components}

Substituting \refeqn{ai0} into \refeqn{dotwi} and using \refeqn{delxiperp}, the equation of motion for the $i$-link is given by
\begin{align}
\dot\omega_i & = \frac{1}{l_i}\hat q_i a_i-\frac{1}{m_il_i}\hat q_i (u_i^\perp+\Delta_{x_i}^\perp).\label{eqn:widotf0}
\end{align}
Here, the normal component of the control input $u_i^\perp$ is chosen such that $q_i\rightarrow q_{i_d}$ as $t\rightarrow\infty$. Control systems for the unit-vectors on the two-sphere have been studied in~\cite{BulLew05,Wu12}. In this paper, we adopt the control system developed in terms of the angular velocity in~\cite{Wu12}, and we augment it with an adaptive control term to handle the disturbance $\Delta_{x_i}^\perp$.

For the given desired direction of each link, its desired angular velocity is obtained from the kinematics equation as
\begin{align*}
\omega_{i_d} = q_{i_d}\times \dot q_{i_d}.%\label{eqn:wid}
\end{align*}
Define the direction and the angular velocity tracking error vectors for the $i$-th link, namely $e_{q_i},e_{\omega_i}\in\Re^3$ as 
\begin{align*}
e_{q_i} & = q_{i_d}\times q_i,\\
e_{\omega_i} & = \omega_i + \hat q_i^2\omega_{i_d}.
\end{align*}
For positive constants $k_{q},k_{\omega}\in\Re$, the normal component of the control input is chosen as
\begin{align}
u_i^\perp & = m_il_i\hat q_i \{-k_q e_{q_i} -k_{\omega}e_{\omega_i} -(q_i\cdot\omega_{i_d})\dot q_i -\hat q_i^2\dot\omega_d\}\nonumber\\
&\quad - m_i\hat q_i^2 a_i- \bar\Delta_{x_i}^\perp,\label{eqn:uiperp}
\end{align}
where $\bar\Delta_{x_i}^\perp\in\Re^3$ corresponds to the component of the estimate $\bar\Delta_{x_i}$ that is projected onto the plane normal to $q_i$, i.e., $\bar\Delta_{x_i}^\perp=-\hat q_i^2 \bar\Delta_{x_i}$.

Note that the expression of $u_i^\perp$ is perpendicular to $q_i$ by definition. Substituting \refeqn{uiperp} into \refeqn{widotf0}, and rearranging by the facts that the matrix $-\hat q_i^2$ corresponds to the orthogonal projection to the plane normal to $q_i$ and $\hat q_i^3=-\hat q_i$, we obtain
\begin{align}
\dot\omega_i & = -k_q e_{q_i} -k_{\omega}e_{\omega_i} -(q_i\cdot\omega_{i_d})\dot q_i -\hat q_i^2\dot\omega_d\nonumber\\
&\quad - \frac{1}{m_il_i}\hat q_i \tilde\Delta_{x_i}^\perp,\label{eqn:dotwif}
\end{align}
where the estimation error is defined as $\tilde\Delta_{x_i}^\perp=\Delta_{x_i}^\perp-\bar\Delta_{x_i}^\perp\in\Re^3$.

In short, the control force for the simplified dynamic model is given by
\begin{align}
u_i = u_i^\parallel + u_i^\perp.\label{eqn:ui}
\end{align}

\subsection{Design of Adaptive Law}\label{sec:AL}
The above control inputs require estimates of the uncertainties. They are updated according to the following adaptive laws,
\begin{align}
\dot{\bar \Delta}_{x_0} & = \frac{h_{x_0}}{m_0}(\dot e_{x_0}+c_x e_{x_0}),\label{eqn:hatDx0_dot}\\
\dot{\bar \Delta}_{R_0} & = h_{R_0}(e_{\Omega_0}+ c_R e_{R_0}),\label{eqn:hatDR0_dot}\\
\dot{\bar \Delta}_{x_i} & = h_{x_i}q_iq_i^T\{\frac{1}{m_0}(\dot e_{x_0}+c_x e_{x_0})
-R_0\hat\rho_i(e_{\Omega_0}+c_R e_{R_0})\}\nonumber\\
&\quad +  \frac{h_{x_i}}{m_il_i}\hat q_i (e_{\omega_i} + c_q e_{q_i}),\label{eqn:hatDxi_dot}
\end{align}
for positive constants $c_x,c_R,c_q\in\Re$ and adaptive gains $h_{x_0},h_{R_0},h_{x_i}\in\Re$. Note that the adaptive law for $\Delta_{x_i}$ is composed of two parts: the first part that is parallel to $q_i$ determined by the tracking errors of the payload, and the second part that is normal to $q_i$ defined by the dynamics of the link.

The resulting stability properties are summarized as follows.

\begin{prop}\label{prop:SDM}
Consider the simplified dynamic model defined by \refeqn{ddotx0}-\refeqn{dotwi}. For given tracking commands $x_{0_d},R_{0_d}$, a control input is designed as \refeqn{ui}-\refeqn{hatDxi_dot}. Then, there exist the values of controller gains and controller parameters such that the following properties are satisfied.

\begin{itemize}
\item[(i)] The zero equilibrium of tracking errors $(e_{x_0},\dot e_{x_0},e_{R_0},e_{\Omega_0},e_{q_i},e_{\omega_i})$ and the estimation errors $(\bar\Delta_{x_0},\bar\Delta_{R_0},\bar\Delta_{x_i})$ is stable in the sense of Lyapunov.
\item[(ii)] The tracking errors asymptotically coverage to zero.
\item[(iii)] The estimation errors are uniformly bounded.
\end{itemize}
\end{prop}
\begin{proof}
Due to the page limit, the proof is relegated to~\cite{LeeACC151ext}.
\end{proof}

%At \refeqn{qid}, the negative sign appeared to make the tension at each cable positive when $q_i=q_{i_d}$.  Assuming that the tracking errors $e_{x_0},\dot e_{x_0},e_{R_0},e_{\Omega_0}$ and the variables $\ddot x_{0_d},\Omega_{0_d},\dot\Omega_{0_d}$ obtained from the desired trajectories are sufficiently small, this guarantees that quadrotors remain above the payload. If desired, the negative sign at \refeqn{qid} can be eliminated to place quadrotors below the payload, resulting in a tracking control of an inverted rigid body multi-link pendulum, that can be considered as a generalization of a flying spherical  inverted spherical pendulum illustrated in~\cite{LeeSrePICDC13}.

%The proposed control system guarantees stability in the sense of Lyapunov and asymptotic convergence of tracking errors variables, but as the convergence of estimation error is not guaranteed, asymptotic stability is not achieved. However, in the absence of the disturbances, we can achieve stronger exponential stability by eliminating the adaptive parts of the proposed control system via setting $\bar\Delta_{x_0}=\bar\Delta_{R_0}=\bar\Delta_{x_i}=0$ and $h_{x_0},h_{R_0},h_{x_i}=0$~\cite{LeePICDC14}.

\section{Control System Design for Full Dynamic Model}\label{sec:FDM}

The control system designed at the previous section is based on a simplifying assumption that each quadrotor can generate a thrust along any arbitrary direction instantaneously. However, the dynamics of quadrotor is underactuated since the direction of the total thrust is always parallel to its third body-fixed axis, while the magnitude of the total thrust can be arbitrarily changed. This can be directly observed from the expression of the total thrust, $u_i = -f_i R_i e_3$, where $f_i$ is the total thrust magnitude, and $R_ie_3$ corresponds to the direction of the third body-fixed axis. Whereas, the rotational attitude dynamics is fully actuated by the control moment $M_i$.

Based on these observations, the attitude of each quadrotor is controlled such that the third body-fixed axis becomes parallel to the direction of the ideal control force $u_i$ designed in the previous section within a finite time. More explicitly, the desired attitude of each quadrotor is constructed as follows. The desired direction of the third body-fixed axis of the $i$-th quadrotor, namely $b_{3_i}\in\Sph^2$ is given by
\begin{align}
b_{3_i} = -\frac{u_i}{\|u_i\|}.\label{eqn:b3i}
\end{align}
This provides two-dimensional constraint on the three-dimensional desired attitude of each quadrotor, and there remains one degree of freedom. To resolve it, the desired direction of the first body-fixed axis $b_{1_i}(t)\in\Sph^2$ is introduced as a smooth function of time~\cite{LeeLeoPICDC10}. This corresponds to controlling the additional one dimensional yawing angle of each quadrotor. From these, the desired attitude of the $i$-th quadrotor is given by
\begin{align*}
R_{i_c} = \begin{bmatrix}-\frac{(\hat b_{3_i})^2 b_{1_i}}{\|(\hat b_{3_i})^2 b_{1_i}\|}, &
\frac{\hat b_{3_i}b_{1_i}}{\|\hat b_{3_i}b_{1_i}\|},& b_{3_i}\end{bmatrix},
\end{align*}
which is guaranteed to be an element of $\SO$. The desired angular velocity is obtained from the attitude kinematics equation, $\Omega_{i_c} = (R_{i_c}^T\dot R_{i_c})^\vee\in\Re^3$.

In the prior work described in~\cite{LeePICDC14}, the attitude of each quadrotor is controlled such that the equilibrium $R_i=R_{i_c}$ becomes exponentially stable, and the stability of the combined full dynamic model is achieved via singular perturbation theory~\cite{Kha96}. However, we can not follow such approach in this paper, as the presented adaptive control system guarantees only the asymptotical convergence of the tracking error variables due to the disturbances, thereby making is challenging to apply the singular perturbation theory. Here, we design the attitude controller of each quadrotor such that $R_i$ becomes equal to $R_{i_c}$ within a finite time via finite-time stability theory~\cite{BhaBerSJCO00,YuYuA05,WuRadAA11}.

Define the tracking error vectors $e_{R_i},e_{\Omega_i}\in\Re^3$ for the attitude and the angular velocity of the $i$-th quadrotor as
\begin{align*}
e_{R_i} = \frac{1}{2}(R_{i_c}^T R_i -R_i^T R_{i_c})^\vee,\quad
e_{\Omega_i} = \Omega_i - R_i^T R_{i_c}\Omega_{i_c}.
\end{align*}
The time-derivative of $e_{R_i}$ can be written as~\cite{LeeLeoPICDC10}
\begin{align}
\dot e_{R_i} & = \frac{1}{2}(\trs{R_i^T R_{i_c}}I-R_i^T R_{i_c})e_{\Omega_i}\triangleq E(R_i,R_{i_c})e_{\Omega_i}.\label{eqn:eRi_dot}
\end{align}
For $0< r <1$, define $S:\Re\times\Re^3\rightarrow\Re^3$ as
\begin{align*}
S(r,y) = \begin{bmatrix}|y_1|^r\mathrm{sgn}(y_1),&
|y_2|^r\mathrm{sgn}(y_2),&
|y_3|^r\mathrm{sgn}(y_3)\end{bmatrix}^T,
\end{align*}
where $y=[y_1,y_2,y_3]^T\in\Re^3$, and $\mathrm{sgn}(\cdot)$ denotes the sign function. For positive constants $k_{R},l_{R}$, the terminal sliding surface $s_i\in\Re^3$ is designed as
\begin{align}
s_i = e_{\Omega_i} + k_{R} e_{R_i} + l_{R}S(r,e_{R_i}).\label{eqn:si}
\end{align}
We can show that when confined to the surface of $s_i\equiv 0$, the tracking errors become zero in a finite time. To reach the sliding surface, for positive constants $k_s,l_s$, the control moment is designed as
\begin{align}
M_i & = -k_s s_i - l_s S(r,s_i)+\Omega_i\times J_i\Omega_i\nonumber\\
&\quad - (k_R J_i+ l_s r J_i\mathrm{diag}_j[|e_{R_{i_j}}|^{r-1}]) E(R_i,R_{c_i}) e_{\Omega_i}\nonumber\\
&\quad  -J_i(\hat\Omega_i R_i^T R_{i_c} \Omega_{i_c} - R_i^T R_{i_c}\dot\Omega_{i_c}).\label{eqn:Mi}
\end{align}

The thrust magnitude is chosen as the length of $u_i$, projected on to $-R_ie_3$,
\begin{align}
f_i & = - u_i\cdot R_i e_3,\label{eqn:fi}
\end{align}
which yields that the thrust of each quadrotor becomes equal to its desired value $u_i$ when $R_i=R_{i_c}$.

Stability of the corresponding controlled systems for the full dynamic model can be shown by using the fact that the full dynamic model becomes exactly same as the simplified dynamic model within a finite time.

\begin{prop}\label{prop:FDM}
Consider the full dynamic model defined by \refeqn{ddotx0}-\refeqn{dotWi}. For given tracking commands $x_{0_d},R_{0_d}$ and the desired direction of the first body-fixed axis $b_{1_i}$, control inputs for quadrotors are designed as \refeqn{Mi} and \refeqn{fi}. Then, there exists controller parameters such that the tracking error variables 
$(e_{x_0},\dot e_{x_0},e_{R_0},e_{\Omega_0},e_{q_i},e_{\omega_i})$ asymptotically converge to zero, and the estimation errors are uniformly bounded.
\end{prop}
\begin{proof}
See Appendix \ref{sec:prfFDM}.
\end{proof}

This implies that the payload asymptotically follows any arbitrary desired trajectory both in translations and rotations in the presence of uncertainties. In contrast to the existing results in aerial transportation of a cable suspended load, it does not rely on any simplifying assumption that ignores the coupling between payload, cable, and quadrotors. Also, the presented global formulation on the nonlinear configuration manifold avoids singularities and complexities that are inherently associated with local coordinates. As such, the presented control system is particularly useful for agile load transportation involving combined translational and rotational maneuvers of the payload in the presence of uncertainties.

\section{Numerical Example}\label{sec:NE}
We consider a numerical example where three quadrotors ($n=3$) transport a rectangular box along a figure-eight curve. More explicitly, the mass of the payload is $m_0=1.5\,\mathrm{kg}$, and its length, width, and height are $1.0\,\mathrm{m}$, $0.8\,\mathrm{m}$, and $0.2\,\mathrm{m}$, respectively. Mass properties of three quadrotors are identical, and they are given by
\begin{gather*}
m_i=0.755\,\mathrm{kg},\quad J_i=\mathrm{diag}[0.0820,\, 0.0845,\, 0.1377]\,\mathrm{kgm^2}.
\end{gather*}
The length of cable is $l_i=1\,\mathrm{m}$, and they are attached to the following points of the payload.
\begin{gather*}
\rho_1 = [0.5,\,0,\,-0.1]^T,\\
\rho_1 = [-0.5,\,0.4,\,-0.1]^T,\quad
\rho_3 = [-0.5,\,-0.4,\,-0.1]^T\,\mathrm{m}.
\end{gather*}
In other words, the first link is attached to the center of the top, front edge, and the remaining two links are attached to the vertices of the top, rear edge (see Figure \ref{fig:DM}).

\begin{figure}
\centerline{
	\subfigure[3D perspective]{
		\includegraphics[width=0.95\columnwidth]{Quad_N_Rgbd_Adaptive_3d0}}
}
\centerline{
	\subfigure[Top view]{
	\setlength{\unitlength}{0.1\columnwidth}\scriptsize
	\begin{picture}(9.5,3.2)(0,0)
	\put(0,0){\includegraphics[width=0.95\columnwidth]{Quad_N_Rgbd_Adaptive_3d2}}
	\put(8.4,3.2){$t=0$}
	\put(6.0,3.2){$t=3.3$}
	\put(2.2,-0.2){$t=6.6$}
	\put(0.2,2.3){$t=10$}
	\put(2.2,3.2){$t=13.3$}
	\put(5.8,-0.2){$t=16.6$}
	\put(7.9,0.4){$t=20$}
	\end{picture}}	
}
\centerline{
	\subfigure[Side view]{
		\includegraphics[width=0.95\columnwidth]{Quad_N_Rgbd_Adaptive_3d1}}
}
\caption{Snapshots of controlled maneuver (red:desired trajectory, blue:actual trajectory). A short animation illustrating this maneuver is available at {\href{http://youtu.be/nOWErfdzZLU}{http://youtu.be/nOWErfdzZLU}}.}\label{fig:SS}
\end{figure}

The desired trajectory of the payload is chosen as
\begin{align*}
x_{0_d}(t) = [1.2\sin (0.2\pi t),\, 4.2\cos (0.1\pi t),\, -0.5]^T\,\mathrm{m}.
\end{align*}
The desired attitude of the payload is chosen such that its first axis is tangent to the desired path, and the third axis is parallel to the direction of gravity, it is given by
\begin{align*}
R_{0_d}(t) = \begin{bmatrix}\frac{\dot x_{0_d}}{\|\dot x_{0_d}\|}&\frac{\hat e_3\dot x_{0_d}}{\|\hat e_3\dot x_{0_d}\|}& e_3\end{bmatrix}.
\end{align*}
Initial conditions are chosen as
\begin{gather*}
x_0(0)=[1,\,4.8,\,0]^T\,\mathrm{m},\quad v_0(0)=0_{3\times 1}\,\mathrm{m/s},\\
q_i(0)= e_3,\; \omega_i(0)=0_{3\times 1},\;
R_i(0)=I_{3\times 3},\; \Omega_i(0)=0_{3\times 1}.
\end{gather*}
The uncertainties are specified as
\begin{align*}
\Delta_{x_0}=[1,\, 3,\, -2.5]^T,\quad \Delta_{R_0}=[-0.5,\, 0.1,\, -1.5]^T,\\
\Delta_{x_i}=[0.5,\, -0.2,\, 0.3]^T,\quad \Delta_{R_i}=[0.2,\, 0.3,\, -0.7]^T.
\end{align*}

\begin{figure}
\centerline{
	\subfigure[Position of payload ($x_0$:blue, $x_{0_d}$:red)]{
		\includegraphics[width=0.48\columnwidth]{Quad_N_Rgbd_Adaptive_x0}}
	\hfill
	\subfigure[Attitude tracking error of payload $\Psi_0=\frac{1}{2}\|R_0-R_{0_d}\|^2$]{
		\includegraphics[width=0.48\columnwidth]{Quad_N_Rgbd_Adaptive_Psi0}}
}
\centerline{
	\subfigure[Link direction error $\Psi_{q_i}=\frac{1}{2}\|q_i-q_{i_d}\|^2$]{
		\includegraphics[width=0.48\columnwidth]{Quad_N_Rgbd_Adaptive_Psiq}}
	\hfill
	\subfigure[Attitude tracking error of quadrotors $\Psi_i=\frac{1}{2}\|R_i-R_{i_d}\|^2$]{
		\includegraphics[width=0.48\columnwidth]{Quad_N_Rgbd_Adaptive_Psi}}
}
\centerline{
	\subfigure[Tension at links]{
		\includegraphics[width=0.45\columnwidth]{Quad_N_Rgbd_Adaptive_Tension}}
	\hfill
	\subfigure[Control input for quadrotors $f_i,M_i$]{
		\includegraphics[width=0.48\columnwidth]{Quad_N_Rgbd_Adaptive_fM}}
}
\caption{Simulation results for tracking errors and control inputs. (for figures (c)-(f): $i=1$:blue, $i=2$:green, $i=3$:red)}\label{fig:SR}
\end{figure}

The corresponding simulation results are presented at Figures \ref{fig:SS} and \ref{fig:SR}. Figure \ref{fig:SS} illustrates the desired trajectory that is shaped like a figure-eight curve around two obstacles represented by cones, and the actual maneuver of the payload and quadrotors. Figure \ref{fig:SR} shows tracking errors for the position and the attitude of the payload, tracking errors for the link directions and the attitude of quadrotors, as well as tension and control inputs. These illustrate excellent tracking performances of the proposed control system.

\appendix
%
%
%
%
%\paragraph{Error Dynamics}
%
%From \refeqn{ddotx0} and \refeqn{mui}, the dynamics of the position tracking error is given by
%\begin{align*}
%m_0\ddot e_{x_0} & = m_0(ge_3-\ddot x_{0_d}) + \Delta_{x_0} + \sum_{i=1}^n q_iq_i^T \mu_{i_d}+ \Delta_{x_i}^\parallel.
%\end{align*}
%From \refeqn{muid0} and \refeqn{Fd}, this can be rearranged as
%\begin{align}
%\ddot e_{x_0} & = ge_3-\ddot x_{0_d} + \frac{1}{m_0}F_d  +Y_x+\frac{1}{m_0}(\tilde\Delta_{x_0}+\sum_{i=1}^n \tilde\Delta_{x_i}^\parallel),\nonumber\\
%& = -k_{x_0}e_{x_0} - k_{\dot x_0} \dot e_{x_0} +\frac{1}{m_0}(\tilde\Delta_{x_0}+\sum_{i=1}^n \tilde\Delta_{x_i}^\parallel) +Y_x,
%\label{eqn:ddotex0}
%\end{align}
%where $\tilde\Delta_{x_i}^\parallel= \Delta_{x_i}^\parallel-\bar\Delta_{x_i}^\parallel\in\Re^{3}$ corresponds to the parallel component of the estimation error. The last term $Y_x\in\Re^3$ represents the error caused by the difference between $q_i$ and $q_{i_d}$, and it is given by
%\begin{align*}
%Y_x=\frac{1}{m_0}\sum_{i=1}^n (q_iq_i^T -I) \mu_{i_d}.
%\end{align*}
%We have $\mu_{i_d}=q_{i_d}q_{i_d}^T\mu_{i_d}$ from \refeqn{qid}. Using this, the error term can be written in terms of $e_{q_i}$ as
%\begin{align*}
%Y_x  & = \frac{1}{m_0}\sum_{i=1}^n %q_iq_i^Tq_{i_d}q_{i_d}^T\mu_{i_d} -q_{i_d}q_{i_d}^T\mu_{i_d}\\
% (q_{i_d}^T\mu_{i_d})\{(q_i^Tq_{i_d})q_i-q_{i_d}\}\nonumber\\
%% = (q_{i_d}^T\mu_{i_d}) \{q_i\times (q_i\times q_{i_d})\}
%& = -\frac{1}{m_0}\sum_{i=1}^n (q_{i_d}^T\mu_{i_d}) \hat q_i e_{q_i}.%\label{eqn:Yx}
%\end{align*}
%Using \refeqn{muid}, an upper bound of $Y_x$ can be obtained as
%\begin{align*}
%\|Y_x\| & \leq \frac{1}{m_0} \sum_{i=1}^n \|\mu_{i_d}\|\|e_{q_i}\|\leq  \sum_{i=1}^n\gamma (\|F_d\|+\|M_d\|)\|e_{q_i}\|,
%\end{align*}
%where $\gamma= \frac{1}{m_0\sqrt{\lambda_{m}[\mathcal{P}\mathcal{P}^T]}}$. From \refeqn{Fd} and \refeqn{Md}, this can be further bounded by
%\begin{align}
%\|Y_x\|  \leq & \sum_{i=1}^n \{\beta(k_{x_0}\|e_{x_0}\|+k_{\dot x_0}\|\dot e_{x_0}\|)\nonumber\\
%&\quad  + \gamma(k_{R_0} \|e_{R_0}\| + k_{\Omega_0}\|e_{\Omega_0}\|) + B\}\|e_{q_i}\|,\label{eqn:YxB}
%\end{align}
%where $\beta=m_0\gamma$, and the constant $B$ is determined by the given desired trajectories of the payload and \refeqn{D}, which defines the domain $D$ of the error variables that the presented stability proof is considered. Throughout the remaining parts of the proof, any bound that can be obtained from $x_{0_d},R_{0_d}$ or \refeqn{D} is denoted by $B$ for simplicity. In short, the position tracking error dynamics of the payload can be written as \refeqn{ddotex0}, where the error term is bounded by \refeqn{YxB}.
%
%Similarly, we find the attitude tracking error dynamics for the payload as follows. Using \refeqn{dotW0}, \refeqn{Md}, and \refeqn{mui}, the time-derivative of $J_0 e_{\Omega_0}$ can be written as
%\begin{align}
%J_0\dot e_{\Omega_0} & = (J_0e_{\Omega_0}+d)^\wedge e_{\Omega_0} - k_{R_0}e_{R_0}-k_{\Omega_0}e_{\Omega_0}\nonumber\\
%&\quad + \tilde\Delta_{R_0} + \sum_{i=1}^n \hat\rho_i R_0^T\tilde\Delta_{x_i}^\parallel+Y_R,\label{eqn:doteW0}
%\end{align}
%where $d=(2J_0-\trs{J_0}I)R_0^T R_{0_d}\Omega_{0_d}\in\Re^3$~\cite{GooLeePECC13} that is bounded, and $\tilde\Delta_{R_0}\in\Re^{3}$ denotes the estimation error given by $\tilde\Delta_{R_0} = \Delta_{R_0}-\hat\Delta_{R_0}$. The error term in the attitude dynamics of the payload, namely $Y_R\in\Re^3$ is given by
%\begin{align*}
%Y_R & = \sum_{i=1}^n\hat\rho_i R_0^T (q_iq_i^T - I)\mu_{i_d}
%= -\sum_{i=1}^n\hat\rho_i R_0^T (q_{i_d}^T \mu_{i_d}) \hat q_i e_{q_i}.%\label{eqn:YR}
%\end{align*}
%Similar with \refeqn{YxB}, an upper bound of $Y_R$ can be obtained as
%\begin{align}
%\|Y_R\|  \leq & \sum_{i=1}^n\{\delta_i (k_{x_0}\|e_{x_0}\|+k_{\dot x_0}\|\dot e_{x_0}\|)\nonumber\\
%&\quad  + \sigma_i(k_{R_0} \|e_{R_0}\| + k_{\Omega_0}\|e_{\Omega_0}\|) + B\}\|e_{q_i}\|,\label{eqn:YRB}
%\end{align}
%where $\delta_i = m_0\frac{\|\hat\rho_i\|}{\sqrt{\lambda_{m}[\mathcal{P}\mathcal{P}^T]}},\sigma_i=\frac{\delta_i}{m_0}\in\Re$. 
%
%
%Next, from \refeqn{dotwif}, the time-derivative of the angular velocity error, projected on to the plane normal to $q_i$ is given as
%\begin{align}
%-\hat q_i^2 \dot e_{\omega_i} %& = -\hat q_i^2(\dot\omega_i + (\dot q_i\cdot\omega_{i_d}) q_i
%%+(q_i\cdot\dot\omega_{i_d}) q_i+( q_i\cdot\omega_{i_d}) \dot q_i - \dot \omega_{i_d})
%& = \dot\omega +( q\cdot\omega_d) \dot q +\hat q^2 \dot \omega_d\nonumber\\
%& = -k_q e_{q_i} - k_\omega e_{\omega_i}-\frac{1}{m_il_i}\hat q_i \tilde\Delta_{x_i}^\perp.\label{eqn:dotewi}
%\end{align}
%In summary, the error dynamics of the simplified dynamic model are given by \refeqn{ddotex0}, \refeqn{doteW0} and \refeqn{dotewi}.
%
%\paragraph{Stability Proof}
%
%Define an attitude configuration error function $\Psi_{R_0}$ for the payload as
%\begin{align*}
%\Psi_{R_0} = \frac{1}{2}\trs{I-R_{0_d}^T R_0},
%\end{align*}
%which is positive-definite about $R_0=R_{0_d}$, and $\dot\Psi_{R_0} = e_{R_0}\cdot e_{\Omega_0}$~\cite{LeeLeoPICDC10,GooLeePECC13}. We also introduce a configuration error function $\Psi_{q_i}$ for each link that is positive-definite about $q_i=q_{i_d}$ as
%\begin{align*}
%\Psi_{q_i} = 1-q_{i}\cdot q_{i_d}.
%\end{align*}
%For positive constants $e_{x_{\max}}, \psi_{R_0}, \psi_{q_i}, B_\delta\in\Re$, consider the following open domain containing the zero equilibrium of tracking error variables:
%\begin{align}
%D=\{&(e_{x_0},\dot e_{x_0}, e_{R_0}, e_{\Omega_0}, e_{q_i}, e_{\omega_i},\tilde\Delta_{x_0},\tilde\Delta_{R_0},\tilde\Delta_{x_i})\nonumber\\
%& \in(\Re^3)^4\times (\Re^3\times\Re^3)^n\times (\Re^3)^2\times \Re^{3n}\,|\,\nonumber\\
%& \|e_{x_0}\|< e_{x_{\max}},\,\Psi_{R_0}< \psi_{R_0}<1,\,\Psi_{q_i}< \psi_{q_i}<1,\nonumber\\
%& \|\tilde\Delta_{x_0}\| < B_\delta,\, \|\tilde\Delta_{R_0}\|< B_\delta,\, \|\tilde\Delta_{x_i}\| < B_\delta\}.\label{eqn:D}
%\end{align}
%In this domain, we have $\|e_{R_0}\|=\sqrt{\Psi_{R_0}(2-\Psi_{R_0})} \leq \sqrt{\psi_{R_0}(2-\psi_{R_0})} \triangleq \alpha_0 < 1$, and $\|e_{q_i}\|=\sqrt{\Psi_{q_i}(2-\Psi_{q_i})} \leq \sqrt{\psi_{q_i}(2-\psi_{q_i})} \triangleq \alpha_i < 1$. It is assumed that $\psi_{q_i}$ is sufficiently small such that $n\alpha_i\beta < 1$. 
%
%We can show that the configuration error functions are quadratic with respect to the error vectors in the sense that
%\begin{gather*}
%\frac{1}{2}\|e_{R_0}\|^2 \leq \Psi_{R_0} \leq \frac{1}{2-\psi_{R_0}} \|e_{R_0}\|^2,\\
%\frac{1}{2}\|e_{q_i}\|^2 \leq \Psi_{q_i} \leq \frac{1}{2-\psi_{q_i}} \|e_{q_i}\|^2,
%\end{gather*}
%where the upper bounds are satisfied only in the domain $D$. 
%
%Define
%\begin{align*}
%\mathcal{V}_0 & = \frac{1}{2}\|\dot e_{x_0}\|^2 + \frac{1}{2}k_{x_0}\|e_{x_0}\|^2   +c_x e_{x_0}\cdot \dot e_{x_0}\\
%&\quad + \frac{1}{2}e_{\Omega_0}\cdot J_0\Omega_0 + k_{R_0}\Psi_{R_0} + c_R e_{R_0}\cdot J_0e_{\Omega_0}\\
%&\quad+ \sum_{i=1}^n \frac{1}{2}\|e_{\omega_i}\|^2 + k_q\Psi_{q_i} + c_q e_{q_i}\cdot e_{\omega_i},
%\end{align*}
%where $c_x,c_R,c_q$ are positive constants. This is composed of tracking error variables only, and we define another function for the estimation errors of the adaptive laws as
%\begin{align*}
%\mathcal{V}_a & = \frac{1}{2h_{x_0}}\|\tilde \Delta_{x_0}\|^2
%+\frac{1}{2h_{R_0}}\|\tilde \Delta_{R_0}\|^2
%+ \sum_{i=1}^n \frac{1}{2h_{x_i}} \|\tilde\Delta_{x_i}\|^2.
%\end{align*}
%The Lyapunov function for the complete simplified dynamic model is chosen as $\mathcal{V}=\mathcal{V}_0+\mathcal{V}_a$.
%
%Let $z_{x_0} = [\|e_{x_0}\|,\|\dot e_{x_0}\|]^T$, $z_{R_0}=[\|e_{R_0}\|,\|e_{\Omega_0}\|]^T$, $z_{q_i}= [\|e_{q_i}\|,\|e_{\omega_i}\|]^T\in\Re^2$. The first part of the Lyapunov function $\mathcal{V}_0$ satisfies
%\begin{align*}
%z_{x_0}^T &\underline{P}_{x_0}z_{x_0} + 
%z_{R_0}^T \underline{P}_{R_0}z_{R_0} + 
%\sum_{i=1}^n z_{q_i}^T \underline{P}_{q_i}z_{q_i}\leq\mathcal{V}_0\\
%&\leq z_{x_0}^T \overline{P}_{x_0}z_{x_0} + 
%z_{R_0}^T \overline{P}_{R_0}z_{R_0} + 
%\sum_{i=1}^n z_{q_i}^T \overline{P}_{q_i}z_{q_i},
%\end{align*}
%where the matrices $\underline{P}_{x_0},\underline{P}_{R_0},\underline{P}_{q_i},\underline{P}_{x_0},\underline{P}_{R_0},\underline{P}_{q_i}\in\Re^{2\times 2}$ are given by
%\begin{alignat*}{2}
%\underline{P}_{x_0}&=\frac{1}{2}\begin{bmatrix} k_{x_0} & -c_x\\-c_x & 1\end{bmatrix},&\;
%\overline{P}_{x_0}&=\frac{1}{2}\begin{bmatrix} k_{x_0} & c_x\\c_x & 1\end{bmatrix},\\
%\underline{P}_{R_0}&=\frac{1}{2}\begin{bmatrix} 2k_{R_0} & -c_R\overline\lambda\\-c_R\overline\lambda & \underline\lambda\end{bmatrix},&
%\overline{P}_{R_0}&=\frac{1}{2}\begin{bmatrix} \frac{2k_{R_0}}{2-\psi_{R_0}} & c_R\overline\lambda\\c_R\overline\lambda & \overline\lambda\end{bmatrix},\\
%\underline{P}_{q_i}&=\frac{1}{2}\begin{bmatrix} 2k_{q} & -c_q\\-c_q & 1\end{bmatrix},&
%\overline{P}_{q_i}&=\frac{1}{2}\begin{bmatrix} \frac{2k_{q}}{2-\psi_{q_i}} & c_q\\c_q & 1\end{bmatrix},
%\end{alignat*}
%where $\underline\lambda=\lambda_{m}[J_0]$ and $\overline\lambda=\lambda_{M}[J_0]$. If the constants $c_x,c_{R_0},c_q$ are sufficiently small, all of the above matrices are positive-definite. As the second part of the Lyapunov function $\mathcal{V}_a$ is already given as a quadratic form, it is straightforward to see that the complete Lyapunov function $\mathcal{V}$ is positive-definite and decrescent. 
%
%The time-derivative of the Lyapunov function along the error dynamics \refeqn{ddotex0}, \refeqn{doteW0}, and \refeqn{dotewi} is given by
%\begin{align}
%\dot{\mathcal{V}} & = -(k_{\dot x_0}-c_x) \|\dot e_{x_0}\|^2 
%-c_xk_{x_0} \|e_{x_0}\|^2 -c_xk_{\dot x_0} e_{x_0}\cdot \dot e_{x_0}\nonumber\\
%&\quad+ (c_x e_{x_0}+\dot e_{x_0})\cdot Y_x -k_{\Omega_0}\|e_{\Omega_0}\|^2 + c_R \dot e_R\cdot J_0e_{\Omega_0}\nonumber\\
%&\quad -c_R k_{R_0}\|e_{R_0}\|^2 + c_R e_{R_0}\cdot ((J_0e_{\Omega_0}+d)^\wedge e_{\Omega_0}-k_{\Omega_0} e_{\Omega_0})\nonumber\\
%&\quad + (e_{\Omega_0}+c_R e_{R_0}) \cdot Y_R\nonumber\\
%&\quad +\sum_{i=1}^n -(k_\omega-c_q) \|e_{\omega_i}\|^2 -c_qk_q \|e_{q_i}\|^2 -c_q k_\omega e_{q_i}\cdot e_{\omega_i}\nonumber\\
%&\quad + \frac{1}{m_0}(\dot e_{x_0}+c_xe_{x_0})\cdot (\tilde\Delta_{x_0} +  \sum_{i=1}^n \tilde\Delta_{x_i}^\parallel) \nonumber\\
%&\quad +(e_{\Omega_0}+ c_Re_{R_0})\cdot (\tilde\Delta_{R_0}+\sum_{i=1}^n \hat\rho_i R_0^T \tilde\Delta_{x_i}^\parallel)\nonumber\\
%&\quad -\parenth{\sum_{i=1}^n ( e_{\omega_i} + c_qe_{q_i})\cdot \frac{\hat q_i}{m_il_i} \tilde\Delta_{x_i}^\perp}-\frac{1}{h_{x_0}}\tilde\Delta_{x_0}\cdot \dot{\bar\Delta}_{x_0}\nonumber\\
%&\quad 
%-\frac{1}{h_{R_0}}\tilde\Delta_{R_0}\cdot\dot{\bar\Delta}_{R_0}
%-\sum_{i=1}^n\frac{1}{h_{x_i}}(\tilde\Delta_{x_i}^\parallel\cdot \dot{\bar\Delta}_{x_i}^\parallel+\tilde\Delta_{x_i}^\perp\cdot \dot{\bar\Delta}_{x_i}^\perp),
%\label{eqn:dotV0}
%\end{align}
%where the last term has been obtained using the facts that $\tilde\Delta_{x_i}^\perp\cdot \dot{\bar \Delta}_{x_i}^\parallel
%=\tilde\Delta_{x_i}^\parallel\cdot \dot{\bar \Delta}_{x_i}^\perp=0$. Substituting the adaptive laws given by \refeqn{hatDx0_dot}, \refeqn{hatDR0_dot}, and \refeqn{hatDxi_dot} into \refeqn{dotV0}, the expressions at the last four lines of \refeqn{dotV0} that are dependent of estimation errors vanish.
%
%An upper bound of the remaining expressions of $\dot{\mathcal{V}}$ can be obtained as follows.
%Since $\|e_{R_0}\|\leq 1$, $\| \dot e_{R_0}\|\leq \|e_{\Omega_0}\|$ and $\|d\|\leq B$, 
%\begin{align}
%\dot{\mathcal{V}} & \leq -(k_{\dot x_0}-c_x) \|\dot e_{x_0}\|^2 
%-c_xk_{x_0} \|e_{x_0}\|^2 -c_xk_{\dot x_0} e_{x_0}\cdot \dot e_{x_0}\nonumber\\
%&\quad+ (c_x e_{x_0}+\dot e_{x_0})\cdot Y_x -(k_{\Omega_0}-2c_R\overline\lambda)\|e_{\Omega_0}\|^2 \nonumber\\
%&\quad -c_R k_{R_0}\|e_{R_0}\|^2 + c_R(k_{\Omega_0}+B) \|e_{R_0}\|\|e_{\Omega_0}\|\nonumber\\
%&\quad + (e_{\Omega_0}+c_R e_{R_0}) \cdot Y_R\nonumber\\
%&\quad +\sum_{i=1}^n -(k_\omega-c_q) \|e_{\omega_i}\|^2 -c_qk_q \|e_{q_i}\|^2 -c_q k_\omega e_{q_i}\cdot e_{\omega_i}.
%\label{eqn:dotV2}
%\end{align}
%From \refeqn{YxB}, an upper bound of the fourth term of the right-hand side is given by
%\begin{align}
%\|&(c_x e_{x_0}+\dot e_{x_0})\cdot Y_x\|  \leq  \nonumber\\
%&\sum_{i=1}^n\alpha_i\beta( c_xk_{x_0}\|e_{x_0}\|^2 + c_x k_{\dot x_0}\|e_{x_0}\|\|\dot e_{x_0}\| +k_{\dot x_0} \|\dot e_{x_0}\|^2)
%\nonumber\\
%&+\{c_xB \|e_x\|+(\beta k_{x_0}e_{x_{\max}}+B) \|\dot e_{x_0}\| \} \|e_{q_i}\|\nonumber\\
%&+\alpha_i\gamma(c_x\|e_{x_0}\|+\|\dot e_{x_0}\|)(k_{R_0}\|e_{R_0}\|+k_{\Omega_0}\|e_{\Omega_0}\|).
%\label{eqn:YxB2}
%\end{align}
%Similarly, using \refeqn{YRB},
%\begin{align}
%\|&(c_R e_{R_0}+e_{\Omega_0})\cdot Y_R\|\leq \nonumber\\
%&\sum_{i=1}^n \alpha_i\sigma_i( c_R k_{R_0}\|e_{R_0}\|^2 + c_R k_{\Omega_0}\|e_{R_0}\|\|e_{\Omega_0}\| + k_{\Omega_0}\|e_{\Omega_0}\|^2)\nonumber\\
%&+\{c_R B\|e_{R_0}\| + (\alpha_0\sigma_ik_{R_0}+B)\|e_{\Omega_0}\|\}\|e_{q_i}\|\nonumber\\
%&+\alpha_i\delta_i(c_R \|e_{R_0}\|+\|e_{\Omega_0}\|)(k_{x_0}\|e_{x_0}\|+k_{\dot x_0}\|\dot e_{x_0}\|).
%\end{align}
%Substituting these into \refeqn{dotV2} and rearranging, 
%\begin{align}
%\dot{\mathcal{V}} \leq \sum_{i=1}^n - z_i^T W_i z_i,\label{eqn:U0}
%\end{align}
%where $z_i=[\|z_{x_0}\|,\, \|z_{R_0}\|,\,\ \|z_{q_i}\|]^T\in\Re^3$, and the matrix $W_i\in\Re^{3\times 3}$ is defined as
%\begin{align}
%W_i = 
%\begin{bmatrix} \lambda_m[W_{x_i}] & -\frac{1}{2}\|W_{xR_i}\| & -\frac{1}{2}\|W_{xq_i}\|\\
%-\frac{1}{2}\|W_{xR_i}\| & \lambda_m[W_{R_i}] & -\frac{1}{2}\|W_{Rq_i}\| \\
%-\frac{1}{2}\|W_{xq_i}\| & -\frac{1}{2}\|W_{Rq_i}\| & \lambda_m[W_{q_i}]
%\end{bmatrix},\label{eqn:Wi}
%\end{align}
%where the sub-matrices are given by
%\begin{gather*}
%W_{x_i} = \frac{1}{n}\begin{bmatrix}
%c_x k_{x_0} (1-n\alpha_i\beta) & -\frac{c_x k_{\dot x_0}}{2}(1+n\alpha_i\beta)\\
%-\frac{c_x k_{\dot x_0}}{2}(1+n\alpha_i\beta) & k_{\dot x_0}(1-n\alpha_i\beta)-c_x 
%\end{bmatrix},\\
%W_{R_i} = \frac{1}{n}\begin{bmatrix}
%c_R k_{R_0} (1-n\alpha_i\sigma_i) & -\frac{c_R }{2}(k_{\Omega_0}+B+n\alpha_i\sigma_i)\\
%-\frac{c_R }{2}(k_{\Omega_0}+B+n\alpha_i\sigma_i) & k_{\Omega_0}(1-n\alpha_i\sigma_i)-2c_R\overline\lambda
%\end{bmatrix},\\
%W_{q_i}=\begin{bmatrix}
%c_q k_q & -\frac{c_qk_\omega}{2}\\
%-\frac{c_qk_\omega}{2} & k_\omega-c_q\\
%\end{bmatrix},\\
%W_{xR_i}= \alpha_i\begin{bmatrix}
%\gamma c_x k_{R_0}+\delta_i c_Rk_{x_0} 
%& \gamma c_x k_{\Omega_0}+\delta_i k_{x_0}\\
%\gamma k_{R_0}+\delta_ic_Rk_{\dot x_0}  &  
%\gamma k_{\Omega_0} + \delta_i k_{\dot x_0}
%\end{bmatrix},\\
%W_{xq_i}= \begin{bmatrix}
%c_x B & 0\\
%\beta k_{x_0}e_{x_{\max}}+B & 0 
%\end{bmatrix},\quad
%W_{xR_i}= \begin{bmatrix}
%c_R B & 0\\
%\alpha_0\sigma_i k_{R_0} +B & 0 
%\end{bmatrix}.
%\end{gather*}
%
%%On the other hand, substituting the adaptive laws given by \refeqn{hatDx0_dot}, \refeqn{hatDxi_dot}, and \refeqn{hatDR0_dot} into \refeqn{dotV0}, the second part of $\dot{\mathcal{V}}$ that is dependent of the estimation errors is given by
%%\begin{align}
%%\mathcal{U}_a & = \tilde\Delta_{x_0} \cdot (I_{3\times 3}-\mathrm{Proj}[\bar \Delta_{x_0},\phi_{x_0}]) \phi_{x_0}\nonumber\\
%%&\quad + \sum_{i=1}^n\tilde\Delta_{x_i} \cdot (I_{3\times 3}-\mathrm{Proj}[\bar \Delta_{x_i},\phi_{x_i}])\phi_{x_i}\nonumber\\
%%&\quad + \tilde\Delta_{R_0} \cdot (I_{3\times 3}-\mathrm{Proj}[\bar \Delta_{R_0},\phi_{R_0}])\phi_{R_0}.
%%\label{eqn:Ua}
%%\end{align}
%%Note that from the definition of the projection operator given at \refeqn{Proj}, for some $\tilde\Delta,\bar\Delta,\phi\in\Re^3$, we have
%%\begin{align*}
%%&\tilde\Delta \cdot (I_{3\times 3}-\mathrm{Proj}[\bar \Delta,\phi]) \phi\\
%%&=\begin{cases}
%%0\;	&\text{if $\|\bar\Delta\| \leq B_{\delta}$}\\
%%    &\text{or $\|\bar\Delta\| = B_{\delta}$ and $\bar\Delta^T\phi \leq 0$,}\\
%%\frac{\tilde\Delta^T\bar\Delta \bar\Delta^T\phi}{\bar\Delta^T\bar\Delta}
%%&\text{if $\|\bar\Delta\| = B_{\delta}$ and $\bar\Delta^T\phi > 0$},
%%\end{cases}
%%\end{align*}
%%For the second case, we have $\tilde\Delta^T\bar\Delta=(\Delta -\bar\Delta)^T\bar\Delta \leq 0$ due to the assumption that $\|\Delta\|\leq B_{\delta}$ and $\|\bar\Delta\|=B_\delta$. This implies that $\tilde\Delta \cdot (I_{3\times 3}-\mathrm{Proj}[\bar \Delta,\phi]) \phi\leq 0$ always for any $\tilde\Delta,\bar\Delta,\phi\in\Re^3$. Therefore, by applying this to \refeqn{Ua} repeatedly, we obtain
%%\begin{align}
%%\mathcal{U}_a \leq 0.\label{eqn:Uaa}
%%\end{align}
%%Therefore, the time-derivative of the Lyapunov function satisfies
%%\begin{align}
%%\dot{\mathcal{V}} = \mathcal{U}_0+\mathcal{U}_a \leq \sum_{i=1}^n - z_i^T W_i z_i.\label{eqn:dotV}
%%\end{align}
%
%If the constants $c_x,c_R,c_q$ that are independent of the control input are sufficiently small, the matrices $W_{x_i},W_{R_i},W_{q_i}$ are positive-definite. Also, if the error in the direction of the link is sufficiently small relative to the desired trajectory, we can choose the controller gains such that the matrix $W_i$ is positive-definite, which follows that the zero equilibrium of tracking errors is stable in the sense of Lyapunov, and all of the tracking error variables $z_i$ and the estimation error variables are uniformly bounded, i.e., $e_{x_0},\dot e_{x_0},e_{R_0},e_{\Omega_0},e_{q_i},e_{\omega_i},\tilde\Delta_{x_0},\tilde\Delta_{R_0},\tilde\Delta_{x_i}\in\mathcal{L}_\infty$. These also imply that $e_{x_0},\dot e_{x_0},e_{R_0},e_{\Omega_0},e_{q_i},e_{\omega_i}\in\mathcal{L}_2$ from \refeqn{U0}, and that $\dot e_{x_0},\ddot e_{x_0},\dot e_{R_0},\dot e_{\Omega_0},\dot e_{q_i},\dot e_{\omega_i}\in\mathcal{L}_\infty$. According to Barbalat's lemma~\cite{IoaSun95}, all of the tracking error variables $e_{x_0},\dot e_{x_0},e_{R_0},e_{\Omega_0},e_{q_i},e_{\omega_i}$ and their time-derivatives asymptotically converge to zero. 
%
\subsection{Proof of Proposition \ref{prop:FDM}}\label{sec:prfFDM}

We first show that the attitude of the $i$-th quadrotor becomes exactly equal to its desired value within a finite time, i.e., $R_i(t)=R_{i_c}(t)$ for any $t\geq T$ for some $T>0$. This is achieved by finite-time stability theory~\cite{BhaBerSJCO00}. This proof is composed of two parts: (i) $s_i(t)=0$ for any $t>T_s$ for some $T_s<\infty$; (ii) when the state is confined to the surface defined by $s_i=0$, we have $e_{R_i}(t)=e_{\Omega_i}(t)=0$ for any $t>T_R$ for some $T_R<\infty$. From now on, we drop the subscript $i$ for simplicity, as the subsequent development is identical for all quadrotors.

From~\cite{LeeLeoPICDC10}, the error dynamics for $e_\Omega$ is given by
\begin{align*}
%\dot{e}_{R} & = \frac{1}{2}(\tr{R^T R_{c}}I-R^T R_{c}) e_{\Omega}\triangleq E(R,R_c)e_\Omega,\label{eqn:eRi_dot}\\
J\dot e_{\Omega} & = -\Omega\times \Omega +M +\Delta_R+ J(\hat\Omega R^T R_c \Omega_c - R^T R_c\dot\Omega_c).%\label{eqn:JeWi_dot}
\end{align*}
Substituting \refeqn{Mi},
\begin{align}
J\dot e_{\Omega} & = -k_s s - l_s S(r,s)+\Delta_R-B_\delta\frac{s}{\|s\|}\nonumber\\
&\quad - (k_R J+ l_s r J\mathrm{diag}_j[|e_{R_{j}}|^{r-1}]) E(R,R_{c})e_\Omega  .\label{eqn:JeWi_dot}
\end{align}

Let a Lyapunov function be
\begin{align*}
\mathcal{W} & = \frac{1}{2} s\cdot Js.
\end{align*}
From \refeqn{si} and \refeqn{eRi_dot}, its time-derivative is given by
\begin{align*}
\dot{\mathcal{W}} & = s\cdot\{J\dot e_\Omega + (k_R J + l_s r J\mathrm{diag}_j[|e_{R_j}|^{r-1}]) E(R,R_c) e_\Omega\}.
\end{align*}
Substituting \refeqn{JeWi_dot} and \refeqn{Mi}, and using \refeqn{Bdelta}, it reduces to
\begin{align*}
\dot{\mathcal{W}} & = s\cdot\{-k_s s -l_s S(r,s)+\Delta_R - \frac{s}{\|s\|}B_\delta\}\\
& \leq -k_s \|s\|^2 - l_s \sum_{j=1}^n |s_j|^{r+1}+B_\delta \|s\| -B_\delta \|s\| \\
& \leq -k_s\|s\|^2 - l_s \|s\|^{r+1},
\end{align*}
where the last inequality is obtained from the fact that $\|x\|^\alpha\leq \sum_{i=1}^n|x_i|^\alpha$ for any $x=[x_1,\ldots,x_n]^T$ and $0<\alpha<2$~\cite[Lemma 2]{WuRadAA11}. Therefore, 
\begin{align*}
\dot{\mathcal{W}} & \leq -\epsilon_1 \mathcal{W}  -\epsilon_2\mathcal{W}^{(r+1)/2},
\end{align*}
where $\epsilon_1=\frac{2k_s}{\lambda_{M}[J]}$ and $\epsilon_2=l_s(\frac{2}{\lambda_{M}[J]})^{(r+1)/2}$. This implies that $s(t)=0$ for any $t \geq T_s$, where the settling time $T_s$ satisfies
\begin{align*}
T_s\leq \frac{2}{\epsilon_1(1-r)}\ln \frac{\epsilon_1 \mathcal{W}(0)^{(1-r)/2}+\epsilon_2}{\epsilon_2},
\end{align*}
according to~\cite[Remark 2]{YuYuA05}.

Next, consider the second part of the proof when $s=0$. Let a configuration error function for the attitude of a quadrotor be
\begin{align*}
\Psi_R =\frac{1}{2}\trs{I-R_c^T R}.
\end{align*}
Consider a domain give by $D_R=\{(R,\Omega)\in\SO\times\Re^3\,|\, \Psi_R < \psi_R < 2\}$. It has been shown that the following inequality is satisfied in the domain,
\begin{align}
\frac{1}{2}\|e_R\|^2 \leq \Psi_R \leq \frac{1}{2-\psi_R} \|e_R\|^2.\label{eqn:PsiB}
\end{align}
Therefore, it is positive-definite about $e_R=0$. The time-derivative of $\Psi_R$ is given by $\dot\Psi_R= e_R\cdot e_\Omega$. Therefore, when $s=0$, we have
\begin{align*}
\dot\Psi_R & = 
-k_{R} \|e_{R}\|^2 - l_{R} \sum_{j=1}^n |e_{R_j}|^{r+1}\\
& \leq 
-k_{R} \|e_{R}\|^2 - l_{R} \|e_R\|^{r+1},
\end{align*}
Substituting \refeqn{PsiB}, we obtain 
\begin{align*}
\dot\Psi_R & \leq -\epsilon_3\Psi_R - \epsilon_4\Psi_R^{(r+1)/2},
\end{align*}
where $\epsilon_3=\frac{k_R}{2-\psi_R}$ and $\epsilon_4=\frac{l_R }{(2-\psi_R)^{(r+1)/2}}$.
This implies that $e_R(t)=e_\Omega(t)=0$ for any $t \geq T_R$, where the settling time $T_R$ satisfies
\begin{align*}
T_R\leq \frac{2}{\epsilon_3(1-r)}\ln \frac{\epsilon_3 \Psi_R(0)^{(1-r)/2}+\epsilon_4}{\epsilon_4}.
\end{align*}

In summary, whenever $t\geq T^*\triangleq \max  \{T_s,T_R\}$, it is guaranteed that $R_i(t)=R_{i_c}(t)$ for the $i$-th quadrotor. Next, we consider the \textit{reduced system}, which corresponds to the dynamics of the payload and the rotational dynamics of the links when $R_i(t)\equiv R_{i_c}(t)$. From \refeqn{fi} and \refeqn{b3i}, the control force of quadrotors when $R_i=R_{i_c}$ is given by
\begin{align*}
-f_i \cdot R_i e_3 = (u_i\cdot R_{c_i} e_3) R_{c_i} e_3 = 
(u_i\cdot -\frac{u_i}{\|u_i\|}) -\frac{u_i}{\|u_i\|} = u_i.
\end{align*}
Therefore, the reduced system is given by the controlled dynamics of the simplified model.

If the controller gains $k_R,l_R,k_s,l_s$ are selected large such that $T^*$ is sufficiently small, the solution stays inside of the domain $D$, where the stability results of Proposition 1 hold, during $0\leq t<T^*$. After $t\geq T^*$, the controlled system corresponds to the controlled system of the simplified dynamic model, and from Proposition \ref{prop:SDM}, the tracking errors asymptotically coverage to zero, and the estimation error are uniformly bounded.

\bibliography{/Users/tylee/Documents/BibMaster}
\bibliographystyle{IEEEtran}

\end{document}

\end{document}

### 2018-06-09 Conference/ICRA 2015/Workshops/ICRA15Workshops.tex
added Conference folder

--- 2018-06-09 Conference/AAS 2015 DA/ACC 2015/Generalized CJPDAF Derivation.tex
@@
 \usepackage{subfigure}
-
+\usepackage[]{algorithm2e}
 \usepackage{amsmath}
@@
 
-\section{Optimal Joint Probabilistic Data Assocation Filter (O-JPDAF)}
+\section*{ICRA15 Workshops}
 
-%\section*{Coalescence Avoiding Optimal Joint Probabilistic Data Association Filter (C-JPDAF)}
-%
-%The difference between the JPDAF and the C-JPDAF is that the JPDAF is optimal in the sence that estimate uncertainty is minimized.
-However, the C-JPDAF minimizes a weighted sum of uncertainty and similarity to decrease track coalescence, from which the JPDAF suffers when estimates are in close proximity over long periods of time.
-
-% Paper copy & paste starts below:
-
-\subsection{Flow Update}
-Similar to Kalman filters, the JPDAF is composed of a flow update and a measurement update.
-During the flow update (or prediction step), the estimated state $\hat x_{i_k}$ and its covariance $P_{i_k}$ for each track $i=\{1,2,...,t\}$ of $t$ tracks at time step $k$ is updated as follows:
-\begin{gather}
-\hat x_{i_{k}} = f_{i_{k-1}}( \hat x_{i_{k-1}}),\\
-P_{i_{k}} = F_{i_{k-1}} P_{i_{k-1}} F_{i_{k-1}}^T + Q_{i_{k-1}}, \quad F_{i_{k-1}}=\frac{\partial f}{\partial x}\bigg|_{x=\hat x_{i_{k}} }
-\end{gather}
-where the uncertainty of the process is represented with covariance $Q_{i_{k-1}}$.
-These equations are applied between measurements to provide predictions of the states and their uncertainties. % paper change
-
-\subsection{Measurement Update}
-
-The measurement serves to update the state estimate.
-Because all equations in this section take place at the $k$ time step, this subscript is removed from the equations.
-Let  the a priori value of a variable and its a posteriori value be denoted by the superscript $-$ and $+$, respectively.
-Let subscript $j$ indicate a measurement index such that the set $j=\{1,2,...,r\}$ for the number of returns $r$ (or measurements).
-For a given measurement $z_j$, its measurement residual or innovation for the $i$-th object is given by
-\begin{align*}
-e_{ij} = z_j - \hat z_i
-\end{align*}
-where $\hat z_i$ is the predicted measurement of the linearized equation
-\begin{align}
-\hat z_i = H_i \hat x_{i}^-
-\end{align}
-such that the measurement is a function of state
-\begin{align}
-z=h(x), \quad H_i=\frac{\partial h}{\partial x}\bigg|_{x=\hat x_{i_{k}}^- }.
-\end{align}•
-Given the uncertainty of the measurements represented with covariance $R_i$, the innovation covariance $S_{i}$ is
-\begin{align}
-S_{i}=H_{i}P_{i}^{-}H_{i}^T+R_{i}.
-\end{align}
-
-% [Eq. 1.4.16-(1-10)]
-
-\subsubsection{Likelihood of Association}
-The computation required with a standard JPDA is inconsistent with the capabilities of the computational resources to account for an arbitrary number of objects to be tracked~\cite[Eq.
-1.(1-2)]{Bar1990}.
-A modified version of the JPDA, commonly known as the "Cheap JPDA", provides a simpler approximation of the likelihood $\beta_{ij}$ that a return $j$ (or measurement) belongs to a particular track $i$ (or object).
-We define $G_{ij}$, which takes a very similar form as the Gaussian probability density function
-\begin{align}
-G_{ij}=\frac1{\sqrt{\det{S_{ij}}}}\exp{\{-\frac12e_{ij}^TS_{ij}^{-1}e_{ij}\}}
-\end{align}
-which is used to calculate the probability of association $\beta_{ij}$ as well as intermediate variables $\mathcal{S}_{ti}$ and $\mathcal{S}_{ri}$
-\begin{align}
-\beta_{ij}=\frac{G_{ij}}{\mathcal{S}_{ti}+\mathcal{S}_{ri}-G_{ij}+B}, 
-\quad 
-\mathcal{S}_{ti}=\sum\limits_{j}G_{ij}, 
-\quad 
-\mathcal{S}_{ri}=\sum\limits_{i}G_{ij}
-\end{align}
-where $B$ is a correction factor to account for the nonunity probability of detection as well as clutter and false alarms. Unless there are reasons that either phenomena may occur, we assume $B=0$.
-
-\subsubsection{State and Covariance Update Equations}
-The weighted sum of all innovations associated with the object $i$ is defined as ${\bf{e}}_{i}={\sum\limits_{j} \beta_{ij}e_{ij}}$.
-The resulting a posteriori estimate of the state is given by~\cite[Eq. 6-(25-26)]{TrackDataAssoc}
-\begin{align}
-\hat x^+_{i}=&\hat x^-_{i}+K_{i}{\bf{e}}_{i}\label{eqn:KalEst}
-\end{align}
-where $K_i$ is a gain chosen to minimize the a posteriori uncertainty of the state estimate; $K_i$ is derived below optimally in the sence that the a posteriori uncertainties are minimized with respect to the generalized linearized equations of a JPDAF.
-Of course, the estimator cannot be optimal if the system is nonlinear, because the linearized equations to not fully represent the dynamics of a nonlinear system.
-
-We define the likelihood that object $i$ is not detected is $\beta_{0,i}=1-{\sum\limits_{j} \beta_{ij}}$.
-The resulting a posteriori covariance can be written as [Tracking and Data Association, pg.
-326, equations D-32 and D-33]
-\begin{align}
-\label{eqn:JPDAFCov}
-P^+_{i}=& \beta_{0,i}P^-_{i}+(1-\beta_{0,i})P_i^c+\tilde P
-\\
-\tilde P=&K_i\left({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_{i}^T\label{eqn:tildeP}
-\end{align}
-where $P_i^c$ is defined as the a posteriori covariance of object $i$ if only a single measurement serves to update this track.
-%From [Tracking and Data Association, pg. 326, equation D-27], which is referenced in numerous publications, $P_i^c$ is calculated
-%\begin{align}
-%\label{eqn:PcKalman1}
-%P_i^c=(I-K_{i,Kalman}H_i)P_i^-
-%\end{align}
-%or equivalently
-%\begin{align}
-%\label{eqn:PcKalman2}
-%P_i^c=P_i^--K_{i,Kalman}SK_{i,Kalman}^T
-%\end{align}
-%where Kalman gain $K_{i,Kalman}=P_iH_i^TS_i^{-1}$ is used.
-In fact, (\ref{eqn:PcKalman1}) and (\ref{eqn:PcKalman2}) are derived from equations that require $K_{i,Kalman}$ be used.
-Below we show that $K_{i,Kalman}$ cannot guarantee that $\tr{P^+_{i}}$ is minimized:
-%\begin{align}
-%P^+_{i}=& \beta_{0,i}P^-_{i}+(1-\beta_{0,i})\left(P_i^--K_{i,Kalman}SK_{i,Kalman}^T\right)+K_{i,Kalman}\left({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_{i,Kalman}^T\\
-%=&P^-_{i}-K_{i,Kalman}\left((1-\beta_{0,i})S-({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T)\right)K_{i,Kalman}^T\\
-%=&P^-_{i}-K_{i,Kalman}S^*K_{i,Kalman}^T, \quad S^*=(1-\beta_{0,i})S-({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T).
-%\end{align}
-%We define the cost function $J^*=\tr{K_{i,Kalman}}$ to test for optimality with respect to $K_i$ assuming that $K_i=K_{i,Kalman}$ to show if this is true:
-%\begin{align}
-%0&=\frac{\partial\tr{P^+_i}}{\partial K_i}
-%\\
-%0&=-2(1-\beta_{0,i})\frac{\partial\tr{K_iH_iP^-_i}}{\partial K_i}+\frac{\partial\tr{K_iS^*K_i^T}}{\partial K_i}
-%\\
-%0&=-2(1-\beta_{0,i})P^-_iH_i^T+2K_iE_i
-%\\
-%K_iE_i&=(1-\beta_{0,i})P^-_iH_i^T
-%\\
-%K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1}\label{eqn:GainK}
-%\end{align}
-We can define $P_i^c$ without any assumptions on $K_i$ (i.e.
-$K_i$ is not necessarily the Kalman gain, so the well-known Kalman covariance update equations are not valid here) [Applied Optimal Estimation, pg.
-108, equation 4.2-12]
-\begin{align}
-\begin{split}
-P^c_i&=(I-K_iH_i)P^-_i(I-K_iH_i)^T+K_iR_iK_i^T
-\\
-&=P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i(H_iP^-_iH_i^T+R_i)K_i^T.
-\label{eqn:GenCov}
-\end{split}
-\end{align}
-Substituting \refeqn{tildeP} and \refeqn{GenCov} into \refeqn{JPDAFCov},
-\begin{align}
-\begin{split}
-P^+_{i}=&\beta_{0,i}P^-_{i}+(1-\beta_{0,i})\left(P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i(H_iP^-_iH_i^T+R_i)K_i^T\right)\\&+K_i\left({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_{i}^T
-\end{split}
-\\
-\begin{split}
-P^+_{i}=&P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\\&+K_i\left((1-\beta_{0,i})(H_iP^-_iH_i^T+R_i)+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\right)K_i^T
-\end{split}
-\\
-\begin{split}
-P^+_{i}=&P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_iE_iK_i^T,\\
-&E_i=(1-\beta_{0,i})(H_iP^-_iH_i^T+R_i)+{\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}}{\bf{e}}_{i}^T\label{eqn:CovExpanded}
-\end{split}
-\end{align}
-where $E_i=E_i^T$ and may be calculated without knowledge of the gain matrix $K_i$, and is invertible because $P^-_i$ and $R_i$ are positive-definite and $(\beta_{ij}e_{ij}e_{ij}^T-{\bf{e}}_{i}{\bf{e}}_{i}^T)$ are positive-semidefinite, shown below.
-
-Consider that object $i$ is not measured (likelihood $\beta_{0,i}$), then this portion of the innovation update $e_0=0_{m\times1}$ because no innovation exists where measurments are not available.
-Then for all innovations in the set $j^*=\{0,1, 2,...,r\}$, the Total Probability Theorem holds true for object $i$:
-\begin{align}
-\sum\limits_{j^*}\beta_{ij^*}=1.
-\end{align}
-From~\cite[Eq. 1.4.16-(1-10)]{ShaRonThi2001}, the covariance spread in the means term of the mixture of $e$ is defined as (note that ${\bf e}_{i}$ remains unchanged)
-\begin{align}
-\tilde P_e=\sum\limits_{j^*}\beta_{ij^*}(e_{ij^*}-{\bf e}_{i})(e_{ij^*}-{\bf e}_{i})^T
-\end{align}
-which is positive semi-definite. Manipulating this equation,
-\begin{align}
-\tilde P_e&=\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}e_{ij^*}^T
--\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}{\bf e}_{i}^T
--{\bf e}_{i}\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}^T
-+{\bf e}_{i}{\bf e}_{i}^T
-\\
-\tilde P_e&=\sum\limits_{j^*}\beta_{ij^*}e_{ij^*}e_{ij^*}^T-{\bf e}_{i}{\bf e}_{i}^T
-\\
-\tilde P_e&=\sum\limits_{j}\beta_{ij}e_{ij}e_{ij}^T-{\bf e}_{i}{\bf e}_{i}^T\geq0
-\label{eqn:CovSpreadPSD}
-\end{align}
-because $e_{ij^*}\equiv0$. Therefore $E_i$ is invertible $\forall\ \beta_{0,i}<1$.
-
-The only other case where $E_i$ is not invertible is when $\beta_{0,i}=1$, or when there is full confidence that a particular object is not observed by any measurement.
-This case is evaluated after the gain matrix is derived.
-
-The goal of the choosing $K_i$ is to minimize the uncertainty of the object state.
-With a Kalman filter, this goal is accomplished by minimizing $\tr{P^+_i}$ with respect to $K_i$ such that the total number of tracks $t=1$; the same approach may be applied to the JPDAF, except the a posteriori covariance equation has changed, so $K_i$ changes as well ($t=1$ is not guaranteed):
-\begin{align}
-0&=\frac{\partial\tr{P^+_i}}{\partial K_i}
-\\
-0&=-2(1-\beta_{0,i})\frac{\partial\tr{K_iH_iP^-_i}}{\partial K_i}+\frac{\partial\tr{K_iE_iK_i^T}}{\partial K_i}
-\\
-0&=-2(1-\beta_{0,i})P^-_iH_i^T+2K_iE_i\label{eqn:Uncertainty0}
-\\
-K_iE_i&=(1-\beta_{0,i})P^-_iH_i^T
-\\
-K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1}\label{eqn:GainK}
-\end{align}
-such that \refeqn{GainK} may be substituted into \refeqn{CovExpanded}
-\begin{align}
-\begin{split}
-P^+_{i}=&P^-_{i}-(1-\beta_{0,i})^2(P^-_iH_i^TE_i^{-1}H_iP^-_i+P^-_iH_i^TE_i^{-1}H_iP^-_i)
-\\
-&+(1-\beta_{0,i})^2P^-_iH_i^TE_i^{-1}E_iE_i^{-1}H_iP^-_i
-\end{split}
-\\
-P^+_{i}=&P^-_{i}-(1-\beta_{0,i})^2P^-_iH_i^TE_i^{-1}H_iP^-_i
-\end{align}
-which may be simplified in two ways
-\begin{align}
-P^+_{i}=&P^-_{i}-K_iE_iK_i^T,\label{eqn:JPDAFPostCovSymmetricUpdate}
-\\
-P^+_{i}=&\left(I-(1-\beta_{0,i})K_iH_i\right)P^-_i.\label{eqn:JPDAFPostCovSimpleUpdate}
-\end{align}
-
-Consider the singularity when $E_i=0$ if no association is made to a track, or where $(1-\beta_{0,i})=\sum\limits_{j}\beta_j=0$.
-While it is clear from \refeqn{KalEst} and \refeqn{CovExpanded} that the updates are independent of $K_i$ at this singularity, we show that $K_i$ is not growing without bound nearby the singularity.
-Hence, we evaluate the limit at this singularity to ensure it is not infinite.
-Note that if $\sum\limits_{j}\beta_j\rightarrow0\Rightarrow \beta_j\rightarrow0\ \forall j$ because of the bounds $\sum\limits_{j}\beta_j\geq\beta_j\geq0\ \forall j$.
-For clarity, let $\beta^*=\sum\limits_{j}\beta_j$ and let $\tilde\beta_p=\sum\limits_{j,j\neq p}\beta_j/\sum\limits_{j}\beta_j$ and we drop the subscript $i$ because this analysis focuses on a single track.
-First we evaluate the summations inside $E$, then consider the effect on $K$:
-\begin{align}
-\begin{split}
-&\sum\limits_{j} \beta_{j}e_{j}e_{j}^T=(\beta_1e_1e_1^T+\beta_2e_2e_2^T+...)
-\\
-&=(\beta_1+\beta_2+...)
-(e_1e_1^T-\tilde\beta_1e_1e_1^T+
-e_2e_2^T-\tilde\beta_2e_2e_2^T+...)
-\\
-&=\beta^*(\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T)
-\end{split}
-\\
-\begin{split}
-&{{\bf{e}}}{\bf{e}}^T=\sum\limits_{j} \beta_{j}e_{j}\sum\limits_{j} \beta_{j}e_{j}^T=
-(\beta_1e_1\sum\limits_{j} \beta_{j}e_{j}^T+\beta_2e_2\sum\limits_{j} \beta_{j}e_{j}^T+...)
-\\
-&=(\beta_1+\beta_2+...)(e_1\sum\limits_{j} \beta_{j}e_{j}^T-\tilde\beta_1e_1\sum\limits_{j} \beta_{j}e_{j}^T+
-e_2\sum\limits_{j} \beta_{j}e_{j}^T-\tilde\beta_2e_2\sum\limits_{j} \beta_{j}e_{j}^T+...)
-\\
-&=\beta^*(\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T)
-\end{split}
-\\
-\begin{split}
-&\lim_{\beta^* \to 0}K=\lim_{\beta^* \to 0}\beta^*P^-H^T
-\left(\beta^*[S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T-\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T+\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T]\right)^{-1}
-\\
-&=P^-H^T\lim_{\beta^* \to 0}
-\left(S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T-\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T+\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T\right)^{-1}
-\label{eqn:OptGainNoSingularity}
-\end{split}
-\end{align}
-which not an infinite limit because of the following bounded conditions:
-\begin{align}
-\begin{split}
-&\lim_{\beta^* \to 0}\tilde\beta_p=\lim_{\beta^* \to 0}\frac{\beta_1+\beta_2+...+\beta_{p-1}+0+\beta_{p+1}+...}{\beta_1+\beta_2+...+\beta_{p-1}+\beta_p+\beta_{p+1}+...}\ \forall p\in j
-\\
-&=1-\lim_{\beta^* \to 0}\frac{\beta_p}{\beta^*},\quad 0\leq\beta_p\leq\beta^*\leq1\Rightarrow0\leq\lim_{\beta^* \to 0}\frac{\beta_p}{\beta^*}\leq1\Rightarrow0\leq\lim_{\beta^* \to 0}\tilde\beta_p\leq1.
-\end{split}
-\end{align}
-and therefore \refeqn{OptGainNoSingularity} may be solved
-\begin{align}
-\lim_{\beta^* \to 0}K&=P^-H^T
-\left(S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}{B_j}e_{j}e_{j}^T\right)^{-1}
-\\
-&=P^-H^T
-\left(S+\sum\limits_{j}(1-B_j)e_{j}e_{j}^T\right)^{-1}
-\label{eqn:OptGainNoSingularitySolved}
-\end{align}
-for $0\leq B_j\leq1\ \forall j$ and it can be easily seen that the sum of the terms inside the parentheses are invertible.
-% and we know that $S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}{B_j}e_{j}e_{j}^T$ is invertible because $S$ is positive definite and the remaining terms are a simplified form of \refeqn{CovSpreadPSD}.
-From a computational standpoint, for every $\beta^*_i\approx0$, the effect of the measurement update should be negligible, so we may choose $K_i=0_{n\times m}$ near this singularity as it will have little effect on the filter performance.
-Choosing $K_i$ this way results in no difference between the a priori and a posteriori update results (no measurement update where no measurement is observed).
-%\begin{align}
-%\begin{split}
-%\lim_{\beta^* \to 0}K&=\lim_{\beta^* \to 0}P^-H^T
-%\left(S+\sum\limits_{j}e_{j}e_{j}^T-\sum\limits_{j}\tilde\beta_je_{j}e_{j}^T-\sum\limits_{j}e_{j}\sum\limits_{j}\beta_je_{j}^T+\sum\limits_{j}\tilde\beta_je_{j}\sum\limits_{j}\beta_je_{j}^T\right)^{-1}
-%\\
-%&asfd
-%\end{split}
-%\end{align}
-
-We show that the Kalman gain $K_{Kalman}$ and the optimal JPDAF gain $K_{O-JPDAF}$ are not always the same with a proof by contradiction.
-Consider that these gains are equivalent (object subscript $i$ is omitted for clarity):
-\begin{align}
-&K_{Kalman}=K_{O-JPDAF}
-\\
-&P^-H^TS^{-1}=(1-\beta_{0})P^-H^TE^{-1}
-\\
-&P^-H^TS^{-1}E=(1-\beta_{0})P^-H^T
-\\
-&P^-H^TS^{-1}\left((1-\beta_{0})S+{\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T\right)=(1-\beta_{0})P^-H^T
-\\
-&(1-\beta_{0})P^-H^T+P^-H^TS^{-1}\left({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T\right)=(1-\beta_{0})P^-H^T
-\\
-&K_{Kalman}\left({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T\right)=0.\label{eqn:ContradictionResult}
-\end{align}
-
-The final statement implies that the spread in the innovations $\tilde P\equiv0$, which is not assumed in the JPDAF.
-The result of \refeqn{ContradictionResult} may only occur if $K_{Kalman}=0$, the probabilities that each measurement $j$ belongs to each object $i$ is exactly $0$ or $1$, or $K_{Kalman}\perp\left({\sum\limits_{j} \beta_{j}e_{j}e_{j}^T-{\bf{e}}}{\bf{e}}^T\right)$ .
-None of these conditions are guaranteed, so it follows that $K_{O-JPDAF}=K_{Kalman}$ is not guaranteed.
-
-While many of the equations of the O-JPDAF and JPDAF are identical, the changes from the JPDAF are shown in \refeqn{GainK} and \refeqn{JPDAFPostCovSymmetricUpdate} or \refeqn{JPDAFPostCovSimpleUpdate}.
-Because these changes yield an update to the states that minimize a posteriori uncertainty specifically suited for the JPDAF, we refer to this technique as the optimal JPDAF, or O-JPDAF for short.
-
-\subsection{Numerical Example}
-We consider an example where four satellites in LEO are in the field of view of a ground sensor.
-The sensor measures range (km), range rate (km/s), and two angles (rad) according to local right-ascension and declination convensions with a $10\%$ chance that an object is not measured or an additional false alarm comes into the field of view.
-The conventional JPDAF and the O-JPDAF algorithms are compared with identical initial condions and measurements.
-
-\begin{figure}
-\centerline{
-	\subfigure[\;JPDAF Radii]
-		{\includegraphics[width=0.45\columnwidth]{JPDAF_Radii_Tracks1234}}
-	\subfigure[\;JPDAF Uncertainty]
-		{\includegraphics[width=0.45\columnwidth]{JPDAF_UncertaintyCosts_Tracks1234}}
-	}
-\centerline{
-	\subfigure[\;O-JPDAF Radii]
-		{\includegraphics[width=0.45\columnwidth]{O-JPDAF_Radii_Tracks1234}}
-	\subfigure[\;O-JPDAF Uncertainty]
-		{\includegraphics[width=0.45\columnwidth]{O-JPDAF_UncertaintyCosts_Tracks1234}}
-	}
-\centerline{
-	\subfigure[\;Comparison of Radii of the First Two Tracks]
-		{\includegraphics[width=0.45\columnwidth]{ComparisonRadii_Tracks12}}
-	\subfigure[\;Comparison of Uncertainties of the First Two Tracks]
-		{\includegraphics[width=0.45\columnwidth]{ComparisonUncertaintyCosts_Tracks12}}
-	}
-\caption{The orbital radii and costs are compared among the algorithms to show that their is a small improvement in the mean of the estimates with the O-JPDAF and large improvement in the uncertainty in the estimates.
-Addtitionally, both algorithms handled an uncertain number of measurements and false alarms in these simulations.
-}\label{fig:hoverRegions}
-\end{figure}
-
-%\\
-%P^+_{i}=&P^-_{i}+K_i\{(1-\beta_{0,i})S_i+({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e_{ij}}}}{\bf{e_{ij}}}^T)\}K_{i}^T
-%\\
-%P^+_{i}=&P^-_{i}+K_iS^*K_i^T, \quad S^*=(1-\beta_{0,i})S_i+({\sum\limits_{j} \beta_{ij}e_{ij}e_{ij}^T-{\bf{e_{ij}}}}{\bf{e_{ij}}}^T)
-
-%
-%
-\section{Coalescence Avoiding Optimal Joint Probabilistic Data Association Filter (C-JPDAF)}% Or another name that you choose
-\label{C-JPDAF}
-
-The O-JPDAF described in the previous section is effective in handling clutter or missed detection of multiple objects.
-However, it is prone to coalescing estimates, especially with objects in close-proximity.
-Here, we present a coalescence-avoiding optimal JPDAF, hereafter C-JPDAF for short, to avoid coalescence of the estimates.
-The main idea is choosing the estimator gain such that a weighted sum of the size of uncertainty and a measure of similarity is minimized.
-
-A similarity measure is introduced first, and an optimization problem is formulated. Throughout this section, the subscript $k$ denoting the time step is omitted for simplicity.
-
-\subsection{Matusita's Similarity Measure}
-
-Coalescence of neighboring estimates only occurs when these estimates are very similar.
-We evaluate the Matusita's measure of similarity because it takes the form of a differentiable analytic index.
-For two Gaussian probability density functions, $f_1(x)$ and $f_2(x)$, Matusita's measure is defined as
-\begin{align*}
-\zeta = \int \sqrt{f_1(x)f_2(x)}dx,
-\end{align*}
-which is between 0 and 1, and it measures the similarity between two densities ~\cite{Coal_k}.
-
-From~\cite{Coal_k}, using the assumption that the a posteriori states $\hat x_\mathcal{A}^+$ and $\hat x_\mathcal{B}^+$ are Gaussian, the Matusita's measure between two objects can be written as
-\begin{align}
-\zeta=\zeta_1\zeta_2,
-\end{align}
-where
-\begin{align}
-\label{Q}
-\zeta_1=&\frac{2^\frac{n}{2}|P_\mathcal{A}^+|^{\frac{1}{4}}|P_\mathcal{B}^+|^{\frac{1}{4}}}{|P_\mathcal{A}^++P_\mathcal{B}^+|^{\frac{1}{2}}},
-\end{align}
-\begin{align}
-\label{R}
-\zeta_2=&\exp{\left\{-\frac{1}{4}(\hat x^+_\mathcal{A}-\hat x^+_\mathcal{B})^T(P_\mathcal{A}^++P_\mathcal{B}^+)^{-1}(\hat x^+_\mathcal{A}-\hat x^+_\mathcal{B})\right\}}.
-\end{align}
-Here $\zeta_1$ measures the differences in the covariances and $\zeta_2$ represents the spread in the means.
-
-%Coalescence is well known to occur when estimates share measurements over long periods of time; estimates with largely different state covariances do not tend to coalesce
-%; consider two satellites $1$ and $2$ such that $S_1>>S_2$. Inside the measurement space, if a measurement falls nearby $\hat z_1$, then estimate with low measurement uncertainty...
-	% Introduce the definition of M-measure
-	% Introduce the definition of Bhattacharyya distance
-
-\subsection{Optimal Coalescence Avoidance}
-
-Define a cost function of coalescence avoiding filtering as
-
-\begin{align}
-J=J_P+cJ_S
-\end{align}
-where $J_P$ is an uncertainty cost that measures the confidence level of estimation and $J_S$ is a similarity cost that measures degree of coalescence.
-The constant $c$ is a positive weighting factor between $J_P$ and $J_S$.
-
-In the proposed C-JPDAF, the a posteriori estimate of the state is given in the form of O-JPDA given by \refeqn{GainK}, but the estimator gain $K_i$ is selected to minimize the total cost $J$.
-In short, the estimator gain is chosen in the consideration of both uncertainty and coalescence.
-These are illustrated at Figure \ref{fig:CostTrends}.
-
-\begin{figure}
-\setlength{\unitlength}{0.1\columnwidth}
-\centerline{\footnotesize\selectfont
-\begin{picture}(8.5,7.3)(-0.5,-0.8)
-\put(-0.5,3){\rotatebox{90}{Cost}}
-\put(0,0){\includegraphics[width=0.8\columnwidth]{ECC14_fig1.pdf}}
-\put(6.2,0.2){\shortstack[c]{Uncertainty\\ cost: $J_P$}}
-\put(6.2,2.8){\shortstack[c]{Similarity\\ cost: $J_S$}}
-\put(5.7,5.2){\shortstack[c]{Weighted sum:\\ $J=J_P+cJ_S$}}
-\put(4.98,0.78){\circle*{0.1}}
-\put(2.48,1.57){\circle*{0.1}}
-\put(5.0,-0.3){$K_{O-JPDAF}$}
-\put(2.5,-0.3){$K_{C-JPDAF}$}
-\put(3.1,-0.8){Estimator gain}
-\end{picture}
-}
-\caption{Illustration of optimal coalescence avoidance: several costs are illustrated with respect to an element of an estimator gain $K$ (red: uncertainty cost $J_P$, blue: similarity cost $J_S$, black: weighted sum).
-The estimator gain $K_{O-JPDAF}$ of O-JPDAF is selected to minimize the uncertainty cost.
-In the proposed approach, the estimator gain $K_{C-JPDAF}$ is chosen such that the weighted sum of the uncertainty cost and the similarity cost is minimized.
-In short, at the expense of increased uncertainty at a single step, the proposed approach prevents coalescence.
-The increased uncertainty at each step is rewarded by reducing the overall estimation error drastically by maintaining correct data association throughout crossing tracks in close proximity.}
-\label{fig:CostTrends}
-\end{figure}
-
-More explicitly, the uncertainty cost $J_P$ is defined as the sum of the trace of each a posteriori covariance:
-\begin{align}
-J_P=\sum\limits_{i}^{t}\tr{P^+_i},\label{eqn:JP}
-\end{align}
-where $P^+_{i}$ is obtained by \refeqn{JPDAFCov}. The similarity cost is chosen as the second part of the Matusita's measure that represents the weighted-spread of the mean values:
-\begin{align}
-J_S&=\sum\limits_{\mathcal{A}}^{t}\sum\limits_{\mathcal{B}}^{t}\exp (-au_{\mathcal{A}\mathcal{B}})\bigg|_{\mathcal{A}\neq\mathcal{B}}\\
-u_{\mathcal{A}\mathcal{B}} & = \begin{cases} (\hat x_{\mathcal{A}}^+-\hat x^+_{\mathcal{B}})^T(P^+_{\mathcal{A}}+P^+_{\mathcal{B}})^{-1}(\hat x^+_{\mathcal{A}}-\hat x^+_{\mathcal{B}})\label{eqn:u}, & \mbox{if } \zeta_2^->\zeta_{2,min} \\ 0 & \mbox{otherwise}  \end{cases}
-\end{align}
-%%%% \norm{\hat x_{\mathcal{A}}^+-\hat x^+_{\mathcal{B}}}
-% OR
-%\begin{align}
-%J_S&=\sum\limits_{\mathcal{A}}^{t}\sum\limits_{\mathcal{B}}^{t}\exp (-au_{\mathcal{A}\mathcal{B}})\bigg|_{\mathcal{A}\neq\mathcal{B}}\\
-%u_{\mathcal{A}\mathcal{B}} & = (\hat x_{\mathcal{A}}^+-\hat x^+_{\mathcal{B}})^T(P^+_{\mathcal{A}}+P^+_{\mathcal{B}})^{-1}(\hat x^+_{\mathcal{A}}-\hat x^+_{\mathcal{B}})
-%\end{align}
-where $\zeta_2^-$ is calculated using $\hat x^-_{\mathcal A}$ and $\hat x^-_{\mathcal B}$ as the means and $P^-_\mathcal{A}$ and $P^-_\mathcal{B}$ as the covariances, $\zeta_{2,min}$ defines a region of separation, and positive constant $a$ serves to amplify or attenuate the function as well as alter its shape.
-
-Necessary conditions for optimality are given by
-\begin{align}
-\deriv{J}{K_i} = \deriv{J_P}{K_i} + c \deriv{J_S}{K_i} =0,\label{eqn:NCO}
-\end{align}
-where $\deriv{J_S}{K_i}$ may be nonzero. These are solved for the set of all gains $\mathcal{K}=\{K_1,K_2,...,K_t\}$.
-
-Next, we find the derivative of each cost. The derivative of $J_P$ is taken from \refeqn{Uncertainty0}
-\begin{align}
-\label{eqn:CostP}
-\frac{\partial J_{P}}{\partial K_{i}}&=2K_iE_i-2(1-\beta_{0,i})P^-_iH_i^T
-\end{align}
-Note that $J_{P_\mathcal{X}}$ does not depend on $K_\mathcal{Y}$ for $\mathcal{X}\neq \mathcal{Y}$.
-
-%
-%To appropriately define the similarity cost, we only consider the $R$ term from Equation \ref{R} because coalescence is well known to occur when estimates are within close proximity.
-For simplicity, we define ${P^+_{12}}^{-1}$ as
-%\begin{align}
-%{P^+_{12}}^{-1}=(P_1^++P_2^+)^{-1}
-%\end{align}
-%and define define $u$ as
-We expand the nonzero portion of \refeqn{u} so it may be differentiated with respect to $K_\mathcal{A}$,
-\begin{align}
-\begin{split}
-u=&(\hat x^+_\mathcal{A}-\hat x^+_\mathcal{B})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^+_\mathcal{A}-\hat x^+_\mathcal{B})
-\\
-=&\hat x^{+T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^+_\mathcal{A}
--2\hat x^{+T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^+_\mathcal{A}
-+\hat x^{+T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^+_\mathcal{B}
-\\
-=&(\hat x^-_\mathcal{A}+K_\mathcal{A}{\bf e}_\mathcal{A})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^-_\mathcal{A}+K_\mathcal{A}{\bf e}_\mathcal{A})%
--2(\hat x^-_\mathcal{B}+K_\mathcal{B}{\bf e}_\mathcal{B})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^-_\mathcal{A}+K_\mathcal{A}{\bf e}_\mathcal{A})
-\\
-&+(\hat x^-_\mathcal{B}+K_\mathcal{B}{\bf e}_\mathcal{B})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^-_\mathcal{B}+K_\mathcal{B}{\bf e}_\mathcal{B})
-\\
-=&\hat x^{-T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^-_\mathcal{A}
-+2\hat x^{-T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
-+{\bf e}_\mathcal{A}^TK_\mathcal{A}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
-\\
-&+\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^-_\mathcal{B}
-+2\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B}
-+{\bf e}_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B}
-\\
-&-2\{\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^-_\mathcal{A}
-+\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
-+\hat x^{-T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B}
-+{\bf e}_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}\}
-\\
-=&\hat x_\mathcal{A}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{A}^-
--2\hat x_\mathcal{A}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{B}^-
-\\
-&+2(\hat x_\mathcal{A}^{-}-\hat x_\mathcal{B}^{-})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
-+\hat x_\mathcal{B}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{B}^-
-\\
-&+2(\hat x_\mathcal{B}^{-}-\hat x_\mathcal{A}^{-})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B}
-+{\bf e}_\mathcal{A}^TK_\mathcal{A}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
-\\
-&-2{\bf e}_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
-+{\bf e}_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B},
-\label{eqn:uexpanded}
-\end{split}
-\end{align}
-where ${P^+_{\mathcal{A}\mathcal{B}}}^{-1}=(P_\mathcal{A}^++P_\mathcal{B}^+)^{-1}$.
-
-%which is the most reduced form to differentiate with respect to $K_1$ and $K_2$.
-
-\newcommand{\Iij}{\ensuremath{\mathbf{1}_{ij}}}
-
-Let $\Iij\in \Re^{n\times m}$ be defined such that its $i,j$-th element is one and other elements are zero.
-The derivative of $(P_{\mathcal{A}\mathcal{B}}^+)^{-1}$ with respect to the $i,j$-th element of $K_\mathcal{A}$ is defined as $\mathcal{P}_{\mathcal{A}\mathcal{B}_{ij}}\in\Re^{n\times n}$:
-%such that $ij1(i,j)=1$ and $0$ elsewhere, we define $P^*_{12}$ as
-\begin{align}
-\begin{split}
-\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}&=\frac{\partial {P^+_{\mathcal{A}\mathcal{B}}}^{-1}}{\partial K_{\mathcal{A},ij}}
-\\
-&=-{P_{\mathcal{A}\mathcal{B}}^+}^{-1}\frac{\partial}{\partial K_{\mathcal{A},ij}}
-\left(
-P^-_\mathcal{A}-(1-\beta_{0,\mathcal{A}})(K_\mathcal{A}H_\mathcal{A}P^-_\mathcal{A}+P^-_\mathcal{A}H_\mathcal{A}^TK_\mathcal{A}^T)+K_\mathcal{A}E_\mathcal{A}K_\mathcal{A}^T
-\right)
-{P_{\mathcal{A}\mathcal{B}}^+}^{-1}
-\\
-&={P_{\mathcal{A}\mathcal{B}}^+}^{-1}[(1-\beta_{0,\mathcal{A}})(\Iij H_\mathcal{A}P_\mathcal{A}^-+P_\mathcal{A}^-H_\mathcal{A}^T\Iij^T)-\Iij E_\mathcal{A}K_\mathcal{A}^T -K_\mathcal{A}E_\mathcal{A}\Iij^T]{P_{\mathcal{A}\mathcal{B}}^+}^{-1}.
-\end{split}
-\end{align}
-Using this, the derivative of $u_{\mathcal{A}\mathcal{B}}$ with respect to the $i,j$-th element of $K_\mathcal{A}$ can be written as
-\begin{align}
-\begin{split}
-\frac{\partial u_{\mathcal{A}\mathcal{B}}}{\partial K_{\mathcal{A},ij}}
-=&\frac{\partial}{\partial K_{\mathcal{A},ij}}\{\hat x_\mathcal{A}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{A}^-
--2\hat x_\mathcal{A}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{B}^-
-\\
-&+2(\hat x_\mathcal{A}^{-}-\hat x_\mathcal{B}^{-})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
-+\hat x_\mathcal{B}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{B}^-
-\\
-&+2(\hat x_\mathcal{B}^{-}-\hat x_\mathcal{A}^{-})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B}
-+{\bf e}_\mathcal{A}^TK_\mathcal{A}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
-\\
-&-2{\bf e}_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
-+{\bf e}_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}{\bf e}_\mathcal{B}\}
-\\
-=& \hat x_\mathcal{A}^{-T}\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}\hat x_\mathcal{A}^-
--2\hat x_\mathcal{A}^{-T}\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}\hat x_\mathcal{B}^-
-\\
-&+2(\hat x_\mathcal{A}^{-}-\hat x_\mathcal{B}^{-})^T
-[\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}K_\mathcal{A}+{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\Iij]
-{\bf e}_\mathcal{A}
-\\
-&+\hat x_\mathcal{B}^{-T}\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}\hat x_\mathcal{B}^-
-+2(\hat x_\mathcal{B}^{-}-\hat x_\mathcal{A}^{-})^T\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}K_\mathcal{B}{\bf e}_\mathcal{B}
-\\
-&+{\bf e}_\mathcal{A}^T[\Iij^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}
-+K_\mathcal{A}^T\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}K_\mathcal{A}
-\\
-&+K_\mathcal{A}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\Iij]{\bf e}_\mathcal{A}
--2{\bf e}_\mathcal{B}^TK_\mathcal{B}^T[\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}K_\mathcal{A}
-\\
-&+{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\Iij]{\bf e}_\mathcal{A}
-+{\bf e}_\mathcal{B}^TK_\mathcal{B}^T\mathcal{P}_{\mathcal{A}\mathcal{B}_{\mathcal{A},ij}}K_\mathcal{B}{\bf e}_\mathcal{B}.
-\end{split}
-\end{align}
-From this, the derivation of the similarity cost $J_S$ with respect to the $i,j$-th element of $K_\mathcal{A}$ is given by
-\begin{align}
-\frac{\partial J_{S}}{\partial K_{\mathcal{A},ij}}=-a\exp(-au_{\mathcal{A}\mathcal{B}})\frac{\partial u_{\mathcal{A}\mathcal{B}}}{\partial K_{\mathcal{A},ij}}.\label{eqn:JSK}
-\end{align}
-
-In summary, \refeqn{CostP} and \refeqn{JSK} are substituted into \refeqn{NCO}, and it is solved to obtain the set of optimal gains $\mathcal{K}$.
-Then, the algorithm of C-JPDAF is completed by updating the state estimate and covariance by using \refeqn{KalEst} and \refeqn{JPDAFCov}, respectively.
-
-As the optimality condition is a nonlinear function of estimator gains, there is no closed-form solution.
-Hence, we require a nonlinear equation solver.
-Note that ${P_\mathcal{A}^+}^{-1}$, ${P_\mathcal{B}^+}^{-1}$, and ${P_{\mathcal{A}\mathcal{B}}^+}^{-1}$ are symmetric inverses and must only be calculated once for each $(\mathcal{A},\mathcal{B})$ where the similarity measure is nontrivial.
-When $c=0$, we can show that the optimal gain reduces to that of the O-JPDAF, which is used as an initial guess of numerical solutions.
-
-%\subsection{Suboptimal Solution}
-%Because optimal solution of \ref{eqn:JSK} does not have a closed-form solution, a close initial guess to the actual solution can decrease the computational requirements of the problem.
-Therefore for objects $1$ and $2$, we can minimize the two equations:
-%\begin{align}
-%\frac{\partial J}{\partial K_1}
-%&=\frac{\partial J_{P}}{\partial K_1}+c\frac{\partial J_{S}}{\partial K_1}
-%\approx2\beta_1({K_1S_1-P_1^-H_1^T})-ac\exp(-au^-)\frac{\partial u^*}{\partial K_1}
-%\label{eqn:subopt1}\\
-%\frac{\partial J}{\partial K_2}
-%&=\frac{\partial J_{P}}{\partial K_2}+c\frac{\partial J_{S}}{\partial K_2}
-%\approx2\beta_2({K_2S_2-P_2^-H_2^T})-ac\exp(-au^-)\frac{\partial u^*}{\partial K_2}
-%\label{eqn:subopt2}
-%\end{align}
-%where
-%\begin{align}
-%u^-&= (\hat x^-_1-\hat x^-_2)^T(P^-_{1}+P^-_{2})^{-1}(\hat x^-_1-\hat x^-_2)\label{eqn:uminus}\\
-%u^*&= (\hat x^+_1-\hat x^+_2)^T(P^-_{1}+P^-_{2})^{-1}(\hat x^+_1-\hat x^+_2)\label{eqn:uminus}
-%\end{align}
-%so that we define a new variable $\Omega=ac\exp(-au^-)$, which may be treated as constant in \ref{eqn:subopt1} and \ref{eqn:subopt2}. We modify \ref{eqn:uexpanded} to solve for $u^*$,
-%\begin{align}
-%\label{eqn:ustarexpanded}
-%u^*=&\{\hat x_1^{-T}{P^-_{12}}^{-1}\hat x_1^-
-%-2\hat x_1^{-T}{P^-_{12}}^{-1}\hat x_2^-\nonumber\\
-%&+2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T{P^-_{12}}^{-1}K_1e_1
-%+\hat x_2^{-T}{P^-_{12}}^{-1}\hat x_2^-\nonumber\\
-%&+2\beta_2(\hat x_2^{-}-\hat x_1^{-})^T{P^-_{12}}^{-1}K_2e_2
-%+\beta_1^2e_1^TK_1^T{P^-_{12}}^{-1}K_1e_1\nonumber\\
-%&-2\beta_1\beta_2e_2^TK_2^T{P^-_{12}}^{-1}K_1e_1
-%+\beta_2^2e_2^TK_2^T{P^-_{12}}^{-1}K_2e_2\}
-%\end{align}
-%where ${P^-_{12}}^{-1}=(P_1^-+P_2^-)^{-1}$, which is fixed with respect to $K_1$ and $K_2$. We take the derivative of \ref{eqn:ustarexpanded} with respect to $K_1$ and $K_2$
-%\begin{align}
-%\begin{split}
-%\frac{\partial u^*}{\partial K_1}=&\frac{\partial}{\partial K_1}\left(2\beta_1(\hat x_1^{-}-\hat x_2^{-})^T{P^-_{12}}^{-1}K_1e_1
-%+\beta_1^2e_1^TK_1^T{P^-_{12}}^{-1}K_1e_1
-%-2\beta_1\beta_2e_2^TK_2^T{P^-_{12}}^{-1}K_1e_1\right)\\
-%=&2\beta_1{P^-_{12}}^{-1}(\hat x_1^{-}-\hat x_2^{-})e_1^T
-%+2\beta_1^2{P^-_{12}}^{-1}K_1e_1e_1^T
-%-2\beta_1\beta_2{P^-_{12}}^{-1}K_2e_2e_1^T.
-%\label{eqn:ustarderivexpanded}
-%\end{split}
-%\end{align}
-%Let $\lambda$ be an eigenvalue of ${P^-_{12}}^{-1}$ and $q$ be the associated eigenvector, i.e.
-${P^-_{12}}^{-1}q=\lambda q\Rightarrow q^T{P^-_{12}}^{-1}=\lambda q^T$ because ${P^-_{12}}^{-1}$ is symmetric.
-%Then the derviative of \ref{eqn:ustarexpanded} with resepect to $K_1$ and $K_2$ and setting this to zero yields
-%\begin{align}
-%0&=2\beta_1({K_1S_1-P_1^-H_1^T})-\Omega\frac{\partial u^*}{\partial K_1}\\
-%0&=2\beta_1({K_1S_1-P_1^-H_1^T})-2\beta_1\Omega\left(
-%{P^-_{12}}^{-1}(\hat x_1^{-}-\hat x_2^{-})e_1^T
-%+\beta_1{P^-_{12}}^{-1}K_1e_1e_1^T
-%-\beta_2{P^-_{12}}^{-1}K_2e_2e_1^T\right)\\
-%0&=K_1S_1-P_1^-H_1^T
-%-\Omega{P^-_{12}}^{-1}(\hat x_1^{-}-\hat x_2^{-})e_1^T
-%-\Omega\beta_1{P^-_{12}}^{-1}K_1e_1e_1^T
-%+\Omega\beta_2{P^-_{12}}^{-1}K_2e_2e_1^T\\
-%0&=q^TK_1S_1-q^TP_1^-H_1^T
-%-\Omega q^T{P^-_{12}}^{-1}(\hat x_1^{-}-\hat x_2^{-})e_1^T
-%-\Omega\beta_1q^T{P^-_{12}}^{-1}K_1e_1e_1^T
-%+\Omega\beta_2q^T{P^-_{12}}^{-1}K_2e_2e_1^T\\
-%0&=q^TK_1S_1-q^TP_1^-H_1^T
-%-\Omega q^T\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T
-%-\Omega\beta_1q^T\lambda K_1e_1e_1^T
-%+\Omega\beta_2q^T\lambda K_2e_2e_1^T\\
-%0&=q^T\left\{K_1S_1-P_1^-H_1^T
-%-\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T
-%-\Omega\beta_1\lambda K_1e_1e_1^T
-%+\Omega\beta_2\lambda K_2e_2e_1^T\right\},
-%\quad q\neq0\\
-%0&=K_1(S_1-\Omega\beta_1\lambda e_1e_1^T)
-%-P_1^-H_1^T
-%-\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T
-%+\Omega\beta_2\lambda K_2e_2e_1^T\\
-%K_1&=\left(P_1^-H_1^T
-%+\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T
-%-\Omega\beta_2\lambda K_2e_2e_1^T\right)
-%(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}
-%\label{eqn:K1coupled}
-%\end{align}
-%and similarly
-%\begin{align}
-%K_2&=\left(P_2^-H_2^T
-%+\Omega\lambda(\hat x_2^{-}-\hat x_1^{-})e_2^T
-%-\Omega\beta_1\lambda K_1e_1e_2^T\right)
-%(S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}
-%\label{eqn:K2coupled}.
-%\end{align}
-%The equations \ref{eqn:K1coupled} and \ref{eqn:K2coupled} may be substituted into each other to seclude $K_1$ and $K_2$
-%\begin{align}
-%\begin{split}
-%K_1&=(P_1^-H_1^T
-%+\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T
-%-\Omega\beta_2\lambda[P_2^-H_2^T
-%+\Omega\lambda(\hat x_2^{-}-\hat x_1^{-})e_2^T\\
-%&-\Omega\beta_1\lambda K_1e_1e_2^T]
-%(S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}e_2e_1^T)
-%(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}
-%\end{split}
-%\end{align}
-%which may be rearranged to
-%\begin{align}
-%\begin{split}
-%&K_1\left(I-\Omega^2\lambda^2\beta_1\beta_2e_1e_2^T(S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}e_2e_1^T
-%(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}\right)=(P_1^-H_1^T
-%+\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T\\
-%&-\Omega\beta_2\lambda[P_2^-H_2^T+\Omega\lambda(\hat x_2^{-}-\hat x_1^{-})e_2^T](S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}e_2e_1^T)
-%(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}.
-%\end{split}
-%\end{align}
-%Solving for $K_1$,
-%\begin{align}
-%\begin{split}
-%K_1=&(P_1^-H_1^T
-%+\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T
-%-\Omega\beta_2\lambda[P_2^-H_2^T+\Omega\lambda(\hat x_2^{-}-\hat x_1^{-})e_2^T]
-%(S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}
-%e_2e_1^T)
-%(S_1-\\
-%&\Omega\beta_1\lambda e_1e_1^T)^{-1}\left(I-\Omega^2\lambda^2\beta_1\beta_2e_1e_2^T(S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}e_2e_1^T
-%(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}\right)^{-1}
-%\end{split}
-%\end{align}
-%and similarly
-%\begin{align}
-%\begin{split}
-%K_2=&(P_2^-H_2^T
-%+\Omega\lambda(\hat x_2^{-}-\hat x_1^{-})e_2^T
-%-\Omega\beta_1\lambda[P_1^-H_1^T+\Omega\lambda(\hat x_1^{-}-\hat x_2^{-})e_1^T]
-%(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}
-%e_1e_2^T)
-%(S_2-\\
-%&\Omega\beta_2\lambda e_2e_2^T)^{-1}\left(I-\Omega^2\lambda^2\beta_2\beta_1e_2e_1^T(S_1-\Omega\beta_1\lambda e_1e_1^T)^{-1}e_1e_2^T
-%(S_2-\Omega\beta_2\lambda e_2e_2^T)^{-1}\right)^{-1}.
-%\end{split}
-%\end{align}
-
-\bibliography{BibMaster}
-\bibliographystyle{IEEEtran}
+\begin{itemize}
+\item Robotic Vision: Challenges and Opportunities
+\begin{itemize}
+\item Vision goals: how many objects can I recognize and how well?
+\item Robotics goals: what can we do with those objects?
+\item Deep reinforcement of Learning
+\begin{itemize}
+\item Deep learning (2012) - 8 layer neural network
+\item $0.216$ trillion sets of training data--could only be done with large computing power, linearized functions
+\item Reinforcement Learning: find a control policy to optimize the ability to accomplish a goal
+\item Training: guided policy search: either removing the nonlinearities or a complicated policy simplifies the process
+\end{itemize}
+\item The 100/100 Tracking Challenge
+\begin{itemize}
+\item Assume the environment and everything inside it is known
+\item Goal: track 100\% of the objects with 100\% accuracy
+\item Modeling non-rigid objects
+\item Fairly decent tracking down to the joints of an object (hand, body)
+\item Detecting articulated objects: data association challenge
+\item Challenges: robust data association and occlusions for very fast motion, activity and task recognition
+\end{itemize}
+\item Geometry vs. Appearance
+\begin{itemize}
+\item Geometry: properties and relations in wold space--map representations (occupancy grids), sparse point clouds, landmark-based SLAM, dense SLAM
+\item Appearance: the was something looks in sensor space (typical definitions)--feature detectors, bag of words, pixel trackers, artificial landmarks (QR codes)
+\item Lots of ways to relate the two, but rarely are they considered together
+\item Why? physics of the sensors that relates geometry and appearance is difficult
+\item Removing illumination does great things
+\end{itemize}
+\item Natural Vision for Robotic Vision: Pillage or Pass Over?
+\begin{itemize}
+\item Mouses have much blurrier vision than humans, although they can quickly learn how to navigate and do tasks
+\item There is a large disconnect between measuring vision as a whole and the actual tasks of animals
+\item Cameras are getting so cheap that mostly nature-based approaches are not very useful now
+\end{itemize}
+\item Visual Navigation
+\begin{itemize}
+\item Self-driving vehicles have a perception problem
+\item Occlusion is incredibly difficult
+\item Signals become very difficult (waving you through an intersection)
+\item Signs may be blocked, but must be understood
+\item Challenges: maintaining maps, dealing with adverse weather, dealing with people, robust computer vision
+\item We need object-based understanding
+\end{itemize}
+\end{itemize}
+\newpage
+\item Tutorial: Industrial Applications with Large-Scale 3D Poin Cloud Processing
+\begin{itemize}
+\item Basic principles, solving problems of modern sensor data processing
+\item 3DTK: 3D Toolkit, OpenSource/GPL
+\item How to save a point cloud?
+\begin{itemize}
+\item vector of (x,y,z) as a text file
+\item range/intensity image: 2D array of kinect-like sensors, but there is distortion of a spherical sensor space onto a 2D image plane: projecting each 3D point to an image plane, but you cannot map a complete sphere into a 2D image
+\item oc-tree: occupancy trees break down the space into blocks, where free space is removed and those being measured are split to increase resolution in this region
+\item k-d tree: binary, heavily used in nearest-neighbor-like scenarios
+\end{itemize}
+\item
+\end{itemize}
+\end{itemize}
 
 \end{document}
-
-%\begin{align}
-%\begin{split}
-%u=&(\hat x^+_\mathcal{A}-\hat x^+_\mathcal{B})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^+_\mathcal{A}-\hat x^+_\mathcal{B})
-%\end{split}
-%\\
-%\begin{split}
-%=&\hat x^{+T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^+_\mathcal{A}
-%-2\hat x^{+T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^+_\mathcal{A}
-%+\hat x^{+T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^+_\mathcal{B}
-%\end{split}
-%\\
-%\begin{split}
-%=&(\hat x^-_\mathcal{A}+K_\mathcal{A}{\bf e}_\mathcal{A})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^-_\mathcal{A}+K_\mathcal{A}{\bf e}_\mathcal{A})%
-%-2(\hat x^-_\mathcal{B}+\beta_\mathcal{B}K_\mathcal{B}e_\mathcal{B})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^-_\mathcal{A}+K_\mathcal{A}{\bf e}_\mathcal{A})
-%\\
-%&+(\hat x^-_\mathcal{B}+\beta_\mathcal{B}K_\mathcal{B}e_\mathcal{B})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}(\hat x^-_\mathcal{B}+\beta_\mathcal{B}K_\mathcal{B}e_\mathcal{B})
-%\end{split}
-%\\
-%\begin{split}
-%=&\hat x^{-T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^-_\mathcal{A}+2\hat x^{-T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\beta_\mathcal{A}K_\mathcal{A}e_\mathcal{A}
-%+e_\mathcal{A}^TK_\mathcal{A}^T\beta_\mathcal{A}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}{\bf e}_\mathcal{A}
-%\\
-%&+\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^-_\mathcal{B}+2\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\beta_\mathcal{B}K_\mathcal{B}e_\mathcal{B}+e_\mathcal{B}^TK_\mathcal{B}^T\beta_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\beta_\mathcal{B}K_\mathcal{B}e_\mathcal{B}
-%\\
-%&-2\{\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x^-_\mathcal{A}+\hat x^{-T}_\mathcal{B}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\beta_\mathcal{A}K_\mathcal{A}e_\mathcal{A}+\hat x^{-T}_\mathcal{A}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\beta_\mathcal{B}K_\mathcal{B}e_\mathcal{B}+e_\mathcal{B}^TK_\mathcal{B}^T\beta_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\beta_\mathcal{A}K_\mathcal{A}e_\mathcal{A}\}
-%\end{split}
-%\\
-%\begin{split}
-%=&\{\hat x_\mathcal{A}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{A}^-
-%-2\hat x_\mathcal{A}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{B}^-
-%\\
-%&+2\beta_\mathcal{B}(\hat x_\mathcal{A}^{-}-\hat x_\mathcal{B}^{-})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}e_\mathcal{A}
-%+\hat x_\mathcal{B}^{-T}{P^+_{\mathcal{A}\mathcal{B}}}^{-1}\hat x_\mathcal{B}^-
-%\\
-%&+2\beta_\mathcal{B}(\hat x_\mathcal{B}^{-}-\hat x_\mathcal{A}^{-})^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}e_\mathcal{B}
-%+\beta_\mathcal{A}^2e_\mathcal{A}^TK_\mathcal{A}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}e_\mathcal{A}
-%\\
-%&-2\beta_\mathcal{A}\beta_\mathcal{B}e_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{A}e_\mathcal{A}
-%+\beta_\mathcal{B}^2e_\mathcal{B}^TK_\mathcal{B}^T{P^+_{\mathcal{A}\mathcal{B}}}^{-1}K_\mathcal{B}e_\mathcal{B}\},
-%\label{eqn:uexpanded}
-%\end{split}
-%\end{align}

### 2018-06-09 Conference/ICRA 2018.1/ICUAS18.tex
added Conference folder

\documentclass[letterpaper, 10pt, conference]{ieeeconf}
\IEEEoverridecommandlockouts \overrideIEEEmargins
\usepackage{amsmath,amssymb,url,times,subfigure,graphicx,theorem}
\usepackage{graphicx,subfigure,hyperref}
\usepackage{color,comment}
\usepackage{csquotes}
%\usepackage{refcheck}

\usepackage{epic,eepic}

%\usepackage[]{algorithm2e}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}

%\usepackage{algorithmicx}
%\usepackage{algpseudocode}

%\usepackage{algorithm}
%\usepackage{algpseudocode}
%\usepackage{pifont}

%\usepackage{amsmath}
%\usepackage{algorithm}
%\usepackage[noend]{algpseudocode}
%
%\makeatletter
%\def\BState{\State\hskip-\ALG@thistlm}
%\makeatother

%\definecolor{red}{rgb}{0,0,0}
\graphicspath{{../../Fig/}}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\abs}[1]{\ensuremath{\left| #1 \right|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\ip}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mbox{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mbox{tr}\ensuremath{\!\bracket{#1}}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathbf{d}}}
\newcommand{\pair}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\!\langle #1 \rangle\!\rangle}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\ad}{\ensuremath{\mathrm{ad}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

\title{\LARGE \bf
Bidding-Based Autonomous Exploration of Cooperative Robots Building a 3D Probabilistic Occupancy Grid Map}

\author{Evan Kaufman and Taeyoung Lee%
\thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 {\tt \{evankaufman,tylee\}@gwu.edu}}
\thanks{This research has been supported by the U.S. Naval Research Laboratory Base Program Work Unit ``Intelligent Microflyer'' and in part by NSF under the grants CMMI-1243000, CMMI-1335008, and CNS-1337722.}
}

\newcommand{\EditEK}[1]{{\color{blue}\protect #1}}
%\renewcommand{\EditTL}[1]{{\protect #1}}

\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{cor}{Corollary}
\newtheorem{remark}{Remark}

\begin{document}
\allowdisplaybreaks

\maketitle \thispagestyle{empty} \pagestyle{empty}

\begin{abstract}
This paper presents a bidding-based cooperative robotic exploration strategy to generate accurate occupancy grid maps with multiple vehicles. Using recent developments in 3D occupancy grid mapping and autonomous exploration, we propose an auction-based framework to build maps from various vantage points. A central executive reviews individual robot bids to maximize map information gain, and these bids are updated to prevent robots from capturing the same grid cells or colliding, which is further optimized with a receding horizon framework. Furthermore, considering a multi-agent system capable of large environmental coverage, several computational enhancements are presented to improve the scalability of the mapping and exploration algorithms. This multi-vehicle exploration approach is demonstrated with a numerical simulation where three robots simultaneously explore a large building floor plan cooperatively.
\end{abstract}

\section{Introduction}

% Multivehicle MAPPING in 3D 
% what is mapping?
% Occupancy grid mapping this way (ACC, JIRS, IRC)
% Designed to work with multiple sensors
% 
% Exploration:
% Frontier, entropy-based
% Recent papers (SIMPAR, JIRS, IRC)
% Main idea: maximizing a reward
% Multivehicle: 
%

Robotic mapping is the process of generating maps representing the environment surrounding robots.
Mapping serves an important role in tasks such as search-and-rescue, surveillance, and robotic cleaning.
Several grid-based map representations have been studied, e.g.,~\cite{WolSuk05,MeyBeiBur12,TanThoWolBus14}, because they provide abundant occupancy knowledge about the environment and are computationally tractable.

Most grid-based mapping representations are designed for 2D, while others are extended to capture a full 3D environment. In particular, occupancy probabilities have been assigned to a uniform 3D grid in~\cite{Andert09,PirRutBisSch11}. Another highly-popular variation of occupancy grid mapping is the Octomap representation~\cite{WurHorBenStaBur10,HorWurBenStaBur13}, which has added features such as extending the mapping limits and changing cell sizes. Most recently,~\cite{KauLeeAiMos16,KauTakAiLee17} proposed a systematic method to build an accurate probabilistic map using depth sensor stochastic properties directly, which is extended to 3D in~\cite{KauTakAiLee18}.

Next consider how the motions of cooperating robots impact the knowledge of a 3D map, known as autonomous exploration. 
A common approach is known as frontier-based exploration, proposed for 2D environments in~\cite{Yam97,Yam98} and extended to 3D  using the Octomap representation in~\cite{ZhuDinLinWu15,SenWan16,KleDor13}, and similarly using a visibility metric in~\cite{SawKriSri09}. This approach moves robots toward the boarder between certain and uncertain spaces, though these systematic actions are not based on any consideration of the future probabilistic map uncertainty.

Alternatively, there have been exploration techniques that are based on a quantitative measure of uncertainty known as Shannon's entropy~\cite{StaGriBur05}, and it is used to acquire expected map information gain in~\cite{KauAiLee16,KauTakAiLee17} in 2D. Here, the robot selects actions to learn the largest amount about the map using expected entropy \emph{change}. This concept is extended to 3D environments in~\cite{KauTakAiLee18}, where important properties of the 3D occupancy grid are projected onto a 2D map to eliminate computational costs. However, this approach governs robotic motion of just a single vehicle.

In this paper, we propose to expand the uncertainty-based exploration technique for multiple robots mapping cooperatively. However, optimizing map coverage simultaneously from multiple vehicles becomes very complicated and computationally intractable. Instead, an auction-based approach addresses the computational restrictions effectively. Bidding-based approaches were introduced with~\cite{SimApfBurFoxMooThrYou00,DiaSte00}, and extended to several other robotic problems~\cite{GerMaj02,ZloSteDiaTha02,SarBal05,ChoBruHow09} exploring decentralization and task allocation with varying robot capabilities. In~\cite{SimApfBurFoxMooThrYou00}, a central executive assigns tasks to robots for maximizing overall utility of the group. This is accomplished by minimizing a coverage overlap among the members. Robots submit bids based on expected coverage of uncertain spaces and travel distance. The robot winning the auction is awarded the task, and the remaining bids are discounted to prevent robots from covering the same area, thereby coordinating the mapping efforts.

% probabilistic map uncertainty to generate bids

Here, we propose an alternative multi-vehicle coordinated exploration approach that uses a central executive, focussed on map information gain maximization. Instead of estimating coverage overlap among robots, the proposed approach predicts how robotic actions are expected to improve map information gain cooperatively in a probabilistic fashion. A series of auctions determines where robots should travel. During each auction, robots submit bids based on how their sensors are expected to improve the quality of the occupancy grid map. Between auctions, bids are modified to reflect the impact of other robot members on the map. Because this approach is directly based on cooperative improvement of map information, considering coverage overlap among the robots becomes unnecessary. This multi-vehicle exploration optimization is further augmented with travel time considerations and collision avoidance among robots. In short, the proposed multi-vehicle exploration scheme employs a bidding framework that maximizes map information while considering multiple vehicles cooperating together.

%Expected changes in Shannon's entropy of the probabilistic map in conjunction with robot travel distances provide sufficient information for the bids from several candidate poses. Then, a novel bid discounting method provides a means for multi-vehicle coordination. This discounting method is composed of two parts: first, the expected information gains within the neighborhood of winning bids are updated based on how the winning bid is expected to change the map, thereby avoiding coverage overlapping among robots. Second, bids are further discounted with a cost associated with Euclidean distance to avoid collisions between exploring robots. In short, the robot locations and expected entropy of the occupancy grid map are considered together to avoid coverage overlaps or collisions.

Furthermore, this bidding-based exploration approach is integrated with a receding horizon framework to optimize map information and collision-avoidance in a dynamic environment. The receding horizon framework serves to update the exploration strategy with a possibly-changing occupancy grid map, which directly impacts the viability of pose candidates. Additionally, a receding horizon provides an optimal feedback policy that can handle dynamic obstacles, such as collision-avoidance with cooperative robots.

Finally, we present a numerical simulation of three quadrotors mapping and exploring a floor plan of a building. Since the environment is particularly large, this motivates several algorithm modifications to promote operational scalability, where communication among the various processes is focussed on map changes and multithreading is applied where necessary, among other computational enhancements.

The paper is organized as follows. The mapping and exploration problems are formulated in Section \ref{sec:ProbDef}. The method for auctioning exploration bids with receding horizons is presented in Section \ref{sec:BiddingExploration}. A mapping and exploration simulation with a scalable solution is shown in Section \ref{sec:NumericalSim}, followed by conclusions.

\section{Problem Formulation}
\label{sec:ProbDef}

This sections summarizes the key equations providing the exact solutions to occupancy grid mapping and expected entropy change proposed in~\cite{KauLeeAiMos16,KauTakAiLee17,KauTakAiLee18,KauAiLee16}.

\subsection{Probabilistic Occupancy Grid Mapping}
\label{subsec:POGM}
Let map $m$ be an occupancy grid map decomposed into $n_m$ evenly-spaced cubic grid cells with fixed edge length $\alpha$. The $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,\ldots,n_m}$, that is defined as $\mathbf{m}_i=1$ when occupied, and $\mathbf{m}_i=0$ when free. The location and size of each grid cell is assumed known, where a smaller cell size better represents a space, but increases computation and memory. Therefore, a map $m$ is defined by $\{\mathbf{m}_1,\ldots, \mathbf{m}_{n_m}\}$ ($2^{n_m}$ possible maps).

Another random variable is defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience, which is the event that the $i$-th cell is free, i.e., $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$. The random variables composing $m$ are assumed mutually independent, 
\begin{align}
P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i).
\end{align}

% {R ? R3×3 | RT R = I, det R = 1}
Consider a range sensor that provides scans of the surrounding environment in order to identify the closest occupied space. Let the measurement scan at the $t$-th time step be $Z_t$, where $z\in Z_t$ is a single measurement ray. Let the $t$-th pose be denoted by $X_t=\braces{x_t,R_t}$ with location $x_t\in\Re^3$ and attitude $R_t\in\SO=\braces{R\in\Re^{3\times3}|R^\T R=I,\det{R}=1}$, where $X_t$ is assumed to be known, i.e., perfect localization is considered. Given the sensor probabilistic properties $p(z|X_t,m)$, commonly referred to as the \emph{forward sensor model}, the goal is finding the \emph{inverse sensor model}, $P(m|z,X_{1:t},Z_{1:t-1})$. To obtain this probability for all cells,~\cite{KauLeeAiMos16,KauTakAiLee17} introduced a reduced map $r\subset m$ composed of $n_r$ cells along $z$, where cells are indexed by increasing distance from $x_t$, and $\mathbf{r}_{k+}$ refers to the event that the $k$-th cell of map $r$ is the closest occupied space. Then, the inverse sensor model for the $k$-th cell is
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})&=\eta\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1}),
\end{align}
where $\eta$ is the normalizer,
\begin{align}
\label{eqn:allEta}
\eta
&=
\bigg[\sum_{i=1}^{n_{r}+1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\} p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_i^-\bigg]^{-1},
\end{align}
such that the probability summation of the entire space is $1$, and the unnormalized probability of the inverse sensor model is defined as
\begin{align}
\label{eqn:Unnormalized}
\tilde P(\mathbf{r}_{k}|z,X_{1:t},&Z_{1:t-1})\nonumber\\
&=\mathbf{P}_k^-
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]\nonumber\\
&\quad + \bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{k+},X_t)\mathbf{P}_k^-,
\end{align}
where a priori probability is $\mathbf{P}_k^-=P(\mathbf{r}_{k}|X_{1:t-1},Z_{1:t-1})$, its complement is $\bar{\mathbf{P}}_k^-=1-\mathbf{P}_k^-$, $P(\bar{\mathbf{r}}_{0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience, and $p(z|\mathbf{r}_{(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading.The proofs of \refeqn{RayISMAnswer}--\refeqn{allEta} are given in~\cite{KauLeeAiMos16}. Since the terms of \refeqn{Unnormalized} are easily obtained and several are used repeatedly to obtain \refeqn{allEta}, the computational cost of \refeqn{RayISMAnswer} is trivial and can be applied in real-time. This process is repeated for all measurements composing scan $Z_t$.

\subsection{Predictive Entropy-Based Autonomous Exploration}

Since occupancy grid mapping provides a probabilistic representation of surrounding space, this mapping scheme holds probabilistic information about the uncertainty of the map. Shannon's entropy is commonly used as a measure of uncertainty, such as in~\cite{StaGriBur05}. Given the probabilities of each grid cell of map $m$, Shannon's entropy of an individual cell and the entire map are defined as
\begin{align}
\label{eqn:ShannonsEntropyCell}
H(P(\mathbf{m}_i))&=-P(\mathbf{m}_i)\log{P(\mathbf{m}_i})-P(\bar{\mathbf{m}}_i)\log{P(\bar{\mathbf{m}}_i}),
\\
\label{eqn:ShannonsEntropyMap}
H(P(m))&=\sum_{i=1}^{n_m}H(P(\mathbf{m}_i)),
\end{align}
respectively.
The entropy of the $i$-th grid cell is maximized when $P(\mathbf{m}_i)=0.5$ (greatest uncertainty) and minimized as $P(\mathbf{m}_i)$ approaches $0$ or $1$ (smallest uncertainty). %; thus Shannon's entropy is a measure of uncertainty.

Let an unvisited candidate future pose and its measurement scan be $X_c=\braces{x_c,R_c}$ and $Z_c$, respectively, such that $c\in\mathcal C$ where $\mathcal C=\braces{1,2,...,n_c}$ is the set of all $n_c$ candidates. Since the robot has not visited $X_c$, any change to the probabilistic map must be predicted. Much like the probabilistic mapping, this is achieved ray-by-ray~\cite{KauAiLee16,KauTakAiLee17}. Since all grid cell probabilities are conditioned on the history of poses $X_{1:t}$ and measurement scans $Z_{1:t}$, these are removed from the remaining equations for simplicity. For candidate ray $z_c$, the expected entropy is
\begin{align}
\label{eqn:DiscExpEntropyRay}
&\text{E}[H(P(m|x_c,z_{c}))]=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(m|x_c,z_{c,k}))P(z_{c,k}|x_c)\bigg\},
\end{align}
where subscript $k$ refers to the event that the candidate \emph{measurement} captures the $k$-th cell along the ray, i.e., $z_{c,k}$ is the distance from $x_c$ to the $k$-th grid cell. The first term of the summation of \refeqn{DiscExpEntropyRay}, namely $H(P(m|x_c,z_{c,k}))$, is obtained with the inverse sensor model \refeqn{RayISMAnswer}--\refeqn{allEta} and entropy definitions \refeqn{ShannonsEntropyCell}, \refeqn{ShannonsEntropyMap}. The second term is found using \refeqn{allEta} with
\begin{align}
\label{eqn:ProbMeas}
P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}},
\end{align}
where $\eta_{c,k}$ refers to the normalizer calculated with $z_{c,k}$.
Then, the expected information gain for candidate pose $X_c$ is the negative entropy change,
\begin{align}
%\label{eqn:objectiveFunOptimized}
%X_c^*&=\argmax_{X_c}{\ \mathcal I(X_c)},
%\\
\label{eqn:expectedInfoGainRay}
\mathcal I(X_c)&=H(P(m|X_{1:t},Z_{1:t}))-\text{E}\left[H(P(m|X_c,Z_c))\right],
\end{align}
where \refeqn{DiscExpEntropyRay} is used repeatedly for several sample rays that compose scan $Z_c$ as shown in~\cite{KauAiLee16}. Assuming the sample rays are spaced sufficiently apart, each ray tends to intersect a different set of grid cells, so these rays composing $Z_c$ are considered independently.

\section{Bidding-Based Autonomous Exploration}
\label{sec:BiddingExploration}

Here, we formulate a cooperative exploration problem with multiple robots using sequential auctions. The first auction is based on the expected information gain presented in the prior section, scaled by travel distance. Once the winner of the first auction is determined, the map for bidding is updated based on the expected measurements of the first robot. The second auction for the remaining robots is determined by the information gain from the revised map that is further scaled by the travel distance and a penalty for collision-avoidance. This process is repeated until all robots are assigned tasks. Then, the robots traverse toward their optimal poses while the bidding process repeats using a receding horizon framework.

%The expected information gain of each robot can be obtained from the computation framework presented in the prior section. As such, we consider the cost of travel and multi-vehicle coordination in this section.

%We assume that all robots mapping the environment use the same set of depth sensors to update the grid cells. Then, \refeqn{expectedInfoGainRay} is valid for all robots, and the total candidate information gains outlined in~\cite{KauAiLee16} need only be calculated once during a set of auctions. The only aspect differentiating robot bids is travel distance. The completed bids, and discounting for multi-vehicle coordination, are described next.

\subsection{Objective Function for the First Auction}

Instead of considering the full 3D map during entropy calculations, which yields a more complicated and computationally expensive information gain expectation, we project the 3D map onto a simpler 2D map, namely $m_\text{2D}$, located at the robot altitude according to~\cite{KauTakAiLee18}. The main idea is that several 3D cells in each dimension serve to update a single 2D projected cell probability, accounting for collision and entropy. This approach conservatively and efficiently captures small occupied spaces around the robot height. However, this approach is not suitable for capturing floors and ceilings or exploring complex 3D environments such multi-level buildings or complex terrain. Since these features are not considered for exploration here, we apply the map projection technique to the multi-vehicle autonomous exploration problem.

%. Consider a candidate pose $X_c$. Even though $\mathcal I(X_c)$ is fixed because the set $\mathcal I=\braces{\mathcal I_1,\mathcal I_2,\ldots,\mathcal I_{n_c}}$ is fixed, the travel time for each robot can differ greatly. This motivates a unified approach to differentiate bids for the same candidate pose from different robots.

%The expected information gain $\mathcal I(X_c)$ from \refeqn{expectedInfoGainRay} is identical for all robots. Therefore the same set $\mathcal I=\braces{\mathcal I_1,\mathcal I_2,\ldots,\mathcal I_{n_c}}$ can be applied for all robots. However

%; in some cases, certain robots, or even all robots, cannot reach $X_c$

Next, we describe how travel time is integrated into exploration. Considering that distances from current robot poses to candidate poses may differ greatly, accounting for these varying travel times properly is essential for exploration time efficiency. Travel distances are computed using Dijkstra's search to provide collision-free waypoints for each robot. There are two steps: first, generate a cost map from the robot location to each location on the 2D projected map. This provides the collision-free distances to all reachable locations on the map, which is different than the Euclidean distance in general. Second, the waypoints from a candidate pose to the current robot pose are easily obtained along the cost map using steepest descent.

%    if x(k) < d
%        f(k) = 0.5*fMax*(1-cos(pi*x(k)/d));
%    else
%        f(k) = fDiff*exp(-1e-2*(x(k)-d)^2)+fFar;
%    end

%Dijkstra's search demonstrates that the distances from each robot to $X_c$ are different in general, so travel times to $X_c$ vary. This motivates a unified approach to account for travel time, which promotes local travel and discourages robots from traversing across the map when possible.

Here, we define a bump function to account for varying travel distances from Dijkstra's search. Let the distance along the collision-free path from the $k$-th robot pose, namely $X_k$, to the $c$-th candidate pose, namely $X_c$, be denoted $d(X_k,X_c,m_\text{2D})\geq0$, which is taken from the cost map belonging to $X_k$. A continuous bump function is defined to account for traveling costs, and is composed of two parts.

The first part of the bump function corresponds to short trajectories. Consider $d_\text{opt}$, the time-optimal distance that a robot may travel at full speed during the time between exploration updates. Here, we consider the case when $d(X_k,X_c,m_\text{2D})\leq d_\text{opt}$, i.e., the robot can completely traverse this distance during the allotted travel time. If $d(X_k,X_c,m_\text{2D})\ll d_\text{opt}$, then the robot is not moving at its full potential. This can be wasteful, because the robot tends to capture larger regions while moving. Therefore, it is desirable for $d(X_k,X_c,m_\text{2D})\rightarrow d_\text{opt}$, which corresponds to maximizing map coverage without time cost. The first part of the bump function is sinusoidal and is optimized at $d_\text{opt}$ to maximize robotic movement as
\begin{align}
\mathcal B_1(d)=\frac12\mathcal B_\text{max}\left(1-\cos{\frac{d\pi}{d_\text{opt}}}\right),
\end{align}
where $\mathcal B_\text{max}>0$ is the maximum value when $d(X_k,X_c,m_\text{2D})=d_\text{opt}$.

The second part of the bump function is defined over the domain where $d(X_k,X_c,m_\text{2D})>d_\text{opt}$ to minimize time required for $X_k$ to arrive at $X_c$, prioritizing local movements rather than traversing across the environment. This choice is beneficial to generating accurate local maps before exploring new regions. The second half is defined to be nonzero everywhere and strictly decreasing as 
\begin{align}
\mathcal B_2(d)=(\mathcal B_\text{max}-\mathcal B_\text{far})\exp\braces{-\beta(d_\text{opt}-d)^2}+\mathcal B_\text{far},
\end{align}
where $\mathcal B_\text{max}>\mathcal B_\text{far}>0$ guarantees $\mathcal B_2>0$ such that $\mathcal B_2\rightarrow \mathcal B_\text{far}$ as $d(X_k,X_c,m_\text{2D})\rightarrow\infty$ and $\beta>0$ assigns the rate of functional decrease relative to $\mathcal B_\text{max}$ and $\mathcal B_\text{far}$. Then, the complete bump function is defined as
\begin{align}
\label{eqn:bumpFun}
\mathcal B(d)=
\begin{cases}
    \mathcal B_1(d),		& \text{if }d\leq d_\text{opt},\\
    \mathcal B_2(d),         & \text{otherwise},
\end{cases}
\end{align}
which is illustrated in Fig. \ref{fig:nonzeroBumpFun}. % In short, the bump function prioritizes short movements to capture the local environment over long traversals for exploration time efficiency.

	\begin{figure}
		\centerline{
			\includegraphics[width=0.9\columnwidth]{NonZeroBump.pdf}
		}
		\caption{Bump Function to Optimize Robot Exploration Time}
		\label{fig:nonzeroBumpFun}
	\end{figure}

Then, using the information gain from \refeqn{expectedInfoGainRay} and considering the impact from travel distance from \refeqn{bumpFun}, the objective function for a single agent with respect to $X_{c}$ is
\begin{align}
\label{eqn:CandidateBidSingle}
\text{Obj}_\text{single}(X_k,X_c,m_\text{2D})&=\mathcal B(d(X_k,X_c,m_\text{2D}))\mathcal I(X_{c},m_\text{2D}).
%X_{k,c}^*&=\argmax_{X_c}{\ \mathcal B_{k,c}\mathcal I(X_c)},
\end{align}
Then, the optimal pose selection for the $k$-th agent is
\begin{align}
\label{eqn:OptPoseSingle}
X^*_{k}=\argmax_{X_c}\text{Obj}_\text{single}(X_k,X_c,m_\text{2D}).
\end{align}
Therefore, the bump function from \refeqn{bumpFun} prioritizes robotic movement within the local surrounding space before the robot moves across the map to distant regions, while still considering faraway pose candidates.

The first auction includes bids from all robots, and proceeds as follows. Let the set of all $n_R$ robots be denoted $\mathcal R=\braces{R_1,R_2,\dots,R_{n_R}}$ such that the $k$-th robot, namely $R_k$, bids $\text{Obj}_\text{single}(X^*_k,X_c,m_\text{2D})$ from \refeqn{CandidateBidSingle} and \refeqn{OptPoseSingle}. This is repeated for all $k$ such that $R_k\in\mathcal R$. The robot with the largest bid wins the first auction at index $k^*$, and is assigned to travel to $X^*_{k^*}$. Then, the remaining robots must account for $R_{k^*}$ in subsequent auctions, described next.

\subsection{Subsequent Auctions}

The robots unable to win the first auction are updated to reflect the expected impact of $R_{k^*}$. This robot is expected to modify the probabilistic occupancy grid map, which may significantly change the expected information gains and collision properties of candidate poses nearby $X^*_{k^*}$. The process of auctioning and modifying bids is repeated, removing the winning robot from consideration after each auction, until no robots remain. Between auctions, updating the map with expected measurements from auction-winning robots discourages the remaining robots from entering the same regions and capturing the same cells. As such, there is no need to consider the coverage overlap explicitly: it is avoided in a systematic way as increased coverage overlap would reduce the overall information gain.

% Furthermore, an additional term serves to avoid collisions between robots.

%When agents submit their first candidate bid, namely $\text{Obj}_\text{single}(X^*_k,X_c,m_\text{2D})$ from \refeqn{CandidateBidSingle} and \refeqn{OptPoseSingle}, this serves to maximize the expected information gain of the map while accounting for distance. However, this bid alone disregards the impact of other robots. Hence, using bids generated this way repeatedly would lead to uncoordinated robot efforts.

More specifically, we coordinate robotic efforts by modifying bids to prevent robots from updating the same grid cells and to avoid collisions among robots. The goal is to find the winning bids $\mathcal X^*=\braces{X^*_{1},X^*_{2},\dots,X^*_{n_R}}$ corresponding to each robot from $\mathcal R$. During the auctioning process, let $\mathcal W\subset\mathcal R$ be the set of robots that have already won auctions by producing the largest objective function bid. After the first auction described above, $\mathcal W=\braces{R_{k^*}}$; after all auctions are complete, $\mathcal W=\mathcal R$. Between auctions, the candidate poses are modified for information gain and collision-avoidance, coordinating the multi-vehicle exploration.

% The main idea is that accounting for the expected change in the occupancy grid map from an auction-winning robot provides a multi-vehicle exploration policy that avoids excessive coverage of the same grid cells.

The first step between auctions is to modify the expected map information gain for efficient map coverage.  Once a robot wins a bid, the measurement ray expected values serve to update a temporary copy of $m_\text{2D}$, namely $m_\text{2D,copy}$, and those candidates located in a local neighborhood (twice the radius of a maximum sensor reading) of the winning candidate from the prior auction are recalculated, thereby coordinating total map information gain with auction-winning robots. Using the same local map notation from \refeqn{allEta} and \refeqn{Unnormalized} along a measurement ray, the expected measurement value is
\begin{align}
\label{eqn:ExpectedMeasRay}
\text{E}[z]=\sum_{k=1}^{n_{r}}\bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}\mathbf{P}_k^-z_k,
\end{align}
where $z_k$ denotes the distance from the robot sensor to the $k$-th cell along the measurement ray. Then $\text{E}[z]$ is substituted into \refeqn{RayISMAnswer}--\refeqn{Unnormalized} to modify $m_\text{2D,copy}$, and expected map information gains are recomputed with \refeqn{ProbMeas} and \refeqn{expectedInfoGainRay}, while the bump function \refeqn{bumpFun} remains the same.

The second step focusses on collision-avoidance between robots based on their proximity. Let $\rho_\text{max}>0$ be a fixed maximum radius to consider collision-avoidance and $\rho_{i,c}\geq0$ be the Euclidean distance from the $i$-th already-assigned pose $X^*_i$ to the $c$-th candidate pose $X_c$, i.e.,
\begin{align}
\rho_{i,c}&=\norm{x^*_i-x_c},
\end{align}
where $x^*_i$ and $x_c$ are the pose locations of $X^*_i$ and $X_c$, respectively. Then the collision-avoidance factor for the $k$-th robot such that $R_k\notin\mathcal W$ is,
\begin{align}
\label{eqn:CollisionAvoidanceAmongRobots}
\mathbf C(X_c)&=
\begin{cases}
    \prod_{i|\mathcal{R}_i\in\mathcal W} \left(\frac{\rho_{i,c}}{\rho_\text{max}}\right)^2,		& \text{if }\rho_{i,c}<\rho_\text{max},\\
    1,              				& \text{otherwise},
\end{cases}
\end{align}
where this product serves to decrease the value of bids for candidate poses in close proximity with already-assigned poses to avoid collisions between robots. The multi-vehicle objective function for the $k$-th robot accounting for collision-avoidance is
\begin{align}
\label{eqn:CandidateBidMulti}
\text{Obj}_\text{multi}&(X_k,X_c,m_\text{2D,copy})
\nonumber\\&=\mathbf C(X_c)\mathcal B(d(X_k,X_c,m_\text{2D}))\mathcal I(X_{c},m_\text{2D,copy}),
\end{align}
where its optimal pose during this auction is
\begin{align}
\label{eqn:OptPoseMulti}
X^*_{k}=\argmax_{X_c}\text{Obj}_\text{multi}&(X_k,X_c,m_\text{2D,copy}),
\end{align}
where $\text{Obj}_\text{multi}(X^*_k,X_c,m_\text{2D,copy})$ is the bid for the $k$-th robot.
Optimal pose selection and bidding is repeated for all robots not belonging to $\mathcal W$. The largest bid among these wins the auction and is tasked with moving to the associated pose candidate, and then this robot is included with $\mathcal W$ to avoid further consideration. Completing $n_R$ auctions produces $\mathcal X^*$.

%In short, the robot with the largest bid wins the first auction, then discounts all candidate poses within a small neighborhood of the auction-winning robot, and this robot is removed from further auctions. This process is repeated, removing a robot from consideration each auction, until just a single robot remains.

The proposed approach uses auctions and bid modifications based on total map information gain and collision-avoidance, promoting exploration of different spaces without explicit consideration of coverage overlaps. The pseudocode for the bidding process is shown with Algorithm \ref{alg:bidding}, and a high level diagram of the entire process is illustrated with \ref{fig:MultiVehicleAutonomousExplorationDiagram}.

\begin{algorithm}
	Function: $Bidding(\mathcal X_r,\mathcal I(X_c)\forall c\in\mathcal C,\rho_\text{max})$\;
	Initialize $k=0$, $\mathcal W=0_{n_R\times1}$, $B=0_{n_R\times1}$\;
	Update candidate expected information gains using \refeqn{DiscExpEntropyRay}\;
	\For{$i=n_R,n_R-1,\ldots,1$}{
		\For{$j=1,2,\ldots,n_R$}{
			\If{$\mathcal W(j)==0$}{
				\If{$i==n_R$}{
					Maximize bid \refeqn{CandidateBidSingle} for $X^*_{j}$ from \refeqn{OptPoseSingle}\;
				}
				\Else{
					Update $\mathcal I(X_c)$ close to $\mathcal X^*(k)$\;
					\For{$c=1,2,\ldots,n_c$}{
						Find Euclidean distance $\rho_{k,c}$\;
					Get $\mathbf C(X_c)$ with \refeqn{CollisionAvoidanceAmongRobots}\;
					}
					Maximize bid \refeqn{CandidateBidMulti} for $X^*_{j}$ from \refeqn{OptPoseMulti}\;
				}
				Insert the $j$-th maximum bid into $B(j)$\;
			}
		}
		$k$: index maximizing $B$ with corresponding $X^*_{k}$\;
		$\mathcal X^*(k)=X^*_{k}$\;
		$\mathcal W(k)=1$\;
		$B(k)=0$\;
	}

	Return $\mathcal X^*$\;
\caption{Robot Task Bidding}
\label{alg:bidding}
\end{algorithm}

	\begin{figure}
		\centerline{
			\includegraphics[width=0.9\columnwidth]{MultiVehicleAutonomousExplorationDiagram.png}
		}
		\caption{Multi-Vehicle Autonomous Exploration Diagram}
		\label{fig:MultiVehicleAutonomousExplorationDiagram}
	\end{figure}

%              C++
%              if(distToOptimalPose < explore.discountRadius)
%                proximityDiscount = pow(distToOptimalPose/explore.discountRadius, 2);
%              else
%                proximityDiscount = 1.0;

%\begin{align}
%\label{eqn:bumpFun}
%\text{bump}(d_\text{cell})= 
%\begin{cases}
%    e^{1-\frac1{1-(d_\text{cell}/d_\text{max})^2}},			& \text{if }-d_\text{max}<d_\text{cell}<d_\text{max}\\
%    0,              				& \text{otherwise}
%\end{cases},
%\end{align}

\subsection{Receding Horizon Framework}
This algorithm is further aided by following a receding horizon framework for improved information gain maximizations and collision-avoidance with dynamic obstacles such as other robots. A receding horizon simply repeats the autonomous exploration steps as quickly as possible, frequently before a robot reaches its desired pose. Since exploration optimizations can only occur during exploration updates, a receding horizon maximizes the rate at which these updates occur. These updates require time for computation, over which the robot can move $d_\text{opt}$, used to optimize the robot travel distances with \refeqn{bumpFun}. Hence, the receding horizon framework serves to repeat optimizations as quickly as possible with maximal robotic movement while exploring a changing occupancy grid.

Furthermore, a receding horizon framework enhances autonomous exploration in two other ways. First, the occupancy grid map is constantly updated while a robot traverses a trajectory, so the expected information gains change as well. Since the bids depend on the occupancy grid, $\mathcal X^*$ is updated accordingly. This prevents robots from completing trajectories that have become unnecessary as new terrain becomes captured. Second, the receding horizon prevents robots from colliding with each other when they become too close together (e.g., crossing paths, traversing a tight passage). The cost map for each robot is updated, and the location of other robots are considered as collision-prone space. Hence, the robot trajectories become better separated due to rapid updates of the cost maps with the receding horizon.

\section{Numerical Results}
\label{sec:NumericalSim}

\subsection{Software Structure}

The algorithms for mapping, exploration, and visualization are designed for the Robot Operating System (ROS). This framework allows the node that computes 3D mapping probabilities to easily transfer these to an exploration node and a visualization node. The exploration node is separated into three threads running in parallel. The first thread constantly updates the 2D projected map according to~\cite{KauTakAiLee18}. The second thread computes the information gains, runs auctions, and generates trajectories. The third thread broadcasts a message for visualizing the 2D projected map. This structure is chosen because the 2D map cannot afford to miss messages, and the last two may produce time bottlenecks and are independent of each other. The final node is for 3D map visualization, and it creates a message to the ROS visualization package, Rviz, to produce 3D cells with varying opacities corresponding to cell occupancy probabilities.

The multi-vehicle exploration problem is typically applied to large environments, so several scalability precautions are taken to avoid large communications and computations. Most importantly, only the map \emph{changes} are communicated from the mapping node to the exploration and visualization nodes using a custom message type. This change drastically decreases the time needed to generate or interpret the message. This is accomplished by finding a rectangular prism of the minimum and maximum limits of cells that may have changed during a particular measurement scan update. Publishing and subscribing to map changes provides an efficient means to transfer mapping information, independent of map size.

Another computational improvement key to autonomous exploration success is that only important and changing map information gains are updated. The expected information gain for a single candidate is updated ray-by-ray, as demonstrated in~\cite{KauTakAiLee17,KauTakAiLee18}. For every measurement ray, only the top $n_p$ cells with the largest detection probabilities are considered. Since \refeqn{ProbMeas} is embedded in \refeqn{DiscExpEntropyRay}, the computational complexity for a ray with $n_{r}$ cells is $\mathcal O(n_{r}^2)$. Reducing the considered cells to $n_p$ limits the computation to $\mathcal O(n_{p}^2)$ where $n_{p}\ll n_r$ in general. Furthermore, finding these $n_p$ cells is computationally inexpensive; using a selection algorithm of the $n_{r}$ cells and a sorting algorithm of the $n_{p}$ cells, the computational complexity is amortized to $\mathcal O(n_r+n_p\log(n_p))$. This entropy updating scheme scales well to varying sensor ranges and cell sizes. Furthermore, the information gains of poses far from any robot trajectory need not be updated between receding horizons.

\subsection{Parameters and Resulting Maps}
The 3D mapping and exploration algorithms for three robots are simulated with several parameters, which are the same as those presented in~\cite{KauTakAiLee18} except the map limits are extended to $-55.0$m to $55.0$m in the x- and y-directions, and $0.0$m to $1.5$m in the z-direction, producing a total of 45,255,504 grid cells with edge length $0.075$m (dimensioned $1468\times1468\times21$). For the bump function, $\mathcal B_\text{max}=1$, $\mathcal B_\text{far}=0.1$, and $\beta=0.01$ are chosen. For bidding $\rho_\text{max}=3.0$m is chosen.

In the simulation, three quadrotors explore the simulated environment of The George Washington University's (GWU) Science and Engineering Hall (SEH) for $10$ minutes while taking measurements. Each quadrotor is given an Asus Xtion depth sensor and Hokuyo LIDAR, where the sensor properties for the Asus Xtion are used with exploration expectations because they provide the vast majority of measurements. The building floor plan and Gazebo simulated environment are shown in Fig. \ref{fig:SEHEnvironment}. The 3D occupancy grid maps are shown in Fig. \ref{fig:sim3DMap}, and the corresponding 2D projected maps for exploration are shown in Fig. \ref{fig:sim2Dmaps} using the ROS visualization package Rviz. The full video of both maps is available at https://youtu.be/OMn2c453oik.

	\begin{figure}
		\centerline{
			\subfigure[Floor Plan]{\includegraphics[height=0.5\columnwidth]{SEH_cropped.png}}
			\hspace*{0.025\columnwidth}
			\subfigure[Gazebo Environment]{\includegraphics[height=0.5\columnwidth]{gazeboSEH.png}}
		}
		\caption{Simulated Environment of SEH}
			\label{fig:SEHEnvironment}
	\end{figure}

	\begin{figure}
	% trim={<left> <lower> <right> <upper>}
		\centerline{
			\subfigure[$0$ min]{\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, height=0.45\columnwidth]{multi_discount_raw_3D_0min_ss.jpg}}
			\hspace*{0.025\columnwidth}
			\subfigure[$1$ min]{\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, height=0.45\columnwidth]{multi_discount_raw_3D_1min_ss.jpg}}
		}
		\vspace*{0.025\columnwidth}
		\centerline{
			\subfigure[$2$ min]{\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, height=0.45\columnwidth]{multi_discount_raw_3D_2min_ss.jpg}}
			\hspace*{0.025\columnwidth}
			\subfigure[$3$ min]{\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, height=0.45\columnwidth]{multi_discount_raw_3D_3min_ss.jpg}}
		}
		\vspace*{0.025\columnwidth}
		\centerline{
			\subfigure[$4$ min]{\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, height=0.45\columnwidth]{multi_discount_raw_3D_4min_ss.jpg}}
			\hspace*{0.025\columnwidth}
			\subfigure[$5$ min]{\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, height=0.45\columnwidth]{multi_discount_raw_3D_5min_ss.jpg}}
		}
		\vspace*{0.025\columnwidth}
		\centerline{
			\subfigure[$6$ min]{\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, height=0.45\columnwidth]{multi_discount_raw_3D_6min_ss.jpg}}
			\hspace*{0.025\columnwidth}
			\subfigure[$7$ min]{\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, height=0.45\columnwidth]{multi_discount_raw_3D_7min_ss.jpg}}
		}
		\vspace*{0.025\columnwidth}
		\centerline{
			\subfigure[$8$ min]{\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, height=0.45\columnwidth]{multi_discount_raw_3D_8min_ss.jpg}}
			\hspace*{0.025\columnwidth}
			\subfigure[$9$ min]{\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, height=0.45\columnwidth]{multi_discount_raw_3D_9min_ss.jpg}}
		}
		\vspace*{0.025\columnwidth}
		\caption{3D Occupancy Grid Map of SEH Second Floor}
			\label{fig:sim3DMap}
	\end{figure}

	\begin{figure}
	% trim={<left> <lower> <right> <upper>}
		\centerline{
			\subfigure[$0$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_0min.jpg}}
			\hspace*{0.025\columnwidth}
			\subfigure[$1$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_1min.jpg}}
		}
		\vspace*{0.025\columnwidth}
		\centerline{
			\subfigure[$2$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_2min.jpg}}
			\hspace*{0.025\columnwidth}
			\subfigure[$3$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_3min.jpg}}
		}
		\vspace*{0.025\columnwidth}
		\centerline{
			\subfigure[$4$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_4min.jpg}}
			\hspace*{0.025\columnwidth}
			\subfigure[$5$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_5min.jpg}}
		}
		\vspace*{0.025\columnwidth}
		\centerline{
			\subfigure[$6$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_6min.jpg}}
			\hspace*{0.025\columnwidth}
			\subfigure[$7$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_7min.jpg}}
		}
		\vspace*{0.025\columnwidth}
		\centerline{
			\subfigure[$8$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_8min.jpg}}
			\hspace*{0.025\columnwidth}
			\subfigure[$9$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_9min.jpg}}
		}
		\vspace*{0.025\columnwidth}
		\caption{2D Projected Occupancy Grid Map of SEH Second Floor}
			\label{fig:sim2Dmaps}
	\end{figure}

The resulting 3D occupancy grid shows a nearly-completed map after three quadrotors cooperatively explored over $10$ minutes, which is quantified with total map entropy from \refeqn{ShannonsEntropyMap}, shown in Fig. \ref{fig:simH}. An important observation is that the majority of map space was unreachable by any robot, so about $1.2\times10^7$ of the entropy metric could not be improved, regardless of mapping or exploration strategy.

Special attention had to be placed on parameter selection for the bump function and collision-avoidance. For the bump function, increasing the parameter $\beta$ narrows the bump, placing a stronger influence on nearby candidates. However, this choice corresponds to the bump function decreasing at a negligible rate beyond a close neighborhood of the robot, i.e., a large range of distances yield $\mathcal B\approx\mathcal B_\text{far}$. Hence, there is a tradeoff between prioritization of local movements and being able to differentiate the effects of distance on large trajectories.

When robots competed for future poses in a centralized auction framework, the repeated optimizations provided an effective multi-vehicle exploration policy. Between auctions, only a small set of candidates within the neighborhood of the last auction-winning candidate require consideration when updating expected information gains with \refeqn{ExpectedMeasRay} and avoiding collisions between robots using \refeqn{CollisionAvoidanceAmongRobots}, both of which modify the optimization of \refeqn{OptPoseMulti}. Since bidding is applied efficiently, its computation time is negligible compared with other parts of the exploration update such as computing the initial candidate information gains and robot cost maps.

% discounting function \refeqn{discount} has an important parameter, namely $\rho_\text{max}$, which determines the size of the circular region around the robot where discounts are applied. Once again, the parameter introduces a tradeoff; small values of $\rho_\text{discount}$ increase the probability of coverage overlaps and collisions, but larger values only promote a single vehicle from viewing a particular uncertain space. The authors found that lower values of $\rho_\text{max}$ (in this case $3$m was chosen) performed better because the receding horizon updating framework in conjunction with bidding discounting quickly exposes coverage overlaps or possible future collisions. Furthermore, a small $\rho_\text{max}$ has less impact on \refeqn{CandidateBidMulti}, and thus the exploration algorithm is closer to optimal.

% discussion on bump function and discount

	\begin{figure}
		\centerline{
			\includegraphics[width=0.9\columnwidth]{multi_H.pdf}
		}
		\caption{Total 3D Map Cell Entropy}
		\label{fig:simH}
	\end{figure}

\section{Conclusions}
\label{sec:Conclusions}

This paper introduced a bidding-based solution to the multi-vehicle exploration problem using repeated optimizations of expected map entropy changes and travel distances. The proposed approach integrates recent developments in obtaining occupancy grid probabilities and entropies into several auctions to assign tasks to robots. The bidding process relies on expected map information gains and travel distances, and modifies bids between auctions to promote robotic cooperation. Information gains are modified by updating a temporary expected map. Then, an additional term prevents collisions among robots. Furthermore, a receding horizon constantly updates the bidding process and improves collision avoidance within a dynamic environment. The efficacy of the approach is shown with a numerical simulation, with several computational enhancements for scalability of the algorithms.% Future work includes capturing more features with a 3D exploration map and employing a distributed framework for mapping and exploration.

\bibliography{../../BibSources}% master source for all publications always 2 directories up
\bibliographystyle{IEEEtran}

\end{document}

\end{document}

### 2018-06-09 Conference/IRC 2018/IRC18.tex
added Conference folder

--- 2018-06-09 Misc/IRC18_backup.tex
@@
 
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
-%\usepackage[tight,footnotesize]{subfigure}
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\usepackage[caption=false]{caption}
-%\usepackage[font=footnotesize]{subfig}
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
 % correct bad hyphenation here
@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Autonomous Quadrotor 3D Mapping and Exploration Using Exact Probabilities and Entropies}
-
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-
-%\author{Evan Kaufman, Taeyoung Lee, and Zhuming Ai%, and Ira S. Moskowitz%
-%\thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 {\tt \{evankaufman,tylee\}@gwu.edu}}
-%\thanks{Zhuming Ai, Information Management \& Decision Architectures, U.S. Naval Research Laboratory,  Washington, DC 20375}
-%\thanks{This research has been supported by the U.S.
-Naval Research Laboratory Base Program ``Intelligent Microflyer'' and in part by NSF under the grants CMMI-1243000, CMMI-1335008, and CNS-1337722.}
-%}
-
-%\author{\IEEEauthorblockN{Evan Kaufman, Kuya Takami, Taeyoung Lee}
-%\IEEEauthorblockA{Department of Mechanical and Aerospace Engineering\\The George Washington University\\
-%Washington, DC 20052\\
-%Email: \{evankaufman,kuya,tylee\}@gwu.edu}
-%\and
-%\IEEEauthorblockN{Zhuming Ai}
-%\IEEEauthorblockA{Information Management \& Decision Architectures
-%              \\ U.S. Naval Research Laboratory \\
-%              Washington, DC 20375}
-%}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
+\title{Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities}
 
@@
 
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
 \maketitle
@@
 \begin{abstract}
-%\boldmath
-Robotic mapping and exploration has received significant attention recently because exploring uncertain environments can pose danger to humans, and robots can produce highly detailed maps.
+%Robotic mapping and exploration have received significant attention recently because exploring uncertain environments can pose danger to humans, and robots can produce highly detailed maps.
 Commonly uncertain environments contain complex geometries and robots frequently face restrictive terrains, motivating a 3D representation of mapping from a vehicle capable of hover.
-This paper focusses on the popular occupancy grid mapping representation, with the goal of determining occupancy probabilities of evenly-spaced grid cells in 3D.
-With recent developments from the authors using the stochastic properties of depth sensors directly with a rigorous, probabilistic formulation, this paper extends this precise mapping technique to 3D for real-time implementation with realistic sensor capabilities.
-Furthermore, important aspects of the 3D map are projected onto 2D maps, where a predicted level of map uncertainty, known as Shannon's entropy, provides an exploration policy that governs robotic motion.
-Both mapping and exploration algorithms produce important information that impacts each other and are demonstrated simultaneously with numerical simulations and experiments in real-time.
+
+This paper deals with the aerial exploration for an unknown three-dimensional environment, where Bayesian probabilistic mapping is integrated with a stochastic motion planning scheme to minimize the map uncertainties in an optimal fashion.
+We utilize the popular occupancy grid mapping representation, with the goal of determining occupancy probabilities of evenly-spaced grid cells in 3D with sensor fusion from multiple depth sensors with realistic sensor capabilities.
+The 3D exploration problem is decomposed into 3D mapping and 2D motion planning for efficient real-time implementation.
+This is achieved by projecting important aspects of the 3D map onto 2D maps, where a predicted level of map uncertainty, known as Shannon's entropy, provides an exploration policy that governs robotic motion.
+Both mapping and exploration algorithms are demonstrated with both numerical simulations and quadrotor flight experiments.
 \end{abstract}
@@
 
-Robotic mapping is the process of generating maps, which represent the environment surrounding a robot. 
-Effectively modeling this environment is crucial when building an understanding of an initially uncertain environment.
-Robotic mapping serves an important role in tasks such as search-and-rescue, surveillance, and robotic cleaning.
-Several grid-based map representations have been studied, e.g.,~\cite{WolSuk05,MeyBeiBur12,TanThoWolBus14}, because they provide abundant occupancy knowledge of the environment are computationally tractable.
+Robotic mapping is the process of generating maps representing the environment surrounding a robot.
+Mapping serves an important role in tasks such as search-and-rescue, surveillance, and robotic cleaning.
+Several grid-based map representations have been studied, e.g.,~\cite{WolSuk05,MeyBeiBur12,TanThoWolBus14}, because they provide abundant occupancy knowledge about the environment and are computationally tractable.
 
@@
 In particular, occupancy probabilities have been assigned to a uniform 3D grid in~\cite{Andert09,PirRutBisSch11}.
-Another highly-popular variation of occupancy grid mapping is the Octomap representation~\cite{WurHorBenStaBur10,HorWurBenStaBur13}, which provides powerful open-source software tools for representing free and occupied spaces nearby the robot, with the capability of extending the mapping limits and changing cell sizes.
-However, Octomaps produce cell occupancy probabilities that do not directly depend on the depth sensor stochastic properties, known as the forward sensor model.
-Ideally, the probability density provided by the forward sensor model should be used explicitly when determining occupancy probabilities of grid cells conditioned on the robot pose and measurements, known as the inverse sensor model.
-Instead, these approaches mentioned above determine occupancy probabilities using an approximate function to replace the true inverse sensor model.
-Additionally, this approach replaces probabilities with log-odds ratios, then combines past and current information with a potentially harmful Markov assumption.
+Another highly-popular variation of occupancy grid mapping is the Octomap representation~\cite{WurHorBenStaBur10,HorWurBenStaBur13}, which has added features such as extending the mapping limits and changing cell sizes.
 
-Alternatively, another popular approach is to simulate maps, robot poses, and measurements and use a learning algorithm, to approximate the inverse sensor model~\cite{Thr01,ThrBurFox05}, which is implemented in 3D with~\cite{SouMaiGon12}.
-These approaches are undesirable in practice due to complexities associated with implementing a learning algorithm.
-For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples, or how many samples are required to obtain a reasonable approximation.
-Furthermore, it is challenging to apply any learning algorithm over the large dimensional space composed of maps, poses, and measurements.
+However, these approaches are based on an approximate solution of the inverse sensor model, to determine the occupancy probabilities based on the depth sensor measurements.
+As such, the prior belief of the environment and the stochastic properties of the sensor are not properly accounted.
+Another popular approach for the inverse sensor model involves a simulated learning technique~\cite{Thr01,ThrBurFox05,SouMaiGon12}, but this is not suitable for the higher dimensional configurations composed of maps, poses, and measurement due to the large number of required samples.
+None of these approaches use the sensor properties to directly obtain a Bayesian solution to occupancy probability.
 
-The approximate nature of the inverse sensor model motivated the authors of this paper to design computationally-efficient algorithm to construct the exact inverse sensor model that follows a Bayesian framework~\cite{KauLeeAiMos16,KauTakAiLee17}.
-The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancies of the other cells blocked by it are irrelevant to the forward sensor model, and this model is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model.
-This method combines a priori probabilities of occupancy and multiple range measurements according to a Bayesian framework to obtain more accurate maps.
-As such, it contrasts from the existing framework based on log-odds ratios that impose an additional Markov assumption or particle-filtering methods such as~\cite{TanThoWolBus14} that require particle weighting and resampling to obtain a variation on a Bayesian solution.
-This paper proposes extending this algorithm to 3D with realistic sensor limitations.
-We demonstrate the efficacy of the proposed algorithm using two different depth sensors simultaneously to generate a 3D occupancy grid map with accurate probabilities in real-time with numerical simulations and experiments.
+%        1.      sensor fusion in mapping
+The approximate nature of the occupancy probability motivates an exact Bayesian solution that can be applied in 3D.
+This paper proposes a computationally-efficient algorithm to obtain the exact solution to the inverse sensor problem to generate probabilities in a 3D environment, following the framework of~\cite{KauLeeAiMos16,KauTakAiLee17}.
+The proposed approach integrates all of the stochastic information available for the environment, such as the prior map estimate and multiple sensor measurements, according to the Bayesian framework, to obtain the exact 3D occupancy probabilities.
+As such, it provides a unified framework for sensor fusion between multiple sensors with various properties, such as sensor accuracy, and fields of view.
 
-The second topic of this paper deals with autonomous exploration.
-A common approach to solve the autonomous exploration problem is known as frontier-based exploration, proposed for 2D environments in~\cite{Yam97,Yam98}.
-This is the process of executing actions that move the robot toward the closest boundary between visited and unvisited space, known as a frontier.
-Then the robot takes measurements at this location such that the mapped territory expands, and thus the new frontiers are pushed back. This process is repeated until the map is well-known.
-This frontier-based approach is extended to 3D with~\cite{ZhuDinLinWu15,SenWan16,KleDor13} using the Octomap representation, and similarly in~\cite{SawKriSri09} using a visibility metric.
-Frontier-based exploration assumes that repeatedly moving toward the closest frontier and taking measurements are the best actions to gain new information about the map, though these systematic actions are not based on any consideration of the future uncertainty of the probabilistic map or optimality.
+%        2.
+proposing a method to combine 3D mapping and 2D exploration: Also, we need to motivate why we are not doing full 3D exploration.
+Stress that we are proposing a computationally friendly approach to perform 3D mapping with projection using the fact there is less variation in the floor plan along the vertical direction.
+The second topic of this paper deals with autonomous exploration, which provides the policies that govern robotic motion.
+A common approach is known as frontier-based exploration, proposed for 2D environments in~\cite{Yam97,Yam98} and extended to 3D  using the Octomap representation in~\cite{ZhuDinLinWu15,SenWan16,KleDor13}, and similarly using a visibility metric in~\cite{SawKriSri09}.
+This approach moves robots toward the boarder between certain and uncertain spaces, though these systematic actions are not based on any consideration of the future probabilistic map uncertainty.
 
-Alternatively, there have been exploration techniques that are based on a quantitative measure of uncertainty known as Shannon's entropy~\cite{StaGriBur05}, an uncertainty metric based on grid cell occupancy probabilities.
-However, the expected entropy of an action is not directly calculated.
-Instead, this approach assumes that expected entropy is equivalent to entropy based on expected measurement scans; a Rao-Blackwellized particle filter estimates the robot pose and map, where expected measurements are drawn from each particle when finding the expected information gains, then combined with the weighting on each particle.
-However, assuming an expected measurement is sufficient to predict expected information is subject to inaccuracies due to the nonlinear nature of object detection probability and entropy.
-For these reasons, the authors have proposed a direct calculation of expected information gain from occupancy probabilities and stochastic sensor models~\cite{KauAiLee16,KauTakAiLee17}.
-In this paper, this concept is extended to 3D, where grid cells above and below a predetermined exploration height are projected onto 2D maps for entropy and collision calculations.
-In this sense, the proposed approach is entropy-based in 3D similar to~\cite{JohStaPfaBur07} where uncertainties are projected onto a 2D map, except that the entropy measures are based on map probabilities and sensor properties directly.
+Alternatively, there have been exploration techniques that are based on a quantitative measure of uncertainty known as Shannon's entropy~\cite{StaGriBur05}, and it is used to acquire expected map information gain in~\cite{KauAiLee16,KauTakAiLee17}.
+This paper focusses on extending these concepts to 3D in a numerically efficient manner.
+The prescribed 3D exact inverse sensor model is critical in computing the expected information gain precisely.
+There are several challenges with obtaining expected information gain of 3D measurement scans, particularly due to the extended computational load.
+Considering the typically low variation in floor plans along the vertical direction, this paper proposes a numerically efficient scheme that projects stochastic properties of the 3D probabilistic occupancy grid to 2D spaces for optimal information gain and collision avoidance.
+This exploits the important characteristics of 3D maps.
+This process of combining 3D mapping and 2D exploration is completed in a computationally-efficient manner.
+
+%        3.      hardware/software for aerial exploration experiment.
+Finally, we present both numerical simulations and flight experiments for the proposed exploration schemes.
+The multiple tasks of sensor data acquisition, mapping, exploration, flight control, and communications are implemented with Python/C++, to be executed simultaneously in a ROS environment.
+They are verified with a quadrotor unmanned aerial vehicle operated by the NVidia Jetson computing module with an ASUS Xtion infrared depth sensor and Hokuyo laser scanner.
 
@@
 A numerical example is demonstrated in Section IV.
-Experimental results are shown in Section V, followed by conclusions.
+Experimental results are shown in Section V, followed by conclusions and future work.
+
+%% Prior to TLee edits:
+%Robotic mapping is the process of generating maps representing the environment surrounding a robot.
+%Mapping serves an important role in tasks such as search-and-rescue, surveillance, and robotic cleaning.
+%Several grid-based map representations have been studied, e.g.,~\cite{WolSuk05,MeyBeiBur12,TanThoWolBus14}, because they provide abundant occupancy knowledge about the environment and are computationally tractable.
+%
+%Most grid-based mapping representations are designed for 2D, while others are extended to capture a full 3D environment.
+In particular, occupancy probabilities have been assigned to a uniform 3D grid in~\cite{Andert09,PirRutBisSch11}.
+Another highly-popular variation of occupancy grid mapping is the Octomap representation~\cite{WurHorBenStaBur10,HorWurBenStaBur13}, which has added features such as extending the mapping limits and changing cell sizes.
+However, Octomaps, and the 3D uniform grid-based approaches mentioned above, produce cell occupancy probabilities that do not directly depend on the depth sensor stochastic properties.
+Instead, these approaches determine occupancy probabilities using an approximate function.
+Another method introduced in~\cite{Thr01,ThrBurFox05} and implemented in 3D with~\cite{SouMaiGon12} uses a simulated learning technique to find these probabilities, which may be difficult over the large dimensional space composed of maps, poses, and measurements.
+An additional approach such as~\cite{TanThoWolBus14} applies a particle filter to estimate the map.
+None of these approaches use the sensor properties to directly obtain a Bayesian solution to occupancy probability.
+%
+%%        1.      sensor fusion in mapping
+%The approximate nature of the occupancy probability motivates an exact Bayesian solution that can be applied in 3D.
+This paper proposes a computationally-efficient algorithm that follows the framework of~\cite{KauLeeAiMos16,KauTakAiLee17} to generate probabilities in a 3D environment.
+Since this approach only requires the origin and direction of sensor measurements to be known, the algorithm is designed to work for any number of sensors.
+Sensor fusion from several sources can provide important information about the environment with different levels of sensor certainty, fields of view, and vantage points.
+%
+%%        2.
+proposing a method to combine 3D mapping and 2D exploration: Also, we need to motivate why we are not doing full 3D exploration.
+Stress that we are proposing a computationally friendly approach to perform 3D mapping with projection using the fact there is less variation in the floor plan along the vertical direction.
+%The second topic of this paper deals with autonomous exploration, which provide the policies that govern robotic motion.
+A common approach is known as frontier-based exploration, proposed for 2D environments in~\cite{Yam97,Yam98} and extended to 3D with~\cite{ZhuDinLinWu15,SenWan16,KleDor13} using the Octomap representation, and similarly in~\cite{SawKriSri09} using a visibility metric.
+This approach moves robots toward the boarder between certain and uncertain space, though these systematic actions are not based on any consideration of the future probabilistic map uncertainty.
+Alternatively, there have been exploration techniques that are based on a quantitative measure of uncertainty known as Shannon's entropy~\cite{StaGriBur05}, an uncertainty metric based on grid cell occupancy probabilities.
+Shannon's entropy is used to acquire expected map information gain in~\cite{KauAiLee16,KauTakAiLee17}, and this paper focusses on extending these concepts to 3D.
+There are several challenges with obtaining expected information gain of 3D measurement scans, particularly as they relate to computational load.
+Considering the typically low variation in floor plans that robots explore during indoor explorations, this paper proposes two methods to generate 2D exploration maps that exploit the important characteristics of 3D maps.
+This process of combining 3D mapping and 2D exploration is completed in a computationally-efficient manner.
+%
+%%        3.      hardware/software for aerial exploration experiment.
+%Finally, this paper comments on the hardware and software necessary to implement these challenging algorithms for an aerial mapping and exploration experiment.
+This paper provides the basic software structure and features for mapping and generating exploration trajectories on a host computer while running depth sensors and a flight controller onboard a mobile robot.
+%
+%The paper is organized as follows.
+The problem is formulated in Section II.
+Special considerations for mapping and exploration in 3D are established in Section III.
+A numerical example is demonstrated in Section IV.
+Experimental results are shown in Section V, followed by conclusions and future work.
 
@@
 
-The key differences between the approaches presented in this paper and others is how map cell occupancy probability and measurement ray entropy are calculated directly from a priori cell probabilities and stochastic properties of the sensors involved.
+The key differences between the approaches presented in this paper and others relate to how map cell occupancy probabilities and measurement ray entropies are calculated directly from a priori cell probabilities and stochastic properties of the sensors involved.
 These are defined next.
@@
 Let the measurement scan at the $t$-th time step be $Z_t$, where $z\in Z_t$ is a single measurement ray.
-Let the $t$-th pose be denoted by $X_t=\braces{x_t,R_t}$ such that $x_t\in\Re^3$ and $R_t\in\SO=\braces{R\in\Re^{3\times3}|R^\T R=I,\det{R}=1}$, where $X_t$ is assumed to be known.
+Let the $t$-th pose be denoted by $X_t=\braces{x_t,R_t}$ with location $x_t\in\Re^3$ and attitude $R_t\in\SO=\braces{R\in\Re^{3\times3}|R^\T R=I,\det{R}=1}$, where $X_t$ is assumed to be known.
 Given the sensor probabilistic properties $p(z|X_t,m)$, commonly referred to as the \emph{forward sensor model}, the goal is finding the \emph{inverse sensor model}, $P(m|z,X_{1:t},Z_{1:t-1})$.
-To obtain this probability for all cells,~\cite{KauLeeAiMos16,KauTakAiLee17} introduced a reduced map $r\subset m$ composed of $n_r$ cells along $z$, where cells are indexed by increasing distance from $x_t$, and $\mathbf{r}_{k+}$ refers to the event that the $k$-th cell is the closest occupied space.
+To obtain this probability for all cells,~\cite{KauLeeAiMos16,KauTakAiLee17} introduced a reduced map $r\subset m$ composed of $n_r$ cells along $z$, where cells are indexed by increasing distance from $x_t$, and $\mathbf{r}_{k+}$ refers to the event that the $k$-th cell of map $r$ is the closest occupied space.
 Then, the inverse sensor model for the $k$-th cell is
@@
 \begin{align}
-P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}}.
+P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}},
 \end{align}
-
+where $\eta_{c,k}$ refers to the normalizer calculated with $z_{c,k}$.
 The goal of entropy-based autonomous exploration is to select the optimal pose $X^*_c$ that maximizes the expected information gain,
@@
 This method is easily extended to 3D space, and the inverse sensor model is applied identically.
+Since the map is updated ray-by-ray (each measurement ray is considered once by itself), this approach is extended for measurement fusion of any number of sensors; the map is updated by each ray of a scan sequentially, then the map is updated by the rays of the next scan in queue, which might have different characteristics.
+There are several reasons why a robot might be equipped with multiple sensors, e.g., different fields-of-view or depth sensor properties.
+The proposed 3D mapping approach can handle measurements from different sources with different properties.
 
@@
 
-Additionally, flying vehicles are frequently subject to fast movements, and the time stamps for pose estimates and depth sensor readings are not guaranteed to align.
+Additionally, aerial vehicles are frequently subject to fast movements, and the time stamps for pose estimates and depth sensor readings are not guaranteed to align.
 For example, the most recent pose estimate and measurement scan may have significantly different time stamps while the robot is rapidly rotating.
@@
 In particular, the expected entropy computational order of a ray that intersects $n_r$ cells is $(n_r+1)^2$.
-Since most quadrotor flight is close to level, i.e., the third body-fixed axis (up/down) of the inertial and quadrotor body-fixed frames are nearly aligned, the proposed exploration uses 2D occupancy grid maps located at the desired exploration height.
+Since most quadrotor flight is close to hover and floor plans exhibit little variation in the vertical direction, the proposed exploration uses 2D occupancy grid maps located at the desired exploration height.
 This approach decreases $n_r$ and simplifies the exploration problem with 3D space.
-These 2D maps project important properties of the 3D map, described next.
+These 2D maps project important properties of the 3D map.
 
@@
 However, this does not imply that the captured space is \emph{entirely} free or occupied (e.g.
-a floor with free space above it, an object that does not breach the upper map boundary), so applying \refeqn{ProbEntropyMap} alone could arbitrarily set $P(\mathbf{E}_k)$ close to $0$ or $1$.
+an object on the ground that does not breach the upper map boundary), so applying \refeqn{ProbEntropyMap} alone could arbitrarily set $P(\mathbf{E}_k)$ close to $0$ or $1$.
 This can have damaging effects on the expected information of cells beyond the $k$-th cell.
@@
 
-The 3D mapping and exploration algorithms are simulated using the first proposed approach with separate collision and entropy 2D projected maps.
-The Robot Operating System (ROS) provides an excellent framework for several programs to communicate easily.
-Gazebo serves as the simulator, which provides plugins to simulate an Asus Xtion color depth sensor and a Hokuyo LIDAR.
-It also provides position and attitude transforms from the world to the quadrotor body.
-Using fixed transforms between the quadrotor body and its onboard sensors, an original mapping node uses a variation of ROS message filters, known as transform message filters, to synchronize the sensor poses with the sensor reading time stamps.
-The node is designed to work for any number of sensors and uses the sensor properties directly when solving the inverse sensor model.
-In particular, the Xtion and Hokuyo are modeled using normalized Gaussian and uniform distributions for the sensor model, with a Gaussian probability of $0.9$ with $\sigma_\text{Xtion}=0.25$ and $\sigma_\text{Hokuyo}=0.1$, respectively, which are greater than their specified values to account for pose uncertainty and distortion with the Xtion.
+\subsection{Software Structure}
+%        A.      Software Structures
+%                list and explain ROS nodes
+
+The Robot Operating System (ROS) provides an excellent framework for several programs, referred to as \emph{nodes}, to communicate easily.
+These are mostly coded in C++ for speed, except for a few Python scripts for a GUI and trivial message conversions, and launch/yaml scripts for running nodes and setting parameters.
+
+Gazebo serves as the simulator node, with plugins to simulate an Asus Xtion color depth sensor and a Hokuyo LIDAR.
+Gazebo also provides position and attitude transforms from the world to the quadrotor body, and the ability to set robot model states directly.
+
+Using fixed transforms between the quadrotor body and its onboard sensors, an original mapping node uses a variation of ROS message filters, known as transform message filters, to synchronize the sensor poses with sensor readings.
+These serve as inputs for ray casting in 3D, which provides the cell probabilities and depths along the ray through geometry.
+Then, the inverse sensor model serves to update the map ray-by-ray.
+The node is designed to work for any number of sensors, specified by sensor parameters in a yaml configuration file.
+
+There are two processes for exploration.
+The first process is projecting the 3D map onto 2D maps; the same code under different conditional parameters produces two nodes, for collision and entropy maps.
+The second process subscribes to these two 2D maps; the collision map serves as input for candidate pose consideration and Dijstra's search, and the entropy map is used to predict expected information gain.
+This node provides path messages, composed of desired poses $0.01$sec apart, that are easily visualized with the ROS visualization package, Rviz.
+
+Finally, a path-interpreter node subscribes to the path messages and interpolates the pose at the current ROS time step. This pose is converted into a message that sets the Gazebo quadrotor state.
+
+\subsection{Simulation Results}
+%        B.      Simulation Results
+%                a little bit more interpretations for the results
+
+The 3D mapping and exploration algorithms are simulated with several parameters.
+In particular, the Xtion and Hokuyo are modeled using normalized Gaussian and uniform distributions for the sensor model, with a Gaussian probability of $0.9$ with $\sigma_\text{Xtion}=0.25$m and $\sigma_\text{Hokuyo}=0.1$m, respectively, which are greater than their specified values to account for pose uncertainty and distortion with the Xtion.
 Since the Xtion tends to distort depths significantly beyond $4$m, this maximum reading considered for both the Xtion and Hokuyo, and the uniform distribution (probability $0.1$) covers the sensor ranges from $0.5$m to $4$m.
 A cell size of $\alpha=0.075$m is selected and the map limits are $-4$m to $4$m in the x-direction, $-6$m to $6$m in the y-direction, and $-0.15$m to $1.5$m in the z-direction.
-The initial probability of cells was selected at $P_\text{init}=0.1$.
-
-The other original node is for autonomous exploration.
-For this section, separate 2D collision and entropy maps are used.
+The initial probability of cells is selected at $P_\text{init}=0.1$.
 For the collision map, the cell edge factor is $k_C=3$.
@@
 \begin{align}
+\label{eqn:bumpFun}
 \text{bump}(d_\text{cell})= 
@@
 For this experiment, $d_\text{max}$ is chosen as the maximum value of $d_\text{cell}$ on the cost map.
-This serves to prioritize understanding the local surrounding space before the robot moves across the map to distant regions.
+This serves to prioritize understanding the local surrounding space before the robot moves across the map to distant regions, while still considering all of the pose candidates.
 
-In the simulation, the robot explores the simulated environment for $5$min while taking measurements.
+In the simulation, the robot explores the simulated environment for $8$min while taking measurements.
 The 3D occupancy grid map and true environment are shown in Fig.
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{SimEntropy.pdf}
+	\includegraphics[width=0.4\textwidth]{entropy_gazebo_aspect3by1.png}
 	\caption{Simulated Environment Map Entropy}
@@
 The two-map approach exhibits advantages and disadvantages.
-The exploration algorithm governs robotic motion to poses that were highly beneficial to the 3D occupancy grid of the local environment.
+The exploration algorithm successfully avoided collisions and gave way to a fairly complete 3D occupancy grid.
 The floor, which is not always visible, is properly estimated in most places.
-Conversely, this attention to detail has negative consequences that encourage the robot to repeatedly look at the same regions for different vantage points.
-Occasionally an entropy map cell, composed of all cells above the floor single-stacked vertically, is rarely captured, which yields large expected information gains, even though other regions might warrant observation more.
-This is because there is no explicit consideration for the \emph{number} of 3D cells with high uncertainty to projected onto the entropy map, only a single probability associated with the maximum entropy belonging to that set.
-In short, this algorithms successfully explores the space collision-free while producing a complete map, but over-attention to just a few uncertain cells has negative consequences on the exploration time.
+Conversely, this attention to detail has negative consequences that encourage the robot to repeatedly look at the same regions from different vantage points.
+Since an entropy map cell is composed from many 3D grid cells single-stacked vertically, if just one of these 3D map cells is not captured or hardly captured, the entropy map cell will be recorded as quite uncertain, shown covering numerous locations in Fig.
+\ref{fig:sim2Dmaps}(c).
+The robot may attempt to measure this space again, even when other regions might more strongly warrant observation.
+These periods of sluggish entropy decrease are shown with flatter sections of Fig.
+\ref{fig:simH}.
+In short, this algorithm successfully explores the space collision-free while producing a complete map, but over-attention to just a few uncertain cells has negative consequences on the exploration speed.
 
-\section{Experimental Results}
+\section{Experimental Resutls}
 
-The experimental setup follows similar conditions with a few key differences.
-Most importantly, the second proposed exploration approach, namely a collision and entropy combination map, replaces the two-map approach.
-Additionally, instead of using a simulated environment such as Gazebo, where the robot state can be set and measured directly, a Vicon motion capture system provides pose data and nonlinear geometric flight controller~\cite{GooDaeLee13} provides the necessary forces and moments to follow the desired exploration trajectories, run onboard a Jetson TX2 computer.
-The Asus Xtion and Hokuyo Lidar are configured the same was as simulation, but their data is streamed from the Jetson TX2 via Wifi to a host computer running the mapping and exploration nodes with an Intel Core i7-6800K CPU (12$\times$3.40GHz).
+\subsection{Exploration Environment}
 
 The Laboratory for Autonomous Systems Research (LASR) at the U.S.
-Naval Research Laboratory (NRL) provided a large experimental space for testing.
-An experimental setup, involving Ram Board suspended from stanchions resembled a building floor plan, as well as miscellaneous objects like trash cans and tables.
-The map limits were $0.5$m to $10.3$m in the x-direction, $-8.3$m to $3.5$m in the y-direction, and $-0.15$m to $1.5$m in the z-direction.
+Naval Research Laboratory (NRL) provides a large experimental space for testing.
+An experimental setup resembles a building floor plan, which involves tan cardboard suspended from $1$m tall stanchions for a perimeter wall, an internal wall from $80/20$ supports with grey sheeting inside, and a floor that reflects some of the florescent lights above.
+Additionally, there are two wooden table desks with chairs and miscellaneous objects on top, and three plastic trash cans of varying sizes and colors.
+The map limits are designed to slightly exceed the experimental space: $0.5$m to $10.3$m in the x-direction (positive east), $-8.3$m to $3.5$m in the y-direction (positive north), and $-0.15$m to $1.5$m in the z-direction (positive vertically up).
 Pictures of the environment are shown in Fig.
-\ref{fig:exp3DEnvironment}, and the resulting 3D maps at various time steps are shown in Fig.
-\ref{fig:exp3DMap}.
-The 2D projected maps (x-direction upward) at corresponding time steps are shown in Figure \ref{fig:exp2DMap}, where trajectories are shown with curves to candidate poses, shown with arrows.
-The total map entropy as a function of time is shown in Figure \ref{fig:expH}.
+\ref{fig:exp3DEnvironment}.
 
-% trim={<left> <lower> <right> <upper>}
 \begin{figure}[!t]
@@
     	\end{subfigure}
-	\caption{3D Environment}
+	\caption{3D Experimental Environment}
 	\label{fig:exp3DEnvironment}
@@
 
+%        A.      Exploration Environment
+%                perhaps adding size, feature and materials
+
+\subsection{Hardware Structure}
+%        B.      Hardware Structure
+
+Instead of using a simulated environment such as Gazebo, where the robot state can be set and measured directly, a Vicon motion capture system produces pose data used in mapping and exploration transforms.
+An Asus Xtion and Hokuyo LIDAR are streamed from an onboard NVidia Jetson TX2 via Wifi to a host computer running the mapping and exploration nodes with an Intel Core i7-6800K CPU (12$\times$3.40GHz).
+The exploration trajectories are sent back to the Jetson TX2 over Wifi, where a nonlinear geometric flight controller~\cite{GooDaeLee13} is run onboard the Jetson to follow these trajectories.
+This configuration avoids large processing tasks onboard the robot.
+
+\subsection{Experimental Results}
+%        C.      Experiments Results
+%                a little bit more interpretations and figures for the results
+
+The experimental setup follows similar conditions as the simulated results with a few key differences.
+Most importantly, the second proposed exploration approach, namely a collision and entropy combination map, replaces the two-map approach used in simulation.
+The resulting 3D maps at various time steps are shown in Fig.
+\ref{fig:exp3DMap}.
+The 2D projected maps (x-direction upward) at corresponding time steps are illustrated in Figure \ref{fig:exp2DMap}, where trajectories are shown with curves to candidate poses are shown with arrows.
+The total map entropy as a function of time is shown in Figure \ref{fig:expH}.
+
+% trim={<left> <lower> <right> <upper>}
+
 \begin{figure}[!t]
 \centering
-    	\begin{subfigure}[t]{0.44\columnwidth}
+    	\begin{subfigure}[t]{0.88\columnwidth}
            	\centering
@@
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.44\columnwidth}
+    	\begin{subfigure}[t]{0.88\columnwidth}
            	\centering
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{ExpEntropy.pdf}
+	\includegraphics[width=0.4\textwidth]{entropy_experiment_aspect3by1.png}
 	\caption{Experimental Environment Map Entropy}
@@
 
-The projected map used for both collision and exploration demonstrated several benefits.
-The robot measured walls and objects with the 3D map, and nicely represented these as occupied space with the 2D projected map.
-Additionally, the robot explored the space much faster than with two separate maps, primarily because the exploration experiment avoided several
+The projected map used for both collision and exploration has several benefits.
+The robot measures walls and objects with the 3D map, and nicely represents these as occupied spaces with the 2D projected map.
+Additionally, the robot explores the space much faster than with two separate maps, primarily because the robot avoided repeatedly exploring spaces as with the entropy-only map in Section IV.
+This decision leads to a tradeoff; with the experiment shown here, the robot exploration did not consider low cells, particularly those near the ground.
+As a result, some floor regions are not captured by the 3D occupancy grid even though cells above the floor are well-known.
 
-\begin{itemize}
-	\item Important result: free space and walls correctly estimated, but features of the floor could be improved
-\end{itemize}
+ The 3D mapping works very well in most scenarios, but exploration requires some tuning.
+Most importantly, the value of $d_\text{max}$ in \refeqn{bumpFun} has a major impact.
+A value of $5$m is selected to prioritize local exploration trajectories first.
+This value only increases if the maximum value of $d_\text{cell}$ exceeds $d_\text{max}$, in which $d_\text{max}$ is reassigned to $d_\text{cell}$.
 
 % ATTENTION: left out patched segment size and polynomial order
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation Results}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command, the
-% \label for the overall figure must come after \caption.
-% \hfil must be used as a separator to get equal spacing.
-% The subfigure.sty package works much the same way, except \subfigure is
-% used instead of \subfloat.
-%
-%\begin{figure*}[!t]
-%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
-%\label{fig_second_case}}}
-%\caption{Simulation results}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-
-% An example of a floating table. Note that, for IEEE style tables, the 
-% \caption command should come BEFORE the table. Table text will default to
-% \footnotesize as IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that IEEE does not put floats in the very first column - or typically
-% anywhere on the first page for that matter. Also, in-text middle ("here")
-% positioning is not used. Most IEEE journals/conferences use top floats
-% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
-% command of the stfloats package.
-
 \section{Conclusions and Future Work}
 
-\begin{itemize}
-	\item Paper demonstrated results for mapping and exploration can be successfully executed in 3D
-	\item Note on differences with projected maps
-	\item Future work: projected maps can serve multiple robots working cooperatively
-	\item Future work: ray casting in 3D allows entropy calculation and exploration in 3D
-\end{itemize}
+This paper covers the important aspects of exploring and mapping a 3D environment using exact inverse sensor models and expected entropies.
+The 3D mapping produces occupancy probabilities with depth sensor fusion from an arbitrary number sensors.
+Due to the computational difficulties of exploration in 3D, a 2D version of exploration is combined with 3D mapping.
+These are demonstrated with numerical simulations and experiments with an aerial vehicle.
+
+There are two important future directions.
+One direction is a more complete 3D exploration considering the map expected information gain via measurement rays cast in 3D.
+This approach could better capture cells on the ground or vertically-nonuniform objects, but careful consideration must be placed on computational restrictions.
+Secondly, the 3D occupancy grid may be used by multiple robots for faster exploration.
+
+%Our contribution will be
+%        1.      proposing a method to combine 3D mapping and 2D exploration
+%        2.      sensor fusion in mapping
+%        3.      hardware/software for aerial exploration experiment.
+%
+%I suggest that the last few paragraphs are revised to emphasize the above.
+Perhaps, one paragraph for each of the above will be ideal.
+Also, we need to motivate why we are not doing full 3D exploration.
+Stress that we are proposing a computationally friendly approach to perform 3D mapping with projection using the fact there is less variation in the floor plan along the vertical direction.
 
@@
 % http://www.michaelshell.org/tex/ieeetran/bibtex/
-\bibliography{../../BibSources}
+%\bibliography{BibSources}% accessible locally only
+\bibliography{../../BibSources}% master source for all publications
 \bibliographystyle{IEEEtran}

### 2018-06-09 Conference/IRC 2018/IRC18_final.tex
added Conference folder

--- 2018-06-09 Conference/IRC 2018/IRC18.tex
@@
 
+% asdf Author Comments:
+
+%----------------------- REVIEW 1 ---------------------
+%PAPER: 45
+%TITLE: Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities
+%AUTHORS: Evan Kaufman, Kuya Takami, Zhuming Ai and Taeyoung Lee
+%
+%Overall evaluation: 2 (accept)
+%
+%----------- Overall evaluation -----------
+%Paper is well defined and explained.
+The literature review is also satisfactory.
+It could be better, if you give further information about how different sensors (Hokuyo, Xtion) fused together to accomplish the mapping task.
+%
+%
+%----------------------- REVIEW 2 ---------------------
+%PAPER: 45
+%TITLE: Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities
+%AUTHORS: Evan Kaufman, Kuya Takami, Zhuming Ai and Taeyoung Lee
+%
+%Overall evaluation: 0 (borderline paper)
+%
+%----------- Overall evaluation -----------
+%In this paper, the authors present a framework for autonomously exploring and mappping an unknown environment from a Quadrotor.
+The vehicle exploits an inverse sensor model for assigning a probability (that the cell has been explored) to each cell of the map.
+It also considers both the entropy information of the map and the collision probabilities when computing the next optimal exploration pose.
+%
+%The paper is overall undestandable, but  some technical sections: 2.a, 2.b, 3.b, 3.c are not easily readable. Providing some explanations to the formule would help the reader to the interpretation.
+%
+%Some other comments:
+%1) Passing from a full 3D map to an "augmented" 2D map is an interesting and smart trick for reducing the computational cost. In spite of this: "Typically,
+%computers small enough to be carried onboard flying robots
+%lack the memory or processing capabilities to update massive
+%maps quickly. Therefore, an onboard computer can simply
+%stream the sensor data, and a more powerful computer can
+%run the mapping."  The authors should provide some references or giving more details.
+The computational time depends on several variables, and this seems to be a simplistic motivation for explaining the presence of a ground station for the map management.
+%
+%2) The ROS message filtering tool just provides in the same callback the temporally closer measurements, but it does not solve the synchronization problem.
+The authors probably need some interpolation method or something similar.
+Please explain.
+%
+%3) "Since most quadrotor flight is close to hover and floor plans
+%exhibit little variation in the vertical direction, the proposed
+%exploration uses 2D occupancy grid maps located at the
+%desired exploration height." Same as (1). It depends on the context. Just by using this assumption, the proposed approach can work in some controlled scenarios with controlled quadrotor motion.
+%
+%4) The authors basically exploit the two map-based approach in simulation, while the composed map approach in the real environment.
+Probably testing both approaches in both scenarios would be more interesting.
+For instance: does the two-map based approach lead to the same slow convergence also in the real case?
+%
+%5) Main Issue: Following the problem formulation, the inverse sensor model exploits the uncertainties coming from the sensor measurements Z and from the current poses of the robot X.
+In Simulation, the authors just use the pose provided by the Gazebo Ground truth, while in the real environment the authors rely upon a Vicon, which provides the pose with very little uncertainty.
+In both cases, it seems that the authors do not introduce any noise or uncertainty in X.
+In this way, the experimental section cannot prove entirely the effectiveness of the proposed approach, and the reader cannot figure out if the approach can actually handle real environments.
+%
+%In my opinion, the paper deals with an interesting problem, but it also presents some questionable assumptions.
+Moreover, the experimental sections do not entirely prove the effectiveness of the proposed approach.
+
 \begin{abstract}
@@
 
-%% Prior to TLee edits:
-%Robotic mapping is the process of generating maps representing the environment surrounding a robot.
-%Mapping serves an important role in tasks such as search-and-rescue, surveillance, and robotic cleaning.
-%Several grid-based map representations have been studied, e.g.,~\cite{WolSuk05,MeyBeiBur12,TanThoWolBus14}, because they provide abundant occupancy knowledge about the environment and are computationally tractable.
-%
-%Most grid-based mapping representations are designed for 2D, while others are extended to capture a full 3D environment.
-In particular, occupancy probabilities have been assigned to a uniform 3D grid in~\cite{Andert09,PirRutBisSch11}.
-Another highly-popular variation of occupancy grid mapping is the Octomap representation~\cite{WurHorBenStaBur10,HorWurBenStaBur13}, which has added features such as extending the mapping limits and changing cell sizes.
-However, Octomaps, and the 3D uniform grid-based approaches mentioned above, produce cell occupancy probabilities that do not directly depend on the depth sensor stochastic properties.
-Instead, these approaches determine occupancy probabilities using an approximate function.
-Another method introduced in~\cite{Thr01,ThrBurFox05} and implemented in 3D with~\cite{SouMaiGon12} uses a simulated learning technique to find these probabilities, which may be difficult over the large dimensional space composed of maps, poses, and measurements.
-An additional approach such as~\cite{TanThoWolBus14} applies a particle filter to estimate the map.
-None of these approaches use the sensor properties to directly obtain a Bayesian solution to occupancy probability.
-%
-%%        1.      sensor fusion in mapping
-%The approximate nature of the occupancy probability motivates an exact Bayesian solution that can be applied in 3D.
-This paper proposes a computationally-efficient algorithm that follows the framework of~\cite{KauLeeAiMos16,KauTakAiLee17} to generate probabilities in a 3D environment.
-Since this approach only requires the origin and direction of sensor measurements to be known, the algorithm is designed to work for any number of sensors.
-Sensor fusion from several sources can provide important information about the environment with different levels of sensor certainty, fields of view, and vantage points.
-%
-%%        2.
-proposing a method to combine 3D mapping and 2D exploration: Also, we need to motivate why we are not doing full 3D exploration.
-Stress that we are proposing a computationally friendly approach to perform 3D mapping with projection using the fact there is less variation in the floor plan along the vertical direction.
-%The second topic of this paper deals with autonomous exploration, which provide the policies that govern robotic motion.
-A common approach is known as frontier-based exploration, proposed for 2D environments in~\cite{Yam97,Yam98} and extended to 3D with~\cite{ZhuDinLinWu15,SenWan16,KleDor13} using the Octomap representation, and similarly in~\cite{SawKriSri09} using a visibility metric.
-This approach moves robots toward the boarder between certain and uncertain space, though these systematic actions are not based on any consideration of the future probabilistic map uncertainty.
-Alternatively, there have been exploration techniques that are based on a quantitative measure of uncertainty known as Shannon's entropy~\cite{StaGriBur05}, an uncertainty metric based on grid cell occupancy probabilities.
-Shannon's entropy is used to acquire expected map information gain in~\cite{KauAiLee16,KauTakAiLee17}, and this paper focusses on extending these concepts to 3D.
-There are several challenges with obtaining expected information gain of 3D measurement scans, particularly as they relate to computational load.
-Considering the typically low variation in floor plans that robots explore during indoor explorations, this paper proposes two methods to generate 2D exploration maps that exploit the important characteristics of 3D maps.
-This process of combining 3D mapping and 2D exploration is completed in a computationally-efficient manner.
-%
-%%        3.      hardware/software for aerial exploration experiment.
-%Finally, this paper comments on the hardware and software necessary to implement these challenging algorithms for an aerial mapping and exploration experiment.
-This paper provides the basic software structure and features for mapping and generating exploration trajectories on a host computer while running depth sensors and a flight controller onboard a mobile robot.
-%
-%The paper is organized as follows.
-The problem is formulated in Section II.
-Special considerations for mapping and exploration in 3D are established in Section III.
-A numerical example is demonstrated in Section IV.
-Experimental results are shown in Section V, followed by conclusions and future work.
-
 \section{Problem Formulation}
@@
 and it is independent of the cell index $k$.
+The proofs of \refeqn{RayISMAnswer}--\refeqn{allEta} are given in~\cite{KauLeeAiMos16}.
 Since the terms of \refeqn{Unnormalized} are easily obtained and several are used repeatedly to obtain \refeqn{allEta}, the computational cost of \refeqn{RayISMAnswer} is trivial and can be applied in real-time.
@@
 There are several reasons why a robot might be equipped with multiple sensors, e.g., different fields-of-view or depth sensor properties.
-The proposed 3D mapping approach can handle measurements from different sources with different properties.
+For example, a robot may be equipped with a 2D LIDAR with highly accurate measurements, but a limited single-plane field-of-view.
+Then, the robot can compensate for the LIDAR limitations with an IR depth sensor that provides many more measurement rays with a 3D scan, where each ray has lower accuracy than the LIDAR rays.
+Since each sensor model is explicitly considered in the cell probability calculations, measurements from both sensors may serve to update the same probabilistic map.
+However, the impact from a LIDAR measurement with high accuracy on the probabilistic map will be greater than that of the IR depth sensor with lower accuracy.
+Since sensor properties are considered explicitly, the proposed 3D mapping approach can fuse the impact of measurements on the probabilistic map from different sources with different properties.
+% EditEK 1: example of 2 sensor measurements fused, starting with "For example, a robot may be equipped..." until the end of the paragraph.
 
@@
 The computational order of ray casting and the inverse sensor model are proportional to the number of grid cells along the ray $n_r$, so careful consideration must be placed on sensor limits and rays per scan that can be considered.
-Typically, computers small enough to be carried onboard flying robots lack the memory or processing capabilities to update massive maps quickly.
+Typically, computers small enough to be carried onboard flying robots lack the processing capabilities to update massive maps quickly, particularly because the ray-by-ray updating method requires many computations in series with large measurement scans.
 Therefore, an onboard computer can simply stream the sensor data, and a more powerful computer can run the mapping.
+% EditEK 2: revised the sentence from:
+% "Typically, computers small enough to be carried onboard flying robots lack the memory or processing capabilities to update massive maps quickly."
+% to:
+% Typically, computers small enough to be carried onboard flying robots lack the processing capabilities to update massive maps quickly, particularly because the ray-by-ray updating method requires many computations in series with large measurement scans.
 
@@
 Therefore, pose estimates and sensor readings must be paired according to their time stamps.
-A useful tool that easily satisfies this goal is message filtering with the Robot Operating System (ROS), where stamped poses and measurement scans are synchronized using recently received data.
+A useful tool that easily satisfies this goal is the Message Filters package with the Robot Operating System (ROS), where stamped poses and measurement scans are synchronized using recently received data.
+% EditEK 3: changed "message filtering" to "the Message Filters package"
 
@@
 Since most quadrotor flight is close to hover and floor plans exhibit little variation in the vertical direction, the proposed exploration uses 2D occupancy grid maps located at the desired exploration height.
+These assumptions correspond to rooms and hallways with non-complex geometries.
 This approach decreases $n_r$ and simplifies the exploration problem with 3D space.
 These 2D maps project important properties of the 3D map.
+% Edit EK 4: Added the sentence: "These assumptions correspond to rooms and hallways with non-complex geometries."
 
@@
 
-Gazebo serves as the simulator node, with plugins to simulate an Asus Xtion color depth sensor and a Hokuyo LIDAR.
+Gazebo is a 3D dynamic simulator~\cite{KaeHow04}, with plugins to simulate an Asus Xtion color depth sensor and a Hokuyo LIDAR.
 Gazebo also provides position and attitude transforms from the world to the quadrotor body, and the ability to set robot model states directly.
@@
 The first process is projecting the 3D map onto 2D maps; the same code under different conditional parameters produces two nodes, for collision and entropy maps.
-The second process subscribes to these two 2D maps; the collision map serves as input for candidate pose consideration and Dijstra's search, and the entropy map is used to predict expected information gain.
-This node provides path messages, composed of desired poses $0.01$sec apart, that are easily visualized with the ROS visualization package, Rviz.
+The second process subscribes to these two 2D maps; the collision map serves as input for candidate pose consideration and Dijkstra's search, and the entropy map serves to predict information gain.
+This node provides path messages, composed of desired poses $0.01$sec apart, displayed with a ROS visualization package, Rviz.
 
@@
 	\begin{subfigure}[t]{0.24\columnwidth}
+	\vspace*{0.15\columnwidth}
            	\centering
@@
     	\begin{subfigure}[t]{0.24\columnwidth}
+	\vspace*{0.15\columnwidth}
            	\centering
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{entropy_gazebo_aspect3by1.png}
+	\includegraphics[width=0.45\textwidth]{entropy_gazebo_aspect3by1.png}
 	\caption{Simulated Environment Map Entropy}
@@
 
-\section{Experimental Resutls}
+\section{Experimental Results}
 
@@
 \centering
-    	\begin{subfigure}[t]{0.44\columnwidth}
+    	\begin{subfigure}[t]{0.95\columnwidth}
            	\centering
@@
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.44\columnwidth}
+    	\begin{subfigure}[t]{0.95\columnwidth}
+	\vspace*{0.03\columnwidth}
            	\centering
@@
 \centering
-    	\begin{subfigure}[t]{0.88\columnwidth}
+    	\begin{subfigure}[t]{0.95\columnwidth}
            	\centering
@@
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.88\columnwidth}
+    	\begin{subfigure}[t]{0.95\columnwidth}
+		\vspace*{0.03\columnwidth}
            	\centering
@@
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.88\columnwidth}
+    	\begin{subfigure}[t]{0.95\columnwidth}
+	\vspace*{0.03\columnwidth}
            	\centering
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{entropy_experiment_aspect3by1.png}
+	\includegraphics[width=0.45\textwidth]{entropy_experiment_aspect3by1.png}
 	\caption{Experimental Environment Map Entropy}

### 2018-06-09 Conference/IRC 2018/Presentation/IRC.1.tex
added Conference folder

\documentclass[11pt,professionalfonts,hyperref={pdftex,pdfpagemode=none,pdfstartview=FitH}]{beamer}
%\usepackage{times}
%\usefonttheme{serif}
%\usepackage{helvet}
%\usepackage{amsmath,amssymb}
\usepackage{graphicx,multirow}
%\usepackage[scriptsize]{subfigure}
\usepackage{subcaption}
%\usepackage{media9}
\usepackage{movie15}
\usepackage{hyperref}

%\usepackage{warmread}
%\usepackage[all,import]{xy}

%\renewcommand\mathfamilydefault{\rmdefault}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand{\refprop}[1]{Proposition \ref{prop:#1}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

\definecolor{mygray}{gray}{0.9}

\mode<presentation> {
  \usetheme{Warsaw}
  \usefonttheme{serif}
  \setbeamercovered{transparent}
}

\newcommand{\mypaper}{}

\setbeamertemplate{footline}%{split theme}
{%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=.85\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.15\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm]{title in head/foot}
    \usebeamerfont{title in head/foot}\mypaper\hfill \insertframenumber/\inserttotalframenumber
%    \usebeamerfont{title in head/foot}\hfill \insertframenumber/\inserttotalframenumber
  \end{beamercolorbox}}%
  \vskip0pt%
} \setbeamercolor{box}{fg=black,bg=yellow}

\title[Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities]{\large Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities}

\author{\vspace*{-0.3cm}}

\institute{\footnotesize
{\normalsize Evan Kaufman, Kuya Takami,\\Zhuming Ai, and Taeyoung Lee}\vspace*{0.2cm}\\
  Mechanical and Aerospace Engineering\\The George Washington University\\Washington, DC, USA
}
%{\normalsize Evan Kaufman, Kuya Takami \\Zhuming Ai, and Taeyoung Lee}\vspace*{0.2cm}\\
%  Mechanical and Aerospace Engineering\\ George Washington University}

\date{}

\definecolor{tmp}{rgb}{0.804,0.941,1.0}
\setbeamercolor{numerical}{fg=black,bg=tmp}
\setbeamercolor{exact}{fg=black,bg=red}

\newtheorem{prop}{Proposition}

\renewcommand{\emph}[1]{\textit{\textbf{\color{blue}{#1}}}}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section*{}
\subsection*{Introduction}

\begin{frame}
\frametitle{Motivation}
\begin{itemize}
    \item US Naval Research Lab's Project ``Intelligent Microflyer''
    \begin{itemize}
    	\item Goal: send their vehicles into uncertain spaces and autonomously explore to generate a map
	\item The vehicle must fly and hover
    	\item The controlled trajectory should be collision-free
    \end{itemize}
\vspace*{0.0cm}\pause
\item General Benefits
\begin{itemize}
	\item Exploring dangerous zones with physical or health hazards
	\item Acquiring intel on enemy territories
\end{itemize}
\vspace*{0.0cm}\pause
\item Mapping and Exploration in 3D
\begin{itemize}
	\item Generate a clear map of 3D space
	\item Explore at a fixed altitude with collision-free robot paths based on the 3D map
	\item Control the vehicle to track these trajectories
	\item All processes done in real-time
\end{itemize}
\end{itemize}

%\only<1->{
%\begin{figure}
%\centerline{
%    \includegraphics[height=2.1cm]{ogm_ex3.jpeg}\hspace*{0.1cm}
%\hspace*{0.5cm}
%    \includegraphics[height=2.1cm]{ogm_ex1.jpg}\hspace*{0.1cm}
%\hspace*{0.5cm}
%    \includegraphics[height=2.1cm]{ogm_ex2.png}\hspace*{0.1cm}
%}
%\end{figure}}

\end{frame}

\begin{frame}
\frametitle{Approach}
\begin{itemize}
    \item Occupancy Grid Mapping
    \begin{itemize}
    	\item Simple and well-known grid-based representation
	\item Decomposes space into evenly-spaced cells that are either \emph{occupied} or \emph{free}
    	\item Probabilistic: goal is to find the \emph{probability} that the cells are occupied
	\item Useful for collision-avoidance and graph searching
    \end{itemize}
\vspace*{0.0cm}\pause
\item Entropy-Based Exploration
\begin{itemize}
	\item \emph{Shannon's entropy}: \begin{align*}H(P)=-P\log{P}-(1-P)\log{(1-P)}\end{align*}
	\item Maximized with $P=0.5$, decreases as $P$ approaches $0$ or $1$
	\item Main idea: choose robotic motions to decrease the entropy of grid cells within an occupancy grid map
	\item \emph{Accurate} occupancy grid probabilities are highly important to accurate entropies
\end{itemize}
\end{itemize}

\end{frame}

\section*{}
\subsection*{Occupancy Grid Mapping}

\begin{frame}
\frametitle{Occupancy Grid Mapping}
\begin{itemize}
	\item Forward Sensor Model
	\vspace*{1cm}
	\item Inverse Sensor Model
	\vspace*{1cm}
	\item Efficient Grouping Solution
	\vspace*{1cm}
	\item 3D Ray Casting
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Problem Definition}
%\framesubtitle{Problem Definition}

\begin{itemize}
	\item The Map and the Robot
	\begin{itemize}
	\item Map $m$ is composed of $n_m$ grid cells with known location and size
	\item The $i$-th grid cell $\mathbf{m}_i$ is a \emph{static binary} random variable, independent of other grid cells: $P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i)$
	\item \emph{Pose} $X_t$ is known, containing robot \emph{position} and \emph{attitude}
	\end{itemize}
\vspace*{0.0cm}\pause
\end{itemize}
\begin{minipage}[t]{7.0cm}
\begin{itemize}
	\item Depth Measurements
	\begin{itemize}
	\item Each measurement origin and direction is known \emph{deterministically}
	\item A measurement \emph{scan} $Z_t=\braces{z_{t,1},z_{t,2},\ldots,z_{t,n_z}}$ contains $n_z$ measurement \emph{rays} (depths)% at the $t$-th time step, and the history of measurement scans $Z_{1:t}$ is known
\item The \emph{forward sensor model} is known from the sensor properties
\end{itemize}
\end{itemize}
\end{minipage}
\begin{minipage}[t]{3.0cm}
%The \emph{forward sensor model} is the probability density distribution $p(z_{t,l}|m,X_{t})$  known from the sensor properties
\hspace*{0.25cm}
\begin{figure}[!htbp]
\vspace*{-0.25cm}
%\vspace*{0.25cm}
\centerline{
    \includegraphics[width=3.5cm]{BeamModel.png}\hspace*{0.1cm}
    }
\vspace*{0.25cm}
\centerline{
    \includegraphics[width=2.5cm]{1D_True_Grid.png}\hspace*{0.1cm}
%    \hspace*{0.75cm}
    }
{Beam Model for Range Finders}
\end{figure}

\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Inverse Sensor Model}

\begin{itemize}
	\item Inverse Sensor Model
	\begin{align*}
&P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\eta_{t,l}\sum_{m\in\mathcal{M}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
\end{align*}
	\begin{itemize}
	\item Given $n$ grid cells: $\mathcal O(2^n)$ is \emph{computationally intractable}, motivating a different solution
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Inexact Solutions}

\begin{minipage}[t]{7.0cm}
\begin{itemize}
	\item Approximate a function for the inverse sensor model based on intuition
	\begin{itemize}
		\item Simple to implement, but mathematically inaccurate
	\end{itemize}
\end{itemize}
\end{minipage}
\begin{minipage}[t]{3.0cm}
\begin{figure}[!htbp]
\centerline{
    \vspace*{-0.5cm}
%	\hspace*{0.25cm}
    \includegraphics[width=4.0cm]{Approx_ISM_shortened.png}\hspace*{0.1cm}
    }
\centerline{
    \includegraphics[width=3.5cm]{1D_True_Grid.png}
%    \hspace*{0.75cm}
    }
		\end{figure}
\end{minipage}
\vspace*{-0.5cm}
\vspace*{0.0cm}\pause
\begin{itemize}
	\item Find a solution through learning
	\begin{itemize}
		\item Simulate maps, poses, and measurements and use learning to obtain an inverse sensor model
		\item Complicated, unclear how these parameters are chosen
	\end{itemize}
	\vspace*{0.0cm}\pause
	\item Goal: design a \emph{simple} and \emph{accurate} occupancy grid mapping method avoiding log-odds ratio assumptions
\end{itemize}
%
%	\begin{minipage}[width=1cm]
%		Approximate a function for the inverse sensor model based on intuition
%	\end{minipage}
%	\begin{minipage}[width=1cm]
%		Approximate a function for the inverse sensor model based on intuition2
%	\end{minipage}
%\begin{itemize}
%	\item Heuristic Approach
%	\begin{itemize}
%	\item asdf
%\begin{figure}
%\centerline{
%    \includegraphics[width=3.5cm]{ISM_Approx_Single_Ray.png}\hspace*{0.1cm}
%    }
%\centerline{
%    \includegraphics[width=3.5cm]{1D_True_Grid.png}\hspace*{0.1cm}
%%    \hspace*{0.75cm}
%    }
%\end{figure}
%	\item Simple to implement, but mathematically inaccurate
%	\end{itemize}
%\vspace*{0.3cm}\pause
%\begin{itemize}
%	\item Learning
%	\begin{itemize}
%	\item Idea
%	\item Complicated, not clear how parameters are chosen
%	\end{itemize}
%\vspace*{0.3cm}\pause
%	\item Goal: design a simple and accurate method for occupancy grid mapping that avoids log-odds ratio assumptions
%\end{itemize}

\end{frame}

\section*{}
\subsection*{Probabilistic Mapping}

\begin{frame}
\frametitle{New Approach: Grouping}

\begin{itemize}
    \item Main idea: make use of occupancy grid mapping \emph{assumptions} and extract \emph{patterns} from probabilistic properties to find a \emph{computationally-efficient} solution
	\begin{itemize}
		\item Since the origin and direction of each measurement ray is known deterministically, the set of grid cells that the ray intersects is \emph{known through geometry}
		\item A depth reading follows the forward sensor model, which \emph{only depends} on the first occupied grid cell along the measurement ray
	\end{itemize}
\end{itemize}
\setcounter{subfigure}{0}
% uncomment
\begin{figure}
  \centering
  \begin{subfigure}[t]{.45\linewidth}
    \centering\includegraphics[width=\linewidth]{rkplus_1.png}
  \end{subfigure}
  \begin{subfigure}[t]{.45\linewidth}
    \centering\includegraphics[width=\linewidth]{rkplus_2.png}
  \end{subfigure}
    \begin{subfigure}[t]{.45\linewidth}
    \centering\includegraphics[width=\linewidth]{rkplus_3.png}
  \end{subfigure}
  \begin{subfigure}[t]{.45\linewidth}
    \centering\includegraphics[width=\linewidth]{rkplus_4.png}
  \end{subfigure}
\end{figure}
%\begin{figure}[!htbp]
%\vspace*{-0.25cm}
%\centerline{
%	\subfigure{
%	\includegraphics[width=4.0cm]{rkplus_1.png}\hspace*{-0.5cm}}
%		\subfigure{
%	\includegraphics[width=4.0cm]{rkplus_2.png}}
%}
%\centerline{
%	\subfigure{
%	\includegraphics[width=4.0cm]{rkplus_3.png}\hspace*{-0.5cm}}
%		\subfigure{
%	\includegraphics[width=4.0cm]{rkplus_4.png}}
%}
%\end{figure}

%% GENERAL FIGURE
%\begin{figure}
%  \centering
%  \begin{subfigure}[t]{.3\linewidth}
%    \centering\includegraphics[width=.5\linewidth]{rkplus_1.png}
%    \caption{This is a sub-caption. This is a sub-caption. This is a sub-caption}
%  \end{subfigure}
%  \begin{subfigure}[t]{.3\linewidth}
%    \centering\includegraphics[width=.5\linewidth]{rkplus_2.png}
%    \caption*{This is a sub-caption.}
%  \end{subfigure}
%\end{figure}

\end{frame}

\begin{frame}
\frametitle{Bayesian Update}

\begin{itemize}
	\item Unnormalized Probability:
	\begin{align*}
\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})
&=\mathbf{P}_k^-
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]\nonumber\\
&\qquad + \bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{k+},X_t)\mathbf{P}_k^-
\end{align*}
	\item Normalizer:
	\begin{align*}
\label{eqn:allEta}
\eta
&=
\bigg[\sum_{i=1}^{n_{r}+1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\} p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]^{-1}
\end{align*}
	\item Solution: $P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})=\eta\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})$
	\item \emph{Linear Complexity} ($\mathcal{O}(n)$ for \emph{all} $n$ cells in FOV)
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Complete Update}

\begin{itemize}
	\item 2D and 3D Ray Casting: use geometry to determine which cells a ray intersects and at what depths
	\item Update the grid cells ray-by-ray
\end{itemize}

%% GENERAL FIGURE
\begin{figure}
  \centering
  \begin{subfigure}[t]{.3\linewidth}
    \centering\includegraphics[height=.5\linewidth]{RayCastIllustration.png}
    \caption*{2D Ray Casting}
  \end{subfigure}
  \begin{subfigure}[t]{.3\linewidth}
    \centering\includegraphics[height=.5\linewidth]{test_setup_bottom_left_cropped.jpg}
    \caption*{2D Environment}
  \end{subfigure}
  \begin{subfigure}[t]{.3\linewidth}
    \centering\includegraphics[height=.5\linewidth]{t_end.jpg}
    \caption*{2D Map}
  \end{subfigure}
  \end{figure}
  \begin{figure}
  \centering
%  \vspace*{0.3\linewidth}
  \begin{subfigure}[t]{.3\linewidth}
    \centering\includegraphics[height=.5\linewidth]{experiment_south.jpg}
    \caption*{3D Environment}
  \end{subfigure}
  \hspace*{0.1\linewidth}
  \begin{subfigure}[t]{.3\linewidth}
    \centering\includegraphics[height=.5\linewidth]{experiment_ogm3D_2min47sec.jpg}
    \caption*{3D Map}
  \end{subfigure}
  %
%  \begin{subfigure}[t]{.3\linewidth}
%    \centering\includegraphics[width=.5\linewidth]{rkplus_2.png}
%    \caption*{This is a sub-caption.}
%  \end{subfigure}
\end{figure}

% RayCastIllustration.png 
% uncomment
%\begin{figure}[!htbp]
%\vspace*{-0.25cm}
%\centerline{
%	\subfigure[Ray Casting]{
%	\includegraphics[width=4.0cm]{rkplus_1.png}\hspace*{-0.5cm}}
%		\subfigure{
%	\includegraphics[width=4.0cm]{rkplus_2.png}}
%}
%\end{figure}
\end{frame}

\section*{}
\subsection*{Entropy-Based Exploration}

\begin{frame}
\frametitle{Entropy-Based Exploration}
\begin{itemize}
	\item Existing Approaches
	\vspace*{1cm}
	\item Map Entropy Change
	\vspace*{1cm}
	\item Optimal Pose Selection
	\vspace*{1cm}
	\item Collision-Free Trajectory
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Existing Exploration Approaches}

\begin{itemize}
	\item Frontier-Based Exploration
	\begin{itemize}
		\item A robot identifies boundaries of free and uncertain cells
		\item The robot moves toward these frontiers, thereby pushing back the boundaries
		\item Heuristic, suboptimal, ad hoc (particularly in 3D)
	\end{itemize}
	\vspace*{0.0cm}\pause
	\item Entropy-Based Approaches
	\begin{itemize}
		\item Probabilities from the occupancy grid map are approximate
		\item Usage of ``hallucination measurements'': assume that $\text{E}[H(P(\mathbf{m}_i|z))]\approx H(P(\mathbf{m}_i|\text{E}[z]))$
	\end{itemize}
	\vspace*{0.0cm}\pause
	\item Proposed Approach
	\begin{itemize}
		\item Use occupancy grid map probabilities
		\item Solve $\text{E}[H(P)]$ and select future poses to maximize map information
		\item Extract important 3D map information into a simplified 2D projected map
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Expected Information Gain}

\begin{itemize}
	\item Main Idea: determine the benefit of future poses based on how their measurements are expected to improve the map
	\vspace*{0.0cm}\pause
	\item Goal: decrease total map entropy:
	\begin{align*}
		H(P(m))&=\sum_{i=1}^{n_m}H(P(\mathbf{m}_i))
	\end{align*}
	\vspace*{0.0cm}\pause
	\item Measurement Ray Expected Information Gain:
	\begin{align*}
		\text{E}[H(P(m|x_c,z_{c}))]&=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(m|x_c,z_{c,k}))P(z_{c,k}|x_c)\bigg\}
		\\
		P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}}
		\\
		\mathcal I(x_c,z_{c})&=H(P(m))-\text{E}[H(P(m|x_c,z_{c}))]
	\end{align*}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Expected Information Gain Computation}

\begin{itemize}
	\item \emph{Squared Complexity}: $\mathcal O(n_r^2)$ for $n_r$ cells along the ray
	\vspace*{0.0cm}\pause
	\item Two Approaches:
	\begin{itemize}
		\vspace*{0.0cm}\pause
		\item Decrease the cells considered by selecting $n_p$ most probable cells: $\mathcal O(n_r^2)$ reduces to $\mathcal O(n_p^2)$
		\vspace*{0.0cm}\pause
		\item Use 3D cells in a 2D \emph{projection}
\begin{figure}
  \centering
  \begin{subfigure}[t]{.3\linewidth}
    \centering\includegraphics[height=.75\linewidth]{experiment_ogm3D_2min47sec.jpg}
    \caption*{\ \ \ \ \ \ \ \ \ \ \ \ \ \ 3D Map}
  \end{subfigure}
  \hspace*{0.25\linewidth}
  \begin{subfigure}[t]{.3\linewidth}
    \centering\includegraphics[height=.75\linewidth]{experiment_2min47sec.jpg}
    \caption*{2D Projection}
  \end{subfigure}
\end{figure}
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Map Projection}

\begin{itemize}
	\item Main Idea: choose a cell probability policy incorporating entropy and collision avoidance
		\vspace*{0.0cm}\pause
	\item Consider the 3D cells of $m_C\subset m$ composing the 2D projection:
	\begin{itemize}
		\item If any cell probability $P(\mathbf{m}_i)>P_\text{thresh}$ where $P_\text{init}<P_\text{thresh}<1$, then the projected cell probability is $\text{max}(P(\mathbf{m}_i))$
		\item Otherwise, the projected cell probability is $\text{min}(P(\mathbf{m}_i))$
	\end{itemize}
		\vspace*{0.0cm}\pause
	\item This approach is fairly-conservative for collision avoidance
	\item Relies on the first 3D probabilities that are changed, indicating the true occupancy of the space, even with sparse measurements
	\item Contains information above/below the projected height
	\item Simple and fast: avoids large computations with squared complexities
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Pose Selection}

\begin{itemize}
	\item Main Idea: select the future pose that maximizes map information gain while accounting for travel distance
	\vspace*{0.0cm}\pause
	\item Attitude Selection: choose the attitude that covers the candidate measurement rays with largest objective function summation
\begin{figure}
\centerline{
	\includegraphics[height=0.35\linewidth]{ExampleOptimalPose.png}
%\begin{picture}(0,0)(0,0)
%\setlength{\unitlength}{0.1\linewidth}\scriptsize
%\put(4.4,2.1){\color{blue}$X_t$}
%\put(5.9,1.9){\color{red}$x_1$}
%\put(2.9,2.0){\color{red}$x_5$}
%\put(4.4,3.6){\color{red}$x_3$}
%\put(4.4,0.6){\color{red}$x_7$}
%\put(5.4,3.0){\color{red}$x_2$}
%\put(3.3,3.1){\color{red}$x_4$}
%\put(3.3,1.0){\color{red}$x_6$}
%\put(5.4,1.0){\color{red}$x_8$}
%\put(6.8,3.1){\color{red}$z_{2,1}$}
%\put(4.5,3.1){\color{red}$z_{2,5}$}
%\put(5.6,4.1){\color{red}$z_{2,3}$}
%\put(5.6,2.2){\color{red}$z_{2,7}$}
%\put(6.5,3.7){\color{red}$z_{2,2}$}
%\put(4.9,3.8){\color{red}$z_{2,4}$}
%\put(5.0,2.5){\color{red}$z_{2,6}$}
%\put(6.5,2.5){\color{red}$z_{2,8}$}
%\put(6.1,2.9){$X_c^*$}
%\end{picture}
}
\vspace*{-0.05\linewidth}
\end{figure}
	\item The attitude $R_c$ is selected for all candidates: the expected information gain becomes $\mathcal I(X_c)$ where $X_c=\braces{x_c,R_c}$
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Pose Selection}

\begin{itemize}
	\item Travel Distance: generate a cost map from the occupancy grid to obtain distances to each candidate pose $d(x_c)$
	\vspace*{0.0cm}\pause
	\item Bump Function: use a bump function \begin{align*}f(d)=\exp\braces{-\frac{d^2}{2\sigma^2}}+\gamma, \quad \sigma>0, \quad \gamma>0\end{align*} to incentivize the robot to explore more locally until nearby spaces are well-known
	\begin{figure}
	\vspace*{-0.010\linewidth}
\centering
	\includegraphics[width=0.8\linewidth]{GaussianBumpFunFlat.png}
		\vspace*{-0.015\linewidth}
\end{figure}
	% C++: exp(-0.5*distAway*distAway/sigSqrd)+minFunVal GaussianBumpFunFlat.png
	\vspace*{0.0cm}\pause
	\item Select the pose that maximizes the product \begin{align*}X^*=\argmax_{X_c}{\mathcal I(X_c)f(d(x_c))}\end{align*}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Collision-Free Path}

\begin{itemize}
	\item Using the cost map from before, \emph{Dijkstra's algorithm} is completed by finding the waypoints from the desired candidate back to the current location
	\vspace*{0.0cm}\pause
	\item These waypoints serve as input to a \emph{constrained polynomial least squares trajectory}
	\begin{itemize}
		\item Each trajectory is solved independently as a function of time, i.e. $x(t)$, $y(t)$
		\item Starting and terminal locations are fixed
		\item Segments patched together have the same position and velocity at the connection
	\end{itemize}
	\vspace*{0.0cm}\pause
	\item A \emph{geometric controller} tracks the trajectory without \emph{singularities} or \emph{ambiguities}
	\vspace*{0.0cm}\pause
	\item The mapping, exploration, and control run in \emph{real-time}
\end{itemize}

\end{frame}

\section*{}
\subsection*{Results}

\begin{frame}
\frametitle{Experimental Setup}
\begin{itemize}
        	\item Hardware
	\begin{itemize}
		\item NVIDIA Jetson Ubuntu computer-on-module
		\item Asus Xtion IR color depth sensor
		\item Hokuyo LIDAR
		\item VectorNav IMU
	\end{itemize}
	\item Software
	\begin{itemize}
		\item Robot Operating System (ROS)
		\item Vicon Tracker (mocap\_vicon)
	\end{itemize}
\end{itemize}
\begin{figure}
  \centering
  \begin{subfigure}[t]{.4\linewidth}
    \centering\includegraphics[height=.5\linewidth]{quad_top.png}
    \caption*{Quadrotor Top}
  \end{subfigure}
  \begin{subfigure}[t]{.4\linewidth}
    \centering\includegraphics[height=.5\linewidth]{quad_bottom.png}
    \caption*{Quadrotor Bottom}
  \end{subfigure}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Experimental Video}
%\begin{itemize}
%        	\item \href{https://www.youtube.com/watch?v=I_1rXV2XRqk&feature=share}{Video Link}
%\end{itemize}

\begin{figure}[ht]
\includemovie[poster]{0.8\linewidth}{0.45\linewidth}{../../../../Movies/nrl_experiment_narrated.mp4}
\end{figure}

%\only<1->{
%\begin{figure}
%\centering
%\includemovie[poster]{5cm}{6cm}{Farhad_AG.mp4}
%\end{figure}}

%\centering
%\includemedia[
%  width=0.8\linewidth,
%  totalheight=0.45\linewidth,
%  activate=pageopen
%]{\fbox{}}{https://www.youtube.com/watch?v=I_1rXV2XRqk&feature=share}

\end{frame}

\section*{}
\subsection*{Conclusions and Future Work}

\begin{frame}
\frametitle{Conclusions}
\begin{itemize}
        	\item Proposed and tested an occupancy grid mapping technique that uses the \emph{exact probabilistic solution}
	\vspace*{0.0cm}\pause
	\item Computational cost is reduced \emph{substantially} for \emph{real-time implementation} using probabilisitic properties and exploiting mathematical patterns
	\vspace*{0.0cm}\pause
	\item The occupancy grid mapping results are extended to autonomous exploration via expected map entropy changes
	\vspace*{0.0cm}\pause
	\item Dijkstra's search and a constrained polynomial least squares trajectory serves as input for a geometric controller
	\vspace*{0.0cm}\pause
	\item Current and Future Work:
	\begin{itemize}
		\item Multi-vehicle problem: auction-based cooperative approach with receding horizons
		\item Expected entropy of rays through the 3D grid cells
	\end{itemize}
\end{itemize}
\end{frame}

\end{document}

### 2018-06-09 Conference/SCITECH19/SCITECH19.tex
added Conference folder

\documentclass[conf]{new-aiaa}
%\documentclass[journal]{new-aiaa} for journal papers
\usepackage[utf8]{inputenc}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage{longtable,tabularx}
\setlength\LTleft{0pt}

\usepackage{amssymb,url,times}%,subfigure}% amsthm is the one!
\usepackage{caption,subcaption,hyperref}
\usepackage{color,comment}

\newtheorem{prop}{Proposition}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\abs}[1]{\ensuremath{\left| #1 \right|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\ip}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Figure \ref{fig:#1}}
\newcommand{\tr}[1]{\mbox{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mbox{tr}\ensuremath{\!\bracket{#1}}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathbf{d}}}
\newcommand{\pair}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\!\langle #1 \rangle\!\rangle}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\ad}{\ensuremath{\mathrm{ad}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\graphicspath{{../../Fig/}}

\title{Autonomous Aerial for Topological Mapping of Mars Environments}

\author{Evan Kaufman\footnote{Graduate Research Assistant} and Taeyoung Lee\footnote{Associate Professor, AIAA Member Grade (if any)}}
\affil{Department of Mechanical and Aerospace Engineering, The George Washington University, Washington, DC, 20052}
% TODO: funding source

\begin{document}

\maketitle

\begin{abstract}
This paper proposes a 3D autonomous exploration approach designed for a flying vehicle while mapping the complex environment of the surface of Mars. Shannon's entropy serves as a measure of map uncertainty, where the robot goal is to minimize entropy, thereby maximizing map information. We propose a novel technique to predict future map entropy from depth sensors capable of scanning 3D spaces, and formulate autonomous exploration as an optimization problem based on map information gain and collision-free travel costs. The proposed approach is demonstrated with numerical simulations using contour data from the surface of Mars.
\end{abstract}

%\section{Nomenclature}
%
%{\renewcommand\arraystretch{1.0}
%\noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
%$A$  & amplitude of oscillation \\
%$a$ &    cylinder diameter \\
%$C_p$& pressure coefficient \\
%$Cx$ & force coefficient in the \textit{x} direction \\
%$Cy$ & force coefficient in the \textit{y} direction \\
%c   & chord \\
%d$t$ & time step \\
%$Fx$ & $X$ component of the resultant pressure force acting on the vehicle \\
%$Fy$ & $Y$ component of the resultant pressure force acting on the vehicle \\
%$f, g$   & generic functions \\
%$h$  & height \\
%$i$  & time index during navigation \\
%$j$  & waypoint index \\
%$K$  & trailing-edge (TE) nondimensional angular deflection rate
%\end{longtable*}}

\section{Introduction}

The topography of Mars has several benefits in future mission planning for landing spacecraft and  image distortion correction among other research objectives. The Mars Orbiting Laser Altimeter (MOLA) mission greatly improved the topography map of the surface, but was limited to an uncertainty of roughly eight meters~\cite{MOLA99}. The only precise topography knowledge about the surface comes from four rovers, two of which are inactive, and have covered only a small portion of the surface area of Mars.

Mapping and exploring Mars presents several challenges. Existing rovers are ground vehicles, incapable of traversing steep or dangerous terrain. Flying vehicles avoid this restriction, but special attention must be placed on their motion planning in this uncertain environment. Humans cannot directly give commands to the vehicles due to the signal time delay between Mars and Earth. Even if a humans are present on Mars, they can only provide sub-optimal motion commands on a small scale. These challenges motivate an autonomous approach to 3D mapping and exploration in uncertain and complex environments for a flying vehicle near the surface of Mars.

The study of 3D mapping and autonomous exploration has received substantial attention for applications on Earth. Occupancy grid mapping in 3D is a popular approach where an environment is decomposed into evenly-spaced cubic grid cells, where each cell is assumed as occupied or free~\cite{ThrBurFox05,WurHorBenStaBur10}. The goal is finding the probability of each cell being occupied. An exact solution of this probability was recently found by the authors using the stochastic properties of onboard sensors directly~\cite{KauLeeAiMos16,KauTakAiLee17}. Using an exact occupancy grid mapping algorithm in 3D, the topological information of the surface of Mars can be accurately computed in real-time.

Next, the robot must determine motions based on how much mapping knowledge it expects to gain, known as autonomous exploration. The most popular approach to this problem is known as frontier-based exploration~\cite{ZhuDinLinWu15,SenWan16,KleDor13}. The main idea is that the robot moves through previously-mapped free space toward the boundary between free space and uncertain space, known as a frontier. The robot takes measurements of this frontier, thereby pushing back the boundary. This process is repeated until the map is well-known. Despite the intuitive nature of this approach, there is no measure or prediction of how these robotic actions improve map information. Thus, frontier-based approaches rely on a heuristic rule, and are inherently suboptimal.

In contrast, entropy-based autonomous exploration approaches seek to minimize future map uncertainty. Shannon's entropy is a measure of map uncertainty; since minimizing future map uncertainty is equivalent to maximizing map information gain, these approaches select robotic actions to minimize entropy~\cite{StaGriBur05}. In~\cite{KauAiLee16}, an exact solution to expected entropy was developed, and in~\cite{KauTakAiLee18}, this concept is extended to 3D office-like environments assuming the environment can be approximated as vertically-uniform. Similarly, \cite{MauDakPet14} simplifies tasks by decomposing the 3D space into rooms. While these approaches can be effective for buildings, they are incompatible with the surface of Mars, which has complicated geological features that require careful consideration.

In this paper, we propose a 3D autonomous exploration algorithm that considers the 3D probabilistic occupancy grid map and sensor properties explicitly. The complicated and uncertain geometry of the surface of Mars is captured by aerial mapping because this approach can cover a large area effectively. An onboard depth sensor, such as a 3D laser scanner, gathers range measurements to determine the occupancy probabilities of 3D grid cells. Then, collision-free future pose candidates are considered. The predicted negative entropy changes from measurements passing through 3D space are maximized while also considering travel costs. The robot then uses Dijkstra's search in 3D for collision-free motion planning, and the complete process follows a receding-horizon framework for consistently using updated information. The proposed optimal approach is designed specifically for real-time exploration of the surface of Mars, where the flying robot mostly maps occupied space below the vehicle, but is also capable of looking ahead to capture steep and quickly-changing terrain. In short, the proposed autonomous exploration approach predicts future map uncertainty in 3D space to determine optimal actions that maximize map information gain while avoiding collisions.

The paper is organized as follows. The mapping probability and expected entropy calculation are defined in Section \ref{sec:ProbDef}. The proposed 3D exploration for information gain maximization is covered in Section \ref{sec:Explore3D}. Numerical simulations are shown in Section \ref{sec:NumExamples}, followed by conclusions.

\section{Problem Definition}
\label{sec:ProbDef}

In this section, we define the 3D occupancy grid map probability and expected map entropy from a future measurement.

\subsection{Probabilistic Occupancy Grid Mapping in 3D}

First we define the map, measurements, and poses as follows. Let 3D map $m$ be composed of $n_m$ evenly-spaced cubic grid cells with edge length $\alpha$. For index $i\in\braces{1,2,\ldots,n_m}$, let $\mathbf{m}_i$ and $\bar{\mathbf{m}}_i$ be the events that the $i$-th grid cell are occupied and free, respectively, where each cell occupancy is assumed static, binary, and independent of other cells. At the $t$-th time step, we assume a known current pose $X_t=\braces{x_t,R_t}$ such that $x_t\in\Re^3$ is a 3D location and $R_t\in\SO$ is the 3D attitude, and history of poses $X_{1:t-1}$ is assumed known as well. Considering that most modern sensors return scans with multiple measurement rays, let measurement ray $z\in\Sph^2$ belong the the most recent scan $Z_t$, and let $Z_{1:t-1}$ be the history of measurement scans.

Next, we present recent contributions to occupancy probability determination from~\cite{KauLeeAiMos16,KauTakAiLee17}, which uses a simplified result of a complicated Bayesian probability. Let a temporary reduced map $r\subset m$ be composed of $n_r$ cells along measurement $z$, indexed by increasing distance from the current pose $X_t$. The cells of $r$ are easily acquired through 3D ray casting: the measurement ray vector from $x_t$ intersects 3D grid cells, and the associated depths are calculated with Euclidean distance. Let $\mathbf{r}_{k+}$ be the event that the $k$-th cell of reduced map $r$ is the closest occupied space, and the \emph{forward sensor model} $p(z|\mathbf{r}_{k+},X_t)$ is known from the stochastic properties of the depth sensor (e.g. a Gaussian distribution). Then, the probability of $\mathbf{r}_k$ based on $z$ is commonly referred to as the \emph{inverse sensor model}, which can be written compactly as
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})&=\eta\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1}),
\end{align}
where $\eta$ serves as a normalizer,
\begin{align}
\label{eqn:allEta}
\eta
&=
\bigg[\sum_{i=1}^{n_{r}+1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\} p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_i^-\bigg]^{-1},
\end{align}
and the probability before normalizing is
\begin{align}
\label{eqn:Unnormalized}
\tilde P(\mathbf{r}_{k}|z,X_{1:t},&Z_{1:t-1})=\mathbf{P}_k^-
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]
+ \bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{k+},X_t)\mathbf{P}_k^-,
\end{align}
where a priori probability is $\mathbf{P}_k^-=P(\mathbf{r}_{k}|X_{1:t-1},Z_{1:t-1})$, its complement is $\bar{\mathbf{P}}_k^-=1-\mathbf{P}_k^-$, $P(\bar{\mathbf{r}}_{0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience, and $p(z|\mathbf{r}_{(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading. By using repeated terms in \refeqn{allEta} and \refeqn{Unnormalized}, this equation has linear complexity with $n_r$ for all cells in reduced map $r$, amortized to $\mathcal{O}(1)$ for each cell, which is easily computed in real-time.

In short, an exact solution to occupancy probability can be computed from the most recent measurement in real time provided the stochastic sensor properties and a priori probabilities. This solution avoids common approximations or machine learning, and simplifies future measurement predictions, described next with expected entropy calculation.

\subsection{Expected Measurement Ray Entropy}

Here we define entropy as a measure of 3D occupancy grid map uncertainty. Shannon's entropy~\cite{StaGriBur05} for the $i$-th cell and the complete map are
\begin{align}
\label{eqn:ShannonsEntropyCell}
H(P(\mathbf{m}_i))&=-P(\mathbf{m}_i)\log{P(\mathbf{m}_i})-P(\bar{\mathbf{m}}_i)\log{P(\bar{\mathbf{m}}_i}),
\\
\label{eqn:ShannonsEntropyMap}
H(P(m))&=\sum_{i=1}^{n_m}H(P(\mathbf{m}_i)),
\end{align}
respectively. This serves as an effective measure because \refeqn{ShannonsEntropyCell} is minimized for $P(\mathbf{m}_i)=0$ and $P(\mathbf{m}_i)=1$ (smallest uncertainty), but is maximized when $P(\mathbf{m}_i)=0.5$ (largest uncertainty).

Next, we define a candidate future pose. Let candidate location $x_c\in\Re^3$ be the 3D location of a future pose, and $R_c\in\SO$ be the candidate attitude, such that the candidate future pose is $X_c=\braces{x_c,R_c}$. Consider measurement $z_c$, which is a possible measurement from pose $X_c$. Since candidate location $x_c$ and the grid cell locations are known, the distance from the robot to the $k$-th cell, namely $z_{c,k}$ is easily solved with 3D ray casting.

Finally, we present the fundamental equations to predict map information gain from a potential future measurement based on~\cite{KauAiLee16}. The measurement ray $z_c$ can only change the cell entropies along its ray, and therefore the reduced map $r\subset m$ is used, similar to mapping. The history of poses and measurements are required for subsequent equations as they affect a priori probabilities, but are removed below for simplicity. The entropy after $z_c$ updates the map is predicted as
\begin{align}
\label{eqn:DiscExpEntropyRay}
&\text{E}[H(P(r|x_c,z_{c}))]=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(\mathbf{r}_k|x_c,z_{c,k}))P(z_{c,k}|x_c)\bigg\}.
\end{align}
The first term of the summand, namely $H(P(\mathbf{r}_k|x_c,z_{c,k}))$, is simply the entropy \refeqn{ShannonsEntropyCell} of the inverse sensor model \refeqn{RayISMAnswer}--\refeqn{Unnormalized}. The second term of the summand, namely $P(z_{c,k}|x_c)$, can be solved by combining \refeqn{allEta} as
\begin{align}
\label{eqn:ProbMeas}
P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}},
\end{align}
where $\eta_{c,k}$ refers to the normalizer based on the measurement $z_{c,k}$.

In short, the expected entropy along a ray is calculated by decomposing the measurement into possible range values with 3D ray casting. Then, we find the entropy from each range value, multiplied by the probability of that outcome using terms from the inverse sensor model. The computational complexity is $\mathcal O(n_r^2)$, but this can be reduced substantially by neglecting highly-improbable measurement outcomes, shown in~\cite{KauAiLee16}.

\section{Autonomous Exploration in Complex 3D Environments}
\label{sec:Explore3D}

In this section, we propose how multiple measurement rays in 3D can determine map information gain from a pose candidate. Then, we show how a collision-free trajectory is determined with a 3D Dijkstra's search. Finally, we combine 3D information gain and travel costs into a single optimization.

\subsection{Map Information Gain in 3D}

The expected entropy for an arbitrary 1D ray spanning 3D space can be calculated with \refeqn{DiscExpEntropyRay} and \refeqn{ProbMeas} in a computationally tractable manner. However, computing the expected entropy from multiple rays simultaneously has exponential complexity, and is therefore computationally intractable. Additionally, numerous rays of a single scan commonly intersect the same grid cells, making consideration of \emph{every} measurement ray unnecessary.

Instead of considering expected entropy from a complete scan, we propose a real-time solution that selects sample measurements to determine an optimal attitude at each candidate pose and the associated map information gain. We assume the vehicle is capable of level flight, so the third axes of the world and body are aligned (Fig. \ref{fig:transforms}), so $R_c$ can be expressed as
\begin{align*}
R_c=\begin{bmatrix}
\cos\psi & -\sin\psi & 0
\\
\sin\psi & \cos\psi & 0
\\
0 & 0 & 1
\end{bmatrix}, \quad 0\leq\psi<2\pi,
\end{align*}
where $\psi$ represents the angle about the third body-fixed axis. The direction of a 3D measurement may also have a nonzero component in the vertical direction. This is achieved by rotating the sensor frame angle $\theta$ about the second sensor-fixed axis,
\begin{align*}
R_{c,\text{sensor}}=\begin{bmatrix}
\cos\theta & 0 & \sin\theta
\\
0 & 1 & 0
\\
-\sin\theta & 0 & \cos\theta
\end{bmatrix}, \quad -\frac{\pi}{2}\leq\theta\leq\frac{\pi}{2},
\end{align*}
where $\Theta$ is the angular range for sensor field-of-view in the vertical direction. Combining these two rotation matrices, the measurement $z$ with depth $\norm{z}$ is expressed with respect to a frame fixed to Mars as
\begin{align}
z(\psi,\theta)=\begin{bmatrix}\cos\theta\cos\psi & -\sin\psi & \cos\psi\sin\theta
\\
\cos\theta\sin\psi & \cos\psi & \sin\psi\sin\theta
\\
-\sin\theta & 0 & \cos\theta
\end{bmatrix}
\begin{bmatrix}
\norm{z} \\ 0 \\ 0
\end{bmatrix}
=
\begin{bmatrix}
\cos\theta\cos\psi
\\
\cos\theta\sin\psi
\\
-\sin\theta
\end{bmatrix}\norm{z}.
\end{align}
In short, unit vectors are acquired via Euler angles within certain sensor limits.

% trim={<left> <lower> <right> <upper>}

\begin{figure}[!t]
	\centering
	\begin{subfigure}[t]{0.3\columnwidth}
           	\centering
          	\includegraphics[trim = {21cm 5cm 20cm 5cm}, clip, height=1.0\textwidth]{mars_general.png}
        		\caption{General View}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.3\columnwidth}
           	\centering
          	\includegraphics[trim = {25cm 6cm 16cm 4cm}, clip, height=1.0\textwidth]{mars_psi.png}
        		\caption{$\psi$ Rotation}
    	\end{subfigure}
	\begin{subfigure}[t]{0.3\columnwidth}
           	\centering
          	\includegraphics[trim = {23cm 4cm 17cm 6cm}, clip, height=1.0\textwidth]{mars_theta.png}
        		\caption{$\theta$ Rotation}
    	\end{subfigure}
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
\caption{The three frames are shown with three axes (red: first axis, green: second axis, blue: third axis). The Mars frame (short and thick) is fixed to the planet. The body-fixed frame (medium thickness and length) is fixed to the camera frame (long and narrow).}
\label{fig:transforms}
\end{figure}

Next, we show how multiple 3D measurements provide an estimate for information gain. Let $\Psi$ and $\Theta$ be the angular ranges for sensor field-of-view (FOV) in the horizontal and vertical directions, respectively. Within the sensor FOV, we select sample measurements for evaluating their expected information gains. Let the number of sample measurements be $n_\psi$ and $n_\theta$ for horizontal and vertical rotations, respectively. Since $R_c$ and the FOV are known, the set of sample measurement rays, namely $\mathcal{Z}(R_c)$, is known as well. Then, the expected information gain for a single measurement ray $z$ from candidate position $x_c$ and attitude $R_c$ is the negative expected entropy change,
\begin{align}
\label{eqn:Iray}
\mathcal{I}_\text{ray}(x_c,R_c,\psi,\theta)=
\begin{cases}
H(P(r))-\text{E}[H(P(r|x_c,z(\psi+\tilde{\psi},\theta+\tilde{\theta})))], & \mbox{if } z(\psi+\tilde{\psi},\theta+\tilde{\theta})\in\mathcal{Z}(R_c), \\ 
0,                                                                                    & \mbox{otherwise},
\end{cases}
\end{align}
where the mounting yaw $\tilde{\psi}$ and pitch $\tilde{\theta}$ of the sensor is a fixed rotation in the horizontal and vertical directions, respectively in that order, such that $-\pi<\tilde{\psi}\leq\pi$ and $-\frac{\pi}{2}+\frac\Theta2\leq\tilde{\theta}\leq\frac{\pi}{2}-\frac\Theta2$. A positive $\tilde{\theta}$ corresponds to rotating the depth sensor from forward to downward toward the surface of Mars. The expected information gain for the full scan is the summation of expected information gains for individual rays
\begin{align}
\label{eqn:Iscan}
\mathcal{I}_\text{scan}(x_c,R_c)=\sum_{i_\psi=0}^{n_\psi-1}\sum_{i_\theta=0}^{n_\psi-1}\mathcal{I}_\text{ray}\left(x_c,R_c,\frac{\Psi i_\psi}{n_\psi},\frac{\Theta i_\theta}{n_\theta}\right).
\end{align}

Finally, \refeqn{Iscan} is computed for $n_\psi$ possible attitudes when the robot is at location $x_c$, one corresponding to each yaw angle $\psi$. Then, the optimal attitude at $x_c$ is
\begin{align}
\label{eqn:RcStar}
R_c^*=\argmax_{R_c} \mathcal{I}_\text{scan}(x_c,R_c).
\end{align}

In short, the expected information gain from each possible measurement ray is obtained using predicted entropy from \refeqn{DiscExpEntropyRay} and \refeqn{ProbMeas}. Then the expected information gain from all possible scans at a candidate location is calculated from \refeqn{Iray} and \refeqn{Iscan}. Finally, the optimal attitude at this candidate location is found using \refeqn{RcStar}.

\subsection{Collision-Free Trajectory in 3D}

Next, we present a fairly straightforward application of Dijkstra's search to 3D occupancy grid mapping. First we reduced the number of grid cells to cubic blocks large enough for a robot to fit completely inside. Based on the occupancy probability of these blocks and their neighbors, we determine which cells are reachable, and what are their travel costs by building a cost map based on Dijkstra's algorithm.

Here we describe how a reduced map is generated for collision avoidance and motion planning. If the largest edge-to-edge distance of the robot, namely $\rho$, is not larger than the 3D grid cell size $\alpha$, this step is unnecessary. Let $k\geq1$ be an integer such that $k\alpha\geq\rho$ and $(k-1)\alpha<\rho$, i.e., $k$ is the minimum number of grid cells in each dimension capable of fully-enclosing the robot inside this cube. Then, we decompose the complete probabilistic 3D into map $m_\text{reduced}$, which is composed of larger cubic cells that encompass $k^3$ cells from map $m$. Let $\mathbf{m}_{\text{reduced},k}$ denote the $k$-th cell of $m_\text{reduced}$ being occupied, and let $a_k\subset m$ be those $k^3$ cells from $m$ composing this larger cell. Then, we apply an effective rule for collision-avoidance from~\cite{KauTakAiLee18} to obtain the probability of $\mathbf{m}_{\text{reduced},k}$,

\begin{align}
\label{eqn:Proj3DMapComb}
P(\mathbf{m}_{\text{reduced},k})= 
\begin{cases}
    \max_{i\in a_k}{P(\mathbf{m}_i)},	&\text{if} \ \max_{i\in a_k}{P(\mathbf{m}_i)}\geq P_\text{thresh},\\
    \min_{i\in a_k}{P(\mathbf{m}_i)},	& \text{otherwise},
\end{cases}
\end{align}
where initial probability is $0<P_\text{init}<1$ and threshold the threshold probability is constrained to $P_\text{init}<P_\text{thresh}<1$. This approach uses probabilities that have changed from $P_\text{init}$, and favors occupied cells to avoid risking collisions.

Next, we show how $\mathbf{m}_{\text{reduced},k}$ is used in Dijkstra's search~\cite{Dij59}. Defining $0<P_\text{coll}<1$ as the acceptable probability of collision, every grid cell of $m_\text{reduced}$ is considered a safe and unvisited node if its occupancy probability, and the occupancy probabilities of its neighbors (sharing a face, edge, or corner), are below $P_\text{coll}$. A cost map is built from the starting robot location by neighboring nodes, where the cost to travel to another node is based on Euclidian distance: a cell sharing a face is $\alpha$ away, a cell sharing an edge is $\sqrt2\alpha$ away, and a cell sharing a corner is $\sqrt3\alpha$ away. Generating a cost map for all reachable cells in $\mathbf{m}_{\text{reduced},k}$ provides a collision-free travel cost for all reachable candidate pose locations. Once an optimal pose is selected, the path is easily found from the cost map using steepest descent.

\subsection{Optimal 3D Pose}

Here we present how an optimal pose is selected based on information gains at optimal attitudes and known collision-free travel costs. First, we show a function to account for travel costs, and then we account for these costs inside the optimization.

The proposed autonomous exploration follows a receding-horizon framework, meaning that optimal poses and trajectories are computed as quickly as possible, even if the robot has not reached the prior optimal pose. However, the computation is not instantaneous; let the distance $d_\text{opt}$ denote the distance that the robot can travel between optimal pose updates. Then, we define a \emph{bump function} loosely based on~\cite{Joh06}, defined as
\begin{align}
\label{eqn:BumpFunRecedingHorizon}
\mathcal B(d)=
\begin{cases}
f_\text{max},												& \text{if }0\leq d\leq d_\text{opt}
\\
(f_\text{max}-f_\text{far})\exp\braces{-\beta(d_\text{opt}-d)^2}+f_\text{far},	& \text{if }d>d_\text{opt}
\end{cases},
\end{align}
where $d$ is a travel distance along the cost map, $f_\text{max}>f_\text{far}>0$, and $\beta>0$ is an exponential decay rate of the bump, which is illustrated in Fig. \ref{fig:recedingHorizonBumpFun}. The main idea is that when travel costs are low, the bump function is large, but the bump decreases with increasing travel cost.

	\begin{figure}
	\vspace*{-0.3\columnwidth}
		\centerline{
			\includegraphics[width=0.6\columnwidth]{recedingHorizonBumpFunFlattened.pdf}
		}
	\vspace*{-0.25\columnwidth}
		\caption{The bump function is designed to prioritize local trajectories to avoid traversals across the map. However, there is no time wasted up to $d_\text{opt}$ because this distance may be achieved in the minimum exploration computation time.}
		\label{fig:recedingHorizonBumpFun}
	\end{figure}

Finally, we incorporate expected information gain and travel costs into a unified optimization. Given $d(x_c)$ from Dijkstra's cost map, the optimal candidate location is
\begin{align}
x_c^*=\argmax_{x_c}\mathcal{I}_\text{scan}(x_c,R_c^*)\mathcal B(d(x_c)),
\end{align}
and therefore the optimal pose is $X_c^*=\braces{x_c^*,R_c^*}$.

In conclusion, every reachable candidate pose is considered for its optimal attitude. The travel costs, acquired from Dijkstra's search, serve as input for the bump function, which prioritizes local trajectories to avoid costly map traversals. Then, the optimal candidate location is selected to maximize a combination of expected information gain and travel cost.

%\begin{itemize}
%	\item Describe the entropy calculation of expected information gain of all ray combinations is computationally intractable
%	\item Proposed strategy: multiple rays in 3D, and optimal attitude covers the rays with the greatest summation of information gains
%	\item Dijkstra's search in 3D, polynomial trajectories
%	\item Complete optimization
%	\item Concluding paragraph
%\end{itemize}

\section{Numerical Examples}
\label{sec:NumExamples}

\begin{itemize}
	\item Mars Environment: simulation environment generation, depth sensor facing slightly downward
	\item Software structure: ROS nodes, full map/reduced map for entropy options
	\item Numerical Results: reduced map used for entropy, full map used for entropy, Figs: RGB/heightmap/Gazebo, 3D colored grid cells with candidates (both full and reduced maps), entropy over time (full and reduced on same plot)
\end{itemize}
% software structure
% Numerical Result
  % reduced map used for entropy
  % full map used for entropy

\section{Conclusions}

\bibliography{../../BibSources}% master source for all publications

\end{document}

### 2018-06-09 Conference/SIMPAR 2016/CDC 2016/CDC16.tex
added Conference folder

\documentclass[letterpaper, 10pt]{ieeeconf}
\IEEEoverridecommandlockouts \overrideIEEEmargins
\usepackage{amsmath,amssymb,url,times,subfigure,graphicx,theorem}
\usepackage{graphicx,subfigure,hyperref}
\usepackage{color,comment}
\usepackage{csquotes}

%\usepackage{refcheck}

%\usepackage{epic,eepic}

\usepackage[]{algorithm2e}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\abs}[1]{\ensuremath{\left| #1 \right|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\ip}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mbox{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mbox{tr}\ensuremath{\!\bracket{#1}}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\SO}{\ensuremath{\mathsf{S\mathcal O(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathbf{d}}}
\newcommand{\pair}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\!\langle #1 \rangle\!\rangle}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\ad}{\ensuremath{\mathrm{ad}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

\title{\LARGE \bf
Autonomous Exploration by Expected Information Gain from Probabilisitic Occupancy Grid Mapping
%Autonomous Exploration by Predictive Information Gain from Probabilisitic Occupancy Grid Mapping
%Autonomous Exploration by Predictive Occupancy Grid Mapping
%Autonomous Exploration by Predicting Occupancy Grid Maps
}

\author{Evan Kaufman, Taeyoung Lee, and Zhuming Ai%, and Ira S. Moskowitz%
\thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 {\tt \{evankaufman,tylee\}@gwu.edu}}
\thanks{Zhuming Ai, Information Management \& Decision Architectures, U.S. Naval Research Laboratory,  Washington, DC 20375}
\thanks{This research has been supported by the U.S. Naval Research Laboratory Base Program ``Intelligent Microflyer'' and in part by NSF under the grants CMMI-1243000, CMMI-1335008, and CNS-1337722.}
}

\newcommand{\EditTL}[1]{{\color{red}\protect #1}}

\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{cor}{Corollary}
\newtheorem{remark}{Remark}
\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Occupancy grid maps are spatial representations of environments, where the space of interest is decomposed into a number of cells that are considered either occupied or free. This paper focuses on exploring occupancy grid maps by predicting the uncertainty of the map. Based on recent improvements in computing occupancy probability, this paper presents a novel approach for selecting robot poses designed to maximize expected map information gain represented by the change in entropy. This result is simplified with several approximations to develop an algorithm suitable for real-time implementation. The predicted information gain proposed in this paper, in conjunction with an existing motion planner to avoid obstacles, yield an effective autonomous exploration strategy, which is illustrated by numerical examples. 
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Robotic mapping is the process of generating maps, which represent the environment in close proximity of a robot. The popular occupancy grid mapping representation provides information about the free and occupied space as the robot measures its surroundings~\cite{ThrBurFox05}. Occupancy grid mapping is commonly used while estimating robot location and orientation, known as simultaneous localization and mapping (SLAM). Several SLAM approaches are applied to a variety of problems when mobile robots traverse through an environment to generate a map, but the trajectories are assumed given (e.g.~\cite{ThrBurFox05,DurBai06,CheChe09}). A major problem is that an effective trajectory would be unavailable knowing little or nothing about an environment. 
Autonomous exploration solves this problem, by choosing robotic motion through uncertain space, while building a map that is used for subsequent navigation~\cite{Yam97}.
In this paper, a robot chooses a trajectory designed to maximize information gain of the occupancy grid map.

A common approach to solve the autonomous exploration problem is known as frontier-based exploration, such as ~\cite{Yam97,Yam98}. This is the process of executing actions that move the robot toward the closest boundary between visited and unvisited space, known as a frontier.
Then the robot takes measurements at this location such that the mapped territory expands, and thus the new frontiers are pushed back. This process is repeated until the map is well-known.
Frontier-based exploration assumes that repeatedly moving toward the closest frontier and taking measurements are the best actions to gain new information about the map, though these systematic actions are not based on any consideration of the future uncertainty of the probabilistic map or optimality.
%there is no guarantee that these actions satisfy any optimality criteria.

%The uncertainty of an occupancy grid map is commonly measured with Shannon's entropy, which is a metric that measures map uncertainty.

Approaches aimed at determining the future uncertainty of an occupancy grid map commonly suffer from several approximations.
%Determining future occupancy grid map uncertainty requires predictive estimations, which suffer from several approximations. 
Most evident is that all approaches rely on inaccurate probabilistic occupancy grids. This mapping representation is typically based on heuristic approximations (e.g.~\cite{MorElf85,Elf89,ChoLynHutKanBurKavThr05,And09,PirRutBisSch11}) or simulated learning techniques \cite{ThrBurFox05,Thr01}. Map-based information gain is commonly determined with Shannon's entropy~\cite{CarDamKumCas15}, an uncertainty metric based on grid cell occupancy probability. Since this probability is heuristic or learned, any measure of entropy is subject to errors from approximated probability.
Furthermore, the expected entropy of an action is not directly calculated~\cite{CarDamKumCas15,StaGriBur05}. Instead, these particle filter-based approaches assume that expected entropy is equivalent to entropy based on expected measurement scans~\cite{JohStaPfaBur07}. This assumption is subject to inaccuracies due to the nonlinear nature of Shannon's entropy. %Furthermore, these approaches typically generate maps with a particle filter, which commonly require some tuning and may consume large amounts of onboard memory.

In this paper, we propose an autonomous exploration technique based on expected entropy change that avoids the aforementioned issues. 
%The occupancy grid mapping scheme is based on~\cite{KauLeeAiMos16}, which provides the exact solution of occupancy grid mapping of a single measurement ray. 
We develop a solution for expected entropy gain based on exact occupancy grid mapping~\cite{KauLeeAiMos16}. This solution directly calculates expected entropy from a probabilistic map based on similar assumptions of those applied to occupancy grid mapping.
Though more accurate than several other approaches, the proposed solution requires large computational resources, making real-time implementation difficult in certain scenarios.
Thus, we provide the tools to approximate the exact solution that reduce the computational requirements with realistic assumptions.
The motion of the robot is formulated as an optimization problem, where the expected entropy change is minimized, or equivalently the expected information gain is maximized.
In short, this paper presents an accurate and computationally-efficient approach to predict map information gain for autonomous exploration with a novel structure that avoids several common assumptions and approximations.

The paper is organized as follows. The problem is formulated in Section II. The exact solution to the expected uncertainty due to an individual 1D measurement is derived in Section III, along with approximations to the solution. The concept of entropy is extended to 2D autonomous exploration with a numerical example in Section IV, followed by conclusions.

%\begin{itemize}
%%	\item Introductory Paragraph
%%	\begin{itemize}
%%		\item Motivation for mapping
%%		\item Background on occupancy grid mapping
%%		\item Background on SLAM and shortfall of SLAM: trajectory is assumed known
%%	\end{itemize}
%	\item Related research
%	\begin{itemize}
%		\item Motion planning: moving toward a goal assuming a map is known
%		\item If map is unknown: frontier-based autonomous exploration
%		\item Problematic assumption: moving toward uncertain space will yield high information gain
%	\end{itemize}
%	\item Proposed approach
%	\begin{itemize}
%		\item Description of Shannon Entropy
%		\item Focus on how we expect the entropy to \emph{change} rather than just its current value
%		\item Work builds largely off contributions in exact inverse sensor models
%		\item Goal: design a computationally-efficient approach to choose a robot pose based on predicting map information gain
%	\end{itemize}
%\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Problem Formulation}

\subsection{Occupancy Grid Mapping}

Let a map $m$ be decomposed into $n_m$ evenly-spaced 2D grid cells, where the $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,...,n_m}$, that is defined as $\mathbf{m}_i=1$ when occupied, and $\mathbf{m}_i=0$ when free. The location and size of each grid cell is assumed known. Therefore, a map $m$ is defined by $\{\mathbf{m}_1,\mathbf{m}_2,\ldots, \mathbf{m}_{n_m}\}$, and there are $2^{n_{m}}$ possible maps.

Another random variable is defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience. The probability that the $i$-th cell is occupied is $P(\mathbf{m}_i)$, and the probability that it is free is $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$. The random variables $\mathbf{m}_i$ are assumed as mutually independent variables, i.e.,
\begin{align}
P(m)=P(\mathbf{m}_1,\mathbf{m}_2,...,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i).
\end{align}

%Occupancy grid mapping produces a probabilistic map; none of the $n_m$ grid cell occupancies are completely certain.
Occupancy grid mapping is the process of determining these probabilities based on robot poses and measurement scans. 
More explicitly, let $t$ be a discrete variable for time. The $t$-th pose is denoted $X_t=\braces{x_t,R_t}$, where the planar position and direction of the robot at $t$ are denoted by $x_t\in\Re^2$ and $R_t\in\Sph^1=\braces{q\in\Re^2|\norm{q}=1}$, i.e., the attitude corresponds to a direction on a 2D plane. 
%the robot position $x_t\in\Re^2$ and attitude $R_t\in\Sph^1$ are confined to a 2D plane such that $\Sph^1=\braces{q\in\Re^2|\norm{q}=1}$.
 Let $X_{1:t}$ denote the history of poses from the initial time to the current time, i.e., $X_{1:t}=\{X_1,X_2,\ldots, X_t\}$. At each pose, the robot receives a 2D measurement \emph{scan}, composed of $n_z$ measurement \emph{rays}. These rays are 1D depth measurements of known direction from the current pose to the closest occupied space, subject to a known \emph{forward sensor model} $p(z_{t,l}|m,X_t)$, where $z_{t,l}$ is the $l$-th measurement ray of the $t$-th scan $Z_t=\braces{z_{t,1},z_{t,2},\ldots,z_{t,n_z}}$ and $Z_{1:t}$ is the measurement history. 
%More explicitly, let the $t$-th measurement \emph{scan} be composed of $n_z$ measurement \emph{rays}, i.e., $Z_t=\braces{z_{t,1},z_{t,2},\ldots,z_{t,n_z}}$, so the measurement scan history is $Z_{1:t}=\{Z_1,Z_2,\ldots, Z_t\}$.

The goal of occupancy grid mapping is to obtain the \emph{inverse sensor model} $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$ for all grid cells inside the sensor field of view.
Suppose that the inverse sensor model before $t$-th time step is given, namely $P(m|X_{1:t-1},Z_{1:t-1})$, and the $l$-th measurement ray $z_{t,l}$ is available from pose $X_t$.
The inverse sensor model due to $z_{t,l}$ at the $t$-th time step follows Bayes' rule,
\begin{align}
\label{eqn:BayesRuleRayISM}
P(&m|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\frac{p(z_{t,l}|m,X_{1:t},Z_{1:t-1})P(m|X_{1:t-1},Z_{1:t-1})}{p(z_{t,l}|X_{1:t},Z_{1:t-1})}\nonumber
\\
&=\eta_{t,l}p(z_{t,l}|m,X_{1:t},Z_{1:t-1})P(m|X_{1:t-1},Z_{1:t-1}),
\end{align}
where $\eta_{t,l}$ is a normalizing constant independent of $m$, and we have used the fact that $X_t$ carries no information about $m$ without $z_{t,l}$. The issue is that $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ is perceived as intractable to calculate because the exact solution considers every combination of map $m$ to obtain the normalizer $\eta_{t,l}$, and therefore this approach yields $\mathcal O(2^{n_m})$ operations per grid cell. This perception is why heuristic solutions based on learning or intuition substitute the inverse sensor model. Recently, \refeqn{BayesRuleRayISM} was solved in a computationally-efficient manner to obtain an exact solution in real-time~\cite{KauLeeAiMos16}, which is summarized next.

Consider a reduced map, namely $r_l$, composed of only $n_{r,l}$ grid cells that the $l$-th measurement ray intersects, indexed by increasing distance from the robot. Let the $k$-th cell of $r_l$, namely $\mathbf{r}_{l,k}$, correspond to $\mathbf{m}_i$, i.e., $\mathbf{r}_k=\mathbf{m}_i$ and $\bar{\mathbf{r}}_k=\bar{\mathbf{m}}_i$.
From~\cite{KauLeeAiMos16}, the inverse sensor model of the individual grid cell $\mathbf{m}_i$ is
\begin{align}
\label{eqn:InvSenModWithProbDens}
&P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\eta_{t,l}
P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\
&\quad\times \bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\nonumber\\
&\quad + \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,k+},X_t)
\bigg],
\end{align}
where $\mathbf{r}_{l,k+}$ corresponds to all $r_l$ such that cells with index less than $k$ are free, the $k$-th cell is occupied, and cells with greater index are irrelevant. Put differently, the $k$-th cell is the closest occupied cell along the $l$-th ray. The normalizer is defined as
\begin{align}
\label{eqn:allEta}
\eta_{t,l}
&=
\bigg[\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1}.
\end{align}
This approach to obtain the inverse sensor model is advantageous because of computational efficiency and accuracy, obtaining the exact solution in $\mathcal O(n_{r,l}+1)$ operations for \emph{all} grid cells visible by this measurement ray. In occupancy grid mapping, one may use \refeqn{InvSenModWithProbDens} and \refeqn{allEta} in real-time to avoid common approximations or learned solutions.

\subsection{Shannon's Entropy and Map Information}

Since occupancy grid mapping provides a probabilistic representation of surrounding space, this mapping scheme holds probabilistic information about the uncertainty of the map. Shannon's entropy is commonly used as a measure of uncertainty, such as in~\cite{CarDamKumCas15,StaGriBur05}. Given the probabilities of each grid cell of map $m$, Shannon's entropy is defined as the sum of the entropy of each cell,
\begin{align}
\label{eqn:ShannonsEntropyDef}
H(&P(m))=\nonumber\\&-\sum_{i=1}^{n_m}\braces{
P(\mathbf{m}_i)\log{P(\mathbf{m}_i})+P(\bar{\mathbf{m}}_i)\log{(P(\bar{\mathbf{m}}_i}))}.
\end{align}
The entropy of the $i$-th grid cell is maximized when $P(\mathbf{m}_i)=0.5$ (greatest uncertainty) and minimized when $P(\mathbf{m}_i)\in\braces{0,1}$ (smallest uncertainty). %; thus Shannon's entropy is a measure of uncertainty.

%\begin{itemize}
%	\item Define the entropy, and describe uncertainty metric meaning
%	\item Entropy can be calculated much more accurately than prior work now that a computationally-efficient exact inverse sensor model is available
%\end{itemize}

\subsection{Autonomous Exploration}

The goal of autonomous exploration is to determine future poses that maximize the map information gain, or equivalently minimize the map entropy. 
%Let the objective function from the first time step $t=1$ to the $t_f$-th time step be
%\begin{align}
%\label{eqn:ObjFun}
%\mathcal I(X_{2:t_f})&=H(P(m|X_{1},Z_{1}))\nonumber\\&\quad-\text{E}\left[H(P(m|X_{1:t_f},Z_{1:t_f}))\right].
%\end{align}
Since autonomous exploration is conducted in uncertain environments, a complete trajectory cannot be known before the robot takes measurements of the entire reachable space.
Instead, an objective function governs the motion between intermediate poses, building an occupancy grid map along the way.

Suppose at time $t$, $P(m|X_{1:t},Z_{1:t})$ is given. Let $X_c$ be the pose of the robot after the $t$-th time step.
The expected information gain $\mathcal I(X_c)$ is defined as the negative change
\begin{align}
\label{eqn:ObjFun}
\mathcal I(X_c)&=H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\quad-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right],
\end{align}
where $H(P(m|X_{1:t},Z_{1:t}))$ may be computed with \refeqn{ShannonsEntropyDef} and $\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right]$ is the expected entropy when the robot is located at $X_c$, where the expectation is taken over $Z_c$.
Since autonomous exploration is formulated as a trajectory optimization, the goal is to repeatedly choose optimal $X_c^*$ that maximizes the objective function
\begin{align}
\label{eqn:Objective}
X_c^*=\argmax_{X_c}{\mathcal I(X_c)},
\end{align}
where $X_c^*$ satisfies the inequality constraint to avoid collisions,
\begin{align}
\label{eqn:CollisionInequalityConstraint}
P_\text{collision}(X)=1-\prod_{i\in\mathcal C_X}(P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})\leq\beta,
\end{align}
where $\mathcal C_X$ is the set of grid cells falling inside a volume of preselected size around $X$ that may cause collision and $\beta>0$ is a small acceptable probability of collision.
Once the robot has translated to $X_c^*$, the above process is repeated.
%Once the trajectory satisfying \refeqn{CollisionInequalityConstraint} between $X_t$ and $X_c$ is completed, the greedy process is repeated with an updated map.

\section{1D Ray Expected Information Gain}

In this section, we present a method to determine the expected entropy from an individual measurement ray from a known pose. This method is based on the exact solution to occupancy grid mapping from~\cite{KauLeeAiMos16}.

\subsection{Single Ray Expected Value of Entropy}
We consider the expected entropy due to the $l$-th ray $z_{c,l}$ at position $x_c$,
\begin{align}
\label{eqn:HRayInt}
\text{E}[H(P(m|&X_{1:t},Z_{1:t},x_c,z_{c,l}))]
\nonumber\\&
=\int_{z_\text{min}}^{z_\text{max}}
H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))\nonumber\\&\qquad\times p(z_{c,l}|X_{1:t},Z_{1:t},x_c)
dz_{c,l}.
\end{align}
%where $z_\text{min}$ and $z_\text{max}$ are sensor limits. Considering the continuous space of possible measurements of $z_{c,l}$ would make \refeqn{HRayInt} intractable to calculate. Thus, w
We discretize the measurement ray space such that $z_{c,l}$ falls on points along the measurement ray intersecting with grid cell edges, similar with how occupancy grid mapping decomposes continuous space into grid cells of length $\alpha$.
The discretized expected value of \refeqn{HRayInt} is
\begin{align}
\label{eqn:DiscExpEntropyRay}
\text{E}[H(P(m|&X_{1:t},Z_{1:t},x_c,z_{c,l}))]
\nonumber\\&=\sum_{k=1}^{n_{r,l}+1}\bigg\{H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l,k}))\nonumber\\&\qquad\times P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)\bigg\},
\end{align}
where index $z_{c,l,k}$ is the distance from $x_c$ to the $k$-th grid cell of the $l$-th reduced map $r_l$, known from geometry.

Standing alone, the term $P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)$ from \refeqn{DiscExpEntropyRay} has a convoluted meaning because the depth $z_{c,l,k}$ does not directly depend on the map. However, we present a method to obtain this discretized probability. %Despite how this term is replaced with a normalizer in ~\cite{ThrBurFox05} replaces
Following the assumption that $z_{c,l}$ is discretized to known distances, the probabilities are proportional to their densities, as the area under the density curve is infinitesimal and fixed. Accounting for all cases, the probability is
\begin{align}
\label{eqn:ProbWithDelta}
P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)=\frac{p(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)}{\sum_{i=1}^{n_{r,l}+1}p(z_{c,l,i}|X_{1:t},Z_{1:t},x_c)}.
\end{align}
Conveniently, this density corresponds to the inverse normalizer defined in \refeqn{BayesRuleRayISM},
\begin{align}
\label{eqn:DiscretizedProb}
p(z_{c,l,k}|X_{1:t},X_c,Z_{1:t})=\eta_{c,l,k}^{-1},
\end{align}
where $\eta_{c,l,k}^{-1}$ is defined in \refeqn{allEta}. However, instead of evaluating the $t$-th pose $X_t$ and measured ray $z_{t,l}$, we simply consider $X_c$ and ray depth $z_{c,l,k}$ based on histories $X_{1:t}$ and $Z_{1:t}$ to obtain the inverse normalizer,
\begin{align}
\label{eqn:allEtaAtC}
\eta_{c,l,k}^{-1}
&=
\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t},Z_{1:t})\bigg\}\nonumber\\&\qquad\times p(z_{c,l,k}|\mathbf{r}_{l,i+},X_c)P(\mathbf{r}_{l,i}|X_{1:t},Z_{1:t}).
\end{align}
By substituting \refeqn{ProbWithDelta} and \refeqn{DiscretizedProb} into \refeqn{DiscExpEntropyRay}, the expected entropy at $X_c$ from individual measurement ray $z_{c,l}$ is
\begin{align}
\label{eqn:HRayComplete}
&\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]=\left(\sum_{i=1}^{n_{r,l}+1}\eta_{c,l,i}^{-1}\right)^{-1}\nonumber\\&\quad\quad\times\sum_{k=1}^{n_{r,l}+1}\bigg\{H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l,k}))\eta_{c,l,k}^{-1}\bigg\},
\end{align}
where $\eta_{c,l,k}^{-1}$ is taken from \refeqn{allEtaAtC}.
Here, \refeqn{HRayComplete} provides the expected entropies for those cells inside the field of view of $r_l$. %; cells outside this reduced map are trivially changed by $0$ due to $z_{c,l}$.
The entropies of cells outside the field of view remain unchanged.

%, and $z_{c,l,n_{r,l}+1}=z_\text{max}$ such that
%\begin{align}
%\label{eqn:SumDiscretizedMeasRay}
%\sum_{k=1}^{n_{r,l}+1}P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)=1,
%\end{align}
%to account for every case on the reduced map.
%
%It is commonly thought that the probability of a future measurement, in this case $P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)$, is too difficult to calculate. As part of the inverse sensor model, the probability density $p(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)$ is absorbed in the normalizer, and typically neglected from further investigation~\cite{ThrBurFox05}. However, the $k$-th normalizer, namely $\eta_{c,l,k}$, is easily obtained with \refeqn{allEta}, and this holds important information.
%From \refeqn{BayesRuleRayISM}, evaluating $z_{c,l}$ now after the $t$-th time step we have
%\begin{align}
%\label{eqn:DiscretizedProb}
%p(z_{c,l,k}|X_{1:t},X_c,Z_{1:t})={\eta_{c,l,k}}^{-1}.
%\end{align}
%%\begin{align}
%%\label{eqn:DiscretizedProb}
%%P(z_{c,l,k}|X_{1:t},X_c,Z_{1:t})=\frac{\delta z}{\eta_{c,l,k}},
%%\end{align}
%%from \refeqn{BayesRuleRayISM} where $\delta z$ represents an infinitisimal width such that $\delta z\ll\alpha$. 
%Following the assumption that $z_{c,l}$ is restricted to cell edges and a maximum reading, the probabilities are proportional to their densities, as the area under the density curve is infinitesimal and fixed. Accounting for all cases, the probability is
%\begin{align}
%\label{eqn:ProbWithDelta}
%P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)=\frac{p(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)}{\sum_{i=1}^{n_{r,l}+1}p(z_{c,l,i}|X_{1:t},Z_{1:t},x_c)},
%\end{align}
%which may be written in terms of normalizers as
%\begin{align}
%\label{eqn:PRayFromEta}
%P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)=\frac{\eta_{c,l,k}^{-1}}{\sum_{i=1}^{n_{r,l}+1}{\eta_{c,l,i}}^{-1}}.
%\end{align}

%The expected entropy from candidate pose $X_c$ with measurement ray $z_{c,l}$ is obtained with \refeqn{DiscExpEntropyRay} and \refeqn{PRayFromEta}, where the normlizers come from \refeqn{allEta}. The ray inverse sensor model \refeqn{InvSenModWithProbDens} provides the probabilities of the reduced map $r_l$, and all cells outside of $r_l$ remain unchanged. Thus, grid cells outside $r_l$ need not be considered in either term of the objective function \refeqn{ObjFun} with respect to $x_c$ and $z_{c,l}$.
%
%Let $k_\text{c}$ be an index corresponding to a cell along a measurement ray and let $k_\text{m}$ be an index corresponding to a measurement along a measurement ray. 
%Let the objective function of the $l$-th measurement ray \emph{only} be denoted $\mathcal I_\text{ray}(x_c,z_{c,l})$.
%Then, the objective function from \refeqn{ObjFun} is simplified to include just the grid cells intersecting along the ray as
%\begin{align}
%\label{eqn:ObjFunRayComplete}
%&\mathcal I_\text{ray}(x_c,z_{c,l})\nonumber\\&=\sum_{k_\text{c}=1}^{n_{r,l}+1}\bigg[H(P(\mathbf{r}_{c,l,k_\text{c}}|X_{1:t},Z_{1:t}))\nonumber\\
%&\quad
%-\sum_{k_\text{m}=1}^{n_{r,l}+1}\bigg( H(P(\mathbf{r}_{c,l,k_\text{c}}|X_{1:t},Z_{1:t},x_c,z_{c,l,k_\text{m}}))\nonumber\\
%&\quad\quad\quad\quad\quad\quad\quad\times P(z_{c,l,k_\text{m}}|X_{1:t},Z_{1:t},x_{c})\bigg)
%\bigg].
%\end{align}

%\subsection{Full Scan Expected Value of Entropy}

\subsection{Approximation of Expected Ray Entropy}

The order of computation for each measurement ray is $\mathcal O((n_{r,l}+1)^2)$ since the summations of \refeqn{allEtaAtC} are embedded in \refeqn{HRayComplete}. Such computational requirements might be too cumbersome for some processors, especially if the measurement ray is expected to intersect many grid cells. Many of those intersections provide negligible information since the probability of the $l$-th measurement ray capturing certain cell depths is close to zero.

The approximation of expected ray entropy provides a method to reduce the computation of \refeqn{allEtaAtC} substantially. This goal is achieved by systematically selecting a smaller set of grid cells to consider over the summations of \refeqn{allEtaAtC} and \refeqn{HRayComplete}.
The smaller set is determined by the probability that each cell is captured by the measurement ray, known as the detection probability. This can be found recursively as
%These cells correspond to the $\hat n$ cells with the greatest probabilities of being captured by the measurement ray, known as the detection probability. This can be found recursively as
%
%select the terms that contribute
%We systematically select a limited number of terms that affect
%
%
%The large order of computation is reduced substantially by lowering the summation limits of \refeqn{allEtaAtC} and \refeqn{HRayComplete}. Consider the $\hat{n}$ intersections of $r_l$ with the greatest probability of returning a measurement. The detection probability is found recursively as
\begin{align}
\label{eqn:ProbOfFirstCell}
&P(\mathbf{r}_{l,k+}|X_{1:t},Z_{1:t})\nonumber\\&=\bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t},Z_{1:t})\bigg\}P(\mathbf{r}_{l,k}|X_{1:t},Z_{1:t}),
\end{align}
which is the probability that $\mathbf{r}_{l,k}$ is the closest occupied grid cell based on past poses and measurement scans, independent of cells beyond the $k$-th cell from $x_c$.
Let $\hat n>0$ be a fixed number of grid cells for all rays such that $\hat n\leq n_{r,l}+1$ for all $l$.
Let $\hat{r}_{l}$ correspond to the grid cells that yield the $\hat{n}$ maximum values \refeqn{ProbOfFirstCell} (the $\hat n$ most likely ray detections), indexed by increasing distance from candidate location $x_c$.
By replacing the reduced map $r_l$ with $\hat{r}_l$ and changing the summation limits to $\braces{1,2,...,\hat n}$ in \refeqn{allEtaAtC} and \refeqn{HRayComplete}, the order of computation is reduced to $\mathcal O({\hat{n}}^2)$.
Even though the value of $n_{r,l}$ is different among various rays in general, $\hat n$ is fixed, so the computational order is fixed as well.
In short, this method reduces the required computation substantially by systematically neglecting those grid cells with little effect.
%
%
%Then, the objective function \refeqn{ObjFunRayComplete} from this measurement ray is approximated as
%\begin{align}
%\label{eqn:ObjFunRayApprox}
%&\mathcal I_\text{ray}_\text{approx}(x_c,z_{c,l})\nonumber\\&=\sum_{k_\text{c}=1}^{\hat{n}}\bigg[H(P(\hat{\mathbf{r}}_{c,l,k_\text{c}}|X_{1:t},Z_{1:t}))\nonumber\\
%&\quad
%-\sum_{k_\text{m}=1}^{\hat{n}}\bigg( H(P(\hat{\mathbf{r}}_{c,l,k_\text{c}}|X_{1:t},Z_{1:t},x_c,\hat{z}_{c,l,k_\text{m}}))\nonumber\\
%&\quad\quad\quad\quad\quad\quad\quad\times P(\hat{z}_{c,l,k_\text{m}}|X_{1:t},Z_{1:t},x_{c})\bigg)
%\bigg],
%\end{align}
%where $\hat{z}_{c,l,k_\text{m}}$ is the expected measurement distance from $x_c$ to $\hat{\mathbf{r}}_{k_\text{m}}$, known from geometry. This measurement probability comes from \refeqn{PRayFromEta} over a smaller set,
%\begin{align}
%\label{eqn:ApproxSimplifiedPRayFromEta}
%P(\hat{z}_{c,l,k_\text{m}}|X_{1:t},x_{c},Z_{1:t})=\frac{\hat{\eta}_{c,l,k_\text{m}}^{-1}}{
%\sum_{i=1}^{\hat{n}}
%\hat{\eta}_{c,l,i}^{-1}
%},
%\end{align}
%where the normalizer is extended from \refeqn{allEta} as
%\begin{align}
%\label{eqn:SimplifiedAllEta}
%\hat{\eta}_{c,l,k_\text{m}}
%&=
%\bigg[\sum_{i=1}^{\hat{n}}\bigg\{\prod_{j=0}^{i-1}P(\hat{\bar{\mathbf{r}}}_{c,l,j}|X_{1:t},Z_{1:t})\bigg\}\nonumber\\&\times p(\hat{z}_{c,l,k_\text{m}}|\hat{\mathbf{r}}_{c,l,i+},x_c)P(\hat{\mathbf{r}}_{l,i}|X_{1:t},Z_{1:t})\bigg]^{-1}, 
%\end{align}
%and the inverse sensor model from \refeqn{InvSenModWithProbDens} is
%\begin{align}
%\label{eqn:SimplifiedInvSenModWithProbDens}
%&P(\hat{\mathbf{r}}_{c,l,k_\text{c}}|X_{1:t},Z_{1:t},x_c,\hat{z}_{c,l,k_\text{m}})\nonumber
%\\
%&=\hat{\eta}_{c,l,k_\text{m}}
%P(\mathbf{r}_{c,l,k_\text{c}}|X_{1:t},Z_{1:t})\nonumber\\
%&\quad\times \bigg[\sum_{i=1}^{k_\text{c}-1}\bigg\{\prod_{j=0}^{i-1}P(\hat{\bar{\mathbf{r}}}_{c,l,j}|X_{1:t},Z_{1:t})\bigg\}\nonumber\\
%&\quad\times p(z_{c,l,k_\text{m}}|\hat{\mathbf{r}}_{c,l,i+},X_t)P(\hat{\mathbf{r}}_{c,l,i}|X_{1:t},Z_{1:t})\nonumber\\
%&\quad + \bigg\{\prod_{j=0}^{k_\text{c}-1}P(\hat{\bar{\mathbf{r}}}_{c,l,j}|X_{1:t},Z_{1:t})p(z_{c,l,k_\text{m}}|\mathbf{r}_{l,k+},x_c)\bigg\}
%\bigg].
%\end{align}
%In short, the $\hat{n}$ grid cells with greatest detection probabilities from \refeqn{ProbOfFirstCell} serve as the only indices to obtain the approximate objective function $\mathcal I_\text{ray}_\text{approx}(x_c,z_{c,l})$ from \refeqn{ObjFunRayApprox}, which uses \refeqn{ApproxSimplifiedPRayFromEta}--\refeqn{SimplifiedInvSenModWithProbDens}. While a small amount of information is lost by neglecting grid cells that are unlikely to be detected, the required computation is reduced to $\mathcal O(\hat{n}^2)$ where $\hat{n}\ll n_{r,l}+1$. This approximation is justified with Monte Carlo numerical results.
%

\subsection{1D Ray Expected Information Gain Algorithm}

We present an algorithm pseudocode providing the necessary steps to obtain the objective function for a single measurement ray (Table \ref{tab:RayExpectedEntropyGain}). The variable $a_\text{temp}$ is an intermediate variable designed to avoid repeated calculations.
Since this algorithm operates as a function, fixed indices and condition variables are removed for simplification. It can be noted that if $\hat n=n_{r,l}+1$, the ray objective function is computed without approximation.

\begin{table}
\begin{tabular}{ l }
  Function: $\mathcal I_\text{ray}=\text{RayExpInfoGain}(x,P(r),z_{1:n_{r,l}})$\\
  Define $P(\bar{\mathbf{r}}_{0})=P(\bar{\mathbf{r}}_{n_{r,l}+1})=1$;\\
  for $k = 1,2,\ldots,n_{r,l}+1$\\
   \ \ $P(\mathbf{r}_{k+}) = P(\bar{\mathbf{r}}_{0:k-1})P(\mathbf{r}_{k})$;\\
   \ \ $P(\bar{\mathbf{r}}_{0:k})=P(\bar{\mathbf{r}}_{0:k-1})P(\bar{\mathbf{r}}_{k})$;\\
  end for\\
%  $P(\mathbf{r}_{(n_{r}+1)+}) = P(\bar{\mathbf{r}}_{0:n_{r}})$;\\
  Find $\hat r\subset r$ corresponding to the $\hat n$ greatest values of\\
  $\braces{P(\mathbf{r}_{1+}),P(\mathbf{r}_{2+}),\ldots,P(\mathbf{r}_{(n_{r,l}+1)+})}$;\\
  Define $P(\hat{\bar{\mathbf{r}}}_{0})=1$;\\
  for $k = 1,2,\ldots,\hat{n}$\\
   \ \ $P(\hat{\mathbf{r}}_{k+}) = P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\mathbf{r}}_{k})$;\\
   \ \ $P(\hat{\bar{\mathbf{r}}}_{0:k})=P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\bar{\mathbf{r}}}_{k})$;\\
  end for\\
  for $k_\text{m}=1,2,\ldots,\hat n$\\
   \ \ Initialize $\eta^{-1}_{k_\text{m}}=0$;\\
   \ \ for $k_\text{c}=1,2,\ldots,\hat n$\\
   \ \ \ \ $a_\text{temp}=P(\hat{\mathbf{r}}_{k_\text{c}+})p(z_{k_\text{m}}|\hat{\mathbf{r}}_{k_\text{c}+},x)$;\\
   \ \ \ \ $\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=P(\hat{\mathbf{r}}_{k_\text{c}})\eta^{-1}_{k_\text{m}}+a_\text{temp}$;\\
   \ \ \ \ $\eta^{-1}_{k_\text{m}}=\eta^{-1}_{k_\text{m}}+a_\text{temp}$;\\
   \ \ end for\\
   \ \ $P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=\eta_{k_\text{m}}\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})$ for all $k_\text{c}=1,2,\ldots,\hat n$;\\
   end for\\ 
   $P(z_{k_\text{m}}|x)=\frac{\eta^{-1}_{k_\text{m}}}{\sum_{i=1}^{\hat n}\eta^{-1}_{i}}$ for $k_\text{m}=1,2,\ldots,\hat n$;\\

   Initialize $\mathcal I_\text{ray}=0$;\\
   for $k_\text{c}=1,2,\ldots,\hat n$\\
   \ \ $\mathcal I_\text{ray}=\mathcal I_\text{ray}+H(P(\hat{\mathbf{r}}_{k_\text{c}}))$;\\
   \ \ for $k_\text{m}=1,2,\ldots,\hat n$\\
   \ \ \ \ $\mathcal I_\text{ray} = \mathcal I_\text{ray}-H(P(\hat{\mathbf{r}}_{k_\text{c}}|x_c,z_{k_\text{m}}))P(z_{k_\text{m}}|x_{c})$;\\
   \ \ end for\\
  end for\\
  Return: $\mathcal I_\text{ray}$\\
\end{tabular}
\caption{Expected Information Gain from a Measurement Ray}
\label{tab:RayExpectedEntropyGain}
\end{table}

\subsection{Numerical Justification for Approximation}

The purpose of this numerical example is to provide evidence that the approximations are reasonable and speed up the algorithm substantially.
Since a measurement ray produces a depth measurement in a single direction, we only consider a 1D map where the grid cells have spacing $\alpha=0.2$ meters, and the properties of the range sensor are based off the Microsoft Kinect~\cite{PirRutBisSch11,KhoElb12} with maximum reading depth $z_\text{max}=4$m ($20$ grid cells inside the sensor FOV). The goal is to compare the expected entropy $\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$ from \refeqn{HRayComplete} and with an approximation $\text{E}[H_\text{approx}(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$, which only considers $\hat n$ grid cells with highest detection probability \refeqn{ProbOfFirstCell}.

We consider $100$ probabilistic maps to obtain Monte Carlo results to evaluate the approximate entropy. In every Monte Carlo trial, each grid cell has an $80\%$ chance of being free, and the remaining $20\%$ of grid cells are given an a priori probability uniformly distributed between $0$ and $1$. 
Several metrics serve to evaluate $\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$ with $\text{E}[H_\text{approx}(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$. The median expected entropy change is $\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]-H(P(m|X_{1:t},Z_{1:t}))=-0.83792$. The error for the $100$ Monte Carlo cases is defined simply as
\begin{align}
e_{H}&=\frac1{100}\sum_{i=1}^{100}\text{abs}\bigg(\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]\nonumber\\&\qquad\quad-\text{E}[H_\text{approx}(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]\bigg).
\end{align}
The Monte Carlo trials are repeated for $\hat n=\braces{1,2,\ldots,10}$ and the results are plotted in Figure \ref{fig:ApproxJust}.
This example shows a typical case when the summation limits generated from \refeqn{ProbOfFirstCell} have only small effects on \refeqn{HRayComplete}, while providing very large improvements in reducing computation.

\begin{figure}
\centerline{
	\subfigure[Entropy Error]
		{\includegraphics[height=0.35\columnwidth]{JustifyApprox_eH.pdf}
	}	
	\subfigure[Computation Ratio]
		{\includegraphics[height=0.35\columnwidth]{JustifyApprox_tRatio.pdf}
	}
}
\caption{The entropy error is decreased at the cost of increasing computation time in this Monte Carlo 1D measurement ray expected entropy case.}
\label{fig:ApproxJust}
\end{figure}

\section{Autonomous Exploration}

In this section, we develop an autonomous exploration scheme on a 2D map based on choosing movements that minimize entropy and avoid collisions.

\subsection{Pose Selection Optimization}

The robot pose at a future time step is selected to maximize \refeqn{ObjFun} over the search space $\Re^2\times\Sph^1$. We reduce the search by discretizing the search space in both position and attitude as follows. First, we choose $n_c$ candidate locations evenly-spaced along a circle of radius $\delta$, centered around the current location $x_t$, where the $c$-th candidate location is denoted $x_c\in\braces{x_1,x_2,\ldots,x_{n_c}}$. Any location that violates \refeqn{CollisionInequalityConstraint} is excluded to avoid collisions. For each collision-free candidate location, we consider $n_d$ evenly-spaced directions. The rays and attitudes at $x_c$ are denoted $\braces{z_{c,1},z_{c,2},\ldots,z_{c,n_d}}$ and $\braces{R_{c,1},R_{c,2},\ldots,R_{c,n_d}}$, respectively, where the scan with attitude $R_{c,d}$ might cover several ray directions depending on the sensor field of view (FOV). For each candidate location, we choose optimal attitude $R_c^*$ as the summation of the expected entropy changes covered by the scan,
\begin{align}
\label{eqn:FindRc}
&R_c^*=\argmax_{R_{c,d}}\sum_{z_{c,i}\in R_{c,d}\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\qquad\quad\quad\quad\quad-\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,i}))]\bigg).
%\mathcal I_\text{ray}(x_c,z_{c,i})\right).
\end{align}

Therefore, the optimal attitude $R_c^*$ is given as a function of the location $x_c$. The objective function \refeqn{ObjFun} is computed by a summation about the $n_d$ rays as %\refeqn{Objective},
\begin{align}
\label{eqn:ObjFunApprox}
\mathcal I(x_c,R_c^*(x_c))&\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*(x_c),z_{c,i}))\right]\bigg),
\\
x_c^*&=\argmax_{x_c}\mathcal I(x_c,R_c^*(x_c)).
\end{align}
%such that $X_c^*=\braces{x_c^*,R_c^*}$ is the optimal candidate over the discretized search space. Figure \ref{fig:OptProcess} illustrates this process.

\begin{figure}
\centerline{
	\includegraphics[height=0.35\columnwidth]{ExampleOptimalPose.png}
}
\begin{picture}(0,0)(0,0)
\setlength{\unitlength}{0.1\columnwidth}\scriptsize
\put(4.4,2.1){\color{blue}$X_t$}
\put(5.9,1.9){\color{red}$x_1$}
\put(2.9,2.0){\color{red}$x_5$}
\put(4.4,3.6){\color{red}$x_3$}
\put(4.4,0.6){\color{red}$x_7$}
\put(5.4,3.0){\color{red}$x_2$}
\put(3.3,3.1){\color{red}$x_4$}
\put(3.3,1.0){\color{red}$x_6$}
\put(5.4,1.0){\color{red}$x_8$}
\put(6.8,3.1){\color{red}$z_{2,1}$}
\put(4.5,3.1){\color{red}$z_{2,5}$}
\put(5.6,4.1){\color{red}$z_{2,3}$}
\put(5.6,2.2){\color{red}$z_{2,7}$}
\put(6.5,3.7){\color{red}$z_{2,2}$}
\put(4.9,3.8){\color{red}$z_{2,4}$}
\put(5.0,2.5){\color{red}$z_{2,6}$}
\put(6.5,2.5){\color{red}$z_{2,8}$}
\put(6.1,2.9){$X_c^*$}
\end{picture}
\caption{Initially, a robot (blue circle) views down and left (blue sector). Then, the robot considers $n_c=8$ poses (red circles) as possible candidates for where to move next. For each candidate, $n_d=8$ directions are considered (red lines, only displayed on candidate location $c=2$). Then, the expected entropy of each ray is calculated. The scan (red sector) covering those rays with the largest expected entropy decreases is chosen for each candidate, and the best candidate $X_c^*$ (black circle: location, black sector: scan) is chosen to maximize information gain. Finally, a wavefront planner provides the collision-free motion between $X_t$ and $X_c^*$, and the process is repeated.
}
\label{fig:OptProcess}
\end{figure}

%In various modern sensors, such as the Kinect and LIDAR, several measurement rays are available simultaneously at any given scan. One may assume that each measurement ray provides independent information about the expected entropy change. Thus, one may add \refeqn{ObjFun} for each measurement ray. However, this assumption is often invalid in scenarios when a single grid cell is intersected by more than one measurement ray. For example, consider that the $i$-th cell is uncertain before considering $X_c$ as a future destination, i.e., $P(\mathbf{m}_i|X_{1:t},Z_{1:t})=0.5$ where H(0.5)=0.6931. Then, $\mathbf{m}_i$ is intersected by two candidate measurement rays $z_{c,1}$ and $z_{c,2}$, each with an expected entropy decrease of $0.4$. If summed, the total expected entropy value of this cell would be $-0.1069$, which is an impossible value because entropy is lower-bounded by $0$. The assumption is more flagrantly violated with increasing numbers of measurement rays passing through individual grid cells, motivating an alternative solution.
%
%Instead, when considering the $c$-th candidate, one may preselect a sample of evenly-spaced measurement rays, namely $\mathcal N=\braces{1,2,\ldots,n_d}$, deemed representative of the future scan. The spacing of these rays depends on the number of directions considered $n_d$, %which is chosen based on the desired predicted scan resolution and the spacing of the map grid cells, 
%where $n_d<n_z$ in general. Let these rays be denoted $\braces{z_{c,1},z_{c,2},\ldots,z_{c,n_d}}$, where the $d$-th direction satisfies $\braces{z_{c,d}\in\Re^2,\norm{z_{c,d}}=1}$. The goal is to consider enough rays that most grid cells covered by the scan are accounted by the rays of $\mathcal N$, while avoiding the case when multiple ray directions pass through any individual grid cell. In particular, one may choose a location for a pose, then find the expected entropy changes from various measurement rays. The $c$-th attitude $R_c^*$ is chosen such that the scan covers those measurement rays in the field of view (FOV) with the greatest sum of expected entropy decreases,
%\begin{align}
%\label{eqn:FindRc}
%R_c^*=\argmax_{R_c}\left(\sum_{z_{c,d}\in R_c \text{ FOV}}\mathcal I_\text{ray}(x_c,z_{c,l})\right).
%\end{align}
%In short, measurement rays in $n_d$ evenly-spaced directions around location $x_c$ are evaluated with \refeqn{DiscExpEntropyRay}. The attitude $R_c^*$ from \refeqn{FindRc} is determined from the maximum summation of ray expected information gains based off the $\mathcal N$ selected set, so $X_c=\braces{x_c,R_c^*}$ is selected as the candidate at $x_c$ to be compared with other candidates.

%\subsection{Motion Planning and Algorithm}
%
%The candidate pose locations may be determined several ways. For example in~\cite{CarDamKumCas15}, poses near frontiers as described in~\cite{Yam98} are selected as candidates. This choice is based on the intuition that poses near frontiers yield high expected entropy decreases, which is unproven. While frontier poses might produce effective candidates, they present a bias of candidates dependent on the geometry of the map according the robot knowledge. Thus, we choose pose candidates evenly-spaced in a circular pattern around the current location of the robot. Only should a candidate violate \refeqn{CollisionInequalityConstraint}, this candidate need not be considered.
%
%
%%Autonomous exploration is greedy, meaning that 
%We provide an algorithm pseudocode for autonomous exploration with occupancy grid applications. Consider $n_c$ candidate locations, namely $\braces{x_1,x_2,\ldots,x_{n_c}}$, evenly spaced around the current pose in a circular pattern of radius $\delta$, where the $c$-th candidate location is $x_{c}\in\Re^2$. %For each candidate location, consider $n_d$ evenly-spaced directions of the range sensor, namely $\braces{d_1,d_2,\ldots,d_{n_d}}$, where the $d$-th direction satisfies $\braces{d\in\Re^2,\norm{d}=1}$.

The objective function of the optimal pose $X_c^*=\braces{x_c^*,R_c^*}$ must satisfy $\mathcal I(X_c^*)\geq\mathcal I_\text{min}$ to avoid motions where little information is expected to be gained. 
If this condition is not satisfied, then $n_c$ and $\delta$ are multiplied by a scale function $\lambda>1$, candidates are chosen again, and $\mathcal I(X_c^*)$ is reevaluated.
Once the condition is satisfied, the robot must move to this pose while avoiding collisions subject to the inequality constraint \refeqn{CollisionInequalityConstraint}.

We choose the wavefront planner~\cite{ZelJarByrYut93} because it provides a relatively simple motion planning strategy from $x_t$ to $x_c^*$ that avoids local minima. Once the robot completes this motion, the entire process is repeated. This is summarized in the autonomous exploration pseudocode in Table \ref{tab:Alg_AutomousExploration}.

%The first pseudocode in table \ref{tab:Alg_AutomousExploration} outlines how a candidate pose is selected given an objective function. The second pseudocode in table \ref{tab:Alg_ObjFun}

\begin{table}
\begin{tabular}{ l }
%  Define $n_c$: number of candidate locations\\
%  Define $n_d$: number of evenly-spaced measurement rays per candidate\\
  Given: $t=1$, $X_{1}$, $Z_{1}$, $P(m|X_{1},Z_{1})$ from the inverse sensor model\\
  TrajectoryComplete = false;\\
  while TrajectoryComplete = false\\
  \ \ if $X_{t+1}$ is unknown\\
  \ \ \ \ MotionPlanning = false;\\
  \ \ \ \ while MotionPlanning = false\\
  \ \ \ \ \ \ for $c = 1,2,\ldots,n_c$\\
  \ \ \ \ \ \ \ \ Candidate location: $x_{c}$ from evenly-spaced circle around\\
  \ \ \ \ \ \ \ \ $X_t$ with radius $\delta$;\\
  \ \ \ \ \ \ \ \ for $d=1,2,\ldots,n_d$\\
  \ \ \ \ \ \ \ \ \ \ Obtain reduced map $r_{c,d}$ with $n_{r_{c,d}}$ intersections\\
  \ \ \ \ \ \ \ \ \ \ through grid cells with depths $z_{c,d,1:n_{r,d}}$ from geometry;\\
  \ \ \ \ \ \ \ \ \ \ $\mathcal I_\text{ray}(x_c,z_{c,d})=\text{RayExpInfoGain}(x_c,P(r_{c,d}),z_{c,d,1:n_{r,d}})$;\\
  \ \ \ \ \ \ \ \ end for\\
  \ \ \ \ \ \ \ \ $R_{c}^*=\argmax_{R_c}\left(\sum_{z_{c,i}\in R_c \text{ FOV}}\mathcal I_\text{ray}(x_c,z_{c,i})\right)$;\\%\argmax_{R}{\sum_{d=1}^{n_d}\mathcal I_{\text{ray},c,d}|d\in\text{FOV}(R))}$;\\
  \ \ \ \ \ \ \ \ $\mathcal I(x_c,R_c^*)=\sum_{z_{c,i}\in R_c^* \text{ FOV}}\mathcal I_\text{ray}(x_c,z_{c,i})$;\\%\sum_{d\in \text{FOV}(R_c^*)}\mathcal I_{\text{ray}_{c,d}}$;\\
  \ \ \ \ \ \ end for\\
  \ \ \ \ \ \ $x_c^*=\argmax_{x_c}\mathcal I(x_c,R_c^*)$;\\
  \ \ \ \ \ \ $X_c^*=\braces{x_c^*,R_c^*}$;\\
  \ \ \ \ \ \ if $\mathcal I(X_c^*)<\mathcal I_\text{min}$\\
  \ \ \ \ \ \ \ \ $n_c=\text{round}(\lambda n_c)$;\\
  \ \ \ \ \ \ \ \ $\delta=\lambda \delta$;\\
  \ \ \ \ \ \ else\\
  \ \ \ \ \ \ \ \ Obtain $X_{t+1},X_{t+2},\ldots,X_{c}^*$ using a wavefront\\  \ \ \ \ \ \ \ \ planner and minimum rotation;\\
  \ \ \ \ \ \ \ \ Reset $n_c$ and $\delta$ to initial values;\\
  \ \ \ \ \ \ \ \ MotionPlanning = true;\\  
  \ \ \ \ \ \ end if\\
  \ \ \ \ end while\\
%  $X_c^*=\argmax_{X}$\mathcal I_{\text{ray},c
%  \ \ $\mathcal D\subset\braces{1,2,\ldots,n_d}$ represents scan coverage of candidate rays\\
%  \ \ The $c$-th pose candidate: $X_{c}=\braces{x_{c},R_{c}}$;\\
%   end for\\
%  $X_c^*=\argmin_{X_c}({\Delta H(X_c)}|X_c\in\braces{X_1,X_2,\ldots,X_{n_c}})$;\\
%  Motion planner generates a collision-free path between $X_t$ to $X_c^*$\\
%  Move to $X_{t+1}$, obtain $Z_{t+1}$, and find $P(m|X_{1:t+1},Z_{1:t+1})$;
  \ \ end if\\
  \ \ t = t+1;\\
  \ \ Obtain $P(m|X_{1:t},Z_{1:t})$ from the inverse sensor model;\\
  \ \ if map $m$ is completely explored\\
  \ \ \ \ TrajectoryComplete = true;\\
  \ \ end if\\
  end while

\end{tabular}
\caption{Autonomous Exploration via Expected Information Gain}
\label{tab:Alg_AutomousExploration}
\end{table}

\subsection{Numerical Example}

%Maximizing the proposed expected information gain serves as the objective for autonomous exploration. A robot begins in a completely uncertain environment, except for the grid cells inside the immediate vicinity the robot location. 
The robot models its surroundings with an occupancy grid with $15000$ cells where grid cell edges are $\alpha=0.2$m, composing a map with dimensions $30\text{m}\times20\text{m}$. The initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) for grid cells covered by the circular robot of radius $0.1$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
At each time step, the robot receives a measurement scan, where the probabilistic properties of the sensor are taken from~\cite{PirRutBisSch11,KhoElb12}. Then, the $n_c=8$ evenly-spaced candidate locations about a circle of radius $\delta=0.5$m around the current pose location are considered, where $n_d=32$ measurement rays are evenly-spaced about the candidate location.  When no current candidates yield expected information gains above $\mathcal I_\text{min}=2$, $\lambda=1.25$ is multiplied to the current $n_c$ and $\delta$ values. The motion of the robot is restricted to movement on grid cells satisfying \refeqn{CollisionInequalityConstraint} with $\beta=0.01$. The wavefront motion planner guides the robot through the environment without collision. The results are illustrated in Figure \ref{fig:AutonomousExploration}.

\begin{figure}
\centerline{
	\subfigure[Map at $t=50$sec]
		{\includegraphics[height=0.35\columnwidth]{CDC16_t50.png}
	}	
	\subfigure[Map at $t=100$sec]
		{\includegraphics[height=0.35\columnwidth]{CDC16_t100.png}
	}
}
\centerline{
	\subfigure[Map at $t=150$sec]
		{\includegraphics[height=0.35\columnwidth]{CDC16_t150.png}
	}	
	\subfigure[Map at $t=200$sec]
		{\includegraphics[height=0.35\columnwidth]{CDC16_t200.png}
	}
}
\centerline{
	\subfigure[Map at $t=250$sec]
		{\includegraphics[height=0.35\columnwidth]{CDC16_t250.png}
	}	
	\subfigure[Map at $t=300$sec]
		{\includegraphics[height=0.35\columnwidth]{CDC16_t300.png}
	}
}
\centerline{
	\subfigure[Map at $t=350$sec]
		{\includegraphics[height=0.35\columnwidth]{CDC16_t350.png}
	}	
	\subfigure[Map at $t=400$sec]
		{\includegraphics[height=0.35\columnwidth]{CDC16_t400.png}
	}
}\centerline{
	\subfigure[Map at $t=450$sec]
		{\includegraphics[height=0.35\columnwidth]{CDC16_t450.png}
	}	
	\subfigure[Map at $t=500$sec]
		{\includegraphics[height=0.35\columnwidth]{CDC16_t500.png}
	}
}
\caption{A robot (green circle, green marker of its previous pose locations) measures a room with a Kinect depth sensor. The robot moves to maximize its information gain. In the end, the robot is returning to the left room to gain more information about the uncertain regions it originally left behind.% The algorithms update the grid cells from uncertain (grey, $0.5$ occupancy probability) to either occupied (black, $1$ occupancy probability) or free (white, $0$ occupancy probability). Measured cells have occupancy probabilities closer to $0$ or $1$ as more information becomes available, and unmeasured cells remain near $0.5$, or uncertain.
%The exact inverse sensor model approach generates a more accurate result.
}
\label{fig:AutonomousExploration}
\end{figure}

Knowing only that the robot is inside free space at the beginning, the robot carefully navigates the environment while avoiding collisions. The robot motion is governed by a policy that maximizes the map information gain within its set of pose candidates, where the $\hat n=6$ is chosen to approximate the expected entropy of each ray. The median computation time for a single update is $0.7087$sec, though the mean is $1.3263$sec. This discrepancy is due to high-computation outliers corresponding to cases when the robot is locally surrounded by a highly-certain environment; the robot must to traverse a large distance to learn information about another part of the map. Such a task requires evaluating many more pose candidates and motion planning over a larger terrain.

Throughout the numerical example, the robot must make several important decisions on which actions to take. These actions are evaluated based on their expected information gains, not frontiers a predicted measurement scans. If the obstacles were known a priori, the motion planning problem could be solved with a smoother path; however, the autonomous exploration is based on only the information of the map that the robot generates, so the motion planning must be reevaluated repeatedly. Even with these limitations, the robot explores the vast majority of reachable space in the $500$sec period.

\section{Conclusions}

We developed a novel approach to autonomous exploration through an occupancy grid map based on a direct solution to expected entropy gain. Unlike common techniques based on finding frontiers or predicting measurement scans with particle filters, the proposed approach is based on an expected value of the entropy directly. The approach is approximated with large improvements in computation and small losses in accuracy. A numerical example shows how a robot navigates an uncertain map governed by the proposed autonomous exploration algorithm.

%There are two important steps for future work. The first is to test the algorithm on a mobile robot navigating an uncertain environment while equipped with a Kinect depth sensor. The second is to extend the occupancy grid mapping and autonomous exploration to a multi-agent scenario.

\section*{Acknowledgment}
\addcontentsline{toc}{section}{Acknowledgment}
The authors would like to acknowledge Dr. Ira S. Moskowitz for helpful discussions on this topic.
%his helpful comments that served to improve the clarity of the theoretical contributions and writing of this paper.

\bibliography{BibSources}
\bibliographystyle{IEEEtran}

%\begin{thebibliography}{99}
%
%\bibliography{BibSources}
%\bibliographystyle{IEEEtran}
%
%\end{thebibliography}

% Henry Carrillo

%\begin{thebibliography}{99}
%
%\bibitem{c1}
%J.G.F. Francis, The QR Transformation I, {\it Comput. J.}, vol. 4, 1961, pp 265-271.
%
%\bibitem{c2}
%H. Kwakernaak and R. Sivan, {\it Modern Signals and Systems}, Prentice Hall, Englewood Cliffs, NJ; 1991.
%
%\bibitem{c3}
%D. Boley and R. Maier, "A Parallel QR Algorithm for the Non-Symmetric Eigenvalue Algorithm", {\it in Third SIAM Conference on Applied Linear Algebra}, Madison, WI, 1988, pp. A20.
%
%\end{thebibliography}

\end{document}

### 2018-06-09 Conference/SIMPAR 2016/CDC 2016/Related Work/BackgroundResearchSummary.tex
added Conference folder

\documentclass[10pt]{article}
\usepackage[letterpaper,text={6.5in,8.7in},centering]{geometry}
\usepackage{epic,eepic}
\usepackage{amssymb,amsmath,times,subfigure,graphicx,theorem}
%\usepackage{warmread}
%\usepackage[all,import]{xy}
%\usepackage{eepic}
\usepackage{subfigure}
\usepackage[]{algorithm2e}
\usepackage{amsmath}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Figure \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}

\renewcommand{\baselinestretch}{1.2}
\date{}

\renewcommand{\thesubsection}{\arabic{subsection}. }
\renewcommand{\thesubsubsection}{\arabic{subsection}.\arabic{subsubsection} }

\theoremstyle{plain}\theorembodyfont{\normalfont}
\newtheorem{prob}{Question}[section]
%\renewcommand{\theprob}{\arabic{section}.\arabic{prob}}
\renewcommand{\theprob}{\arabic{prob}}

\newenvironment{subprob}%
{\renewcommand{\theenumi}{\alph{enumi}}\renewcommand{\labelenumi}{(\theenumi)}\begin{enumerate}}%
{\end{enumerate}}%

\begin{document}

\section*{Background Reasearch for CDC 2016}

\subsection*{Summary Points}
\begin{itemize}
	\item Autonomous exploration governed by entropy:
	\begin{itemize}
		\item Frontier-based: move toward uncertain grid cells
		\item Expected entropy based on expected measurements: unclear how those are chosen
	\end{itemize}
	\item Other common problem: choosing movements based on map uncertainty AND localization uncertainty
\end{itemize}

\subsection*{Reference Information}
\begin{itemize}
	\item Autonomous Robotic Exploration Using Occupancy Grid Maps and Graph SLAM Based on Shannon and Renyi Entropy (CarDamKumCas15)
	\begin{itemize}
		\item Graph-based SLAM, occupancy grid
		\item tradeoff: localization and mapping
		\item The 
		\item Contribution: information-theoretic utility function: fuse uncertainties without manual tuning
		\item When identifying pose candidates, frontier-based exploration is typically used (Yam98, [2] in this paper)
		\item How do we get ``hallucinated'' measurements $\hat z$? Commonly through approximate ray-casting (no explanation)
		\item $\hat z$ obtained through ray casting ``This reduces the computational complexity by not considering all possible combinations of measurements for an action a.'' See (BurMooStaSch05, [1] in this paper) and (StaGriBur05, [3] in this paper) below.
	\end{itemize}
	\item A Frontier-Based Approach for Autonomous Exploration (Yam97)
	\begin{itemize}
		\item Define exploration: act of moving through an unknown environment while building a map that can be used for subsequent navigation
		\item Defined open, occupied, and unknown cells vaguely (no focus on the probability values)
		\item Process analogous to edge detection and region extraction in computer vision to find boundaries
		\item Any frontier above a minimum size is considered a frontier
		\item Robot goes to nearest accessible available frontier
		\item Then the robot does a $360^\text{o}$ sweep
	\end{itemize}
	\item Frontier-based Exploration Using Multiple Robots (Yam98)
	\begin{itemize}
		\item Central idea is a false assumption: ``To gain the most new information about the world, move to the boundary between open space and uncharted territory.''
		\item Paper is just about finding frontiers
	\end{itemize}
	\item Coordinated Multi-Robot Exploration (BurMooStaSch05)
	\begin{itemize}
		\item About optimizing multi-robot exploration with simultaneous coverage
		\item Moving toward frontiers: ``When exploring an unknown environment we are especially interested in frontier cells [63]. As a frontier cell we denote each already explored cell that is an immediate neighbor of an unknown, unexplored cell. If we direct a robot to such a cell, we can expect that it gains information about the unexplored area when it arrives at its target location." and "However, if there already is a robot that moves to a particular frontier cell, the utility of that cell can be expected to be lower for other robots."
		\item Not much on how we get expected measurements; only that ray-casting may be used to find where obstacles exist between frontier cells
	\end{itemize}	
	\item Information Gain-based Exploration Using Rao-Blackwellized Particle Filters (StaGriBur05)
	\begin{itemize}
		\item Rao-Blackwellized particle filter to simultaneously consider map and pose uncertainy to choose actions
		\item Consider uncertainty about the map (in paper: [2], [14], [23], [26], check: [27], [29])
		\item Not too useful because we don't use partical filters and $\hat z$ is based on that.
	\end{itemize}	
	\item Autonomous Exploration for 3D Map Learning (JohStaPfaBur07)
	\begin{itemize}
		\item About the tradeoff between poser uncertainty and expected information gain
		\item Again they look at the difference in entropy with expected measurements, but fail to explain how such expectations are calculated
	\end{itemize}	
	\item Dense Entropy Decrease Estimation for Mobile Robot Exploration (ValAnd14)
	\begin{itemize}
		\item Not very helpful
	\end{itemize}	
	\item Mapping and Exploration with Mobile Robots using Coverage Maps (StaBur03)
	\begin{itemize}
		\item ``we apply a ray-tracing technique similar to Moravec and Elfes [13] using the current maximum likelihood coverage map.''
	\end{itemize}
	\item  Probabilisitic Robotics (ThrBurFox05)
	\begin{itemize}
		\item Says that expected entropy is needed, but does not detail how this is done
		\item Totally based on particle filters
	\end{itemize}
	\item Principles of Robotic Motion (ChoLynHutKanBurKavThr05)
	\begin{itemize}
		\item Motion planning makes reference to the speed of execution (or some objective function along the way), whereas path planning does not
		\item Not one mention of entropy or information gain of the map
		\item Well-known, but not a great resource
	\end{itemize}	
	\item  ()
	\begin{itemize}
		\item 
	\end{itemize}
\end{itemize}

\end{document}

### 2018-06-09 Conference/SIMPAR 2016/SIMPAR16 Presentation.tex
added Conference folder

\documentclass[11pt,professionalfonts,hyperref={pdftex,pdfpagemode=none,pdfstartview=FitH}]{beamer}
%\usepackage{times}
%\usefonttheme{serif}
%\usepackage{helvet}
%\usepackage{amsmath,amssymb}
\usepackage{graphicx,multirow}

\usepackage{movie15}

\usepackage{caption}
\usepackage{subcaption}
\captionsetup{compatibility=false}

%\usepackage{warmread}
%\usepackage[all,import]{xy}

%\renewcommand\mathfamilydefault{\rmdefault}

% JIRS includes
%\usepackage{graphicx}
%\usepackage{amsmath,amssymb,url,times}%,subfigure}% amsthm is the one!
%\usepackage{caption,subcaption,hyperref}
%\usepackage{color,comment}
%\usepackage{curves,pgfgantt}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand{\refprop}[1]{Proposition \ref{prop:#1}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

\definecolor{mygray}{gray}{0.9}

\mode<presentation> {
  \usetheme{Warsaw}
  \usefonttheme{serif}
  \setbeamercovered{transparent}
}

\newcommand{\mypaper}{}

\setbeamertemplate{footline}%{split theme}
{%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=.9\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.1\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{title in head/foot}
    \usebeamerfont{title in head/foot}\mypaper\hfill \insertframenumber/\inserttotalframenumber
  \end{beamercolorbox}}%
  \vskip0pt%
} \setbeamercolor{box}{fg=black,bg=yellow}

\title[Autonomous Exploration by Expected Information Gain from Prob. Occupancy Grid Mapping]
{\large Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping}

\author{\vspace*{-0.3cm}}

\institute{\footnotesize
{\normalsize Evan Kaufman, Taeyoung Lee, and Zhuming Ai}\vspace*{0.2cm}\\
  Mechanical and Aerospace Engineering\\ George Washington University \vspace*{0.6cm}\\ {\normalsize Special Thanks to Ira. S. Moskowitz and Mark Livingston} \vspace*{0.2cm}\\Information Management \& Decision Architectures\\U.S. Naval Research Laboratory}

\date{}

\definecolor{tmp}{rgb}{0.804,0.941,1.0}
\setbeamercolor{numerical}{fg=black,bg=tmp}
\setbeamercolor{exact}{fg=black,bg=red}

\newtheorem{prop}{Proposition}

\renewcommand{\emph}[1]{\textit{\textbf{\color{blue}{#1}}}}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section*{}
\subsection*{Background and Motivation}

\begin{frame}
\frametitle{Motivation}
\begin{itemize}
	\item Building Maps and Exploration
	\begin{itemize}
		\item Search and rescue: quickly building a map and finding targets in places with physical or health hazards
		\item Surveillance: gather information of enemy locations
    		\item Convenience: autonomous robotic cleaning
	\end{itemize}
	\pause
	\item Goals
	\begin{itemize}
		\item Build \emph{accurate} probabilistic maps
		\item Use map properties to \emph{predict} how the \emph{map information} will change with various actions
		\item Motion planning for \emph{collision-free} trajectories
		\item Generate algorithms capable of \emph{real-time} implementation
	\end{itemize}
\end{itemize}

\only<1->{
\begin{figure}
\centerline{
    \includegraphics[height=2.1cm]{ogm_ex3.jpeg}\hspace*{0.1cm}
\hspace*{0.5cm}
    \includegraphics[height=2.1cm]{ogm_ex1.jpg}\hspace*{0.1cm}
\hspace*{0.5cm}
    \includegraphics[height=2.1cm]{ogm_ex2.png}\hspace*{0.1cm}
}
\end{figure}}
\end{frame}

\begin{frame}
\frametitle{Background: Occupancy Grid Mapping}
	\begin{itemize}
		\item Environment decomposed into evenly-space grid cells
		\item \emph{Binary} random variable: cells are either \emph{occupied} or \emph{free}
		\item Sensor reading $z$, the \emph{forward sensor model} is $p(z|m,X)$ is the measurement probability density, assuming the stochastic properties are known
		\begin{figure}[!htbp]
		\vspace*{-0.0cm}
		%\vspace*{0.25cm}
		\centerline{
	  	  \includegraphics[width=7cm]{BeamModel.png}\hspace*{0.1cm}
	    	}
		\vspace*{0.25cm}
		\centerline{
	    	\includegraphics[width=5cm]{1D_True_Grid.png}\hspace*{0.1cm}
		%    \hspace*{0.75cm}
	    	}
		{Beam Model for Range Finders}
	\end{figure}
	    	\item Goal: generate a \emph{probabilistic} map $m$ given robot pose $X$
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Background: Occupancy Grid Mapping}
	\begin{itemize}
		\item Cell occupancy probabilities obtained using the \emph{inverse sensor model}, $P(m|z,X)$, which gives grid cell probabilities based on the latest measurement $z$
		\item The exact inverse sensor model was believed to require $\mathcal O(2^n)$ for \emph{each} grid cell within the sensor field of view, so popular substitutes were \emph{heuristic} or \emph{learned} solutions
		\item Recently, the exact inverse sensor model was solved in real-time by grouping map outcomes
		\begin{figure}[!htbp]
		\vspace*{0.5cm}
		\centerline{
			\includegraphics[width=4.0cm]{rkplus_1.png}\hspace*{-0.5cm}
			\includegraphics[width=4.0cm]{rkplus_2.png}}
		 \centerline{
		\includegraphics[width=4.0cm]{rkplus_3.png}\hspace*{-0.5cm}
		\includegraphics[width=4.0cm]{rkplus_4.png}}
		\end{figure}
		\item By grouping, the inverse sensor model requires $\mathcal O(n+1)$, compared with $\mathcal O(2^n)$: applied in \emph{real-time}
	\end{itemize}
\end{frame}

% check here
%\begin{frame}
%\frametitle{Exact Occupancy Grid Mapping: Hallway Comparison}
%	\begin{itemize}
%		\item Comparison between approximate and exact results
%		\begin{figure}[!htbp]
%		    \centering
%		    \begin{subfigure}{.45\textwidth}
%		        \centering
%		        \includegraphics[width=\textwidth]{map_comparison.png}
%		        \caption*{Occupancy Grid Map}
%		    \end{subfigure}
%		    \hspace*{-.05\textwidth}
%		    \begin{subfigure}{.45\textwidth}
%		        \centering
%		        \includegraphics[width=\textwidth]{map_zoom.png}
%		        \caption*{Zoomed In Map}
%		    \end{subfigure}
%		\end{figure}
%	\end{itemize}
%\end{frame}

\begin{frame}
\frametitle{Exact Occupancy Grid Mapping: Hallway Comparison}
\begin{minipage}[t]{0.4\textwidth}
\begin{itemize}
	\item Occupancy grid mapping using an approximate and the exact solution are compared
	\item Both algorithms subject to the same pose and measurement sets 
	\item The exact algorithm yields a more \emph{certain} map shown with entropy $H$ (blue: certain, red: uncertain)
\end{itemize}
\end{minipage}
	\begin{minipage}[t]{0.5\textwidth}
		\begin{figure}[!htbp]
		    \centering
		    \begin{subfigure}{0.5\textwidth}
		        \centering
		        \includegraphics[width=.8\textwidth]{AISM_Image_19.pdf}
		        \caption*{Approx. $P(m)$}
		    \end{subfigure}
		    \hspace*{-0.1\textwidth}
		    \begin{subfigure}{0.5\textwidth}
		        \centering
		        \includegraphics[width=.8\textwidth]{EISM_Image_19.pdf}
		        \caption*{Exact $P(m)$}
		    \end{subfigure}
		\end{figure}
		\vspace*{-.1\textwidth}
		\begin{figure}[!htbp]
		    \centering
		    \begin{subfigure}{0.5\textwidth}
		        \centering
		        \includegraphics[width=.8\textwidth]{AISM_Image_inf_19.pdf}
		        \caption*{Approx. $H$}
		    \end{subfigure}
		    \hspace*{-0.1\textwidth}
		    \begin{subfigure}{0.5\textwidth}
		        \centering
		        \includegraphics[width=.8\textwidth]{EISM_Image_inf_19.pdf}
		        \caption*{Exact $H$}
		    \end{subfigure}
		\end{figure}
	\end{minipage}

%	\begin{itemize}
%		\item The exact algorithm yields a more \emph{certain} map from the same measurement set
%		\begin{figure}[!htbp]
%		    \centering
%		    \begin{subfigure}{0.5\textwidth}
%		        \centering
%		        \includegraphics[width=.5\textwidth]{AISM_Image_inf_19.pdf}
%		        \caption*{Approximate}
%		    \end{subfigure}
%		    \hspace*{-0.1\textwidth}
%		    \begin{subfigure}{0.5\textwidth}
%		        \centering
%		        \includegraphics[width=0.5\textwidth]{EISM_Image_inf_19.pdf}
%		        \caption*{Exact}
%		    \end{subfigure}
%		\end{figure}
%	\end{itemize}
\end{frame}

\section*{}
\subsection*{Autonomous Exploration}

\begin{frame}
\frametitle{Autonomous Exploration Description}
\begin{itemize}
        \item The environment is initially unknown, so the robot path is initially uncertain
	\item Simultaneous localization and mapping assumes robotic actions are provided
	\item Human interaction is common to guide the robot
	\item Goal: develop effective policies to govern robotic motion that generates a map accurately and efficiently
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Background: Autonomous Exploration}
\begin{itemize}
        	\item Uncertainty-Based Exploration
	\begin{itemize}
		\item Main idea: choose robot motions that minimize the uncertainty of the occupancy grid map
		\item Equivalently maximize map information gain
	\end{itemize}
\end{itemize}
\begin{minipage}[t]{7.0cm}
\begin{itemize}
	\item Entropy
	\begin{itemize}		
		\item \emph{Shannon's entropy} serves as an uncertainty measure
		\begin{align*}
			H(P)=-P\log P-\bar{P}\log \bar{P}%-(1-P)\log(1-P)
		\end{align*}
		with cell occupancy probability $P$ and complement $\bar{P}=1-P$
		\item Entropy $H$ is maximized when $P=0.5$ and minimized as $P\rightarrow0$ or $P\rightarrow1$
	\end{itemize}
\end{itemize}
\end{minipage}
\begin{minipage}[t]{3.0cm}
\vspace*{0.5cm}
\begin{figure}[!htbp]
	\centerline{
		\hspace*{1.25cm}
   		\includegraphics[width=5.0cm]{H_Plotted_square.pdf}%\hspace*{0.1cm}
	}
\end{figure}
\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Problem Definition}
\begin{itemize}
        	\item Map Information Gain Maximization
	\begin{itemize}
		\item Goal: choose a set of optimal future poses to maximize map information gain
		\pause
		\item Challenge: the complete set cannot be calculated accurately with an initially highly-uncertain environment
		\pause
		\item Solution: choose short-term optimal pose goals, reevaluating an updated map for the subsequent optimizations
		\pause
		\item The information gain considering candidate future pose $X_c$:% is
		\begin{align*}
			\mathcal I(X_c)&=H(P(m|X_{1:t},Z_{1:t}))-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right]
		\end{align*}
		\pause
		\item Optimal pose $X^*_c$ satisfies
		\begin{align*}
			X_c^*=\argmax_{X_c}{\mathcal I(X_c)}
		\end{align*}
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Related Work}
\begin{itemize}
        	\item Frontier-Based Exploration
	\begin{itemize}
		\item Main idea: robot moves toward boundaries between certain and uncertain space, takes measurements, then pushes back the boundaries
		\item Typically not based directly on grid cell entropy
		\item Heuristic
	\end{itemize}
	\pause
	\item Entropy-Based Exploration
	\begin{itemize}
		\item Cell occupancy probabilities are inaccurate: inverse sensor model is heuristic or learned, and frequently estimated with Rao-Blackwellized particle filters
		\item Inaccurate probabilities imply inaccurate entropy measures
		\item Approaches use ``hallucination measurements'' to predict how future measurements \emph{might} affect the map uncertainty,
		\begin{align*}
			\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right]\approx H(P(m|X_{1:t},Z_{1:t},X_c,\text{E}\left[Z_c\right]))
		\end{align*}
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Direct Entropy Calculation}
\begin{itemize}
        \item Instead of relying on frontiers or making assumptions about measurements, the expected map entropy is \emph{calculated directly} by the expected value:
	\begin{align*}
		\text{E}[H(P(m|X,z))]=\int_{z_\text{min}}^{z_\text{max}}
		H(P(m|X,z))p(z|X)dz,
	\end{align*}
	\pause
	\item The integral is discretized by intersections between the measurement and grid cells:
	\begin{align*}
		\text{E}[H(P(m|X,z))]=\sum_{k=1}^{n+1}\bigg\{H(P(m|X,z_k))P(z_{k}|X)\bigg\}
	\end{align*}
	\pause
	\item Entropy $H(P(m|X,z_k))$ is obtained with the exact inverse sensor model and definition of entropy
	\item Probability $P(z_{k}|X)$ is obtained with normalizing values from the exact inverse sensor model
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Direct Entropy Calculation: Computational Challenges}
\begin{itemize}
        \item Problem: $\text{E}[H(P(m|X,z))]$ requires $\mathcal O((n+1)^2)$ for $n+1$ possible measurement outcomes: measurement detects any of the $n$ cells or returns a maximum reading
        \pause
        \item Solution: select the $\hat n$ most probable outcomes for $\mathcal O(\hat n^2)$
        \pause
        \item Monte Carlo simulations ($100$ trials) to determine the expected entropy of a measurement ray compare the exact and approximate resulting accuracies and times
\end{itemize}

		\begin{figure}[!htbp]
		    \centering
		    \begin{subfigure}{0.5\textwidth}
		        \centering
		        \includegraphics[width=.8\textwidth]{simpar_approx_err.pdf}
		        \caption*{Error}
		    \end{subfigure}
		    \hspace*{-0.1\textwidth}
		    \begin{subfigure}{0.5\textwidth}
		        \centering
		        \includegraphics[width=.8\textwidth]{simpar_approx_t.pdf}
		        \caption*{Time Ratio}
		    \end{subfigure}
		\end{figure}

\end{frame}

\begin{frame}
\frametitle{Pose Selection}

\begin{minipage}[t]{7.0cm}
\begin{itemize}
	\item Several locations serve as candidates for the robot to move toward
	\item The optimal attitude at each location maximized expected information gain
	\item The pose (location and attitude) with the largest overall expected information gain is selected
	\item Other costs, such as distance or travel time, can be included in the optimization
\end{itemize}
\end{minipage}
\begin{minipage}[t]{3.0cm}
\vspace*{0.5cm}
\begin{figure}[!htbp]
	\centerline{
		\hspace*{1.0cm}
   		\includegraphics[width=5.0cm]{pose_selection}%\hspace*{0.1cm}
	}
\end{figure}
\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Collision-Free Path}

\begin{itemize}
	\item \emph{Waypoints}
	\begin{itemize}
		\item Using the grid map as the graph, \emph{Dijkstra's algorithm} provides collision-free waypoints for the robot path
		\item Probability of collision at pose $X$ must be within threshold $\beta>0$,
		\begin{align*}
			P_\text{collision}(X)=1-\prod_{i\in\mathcal C_X}P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})\leq\beta
		\end{align*}
	\end{itemize}
	\pause
	\item \emph{Constrained Polynomial Least Squares}
	\begin{itemize}
		\item Smooth trajectory as a function of time minimizes the squared error between the curve and the waypoints
		\item Constrained to beginning and terminal poses
		\item For complicated trajectories, splines are patched together by constraining position and velocity at the connection
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Numerical Example}
\begin{itemize}
        	\item A benchmark example floor plan is explored and mapped using the proposed mapping and exploration algorithms
	\item Simulations in Robot Operating System (ROS)
\end{itemize}
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{intel_clean.png}
    \caption*{Modified Intel Research Lab floor plan form the SLAM benchmark}
\end{figure}

\end{frame}

%% VIDEO 1: Simulation %%
\begin{frame}
%\href{run:JIRS16_simulation_speedup10x_then_speedup100x.mov}{Simulation Video Link}
\frametitle{Numerical Results}
	\centering{
	\includemovie[poster,autoplay]{\textwidth}{0.66667\textwidth}{JIRS16_simulation_speedup10x_then_speedup100x.mov}}
\end{frame}

\begin{frame}
\frametitle{Numerical Results}

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[t]{0.2\columnwidth}
        \centering
        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{0min.png}
        \caption*{$t=0$ min}
        \label{fig:IRL0min}
    \end{subfigure}
    \begin{subfigure}[t]{0.2\columnwidth}
        \centering
        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{5min.png}
        \caption*{$t=5$ min}
        \label{fig:IRL5min}
    \end{subfigure}
    \begin{subfigure}[t]{0.2\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{10min.png}
        \caption*{$t=10$ min}
        \label{fig:IRL10min}
    \end{subfigure}
    \begin{subfigure}[t]{0.2\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{15min.png}
        \caption*{$t=15$ min}
        \label{fig:IRL15min}
    \end{subfigure}
    \begin{subfigure}[t]{0.2\columnwidth}
         \centering
         \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{20min.png}
        \caption*{$t=20$ min}
        \label{fig:IRL20min}
    \end{subfigure}
    \begin{subfigure}[t]{0.2\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{25min.png}
        \caption*{$t=25$ min}
        \label{fig:IRL25min}
    \end{subfigure}
    \begin{subfigure}[t]{0.2\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{30min.png}
        \caption*{$t=30$ min}
        \label{fig:IRL30min}
    \end{subfigure}
    \begin{subfigure}[t]{0.2\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{35min.png}
        \caption*{$t=35$ min}
        \label{fig:IRL35min}
    \end{subfigure}
%    \begin{subfigure}[t]{0.2\columnwidth}
%           \centering
%           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{40min.png}
%        \caption*{$t=40$ min}
%        \label{fig:IRL40min}
%    \end{subfigure}
\vspace*{-0.25cm}
    \caption*{A robot (green) measures a room in ROS Stage simulator using modified Intel Research Lab floor plan using the proposed mapping and exploration algorithms. \href{run:JIRS16_simulation_speedup10x_then_speedup100x.mov}{(Simulation Video Link)}
}
    \label{fig:IRL}
\end{figure}
\href{run:JIRS16_simulation_speedup10x_then_speedup100x.mov}{Simulation Video Link}

\end{frame}

\begin{frame}
\frametitle{Experimental Example}
\begin{itemize}
        	\item A pioneer ground vehicle explores a room composed of Styrofoam walls
	\item Experiments use ROS, a Microsoft Kinect depth sensor, and Vicon Tracker for pose information
\end{itemize}
\begin{figure}
	\centering
    	\begin{subfigure}[b]{0.28\textwidth}
        		\includegraphics[width=\textwidth]{pioneer.png}
        		\caption*{Pioneer}
    	\end{subfigure}    	
	\hspace*{0.04\textwidth}
	\begin{subfigure}[b]{0.28\textwidth}
        		\includegraphics[width=\textwidth]{test_setup_1.jpg}
        		\caption*{Bottom-left view}
    	\end{subfigure}
	\hspace*{0.03\textwidth}
	\begin{subfigure}[b]{0.28\textwidth}
        		\includegraphics[width=\textwidth]{test_setup_2.jpg}
        		\caption*{Bottom-right view}
    	\end{subfigure}
\end{figure}

\end{frame}

%% VIDEO 2: Experiment %%
\begin{frame}
%\href{run:JIRS16_experiment_SideBySide_speedup8x.mov}{Experimental Video Link}
\frametitle{Experimental Results}
	\centering{
	\includemovie[poster,autoplay]{\textwidth}{0.5625\textwidth}{JIRS16_experiment_SideBySide_speedup8x.mov}}
\end{frame}

\begin{frame}
\frametitle{Experimental Results}

\begin{figure}
	\centering{
    	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_start.jpg}
        		\caption*{$t=0.0$min}
        		\label{fig:Experiment_ogm_t0}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_0p5min.jpg}
        		\caption*{$t=0.5$min}
        		\label{fig:Experiment_ogm_t0p5}
    	\end{subfigure}    
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1min.jpg}
        		\caption*{$t=1.0$min}
        		\label{fig:Experiment_ogm_t1}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1p5min.jpg}
        		\caption*{$t=1.5$min}
        		\label{fig:Experiment_ogm_t1p5}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2min.jpg}
        		\caption*{$t=2.0$min}
        		\label{fig:Experiment_ogm_t2}
    	\end{subfigure}
		\vspace*{0.05\textwidth}

	}
	\vspace*{-1cm}
	\centering{
    	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2p5min.jpg}
        		\caption*{$t=2.5$min}
        		\label{fig:Experiment_ogm_t2p5}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3min.jpg}
        		\caption*{$t=3.0$min}
        		\label{fig:Experiment_ogm_t3}
    	\end{subfigure}    
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3p5min.jpg}
        		\caption*{$t=3.5$min}
        		\label{fig:Experiment_ogm_t3p5}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4min.jpg}
        		\caption*{$t=4.0$min}
        		\label{fig:Experiment_ogm_t4}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4p5min.jpg}
        		\caption*{$t=4.5$min}
        		\label{fig:Experiment_ogm_t4p5}
    	\end{subfigure}
	}
\end{figure}
\vspace{-0.5cm}

\href{run:JIRS16_experiment_SideBySide_speedup8x.mov}{\tiny Experimental Video Link}

\end{frame}

% Conclusions and Future directions
%
%\section*{}
%
%\begin{frame}
%\frametitle{}
%%\center
%\center{\bf \color{blue} Future Direction:\\Multi-Robot Autonomous Exploration}
%\end{frame}
%
%\subsection*{Future Direction: Multi-Robot Autonomous Exploration}
%
%\begin{frame}
%\frametitle{Proposed Topics}
%\begin{itemize}
%        \item The ray-by-ray exact occupancy grid mapping approach in centralized probabilistic mapping
%	\item Multiple robots choosing optimal motions considering each other
%	\item Robots avoiding collisions with each other
%	\item Coverage optimization of robots from different poses
%\end{itemize}
%
%\end{frame}
%
%\begin{frame}
%\frametitle{Proposed Future Steps}
%\begin{itemize}
%        	\item Mapping:
%	\begin{itemize}
%		\item Restructure algorithms for centralized map generation
%		\item Communicate with multiple measurement sources
%	\end{itemize}
%	\item Exploration:
%	\begin{itemize}
%		\item Simplify enormous computation: bidding-based approach
%		\item Temporary theoretical map building
%		\item Collision avoidance between robots
%	\end{itemize}
%\end{itemize}
%
%\end{frame}

\section*{}
\subsection*{Conclusions}

\begin{frame}
\frametitle{Conclusions}
\begin{itemize}
	\item Map Information
	\begin{itemize}
		\item Exact occupancy grid mapping concepts are extended to map uncertainty to accomplish \emph{autonomous exploration}
		\item Shannon's entropy is calculated \emph{directly} by discretization and using terms from mapping
	\end{itemize}
	\pause
	\item Exploration
	\begin{itemize}
		\item Optimal pose selection considers both robot location and orientation
		\item Dijkstra's algorithm and constrained polynomial least squares provide a \emph{collision-free trajectory}
		\item The efficacy of the approach is verified with numerical and experimental results
	\end{itemize}
	\pause
	\item Future Steps
	\begin{itemize}
		\item 3D mapping and exploration
		\item Multi-vehicle exploration
	\end{itemize}
\end{itemize}
\end{frame}

\end{document}

### 2018-06-09 Conference/SIMPAR 2016/SIMPAR2016 (1).tex
added Conference folder

--- 2018-06-09 Conference/SIMPAR 2016/CDC 2016/CDC16.tex
@@
-\documentclass[letterpaper, 10pt]{ieeeconf}
+\documentclass[letterpaper, 10pt, conference]{ieeeconf}
 \IEEEoverridecommandlockouts \overrideIEEEmargins
@@
 \usepackage{csquotes}
-
-%\usepackage{refcheck}
-
-%\usepackage{epic,eepic}
-
-\usepackage[]{algorithm2e}
 
@@
 \title{\LARGE \bf
-Autonomous Exploration by Expected Information Gain from Probabilisitic Occupancy Grid Mapping
+Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping
 %Autonomous Exploration by Predictive Information Gain from Probabilisitic Occupancy Grid Mapping
@@
 \newcommand{\EditTL}[1]{{\color{red}\protect #1}}
+\newcommand{\EditEK}[1]{{\color{blue}\protect #1}}
 
@@
 This result is simplified with several approximations to develop an algorithm suitable for real-time implementation.
-The predicted information gain proposed in this paper, in conjunction with an existing motion planner to avoid obstacles, yield an effective autonomous exploration strategy, which is illustrated by numerical examples.
+The predicted information gain proposed in this paper governs an effective autonomous exploration strategy when applied in conjunction with an existing motion planner to avoid obstacles, which is illustrated by numerical examples.
+%The predicted information gain proposed in this paper, in conjunction with an existing motion planner to avoid obstacles, yield an effective autonomous exploration strategy, which is illustrated by numerical examples.
 \end{abstract}
@@
 Autonomous exploration solves this problem, by choosing robotic motion through uncertain space, while building a map that is used for subsequent navigation~\cite{Yam97}.
-In this paper, a robot chooses a trajectory designed to maximize information gain of the occupancy grid map.
+In this paper, robotic motion is chosen to increase knowledge about the map; a robot chooses a trajectory designed to maximize information gain of the occupancy grid map.
 
@@
 Though more accurate than several other approaches, the proposed solution requires large computational resources, making real-time implementation difficult in certain scenarios.
-Thus, we provide the tools to approximate the exact solution that reduce the computational requirements with realistic assumptions.
+Thus, we provide the tools to approximate the exact solution that reduces the computational requirements with realistic assumptions.
 The motion of the robot is formulated as an optimization problem, where the expected entropy change is minimized, or equivalently the expected information gain is maximized.
-In short, this paper presents an accurate and computationally-efficient approach to predict map information gain for autonomous exploration with a novel structure that avoids several common assumptions and approximations.
+Compared with the recent results (e.g.
+[12], [13], [16], [17]) that employ Rao-Blackwellized particle filters to deal with localization and autonomous exploration simultaneously with an approximate entropy change, this paper is based on an accurate solution to map information gain, under the assumption that the robot pose is known.
+In short, this paper presents an accurate and computationally-efficient approach to predict
+map information gain for autonomous exploration with a novel structure that avoids common assumptions of the inverse sensor model and approximations in the entropy change.
 
@@
 Put differently, the $k$-th cell is the closest occupied cell along the $l$-th ray.
-The normalizer is defined as
+The normalizer used in \refeqn{InvSenModWithProbDens} is defined as
 \begin{align}
@@
 We consider $100$ probabilistic maps to obtain Monte Carlo results to evaluate the approximate entropy.
+In every Monte Carlo trial, each grid cell has an $80\%$ chance of being free and a $20\%$ chance of receiving an a priori probability uniformly distributed between $0$ and $1$.
+%We consider $100$ probabilistic maps to obtain Monte Carlo results to evaluate the approximate entropy.
 In every Monte Carlo trial, each grid cell has an $80\%$ chance of being free, and the remaining $20\%$ of grid cells are given an a priori probability uniformly distributed between $0$ and $1$.
@@
 	\subfigure[Entropy Error]
-		{\includegraphics[height=0.35\columnwidth]{JustifyApprox_eH.pdf}
+		{\includegraphics[height=0.35\columnwidth]{simpar_approx_err.pdf}
 	}	
 	\subfigure[Computation Ratio]
-		{\includegraphics[height=0.35\columnwidth]{JustifyApprox_tRatio.pdf}
+		{\includegraphics[height=0.35\columnwidth]{simpar_approx_t.pdf}
 	}
@@
 We reduce the search by discretizing the search space in both position and attitude as follows.
-First, we choose $n_c$ candidate locations evenly-spaced along a circle of radius $\delta$, centered around the current location $x_t$, where the $c$-th candidate location is denoted $x_c\in\braces{x_1,x_2,\ldots,x_{n_c}}$.
+First, we choose $n_c$ candidate locations evenly-spaced along a circle of radius $\delta$, centered around the current location $x_t$, where the $c$-th candidate location is denoted $x_c\in\braces{x_1,x_2,\ldots,x_{n_c}}$, illustrated in Figure \ref{fig:OptProcess}.
 Any location that violates \refeqn{CollisionInequalityConstraint} is excluded to avoid collisions.
@@
 \begin{figure}
+\vspace*{0.08\columnwidth}
 \centerline{
@@
 \caption{Initially, a robot (blue circle) views down and left (blue sector).
-Then, the robot considers $n_c=8$ poses (red circles) as possible candidates for where to move next.
-For each candidate, $n_d=8$ directions are considered (red lines, only displayed on candidate location $c=2$).
+Considering $n_c=8$ poses (red circles) as possible candidate locations to move next, $n_d=8$ sensor directions for each candidate are considered (red lines, only displayed on candidate location $c=2$).
 Then, the expected entropy of each ray is calculated.
 The scan (red sector) covering those rays with the largest expected entropy decreases is chosen for each candidate, and the best candidate $X_c^*$ (black circle: location, black sector: scan) is chosen to maximize information gain.
-Finally, a wavefront planner provides the collision-free motion between $X_t$ and $X_c^*$, and the process is repeated.
+Finally, Dijkstra's algorithm provides the collision-free motion between $X_t$ and $X_c^*$, and the process is repeated.
 }
@@
 Once the condition is satisfied, the robot must move to this pose while avoiding collisions subject to the inequality constraint \refeqn{CollisionInequalityConstraint}.
+In certain robotic applications, other cost parameters might be included in the optimization, e.g., driving time.
 
-We choose the wavefront planner~\cite{ZelJarByrYut93} because it provides a relatively simple motion planning strategy from $x_t$ to $x_c^*$ that avoids local minima.
+We choose Dijkstra's algorithm~\cite{Dij59} because it provides a relatively simple motion planning strategy from $x_t$ to $x_c^*$ that avoids local minima.
 Once the robot completes this motion, the entire process is repeated.
@@
 \begin{table}
+\vspace*{0.02\columnwidth}
 \begin{tabular}{ l }
@@
   \ \ \ \ \ \ else\\
-  \ \ \ \ \ \ \ \ Obtain $X_{t+1},X_{t+2},\ldots,X_{c}^*$ using a wavefront\\  \ \ \ \ \ \ \ \ planner and minimum rotation;\\
+  \ \ \ \ \ \ \ \ Obtain $X_{t+1},X_{t+2},\ldots,X_{c}^*$ using Dijkstra's algorithm\\  \ \ \ \ \ \ \ \ planner and minimum rotation;\\
   \ \ \ \ \ \ \ \ Reset $n_c$ and $\delta$ to initial values;\\
@@
 The motion of the robot is restricted to movement on grid cells satisfying \refeqn{CollisionInequalityConstraint} with $\beta=0.01$.
-The wavefront motion planner guides the robot through the environment without collision.
+Dijkstra's algorithm for robotic motion planning guides the robot through the environment without collision.
 The results are illustrated in Figure \ref{fig:AutonomousExploration}.
@@
 The robot motion is governed by a policy that maximizes the map information gain within its set of pose candidates, where the $\hat n=6$ is chosen to approximate the expected entropy of each ray.
-The median computation time for a single update is $0.7087$sec, though the mean is $1.3263$sec.
-This discrepancy is due to high-computation outliers corresponding to cases when the robot is locally surrounded by a highly-certain environment; the robot must to traverse a large distance to learn information about another part of the map.
+When running the exploration algorithm in Robot Operating System (ROS), the mean computation time is $0.0194$ seconds to determine the optimal future pose and complete Dijkstra's algorithm on the occupancy grid.
+The computation times for this map reached roughly $1$ second at maximum, but this corresponds to cases when the robot is locally surrounded by a highly-certain environment; the robot must traverse a large distance to learn information about another part of the map.
+Such a task requires evaluating many more pose candidates and motion planning over a larger terrain.
+Thus, the efficient computation makes the proposed autonomous exploration algorithm an effective strategy for real-time implementation.
+%The median computation time for a single update is $0.7087$sec, though the mean is $1.3263$sec. 
+%This discrepancy is due to high-computation outliers corresponding to cases when the robot is locally surrounded by a highly-certain environment; the robot must to traverse a large distance to learn information about another part of the map.
 Such a task requires evaluating many more pose candidates and motion planning over a larger terrain.
@@
 If the obstacles were known a priori, the motion planning problem could be solved with a smoother path; however, the autonomous exploration is based on only the information of the map that the robot generates, so the motion planning must be reevaluated repeatedly.
-Even with these limitations, the robot explores the vast majority of reachable space in the $500$sec period.
+Even with these limitations, the robot explores the vast majority of reachable space in the $500$ second period.
 

### 2018-06-09 Conference/SIMPAR 2016/SIMPAR2016.tex
added Conference folder

--- 2018-06-09 Conference/SIMPAR 2016/SIMPAR2016 (1).tex
@@
 Occupancy grid mapping is commonly used while estimating robot location and orientation, known as simultaneous localization and mapping (SLAM).
-Several SLAM approaches are applied to a variety of problems when mobile robots traverse through an environment to generate a map, but the trajectories are assumed given (e.g.~\cite{ThrBurFox05,DurBai06,CheChe09}).
-A major problem is that an effective trajectory would be unavailable knowing little or nothing about an environment.
+Several SLAM approaches are applied to a variety of problems when mobile robots traverse an environment to generate a map, but the robot actions assumed given (e.g.~\cite{ThrBurFox05,DurBai06,CheChe09}).
+A major problem is that generating an effective trajectory for the robot to follow would be unavailable knowing little or nothing about an environment.
 Autonomous exploration solves this problem, by choosing robotic motion through uncertain space, while building a map that is used for subsequent navigation~\cite{Yam97}.
-In this paper, robotic motion is chosen to increase knowledge about the map; a robot chooses a trajectory designed to maximize information gain of the occupancy grid map.
+In this paper, robotic motion is chosen to increase knowledge about the map; a robot chooses a trajectory designed to maximize the map information gain. % of the occupancy grid map.
 
@@
 This mapping representation is typically based on heuristic approximations (e.g.~\cite{MorElf85,Elf89,ChoLynHutKanBurKavThr05,And09,PirRutBisSch11}) or simulated learning techniques \cite{ThrBurFox05,Thr01}.
-Map-based information gain is commonly determined with Shannon's entropy~\cite{CarDamKumCas15}, an uncertainty metric based on grid cell occupancy probability.
+Map-based information gain is commonly determined with Shannon's entropy~\cite{StaGriBur05}, an uncertainty metric based on grid cell occupancy probability.
 Since this probability is heuristic or learned, any measure of entropy is subject to errors from approximated probability.
-Furthermore, the expected entropy of an action is not directly calculated~\cite{CarDamKumCas15,StaGriBur05}.
-Instead, these particle filter-based approaches assume that expected entropy is equivalent to entropy based on expected measurement scans~\cite{JohStaPfaBur07}.
+Furthermore, the expected entropy of an action is not directly calculated.
+Instead, these particle filter-based approaches assume expected entropy is equivalent to entropy based on expected measurement scans~\cite{JohStaPfaBur07}.
 This assumption is subject to inaccuracies due to the nonlinear nature of Shannon's entropy.
@@
 In short, this paper presents an accurate and computationally-efficient approach to predict
-map information gain for autonomous exploration with a novel structure that avoids common assumptions of the inverse sensor model and approximations in the entropy change.
+map information gain for autonomous exploration with a novel structure that avoids common assumptions of the inverse sensor model and approximations of entropy change.
 
@@
 The problem is formulated in Section II.
-The exact solution to the expected uncertainty due to an individual 1D measurement is derived in Section III, along with approximations to the solution.
+The exact solution to the expected uncertainty due to an individual 1D measurement is derived in Section III, along with optional approximations.
 The concept of entropy is extended to 2D autonomous exploration with a numerical example in Section IV, followed by conclusions.
@@
 \end{align}
-where $\eta_{t,l}$ is a normalizing constant independent of $m$, and we have used the fact that $X_t$ carries no information about $m$ without $z_{t,l}$.
-The issue is that $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ is perceived as intractable to calculate because the exact solution considers every combination of map $m$ to obtain the normalizer $\eta_{t,l}$, and therefore this approach yields $\mathcal O(2^{n_m})$ operations per grid cell.
-This perception is why heuristic solutions based on learning or intuition substitute the inverse sensor model.
-Recently, \refeqn{BayesRuleRayISM} was solved in a computationally-efficient manner to obtain an exact solution in real-time~\cite{KauLeeAiMos16}, which is summarized next.
+where $\eta_{t,l}$ is a normalizing constant independent of $m$, and we used the fact that $X_t$ carries no information about $m$ without $z_{t,l}$.
+However, $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ is perceived as intractable because the exact solution considers every combination of map $m$ to obtain the normalizer $\eta_{t,l}$, yielding $\mathcal O(2^{n_m})$ per grid cell.
+This is why heuristic solutions based on learning or intuition substitute the inverse sensor model.
+Recently, \refeqn{BayesRuleRayISM} was solved computationally-efficiently to obtain an exact solution in real-time~\cite{KauLeeAiMos16}, summarized next.
 
@@
 Since occupancy grid mapping provides a probabilistic representation of surrounding space, this mapping scheme holds probabilistic information about the uncertainty of the map.
-Shannon's entropy is commonly used as a measure of uncertainty, such as in~\cite{CarDamKumCas15,StaGriBur05}.
+Shannon's entropy is commonly used as a measure of uncertainty, such as in~\cite{StaGriBur05}.
 Given the probabilities of each grid cell of map $m$, Shannon's entropy is defined as the sum of the entropy of each cell,
@@
 %\end{align}
-Since autonomous exploration is conducted in uncertain environments, a complete trajectory cannot be known before the robot takes measurements of the entire reachable space.
+Since autonomous exploration is conducted in uncertain environments, a complete trajectory cannot be known before a priori.
 Instead, an objective function governs the motion between intermediate poses, building an occupancy grid map along the way.
@@
 \end{align}
-where $H(P(m|X_{1:t},Z_{1:t}))$ may be computed with \refeqn{ShannonsEntropyDef} and $\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right]$ is the expected entropy when the robot is located at $X_c$, where the expectation is taken over $Z_c$.
+where $H(P(m|X_{1:t},Z_{1:t}))$ may be computed with \refeqn{ShannonsEntropyDef} and $\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right]$ is the expected entropy with robot pose $X_c$ and the expectation is taken over $Z_c$.
 Since autonomous exploration is formulated as a trajectory optimization, the goal is to repeatedly choose optimal $X_c^*$ that maximizes the objective function
@@
 \end{align}
-where $X_c^*$ satisfies the inequality constraint to avoid collisions,
+avoiding collisions by satisfying the inequality constraint
 \begin{align}
@@
 
-In this section, we present a method to determine the expected entropy from an individual measurement ray from a known pose.
-This method is based on the exact solution to occupancy grid mapping from~\cite{KauLeeAiMos16}.
+The expected entropy from an individual measurement ray from a known pose is presented, based on the exact solution to occupancy grid mapping from~\cite{KauLeeAiMos16}.
 
 \subsection{Single Ray Expected Value of Entropy}
-We consider the expected entropy due to the $l$-th ray $z_{c,l}$ at position $x_c$,
+We consider the expected entropy due to the $l$-th measurement ray $z_{c,l}$ at position $x_c$,
 \begin{align}
@@
 =\int_{z_\text{min}}^{z_\text{max}}
-H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))\nonumber\\&\qquad\times p(z_{c,l}|X_{1:t},Z_{1:t},x_c)
+H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))\nonumber\\&\qquad\quad\times p(z_{c,l}|X_{1:t},Z_{1:t},x_c)
 dz_{c,l}.
@@
 %where $z_\text{min}$ and $z_\text{max}$ are sensor limits. Considering the continuous space of possible measurements of $z_{c,l}$ would make \refeqn{HRayInt} intractable to calculate. Thus, w
-We discretize the measurement ray space such that $z_{c,l}$ falls on points along the measurement ray intersecting with grid cell edges, similar with how occupancy grid mapping decomposes continuous space into grid cells of length $\alpha$.
+We discretize the measurement ray space such that $z_{c,l}$ falls on points along the measurement ray intersecting with grid cell edges.
 The discretized expected value of \refeqn{HRayInt} is
@@
 \text{E}[H(P(m|&X_{1:t},Z_{1:t},x_c,z_{c,l}))]
-\nonumber\\&=\sum_{k=1}^{n_{r,l}+1}\bigg\{H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l,k}))\nonumber\\&\qquad\times P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)\bigg\},
+\nonumber\\&=\sum_{k=1}^{n_{r,l}+1}\bigg\{H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l,k}))\nonumber\\&\qquad\qquad\times P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)\bigg\},
 \end{align}
@@
 \end{align}
-Conveniently, this density corresponds to the inverse normalizer defined in \refeqn{BayesRuleRayISM},
+Conveniently, this density corresponds to the inverse normalizer defined in \refeqn{BayesRuleRayISM} as the denominator from Bayes' rule,
 \begin{align}
@@
 where $\eta_{c,l,k}^{-1}$ is taken from \refeqn{allEtaAtC}.
-Here, \refeqn{HRayComplete} provides the expected entropies for those cells inside the field of view of $r_l$. %; cells outside this reduced map are trivially changed by $0$ due to $z_{c,l}$.
+Here, \refeqn{HRayComplete} provides the expected entropies for cells inside the field of view of $r_l$. %; cells outside this reduced map are trivially changed by $0$ due to $z_{c,l}$.
 The entropies of cells outside the field of view remain unchanged.
@@
 The order of computation for each measurement ray is $\mathcal O((n_{r,l}+1)^2)$ since the summations of \refeqn{allEtaAtC} are embedded in \refeqn{HRayComplete}.
-Such computational requirements might be too cumbersome for some processors, especially if the measurement ray is expected to intersect many grid cells.
-Many of those intersections provide negligible information since the probability of the $l$-th measurement ray capturing certain cell depths is close to zero.
+However, several of those intersections provide negligible information since the probability of the $l$-th measurement ray capturing certain cell depths is close to zero.
 
-The approximation of expected ray entropy provides a method to reduce the computation of \refeqn{allEtaAtC} substantially.
-This goal is achieved by systematically selecting a smaller set of grid cells to consider over the summations of \refeqn{allEtaAtC} and \refeqn{HRayComplete}.
-The smaller set is determined by the probability that each cell is captured by the measurement ray, known as the detection probability. This can be found recursively as
-%These cells correspond to the $\hat n$ cells with the greatest probabilities of being captured by the measurement ray, known as the detection probability. This can be found recursively as
+The approximation of expected ray entropy provides a method to reduce the computation of \refeqn{allEtaAtC} substantially by systematically selecting a smaller set of grid cells to consider over the summations of \refeqn{allEtaAtC} and \refeqn{HRayComplete}.
+The smaller set is determined by the detection probability, or the probability that a particular cell is captured by the measurement ray, %These cells correspond to the $\hat n$ cells with the greatest probabilities of being captured by the measurement ray, known as the detection probability.
+This can be found recursively as
 %
@@
 Let $\hat n>0$ be a fixed number of grid cells for all rays such that $\hat n\leq n_{r,l}+1$ for all $l$.
-Let $\hat{r}_{l}$ correspond to the grid cells that yield the $\hat{n}$ maximum values \refeqn{ProbOfFirstCell} (the $\hat n$ most likely ray detections), indexed by increasing distance from candidate location $x_c$.
+Let $\hat{r}_{l}$ correspond to grid cells that yield the $\hat{n}$ maximum values \refeqn{ProbOfFirstCell} (the $\hat n$ most likely ray detections), indexed by increasing distance from candidate location $x_c$.
 By replacing the reduced map $r_l$ with $\hat{r}_l$ and changing the summation limits to $\braces{1,2,...,\hat n}$ in \refeqn{allEtaAtC} and \refeqn{HRayComplete}, the order of computation is reduced to $\mathcal O({\hat{n}}^2)$.
-Even though the value of $n_{r,l}$ is different among various rays in general, $\hat n$ is fixed, so the computational order is fixed as well.
+Even though $n_{r,l}$ is different among various rays in general, $\hat n$ is fixed, which fixes the computational order.
 In short, this method reduces the required computation substantially by systematically neglecting those grid cells with little effect.
+%When $\hat n=n_{r,l}+1$, the ray objective function is computed without approximation.
 %
@@
 The variable $a_\text{temp}$ is an intermediate variable designed to avoid repeated calculations.
-Since this algorithm operates as a function, fixed indices and condition variables are removed for simplification.
-It can be noted that if $\hat n=n_{r,l}+1$, the ray objective function is computed without approximation.
+Fixed indices and condition variables are removed for simplification.
 
@@
 
-The purpose of this numerical example is to provide evidence that the approximations are reasonable and speed up the algorithm substantially.
-Since a measurement ray produces a depth measurement in a single direction, we only consider a 1D map where the grid cells have spacing $\alpha=0.2$ meters, and the properties of the range sensor are based off the Microsoft Kinect~\cite{PirRutBisSch11,KhoElb12} with maximum reading depth $z_\text{max}=4$m ($20$ grid cells inside the sensor FOV).
-The goal is to compare the expected entropy $\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$ from \refeqn{HRayComplete} and with an approximation $\text{E}[H_\text{approx}(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$, which only considers $\hat n$ grid cells with highest detection probability \refeqn{ProbOfFirstCell}.
+The purpose of this numerical example is to provide evidence that the approximations are reasonable and reduce computation substantially.
+Since a measurement ray produces a depth measurement in a single direction, we only consider a 1D map where the grid cells have edge length $\alpha=0.2$ meters.
+We apply properties of the Microsoft Kinect depth sensor~\cite{PirRutBisSch11,KhoElb12} with maximum reading depth $z_\text{max}=4$m ($20$ grid cells inside the sensor FOV).
+The goal is to compare the expected entropy $\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$ from \refeqn{HRayComplete} and with an approximation $\text{E}[H_\text{approx}(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$, only considering $\hat n$ grid cells with highest detection probability \refeqn{ProbOfFirstCell}.
 
@@
 Any location that violates \refeqn{CollisionInequalityConstraint} is excluded to avoid collisions.
-For each collision-free candidate location, we consider $n_d$ evenly-spaced directions.
+For each candidate location, we consider $n_d$ evenly-spaced directions.
 The rays and attitudes at $x_c$ are denoted $\braces{z_{c,1},z_{c,2},\ldots,z_{c,n_d}}$ and $\braces{R_{c,1},R_{c,2},\ldots,R_{c,n_d}}$, respectively, where the scan with attitude $R_{c,d}$ might cover several ray directions depending on the sensor field of view (FOV).
-For each candidate location, we choose optimal attitude $R_c^*$ as the summation of the expected entropy changes covered by the scan,
+At $x_c$, we choose optimal attitude $R_c^*$ as the summation of the expected entropy changes covered by the scan,
 \begin{align}
@@
 \label{eqn:ObjFunApprox}
-\mathcal I(x_c,R_c^*(x_c))&\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*(x_c),z_{c,i}))\right]\bigg),
+\mathcal I(x_c,R_c^*)&\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\ -\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*,z_{c,i}))\right]\bigg),
 \\
-x_c^*&=\argmax_{x_c}\mathcal I(x_c,R_c^*(x_c)).
+x_c^*&=\argmax_{x_c}\mathcal I(x_c,R_c^*).
 \end{align}
@@
 
-The objective function of the optimal pose $X_c^*=\braces{x_c^*,R_c^*}$ must satisfy $\mathcal I(X_c^*)\geq\mathcal I_\text{min}$ to avoid motions where little information is expected to be gained. 
+The objective function of the optimal pose $X_c^*=\braces{x_c^*,R_c^*}$ must satisfy $\mathcal I(X_c^*)\geq\mathcal I_\text{min}$ to avoid motions where little increased information is expected. 
 If this condition is not satisfied, then $n_c$ and $\delta$ are multiplied by a scale function $\lambda>1$, candidates are chosen again, and $\mathcal I(X_c^*)$ is reevaluated.
@@
 A robot begins in a completely uncertain environment, except for the grid cells inside the immediate vicinity the robot location.
-The robot models its surroundings with an occupancy grid with $15000$ cells where grid cell edges are $\alpha=0.2$m, composing a map with dimensions $30\text{m}\times20\text{m}$.
+The robot models its surroundings with an occupancy grid with $15000$ cells where grid cell edges have length $\alpha=0.2$m, composing a map with dimensions $30\text{m}\times20\text{m}$.
 The initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) for grid cells covered by the circular robot of radius $0.1$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
@@
 Then, the $n_c=8$ evenly-spaced candidate locations about a circle of radius $\delta=0.5$m around the current pose location are considered, where $n_d=32$ measurement rays are evenly-spaced about the candidate location.
-When no current candidates yield expected information gains above $\mathcal I_\text{min}=2$, $\lambda=1.25$ is multiplied to the current $n_c$ and $\delta$ values.
-The motion of the robot is restricted to movement on grid cells satisfying \refeqn{CollisionInequalityConstraint} with $\beta=0.01$.
-Dijkstra's algorithm for robotic motion planning guides the robot through the environment without collision.
+When no current candidates yield expected information gains above $\mathcal I_\text{min}=2$, $\lambda=1.25$ is multiplied to $n_c$ and $\delta$.
+The motion of the robot is restricted to movement along cells satisfying \refeqn{CollisionInequalityConstraint} with $\beta=0.01$.
+Dijkstra's algorithm provides collision-free motion planning.
+%for robotic motion planning guides the robot through the environment without collision.
 The results are illustrated in Figure \ref{fig:AutonomousExploration}.
@@
 	\subfigure[Map at $t=50$sec]
-		{\includegraphics[height=0.35\columnwidth]{CDC16_t50.png}
+		{\includegraphics[height=0.3\columnwidth]{CDC16_t50.png}
 	}	
+	\hspace*{0.01\columnwidth}
 	\subfigure[Map at $t=100$sec]
-		{\includegraphics[height=0.35\columnwidth]{CDC16_t100.png}
+		{\includegraphics[height=0.3\columnwidth]{CDC16_t100.png}
 	}
@@
 	\subfigure[Map at $t=150$sec]
-		{\includegraphics[height=0.35\columnwidth]{CDC16_t150.png}
+		{\includegraphics[height=0.3\columnwidth]{CDC16_t150.png}
 	}	
+	\hspace*{0.01\columnwidth}
 	\subfigure[Map at $t=200$sec]
-		{\includegraphics[height=0.35\columnwidth]{CDC16_t200.png}
+		{\includegraphics[height=0.3\columnwidth]{CDC16_t200.png}
 	}
@@
 	\subfigure[Map at $t=250$sec]
-		{\includegraphics[height=0.35\columnwidth]{CDC16_t250.png}
+		{\includegraphics[height=0.3\columnwidth]{CDC16_t250.png}
 	}	
+	\hspace*{0.01\columnwidth}
 	\subfigure[Map at $t=300$sec]
-		{\includegraphics[height=0.35\columnwidth]{CDC16_t300.png}
+		{\includegraphics[height=0.3\columnwidth]{CDC16_t300.png}
 	}
@@
 	\subfigure[Map at $t=350$sec]
-		{\includegraphics[height=0.35\columnwidth]{CDC16_t350.png}
+		{\includegraphics[height=0.3\columnwidth]{CDC16_t350.png}
 	}	
+	\hspace*{0.01\columnwidth}
 	\subfigure[Map at $t=400$sec]
-		{\includegraphics[height=0.35\columnwidth]{CDC16_t400.png}
+		{\includegraphics[height=0.3\columnwidth]{CDC16_t400.png}
 	}
@@
 	\subfigure[Map at $t=450$sec]
-		{\includegraphics[height=0.35\columnwidth]{CDC16_t450.png}
+		{\includegraphics[height=0.3\columnwidth]{CDC16_t450.png}
 	}	
+	\hspace*{0.01\columnwidth}
 	\subfigure[Map at $t=500$sec]
-		{\includegraphics[height=0.35\columnwidth]{CDC16_t500.png}
+		{\includegraphics[height=0.3\columnwidth]{CDC16_t500.png}
 	}
 }
-\caption{A robot (green circle, green marker of its previous pose locations) measures a room with a Kinect depth sensor.
-The robot moves to maximize its information gain.
-In the end, the robot is returning to the left room to gain more information about the uncertain regions it originally left behind.% The algorithms update the grid cells from uncertain (grey, $0.5$ occupancy probability) to either occupied (black, $1$ occupancy probability) or free (white, $0$ occupancy probability).
+\caption{A robot (green circle, green marker of previous path) measures a room with a Kinect depth sensor to maximize map information gain.
+%The robot moves to maximize its information gain.
+%In the end, the robot is returning to the left room to gain more information about the uncertain regions it originally left behind.% The algorithms update the grid cells from uncertain (grey, $0.5$ occupancy probability) to either occupied (black, $1$ occupancy probability) or free (white, $0$ occupancy probability).
 Measured cells have occupancy probabilities closer to $0$ or $1$ as more information becomes available, and unmeasured cells remain near $0.5$, or uncertain.
@@
 When running the exploration algorithm in Robot Operating System (ROS), the mean computation time is $0.0194$ seconds to determine the optimal future pose and complete Dijkstra's algorithm on the occupancy grid.
-The computation times for this map reached roughly $1$ second at maximum, but this corresponds to cases when the robot is locally surrounded by a highly-certain environment; the robot must traverse a large distance to learn information about another part of the map.
-Such a task requires evaluating many more pose candidates and motion planning over a larger terrain.
-Thus, the efficient computation makes the proposed autonomous exploration algorithm an effective strategy for real-time implementation.
+The computation times for this map reached roughly $1$ second at maximum, corresponding to rare cases when the robot is locally surrounded by a highly-certain environment; this task requires evaluating many more pose candidates and motion planning over a larger terrain.
+%Thus, the efficient computation makes the proposed autonomous exploration algorithm an effective strategy for real-time implementation.
 %The median computation time for a single update is $0.7087$sec, though the mean is $1.3263$sec. 
@@
 
-Throughout the numerical example, the robot must make several important decisions on which actions to take.
-These actions are evaluated based on their expected information gains, not frontiers a predicted measurement scans.
-If the obstacles were known a priori, the motion planning problem could be solved with a smoother path; however, the autonomous exploration is based on only the information of the map that the robot generates, so the motion planning must be reevaluated repeatedly.
+Throughout the numerical example, the robot chooses numerous actions, based directly on their expected information gains, not frontiers or predicted measurement scans.
+If the obstacles were known a priori, the motion planning problem would yield a simpler path; however, the autonomous exploration is based on only the information of the map that the robot generates, so the motion planning must be reevaluated repeatedly.
 Even with these limitations, the robot explores the vast majority of reachable space in the $500$ second period.
@@
 We developed a novel approach to autonomous exploration through an occupancy grid map based on a direct solution to expected entropy gain.
-Unlike common techniques based on finding frontiers or predicting measurement scans with particle filters, the proposed approach is based on an expected value of the entropy directly.
+Unlike common techniques based on frontiers or predicting measurement scans, the proposed approach is based on an expected value of the entropy directly.
 The approach is approximated with large improvements in computation and small losses in accuracy.

### 2018-06-09 Conference/SIMPAR 2016/SIMPAR2016_after_reviewer_edits_before_shortening.tex
added Conference folder

--- 2018-06-09 Conference/SIMPAR 2016/SIMPAR2016 (1).tex
@@
 Occupancy grid mapping is commonly used while estimating robot location and orientation, known as simultaneous localization and mapping (SLAM).
-Several SLAM approaches are applied to a variety of problems when mobile robots traverse through an environment to generate a map, but the trajectories are assumed given (e.g.~\cite{ThrBurFox05,DurBai06,CheChe09}).
-A major problem is that an effective trajectory would be unavailable knowing little or nothing about an environment.
+Several SLAM approaches are applied to a variety of problems when mobile robots traverse an environment to generate a map, but the robot actions assumed given (e.g.~\cite{ThrBurFox05,DurBai06,CheChe09}).
+A major problem is that generating an effective trajectory for the robot to follow would be unavailable knowing little or nothing about an environment.
 Autonomous exploration solves this problem, by choosing robotic motion through uncertain space, while building a map that is used for subsequent navigation~\cite{Yam97}.
@@
 Throughout the numerical example, the robot must make several important decisions on which actions to take.
-These actions are evaluated based on their expected information gains, not frontiers a predicted measurement scans.
+These actions are evaluated based directly on their expected information gains, not frontiers or predicted measurement scans.
 If the obstacles were known a priori, the motion planning problem could be solved with a smoother path; however, the autonomous exploration is based on only the information of the map that the robot generates, so the motion planning must be reevaluated repeatedly.

### 2018-06-09 Conference/AAS 2015 DA/Presentation/Figs/figSat.m

@@ -1,0 +1,107 @@
+clear all;
+close all;
+
+a=figure('units','normalized','position',[0.1 0.1 0.7 0.7],'Color','w','renderer','opengl');
+
+
+load('topo.mat','topo','topomap1');
+[x,y,z] = sphere(100);
+props.AmbientStrength = 0.1;
+props.DiffuseStrength = 1;
+props.SpecularColorReflectance = .5;
+props.SpecularExponent = 20;
+props.SpecularStrength = 1;
+props.FaceColor= 'texture';
+props.EdgeColor = 'none';
+props.FaceLighting = 'phong';
+props.Cdata = topo;
+radE=8;
+surface(radE*x,radE*y,radE*z,props);
+
+
+N=20;
+a=linspace(0,2*pi,N)';
+vertice.upper=[sin(a) 1*ones(N,1) cos(a)];
+vertice.lower=[sin(a) -1*ones(N,1) cos(a)];
+W=1.6;
+l=5;
+d=0.1;
+vertice.plate=[-l -d W/2;
+    l -d W/2;
+    l d -W/2;
+    -l d -W/2];
+W=0.2;
+l=3;
+d=0.5;
+vertice.pin=[0 W/2 -l;
+    0 W/2 0;
+    0 -W/2 0;
+    0 -W/2 -l];
+
+vertice=[vertice.upper;vertice.lower;0 1 0;0 -1.0 0;vertice.plate;vertice.pin];
+faces=[1:N-1 1:N-1 N+1:2*N-1 2*N+3 2*N+7;
+    2:N 2:N N+2:2*N 2*N+4 2*N+8;
+    N+2:2*N (2*N+1)*ones(1,N-1) (2*N+2)*ones(1,N-1) 2*N+5 2*N+9;
+    N+1:2*N-1 (2*N+1)*ones(1,N-1) (2*N+2)*ones(1,N-1) 2*N+6 2*N+10]';
+set(gca,'Position',[0.025 0.025 0.95 0.95],'XColor',[1 1 1],'YColor',[1 1 1],'ZColor',[1 1 1]);
+light('Position',[-100 -200 20],'Style','Local');
+
+r0=14;
+orbit.x=r0*cos(linspace(0,2*pi,100));
+orbit.z=r0*sin(linspace(0,2*pi,100));
+line(orbit.x,0*orbit.x,orbit.z,'Color',[0.1 0.1 0.1]);
+
+theta=20;
[…]

### 2018-06-09 Conference/AAS 2015 OD EKF/EKF_TBP.m

@@ -1,0 +1,244 @@
+function EKF_TBP
+% Extended Kalman filter for the nonlinear relative orbit
+global mu n a
+close all;
+
+%%% Initial conditions and Parameters
+
+% Fixed parameters
+mu=398600;  % Earth gravitational parameter
+RE=6378;    % Earth radius
+
+% Orbital properties of chief
+a=7100;         % Orbital radius of the chief
+n=sqrt(mu/a^3); % Mean motion
+
+% Simulation time
+T=2*pi/n;     % Orbital period of the chief
+tf=2*T;      % Terminal simulation time
+N=2*6000+1;     % Total number of time steps
+t=linspace(0,tf,N);  % Time vector
+
+% Simulation title
+filename='EKF_IOD_0A';
+disp(['Case: ' filename]);
+
+% Initial condition
+xx0=[ -7.100116477545843e-02
+    -9.913332955272308e-04
+    1.307672689210581e-01
+    -2.329550204657181e-06
+    -9.913316565799100e-04
+    1.307672203750978e-01];
+
+% Initial estimate
+xx0_hat=[ -7.100116477545843e-02
+    -9.913332955272308e-04
+    1.307672689210581e-01
+    -2.329550204657181e-06
+    -9.913316565799100e-04
+    1.307672203750978e-01];
+
+% Covariance matrices
+P0=diag([50^2 50^2 50^2 (50*n)^2 (50*n)^2 (50*n)^2]);   % Initial covariance
+Qk=10^0*diag([1e-8 1e-8 1e-8 1e-10 1e-10 1e-10]);    % Covariance of process noise
+Rk=tand(2)^2*eye(3);    % Covariance of the measuremet
+
+nu=(randn(N,3)*chol(Rk))'; % Generate measurement errors
+mea_interval_steps=1;   % Numer of time steps between measurements
+
+%%% EKF
+
+% Initialize variables
+
+y=zeros(3,N);y_hat=zeros(3,N);y_mea=zeros(3,N);
+xx=zeros(6,N);xx_hat=zeros(6,N);
+P=zeros(6,6,N);
+
+k=1;
+xx(:,k)=xx0;            % true state vector
[…]

### 2018-06-09 Conference/AAS 2015 OD EKF/OM.m

@@ -1,0 +1,377 @@
+function OM
+% Observability criteria and measure for the nonlinear relative orbit
+global mu n a
+close all;
+
+%%% Initial conditions and Parameters
+
+% Fixed parameters
+mu=398600;      % Earth gravitational parameter
+RE=6378;        % Earth radius
+
+% Orbital properties of chief
+a=7100;
+n=sqrt(mu/a^3); % Mean motion
+
+% Simulation time
+T=2*pi/n;       % Orbital period of the chief
+tf=2*T;        % Terminal simulation time
+N=2*6000+1;    % Total number of time steps
+t=linspace(0,tf,N);  % Time vector
+
+% Simulation title
+filename='OM_IOD_0A';
+disp(['Case: ' filename]);
+
+% Initial condition
+xx0=[ -7.100116477545843e-02
+    -9.913332955272308e-04
+    1.307672689210581e-01
+    -2.329550204657181e-06
+    -9.913316565799100e-04
+    1.307672203750978e-01];
+
+
+%%% Observability Test for the Linear HCW Model
+disp('Observablity test for the linear HCW model... ');
+ode_options=odeset('RelTol',1e-6,'AbsTol',1e-6);    % Options for ode45
+[t xx_HCW]=ode45(@eom_HCW,t,xx0,ode_options);
+xx_HCW=xx_HCW';     % State vectors for the HCW model (6 x N)
+
+for k=1:N
+    O_H_svd(:,k)=ObservabilityTest_HCW(xx_HCW(:,k));
+end
+
+%%% Observability Test for the Nonlinear Model
+disp('Observablity test for the nonlinear model... ');
+[t xx_TBP]=ode45(@eom_RelTBP,t,xx0,ode_options);
[…]

### 2018-06-09 Conference/AAS 2015 OD EKF/Presentation/Figs/ObservabilityMeasure3b_ISSFD14.m

@@ -1,0 +1,451 @@
+function ObservabilityMeasure3b
+global mu n a
+close all;
+
+%%% Initial conditions and Parameters
+
+% Fixed parameters
+mu=398600;  % Earth gravitational parameter
+RE=6378;    % Earth radius
+
+% Orbital properties of chief
+a=RE+500;   % Orbital radius of the chief
+n=sqrt(mu/a^3); % Mean motion
+[r_vec_chief0, v_vec_chief0]=oe2rv(sqrt(mu*a),0,0,0,0,0); 
+
+% Simulation time
+T=2*pi/n;   % Orbital period of the chief
+tf=20*T;     % Terminal simulation time
+N=10001;     % Total number of time steps
+t=linspace(0,tf,N);  % Time vector
+
+% Initial conditions
+
+% r0=[1.5 -0.5 0.7]';     % Initial relative position
+% v0=[0.05 -2*n -0.1]';   % Initial relative velocity
+% xx0=[r0;v0];            % Initial state vector
+
+e_deputy=0.2;
+for i_deputy=50;%[0 10 20 30 40 50];%*pi/180;
+    filename=['ISSFD14_i_' num2str(i_deputy)];
+    i_deputy=i_deputy*pi/180;
+    
+    % Orbital properties of deputy
+    a_deputy=(T*sqrt(mu)/2/pi)^(2/3);
+    h_deputy=sqrt(mu*a_deputy*(1-e_deputy^2));
+    %i_deputy=0;
+    [r_vec_deputy0, v_vec_deputy0]=oe2rv(h_deputy,e_deputy,0,0,0,i_deputy);
+    
+    r0=r_vec_deputy0-r_vec_chief0;
+    v0=v_vec_deputy0-v_vec_chief0-cross([0 0 n]',r0);
+    xx0=[r0;v0];
+    
+    % % % Verification of initial conditions
+    ode_options=odeset('RelTol',1e-6,'AbsTol',1e-6);    % Options for ode45
+    [~, xx_chief]=ode45(@eom_TBP,t,[r_vec_chief0; v_vec_chief0],ode_options);
[…]

### 2018-06-09 Conference/ASC 2014/AdmissibleRegions.m

@@ -1,0 +1,58 @@
+clear all;
+close all;
+
+figure;
+[x, y, z] = sphere(1000);
+% set(gca, 'Color',[0.3 0.1 0.3]);
+hold on;
+
+patch(surf2patch(x,y,z,z), 'FaceColor','r', 'EdgeColor','r');
+view(3);
+% l = 0.5;% completely indadmissible
+l = 0.7;% non-connecting
+% l = 0.75;% connecting
+% l = 1.1;% completely admissible
+topsurf = l*[-1 1 1 -1; -1 -1 1 1; 1 1 1 1];
+bottomsurf = l*[-1 1 1 -1; -1 -1 1 1; -1 -1 -1 -1];
+backsurf = l*[1 1 1 1; -1 -1 1 1; -1 1 1 -1];
+frontsurf = l*[-1 -1 -1 -1; -1 -1 1 1; -1 1 1 -1];
+leftsurf = l*[-1 -1 1 1; 1 1 1 1; -1 1 1 -1];
+rightsurf = l*[-1 -1 1 1; -1 -1 -1 -1; -1 1 1 -1];
+patch(topsurf(1,:), topsurf(2,:), topsurf(3,:), 'b', 'EdgeColor','k');
+patch(bottomsurf(1,:), bottomsurf(2,:), bottomsurf(3,:), 'b', 'EdgeColor','k');
+patch(backsurf(1,:), backsurf(2,:), backsurf(3,:), 'b', 'EdgeColor','k');
+patch(frontsurf(1,:), frontsurf(2,:), frontsurf(3,:), 'b', 'EdgeColor','k');
+patch(leftsurf(1,:), leftsurf(2,:), leftsurf(3,:), 'b', 'EdgeColor','k');
+patch(rightsurf(1,:), rightsurf(2,:), rightsurf(3,:), 'b', 'EdgeColor','k');
+
+set(gca, 'XTick', [], 'YTick', [], 'ZTick', []);
+axis equal;
+axis off;
+
+% figure;
+% 3D Graphics: Cube
+% Dr. P.Venkataraman
+% 
+% set(gcf,'Menubar','none','Name','Cube', ...
+%     'NumberTitle','off','Position',[10,350,300,200], ...
+%     'Color',[0.3 0.1 0.3]);
+% % the cube
+% h(1) = axes('Position',[0.2 0.2 0.6 0.6]);
+% vert = [1 1 1; 1 2 1; 2 2 1; 2 1 1 ; ...
[…]

### 2018-06-09 Conference/ASC 2014/EAngles321.m

@@ -1,0 +1,15 @@
+function R = EAngles321(angles)
+phi = angles(1,1);
+theta = angles(2,1);
+psi = angles(3,1);
+R11 = cos(theta)*cos(psi);
+R12 = -cos(phi)*sin(psi)+sin(phi)*sin(theta)*cos(psi);
+R13 = sin(phi)*sin(psi)+cos(phi)*sin(theta)*cos(psi);
+R21 = cos(theta)*sin(psi);
+R22 = cos(phi)*cos(psi)+sin(phi)*sin(theta)*sin(psi);
+R23 = -sin(phi)*cos(psi)+cos(phi)*sin(theta)*sin(psi);
+R31 = -sin(theta);
+R32 = sin(psi)*cos(theta);
+R33 = cos(phi)*cos(theta);
+R = [R11 R12 R13;R21 R22 R23;R31 R32 R33];
+end

### 2018-06-09 Conference/ASC 2014/UAV_Patch.m

@@ -1,0 +1,104 @@
+function UAV_Patch
+clear all;
+close all;
+
+figure('color','w');
+view(3);
+hold on;
+%% Rods:
+nRod = 100;
+hRod = 10;
+rRod = .18;
+EAnglesRod1 = [0;3*pi/4;0];
+EAnglesRod2 = [pi;pi/4;pi/2];
+EAnglesRod3 = [pi/4;pi/2;pi/2];
+[facesRod verticesRod1] = cylrod(rRod,-hRod,hRod,nRod,EAnglesRod1);
+[facesRod verticesRod2] = cylrod(rRod,-hRod,hRod,nRod,EAnglesRod2);
+[facesRod verticesRod3] = cylrod(rRod,-hRod,hRod,nRod,EAnglesRod3);
+patch('faces',facesRod,'vertices',verticesRod1,'FaceColor',[.5 .5 .5],'EdgeColor',[.5 .5 .5]);
+patch('faces',facesRod,'vertices',verticesRod2,'FaceColor',[.5 .5 .5],'EdgeColor',[.5 .5 .5]);
+patch('faces',facesRod,'vertices',verticesRod3,'FaceColor',[.5 .5 .5],'EdgeColor',[.5 .5 .5]);
+xlabel('x')
+ylabel('y')
+zlabel('z')
+axis(5*[-10 10 -10 10 -10 10]);
+
+%% Propellers
+nProp = 100;
+hProp = .5;
+rProp = 4;
+EAnglesProp1 = [0;0;0];
+EAnglesProp2 = [0;0;0];
+EAnglesProp3 = [0;pi/2;0];
+EAnglesProp4 = [0;pi/2;0];
+EAnglesProp5 = [0;pi/2;pi/2];
+EAnglesProp6 = [0;pi/2;pi/2];
+[facesProp verticesProp1] = cyl(rProp,0,hProp,nProp,EAnglesProp1);
+verticesProp1(:,1) = verticesProp1(:,1)+hRod/sqrt(2)*ones(2*nProp+2,1);
+verticesProp1(:,2) = verticesProp1(:,2)+hRod/sqrt(2)*ones(2*nProp+2,1);
+hprop(1)=patch('faces',facesProp,'vertices',verticesProp1,'FaceColor',[0 0 1]);
+[facesProp verticesProp2] = cyl(rProp,-hProp,0,nProp,EAnglesProp2);
+verticesProp2(:,1) = verticesProp2(:,1)-hRod/sqrt(2)*ones(2*nProp+2,1);
[…]

### 2018-06-09 Conference/ASC 2014/cyl~.m

@@ -1,0 +1,16 @@
+function [faces vertices] = cyl(r,h1,h2,n,EAngles)
+theta = linspace(0,2*pi,n);
+x = r*cos(theta);
+y = r*sin(theta);
+top = transpose(EAngles321(EAngles)*[x',y',h2*ones(n,1)]');
+bottom = transpose(EAngles321(EAngles)*[x',y',h1*ones(n,1)]');
+vertices = [bottom;top;0 0 h1; 0 0 h2];
+faces = [(1:n)',(n+1:2*n)',((n+1:2*n)+1)'];
+faces(end,end) = 1;
+faces = [faces;(1:n)',(2:n+1)',((n+1:2*n)+1)'];
+faces(end,2) = 1;
+faces(end,3) = n+1;
+faces=[faces; (1:n-1)' (2:n)' (2*n+1)*ones(n-1,1)];
+faces=[faces; (n+1:2*n-1)' (n+2:2*n)' (2*n+2)*ones(n-1,1)];
+
+end

### 2018-06-09 Conference/ASC 2014/generate_figures.m

@@ -1,0 +1,28 @@
+clc
+clear all
+close all
+
+x = 0:30;
+xnew = 130*x/30+1300;
+y1 = .3*x.^2 -.2*x + 80;
+y2 = -.01*x.^3 + .4*x.^2 -10;
+ytop = 300*ones(size(x));
+ybottom = zeros(size(x));
+line = ((220-20)/30)*x + 20;
+
+
+shadedplot(xnew,y1,y2,'g')
+hold on
+shadedplot(xnew,ytop,y1,'r')
+hold on
+shadedplot(xnew,y2,ybottom,'r')
+hold on
+plot(xnew,line,'k','LineWidth',3)
+plot(xnew,y1,'b','LineWidth',2)
+plot(xnew,y2,'b','LineWidth',2)
+axis([1300 1430 0 220])
+text(1320,180,'Low Efficiency','FontSize',16)
+text(1380,30,'Low Efficiency','FontSize',16)
+text(1360,105,'Line of High Efficiency','FontSize',16)
+xlabel('Servo Command','FontSize',16)
+ylabel('BLDC Motor Command','FontSize',16)

### 2018-06-09 Journal/JAS Data Association/JAS Data Assoc Reviewer Answers.txt
journal added

Response to the Reviewers’ Comments for JASS-D-15-00036

The authors would like to thank for the reviewers for their thoughtful comments, which are aimed toward improving the quality of the paper and the clarity and impact of the results. In accordance with the comments and suggestions, the paper has been revised as follows.

Reviewer 1
Reviewer #1: “This is an excellently written paper. The literature review seems to be complete and the mathematical derivation correct. Results of the paper are backed by numerical examples demonstrating the viability of the solution.”

The authors appreciate the comments.

Reviewer 2
Reviewer #2: “As a minor editorial comment, avoid the use of "extensively" in the abstract and when referring to the simulations.  A handful of simulations cannot really be considered extensive in the context of space object tracking.”
This is a good point. The word “extensively” is replaced with “in various cases”.

Introduction, Second Paragraph:
Reviewer #2: “Regarding the sentence ‘A hard decision is when estimates are fully associated with measurements according to some metric between the estimates and the measurements.’  Unless all cases truly involve the use of a metric and all of the properties of a metric are satisfied, this sentence needs to be revised.  Additionally, the use of ‘estimates’ is somewhat confusing here.  It is not clear if this refers to estimated states or estimated measurements.  While the latter may be obviously true to a knowledgeable reader, it should still be made more clear.”

The sentence has been updated as follows: “When the association between an object of interest and a measurement is assumed correct, this is referred to as a hard decision.”

Reviewer #2: “The nearest neighbor filter does not necessarily use the Mahalanobis distance.  It could easily use a Euclidean distance.  However, as the sentence "For example, a common..." reads, the NNF requires the use of the Mahalanobis distance.”

Thank you for this point. It now reads: “For example, a common heuristic approach is known as the nearest neighbor filter (NNF), which uses a distance measure between each measurement and the predicted measurements of each potential object to determine likely associations.”

Introduction, Fourth Paragraph:
Reviewer #2: “It is stated that ‘Kalman filters are considered optimal in the sense that the estimator gain is selected to minimize the posterior uncertainties.’  This is misleading.  The Kalman gain is selected so as to minimize the posterior mean square error.  This is then related to the trace of the posterior covariance matrix.”

While it is correct that Kalman filters minimize the posterior mean square error (the length of the estimation error vector), the trace of the posterior covariance matrix is widely accepted as a measure of estimate uncertainty. This is because the diagonal elements of the covariance matrix are composed of state vector variances, so their summation is a measure of state vector uncertainty. In this sense, this uncertainty measure is minimized. The sentence is revised as follows:

“More explicitly, Kalman filters are considered optimal in the sense that the estimator gain is selected to minimize the mean square error of the estimated state, or equivalently a measure of posterior uncertainty.”

Introduction, Sixth Paragraph:
Reviewer #2: “What is meant in the first sentence when it is stated that ‘objects are in close-proximity such that neighboring objects share measurements’?  This sentence needs to be reworded.  If a graphic would aid in explaining this, please develop and use one.  Alternatively, it should be restated to read as something along the lines of multiple objects state estimates are consistent with one measurement.  In any event, this statement is unclear.”

Perhaps we attempted to say too much with this single sentence. It is now composed of three sentences as follows:

“The second part of this paper deals with a common pitfall of soft decision data association techniques, known as coalescence.
When objects are in close-proximity, a single measurement may be consistent with more than one object, forcing a soft decision.
When these objects share measurements, their updated states tend to converge toward each other.”

Introduction, Seventh Paragraph:
Reviewer #2: “What is meant by ‘measurements with a much different interpretation that just the object states’?  The entire last part of this paragraph is much too overstated, but this part absolutely makes no sense.  Unless it is justified what this means, remove it.”

The purpose of this paragraph is to introduce the second contribution of this paper (the C-JPDAF), and differentiate the contributions of this paper from prior work. We are not sure how the second part of the paragraph is overstated. The last sentence is simplified as follows:

“Therefore, we generalize the C-JPDAF algorithm in this paper to handle any number of objects and measurements, missed detections, measurements originating from extraneous clutter, and various types of measurements.”

Section 2:
Reviewer #2: “After Eq. (3), N[mu,sigma] should really have sigma as a covariance matrix since that is what appears in Eq. (3) and Eq. (4).”

This is a good comment. The sentence is updated to reflect your comment:

“…where N[mu,P] denotes the Gaussian distribution with mean mu and covariance matrix P.”

Reviewer #2: “The paragraph following Eq. (4) seems to be tied to nothing.  Revise to include more information, to have it be more thoroughly developed, or move it to a place where it belongs better.”

The paragraph is rewritten to flow better with prior paragraphs and prepare for later sections as follows:

“The Gaussian states are tracked with measurements; however, the measurement origins are unknown. Suppose that these measurements are unordered and possibly include spurious measurements and missed detections. Data association and filtering serves to estimate the state of the systems when there exist uncertainties in the system processes, the measurements, and the measurement originations. In this paper, the popular joint probabilistic data association filtering is improved in two distinct ways in the subsequent sections.”

Reviewer #2: “In the introductory paragraph, it is stated that ‘uncertainties are minimized’ and ‘JPDAF fails to minimize these uncertainties’.  The uncertainties are not minimized.  The posterior mean square error is minimized.  All instances of this throughout the paper must be fixed.”

See the comment about this issue above. The last sentence of this paragraph is rewritten as two sentences:

“As a result of the changes described in this section, the M-JPDAF serves to update state estimates from measurements in a soft decision approach that minimizes the sum of the posterior state vector variances. This a measure of state uncertainty and minimizing this measure is equivalent to minimizing the length of the estimation error vector; the conventional JPDAF fails to minimize this measure.”

Section 3.1:
Reviewer #2: “Please revise the first sentence that reads ‘During the flow update…’” and “The a priori estimated state is not found with Eq. (1).  It is found through the expected value of Eq. (1).  This is a very important distinction that needs to be made, and it is not clear at all why the governing equation for the estimated state has been omitted from the paper.”

The sentence is restructured, and the expected value is used (expected value of the noise is 0). Please see Section 3.1 of the paper for the updates, as they are rather complex for this text response.

Section 3.2:
Reviewer #2: “Prior to Eq. (8) is is stated that the ith covariance matrix represents the uncertainty.  No, it represents the covariance, not the uncertainty.  Also, given the appearance of z_j in Eq. (7), it may help to be even more explicit that Eq. (8) is effectively taking E{(e_ii)(e_ii)^{T}}.”

The covariance matrix is how we represent uncertainty of a multidimensional variable, much like a variance represents the uncertainty of a scalar. The authors agree on this point.

To the other point, we included E{(e_ii)(e_ii)^{T}} as an intermediate step inside the equation.

Reviewer #2: “Prior to Eq. (9), it is stated that G_{ij} is a simplified scaled Mahalanobis distance.  What makes it simplified?  It is also more than just scaled, as well.  It is scaled, transformed, then scaled again.  After Eq. (9), it is stated that ‘a close approximation of the probability’.  Explain the restrictions under which is a close approximation.  This should be an easy explanation.”

The writers made an error here: it is a scaled density function, not a Mahalanobis distance, and the error is fixed. Thank you for the comment. The only restriction on the approximation is (added) “where this approximation is valid when extraneous measurements and missed detections are rare.”

Reviewer #2: “Following Eq. (11), it is stated that JPDAF choses K_i as the traditional Kalman gain, ‘based on an incorrect assumption that the Kalman gain yields an optimal estimator for the JPDAF’.  Cite the sources where it is explicitly selected for this reason, or do not state that it is chosen based on the preceding assumption.”

We removed this. However, no source would explicitly state that they choose a sub-optimal gain matrix, but those sources apply their version of soft decision data association to Kalman filters or extended Kalman filters without considering how the two affect each other.

Reviewer #2: “How is Eq. (3.2) found?  Please provide details.  Traditionally, the covariance should following from analysis of the posterior error.  Under what conditions does Eq. (12) hold?  What are the uncorrelated assumptions going into this equation?”

What do you mean by Eq. (3.2)? If you are referring to Eq. (12)-(15), these come from [3], where Eq. (15) is written generally (for any K_i, not just the Kalman gain) according to the form found in [11] as it is referenced. The only conditions when Eq. (13) (formerly 12) holds is when at least one measurement may detect the i-th estimate, as well as assumptions stated earlier in the paper. We are confused by the “uncorrelated assumptions” part of this comment.

Reviewer #2: “Also following Eq. (11), there is another case of referring to minimization of the ‘posterior uncertainty’.  Once again, this is not an accurate statement.”

Please see the above responses to your comments and earlier revisions.

Reviewer #2: “Why, after Eq. (14) is the Joseph form referred to?  This comes out of nowhere and does not make any connection to any of the work.  Please revise to make this more clear.”

The term “Joseph Form” is commonly used for this general form. More on this structure can be found in the reference [11] provided there, and instead of the statement in parenthesis, we have:

“…which is commonly referred to as the Joseph form, valid for any gain matrix K_i [11].”

Reviewer #2: “Preceding Eq. (17), it is stated that the cost function is chosen to be an ‘uncertainty cost function’ ‘as with the Kalman filter’.  This is incorrect.  The Kalman filter cost function is a mean square error cost function.”

Once again, please see the above responses.

Reviewer #2: “After the proof of Proposition 1, it would be beneficial to discuss any other comparisons to the usual Kalman filter; in particular, does Eq. (15) hold for any linear gain?  This would be the analog to the Joseph form of the covariance update, but it would advantageous to give more details for the interested reader.”

Thanks for the idea. From the added portion two comments ago, Eq. (16) (formerly 15) is based on any gain. Additionally, the following sentence is added after Eq. (17):

“Because (16) is based on the Jordan form, this equation yields an accurate $P^+_i$ for any gain $K_i$.”

Section 3.2, Second to Last Paragraph:
Reviewer #2: “Once more, the cost function is a mean square error cost function.  An ‘uncertainty’ cost function is not an adequate description in this case, nor is it really appropriate.”

Once again, please see the above responses.

Section 4.1, First Paragraph:
Reviewer #2: “Matusita's measure is introduced, and reference is given to the fact that it is computed for Gaussian distributions, but the introduction here reads a bit weakly.  It would be better if a few more details regarding Matusita's measure were presented so that the reader has a firm understanding.  Additionally, why is this measure selected?  Is it possible to consider other measures?  What are the ramifications of one measure over another?”

The beginning of the paragraph is revised as follows to reflect the comments:

“The similarity between multiple multivariate Gaussian distributions can be represented by the overlapping regions of their probability density functions. Since there exists no analytic solution for this overlap, we choose Matusita's measure because it provides a scalar approximation of the overlap of multivariate Gaussian distributions [18], [19]. Furthermore, this measure is easy to compute and is easily differentiable unlike the Bhattacharyya measure [22]. More explicitly…”

Section 4.1, Prior to Eq. (23):
Reviewer #2: “It is stated that a > 0 \in R, but this could be more simply stated as a \in R^{+}.”

This detail is updated accordingly.

Section 4.2, After Eq. (23):
Reviewer #2: “It should be stated that all unique pairs are considered instead of all possible pairs since Eq. (23) explicitly excludes all of the pairings where both inputs are the same.”

Thank you for the comment. The paper is updated with “unique” pairs.

Section 4.2:
Reviewer #2: “If the paper is to be published in color, the captions to Figs. 1 and 2 are acceptable; however, it would be easy enough not to publish in color.  In this case, the captions to the aforementioned figures should be updated to avoid the use of descriptions based on color.”

We agree with that comment. We can contact the editor about this potential issue.

Section 5:
Reviewer #2: “Please review the first paragraph.  There are several poorly worded phrases here that should be revised, e.g. ‘We consider satellite the low earth orbit’.”

Reviewer #2: “ALL of the theoretical development considered linear systems, but the application is to nonlinear systems.  This is a completely illogical disconnect.  At least some mention of the fact that the linear theory is being applied to a nonlinear system, as well as a discussion regarding the possible detrimental outcomes, should be given.  This is a rather major issue that must be corrected.  There is certainly nothing inconsistent here, when one considers the application of the EKF; however, the EKF is still developed using first-order expansions, whereas no such expansions are considered in the body of this work.”

Both comments are address with major changes to the first paragraph of Section 5:

“We consider satellites in the low earth orbit (LEO) and in the geosynchronous orbit (GEO) to compare the performances of the conventional JPDAF, the M-JPDAF, and the C-JPDAF. The satellites are subject to the dynamics of the two-body problem. Since the system dynamics $\dot x=f(x)$ is nonlinear, the linearized matrix $F_{i_{k-1}}=\deriv{f}{x_i}\bigg|_{x_i=\hat x_{i_{k-1}}^+}$ is used with (5) and (6). Similarly, the measurement dynamics $z_i = h(x_i)$ may experience nonlinearities (e.g. range, range rate, bearing), the linearized matrix $H_i=\deriv{h_i}{x_i}\bigg|_{x_i=\hat x_{i}^-}$ is used with (20, (22), (27), (30), and (33).

Reviewer #2: “Table 1 and all of the subsequent references to the parameters in Table 1 are very hard to follow from the reader's perspective.  This makes it hard to understand exactly the configuration of each of the subsequent simulations without constantly referring back to the table.  Even then, it is not very clear.  Please revise this so that a reader may more logically follow the setup.”

The results of perturbed cases are not of fundamental importance on how the algorithms work generally, but they show how specific parameters affect data association and filtering performance for interested readers. The paragraph describing the scenarios and Table 1 is revised as follows:

“In the following scenarios, several cases are simulated with (35) and measurements are generated according to (38). For all cases and time steps, measurements may originate from satellites, but this origin is not given to the data association algorithm. Additionally, there is a 10% chance that a measurement is either deleted or an extraneous measurement is added at each time step. Further descriptions of the cases are available below, and the resulting accuracy is measured with RMS error, the uncertainty is summarized with $\tr{J_P}$, and the computation time is measured as well, tabulated in Tables 2-5. Scenarios A1 and B1 are data association and filtering examples in LEO, while scenarios A2 and B2 are GEO examples. For readers with interest in how the algorithms handle perturbations, various parameters are slightly changed with $\epsilon\in\Re$, summarized in Table 1. The metrics from the perturbed cases are summarized in the subsequent tables as well.”

Section 5, Preceding Eq. (37):
Reviewer #2: “It is stated that z_j \in R^4, but it is clearly not always the case.  This needs to be revised.”

Thank you for picking that up. The introduction of z_j is revised.

Section 5, Paragraph Preceding Table 2:
Reviewer #2: “The discussion about coalescence that is given in this paragraph only serves to confuse the reader.  It would be advisable to push off all of this discussion to the later sections where coalescence is actually dealt with.  As it is currently, the insights that are to be gained from the simulation results are somewhat lost.  The same goes for the caption of Fig. 3.”

That clause on coalescence is removed and only, “In some cases, the JPDAF suffers from track swapping, which is when measurements are consistently associated incorrectly” is added, where a new subsection with Monte Carlo results covers this in further detail. The Figure 3 title is revised slightly as well to avoid coalescence discussion until later as well.

Section 5.2, Scenario A.2:
Reviewer #2: “In the first paragraph of this discussion, reference is given to ‘The second set of cases’.  This is confusing since this is the third set of cases.  Please revise.”

We reordered the results before submission and missed this detail. It now reads:

“These cases serve to compare the JPDAF with the M-JPDAF…”

Reviewer #2: “It is stated that ‘Unlike the last cases, the A2 cases consider small initial covariance matrices’.  However, case B1 considered the exact same initial covariance matrix.  Is the initial covariance on Case B1 incorrect?  If not, this needs to be revised to reflect the actual differences between B1 and A2.”

Again, this was an error with reordering. The “unlike the last cases” part is removed.

Reviewer #2: “What are the units of rad given in Table 4?  It appears that they may be Earth radii, but it is very unclear.”

We changed the metric from “RMS Radial Err.” to “RMS Angular Err.” in Tables 4 and 5.

Section 5.2, Scenario B.2:
Reviewer #2: “In the paragraph preceding Table 5, it is stated that the ‘computation cost of the C-JPDAF is roughly 30 times that of either other algorithm’.  What are the implications for tracking the catalog of 20,000+ objects?  It seems that the cost may become overwhelming; however, if one takes into account the fact that coalescing objects are perhaps rare, the increase in computational complexity may be somewhat mitigated.  As it stands, the reader has the inclination to believe that the C-JPDAF algorithm will not be applicable to the full space object tracking problem, but this should be clarified in some way so that the method is not immediately ruled out.”

This is exactly correct, and a good point that we thought about as well, but did not include until now at the end of the paragraph:

“In applications with a large number of objects to track, the C-JPDAF need not be applied on most object; only those close neighbors where the similarity among the estimates yield coalescence with the JPDAF or the M-JPDAF. This way, a data association scheme with an analytic solution, such as the JPDAF or the M-JPDAF, could be applied to the remaining estimates to mitigate the large computational requirements of the C-JPDAF.”

Appendix A and B:
Reviewer #2: “Some notational liberties are taken here.  A careful explanation of these should be made so that the reader can thoroughly follow the developments that are presented.  As it is now, it is bit harder to follow that is necessary.”

A little information on the indexing of variables is included. Otherwise, the notation is consistent with the paper.

Reviewer 3
High Level Comments

Reviewer #3: “The first derivation on M-JPDAF looks good except for eqn (12).  I do not see where this equation comes from.  I am not convinced it is wrong, but this must be better developed or better cited.”

We understand the confusion. This comes from reference [3], which is better cited now.

Reviewer #3: “Eqn (22) is very close to the actual probability that two PDFs represent the same object.  This paper would be much stronger if the cost function in eqn (30-32) was tied in actual concrete probabilities rather than a heuristic.”

This is a good observation, but this measure also takes into account different covariance matrices among those Gaussian distributions. The measure is designed to compute similarity in a way that is easily differentiable. This measure satisfies these goals better than other measures, and there no known analytic solution for the similarity between many multivariate Gaussians.

Reviewer #3: “More in depth discussion of the weighting factor ‘c’ should be included.  Maybe a quick derivation of what it should be, based on a particular example?”
Reviewer #3: “More fundamentally, this problem is a multi-objective optimization which means there exists a Pareto frontier, which should be explored.  By choosing a specific c, you are making an assumption of a particular design.  This is not necessarily wrong, but definitely warrants further discussion, from a multi-objective optimization standpoint.”

Thank you for both comments. The actual calculation of an optimal “c” or “a” (in the similarity measure) are beyond the scope of this paper. And yes, there exists a Parento frontier and the choice of “c” depends on the type of scenario (for example, the B1 and B2 results have different cost function parameters). A little more explanation on this is included after eqn (25):

“The estimator gain $K_i$ is selected to minimize the above cost $\mathbf{J}$. By considering similarity to obtain the estimator gain, we mitigate some coalescence at the expense of increasing estimate uncertainty according to the magnitude of $c$. The choice of $c$ depends on the type of scenario and the parameter $a$ composing the similarity cost $J_S$.”

Reviewer #3: “The C-JPDAF is referred to as a ‘coalescence-avoiding optimal JPDAF’.   While I don't disagree that you defined a cost function and optimized it, there is no proof that this correlates to a good filter.  The cost function J_s seems to be completely heuristic.  By simply combing two cost function with a weight parameter, the new cost function in eqn (24) is generally not optimal with respect to either J_P or J_S, so calling it ‘optimal coalescence avoiding’ is not true.”

The C-JPDAF is only optimal with respect to the combined cost, where the $J_S$ component is based off of prior work on multivariate Gaussian similarity, which yields to sharing measurements, and ultimately coalescence. The authors did not see a mathematical method to show that minimizing this weighted cost function yielded a good filter, but showed its performance with numerical simulations.

Reviewer #3: “The C-JPDAF, by virtue of the cost function in eqn (24), seems to define an inherently biased estimator.  The J_S term introduces a bias.  Biases are a very bad property for an estimator.  This bias may also cause the filter to fail completely, and no discussion on how to avoid this is included.”
Reviewer #3: “The C-JPDAF as derived appears to have major flaws involved in cross tagging (aka track swapping).  Cross tagging happens when two objects become identical in the measurement space, implying inherent unobservablity.  This implies that during the unobservability, you cannot distinguish the tracks. The C-JPDAF attempts to solve this problem by adding a repulsive term to the cost function.  It seems to me that if two objects cross over each other similar to Fig (3), this repulsive term may serve to encourage cross tagging.  In general, there may be many unintended consequences of this filter, which is dangerous.”

These are good points. So too does coalescence produce a biased estimator with the JPDAF or the M-JPDAF, and there is no guarantee that cross-tagging will be avoided. The following is added as the second-to-last paragraph of Section 5 to address these issues:

“Furthermore, coalescence serves as a bias to the soft decision data association algorithms, where the bias is toward nearby tracks. The C-JPDAF also has an inherent bias by including $J_S$ from (31), except this bias is away from the nearby tracks. While there is no guarantee that these biases cancel, their opposite effects can reduce estimation error and track swapping. If (25) is poorly tuned such that $J_S$ is too strong, its repulsive nature may yield increased estimation error and track swapping.”

Reviewer #3: “There is no proof that the C-JPDAF solves the coalescence problem.  A heuristic cost function and 6 simulated test cases are not enough proof.”

We never claim that the coalescence problem is solved. This paper presents a method to mitigate some coalescence.

Reviewer #3: “Everything that is claimed as a contribution in this paper is mathematically correct (e.g. a cost function is correctly defined and optimized over). This filter is optimal with respect to a cost function and may avoid coalescence in certain test cases.  However, this filter may also fail in certain test cases, cause problems with cross tagging, and not actually solve coalescence.  If you are convinced my assessment is incorrect, then you must prove these properties mathematically.”

Your assessment is correct.

Reviewer #3: “Results section shows a few test cases and claims them as representative.  Given that the C-JPDAF is completely heuristic, simulation results should be comprehensive, to at least back up numerically that claimed properties of the filter. The work is in filtering which is fundamentally stochastic.  The results need to be run over many different test cases.  Some measure of success, most likely in terms of percent success should be shown.  When does this method break down and how does it fail? If this is to be accepted, Monte Carlo level results are needed to give evidence in support of claims.”

The C-JPDAF is not completely heuristic; it solves an optimization problem, where the cost function itself can be considered heuristic, as many are. This is contrary to other coalescence-avoiding techniques that simply remove possible associations or artificially change estimation parameters known to cause coalescence.

Regarding the Monte Carlo results, running 100 trials of each case with different noise highlighted some important points. These are included under the paragraph “Monte Carlo Results of All Scenarios”.

Technical Detail Comments

Reviewer #3: “Eqn (8), you have R_i as dependent on the object rather than dependent on the measurement.  Because you don't a priori know which objects produce which measurements, and you make no ‘hard decision’, the measurement noise can only be a function of the measurement itself, ‘R_j’”

Note that this is the “a priori innovation covariance matrix”, which has nothing to do with a particular $z_j$, but rather only the expected noise of a correctly associated measurement originating from the $i$-th estimate. The term $R_i$ is indexed with the objects because it might be different among various object estimates, e.g., an object much closer to a sensor might have a measurement with less expected noise than with an object much farther away.

Reviewer #3: “Eqn (10), the notation is very confusing.  Do not use i and j as both global markers, and summation counters in the same equation.”

Thank you for noticing this. The summation counter index is replaced with $l$.

Reviewer #3: “Eqn (11), the e_i is boldface, while up until now, no variable type has been used as bold.  Please use consistent notation.”

The boldface has been removed from this variable.

Reviewer #3: “Eqn (13) and throughout the derivation, random variables don't appear to be correctly handled. e_{ij} is a function of z_j.  z_j is a random variable which means e_{ij} is a random variable.  ~P_i should not be a random variable.”

It is true that $z_j$ and $e_{ij}$ are random variables, but so too is $\tilde P_i$ (see reference [3]), because this positive semi-definite portion of the posterior covariance matrix depends on the innovation spread composed of random variables. This term is specific to soft decision updates only, so this type of term is not found in a nearest neighbor Kalman filter, for example.

Reviewer #3: “Section 4.1: There are some discrepancies between multivariate probability distributions and Gaussian distributions.  You describe Matusita's method as a measure for any two PDFs, but then define it for specifically, Gaussian variables.  Use consistent language.”

The measure is only valid for Gaussian distributions, and both the paper text and figure is updated to reflect that.

Reviewer #3: “In section 4.1, you use the term ‘similarity measure’.  Measure is a mathematical term with a very specific meaning.  Please clarify or use different terminology.”

The term is a “measure” because it is a systematic way of assigning a number to the similarity of Gaussian random variables.

Reviewer #3: “Between Eqn (25) and (26) you make the claim that P_i^+ does not depend on any gain except K_i.  Your posterior distribution does depend on implicitly on previous gains.  Clarify or correct this statement.”

As stated at the beginning of Section 3.2, the remaining analysis takes place at the $k$-th time step. The sentences are combined as:

“…where $P^+_i$ is obtained by (13), dependent on the gain $K_i$ at the current time step.”

Reviewer #3: “I am not convinced the weighting factor, ‘a’, cannot be taken out of eqn (30) and combined with ‘c’.  Same as with ‘c’, a discussion of what ‘a’ should be must be included.”

The weighting factor $a$ is part of a product inside the exponent, so it cannot be easily removed; $a$ serves to change the size and shape of the similarity cost, whereas $c$ only affects the size. The added portion from a previous comment above discusses this briefly:

“The choice of $c$ depends on the type of scenario and the parameter $a$ shaping the similarity cost $J_S$.”

Reviewer #3: “The C-JPDAF appear to have a global minimum, but in 6 dimensional state space it will most likely be difficult to determine.  More discussion on ways to solve this problem would be useful.”

It is difficult to determine. A little more discussion on this with the Monte Carlo simulations is included per your suggestion.

Reviewer #3: “The thorough discussion of Keplarian orbital dynamics around eqn (34-35) is largely unnecessary considering the audience.”

This is true considering the majority of the audience. However, the contributions from this paper are easily extended to other data association problems, such as vision-base tracking of people or automobiles. Because other audiences do not necessarily have an aerospace background, we did not want to alienate this audience from understanding the results.

Additional AE Comments for the Author
“Please provide sufficient detail of the version of the JPDAF considered in this paper, list the assumptions about the detection probability, false measurements, measurement resolution (i.e., can one measurement originate from two objects?), and object size (i.e., can two or more measurements originate from one object?) under which that the M-JPDAF and C-JPDAF are derived, and explain why neither a missed detection model nor a false measurement model is needed for determining the association probability in equation (10).”

To address these issues, the following is revised after equation (11) (formerly 10):

“…where $B=0$ corresponds to rare extraneous measurements and missed detections. When these are prevalent, a fixed value of $B$ found from simulations was shown to perform well in various applications [1]. In this paper, we assume that at most a single measurement may originate from a single object, and missed detections and measurements originating from extraneous clutter are considered insignificant, so $B=0$ is used to determine joint association probability.”

“The statement that the JPDAF is interior to the M-JPDAF needs further analysis. Equation (12) should be re-derived. When K_i is not the optimal gain from the Kalman filter, the covariance or mean square matrix of the estimate given by equation (11) may take a different form than equation (12). For targets described by linear Gaussian systems, if the prior pdf of the state of target i is Gaussian, the posterior pdf of the state of target i is a Gaussian mixture. In the JPDAF, the Gaussian mixture is approximated by a Gaussian pdf, but the two pdfs have the same mean and covariance. Thus, the JPDAF estimate given by equation (11), with K_i the optimal gain from the Kalman filter, is the posterior mean and therefore also the minimum mean square error estimator. Why the posterior mean is worse than the C-JPDAF estimate should be explained.”

<Answer to come>

### 2018-06-09 Journal/JAS Data Association/JAS Data Association After Rev1 (1)/JAS Data Association Old/JAS Data Association NEW Derivation (1)/JAS_DataAssoc.tex
journal added

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%%!PS-Adobe-3.0 EPSF-3.0
%%%BoundingBox: 19 19 221 221
%%%CreationDate: Mon Sep 29 1997
%%%Creator: programmed by hand (JK)
%%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 linetof
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

\usepackage{bm}
\usepackage{amssymb,amsmath,times,subfigure,tabularx,booktabs,colortbl,multirow,threeparttable}
\usepackage{subfigure}
%\usepackage[notref,notcite]{showkeys}  % use this to temporarily show labels
\usepackage[colorlinks=true, pdfstartview=FitV, linkcolor=black, citecolor= black, urlcolor= black]{hyperref}
%\usepackage{overcite}
\usepackage{footnpag}			      	% make footnote symbols restart on each page

\usepackage{color}
\newcommand{\hilight}[1]{\colorbox{green}{#1}}

\newtheorem{prop}{Proposition}

\newcommand{\abs}{\mathop{\mathrm{abs}}\nolimits}
\newcommand{\diag}{\mathop{\mathrm{diag}}\nolimits}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\ave}[1]{\mathrm{E}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}

\newcommand{\x}{\ensuremath{\mathbf{x}}}
\renewcommand{\r}{\mathbf{r}}
\renewcommand{\u}{\mathbf{u}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\Ibd}{\ensuremath{\mathbf{1}_{bd}}}

\newcommand{\bfi}{\bfseries\itshape\selectfont}

\newcommand{\EditTL}[1]{{\color{red}\protect #1}}
%\renewcommand{\EditTL}[1]{{\protect #1}}

\begin{document}

\title{Minimum Uncertainty JPDA Filters and Coalescence Avoidance for Multiple Object Tracking%with Performance Evaluations
\thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Evan Kaufman         \and
		T. Alan Lovell         \and
        	Taeyoung Lee
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{E. Kaufman \at
              Dept. of Mechanical and Aerospace Engineering \\
		 The George Washington University \\
		 Washington DC 20052 \\
              Tel.: 216-346-0946\\
              \email{evankaufman@gwu.edu}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           T. A. Lovell \at
              Air Force Research Laboratory, Space Vehicles Directorate \\
		 Kirtland AFB, NM\\
		 \email{AFRL.RVSV@kirtland.af.mil}
	    \and
		 T. Lee \at
              Dept. of Mechanical and Aerospace Engineering \\
		 The George Washington University \\
		 Washington DC 20052 \\
              Tel.: 202-994-8710\\
		 Fax: 202-994-0238\\
              \email{tylee@gwu.edu}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
Two variations of the joint probabilistic data association filter (JPDAF) are derived and simulated in various cases in this paper.
First, an analytic solution for an optimal gain that minimizes posterior estimate uncertainty is derived, referred to as the minimum uncertainty JPDAF (M-JPDAF).
Second, the coalescence-avoiding optimal JPDAF (C-JPDAF) is derived, which removes coalescence by minimizing a weighted sum of the posterior uncertainty and a measure of similarity between estimated probability densities.
Both novel algorithms are tested in much further depth than any prior work to show how the algorithms perform in various scenarios.
In particular, the M-JPDAF more accurately tracks objects than the conventional JPDAF in all simulated cases.
When coalescence degrades the estimates at too great of a level, and the C-JPDAF is often superior at removing coalescence when its parameters are properly tuned.
\keywords{Data Association \and JPDAF \and Minimum Uncertainty \and Coalescence}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}

Data association has been extensively studied with various applications in estimation when measurements are not necessarily originating from a single object.
We consider applications wherein multiple objects of interest enter the field of view of a sensor; however, the measurement origins are unknown.
To properly track these objects, we require estimation techniques that associate objects with measurements.
When the objects are in very close proximity, the association uncertainties grow, and an effective algorithm is necessary to maintain close estimations of the objects~\cite{KauLovLee14}.

A variety of data association techniques are considered for close-proximity trajectories with varying performance.
Multiple hypothesis tracking (MHT) and probability hypothesis density (PHD) techniques involve a high computational load, especially when objects are in close proximity, making real-time implementation more difficult~\cite{MHT1,PHD1,PHD2}.
Many other techniques are based on the Kalman filter (KF) or the extended KF (EKF), which provide a framework for recursive data association techniques for easier real-time implementation.
There are two popular strategies among recursive approaches, namely \emph{hard decisions} and \emph{soft decisions}~\cite{JPDAF1}.
When the association between an object of interest and a measurement is assumed correct, this is referred to as a hard decision.
%A hard decision is when estimates are fully associated with measurements according to some metric between the estimates and the measurements.
For example, a common heuristic approach is known as the nearest neighbor filter (NNF), which uses a distance measure between each measurement and the predicted measurements of each potential object to determine likely associations.
%For example, a common heuristic approach is known as the nearest neighbor filter (NNF), which uses the Mahalanobis distance between each measurement and the predicted measurements of each potential object to determine likely associations.
The measurement with the smallest Mahalanobis distance is considered as the correct one~\cite{NN2}.
However, when estimates are in close proximity, there is a high likelihood of making incorrect associations, which may be detrimental to the estimated states, so this technique is avoided in the analysis of this paper.

Alternatively with a soft decision technique such as the joint probabilistic data association filter (JPDAF), measurements are shared among the estimates. With the JPDAF, the likelihoods of association between measurements and estimates serve as weighting factors inside the measurement update portion of the filtering~\cite{JPDAF1,TrackDataAssoc}.
Since all of the potential candidates for association to a track are combined in a single statistically most probable update, the JPDAF is particularly advantageous in multi-object scenarios.
This is in contrast to other hard decision approaches that may cause significant information loss in the selection process.
As such, the JPDAF has been applied to various data association problems, such as tracking unresolved object groups~\cite{SveUlmDan10} and scenarios where objects generate multiple measurements~\cite{HabThaThaMalKir13}, and it has been generalized to nonlinear non-Gaussian systems~\cite{WeiJinYou05}.

%The JPDAF is particularly advantageous in multi-object scenarios because when objects share measurements, the soft decisions prevent loss of information, which occurs with hard decision approaches.

However, the popular JPDAF approaches suffer from a common pitfall, namely sub-optimality.
More explicitly, Kalman filters are considered optimal in the sense that the estimator gain is selected to minimize the mean square error of the estimated state, or equivalently a measure of posterior uncertainty.
%More explicitly, Kalman filters are considered optimal in the sense that the estimator gain is selected to minimize the posterior uncertainties.
In the current JPDAF applications, the Kalman gain is directly used in the measurement update processes, indeliberately assuming that the Kalman gain designed for a single object tracking is also optimal for multiple object tracking problems.

%The a posteriori uncertainties of the estimates are changed because of an additional term representing the spread in the innovations; the Kalman gain serves as a close estimate to an optimal gain in the sense that in nearly minimizes this a posteriori gain as derived by the linearized equations of the JPDAF.
%The JPDAF is based on the Kalman gain, which minimizes the posterior uncertainty of individual tracks. However, the Kalman gain is not guaranteed to \emph{exactly} minimize the posterior uncertainties as they are defined for the JPDAF.

% In fact, the JPDAF is applied to multi-object tracking applications widely with the incorrect assumption that the Kalman gain yields an optimal estimator.
%The unnecessarily large uncertainty of the suboptimal Kalman gain serves as motivation to derive an optimal gain that minimizes the a posteriori uncertainty of the linearized JPDAF equations.

The first part of this paper reexamines the current derivation of the JPDAF carefully, and it shows that the Kalman gain is indeed not optimal for probabilistic data association filtering. Further, this paper derives the \textit{optimal} gain that actually minimizes the posterior uncertainties of JPDAF.  
%In this paper, we show the derivation of the optimal estimation gain that minimizes the magnitude of the posterior covariance matrix that maintains the framework of the JPDAF.
The resulting optimal JPDAF is referred to as the minimum uncertainty JPDAF or M-JPDAF in this paper, and it is shown that the optimal M-JPDAF performs substantially better than the suboptimal JPDAF through various numerical examples.

%and where this new algorithm is neither more complicated nor more computationally expensive than the conventional JPDAF. We show that decreasing estimate uncertainty to its minimum value increases the confidence of associating measurements to estimates through various numerical examples.

%The second part of this paper is focused on the cases when the objects are in close-proximity such that neighboring objects share measurements, and the estimates are subject to coalescence.
The second part of this paper deals with a common pitfall of soft decision data association techniques, known as coalescence.
When objects are in close-proximity, a single measurement may be consistent with more than one object, forcing a soft decision.
When these objects share measurements, their updated states tend to converge toward each other.
Most techniques to avoid track coalescence involve pruning the tracks, or designing a set of rules that limit the possible outcomes to likely scenarios. These rules typically involve exact nearest neighbor probabilistic data association (ENNPDA), which applies a hard decision to the most likely associated estimate. Pruning algorithms incorporate the ENNPDA algorithm because estimates do not share measurements, making it insensitive to track coalescence~\cite{Coal1}.
For example in~\cite{Fitzgerald}, the algorithm involves a set of rules to prune the tracks that share measurements.
The algorithm considers the possible combinations of estimate-to-measurement associations over a time window and applies a logic that assigns estimates to likely scenarios; only a small set of possible scenarios are considered and lower likelihood scenarios are deleted~\cite{Coal_d,Coal_e,Coal_c}.
However, track pruning assumes that low likelihood association scenarios are incorrect, which is not necessarily true.
Furthermore, removing the weighting scheme of the JPDAF causes the estimates to lose information.

Therefore, we propose a new coalescence-avoiding optimal JPDAF, referred to as the C-JPDAF, which is a systematic approach to remove coalescence by minimizing the weighted sum of the uncertainty and a similarity index between estimates. The prior work from~\cite{KauLovLee14} covers the concept in simple cases, assuming that only two objects are considered with a single measurement at each time step, and that the state and measurement are very similar. These assumptions are unrealistic in most scenarios.
%Therefore, we generalize the C-JPDAF algorithm in this paper to handle any number of objects and measurements, missed detections and measurements originating from extraneous clutter, and measurements with a much different interpretation than just the object states.
Therefore, we generalize the C-JPDAF algorithm in this paper to handle any number of objects and measurements, missed detections, measurements originating from extraneous clutter, and various types of measurements.

In short, the main contributions of this paper are (i) deriving the optimal JPDAF that minimizes the posterior uncertainties in estimation, and (ii) proposing a systematic approach to avoid data coalescence of multiple objects in close-proximity. The desirable features of the proposed approaches are illustrated by extensive numerical simulations with satellites in orbits.

This paper is organized as follows.
The problem is formulated in Section \ref{sec:PF}.
The M-JPDAF is derived in Section \ref{sec:M-JPDAF} in a similar fashion as the Kalman filter.
The coalescence problem is examined and the C-JPDAF is derived in Section \ref{sec:C-JPDAF}.
Numerous multi-object tracking numerical examples serve to compare the performances of the JPDAF, M-JPDAF, and C-JPDAF in Section \ref{sec:NE}.
Then, conclusions are summarized in Section \ref{sec:Conclusion}.

%This paper serves two major purposes: first, it provides a thorough derivation and relevant proofs for the M-JPDAF and C-JPDAF for a generalized system of estimates where measurement origins are uncertain. Second, the algorithms undergo extensive numerical simulations with orbiting satellites that provide insight into the effects of various parameters on the algorithms.

\section{Problem Formulation}
\label{sec:PF}
%This section describes the dynamics of objects and the measurement model considered in this paper and a filtering problem is formulated. 
%
%
%\subsection{Dynamics of Objects and Measurements}

%
% Linearized equations of motion: assumed known
% Assume the Gaussian noise model and define unknown random variables $w$ (process noise) and $v$ (measurement noise)
% Initial conditions are assumed known, define $\hat x_{i,0}$, $P^+_{i,0}$

%Define the set of all objects $I=\{1,2,...,t\}$ such that the object index $i\in I$ where the total number of objects $t$ is assumed known.
Assume that the number of objects being tracked $n_t$ is known.
%Consider that the number of objects being tracked $n_t$ is assumed known. 
%Consider $n_t$ objects indexed by $i\in I=\{1,2,...,n_t\}$, where \EditTL{the number of objects} $n_t$ is assumed known.
The dynamics of the $i$-th object are modeled as a linear stochastic discrete-time system described by
\begin{align}
x_{i_{k}} & = F_{i_{k-1}} x_{i_{k-1}} + w_{i_{k}},\label{eqn:xkp}\\
z_{i_k} & = H_{i_k} x_{i_k} + v_{i_k},\label{eqn:zlin}
\end{align}
where $x_{i_k}\in\Re^n$ is the state vector, $z_{i_k}\in\Re^m$ is the output vector, and  the matrices $F_{i_{k-1}}\in\Re^{n\times n}$ and $H_{i_k}\in\Re^{m\times n}$ describe system dynamics and outputs.
The process noise and the measurement noise are denoted by $w_{i_k}\in\Re^n$ and $v_{i_k}\in\Re^m$, respectively, where they are assumed to be zero-mean, mutually independent white, Gaussian noise sequences with known covariance matrices $Q_{i_k}\in\Re^{n\times n}$ and $R_{i_k}\in\Re^{m\times m}$, respectively, i.e.,
\begin{align}
w_{i_k} \sim \mathcal{N}[0,Q_{i_k}],\quad
v_{i_k} \sim \mathcal{N}[0,R_{i_k}],
\end{align}
where $\mathcal{N}[\mu,P]$ denotes the Gaussian distribution with mean $\mu$ and covariance matrix $P$.

Each initial state $\bar x_{i_0}\in\Re^n$ and covariance $P_{i_0}\in\Re^{n\times n}$ are also assumed to be mutually independent with noise. They are Gaussian with known mean values and covariance matrices:
\begin{align}
x_{i_0} \sim \mathcal{N}[\bar x_{i_0}, P_{i_0}].\label{eqn:xi0}
\end{align}

The Gaussian states are tracked with measurements; however, the measurement origins are unknown. Suppose that these measurements are unordered and possibly include spurious measurements and missed detections. Data association and filtering serves to estimate the state of the systems when there exist uncertainties in the system processes, the measurements, and the measurement originations. In this paper, the popular joint probabilistic data association filtering is improved in two distinct ways in the subsequent sections.

%Suppose that there are unordered measurements that possibly include spurious measurements and missed detections.
%Data association and filtering serves to estimate the state of the systems when there exist uncertainties in measurement origination.
%In this paper, the popular joint probabilistic data association filtering is improved in two distinct ways as follows.

\section{Minimum Uncertainty Probabilistic Data Association Filter (M-JPDAF)}
\label{sec:M-JPDAF}

In this section, we derive the M-JPDAF with a similar framework to that of the JPDAF.
Differences between the JPDAF and M-JPDAF occur in the measurement update where the gain is derived, which changes the posterior covariance equation.
%As a result of the changes described in this section, the M-JPDAF serves to update state estimates from measurements in a soft decision approach such that the state estimate uncertainties are minimized; the conventional JPDAF fails to minimize these uncertainties.
As a result of the changes described in this section, the M-JPDAF serves to update state estimates from measurements in a soft decision approach that minimizes the sum of the posterior state vector variances. This a measure of state uncertainty and minimizing this measure is equivalent to minimizing the length of the estimation error vector; the conventional JPDAF fails to minimize this measure.

\subsection{Flow Update}

The flow update serves to predict the how a system evolves between measurements. The system state obtains its $k$-th a priori state estimate $\hat x_{i_k}^-\in\Re^n$ from the posterior $(k-1)$-th state estimate $\hat x_{i_{k-1}}^+\in\Re^n$ with the expected value of \refeqn{xkp},
\begin{align}
\label{eqn:xestapriori}
\hat x_{i_{k}}^- & = F_{i_{k-1}} \hat x_{i_{k-1}}^+,
\end{align}
where $\mathrm{E}[w_{i_{k}}]=0$. The $k$-th a priori covariance matrix $P_{i_k}^-\in\Re^{n\times n}$ is calculated from the $(k-1)$-th posterior covariance matrix $P_{i_{k-1}}^+\in\Re^{n\times n}$ for all $i$ as
\begin{gather}
\label{eqn:Papriori}
%\hat x_{i_{k}}^- = f(\hat x_{i_{k-1}}^+),\\
P_{i_{k}}^- = F_{i_{k-1}} P_{i_{k-1}}^+ F_{i_{k-1}}^T + Q_{i_{k}}.
\end{gather}
%These equations provide predictions of the states and their uncertainties.
% Equations for a priori state estimates and covariances

\subsection{Measurement Update}

Suppose that there are $n_r$ measurements at the $k$-th time step where $n_r$ is finite and nonzero and may be greater or less than the number of objects $n_t$ due to extraneous clutter or missed detections.
Because all equations in this subsection take place at the $k$-th time step, this subscript is removed from the equations.
For the $i$-th object, the predicted measurement $\hat z_i\in\Re^m$ is based on the a priori state estimate $\hat x_{i}^-$ according to \refeqn{zlin},
%\footnote{Note that for nonlinear systems where $z_i = h_i(x_i)$, then $\hat z_i = h_i(\hat x_{i}^-)$ replaces \refeqn{zlin} and the linearized matrix $H_i=\deriv{h_i}{x_i}\bigg|_{x_i=\hat x_{i}^-}$ is used for the remaining equations.\label{fn:nonlinsys}}
\begin{align}
\label{eqn:zhatlin}
\hat z_i = H_i\hat x_{i}^-.
\end{align}
The innovation, often referred to as the residual, $e_{i,j}\in\Re^m$ is defined as the difference between the $j$-th measurement and the predicted measurement of the $i$-th object, and it is given by
\begin{align}
e_{i,j} = z_j - \hat z_i.\label{eqn:eij}
\end{align}
%where the true measurement is expected to be
%\begin{align}
%z_i=h(x_i)+v_i, \quad H_i=\frac{\partial h}{\partial x}\bigg|_{x=\hat x_{i}^- }.
%\end{align}
%Given the uncertainty of the measurements represented with covariance matrix $R_i$, t
For the $j$-th measurement, the $i$-th a priori innovation covariance matrix $S_{i,j}\in\Re^{m\times m}$ represents the uncertainty between the $i$-th measurement estimate and its true measurement, defined as
%The a priori innovation covariance matrix $S_{i}\in\Re^{m\times m}$ for all innovations corresponding to the $i$-th object is defined as \EditTL{(Is it for the covariance of $e_{ii}$? Not sure.. What is the meaning of `the covariance matrix for all innovations...')}
\begin{align}
S_{i,j}&=\mathrm{E}[e_{i,j} e_{i,j}^T]=\mathrm{E}[\left((H_ix_i+v_i)-H_i\hat x_{i}^-\right)\left((H_ix_i+v_i)-H_i\hat x_{i}^-\right)^T]\nonumber\\
&=H_{i}P_{i}^{-}H_{i}^T+R_{j},\label{eqn:S}
\end{align}
where it is commonly assumed that $R_j$ is uniform for all measurements, but the subsequent equations make no such assumption.

% Likelihoods of association are calculated from the innovations and innovation covariances only
% Two techniques are used to obtain the likelihoods of association, both of which may be applied to the JPDAF, M-JPDAF, and C-JPDAF
	% "Cheap JPDAF": inexpensive calculation, present the four equations and two references
	% True likelihood of association: infeasible for many real-time applications, calculated in the appendix

%Define the set $J=\{1,2,...,r\}$ such that the measurement index $j\in J$ where $r$ is a variable number of measurement returns such that $r\geq1$.
%\paragraph{Measurement Origination Model}\

%Suppose that there are $n_r$ measurements indexed by $j\in J=\{1,2,...,n_r\}$ \EditTL{at each instant. The number of measurement $n_r$ is assumed to be varied, but it is less than $n_t$???.

Next, we define a measurement origination model. Let $\beta_{i,j}\in\Re$ be the probability that the $j$-th measurement originates from the $i$-th object. 
%Let $\beta_{i,j}$ be the probability that the $i$-th system is associated with measurement $j$.
%Two techniques may be applied to solve for every $\beta_{i,j}$, both of which require the innovation and its covariance (defined later) and information on detection probabilities, which are assumed known. The first technique is applied in the standard JPDAF that calculates the true association probabilities from~\cite[Eq. 9.(45-46)]{TrackDataAssoc}. However, the computation required with a standard JPDAF is inconsistent with the capabilities of the computational resources to account for an arbitrary number of objects to be tracked~\cite{Bar1990}.
Several approaches have been proposed to approximate the association probability $\beta_{i,j}$ for the JPDAF.
This paper adopts the expressions given in~\cite{Bar1990}, and they are summarized as follows.

% [Eq. 1.(1-2)]

%A modified version of the JPDA, commonly known as the ``Cheap JPDA'', provides a simpler approximation of the likelihood $\beta_{i,j}$.
%The equations below are adopted from~\cite[Eq. 1.(1-2)]{Bar1990}.

Let $G_{i,j}\in\Re$ be a simplified scaled density function between the $i$-th measurement estimate and the $j$-th measurement,
% variable which takes a very similar form as the Gaussian probability density function
\begin{align}
G_{i,j}=\frac1{\sqrt{\det{S_{i,j}}}}\exp{\{-\frac12e_{i,j}^TS_{i,j}^{-1}e_{i,j}\}}.
\end{align}
%This is used to calculate intermediate variables $\mathcal{S}_{ti}$ and $\mathcal{S}_{ri}$
%\begin{align}
%\mathcal{S}_{ti}=\sum\limits_{j\in J}G_{i,j}, 
%\quad 
%\mathcal{S}_{ri}=\sum\limits_{i\in I}G_{i,j}.
%\end{align}
Then a close approximation of the probability that the $i$-th system is associated with the $j$-th measurement is
\begin{align}
\label{eqn:betaDef}
\beta_{i,j}=\frac{G_{i,j}}{\sum\limits_{l=1}^{n_r}G_{il}+\sum\limits_{l=1}^{n_t}G_{lj}-G_{i,j}+B},
\end{align}
where $B=0$ corresponds to rare extraneous measurements and missed detections. When these are prevalent, a fixed value of $B$ found from simulations was shown to perform well in various applications~\cite{Bar1990}. In this paper, we assume that at most a single measurement may originate from a single object, and missed detections and measurements originating from extraneous clutter are considered insignificant, so $B=0$ is used to determine joint association probability.

%this approximation is valid when extraneous measurements and missed detections are rare. Even if missed detections and measurements originating from extraneous clutter is present, \refeqn{betaDef} performs well~\cite{Bar1990}.

%\EditTL{(The above writing has several issues that the reviewers of ACC15 mentioned. (i) There is no indication on the dimension of variables. For example is $G_{i,j}$ a scalar or a vector; (ii) there are several variables, such as $e_{i,j}$ or $S_i$ that were used before defining them explicitly. Saying that it is the innovation and the a priori innovation covariance is not sufficient. Ideally, we have to define a variable before using it; (iii) The definition of $\mathcal{S}_{ti}$ and $\mathcal{S}_{ri}$ is unclear. Is $t$ also an indexed variable?; (iv) Do you really have to introduce the set $I$ and $J$. It is causing more confusion as several other variations of $J$ will appear later in this paper. Would it be more concise to rewrite it as $\sum_{j=1}^{n_r}$?; (v) Do we really have to introduce the variables $\mathcal{S}_{ti}$ and $\mathcal{S}_{ri}$. If they are used here only, there is no need to define them. It's better to substitute (6) into (7). Usually, a new variable or symbol is defined, only if it has an important meaning, or it is used repeatedly later; (vi) how is $B$ defined, and what's the dimension of it. Give the explicit definition, or at least cite something. The given explanation raises more questions. }

The weighted sum of all innovations associated with the object $i$ is defined as ${\bar{e}}_{i}={\sum\limits_{j=1}^{n_r} \beta_{i,j}e_{i,j}}\in\Re^m$.
The innovations serve to update the resulting posterior state estimate $\hat x^+_{i}$ with the gain matrix $K_i\in\Re^{n\times m}$~\cite{TrackDataAssoc},
%~\cite[Eq. 6-(25-26)]{TrackDataAssoc}
\begin{align}
\hat x^+_{i}= \hat x^-_{i}+K_{i}{\bar{e}}_{i}.\label{eqn:KalEst}
\end{align}
The conventional JPDAF chooses $K_i=P^-_iH_i^TS_i^{-1}$ (assuming $S_i$ is fixed for all measurements) as the gain from the Kalman filter.
In this paper, we show that the Kalman gain does not actually minimize the posterior uncertainty in probabilistic data association filtering, and we derive the new \textit{optimal} gain $K_i$.

Let the probability that $i$-th object is \textit{not} associated with any measurement be $\beta_{i,0}=1-{\sum_{j=1}^{n_r} \beta_{i,j}}\in\Re$. The case when $\beta_{i,0}=1$ implies that the $i$-th estimate is \emph{not} associated with any measurement completely, i.e., the $i$-th object is not measured by any sensor with the probability of one. In such trivial cases, there is no measurement update, i.e., $\hat x^+_i=\hat x^-_i$.

Next we consider when $\beta_{i,0}\neq1$, i.e., $0\leq\beta_{i,0}<1$.
Let the a priori and a posteriori errors be defined as
\begin{align}
\label{eqn:ErrFuns}
\tilde x_i^-=\hat x_i^--x_i,
\quad
\tilde x_i^+=\hat x_i^+-x_i,
\end{align}
respectively.
Then, using \refeqn{zlin}, \refeqn{zhatlin}, \refeqn{eij}, \refeqn{ErrFuns}, and the difference between \refeqn{KalEst} and \refeqn{xestapriori}, we have
\begin{align}
\tilde x_i^+&=
%\hat x_i^+-\hat x_i^-+\tilde x_i^-
%\nonumber\\&=
K_i\sum_{j=1}^{n_r}\beta_{i,j}(z_j-\hat z_i)+\tilde x_i^-
\nonumber\\&=
%K_i\sum_{j=1}^{n_r}\beta_{i,j}(H_ix_i+v_j-H_i\hat x_i^-)+\tilde x_i^-
%\nonumber\\&=
%-K_iH_i\tilde x_i^-\sum_{j=1}^{n_r}\beta_{i,j}+K_i\sum_{j=1}^{n_r}\beta_{i,j}v_j+\tilde x_i^-
%\nonumber\\&=
(I-(1-\beta_{i,0})K_iH_i)\tilde x_i^-+K_i\sum_{j=1}^{n_r}\beta_{i,j}v_j.
\end{align}

The resulting posterior covariance $P^+_{i}\in\Re^{n\times n}$ is defined as
%~\cite[Eq. D-32 and D-33]{TrackDataAssoc}
\begin{align}
\label{eqn:JPDAFCov}
P_i^+&=\text{E}[\tilde x_i^+\tilde x_i^{+T}]
%\nonumber\\&=
%\text{E}[
%\{(I-(1-\beta_{i,0})K_iH_i)\tilde x_i^-+K_i\sum_{j=1}^{n_r}\beta_{i,j}v_j\}\{(I-(1-\beta_{i,0})K_iH_i)\tilde x_i^-+K_i\sum_{j=1}^{n_r}\beta_{i,j}v_j\}^T
%]
%\nonumber\\&=
%(I-(1-\beta_{i,0})K_iH_i)\text{E}[\tilde x_i^-\tilde x_i^{-T}](I-(1-\beta_{i,0})K_iH_i)^T
%+
%K_i\text{E}[\sum_{j=1}^{n_r}\beta_{i,j}v_j\sum_{j=1}^{n_r}\beta_{i,j}v_j^T]K_i^T
\nonumber\\&=
(I-(1-\beta_{i,0})K_iH_i)P_i^-(I-(1-\beta_{i,0})K_iH_i)^T
+
K_i(\sum_{j=1}^{n_r}\beta^2_{i,j}R_j)K_i^T
\nonumber\\&=
P_i^--(1-\beta_{i,0})
(K_iH_iP_i^-+P_i^-H_i^TK_i^T)
+
K_iE_iK_i^T.
\end{align}
where $E_i\in\Re^{m\times m}$ is defined for convenience as
\begin{align}
\label{eqn:E}
E_i=\sum_{j=1}^{n_r}\beta^2_{i,j}R_j+(1-\beta_{i,0})^2H_iP_i^-H_i^T.
\end{align}
%$\tilde P_i\in\Re^{n\times n}$ is the portion of the $i$-th state covariance matrix due to the spread in the innovations
%\begin{align}
%\tilde P_i=&\ K_i\parenth{{\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T}K_{i}^T,\label{eqn:tildeP}
%\end{align}
%and $P_i^c\in\Re^{n\times n}$ corresponds to the $i$-th posterior covariance if only a single measurement serves to update the $i$-th estimate~\cite{TrackDataAssoc}%, and it can be written as
%%We can define $P_i^c$ without any assumptions on $K_i$ (i.e., $K_i$ is not necessarily the Kalman gain, so the well-known Kalman covariance update equations are not valid here) as~\cite[Eq. 4.2-12]{OptEst1}
%\begin{align}
%P^c_i=&\ (I-K_iH_i)P^-_i(I-K_iH_i)^T+K_iR_iK_i^T\nonumber
%\\
%=&\ P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i(H_iP^-_iH_i^T+R_i)K_i^T,
%\label{eqn:GenCov}
%\end{align}
%which is commonly referred to as the \emph{Joseph form}, valid for \emph{any} gain matrix $K_i$~\cite{OptEst1}.
%Substituting \refeqn{tildeP} and \refeqn{GenCov} into \refeqn{JPDAFCov},% and rearranging,
%\begin{align}
%P^+_{i}=&\ \beta_{i,0}P^-_{i}+(1-\beta_{i,0})[P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\nonumber
%\\
%&\ +K_i(H_iP^-_iH_i^T+R_i)K_i^T]+K_i\left({\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right)K_{i}^T\nonumber
%\\
%%=&\ P^-_{i}-(1-\beta_{i,0})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i\left((1-\beta_{i,0})S_i+{\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right)K_i^T\nonumber%asdf
%%\\
%=&\ P^-_{i}-(1-\beta_{i,0})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_iE_iK_i^T,\label{eqn:CovExpanded}
%\end{align}
%where the symmetric matrix $E_i=E_i^T\in\Re^{m\times m}$ is defined as
%\begin{align}
%E_i=(1-\beta_{i,0})S_i+{\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T,\label{eqn:E}
%\end{align}
%and it corresponds to an innovation covariance that accounts for the possibility that more than a single measurement may update the $i$-th estimate.
Since \refeqn{JPDAFCov} and \refeqn{E} make no assumptions on $K_i$, this form is valid for \emph{any} gain $K_i$.

As with the Kalman filter, we choose the uncertainty cost function $J_{P_i}\in\Re$ to be the trace of the posterior covariance,
\begin{align}
\label{eqn:JpIntro}
J_{P_i}=\tr{P_i^+}.
\end{align}
We choose the gain to minimize $J_{p_i}$. The necessary condition for a local minimum is given by
\begin{align}
0=\frac{\partial\tr{P^+_i}}{\partial K_i}=-2(1-\beta_{i,0})P^-_iH_i^T+2K_iE_i.\label{eqn:Uncertainty0}
\end{align}
Therefore the optimal gain is selected as
\begin{align}
K_i&=(1-\beta_{i,0})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
\end{align}
%As with the Kalman filter, we choose the estimation gain to minimize the magnitude of the posterior covariance measured by $\tr{P_i^+}$.
%The first derivative necessary condition for optimality requires that
%\begin{align}
%0&=\frac{\partial\tr{P^+_i}}{\partial K_i},\nonumber
%\\
%&=-2(1-\beta_{i,0})\frac{\partial\tr{K_iH_iP^-_i}}{\partial K_i}+\frac{\partial\tr{K_iE_iK_i^T}}{\partial K_i},\nonumber
%\\
%&=-2(1-\beta_{i,0})P^-_iH_i^T+2K_iE_i.\label{eqn:Uncertainty0}
%\end{align}
%%Therefore the optimal gain is given by
%%\begin{align}
%%K_i&=(1-\beta_{i,0})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
%%\end{align}
%Solving \refeqn{Uncertainty0} to satisfy necessary condition for optimality,
%\begin{align}
%K_i&=(1-\beta_{i,0})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
%\end{align}
%The Hessian

Finally, \refeqn{GainK} is substituted into \refeqn{JPDAFCov} to obtain a simplified posterior state covariance matrix
%\begin{align}
%\label{eqn:PNotSimplified}
%P^+_{i}&=P^-_{i}-(1-\beta_{i,0})^2P^-_iH_i^TE_i^{-1}H_iP^-_i,
%\end{align}
%which may be simplified in two ways
\begin{align}
P^+_{i}&=P^-_{i}-K_iE_iK_i^T, \quad \mbox{or equivalently}\label{eqn:JPDAFPostCovSymmetricUpdate}
\\
P^+_{i}&=\left(I-(1-\beta_{i,0})K_iH_i\right)P^-_i,\label{eqn:JPDAFPostCovSimpleUpdate}
\end{align}
where these can \emph{only} be used with \refeqn{GainK}.
Several properties of the proposed optimal gain are summarized as follows.
\begin{prop}
The optimal gain designed in \refeqn{GainK} satisfies the following properties.
\begin{itemize}
\item[(i)] The gain $K_i$ globally minimizes the cost \refeqn{JpIntro}, and the corresponding minimum value is given by
\begin{align*}
J^*_{P_i} = \tr{P^-_{i}-K_iE_iK_i^T}.
\end{align*}
\item[(ii)] The matrix $E_i$ given in \refeqn{E} is invertible, or equivalently, the gain $K_i$ is well-defined, except when $\beta_{i,0}=1$, corresponding to the trivial case when $\hat x_i^+=\hat x_i^-$.
% the $i$-th estimate does not receive a measurement update.
% to the highly unlikely case that the $i$-th object is not measured by full certainly. Furthermore, the limit of the gain is given by
%\begin{align*}
%\lim_{\beta_{i,0}\rightarrow 1} K_i= 0.
%\end{align*}
\item[(iii)] The proposed gain $K_i$ is different from the common Kalman gain, namely $K_{K_i}=P_{i}^{-}H^T S_i^{-1}$, in general. Even when the measurement matrix is fixed, i.e., $R_{i,j}=R_i$, each association probability is restricted to $\beta_{i,j}\in\{0,1\}$, which is not satisfied in general.
%\begin{align*}
%K_{K_i}\Sigma_i=0, \quad \Sigma_i&=\sum_{j=1}^{n_r}\beta^2_{i,j}R_j+(1-\beta_{i,0})^2H_iP_i^-H_i^T,%{\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}_i}}{\bar{e}_i}^T,
%\end{align*}
%or every $\beta_{i,j}$ is exactly $0$ or $1$, which corresponds to filtering when no measurement origin uncertainty exists.
\item[(iv)] The uncertainty cost of the $i$-th estimate from the Kalman gain satisfies $J_{P_{i,K}}\geq J^*_{P_i}$ because
\begin{align*}
J_{P_{i,K}}-J^*_{P_i}&=\tr{K_{K_i}(\beta_{i,0}E_i+\tilde\beta R_i)E_i^{-1}(\beta_{i,0}E_i+\tilde\beta R_i)K_{K_i}^T}\geq0, \\ \tilde\beta_i&=\left(
\sum_{j=1}^{n_r}\beta_{i,j}
\right)^2
-\sum_{j=1}^{n_r}\beta_{i,j}^2.
\end{align*}
\end{itemize}
\end{prop}
\begin{proof}
See the Appendix A.

The Kalman gain employed by JPDAF is optimal if there exists no measurement uncertainty. This corresponds to applying a hard decision, where this is the special case when $E_i=S_i$ from \refeqn{E} and $K_i=P^-_iH_i^TS_i^{-1}$ from \refeqn{GainK} is the Kalman gain~\cite{OptEst1}.
Since the JPDAF may apply soft decisions, it follows that the Kalman gain is suboptimal for the JPDAF in general.
%These conclusions come from comparing the derived optimal gain and the Kalman gain, and the differences in the uncertainty cost functions associated with these gains, shown later in this paper.
%
%
%(You may reuse...
%Note that $E_i$ is independent of $K_i$ and it is invertible when $\beta_{i,0}\neq1$ since $S_i$ is positive-definite and
%
%$(\beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}{\bar{e}}_{i}^T)$ is positive-semidefinite, as shown in the Appendix A. 
%The case when $\beta_{i,0}=1$ corresponds with the $i$-th object \emph{not} being measured with full certainty, and such cases is analyzed in Appendix B.)
\end{proof}

In short, the differences between the M-JPDAF and the conventional JPDAF are shown in \refeqn{GainK} and \refeqn{JPDAFPostCovSymmetricUpdate} or \refeqn{JPDAFPostCovSimpleUpdate}.
As the proposed optimal gain minimizes the posterior uncertainties, the corresponding M-JPDAF performs better than the conventional JPDAF, as illustrated by numerical examples later.

\section{Coalescence-Avoiding Optimal JPDAF (C-JPDAF)}
\label{sec:C-JPDAF}

%Coalescence of the JPDAF (or variations thereof such as the M-JPDAF described in the last section) is the outcome of estimates becoming closer to each other as a result of sharing measurements. This occurs because the signs of the innovations typically cause updates that draw measurement estimates towards the observed measurements; the net effect tends to cause estimates to become more similar to each other (e.g. a large number of estimates in close proximity, estimates with large uncertainties).
%The goal of the C-JPDAF is to minimize a cost that includes a similarity index in addition to uncertainty.
When multiple objects are in close proximity, a single measurement may be shared by several objects with similar association probabilities.
Probabilistic data association filtering with shared measurements may cause the estimates to become more similar to each other, thereby causing coalescence in estimated data, i.e., the estimates of several objects tend to become identical.

%\EditTL{(Since this is a journal paper, there is no need to emphasize how this paper is different from~\cite{KauLovLee14}. You may rewrite this paragraph such that it just summarize the key idea and the key assumptions of the proposed approach).}
In this paper, the approach to remove coalescence in the probabilistic data association filtering problem explicitly considers a measure of similarity among the estimates.
The filter is designed to minimize a weighted sum of uncertainty and similarity costs.
This concept is explained and very simple numerical examples are presented in~\cite{KauLovLee14}.
These are based on the assumptions that during every filter update, exactly two objects are present in the field of view of a sensor, and only a single measurement at each time step is available.
%This measurement originates from one of the two objects with equal probability that it originated from either object.
%These assumptions severely limit the scenarios where this algorithm can be useful.
In~\cite{KauLovLeeAASDA15}, these restrictive assumptions are removed so that any number of objects may be tracked with an arbitrary number of measurements.
Furthermore, the similarity measure is more effectively derived in terms of the measurement space (measurements, measurement estimates, and innovation uncertainties), because this space is where coalescence is most evident.
%This measurement originates from one of the two objects with equal probability that it originated from either object.
%These assumptions severely limit the scenarios where this algorithm can be useful.
%Therefore in this paper, the derivation and numerical examples do not operate on such restrictive assumptions.
%Additionally, the similarity measure is slightly changed to be in terms of measurements and measurement estimates (measurement space), so the technique may be applied to a wider variety of systems.
%This approach to the problem further contrasts with~\cite{KauLovLee14}, which derives the similarity index in terms of the state variables (state space).

\subsection{Matusita's Similarity Measure of Multiple Distributions}

%We choose Matusita's measure to determine the degree of similarity among multiple Gaussian estimates because this measure is easy to compute and is easily differentiable compared to other similarity measures (e.g.~\cite{Bhatt},~\cite{MacA}).
%Matusita's measure represents the degree of similarity between two multivariate Gaussian probability distributions~\cite{Mat55},~\cite{Coal_k}.
The similarity between multiple multivariate Gaussian distributions can be represented by the overlapping regions of their probability density functions. Since there exists no analytic solution for this overlap, we choose Matusita's measure because it provides a scalar approximation of the overlap of multivariate Gaussian distributions~\cite{Mat55},~\cite{Coal_k}. Furthermore, this measure is easy to compute and is easily differentiable unlike the Bhattacharyya measure~\cite{Bhatt97}.
More explicitly, let $y_i$ be the $i$-th Gaussian random variable, satisfying $y_i \sim \mathcal{N}[\hat y_i,{\mathbf P}_i]$ for $1\leq i \leq n_t$. The portion of the Matusita's similarity measure representing the spread in the means between $y_p$ and $y_q$ for some $1\leq p,q\leq n_t$ with $p\neq q$ is given by,
\begin{align}
\zeta_{pq}=\exp \{-\frac14(\hat y_{p}- \hat y_{q})^T({\mathbf P}_{p}
+{\mathbf P}_{q})^{-1}(\hat y_{p}-\hat y_{q})\}.\label{eqn:Mat2Est}
\end{align}

%Let $y_i$ be a random variable corresponding to the $i$-th system such that $y_i \sim \mathcal{N}[\hat y_i,{\mathbf P}_i]$, and let indices $p,q\leq n_t$ such that $p\neq q$.
%From~\cite{KauLovLee14}, the portion of the Matusita's similarity measure $\zeta_{pq}\in\Re$ corresponding to the spread in the estimate means is given by
%\begin{align}
%\zeta_{pq}=\exp \{-a(\hat y_{p}- \hat y_{q})^T({\mathbf P}_{p}
%+{\mathbf P}_{q})^{-1}(\hat y_{p}-\hat y_{q})\},\label{eqn:Mat2Est}
%\end{align}
%where $a\in\Re$ such that $a>0$ shapes the similarity measure.

%However, the distributions of more than two estimates may exhibit a nonzero overlap.

We generalize it into a similarity measure for \emph{any} number of Gaussian distributions shaped by constant $a\in\Re^+$,%by summing \emph{all possible} combinations of Matusita's similarity measures% $\zeta\in\Re$, illustrated in Figure \ref{fig:SimMeas}, expressed in a general sense as
\begin{align}
\label{eqn:zetaGeneral}
\zeta=\sum\limits_{\substack{p,q=1\\p\neq q}}^{n_t} \exp \{-a(\hat y_{p}- \hat y_{q})^T({\mathbf P}_{p}
+{\mathbf P}_{q})^{-1}(\hat y_{p}-\hat y_{q})\},
\end{align}
which is the summation of \refeqn{Mat2Est} between all unique pairs.
The structure of \refeqn{zetaGeneral} is maintained to generate a cost function for optimal coalescence avoidance.
\begin{figure}
\centerline{
		{\includegraphics[width=.7\columnwidth]{GaussianDistOverlap1D.png}}
	}
\caption{A similarity index between three Gaussian distributions is illustrated as summation of overlapping regions of two distributions (green) or three distributions (brown).
This concept is easily extended to multivariate Gaussian distributions.
}\label{fig:SimMeas}
\end{figure}

\subsection{Optimal Coalescence Avoidance}

%\EditTL{(Not sure we need this paragraph) Soft decision algorithms such as the JPDAF and M-JPDAF, which only consider estimate uncertainties when determining estimator gains, yield coalescence among the estimates. However, including a similarity index in an optimal fashion to determine each gain matrix serves to remove this coalescence.}

The proposed coalescence avoiding filtering algorithm is constructed by minimizing the cost function $\mathbf{J}\in\Re$, defined as as the sum of the uncertainty cost $J_P\in\Re$ and similarity cost $J_S\in\Re$ with positive weighting factor $c\in\Re$,
\begin{align}
\label{eqn:CostGen}
\mathbf{J}=J_P+cJ_S.
\end{align}
%The estimator gain $K_i$ is selected to minimize the above cost $\mathbf{J}$. In short, the estimator gain is chosen in the consideration of both uncertainty and coalescence, illustrated in Figure \ref{fig:CostTrends}.
The estimator gain $K_i$ is selected to minimize the above cost $\mathbf{J}$. By considering similarity to obtain the estimator gain, we mitigate some coalescence at the expense of increasing estimate uncertainty according to the magnitude of $c$.
The choice of $c$ depends on the type of scenario and the parameter $a$ shaping the similarity cost $J_S$.

\begin{figure}
\setlength{\unitlength}{0.06\columnwidth}
\centerline{\footnotesize\selectfont
\begin{picture}(8.5,7.3)(-0.5,-0.8)
\put(-0.5,3){\rotatebox{90}{Cost}}
\put(0,0){\includegraphics[width=0.48\columnwidth]{ECC14_fig1.pdf}}
\put(5.3,1.2){\shortstack[c]{Uncertainty\\ cost: $J_P$}}
\put(6.2,2.6){\shortstack[c]{Similarity\\ cost: $J_S$}}
\put(5.2,5.2){\shortstack[c]{Weighted sum:\\ $\mathbf{J}=J_P+cJ_S$}}
\put(4.98,0.78){\circle*{0.1}}
\put(2.48,1.57){\circle*{0.1}}
\put(4.8,-0.4){$K_{M-JPDAF}$}
\put(2.3,-0.4){$K_{C-JPDAF}$}
\put(3.1,-0.8){Estimator gain}
\end{picture}
}
\caption{Illustration of optimal coalescence avoidance: several costs are illustrated with respect to an element of an estimator gain $K$ (red: uncertainty cost $J_P$, blue: similarity cost $J_S$, black: weighted sum). The estimator gain $K_{M-JPDAF}$ is selected to minimize the uncertainty cost. The estimator gain $K_{C-JPDAF}$ is chosen such that the weighted sum of the uncertainty cost and the similarity cost is minimized. In short, at the expense of increased uncertainty at a single step, the proposed approach prevents coalescence. The increased uncertainty at each step is rewarded by reducing the overall estimation error drastically by maintaining correct data association during close proximity tracking scenarios or missions.}
\label{fig:CostTrends}
\end{figure}

\paragraph*{Uncertainty Cost}\ 
The uncertainty cost $J_P$ is defined as the sum of the trace of each posterior covariance as% the summation of \refeqn{JpIntro}:
\begin{align}
J_P=\sum\limits_{i}\tr{P^+_i},\label{eqn:JP}
\end{align}
where $P^+_{i}$ is obtained by \refeqn{JPDAFCov}, dependent on the gain $K_i$ at the current time step.
The derivative of $J_P$ is taken from \refeqn{Uncertainty0}
\begin{align}
\label{eqn:CostP}
\frac{\partial J_{P}}{\partial K_{i}}&=2K_iE_i-2(1-\beta_{i,0})P^-_iH_i^T.
\end{align}

\paragraph*{Similarity Cost}\ 
Any shared measurement $z_j$ on measurement estimate $\hat z_i$ yields an innovation $e_{i,j}$ according to \refeqn{eij}.
During the measurement update, this innovation typically has the net effect of drawing $\hat z_i$ towards $z_j$ if $\beta_{i,j}$ is nonzero.
This occurs for all estimates and all measurements; coalescence results when measurement estimates are drawn towards common measurements.
Therefore, coalescence of the JPDAF or M-JPDAF occurs inside the measurement space (e.g. if a sensor provides ranges and bearings, these are in the measurement space as opposed to a Cartesian state vector inside the state space).

Contrary to~\cite{KauLovLee14} where the similarity index is in terms of state space variables, the coalescence avoidance presented in this paper takes place inside the measurement space to most directly remove the coalescence.
Let the posterior measurement estimate be $\hat z_i^+\in\Re^m$, and define the innovation $e_{i}^+\in\Re^m$ as
\begin{align}
\hat z_i^+&=H_i\hat x_i^+,
\\
e_{i}^+&=z_i-\hat z_i^+=(H_ix_i+v_i)-H_i\hat x_{i}^+=H_i\tilde x_i^++v_i,
\end{align}
where $\tilde x_i^+=x_i-\hat x_{i}^+\in\Re^n$ denotes the error between the true state and the posterior estimate. The posterior innovation covariance $S^+_i\in\Re^{m\times m}$ is given by
\begin{align}
S^+_i=\mathrm{E}[e_i^+e_i^{+T}]=\mathrm{E}[(H_i\tilde x_i^++v_i)(H_i\tilde x_i^++v_i)^T]
=H_iP^+_iH_i^T+R_i.\label{eqn:Spi}
\end{align}
From these, the similarity cost $J_S$ is defined according to the format of \refeqn{zetaGeneral} as
%which are used to define $u_{p q}$ between $\hat z^+_{p}$ and $\hat z^+_{q}$:
\begin{align}
\label{eqn:Js}
J_S=\sum\limits_{\substack{p,q=1,\\p\neq q}}^{n_t}\exp (-au_{pq}),
\end{align}
where $u_{pq}\in\Re$ is a function of the innovation covariance summation $S^+_{pq}=S^+_{p}+S^+_{q}\in\Re^{m\times m}$,
\begin{align}
u_{pq} & = (\hat z_{p}^+-\hat z^+_{q})^T(S^+_{pq})^{-1}(\hat z^+_{p}-\hat z^+_{q})\label{eqn:U}.
\end{align}
%Expanding \refeqn{U} with respect to the $p$-th system,
Substituting the expression for $S_p^+$ obtained by \refeqn{Spi},
\begin{align}
u_{pq}=&\ 
(H_pK_p{\bar{e}}_p-\hat z^+_{q})^T
(H_pP^+_{p}H_p^T+R_p+S^+_q)^{-1}
(H_pK_p{\bar{e}}_p-\hat z^+_{q}).
\label{eqn:Up}
\end{align}

\begin{prop}
Let $b$, $d$, and $p$ be integers such that $b\in\{1,2,...,n\}$,  $d\in\{1,2,...,m\}$, and $p\in\{1,2,...,n_t\}$ such that the $(b,d)$-th element of the $p$-th gain matrix is written as $[K_p]_{bd}\in\Re$.
Also, let $\Ibd\in \Re^{n\times m}$ be defined such that its $(b,d)$-th element is one and other elements are zero.
Then the following partial derivative may be obtained,
\begin{align*}
\deriv{J_S}{[K_p]_{bd}}&=\ -a\sum\limits_{q,q\neq p}\exp (-au_{pq})
\{
2(H_p\Ibd{\bar{e}}_p)^T
{S_{pq}^+}^{-1}
\nonumber
\\
& \quad
-(H_pK_p{\bar{e}}_p-\hat z^+_{q})^T
{S^+_{pq}}^{-1}
H_p
[(1-\beta_{0,p})(\Ibd H_pP_p^-
\nonumber
\\
& \quad
+P_p^-H_p^T\Ibd^T)-\Ibd E_pK_p^T\nonumber
\\
& \quad -K_pE_p\Ibd^T]
H_p^T
{S^+_{pq}}^{-1}
\}(H_pK_p{\bar{e}}_p-\hat z^+_{q}).
%\\
%\mathcal{S}_{pq,p,bd}&=-{S^+_{pq}}^{-1}
%H_p
%[(1-\beta_{0,p})(\Ibd H_pP_p^-+P_p^-H_p^T\Ibd^T)-\Ibd E_pK_p^T\nonumber
%\\
%& \quad -K_pE_p\Ibd^T]
%H_p^T
%{S^+_{pq}}^{-1}.
\end{align*}
\end{prop}
\begin{proof}
See the Appendix B.
\end{proof}

Necessary conditions for optimality are given by
\begin{align}
\deriv{\mathbf{J}}{K_i} = \deriv{J_P}{K_i} + c \deriv{J_S}{K_i} =0,\label{eqn:NCO}
\end{align}
where $\deriv{J_P}{K_i}$ and $\deriv{J_S}{K_i}$ may be nonzero, so the result may be different than \refeqn{GainK}, which only considers state estimate uncertainty in the optimization.
In short, \refeqn{CostP} and \refeqn{JSK} are substituted into \refeqn{NCO} and repeated for all $p$ to obtain $n_t$ optimal gain matrices.
Then, the C-JPDAF is completed by updating the state estimate and covariance by using \refeqn{KalEst} and \refeqn{JPDAFCov}, respectively.

As the optimality condition is a nonlinear function of estimator gains, there is no closed-form solution.
Hence, we require a nonlinear equation solver.
When $c=0$, we can show that the optimal gain reduces to that of the M-JPDAF, which is used as an initial guess during iterative numerical methods.
%Therefore, if the number of estimates exhibiting significant similarity is $n_t^*\leq n_t$ and $\mathcal C$ is the number of times \refeqn{NCO} is solved depending on the numerical accuracy, then the required iterations is $n_t^*(n_t^*-1)\mathcal C$, where some variables may be reused (e.g. $S^+_{pq}\equiv S^+_{qp}$) and often the effect of \refeqn{derivu} is negligible with large estimate separations.

\section{Numerical Examples}
\label{sec:NE}

We consider satellites in the low earth orbit (LEO) and in the geosynchronous orbit (GEO) to compare the performances of the conventional JPDAF, the M-JPDAF, and the C-JPDAF.
The satellites are subject to the dynamics of the two-body problem.
Since the system dynamics $\dot x=f(x)$ is nonlinear, the linearized matrix $F_{i_{k-1}}=\deriv{f}{x_i}\bigg|_{x_i=\hat x_{i_{k-1}}^+}$ is used with \refeqn{xestapriori} and \refeqn{Papriori}.
Similarly, the measurement dynamics $z_i = h(x_i)$ may experience nonlinearities (e.g. range, range rate, bearing), the linearized matrix $H_i=\deriv{h_i}{x_i}\bigg|_{x_i=\hat x_{i}^-}$ is used with \refeqn{GainK}, \refeqn{JPDAFPostCovSimpleUpdate}, \refeqn{CostP}, \refeqn{Spi}, and \refeqn{Up}.

Let $\vec r=\begin{bmatrix}x_1, & x_2, & x_3\end{bmatrix}^T\in\Re^3$ be the position vector of the $i$-th satellite represented with respect to the Earth-Centered Inertial (ECI) frame.
The equation of motion corresponding to the two-body problem is given by
\begin{align}
\label{eqn:NonLin2BP}
\ddot{\vec r}&=-\frac{\mu}{r^3}\vec r=-\mu(\vec r^T\vec r)^{-\frac32}\vec r,
\end{align}
where $\mu=398,600\ \frac{km^3}{sec^2}$ is the gravitational constant of the earth if position is estimated in kilometers and time is measured in seconds~\cite{Val01}.

Let $\delta$ denote an infinitesimal change of a state variable.
The linearized equations of state variable $\vec x=\begin{bmatrix}\vec r^T, & \dot{\vec r}^T\end{bmatrix}^T\in\Re^6$ are based on \refeqn{NonLin2BP} and given by
\begin{align}
\begin{bmatrix}
\delta\dot{\vec r} \\ \delta\ddot{\vec r}
\end{bmatrix}
%\delta\dot{\vec x}
=
\begin{bmatrix}
0_{3\times3}, & I_{3\times3} \\
\mu\left[3\vec r({\vec r}^T\vec r)^{-\frac52}{\vec r}^T-({\vec r}^T\vec r)^{-\frac32}I_{3\times3}\right], & 0_{3\times3}
\end{bmatrix}
\begin{bmatrix}
\delta\vec r \\ \delta\dot{\vec r}
\end{bmatrix}
%\delta\vec x
.
\end{align}
%Defining the above square matrix as $A_{lin}$, the linearized state transition matrix over the time step $dt$ is defined as
%\begin{align}
%F=\expm{(A_{lin}dt)}
%\end{align}
%where $\expm$ denotes the matrix exponential function.

The process uncertainty may be subject to a variety of perturbations beyond two-body forces; in the following simulations, let these uncertainties be summarized with the standard deviation of position $\sigma_r\in\Re$ and velocity $\sigma_{\dot r}\in\Re$ such that the process covariance is defined as $Q=\diag[\sigma_r^2I_{3\times3}, \sigma_{\dot r}^2I_{3\times3}]\in\Re^{6\times6}$, where $\sigma_r$ and $\sigma_{\dot r}$ take several values depending on the case considered.
%=\diag[10^{-6}I_{3\times3}, 10^{-8}I_{3\times3}]$.

For all cases, the satellites are measured with a phased-array radar (LEO) or telescope (GEO) at the location $\vec r_s\in\Re^3$ on the equator in the ECI frame, which has nonzero velocity due to the rotation of the earth,
\begin{align}
\vec r_s=R_E
\begin{bmatrix}
\cos\frac{2\pi t}T, & \sin\frac{2\pi t}T, & 0
\end{bmatrix}^T,
\quad
\dot{\vec r}_s=\frac{2\pi tR_E}T
\begin{bmatrix}
-\sin\frac{2\pi t}T,  & \cos\frac{2\pi t}T, & 0
\end{bmatrix}^T,
\end{align}
where $R_E=6378$ km is the radius of the earth, time of the day $t$ starts at $0$ seconds, and the time period of the day is $T=86164.090518$ seconds.
Here, the measurements consist of the range $\rho$ between the sensor and the satellite (km), the range rate $\dot\rho$ (km/sec), and topocentric right ascension $\alpha$ and declination $\delta$ (rad)~\cite{Val01}, i.e., the $j$-th measurement $z_j$ ($z_j\in\Re^4$ in LEO and $z_j\in\Re^2$ in GEO) is
\begin{align}
\label{eqn:zj}
z_j=\begin{cases}
                        \begin{bmatrix}\rho, & \dot\rho, & \alpha, & \delta\end{bmatrix}^T \text{for LEO satellites} \\
                        \begin{bmatrix}\alpha, & \delta\end{bmatrix}^T \qquad\  \text{for GEO satellites}
                    \end{cases}
.
\end{align}
%The standard deviations $\sigma_\rho$, $\sigma_\alpha$, and $\sigma_\delta$ are based on the Mahe Ground Station uncertainty data~\cite{VerSauSco04} and $\sigma_{\dot\rho}$ from the examples from~\cite{KurAriAriEfe10}.

%For each following scenario, several cases are simulated where various parameters are slightly changed with $\epsilon\in\Re$, summarized in Table \ref{tab:CaseDef}, and measurement are simulated with \refeqn{NonLin2BP} and \refeqn{zj}.
In the following scenarios, several cases are simulated with \refeqn{NonLin2BP} and measurements are generated according to \refeqn{zj}.
For all cases and time steps, measurements may originate from satellites, but this origin is not given to the data association algorithm.
Additionally, there is a $10\%$ chance that a measurement is either deleted or an extraneous measurement is added at each time step.
Further descriptions of the cases are available below, and the resulting accuracy is measured with RMS error, the uncertainty is summarized with $\tr{J_P}$, and the computation time is measured as well, tabulated in Tables \ref{tab:A1}--\ref{tab:B2}.
Scenarios A1 and B1 are data association and filtering examples in LEO, while scenarios A2 and B2 are GEO examples.
For readers with interest in how the algorithms handle perturbations, various parameters are slightly changed with $\epsilon\in\Re$, summarized in Table \ref{tab:CaseDef}.
The metrics from the perturbed cases are summarized in the subsequent tables as well.

\begin{table}
\begin{center}
\caption{Case Definitions} \label{tab:CaseDef}
\begin{threeparttable}[h]
\begin{tabularx}{.65\textwidth}
{
>{$}c<{$} |
>{$}c<{$}
}
\toprule
\multirow{2}{*}{Symbol} & \multirow{2}{*}{Explanation}\\
\\
\midrule
\multirow{1}{*}{A} &  \multirow{1}{*}{Comparison of JPDAF and M-JPDAF}
\\
\multirow{1}{*}{B} &  \multirow{1}{*}{Comparison of JPDAF, M-JPDAF, and C-JPDAF}
\\
\multirow{1}{*}{1} &  \multirow{1}{*}{Orbits in LEO}
\\
\multirow{1}{*}{2} &  \multirow{1}{*}{Orbits in GEO}
\\
\multirow{1}{*}{a} &  \multirow{1}{*}{$\sigma^2_{\rho,a} = \epsilon\sigma^2_{\rho}$}
\\
\multirow{1}{*}{b} &  \multirow{1}{*}{$\sigma^2_{\dot \rho,b} = \epsilon\sigma^2_{\dot \rho}$}
\\
\multirow{1}{*}{c} &  \multirow{1}{*}{$\sigma^2_{\rho,c} = \epsilon\sigma^2_{\rho}$, $\sigma^2_{\dot \rho,c} = \epsilon\sigma^2_{\dot \rho}$}
\\
\multirow{1}{*}{d} &  \multirow{1}{*}{$\sigma^2_{\alpha,d} = \epsilon\sigma^2_{\alpha}$, $\sigma^2_{\delta,d} = \epsilon\sigma^2_{\delta}$}
\\
\multirow{1}{*}{e} &  \multirow{1}{*}{$\sigma^2_{\mathbf{\rho},e} = \epsilon\sigma^2_{\mathbf{\rho}}$ (scenario 1 only)}
\\
\multirow{1}{*}{f} &  \multirow{1}{*}{$\sigma^2_{\mathbf{\dot\rho},f} = \epsilon\sigma^2_{\mathbf{\dot\rho}}$ (scenario 1 only)}
\\
\bottomrule
\end{tabularx}
\end{threeparttable}
\end{center}
\end{table}

\subsection{Data Association and Filtering in LEO}

Various cases in LEO are simulated to compare the data association and filtering algorithms, namely the JPDAF, M-JPDAF, and C-JPDAF.
The measurement uncertainty is $R=\diag[\sigma_\rho^2, \sigma_{\dot\rho}^2, \sigma_\alpha^2, \sigma_\delta^2]=\\\diag[(0.15\ \text{km})^2, (0.005\ \text{km/s})^2, (\frac{0.01\pi}{180}\ \text{rad})^2, (\frac{0.01\pi}{180}\ \text{rad})^2]\in\Re^{4\times4}$ where these standard deviations are selected from the Mahe Ground Station uncertainty data~\cite{VerSauSco04} and ~\cite{KurAriAriEfe10}.

\paragraph*{Scenario A1: Comparison of the JPDAF and M-JPDAF with two LEO satellites in close proximity}\

The first set of cases compares the JPDAF with the M-JPDAF for tracking two nearby satellites passing over the sensor field of view in LEO. The initial conditions of the first two orbits are chosen as
\begin{align}
x_{1,A1}&=x_{ref,A1}+x_{pert,A1}, \quad x_{2,A1}=x_{ref,A1}-x_{pert,A1},\nonumber
\\
x_{ref,A1}&=\begin{bmatrix}7000\ \text{km}, & 0\ \text{km}, & 1000\ \text{km}, & 0\ \text{km/s}, & 7.546\ \text{km/s}, & 0\ \text{km/s}\end{bmatrix}^T,
\nonumber\\
x_{pert,A1}&=\begin{bmatrix}
0.2\ \text{km}, & 0\ \text{km}, & 0\ \text{km}, & 0\ \text{km/s}, & -0.105\ \text{km/s}, & 0\ \text{km/s}
\end{bmatrix}^T.\nonumber
\end{align}
In these cases, the process covariance is $Q=\diag[10^{-6}I_{3\times3}\ \text{km}^2, 10^{-8}I_{3\times3}\ \text{km/s}^2]$, the estimates have initial covariances matrix $P_{0,A1}=10^7\times Q$, time steps of $1$ second over $100$ seconds, and $\epsilon=10$.
These scenarios are chosen because the estimates exhibit large initial uncertainties, so there are soft decisions with large measurements shared and the portion of the posterior covariance affected by \refeqn{E} is significant.
The gain of the M-JPDAF accounts for this portion of the state covariance matrix better than the gain for the conventional JPDAF in every single case.

Given identical parameters and measurements, the RMS errors of the position and the mean uncertainty cost $J_P$ defined in \refeqn{JP} are lower in \emph{every} case when the M-JPDAF is applied compared with when the conventional JPDAF is applied.
In fact in two cases, namely A1b and A1c, the JPDAF is incapable of decreasing the estimate uncertainty as quickly as the M-JPDAF.
In some cases, the JPDAF suffers from track swapping, which is when measurements are consistently associated incorrectly.
%; the JPDAF initially suffers from such great coalescence that the estimates become associated with incorrect measurements consistently, known as track swapping.
Furthermore, the computational time is roughly the same between the JPDAF and M-JPDAF, so the results suggest that the improvement in accuracy is at no increase in computation.
These results are tabulated in Table \ref{tab:A1}, and some cases are plotted in Figure \ref{fig:A1}.

\begin{center}
\begin{threeparttable}[h]
\caption{Cases of A1} \label{tab:A1}
\begin{tabularx}{\textwidth}
{
>{$}c<{$} |
*{2}{>{$}c<{$}} |
*{2}{>{$}c<{$}} |
*{2}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multicolumn{2}{c}{\multirow{1}{*}{RMS Position Err. (km)}} & \multicolumn{2}{c}{\multirow{1}{*}{Time (sec)}} & \multicolumn{2}{c}{\multirow{1}{*}{Mean $J_P$}} \\
 & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF}
\\
\midrule
\text{A1}  & 0.12026 & 0.10646 & 0.582 & 0.561 & 0.5125 & 0.49054 \\
\text{A1a} & 0.17910 & 0.17157 & 0.558 & 0.561 & 4.6154 & 4.5607 \\
\text{A1b} & \hilight{2.63710} & 0.14150 & 0.562 & 0.559 & 1.9718 & 0.68328 \\
\text{A1c} & \hilight{2.63520} & 0.19559 & 0.570 & 0.579 & 5.7223 & 4.8463 \\
\text{A1d} & 0.20204 & 0.18865 & 0.573 & 0.578 & 0.59901 & 0.57227 \\
\text{A1e} & 0.14541 & 0.13361 & 0.576 & 0.574 & 0.53064 & 0.50884 \\
\text{A1f} & 0.12175 & 0.10795 & 0.576 & 0.560 & 0.51338 & 0.49086 \\
\bottomrule
\end{tabularx}
{\small
\begin{tablenotes}
    \item \hilight{Track swapping occurs}
  \end{tablenotes}}
\end{threeparttable}
\end{center}

\begin{figure}
{
%\hspace*{0.05\columnwidth}
\centerline{
	\subfigure[\;Case A1: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1JPDAF.pdf}}
	\subfigure[\;Case A1: M-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1MUJPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case A1b: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1bJPDAF.pdf}}
	\subfigure[\;Case A1b: M-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1bMUJPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case A1d: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1dJPDAF.pdf}}
	\subfigure[\;Case A1d: M-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1dMUJPDAF.pdf}}
	}
}
\caption{The orbital radii of three cases are shown where the dashed plots are the true trajectories and the solid plots are the estimated trajectories.
In all of these cases, the M-JPDAF is more accurate and avoids track swapping (Case A1b in particular).
% experiences less coalescence than the JPDAF.
%In particular, when the JPDAF is used in Case A1b, the initial coalescence causes track swapping.
Additionally, Case A1d experiences the most inaccurate estimation without track swapping with very noisy angle measurements, and the M-JPDAF has lower RMS error when compared to the JPDAF.
}\label{fig:A1}
\end{figure}

\paragraph*{Scenario B1: Comparison of the JPDAF, M-JPDAF, and C-JPDAF with three LEO satellites in close proximity}\

The JPDAF, M-JPDAF, and C-JPDAF algorithms are compared for three satellites in LEO with initial conditions
\begin{align}
x_{1,B1}&=x_{ref,B1}+x_{pert,B1}, \quad x_{2,B1}=x_{ref,B1}-x_{pert,B1}, \quad x_{3,B1}=x_{ref,B1},\nonumber
\\
x_{ref,B1}&=\begin{bmatrix}7000\ \text{km}, & 0\ \text{km}, & 0\ \text{km}, & 0\ \text{km/s}, & 7.546\ \text{km/s}, & 0\ \text{km/s}\end{bmatrix}^T,
\nonumber\\
x_{pert,B1}&=\begin{bmatrix}
0.025\ \text{km}, & 0\ \text{km}, & 0\ \text{km}, & 0\ \text{km/s}, & 0\ \text{km/s}, & 0\ \text{km/s}
\end{bmatrix}^T,\nonumber
\end{align}
where $Q=\diag[10^{-6}I_{3\times3}\ \text{km}^2, 10^{-8}I_{3\times3}\ \text{km/s}^2]$, $P_{0,B1}=Q$, and $\epsilon=1.1$ for time steps of $10$ seconds over a $2000$ second time period.
This corresponds to a phased-array radar observing observing a slow separation of LEO objects in a near-circular orbit; because the satellite states are in close proximity over a large time period, the JPDAF and M-JPDAF are subject to severe coalescence.
The C-JPDAF is designed to avoid this coalescence, where the parameters $a=5\times10^{-1}$ and $c=2.3\times10^{-4}$ from \refeqn{zetaGeneral} and \refeqn{CostGen} are used in these scenarios from trial and error.
The accuracy of the C-JPDAF is largely improved from that of the JPDAF and M-JPDAF; however, the computational cost of the C-JPDAF is roughly $60$ times that of either other algorithm.
The position RMS errors and computation times are tabulated in Table \ref{tab:B1} and example plots from these cases are shown in Figure \ref{fig:B1}.

\begin{center}
\begin{threeparttable}[h]
\caption{Case B1} \label{tab:B1}
\begin{tabularx}{\textwidth}
{
>{$}c<{$} |
*{3}{>{$}c<{$}} |
*{3}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multicolumn{3}{c}{\multirow{1}{*}{RMS Position Err. (km)}} & \multicolumn{3}{c}{\multirow{1}{*}{Time (sec)}} \\
 & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{C-JPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{C-JPDAF}
\\
\midrule
\text{B1}   	& 0.14432 & 0.091133 & 0.057312 & 1.74 & 1.67 & 101 \\
\text{B1a} 	& 0.14432 & 0.091134 & 0.057307 & 1.69 & 1.73 & 101 \\
\text{B1b} 	& 0.14694 & 0.094506 & 0.055145 & 1.67 & 1.68 & 101 \\
\text{B1c} 	& 0.14693 & 0.094507 & 0.055136 & 1.68 & 1.67 & 101 \\
\text{B1d} 	& 0.14400 & 0.088933 & 0.058911 & 1.66 & 1.66 & 103 \\
\text{B1e} 	& 0.14408 & 0.090823 & 0.059919 & 1.67 & 1.67 & 100 \\
\text{B1f} 	& 0.14436 & 0.091212 & 0.064299 & 1.73 & 1.72 & 98.7 \\
\bottomrule
\end{tabularx}
\begin{tabularx}{.52\textwidth}
{
>{$}c<{$} |
*{3}{>{$}c<{$}}
}
\multirow{2}{*}{Case} & \multicolumn{3}{c}{\multirow{1}{*}{Mean $J_P$}} \\
 & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{C-JPDAF}
\\
\midrule
\text{B1}   	    	& 0.24559 & 0.17819 & 0.17958 \\
\text{B1a} 		& 0.24559 & 0.17819 & 0.17959 \\
\text{B1b} 		& 0.25349 & 0.18386 & 0.18510 \\
\text{B1c} 		& 0.25350 & 0.18386 & 0.18511 \\
\text{B1d} 		& 0.26058 & 0.18944 & 0.19177 \\
\text{B1e} 		& 0.24626 & 0.17861 & 0.1794 \\
\text{B1f} 		& 0.24561 & 0.17820 & 0.18009\\
\bottomrule
\end{tabularx}
\end{threeparttable}
\end{center}

\begin{figure}
{
%\hspace*{0.05\columnwidth}
\centerline{
	\subfigure[\;Case B1: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1JPDAF.pdf}}
	\subfigure[\;Case B1: M-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1MUJPDAF.pdf}}
	\subfigure[\;Case B1: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1C-JPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case B1c: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1cJPDAF.pdf}}
	\subfigure[\;Case B1c: M-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1cMUJPDAF.pdf}}
	\subfigure[\;Case B1c: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1cC-JPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case B1f: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1fJPDAF.pdf}}
	\subfigure[\;Case B1f: M-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1fMUJPDAF.pdf}}
	\subfigure[\;Case B1f: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1fC-JPDAF.pdf}}
	}
}
\caption{The orbital radii of three cases are shown where the dashed plots are the true trajectories and the solid plots are the estimated trajectories.
In all cases the M-JPDAF is more accurate and experiences less coalescence than the JPDAF, but both algorithms suffer from coalescence.
The C-JPDAF systematically removes this coalescence from the updates at the cost of slightly larger covariance matrices and a large increase in required computational resources.
%Between Cases B1 and B1c, the process position uncertainty is increased, which last little effect on the estimates since this has little
The conditions that affect the C-JPDAF most are with Case B1f where the range-rate measurements have the greatest uncertainty.
}\label{fig:B1}
\end{figure}

Among the variations of the B1 cases, the range-rate measurement uncertainty (Case B1f) has the largest effect on tracking with the C-JPDAF.
The velocities of satellites in LEO are very large, so a small increase in $\sigma_{\dot\rho}$ corresponds to a large decrease in $J_S$, i.e., the measurement sets for the satellites will be less similar.  This is further shown with the smallest computational cost among the B1 cases.
Hence, the similarity optimization is very sensitive to velocity measurement uncertainty in these cases.
%This can be explained because in the B1 cases, the ranges and bearings of each track are close and changing, but the difference between the velocities of the three satellites remain largely constant after about $500$ seconds.
%As long as the estimates are able to stay sufficiently far apart, as the C-JPDAF is able to accomplish, the component of the velocity measurement in the radial direction serves as a distinguishing portion of the full measurement, so that the correct association between measurement and estimate is more heavily weighted, and hence better tracking is achieved.
%The quality of the range-rate measurement is degraded in Case B1f, and hence the C-JPDAF algorithm is not able to as accurately track each estimate when compared with the other B1 cases.

\subsection{Data Association and Filtering in GEO}

The data association and filtering algorithms are simulated in various GEO scenarios.
Angles-only measurements are available with higher accuracy; namely the topocentric right ascension and declination are chosen as $\sigma_{\alpha,GEO}=\sigma_{\delta,GEO}=10^{-2}\sigma_\alpha$ and the $j$-th measurement $z_{j,GEO}\in\begin{bmatrix}\alpha, & \delta\end{bmatrix}^T\in\Re^2$ with covariance $R=\diag[\sigma_{\alpha,GEO}^2, \sigma_{GEO,\delta}^2]\in\Re^{2\times2}$.% YOU MAY HAVE TO REMOVE "\\" WITH ANY EDITING CHANGES

\paragraph*{Scenario A2: Comparison of the JPDAF and M-JPDAF with two near-GEO satellites in close proximity}\

These cases serve to compare the JPDAF with the M-JPDAF for tracking two nearby satellites passing over the sensor field of view in GEO, where the orbital radius is $r_{GEO}=4.2164\times10^{4}\ \text{km}$. The initial conditions of the first two orbits are chosen as
\begin{align*}
x_{1,A2}&=\begin{bmatrix}r_{GEO}, & 0\ \text{km}, & 0\ \text{km}, & 0\ \text{km/s}, & \sqrt{\frac{\mu}{r_{GEO}}}, & 0\ \text{km/s}\end{bmatrix}^T,
\\
x_{2,A2}&=x_{1,A2}+\begin{bmatrix}
0\ \text{km}, & 0\ \text{km}, & -1\ \text{km}, & 0\ \text{km/s}, & 0\ \text{km/s}, & 4\times10^{-4}\ \text{km/s}
\end{bmatrix}^T,
\end{align*}
or where $x_{1,A2}$ remains exactly in GEO while $x_{2,A2}$ starts slightly perturbed from a GEO orbit with a nonzero inclination such that the two satellites nearly collide as they pass by in close proximity.
The A2 cases consider small initial covariance matrices $P_{0,A2}= Q$ where $Q=\diag[10^{-6}I_{3\times3}, 10^{-8}I_{3\times3}]$, time steps of $10$ seconds over $5000$ seconds, and $\epsilon=1.1$.
%These scenarios are chosen because the estimates exhibit large initial uncertainties, so the spread in the innovations affects the updates.
Even though the initial conditions are largely different from those of the A1 cases, the A2 cases display the same trends, where the computational costs among the algorithms are roughly the same but the tracking abilities of the M-JPDAF \emph{always} yield estimates closer to the truth with less uncertainty than those of the JPDAF.
These results are tabulated in Table \ref{tab:A2}, and some cases are plotted in Figure \ref{fig:A2}.

\begin{center}
\begin{threeparttable}[h]
\caption{Case A2} \label{tab:A2}
\begin{tabularx}{\textwidth}
{
>{$}c<{$} |
*{2}{>{$}c<{$}} |
*{2}{>{$}c<{$}} |
*{2}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multicolumn{2}{c}{\multirow{1}{*}{RMS Angular Err. ($10^{-6}$rad)}} & \multicolumn{2}{c}{\multirow{1}{*}{Time (sec)}} & \multicolumn{2}{c}{\multirow{1}{*}{Mean $J_P$}} \\
 & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF}
\\
\midrule
\text{A2}  & 1.2140 & 0.98784 & 2.77 & 2.73 & 16.5052 & 16.5015 \\
\text{A2a} & 1.2141 & 0.98780 & 2.74 & 2.74 & 16.5054 & 16.5017 \\
\text{A2b} & \hilight{11.728} & 1.0143 & 2.73 & 2.74 & 18.1526 & 18.1477 \\
\text{A2c} & \hilight{11.728} & 1.0142 & 2.73 & 2.75 & 18.1528 & 18.1479 \\
\text{A2d} & \hilight{11.700} & 1.0332 & 2.82 & 2.77 & 16.511 & 16.5054 \\
\bottomrule
\end{tabularx}
{\small
\begin{tablenotes}
    \item \hilight{Track swapping occurs}
  \end{tablenotes}}
\end{threeparttable}
\end{center}

\begin{figure}
{
%\hspace*{0.05\columnwidth}
\centerline{
	\subfigure[\;Case A2: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2JPDAF.pdf}}
	\subfigure[\;Case A2: M-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2MUJPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case A2a: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2aJPDAF.pdf}}
	\subfigure[\;Case A2a: M-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2aMUJPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case A2b: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2bJPDAF.pdf}}
	\subfigure[\;Case A2b: M-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2bMUJPDAF.pdf}}
	}
}
\caption{The orbital radii of three cases are shown where the dashed plots are the true trajectories and the solid plots are the estimated trajectories and $\alpha_{diff}$ is the difference between $\alpha_{1}$ and the measurement estimates of the two tracks.
Cases A2 and A2a serve as examples where the coalescence during crossing periods with the JPDAF yields much more similar estimates than with the M-JPDAF.
In Case A2b, this coalescence is so severe that track swapping occurs with the JPDAF, but this is avoided with the M-JPDAF.
}\label{fig:A2}
\end{figure}

\paragraph*{Scenario B2: Comparison of the JPDAF, M-JPDAF, and C-JPDAF with two GEO satellites in close proximity}\

The JPDAF, M-JPDAF, and C-JPDAF algorithms are compared for two satellites in GEO with initial conditions
\begin{align}
\begin{split}
x_{1,B2}&=f_{GEO}(0^\circ), \quad x_{2,B2}=f_{GEO}({10^{-4}}^\circ), \quad \mbox{where}
\\
f_{GEO}(\theta)&=r_{GEO}\begin{bmatrix}\cos\theta, & \sin\theta, & 0, & -\sqrt{\frac{\mu}{r_{GEO}^3}}\sin\theta, & \sqrt{\frac{\mu}{r_{GEO}^3}}\cos\theta, & 0\end{bmatrix}^T,
\end{split}
\end{align}
where $Q=\diag[10^{-8}I_{3\times3}\ \text{km}^2, 10^{-10}I_{3\times3}\ \text{km/s}^2]$, $P_{0,B}=Q$, and $\epsilon=1.1$ for time steps of $10$ seconds over a $2000$ second time period.
This corresponds to a scenario where two satellites orbit steadily nearby in GEO.
For the C-JPDAF, the parameters $a=1.5\times10^{2}$ and $c=1\times10^{10}$ from \refeqn{zetaGeneral} and \refeqn{CostGen} are used in these scenarios from trial and error.
In theory, topocentric right ascension measurements should be the same distance apart throughout the cases and the declination should always be zero.
Much like the B1 cases, the accuracy of the C-JPDAF yields a large improvement from that of the JPDAF and M-JPDAF; however, the increase in computational cost of the C-JPDAF is roughly $30$ times that of either other algorithm, which is about half of the increase found with the B1 cases.
This is because the similarity measure is less complex with the B2 cases than with the B1 cases due to the number of estimates and the size of the measurement space.
The angular RMS errors and computation times are tabulated in Table \ref{tab:B2} and some example plots from these cases are shown in Figure \ref{fig:B2}. In applications with a large number of objects to track, the C-JPDAF need not be applied on most object; only those close neighbors where the similarity among the estimates yield coalescence with the JPDAF or the M-JPDAF. This way, a data association scheme with an analytic solution, such as the JPDAF or the M-JPDAF, could be applied to the remaining estimates to mitigate the large computational requirements of the C-JPDAF.

\begin{center}
\begin{threeparttable}[h]
\caption{Case B2} \label{tab:B2}
\begin{tabularx}{0.98\textwidth}
{
>{$}c<{$} |
*{3}{>{$}c<{$}} |
*{3}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{Case} & \multicolumn{3}{c}{\multirow{1}{*}{RMS Angular Err. ($10^{-6}$rad)}} & \multicolumn{3}{c}{\multirow{1}{*}{Time (sec)}} \\
 & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{C-JPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{C-JPDAF}
\\
\midrule
\text{B2}   	& 1.111 & 0.9478 & 0.5671 & 1.15 & 1.12 & 34.3 \\
\text{B2a} 	& 1.111 & 0.9478 & 0.5635 & 1.15 & 1.15 & 34.8 \\
\text{B2b} 	& 1.131 & 0.9829 & 0.5739 & 1.13 & 1.13 & 35.3 \\
\text{B2c} 	& 1.131 & 0.9829 & 0.5644 & 1.12 & 1.13 & 34.8 \\
\text{B2d} 	& 1.112 & 0.9544 & 0.5614 & 1.11 & 1.12 & 35.2 \\
\bottomrule
\end{tabularx}
\begin{tabularx}{.57\textwidth}
{
>{$}c<{$} |
*{3}{>{$}c<{$}}
}
\multirow{2}{*}{Case} & \multicolumn{3}{c}{\multirow{1}{*}{Mean $J_P$}} \\
 & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{C-JPDAF}
\\
\midrule
\text{B2}   	    	& 0.0076169 & 0.0073684 & 0.0073846 \\
\text{B2a} 		& 0.007617 & 0.0073685 & 0.0073879 \\
\text{B2b} 		& 0.0083158 & 0.0080603 & 0.0080799 \\
\text{B2c} 		& 0.0083159 & 0.0080604 & 0.0080812 \\
\text{B2d} 		& 0.0076643 & 0.0074103 & 0.0074331 \\
\bottomrule
\end{tabularx}
\end{threeparttable}
\end{center}

\begin{figure}
{
%\hspace*{0.05\columnwidth}
\centerline{
	\subfigure[\;Case B2: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2JPDAF.pdf}}
	\subfigure[\;Case B2: M-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2MUJPDAF.pdf}}
	\subfigure[\;Case B2: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2C-JPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case B2b: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2bJPDAF.pdf}}
	\subfigure[\;Case B2b: M-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2bMUJPDAF.pdf}}
	\subfigure[\;Case B2b: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2bC-JPDAF.pdf}}
	}
\centerline{
	\subfigure[\;Case B2d: JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2dJPDAF.pdf}}
	\subfigure[\;Case B2d: M-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2dMUJPDAF.pdf}}
	\subfigure[\;Case B2d: C-JPDAF]
		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2dC-JPDAF.pdf}}
	}
}
\caption{In the various cases, the C-JPDAF is always able to produce more distinguishable estimates.
The perturbations among the cases yield slightly different results, but none are drastically different.
}\label{fig:B2}
\end{figure}

The most important result is that the C-JPDAF consistently avoids track-swapping, whereas the JPDAF and M-JPDAF do not.
However unlike the B1 cases, the B2 cases do not have any parameters that affect the coalescence-avoiding optimization significantly.
Therefore, all variations from Case B2 all yield very close results.

Furthermore, coalescence serves as a bias to the soft decision data association algorithms, where the bias is toward nearby tracks. The C-JPDAF also has an inherent bias by including $J_S$ from \refeqn{Js}, except this bias is away from the nearby tracks. While there is no guarantee that these biases cancel, their opposite effects can reduce estimation error and track swapping. If \refeqn{CostGen} is poorly tuned such that $J_S$ is too strong, its repulsive nature may yield increased estimation error and track swapping.

An interesting observation from the data is the nonzero estimated declination, which would be zero without measurement or process noise because the GEO orbit lies on the equatorial plane.
Rather than observing independent errors among the two estimates, the declination errors are correlated and coalescing together.
The correlation of the sets of measurement estimates is easily explained by the soft decisions applied by all of the compared algorithms; measurement updates are largely shared among the estimates, so uncertainties affect the two estimates in a similar way.
Coalescence occurs when measurement updates are shared, shown in all cases in this paper, but the declination errors are less similar with the C-JPDAF.

\paragraph*{Monte Carlo Results of All Scenarios}\

Since data association and filtering are fundamentally stochastic, Cases A1, A2, B1, and B2 are repeated $100$ times with different noise to obtain mean metrics and number of failures.
For these cases, a failure is defined as at least one of the following: (i) track swapping (when estimates are consistently incorrectly associated with measurements from other estimates), (ii) when the estimates become ambiguous with respect to each other (their position relative to each other is consistently closer than half of the true estimate separation), or (iii) a numerical optimization does not converge (C-JPDAF only).
The following Monte Carlo results are tabulated in Table \ref{tab:MonteCarlo}.

\begin{center}
\begin{threeparttable}[h]
\caption{Monte Carlo Results} \label{tab:MonteCarlo}
\begin{tabularx}{0.85\textwidth}
{
>{$}c<{$} | >{$}c<{$} | >{$}c<{$} | >{$}c<{$} | >{$}c<{$} | >{$}c<{$}
%*{1}{>{$}c<{$}} |
%*{2}{>{$}c<{$}} |
%*{2}{>{$}c<{$}}
}
\toprule
\multirow{1}{*}{Approach} & \multirow{1}{*}{Case} & \multirow{1}{*}{RMS Err.} & \multirow{1}{*}{Time (sec)} & \multirow{1}{*}{Mean $J_P$} & \multirow{1}{*}{Failures}
\\
\midrule
\text{JPDAF}        & A1 & 0.12181 & 0.564 & 0.51843 & 0 \\
\text{M-JPDAF}    & A1 & 0.10791 & 0.563 & 0.49490 & 0 \\
\midrule
\text{JPDAF}        & A2 & 8.66749 & 2.764 & 16.5046 & 69 \\
\text{M-JPDAF}    & A2 & 4.76320 & 2.763 & 16.5023 & 36 \\
\midrule
\text{JPDAF}        & B1 & 0.14520 & 1.668 & 0.24859 & 100 \\
\text{M-JPDAF}    & B1 & 0.11004 & 1.671 & 0.17791 & 100 \\
\text{C-JPDAF}\textsuperscript{*}    & B1 & 0.09752 & 117.115 & 0.18032 & 20 \\
\midrule
\text{JPDAF}        & B2 & 1.01311 & 1.107 & 0.00762 & 100 \\
\text{M-JPDAF}    & B2 & 0.84104 & 1.107 & 0.00736 & 100 \\
\text{C-JPDAF}    & B2 & 0.57102 & 31.360 & 0.00738 & 0 \\
\end{tabularx}
{\small
\begin{tablenotes}
	\item RMS error, time, and mean $J_P$ are the mean values of $100$ trials.
	\item Failures are the number of failures out of $100$ trials.
	\item The RMS error unit for Cases A1 and B1 is km, and the RMS error unit for Cases A2 and B2 is $10^{-6}$ rad.
	\item In Case B1 with the C-JPDAF (denoted with \textsuperscript{*}), the $20$ failures occurred when the optimal gains failed to converge, and the metrics are not based on the failed trials.
%    \item *: track swapping
%    \item $\dagger$: ambiguous
  \end{tablenotes}}
\end{threeparttable}
\end{center}

In Case A1, neither the JPDAF nor the M-JPDAF experience failures.
However, these data support that the error and uncertainty cost are reduced on average, without an increase in computation time.
The simulated results of Case A2 also serves to compare these two algorithms.
The computation time and uncertainty cost follow the same trends as Case A1.
However, both algorithms suffer from track swapping, but this happens roughly twice as often with the JPDAF than it does with the M-JPDAF. As a result, the RMS error of the JPDAF is almost twice that of the M-JPDAF.

All three algorithms are compared with Cases B1 and B2. Case B1 is designed as a scenario where both the JPDAF and M-JPDAF yield object estimates with ambiguous estimates, which is considered a failure all $100$ times.
However, when the C-JPDAF converges properly, this failure is mitigated and the RMS error is reduced.
Since the C-JPDAF is solved numerically rather than analytically, a global minimum is not always obtained in four-dimensional measurement space.
The C-JPDAF algorithm converges well every time with Case B2. The C-JPDAF solves a simpler optimization problem, where the number of estimates is reduced from three to two and the measurement space is reduced by two dimensions with angles-only measurements (no range or range-rate).
In every trial, the optimized cost of the C-JPDAF avoids the track swapping and ambiguity of the JPDAF and M-JPDAF algorithms, and in consequence the C-JPDAF RMS error is reduced as well.

\section{Conclusion}
\label{sec:Conclusion}
The conventional JPDAF is a well-known algorithm with various applications in data association.
The soft decision approach makes the algorithm robust to missed detections and other clutter.
This paper enhances the JPDAF in two different ways.
First, the conventional JPDAF employs a Kalman gain, which is suboptimal because it does not minimize the trace of the state covariance matrix.
A minimum uncertainty JPDAF (M-JPDAF) is derived with a new gain that minimizes the full posterior uncertainty, which includes an important spread in the innovations state covariance matrix term.
This gain is computationally comparable to the JPDAF, and shows superior performance in estimate accuracy when multiple measurements serve to update the estimates.

Second, both the JPDAF and the M-JPDAF involve measurement sharing, which tends to coalesce estimates in close proximity.
%From~\cite{KauLovLee14}, t
The C-JPDAF is generalized for any number of estimates and measurements and the similarity cost is found inside the measurement space where coalescence occurs.
When using the C-JPDAF, there must be a trade-off between uncertainty and coalescence based on a weighted cost function, and powerful computational resources are also required.

The results in this paper show how the JPDAF, M-JPDAF, and C-JPDAF perform in LEO and GEO scenarios with various changes in parameters.
Numerous simulations show that the M-JPDAF operates at a similar computational cost to the JPDAF with a consistent improvement in accuracy.
In some scenarios where coalescence of the estimates is expected to be severe, the C-JPDAF can better track objects at the cost of increased computation.

Future work includes determining a systematic method of choosing the parameters for the C-JPDAF and deriving a suboptimal solution for the C-JPDAF that does not involve the large computation of the optimal solution.

\begin{appendix}
\label{append}

\subsection*{Appendix A: M-JPDAF Proofs} \label{NewPartOfEIsPSD}
Note that all object-related variables refer to the $i$-th object, so this subscript is omitted from indexing the object. Its usage in Appendix A is referring to matrix rows.
\begin{itemize}
\item[(i)] From \refeqn{JPDAFCov} and \refeqn{JpIntro} the uncertainty cost for any gain $K$ for each estimate is given by
\begin{align}
\label{eqn:JpAppend}
J_P=\tr{P^--(1-\beta_{0})(KHP^-+P^-H^TK^T)+KEK^T}.
\end{align}
It can be easily seen that the second derivative of $J_P$ is only a function of the third term
\begin{align}
\tr{KEK^T}=\sum\limits_{i=1}^nk_i^TEk_i
\end{align}
where $k_i$ corresponds to the $i$-th row of $K$.
Taking derivatives with respect to $k_i$,
\begin{align}
\deriv{(k_i^TEk_i)}{k_i}=2Ek_i,
\quad
\frac{\partial^2(k_i^TEk_i)}{\partial k_i^2}=2E,
\end{align}
which is positive-definite from (ii).
As the Hessian is positive-definite, the given optimal gain globally minimizes the cost.
Using \refeqn{JPDAFPostCovSymmetricUpdate}, the optimal cost is given by,
%Therefore the only stationary point when $K=(1-\beta_{0})P^-H^TE^{-1}$ globally minimizes $J_P$, which may be simplified with \refeqn{JPDAFPostCovSymmetricUpdate} to obtain the minimum cost $J_P^*\in\Re$
\begin{align}
\label{eqn:OptCost}
J^*_{P} = \tr{P^--KEK^T}.
\end{align}
\item[(ii)] Consider the nontrivial case when a measurement update is possible, i.e., $0\leq\beta_{0}<1$. Then from \refeqn{E}, $E$ is the summation of two positive-definite terms,
\begin{align}
E=\sum_{j=1}^{n_r}\beta^2_{i}R_j+(1-\beta_{0})^2HP^-H^T,
\end{align}
so $E$ is positive-definite. The trivial case when $\beta_0=1$ corresponds to no measurement update, so the positive-definiteness of $E$ is irrelevant.
%Suppose that an object is not measured.
%Let this portion of the innovation update be $e_0=0_{m\times1}$ because no innovation exists where measurements are not available.
%The Total Probability Theorem holds true for object $i$:
%\begin{align}
%\sum\limits_{j=0}^{n_r}\beta_{j}=1.
%\end{align}
%Consider the positive-semidefinite matrix $\Sigma\in\Re^{m\times m}$ from~\cite[Eq. 1.4.16-(1-10)]{ShaRonThi2001},% the $i$-th covariance spread in the means term is defined as% (note that ${\bf e}_{i}$ remains unchanged)
%\begin{align}
%\Sigma=\sum\limits_{j=0}^{n_r}\beta_{j}(e_{j}-{\bar{e}})(e_{j}-{\bar{e}})^T,
%\end{align}
%which may be manipulated as follows,
%\begin{align}
%\Sigma=\sum\limits_{j=0}^{n_r}\beta_{j}e_{j}e_{j}^T
%-\sum\limits_{j=0}^{n_r}\beta_{j}e_{j}{\bar{e}}^T
%-{\bar{e}}\sum\limits_{j=0}^{n_r}\beta_{j}e_{j}^T
%+{\bar{e}}{\bar{e}}^T
%=\sum\limits_{j=1}^{n_r}\beta_{j}e_{j}e_{j}^T-{\bar{e}}{\bar{e}}^T,
%\label{eqn:CovSpreadPSD}
%\end{align}
%because $e_{0}\equiv0$.
%From \refeqn{E},
%\begin{align}
%E=(1-\beta_{0})S+\Sigma,
%\end{align}
%which yields a positive-definite $E$ because $S$ is positive-definite, $\Sigma$ is positive-semidefinite, and $0\leq\beta_{0}<1$.
%Therefore, $E$ is invertible.
\item[(iii)] The JPDAF applies the Kalman gain, which assumes a fixed measurement covariance $R$, i.e., $R=R_j$ for all $j\in\{1,2,...,n_r\}$.
By manipulating \refeqn{E} and using the definition of \refeqn{S}, $S$ and $E$ are related using $\tilde\beta\in\Re$ as
\begin{align}
\label{eqn:EandS}
%S=\frac{E+\tilde\beta R}{(1-\beta_0)^2}
E=(1-\beta_0)^2S-\tilde\beta R
, 
\quad
%S=\frac{E+\tilde\beta R}{(1-\beta_0)^2},
%\quad 
\tilde\beta=
\left(
\sum_{j=1}^{n_r}\beta_j
\right)^2
-\sum_{j=1}^{n_r}\beta_j^2,
\end{align}
where $0\leq\tilde\beta<1$. The Kalman gain is defined as $K_{K}=P^-H^TS^{-1}\in\Re^{n\times m}$~\cite{TrackDataAssoc} and the proposed gain is $K=(1-\beta_0)P^-H^TE^{-1}$.
%The Kalman filter gain $K_{K}=P^-H^TS^{-1}\in\Re^{n\times m}$ from~\cite{TrackDataAssoc} and the M-JPDAF gain $K_{M}=(1-\beta_{0})P^-H^TE^{-1}\in\Re^{n\times m}$ from \refeqn{GainK} are not always equivalent with a proof by contradiction.
Suppose that these gains are identical, i.e., $K=K_K$.
From \refeqn{S} and \refeqn{E}, having identical gains implies
%\begin{align}
%\label{eqn:SequalsE}
%S&=(1-\beta_0)^{-1}E
%\nonumber\\
%R+HP^-H^T&=(1-\beta_0)^{-1}\sum_{j=1}^{n_r}\beta_j^2R_j+(1-\beta_0)HP^-H^T
%.
%\end{align}
\begin{align}
\label{eqn:SequalsE}
K_KE&=KE
%\nonumber\\
%P^-H^TS^{-1}((1-\beta_0)^2S-\tilde\beta R)&=(1-\beta_0)P^-H^T
\nonumber\\
(1-\beta_0)^2P^-H^T-\tilde\beta P^-H^TS^{-1}R&=(1-\beta_0)P^-H^T
.
\end{align}
%This implies
%\begin{align}
%(1-\beta_0)((1-\beta_0)P^-H^T-P^-H^T)&=\tilde\beta P^-H^TS^{-1}R
%\nonumber\\
%-\beta_0(1-\beta_0)P^-H^T&=\tilde\beta P^-H^TS^{-1}R.
%\end{align}
%This implies
Rearranging \refeqn{SequalsE} yields
\begin{align}
\label{eqn:KalMUareSame}
P^-H^T(\beta_0(1-\beta_0)I+\tilde\beta S^{-1}R)=0,
\end{align}
which is satisfied with the case $\tilde\beta=0$. This case implies that $\beta_j=1$ for exactly one $j\in\{1,2,...,n_r\}$. This case corresponds to a hard decision (unity probability between a single measurement and single estimate) where $\beta_0=0$. This case is not satisfied in general, so it follows that the Kalman gain is different in general from the optimal gain proposed in this paper.

%Since the measurement covariance $R$ is independent of a priori covariance $P^-$ and measurement matrix $H$, \refeqn{SequalsE} is only true if $R=R_j$ is fixed for all measurements and exactly one association is unity, i.e., $\beta_j=1$ for $1\leq j\leq n_r$ (excluding the trivial case $\beta_0=1$), where all other association probabilities are zero.
%Neither of these conditions are satisfied in general, so it follows that the Kalman gain is different in general from the optimal gain proposed in this paper.
%Multiplying both sides by $E$ and using \refeqn{E},
%% are equivalent and multiplying both gains by $E$ from \refeqn{E},
%\begin{align}
%KE&=K_{K}E\nonumber
%\\
%(1-\beta_{0})P^-H^T&=
%(1-\beta_{0})P^-H^T+K_{K}\Sigma.
%%({\sum_{j=1}^{n_r} \beta_{j}e_{j}e_{j}^T-{\bar{e}}}{\bar{e}}^T).
%\label{eqn:KalMUareSame}
%\end{align}
%Therefore if \refeqn{KalMUareSame} is true, then $K_{K}\Sigma=0$, which may only occur if the probabilities that each measurement belongs to an object are exactly $0$ or $1$ or $K_{K}\perp\Sigma$.
%None of these conditions are satisfied in general, so it follows that the Kalman gain is different in general from the optimal gain proposed in this paper.
\item[(iv)] From \refeqn{JpAppend}, the uncertainty cost with the Kalman gain $J_{P_{K}}\in\Re$ can be written as
\begin{align}
\label{eqn:CostKal}
J_{P_{K}}
&=
\tr{P^--2(1-\beta_{0})P^-H^TS^{-1}HP^-+P^-H^TS^{-1}ES^{-1}HP^{-1}}\nonumber
\\
%&=
%\tr{P^-+P^-H^TS^{-1}\left(ES^{-1}-2(1-\beta_{0})I\right)HP^-}
%\\
&=
\tr{P^--P^-H^TS^{-1}\left(2(1-\beta_0)S-E\right)S^{-1}HP^-}
.
%\nonumber
%\\
%&=
%\tr{P^-+(1-\beta_{0})^2P^-H^TS^{-1}\left(\frac1{(1-\beta_{0})^2}\left((1-\beta_{i,0})S_i+{\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right) S^{-1}-\frac{2I}{1-\beta_{0}}\right)HP^-}\nonumber
%\\
%&=
%\tr{P^--(1-\beta_{0})^2P^-H^TS^{-1}\left(\frac{1}{1-\beta_{0}}I-\frac1{(1-\beta_{0})^2}\left({\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right) S^{-1}\right)HP^-}
%\nonumber
%\\
%&=
%\tr{P^--(1-\beta_{0})P^-H^TS^{-1}HP^-+P^-H^TS^{-1}\left({\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right) S^{-1}HP^-}\nonumber
%\\
%&=
%\tr{P^--(1-\beta_{0})P^-H^TS^{-1}HP^-+K_K\left({\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right) K_K^T}
\end{align}
Similarly, the optimal cost $J^*_{P}$ may be rewritten as
\begin{align}
\label{eqn:CostOpt}
J^*_{P}
&=
\tr{P^--(1-\beta_{0})^2P^-H^TE^{-1}HP^-}.
%\\
%&=
%\tr{P^--(1-\beta_{0})P^-H^T\left((1-\beta_{0})E^{-1}\right)HP^-}
\end{align}
From \refeqn{CostKal} and \refeqn{CostOpt} and using the relationship between $E$ and $S$ from \refeqn{EandS} repeatedly, the difference $J_{P_K}-J^*_P$can be written as
\begin{align}
J_{P_K}-J^*_P
&=
\tr{P^-H^TS^{-1}\left((1-\beta_{0})^2SE^{-1}S-2(1-\beta_{0})S+E\right)S^{-1}HP^-}
\nonumber
\\
&=\frac1{(1-\beta_0)^2}
\tr{K_K(\beta_0E+\tilde\beta R)E^{-1}(\beta_0E+\tilde\beta R)K_K^T}
%\nonumber
%\\
%&
%=
%\tr{K_K\Sigma E^{-1}\Sigma K_K^T}
\nonumber
\\
&
=\sum\limits_{i=1}^n\nu_i^TE^{-1}\nu_i\geq0,
\end{align}
where $\nu_i$ corresponds to the $i$-th row of $K_K(\beta_0E+\tilde\beta R)$ and $E^{-1}$ is positive-definite because $E$ is positive-definite from (ii).
This implies the inequality $J_{P_{K}}\geq J^*_{P}$.
The only times when $K_K$ becomes optimal are if there exists no measurement origin uncertainty, i.e., $\Sigma=0$, or $K_K\perp\Sigma$, and neither scenario is satisfied in general.
These results are consistent with those obtained from \refeqn{KalMUareSame}.

%For , the following inequality must hold true:
%\begin{align}
%\tr{P^-H^T\left(E^{-1}-\frac{2}{1-\beta_{0}}S^{-1}+\frac1{(1-\beta_{0})^2}S^{-1}ES^{-1}\right)HP^-}\geq0.
%\end{align}
%Let $\Sigma=\sum\limits_{j=1}^{n_r}\beta_{j}e_{j}e_{j}^T-{\bar{e}}{\bar{e}}^T$, which is shown to be positive-semidefinite from \refeqn{CovSpreadPSD}.
%Then by using variations of the definition of $E$ from \refeqn{E} repeatedly, the following expression is obtained,
%\begin{align}
%&
%(1-\beta_{0})^2E^{-1}-2(1-\beta_{0})S^{-1}+S^{-1}ES^{-1}\geq0
%\nonumber
%\\
%&
%(1-\beta_{0})^2E^{-1}-2(1-\beta_{0})S^{-1}+S^{-1}((1-\beta_0)S+\Sigma)S^{-1}\geq0
%\nonumber
%\\
%&
%(1-\beta_{0})E^{-1}(1-\beta_{0})-(1-\beta_{0})S^{-1}+S^{-1}\Sigma S^{-1}\geq0
%\nonumber
%\\
%&
%S^{-1}(E-\Sigma)E^{-1}(E-\Sigma)S^{-1}-(1-\beta_{0})S^{-1}(S+\frac1{(1-\beta_{0})}\Sigma)S^{-1}+2S^{-1}\Sigma S^{-1}\geq0
%\nonumber
%\\
%&
%S^{-1}(E-2\Sigma+\Sigma E^{-1}\Sigma)S^{-1}-S^{-1}ES^{-1}+2S^{-1}\Sigma S^{-1}\geq0
%\nonumber
%\\
%&
%(1-\beta_{0})^2E^{-1}-2(1-\beta_{0})S^{-1}+S^{-1}ES^{-1}=S^{-1}\Sigma E^{-1}\Sigma S^{-1}\geq0.
%\end{align}
\end{itemize}

\subsection*{Appendix B: Derivatives to obtain the C-JPDAF}

The derivative of the similarity cost is obtained as follows.
Let $\Ibd\in \Re^{n\times m}$ be defined such that its $(b,d)$-th element is one and other elements are zero, and let $[K_p]_{bd}\in\Re$ be the $(b,d)$-th element of $K_p$.
Then we have
% define $\mathcal{S}_{pq,p,bd}\in\Re^{m\times m}$ as
\begin{align}
%\mathcal{S}_{pq,p,bd}=&\ 
\deriv{(S^+_{pq})^{-1}}{[K_p]_{bd}}
=-(S^+_{pq})^{-1}
H_p
\deriv{P^+_{p}}{[K_p]_{bd}}
H_p^T
(S^+_{pq})^{-1}
%\nonumber
%\\
%=&\ -{S^+_{pq}}^{-1}
%H_p
%[(1-\beta_{0,p})(\Ibd H_pP_p^-+P_p^-H_p^T\Ibd^T)-\Ibd E_pK_p^T\nonumber
%\\
%&\ -K_pE_p\Ibd^T]
%H_p^T
%{S^+_{pq}}^{-1}
,\label{eqn:SS1}
\end{align}
where $\deriv{P^+_{p}}{[K_p]_{bd}}\in\Re^{n\times n}$ is
\begin{align}
\deriv{P^+_{p}}{[K_p]_{bd}}
=
[(1-\beta_{0,p})(\Ibd H_pP_p^-+P_p^-H_p^T\Ibd^T)-\Ibd E_pK_p^T -K_pE_p\Ibd^T].\label{eqn:SSmidpart}
\end{align}
Using \refeqn{SS1} and \refeqn{SSmidpart}, we have
\begin{align}
%\mathcal{S}_{pq,p,bd}
\deriv{(S^+_{pq})^{-1}}{[K_p]_{bd}}&=-{S^+_{pq}}^{-1}
H_p
[(1-\beta_{0,p})(\Ibd H_pP_p^-+P_p^-H_p^T\Ibd^T)-\Ibd E_pK_p^T\nonumber
\\
& \quad -K_pE_p\Ibd^T]
H_p^T
{S^+_{pq}}^{-1}.
\label{eqn:SS2}
\end{align}

From \refeqn{Up} and \refeqn{SS2}, $\deriv{u_{pq}}{[K_p]_{bd}}\in\Re$ is obtained as
\begin{align}
\deriv{u_{pq}}{[K_p]_{bd}}
&=
(H_p\Ibd{\bar{e}}_p)^T
{S_{pq}^+}^{-1}
(H_pK_p{\bar{e}}_p-\hat z^+_{q})
\nonumber
\\
& \quad
+(H_pK_p{\bar{e}}_p-\hat z^+_{q})^T
\deriv{(S^+_{pq})^{-1}}{[K_p]_{bd}}
%\mathcal{S}_{pq,p,bd}
(H_pK_p{\bar{e}}_p-\hat z^+_{q})\nonumber
\\
& \quad +
(H_pK_p{\bar{e}}_p-\hat z^+_{q})^T
{S_{pq}^+}^{-1}
(H_p\Ibd{\bar{e}}_p)\nonumber
\\
%=&\ \{
%2(H_p\Ibd{\bar{e}}_p)^T
%{S_{pq}}^{-1} +(H_pK_p{\bar{e}}_p-\hat z^+_{q})^T
%\mathcal{S}_{pq,p,bd}
%\}(H_pK_p{\bar{e}}_p-\hat z^+_{q})
%\nonumber
%\\
&=
\{
2(H_p\Ibd{\bar{e}}_p)^T
{S_{pq}^+}^{-1}
-(H_pK_p{\bar{e}}_p-\hat z^+_{q})^T
{S^+_{pq}}^{-1}
H_p
[(1-\beta_{0,p})(\Ibd H_pP_p^-
\nonumber
\\
& \quad
+P_p^-H_p^T\Ibd^T)-\Ibd E_pK_p^T-K_pE_p\Ibd^T]
H_p^T
{S^+_{pq}}^{-1}
\}(H_pK_p{\bar{e}}_p-\hat z^+_{q}),\label{eqn:derivu}
\end{align}

Then from \refeqn{U} and \refeqn{derivu}, we obtain $\deriv{J_S}{[K_p]_{bd}}\in\Re$ as
\begin{align}
\label{eqn:JSK}
\deriv{J_S}{[K_p]_{bd}}&=\ -a\sum\limits_{q,q\neq p}\exp (-au_{pq})\deriv{u_{pq}}{[K_p]_{bd}}
\nonumber
\\
&=\ -a\sum\limits_{q,q\neq p}\exp (-au_{pq})
\{
2(H_p\Ibd{\bar{e}}_p)^T
{S_{pq}^+}^{-1}
\nonumber
\\
& \quad
-(H_pK_p{\bar{e}}_p-\hat z^+_{q})^T
{S^+_{pq}}^{-1}
H_p
[(1-\beta_{0,p})(\Ibd H_pP_p^-
\nonumber
\\
& \quad
+P_p^-H_p^T\Ibd^T)-\Ibd E_pK_p^T\nonumber
\\
& \quad -K_pE_p\Ibd^T]
H_p^T
{S^+_{pq}}^{-1}
\}(H_pK_p{\bar{e}}_p-\hat z^+_{q}).
\end{align}

%\subsection*{Appendix B: The singularity of $E$ is not an infinite limit}
%Consider the singularity when $E_i=0$, which corresponds to no measurements being associated with the $i$-th estimate.
%Let $\beta^*\in\Re$ be the summation of association probabilities of the $i$-th object.
%The singularity of $E_i$ only occurs when $\beta^*=(1-\beta_{i,0})=\sum_{j=1}^{n_r}\beta_j=0$.
%It is clear from \refeqn{KalEst} and \refeqn{CovExpanded} that the updates are independent of $K_i$ exactly at this singularity.
%We show that $K_i$ is \emph{not} growing without bound nearby this singularity, shown by evaluating the limit at this singularity to ensure boundedness.
%Note that if $\sum_{j=1}^{n_r}\beta_j\rightarrow0\Rightarrow \beta_j\rightarrow0\ \forall\ j$ because of the bounds $\sum_{j=1}^{n_r}\beta_j\geq\beta_j\geq0\ \forall \ j$.
%For clarity, let $\tilde\beta_j=(\beta^*-\beta_j)/\beta^*\in\Re$ and drop the subscript $i$ because this analysis focuses on a single track.
%%For clarity, let $\tilde\beta_p=\sum_{j=1,j\neq p}^{n_r}\beta_j/\sum_{j=1}^{n_r}\beta_j\in\Re$ and drop the subscript $i$ because this analysis focuses on a single track.
%
%First, we factor $\beta^*$ from $\sum_{j}^{n_r} \beta_{j}e_{j}e_{j}^T$,
%\begin{align}
%\sum_{j}^{n_r} \beta_{j}e_{j}e_{j}^T=(\beta_1e_1e_1^T+\beta_2e_2e_2^T+\cdots)
%%\\
%%&=\beta^*
%%(e_1e_1^T-\tilde\beta_1e_1e_1^T+
%%e_2e_2^T-\tilde\beta_2e_2e_2^T+\cdots),\nonumber
%=\beta^*(\sum_{j=1}^{n_r}e_{j}e_{j}^T-\sum_{j=1}^{n_r}\tilde\beta_je_{j}e_{j}^T).
%\end{align}
%Similarly, we factor $\beta^*$ from ${{\bar{e}}}{\bar{e}}^T$,
%\begin{align}
%{{\bar{e}}}{\bar{e}}^T
%%&=\sum_{j=1}^{n_r} \beta_{j}e_{j}\sum_{j=1}^{n_r} \beta_{j}e_{j}^T\nonumber%asdf
%%\\
%%&=
%%(\beta_1e_1\sum_{j=1}^{n_r} \beta_{j}e_{j}^T+\beta_2e_2\sum_{j=1}^{n_r} \beta_{j}e_{j}^T+\cdots),\nonumber
%%\\
%%&=\beta^*(e_1\sum_{j=1}^{n_r} \beta_{j}e_{j}^T-\tilde\beta_1e_1\sum_{j=1}^{n_r} \beta_{j}e_{j}^T\nonumber
%%\\
%%&\ \ \ +e_2\sum_{j=1}^{n_r} \beta_{j}e_{j}^T-\tilde\beta_2e_2\sum_{j=1}^{n_r} \beta_{j}e_{j}^T+\cdots),\nonumber
%%\\
%=\beta^*(\sum_{j=1}^{n_r}e_{j}\sum_{j=1}^{n_r}\beta_je_{j}^T-\sum_{j=1}^{n_r}\tilde\beta_je_{j}\sum_{j=1}^{n_r}\beta_je_{j}^T).
%\end{align}
%Now that the summations that compose $E$ are evaluated, their effects on $K$ are analyzed,
%\begin{align}
%%\lim_{\beta^* \to 0}K&=\lim_{\beta^* \to 0}\beta^*P^-H^T
%%(\beta^*[S+\sum_{j=1}^{n_r}e_{j}e_{j}^T
%%-\sum_{j=1}^{n_r}\tilde\beta_je_{j}e_{j}^T-\sum_{j=1}^{n_r}e_{j}\sum_{j=1}^{n_r}\beta_je_{j}^T\nonumber
%%\\
%%&\quad +\sum_{j=1}^{n_r}\tilde\beta_je_{j}\sum_{j=1}^{n_r}\beta_je_{j}^T])^{-1}\nonumber
%%\\
%%&=P^-H^T\lim_{\beta^* \to 0}
%%(S+\sum_{j=1}^{n_r}e_{j}e_{j}^T-\sum_{j=1}^{n_r}\tilde\beta_je_{j}e_{j}^T-\sum_{j=1}^{n_r}e_{j}\sum_{j=1}^{n_r}\beta_je_{j}^T\nonumber
%%\\
%%&\quad +\sum_{j=1}^{n_r}\tilde\beta_je_{j}\sum_{j=1}^{n_r}\beta_je_{j}^T)^{-1}
%\lim_{\beta^* \to 0}K=\lim_{\beta^* \to 0}\beta^*P^-H^T
%\left(\beta^*[S+\sum_{j=1}^{n_r}(1-\tilde\beta_j)e_{j}e_{j}^T
%-\sum_{j=1}^{n_r}(1-\tilde\beta_j)e_{j}\sum_{j=1}^{n_r}\beta_je_{j}^T]\right)^{-1},
%\label{eqn:OptGainNoSingularity}
%\end{align}
%where $\beta^*\to0\Rightarrow\beta_j\to0$ because $0\leq\beta_j\leq\beta^*\leq1$ for any measurement $j$. Therefore,
%\begin{align}
%\lim_{\beta^* \to 0}K=P^-H^T
%\left(S+\sum_{j=1}^{n_r}\left[\lim_{\beta^* \to 0}\frac{\beta_j}{\beta^*}\right]e_{j}e_{j}^T\right)^{-1},
%\end{align}
%which is a finite limit because $S$ is positive-definite, $e_je_j^T$ is positive-semidefinite, and $0\leq\lim\limits_{\beta^* \to 0}\frac{\beta_j}{\beta^*}\leq1$ based on the previous inequality.
%From a computational standpoint, for every $\beta^*_i\approx0$, the effect of the measurement update should be negligible, so we may choose $K_i=0_{n\times m}$ near this singularity as it will have little effect on the filter performance; this typically occurs when the object is not measured.
%Choosing this $K_i$ results in no difference between the a priori and a posteriori state or covariance matrix.

%\subsection*{Appendix C: The optimal gain yields an absolute minimum to the uncertainty cost}

%%%%% Direct method:
%From \refeqn{CovExpanded} and \refeqn{JpIntro} the uncertainty cost for each estimate is given by
%\begin{align}
%\label{eqn:JpAppendixC}
%J_p(K)=\tr{P^--(1-\beta_{0})(KHP^-+P^-H^TK^T)+KEK^T}
%\end{align}
%where the uncertainty cost associated with the M-JPDAF gain $K_{M-JPDAF}=(1-\beta_{0})P^-H^TE^{-1}$ can be easily found using \refeqn{GainK} and \refeqn{PNotSimplified} as
%\begin{align}
%\label{eqn:JpOpt}
%J_{p}(K_{M-JPDAF})=\tr{P^--(1-\beta_{0})^2P^-H^TE^{-1}HP^-
%}.
%\end{align}
%Let $\epsilon\in\Re^{m\times n}$ be an arbitrary matrix such that at least one of its elements is nonzero.
%Consider $\tilde K\in\Re^{n\times m}$ is any gain matrix that is \emph{not} $K_{M-JPDAF}$, i.e., $\tilde K=K_{M-JPDAF}+\epsilon^T=(1-\beta_{0})(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)E^{-1}$.
%Then the posterior covariance from \refeqn{CovExpanded} is
%\begin{align}
%\label{eqn:CovSubOpt}
%P^+(\tilde K)=&\ P^--(1-\beta_{0})^2
%\left(
%(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)E^{-1}HP^-
%+P^-H^TE^{-1}(HP^-+\frac1{(1-\beta_{0})}E\epsilon)
%\right)\nonumber
%\\
%&+(1-\beta_{0})^2\left(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE\right)E^{-1}EE^{-1}\left(HP^-+\frac1{(1-\beta_{0})}E\epsilon\right)\nonumber
%\\
%=&\ P^-+(1-\beta_{0})^2
%\left(
%P^-H^TE^{-1}HP^--P^-H^TE^{-1}HP^--P^-H^TE^{-1}HP^-
%\right)\nonumber
%\\
%&+(1-\beta_{0})(\epsilon^THP^-+P^-H^T\epsilon-\epsilon^THP^--P^-H^T\epsilon)+\epsilon^TE\epsilon\nonumber
%\\
%=&\ P^--(1-\beta_{0})^2P^-H^TE^{-1}HP^-+\epsilon^TE\epsilon
%\end{align}
%
%If $J_{p}(\tilde K)-J_{p}(K_{M-JPDAF})>0\ \forall\ \epsilon$, then $K_{M-JPDAF}$ is the \emph{only} gain matrix that yields an absolute minimum of the cost function $J_{p}$.
%We solve for this inequality with \refeqn{JpOpt} and \refeqn{CovSubOpt},
%\begin{align}
%J_{p}(\tilde K)-J_{p}(K_{M-JPDAF})=&\ \tr{P^--(1-\beta_{0})^2P^-H^TE^{-1}HP^-+\epsilon^TE\epsilon}\nonumber
%\\
%&-\tr{P^--(1-\beta_{0})^2P^-H^TE^{-1}HP^-}\nonumber
%\\
%=&\ \tr{\epsilon^TE\epsilon}\nonumber
%\\
%=&\ \sum\limits_{i=1}^n\epsilon_i^TE\epsilon_i>0
%\end{align}
%because $E$ is a positive-definite symmetric matrix (see Appendices A and B) and where $\epsilon_i\in\Re^{m\times1}$ corresponds to the $i$-th column of $\epsilon$, where at least a single column contains at least a single nonzero element.
%%\begin{align}
%%J_{p}(\tilde K)&=J_{p}((1-\beta_{0})P^-H^TE^{-1}+\epsilon^T)\nonumber
%%\\
%%&=J_{p}((1-\beta_{0})(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)E^{-1}),\nonumber
%%\\
%%&=\tr{P^-}\nonumber
%%\\
%%&\ \ -(1-\beta_{0})^2\tr{(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)HP^-+P^-H^T(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)^T}\nonumber
%%\\
%%&\ \ +(1-\beta_{0})^2\tr{(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)E(P^-H^T+\frac1{(1-\beta_{0})}\epsilon^TE)^T}\nonumber
%%\end{align}

%\subsection*{Appendix D: The optimal gain and Kalman gain are not equivalent}

\end{appendix}

%\section{Section title}
%\label{sec:1}
%Text with citations \cite{RefB} and \cite{RefJ}.
%\subsection{Subsection title}
%\label{sec:2}
%as required. Don't forget to give each section
%and subsection a unique label (see Sect.~\ref{sec:1}).
%\paragraph{Paragraph headings} Use paragraph headings as needed.
%\begin{equation}
%a^2+b^2=c^2
%\end{equation}

% For one-column wide figures use
%\begin{figure}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:1}       % Give a unique label
%\end{figure}
%
% For two-column wide figures use
%\begin{figure*}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics[width=0.75\textwidth]{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:2}       % Give a unique label
%\end{figure*}
%
%% For tables use
%\begin{table}
%% table caption is above the table
%\caption{Please write your table caption here}
%\label{tab:1}       % Give a unique label
%% For LaTeX tables use
%\begin{tabular}{lll}
%\hline\noalign{\smallskip}
%first & second & third  \\
%\noalign{\smallskip}\hline\noalign{\smallskip}
%number & number & number \\
%number & number & number \\
%\noalign{\smallskip}\hline
%\end{tabular}
%\end{table}

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
\bibliography{BibSources}   % name your BibTeX data base

%% Non-BibTeX users please use
%\begin{thebibliography}{}
%%
%% and use \bibitem to create references. Consult the Instructions
%% for authors for reference list style.
%%
%\bibitem{RefJ}
%% Format for Journal Reference
%Author, Article title, Journal, Volume, page numbers (year)
%% Format for books
%\bibitem{RefB}
%Author, Book title, page numbers. Publisher, place (year)
%% etc
%\end{thebibliography}

\end{document}
% end of file template.tex

### 2018-06-09 Journal/JAS Data Association/JAS Data Association Before Rev 1/JAS_DataAssoc.tex
journal added

--- 2018-06-09 Journal/JAS Data Association/JAS Data Association After Rev1 (1)/JAS Data Association Old/JAS Data Association NEW Derivation (1)/JAS_DataAssoc.tex
@@
 First, an analytic solution for an optimal gain that minimizes posterior estimate uncertainty is derived, referred to as the minimum uncertainty JPDAF (M-JPDAF).
-Second, the coalescence-avoiding optimal JPDAF (C-JPDAF) is derived, which removes coalescence by minimizing a weighted sum of the posterior uncertainty and a measure of similarity between estimated probability densities.
+Second, the coalescence-avoiding JPDAF (C-JPDAF) is derived, which removes coalescence by minimizing a weighted sum of the posterior uncertainty and a measure of similarity between estimated probability densities.
 Both novel algorithms are tested in much further depth than any prior work to show how the algorithms perform in various scenarios.
@@
 There are two popular strategies among recursive approaches, namely \emph{hard decisions} and \emph{soft decisions}~\cite{JPDAF1}.
-When the association between an object of interest and a measurement is assumed correct, this is referred to as a hard decision.
+When the association between an object of interest and a measurement is assumed correct and certain, this is referred to as a hard decision.
 %A hard decision is when estimates are fully associated with measurements according to some metric between the estimates and the measurements.
-For example, a common heuristic approach is known as the nearest neighbor filter (NNF), which uses a distance measure between each measurement and the predicted measurements of each potential object to determine likely associations.
+For example, a common heuristic approach is known as the nearest neighbor filter (NNF), which uses a distance (e.g.
+Euclidean, Mahalanobis) between each measurement and the predicted measurements of each potential object to determine likely associations.
 %For example, a common heuristic approach is known as the nearest neighbor filter (NNF), which uses the Mahalanobis distance between each measurement and the predicted measurements of each potential object to determine likely associations.
-The measurement with the smallest Mahalanobis distance is considered as the correct one~\cite{NN2}.
+The measurement with the smallest distance is considered as the correct one~\cite{NN2}.
 However, when estimates are in close proximity, there is a high likelihood of making incorrect associations, which may be detrimental to the estimated states, so this technique is avoided in the analysis of this paper.
@@
 However, the popular JPDAF approaches suffer from a common pitfall, namely sub-optimality.
-More explicitly, Kalman filters are considered optimal in the sense that the estimator gain is selected to minimize the mean square error of the estimated state, or equivalently a measure of posterior uncertainty.
+More explicitly, Kalman filters are considered optimal in the sense that the estimator gain is selected to minimize the mean square error of the estimated state, or equivalently a measure of posterior uncertainty represented by the trace of the covariance.
 %More explicitly, Kalman filters are considered optimal in the sense that the estimator gain is selected to minimize the posterior uncertainties.
-In the current JPDAF applications, the Kalman gain is directly used in the measurement update processes, indeliberately assuming that the Kalman gain designed for a single object tracking is also optimal for multiple object tracking problems.
+In the current JPDAF applications, the Kalman gain is directly used in the measurement update processes, indeliberately assuming that the Kalman gain designed for a single object tracking is also optimal for multiple object tracking problems with soft decisions.
 
@@
 When objects are in close-proximity, a single measurement may be consistent with more than one object, forcing a soft decision.
-When these objects share measurements, their updated states tend to converge toward each other.
+When these objects are estimated by the same measurements, their updated states tend to converge toward each other.
 Most techniques to avoid track coalescence involve pruning the tracks, or designing a set of rules that limit the possible outcomes to likely scenarios.
@@
 
-Therefore, we propose a new coalescence-avoiding optimal JPDAF, referred to as the C-JPDAF, which is a systematic approach to remove coalescence by minimizing the weighted sum of the uncertainty and a similarity index between estimates.
-The prior work from~\cite{KauLovLee14} covers the concept in simple cases, assuming that only two objects are considered with a single measurement at each time step, and that the state and measurement are very similar.
+Therefore, we propose a new coalescence-avoiding JPDAF, referred to as the C-JPDAF, which is a systematic approach to remove coalescence by minimizing the weighted sum of the uncertainty and a similarity index between estimates.
+The prior work from~\cite{KauLovLee14} covers the concept in simple cases, assuming that only two objects are considered with a single measurement at each time step, and that the measurement is one of the state variables.
 These assumptions are unrealistic in most scenarios.
 %Therefore, we generalize the C-JPDAF algorithm in this paper to handle any number of objects and measurements, missed detections and measurements originating from extraneous clutter, and measurements with a much different interpretation than just the object states.
-Therefore, we generalize the C-JPDAF algorithm in this paper to handle any number of objects and measurements, missed detections, measurements originating from extraneous clutter, and various types of measurements.
-
-In short, the main contributions of this paper are (i) deriving the optimal JPDAF that minimizes the posterior uncertainties in estimation, and (ii) proposing a systematic approach to avoid data coalescence of multiple objects in close-proximity.
+Therefore, we generalize the C-JPDAF algorithm in this paper to handle any number of objects and measurements, missed detections, and measurements originating from extraneous clutter.
+
+In short, the main contributions of this paper are (i) deriving the optimal JPDAF that minimizes the posterior mean square error and uncertainties in estimation, and (ii) proposing a systematic approach to avoid data coalescence of multiple objects in close-proximity.
 The desirable features of the proposed approaches are illustrated by extensive numerical simulations with satellites in orbits.
@@
 
-The Gaussian states are tracked with measurements; however, the measurement origins are unknown.
-Suppose that these measurements are unordered and possibly include spurious measurements and missed detections.
-Data association and filtering serves to estimate the state of the systems when there exist uncertainties in the system processes, the measurements, and the measurement originations.
+%The Gaussian states are tracked with measurements; however, 
+We consider the problem of estimating object states, when the number of objects for tracking is assumed known, but the measurement origins are unknown.
+We assume that a sensor generates measurements, composed of only object detections and spurious readings from clutter.
+Multiple measurements may originate from a single object, or an object may be missed entirely.
+However, we assume that a single measurement does not originate from more than one object.
+%We assume accurate measurement resolution, such that at maximum one measurement may originate from a single object, where this measurement is independent of other objects detected by the same sensor.
+%Suppose that these measurements are not associated with a specific object, and they possibly include spurious measurements or missed detections, where the probability of detecting objects is assumed known.
+Data association and filtering serves to estimate the state of the systems when there exist uncertainties in the system processes, measurements, and measurement origins.
 In this paper, the popular joint probabilistic data association filtering is improved in two distinct ways in the subsequent sections.
@@
 
-In this section, we derive the M-JPDAF with a similar framework to that of the JPDAF.
+In this section, we derive the M-JPDAF with a similar framework to that of the JPDAF, based off of the version from~\cite{TrackDataAssoc},~\cite{JPDAF1} for the update structure, and the association probabilities from~\cite{Bar1990}.
 Differences between the JPDAF and M-JPDAF occur in the measurement update where the gain is derived, which changes the posterior covariance equation.
@@
 \begin{align}
-\label{eqn:zhatlin}
+\label{eqn:zlin}
 \hat z_i = H_i\hat x_{i}^-.
 \end{align}
-The innovation, often referred to as the residual, $e_{i,j}\in\Re^m$ is defined as the difference between the $j$-th measurement and the predicted measurement of the $i$-th object, and it is given by
-\begin{align}
-e_{i,j} = z_j - \hat z_i.\label{eqn:eij}
+The innovation, often referred to as the residual, $e_{ij}\in\Re^m$ is defined as the difference between the $j$-th measurement and the predicted measurement of the $i$-th object, and it is given by
+\begin{align}
+e_{ij} = z_j - \hat z_i.\label{eqn:eij}
 \end{align}
@@
 %Given the uncertainty of the measurements represented with covariance matrix $R_i$, t
-For the $j$-th measurement, the $i$-th a priori innovation covariance matrix $S_{i,j}\in\Re^{m\times m}$ represents the uncertainty between the $i$-th measurement estimate and its true measurement, defined as
+%The $i$-th a priori innovation covariance matrix 
+Let $S_i\in\Re^{m\times m}$ be the covariance matrix of the difference between the $i$-th measurement estimate and its true measurement, defined as
 %The a priori innovation covariance matrix $S_{i}\in\Re^{m\times m}$ for all innovations corresponding to the $i$-th object is defined as \EditTL{(Is it for the covariance of $e_{ii}$?
@@
 \begin{align}
-S_{i,j}&=\mathrm{E}[e_{i,j} e_{i,j}^T]=\mathrm{E}[\left((H_ix_i+v_i)-H_i\hat x_{i}^-\right)\left((H_ix_i+v_i)-H_i\hat x_{i}^-\right)^T]\nonumber\\
-&=H_{i}P_{i}^{-}H_{i}^T+R_{j},\label{eqn:S}
-\end{align}
-where it is commonly assumed that $R_j$ is uniform for all measurements, but the subsequent equations make no such assumption.
+S_{i}&=\mathrm{E}[e_{ij} e_{ij}^T]=\mathrm{E}[\left((H_ix_i+v_i)-H_i\hat x_{i}^-\right)\left((H_ix_i+v_i)-H_i\hat x_{i}^-\right)^T]\nonumber\\
+&=H_{i}P_{i}^{-}H_{i}^T+R_{i}.\label{eqn:S}
+\end{align}
 
@@
 
-Next, we define a measurement origination model. Let $\beta_{i,j}\in\Re$ be the probability that the $j$-th measurement originates from the $i$-th object. 
-%Let $\beta_{i,j}$ be the probability that the $i$-th system is associated with measurement $j$.
-%Two techniques may be applied to solve for every $\beta_{i,j}$, both of which require the innovation and its covariance (defined later) and information on detection probabilities, which are assumed known.
+Next, we define a measurement origination model. Let $\beta_{ij}\in\Re$ be the probability that the $j$-th measurement originates from the $i$-th object. 
+%Let $\beta_{ij}$ be the probability that the $i$-th system is associated with measurement $j$.
+%Two techniques may be applied to solve for every $\beta_{ij}$, both of which require the innovation and its covariance (defined later) and information on detection probabilities, which are assumed known.
 The first technique is applied in the standard JPDAF that calculates the true association probabilities from~\cite[Eq.
@@
 However, the computation required with a standard JPDAF is inconsistent with the capabilities of the computational resources to account for an arbitrary number of objects to be tracked~\cite{Bar1990}.
-Several approaches have been proposed to approximate the association probability $\beta_{i,j}$ for the JPDAF.
+Several approaches have been proposed to approximate the association probability $\beta_{ij}$ for the JPDAF.
 This paper adopts the expressions given in~\cite{Bar1990}, and they are summarized as follows.
@@
 
-%A modified version of the JPDA, commonly known as the ``Cheap JPDA'', provides a simpler approximation of the likelihood $\beta_{i,j}$.
+%A modified version of the JPDA, commonly known as the ``Cheap JPDA'', provides a simpler approximation of the likelihood $\beta_{ij}$.
 %The equations below are adopted from~\cite[Eq. 1.(1-2)]{Bar1990}.
 
-Let $G_{i,j}\in\Re$ be a simplified scaled density function between the $i$-th measurement estimate and the $j$-th measurement,
+Let $G_{ij}\in\Re$ be a scaled density function used repeatedly between the $i$-th measurement estimate and the $j$-th measurement, defined as
 % variable which takes a very similar form as the Gaussian probability density function
 \begin{align}
-G_{i,j}=\frac1{\sqrt{\det{S_{i,j}}}}\exp{\{-\frac12e_{i,j}^TS_{i,j}^{-1}e_{i,j}\}}.
+G_{ij}=\frac1{\sqrt{\det{S_{i}}}}\exp{\{-\frac12e_{ij}^TS_{i}^{-1}e_{ij}\}}.
 \end{align}
@@
 %\begin{align}
-%\mathcal{S}_{ti}=\sum\limits_{j\in J}G_{i,j}, 
+%\mathcal{S}_{ti}=\sum\limits_{j\in J}G_{ij}, 
 %\quad 
-%\mathcal{S}_{ri}=\sum\limits_{i\in I}G_{i,j}.
+%\mathcal{S}_{ri}=\sum\limits_{i\in I}G_{ij}.
 %\end{align}
-Then a close approximation of the probability that the $i$-th system is associated with the $j$-th measurement is
+Then, a close approximation of the probability that the $i$-th system is associated with the $j$-th measurement is (from~\cite{Bar1990})
 \begin{align}
 \label{eqn:betaDef}
-\beta_{i,j}=\frac{G_{i,j}}{\sum\limits_{l=1}^{n_r}G_{il}+\sum\limits_{l=1}^{n_t}G_{lj}-G_{i,j}+B},
-\end{align}
-where $B=0$ corresponds to rare extraneous measurements and missed detections.
-When these are prevalent, a fixed value of $B$ found from simulations was shown to perform well in various applications~\cite{Bar1990}.
-In this paper, we assume that at most a single measurement may originate from a single object, and missed detections and measurements originating from extraneous clutter are considered insignificant, so $B=0$ is used to determine joint association probability.
+\beta_{ij}=\frac{G_{ij}}{\sum\limits_{l=1}^{n_r}G_{il}+\sum\limits_{l=1}^{n_t}G_{lj}-G_{ij}+B},
+\end{align}
+where $B$ accounts for measurements \emph{not} originating from known targets.
+When $B=0$, this corresponds to rare extraneous measurements and missed detections.
+When these are prevalent, tuning $B$ is required, though this is beyond the paper scope.
+
+%When these are prevalent, a fixed value of $B$ in \refeqn{betaDef} was found from simulations to provide a close approximation of $\beta_{ij}$ in various applications, though tuning this parameter is beyond the scope of this paper.
+
+%In this paper, we assume that at most a single measurement may originate from a single object, and missed detections and measurements originating from extraneous clutter are considered insignificant, so $B=0$ is used to determine joint association probability.
 
@@
 (i) There is no indication on the dimension of variables.
-For example is $G_{i,j}$ a scalar or a vector; (ii) there are several variables, such as $e_{i,j}$ or $S_i$ that were used before defining them explicitly.
+For example is $G_{ij}$ a scalar or a vector; (ii) there are several variables, such as $e_{ij}$ or $S_i$ that were used before defining them explicitly.
 Saying that it is the innovation and the a priori innovation covariance is not sufficient.
@@
 
-The weighted sum of all innovations associated with the object $i$ is defined as ${\bar{e}}_{i}={\sum\limits_{j=1}^{n_r} \beta_{i,j}e_{i,j}}\in\Re^m$.
+The weighted sum of all innovations associated with the object $i$ is defined as ${\bar{e}}_{i}={\sum\limits_{j=1}^{n_r} \beta_{ij}e_{ij}}\in\Re^m$.
 The innovations serve to update the resulting posterior state estimate $\hat x^+_{i}$ with the gain matrix $K_i\in\Re^{n\times m}$~\cite{TrackDataAssoc},
@@
 \end{align}
-The conventional JPDAF chooses $K_i=P^-_iH_i^TS_i^{-1}$ (assuming $S_i$ is fixed for all measurements) as the gain from the Kalman filter.
+The conventional JPDAF chooses $K_i=P^-_iH_i^TS_i^{-1}$ as the gain from the Kalman filter~\cite{TrackDataAssoc}.
 In this paper, we show that the Kalman gain does not actually minimize the posterior uncertainty in probabilistic data association filtering, and we derive the new \textit{optimal} gain $K_i$.
 
-Let the probability that $i$-th object is \textit{not} associated with any measurement be $\beta_{i,0}=1-{\sum_{j=1}^{n_r} \beta_{i,j}}\in\Re$.
-The case when $\beta_{i,0}=1$ implies that the $i$-th estimate is \emph{not} associated with any measurement completely, i.e., the $i$-th object is not measured by any sensor with the probability of one.
-In such trivial cases, there is no measurement update, i.e., $\hat x^+_i=\hat x^-_i$.
-
-Next we consider when $\beta_{i,0}\neq1$, i.e., $0\leq\beta_{i,0}<1$.
-Let the a priori and a posteriori errors be defined as
-\begin{align}
-\label{eqn:ErrFuns}
-\tilde x_i^-=\hat x_i^--x_i,
-\quad
-\tilde x_i^+=\hat x_i^+-x_i,
-\end{align}
-respectively.
-Then, using \refeqn{zlin}, \refeqn{zhatlin}, \refeqn{eij}, \refeqn{ErrFuns}, and the difference between \refeqn{KalEst} and \refeqn{xestapriori}, we have
-\begin{align}
-\tilde x_i^+&=
-%\hat x_i^+-\hat x_i^-+\tilde x_i^-
-%\nonumber\\&=
-K_i\sum_{j=1}^{n_r}\beta_{i,j}(z_j-\hat z_i)+\tilde x_i^-
-\nonumber\\&=
-%K_i\sum_{j=1}^{n_r}\beta_{i,j}(H_ix_i+v_j-H_i\hat x_i^-)+\tilde x_i^-
-%\nonumber\\&=
-%-K_iH_i\tilde x_i^-\sum_{j=1}^{n_r}\beta_{i,j}+K_i\sum_{j=1}^{n_r}\beta_{i,j}v_j+\tilde x_i^-
-%\nonumber\\&=
-(I-(1-\beta_{i,0})K_iH_i)\tilde x_i^-+K_i\sum_{j=1}^{n_r}\beta_{i,j}v_j.
-\end{align}
-
-The resulting posterior covariance $P^+_{i}\in\Re^{n\times n}$ is defined as
-%~\cite[Eq. D-32 and D-33]{TrackDataAssoc}
+Let the probability that $i$-th object is \textit{not} associated with any measurement be $\beta_{0,i}=1-{\sum_{j=1}^{n_r} \beta_{ij}}\in\Re$.
+The case when $\beta_{0,i}=1$ implies that the $i$-th estimate is \emph{not} associated with any measurement completely, i.e., the $i$-th object is not measured by any sensor with the probability of one.
+In such trivial cases, there is no measurement update, i.e., $\hat x^+_i=\hat x^-_i$ and $P_i^+=P_i^-$.
+
+Next we consider when $\beta_{0,i}\neq1$, i.e., $0\leq\beta_{0,i}<1$.
+The resulting posterior state covariance $P^+_{i}\in\Re^{n\times n}$ is derived in~\cite[Eq.
+D-32 and D-33]{TrackDataAssoc}, where this term is defined as $P^+_{i}=\text{E}[(x_i-\hat x_i^+)(x_i-\hat x_i^+)^T]$ and the resulting solution is derived from the law of total expectation considering each innovation individually.
+Then, $P^+_{i}$ is simplified to
 \begin{align}
 \label{eqn:JPDAFCov}
-P_i^+&=\text{E}[\tilde x_i^+\tilde x_i^{+T}]
-%\nonumber\\&=
-%\text{E}[
-%\{(I-(1-\beta_{i,0})K_iH_i)\tilde x_i^-+K_i\sum_{j=1}^{n_r}\beta_{i,j}v_j\}\{(I-(1-\beta_{i,0})K_iH_i)\tilde x_i^-+K_i\sum_{j=1}^{n_r}\beta_{i,j}v_j\}^T
-%]
-%\nonumber\\&=
-%(I-(1-\beta_{i,0})K_iH_i)\text{E}[\tilde x_i^-\tilde x_i^{-T}](I-(1-\beta_{i,0})K_iH_i)^T
-%+
-%K_i\text{E}[\sum_{j=1}^{n_r}\beta_{i,j}v_j\sum_{j=1}^{n_r}\beta_{i,j}v_j^T]K_i^T
-\nonumber\\&=
-(I-(1-\beta_{i,0})K_iH_i)P_i^-(I-(1-\beta_{i,0})K_iH_i)^T
-+
-K_i(\sum_{j=1}^{n_r}\beta^2_{i,j}R_j)K_i^T
-\nonumber\\&=
-P_i^--(1-\beta_{i,0})
-(K_iH_iP_i^-+P_i^-H_i^TK_i^T)
-+
-K_iE_iK_i^T.
-\end{align}
-where $E_i\in\Re^{m\times m}$ is defined for convenience as
-\begin{align}
-\label{eqn:E}
-E_i=\sum_{j=1}^{n_r}\beta^2_{i,j}R_j+(1-\beta_{i,0})^2H_iP_i^-H_i^T.
-\end{align}
-%$\tilde P_i\in\Re^{n\times n}$ is the portion of the $i$-th state covariance matrix due to the spread in the innovations
-%\begin{align}
-%\tilde P_i=&\ K_i\parenth{{\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T}K_{i}^T,\label{eqn:tildeP}
-%\end{align}
-%and $P_i^c\in\Re^{n\times n}$ corresponds to the $i$-th posterior covariance if only a single measurement serves to update the $i$-th estimate~\cite{TrackDataAssoc}%, and it can be written as
-%%We can define $P_i^c$ without any assumptions on $K_i$ (i.e., $K_i$ is not necessarily the Kalman gain, so the well-known Kalman covariance update equations are not valid here) as~\cite[Eq.
+P^+_{i}=&\ \beta_{0,i}P^-_{i}+(1-\beta_{0,i})P_i^c+\tilde P_i,
+\end{align}
+where $\tilde P_i\in\Re^{n\times n}$ is the portion of the $i$-th state covariance matrix due to the spread in the innovations
+\begin{align}
+\tilde P_i=&\ K_i\parenth{{\sum_{j=1}^{n_r} \beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T}K_{i}^T,\label{eqn:tildeP}
+\end{align}
+and $P_i^c\in\Re^{n\times n}$ corresponds to the $i$-th posterior covariance if only a single measurement serves to update the $i$-th estimate.
+This covariance term is defined as $P^c_i=(I-K_iH_i)P_i^-$ according to~\cite[Eq.
+D-27]{TrackDataAssoc}, or $P^c_i=P_i^--K_iS_iK_i^T$ according to~\cite[Eq.
+43]{JPDAF1}, where both equivalent versions assume that the Kalman gain is used.
+The derivation of $P^c_i$ is the only occurence in \refeqn{JPDAFCov} where the Kalman gain assumption is enforced.
+Without such an assumption, we have%, and it can be written as
+%We can define $P_i^c$ without any assumptions on $K_i$ (i.e., $K_i$ is not necessarily the Kalman gain, so the well-known Kalman covariance update equations are not valid here) as~\cite[Eq.
 4.2-12]{OptEst1}
-%\begin{align}
-%P^c_i=&\ (I-K_iH_i)P^-_i(I-K_iH_i)^T+K_iR_iK_i^T\nonumber
-%\\
-%=&\ P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i(H_iP^-_iH_i^T+R_i)K_i^T,
-%\label{eqn:GenCov}
-%\end{align}
-%which is commonly referred to as the \emph{Joseph form}, valid for \emph{any} gain matrix $K_i$~\cite{OptEst1}.
-%Substituting \refeqn{tildeP} and \refeqn{GenCov} into \refeqn{JPDAFCov},% and rearranging,
-%\begin{align}
-%P^+_{i}=&\ \beta_{i,0}P^-_{i}+(1-\beta_{i,0})[P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\nonumber
-%\\
-%&\ +K_i(H_iP^-_iH_i^T+R_i)K_i^T]+K_i\left({\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right)K_{i}^T\nonumber
-%\\
-%%=&\ P^-_{i}-(1-\beta_{i,0})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i\left((1-\beta_{i,0})S_i+{\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right)K_i^T\nonumber%asdf
-%%\\
-%=&\ P^-_{i}-(1-\beta_{i,0})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_iE_iK_i^T,\label{eqn:CovExpanded}
-%\end{align}
-%where the symmetric matrix $E_i=E_i^T\in\Re^{m\times m}$ is defined as
-%\begin{align}
-%E_i=(1-\beta_{i,0})S_i+{\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T,\label{eqn:E}
-%\end{align}
-%and it corresponds to an innovation covariance that accounts for the possibility that more than a single measurement may update the $i$-th estimate.
-Since \refeqn{JPDAFCov} and \refeqn{E} make no assumptions on $K_i$, this form is valid for \emph{any} gain $K_i$.
+\begin{align}
+P^c_i=&\ (I-K_iH_i)P^-_i(I-K_iH_i)^T+K_iR_iK_i^T\nonumber
+\\
+=&\ P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i(H_iP^-_iH_i^T+R_i)K_i^T,
+\label{eqn:GenCov}
+\end{align}
+which is commonly referred to as the \emph{Joseph form}, valid for \emph{any} gain matrix $K_i$~\cite{OptEst1}.
+Substituting \refeqn{tildeP} and \refeqn{GenCov} into \refeqn{JPDAFCov},% and rearranging,
+\begin{align}
+P^+_{i}=&\ \beta_{0,i}P^-_{i}+(1-\beta_{0,i})[P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\nonumber
+\\
+&\ +K_i(H_iP^-_iH_i^T+R_i)K_i^T]+K_i\left({\sum_{j=1}^{n_r} \beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right)K_{i}^T\nonumber
+\\
+%=&\ P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i\left((1-\beta_{0,i})S_i+{\sum_{j=1}^{n_r} \beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right)K_i^T\nonumber%asdf
+%\\
+=&\ P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_iE_iK_i^T,\label{eqn:CovExpanded}
+\end{align}
+where the symmetric matrix $E_i=E_i^T\in\Re^{m\times m}$ is defined as
+\begin{align}
+E_i=(1-\beta_{0,i})S_i+{\sum_{j=1}^{n_r} \beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T,\label{eqn:E}
+\end{align}
+and it corresponds to an innovation covariance that accounts for the possibility that more than a single measurement may update the $i$-th estimate.
+Because \refeqn{CovExpanded} is based on the Joseph form, this equation yields $P^+_{i}$ for arbitrary gain $K_i$.
 
@@
 \end{align}
-We choose the gain to minimize $J_{p_i}$. The necessary condition for a local minimum is given by
-\begin{align}
-0=\frac{\partial\tr{P^+_i}}{\partial K_i}=-2(1-\beta_{i,0})P^-_iH_i^T+2K_iE_i.\label{eqn:Uncertainty0}
+We choose the gain to minimize $J_{P_i}$, where the necessary condition for a local minimum is given by
+\begin{align}
+0=\frac{\partial\tr{P^+_i}}{\partial K_i}=-2(1-\beta_{0,i})P^-_iH_i^T+2K_iE_i.\label{eqn:Uncertainty0}
 \end{align}
@@
 \begin{align}
-K_i&=(1-\beta_{i,0})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
+K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
 \end{align}
@@
 %\\
-%&=-2(1-\beta_{i,0})\frac{\partial\tr{K_iH_iP^-_i}}{\partial K_i}+\frac{\partial\tr{K_iE_iK_i^T}}{\partial K_i},\nonumber
-%\\
-%&=-2(1-\beta_{i,0})P^-_iH_i^T+2K_iE_i.\label{eqn:Uncertainty0}
+%&=-2(1-\beta_{0,i})\frac{\partial\tr{K_iH_iP^-_i}}{\partial K_i}+\frac{\partial\tr{K_iE_iK_i^T}}{\partial K_i},\nonumber
+%\\
+%&=-2(1-\beta_{0,i})P^-_iH_i^T+2K_iE_i.\label{eqn:Uncertainty0}
 %\end{align}
@@
 %%\begin{align}
-%%K_i&=(1-\beta_{i,0})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
+%%K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
 %%\end{align}
@@
 %\begin{align}
-%K_i&=(1-\beta_{i,0})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
+%K_i&=(1-\beta_{0,i})P^-_iH_i^TE_i^{-1}.\label{eqn:GainK}
 %\end{align}
@@
 
-Finally, \refeqn{GainK} is substituted into \refeqn{JPDAFCov} to obtain a simplified posterior state covariance matrix
+Finally, \refeqn{GainK} is substituted into \refeqn{CovExpanded} to obtain a simplified posterior state covariance matrix
 %\begin{align}
 %\label{eqn:PNotSimplified}
-%P^+_{i}&=P^-_{i}-(1-\beta_{i,0})^2P^-_iH_i^TE_i^{-1}H_iP^-_i,
+%P^+_{i}&=P^-_{i}-(1-\beta_{0,i})^2P^-_iH_i^TE_i^{-1}H_iP^-_i,
 %\end{align}
@@
 \\
-P^+_{i}&=\left(I-(1-\beta_{i,0})K_iH_i\right)P^-_i,\label{eqn:JPDAFPostCovSimpleUpdate}
-\end{align}
-where these can \emph{only} be used with \refeqn{GainK}.
+P^+_{i}&=\left(I-(1-\beta_{0,i})K_iH_i\right)P^-_i.\label{eqn:JPDAFPostCovSimpleUpdate}
+\end{align}
+
 Several properties of the proposed optimal gain are summarized as follows.
@@
 \end{align*}
-\item[(ii)] The matrix $E_i$ given in \refeqn{E} is invertible, or equivalently, the gain $K_i$ is well-defined, except when $\beta_{i,0}=1$, corresponding to the trivial case when $\hat x_i^+=\hat x_i^-$.
+\item[(ii)] The matrix $E_i$ given in \refeqn{E} is invertible, or equivalently, the gain $K_i$ is well-defined, except when $\beta_{0,i}=1$, which corresponds to the trivial case when $\hat x_i^+=\hat x_i^-$ and $P_i^+=P_i^-$.
 % the $i$-th estimate does not receive a measurement update.
@@
 %\begin{align*}
-%\lim_{\beta_{i,0}\rightarrow 1} K_i= 0.
+%\lim_{\beta_{0,i}\rightarrow 1} K_i= 0.
 %\end{align*}
-\item[(iii)] The proposed gain $K_i$ is different from the common Kalman gain, namely $K_{K_i}=P_{i}^{-}H^T S_i^{-1}$, in general.
-Even when the measurement matrix is fixed, i.e., $R_{i,j}=R_i$, each association probability is restricted to $\beta_{i,j}\in\{0,1\}$, which is not satisfied in general.
-%\begin{align*}
-%K_{K_i}\Sigma_i=0, \quad \Sigma_i&=\sum_{j=1}^{n_r}\beta^2_{i,j}R_j+(1-\beta_{i,0})^2H_iP_i^-H_i^T,%{\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}_i}}{\bar{e}_i}^T,
-%\end{align*}
-%or every $\beta_{i,j}$ is exactly $0$ or $1$, which corresponds to filtering when no measurement origin uncertainty exists.
-\item[(iv)] The uncertainty cost of the $i$-th estimate from the Kalman gain satisfies $J_{P_{i,K}}\geq J^*_{P_i}$ because
+\item[(iii)] The proposed gain $K_i$ is different from the common Kalman gain, namely $K_{K_i}=P_{i}^{-}H^T S_i^{-1}$, in general, unless the following condition is satisfied,
 \begin{align*}
-J_{P_{i,K}}-J^*_{P_i}&=\tr{K_{K_i}(\beta_{i,0}E_i+\tilde\beta R_i)E_i^{-1}(\beta_{i,0}E_i+\tilde\beta R_i)K_{K_i}^T}\geq0, \\ \tilde\beta_i&=\left(
-\sum_{j=1}^{n_r}\beta_{i,j}
-\right)^2
--\sum_{j=1}^{n_r}\beta_{i,j}^2.
+K_{K_i}\Sigma_i=0, \quad \Sigma_i&={\sum_{j=1}^{n_r} \beta_{ij}e_{ij}e_{ij}^T-{\bar{e}_i}}{\bar{e}_i}^T,
+\end{align*}
+or every $\beta_{ij}$ is exactly $0$ or $1$, which corresponds to filtering when no measurement origin uncertainty exists.
+\item[(iv)] The uncertainty cost of the $i$-th estimate from the Kalman gain $J_{P_{i,K}}\geq J^*_{P_i}$ because
+\begin{align*}
+J_{P_{i,K}}-J^*_{P_i}&=\tr{K_{K_i}\Sigma_i E_i^{-1}\Sigma_i K_{K_i}^T}\geq0.
 \end{align*}
@@
 
-The Kalman gain employed by JPDAF is optimal if there exists no measurement uncertainty.
-This corresponds to applying a hard decision, where this is the special case when $E_i=S_i$ from \refeqn{E} and $K_i=P^-_iH_i^TS_i^{-1}$ from \refeqn{GainK} is the Kalman gain~\cite{OptEst1}.
-Since the JPDAF may apply soft decisions, it follows that the Kalman gain is suboptimal for the JPDAF in general.
-%These conclusions come from comparing the derived optimal gain and the Kalman gain, and the differences in the uncertainty cost functions associated with these gains, shown later in this paper.
-%
+The Kalman gain employed by JPDAF is optimal if there exists no measurement uncertainty or when the rows of the Kalman gain are orthogonal to the columns of $\sum_{j=1}^{n_r}\beta_{ij}e_{ij}e_{ij}^T-{\bf\bar{e}_i}{\bf\bar{e}_i}^T$ because this is the special case when $E_i=S_i$ from \refeqn{E} and $K_i=P^-_iH_i^TS_i^{-1}$ from \refeqn{GainK} is the Kalman gain~\cite{OptEst1}.
+Since the JPDAF does not require these assumptions, it follows that the Kalman gain is suboptimal for the JPDAF in general.
+These conclusions come from comparing the derived optimal gain and the Kalman gain, and differences in cost functions associated with these gains, shown later in this paper.
+
 %
 %(You may reuse...
-%Note that $E_i$ is independent of $K_i$ and it is invertible when $\beta_{i,0}\neq1$ since $S_i$ is positive-definite and
-%
-%$(\beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}{\bar{e}}_{i}^T)$ is positive-semidefinite, as shown in the Appendix A. 
-%The case when $\beta_{i,0}=1$ corresponds with the $i$-th object \emph{not} being measured with full certainty, and such cases is analyzed in Appendix B.)
+%Note that $E_i$ is independent of $K_i$ and it is invertible when $\beta_{0,i}\neq1$ since $S_i$ is positive-definite and
+%
+%$(\beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}{\bar{e}}_{i}^T)$ is positive-semidefinite, as shown in the Appendix A. 
+%The case when $\beta_{0,i}=1$ corresponds with the $i$-th object \emph{not} being measured with full certainty, and such cases is analyzed in Appendix B.)
 \end{proof}
@@
 
-\section{Coalescence-Avoiding Optimal JPDAF (C-JPDAF)}
+\section{Coalescence-Avoiding JPDAF (C-JPDAF)}
 \label{sec:C-JPDAF}
@@
 %Matusita's measure represents the degree of similarity between two multivariate Gaussian probability distributions~\cite{Mat55},~\cite{Coal_k}.
-The similarity between multiple multivariate Gaussian distributions can be represented by the overlapping regions of their probability density functions.
+%The similarity between multiple multivariate Gaussian distributions can be represented by the overlapping regions of their probability density functions.
 Since there exists no analytic solution for this overlap, we choose Matusita's measure because it provides a scalar approximation of the overlap of multivariate Gaussian distributions~\cite{Mat55},~\cite{Coal_k}.
-Furthermore, this measure is easy to compute and is easily differentiable unlike the Bhattacharyya measure~\cite{Bhatt97}.
+We choose Matusita's measure because it provides a degree of similarity among multivariate Gaussian distributions~\cite{Mat55},~\cite{Coal_k}. 
+This measure is advantageous because of its straightforward computation and differentiable structure unlike the Bhattacharyya measure~\cite{Bhatt97}.
 More explicitly, let $y_i$ be the $i$-th Gaussian random variable, satisfying $y_i \sim \mathcal{N}[\hat y_i,{\mathbf P}_i]$ for $1\leq i \leq n_t$.
@@
 The structure of \refeqn{zetaGeneral} is maintained to generate a cost function for optimal coalescence avoidance.
-\begin{figure}
-\centerline{
-		{\includegraphics[width=.7\columnwidth]{GaussianDistOverlap1D.png}}
-	}
-\caption{A similarity index between three Gaussian distributions is illustrated as summation of overlapping regions of two distributions (green) or three distributions (brown).
-This concept is easily extended to multivariate Gaussian distributions.
-}\label{fig:SimMeas}
-\end{figure}
+%\begin{figure}
+%\centerline{
+%		{\includegraphics[width=.7\columnwidth]{GaussianDistOverlap1D.png}}
+%	}
+%\caption{A similarity index between three Gaussian distributions is illustrated as the summation of overlapping regions of two distributions (green) or three distributions (brown).
+%This concept is easily extended to multivariate Gaussian distributions.
+%}\label{fig:SimMeas}
+%\end{figure}
 
@@
 By considering similarity to obtain the estimator gain, we mitigate some coalescence at the expense of increasing estimate uncertainty according to the magnitude of $c$.
-The choice of $c$ depends on the type of scenario and the parameter $a$ shaping the similarity cost $J_S$.
+The choice of $c$ depends on the type of scenario and the parameter $a$ shapes the similarity cost $J_S$.
 
@@
 \put(-0.5,3){\rotatebox{90}{Cost}}
-\put(0,0){\includegraphics[width=0.48\columnwidth]{ECC14_fig1.pdf}}
+\put(0,0){\includegraphics[width=0.48\columnwidth]{CT.png}}
 \put(5.3,1.2){\shortstack[c]{Uncertainty\\ cost: $J_P$}}
@@
 \put(5.2,5.2){\shortstack[c]{Weighted sum:\\ $\mathbf{J}=J_P+cJ_S$}}
-\put(4.98,0.78){\circle*{0.1}}
-\put(2.48,1.57){\circle*{0.1}}
-\put(4.8,-0.4){$K_{M-JPDAF}$}
-\put(2.3,-0.4){$K_{C-JPDAF}$}
+%\put(4.98,0.78){\circle*{0.1}}
+%\put(2.48,1.57){\circle*{0.1}}
+%\put(5.00,0.81){\circle*{0.1}}
+%\put(2.47,1.60){\circle*{0.1}}
+\put(4.99,0.81){\circle*{0.1}}
+\put(2.47,1.60){\circle*{0.1}}
+\put(4.8,-0.4){$K_\text{M-JPDAF}$}
+\put(2.3,-0.4){$K_\text{C-JPDAF}$}
 \put(3.1,-0.8){Estimator gain}
@@
 }
-\caption{Illustration of optimal coalescence avoidance: several costs are illustrated with respect to an element of an estimator gain $K$ (red: uncertainty cost $J_P$, blue: similarity cost $J_S$, black: weighted sum).
-The estimator gain $K_{M-JPDAF}$ is selected to minimize the uncertainty cost.
-The estimator gain $K_{C-JPDAF}$ is chosen such that the weighted sum of the uncertainty cost and the similarity cost is minimized.
+\caption{Illustration of optimal coalescence avoidance: several costs are illustrated with respect to an element of an estimator gain $K$ (blue dashed: uncertainty cost $J_P$, red dashed-dotted: similarity cost $J_S$, black solid: weighted sum).
+The estimator gain $K_\text{M-JPDAF}$ is selected to minimize the uncertainty cost.
+The estimator gain $K_\text{C-JPDAF}$ is chosen such that the weighted sum of the uncertainty cost and the similarity cost is minimized.
 In short, at the expense of increased uncertainty at a single step, the proposed approach prevents coalescence.
@@
 \label{eqn:CostP}
-\frac{\partial J_{P}}{\partial K_{i}}&=2K_iE_i-2(1-\beta_{i,0})P^-_iH_i^T.
+\frac{\partial J_{P}}{\partial K_{i}}&=2K_iE_i-2(1-\beta_{0,i})P^-_iH_i^T.
 \end{align}
@@
 \paragraph*{Similarity Cost}\ 
-Any shared measurement $z_j$ on measurement estimate $\hat z_i$ yields an innovation $e_{i,j}$ according to \refeqn{eij}.
-During the measurement update, this innovation typically has the net effect of drawing $\hat z_i$ towards $z_j$ if $\beta_{i,j}$ is nonzero.
+Any shared measurement $z_j$ on measurement estimate $\hat z_i$ yields an innovation $e_{ij}$ according to \refeqn{eij}.
+During the measurement update, this innovation typically has the net effect of drawing $\hat z_i$ towards $z_j$ if $\beta_{ij}$ is nonzero.
 This occurs for all estimates and all measurements; coalescence results when measurement estimates are drawn towards common measurements.
@@
 [(1-\beta_{0,p})(\Ibd H_pP_p^-
++P_p^-H_p^T\Ibd^T)
 \nonumber
@@
 & \quad
-+P_p^-H_p^T\Ibd^T)-\Ibd E_pK_p^T\nonumber
-\\
-& \quad -K_pE_p\Ibd^T]
+-\Ibd E_pK_p^T-K_pE_p\Ibd^T]
 H_p^T
@@
 Hence, we require a nonlinear equation solver.
-When $c=0$, we can show that the optimal gain reduces to that of the M-JPDAF, which is used as an initial guess during iterative numerical methods.
-%Therefore, if the number of estimates exhibiting significant similarity is $n_t^*\leq n_t$ and $\mathcal C$ is the number of times \refeqn{NCO} is solved depending on the numerical accuracy, then the required iterations is $n_t^*(n_t^*-1)\mathcal C$, where some variables may be reused (e.g.
+When $c=0$, we can show that the optimal gain reduces to that of the M-JPDAF, which serves as an initial guess during iterative numerical methods.
+When $c>0$, the gains of the C-JPDAF serve to mitigate some coalescence from data association and filtering.
+In some cases, the C-JPDAF prevents cross tagging, when ambiguous estimates yield consistent incorrect associations.
+However, the C-JPDAF is not proven to solve the problem of coalescence completely, so it is possible that the C-JPDAF may cause cross tagging.
+Filter performance is largely governed by tuning the weighted costs, and several numerical scenarios demonstrate the coalescence-avoiding benefits of the C-JPDAF in the subsequent section.%Therefore, if the number of estimates exhibiting significant similarity is $n_t^*\leq n_t$ and $\mathcal C$ is the number of times \refeqn{NCO} is solved depending on the numerical accuracy, then the required iterations is $n_t^*(n_t^*-1)\mathcal C$, where some variables may be reused (e.g.
 $S^+_{pq}\equiv S^+_{qp}$) and often the effect of \refeqn{derivu} is negligible with large estimate separations.
@@
 We consider satellites in the low earth orbit (LEO) and in the geosynchronous orbit (GEO) to compare the performances of the conventional JPDAF, the M-JPDAF, and the C-JPDAF.
-The satellites are subject to the dynamics of the two-body problem.
-Since the system dynamics $\dot x=f(x)$ is nonlinear, the linearized matrix $F_{i_{k-1}}=\deriv{f}{x_i}\bigg|_{x_i=\hat x_{i_{k-1}}^+}$ is used with \refeqn{xestapriori} and \refeqn{Papriori}.
-Similarly, the measurement dynamics $z_i = h(x_i)$ may experience nonlinearities (e.g.
+Although the theoretical contributions are developed for linear systems, the satellites are subject to the nonlinear dynamics of the two-body problem.
+This motivates that the data association and tracking algorithms be extended to nonlinear systems in the same manner that the extended Kalman filter generalizes the Kalman filter for nonlinear systems.
+Using first-order expansions, linear theory of data association and filtering is applied to nonlinear systems.
+%Since the system dynamics $\dot x=f(x)$ is nonlinear, the linearized matrix $F_{i_{k-1}}=\deriv{f}{x_i}\bigg|_{x_i=\hat x_{i_{k-1}}^+}$ is used with \refeqn{xestapriori} and \refeqn{Papriori}.
+%Similarly, the measurement dynamics $z_i = h(x_i)$ may experience nonlinearities (e.g.
 range, range rate, bearing), the linearized matrix $H_i=\deriv{h_i}{x_i}\bigg|_{x_i=\hat x_{i}^-}$ is used with \refeqn{GainK}, \refeqn{JPDAFPostCovSimpleUpdate}, \refeqn{CostP}, \refeqn{Spi}, and \refeqn{Up}.
@@
 %For each following scenario, several cases are simulated where various parameters are slightly changed with $\epsilon\in\Re$, summarized in Table \ref{tab:CaseDef}, and measurement are simulated with \refeqn{NonLin2BP} and \refeqn{zj}.
-In the following scenarios, several cases are simulated with \refeqn{NonLin2BP} and measurements are generated according to \refeqn{zj}.
-For all cases and time steps, measurements may originate from satellites, but this origin is not given to the data association algorithm.
+In the following scenarios, several cases are simulated with \refeqn{NonLin2BP} and measurements are generated according to \refeqn{zj}, but the measurement origins are not given to the data association algorithm.
 Additionally, there is a $10\%$ chance that a measurement is either deleted or an extraneous measurement is added at each time step.
-Further descriptions of the cases are available below, and the resulting accuracy is measured with RMS error, the uncertainty is summarized with $\tr{J_P}$, and the computation time is measured as well, tabulated in Tables \ref{tab:A1}--\ref{tab:B2}.
-Scenarios A1 and B1 are data association and filtering examples in LEO, while scenarios A2 and B2 are GEO examples.
-For readers with interest in how the algorithms handle perturbations, various parameters are slightly changed with $\epsilon\in\Re$, summarized in Table \ref{tab:CaseDef}.
+Since this probability is relatively small, the association probabilities are computed with \refeqn{betaDef} using $B=0$.
+Further descriptions of the cases are available below, and the resulting accuracy over $n$ time steps is measured with RMS position (kilometers) and angular (radian) errors, defined as
+\begin{align}
+% MATLAB: RMS_r_indiv(k) = (sum((Ri{k}-Rhati{k}).^2)/NumStates).^0.5;
+\mathbf{e}_{RMS,r,i}=\sqrt{\frac{\sum_{k=1}^n\norm{r_{i,k}-\hat r_{i,k}}^2}{n}}
+, \quad
+\mathbf{e}_{RMS,\theta,i}=\sqrt{\frac{\sum_{k=1}^n(\cos^{-1}(r_{i,k}^T\hat r_{i,k}))^2}{n}}
+,
+\end{align}
+respectively, and the subsequent tabulated RMS errors are averaged over all objects.
+The uncertainty is summarized with $\tr{J_P}$, and the computation time is measured as well, tabulated in Tables \ref{tab:I}--\ref{tab:IV}.
+Scenarios I and II are data association and filtering examples in LEO, while scenarios III and IV are GEO examples.
+For readers with interest in how the algorithms handle perturbations, various parameters are slightly changed, tabulated at each scenario description.
 The metrics from the perturbed cases are summarized in the subsequent tables as well.
 
+%\begin{table}
+%\begin{center}
+%\caption{Case Definitions} \label{tab:CaseDef}
+%\begin{threeparttable}[h]
+%\begin{tabularx}{.65\textwidth}
+%{
+%>{$}c<{$} |
+%>{$}c<{$}
+%}
+%\toprule
+%\multirow{2}{*}{Symbol} & \multirow{2}{*}{Explanation}\\
+%\\
+%\midrule
+%\multirow{1}{*}{A} &  \multirow{1}{*}{Comparison of JPDAF and M-JPDAF}
+%\\
+%\multirow{1}{*}{B} &  \multirow{1}{*}{Comparison of JPDAF, M-JPDAF, and C-JPDAF}
+%\\
+%\multirow{1}{*}{1} &  \multirow{1}{*}{Orbits in LEO}
+%\\
+%\multirow{1}{*}{2} &  \multirow{1}{*}{Orbits in GEO}
+%\\
+%\multirow{1}{*}{a} &  \multirow{1}{*}{$\sigma^2_{\rho,a} = \epsilon\sigma^2_{\rho}$}
+%\\
+%\multirow{1}{*}{b} &  \multirow{1}{*}{$\sigma^2_{\dot \rho,b} = \epsilon\sigma^2_{\dot \rho}$}
+%\\
+%\multirow{1}{*}{c} &  \multirow{1}{*}{$\sigma^2_{\rho,c} = \epsilon\sigma^2_{\rho}$, $\sigma^2_{\dot \rho,c} = \epsilon\sigma^2_{\dot \rho}$}
+%\\
+%\multirow{1}{*}{d} &  \multirow{1}{*}{$\sigma^2_{\alpha,d} = \epsilon\sigma^2_{\alpha}$, $\sigma^2_{\delta,d} = \epsilon\sigma^2_{\delta}$}
+%\\
+%\multirow{1}{*}{e} &  \multirow{1}{*}{$\sigma^2_{\mathbf{\rho},e} = \epsilon\sigma^2_{\mathbf{\rho}}$ (scenario 1 only)}
+%\\
+%\multirow{1}{*}{f} &  \multirow{1}{*}{$\sigma^2_{\mathbf{\dot\rho},f} = \epsilon\sigma^2_{\mathbf{\dot\rho}}$ (scenario 1 only)}
+%\\
+%\bottomrule
+%\end{tabularx}
+%\end{threeparttable}
+%\end{center}
+%\end{table}
+
+\subsection{Data Association and Filtering in LEO}
+
+Various cases in LEO are simulated to compare the data association and filtering algorithms, namely the JPDAF, M-JPDAF, and C-JPDAF.
+The process covariance is $Q=\diag[\sigma^2_rI_{3\times3},\dot\sigma^2_rI_{3\times3}]=\diag[10^{-6}I_{3\times3}\ \text{km}^2, 10^{-8}I_{3\times3}\ (\text{km/s})^2]$ and the measurement uncertainty is $R=\diag[\sigma_\rho^2, \sigma_{\dot\rho}^2, \sigma_\alpha^2, \sigma_\delta^2]=\diag[(0.15\ \text{km})^2, (0.005\ \text{km/s})^2,\\ (\frac{0.01\pi}{180}\ \text{rad})^2, (\frac{0.01\pi}{180}\ \text{rad})^2]\in\Re^{4\times4}$ where these standard deviations are selected from the Mahe Ground Station uncertainty data~\cite{VerSauSco04} and~\cite{KurAriAriEfe10}.
+
+\paragraph*{Scenario I: Comparison of the JPDAF and M-JPDAF with two LEO satellites in close proximity}\
+
+The first set of cases compares the JPDAF with the M-JPDAF for tracking two nearby satellites passing over the sensor field of view in LEO. The initial conditions of the first two orbits are chosen as
+\begin{align}
+x_{1,\text{I}}&=x_{ref,\text{I}}+x_{pert,\text{I}}, \quad x_{2,\text{I}}=x_{ref,\text{I}}-x_{pert,\text{I}},\nonumber
+\\
+x_{ref,\text{I}}&=\begin{bmatrix}7000\ \text{km}, & 0\ \text{km}, & 1000\ \text{km}, & 0\ \text{km/s}, & 7.546\ \text{km/s}, & 0\ \text{km/s}\end{bmatrix}^T,
+\nonumber\\
+x_{pert,\text{I}}&=\begin{bmatrix}
+0.2\ \text{km}, & 0\ \text{km}, & 0\ \text{km}, & 0\ \text{km/s}, & -0.105\ \text{km/s}, & 0\ \text{km/s}
+\end{bmatrix}^T.\nonumber
+\end{align}
+The perturbations are summarized in Table \ref{tab:PertCaseI}.
 \begin{table}
 \begin{center}
-\caption{Case Definitions} \label{tab:CaseDef}
+\caption{Perturbations for Scenario I Cases} \label{tab:PertCaseI}
 \begin{threeparttable}[h]
-\begin{tabularx}{.65\textwidth}
+\begin{tabularx}{.75\textwidth}
 {
@@
 \toprule
-\multirow{2}{*}{Symbol} & \multirow{2}{*}{Explanation}\\
+\multirow{2}{*}{Perturbed Case} & \multirow{2}{*}{Perturbation}\\
 \\
 \midrule
-\multirow{1}{*}{A} &  \multirow{1}{*}{Comparison of JPDAF and M-JPDAF}
-\\
-\multirow{1}{*}{B} &  \multirow{1}{*}{Comparison of JPDAF, M-JPDAF, and C-JPDAF}
-\\
-\multirow{1}{*}{1} &  \multirow{1}{*}{Orbits in LEO}
-\\
-\multirow{1}{*}{2} &  \multirow{1}{*}{Orbits in GEO}
-\\
-\multirow{1}{*}{a} &  \multirow{1}{*}{$\sigma^2_{\rho,a} = \epsilon\sigma^2_{\rho}$}
-\\
-\multirow{1}{*}{b} &  \multirow{1}{*}{$\sigma^2_{\dot \rho,b} = \epsilon\sigma^2_{\dot \rho}$}
-\\
-\multirow{1}{*}{c} &  \multirow{1}{*}{$\sigma^2_{\rho,c} = \epsilon\sigma^2_{\rho}$, $\sigma^2_{\dot \rho,c} = \epsilon\sigma^2_{\dot \rho}$}
-\\
-\multirow{1}{*}{d} &  \multirow{1}{*}{$\sigma^2_{\alpha,d} = \epsilon\sigma^2_{\alpha}$, $\sigma^2_{\delta,d} = \epsilon\sigma^2_{\delta}$}
-\\
-\multirow{1}{*}{e} &  \multirow{1}{*}{$\sigma^2_{\mathbf{\rho},e} = \epsilon\sigma^2_{\mathbf{\rho}}$ (scenario 1 only)}
-\\
-\multirow{1}{*}{f} &  \multirow{1}{*}{$\sigma^2_{\mathbf{\dot\rho},f} = \epsilon\sigma^2_{\mathbf{\dot\rho}}$ (scenario 1 only)}
+\multirow{1}{*}{I.a} &  \multirow{1}{*}{$\sigma^2_{r} = 10^{-5}\text{ km}^2$}
+\\
+\multirow{1}{*}{I.b} &  \multirow{1}{*}{$\sigma^2_{\dot r} = 10^{-7}(\text{ km/s})^2$/s}
+\\
+\multirow{1}{*}{I.c} &  \multirow{1}{*}{$\sigma^2_{r} = 10^{-5}\text{ km}^2$, $\sigma^2_{\dot r} = 10^{-7}(\text{ km/s})^2$}
+\\
+\multirow{1}{*}{I.d} &  \multirow{1}{*}{$\sigma^2_{\alpha} = \sigma^2_{\delta} = 3.0462\times10^{-7}\text{ rad}^2$}
+\\
+\multirow{1}{*}{I.e} &  \multirow{1}{*}{$\sigma^2_{\mathbf{\rho}} = 2.25\times10^{-1}\text{ km}^2$}
+\\
+\multirow{1}{*}{I.f} &  \multirow{1}{*}{$\sigma^2_{\mathbf{\dot\rho}} = 2.5\times10^{-4}(\text{ km/s})^2$}
 \\
@@
 \end{table}
-
-\subsection{Data Association and Filtering in LEO}
-
-Various cases in LEO are simulated to compare the data association and filtering algorithms, namely the JPDAF, M-JPDAF, and C-JPDAF.
-The measurement uncertainty is $R=\diag[\sigma_\rho^2, \sigma_{\dot\rho}^2, \sigma_\alpha^2, \sigma_\delta^2]=\\\diag[(0.15\ \text{km})^2, (0.005\ \text{km/s})^2, (\frac{0.01\pi}{180}\ \text{rad})^2, (\frac{0.01\pi}{180}\ \text{rad})^2]\in\Re^{4\times4}$ where these standard deviations are selected from the Mahe Ground Station uncertainty data~\cite{VerSauSco04} and ~\cite{KurAriAriEfe10}.
-
-\paragraph*{Scenario A1: Comparison of the JPDAF and M-JPDAF with two LEO satellites in close proximity}\
-
-The first set of cases compares the JPDAF with the M-JPDAF for tracking two nearby satellites passing over the sensor field of view in LEO. The initial conditions of the first two orbits are chosen as
-\begin{align}
-x_{1,A1}&=x_{ref,A1}+x_{pert,A1}, \quad x_{2,A1}=x_{ref,A1}-x_{pert,A1},\nonumber
-\\
-x_{ref,A1}&=\begin{bmatrix}7000\ \text{km}, & 0\ \text{km}, & 1000\ \text{km}, & 0\ \text{km/s}, & 7.546\ \text{km/s}, & 0\ \text{km/s}\end{bmatrix}^T,
-\nonumber\\
-x_{pert,A1}&=\begin{bmatrix}
-0.2\ \text{km}, & 0\ \text{km}, & 0\ \text{km}, & 0\ \text{km/s}, & -0.105\ \text{km/s}, & 0\ \text{km/s}
-\end{bmatrix}^T.\nonumber
-\end{align}
-In these cases, the process covariance is $Q=\diag[10^{-6}I_{3\times3}\ \text{km}^2, 10^{-8}I_{3\times3}\ \text{km/s}^2]$, the estimates have initial covariances matrix $P_{0,A1}=10^7\times Q$, time steps of $1$ second over $100$ seconds, and $\epsilon=10$.
-These scenarios are chosen because the estimates exhibit large initial uncertainties, so there are soft decisions with large measurements shared and the portion of the posterior covariance affected by \refeqn{E} is significant.
+In these cases, the estimates have initial covariances matrix $P_{0,\text{I}}=10^7\times Q$, and time steps are $1$ second over $100$ seconds.
+These scenarios are chosen because the estimates exhibit large initial uncertainties, so the spread in the innovations term from \refeqn{tildeP} affects the updates.
 The gain of the M-JPDAF accounts for this portion of the state covariance matrix better than the gain for the conventional JPDAF in every single case.
@@
 Given identical parameters and measurements, the RMS errors of the position and the mean uncertainty cost $J_P$ defined in \refeqn{JP} are lower in \emph{every} case when the M-JPDAF is applied compared with when the conventional JPDAF is applied.
-In fact in two cases, namely A1b and A1c, the JPDAF is incapable of decreasing the estimate uncertainty as quickly as the M-JPDAF.
-In some cases, the JPDAF suffers from track swapping, which is when measurements are consistently associated incorrectly.
+%In fact in two cases, namely A1b and A1c, the JPDAF is incapable of decreasing the estimate uncertainty as quickly as the M-JPDAF.
+In two cases, namely I.b and I.c, the JPDAF suffers from track swapping, which is when measurements are consistently associated incorrectly.
 %; the JPDAF initially suffers from such great coalescence that the estimates become associated with incorrect measurements consistently, known as track swapping.
 Furthermore, the computational time is roughly the same between the JPDAF and M-JPDAF, so the results suggest that the improvement in accuracy is at no increase in computation.
-These results are tabulated in Table \ref{tab:A1}, and some cases are plotted in Figure \ref{fig:A1}.
+These results are tabulated in Table \ref{tab:I}, and some cases are plotted in Figure \ref{fig:I}.
 
@@
 \begin{threeparttable}[h]
-\caption{Cases of A1} \label{tab:A1}
+\caption{Results of Scenario I Cases} \label{tab:I}
 \begin{tabularx}{\textwidth}
@@
 \midrule
-\text{A1}  & 0.12026 & 0.10646 & 0.582 & 0.561 & 0.5125 & 0.49054 \\
-\text{A1a} & 0.17910 & 0.17157 & 0.558 & 0.561 & 4.6154 & 4.5607 \\
-\text{A1b} & \hilight{2.63710} & 0.14150 & 0.562 & 0.559 & 1.9718 & 0.68328 \\
-\text{A1c} & \hilight{2.63520} & 0.19559 & 0.570 & 0.579 & 5.7223 & 4.8463 \\
-\text{A1d} & 0.20204 & 0.18865 & 0.573 & 0.578 & 0.59901 & 0.57227 \\
-\text{A1e} & 0.14541 & 0.13361 & 0.576 & 0.574 & 0.53064 & 0.50884 \\
-\text{A1f} & 0.12175 & 0.10795 & 0.576 & 0.560 & 0.51338 & 0.49086 \\
+\text{I}  & 0.12026 & 0.10646 & 0.582 & 0.561 & 0.5125 & 0.49054 \\
+\text{I.a} & 0.17910 & 0.17157 & 0.558 & 0.561 & 4.6154 & 4.5607 \\
+\text{I.b} & \hilight{2.63710} & 0.14150 & 0.562 & 0.559 & 1.9718 & 0.68328 \\
+\text{I.c} & \hilight{2.63520} & 0.19559 & 0.570 & 0.579 & 5.7223 & 4.8463 \\
+\text{I.d} & 0.20204 & 0.18865 & 0.573 & 0.578 & 0.59901 & 0.57227 \\
+\text{I.e} & 0.14541 & 0.13361 & 0.576 & 0.574 & 0.53064 & 0.50884 \\
+\text{I.f} & 0.12175 & 0.10795 & 0.576 & 0.560 & 0.51338 & 0.49086 \\
 \bottomrule
@@
 {
-%\hspace*{0.05\columnwidth}
+%\hspace*{0.05\colu0.33mnwidth}
 \centerline{
-	\subfigure[\;Case A1: JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1JPDAF.pdf}}
-	\subfigure[\;Case A1: M-JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1MUJPDAF.pdf}}
+	\subfigure[\;Case I: JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{A1JPDAF.pdf}}
+	\subfigure[\;Case I.b: JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{A1bJPDAF.pdf}}
+	\subfigure[\;Case I.d: JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{A1dJPDAF.pdf}}
 	}
 \centerline{
-	\subfigure[\;Case A1b: JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1bJPDAF.pdf}}
-	\subfigure[\;Case A1b: M-JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1bMUJPDAF.pdf}}
-	}
-\centerline{
-	\subfigure[\;Case A1d: JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1dJPDAF.pdf}}
-	\subfigure[\;Case A1d: M-JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A1dMUJPDAF.pdf}}
-	}
-}
-\caption{The orbital radii of three cases are shown where the dashed plots are the true trajectories and the solid plots are the estimated trajectories.
-In all of these cases, the M-JPDAF is more accurate and avoids track swapping (Case A1b in particular).
+	\subfigure[\;Case I: M-JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{A1MUJPDAF.pdf}}
+	\subfigure[\;Case I.b: M-JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{A1bMUJPDAF.pdf}}
+	\subfigure[\;Case I.d: M-JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{A1dMUJPDAF.pdf}}
+	}
+}
+\caption{Scenario I: The orbital radii of three cases are shown where the dashed plots are the true trajectories and the solid plots are the estimated trajectories.
+In all of these cases, the M-JPDAF is more accurate and avoids track swapping (Case I.b in particular).
 % experiences less coalescence than the JPDAF.
 %In particular, when the JPDAF is used in Case A1b, the initial coalescence causes track swapping.
-Additionally, Case A1d experiences the most inaccurate estimation without track swapping with very noisy angle measurements, and the M-JPDAF has lower RMS error when compared to the JPDAF.
-}\label{fig:A1}
+Additionally, Case I.d experiences the most inaccurate estimation without track swapping with very noisy angle measurements, and the M-JPDAF has lower RMS error when compared to the JPDAF.
+}\label{fig:I}
 \end{figure}
 
-\paragraph*{Scenario B1: Comparison of the JPDAF, M-JPDAF, and C-JPDAF with three LEO satellites in close proximity}\
+\newpage
+
+\paragraph*{Scenario II: Comparison of the JPDAF, M-JPDAF, and C-JPDAF with three LEO satellites in close proximity}\
 
@@
 \begin{align}
-x_{1,B1}&=x_{ref,B1}+x_{pert,B1}, \quad x_{2,B1}=x_{ref,B1}-x_{pert,B1}, \quad x_{3,B1}=x_{ref,B1},\nonumber
-\\
-x_{ref,B1}&=\begin{bmatrix}7000\ \text{km}, & 0\ \text{km}, & 0\ \text{km}, & 0\ \text{km/s}, & 7.546\ \text{km/s}, & 0\ \text{km/s}\end{bmatrix}^T,
+x_{1,\text{II}}&=x_{ref,\text{II}}+x_{pert,\text{II}}, \quad x_{2,\text{II}}=x_{ref,\text{II}}-x_{pert,\text{II}}, \quad x_{3,\text{II}}=x_{ref,\text{II}},\nonumber
+\\
+x_{ref,\text{II}}&=\begin{bmatrix}7000\ \text{km}, & 0\ \text{km}, & 0\ \text{km}, & 0\ \text{km/s}, & 7.546\ \text{km/s}, & 0\ \text{km/s}\end{bmatrix}^T,
 \nonumber\\
-x_{pert,B1}&=\begin{bmatrix}
+x_{pert,\text{II}}&=\begin{bmatrix}
 0.025\ \text{km}, & 0\ \text{km}, & 0\ \text{km}, & 0\ \text{km/s}, & 0\ \text{km/s}, & 0\ \text{km/s}
-\end{bmatrix}^T,\nonumber
-\end{align}
-where $Q=\diag[10^{-6}I_{3\times3}\ \text{km}^2, 10^{-8}I_{3\times3}\ \text{km/s}^2]$, $P_{0,B1}=Q$, and $\epsilon=1.1$ for time steps of $10$ seconds over a $2000$ second time period.
-This corresponds to a phased-array radar observing observing a slow separation of LEO objects in a near-circular orbit; because the satellite states are in close proximity over a large time period, the JPDAF and M-JPDAF are subject to severe coalescence.
+\end{bmatrix}^T.\nonumber
+\end{align}
+The perturbations are summarized in Table \ref{tab:PertCaseII}.
+\begin{table}
+\begin{center}
+\caption{Perturbations for Scenario II Cases} \label{tab:PertCaseII}
+\begin{threeparttable}[h]
+\begin{tabularx}{.75\textwidth}
+{
+>{$}c<{$} |
+>{$}c<{$}
+}
+\toprule
+\multirow{2}{*}{Perturbed Case} & \multirow{2}{*}{Perturbation}\\
+\\
+\midrule
+\multirow{1}{*}{II.a} &  \multirow{1}{*}{$\sigma^2_{r} = 1.1\times10^{-6}\text{ km}^2$}
+\\
+\multirow{1}{*}{II.b} &  \multirow{1}{*}{$\sigma^2_{\dot r} = 1.1\times10^{-8}(\text{ km/s})^2$}
+\\
+\multirow{1}{*}{II.c} &  \multirow{1}{*}{$\sigma^2_{r} = 1.1\times10^{-6}\text{ km}^2$, $\sigma^2_{\dot r} = 1.1\times10^{-8}(\text{ km/s})^2$}
+\\
+\multirow{1}{*}{II.d} &  \multirow{1}{*}{$\sigma^2_{\alpha} = \sigma^2_{\delta} = 3.3508\times10^{-8}\text{ rad}^2$}
+\\
+\multirow{1}{*}{II.e} &  \multirow{1}{*}{$\sigma^2_{\mathbf{\rho}} = 2.48\times10^{-2}\text{ km}^2$}
+\\
+\multirow{1}{*}{II.f} &  \multirow{1}{*}{$\sigma^2_{\mathbf{\dot\rho}} = 2.75\times10^{-5}(\text{ km/s})^2$}
+\\
+\bottomrule
+\end{tabularx}
+\end{threeparttable}
+\end{center}
+\end{table}
+The initial state covariance is chosen as $P_{0,\text{II}}=Q$ for time steps of $10$ seconds over a $2000$ second time period.
+This corresponds to a phased-array radar observing a slow separation of LEO objects in a near-circular orbit; because the satellite states are in close proximity over a large time period, the JPDAF and M-JPDAF are subject to severe coalescence.
 The C-JPDAF is designed to avoid this coalescence, where the parameters $a=5\times10^{-1}$ and $c=2.3\times10^{-4}$ from \refeqn{zetaGeneral} and \refeqn{CostGen} are used in these scenarios from trial and error.
 The accuracy of the C-JPDAF is largely improved from that of the JPDAF and M-JPDAF; however, the computational cost of the C-JPDAF is roughly $60$ times that of either other algorithm.
-The position RMS errors and computation times are tabulated in Table \ref{tab:B1} and example plots from these cases are shown in Figure \ref{fig:B1}.
+The position RMS errors and computation times are tabulated in Table \ref{tab:II} and example plots from these cases are shown in Figure \ref{fig:II}.
 
@@
 \begin{threeparttable}[h]
-\caption{Case B1} \label{tab:B1}
+\caption{Results of Scenario II Cases} \label{tab:II}
 \begin{tabularx}{\textwidth}
@@
 \midrule
-\text{B1}   	& 0.14432 & 0.091133 & 0.057312 & 1.74 & 1.67 & 101 \\
-\text{B1a} 	& 0.14432 & 0.091134 & 0.057307 & 1.69 & 1.73 & 101 \\
-\text{B1b} 	& 0.14694 & 0.094506 & 0.055145 & 1.67 & 1.68 & 101 \\
-\text{B1c} 	& 0.14693 & 0.094507 & 0.055136 & 1.68 & 1.67 & 101 \\
-\text{B1d} 	& 0.14400 & 0.088933 & 0.058911 & 1.66 & 1.66 & 103 \\
-\text{B1e} 	& 0.14408 & 0.090823 & 0.059919 & 1.67 & 1.67 & 100 \\
-\text{B1f} 	& 0.14436 & 0.091212 & 0.064299 & 1.73 & 1.72 & 98.7 \\
+\text{II}   	& 0.14432 & 0.091133 & 0.057312 & 1.74 & 1.67 & 101 \\
+\text{II.a} 	& 0.14432 & 0.091134 & 0.057307 & 1.69 & 1.73 & 101 \\
+\text{II.b} 	& 0.14694 & 0.094506 & 0.055145 & 1.67 & 1.68 & 101 \\
+\text{II.c} 	& 0.14693 & 0.094507 & 0.055136 & 1.68 & 1.67 & 101 \\
+\text{II.d} 	& 0.14400 & 0.088933 & 0.058911 & 1.66 & 1.66 & 103 \\
+\text{II.e} 	& 0.14408 & 0.090823 & 0.059919 & 1.67 & 1.67 & 100 \\
+\text{II.f} 	& 0.14436 & 0.091212 & 0.064299 & 1.73 & 1.72 & 98.7 \\
 \bottomrule
@@
 \midrule
-\text{B1}   	    	& 0.24559 & 0.17819 & 0.17958 \\
-\text{B1a} 		& 0.24559 & 0.17819 & 0.17959 \\
-\text{B1b} 		& 0.25349 & 0.18386 & 0.18510 \\
-\text{B1c} 		& 0.25350 & 0.18386 & 0.18511 \\
-\text{B1d} 		& 0.26058 & 0.18944 & 0.19177 \\
-\text{B1e} 		& 0.24626 & 0.17861 & 0.1794 \\
-\text{B1f} 		& 0.24561 & 0.17820 & 0.18009\\
+\text{II}   	    	& 0.24559 & 0.17819 & 0.17958 \\
+\text{II.a} 		& 0.24559 & 0.17819 & 0.17959 \\
+\text{II.b} 		& 0.25349 & 0.18386 & 0.18510 \\
+\text{II.c} 		& 0.25350 & 0.18386 & 0.18511 \\
+\text{II.d} 		& 0.26058 & 0.18944 & 0.19177 \\
+\text{II.e} 		& 0.24626 & 0.17861 & 0.1794 \\
+\text{II.f} 		& 0.24561 & 0.17820 & 0.18009\\
 \bottomrule
@@
 \centerline{
-	\subfigure[\;Case B1: JPDAF]
+	\subfigure[\;Case II: JPDAF]
 		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1JPDAF.pdf}}
-	\subfigure[\;Case B1: M-JPDAF]
+	\subfigure[\;Case II.c: JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1cJPDAF.pdf}}
+	\subfigure[\;Case II.f: JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1fJPDAF.pdf}}
+	}
+\centerline{
+	\subfigure[\;Case II: M-JPDAF]
 		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1MUJPDAF.pdf}}
-	\subfigure[\;Case B1: C-JPDAF]
+	\subfigure[\;Case II.c: M-JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1cMUJPDAF.pdf}}
+	\subfigure[\;Case II.f: M-JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1fMUJPDAF.pdf}}
+	}
+\centerline{
+	\subfigure[\;Case II: C-JPDAF]
 		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1C-JPDAF.pdf}}
-	}
-\centerline{
-	\subfigure[\;Case B1c: JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1cJPDAF.pdf}}
-	\subfigure[\;Case B1c: M-JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1cMUJPDAF.pdf}}
-	\subfigure[\;Case B1c: C-JPDAF]
+	\subfigure[\;Case II.c: C-JPDAF]
 		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1cC-JPDAF.pdf}}
-	}
-\centerline{
-	\subfigure[\;Case B1f: JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1fJPDAF.pdf}}
-	\subfigure[\;Case B1f: M-JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1fMUJPDAF.pdf}}
-	\subfigure[\;Case B1f: C-JPDAF]
+	\subfigure[\;Case II.f: C-JPDAF]
 		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B1fC-JPDAF.pdf}}
@@
 }
-\caption{The orbital radii of three cases are shown where the dashed plots are the true trajectories and the solid plots are the estimated trajectories.
+\caption{Scenario II: The orbital radii of three cases are shown where the dashed plots are the true trajectories and the solid plots are the estimated trajectories.
 In all cases the M-JPDAF is more accurate and experiences less coalescence than the JPDAF, but both algorithms suffer from coalescence.
-The C-JPDAF systematically removes this coalescence from the updates at the cost of slightly larger covariance matrices and a large increase in required computational resources.
+The C-JPDAF systematically removes this coalescence from the updates at the cost of covariance matrices with slightly larger traces and a large increase in required computational resources.
 %Between Cases B1 and B1c, the process position uncertainty is increased, which last little effect on the estimates since this has little
-The conditions that affect the C-JPDAF most are with Case B1f where the range-rate measurements have the greatest uncertainty.
-}\label{fig:B1}
+The conditions that affect the C-JPDAF most are with Case II.f where the range-rate measurements have the greatest uncertainty.
+}\label{fig:II}
 \end{figure}
 
-Among the variations of the B1 cases, the range-rate measurement uncertainty (Case B1f) has the largest effect on tracking with the C-JPDAF.
+Among the variations of the II cases, the range-rate measurement uncertainty (Case II.f) has the largest effect on tracking with the C-JPDAF.
 The velocities of satellites in LEO are very large, so a small increase in $\sigma_{\dot\rho}$ corresponds to a large decrease in $J_S$, i.e., the measurement sets for the satellites will be less similar.
-This is further shown with the smallest computational cost among the B1 cases.
+This is further shown with the smallest computational cost among the II cases.
 Hence, the similarity optimization is very sensitive to velocity measurement uncertainty in these cases.
@@
 The data association and filtering algorithms are simulated in various GEO scenarios.
-Angles-only measurements are available with higher accuracy; namely the topocentric right ascension and declination are chosen as $\sigma_{\alpha,GEO}=\sigma_{\delta,GEO}=10^{-2}\sigma_\alpha$ and the $j$-th measurement $z_{j,GEO}\in\begin{bmatrix}\alpha, & \delta\end{bmatrix}^T\in\Re^2$ with covariance $R=\diag[\sigma_{\alpha,GEO}^2, \sigma_{GEO,\delta}^2]\in\Re^{2\times2}$.% YOU MAY HAVE TO REMOVE "\\" WITH ANY EDITING CHANGES
-
-\paragraph*{Scenario A2: Comparison of the JPDAF and M-JPDAF with two near-GEO satellites in close proximity}\
-
-These cases serve to compare the JPDAF with the M-JPDAF for tracking two nearby satellites passing over the sensor field of view in GEO, where the orbital radius is $r_{GEO}=4.2164\times10^{4}\ \text{km}$.
+Angles-only measurements are available with higher accuracy; namely the topocentric right ascension and declination are chosen as $\sigma_{\alpha}=\sigma_{\delta}=(\frac{0.001\pi}{180}\ \text{rad})^2$ and the $j$-th measurement $z_{j}\in\begin{bmatrix}\alpha, & \delta\end{bmatrix}^T\in\Re^2$ with covariance $R=\diag[\sigma_{\alpha}^2, \sigma_{\delta}^2]\in\Re^{2\times2}$.% YOU MAY HAVE TO REMOVE "\\" WITH ANY EDITING CHANGES
+
+\paragraph*{Scenario III: Comparison of the JPDAF and M-JPDAF with two near-GEO satellites in close proximity}\
+
+These cases serve to compare the JPDAF with the M-JPDAF for tracking two nearby satellites passing over the sensor field of view in GEO, where the orbital radius is $r_\text{GEO}=4.2164\times10^{4}\ \text{km}$.
 The initial conditions of the first two orbits are chosen as
 \begin{align*}
-x_{1,A2}&=\begin{bmatrix}r_{GEO}, & 0\ \text{km}, & 0\ \text{km}, & 0\ \text{km/s}, & \sqrt{\frac{\mu}{r_{GEO}}}, & 0\ \text{km/s}\end{bmatrix}^T,
-\\
-x_{2,A2}&=x_{1,A2}+\begin{bmatrix}
+x_{1,\text{III}}&=\begin{bmatrix}r_{GEO}, & 0\ \text{km}, & 0\ \text{km}, & 0\ \text{km/s}, & \sqrt{\frac{\mu}{r_{GEO}}}, & 0\ \text{km/s}\end{bmatrix}^T,
+\\
+x_{2,\text{III}}&=x_{1,\text{III}}+\begin{bmatrix}
 0\ \text{km}, & 0\ \text{km}, & -1\ \text{km}, & 0\ \text{km/s}, & 0\ \text{km/s}, & 4\times10^{-4}\ \text{km/s}
@@
 \end{align*}
-or where $x_{1,A2}$ remains exactly in GEO while $x_{2,A2}$ starts slightly perturbed from a GEO orbit with a nonzero inclination such that the two satellites nearly collide as they pass by in close proximity.
-The A2 cases consider small initial covariance matrices $P_{0,A2}= Q$ where $Q=\diag[10^{-6}I_{3\times3}, 10^{-8}I_{3\times3}]$, time steps of $10$ seconds over $5000$ seconds, and $\epsilon=1.1$.
+or where $x_{1,\text{III}}$ remains exactly in GEO while $x_{2,\text{III}}$ starts slightly perturbed from a GEO orbit with a nonzero inclination such that the two satellites nearly collide as they pass by in close proximity.
+The perturbations are summarized in Table \ref{tab:PertCaseIII}.
+\begin{table}
+\begin{center}
+\caption{Perturbations for Scenario III Cases} \label{tab:PertCaseIII}
+\begin{threeparttable}[h]
+\begin{tabularx}{.75\textwidth}
+{
+>{$}c<{$} |
+>{$}c<{$}
+}
+\toprule
+\multirow{2}{*}{Perturbed Case} & \multirow{2}{*}{Perturbation}\\
+\\
+\midrule
+\multirow{1}{*}{III.a} &  \multirow{1}{*}{$\sigma^2_{r} = 1.1\times10^{-6}\text{ km}^2$}
+\\
+\multirow{1}{*}{III.b} &  \multirow{1}{*}{$\sigma^2_{\dot r} = 1.1\times10^{-8}(\text{ km/s})^2$}
+\\
+\multirow{1}{*}{III.c} &  \multirow{1}{*}{$\sigma^2_{r} = 1.1\times10^{-6}\text{ km}^2$, $\sigma^2_{\dot r} = 1.1\times10^{-8}(\text{ km/s})^2$}
+\\
+\multirow{1}{*}{III.d} &  \multirow{1}{*}{$\sigma^2_{\alpha} = \sigma^2_{\delta} = 3.3508\times10^{-8}\text{ rad}^2$}
+\\
+\bottomrule
+\end{tabularx}
+\end{threeparttable}
+\end{center}
+\end{table}
+The III cases consider small initial covariance matrices $P_{0,\text{III}}= Q$ where $Q$ is defined the same as the LEO cases for time steps of $10$ seconds over $5000$ seconds.
 %These scenarios are chosen because the estimates exhibit large initial uncertainties, so the spread in the innovations affects the updates.
-Even though the initial conditions are largely different from those of the A1 cases, the A2 cases display the same trends, where the computational costs among the algorithms are roughly the same but the tracking abilities of the M-JPDAF \emph{always} yield estimates closer to the truth with less uncertainty than those of the JPDAF.
-These results are tabulated in Table \ref{tab:A2}, and some cases are plotted in Figure \ref{fig:A2}.
+Even though the initial conditions are largely different from those of the I cases, the III cases display the same trends, where the computational costs among the algorithms are roughly the same but the tracking abilities of the M-JPDAF \emph{always} yield estimates closer to the truth with less uncertainty than those of the JPDAF.
+These results are tabulated in Table \ref{tab:III}, and some cases are plotted in Figure \ref{fig:III}.
 
@@
 \begin{threeparttable}[h]
-\caption{Case A2} \label{tab:A2}
+\caption{Results of Scenario III Cases} \label{tab:III}
 \begin{tabularx}{\textwidth}
@@
 \toprule
-\multirow{2}{*}{Case} & \multicolumn{2}{c}{\multirow{1}{*}{RMS Angular Err. ($10^{-6}$rad)}} & \multicolumn{2}{c}{\multirow{1}{*}{Time (sec)}} & \multicolumn{2}{c}{\multirow{1}{*}{Mean $J_P$}} \\
+\multirow{2}{*}{Case} & \multicolumn{2}{c}{\multirow{1}{*}{RMS Ang. Err. ($10^{-6}$rad)}} & \multicolumn{2}{c}{\multirow{1}{*}{Time (sec)}} & \multicolumn{2}{c}{\multirow{1}{*}{Mean $J_P$}} \\
  & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF}
@@
 \midrule
-\text{A2}  & 1.2140 & 0.98784 & 2.77 & 2.73 & 16.5052 & 16.5015 \\
-\text{A2a} & 1.2141 & 0.98780 & 2.74 & 2.74 & 16.5054 & 16.5017 \\
-\text{A2b} & \hilight{11.728} & 1.0143 & 2.73 & 2.74 & 18.1526 & 18.1477 \\
-\text{A2c} & \hilight{11.728} & 1.0142 & 2.73 & 2.75 & 18.1528 & 18.1479 \\
-\text{A2d} & \hilight{11.700} & 1.0332 & 2.82 & 2.77 & 16.511 & 16.5054 \\
+\text{III}  & 1.2140 & 0.98784 & 2.77 & 2.73 & 16.5052 & 16.5015 \\
+\text{III.a} & 1.2141 & 0.98780 & 2.74 & 2.74 & 16.5054 & 16.5017 \\
+\text{III.b} & \hilight{11.728} & 1.0143 & 2.73 & 2.74 & 18.1526 & 18.1477 \\
+\text{III.c} & \hilight{11.728} & 1.0142 & 2.73 & 2.75 & 18.1528 & 18.1479 \\
+\text{III.d} & \hilight{11.700} & 1.0332 & 2.82 & 2.77 & 16.511 & 16.5054 \\
 \bottomrule
@@
 
+%\begin{figure}
+%{
+%%\hspace*{0.05\columnwidth}
+%\centerline{
+%	\subfigure[\;Case III: JPDAF]
+%		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2JPDAF.pdf}}
+%	\subfigure[\;Case III: M-JPDAF]
+%		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2MUJPDAF.pdf}}
+%	}
+%\centerline{
+%	\subfigure[\;Case III.a: JPDAF]
+%		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2aJPDAF.pdf}}
+%	\subfigure[\;Case III.a: M-JPDAF]
+%		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2aMUJPDAF.pdf}}
+%	}
+%\centerline{
+%	\subfigure[\;Case III.b: JPDAF]
+%		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2bJPDAF.pdf}}
+%	\subfigure[\;Case III.b: M-JPDAF]
+%		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2bMUJPDAF.pdf}}
+%	}
+%}
+%\caption{The orbital radii of three cases are shown where the dashed plots are the true trajectories and the solid plots are the estimated trajectories and $\alpha_{diff}$ is the difference between $\alpha_{1}$ and the measurement estimates of the two tracks.
+%Cases III and III.a serve as examples where the coalescence during crossing periods with the JPDAF yields much more similar estimates than with the M-JPDAF.
+%In Case III.b, this coalescence is so severe that track swapping occurs with the JPDAF, but this is avoided with the M-JPDAF.
+%}\label{fig:III}
+%\end{figure}
+
 \begin{figure}
@@
 \centerline{
-	\subfigure[\;Case A2: JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2JPDAF.pdf}}
-	\subfigure[\;Case A2: M-JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2MUJPDAF.pdf}}
+	\subfigure[\;Case III: JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{A2JPDAF.pdf}}
+	\subfigure[\;Case III.a: JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{A2aJPDAF.pdf}}
+	\subfigure[\;Case III.b: JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{A2bJPDAF.pdf}}
 	}
 \centerline{
-	\subfigure[\;Case A2a: JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2aJPDAF.pdf}}
-	\subfigure[\;Case A2a: M-JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2aMUJPDAF.pdf}}
-	}
-\centerline{
-	\subfigure[\;Case A2b: JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2bJPDAF.pdf}}
-	\subfigure[\;Case A2b: M-JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.5\columnwidth]{A2bMUJPDAF.pdf}}
-	}
-}
-\caption{The orbital radii of three cases are shown where the dashed plots are the true trajectories and the solid plots are the estimated trajectories and $\alpha_{diff}$ is the difference between $\alpha_{1}$ and the measurement estimates of the two tracks.
-Cases A2 and A2a serve as examples where the coalescence during crossing periods with the JPDAF yields much more similar estimates than with the M-JPDAF.
-In Case A2b, this coalescence is so severe that track swapping occurs with the JPDAF, but this is avoided with the M-JPDAF.
-}\label{fig:A2}
+	\subfigure[\;Case III: M-JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{A2MUJPDAF.pdf}}
+	\subfigure[\;Case III.a: M-JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{A2aMUJPDAF.pdf}}
+	\subfigure[\;Case III.b: M-JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{A2bMUJPDAF.pdf}}
+	}
+}
+\caption{Scenario III: The orbital radii of three cases are shown where the dashed plots are the true trajectories and the solid plots are the estimated trajectories and $\alpha_{diff}$ is the difference between $\alpha_{1}$ and the measurement estimates of the two tracks.
+Cases III and III.a serve as examples where the coalescence during crossing periods with the JPDAF yields much more similar estimates than with the M-JPDAF.
+In Case III.b, this coalescence is so severe that track swapping occurs with the JPDAF, but this is avoided with the M-JPDAF.
+}\label{fig:III}
 \end{figure}
 
-\paragraph*{Scenario B2: Comparison of the JPDAF, M-JPDAF, and C-JPDAF with two GEO satellites in close proximity}\
+\paragraph*{Scenario IV: Comparison of the JPDAF, M-JPDAF, and C-JPDAF with two GEO satellites in close proximity}\
 
@@
 \begin{split}
-x_{1,B2}&=f_{GEO}(0^\circ), \quad x_{2,B2}=f_{GEO}({10^{-4}}^\circ), \quad \mbox{where}
-\\
-f_{GEO}(\theta)&=r_{GEO}\begin{bmatrix}\cos\theta, & \sin\theta, & 0, & -\sqrt{\frac{\mu}{r_{GEO}^3}}\sin\theta, & \sqrt{\frac{\mu}{r_{GEO}^3}}\cos\theta, & 0\end{bmatrix}^T,
+x_{1,\text{IV}}&=f_\text{GEO}(0^\circ), \quad x_{2,\text{IV}}=f_\text{GEO}({10^{-4}}^\circ), \quad \mbox{where}
+\\
+f_\text{GEO}(\theta)&=r_\text{GEO}\begin{bmatrix}\cos\theta, & \sin\theta, & 0, & -\sqrt{\frac{\mu}{r_\text{GEO}^3}}\sin\theta, & \sqrt{\frac{\mu}{r_\text{GEO}^3}}\cos\theta, & 0\end{bmatrix}^T,
 \end{split}
 \end{align}
-where $Q=\diag[10^{-8}I_{3\times3}\ \text{km}^2, 10^{-10}I_{3\times3}\ \text{km/s}^2]$, $P_{0,B}=Q$, and $\epsilon=1.1$ for time steps of $10$ seconds over a $2000$ second time period.
+where $Q=\diag[\sigma^2_rI_{3\times3}, \dot\sigma^2_rI_{3\times3}]=\diag[10^{-8}I_{3\times3}\ \text{km}^2, 10^{-10}I_{3\times3}\ \text{km/s}^2]$.
+The perturbations are summarized in Table \ref{tab:PertCaseIV}.
+\begin{table}
+\begin{center}
+\caption{Perturbations for Scenario IV Cases} \label{tab:PertCaseIV}
+\begin{threeparttable}[h]
+\begin{tabularx}{.75\textwidth}
+{
+>{$}c<{$} |
+>{$}c<{$}
+}
+\toprule
+\multirow{2}{*}{Perturbed Case} & \multirow{2}{*}{Perturbation}\\
+\\
+\midrule
+\multirow{1}{*}{IV.a} &  \multirow{1}{*}{$\sigma^2_{r} = 1.1\times10^{-8}\text{ km}^2$}
+\\
+\multirow{1}{*}{IV.b} &  \multirow{1}{*}{$\sigma^2_{\dot r} = 1.1\times10^{-10}(\text{ km/s})^2$}
+\\
+\multirow{1}{*}{IV.c} &  \multirow{1}{*}{$\sigma^2_{r} = 1.1\times10^{-6}\text{ km}^2$, $\sigma^2_{\dot r} = 1.1\times10^{-8}(\text{ km/s})^2$}
+\\
+\multirow{1}{*}{IV.d} &  \multirow{1}{*}{$\sigma^2_{\alpha} = \sigma^2_{\delta} = 3.3508\times10^{-8}\text{ rad}^2$}
+\\
+\bottomrule
+\end{tabularx}
+\end{threeparttable}
+\end{center}
+\end{table}
+The initial state covariance is $P_{0,\text{IV}}=Q$, where the time steps are $10$ seconds over a $2000$ second time period.
 This corresponds to a scenario where two satellites orbit steadily nearby in GEO.
@@
 In theory, topocentric right ascension measurements should be the same distance apart throughout the cases and the declination should always be zero.
-Much like the B1 cases, the accuracy of the C-JPDAF yields a large improvement from that of the JPDAF and M-JPDAF; however, the increase in computational cost of the C-JPDAF is roughly $30$ times that of either other algorithm, which is about half of the increase found with the B1 cases.
-This is because the similarity measure is less complex with the B2 cases than with the B1 cases due to the number of estimates and the size of the measurement space.
-The angular RMS errors and computation times are tabulated in Table \ref{tab:B2} and some example plots from these cases are shown in Figure \ref{fig:B2}.
-In applications with a large number of objects to track, the C-JPDAF need not be applied on most object; only those close neighbors where the similarity among the estimates yield coalescence with the JPDAF or the M-JPDAF.
+Much like the II cases, the accuracy of the C-JPDAF yields a large improvement from that of the JPDAF and M-JPDAF; however, the increase in computational cost of the C-JPDAF is roughly $30$ times that of either other algorithm, which is about half of the increase found with the II cases.
+This is because the similarity measure is less complex with the IV cases than with the II cases due to the number of estimates and the size of the measurement space.
+The angular RMS errors and computation times are tabulated in Table \ref{tab:IV} and some example plots from these cases are shown in Figure \ref{fig:IV}.
+In applications with a large number of objects to track, the C-JPDAF need not be applied on most objects; only those close neighbors where the similarity among the estimates yield coalescence with the JPDAF or the M-JPDAF.
 This way, a data association scheme with an analytic solution, such as the JPDAF or the M-JPDAF, could be applied to the remaining estimates to mitigate the large computational requirements of the C-JPDAF.
@@
 \begin{threeparttable}[h]
-\caption{Case B2} \label{tab:B2}
+\caption{Results of Scenario IV Cases} \label{tab:IV}
 \begin{tabularx}{0.98\textwidth}
@@
 \toprule
-\multirow{2}{*}{Case} & \multicolumn{3}{c}{\multirow{1}{*}{RMS Angular Err. ($10^{-6}$rad)}} & \multicolumn{3}{c}{\multirow{1}{*}{Time (sec)}} \\
+\multirow{2}{*}{Case} & \multicolumn{3}{c}{\multirow{1}{*}{RMS Ang. Err. ($10^{-6}$rad)}} & \multicolumn{3}{c}{\multirow{1}{*}{Time (sec)}} \\
  & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{C-JPDAF} & \multirow{1}{*}{JPDAF} & \multirow{1}{*}{M-JPDAF} & \multirow{1}{*}{C-JPDAF}
@@
 \midrule
-\text{B2}   	& 1.111 & 0.9478 & 0.5671 & 1.15 & 1.12 & 34.3 \\
-\text{B2a} 	& 1.111 & 0.9478 & 0.5635 & 1.15 & 1.15 & 34.8 \\
-\text{B2b} 	& 1.131 & 0.9829 & 0.5739 & 1.13 & 1.13 & 35.3 \\
-\text{B2c} 	& 1.131 & 0.9829 & 0.5644 & 1.12 & 1.13 & 34.8 \\
-\text{B2d} 	& 1.112 & 0.9544 & 0.5614 & 1.11 & 1.12 & 35.2 \\
+\text{IV}   	& 1.111 & 0.9478 & 0.5671 & 1.15 & 1.12 & 34.3 \\
+\text{IV.a} 	& 1.111 & 0.9478 & 0.5635 & 1.15 & 1.15 & 34.8 \\
+\text{IV.b} 	& 1.131 & 0.9829 & 0.5739 & 1.13 & 1.13 & 35.3 \\
+\text{IV.c} 	& 1.131 & 0.9829 & 0.5644 & 1.12 & 1.13 & 34.8 \\
+\text{IV.d} 	& 1.112 & 0.9544 & 0.5614 & 1.11 & 1.12 & 35.2 \\
 \bottomrule
@@
 \midrule
-\text{B2}   	    	& 0.0076169 & 0.0073684 & 0.0073846 \\
-\text{B2a} 		& 0.007617 & 0.0073685 & 0.0073879 \\
-\text{B2b} 		& 0.0083158 & 0.0080603 & 0.0080799 \\
-\text{B2c} 		& 0.0083159 & 0.0080604 & 0.0080812 \\
-\text{B2d} 		& 0.0076643 & 0.0074103 & 0.0074331 \\
+\text{IV}   	    	& 0.0076169 & 0.0073684 & 0.0073846 \\
+\text{IV.a} 		& 0.007617 & 0.0073685 & 0.0073879 \\
+\text{IV.b} 		& 0.0083158 & 0.0080603 & 0.0080799 \\
+\text{IV.c} 		& 0.0083159 & 0.0080604 & 0.0080812 \\
+\text{IV.d} 		& 0.0076643 & 0.0074103 & 0.0074331 \\
 \bottomrule
@@
 \centerline{
-	\subfigure[\;Case B2: JPDAF]
+	\subfigure[\;Case IV: JPDAF]
 		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2JPDAF.pdf}}
-	\subfigure[\;Case B2: M-JPDAF]
+	\subfigure[\;Case IV.b: JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2bJPDAF.pdf}}
+	\subfigure[\;Case IV.d: JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2dJPDAF.pdf}}
+	}
+\centerline{
+	\subfigure[\;Case IV: M-JPDAF]
 		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2MUJPDAF.pdf}}
-	\subfigure[\;Case B2: C-JPDAF]
+	\subfigure[\;Case IV.b: M-JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2bMUJPDAF.pdf}}
+	\subfigure[\;Case IV.d: M-JPDAF]
+		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2dMUJPDAF.pdf}}
+	}
+\centerline{
+	\subfigure[\;Case IV: C-JPDAF]
 		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2C-JPDAF.pdf}}
-	}
-\centerline{
-	\subfigure[\;Case B2b: JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2bJPDAF.pdf}}
-	\subfigure[\;Case B2b: M-JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2bMUJPDAF.pdf}}
-	\subfigure[\;Case B2b: C-JPDAF]
+	\subfigure[\;Case IV.b: C-JPDAF]
 		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2bC-JPDAF.pdf}}
-	}
-\centerline{
-	\subfigure[\;Case B2d: JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2dJPDAF.pdf}}
-	\subfigure[\;Case B2d: M-JPDAF]
-		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2dMUJPDAF.pdf}}
-	\subfigure[\;Case B2d: C-JPDAF]
+	\subfigure[\;Case IV.d: C-JPDAF]
 		{\hspace*{0.\columnwidth}\includegraphics[width=0.33\columnwidth]{B2dC-JPDAF.pdf}}
@@
 }
-\caption{In the various cases, the C-JPDAF is always able to produce more distinguishable estimates.
+\caption{Scenario IV: In the various cases, the C-JPDAF is always able to produce more distinguishable estimates.
 The perturbations among the cases yield slightly different results, but none are drastically different.
-}\label{fig:B2}
+}\label{fig:IV}
 \end{figure}
@@
 The most important result is that the C-JPDAF consistently avoids track-swapping, whereas the JPDAF and M-JPDAF do not.
-However unlike the B1 cases, the B2 cases do not have any parameters that affect the coalescence-avoiding optimization significantly.
-Therefore, all variations from Case B2 all yield very close results.
+However unlike the II cases, the IV cases do not have any parameters that affect the coalescence-avoiding optimization significantly.
+Therefore, all variations from Case IV all yield very close results.
 
 Furthermore, coalescence serves as a bias to the soft decision data association algorithms, where the bias is toward nearby tracks.
-The C-JPDAF also has an inherent bias by including $J_S$ from \refeqn{Js}, except this bias is away from the nearby tracks.
+The C-JPDAF also has an inherent bias by including $J_S$ from \refeqn{Js}, except this bias tends to repel neighboring tracks.
 While there is no guarantee that these biases cancel, their opposite effects can reduce estimation error and track swapping.
 If \refeqn{CostGen} is poorly tuned such that $J_S$ is too strong, its repulsive nature may yield increased estimation error and track swapping.
+Furthermore, when estimates are not close, the additional term $J_S$ has minimal effect as it becomes small.
+In short, $J_S$ is active mostly in the vicinity of coalescence to avoid it, and the C-JPDAF performs better than the JPDAF or M-JPDAF as detrimental track swapping is less likely.
 
@@
 
-Since data association and filtering are fundamentally stochastic, Cases A1, A2, B1, and B2 are repeated $100$ times with different noise to obtain mean metrics and number of failures.
+Since data association and filtering are fundamentally stochastic, Cases I--IV are repeated $100$ times with different noise to obtain mean metrics and number of failures.
 For these cases, a failure is defined as at least one of the following: (i) track swapping (when estimates are consistently incorrectly associated with measurements from other estimates), (ii) when the estimates become ambiguous with respect to each other (their position relative to each other is consistently closer than half of the true estimate separation), or (iii) a numerical optimization does not converge (C-JPDAF only).
 The following Monte Carlo results are tabulated in Table \ref{tab:MonteCarlo}.
+
+In Case I, neither the JPDAF nor the M-JPDAF experience failures.
+However, these data support that the error and uncertainty cost are reduced on average, without an increase in computation time.
+The simulated results of Case III also serves to compare these two algorithms, where both are subject to failures.
+The computation time and uncertainty cost follow the same trends as Case I.
+However, both algorithms suffer from track swapping, but this happens roughly twice as often with the JPDAF than it does with the M-JPDAF.
+As a result, the RMS error of the JPDAF is almost twice that of the M-JPDAF.
+
+All three algorithms are compared with Cases II and IV.
+Case II is designed as a scenario where both the JPDAF and M-JPDAF yield object estimates with ambiguous estimates, which is considered a failure all $100$ times.
+However, when the C-JPDAF converges properly, this failure is mitigated and the RMS error is reduced.
+Since the C-JPDAF is solved numerically rather than analytically, a global minimum is not always obtained in four-dimensional measurement space.
+Conversely, the C-JPDAF algorithm converges well every time with Case IV.
+The C-JPDAF solves a simpler optimization problem, where the number of estimates is reduced from three to two and the measurement space is reduced by two dimensions with angles-only measurements (no range or range-rate).
+In every trial, the optimized cost of the C-JPDAF avoids the track swapping and ambiguity of the JPDAF and M-JPDAF algorithms, and in consequence the C-JPDAF RMS error is reduced as well.
 
@@
 \midrule
-\text{JPDAF}        & A1 & 0.12181 & 0.564 & 0.51843 & 0 \\
-\text{M-JPDAF}    & A1 & 0.10791 & 0.563 & 0.49490 & 0 \\
+\text{JPDAF}        & \text{I} & 0.12181 & 0.564 & 0.51843 & 0 \\
+\text{M-JPDAF}    & \text{I} & 0.10791 & 0.563 & 0.49490 & 0 \\
 \midrule
-\text{JPDAF}        & A2 & 8.66749 & 2.764 & 16.5046 & 69 \\
-\text{M-JPDAF}    & A2 & 4.76320 & 2.763 & 16.5023 & 36 \\
+\text{JPDAF}        & \text{II} & 0.14520 & 1.668 & 0.24859 & 100 \\
+\text{M-JPDAF}    & \text{II} & 0.11004 & 1.671 & 0.17791 & 100 \\
+\text{C-JPDAF}\textsuperscript{*}    & \text{II} & 0.09752 & 117.115 & 0.18032 & 20 \\
 \midrule
-\text{JPDAF}        & B1 & 0.14520 & 1.668 & 0.24859 & 100 \\
-\text{M-JPDAF}    & B1 & 0.11004 & 1.671 & 0.17791 & 100 \\
-\text{C-JPDAF}\textsuperscript{*}    & B1 & 0.09752 & 117.115 & 0.18032 & 20 \\
+\text{JPDAF}        & \text{III} & 8.66749 & 2.764 & 16.5046 & 69 \\
+\text{M-JPDAF}    & \text{III} & 4.76320 & 2.763 & 16.5023 & 36 \\
 \midrule
-\text{JPDAF}        & B2 & 1.01311 & 1.107 & 0.00762 & 100 \\
-\text{M-JPDAF}    & B2 & 0.84104 & 1.107 & 0.00736 & 100 \\
-\text{C-JPDAF}    & B2 & 0.57102 & 31.360 & 0.00738 & 0 \\
+\text{JPDAF}        & \text{IV} & 1.01311 & 1.107 & 0.00762 & 100 \\
+\text{M-JPDAF}    & \text{IV} & 0.84104 & 1.107 & 0.00736 & 100 \\
+\text{C-JPDAF}    & \text{IV} & 0.57102 & 31.360 & 0.00738 & 0 \\
+\midrule
 \end{tabularx}
@@
 	\item Failures are the number of failures out of $100$ trials.
-	\item The RMS error unit for Cases A1 and B1 is km, and the RMS error unit for Cases A2 and B2 is $10^{-6}$ rad.
-	\item In Case B1 with the C-JPDAF (denoted with \textsuperscript{*}), the $20$ failures occurred when the optimal gains failed to converge, and the metrics are not based on the failed trials.
+	\item The RMS error unit for Cases I and II is km, and the RMS error unit for Cases III and IV is $10^{-6}$ rad.
+	\item In Case II with the C-JPDAF (denoted with \textsuperscript{*}), the $20$ failures occurred when the optimal gains failed to converge, and the metrics are not based on the failed trials.
 %    \item *: track swapping
@@
 \end{center}
-
-In Case A1, neither the JPDAF nor the M-JPDAF experience failures.
-However, these data support that the error and uncertainty cost are reduced on average, without an increase in computation time.
-The simulated results of Case A2 also serves to compare these two algorithms.
-The computation time and uncertainty cost follow the same trends as Case A1.
-However, both algorithms suffer from track swapping, but this happens roughly twice as often with the JPDAF than it does with the M-JPDAF.
-As a result, the RMS error of the JPDAF is almost twice that of the M-JPDAF.
-
-All three algorithms are compared with Cases B1 and B2.
-Case B1 is designed as a scenario where both the JPDAF and M-JPDAF yield object estimates with ambiguous estimates, which is considered a failure all $100$ times.
-However, when the C-JPDAF converges properly, this failure is mitigated and the RMS error is reduced.
-Since the C-JPDAF is solved numerically rather than analytically, a global minimum is not always obtained in four-dimensional measurement space.
-The C-JPDAF algorithm converges well every time with Case B2.
-The C-JPDAF solves a simpler optimization problem, where the number of estimates is reduced from three to two and the measurement space is reduced by two dimensions with angles-only measurements (no range or range-rate).
-In every trial, the optimized cost of the C-JPDAF avoids the track swapping and ambiguity of the JPDAF and M-JPDAF algorithms, and in consequence the C-JPDAF RMS error is reduced as well.
 
@@
 \subsection*{Appendix A: M-JPDAF Proofs} \label{NewPartOfEIsPSD}
-Note that all object-related variables refer to the $i$-th object, so this subscript is omitted from indexing the object. Its usage in Appendix A is referring to matrix rows.
+Note that all object-related variables refer to the $i$-th object, so this subscript is omitted from indexing the object. Any subsequent usage in Appendix A is when $i$ refers to matrix rows.
 \begin{itemize}
-\item[(i)] From \refeqn{JPDAFCov} and \refeqn{JpIntro} the uncertainty cost for any gain $K$ for each estimate is given by
+\item[(i)] From \refeqn{CovExpanded} and \refeqn{JpIntro} the uncertainty cost for any gain $K$ for each estimate is given by
 \begin{align}
@@
 \end{align}
-\item[(ii)] Consider the nontrivial case when a measurement update is possible, i.e., $0\leq\beta_{0}<1$. Then from \refeqn{E}, $E$ is the summation of two positive-definite terms,
-\begin{align}
-E=\sum_{j=1}^{n_r}\beta^2_{i}R_j+(1-\beta_{0})^2HP^-H^T,
-\end{align}
-so $E$ is positive-definite. The trivial case when $\beta_0=1$ corresponds to no measurement update, so the positive-definiteness of $E$ is irrelevant.
-%Suppose that an object is not measured.
-%Let this portion of the innovation update be $e_0=0_{m\times1}$ because no innovation exists where measurements are not available.
-%The Total Probability Theorem holds true for object $i$:
-%\begin{align}
-%\sum\limits_{j=0}^{n_r}\beta_{j}=1.
-%\end{align}
-%Consider the positive-semidefinite matrix $\Sigma\in\Re^{m\times m}$ from~\cite[Eq.
+\item[(ii)] Suppose that an object is not measured.
+Let this portion of the innovation update be $e_0=0_{m\times1}$ because no innovation exists where measurements are not available.
+The Total Probability Theorem holds true for object $i$:
+\begin{align}
+\sum\limits_{j=0}^{n_r}\beta_{j}=1.
+\end{align}
+Consider the positive-semidefinite matrix $\Sigma\in\Re^{m\times m}$ from~\cite[Eq.
 1.4.16-(1-10)]{ShaRonThi2001},% the $i$-th covariance spread in the means term is defined as% (note that ${\bf e}_{i}$ remains unchanged)
-%\begin{align}
-%\Sigma=\sum\limits_{j=0}^{n_r}\beta_{j}(e_{j}-{\bar{e}})(e_{j}-{\bar{e}})^T,
-%\end{align}
-%which may be manipulated as follows,
-%\begin{align}
-%\Sigma=\sum\limits_{j=0}^{n_r}\beta_{j}e_{j}e_{j}^T
-%-\sum\limits_{j=0}^{n_r}\beta_{j}e_{j}{\bar{e}}^T
-%-{\bar{e}}\sum\limits_{j=0}^{n_r}\beta_{j}e_{j}^T
-%+{\bar{e}}{\bar{e}}^T
-%=\sum\limits_{j=1}^{n_r}\beta_{j}e_{j}e_{j}^T-{\bar{e}}{\bar{e}}^T,
-%\label{eqn:CovSpreadPSD}
-%\end{align}
-%because $e_{0}\equiv0$.
-%From \refeqn{E},
-%\begin{align}
-%E=(1-\beta_{0})S+\Sigma,
-%\end{align}
-%which yields a positive-definite $E$ because $S$ is positive-definite, $\Sigma$ is positive-semidefinite, and $0\leq\beta_{0}<1$.
-%Therefore, $E$ is invertible.
-\item[(iii)] The JPDAF applies the Kalman gain, which assumes a fixed measurement covariance $R$, i.e., $R=R_j$ for all $j\in\{1,2,...,n_r\}$.
-By manipulating \refeqn{E} and using the definition of \refeqn{S}, $S$ and $E$ are related using $\tilde\beta\in\Re$ as
-\begin{align}
-\label{eqn:EandS}
-%S=\frac{E+\tilde\beta R}{(1-\beta_0)^2}
-E=(1-\beta_0)^2S-\tilde\beta R
-, 
-\quad
-%S=\frac{E+\tilde\beta R}{(1-\beta_0)^2},
-%\quad 
-\tilde\beta=
-\left(
-\sum_{j=1}^{n_r}\beta_j
-\right)^2
--\sum_{j=1}^{n_r}\beta_j^2,
-\end{align}
-where $0\leq\tilde\beta<1$. The Kalman gain is defined as $K_{K}=P^-H^TS^{-1}\in\Re^{n\times m}$~\cite{TrackDataAssoc} and the proposed gain is $K=(1-\beta_0)P^-H^TE^{-1}$.
+\begin{align}
+\Sigma=\sum\limits_{j=0}^{n_r}\beta_{j}(e_{j}-{\bar{e}})(e_{j}-{\bar{e}})^T,
+\end{align}
+which may be manipulated as follows,
+\begin{align}
+\Sigma=\sum\limits_{j=0}^{n_r}\beta_{j}e_{j}e_{j}^T
+-\sum\limits_{j=0}^{n_r}\beta_{j}e_{j}{\bar{e}}^T
+-{\bar{e}}\sum\limits_{j=0}^{n_r}\beta_{j}e_{j}^T
++{\bar{e}}{\bar{e}}^T
+=\sum\limits_{j=1}^{n_r}\beta_{j}e_{j}e_{j}^T-{\bar{e}}{\bar{e}}^T,
+\label{eqn:CovSpreadPSD}
+\end{align}
+because $e_{0}\equiv0$.
+From \refeqn{E},
+\begin{align}
+E=(1-\beta_{0})S+\Sigma,
+\end{align}
+which yields a positive-definite $E$ because $S$ is positive-definite, $\Sigma$ is positive-semidefinite, and $0\leq\beta_{0}<1$.
+Therefore, $E$ is invertible.
+\item[(iii)] Let $K_{K}=P^-H^TS^{-1}\in\Re^{n\times m}$~\cite{TrackDataAssoc}.
 %The Kalman filter gain $K_{K}=P^-H^TS^{-1}\in\Re^{n\times m}$ from~\cite{TrackDataAssoc} and the M-JPDAF gain $K_{M}=(1-\beta_{0})P^-H^TE^{-1}\in\Re^{n\times m}$ from \refeqn{GainK} are not always equivalent with a proof by contradiction.
 Suppose that these gains are identical, i.e., $K=K_K$.
-From \refeqn{S} and \refeqn{E}, having identical gains implies
-%\begin{align}
-%\label{eqn:SequalsE}
-%S&=(1-\beta_0)^{-1}E
-%\nonumber\\
-%R+HP^-H^T&=(1-\beta_0)^{-1}\sum_{j=1}^{n_r}\beta_j^2R_j+(1-\beta_0)HP^-H^T
-%.
-%\end{align}
-\begin{align}
-\label{eqn:SequalsE}
-K_KE&=KE
-%\nonumber\\
-%P^-H^TS^{-1}((1-\beta_0)^2S-\tilde\beta R)&=(1-\beta_0)P^-H^T
-\nonumber\\
-(1-\beta_0)^2P^-H^T-\tilde\beta P^-H^TS^{-1}R&=(1-\beta_0)P^-H^T
-.
-\end{align}
-%This implies
-%\begin{align}
-%(1-\beta_0)((1-\beta_0)P^-H^T-P^-H^T)&=\tilde\beta P^-H^TS^{-1}R
-%\nonumber\\
-%-\beta_0(1-\beta_0)P^-H^T&=\tilde\beta P^-H^TS^{-1}R.
-%\end{align}
-%This implies
-Rearranging \refeqn{SequalsE} yields
-\begin{align}
+Multiplying both sides by $E$ and using \refeqn{E},
+% are equivalent and multiplying both gains by $E$ from \refeqn{E},
+\begin{align}
+KE&=K_{K}E\nonumber
+\\
+(1-\beta_{0})P^-H^T&=
+(1-\beta_{0})P^-H^T+K_{K}\Sigma.
+%({\sum_{j=1}^{n_r} \beta_{j}e_{j}e_{j}^T-{\bar{e}}}{\bar{e}}^T).
 \label{eqn:KalMUareSame}
-P^-H^T(\beta_0(1-\beta_0)I+\tilde\beta S^{-1}R)=0,
-\end{align}
-which is satisfied with the case $\tilde\beta=0$.
-This case implies that $\beta_j=1$ for exactly one $j\in\{1,2,...,n_r\}$.
-This case corresponds to a hard decision (unity probability between a single measurement and single estimate) where $\beta_0=0$.
-This case is not satisfied in general, so it follows that the Kalman gain is different in general from the optimal gain proposed in this paper.
-
-%Since the measurement covariance $R$ is independent of a priori covariance $P^-$ and measurement matrix $H$, \refeqn{SequalsE} is only true if $R=R_j$ is fixed for all measurements and exactly one association is unity, i.e., $\beta_j=1$ for $1\leq j\leq n_r$ (excluding the trivial case $\beta_0=1$), where all other association probabilities are zero.
-%Neither of these conditions are satisfied in general, so it follows that the Kalman gain is different in general from the optimal gain proposed in this paper.
-%Multiplying both sides by $E$ and using \refeqn{E},
-%% are equivalent and multiplying both gains by $E$ from \refeqn{E},
-%\begin{align}
-%KE&=K_{K}E\nonumber
-%\\
-%(1-\beta_{0})P^-H^T&=
-%(1-\beta_{0})P^-H^T+K_{K}\Sigma.
-%%({\sum_{j=1}^{n_r} \beta_{j}e_{j}e_{j}^T-{\bar{e}}}{\bar{e}}^T).
-%\label{eqn:KalMUareSame}
-%\end{align}
-%Therefore if \refeqn{KalMUareSame} is true, then $K_{K}\Sigma=0$, which may only occur if the probabilities that each measurement belongs to an object are exactly $0$ or $1$ or $K_{K}\perp\Sigma$.
-%None of these conditions are satisfied in general, so it follows that the Kalman gain is different in general from the optimal gain proposed in this paper.
+\end{align}
+Therefore if \refeqn{KalMUareSame} is true, then $K_{K}\Sigma=0$, which may only occur if the probabilities that each measurement belongs to an object are exactly $0$ or $1$ or $K_{K}\perp\Sigma$.
+None of these conditions are satisfied in general, so it follows that the Kalman gain is different in general from the optimal gain proposed in this paper.
+
 \item[(iv)] From \refeqn{JpAppend}, the uncertainty cost with the Kalman gain $J_{P_{K}}\in\Re$ can be written as
@@
 &=
-\tr{P^--P^-H^TS^{-1}\left(2(1-\beta_0)S-E\right)S^{-1}HP^-}
-.
+\tr{P^--(1-\beta_{0})^2P^-H^T\left(\frac{2}{1-\beta_{0}}S^{-1}-\frac1{(1-\beta_{0})^2}S^{-1}ES^{-1}\right)HP^-}.
 %\nonumber
@@
 %&=
-%\tr{P^-+(1-\beta_{0})^2P^-H^TS^{-1}\left(\frac1{(1-\beta_{0})^2}\left((1-\beta_{i,0})S_i+{\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right) S^{-1}-\frac{2I}{1-\beta_{0}}\right)HP^-}\nonumber
+%\tr{P^-+(1-\beta_{0})^2P^-H^TS^{-1}\left(\frac1{(1-\beta_{0})^2}\left((1-\beta_{0,i})S_i+{\sum_{j=1}^{n_r} \beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right) S^{-1}-\frac{2I}{1-\beta_{0}}\right)HP^-}\nonumber
 %\\
 %&=
-%\tr{P^--(1-\beta_{0})^2P^-H^TS^{-1}\left(\frac{1}{1-\beta_{0}}I-\frac1{(1-\beta_{0})^2}\left({\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right) S^{-1}\right)HP^-}
+%\tr{P^--(1-\beta_{0})^2P^-H^TS^{-1}\left(\frac{1}{1-\beta_{0}}I-\frac1{(1-\beta_{0})^2}\left({\sum_{j=1}^{n_r} \beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right) S^{-1}\right)HP^-}
 %\nonumber
@@
 %&=
-%\tr{P^--(1-\beta_{0})P^-H^TS^{-1}HP^-+P^-H^TS^{-1}\left({\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right) S^{-1}HP^-}\nonumber
+%\tr{P^--(1-\beta_{0})P^-H^TS^{-1}HP^-+P^-H^TS^{-1}\left({\sum_{j=1}^{n_r} \beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right) S^{-1}HP^-}\nonumber
 %\\
 %&=
-%\tr{P^--(1-\beta_{0})P^-H^TS^{-1}HP^-+K_K\left({\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right) K_K^T}
+%\tr{P^--(1-\beta_{0})P^-H^TS^{-1}HP^-+K_K\left({\sum_{j=1}^{n_r} \beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right) K_K^T}
 \end{align}
@@
 \end{align}
-From \refeqn{CostKal} and \refeqn{CostOpt} and using the relationship between $E$ and $S$ from \refeqn{EandS} repeatedly, the difference $J_{P_K}-J^*_P$can be written as
+From \refeqn{CostKal} and \refeqn{CostOpt} and using variations of the definition of $E$ from \refeqn{E} repeatedly, the difference $J_{P_K}-J^*_P$can be written as
 \begin{align}
@@
 &=
-\tr{P^-H^TS^{-1}\left((1-\beta_{0})^2SE^{-1}S-2(1-\beta_{0})S+E\right)S^{-1}HP^-}
+(1-\beta_{0})^2\tr{P^-H^T\left(E^{-1}-\frac{2}{1-\beta_{0}}S^{-1}+\frac1{(1-\beta_{0})^2}S^{-1}ES^{-1}\right)HP^-}
 \nonumber
 \\
-&=\frac1{(1-\beta_0)^2}
-\tr{K_K(\beta_0E+\tilde\beta R)E^{-1}(\beta_0E+\tilde\beta R)K_K^T}
+&=
+\tr{P^-H^TS^{-1}\Sigma E^{-1}\Sigma S^{-1}HP^-}
+\nonumber
+\\
+&
+=
+\tr{K_K\Sigma E^{-1}\Sigma K_K^T}
 %\nonumber
@@
 %&
-%=
-%\tr{K_K\Sigma E^{-1}\Sigma K_K^T}
-\nonumber
-\\
-&
 =\sum\limits_{i=1}^n\nu_i^TE^{-1}\nu_i\geq0,
 \end{align}
-where $\nu_i$ corresponds to the $i$-th row of $K_K(\beta_0E+\tilde\beta R)$ and $E^{-1}$ is positive-definite because $E$ is positive-definite from (ii).
+where $\nu_i$ corresponds to the $i$-th row of $K_K\Sigma$ and $E^{-1}$ is positive-definite because $E$ is positive-definite from (ii).
 This implies the inequality $J_{P_{K}}\geq J^*_{P}$.
@@
 H_p
-[(1-\beta_{0,p})(\Ibd H_pP_p^-+P_p^-H_p^T\Ibd^T)-\Ibd E_pK_p^T\nonumber
-\\
-& \quad -K_pE_p\Ibd^T]
+[(1-\beta_{0,p})(\Ibd H_pP_p^-+P_p^-H_p^T\Ibd^T)-\Ibd E_pK_p^T
+\nonumber
+\\
+& \quad 
+-K_pE_p\Ibd^T]
 H_p^T
@@
 %\mathcal{S}_{pq,p,bd}
-(H_pK_p{\bar{e}}_p-\hat z^+_{q})\nonumber
-\\
-& \quad +
+(H_pK_p{\bar{e}}_p-\hat z^+_{q})
+\nonumber
+\\
+& \quad 
++
 (H_pK_p{\bar{e}}_p-\hat z^+_{q})^T
@@
 & \quad
-+P_p^-H_p^T\Ibd^T)-\Ibd E_pK_p^T\nonumber
-\\
-& \quad -K_pE_p\Ibd^T]
+-\Ibd E_pK_p^T
+%\nonumber
+%\\
+%& \quad 
+-K_pE_p\Ibd^T]
 H_p^T
@@
 %Let $\beta^*\in\Re$ be the summation of association probabilities of the $i$-th object.
-%The singularity of $E_i$ only occurs when $\beta^*=(1-\beta_{i,0})=\sum_{j=1}^{n_r}\beta_j=0$.
+%The singularity of $E_i$ only occurs when $\beta^*=(1-\beta_{0,i})=\sum_{j=1}^{n_r}\beta_j=0$.
 %It is clear from \refeqn{KalEst} and \refeqn{CovExpanded} that the updates are independent of $K_i$ exactly at this singularity.

### 2018-06-09 Journal/JAS Data Association/JAS_DataAssoc (Taeyoung Lee's conflicted copy 2015-12-10).tex
journal added

--- 2018-06-09 Journal/JAS Data Association/JAS Data Association Before Rev 1/JAS_DataAssoc.tex
@@
 We consider the problem of estimating object states, when the number of objects for tracking is assumed known, but the measurement origins are unknown.
-We assume that a sensor generates measurements, composed of only object detections and spurious readings from clutter.
-Multiple measurements may originate from a single object, or an object may be missed entirely.
+More explicitly, a measurement may be originated from a single object or it corresponds to spurious readings from clutter.
+Therefore, multiple measurements may originate from a single object, or an object may be missed entirely.
 However, we assume that a single measurement does not originate from more than one object.
@@
 %\end{align}
-Then, a close approximation of the probability that the $i$-th system is associated with the $j$-th measurement is (from~\cite{Bar1990})
+Then, a close approximation of the probability that the $i$-th system is associated with the $j$-th measurement is
 \begin{align}
@@
 \end{align}
-where $B$ accounts for measurements \emph{not} originating from known targets.
-When $B=0$, this corresponds to rare extraneous measurements and missed detections.
-When these are prevalent, tuning $B$ is required, though this is beyond the paper scope.
-
-%When these are prevalent, a fixed value of $B$ in \refeqn{betaDef} was found from simulations to provide a close approximation of $\beta_{ij}$ in various applications, though tuning this parameter is beyond the scope of this paper.
+where $B=0$ corresponds to rare extraneous measurements and missed detections.
+When these are prevalent, a fixed value of $B$ in \refeqn{betaDef} was found from simulations to provide a close approximation of $\beta_{ij}$ in various applications~\cite{Bar1990}.
 
@@
 \put(-0.5,3){\rotatebox{90}{Cost}}
-\put(0,0){\includegraphics[width=0.48\columnwidth]{CT.png}}
+\put(0,0){\includegraphics[width=0.48\columnwidth]{CostTrends.png}}
 \put(5.3,1.2){\shortstack[c]{Uncertainty\\ cost: $J_P$}}
@@
 %\put(2.48,1.57){\circle*{0.1}}
-%\put(5.00,0.81){\circle*{0.1}}
-%\put(2.47,1.60){\circle*{0.1}}
-\put(4.99,0.81){\circle*{0.1}}
+\put(5.00,0.81){\circle*{0.1}}
 \put(2.47,1.60){\circle*{0.1}}
-\put(4.8,-0.4){$K_\text{M-JPDAF}$}
-\put(2.3,-0.4){$K_\text{C-JPDAF}$}
+\put(4.8,-0.4){$K_{M-JPDAF}$}
+\put(2.3,-0.4){$K_{C-JPDAF}$}
 \put(3.1,-0.8){Estimator gain}
@@
 }
-\caption{Illustration of optimal coalescence avoidance: several costs are illustrated with respect to an element of an estimator gain $K$ (blue dashed: uncertainty cost $J_P$, red dashed-dotted: similarity cost $J_S$, black solid: weighted sum).
-The estimator gain $K_\text{M-JPDAF}$ is selected to minimize the uncertainty cost.
-The estimator gain $K_\text{C-JPDAF}$ is chosen such that the weighted sum of the uncertainty cost and the similarity cost is minimized.
+\caption{Illustration of optimal coalescence avoidance: several costs are illustrated with respect to an element of an estimator gain $K$ (dashed: uncertainty cost $J_P$, dash-dot: similarity cost $J_S$, solid: weighted sum).
+The estimator gain $K_{M-JPDAF}$ is selected to minimize the uncertainty cost.
+The estimator gain $K_{C-JPDAF}$ is chosen such that the weighted sum of the uncertainty cost and the similarity cost is minimized.
 In short, at the expense of increased uncertainty at a single step, the proposed approach prevents coalescence.
@@
 % MATLAB: RMS_r_indiv(k) = (sum((Ri{k}-Rhati{k}).^2)/NumStates).^0.5;
-\mathbf{e}_{RMS,r,i}=\sqrt{\frac{\sum_{k=1}^n\norm{r_{i,k}-\hat r_{i,k}}^2}{n}}
+\mathbf{e}_{RMS,r,i}=\sqrt{\frac{\sum_{k=1}^n(r_{i,k}-\hat r_{i,k})^2}{n}}
 , \quad
@@
 Various cases in LEO are simulated to compare the data association and filtering algorithms, namely the JPDAF, M-JPDAF, and C-JPDAF.
-The process covariance is $Q=\diag[\sigma^2_rI_{3\times3},\dot\sigma^2_rI_{3\times3}]=\diag[10^{-6}I_{3\times3}\ \text{km}^2, 10^{-8}I_{3\times3}\ (\text{km/s})^2]$ and the measurement uncertainty is $R=\diag[\sigma_\rho^2, \sigma_{\dot\rho}^2, \sigma_\alpha^2, \sigma_\delta^2]=\diag[(0.15\ \text{km})^2, (0.005\ \text{km/s})^2,\\ (\frac{0.01\pi}{180}\ \text{rad})^2, (\frac{0.01\pi}{180}\ \text{rad})^2]\in\Re^{4\times4}$ where these standard deviations are selected from the Mahe Ground Station uncertainty data~\cite{VerSauSco04} and~\cite{KurAriAriEfe10}.
+The process covariance is $Q=\diag[\sigma^2_rI_{3\times3},\dot\sigma^2_rI_{3\times3}]=\diag[10^{-6}I_{3\times3}\ \text{km}^2, 10^{-8}I_{3\times3}\ \text{km/s}^2]$ and the measurement uncertainty is $R=\diag[\sigma_\rho^2, \sigma_{\dot\rho}^2, \sigma_\alpha^2, \sigma_\delta^2]=\diag[(0.15\ \text{km})^2, (0.005\ \text{km/s})^2,\\ (\frac{0.01\pi}{180}\ \text{rad})^2, (\frac{0.01\pi}{180}\ \text{rad})^2]\in\Re^{4\times4}$ where these standard deviations are selected from the Mahe Ground Station uncertainty data~\cite{VerSauSco04} and~\cite{KurAriAriEfe10}.
 
@@
 
-These cases serve to compare the JPDAF with the M-JPDAF for tracking two nearby satellites passing over the sensor field of view in GEO, where the orbital radius is $r_\text{GEO}=4.2164\times10^{4}\ \text{km}$.
+These cases serve to compare the JPDAF with the M-JPDAF for tracking two nearby satellites passing over the sensor field of view in GEO, where the orbital radius is $r_{GEO}=4.2164\times10^{4}\ \text{km}$.
 The initial conditions of the first two orbits are chosen as
@@
 \begin{split}
-x_{1,\text{IV}}&=f_\text{GEO}(0^\circ), \quad x_{2,\text{IV}}=f_\text{GEO}({10^{-4}}^\circ), \quad \mbox{where}
-\\
-f_\text{GEO}(\theta)&=r_\text{GEO}\begin{bmatrix}\cos\theta, & \sin\theta, & 0, & -\sqrt{\frac{\mu}{r_\text{GEO}^3}}\sin\theta, & \sqrt{\frac{\mu}{r_\text{GEO}^3}}\cos\theta, & 0\end{bmatrix}^T,
+x_{1,\text{IV}}&=f_{GEO}(0^\circ), \quad x_{2,\text{IV}}=f_{GEO}({10^{-4}}^\circ), \quad \mbox{where}
+\\
+f_{GEO}(\theta)&=r_{GEO}\begin{bmatrix}\cos\theta, & \sin\theta, & 0, & -\sqrt{\frac{\mu}{r_{GEO}^3}}\sin\theta, & \sqrt{\frac{\mu}{r_{GEO}^3}}\cos\theta, & 0\end{bmatrix}^T,
 \end{split}

### 2018-06-09 Journal/JAS Data Association/JAS_DataAssoc copy.tex
journal added

--- 2018-06-09 Journal/JAS Data Association/JAS Data Association Before Rev 1/JAS_DataAssoc.tex
@@
 %\end{align}
-Then, a close approximation of the probability that the $i$-th system is associated with the $j$-th measurement is (from~\cite{Bar1990})
+%
+%
+%Then, a close approximation of the probability that the $i$-th system is associated with the $j$-th measurement is (from~\cite{Bar1990})
+%\begin{align}
+%\label{eqn:betaDef}
+%\beta_{ij}=\frac{G_{ij}}{\sum\limits_{l=1}^{n_r}G_{il}+\sum\limits_{l=1}^{n_t}G_{lj}-G_{ij}+B},
+%\end{align}
+%where $B$ accounts for measurements \emph{not} originating from known targets.
+When $B=0$, this corresponds to rare extraneous measurements and missed detections.
+When these are prevalent, tuning $B$ is required, though this is beyond the paper scope.
+Then, a close approximation of the probability that the $i$-th system is associated with the $j$-th measurement is adopted from~\cite{Bar1990} as
 \begin{align}
@@
 \end{align}
-where $B$ accounts for measurements \emph{not} originating from known targets.
-When $B=0$, this corresponds to rare extraneous measurements and missed detections.
-When these are prevalent, tuning $B$ is required, though this is beyond the paper scope.
-
-%When these are prevalent, a fixed value of $B$ in \refeqn{betaDef} was found from simulations to provide a close approximation of $\beta_{ij}$ in various applications, though tuning this parameter is beyond the scope of this paper.
-
-%In this paper, we assume that at most a single measurement may originate from a single object, and missed detections and measurements originating from extraneous clutter are considered insignificant, so $B=0$ is used to determine joint association probability.
-
-%this approximation is valid when extraneous measurements and missed detections are rare.
-Even if missed detections and measurements originating from extraneous clutter is present, \refeqn{betaDef} performs well~\cite{Bar1990}.
-
-%\EditTL{(The above writing has several issues that the reviewers of ACC15 mentioned.
-(i) There is no indication on the dimension of variables.
-For example is $G_{ij}$ a scalar or a vector; (ii) there are several variables, such as $e_{ij}$ or $S_i$ that were used before defining them explicitly.
-Saying that it is the innovation and the a priori innovation covariance is not sufficient.
-Ideally, we have to define a variable before using it; (iii) The definition of $\mathcal{S}_{ti}$ and $\mathcal{S}_{ri}$ is unclear.
-Is $t$ also an indexed variable?; (iv) Do you really have to introduce the set $I$ and $J$.
-It is causing more confusion as several other variations of $J$ will appear later in this paper.
-Would it be more concise to rewrite it as $\sum_{j=1}^{n_r}$?; (v) Do we really have to introduce the variables $\mathcal{S}_{ti}$ and $\mathcal{S}_{ri}$.
-If they are used here only, there is no need to define them.
-It's better to substitute (6) into (7).
-Usually, a new variable or symbol is defined, only if it has an important meaning, or it is used repeatedly later; (vi) how is $B$ defined, and what's the dimension of it.
-Give the explicit definition, or at least cite something.
-The given explanation raises more questions.
-}
+where $B\geq0$ is the variable that accounts for measurements \emph{not} originating from known targets. When $B=0$, this corresponds to rare extraneous measurements and missed detections. 
+When these are prevalent, a fixed value of $B$ in \refeqn{betaDef} was found from simulations by tuning to provide a close approximation of $\beta_{ij}$.% in various applications.
 
@@
 respectively, and the subsequent tabulated RMS errors are averaged over all objects. 
+The position RMS error metrics are decomposed into range and direction components because range measurements are available in LEO cases and angles-only measurements are available in GEO cases.
 The uncertainty is summarized with $\tr{J_P}$, and the computation time is measured as well, tabulated in Tables \ref{tab:I}--\ref{tab:IV}.

### 2018-06-09 Journal/JAS Data Association/JAS_DataAssoc.tex
journal added

--- 2018-06-09 Journal/JAS Data Association/JAS_DataAssoc copy.tex
@@
 
-In short, the main contributions of this paper are (i) deriving the optimal JPDAF that minimizes the posterior mean square error and uncertainties in estimation, and (ii) proposing a systematic approach to avoid data coalescence of multiple objects in close-proximity.
+%In short, the main contributions of this paper are (i) deriving the optimal JPDAF that minimizes the posterior mean square error and uncertainties in estimation, and (ii) proposing a systematic approach to avoid data coalescence of multiple objects in close-proximity.
+The desirable features of the proposed approaches are illustrated by extensive numerical simulations with satellites in orbits.
+In short, the main contributions of this paper are (i) deriving the optimal JPDAF that minimizes the trace of the posterior state covariance matrix, and (ii) proposing a systematic approach to avoid data coalescence of multiple objects in close-proximity.
 The desirable features of the proposed approaches are illustrated by extensive numerical simulations with satellites in orbits.
@@
 
-The proposed coalescence avoiding filtering algorithm is constructed by minimizing the cost function $\mathbf{J}\in\Re$, defined as as the sum of the uncertainty cost $J_P\in\Re$ and similarity cost $J_S\in\Re$ with positive weighting factor $c\in\Re$,
+The proposed coalescence avoiding filtering algorithm is constructed by minimizing the cost function $\mathcal{J}\in\Re$, defined as as the sum of the uncertainty cost $J_P\in\Re$ and similarity cost $J_S\in\Re$ with positive weighting factor $c\in\Re$,
 \begin{align}
 \label{eqn:CostGen}
-\mathbf{J}=J_P+cJ_S.
+\mathcal{J}=J_P+cJ_S.
 \end{align}
@@
 In short, the estimator gain is chosen in the consideration of both uncertainty and coalescence, illustrated in Figure \ref{fig:CostTrends}.
-The estimator gain $K_i$ is selected to minimize the above cost $\mathbf{J}$.
+The estimator gain $K_i$ is selected to minimize the above cost $\mathcal{J}$.
 By considering similarity to obtain the estimator gain, we mitigate some coalescence at the expense of increasing estimate uncertainty according to the magnitude of $c$.
@@
 \put(6.2,2.6){\shortstack[c]{Similarity\\ cost: $J_S$}}
-\put(5.2,5.2){\shortstack[c]{Weighted sum:\\ $\mathbf{J}=J_P+cJ_S$}}
+\put(5.2,5.2){\shortstack[c]{Weighted sum:\\ $\mathcal{J}=J_P+cJ_S$}}
 %\put(4.98,0.78){\circle*{0.1}}
@@
 \begin{align}
-\deriv{\mathbf{J}}{K_i} = \deriv{J_P}{K_i} + c \deriv{J_S}{K_i} =0,\label{eqn:NCO}
+\deriv{\mathcal{J}}{K_i} = \deriv{J_P}{K_i} + c \deriv{J_S}{K_i} =0,\label{eqn:NCO}
 \end{align}
@@
 % MATLAB: RMS_r_indiv(k) = (sum((Ri{k}-Rhati{k}).^2)/NumStates).^0.5;
-\mathbf{e}_{RMS,r,i}=\sqrt{\frac{\sum_{k=1}^n\norm{r_{i,k}-\hat r_{i,k}}^2}{n}}
+{e}_{RMS,r,i}=\sqrt{\frac{\sum_{k=1}^n\norm{\mathbf{r}_{i,k}-\hat{\mathbf{r}}_{i,k}}^2}{n}}
 , \quad
-\mathbf{e}_{RMS,\theta,i}=\sqrt{\frac{\sum_{k=1}^n(\cos^{-1}(r_{i,k}^T\hat r_{i,k}))^2}{n}}
+{e}_{RMS,\theta,i}=\sqrt{\frac{\sum_{k=1}^n(\cos^{-1}(\mathbf{r}_{i,k}^T\hat{\mathbf{r}}_{i,k}))^2}{n}}
 ,

### 2018-06-09 Journal/JAS Data Association/JAS_DataAssoc_ans1.tex
journal added

\documentclass[11pt]{article}
\usepackage[letterpaper,text={6.5in,8.6in},centering]{geometry}
\usepackage{amssymb,amsmath,times,url}
\usepackage{xr,color}
\usepackage{hyperref}
\usepackage{amssymb,amsmath,times,subfigure,tabularx,booktabs,colortbl,multirow,threeparttable}

\externaldocument[]{JAS_OD_EKF_rev1}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand{\refprop}[1]{Proposition \ref{prop:#1}}

\newcommand{\RNum}[1]{\uppercase\expandafter{\romannumeral #1\relax}}
\newcommand{\RI}{\text{\RNum{1}}}
\newcommand{\RII}{\text{\RNum{2}}}
\newcommand{\RIII}{\text{\RNum{3}}}

\newenvironment{correction}{\begin{list}{}{\setlength{\leftmargin}{1cm}\setlength{\rightmargin}{1cm}}\vspace{\parsep}\item[]``}{''\end{list}}

\begin{document}

%\pagestyle{empty}

\section*{Response to the Reviewers' Comments for JASS-D-15-00036}

The authors would like to thank the reviewers for their thoughtful comments, which are aimed toward improving the quality of the paper and the clarity of the results. In accordance with the comments and suggestions, the paper has been revised as follows.
Since some numbers of equations, tables, and figures have been changed during revisions, these numbers from reviewer comments and responses are updated to reflect the current numbers in the revised paper.

\setlength{\leftmargini}{0pt}
\begin{itemize}\setlength{\itemsep}{2\parsep}

\subsection*{Associate Editor Comments for the Author}

\item {\itshape Please provide sufficient detail of the version of the JPDAF considered in this paper, list the assumptions about the detection probability, false measurements, measurement resolution (i.e., can one measurement originate from two objects?), and object size (i.e., can two or more measurements originate from one object?) under which that the M-JPDAF and C-JPDAF are derived, and explain why neither a missed detection model nor a false measurement model is needed for determining the association probability in equation (11).}

Several changes are made to address these issues. The last paragraph of Section 2 is updated as follows:
\begin{correction}
We consider the problem of estimating object states, when the number of objects for tracking is assumed known, but the measurement origins are unknown. More explicitly, a measurement may be originated from a single object or it corresponds to spurious readings from clutter. Therefore, multiple measurements may originate from a single object, or an object may be missed entirely.
However, we assume that a single measurement does not originate from more than one object.
%We assume accurate measurement resolution, such that at maximum one measurement may originate from a single object, where this measurement is independent of other objects detected by the same sensor.
%Suppose that these measurements are not associated with a specific object, and they possibly include spurious measurements or missed detections, where the probability of detecting objects is assumed known. 
Data association and filtering serves to estimate the state of the systems when there exist uncertainties in the system processes, measurements, and measurement origins. In this paper, the popular joint probabilistic data association filtering is improved in two distinct ways in the subsequent sections.
\end{correction}
The version of the JDPAF is specified at the beginning of Section 3:
\begin{correction}
In this section, we derive the M-JPDAF with a similar framework to that of the JPDAF, based off of the version from [3], [2] for the update structure, and the association probabilities from [1].
\end{correction}
The missed detection and false measurement models are accounted for with variable $B$, where determining its value is beyond the scope of this paper, so the reader is referred to [1]. This part is revised as follows in Section 3.2:
\begin{correction}
...where $B=0$ corresponds to rare extraneous measurements and missed detections. When these are prevalent, a fixed value of $B$ in (11) was found from simulations to provide a close approximation of $\beta_{ij}$ in various applications [1].
\end{correction}
A value of $B=0$ is chosen in the numerical simulations due to the low probabilities of missed detections and false measurements, specified in Section 5.

\item {\itshape The statement that the JPDAF is interior to the M-JPDAF needs further analysis. Equation (13) should be re-derived. When $K_i$ is not the optimal gain from the Kalman filter, the covariance or mean square matrix of the estimate given by equation (12) may take a different form than equation (13). For targets described by linear Gaussian systems, if the prior pdf of the state of target $i$ is Gaussian, the posterior pdf of the state of target $i$ is a Gaussian mixture. In the JPDAF, the Gaussian mixture is approximated by a Gaussian pdf, but the two pdfs have the same mean and covariance. Thus, the JPDAF estimate given by equation (12), with $K_i$ the optimal gain from the Kalman filter, is the posterior mean and therefore also the minimum mean square error estimator. Why the posterior mean is worse than the C-JPDAF estimate should be explained.}

The derivation of the entire expression of (13) would take a lot of spaces, and several derivations would be just repetitive. So, instead of re-deriving the entire expression from scratch, the part of the expressions that are independent of the Kalman gain is referred to a reference, namely [3, Eq. D-32 and D-33], and it is briefly discussed how they are derived in ths manuscript. And, the term $P^c_i$ depending on the Kalman gain is explicilty re-derived at (15), which corresponds to one of the contributions of this paper. The authors believe that any reader who is interested in redeveloping the presented results is able to do so with the revised manuscript.

%
%
%
%
%First, (13) is better cited, with a short summary of how the term is derived (re-deriving this term would take large amounts of space and would not be original). The only place in the derivation of (13) enforcing that the gain be Kalman is with the $P^c_i$ term. This explanation is now added shortly before equation (15), including what (15) would be if the gain were the Kalman gain.

%Second, the JPDAF employs the Kalman gain, which is optimal for single-target tracking, or multi-target tracking with hard decisions. This corresponds to measurements being associated with objects, assuming this association is correct. However, the soft decisions employed by the JPDAF yield a different covariance matrix, so minimizing the mean square error requires a different gain, derived in Section 3. We made several revisions to Section 3 based on earlier comments, so this point should be more clear.

Second, in a single-target tracking scenario, the Kalman gain is optimal in the sense that it minimizes the trace of the posterior covariance matrix $P^+_i$, which is equivalent to minimizing the mean square error of the target estimate. However, this paper shows that the Kalman gain is NOT optimal when there are uncertainties in the measurement origins. More explicitly, the Kalman gain does not minimize the posterior covariance $P^+_i$ that incorporates association uncertainties, given in (16). The proposed gain (20) is obtained by the optimality condition (19) to minimize the posterior covariance with association uncertainties,  and it is distinct from the Kalman gain. Therefore, the proposed M-JPDAF, based off the proposed gain, yields a smaller mean square error, compared with the conventional JPDAF. This is also verified by several numerical examples in Section 5.

Third, the C-JPDAF improves the mean estimate in many cases because the coalescence-avoiding portion prevents track swapping or cross tagging in many cases. The JPDAF and M-JPDAF may coalesce estimates to be unidentifiable, such that when the objects split, the associations might be consistently incorrect, causing large RMS errors. Simulations (including new Monte Carlo results) demonstrate the large estimation errors caused by track swapping for JPDAF and M-JPDAF, and how it is prevented for C-JPDAF.

%%%%%%%%%%%%%%%

\subsection*{Reviewer 1}

\item {\itshape Reviewer \#1: This is an excellently written paper. The literature review seems to be complete and the mathematical derivation correct. Results of the paper are backed by numerical examples demonstrating the viability of the solution.}\\

The authors appreciate the comments.

\end{itemize}

\subsection*{Reviewer 2}

\setlength{\leftmargini}{0pt}
\begin{itemize}\setlength{\itemsep}{2\parsep}

\item {\itshape Reviewer \#2: As a minor editorial comment, avoid the use of ``extensively`` in the abstract and when referring to the simulations.  A handful of simulations cannot really be considered extensive in the context of space object tracking.}

The word ``extensively'' is replaced with ``in various cases''.

\item {\itshape Introduction, Second Paragraph:
Regarding the sentence ``A hard decision is when estimates are fully associated with measurements according to some metric between the estimates and the measurements.''  Unless all cases truly involve the use of a metric and all of the properties of a metric are satisfied, this sentence needs to be revised.  Additionally, the use of ``estimates'' is somewhat confusing here.  It is not clear if this refers to estimated states or estimated measurements.  While the latter may be obviously true to a knowledgeable reader, it should still be made more clear.}

The sentence has been updated as follows:
\begin{correction}
When the association between an object of interest and a measurement is assumed correct and certain, this is referred to as a hard decision.
\end{correction}

%\begin{correction}
%
%\end{correction}

\item {\itshape Introduction, Second Paragraph:
The nearest neighbor filter does not necessarily use the Mahalanobis distance.  It could easily use a Euclidean distance.  However, as the sentence ``For example, a common...`` reads, the NNF requires the use of the Mahalanobis distance.}

It now reads:
\begin{correction}
For example, a common heuristic approach is known as the nearest neighbor filter (NNF), which uses a distance (e.g. Euclidean, Mahalanobis) between each measurement and the predicted measurements of each potential object to determine likely associations.
\end{correction}

%% UNC COMMENT %%

\item {\itshape Introduction, Fourth Paragraph:
It is stated that ``Kalman filters are considered optimal in the sense that the estimator gain is selected to minimize the posterior uncertainties." This is misleading. The Kalman gain is selected so as to minimize the posterior mean square error.  This is then related to the trace of the posterior covariance matrix.
\\
Section 3:
In the introductory paragraph, it is stated that ``uncertainties are minimized" and ``JPDAF fails to minimize these uncertainties".  The uncertainties are not minimized.  The posterior mean square error is minimized.  All instances of this throughout the paper must be fixed.
\\
Section 3.2:
Prior to Eq. (9) it is stated that the $i$-th covariance matrix represents the uncertainty.  No, it represents the covariance, not the uncertainty.  Also, given the appearance of $z_j$ in Eq. (8), it may help to be even more explicit that Eq. (9) is effectively taking $\text{E}[e_{ij}e_{ij}]^{T}$.
\\
Section 3.2:
Preceding Eq. (18), it is stated that the cost function is chosen to be an ``uncertainty cost function" ``as with the Kalman filter".  This is incorrect.  The Kalman filter cost function is a mean square error cost function.
\\
Section 3.2, Second to Last Paragraph:
Once more, the cost function is a mean square error cost function.  An ``uncertainty" cost function is not an adequate description in this case, nor is it really appropriate.
\\
Also following Eq. (12), there is another case of referring to minimization of the ``posterior uncertainty".  Once again, this is not an accurate statement.
}

While it is correct that Kalman filters minimize the posterior mean square error (minimizing the length of the estimation error vector), the trace of the posterior covariance matrix is widely accepted as a measure of estimate uncertainty. This is because the trace of a covariance matrix represents a measure of the volume of the corresponding Gaussian ellipsoid. In this sense, an uncertainty measure is minimized, and therefore, they are equivalent statements. To clarify these, two excerpts referenced in the comments are revised as follows:

%the diagonal elements of the covariance matrix are composed of state vector variances, so their summation is a measure of state vector uncertainty. I
\begin{correction}
More explicitly, Kalman filters are considered optimal in the sense that the estimator gain is selected to minimize the mean square error of the estimated state, or equivalently a measure of posterior uncertainty represented by the trace of the covariance.
\end{correction}

\begin{correction}
In this section, we derive the M-JPDAF with a similar framework to that of the JPDAF, based off of the version from [3], [2] for the update structure, and the association probabilities from [1].
Differences between the JPDAF and M-JPDAF occur in the measurement update where the gain is derived, which changes the posterior covariance equation.
As a result of the changes described in this section, the M-JPDAF serves to update state estimates from measurements in a soft decision approach that minimizes the sum of the posterior state vector variances. This a measure of state uncertainty and minimizing this measure is equivalent to minimizing the length of the estimation error vector; the conventional JPDAF fails to minimize this measure.
\end{correction}

To the other point, we included $\text{E}[e_{ij}e_{ij}]^{T}$ as an intermediate step inside Eq. (9).

\item {\itshape Introduction, Sixth Paragraph:
What is meant in the first sentence when it is stated that ``objects are in close-proximity such that neighboring objects share measurements"?  This sentence needs to be reworded.  If a graphic would aid in explaining this, please develop and use one.  Alternatively, it should be restated to read as something along the lines of multiple objects state estimates are consistent with one measurement.  In any event, this statement is unclear.}

This sentence is clarified as follows:

\begin{correction}
The second part of this paper deals with a common pitfall of soft decision data association techniques, known as coalescence. When objects are in close-proximity, a single measurement may be consistent with more than one object, forcing a soft decision. When these objects are estimated by the same measurements, their updated states tend to converge toward each other.\end{correction}

\item {\itshape Introduction, Seventh Paragraph:
What is meant by ``measurements with a much different interpretation than just the object states"?  The entire last part of this paragraph is much too overstated, but this part absolutely makes no sense.  Unless it is justified what this means, remove it.}

The purpose of this paragraph is to introduce the second contribution of this paper (the C-JPDAF), and differentiate the contributions of this paper from prior work. The last sentence is simplified as follows:

\begin{correction}
Therefore, we generalize the C-JPDAF algorithm in this paper to handle any number of objects and measurements, missed detections, and measurements originating from extraneous clutter.\end{correction}

\item {\itshape Section 2:
After Eq. (3), $\mathcal N[\mu,P]$ should really have $\sigma$ as a covariance matrix since that is what appears in Eq. (3) and Eq. (4).}

This is correct. The sentence is updated as follows:

\begin{correction}
...where $\mathcal N[\mu,P]$ denotes the Gaussian distribution with mean $\mu$ and covariance matrix $P$.
\end{correction}

\item {\itshape Section 2:
The paragraph following Eq. (4) seems to be tied to nothing.  Revise to include more information, to have it be more thoroughly developed, or move it to a place where it belongs better.}

The paragraph is rewritten to flow better with prior paragraphs and prepare for later sections as follows:

\begin{correction}
We consider the problem of estimating object states, when the number of objects for tracking is assumed known, but the measurement origins are unknown.
We assume accurate measurement resolution, such that at maximum one measurement may originate from a single object, where this measurement is independent of other objects detected by the same sensor.
Suppose that these measurements are not associated with a specific object, and they possibly include spurious measurements or missed detections, where the probability of detecting objects is assumed known. Data association and filtering serves to estimate the state of the systems when there exist uncertainties in the system processes, the measurements, and the measurement origins. In this paper, the popular joint probabilistic data association filtering is improved in two distinct ways in the subsequent sections.
\end{correction}

%\item {\itshape Section 3:
%In the introductory paragraph, it is stated that ``uncertainties are minimized" and ``JPDAF fails to minimize these uncertainties".  The uncertainties are not minimized.  The posterior mean square error is minimized.  All instances of this throughout the paper must be fixed.}
%
%See the comment about this issue above. The last sentence of this paragraph is rewritten as two sentences:

%\begin{correction}
%As a result of the changes described in this section, the M-JPDAF serves to update state estimates from measurements in a soft decision approach that minimizes the sum of the posterior state vector variances. This a measure of state uncertainty and minimizing this measure is equivalent to minimizing the length of the estimation error vector; the conventional JPDAF fails to minimize this measure.
%\end{correction}

\item {\itshape Section 3.1:
Please revise the first sentence that reads ``During the flow update...". The a priori estimated state is not found with Eq. (1).  It is found through the expected value of Eq. (1).  This is a very important distinction that needs to be made, and it is not clear at all why the governing equation for the estimated state has been omitted from the paper.}

This part is updated as follows:
\begin{correction}
The flow update serves to predict the how a system evolves between measurements. The system state obtains its $k$-th a priori state estimate $\hat x_{i_k}^-\in\Re^n$ from the posterior $(k-1)$-th state estimate $\hat x_{i_{k-1}}^+\in\Re^n$ with the expected value of (1),
\begin{align*}
\label{eqn:xestapriori}
\hat x_{i_{k}}^- & = F_{i_{k-1}} \hat x_{i_{k-1}}^+,
\end{align*}
where $\mathrm{E}[w_{i_{k}}]=0$.
\end{correction}

%\item {\itshape Section 3.2:
%Prior to Eq. (9) is is stated that the $i$-th covariance matrix represents the uncertainty.  No, it represents the covariance, not the uncertainty.  Also, given the appearance of $z_j$ in Eq. (8), it may help to be even more explicit that Eq. (9) is effectively taking $\text{E}[e_{ij}e_{ij}]^{T}$.}
%
%The covariance matrix is how we represent uncertainty of a multidimensional variable, much like a variance represents the uncertainty of a scalar. The authors agree on this point.

\item {\itshape Section 3.2:
Prior to Eq. (10), it is stated that $G_{ij}$ is a simplified scaled Mahalanobis distance.  What makes it simplified?  It is also more than just scaled, as well.  It is scaled, transformed, then scaled again.  After Eq. (10), it is stated that ``a close approximation of the probability".  Explain the restrictions under which is a close approximation.  This should be an easy explanation.
}

The authors made an error here: it is a scaled density function, not a Mahalanobis distance, and the error is fixed:

\begin{correction}
Let $G_{ij}\in\Re$ be a scaled density function used repeatedly between the $i$-th measurement estimate and the $j$-th measurement, defined as...
\end{correction}

To the second point, the following is included immediately after Eq. (10):

\begin{correction}
...where $B=0$ corresponds to rare extraneous measurements and missed detections. When these are prevalent, a fixed value of $B$ in (11) was found from simulations to provide a close approximation of $\beta_{ij}$ in various applications [1].
\end{correction}

\item {\itshape Section 3.2:
Following Eq. (12), it is stated that JPDAF choses $K_i$ as the traditional Kalman gain, ``based on an incorrect assumption that the Kalman gain yields an optimal estimator for the JPDAF".  Cite the sources where it is explicitly selected for this reason, or do not state that it is chosen based on the preceding assumption.}

Numerous sources apply the JPDAF without considering its optimality, though these sources do not explicitly state this sub-optimality. The sentences are revised as follows:

\begin{correction}
The conventional JPDAF chooses $K_i=P^-_iH_i^TS_i^{-1}$ as the gain from the Kalman filter [3].
In this paper, we show that the Kalman gain does not actually minimize the posterior uncertainty in probabilistic data association filtering, and we derive the new \textit{optimal} gain $K_i$.
\end{correction}

%We removed this. However, no source would explicitly state that they choose a sub-optimal gain matrix, but those sources apply their version of soft decision data association with Kalman filters or extended Kalman filters without considering how the data association and filtering affect each other.

\item {\itshape Section 3.2:
How is Eq. (3.2) found?  Please provide details.  Traditionally, the covariance should following from analysis of the posterior error.  Under what conditions does Eq. (12) hold?  What are the uncorrelated assumptions going into this equation?}

There is no Eq. (3.2). If you are referring to Eq. (12)-(15), these come from [3], where Eq. (15) is written generally (for any $K_i$, not just the Kalman gain) according to the form found in [11] as it is referenced. The only conditions when Eq. (13) holds is when at least one measurement may detect the $i$-th estimate, as well as assumptions stated earlier in the paper. We are confused by the ``uncorrelated assumptions'' part of this comment. Several comments are added throughout Eq. (12)-(15) with regard to the derivation of posterior covariance and sub-optimality of the Kalman gain in general.

%\item {\itshape Section 3.2:
%Also following Eq. (12), there is another case of referring to minimization of the ``posterior uncertainty".  Once again, this is not an accurate statement.}
%
%Please see the above responses to your comments and earlier revisions.

\item {\itshape Section 3.2:
Why, after Eq. (15) is the Joseph form referred to?  This comes out of nowhere and does not make any connection to any of the work.  Please revise to make this more clear.
\\
Section 3.2:
After the proof of Proposition 1, it would be beneficial to discuss any other comparisons to the usual Kalman filter; in particular, does Eq. (16) hold for any linear gain?  This would be the analog to the Joseph form of the covariance update, but it would advantageous to give more details for the interested reader.
}

The expression following Eq. (15) is commonly referred to as the \textit{Joseph form} and it is valid for any value of $K_i$ that is not necessarily the Kalman gain. This part is revised as:
%The term ``Joseph Form'' is commonly used for this general form. More on this structure can be found in the [11] provided there, and instead of the statement in parenthesis, we have:

\begin{correction}
...which is commonly referred to as the Joseph form, valid for any gain matrix $K_i$ [11].
\end{correction}

%\item {\itshape Section 3.2:
%Preceding Eq. (18), it is stated that the cost function is chosen to be an ``uncertainty cost function" ``as with the Kalman filter".  This is incorrect.  The Kalman filter cost function is a mean square error cost function.}
%
%Once again, please see the above responses.

Additionally, the following sentence is added after Eq. (17):

\begin{correction}
Because (16) is based on the Joseph form, this equation yields $P^+_i$ for arbitrary gain $K_i$.
\end{correction}

%\item {\itshape Section 3.2, Second to Last Paragraph:
%Once more, the cost function is a mean square error cost function.  An ``uncertainty" cost function is not an adequate description in this case, nor is it really appropriate.}
%
%Once again, please see the above responses.

\item {\itshape Section 4.1, First Paragraph:
Matusita's measure is introduced, and reference is given to the fact that it is computed for Gaussian distributions, but the introduction here reads a bit weakly.  It would be better if a few more details regarding Matusita's measure were presented so that the reader has a firm understanding.  Additionally, why is this measure selected?  Is it possible to consider other measures?  What are the ramifications of one measure over another?}

The beginning of the paragraph is revised as follows to reflect the comments:

\begin{correction}
We choose Matusita's measure because it provides a degree of similarity among multivariate Gaussian distributions [18], [19]. 
This measure is advantageous because of its straightforward computation and differentiable structure unlike the Bhattacharyya measure [22].
%The similarity between multiple multivariate Gaussian distributions can be represented by the overlapping regions of their probability density functions. Since there exists no analytic solution for this overlap, we choose Matusita's measure because it provides a scalar approximation of the overlap of multivariate Gaussian distributions [18], [19]. Furthermore, this measure is easy to compute and is easily differentiable unlike the Bhattacharyya measure [22]. More explicitly...
\end{correction}

\item {\itshape Section 4.1, Prior to Eq. (24):
It is stated that $a > 0 \in \Re$, but this could be more simply stated as $a \in \Re^{+}$.}

This detail is updated accordingly.

\item {\itshape Section 4.2, After Eq. (24):
It should be stated that all unique pairs are considered instead of all possible pairs since Eq. (23) explicitly excludes all of the pairings where both inputs are the same.}

The paper is updated with ``unique'' pairs.

\item {\itshape Section 4.2:
If the paper is to be published in color, the captions to Figs. 1 and 2 are acceptable; however, it would be easy enough not to publish in color.  In this case, the captions to the aforementioned figures should be updated to avoid the use of descriptions based on color.}

Since the concept of a similarity measure among mutivariate Gaussians is explained further (see above response), Fig. 1 is removed. The other figure is edited for publishing without color.

\item {\itshape Section 5:
Please review the first paragraph.  There are several poorly worded phrases here that should be revised, e.g. ``We consider satellite the low earth orbit".
\\
ALL of the theoretical development considered linear systems, but the application is to nonlinear systems.  This is a completely illogical disconnect.  At least some mention of the fact that the linear theory is being applied to a nonlinear system, as well as a discussion regarding the possible detrimental outcomes, should be given.  This is a rather major issue that must be corrected.  There is certainly nothing inconsistent here, when one considers the application of the EKF; however, the EKF is still developed using first-order expansions, whereas no such expansions are considered in the body of this work.}

Both of the above comments are addressed with major changes to the first paragraph of Section 5:

\begin{correction}
We consider satellites in the low earth orbit (LEO) and in the geosynchronous orbit (GEO) to compare the performances of the conventional JPDAF, the M-JPDAF, and the C-JPDAF.
Although the theoretical contributions are developed for linear systems, the satellites are subject to the nonlinear dynamics of the two-body problem. This motivates that the data association and tracking algorithms be extended to nonlinear systems in the same manner that the extended Kalman filter generalizes the Kalman filter for nonlinear systems. Using first-order expansions, linear theory of data association and filtering is applied to nonlinear systems.
\end{correction}

\item {\itshape Section 5:
Table 1 and all of the subsequent references to the parameters in Table 1 are very hard to follow from the reader's perspective.  This makes it hard to understand exactly the configuration of each of the subsequent simulations without constantly referring back to the table.  Even then, it is not very clear.  Please revise this so that a reader may more logically follow the setup.}

This table has been removed, and a small table is added for each scenario. These tables cover which parameters are perturbed and the magnitude of the perturbation, and the tables are located at the scenario description.

\item {\itshape Section 5, Preceding Eq. (38):
It is stated that $z_j \in \Re^4$, but it is clearly not always the case.  This needs to be revised.}

The introduction of $z_j$ is revised with ``$z_j\in\Re^4$ in LEO and $z_j\in\Re^2$ in GEO''.

\item {\itshape Section 5, Paragraph Preceding Table 2:
The discussion about coalescence that is given in this paragraph only serves to confuse the reader.  It would be advisable to push off all of this discussion to the later sections where coalescence is actually dealt with.  As it is currently, the insights that are to be gained from the simulation results are somewhat lost.  The same goes for the caption of Fig. 2.}

That clause on coalescence is removed and only, ``In some cases, the JPDAF suffers from track swapping, which is when measurements are consistently associated incorrectly'' is added, where a new subsection with Monte Carlo results covers this in further detail. The Figure 2 title is revised slightly as well to avoid coalescence discussion until later.

\item {\itshape Section 5.2, Scenario A.2:
In the first paragraph of this discussion, reference is given to ``The second set of cases".  This is confusing since this is the third set of cases.  Please revise. It is stated that ``Unlike the last cases, the III cases consider small initial covariance matrices".  However, case II considered the exact same initial covariance matrix.  Is the initial covariance on Case II incorrect?  If not, this needs to be revised to reflect the actual differences between II and III.}

These errors were a result of reordering the results section, such that now the LEO cases are first and the GEO cases are second. Further detail on initial covariances are included for clarification.

\item {\itshape Section 5.2, Scenario A.2:
What are the units of rad given in Table 6?  It appears that they may be Earth radii, but it is very unclear.}

The units are radians. We changed the metric from ``RMS Radial Err.'' to ``RMS Ang. Err.'' in Tables 6 and 8, where RMS errors are defined in Eq. (39) along with units.

\item {\itshape Section 5.2, Scenario B.2:
In the paragraph preceding Table 8, it is stated that the ``computation cost of the C-JPDAF is roughly 30 times that of either other algorithm".  What are the implications for tracking the catalog of 20,000+ objects?  It seems that the cost may become overwhelming; however, if one takes into account the fact that coalescing objects are perhaps rare, the increase in computational complexity may be somewhat mitigated.  As it stands, the reader has the inclination to believe that the C-JPDAF algorithm will not be applicable to the full space object tracking problem, but this should be clarified in some way so that the method is not immediately ruled out.}

This point is clarified at the end of the paragraph:

\begin{correction}
In applications with a large number of objects to track, the C-JPDAF need not be applied on most objects; only those close neighbors where the similarity among the estimates yield coalescence with the JPDAF or the M-JPDAF. This way, a data association scheme with an analytic solution, such as the JPDAF or the M-JPDAF, could be applied to the remaining estimates to mitigate the large computational requirements of the C-JPDAF.
\end{correction}

\item {\itshape Appendix A and B:
Some notational liberties are taken here.  A careful explanation of these should be made so that the reader can thoroughly follow the developments that are presented.  As it is now, it is bit harder to follow than is necessary.}

%A little information on the indexing of variables is included. 
Additional information on the indexing of variables has been added to Appendix A.
Otherwise, the notation is consistent with the paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Reviewer 3}

\subsubsection*{High Level Comments}

\item {\itshape Reviewer \#3: The first derivation on M-JPDAF looks good except for eqn (13).  I do not see where this equation comes from.  I am not convinced it is wrong, but this must be better developed or better cited.}

This comes from [3], which is better cited now with a short summary of how the equation is derived:

\begin{correction}
Next we consider when $\beta_{0,i}\neq1$, i.e., $0\leq\beta_{0,i}<1$. The resulting posterior state covariance $P^+_{i}\in\Re^{n\times n}$ is derived in [3, Eq. D-32 and D-33], where this term is defined as $\text{E}[(x_i-\hat x_i^+)(x_i-\hat x_i^+)^T]$ and the resulting solution is derived from the law of total expectation considering each innovation individually. Then, $P^+_{i}$ is simplified to...
\end{correction}

\item {\itshape Eqn (23) is very close to the actual probability that two PDFs represent the same object.  This paper would be much stronger if the cost function in eqn (31-33) was tied in actual concrete probabilities rather than a heuristic.}

This measure also takes into account different covariance matrices among those Gaussian distributions. The measure is designed to compute similarity in a way that is easily differentiable. This measure satisfies these goals better than other measures, and there no known analytic solution for the similarity between many multivariate Gaussians. Revised descriptions on why this measure is chosen can be found in the revised description before (23) and (24).

\item {\itshape More in depth discussion of the weighting factor $c$ should be included.  Maybe a quick derivation of what it should be, based on a particular example?
\\
More fundamentally, this problem is a multi-objective optimization which means there exists a Pareto frontier, which should be explored.  By choosing a specific $c$, you are making an assumption of a particular design.  This is not necessarily wrong, but definitely warrants further discussion, from a multi-objective optimization standpoint.}

The actual calculation of an optimal $c$ or $a$ (in the similarity measure) are beyond the scope of this paper. To compute these parameters, it is correct that there exists a Pareto frontier; the choice of $c$ depends on the type of scenario (for example, the II and IV results have different cost function parameters). A little more explanation on this is included after (25):

\begin{correction}
The estimator gain $K_i$ is selected to minimize the above cost $\mathbf{J}$. By considering similarity to obtain the estimator gain, we mitigate some coalescence at the expense of increasing estimate uncertainty according to the magnitude of $c$. The choice of $c$ depends on the type of scenario and the parameter $a$ composing the similarity cost $J_S$.
\end{correction}

\item {\itshape The C-JPDAF is referred to as a ``coalescence-avoiding optimal JPDAF".   While I don't disagree that you defined a cost function and optimized it, there is no proof that this correlates to a good filter.  The cost function $J_S$ seems to be completely heuristic.  By simply combing two cost function with a weight parameter, the new cost function in eqn (24) is generally not optimal with respect to either $J_P$ or $J_S$, so calling it ``optimal coalescence avoiding" is not true.
\\
The C-JPDAF, by virtue of the cost function in eqn (25), seems to define an inherently biased estimator.  The $J_S$ term introduces a bias.  Biases are a very bad property for an estimator.  This bias may also cause the filter to fail completely, and no discussion on how to avoid this is included. The C-JPDAF as derived appears to have major flaws involved in cross tagging (aka track swapping).  Cross tagging happens when two objects become identical in the measurement space, implying inherent unobservablity.  This implies that during the unobservability, you cannot distinguish the tracks. The C-JPDAF attempts to solve this problem by adding a repulsive term to the cost function.  It seems to me that if two objects cross over each other similar to Fig (2), this repulsive term may serve to encourage cross tagging.  In general, there may be many unintended consequences of this filter, which is dangerous.}

The C-JPDAF is only optimal with respect to the combined cost, where the $J_S$ component is based on multivariate Gaussian similarity. Similarity in the context of the JPDAF leads to sharing of measurements, ultimately causing coalescence.
Optimizing the cost function of the C-JPDAF is based on a compromise between uncertainty or mean square error of each estimate and similarity between the nearby estimates. In many cases, the C-JPDAF avoids cross tagging of the JPDAF or M-JPDAF, which cause large errors. Crossing scenarios may be improved as well (see the numerical simulation of [13]) as the coalescence-avoiding benefit of the C-JPDAF is only effective when objects are in close-proximity such that estimates may cross without becoming ambiguous. In all other situations, the C-JPDAF produces estimates with negligible difference to the M-JPDAF. The weighted cost function of the C-JPDAF is fairly common in literature (e.g. LQR optimal controller).
The terms ``coalescence-avoiding optimal'' might convey to some readers that the optimized parameter is coalescence only, so the name has been changed to ``coalescence-avoiding JPDAF'', so ``C-JPDAF'' is still used as its shorthand name.
Also correctly noted, there is no analytic proof that the filter is ``good'', so the following is added at the last paragraph of Section 4:

\begin{correction}
Filter performance is largely governed by tuning the weighted costs, and several numerical scenarios demonstrate the coalescence-avoiding benefits of the C-JPDAF in the subsequent section.
\end{correction}

The coalescence itself causes the JPDAF and M-JPDAF to be biased estimators, and there is no guarantee that cross tagging will be avoided. 
The following is added as the second-to-last paragraph of Section 5 to address this issues:

\begin{correction}
Furthermore, coalescence serves as a bias to the soft decision data association algorithms, where the bias is toward nearby tracks. The C-JPDAF also has an inherent bias by including $J_S$ from (32), except this bias tends to repel neighboring tracks. While there is no guarantee that these biases cancel, their opposite effects can reduce estimation error and track swapping. If (26) is poorly tuned such that $J_S$ is too strong, its repulsive nature may yield increased estimation error and track swapping.
Furthermore, when estimates are not close, the additional term $J_S$ has minimal effect as it becomes small.
In short, $J_S$ is active mostly in the vicinity of coalescence to avoid it, and the C-JPDAF performs better than the JPDAF or M-JPDAF as detrimental track swapping is less likely.
\end{correction}

\item {\itshape There is no proof that the C-JPDAF solves the coalescence problem.  A heuristic cost function and 6 simulated test cases are not enough proof.
\\
Everything that is claimed as a contribution in this paper is mathematically correct (e.g. a cost function is correctly defined and optimized over). This filter is optimal with respect to a cost function and may avoid coalescence in certain test cases.  However, this filter may also fail in certain test cases, cause problems with cross tagging, and not actually solve coalescence.  If you are convinced my assessment is incorrect, then you must prove these properties mathematically.}

The assessments are correct. The C-JPDAF provides a method to mitigate some coalescence, though it does not prove that the problem is solved completely. The following is added at the end of the last paragraph of Section 3 to clarify these points:

\begin{correction}
When $c>0$, the gains of the C-JPDAF serve to mitigate some coalescence from data association and filtering. In some cases, the C-JPDAF prevents cross tagging, when ambiguous estimates yield consistent incorrect associations. However, the C-JPDAF is not proven to solve the problem of coalescence completely, so it is possible that the C-JPDAF may cause cross tagging.
\end{correction}

\item {\itshape Results section shows a few test cases and claims them as representative.  Given that the C-JPDAF is completely heuristic, simulation results should be comprehensive, to at least back up numerically that claimed properties of the filter. The work is in filtering which is fundamentally stochastic.  The results need to be run over many different test cases.  Some measure of success, most likely in terms of percent success should be shown.  When does this method break down and how does it fail? If this is to be accepted, Monte Carlo level results are needed to give evidence in support of claims.}

The C-JPDAF is not completely heuristic; it solves an optimization problem, where the cost function itself can be considered heuristic, as many cost functions are. This is contrary to other coalescence-avoiding techniques that simply remove possible associations or artificially change estimation parameters known to cause coalescence.

Regarding the Monte Carlo results, running 100 trials of Cases I--IV with different noise (same noise levels) highlighted some important points. These are included under the paragraph ``Monte Carlo Results of All Scenarios'' at the end of the results section, copied below.

\begin{correction}
Since data association and filtering are fundamentally stochastic, Cases I--IV are repeated $100$ times with different noise to obtain mean metrics and number of failures.
For these cases, a failure is defined as at least one of the following: (i) track swapping (when estimates are consistently incorrectly associated with measurements from other estimates), (ii) when the estimates become ambiguous with respect to each other (their position relative to each other is consistently closer than half of the true estimate separation), or (iii) a numerical optimization does not converge (C-JPDAF only).
The following Monte Carlo results are tabulated in Table \ref{tab:MonteCarlo}.

In Case I, neither the JPDAF nor the M-JPDAF experience failures.
However, these data support that the error and uncertainty cost are reduced on average, without an increase in computation time.
The simulated results of Case III also serves to compare these two algorithms, where both are subject to failures.
The computation time and uncertainty cost follow the same trends as Case I.
However, both algorithms suffer from track swapping, but this happens roughly twice as often with the JPDAF than it does with the M-JPDAF. As a result, the RMS error of the JPDAF is almost twice that of the M-JPDAF.

All three algorithms are compared with Cases II and IV. Case II is designed as a scenario where both the JPDAF and M-JPDAF yield object estimates with ambiguous estimates, which is considered a failure all $100$ times.
However, when the C-JPDAF converges properly, this failure is mitigated and the RMS error is reduced.
Since the C-JPDAF is solved numerically rather than analytically, a global minimum is not always obtained in four-dimensional measurement space.
Conversely, the C-JPDAF algorithm converges well every time with Case IV. The C-JPDAF solves a simpler optimization problem, where the number of estimates is reduced from three to two and the measurement space is reduced by two dimensions with angles-only measurements (no range or range-rate).
In every trial, the optimized cost of the C-JPDAF avoids the track swapping and ambiguity of the JPDAF and M-JPDAF algorithms, and in consequence the C-JPDAF RMS error is reduced as well.
\end{correction}

\begin{center}
\begin{threeparttable}[h]
\caption{Monte Carlo Results} \label{tab:MonteCarlo}
\begin{tabularx}{0.65\textwidth}
{
>{$}c<{$} | >{$}c<{$} | >{$}c<{$} | >{$}c<{$} | >{$}c<{$} | >{$}c<{$}
%*{1}{>{$}c<{$}} |
%*{2}{>{$}c<{$}} |
%*{2}{>{$}c<{$}}
}
\toprule
\multirow{1}{*}{Approach} & \multirow{1}{*}{Case} & \multirow{1}{*}{RMS Err.} & \multirow{1}{*}{Time (sec)} & \multirow{1}{*}{Mean $J_P$} & \multirow{1}{*}{Failures}
\\
\midrule
\text{JPDAF}        & \text{I} & 0.12181 & 0.564 & 0.51843 & 0 \\
\text{M-JPDAF}    & \text{I} & 0.10791 & 0.563 & 0.49490 & 0 \\
\midrule
\text{JPDAF}        & \text{II} & 0.14520 & 1.668 & 0.24859 & 100 \\
\text{M-JPDAF}    & \text{II} & 0.11004 & 1.671 & 0.17791 & 100 \\
\text{C-JPDAF}\textsuperscript{*}    & \text{II} & 0.09752 & 117.115 & 0.18032 & 20 \\
\midrule
\text{JPDAF}        & \text{III} & 8.66749 & 2.764 & 16.5046 & 69 \\
\text{M-JPDAF}    & \text{III} & 4.76320 & 2.763 & 16.5023 & 36 \\
\midrule
\text{JPDAF}        & \text{IV} & 1.01311 & 1.107 & 0.00762 & 100 \\
\text{M-JPDAF}    & \text{IV} & 0.84104 & 1.107 & 0.00736 & 100 \\
\text{C-JPDAF}    & \text{IV} & 0.57102 & 31.360 & 0.00738 & 0 \\
\midrule
\end{tabularx}
{\small
\begin{tablenotes}
	\item RMS error, time, and mean $J_P$ are the mean values of $100$ trials.
	\item Failures are the number of failures out of $100$ trials.
	\item The RMS error unit for Cases I and II is km, and the RMS error unit for Cases III and IV is $10^{-6}$ rad.
	\item In Case II with the C-JPDAF (denoted with \textsuperscript{*}), the $20$ failures occurred when the optimal gains failed to converge, and the metrics are not based on the failed trials.
%    \item *: track swapping
%    \item $\dagger$: ambiguous
  \end{tablenotes}}
\end{threeparttable}
\end{center}

\subsubsection*{Technical Detail Comments}

\item {\itshape Eqn (9), you have $R_i$ as dependent on the object rather than dependent on the measurement.  Because you don't a priori know which objects produce which measurements, and you make no ``hard decision", the measurement noise can only be a function of the measurement itself, ``$R_j$"}

Note that this is the ``a priori innovation covariance matrix'', which has nothing to do with a particular $z_j$, but rather only the expected noise of a correctly associated measurement originating from the $i$-th estimate. The term $R_i$ is indexed with the objects because it might be different among various object estimates, e.g., an object much closer to a sensor might have a measurement with less expected noise than with an object much farther away. In short, the term $R_i$ is based on a correct association, where the measurement origin uncertainty is accounted for later.

\item {\itshape Eqn (11), the notation is very confusing.  Do not use i and j as both global markers, and summation counters in the same equation.}

The summation counter index is replaced with $l$.

\item {\itshape Eqn (12), the $e_i$ is boldface, while up until now, no variable type has been used as bold.  Please use consistent notation.}

The boldface has been removed from this variable.

\item {\itshape Eqn (14) and throughout the derivation, random variables don't appear to be correctly handled. $e_{ij}$ is a function of $z_j$.  $z_j$ is a random variable which means $e_{ij}$ is a random variable.  $\tilde P_i$ should not be a random variable.}

It is true that $z_j$ and $e_{ij}$ are random variables, but so too is $\tilde P_i$ (see [3]), because this positive semi-definite portion of the posterior covariance matrix depends on the innovation spread composed of random variables. This term is specific to soft decision updates only, so this type of term is not found in a nearest neighbor Kalman filter, for example.

\item {\itshape Section 4.1: There are some discrepancies between multivariate probability distributions and Gaussian distributions.  You describe Matusita's method as a measure for any two PDFs, but then define it for specifically, Gaussian variables.  Use consistent language.}

The measure is only valid for Gaussian distributions, and this is updated accordingly in the paper.

\item {\itshape In section 4.1, you use the term ``similarity measure''.  Measure is a mathematical term with a very specific meaning.  Please clarify or use different terminology.}

While the term ``measure'' does not come directly from measure theory, this is referred to as a ``measure'' because is represents the degree of similarity. The similarity measure is a positive scalar when the two estimates are the same, and approaches zero as their separation grows. The first sentence of Section 4.1 is revised as follows:

\begin{correction}
We choose Matusita's measure because it provides a degree of similarity among multivariate Gaussian distributions [18], [19]. 
\end{correction}

%The term is a ``measure'' because it is a systematic way of assigning a number. In this case, the number refers to the similarity of multivariate Gaussian random variables.

\item {\itshape Between Eqn (26) and (27) you make the claim that $P_i^+$ does not depend on any gain except $K_i$.  Your posterior distribution does depend on implicitly on previous gains.  Clarify or correct this statement.}

This point is clarified not to explicitly exclude past gains:

\begin{correction}
...where $P^+_i$ is obtained by (13), dependent on the gain $K_i$ at the current time step.
\end{correction}

\item {\itshape I am not convinced the weighting factor, $a$, cannot be taken out of eqn (31) and combined with $c$.  Same as with $c$, a discussion of what $a$ should be must be included.}

The weighting factor $a$ is part of a product inside the exponent, so it cannot be easily removed; $a$ affects the shape of the similarity cost, whereas $c$ only affects the size. The added portion from a previous comment above discusses this briefly:

\begin{correction}
The choice of $c$ depends on the type of scenario and the parameter $a$ shapes the similarity cost $J_S$.
\end{correction}

\item {\itshape The C-JPDAF appear to have a global minimum, but in 6 dimensional state space it will most likely be difficult to determine.  More discussion on ways to solve this problem would be useful.}

The paper provides a necessary condition for local optimality. However, global optimality is beyond the scope of this paper.
%It is in fact difficult to determine. A little more discussion on this with the Monte Carlo simulations is included per your suggestion.

\item {\itshape The thorough discussion of Keplarian orbital dynamics around eqn (35-36) is largely unnecessary considering the audience.}

This is true considering the majority of the audience. However, the contributions from this paper are easily extended to other data association problems, such as vision-based tracking of people or automobiles. Because other audiences do not necessarily have an aerospace background, we did not want to alienate this audience from understanding the results.

%%%%%%%%%%%%%%%%%%%%%%%%%

\end{itemize}

\end{document}

### 2018-06-09 Journal/JAS Data Association/JAS_DataAssoc_rev1.tex
journal added

--- 2018-06-09 Journal/JAS Data Association/JAS Data Association After Rev1 (1)/JAS Data Association Old/JAS Data Association NEW Derivation (1)/JAS_DataAssoc.tex
@@
 %A hard decision is when estimates are fully associated with measurements according to some metric between the estimates and the measurements.
-For example, a common heuristic approach is known as the nearest neighbor filter (NNF), which uses a distance measure between each measurement and the predicted measurements of each potential object to determine likely associations.
+For example, a common heuristic approach is known as the nearest neighbor filter (NNF), which uses a distance measure (e.g.
+Euclidean, Mahalanobis) between each measurement and the predicted measurements of each potential object to determine likely associations.
 %For example, a common heuristic approach is known as the nearest neighbor filter (NNF), which uses the Mahalanobis distance between each measurement and the predicted measurements of each potential object to determine likely associations.
-The measurement with the smallest Mahalanobis distance is considered as the correct one~\cite{NN2}.
+The measurement with the smallest distance is considered as the correct one~\cite{NN2}.
 However, when estimates are in close proximity, there is a high likelihood of making incorrect associations, which may be detrimental to the estimated states, so this technique is avoided in the analysis of this paper.
@@
 %More explicitly, Kalman filters are considered optimal in the sense that the estimator gain is selected to minimize the posterior uncertainties.
-In the current JPDAF applications, the Kalman gain is directly used in the measurement update processes, indeliberately assuming that the Kalman gain designed for a single object tracking is also optimal for multiple object tracking problems.
+In the current JPDAF applications, the Kalman gain is directly used in the measurement update processes, indeliberately assuming that the Kalman gain designed for a single object tracking is also optimal for multiple object tracking problems with soft decisions.
 
@@
 Therefore, we propose a new coalescence-avoiding optimal JPDAF, referred to as the C-JPDAF, which is a systematic approach to remove coalescence by minimizing the weighted sum of the uncertainty and a similarity index between estimates.
-The prior work from~\cite{KauLovLee14} covers the concept in simple cases, assuming that only two objects are considered with a single measurement at each time step, and that the state and measurement are very similar.
+The prior work from~\cite{KauLovLee14} covers the concept in simple cases, assuming that only two objects are considered with a single measurement at each time step, and that the object state and measurement are very similar.
 These assumptions are unrealistic in most scenarios.
@@
 
-In short, the main contributions of this paper are (i) deriving the optimal JPDAF that minimizes the posterior uncertainties in estimation, and (ii) proposing a systematic approach to avoid data coalescence of multiple objects in close-proximity.
+In short, the main contributions of this paper are (i) deriving the optimal JPDAF that minimizes the posterior mean square error and uncertainties in estimation, and (ii) proposing a systematic approach to avoid data coalescence of multiple objects in close-proximity.
 The desirable features of the proposed approaches are illustrated by extensive numerical simulations with satellites in orbits.
@@
 Suppose that these measurements are unordered and possibly include spurious measurements and missed detections.
-Data association and filtering serves to estimate the state of the systems when there exist uncertainties in the system processes, the measurements, and the measurement originations.
+Data association and filtering serves to estimate the state of the systems when there exist uncertainties in the system processes, the measurements, and the measurement origins.
 In this paper, the popular joint probabilistic data association filtering is improved in two distinct ways in the subsequent sections.
@@
 %Given the uncertainty of the measurements represented with covariance matrix $R_i$, t
-For the $j$-th measurement, the $i$-th a priori innovation covariance matrix $S_{i,j}\in\Re^{m\times m}$ represents the uncertainty between the $i$-th measurement estimate and its true measurement, defined as
+The $i$-th a priori innovation covariance matrix $S_i\in\Re^{m\times m}$ represents the uncertainty between the $i$-th measurement estimate and its true measurement, defined as
 %The a priori innovation covariance matrix $S_{i}\in\Re^{m\times m}$ for all innovations corresponding to the $i$-th object is defined as \EditTL{(Is it for the covariance of $e_{ii}$?
@@
 
-%A modified version of the JPDA, commonly known as the ``Cheap JPDA'', provides a simpler approximation of the likelihood $\beta_{i,j}$.
+%A modified version of the JPDA, commonly known as the ``Cheap JPDA'', provides a simpler approximation of the likelihood $\beta_{ij}$.
 %The equations below are adopted from~\cite[Eq. 1.(1-2)]{Bar1990}.
 
-Let $G_{i,j}\in\Re$ be a simplified scaled density function between the $i$-th measurement estimate and the $j$-th measurement,
+Let $G_{ij}\in\Re$ be a simplified scaled density function between the $i$-th measurement estimate and the $j$-th measurement,
 % variable which takes a very similar form as the Gaussian probability density function
 \begin{align}
-G_{i,j}=\frac1{\sqrt{\det{S_{i,j}}}}\exp{\{-\frac12e_{i,j}^TS_{i,j}^{-1}e_{i,j}\}}.
+G_{ij}=\frac1{\sqrt{\det{S_{i}}}}\exp{\{-\frac12e_{ij}^TS_{i}^{-1}e_{ij}\}}.
 \end{align}
@@
 %\begin{align}
-%\mathcal{S}_{ti}=\sum\limits_{j\in J}G_{i,j}, 
+%\mathcal{S}_{ti}=\sum\limits_{j\in J}G_{ij}, 
 %\quad 
-%\mathcal{S}_{ri}=\sum\limits_{i\in I}G_{i,j}.
+%\mathcal{S}_{ri}=\sum\limits_{i\in I}G_{ij}.
 %\end{align}
@@
 \label{eqn:betaDef}
-\beta_{i,j}=\frac{G_{i,j}}{\sum\limits_{l=1}^{n_r}G_{il}+\sum\limits_{l=1}^{n_t}G_{lj}-G_{i,j}+B},
+\beta_{ij}=\frac{G_{ij}}{\sum\limits_{l=1}^{n_r}G_{il}+\sum\limits_{l=1}^{n_t}G_{lj}-G_{ij}+B},
 \end{align}
@@
 \end{align}
-The conventional JPDAF chooses $K_i=P^-_iH_i^TS_i^{-1}$ (assuming $S_i$ is fixed for all measurements) as the gain from the Kalman filter.
+The conventional JPDAF chooses $K_i=P^-_iH_i^TS_i^{-1}$ as the gain from the Kalman filter.
 In this paper, we show that the Kalman gain does not actually minimize the posterior uncertainty in probabilistic data association filtering, and we derive the new \textit{optimal} gain $K_i$.
 
-Let the probability that $i$-th object is \textit{not} associated with any measurement be $\beta_{i,0}=1-{\sum_{j=1}^{n_r} \beta_{i,j}}\in\Re$.
-The case when $\beta_{i,0}=1$ implies that the $i$-th estimate is \emph{not} associated with any measurement completely, i.e., the $i$-th object is not measured by any sensor with the probability of one.
-In such trivial cases, there is no measurement update, i.e., $\hat x^+_i=\hat x^-_i$.
-
-Next we consider when $\beta_{i,0}\neq1$, i.e., $0\leq\beta_{i,0}<1$.
-Let the a priori and a posteriori errors be defined as
-\begin{align}
-\label{eqn:ErrFuns}
-\tilde x_i^-=\hat x_i^--x_i,
-\quad
-\tilde x_i^+=\hat x_i^+-x_i,
-\end{align}
-respectively.
-Then, using \refeqn{zlin}, \refeqn{zhatlin}, \refeqn{eij}, \refeqn{ErrFuns}, and the difference between \refeqn{KalEst} and \refeqn{xestapriori}, we have
-\begin{align}
-\tilde x_i^+&=
-%\hat x_i^+-\hat x_i^-+\tilde x_i^-
-%\nonumber\\&=
-K_i\sum_{j=1}^{n_r}\beta_{i,j}(z_j-\hat z_i)+\tilde x_i^-
-\nonumber\\&=
-%K_i\sum_{j=1}^{n_r}\beta_{i,j}(H_ix_i+v_j-H_i\hat x_i^-)+\tilde x_i^-
-%\nonumber\\&=
-%-K_iH_i\tilde x_i^-\sum_{j=1}^{n_r}\beta_{i,j}+K_i\sum_{j=1}^{n_r}\beta_{i,j}v_j+\tilde x_i^-
-%\nonumber\\&=
-(I-(1-\beta_{i,0})K_iH_i)\tilde x_i^-+K_i\sum_{j=1}^{n_r}\beta_{i,j}v_j.
-\end{align}
-
-The resulting posterior covariance $P^+_{i}\in\Re^{n\times n}$ is defined as
-%~\cite[Eq. D-32 and D-33]{TrackDataAssoc}
+Let the probability that $i$-th object is \textit{not} associated with any measurement be $\beta_{0,i}=1-{\sum_{j=1}^{n_r} \beta_{ij}}\in\Re$.
+The case when $\beta_{0,i}=1$ implies that the $i$-th estimate is \emph{not} associated with any measurement completely, i.e., the $i$-th object is not measured by any sensor with the probability of one.
+In such trivial cases, there is no measurement update, i.e., $\hat x^+_i=\hat x^-_i$ and $P_i^+=P_i^-$.
+
+Next we consider when $\beta_{0,i}\neq1$, i.e., $0\leq\beta_{0,i}<1$.
+The resulting posterior state covariance $P^+_{i}\in\Re^{n\times n}$ is derived in~\cite[Eq.
+D-32 and D-33]{TrackDataAssoc}, where this term is defined as $\text{E}[(x_i-\hat x_i^+)(x_i-\hat x_i^+)^T]$ and the resulting solution is derived from the law of total expectation considering each innovation individually.
+Then, $P^+_{i}$ is simplified to
 \begin{align}
 \label{eqn:JPDAFCov}
-P_i^+&=\text{E}[\tilde x_i^+\tilde x_i^{+T}]
-%\nonumber\\&=
-%\text{E}[
-%\{(I-(1-\beta_{i,0})K_iH_i)\tilde x_i^-+K_i\sum_{j=1}^{n_r}\beta_{i,j}v_j\}\{(I-(1-\beta_{i,0})K_iH_i)\tilde x_i^-+K_i\sum_{j=1}^{n_r}\beta_{i,j}v_j\}^T
-%]
-%\nonumber\\&=
-%(I-(1-\beta_{i,0})K_iH_i)\text{E}[\tilde x_i^-\tilde x_i^{-T}](I-(1-\beta_{i,0})K_iH_i)^T
-%+
-%K_i\text{E}[\sum_{j=1}^{n_r}\beta_{i,j}v_j\sum_{j=1}^{n_r}\beta_{i,j}v_j^T]K_i^T
-\nonumber\\&=
-(I-(1-\beta_{i,0})K_iH_i)P_i^-(I-(1-\beta_{i,0})K_iH_i)^T
-+
-K_i(\sum_{j=1}^{n_r}\beta^2_{i,j}R_j)K_i^T
-\nonumber\\&=
-P_i^--(1-\beta_{i,0})
-(K_iH_iP_i^-+P_i^-H_i^TK_i^T)
-+
-K_iE_iK_i^T.
-\end{align}
-where $E_i\in\Re^{m\times m}$ is defined for convenience as
-\begin{align}
-\label{eqn:E}
-E_i=\sum_{j=1}^{n_r}\beta^2_{i,j}R_j+(1-\beta_{i,0})^2H_iP_i^-H_i^T.
-\end{align}
-%$\tilde P_i\in\Re^{n\times n}$ is the portion of the $i$-th state covariance matrix due to the spread in the innovations
-%\begin{align}
-%\tilde P_i=&\ K_i\parenth{{\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T}K_{i}^T,\label{eqn:tildeP}
-%\end{align}
-%and $P_i^c\in\Re^{n\times n}$ corresponds to the $i$-th posterior covariance if only a single measurement serves to update the $i$-th estimate~\cite{TrackDataAssoc}%, and it can be written as
-%%We can define $P_i^c$ without any assumptions on $K_i$ (i.e., $K_i$ is not necessarily the Kalman gain, so the well-known Kalman covariance update equations are not valid here) as~\cite[Eq.
+P^+_{i}=&\ \beta_{0,i}P^-_{i}+(1-\beta_{0,i})P_i^c+\tilde P_i,
+\end{align}
+where $\tilde P_i\in\Re^{n\times n}$ is the portion of the $i$-th state covariance matrix due to the spread in the innovations
+\begin{align}
+\tilde P_i=&\ K_i\parenth{{\sum_{j=1}^{n_r} \beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T}K_{i}^T,\label{eqn:tildeP}
+\end{align}
+and $P_i^c\in\Re^{n\times n}$ corresponds to the $i$-th posterior covariance if only a single measurement serves to update the $i$-th estimate.
+This covariance term is defined as $P^c_i=(I-K_iH_i)P_i^-$ according to~\cite[Eq.
+D-27]{TrackDataAssoc}, or $P^c_i=P_i^--K_iS_iK_i^T$ according to~\cite[Eq.
+43]{JPDAF1}, where both equivalent versions assume that the Kalman gain is used.
+The derivation of $P^c_i$ is the only occurence in \refeqn{JPDAFCov} where the Kalman gain assumption is enforced.
+Without such an assumption, we have%, and it can be written as
+%We can define $P_i^c$ without any assumptions on $K_i$ (i.e., $K_i$ is not necessarily the Kalman gain, so the well-known Kalman covariance update equations are not valid here) as~\cite[Eq.
 4.2-12]{OptEst1}
-%\begin{align}
-%P^c_i=&\ (I-K_iH_i)P^-_i(I-K_iH_i)^T+K_iR_iK_i^T\nonumber
-%\\
-%=&\ P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i(H_iP^-_iH_i^T+R_i)K_i^T,
-%\label{eqn:GenCov}
-%\end{align}
-%which is commonly referred to as the \emph{Joseph form}, valid for \emph{any} gain matrix $K_i$~\cite{OptEst1}.
-%Substituting \refeqn{tildeP} and \refeqn{GenCov} into \refeqn{JPDAFCov},% and rearranging,
-%\begin{align}
-%P^+_{i}=&\ \beta_{i,0}P^-_{i}+(1-\beta_{i,0})[P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\nonumber
-%\\
-%&\ +K_i(H_iP^-_iH_i^T+R_i)K_i^T]+K_i\left({\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right)K_{i}^T\nonumber
-%\\
-%%=&\ P^-_{i}-(1-\beta_{i,0})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i\left((1-\beta_{i,0})S_i+{\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right)K_i^T\nonumber%asdf
-%%\\
-%=&\ P^-_{i}-(1-\beta_{i,0})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_iE_iK_i^T,\label{eqn:CovExpanded}
-%\end{align}
-%where the symmetric matrix $E_i=E_i^T\in\Re^{m\times m}$ is defined as
-%\begin{align}
-%E_i=(1-\beta_{i,0})S_i+{\sum_{j=1}^{n_r} \beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T,\label{eqn:E}
-%\end{align}
-%and it corresponds to an innovation covariance that accounts for the possibility that more than a single measurement may update the $i$-th estimate.
-Since \refeqn{JPDAFCov} and \refeqn{E} make no assumptions on $K_i$, this form is valid for \emph{any} gain $K_i$.
+\begin{align}
+P^c_i=&\ (I-K_iH_i)P^-_i(I-K_iH_i)^T+K_iR_iK_i^T\nonumber
+\\
+=&\ P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i(H_iP^-_iH_i^T+R_i)K_i^T,
+\label{eqn:GenCov}
+\end{align}
+which is commonly referred to as the \emph{Joseph form}, valid for \emph{any} gain matrix $K_i$~\cite{OptEst1}.
+Substituting \refeqn{tildeP} and \refeqn{GenCov} into \refeqn{JPDAFCov},% and rearranging,
+\begin{align}
+P^+_{i}=&\ \beta_{0,i}P^-_{i}+(1-\beta_{0,i})[P^-_i-(K_iH_iP^-_i+P^-_iH_i^TK_i^T)\nonumber
+\\
+&\ +K_i(H_iP^-_iH_i^T+R_i)K_i^T]+K_i\left({\sum_{j=1}^{n_r} \beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right)K_{i}^T\nonumber
+\\
+%=&\ P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_i\left((1-\beta_{0,i})S_i+{\sum_{j=1}^{n_r} \beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T\right)K_i^T\nonumber%asdf
+%\\
+=&\ P^-_{i}-(1-\beta_{0,i})(K_iH_iP^-_i+P^-_iH_i^TK_i^T)+K_iE_iK_i^T,\label{eqn:CovExpanded}
+\end{align}
+where the symmetric matrix $E_i=E_i^T\in\Re^{m\times m}$ is defined as
+\begin{align}
+E_i=(1-\beta_{0,i})S_i+{\sum_{j=1}^{n_r} \beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}}{\bar{e}}_{i}^T,\label{eqn:E}
+\end{align}
+and it corresponds to an innovation covariance that accounts for the possibility that more than a single measurement may update the $i$-th estimate.
+Because \refeqn{CovExpanded} is based on the Jordan form, this equation yields an accurate $P^+_{i}$ for \emph{any} gain $K_i$.
 
@@
 
-The Kalman gain employed by JPDAF is optimal if there exists no measurement uncertainty.
-This corresponds to applying a hard decision, where this is the special case when $E_i=S_i$ from \refeqn{E} and $K_i=P^-_iH_i^TS_i^{-1}$ from \refeqn{GainK} is the Kalman gain~\cite{OptEst1}.
-Since the JPDAF may apply soft decisions, it follows that the Kalman gain is suboptimal for the JPDAF in general.
-%These conclusions come from comparing the derived optimal gain and the Kalman gain, and the differences in the uncertainty cost functions associated with these gains, shown later in this paper.
-%
+The Kalman gain employed by JPDAF is optimal if there exists no measurement uncertainty or when the rows of the Kalman gain are orthogonal to the columns of $\sum_{j=1}^{n_r}\beta_{ij}e_{ij}e_{ij}^T-{\bf\bar{e}_i}{\bf\bar{e}_i}^T$ because this is the special case when $E_i=S_i$ from \refeqn{E} and $K_i=P^-_iH_i^TS_i^{-1}$ from \refeqn{GainK} is the Kalman gain~\cite{OptEst1}.
+Since the JPDAF does not require these assumptions, it follows that the Kalman gain is suboptimal for the JPDAF in general.
+These conclusions come from comparing the derived optimal gain and the Kalman gain, and the differences in the uncertainty cost functions associated with these gains, shown later in this paper.
+
 %
 %(You may reuse...
-%Note that $E_i$ is independent of $K_i$ and it is invertible when $\beta_{i,0}\neq1$ since $S_i$ is positive-definite and
-%
-%$(\beta_{i,j}e_{i,j}e_{i,j}^T-{\bar{e}}_{i}{\bar{e}}_{i}^T)$ is positive-semidefinite, as shown in the Appendix A. 
-%The case when $\beta_{i,0}=1$ corresponds with the $i$-th object \emph{not} being measured with full certainty, and such cases is analyzed in Appendix B.)
+%Note that $E_i$ is independent of $K_i$ and it is invertible when $\beta_{0,i}\neq1$ since $S_i$ is positive-definite and
+%
+%$(\beta_{ij}e_{ij}e_{ij}^T-{\bar{e}}_{i}{\bar{e}}_{i}^T)$ is positive-semidefinite, as shown in the Appendix A. 
+%The case when $\beta_{0,i}=1$ corresponds with the $i$-th object \emph{not} being measured with full certainty, and such cases is analyzed in Appendix B.)
 \end{proof}
@@
 	}
-\caption{A similarity index between three Gaussian distributions is illustrated as summation of overlapping regions of two distributions (green) or three distributions (brown).
+\caption{A similarity index between three Gaussian distributions is illustrated as the summation of overlapping regions of two distributions (green) or three distributions (brown).
 This concept is easily extended to multivariate Gaussian distributions.
@@
 \end{align}
-In these cases, the process covariance is $Q=\diag[10^{-6}I_{3\times3}\ \text{km}^2, 10^{-8}I_{3\times3}\ \text{km/s}^2]$, the estimates have initial covariances matrix $P_{0,A1}=10^7\times Q$, time steps of $1$ second over $100$ seconds, and $\epsilon=10$.
-These scenarios are chosen because the estimates exhibit large initial uncertainties, so there are soft decisions with large measurements shared and the portion of the posterior covariance affected by \refeqn{E} is significant.
+In these cases, the process covariance is $Q=\diag[10^{-6}I_{3\times3}\ \text{km}^2, 10^{-8}I_{3\times3}\ \text{km/s}^2]$, the estimates have initial covariances matrix $P_{0,A1}=10^7\times Q$, time steps of $1$ second over $100$ seconds, and $\epsilon=10$ for the perturbed cases.
+These scenarios are chosen because the estimates exhibit large initial uncertainties, so the spread in the innovations term from \refeqn{tildeP} affects the updates.
 The gain of the M-JPDAF accounts for this portion of the state covariance matrix better than the gain for the conventional JPDAF in every single case.
@@
 Given identical parameters and measurements, the RMS errors of the position and the mean uncertainty cost $J_P$ defined in \refeqn{JP} are lower in \emph{every} case when the M-JPDAF is applied compared with when the conventional JPDAF is applied.
-In fact in two cases, namely A1b and A1c, the JPDAF is incapable of decreasing the estimate uncertainty as quickly as the M-JPDAF.
-In some cases, the JPDAF suffers from track swapping, which is when measurements are consistently associated incorrectly.
+%In fact in two cases, namely A1b and A1c, the JPDAF is incapable of decreasing the estimate uncertainty as quickly as the M-JPDAF.
+In two cases, namely A1b and A1c, the JPDAF suffers from track swapping, which is when measurements are consistently associated incorrectly.
 %; the JPDAF initially suffers from such great coalescence that the estimates become associated with incorrect measurements consistently, known as track swapping.
@@
 where $Q=\diag[10^{-6}I_{3\times3}\ \text{km}^2, 10^{-8}I_{3\times3}\ \text{km/s}^2]$, $P_{0,B1}=Q$, and $\epsilon=1.1$ for time steps of $10$ seconds over a $2000$ second time period.
-This corresponds to a phased-array radar observing observing a slow separation of LEO objects in a near-circular orbit; because the satellite states are in close proximity over a large time period, the JPDAF and M-JPDAF are subject to severe coalescence.
+This corresponds to a phased-array radar observing a slow separation of LEO objects in a near-circular orbit; because the satellite states are in close proximity over a large time period, the JPDAF and M-JPDAF are subject to severe coalescence.
 The C-JPDAF is designed to avoid this coalescence, where the parameters $a=5\times10^{-1}$ and $c=2.3\times10^{-4}$ from \refeqn{zetaGeneral} and \refeqn{CostGen} are used in these scenarios from trial and error.
@@
 In all cases the M-JPDAF is more accurate and experiences less coalescence than the JPDAF, but both algorithms suffer from coalescence.
-The C-JPDAF systematically removes this coalescence from the updates at the cost of slightly larger covariance matrices and a large increase in required computational resources.
+The C-JPDAF systematically removes this coalescence from the updates at the cost of covariance matrices with slightly larger traces and a large increase in required computational resources.
 %Between Cases B1 and B1c, the process position uncertainty is increased, which last little effect on the estimates since this has little
@@
 The angular RMS errors and computation times are tabulated in Table \ref{tab:B2} and some example plots from these cases are shown in Figure \ref{fig:B2}.
-In applications with a large number of objects to track, the C-JPDAF need not be applied on most object; only those close neighbors where the similarity among the estimates yield coalescence with the JPDAF or the M-JPDAF.
+In applications with a large number of objects to track, the C-JPDAF need not be applied on most objects; only those close neighbors where the similarity among the estimates yield coalescence with the JPDAF or the M-JPDAF.
 This way, a data association scheme with an analytic solution, such as the JPDAF or the M-JPDAF, could be applied to the remaining estimates to mitigate the large computational requirements of the C-JPDAF.
@@
 The following Monte Carlo results are tabulated in Table \ref{tab:MonteCarlo}.
+
+In Case A1, neither the JPDAF nor the M-JPDAF experience failures.
+However, these data support that the error and uncertainty cost are reduced on average, without an increase in computation time.
+The simulated results of Case A2 also serves to compare these two algorithms, where both are subject to failures.
+The computation time and uncertainty cost follow the same trends as Case A1.
+However, both algorithms suffer from track swapping, but this happens roughly twice as often with the JPDAF than it does with the M-JPDAF.
+As a result, the RMS error of the JPDAF is almost twice that of the M-JPDAF.
+
+All three algorithms are compared with Cases B1 and B2.
+Case B1 is designed as a scenario where both the JPDAF and M-JPDAF yield object estimates with ambiguous estimates, which is considered a failure all $100$ times.
+However, when the C-JPDAF converges properly, this failure is mitigated and the RMS error is reduced.
+Since the C-JPDAF is solved numerically rather than analytically, a global minimum is not always obtained in four-dimensional measurement space.
+Conversely, the C-JPDAF algorithm converges well every time with Case B2.
+The C-JPDAF solves a simpler optimization problem, where the number of estimates is reduced from three to two and the measurement space is reduced by two dimensions with angles-only measurements (no range or range-rate).
+In every trial, the optimized cost of the C-JPDAF avoids the track swapping and ambiguity of the JPDAF and M-JPDAF algorithms, and in consequence the C-JPDAF RMS error is reduced as well.
 
@@
 
-In Case A1, neither the JPDAF nor the M-JPDAF experience failures.
-However, these data support that the error and uncertainty cost are reduced on average, without an increase in computation time.
-The simulated results of Case A2 also serves to compare these two algorithms.
-The computation time and uncertainty cost follow the same trends as Case A1.
-However, both algorithms suffer from track swapping, but this happens roughly twice as often with the JPDAF than it does with the M-JPDAF.
-As a result, the RMS error of the JPDAF is almost twice that of the M-JPDAF.
-
-All three algorithms are compared with Cases B1 and B2.
-Case B1 is designed as a scenario where both the JPDAF and M-JPDAF yield object estimates with ambiguous estimates, which is considered a failure all $100$ times.
-However, when the C-JPDAF converges properly, this failure is mitigated and the RMS error is reduced.
-Since the C-JPDAF is solved numerically rather than analytically, a global minimum is not always obtained in four-dimensional measurement space.
-The C-JPDAF algorithm converges well every time with Case B2.
-The C-JPDAF solves a simpler optimization problem, where the number of estimates is reduced from three to two and the measurement space is reduced by two dimensions with angles-only measurements (no range or range-rate).
-In every trial, the optimized cost of the C-JPDAF avoids the track swapping and ambiguity of the JPDAF and M-JPDAF algorithms, and in consequence the C-JPDAF RMS error is reduced as well.
-
 \section{Conclusion}

### 2018-06-09 Journal/JAS OD EKF/JAS OD EKF Old/JAS_OD_EKF[Conflict].tex
journal added

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%%!PS-Adobe-3.0 EPSF-3.0
%%%BoundingBox: 19 19 221 221
%%%CreationDate: Mon Sep 29 1997
%%%Creator: programmed by hand (JK)
%%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

\usepackage{bm}
\usepackage{amssymb,amsmath,times,subfigure,tabularx,booktabs,colortbl,multirow,threeparttable}
\usepackage{subfigure}
%\usepackage[notref,notcite]{showkeys}  % use this to temporarily show labels
\usepackage[colorlinks=true, pdfstartview=FitV, linkcolor=black, citecolor= black, urlcolor= black]{hyperref}
\usepackage{overcite}
\usepackage{footnpag}			      	% make footnote symbols restart on each page

\usepackage{color}
\usepackage{afterpage}
\newcommand{\hilight}[1]{\colorbox{green}{#1}}

\newtheorem{prop}{Proposition}

\newcommand{\abs}{\mathop{\mathrm{abs}}\nolimits}
\newcommand{\diag}{\mathop{\mathrm{diag}}\nolimits}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\ave}[1]{\mathrm{E}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\dderiv}[2]{\ensuremath{\dfrac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}

\newcommand{\x}{\ensuremath{\mathbf{x}}}
\renewcommand{\r}{\mathbf{r}}
\renewcommand{\u}{\mathbf{u}}
\newcommand{\y}{\mathbf{y}}

\newcommand{\bfi}{\bfseries\itshape\selectfont}

\graphicspath{{./Figs/}}

\begin{document}

\title{Nonlinear Observability for Relative Orbit Determination with Angles-Only Measurements
\thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Evan Kaufman         \and
		T. Alan Lovell         \and
        	Taeyoung Lee
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{E. Kaufman \at
              Dept. of Mechanical and Aerospace Engineering \\
		 The George Washington University \\
		 Washington DC 20052 \\
              Tel.: 216-346-0946\\
              \email{evankaufman@gwu.edu}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           T. A. Lovell \at
              Air Force Research Laboratory, Space Vehicles Directorate \\
		 Kirtland AFB, NM\\
		 \email{AFRL.RVSV@kirtland.af.mil}
	    \and
		 T. Lee \at
              Dept. of Mechanical and Aerospace Engineering \\
		 The George Washington University \\
		 Washington DC 20052 \\
              Tel.: 202-994-8710\\
		 Fax: 202-994-0238\\
              \email{tylee@gwu.edu}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
}
\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
This paper presents nonlinear observability criteria for the relative orbital dynamics represented by the solutions of the two-body problem. It is assumed that a chief is on a circular orbit with a prescribed orbital radius, and it measures lines-of-sight toward a deputy only. A differential geometric method, based on the Lie derivatives, is used to derive sufficient conditions for observability of the orbital properties of the deputy. It is shown that under certain geometric conditions on the relative configuration between the chief and the deputy, the nonlinear relative motion is observable from angles-only measurements. The second part of this paper presents a quantitative measure of observability for the relative orbits, and it is formulated by generalizing the observability Gramian of linear dynamic systems. An extended Kalman filter is also developed to numerically illustrate the observability of nonlinear relative orbits with angles-only measurements and to show correspondence between the proposed observability measure with filtering errors explicitly.

%A new nonlinear observability measure is proposed for relative orbit determination when lines-of-sight between satellites are measured only. It corresponds to a generalization of the observability Gramian in linear dynamic systems to the nonlinear relative orbit dynamics represented by the two-body problems. An extended Kalman filter (EKF) is adapted to this problem and is evaluated with various gravitational harmonics and initial orbital determination (IOD) predictions. Extensive results illustrate correspondence between the proposed observability measure with filtering errors. An extensive numerical analysis in realistic scenarios includes satellite propagation of the two-body problem the $J_2$ perturbation effects.
\keywords{Nonlinear Observability \and Relative Orbit Determination \and Angles-Only Measurements}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}

Space-based surveillance or relative navigation is desirable for many spacecraft missions, such as formation control and rendezvous. Spacecraft maneuvers based only on on-board measurements reduce the total operating cost significantly, and it improves safety against communication interruptions with ground stations.  Relative navigation between spacecraft in close-proximity essentially corresponds to space-based orbit determination.  In particular, vision-based navigation and estimation of relative orbits have received attention recently, since optical sensors have the desirable properties of low cost and minimal maintenance, while providing accurate line-of-sight measurements, or angles from a chief to a deputy.

Relative navigation based on angles-only measurements has been investigated in~\cite{WofGelITAES09,Tom11,PatLovPASFMM12}.  The problem is to determine the relative orbit between a chief spacecraft and a deputy spacecraft by using the line-of-sight between the two objects, assuming that the orbit of the chief is prescribed exactly.  Reference \cite{WofGelITAES09} shows that the relative orbit is unobservable from angles-only measurements when linear relative orbital dynamics are assumed, unless there are thrusting maneuvers. Reference \cite{Tom11} investigates observability by using a relative orbit model linearized in terms of spherical coordinates.  Reference~\cite{PatLovPASFMM12} introduces the concept of partial observability to determine a basis vector representing a family of relative orbits, and an initial orbit determination technique is developed for this method. Linear observability analysis is performed numerically for a particular case in~\cite{YimCraPASMM04}.

All of these results are based on linear relative orbital dynamics. It is straightforward to see that the relative orbit is not observable with angles-only measurements through its linearized dynamics. In other words, there are an infinite number of relative orbits that yield the identical line-of-sight measurements, and the orbital distance between a deputy and a chief cannot be determined by angles-only measurements. As such, it is required to study the nonlinear relative orbital dynamics to determine observability with angles-only measurements.

%Recently, observability criteria have been derived for the nonlinear relative orbital dynamics represented by the solutions of the two-body problem without linearization~\cite{LovLeePISSFD14}.

This paper constructs observability criteria for the nonlinear relative orbital dynamics represented by the solutions of the two-body problem without linearization. Assuming that a chief is on a circular orbit with a prescribed orbital radius, nonlinear equations of motion for the relative orbital motion of a deputy with respect to the chief are presented.  A differential geometric method, based on the Lie derivatives of the line-of-sight from the chief to the deputy~\cite{HerKreITAC77}, is applied to obtain sufficient conditions for observability.  It is shown that under certain geometric conditions on the relative configuration between the chief and the deputy, the nonlinear relative motion is observable from angles-only measurements.

Next, we present a new measure of observability for relative orbit determination with angles-only measurements. This is motivated to understand the degree of observability for the relative orbital dynamics, and to identify the class of relative orbits that have stronger or weaker observability. For a given initial condition of relative orbits, the proposed observability measure determines how much the corresponding relative orbits are easy or difficult to estimate, thereby providing a quantitative measure for relative orbit determination.

The last part of this paper is focused on a feasibility study of an extended Kalman filter (EKF) for orbit determination using angles-only measurements in realistic scenarios. The filtering performances of the EKF are studied for several types of relative orbits to show nonlinear observability with angles-only measurements explicitly, and they are compared with the proposed observability measure. Additional cases with $J_2$ perturbations are also considered to identify the effects of orbital perturbations.

In short, the main contributions of this paper are constructing sufficient conditions for the observability of nonlinear relative orbits with angles-only measurements that have not been studied before, and proposing a new quantitative measure of observability.

This paper is organized as follows. The relative orbital dynamics and the orbit determination problem are formulated at Section 2. Observability criteria and a observability measure are presented at Section 3 and Section 4, respectively, which are followed by numerical examples and conclusions.

%The filter is subject to measurement uncertainty and perturbation forces not considered with the two-body problem, on which the filter is based. Then, the performance of the EKF is compared with the proposed observability measure to show the correspondence.

\section{Nonlinear Relative Orbital Dynamics}\label{sec:ND}

Consider two satellites orbiting around the Earth, referred to as a \textit{chief} and a \textit{deputy}. Suppose that the chief satellite is on a circular orbit with a pre-determined orbital radius of $a\in\Re$.

Define a local-vertical, local-horizontal (LVLH) frame as follows. Its origin is located at the chief satellite. The $x$-axis is along the radial direction from the Earth to the chief, and the $y$-axis is along the velocity vector of the chief. The $z$-axis is normal to the orbital plane, and it is parallel to the angular momentum vector of the chief. The LVLH frame is rotating with the angular velocity of $\mathbf{\omega}=[0,0,n]^T\in\Re^3$, where $n=\sqrt{\frac{\mu}{a^3}}$ is the mean motion of the chief satellite, and $\mu$ denotes the gravitational parameter of the Earth. Note that the velocity of the chief is given by $\mathbf{v}_{chief}=[0,na,0]^T\in\Re^3$. Let the relative position of the deputy satellite with respect to the chief be given by $\mathbf{r}=[x,y,z]^T\in\Re^3$ in the LVLH frame.

\subsection{Nonlinear Equations of Motion}

Nonlinear equations of motion for the relative motion of the deputy with respect to the chief can be derived as follows based on Lagrangian mechanics~\cite{LovLeePISSFD14}.

Considering that the LVLH frame is rotating with the angular velocity of $\mathbf{\omega}=[0,0,n]^T$, the inertial velocity of the deputy satellite is given by
\begin{align*}
\mathbf{v}_{deputy}
=
\mathbf{v}_{chief} + \dot{\r} + \mathbf{\omega}\times \mathbf{r} = 
\begin{bmatrix}
0 \\ na \\ 0
\end{bmatrix}
+
\begin{bmatrix}
\dot x \\ \dot y \\ \dot z
\end{bmatrix}
+
\begin{bmatrix}
-n y \\ nx \\ 0
\end{bmatrix}
=
\begin{bmatrix}
\dot x - ny \\ \dot y + nx + na \\ \dot z
\end{bmatrix}.
\end{align*}
Therefore, the (normalized) kinetic energy of the deputy satellite is 
\begin{align*}
T = \frac{1}{2}\|\mathbf{v}_{deputy}\|^2 = \frac{1}{2} \braces{(\dot x -ny)^2+(\dot y + nx +na)^2 + \dot z^2}.
\end{align*}
The location of the Earth from the chief is given by $[-a,0,0]^T$ in the LVLH frame. Therefore, the position vector of the deputy from the center of the Earth is given by 
$\r_a=[x+a,y,z]^T\in\Re^3$. The gravitational potential energy is 
\begin{align*}
U = -\frac{\mu}{\sqrt{(x+a)^2 + y^2 + z^2}} = -\frac{\mu}{\|\r_a\|}.
\end{align*}
From the above equations, the Lagrangian of the deputy satellite is expressed as $L(x,y,z,\dot x,\dot y ,\dot z) = T(x,y,\dot x,\dot y,\dot z)- U(x,y,z)$. Using the Euler-Lagrange equations, given by
\begin{align*}
\frac{d}{dt}\deriv{L}{\dot q}-\deriv{L}{q}=0,
\end{align*}
for $q\in\{x,y,z\}$, we obtain the nonlinear equations of motion for the relative orbit as follows.
\begin{align}
\ddot x - 2n\dot y -n^2 x&=n^2 a - \frac{\mu (x+a)}{((x+a)^2 + y^2 + z^2)^{3/2}},\label{eqn:ddotx}\\
\ddot y + 2n\dot x -n^2 y &=   - \frac{\mu y}{((x+a)^2 + y^2 + z^2)^{3/2}},\label{eqn:ddoty}\\
\ddot z &= - \frac{\mu z}{((x+a)^2 + y^2 + z^2)^{3/2}}.\label{eqn:ddotz}
\end{align}

These can be written as the standard form of the nonlinear state equation,
\begin{align}
\dot{\mathbf{x}} = f (\mathbf{x}), \label{eqn:xxdot}
\end{align}
where the state vector is $\mathbf{x}=[x,y,z,\dot x, \dot y,\dot z]^T\in\Re^N$ with $N=6$, and
\begin{align}
f(\mathbf{x}) 
%= \begin{bmatrix}
%\dot x\\
%\dot y\\
%\dot z\\
%2n\dot y +n^2 (x+a) - \dfrac{\mu (x+a)}{((x+a)^2 + y^2 + z^2)^{3/2}}\\
%-2n\dot x +n^2 y  - \dfrac{\mu y}{((x+a)^2 + y^2 + z^2)^{3/2}}\\
%- \dfrac{\mu z}{((x+a)^2 + y^2 + z^2)^{3/2}}
%\end{bmatrix}
=\begin{bmatrix}
\dot \r \\
-2\omega\times \dot\r - \omega\times(\omega\times \r_a) -\dfrac{\mu \r_a}{\|\r_a\|^3}
\end{bmatrix}.\label{eqn:f}
\end{align}
%where $\r=[x,y,z]^T\in\Re^3$, $\r_a = [x+a,y,z]^T\in\Re^3$, and $\dot\r =[\dot x,\dot y,\dot x]^T\in\Re^3$.

\subsection{Line-of-sight Measurement}

We assume that the line-of-sight from the chief to the deputy is measured by onboard sensors, such as optical sensors. The measurement is represented by the unit-vector along the relative position vector, i.e.,
\begin{align}
\y = h(\mathbf{x}) =\frac{\r}{\|\r\|},\label{eqn:y}
\end{align}
where $\y\in\Re^M$ with $M=3$, and it satisfies $\|\y\|=1$.

\section{Nonlinear Observability Criteria}\label{sec:OC}

Based on the nonlinear dynamic model presented at the previous section, here we derive sufficient conditions for the observability of the relative orbit with line-of-sight measurements.

\subsection{Observability Criteria for Nonlinear Systems}

Observability of nonlinear systems and sufficient conditions for observability are summarized as follows~\cite{HerKreITAC77}. For a given nonlinear dynamic system described with \refeqn{xxdot} and \refeqn{y}, a pair of points $\x_0$ and $\x_1$ is called \textit{indistinguishable} if the outputs of the corresponding solutions starting from each of $\x_0$ and $\x_1$ are identical for a certain time period. Note that $\x_0$ is indistinguishable from $\x_0$ trivially from the definition. The system is \textit{locally weakly observable} at $\x_0$, if there exists an open neighborhood $V$ of $\x_0$ such that for every open neighborhood $V_0$ of $\x_0$ contained in $V$, the only indistinguishable point to $\x_0$ is the point $\x_0$ itself. This implies that $\x_0$ is distinguishable from any other points in a neighborhood of it.

Next, sufficient conditions for observability are presented. The Lie-derivative of the output $h(\x)$ along $f(\x)$, namely $L_f h(\x)$ is defined as follows:
\begin{align*}
L_f h(\x) = \deriv{h(\x)}{\x}f(\x)\in\Re^{M\times 1},
\end{align*}
which corresponds to the directional derivative of $h(\x)$ along $f(\x)$, and $\deriv{h(\x)}{\x}\in\Re^{M\times N}$. For a non-negative integer $i$, the $i$-th order Lie-derivative is defined by induction as $L_f^i h = L_f (L_f^{i-1} h)$ with $L_f^0 h = h$. Define an observability matrix $\mathcal{O}\in\Re^{NM\times N}$ as
\begin{align*}
\mathcal{O}(\x_0) = \deriv{}{\x} \begin{bmatrix} L_f^0 h(\x) \\ L_f^1 h(\x)\\ \vdots \\L_f^{N-1} h(\x)\end{bmatrix}\Bigg|_{\x=\x_0}.
\end{align*}
It has been shown that the system is locally weakly observable at $\x_0$ if the rank of the observability matrix $\mathcal{O}(\x_0) = N$. When applied to linear dynamics, this yields the well-known observability rank condition for linear systems. Note that when there is more than a single measurement, i.e., $M>1$, the observability rank condition can be satisfied without need for computing the higher-order Lie derivatives up to the $N-1$-th order.

\subsection{Observability Criteria for Relative Orbital Dynamics}

\paragraph{Observability Matrix}
We apply the above observability criteria for the nonlinear relative orbital dynamics. The first block of rows of the observability matrix is simply given by
\begin{align*}
\deriv{}{\x} L_f^0 h(\x) = \deriv{\y}{\x}.
\end{align*} 
Similarly, from \refeqn{xxdot}, the second row block is
\begin{align*}
\deriv{}{\x} L_f^1 h(\x) = \deriv{}{\x}\parenth{\deriv{\y}{\x}f(\x)}=\deriv{\dot\y}{\x}.
\end{align*} 
Since $\dot\y$ is determined completely from $\x$, we have
\begin{align*}
\deriv{}{\x} L_f^2 h(\x) = \deriv{}{\x}\parenth{\deriv{\dot\y}{\x}f(\x)}=\deriv{\ddot\y}{\x}.
\end{align*}
Using these and the fact that $\x=[\r^T,\dot \r^T]$, the observability matrix of the relative orbital dynamics can be written as
\begin{align}
\mathcal{O} =
\begin{bmatrix}
\deriv{\y}{\r} & \deriv{\y}{\dot\r}\\
\deriv{\dot\y}{\r} & \deriv{\dot\y}{\dot\r}\\
\deriv{\ddot\y}{\r} & \deriv{\ddot\y}{\dot\r}
\end{bmatrix}
\triangleq
\begin{bmatrix}
\mathcal{O}_{00} & 0_{3\times 3}\\
\mathcal{O}_{10} & \mathcal{O}_{11}\\
\mathcal{O}_{20} & \mathcal{O}_{21}
\end{bmatrix},\label{eqn:OO}
\end{align}
where the submatrices $\mathcal{O}_{ij}\in\Re^{3\times 3}$ are defined for each block, and we used the fact that the measurement $\y$ is independent of $\dot\r$. Here we consider the part of the observability matrix obtained by the second order Lie derivatives of the measurement due to complexity. But, this still provides sufficient conditions for observability with interesting geometric meanings described later.

After straightforward but tedious algebraic manipulations using the following identity repeatedly,
\begin{align*}
\delta \parenth{\frac{1}{\|\mathbf{r}\|^i}} 
& = -i\frac{ \mathbf{r}^T \delta \mathbf{r}}{\|\mathbf{r}\|^{i+2}},
\end{align*}
for any positive integer $i$, we can show that the submatrices $\mathcal{O}_{ij}$ of the observability matrix $\mathcal{O}$ are given by
\begin{align}
\mathcal{O}_{00} & = \deriv{\y}{\r} = \frac{1}{\|\r\|}(I - \y\y^T),\label{eqn:O_00}\\
\mathcal{O}_{10} & = \deriv{\dot\y}{\r} = \deriv{}{\r} \parenth{\deriv{\y}{\r}\dot \r}= -\frac{1}{\|\r\|^2}\braces{
{\dot\r\y^T +\y^T\dot\r I +\y\dot\r^T}
- 3\y\y^T\dot\r \y^T},\\
\mathcal{O}_{11} & = \deriv{\dot\y}{\dot\r} = \deriv{}{\dot\r} \parenth{\deriv{\y}{\r}\dot \r}= \deriv{\y}{\r} = \mathcal{O}_{00},\label{eqn:O11}\\
\mathcal{O}_{20} & = \deriv{\ddot\y}{\r} = -\dfrac{2\dot\r \dot\r^T  +\dot\r^T\dot\r I}{\|\r\|^3}
+3\dfrac{2(\r^T\dot\r) \dot\r\r^T +(\dot\r^T\dot\r)\r\r^T
+(\r^T\dot\r)^2 I
+2(\r^T\dot\r)\r  \dot\r^T
}{\|\r\|^5}\nonumber \\
&\quad- 15 \dfrac{(\r^T\dot\r)^2\r  \r^T}{\|\r\|^7}
-\dfrac{\ddot{\r}\r^T +\r^T\ddot{\r} I +\r \ddot{\r}^T}{\|\r\|^3}
+ 3 \dfrac{(\r^T\ddot{\r})\r \r^T}{\|\r\|^5}\nonumber\\
&\quad+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}},\\
\mathcal{O}_{21} & = \deriv{\ddot\y}{\dot\r} = \deriv{}{\dot\r}\parenth{\deriv{}{\r}\parenth{\deriv{\y}{\r}\dot\r}\dot\r + \deriv{\y}{\r}\ddot\r} = 2 \deriv{}{\r} \parenth{\deriv{\y}{\r}\dot\r} + \deriv{\y}{\r} \deriv{\ddot\r}{\dot\r}
\nonumber\\
&  = 2\mathcal{O}_{10} 
- 2 \mathcal{O}_{00}[\omega]_\times,\label{eqn:O_21}
\end{align}
where $[\omega]_\times\in\Re^{3\times 3}$ is defined as
\begin{align*}
[\omega]_\times = \begin{bmatrix} 0 & -n & 0 \\ n & 0 & 0 \\ 0 & 0 & 0 \end{bmatrix}.
\end{align*}
These expressions have been verified by the Matlab symbolic computation toolbox.

After some algebraic manipulations, we can show that the submatrices satisfy the following identities, 
\begin{align}
\mathcal{O}_{10}\r & = 
%-\frac{1}{\|\r\|}\braces{{\dot\r\y^T +\y^T\dot\r I +\y\dot\r^T}- 3\y\y^T\dot\r \y^T}\y\\
%&=-\frac{1}{\|\r\|}\braces{\dot\r +(\y^T\dot\r)\y +\y(\dot\r^T\y)- 3\y(\y^T\dot\r) }\\
%&=-\frac{1}{\|\r\|}\braces{\dot\r -\y(\y^T\dot\r) }=
-\mathcal{O}_{00}\dot\r,\label{eqn:O10r}\\
\mathcal{O}_{10}\dot\r & = -\frac{1}{\|\r\|^2}\braces{
{2\dot\r(\y^T\dot \r)  +\y(\dot\r^T\dot \r)}
- 3\y(\y^T\dot\r)^2},\label{eqn:O10dotr}\\
\mathcal{O}_{20}\r %& = 
%-\dfrac{2\dot\r \dot\r^T\y  +\dot\r^T\dot\r \y}{\|\r\|^2}
%+3\dfrac{2(\y^T\dot\r) \dot\r\y^T\y +(\dot\r^T\dot\r)\y\y^T\y+(\y^T\dot\r)^2 y +2(\y^T\dot\r)\y \dot\r^T\y}{\|\r\|^2}
%- 15 \dfrac{(\y^T\dot\r)^2\y  \y^T\y}{\|\r\|^2}\nonumber \\
%&\quad -\dfrac{\ddot{\r}\y^Ty +\y^T\ddot{\r} \y +\y \ddot{\r}^T\y}{\|\r\|^1}
%+ 3 \dfrac{(\y^T\ddot{\r})\y \y^T\y}{\|\r\|}
%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
%&=
%-\dfrac{2\dot\r (\dot\r^T\y)  +(\dot\r^T\dot\r) \y}{\|\r\|^2}
%+3\dfrac{2(\y^T\dot\r) \dot\r +(\dot\r^T\dot\r)\y+(\y^T\dot\r)^2 \y +2(\y^T\dot\r)^2\y }{\|\r\|^2}
%- 15 \dfrac{(\y^T\dot\r)^2\y}{\|\r\|^2}\nonumber \\
%&\quad -\dfrac{\ddot{\r} +(\y^T\ddot{\r}) \y +\y (\ddot{\r}^T\y)}{\|\r\|^1}
%+ 3 \dfrac{(\y^T\ddot{\r})\y }{\|\r\|}
%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
%&=
%-\dfrac{2\dot\r (\dot\r^T\y)  +(\dot\r^T\dot\r) \y}{\|\r\|^2}
%+3\dfrac{2(\y^T\dot\r) \dot\r +(\dot\r^T\dot\r)\y+(\y^T\dot\r)^2 \y +2(\y^T\dot\r)^2\y }{\|\r\|^2}
%- 15 \dfrac{(\y^T\dot\r)^2\y}{\|\r\|^2}\nonumber \\
%&\quad -\dfrac{\ddot{\r} +(\y^T\ddot{\r}) \y +\y (\ddot{\r}^T\y)}{\|\r\|^1}
%+ 3 \dfrac{(\y^T\ddot{\r})\y }{\|\r\|}
%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
%&=
%\dfrac{4\dot\r (\dot\r^T\y)  -2(\dot\r^T\dot\r)\y-6(\y^T\dot\r)^2 \y }{\|\r\|^2}
% -\mathcal{O}_{00}\ddot \r
%+\parenth{\dfrac{I}{\|\mathbf{r}\|} - \dfrac{\mathbf{r}\mathbf{r}^T}{\|\mathbf{r}\|^3}}
%\parenth{-[\omega]_\times^2 -\dfrac{\mu I_{3\times 3}}{\|\r_a\|^3} + \dfrac{3\mu\r_a\r_a^T}{\|\r_a\|^5}}\r,\\
& = -2\mathcal{O}_{10}\dot\r +\mathcal{O}_{00}\braces{-\ddot \r +\deriv{\ddot\r}{\r}\r},\label{eqn:O20r}\\
\mathcal{O}_{21}\r & = -2\mathcal{O}_{00} (\dot\r+\omega\times\r),\label{eqn:O21r}\\
\mathcal{O}_{21}\dot\r & = 2\mathcal{O}_{10} \dot\r
- 2 \mathcal{O}_{00}[\omega]_\times\dot\r,\label{eqn:O21rdot}
\end{align}
which are useful to derive the observability criteria.

\paragraph{Observability Rank Condition}

Now we present sufficient conditions that the observability matrix $\mathcal{O}$ has the full rank of six.

\begin{prop}
Define three vectors $\mathbf{v}_{rel}$, $\mathbf{a}_1,\mathbf{a}_2\in\Re^3$ as 
\begin{align}
\mathbf{v}_{rel} & = \dot\r +\omega\times\r,\label{eqn:vrel}\\
\mathbf{a}_1 & = \ddot\r-\deriv{\ddot\r}{\r}\r = -2\omega\times \dot\r - [\omega]_\times^2 a \mathbf{e}_1 -\dfrac{\mu a }{\|\r_a\|^3}\mathbf{e}_1- \dfrac{3\mu\r_a^T\r}{\|\r_a\|^5}\r_a,\label{eqn:a1}\\
\mathbf{a}_2 & = \ddot\r-\deriv{\ddot\r}{\r}\r -\deriv{\ddot\r}{\dot\r}\dot\r = \mathbf{a}_1+2\omega\times\dot\r= - [\omega]_\times^2 a \mathbf{e}_1 -\dfrac{\mu a }{\|\r_a\|^3}\mathbf{e}_1- \dfrac{3\mu\r_a^T\r}{\|\r_a\|^5}\r_a,\label{eqn:a2}
\end{align}
where $\mathbf{e}_1=[1,0,0]^T\in\Re^3$ and $\r_a=[x+a,y,z]^T\in\Re^3$. The nonlinear relative orbital dynamics are locally weakly observable at $\x=[\r^T,\dot\r^T]$ if
%\begin{list}{}{\setlength{\leftmargin}{1cm}\setlength{\itemsep}{0mm}\setlength{\parsep}{0mm}\setlength{\topsep}{0mm}\setlength{\parskip}{0mm}\setlength{\labelwidth}{2cm}}
%\item[(i)] when $\r\times\dot\r =0$, $\r\times\mathbf{a}_1\neq 0$ and $\r^T(\mathbf{v}_{rel}\times \mathbf{a}_1)\neq 0$,
%\item[(ii)] when $\r\times\dot\r \neq 0$, $\r\times\mathbf{a}_2\neq 0$ and $\r^T(\mathbf{v}_{rel}\times \mathbf{a}_2)\neq 0$.
%\end{list}
%\begin{alignat}{2}
%(i)&\; &\text{when $\r\times\dot\r =0$, $\r\times\mathbf{v}_{rel}\neq0$, $\r\times\mathbf{a}_1\neq 0$,  and $\r^T(\mathbf{v}_{rel}\times \mathbf{a}_1)\neq 0$},\label{eqn:cond1}\\
%(ii)& &\text{when $\r\times\dot\r \neq 0$, $\r\times\mathbf{v}_{rel}\neq0$, $\r\times\mathbf{a}_2\neq 0$ and $\r^T(\mathbf{v}_{rel}\times \mathbf{a}_2)\neq 0$.}\label{eqn:cond2}
%\end{alignat}
\begin{itemize}
\item[(i)] when $\r\times\dot\r =0$, 
\begin{gather}
\r\times\mathbf{v}_{rel}\neq0,\; \r\times\mathbf{a}_1\neq 0,\; \text{ and } \r^T(\mathbf{v}_{rel}\times \mathbf{a}_1)\neq 0,\label{eqn:cond1}
\end{gather}
\item[(ii)] when $\r\times\dot\r \neq 0$, 
\begin{gather}
\r\times\mathbf{v}_{rel}\neq0,\; \r\times\mathbf{a}_2\neq 0,\; \text{ and } \r^T(\mathbf{v}_{rel}\times \mathbf{a}_2)\neq 0.\label{eqn:cond2}
\end{gather}
\end{itemize}

\end{prop}

\begin{proof}
We show that if the above conditions are satisfied, then the six columns of the observability matrix are linearly independent. Suppose that for a constant vector $\mathbf{c}=[\mathbf{c}_1^T,\mathbf{c}_2^T]\in\Re^6$, where $\mathbf{c}_1,\mathbf{c}_2\in\Re^3$, we have $\mathcal{O}\mathbf{c}=0$, i.e., 
\begin{align}
\mathcal{O} \mathbf{c} = \begin{bmatrix}
\mathcal{O}_{00} & 0_{3\times 3}\\
\mathcal{O}_{10} & \mathcal{O}_{11}\\
\mathcal{O}_{20} & \mathcal{O}_{21}
\end{bmatrix}
\begin{bmatrix} \mathbf{c}_1 \\ \mathbf{c}_2 \end{bmatrix}
=
\begin{bmatrix}
\mathcal{O}_{00}\mathbf{c}_1\\
\mathcal{O}_{10}\mathbf{c}_1+ \mathcal{O}_{11}\mathbf{c}_2\\
\mathcal{O}_{20}\mathbf{c}_1+ \mathcal{O}_{21}\mathbf{c}_2
\end{bmatrix}
=
\begin{bmatrix}
0_{3\times 1}\\
0_{3\times 1}\\
0_{3\times 1}
\end{bmatrix}.\label{eqn:Oc}
\end{align}

We wish to show that there is no non-zero vector $\mathbf{c}$ satisfying \refeqn{Oc} under the given conditions. At the first three rows of \refeqn{Oc}, we have
\begin{align}
\mathcal{O}_{00} \mathbf{c}_1 = \frac{1}{\|\r\|}(I-\y\y^T) \mathbf{c}_1 = 0.\label{eqn:c10}
\end{align}
The matrix $\mathcal{O}_{00}$ has a one-dimensional null space spanned by $\y$. Therefore, without loss of generality, we can choose $\mathbf{c}_1=0$ or $\mathbf{c}_1=\r$. As we are interested in any non-zero vector satisfying the above equation, let $\mathbf{c}_1=\r$ for the subsequent development.

For the chosen value of $\mathbf{c}_1=\r$, we find the next three rows of \refeqn{Oc} as
\begin{align}
\mathcal{O}_{10} \mathbf{c}_1 +\mathcal{O}_{11} \mathbf{c}_2 & = \mathcal{O}_{10} \r +\mathcal{O}_{11} \mathbf{c}_2  = \mathcal{O}_{00} (-\dot\r + \mathbf{c}_2)\nonumber\\
& = \frac{1}{\|\r\|}(I-\y\y^T)(-\dot\r + \mathbf{c}_2)=0,\label{eqn:row2}
\end{align}
where we have used \refeqn{O11} and \refeqn{O10r}. Next, we consider two cases of \refeqn{row2}, namely (i) when $\r\times\dot\r=0$, and (ii) when $\r\times\dot\r\neq 0$. For each case, we show that $\mathbf{c}=0$ to satisfy \refeqn{Oc} under the given sufficient conditions.

\paragraph{Case (i): $\r\times\dot \r=0$} In this case, $\dot \r$ can be written as $\dot \r = \alpha\r$ for some constant $\alpha$ as $\r$ is parallel to $\dot\r$. Then, \refeqn{row2} reduces to 
\begin{align*}
\frac{1}{\|\r\|}(I-\y\y^T)\mathbf{c}_2=0,
\end{align*}
which implies that $\mathbf{c}_2= c\r$ for an arbitrary constant $c$. For the given choice of $\mathbf{c}=[\r^T,c\r^T]^T$, the last three rows of \refeqn{Oc} are given by
\begin{align*}
\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
& = \mathcal{O}_{20} \r + c\mathcal{O}_{21} \r.
\end{align*}
Using \refeqn{O20r} and \refeqn{O21r}, this can be rewritten as
\begin{align*}
\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
& =
-2\mathcal{O}_{10}\dot\r +\mathcal{O}_{00}\braces{-\ddot \r +\deriv{\ddot\r}{\r}\r}
-2c\mathcal{O}_{00} (\dot\r+\omega\times\r).
\end{align*}
But, from \refeqn{O10dotr}, we can show that $2\mathcal{O}_{10}\dot\r=0$ when $\r$ is parallel to $\dot \r$. Using \refeqn{vrel} and \refeqn{a1}, this further reduces to
\begin{align}
\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
& =
-\mathcal{O}_{00}(\mathbf{a}_1 +2c \mathbf{v}_{rel})\nonumber\\
& = - \frac{1}{\|\r\|}(I-\y\y^T)(\mathbf{a}_1 +2c \mathbf{v}_{rel})=0.\label{eqn:row3i}
\end{align}
The matrix $(I-\y\y^T)$ represents the orthogonal projection of a vector into the plane normal to $\y$. From the second condition of \refeqn{cond1}, namely $\r\times\mathbf{a}_1\neq 0$, we have $(I-\y\y^T)\mathbf{a}_1\neq 0$, which implies that the constant $c$ cannot be simply chosen as $c=0$. Therefore, the only possible case to satisfy the above equation is when $\mathbf{a}_1 +2c \mathbf{v}_{rel}$ is parallel to $\r$ for some values of $c$. However, that is not feasible since the third condition of \refeqn{cond1}, namely $\r^T(\mathbf{v}_{rel}\times\mathbf{a}_1)\neq0$,  implies that the three vectors $\r$, $\mathbf{a}_1$, and $\mathbf{v}_{rel}$ do not belong to a common plane, i.e., there is no constant $c$ such that $\mathbf{a}_1 +2c \mathbf{v}_{rel}$ is parallel to $\r$.

Therefore, there is no $\mathbf{c}\in\Re^6$ satisfying \refeqn{Oc} if $\mathbf{c}_1=\r$, under the given condition \refeqn{cond1}. This implies that $\mathbf{c}_1=0$ from \refeqn{c10}. Substituting this back to \refeqn{Oc}, we have $\mathcal{O}_{11}\mathbf{c}_2=\mathcal{O}_{00}\mathbf{c}_2=0$, which follows that $\mathbf{c}_2=c\r$ for some constant $c$. However, when $\mathbf{c}_2=c\r$, the last three rows of \refeqn{Oc} are given by
\begin{align}
\mathcal{O}_{21}\mathbf{c}_2 = 2c(\mathcal{O}_{10}-\mathcal{O}_{00}[\omega]_\times) \r
= -2c\mathcal{O}_{00}\mathbf{v}_{rel}=0,\label{eqn:row3i1}
\end{align}
where \refeqn{O10r} is used. But, from the first condition of \refeqn{cond1}, we have $\mathcal{O}_{00}\mathbf{v}_{rel}\neq 0$, and therefore $c=0$, i.e., $\mathbf{c}_2=c\r =0$.

In short, under the given condition \refeqn{cond1}, the equation \refeqn{Oc} implies that $\mathbf{c}=0$. Therefore, the six columns of the observability matrix $\mathcal{O}$ are linearly independent.

\paragraph{Case (ii): $\r\times\dot \r\neq0$} Next, we consider the second case of \refeqn{row2}. It implies that $-\dot\r + \mathbf{c}_2$ is parallel to $\r$, or equivalently, $\mathbf{c}_2 = \dot\r + c\r$ for an arbitrary constant $c$. For the given choice of $\mathbf{c}=[\r^T,\dot\r+ c\r^T]^T$, the last three rows of \refeqn{Oc} are given by
\begin{align*}
\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
& = (\mathcal{O}_{20} \r+\mathcal{O}_{21}\dot\r) + c\mathcal{O}_{21} \r.
\end{align*}
From \refeqn{O20r}, \refeqn{O21r}, \refeqn{O21rdot}, this can be rewritten as
\begin{align*}
\mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
= -\mathcal{O}_{00}(\mathbf{a}_2+2c\mathrm{v}_{ref})= -\frac{1}{\|\r\|}(I-\y\y^T)(\mathbf{a}_2+2c\mathbf{v}_{rel})=0.
\end{align*}
By following the same argument given after \refeqn{row3i}, under \refeqn{cond2}, there is no $c$ satisfying the above equation.

This implies that $\mathbf{c}_1=0$. Then, by the same argument given at \refeqn{row3i1}, we have $\mathbf{c}_2=0$. In short, under the given condition \refeqn{cond2}, the equation \refeqn{Oc} implies that $\mathbf{c}=0$. Therefore, the six columns of the observability matrix $\mathcal{O}$ are linearly independent. 
\end{proof}

This theorem shows that the nonlinear relative orbital dynamics is indeed observable from angles-only measurements if certain geometric conditions are satisfied. This generalizes the prior observability study developed for the linear relative orbital dynamics only.

Next, we try to interpret the given observability criteria. In particular, we consider the cases where the given sufficient conditions \refeqn{cond1} and \refeqn{cond2} are violated. For both cases, we have $\r\times\mathbf{v}_{rel}\neq 0$. At \refeqn{vrel}, $\mathbf{v}_{rel}$ corresponds to the relative velocity observed in the inertial frame. Therefore, the given sufficient conditions for observability are violated when the relative velocity vector is parallel to the relative position vector in the inertial frame. This corresponds to the cases when the deputy is moving directly toward or away from the chief, and it is not likely that such relative motion is maintained over a finite time period.

The second condition of each of \refeqn{cond1} and \refeqn{cond2} is satisfied in general, as the expressions for $\mathbf{a}_1$ and $\mathbf{a}_2$ representing certain accelerations are relatively arbitrary at \refeqn{a1} and \refeqn{a2}. Similar to the first condition, there is little chance that the second condition is violated over a finite time period entirely.

The third condition of \refeqn{cond1} implies that three vectors $\r$, $\mathbf{v}_{rel}$ and $\mathbf{a}_1$ do not belong to the same plane. The third condition of \refeqn{cond2} has a similar structure as well. Unlike the prior two conditions, this can be violated over a certain time period if the relative motion is planar, i.e., when $z(t)\equiv 0$ for all $t$.

However, \refeqn{cond1} and \refeqn{cond2} are sufficient conditions for observability, and the fact that any of these condition is not satisfied does not necessarily imply that the considered point is not observable. In such cases, the third or higher order Lie derivatives should be checked to determine observability. This paper is focused on verifying the observability of nonlinear relative orbits via angles-only measurements under certain conditions, and deriving the higher-order Lie derivatives may be a topic for future work.

%The main contribution of this section of the paper is showing that under certain geometric conditions, the nonlinear relative orbital dynamics are indeed observable via angles-only measurements.

\section{Observability Measure of Nonlinear Systems}

The observability criteria derived in the previous section guarantee that the nonlinear relative orbital dynamics are indeed observable under certain geometric conditions. However, it does not provide any information about how strong the observability is.

Several approaches have been proposed for an observability measure of nonlinear systems. 
%The measure of observability has been generalized to nonlinear dynamic systems in order to obtain balanced realizations.
An energy-like observability function is introduced~\cite{SchSCL93,Sch94,NewKriPICDC98} to measure the degree of contribution of an initial state to the output, and it is applied to a pendulum system. But, this approach is only applicable to asymptotically stable equilibrium states. An empirical observability Gramian is introduced~\cite{LalMarPIWC99,LalMarIJRNC02}, which is essentially a covariance matrix of the output computed by a number of sample trajectories.
It has been applied to a balanced realization of chemical processes~\cite{HahEdgJPC03,HahEdgCCE02}. However, this approach is based on the assumption that the output converges to a steady-state value, and it is not clear how  sample trajectories are selected. Therefore, these approaches cannot be applied for the relative orbital dynamics that do not have steady-state responses.

Another nonlinear observability measure is introduced~\cite{KreIdePICDC09}, which is basically the Gramian for the sensitivity of the output with respect to the initial condition:
\begin{align}
\mathcal{W} (\x_0,t_0,t_f) = \int_{t_0}^{t_f} \parenth{\deriv{\y(\tau)}{\x_0}}^T\deriv{\y(\tau)}{\x_0}\,d\tau,\label{eqn:Wo_NL}
\end{align}
where $\y(t)$ denotes the output of the system at $t=\tau$ for the state trajectory $\x(t)$ starting from the initial condition given by $\x(t_0)=\x_0$, and $\deriv{\y}{\x_0}\in\Re^{M\times N}$ is defined such that its $i,j$-th element is $\deriv{\y_i}{\x_{0_j}}$.

In general, the Gramian of a set of time-dependent functions represents the degree of linear independence of those functions over a given period of time. Therefore, the observability Gramian defined at \refeqn{Wo_NL} measures how much the sensitivities of the output with respect to the initial condition are linearly independent of each other. This represents the degree of observability, since for example, if $\deriv{\y}{\x_{0_i}}$ is linearly dependent to $\deriv{\y}{\x_{0_j}}$, then it is difficult to distinguish the effects on $\delta \x_{0_i}$ to the output $\y$ from the effects of $\delta \x_{0_j}$ on $\y$. Therefore, the condition number of $\mathcal{W} (\x_0,t_0,t_f)$ can be used as a measure of observability. When applied to linear dynamic systems, we can show that \refeqn{Wo_NL} reduces to the common observability Gramian for linear dynamic systems.

However, obtaining an analytic expression for the sensitivity of the output with respect to the initial condition is challenging for nonlinear systems in general. Therefore, a numerical approach is considered as follows. The $i,j$--th element of $\mathcal{W}_o$ can be approximated by
\begin{align}
[\mathcal{W}]_{ij} = \frac{1}{4\epsilon^2}\sum_{k=0}^N (\y^{i+}_k-\y^{i-}_k)^T(\y^{j+}_k-\y^{j-}_k) \Delta t_k,
\end{align}
where $\y^{i\pm}_k$ denotes the value of $\y$ at $t=t_k$ obtained from the state trajectory starting from the initial condition of $\x(t_0)=\x_0\pm \epsilon \mathbf{e}_i $ for a positive constant $\epsilon$, and the standard basis $\mathbf{e}_i$ of $\Re^n$.

\section{Numerical Examples}

%Preliminary results between the filter performance of an extended Kalman filter and the nonlinear observability Gramian are shown.

%We consider the relative orbital dynamics between two satellites, where the observability of the deputy relative to the chief satellite is on a prescribed circular orbit when only $\mathbf{y}$ is assumed to be measured.
%Then, an extended Kalman filter is developed for numerical analyses, where the satellites are either subject to two-body forces or with gravitational $J_2$ perturbations included.

To illustrate observability of nonlinear relative orbital dynamics, we develop an extended Kalman
filter and apply it to several types of relative orbits. For each orbit, the observability measure and the corresponding filtering performance are compared.

\subsection{Extended Kalman Filter}

Let the initial estimate of the state vector and its covariance be $\hat \x_0^+\in\Re^6$ and $P^+_0\in\Re^{6\times 6}$, respectively. Extended Kalman filters are composed of the following two updates, namely the flow update and the measurement update.

\paragraph{Flow Update}
The flow update provides an a priori state estimate $\hat \x^-_k$ with covariance matrix $P^-_k$ based on the equations of motion. The a priori state estimate is updated by the nonlinear equations of motion given by \refeqn{xxdot}. The a priori covariance matrix is defined as
\begin{align}
P^-_k=F_{k-1}P^+_{k-1}F_{k-1}^T+Q_k,
\end{align}
where the covariance matrix for the process noise is given by $Q_k\in\Re^{6\times 6}$, and the state transition matrix from the linearized system is given by
\begin{align}
F_{k-1}=\exp\left(
\begin{bmatrix}
0_{3\times3} & I_{3\times3} \\
\left(-[\omega]_\times^2-\frac{\mu I_{3\times3}}{\norm{\r_a}^3}+\frac{3\mu \r_a \r_a^T}{\norm{\r_a}^5}\right) & -2[\omega]_\times
%0 & 0 & 0 & 1 & 0 & 0\\
%0 & 0 & 0 & 0 & 1 & 0\\
%0 & 0 & 0 & 0 & 0 & 1\\
%3*n^2 & 0 & 0 & 0 & 2*n & 0\\
%0 & 0 & 0 & -2*n & 0 & 0\\
%0 & 0 & -n^2 & 0 & 0 & 0
\end{bmatrix}
(t_k-t_{k-1})\right).
\end{align}

\paragraph{Measurement Update}
The linearized measurement matrix $H_k$ is derived from \refeqn{y},
\begin{align}
H_k=
\begin{bmatrix}
\left(\frac1{\norm{\hat \r_{k}^-}}I_{3\times3}-\frac{\hat \r_{k}^-{\hat \r_{k}^{-T}}}{\norm{\hat \r_{k}^-}^3}\right) & 0_{3\times3}
\end{bmatrix},
\end{align}
where $\hat \r_{k}^-$ is the a priori estimate of $\r$ given by the first three elements of $\hat \x_{k}^-$.
Based on \refeqn{y}, the measurement prediction is given by
\begin{align}
\hat{\y}_k=h(\hat \r_{k}^-)=\frac{\hat \r_{k}^-}{\norm{\hat \r_{k}^-}}.
\end{align}
The posterior state estimate and its covariance matrix are given by
\begin{align}
\x^+_k&=\x^-_k+K_k(\y_k-\hat{\y}_k),
\\
P^+_k&=(I_{6\times6}-K_kH_k)P^-_k,
\end{align}
such that $K_k$ is chosen to minimize $\tr{P^+_k}$ with
\begin{align}
K_k=P^-_kH_k^T(H_kP^-_kH_k+R_k)^{-1},
\end{align}
where $R_k\in\Re^{3\times 3}$ denotes the covariance matrix for the measurement noise.

\subsection{Observability with Varying Deputy Inclinations}

It is assumed that the chief is on a circular orbit with the altitude of $500\,\mathrm{km}$ and zero inclination. Here, we study the effects of the deputy inclination on the observability measure and filtering performance. It is motivated by the fact that the given sufficient conditions for observability are violated if the inclination of the deputy is zero as well, since the deputy is on the same orbital plane with the chief. The eccentricity of the deputy is chosen as $e_{deputy}=0.2$ and the semi-major axis of the deputy is chosen such that its orbital period is identical to the chief, i.e.,
\begin{align*}
a_{deputy} = \parenth{\frac{T_{chief}\sqrt{\mu}}{2\pi}}^{2/3}.
\end{align*}
Therefore, the relative orbits are periodic. The inclination of the deputy is varied as $i_{deputy}=0,\,10,\,\cdots,\,50^\circ$.

For the extended Kalman filter, the initial estimate of the state is chosen as two times of its true value, i.e., $\hat\x_0=2\x_0$. This implies that there is a large initial error in the magnitude of the state, which is difficult to estimate accurately using angles-only measurements. The initial covariance of the state is chosen as $P_0=\mathrm{diag}[50^2I_{3\times 3}, (50n)^2I_{3\times 3}]$. The covariance matrices for the process noise and the measurement noise are $Q_k=\mathrm{diag}[10^{-6}I_{3\times 3}, 10^{-8}I_{3\times 3}]$, and $R_k=1.306^2 I_{3\times 3}\,\mathrm{deg}^2$. It is assumed that the line-of-sight is measured at every $\Delta t = 0.47$ seconds. It is simulated for 20 orbits of the chief around the Earth.

To compare the convergence property of each case, the following measures for estimation errors are introduced.
\begin{align*}
e_{dir_k}  & = \cos^{-1}\parenth{\frac{\mathbf{x}(t_k)^T \hat{\mathbf{x}} (t_k)}
{\|\mathbf{x}(t_k)\| \|\hat{\mathbf{x}} (t_k)\|}},\\
e_{mag_k}  & = 
\frac{|\|\mathbf{x}(t_k)\|-\|\hat{\mathbf{x}} (t_k)\||}{\|\mathbf{x}(t_k)\|},
\end{align*}
where the error variable $e_{dir}$ represents the angle between the true state and the estimated state, and the variable $e_{mag}$ corresponds to the normalized error for the magnitude of the state vector. This choice of error variables are motivated by the fact that the lines-of-sight provide direct measurements on the relative direction, but they do not yield any information about the range directly.

The condition number of the proposed observability Gramian and the mean values of the above error variables are summarized at Table \ref{tab:Ei} for each inclination. The second column shows that the condition number of the Gramian decreases as the inclination increases, i.e., it is relatively easier to determine the relative orbit via angles-only measurements if the inclination of the deputy is higher. However, the condition numbers are quite large overall, and it shows that there is only weak observability. At the third  and sixth columns, the mean error in the magnitude of the estimated state decreases as the inclination increases, and it is consistent with the condition number of the observability Gramian. The correlation between the condition number and the estimation error is more apparent at the fourth and seventh columns, which show the error in the magnitude of the estimated state averaged over the last 10 orbits. This is because there is greater transient response in the estimated state when the inclination is larger. The steady-state estimation error represented by the fourth column strictly decreases as the inclination increases. There is no meaningful variation at the fifth and eighth columns, as the estimation errors in the direction are determined mostly by the line-of-sight measurements. In the sixth, seventh, and eighth columns, the $J_2$ perturbation is considered in the chief and deputy orbit process models. The perturbation tends to increase the magnitude and direction errors, especially when the deputy satellite is at large inclinations, but a correlation between a lower condition number of the Gramian and an increase in filter accuracy is still largely maintained.

%% Table before J2
%\begin{table}[h]
%\caption{Observability Gramian and estimation errors for varying deputy inclinations}\label{tab:Ei}
%\begin{center}
%\begin{tabularx}{0.7\textwidth}{>{\centering $}X<{$}*{4}{>{$}c<{$}}}\toprule
%i_{deputy} & \mathrm{cond}[\mathcal{W}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}]\,(rad) \\\midrule
% 0 & 10^{11.7944} &    0.2783  &  0.2328  &  0.0155\\
%10 & 10^{11.5021} &    0.1408  &  0.0872  &  0.0113\\
%20 & 10^{11.0530} &    0.0762  &  0.0355  &  0.0102\\
%30 & 10^{10.7826} &    0.0566  &  0.0214  &  0.0112\\
%40 & 10^{10.7744} &    0.0567  &  0.0190  &  0.0173\\
%50 & 10^{10.7619} &    0.0425  &  0.0079  &  0.0169\\
%\bottomrule
%\end{tabularx}\\
%(${}^\dagger$ averaged for the last 10 orbits)
%\end{center}
%\end{table}

\begin{table}[h]
\caption{Observability Gramian and estimation errors for varying deputy inclinations}
\begin{center}
\begin{tabularx}{0.95\textwidth}
{
>{$}c<{$}
*{1}{>{$}c<{$}} |
*{3}{>{$}c<{$}} |
*{3}{>{$}c<{$}}
}
\toprule
\multirow{2}{*}{$i_{dep}$} & \multirow{2}{*}{$\mathrm{cond}[\mathcal{W}]$} & \multicolumn{3}{c}{\multirow{1}{*}{Two-Body Solution}} & \multicolumn{3}{c}{\multirow{1}{*}{$J_2$ Perturbation}} \\
 & &  \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}] \\\midrule
%{>{\centering $}X<{$}*{7}{>{$}c<{$}}}\toprule
%i_{dep} & \mathrm{cond}[\mathcal{W}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}]\,(rad) \\\midrule
 0 & 10^{11.7944} &    0.2783  &  0.2328  &  0.0155     	& 0.2664 & 0.2329 & 0.0169\\
10 & 10^{11.5021} &    0.1408  &  0.0872  &  0.0113 	& 0.1042 & 0.1013 & 0.0088\\
20 & 10^{11.0530} &    0.0762  &  0.0355  &  0.0102 	& 0.0840 & 0.0837 & 0.0097\\
30 & 10^{10.7826} &    0.0566  &  0.0214  &  0.0112 	& 0.0787 & 0.0784 & 0.0117\\
40 & 10^{10.7744} &    0.0567  &  0.0190  &  0.0173 	& 0.0841 & 0.0849 & 0.0136\\
50 & 10^{10.7619} &    0.0425  &  0.0079  &  0.0169 	& 0.0961 & 0.0960 & 0.0179\\
\bottomrule
\label{tab:Ei}
\end{tabularx}\\
(${}^\dagger$ averaged for the last 10 orbits)
\end{center}
\end{table}

Estimation results are also illustrated at Figures \ref{fig:i0}--\ref{fig:i50}. At Figure \ref{fig:i50}, the magnitude of the estimated state becomes quite close to the true value after four orbits. This confirms the proposed observability of nonlinear relative orbital dynamics with angles-only measurements.

\clearpage\newpage
\begin{figure}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.3\textwidth]{PE_i_0_e.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.3\textwidth]{PE_i_0_xx_norm.pdf}}
}
\caption{Estimation results: $i=0$}\label{fig:i0}
\end{figure}
\begin{figure}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.3\textwidth]{PE_i_30_e.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.3\textwidth]{PE_i_30_xx_norm.pdf}}
}
\caption{Estimation results: $i=30$}\label{fig:i30}
\end{figure}
\begin{figure}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.3\textwidth]{PE_i_30_e_J2.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.3\textwidth]{PE_i_30_xx_norm_J2.pdf}}
}
\caption{Estimation results with $J_2$ perturbations: $i=30$}\label{fig:i30}
\end{figure}
\begin{figure}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.3\textwidth]{PE_i_50_e.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.3\textwidth]{PE_i_50_xx_norm.pdf}}
}
\caption{Estimation results: $i=50$}\label{fig:i50}
\end{figure}
%\begin{figure}
%\centerline{
%	\subfigure[Estimation errors]{
%		\includegraphics[width=0.3\textwidth]{PE_i_50_e_J2.pdf}}
%	\hspace*{0.15\textwidth}
%		\subfigure[Magnitude of state (red:true, blue:estimate)]{
%		\includegraphics[width=0.3\textwidth]{PE_i_50_xx_norm_J2.pdf}}
%}
%\caption{Estimation results with $J_2$ perturbations: $i=50$}\label{fig:i50}
%\end{figure}

\clearpage\newpage

\subsection{Observability with Varying Size of Relative Orbits}

Next, we study the effects of the size of the relative orbits on nonlinear observability. Same as before, the chief is on a circular orbit with the altitude of $500\,\mathrm{km}$ and zero inclination. The initial conditions for the relative position and the velocity are chosen as
\begin{gather*}
\r_0=[r_{x_0},\,0,\,0]^T\,\mathrm{km},\quad
\mathbf{v}_0=[0,\,-2r_{x_0}n,\,0]^T\,\mathrm{km/s},
\end{gather*}
where $r_{x_0}\in\Re$ is the initial relative distance that is varied as $10,\,20,\,50,\,100\,\mathrm{km}$. For the linearized relative orbital dynamics, these initial conditions yield elliptic orbits around the chief, but for the nonlinear relative orbital dynamics, there is drift proportional to the size of the orbit parameterized by $r_{x_0}$. While the resulting orbits are planar such that the third condition of \refeqn{cond1} and \refeqn{cond2} is violated, it is still desirable to understand how the orbital size or drift affects nonlinear observability. For the extended Kalman filter, the initial estimate of the state is chosen as $\hat\x_0=1.5\x_0$, and the other parameters are identical to the prior case.

The corresponding condition number of the observability Gramian and the estimation errors are summarized at Table \ref{tab:Es}. It is shown that as $r_{x_0}$ increases, the condition number decreases and the estimation error for the magnitude of the estimated state decreases.

% Table before J2 is added
\begin{table}[h]
\caption{Observability Gramian and estimation errors for varying size of orbits}\label{tab:Es}
\begin{center}
\begin{tabularx}{0.7\textwidth}{>{\centering $}X<{$}*{4}{>{$}c<{$}}}\toprule
r_{x_0}\,(\mathrm{km}) & \mathrm{cond}[\mathcal{W}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}]\,(\mathrm{rad}) \\\midrule
 10 & 10^{15.8194} & 2.7739  &  2.8535  &  0.0022\\
 20 & 10^{15.2832} &  1.1787  &  1.1093  &  0.0022\\
 50 & 10^{14.4052} &  0.3146  &  0.2284  &  0.0020\\
100 & 10^{12.9984} &  0.1038  &  0.0354  &  0.0016\\
\bottomrule
\end{tabularx}\\
(${}^\dagger$ averaged for the last 10 orbits)
\end{center}
\end{table}

%% With J2
%\begin{table}[h]
%\caption{Observability Gramian and estimation errors for varying size of orbits}\label{tab:Es}
%\begin{center}
%\begin{tabularx}{\textwidth}
%{
%>{$}c<{$}
%*{1}{>{$}c<{$}} |
%*{3}{>{$}c<{$}} |
%*{3}{>{$}c<{$}}
%}
%\toprule
%\multirow{2}{*}{$r_{x_0}\,(\mathrm{km}) $} & \multirow{2}{*}{$\mathrm{cond}[\mathcal{W}]$} & \multicolumn{3}{c}{\multirow{1}{*}{Two-Body Solution}} & \multicolumn{3}{c}{\multirow{1}{*}{$J_2$ Perturbation}} \\
% & &  \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}] \\\midrule
%%{>{\centering $}X<{$}*{7}{>{$}c<{$}}}\toprule
%%i_{dep} & \mathrm{cond}[\mathcal{W}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}]\,(rad) \\\midrule
%10 & 10^{15.8194} & 2.7739  &  2.8535  &  0.0022 & 2.6790 & 1.927 & 0.0028\\
% 20 & 10^{15.2832} &  1.1787  &  1.1093  &  0.0022 & 1.024 & 0.8542 & 0.0027\\
% 50 & 10^{14.4052} &  0.3146  &  0.2284  &  0.0020 & 0.1531 & 0.1473 & 0.0025\\
%100 & 10^{12.9984} &  0.1038  &  0.0354  &  0.0016 & 0.1318 & 0.1197 & 0.0023\\
%\bottomrule
%\end{tabularx}\\
%(${}^\dagger$ averaged for the last 10 orbits)
%\end{center}
%\end{table}

\afterpage{\clearpage
\begin{figure}[p]
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.45\textwidth]{EKF_PE_r0_20_e.pdf}}
	\hspace*{0.05\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.45\textwidth]{EKF_PE_r0_20_xx_norm.pdf}}
}
\caption{Estimation results: $r_{x_0}=20\,\mathrm{km}$}\label{fig:r0_20}
\end{figure}
\begin{figure}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.45\textwidth]{EKF_PE_r0_50_e.pdf}}
	\hspace*{0.05\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.45\textwidth]{EKF_PE_r0_50_xx_norm.pdf}}
}
\caption{Estimation results: $r_{x_0}=50\,\mathrm{km}$}\label{fig:r0_50}
\end{figure}
\begin{figure}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.45\textwidth]{EKF_PE_r0_100_e.pdf}}
	\hspace*{0.05\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.45\textwidth]{EKF_PE_r0_100_xx_norm.pdf}}
}
\caption{Estimation results: $r_{x_0}=100\,\mathrm{km}$}\label{fig:r0_100}
\end{figure}
%\begin{figure}
%\centerline{
%	\subfigure[Estimation errors]{
%		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_100_e_J2.pdf}}
%	\hspace*{0.15\textwidth}
%		\subfigure[Magnitude of state (red:true, blue:estimate)]{
%		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_100_xx_norm_J2.pdf}}
%}
%\caption{Estimation results with $J_2$ perturbations: $r_{x_0}=100\,\mathrm{km}$}\label{fig:r0_100}
%\end{figure}
}

%\section{Numerical Results}
%
%We choose numerous cases of relative motion between chief and deputy satellite orbits.
%With each case, we evaluate the observability measure and process angles-only measurements with the EKF, using both the two-body equations of motion and higher fidelity propagation with $J_2$ effects as the truth models.
%%In particular, the $J_2$ gravitational harmonics will be included in the satellite propagation, so the EKF must handle simulated process uncertainty.
%To begin each simulation, an initial orbit determination (IOD) estimate is determined by various angles-only observations.
%These IOD estimates serve as initial estimates for the EKF.
%Then, we evaluate the performance of the EKF for these scenarios, specifically focusing on the observability measure and issues arising from initial conditions.
%
%Two IOD methods are developed in~\cite{NewLovPra14,PraLovNew14}, referred to as \emph{IOD1} and \emph{IOD2}, and employed to provide initial conditions for $13$ cases, outlined in Table \ref{tab:IODerr}.
%%Two IOD methods are referred to as \emph{IOD1} and \emph{IOD2}, and employed to provide initial conditions for $13$ cases, outlined in Table \ref{tab:IODerr}.
%These cases yield relative orbit trajectories illustrated in Figures \ref{fig:0CasesTraj}, \ref{fig:1CasesTraj}, \ref{fig:2CasesTraj}, \ref{fig:3CasesTraj}, and \ref{fig:4CasesTraj}.
%The most important distinction between the various cases is that in Case $\mathbf{1}$ the two-body solution and the linearized dynamics of the Hill-Clohessy-Wiltshire (HCW) solution are very close and the trajectory remains largely planar with a circular shape.
%
%\begin{figure}[h]
%\centerline{
%	\subfigure[Case 0A]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_0A_traj.pdf}}
%	\hfill
%	\subfigure[Case 0B]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_0B_traj.pdf}}
%	\hfill
%	\subfigure[Case 0C]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_0C_traj.pdf}}
%}
%\caption{The trajectories of the Cases 0A, 0B, and 0C (solid red: HCW, dashed blue: two-body solution) show that the linear and nonlinear solutions are close.
%The resulting trajectories are not cyclic as the trajectories change in the $y$-direction with time.}
%\label{fig:0CasesTraj}
%\end{figure}
%
%\begin{figure}[h]
%\label{fig:CaseTraj}
%\centerline{
%	\subfigure[Case 1]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_1_traj.pdf}}
%}
%\caption{The trajectory of the Case 1 (solid red: HCW, dashed blue: two-body solution) shows that the linear and nonlinear solutions are very close, where both solutions are largely planar and cyclic.}
%\label{fig:1CasesTraj}
%\end{figure}
%
%
%\begin{figure}[h]
%\centerline{
%	\subfigure[Case 2A]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_2A_traj.pdf}}
%	\hfill
%	\subfigure[Case 2B]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_2B_traj.pdf}}
%	\hfill
%	\subfigure[Case 2C]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_2C_traj.pdf}}
%}
%\caption{The trajectories of the Cases 2A, 2B, and 2C (solid red: HCW, dashed blue: two-body solution) show differences between the linear and nonlinear solutions that vary about the $y$-axis, increasing the observabilities.}
%\label{fig:2CasesTraj}
%\end{figure}
%
%\begin{figure}[h]
%\centerline{
%	\subfigure[Case 3A]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_3A_traj.pdf}}
%	\hfill
%	\subfigure[Case 3B]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_3B_traj.pdf}}
%	\hfill
%	\subfigure[Case 3C]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_3C_traj.pdf}}
%}
%\caption{The trajectories of the Cases 3A, 3B, and 3C (solid red: HCW, dashed blue: two-body solution) exhibit large observabilities due to nonlinear solutions that move far away from the linearized solutions.}
%\label{fig:3CasesTraj}
%\end{figure}
%
%\begin{figure}[h]
%\centerline{
%	\subfigure[Case 4A]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_4A_traj.pdf}}
%	\hfill
%	\subfigure[Case 4B]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_4B_traj.pdf}}
%	\hfill
%	\subfigure[Case 4C]{
%		\includegraphics[width=0.3\textwidth]{OM_IOD_4C_traj.pdf}}
%}
%\caption{The trajectories of the Cases 4A, 4B, and 4C (solid red: HCW, dashed blue: two-body solution) experience the largest observabilities due to nonlinear solutions that move farther away from the linearized solutions than any of the trajectories from other cases considered in this paper.}
%\label{fig:4CasesTraj}
%\end{figure}
%
%
%
%The initial error functions between the true orbit and and IOD predictions are defined as
%\begin{align}
%e_{mag,0}=\abs\left(1-{\frac{\norm{\r^+_0}}{\norm{\r_0}}}\right), \quad e_{dir,0}=\cos^{-1}\left(\frac{\r_0}{\norm{\r_0}}\cdot \frac{\r^+_0}{\norm{\r^+_0}}\right).
%\end{align}
%
%\begin{center}
%\begin{threeparttable}[h]
%\caption{Case Observabilities and Initial Errors}
%\begin{tabularx}{0.92\textwidth}
%{
%>{$}c<{$}
%*{1}{>{$}c<{$}} |
%*{2}{>{$}c<{$}} |
%*{2}{>{$}c<{$}}
%}
%\toprule
%\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{\multirow{1}{*}{IOD1}} & \multicolumn{2}{c}{\multirow{1}{*}{IOD2}} \\
%& &  e_{mag,0} & e_{dir,0} & e_{mag,0} & e_{dir,0} \\\midrule
%0A & 10^{11.1486} & 0.0003 & 1.4901\times10^{-8} & 0.0012  & 2.1073\times10^{-8} \\
%0B & 10^{10.9387} & 0.0233 & 0 & 0.0088 & 0 \\
%0C & 10^{10.9851} & 0.0152 & 2.1073\times10^{-8} & 0.0139 & 1.4901\times10^{-8}  \\
%\midrule
%\mathbf{1} & 10^{16.6871} & 0.0001 & 0 & 0 & 0  \\
%\midrule
%2A & 10^{10.6492} & 0.0550 & 1.8150\times10^{-4} & 0.0238 & 1.8150\times10^{-4}  \\
%2B & 10^{10.4163} & 0.9571 & 0.0734                     & 0.9571 & 0.0734  \\
%2C & 10^{10.2155} & 0.0219 & 0.0044 & 0.0065 & 0.0044 \\
%\midrule
%3A & 10^{8.3490} & 0.0191 & 0.0031 & 0.1830 & 0.0031  \\
%3B & 10^{8.5372} & 0.0612 & 0.0533 & 0.2315 & 0.0533  \\
%3C & 10^{8.6586} & 6.8962 & 0.7348 & 5.4921 & 0.7348  \\
%\midrule
%4A & 10^{7.9062} & 8.7347 & 0.4587 & 0.1575 & 0.4587  \\
%4B & 10^{7.9903} & 0.9276 & 0.1819 & 0.0621 & 0.1819  \\
%4C & 10^{8.0645} & 0.1161 & 0.0554 & 0.0354 & 0.0554  \\
%\bottomrule
%\end{tabularx}
%{\small
%\begin{tablenotes}
%    \item Note: $0$ indicates numbers small enough to be below Matlab numerical accuracy.
%  \end{tablenotes}}
%\label{tab:IODerr}
%\end{threeparttable}
%\end{center}
%
%
%The Kalman filter is used in each case. The initial state covariance matrix is chosen as
%\begin{align*}
%P_0^+=\diag[50^2,\ 50^2,\ 50^2,\ (50n)^2,\ (50n)^2,\ (50n)^2],
%\end{align*}
%and the process and measurement covariance matrices are chosen as
%\begin{align*}
%Q_k=\diag[10^{-7},\ 10^{-7},\ 10^{-7},\ 10^{-9},\ 10^{-9},\ 10^{-9}], \quad R_k=\tan^2{2^\circ}I_{3\times3}\ \forall\ k.
%\end{align*}
%%\begin{align*}
%%Q_k=10^q\diag[10^{-8},\ 10^{-8},\ 10^{-8},\ 10^{-10},\ 10^{-10},\ 10^{-10}], \quad R_k=\tan^2{2^\circ}I_{3\times3}\ \forall\ k
%%\end{align*}
%%where $q\in[-1,\ 0,\ 1,\ 2]$ is varied in the simulations, thereby changing the impact of measurement updates inside the Kalman filter.
%
%Each case is simulated for every IOD estimate. The running time is $59540$ seconds (slightly more than $\frac23$ of a day) with time steps of $10$ seconds between measurements. The following RMS errors are collected from each trial:
%\begin{align}
%\bar e_{mag}&=\left(\frac1N\sum_{k=1}^N\left\{1-{\frac{\norm{\r^+_k}}{\norm{\r_k}}}\right\}^2\right)^{\frac12},
%\\
%\bar e_{dir}&=\left(\frac1N\sum_{k=1}^N
%\left\{\cos^{-1}\left(\frac{\r_k}{\norm{\r_k}}\cdot \frac{\r^+_k}{\norm{\r^+_k}}\right)\right\}^{2}
%\right)^{\frac12},
%\end{align}
%for $N$ time steps.
%These metrics are tabulated in Tables \ref{tab:ResultsTBP} and \ref{tab:ResultsJ2}.
%Over the simulated time period, figures were generated for these analyses that  serve to compare the true and estimated state magnitudes and the maximum eigenvalues of the state covariance matrices as a measure of uncertainty for the state estimates.
%Because of paper constraints, a sample of these plots are shown that highlight various aspects of the observability and filter performance.
%
%
%
%%The figure below illustrates preliminary results for angles-only relative orbit estimation with $J_2$ perturbation in the measurements.
%%Note that the EKF performs well in this particular case, as evidenced by the close agreement between the true and estimated trajectory (Figure 1a) and the decrease in the maximum eigenvalue of the covariance matrix over time (Figure 1b).
%%For the various cases to be chosen, we will evaluate the performance of the EKF (using metrics such as those illustrated in Figure 1) and seek a consistent correlation between the filter performance and the new observability measure proposed above.
%
%
%
%%\begin{center}
%%\begin{threeparttable}[h]
%%\caption{Error variables for EKF with varying $Q_k$ Considering Only Two-Body Problem Forces}
%%\begin{tabularx}{\textwidth}
%%{
%%>{$}c<{$}
%%*{1}{>{$}c<{$}} |
%%*{4}{>{$}c<{$}} |
%%*{4}{>{$}c<{$}}
%%}
%%% & \bar e_{mag} & \bar e_{dir}
%%\toprule
%%\multirow{2}{*}{Case} & \multirow{2}{*}{IOD} & \multicolumn{4}{c}{$\bar e_{mag}$} & \multicolumn{4}{c}{$\bar e_{dir}$} \\
%%& & q=-1 & q=0 & q=1 & q=2 & q=-1 & q=0 & q=1 & q=2 \\\midrule
%%0A & IOD 1 & 0.0895 &  0.0611  &  0.3766  &  0.2861  &  0.0430  &  0.0135  &  0.0144  &  0.0183 \\
%%0A & IOD 2 & 0.1793 &  0.2811  &  0.5145  &  0.7716  &  0.0203  &  0.0313  &  0.0691  &  0.1046 \\
%%0B & IOD 1 & 0.3241 &  0.0531  &  0.5186  &  0.0620  &  0.1695  &  0.0143  &  0.0810  &  0.0172 \\
%%0B & IOD 2 & 0.2330 &  0.2115  &  0.3092  &  0.2733  &  0.1007  &  0.0879  &  0.0804  &  0.1048 \\
%%0C & IOD 1 & 0.0973 &   0.0748 &   0.2723 &   0.0607 &   0.0292 &   0.0068  &  0.1362 &   0.0443 \\
%%0C & IOD 2 & 0.1473 &   0.1017 &   0.1707 &   0.3545  &  0.0702  &  0.0517  &  0.0614  &  0.1224 \\
%%\midrule
%%\mathbf{1} & IOD 1 & \hilight{8.922} &   \hilight{2.964}   & \hilight{4.078}  &  \hilight{3.571} &   0.0071  &  0.0089  &  0.0109  &  0.0418 \\
%%\mathbf{1} & IOD 2 & \hilight{13.635} &  \hilight{14.121} &  \hilight{42.486} &  \hilight{42.573} &   0.0113 &   0.0107  &  0.0445  &  0.0115 \\
%%\midrule
%%2A & IOD 1 & 0.2045 &   0.0629 &   0.1592  &  0.1329  &  0.0041 &   0.0041  &  0.0043  &  0.0039 \\
%%2A & IOD 2 & 0.1013 &   0.1797  &  0.1348  &  0.2247  &  0.0050 &   0.0048 &   0.0065 &   0.0061 \\
%%2B & IOD 1 & \hilight{0.9886} &   \hilight{0.9895}   & \hilight{0.9843}  &  \hilight{0.9851}  &  0.0114 &   0.0116  &  0.0114  &  0.0116 \\
%%2B & IOD 2 & \hilight{0.9557} &   \hilight{0.9685} &   \hilight{0.9278}  &  \hilight{0.9431} &   0.0105  &  0.0108 &   0.0113 &   0.0153 \\
%%2C & IOD 1 & 0.0219 &   0.0184  &  0.0258  &  0.0238 &   0.0026 &   0.0031  &  0.0044 &   0.0038 \\
%%2C & IOD 2 & 0.0303 &   0.0287  &  0.0360  &  0.0470  &  0.0044  &  0.0054  &  0.0059 &   0.0062 \\
%%\midrule
%%3A & IOD 1 &  0.2377 &   0.1684  &  0.0486 &   0.0695 &   0.0094 &   0.0070 &   0.0040 &  0.0039 \\
%%3A & IOD 2 &  0.0854  &  0.0874 &   0.0603  &  0.0544  &  0.0035 &   0.0033  &  0.0053  &  0.0042 \\
%%3B & IOD 1 &  0.0724 &   0.0260 &   0.0680  &  0.0289  &  0.0035 &   0.0029  &  0.0033  &  0.0027 \\
%%3B & IOD 2 &  0.0385  &  0.0432  &  0.1459  &  0.0284  &  0.0038  &  0.0037  &  0.0070  &  0.0047 \\
%%3C & IOD 1 &  \hilight{2.2433}  &  \hilight{1.5231} &   \hilight{2.0611} &   \hilight{1.5110}  &  0.1803  &  0.1172  &  0.1595  &  0.1103 \\
%%3C & IOD 2 &  \hilight{4.0635} &   \hilight{1.7717}  & \hilight{10.567}  &  \hilight{4.0277} &   0.4475 &   0.1506 &   0.9011 &   0.4028 \\
%%\midrule
%%4A & IOD 1 &  \hilight{6.0259} &   \hilight{0.0464}  &  \hilight{6.0706}  &  \hilight{0.0325}  &  1.3353  &  0.0087   & 1.3182  &  0.0065 \\
%%4A & IOD 2 &  6.3646 &   0.0173  & 11.8882 &   0.0256 &   1.3423  &  0.0064  &  1.5319 &   0.0066 \\
%%4B & IOD 1 &  \hilight{0.2922}  &  \hilight{0.0258}  &  \hilight{0.2723}  &  \hilight{0.0182}   & 0.0929 &   0.0032  &  0.0789  &  0.0036 \\
%%4B & IOD 2 &  0.1815  &  0.0217  &  0.3155 &   0.0771  &  0.0268   & 0.0034  &  0.1061 &   0.0044 \\
%%4C & IOD 1 &  0.0247  &  \hilight{0.0421}  &  2.3699  &  0.0182  &  0.0032  &  0.0032  &  0.8486 &   0.0028 \\
%%4C & IOD 2 &  0.0200 &   0.0128  &  0.0278  &  0.0299 &   0.0040  &  0.0029  &  0.0034  &  0.0033 \\
%%\bottomrule
%%\end{tabularx}
%%{\small
%%\begin{tablenotes}
%%    \item \hilight{Filter diverges}
%%  \end{tablenotes}}
%%\label{tab:ResultsTBP}
%%\end{threeparttable}
%%\end{center}
%%
%%\begin{center}
%%\begin{threeparttable}[h]
%%\caption{Error variables for EKF with varying $Q_k$ with $J_2$ Perturbations Included}
%%\begin{tabularx}{\textwidth}
%%{
%%>{$}c<{$}
%%*{1}{>{$}c<{$}} |
%%*{4}{>{$}c<{$}} |
%%*{4}{>{$}c<{$}}
%%}
%%\toprule
%%\multirow{2}{*}{Case} & \multirow{2}{*}{IOD} & \multicolumn{4}{c}{$\bar e_{mag}$} & \multicolumn{4}{c}{$\bar e_{dir}$} \\
%%& & q=-1 & q=0 & q=1 & q=2 & q=-1 & q=0 & q=1 & q=2 \\\midrule
%%0A & IOD 1 & 0.0553  &  0.1105  &  0.1918  &  0.2298  &  0.0059  &  0.0062  &  0.0069  &  0.0069 \\
%%0A & IOD 2 & 0.2177  &  0.3227  &  0.3140  &  0.5013  &  0.0082  &  0.0074  &  0.0129  &  0.0085 \\
%%0B & IOD 1 & 0.0473  &  0.1246  &  0.7262  &  0.0655  &  0.0077  &  0.0088  &  0.0505  &  0.0066\\
%%0B & IOD 2 & 0.8872  &  0.4186  &  0.1571  &  0.1561  &  0.0390  &  0.0168  &  0.0086  &  0.0293 \\
%%0C & IOD 1 & 0.0652  &  0.0809  &  0.0900  &  0.1113  &  0.0082  &  0.0079  &  0.0061  &  0.0106 \\
%%0C & IOD 2 & 0.0367  &  0.0204  &  0.1002  &  0.7253  &  0.0145  &  0.0084  &  0.0154  &  0.0579 \\
%%\midrule
%%\mathbf{1} & IOD 1 & \hilight{9.1098}  &  \hilight{2.900}  &  \hilight{4.059}  &  \hilight{3.567}  &  0.0071  &  0.0090  &  0.0109  &   0.0418 \\
%%\mathbf{1} & IOD 2 & \hilight{13.594} & \hilight{14.033} & \hilight{42.381} & \hilight{42.979}  & 0.0113   & 0.0107   & 0.0445   &   0.0115 \\
%%\midrule
%%2A & IOD 1 & 0.2301  &  0.0423  &  0.1978  &  0.1740  &  0.0048  &  0.0049  &  0.0047  &  0.0042 \\
%%2A & IOD 2 & 0.1058  &  0.2164  &  0.1638  &  0.2644  &  0.0051  &  0.0049  &  0.0065  &  0.0061 \\
%%2B & IOD 1 & \hilight{0.9895}  &  \hilight{0.9902}  &  \hilight{0.9853}  &  \hilight{0.9859}  &  0.0115  &  0.0117  &  0.0115  &  0.0117 \\
%%2B & IOD 2 & \hilight{0.9585}  &  \hilight{0.9701}  &  \hilight{0.9307}  &  \hilight{0.9455}  &  0.0106  &  0.0109  &  0.0114  &  0.0153 \\
%%2C & IOD 1 & 0.0481  &  0.0474  &  0.0563  &  0.0415  &  0.0041  &  0.0042  &  0.0047  &  0.0042 \\
%%2C & IOD 2 & 0.0535  &  0.0397  &  0.0536  &  0.0705  &  0.0044  &  0.0056  &  0.0059  &  0.0062 \\
%%\midrule
%%3A & IOD 1 &  0.2462  &  0.1827  &  0.0516  &  0.0699 &   0.0097 &   0.0077  &  0.0041 &   0.0041 \\
%%3A & IOD 2 &  0.1038  &  0.0986  &  0.0643  &  0.0692 &   0.0036  &  0.0034  &  0.0053  &  0.0041 \\
%%3B & IOD 1 &  0.0770  &  0.0310  &  0.0739  &  0.0458 &   0.0038 &   0.0034  &  0.0035  &  0.0031 \\
%%3B & IOD 2 &  0.0461  &  0.0585   & 0.1485 &   0.0391  &  0.0039  &  0.0040  &  0.0071  &  0.0047 \\
%%3C & IOD 1 &  \hilight{2.239}  &  \hilight{1.522}  &  \hilight{2.063}  &  \hilight{1.514}  &  0.1817  &  0.1179  &  0.1614  &  0.1116 \\
%%3C & IOD 2 &  \hilight{4.188}  &  \hilight{1.821}  & \hilight{10.136}  &  \hilight{4.510} &   0.4873 &   0.1592  &  0.9195 &   0.5532 \\
%%\midrule
%%4A & IOD 1 &  \hilight{6.013}  &  \hilight{0.0454}  &  \hilight{6.359}  &  \hilight{0.031}  &  1.3131  &  0.0086  &  1.4235 &   0.0066 \\
%%4A & IOD 2 &  6.3272  &  0.0197  & 12.0062 &  0.0248  &  1.3116  &  0.0065  &  1.5310  &  0.0066 \\
%%4B & IOD 1 &  \hilight{0.2936}  &  \hilight{0.0272}  &  \hilight{0.2773}  &  \hilight{0.0217}  &  0.0935  &  0.0032  &  0.0814  &  0.0038 \\
%%4B & IOD 2 &  0.1853  &  0.0210  &  0.5244  &  0.0785  &  0.0293  &  0.0036  &  0.2305  &  0.0045 \\
%%4C & IOD 1 &  0.0260  &  \hilight{0.0410}  &  2.0105  &  0.0164  &  0.0034  &  0.0031  &  0.9144  &  0.0030 \\
%%4C & IOD 2 &  0.0210  &  0.0145  &  0.0303  &  0.0349  &  0.0042  &  0.0030  &  0.0036  &  0.0034 \\
%%\bottomrule
%%\end{tabularx}
%%{\small
%%\begin{tablenotes}
%%    \item \hilight{Filter diverges}
%%  \end{tablenotes}}
%%\label{tab:ResultsJ2}
%%\end{threeparttable}
%%\end{center}
%
%\begin{center}
%\begin{threeparttable}[h]
%\caption{Error variables for EKF Considering Only Two-Body Problem Forces}
%\begin{tabularx}{.74\textwidth}
%{
%>{$}c<{$}
%*{1}{>{$}c<{$}} |
%*{2}{>{$}c<{$}} |
%*{2}{>{$}c<{$}}
%}
%\toprule
%\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{IOD1} & \multicolumn{2}{c}{IOD2} \\
%& & \bar e_{mag} & \bar e_{dir} & \bar e_{mag} & \bar e_{dir} \\\midrule
%0A & 10^{11.1486} &  0.3766  &  0.0144  &  0.5145  &  0.0691 \\
%0B & 10^{10.9387} &  0.5186  &  0.0810  &  0.3092  &  0.0804\\
%0C & 10^{10.9851} &  0.2723  &  0.1362 &  0.1707 &  0.0614\\
%\midrule
%\mathbf{1} & 10^{16.6871} & \cellcolor{green} 4.078  & \cellcolor{green}  0.0109 & \cellcolor{green} 42.486 & \cellcolor{green}  0.0445 \\
%\midrule
%2A & 10^{10.6492} &  0.1592  &   0.0043  &  0.1348  &   0.0065   \\
%2B & 10^{10.4163} & \cellcolor{green} 0.9843   & \cellcolor{green}  0.0114  & \cellcolor{green}  0.9278 & \cellcolor{green}   0.0113 \\
%2C & 10^{10.2155} & 0.0258 &  0.0044 & 0.0360 &  0.0059 \\
%\midrule
%3A & 10^{8.3490} &  0.0486 & 0.0040   &  0.0603  &  0.0053 \\
%3B & 10^{8.5372} &  0.0680  &  0.0033 &  0.1459 &  0.0070 \\
%3C & 10^{8.6586} & \cellcolor{green}  2.0611 & \cellcolor{green}  0.1595   & \cellcolor{green}  10.567  & \cellcolor{green}   0.9011 \\
%\midrule
%4A & 10^{7.9062} & \cellcolor{green}  6.0706  & \cellcolor{green} 1.3182  &  11.8882   &  1.5319  \\
%4B & 10^{7.9903} & \cellcolor{green}  0.2723  & \cellcolor{green}  0.0789   &  0.3155  &  0.1061 \\
%4C & 10^{8.0645} &  2.3699 &  0.8486   &  0.0278 &  0.0034 \\
%\bottomrule
%\end{tabularx}
%{\small
%\begin{tablenotes}
%    \item \hilight{Filter diverges}
%  \end{tablenotes}}
%\label{tab:ResultsTBP}
%\end{threeparttable}
%\end{center}
%
%\begin{center}
%\begin{threeparttable}[h]
%\caption{Error variables for EKF Considering $J_2$ Perturbations}
%\begin{tabularx}{.74\textwidth}
%{
%>{$}c<{$}
%*{1}{>{$}c<{$}} |
%*{2}{>{$}c<{$}} |
%*{2}{>{$}c<{$}}
%}
%\toprule
%\multirow{2}{*}{Case} & \multirow{2}{*}{$\mathrm{cond}\braces{\mathcal{W}}$} & \multicolumn{2}{c}{IOD1} & \multicolumn{2}{c}{IOD2} \\
%& & \bar e_{mag} & \bar e_{dir} & \bar e_{mag} & \bar e_{dir} \\\midrule
%0A & 10^{11.1486} &  0.1918  &  0.0069  &  0.3140  &  0.0129 \\
%0B & 10^{10.9387} &  0.7262  &  0.0505  &  0.1571  &  0.0086\\
%0C & 10^{10.9851} &  0.0900  &  0.0061 &  0.1002  &  0.0154\\
%\midrule
%\mathbf{1} & 10^{16.6871} & \cellcolor{green} 4.059  & \cellcolor{green}  0.0109 & \cellcolor{green} 42.381 & \cellcolor{green}  0.0445 \\
%\midrule
%2A & 10^{10.6492} &  0.1978  &  0.0047  &  0.1638  &   0.0065   \\
%2B & 10^{10.4163} & \cellcolor{green}  0.9853  & \cellcolor{green}  0.9307 & \cellcolor{green}  0.9278 & \cellcolor{green}   0.0114 \\
%2C & 10^{10.2155} & 0.0563 &  0.0047 & 0.0536 &  0.0059 \\
%\midrule
%3A & 10^{8.3490} &  0.0516 & 0.0041   &  0.0643  &  0.0053 \\
%3B & 10^{8.5372} &  0.0739  &  0.0035 &  0.1485 &  0.0071 \\
%3C & 10^{8.6586} & \cellcolor{green}  2.063 & \cellcolor{green} 0.1614    & \cellcolor{green}  10.136  & \cellcolor{green}   0.9195 \\
%\midrule
%4A & 10^{7.9062} & \cellcolor{green}  6.359  & \cellcolor{green} 1.4235  &  12.0062   &  1.5310  \\
%4B & 10^{7.9903} & \cellcolor{green}  0.2773  & \cellcolor{green}  0.0814   &  0.5244  &  0.2305 \\
%4C & 10^{8.0645} &  2.0105 &  0.9144   &  0.0303 &  0.0036 \\
%\bottomrule
%\end{tabularx}
%{\small
%\begin{tablenotes}
%    \item \hilight{Filter diverges}
%  \end{tablenotes}}
%\label{tab:ResultsJ2}
%\end{threeparttable}
%\end{center}
%
%These data from the simulations show several important aspects and challenges with relative orbit determination involving angles-only measurements.
%First, an important observation from the data is that the IOD guess is critical to the success of the Kalman filter.
%Cases 2B, 3C, 4A, and 4B experience filter divergence when the IOD estimate is largely inaccurate, shown in Tables \ref{tab:IODerr}, \ref{tab:ResultsTBP}, and \ref{tab:ResultsJ2}, even when cases with similar observabilities do not experience filter divergence.
%The Kalman filter, which is based on linear dynamic systems, is modeled as a Markov chain and hence is memoryless.
%Therefore when nothing more than an inaccurate IOD prediction is poorly assumed to contain accurate information about the system, the recursive formulation of the Kalman filter is unable to converge.
%In these diverging cases, a greater level of observability or an algorithm that relies on a larger history of measurements may improve convergence.
%
%
%In contrast, every variation of Case $\mathbf{1}$ has a closer IOD than the IODs of other cases, but experiences the most severe divergence.
%Case $\mathbf{1}$ has a greater condition number of $\mathcal W$, which corresponds with low observability according to the proposed observability measure.
%A series of angles-only measurements may correspond with a large range of possible relative orbits, even with a low variation in those measurements.
%Therefore, small levels of noise in the sensor or uncertainties in the orbit prevent the Kalman filter from converging to an accurate estimate, even with an accurate IOD.
%The state magnitudes of both variations of Case $\mathbf{1}$ are illustrated against various converging cases in Figure \ref{fig:ExampesEachCase} to show how important this observability is to the filter performance.
%
%\begin{figure}[h]
%\centerline{
%	\subfigure[Case 0C, IOD1]{
%		\includegraphics[width=0.35\textwidth]{Case0Cq1IOD1TBPnormX.pdf}}
%	\hfill
%	\subfigure[Case 0C, IOD2]{
%	\includegraphics[width=0.35\textwidth]{Case0Cq1IOD2TBPnormX.pdf}}
%}
%\centerline{
%	\subfigure[Case $\mathbf{1}$, IOD1]{
%		\includegraphics[width=0.35\textwidth]{Case1q1IOD1TBPnormX.pdf}}
%	\hfill
%	\subfigure[Case $\mathbf{1}$, IOD2]{
%	\includegraphics[width=0.35\textwidth]{Case1q1IOD2TBPnormX.pdf}}
%}
%\centerline{
%	\subfigure[Case 2A, IOD1]{
%		\includegraphics[width=0.35\textwidth]{Case2Aq1IOD1TBPnormX.pdf}}
%	\hfill
%	\subfigure[Case 2A, IOD2]{
%	\includegraphics[width=0.35\textwidth]{Case2Aq1IOD2TBPnormX.pdf}}
%}
%\centerline{
%	\subfigure[Case 3A, IOD1]{
%		\includegraphics[width=0.35\textwidth]{Case3Aq1IOD1TBPnormX.pdf}}
%	\hfill
%	\subfigure[Case 3A, IOD2]{
%	\includegraphics[width=0.35\textwidth]{Case3Aq1IOD2TBPnormX.pdf}}
%}
%\centerline{
%	\subfigure[Case 4C, IOD1]{
%		\includegraphics[width=0.35\textwidth]{Case4Cq1IOD1TBPnormX.pdf}}
%	\hfill
%	\subfigure[Case 4C, IOD2]{
%	\includegraphics[width=0.35\textwidth]{Case4Cq1IOD2TBPnormX.pdf}}
%}
%\caption{The magnitudes of the relative states (solid red: true, blue dashed: estimated) are shown for various two-body propagated cases with close IOD estimates where the filter only diverges when the observability is low.}\label{fig:ExampesEachCase}
%\end{figure}
%
%
%Another observation from the data is that the $J_2$ perturbation, which is well-known to have a noticeable impact perturbing orbits from the two-body model for satellites around the earth, yields an insignificant change when compared to the data when only two-body forces are considered in the process model.
%In fact, every case when the filter diverged occurred with both the two-body problem and $J_2$ simulations.
%Examples of these results are tabulated in Tables \ref{tab:ResultsTBP} and \ref{tab:ResultsJ2} and illustrated in Figure \ref{fig:TBPvsJ2}.
%
%
%
%\begin{figure}[h]
%\centerline{
%	\subfigure[State magnitude with the two-body solution]{
%		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1TBPnormX.pdf}}
%	\hfill
%	\subfigure[State magnitude with the $J_2$ solution]{
%		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1HighFidelitynormX.pdf}}
%}
%\centerline{
%	\subfigure[Uncertainty measure with the two-body solution]{
%		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1TBPeigP.pdf}}
%	\hfill
%	\subfigure[Uncertainty measure with the $J_2$ solution]{
%		\includegraphics[width=0.48\textwidth]{Case3Bq1IOD1eigP.pdf}}
%}
%\caption{The magnitudes and maximum eigenvalues of the state covariance matrices of the relative states (solid red: true, blue dashed: estimated) in Case 3B with IOD1 serve as an example where the $J_2$ propagated simulations give very close results to those of the two-body solution.
%}\label{fig:TBPvsJ2}
%\end{figure}

\section{Conclusions}

It is well known that the linearized relative orbital dynamics are not observable with angles-only
measurements. This paper studies the observability for nonlinear relative orbital dynamics, and it shows that the relative orbit is indeed observable under certain geometric conditions. Also, a quantitative observability measure that represents the degree of observability is proposed. They are illustrated by several numerical examples. Future works include constructing observability criteria based on higher-order Lie derivatives of the measurement, and developing nonlinear filtering techniques that can capture the weak coupling between the dynamics of relative distance and the line-of-sight measurements effectively.

%formulates a measure of nonlinear observability, and investigates a less conservative observability criteria by using higher-order Lie derivatives. Numerous relative orbit determination cases are illustrated numerically by an extended Kalman filter, which shows that the degree of observability by the proposed measure and an accurate IOD prediction are useful for predicting filter convergence.

%\section{Section title}
%\label{sec:1}
%Text with citations \cite{RefB} and \cite{RefJ}.
%\subsection{Subsection title}
%\label{sec:2}
%as required. Don't forget to give each section
%and subsection a unique label (see Sect.~\ref{sec:1}).
%\paragraph{Paragraph headings} Use paragraph headings as needed.
%\begin{equation}
%a^2+b^2=c^2
%\end{equation}

% For one-column wide figures use
%\begin{figure}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:1}       % Give a unique label
%\end{figure}
%
% For two-column wide figures use
%\begin{figure*}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics[width=0.75\textwidth]{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:2}       % Give a unique label
%\end{figure*}
%
%% For tables use
%\begin{table}
%% table caption is above the table
%\caption{Please write your table caption here}
%\label{tab:1}       % Give a unique label
%% For LaTeX tables use
%\begin{tabular}{lll}
%\hline\noalign{\smallskip}
%first & second & third  \\
%\noalign{\smallskip}\hline\noalign{\smallskip}
%number & number & number \\
%number & number & number \\
%\noalign{\smallskip}\hline
%\end{tabular}
%\end{table}

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
\bibliography{BibSources}   % name your BibTeX data base

%% Non-BibTeX users please use
%\begin{thebibliography}{}
%%
%% and use \bibitem to create references. Consult the Instructions
%% for authors for reference list style.
%%
%\bibitem{RefJ}
%% Format for Journal Reference
%Author, Article title, Journal, Volume, page numbers (year)
%% Format for books
%\bibitem{RefB}
%Author, Book title, page numbers. Publisher, place (year)
%% etc
%\end{thebibliography}

\end{document}
% end of file template.tex

### 2018-06-09 Journal/JAS OD EKF/JAS OD EKF Old/JAS_OD_EKF_ans1.tex
journal added

\documentclass[11pt]{article}
\usepackage[letterpaper,text={6.5in,8.6in},centering]{geometry}
\usepackage{amssymb,amsmath,times,url}
\usepackage{xr,color}
\usepackage{hyperref}

\externaldocument[]{JAS_OD_EKF_rev1}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand{\refprop}[1]{Proposition \ref{prop:#1}}

\newcommand{\RNum}[1]{\uppercase\expandafter{\romannumeral #1\relax}}
\newcommand{\RI}{\text{\RNum{1}}}
\newcommand{\RII}{\text{\RNum{2}}}
\newcommand{\RIII}{\text{\RNum{3}}}

\newenvironment{correction}{\begin{list}{}{\setlength{\leftmargin}{1cm}\setlength{\rightmargin}{1cm}}\vspace{\parsep}\item[]``}{''\end{list}}

\begin{document}

%\pagestyle{empty}

\section*{Response to the Reviewers' Comments for JASS-D-15-00027}

The authors would like to thank for the reviewers for their thoughtful comments, which are aimed toward improving the quality of the paper and the clarity of the results. In accordance with the comments and suggestions, the paper has been revised as follows.

\subsection*{Reviewer 1}

\setlength{\leftmargini}{0pt}
\begin{itemize}\setlength{\itemsep}{2\parsep}

\item {\itshape Reviewer \#1: This paper addresses the question of observability from angles-only measurements. It is known that in the linear approximation, the motion is not fully observable; the present paper addresses the question of observability in the nonlinear case. This topic has received some attention in the field in recent years. It is appropriate for the Journal of the Astronautical Sciences.

   The authors divide their paper into two parts. The first gives, analytically, sufficient conditions for nonlinear observability. To my knowledge, this is original, and while I did not confirm all the math in detail, there was nothing that jumped out at me as obviously wrong. The second part of the paper is a numerical study of the observability of relative orbits, and an extended Kalman filter to demonstrate this observability and to connect their proposed observability measure and the accuracy of the filtered solution.

   The paper makes a good start on the goal, but feels incomplete. More specifically, it reads like the starts to two different papers were merged into one. The analytical observability analysis is developed over ten pages but the reader is left to figure out the physical/geometric implication of this result. The authors emphasize that they have shown sufficient conditions for observability and that failing the sufficiency test doesn't show a given point isn't observable. Then, the authors jump right into the numerical study without any connection or apparent use of the analytical study. Why bother doing the first part if it doesn't give you what you really want? On the other hand, the numerical analysis seems very short. While the connection to physical and geometric properties of the relative orbit is better shown here, it could be made clearer. The conclusion section says "... a quantitative observability measure...is proposed. It is illustrated by several numerical
examples"; it should have specific numerical conclusions that can be drawn from the presented results, such as those given in section 5. The connection between the two parts should be better drawn out; I suggest the physical/geometric implications of each because readers will be able to better relate the general statements to specific relative motion scenarios.$ $}\\

As pointed out by the reviewer, there are two main contributions in this paper. The first is verifying nonlinear observability of relative orbital dynamics with angles-only measurements, and the second is proposing a computational method to obtain a measure of observability.

The first part is motivated by the fact that no literature has definitely addressed whether the nonlinear relative orbital dynamics are observable with angles-measurements only. In other words, the main objective is to \textit{provide reasonable assurance} that it is actually observable under certain conditions. Since this paper verifies that the nonlinear relative orbital dynamics is observable under certain conditions, the objectives of studying nonlinear observability and providing confidence are indeed achieved in this paper.

As noted by the reviewers, there are certain limitations in the provided results. The presented conditions for observability are based on the second order derivatives, and they only provide sufficient conditions that are not suitable to understand the type of relative orbits that prohibit observability. However, these limitations originate from the differential algebraic approach for nonlinear observability criteria: it requires higher-order derivatives of the measurements; it yields sufficient conditions but it does not provide necessary conditions for observability.

Therefore, new observability criteria for nonlinear dynamic systems should be developed from first principles, to study observability of relative orbits completely and to provide sufficient conditions for unobservability as suggested by the reviewer. Such theoretical study for nonlinear observability is certainly a next step in this work, but is out of the scope of this paper, which is focused on verifying nonlinear observability of relative orbital dynamics.

To clarify these, the second to last paragraph of the introduction has been revised as follows.

\begin{correction}
In short, the main contributions of this paper are constructing sufficient conditions for the observability of nonlinear relative orbits with angles-only measurements that have not been studied before, and proposing a new quantitative measure of observability. The main objective is to provide assurance that relative orbits are observable with angles-only measurements under certain conditions, if the nonlinearity of relative orbital dynamics are properly considered. More comprehensive study for observability of nonlinear relative orbits, such as obtaining necessary conditions for observability or sufficient conditions for unobservability is referred to as future study. 
\end{correction}

Furthermore, the limitations of the presented results are explicitly summarized in the last paragraph of Section 3 as follows.

\begin{correction}
However, \refeqn{cond1} and \refeqn{cond2} are sufficient conditions for observability, and the fact that any of these conditions is not satisfied does not necessarily imply that the considered point is not observable. In such cases, the third or higher order Lie derivatives should be checked to confirm observability. Furthermore, the presented differential geometric approaches to determine observability do not yield necessary conditions for observability. As such, these conditions cannot be used to determine the certain types of relative motions that are not observable with angles-only measurements. This paper is focused on verifying the observability of nonlinear relative orbits via angles-only measurements under certain conditions, and determining observability with higher-order Lie derivatives or necessary conditions for observability is referred to future investigation.
\end{correction}

The relevant future work is also summarized in the conclusions.

\begin{correction}
Future work objectives include constructing observability criteria based on higher-order Lie derivatives of the measurement, and developing necessary conditions for observability to classify the type of relative orbits that are not observable completely. We also plan to develop nonlinear filtering techniques that can capture the weak coupling between the dynamics of relative motion and the line-of-sight measurements effectively, as well as studying the effects of higher-order gravitational perturbations on the observability. 
\end{correction}

The observability measure proposed in Section 4 is motivated by the limitations of the differential geometric approach in Section 3. Since the observability rank condition is defined for a state vector at a specific time, it does not capture the observability with respect to the measurement trajectories over a certain time period. Also, it only provides binary information whether the system is completely observable or not. Therefore, the observability criteria presented in Section 3 are not suitable to determine whether a particular relative orbit is easier to observe compared with other orbits.

The quantitative observability measure in Section 4 addresses these issues. As the observability gramian given in (28) is defined with respect to the measurement trajectory over a certain time period, it actually captures the accumulated effects of the line-of-sight measurements on observability. As such, it can be used to anticipate the evolution of the performance of nonlinear estimators over time. It also yields a certain quantitative measure of observability.

In contrast to the observability criteria in Section 3, the observability measure in Section 4 is obtained numerically.  While it is likely very difficult to extract physical implications that can be applied to various classes of relative orbits from the presented observability measure, the measure can be used to determine whether a certain relative orbit is easier or more difficult to estimate compared with other relative orbits.

Instead, this paper presents the correlation between the performance of extended Kalman filters and the proposed observability measure numerically in Section 5. In particular, the planar relative orbits presented in Section 5.3 correspond to a case where the observability criteria in Section 3 are violated, but the observability measure in Section 4 can be used to study the degree of observability.

In short, the numerical observability measure proposed in Section 4 is complementary to the analytic observability criteria in Section 3. The first two paragraphs of Section 4 have been revised as follows to motivate the observability measure more explicitly.

\begin{correction}
The observability criteria derived in the previous section guarantee that the nonlinear relative orbital dynamics are indeed observable under certain geometric conditions. However, they are limited in the sense that they only provide binary information about observability. Therefore, they cannot be used to determine whether a certain type of orbit is more easily estimated compared with other orbits based on angles-only measurements. Furthermore, as they are defined in terms of the state vector at a specific time, they do not capture the effects of the measurement trajectory over a certain time period on the observability, which often predicts the performance of estimation.

Motivated by the above issues, we propose a quantitative measure that represents the degree of observability for relative orbits in this section. More explicitly, this measure evaluates the observability in a quantitative manner as it determines how effectively the relative orbit can be estimated by line-of-sight measurements. 
\end{correction}

Also, the following paragraph is added at the end of Section 4 to summarize the limitation and the usage of the proposed observability measure explicitly.

\begin{correction}
In contrast to the observability criteria presented in Section 3, the proposed observability measure is obtained numerically. As such, it is not useful to identify the particular class of the relative orbits that yield stronger, or weaker observability. However, it can be used to analyze the degree of observability in a complementary manner when the sufficient conditions of the observability criteria are violated, as illustrated by the following numerical examples.  
\end{correction}

In Section 5.3, it is explicitly stated that the observability criteria proposed in Section 3 cannot be used to investigate the observability of the selected planar orbits, but the observability measure can be used instead.

\begin{correction}
As discussed in Section 3, these relative orbits violate the observability criteria summarized in Proposition 1 since they are planar orbits. Therefore, Proposition 1 cannot be applied to investigate the observability of these orbits. Instead, the observability is analyzed by the observability measure proposed in Section 4 as follows. 
\end{correction}

\item   {\itshape The concept of degree of observability should be explained better; reference is made to "weakly observable" and "how strong the observability is" without ever explaining what that means. A brief review of observability with an explanation of the measure of strength at the beginning would be helpful, given how central this concept is to the paper.}

The concept of the degree of controllability and observability is relatively popular in linear system theory. As discussed above, the first two paragraphs of Section 4 are revised to convey the concept of degree of observability and motivate the observability measure. The subsequent paragraph is also revised as follows to provide additional references for the observability measure.

%paragraph of Section 4 is revised to motivate the observability measure. The subsequent paragraphs are also revised as follows to give the definition of the observability clearly, and to provide additional references for the observability measure.

\begin{correction}
%Motivated by these, we propose a quantitative measure that represents the degree of observability for relative orbits in this section. More explicitly, this evaluates the quality of observability in a quantitative manner as it determines whether the relative orbit can be estimated effectively by line-of-sight measurements. 
%
The problem of assigning physically meaningful measures of the quality of observability has been introduced in~[1,9,10] for linear systems, and in particular, in [10] the determinant, trace, and maximal eigenvalue of the inverse characteristic observability matrix are maximized with respect to certain parameters of a dynamic system. Several approaches have been proposed for an observability measure of nonlinear systems....
\end{correction}

\item    {\itshape The writing is generally quite clear, and I noticed few typographical, spelling, grammar or phrasing problems, in comparison to most papers I review for technical journals. Nevertheless, there are some:

   1) Figures and particularly the legend and scale markings are much too small.\\
   2) p. 2 l. 44 "motivated to understand" awkward.\\
   3) p. 3 l. 10 "...with angles-only measurements that have not been studied before" rephrase as it reads like the last clause modifies the previous one (when it is intended to modify "sufficient conditions for the observability" earlier in the sentence, I presume).\\
   4) p. 4 l. 36 and elsewhere, I suggest replacing "line-of-sight" with "direction vector" for clarity.\\
   5) Eq. (7) needs more space between the rows of the first matrix.\\
   6) p. 10 l. 49 replace "linearly dependent to" with "linearly dependent on".\\
   7) References 2 and 3 "AAS" should be all capitals, and it would help the reader if more details were provided for AAS conference proceedings (publisher, volume title, volume number, page numbers).}

The authors appreciate the detailed comments. All of the above comments are addressed in the revised paper. With regards to the comments about replacing ``line-of-sight'' with ``direction vector'', the authors decided to continue to use the term, ``line-of-sight'' as it is relatively more common in the literature.

%\textbf{EVAN:} can you address the comments 1) and 7)?

\end{itemize}

\subsection*{Reviewer 2}

\setlength{\leftmargini}{0pt}
\begin{itemize}\setlength{\itemsep}{2\parsep}

\item {\itshape Reviewer \#2: This is a well-written paper that uses the appropriate nonlinear tools to analyze the observability of one satellite
using angle only measurements from another.  The treatment is complete, first taking Lie derivatives and giving sufficient
for the observability rank condition to hold, next computing the condition number of the observability gramian and finally comparing
the condition number with the errors of an Extended Kalman Filter.}

The authors appreciate the comment.

\item {\itshape The dimension of the state is six and the dimension of measurement is two so one expects
that only $L^k_f(h)$, $k=0,1,2$ need only be taken into account for verifying the observability rank condition
and this what the authors do.  The system is real analytic so if the observability rank condition holds for one state
and $k=0,1,2$ then it holds for almost all states.  Therefore I don't see much value in considering larger $k$.}

For linear systems, it is well known that the observability matrix is defined in terms of $A^0,\ldots, A^{n-1}$. This is due to the Cayley-Hamilton theorem guaranteeing that $A^n$ is a linear combination of $A^0,\ldots, A^{n-1}$.

The similar results for nonlinear systems had been nonexistent, perhaps until it was shown that for  $n$-dimensional analytic systems, only the Lie derivatives up to the $(n-1)$-th order are required for observability analysis, as suggested by the reviewer. This is shown in

\begin{itemize}
\item[] M. Anguelova, ``Nonlinear Observability and Identifiability: General Theory and a Case Study of a Kinetic Model,'' Ph.D Dissertation, Chalmers University of Technology, 2004.
\end{itemize}

However, according to the above reference, for multiple output systems, the Lie derivatives up to the $(n-1)$-th order is required for \textit{each} output. This can be noted from the multi-output linear systems as well, where the observability matrix defined as $nm\times n$ when the number of the output is $m$.

Therefore, as far as the authors understand, there is a certain value of deriving higher-order Lie derivatives up to $k=5$.

\end{itemize}

\end{document}

### 2018-06-09 Journal/JAS OD EKF/JAS OD EKF Old/JAS_OD_EKF_rev1.tex
journal added

--- 2018-06-09 Journal/JAS OD EKF/JAS OD EKF Old/JAS_OD_EKF[Conflict].tex
@@
 % your LaTeX will extract the file if required
+
 %\begin{filecontents*}{example.eps}
-%%!PS-Adobe-3.0 EPSF-3.0
-%%%BoundingBox: 19 19 221 221
-%%%CreationDate: Mon Sep 29 1997
-%%%Creator: programmed by hand (JK)
-%%%EndComments
+%%%!PS-Adobe-3.0 EPSF-3.0
+%%%%BoundingBox: 19 19 221 221
+%%%%CreationDate: Mon Sep 29 1997
+%%%%Creator: programmed by hand (JK)
+%%%%EndComments
 %gsave
@@
 %\end{filecontents*}
+
 %
@@
 \usepackage[colorlinks=true, pdfstartview=FitV, linkcolor=black, citecolor= black, urlcolor= black]{hyperref}
-\usepackage{overcite}
-\usepackage{footnpag}			      	% make footnote symbols restart on each page
+%\usepackage{overcite}
+%\usepackage{footnpag}			      	% make footnote symbols restart on each page
 
@@
 The second part of this paper presents a quantitative measure of observability for the relative orbits, and it is formulated by generalizing the observability Gramian of linear dynamic systems.
-An extended Kalman filter is also developed to numerically illustrate the observability of nonlinear relative orbits with angles-only measurements and to show correspondence between the proposed observability measure with filtering errors explicitly.
+An extended Kalman filter is also developed to numerically illustrate the observability of nonlinear relative orbits with angles-only measurements and to show correspondence between the proposed observability measure and filtered solution accuracy.
 
@@
 
-Space-based surveillance or relative navigation is desirable for many spacecraft missions, such as formation control and rendezvous.
+%Space-based surveillance or relative navigation is desirable for many spacecraft missions, such as formation control and rendezvous.
 Spacecraft maneuvers based only on on-board measurements reduce the total operating cost significantly, and it improves safety against communication interruptions with ground stations.
@@
 
-Relative navigation based on angles-only measurements has been investigated in~\cite{WofGelITAES09,Tom11,PatLovPASFMM12}.
+Relative orbit determination (ROD) between spacecraft is desirable for many spacecraft missions.
+In scenarios where the spacecraft are in close-proximity, this is commonly referred to as relative navigation.
+When baselines between the spacecraft are longer, a more appropriate term would be space-based space surveillance or space-based orbit determination.
+In particular, vision-based ROD has received attention recently, since optical sensors have the desirable properties of being low cost and minimal maintenance, while providing accurate line-of-sight measurements, or angles, from a chief to a deputy.
+
+ROD based on angles-only measurements has been investigated in \cite{WofGelITAES09,PatLovPASFMM12}.
 The problem is to determine the relative orbit between a chief spacecraft and a deputy spacecraft by using the line-of-sight between the two objects, assuming that the orbit of the chief is prescribed exactly.
 Reference \cite{WofGelITAES09} shows that the relative orbit is unobservable from angles-only measurements when linear relative orbital dynamics are assumed, unless there are thrusting maneuvers.
-Reference \cite{Tom11} investigates observability by using a relative orbit model linearized in terms of spherical coordinates.
 Reference~\cite{PatLovPASFMM12} introduces the concept of partial observability to determine a basis vector representing a family of relative orbits, and an initial orbit determination technique is developed for this method.
@@
 
-All of these results are based on linear relative orbital dynamics.
-It is straightforward to see that the relative orbit is not observable with angles-only measurements through its linearized dynamics.
-In other words, there are an infinite number of relative orbits that yield the identical line-of-sight measurements, and the orbital distance between a deputy and a chief cannot be determined by angles-only measurements.
+All of the above results are based on linear relative orbital dynamics.
+%It is straightforward to see that the relative orbit is not observable with angles-only measurements through its linearized dynamics. In other words,
+Under these assumptions, there are an infinite number of relative orbits that yield identical line-of-sight measurements, and therefore the orbital distance between a deputy and a chief cannot be determined by angles-only measurements.
 As such, it is required to study the nonlinear relative orbital dynamics to determine observability with angles-only measurements.
@@
 
-This paper constructs observability criteria for the nonlinear relative orbital dynamics represented by the solutions of the two-body problem without linearization.
+This paper constructs observability criteria for the nonlinear relative orbital dynamics represented by the two-body problem without linearization.
 Assuming that a chief is on a circular orbit with a prescribed orbital radius, nonlinear equations of motion for the relative orbital motion of a deputy with respect to the chief are presented.
-A differential geometric method, based on the Lie derivatives of the line-of-sight from the chief to the deputy~\cite{HerKreITAC77}, is applied to obtain sufficient conditions for observability.
-It is shown that under certain geometric conditions on the relative configuration between the chief and the deputy, the nonlinear relative motion is observable from angles-only measurements.
+A differential geometric method, based on the Lie derivatives of the line-of-sight from the chief to the deputy~\cite{HerKreITAC77}, is applied to obtain sufficient conditions for observability, under the assumptions of these equations.
+It is shown that under certain geometric conditions on the relative configuration between the chief and the deputy, the relative motion is observable from angles-only measurements if nonlinear dynamics are assumed.
 
 Next, we present a new measure of observability for relative orbit determination with angles-only measurements.
-This is motivated to understand the degree of observability for the relative orbital dynamics, and to identify the class of relative orbits that have stronger or weaker observability.
+The objectives are to understand the degree of observability for the relative orbital dynamics, and to identify the class of relative orbits that have stronger or weaker observability.
 For a given initial condition of relative orbits, the proposed observability measure determines how much the corresponding relative orbits are easy or difficult to estimate, thereby providing a quantitative measure for relative orbit determination.
@@
 
-In short, the main contributions of this paper are constructing sufficient conditions for the observability of nonlinear relative orbits with angles-only measurements that have not been studied before, and proposing a new quantitative measure of observability.
+In short, the main contributions of this paper are constructing sufficient conditions for the observability of nonlinear relative orbits with angles-only measurements, and proposing a new quantitative measure of observability.
+The main objective is to provide assurance that relative orbits are observable with angles-only measurements under certain conditions, if the nonlinearity of relative orbital dynamics are properly considered.
+More comprehensive study for observability of nonlinear relative orbits, such as obtaining necessary conditions for observability or sufficient conditions for unobservability, is deferred for future study.
 
 This paper is organized as follows.
-The relative orbital dynamics and the orbit determination problem are formulated at Section 2.
-Observability criteria and a observability measure are presented at Section 3 and Section 4, respectively, which are followed by numerical examples and conclusions.
+The relative orbital dynamics and the orbit determination problem are formulated in Section 2.
+Observability criteria and an observability measure are presented in Section 3 and Section 4, respectively, which are followed by numerical examples and conclusions.
 
@@
 The LVLH frame is rotating with the angular velocity of $\mathbf{\omega}=[0,0,n]^T\in\Re^3$, where $n=\sqrt{\frac{\mu}{a^3}}$ is the mean motion of the chief satellite, and $\mu$ denotes the gravitational parameter of the Earth.
-Note that the velocity of the chief is given by $\mathbf{v}_{chief}=[0,na,0]^T\in\Re^3$.
+Note that the inertial velocity of the chief expressed in LVLH coordinates is given by $\mathbf{v}_{chief}=[0,na,0]^T\in\Re^3$.
 Let the relative position of the deputy satellite with respect to the chief be given by $\mathbf{r}=[x,y,z]^T\in\Re^3$ in the LVLH frame.
@@
 
-Considering that the LVLH frame is rotating with the angular velocity of $\mathbf{\omega}=[0,0,n]^T$, the inertial velocity of the deputy satellite is given by
+Considering that the LVLH frame is rotating with the angular velocity of $\mathbf{\omega}=[0,0,n]^T$, the inertial velocity of the deputy satellite expressed in LVLH coordinates is given by
 \begin{align*}
@@
 \end{align}
-where $\y\in\Re^M$ with $M=3$, and it satisfies $\|\y\|=1$.
+where $\y\in\Re^3$, and it satisfies $\|\y\|=1$.
 
@@
 
-Based on the nonlinear dynamic model presented at the previous section, here we derive sufficient conditions for the observability of the relative orbit with line-of-sight measurements.
+Based on the nonlinear dynamic model presented in the previous section, here we derive sufficient conditions for the observability of the relative orbit with line-of-sight measurements.
 
@@
 
-Observability of nonlinear systems and sufficient conditions for observability are summarized as follows~\cite{HerKreITAC77}.
+Observability of nonlinear systems and sufficient conditions for observability are summarized as follows~\cite{HerKreITAC77,Nijvan90}.
 For a given nonlinear dynamic system described with \refeqn{xxdot} and \refeqn{y}, a pair of points $\x_0$ and $\x_1$ is called \textit{indistinguishable} if the outputs of the corresponding solutions starting from each of $\x_0$ and $\x_1$ are identical for a certain time period.
@@
 The system is \textit{locally weakly observable} at $\x_0$, if there exists an open neighborhood $V$ of $\x_0$ such that for every open neighborhood $V_0$ of $\x_0$ contained in $V$, the only indistinguishable point to $\x_0$ is the point $\x_0$ itself.
-This implies that $\x_0$ is distinguishable from any other points in a neighborhood of it.
+This implies that $\x_0$ is distinguishable from any other points in a certain neighborhood of it.
 
@@
 For a non-negative integer $i$, the $i$-th order Lie-derivative is defined by induction as $L_f^i h = L_f (L_f^{i-1} h)$ with $L_f^0 h = h$.
-Define an observability matrix $\mathcal{O}\in\Re^{NM\times N}$ as
-\begin{align*}
-\mathcal{O}(\x_0) = \deriv{}{\x} \begin{bmatrix} L_f^0 h(\x) \\ L_f^1 h(\x)\\ \vdots \\L_f^{N-1} h(\x)\end{bmatrix}\Bigg|_{\x=\x_0}.
-\end{align*}
-It has been shown that the system is locally weakly observable at $\x_0$ if the rank of the observability matrix $\mathcal{O}(\x_0) = N$.
+For a positive integer $k$, define a $k$-th order observability matrix $\mathcal{O}\in\Re^{kM\times N}$ as
+\begin{align*}
+\mathcal{O}(\x_0) = \deriv{}{\x} \begin{bmatrix} L_f^0 h(\x) \\ L_f^1 h(\x)\\ \vdots \\L_f^{k} h(\x)\end{bmatrix}\Bigg|_{\x=\x_0}.
+\end{align*}
+It has been shown that the system is locally weakly observable at $\x_0$ if the rank of the observability matrix $\mathcal{O}(\x_0) = N$ for some $k$~\cite{HerKreITAC77,Nijvan90}.
 When applied to linear dynamics, this yields the well-known observability rank condition for linear systems.
-Note that when there is more than a single measurement, i.e., $M>1$, the observability rank condition can be satisfied without need for computing the higher-order Lie derivatives up to the $N-1$-th order.
+
+In general, the order $k$ of the observability should be increased until the rank condition is satisfied.
+Note that when there is more than a single measurement type for each measurement time, i.e., $M>1$, it is possible to satisfy the observability rank condition without need for computing the higher-order Lie derivatives up to the $N-1$-th order.
 
@@
 \begin{bmatrix}
-\deriv{\y}{\r} & \deriv{\y}{\dot\r}\\
-\deriv{\dot\y}{\r} & \deriv{\dot\y}{\dot\r}\\
+\deriv{\y}{\r} & \deriv{\y}{\dot\r}\\[0.4em]
+\deriv{\dot\y}{\r} & \deriv{\dot\y}{\dot\r}\\[0.4em]
 \deriv{\ddot\y}{\r} & \deriv{\ddot\y}{\dot\r}
@@
 \mathcal{O}_{20} & \mathcal{O}_{21}
-\end{bmatrix},\label{eqn:OO}
+\end{bmatrix}\in\Re^{9\times 6},\label{eqn:OO}
 \end{align}
 where the submatrices $\mathcal{O}_{ij}\in\Re^{3\times 3}$ are defined for each block, and we used the fact that the measurement $\y$ is independent of $\dot\r$.
-Here we consider the part of the observability matrix obtained by the second order Lie derivatives of the measurement due to complexity.
-But, this still provides sufficient conditions for observability with interesting geometric meanings described later.
+Here we consider the observability matrix obtained by up to the second order Lie derivatives of the measurement due to complexity.
+But, this still provides sufficient conditions for observability under certain geometric conditions described later.
 
@@
 
-We wish to show that there is no non-zero vector $\mathbf{c}$ satisfying \refeqn{Oc} under the given conditions. At the first three rows of \refeqn{Oc}, we have
+We wish to show that there is no non-zero vector $\mathbf{c}$ satisfying \refeqn{Oc} under the given conditions. In the first three rows of \refeqn{Oc}, we have
 \begin{align}
@@
 This implies that $\mathbf{c}_1=0$ from \refeqn{c10}.
-Substituting this back to \refeqn{Oc}, we have $\mathcal{O}_{11}\mathbf{c}_2=\mathcal{O}_{00}\mathbf{c}_2=0$, which follows that $\mathbf{c}_2=c\r$ for some constant $c$.
+Substituting this back into \refeqn{Oc}, we have $\mathcal{O}_{11}\mathbf{c}_2=\mathcal{O}_{00}\mathbf{c}_2=0$, which follows that $\mathbf{c}_2=c\r$ for some constant $c$.
 However, when $\mathbf{c}_2=c\r$, the last three rows of \refeqn{Oc} are given by
@@
 \mathcal{O}_{20} \mathbf{c}_1 +\mathcal{O}_{21} \mathbf{c}_2 
-= -\mathcal{O}_{00}(\mathbf{a}_2+2c\mathrm{v}_{ref})= -\frac{1}{\|\r\|}(I-\y\y^T)(\mathbf{a}_2+2c\mathbf{v}_{rel})=0.
+= -\mathcal{O}_{00}(\mathbf{a}_2+2c\mathbf{v}_{ref})= -\frac{1}{\|\r\|}(I-\y\y^T)(\mathbf{a}_2+2c\mathbf{v}_{rel})=0.
 \end{align*}
@@
 This implies that $\mathbf{c}_1=0$.
-Then, by the same argument given at \refeqn{row3i1}, we have $\mathbf{c}_2=0$.
+Then, by the same argument given in \refeqn{row3i1}, we have $\mathbf{c}_2=0$.
 In short, under the given condition \refeqn{cond2}, the equation \refeqn{Oc} implies that $\mathbf{c}=0$.
@@
 
-This theorem shows that the nonlinear relative orbital dynamics is indeed observable from angles-only measurements if certain geometric conditions are satisfied.
-This generalizes the prior observability study developed for the linear relative orbital dynamics only.
+This theorem shows that, based on the Lie-derivative sufficiency test for observability, the nonlinear relative orbital dynamics is indeed observable from angles-only measurements if certain geometric conditions are satisfied.
+This generalizes the prior observability study developed for the linear relative orbital dynamics only, such as reference \cite{WofGelITAES09}, and it provides sufficient conditions for nonlinear observability.
 
@@
 For both cases, we have $\r\times\mathbf{v}_{rel}\neq 0$.
-At \refeqn{vrel}, $\mathbf{v}_{rel}$ corresponds to the relative velocity observed in the inertial frame.
+In \refeqn{vrel}, $\mathbf{v}_{rel}$ corresponds to the relative velocity observed in the inertial frame.
 Therefore, the given sufficient conditions for observability are violated when the relative velocity vector is parallel to the relative position vector in the inertial frame.
@@
 
-The second condition of each of \refeqn{cond1} and \refeqn{cond2} is satisfied in general, as the expressions for $\mathbf{a}_1$ and $\mathbf{a}_2$ representing certain accelerations are relatively arbitrary at \refeqn{a1} and \refeqn{a2}.
-Similar to the first condition, there is little chance that the second condition is violated over a finite time period entirely.
-
-The third condition of \refeqn{cond1} implies that three vectors $\r$, $\mathbf{v}_{rel}$ and $\mathbf{a}_1$ do not belong to the same plane.
+The second condition of each of \refeqn{cond1} and \refeqn{cond2} is satisfied in general, as the expressions for $\mathbf{a}_1$ and $\mathbf{a}_2$ representing certain accelerations are relatively arbitrary in \refeqn{a1} and \refeqn{a2}.
+Similar to the first condition, there is little chance that the second condition is violated over a finite time period.
+
+The expression $\r^T(\mathbf{v}_{rel}\times \mathbf{a}_1)$ in the third condition of \refeqn{cond1} corresponds to the (signed) volume of a parallelepiped whose sides are given by the vectors $\r$, $\mathbf{v}_{ref}$, and $\mathbf{a}_1$.
+Therefore, the third condition is violated if all of the three vectors are coplanar.
+This may happen when there is no out-of-plane motion of the deputy, i.e., the orbital plane of the deputy is identical to the chief, and $z(t)\equiv 0$ for all $t$.
+As such, unlike the prior two conditions, the third condition is violated always for any coplanar relative motions.
 The third condition of \refeqn{cond2} has a similar structure as well.
-Unlike the prior two conditions, this can be violated over a certain time period if the relative motion is planar, i.e., when $z(t)\equiv 0$ for all $t$.
-
-However, \refeqn{cond1} and \refeqn{cond2} are sufficient conditions for observability, and the fact that any of these condition is not satisfied does not necessarily imply that the considered point is not observable.
-In such cases, the third or higher order Lie derivatives should be checked to determine observability.
-This paper is focused on verifying the observability of nonlinear relative orbits via angles-only measurements under certain conditions, and deriving the higher-order Lie derivatives may be a topic for future work.
+
+However, \refeqn{cond1} and \refeqn{cond2} are sufficient conditions for observability, and the fact that any of these conditions is not satisfied does not necessarily imply that the considered point is not observable.
+In such cases, the third or higher order Lie derivatives should be checked to confirm observability.
+Furthermore, the presented differential geometric approaches to determine observability do not yield necessary conditions for observability.
+As such, these conditions cannot be used to determine the certain types of relative motions that are not observable with angles-only measurements.
+This paper is focused on verifying the observability of nonlinear relative orbits via angles-only measurements under certain conditions, and determining observability with higher-order Lie derivatives or necessary conditions for observability is deferred for future investigation.
 
@@
 The observability criteria derived in the previous section guarantee that the nonlinear relative orbital dynamics are indeed observable under certain geometric conditions.
-However, it does not provide any information about how strong the observability is.
-
+However, they are limited in the sense that they only provide binary information about observability.
+Therefore, they cannot be used to determine whether a certain type of orbit is more easily estimated compared with other orbits based on angles-only measurements.
+Furthermore, as they are defined in terms of the state vector at a specific time, they do not capture the effects of the measurement trajectory over a certain time period on the observability, which often predicts the performance of estimation.
+
+Motivated by the above issues, we propose a quantitative measure that represents the degree of observability for relative orbits in this section.
+More explicitly, this measure evaluates the observability in a quantitative manner as it determines how effectively the relative orbit can be estimated by line-of-sight measurements.
+
+The problem of assigning physically meaningful measures of the quality of observability has been introduced in~\cite{BroPNEC66,MonITAC67,MulWebA72} for linear systems, and in particular, in \cite{MulWebA72} the determinant, trace, and maximal eigenvalue of the inverse characteristic observability matrix are maximized with respect to certain parameters of a dynamic system.
 Several approaches have been proposed for an observability measure of nonlinear systems.
 %The measure of observability has been generalized to nonlinear dynamic systems in order to obtain balanced realizations.
-An energy-like observability function is introduced~\cite{SchSCL93,Sch94,NewKriPICDC98} to measure the degree of contribution of an initial state to the output, and it is applied to a pendulum system.
+An energy-like observability function is introduced in references~\cite{SchSCL93,Sch94,NewKriPICDC98} to measure the degree of contribution of an initial state to the output, and it is applied to a pendulum system.
 But, this approach is only applicable to asymptotically stable equilibrium states.
-An empirical observability Gramian is introduced~\cite{LalMarPIWC99,LalMarIJRNC02}, which is essentially a covariance matrix of the output computed by a number of sample trajectories.
+An empirical observability Gramian is introduced in references~\cite{LalMarPIWC99,LalMarIJRNC02}, which is essentially a covariance matrix of the output computed by a number of sample trajectories.
 It has been applied to a balanced realization of chemical processes~\cite{HahEdgJPC03,HahEdgCCE02}.
@@
 
-Another nonlinear observability measure is introduced~\cite{KreIdePICDC09}, which is basically the Gramian for the sensitivity of the output with respect to the initial condition:
+Another nonlinear observability measure is introduced in reference~\cite{KreIdePICDC09}, which is basically the Gramian for the sensitivity of the output with respect to the initial condition:
 \begin{align}
@@
 In general, the Gramian of a set of time-dependent functions represents the degree of linear independence of those functions over a given period of time.
-Therefore, the observability Gramian defined at \refeqn{Wo_NL} measures how much the sensitivities of the output with respect to the initial condition are linearly independent of each other.
-This represents the degree of observability, since for example, if $\deriv{\y}{\x_{0_i}}$ is linearly dependent to $\deriv{\y}{\x_{0_j}}$, then it is difficult to distinguish the effects on $\delta \x_{0_i}$ to the output $\y$ from the effects of $\delta \x_{0_j}$ on $\y$.
+Therefore, the observability Gramian defined in \refeqn{Wo_NL} measures how much the sensitivities of the output with respect to the initial condition are linearly independent of each other.
+This represents the degree of observability, since for example, if $\deriv{\y}{\x_{0_i}}$ is linearly dependent on $\deriv{\y}{\x_{0_j}}$, then it is difficult to distinguish the effects on $\delta \x_{0_i}$ to the output $\y$ from the effects of $\delta \x_{0_j}$ on $\y$.
 Therefore, the condition number of $\mathcal{W} (\x_0,t_0,t_f)$ can be used as a measure of observability.
@@
 \begin{align}
-[\mathcal{W}]_{ij} = \frac{1}{4\epsilon^2}\sum_{k=0}^N (\y^{i+}_k-\y^{i-}_k)^T(\y^{j+}_k-\y^{j-}_k) \Delta t_k,
-\end{align}
-where $\y^{i\pm}_k$ denotes the value of $\y$ at $t=t_k$ obtained from the state trajectory starting from the initial condition of $\x(t_0)=\x_0\pm \epsilon \mathbf{e}_i $ for a positive constant $\epsilon$, and the standard basis $\mathbf{e}_i$ of $\Re^n$.
+[\mathcal{W}]_{ij} = \frac{1}{4\epsilon^2}\sum_{k=0}^{N_f} (\y^{i+}_k-\y^{i-}_k)^T(\y^{j+}_k-\y^{j-}_k) \Delta t_k,
+\end{align}
+where $\y^{i\pm}_k$ denotes the value of $\y$ at $t=t_k$ obtained from the state trajectory starting from the initial condition of $\x(t_0)=\x_0\pm \epsilon \mathbf{e}_i $ for a positive constant $\epsilon$, and $\mathbf{e}_i$ of $\Re^n$ is the standard basis.
+The positive integer $N_f$ is chosen such that $t_{N_f}=t_f$.
+
+In contrast to the observability criteria presented in Section 3, the proposed observability measure is obtained numerically.
+As such, it is not useful to identify the particular class of the relative orbits that yield stronger, or weaker observability.
+However, it can be used to analyze the degree of observability in a complementary manner when the sufficient conditions of the observability criteria are violated, as illustrated by the following numerical examples.
 
@@
 
-It is assumed that the chief is on a circular orbit with the altitude of $500\,\mathrm{km}$ and zero inclination.
+It is assumed that the chief is on a circular orbit with an altitude of $500\,\mathrm{km}$ and zero inclination.
 Here, we study the effects of the deputy inclination on the observability measure and filtering performance.
@@
 \begin{align*}
-a_{deputy} = \parenth{\frac{T_{chief}\sqrt{\mu}}{2\pi}}^{2/3}.
-\end{align*}
-Therefore, the relative orbits are periodic. The inclination of the deputy is varied as $i_{deputy}=0,\,10,\,\cdots,\,50^\circ$.
-
-For the extended Kalman filter, the initial estimate of the state is chosen as two times of its true value, i.e., $\hat\x_0=2\x_0$.
+a_{deputy} = a_{chief} = \parenth{\frac{T_{chief}\sqrt{\mu}}{2\pi}}^{2/3}.
+\end{align*}
+Therefore, the relative orbits are periodic.
+The inclination of the deputy is varied as $i_{deputy}=0,\,10,\,\cdots,\,50^\circ$, where the $i_{deputy}=0$ case corresponds to coplanar orbits such that the third condition of \refeqn{cond1} and \refeqn{cond2} is violated.
+
+For the extended Kalman filter, the initial estimate of the state is chosen as two times its true value, i.e., $\hat\x_0=2\x_0$.
 This implies that there is a large initial error in the magnitude of the state, which is difficult to estimate accurately using angles-only measurements.
-The initial covariance of the state is chosen as $P_0=\mathrm{diag}[50^2I_{3\times 3}, (50n)^2I_{3\times 3}]$.
-The covariance matrices for the process noise and the measurement noise are $Q_k=\mathrm{diag}[10^{-6}I_{3\times 3}, 10^{-8}I_{3\times 3}]$, and $R_k=1.306^2 I_{3\times 3}\,\mathrm{deg}^2$.
+The initial covariance of the state is chosen as $P_0=\mathrm{diag}[(50\,\mathrm{km})^2I_{3\times 3}, (50n\,\mathrm{km/s})^2I_{3\times 3}]$.
+The covariance matrices for the process noise and the measurement noise are $Q_k=\mathrm{diag}[10^{-6}I_{3\times 3}, 10^{-8}I_{3\times 3}]$, and \\$R_k=1.306^2 I_{3\times 3}\,\mathrm{deg}^2$.
 It is assumed that the line-of-sight is measured at every $\Delta t = 0.47$ seconds.
 It is simulated for 20 orbits of the chief around the Earth.
+Throughout this paper, the units are chosen as $\mathrm{km}$, $\mathrm{sec}$, and $\mathrm{rad}$, if not specified otherwise.
 
 To compare the convergence property of each case, the following measures for estimation errors are introduced.
+%\begin{align*}
+%e_{dir_k}  & = \cos^{-1}\parenth{\frac{\mathbf{x}(t_k)^T \hat{\mathbf{x}} (t_k)}
+%{\|\mathbf{x}(t_k)\| \|\hat{\mathbf{x}} (t_k)\|}},\\
+%e_{mag_k}  & = 
+%\frac{|\|\mathbf{x}(t_k)\|-\|\hat{\mathbf{x}} (t_k)\||}{\|\mathbf{x}(t_k)\|},
+%\end{align*}
 \begin{align*}
@@
 \end{align*}
-where the error variable $e_{dir}$ represents the angle between the true state and the estimated state, and the variable $e_{mag}$ corresponds to the normalized error for the magnitude of the state vector.
-This choice of error variables are motivated by the fact that the lines-of-sight provide direct measurements on the relative direction, but they do not yield any information about the range directly.
-
-The condition number of the proposed observability Gramian and the mean values of the above error variables are summarized at Table \ref{tab:Ei} for each inclination.
+where the variable $e_{dir}$ represents an angular error between the true state and the estimated state, and the variable $e_{mag}$ corresponds to the normalized error for the magnitude of the state vector.
+This choice of error variables is motivated by the fact that the lines-of-sight provide direct measurements on the relative direction, but they do not yield any information about the range directly.
+As such, it is expected that the direction errors remain small, but the magnitude error is determined indirectly through the nonlinear observability representing the weak coupling between the dynamics of the direction and the dynamics of the magnitude.
+
+The condition number of the proposed observability Gramian and the mean values of the above error variables over all measurement times are summarized in Table \ref{tab:Ei} for each inclination.
 The second column shows that the condition number of the Gramian decreases as the inclination increases, i.e., it is relatively easier to determine the relative orbit via angles-only measurements if the inclination of the deputy is higher.
 However, the condition numbers are quite large overall, and it shows that there is only weak observability.
-At the third  and sixth columns, the mean error in the magnitude of the estimated state decreases as the inclination increases, and it is consistent with the condition number of the observability Gramian.
-The correlation between the condition number and the estimation error is more apparent at the fourth and seventh columns, which show the error in the magnitude of the estimated state averaged over the last 10 orbits.
+In the third and sixth columns, the mean error in the magnitude of the estimated state decreases as the inclination increases, and it is consistent with the condition number of the observability Gramian.
+The correlation between the condition number and the estimation error is more apparent in the fourth and seventh columns, which show the error in the magnitude of the estimated state averaged over the last 10 orbits.
 This is because there is greater transient response in the estimated state when the inclination is larger.
 The steady-state estimation error represented by the fourth column strictly decreases as the inclination increases.
-There is no meaningful variation at the fifth and eighth columns, as the estimation errors in the direction are determined mostly by the line-of-sight measurements.
+
+There is no meaningful variation in the fifth and eighth columns for $e_{dir}$, as the estimation errors in the direction are determined mostly by the line-of-sight measurements.
 In the sixth, seventh, and eighth columns, the $J_2$ perturbation is considered in the chief and deputy orbit process models.
@@
 
-Estimation results are also illustrated at Figures \ref{fig:i0}--\ref{fig:i50}.
-At Figure \ref{fig:i50}, the magnitude of the estimated state becomes quite close to the true value after four orbits.
+Estimation results are also illustrated in Figures \ref{fig:i0}--\ref{fig:i50}.
+In Figure \ref{fig:i50}, the magnitude of the estimated state becomes quite close to the true value after four orbits.
 This confirms the proposed observability of nonlinear relative orbital dynamics with angles-only measurements.
 
-\clearpage\newpage
+%\clearpage\newpage%asdf
+\begin{figure}
+\centerline{
+	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
+		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_i_0.pdf}}
+%	\hspace*{0.15\textwidth}
+		\subfigure[Orbital trajectory in the LVLH frame]{
+		\includegraphics[width=0.5\textwidth]{OT_traj_PE_i_0.pdf}}
+}
+\centerline{
+	\subfigure[Estimation errors]{
+		\includegraphics[width=0.5\textwidth]{PE_i_0_e.pdf}}
+%	\hspace*{0.15\textwidth}
+		\subfigure[Magnitude of state (red:true, blue:estimate)]{
+		\includegraphics[width=0.5\textwidth]{PE_i_0_xx_norm.pdf}}
+}
+\caption{Estimation results: $i=0$}\label{fig:i0}
+\end{figure}
+
+\begin{figure}
+\centerline{
+	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
+		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_i_30.pdf}}
+%	\hspace*{0.15\textwidth}
+		\subfigure[Orbital trajectory in the LVLH frame]{
+		\includegraphics[width=0.5\textwidth]{OT_traj_PE_i_30.pdf}}
+}
+\centerline{
+	\subfigure[Estimation errors]{
+		\includegraphics[width=0.5\textwidth]{PE_i_30_e.pdf}}
+%	\hspace*{0.15\textwidth}
+		\subfigure[Magnitude of state (red:true, blue:estimate)]{
+		\includegraphics[width=0.5\textwidth]{PE_i_30_xx_norm.pdf}}
+}
+\caption{Estimation results: $i=30$}\label{fig:i30}
+\end{figure}
+
 \begin{figure}
@@
 	\subfigure[Estimation errors]{
-		\includegraphics[width=0.3\textwidth]{PE_i_0_e.pdf}}
-	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.5\textwidth]{PE_i_30_e_J2.pdf}}
+%	\hspace*{0.15\textwidth}
 		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.3\textwidth]{PE_i_0_xx_norm.pdf}}
-}
-\caption{Estimation results: $i=0$}\label{fig:i0}
+		\includegraphics[width=0.5\textwidth]{PE_i_30_xx_norm_J2.pdf}}
+}
+\caption{Estimation results with $J_2$ perturbations: $i=30$}\label{fig:i30}
 \end{figure}
+
 \begin{figure}
 \centerline{
-	\subfigure[Estimation errors]{
-		\includegraphics[width=0.3\textwidth]{PE_i_30_e.pdf}}
-	\hspace*{0.15\textwidth}
-		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.3\textwidth]{PE_i_30_xx_norm.pdf}}
-}
-\caption{Estimation results: $i=30$}\label{fig:i30}
-\end{figure}
-\begin{figure}
+	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
+		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_i_50.pdf}}
+%	\hspace*{0.15\textwidth}
+		\subfigure[Orbital trajectory in the LVLH frame]{
+		\includegraphics[width=0.5\textwidth]{OT_traj_PE_i_50.pdf}}
+}
 \centerline{
 	\subfigure[Estimation errors]{
-		\includegraphics[width=0.3\textwidth]{PE_i_30_e_J2.pdf}}
-	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.5\textwidth]{PE_i_50_e.pdf}}
+%	\hspace*{0.15\textwidth}
 		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.3\textwidth]{PE_i_30_xx_norm_J2.pdf}}
-}
-\caption{Estimation results with $J_2$ perturbations: $i=30$}\label{fig:i30}
-\end{figure}
-\begin{figure}
-\centerline{
-	\subfigure[Estimation errors]{
-		\includegraphics[width=0.3\textwidth]{PE_i_50_e.pdf}}
-	\hspace*{0.15\textwidth}
-		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.3\textwidth]{PE_i_50_xx_norm.pdf}}
+		\includegraphics[width=0.5\textwidth]{PE_i_50_xx_norm.pdf}}
 }
@@
 
-\clearpage\newpage
+%\clearpage\newpage%asdf
 
@@
 
-Next, we study the effects of the size of the relative orbits on nonlinear observability.
-Same as before, the chief is on a circular orbit with the altitude of $500\,\mathrm{km}$ and zero inclination.
+Next, we study the effects of varying the size of the relative orbits proportionally on nonlinear observability.
+As before, the chief is on a circular orbit with the altitude of $500\,\mathrm{km}$ and zero inclination.
 The initial conditions for the relative position and the velocity are chosen as
@@
 where $r_{x_0}\in\Re$ is the initial relative distance that is varied as $10,\,20,\,50,\,100\,\mathrm{km}$.
-For the linearized relative orbital dynamics, these initial conditions yield elliptic orbits around the chief, but for the nonlinear relative orbital dynamics, there is drift proportional to the size of the orbit parameterized by $r_{x_0}$.
-While the resulting orbits are planar such that the third condition of \refeqn{cond1} and \refeqn{cond2} is violated, it is still desirable to understand how the orbital size or drift affects nonlinear observability.
+For the linearized relative orbital dynamics of the two-body problem, these initial conditions yield elliptic orbits around the chief, but for the nonlinear relative orbital dynamics, there is drift proportional to the size of the orbit parameterized by $r_{x_0}$.
+While the resulting orbits are coplanar such that the third condition of \refeqn{cond1} and \refeqn{cond2} is violated, it is still desirable to understand how the orbital size or drift affects nonlinear observability.
 For the extended Kalman filter, the initial estimate of the state is chosen as $\hat\x_0=1.5\x_0$, and the other parameters are identical to the prior case.
 
-The corresponding condition number of the observability Gramian and the estimation errors are summarized at Table \ref{tab:Es}.
+As discussed in Section 3, these relative orbits violate the observability criteria summarized in Proposition 1 since they are coplanar orbits.
+Therefore, Proposition 1 cannot be applied to investigate the observability of these orbits.
+Instead, the observability is analyzed by the observability measure proposed in Section 4 as follows.
+
+The corresponding condition number of the observability Gramian and the estimation errors are summarized in Table \ref{tab:Es}.
 It is shown that as $r_{x_0}$ increases, the condition number decreases and the estimation error for the magnitude of the estimated state decreases.
-
-% Table before J2 is added
+This is expected since the nonlinear relative orbital dynamics have stronger effects as the separation between the deputy and the chief increases.
+Figures \ref{fig:r0_20}-\ref{fig:r0_100} illustrate the corresponding orbital trajectories and the estimation errors.
+In particular, in Figure \ref{fig:r0_100}.(d) when the separation between the chief and the deputy is greatest, it is shown that the magnitude of the estimated state vector follows the large variation of the magnitude of the true state accurately.
+
+%% Table before J2 is added
+%\begin{table}[h]
+%\caption{Observability Gramian and estimation errors for varying size of orbits}\label{tab:Es}
+%\begin{center}
+%\begin{tabularx}{0.7\textwidth}{>{\centering $}X<{$}*{4}{>{$}c<{$}}}\toprule
+%r_{x_0}\,(\mathrm{km}) & \mathrm{cond}[\mathcal{W}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}]\,(\mathrm{rad}) \\\midrule
+% 10 & 10^{15.8194} & 2.7739  &  2.8535  &  0.0022\\
+% 20 & 10^{15.2832} &  1.1787  &  1.1093  &  0.0022\\
+% 50 & 10^{14.4052} &  0.3146  &  0.2284  &  0.0020\\
+%100 & 10^{12.9984} &  0.1038  &  0.0354  &  0.0016\\
+%\bottomrule
+%\end{tabularx}\\
+%(${}^\dagger$ averaged for the last 10 orbits)
+%\end{center}
+%\end{table}
+
 \begin{table}[h]
@@
 \begin{center}
-\begin{tabularx}{0.7\textwidth}{>{\centering $}X<{$}*{4}{>{$}c<{$}}}\toprule
-r_{x_0}\,(\mathrm{km}) & \mathrm{cond}[\mathcal{W}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}]\,(\mathrm{rad}) \\\midrule
- 10 & 10^{15.8194} & 2.7739  &  2.8535  &  0.0022\\
- 20 & 10^{15.2832} &  1.1787  &  1.1093  &  0.0022\\
- 50 & 10^{14.4052} &  0.3146  &  0.2284  &  0.0020\\
-100 & 10^{12.9984} &  0.1038  &  0.0354  &  0.0016\\
+\begin{tabularx}{\textwidth}
+{
+>{$}c<{$}
+*{1}{>{$}c<{$}} |
+*{3}{>{$}c<{$}} |
+*{3}{>{$}c<{$}}
+}
+\toprule
+\multirow{2}{*}{$r_{x_0}\,(\mathrm{km}) $} & \multirow{2}{*}{$\mathrm{cond}[\mathcal{W}]$} & \multicolumn{3}{c}{\multirow{1}{*}{Two-Body Solution}} & \multicolumn{3}{c}{\multirow{1}{*}{$J_2$ Perturbation}} \\
+ & &  \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}] \\\midrule
+%{>{\centering $}X<{$}*{7}{>{$}c<{$}}}\toprule
+%i_{dep} & \mathrm{cond}[\mathcal{W}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}]\,(rad) \\\midrule
+10 & 10^{15.8194} & 2.7739  &  2.8535  &  0.0022 & 2.6790 & 1.927 & 0.0028\\
+ 20 & 10^{15.2832} &  1.1787  &  1.1093  &  0.0022 & 1.024 & 0.8542 & 0.0027\\
+ 50 & 10^{14.4052} &  0.3146  &  0.2284  &  0.0020 & 0.1531 & 0.1473 & 0.0025\\
+100 & 10^{12.9984} &  0.1038  &  0.0354  &  0.0016 & 0.1318 & 0.1197 & 0.0023\\
 \bottomrule
@@
 
-%% With J2
-%\begin{table}[h]
-%\caption{Observability Gramian and estimation errors for varying size of orbits}\label{tab:Es}
-%\begin{center}
-%\begin{tabularx}{\textwidth}
-%{
-%>{$}c<{$}
-%*{1}{>{$}c<{$}} |
-%*{3}{>{$}c<{$}} |
-%*{3}{>{$}c<{$}}
-%}
-%\toprule
-%\multirow{2}{*}{$r_{x_0}\,(\mathrm{km}) $} & \multirow{2}{*}{$\mathrm{cond}[\mathcal{W}]$} & \multicolumn{3}{c}{\multirow{1}{*}{Two-Body Solution}} & \multicolumn{3}{c}{\multirow{1}{*}{$J_2$ Perturbation}} \\
-% & &  \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}] \\\midrule
-%%{>{\centering $}X<{$}*{7}{>{$}c<{$}}}\toprule
-%%i_{dep} & \mathrm{cond}[\mathcal{W}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}] & \mathrm{E}[e_{mag}] & \mathrm{E}[e_{mag}]^{\dagger} & \mathrm{E}[e_{dir}]\,(rad) \\\midrule
-%10 & 10^{15.8194} & 2.7739  &  2.8535  &  0.0022 & 2.6790 & 1.927 & 0.0028\\
-% 20 & 10^{15.2832} &  1.1787  &  1.1093  &  0.0022 & 1.024 & 0.8542 & 0.0027\\
-% 50 & 10^{14.4052} &  0.3146  &  0.2284  &  0.0020 & 0.1531 & 0.1473 & 0.0025\\
-%100 & 10^{12.9984} &  0.1038  &  0.0354  &  0.0016 & 0.1318 & 0.1197 & 0.0023\\
-%\bottomrule
-%\end{tabularx}\\
-%(${}^\dagger$ averaged for the last 10 orbits)
-%\end{center}
-%\end{table}
-
 \afterpage{\clearpage
@@
 \centerline{
+	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
+		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_r0_JAS_20}}
+%	\hspace*{0.15\textwidth}
+		\subfigure[Orbital trajectory in the LVLH frame]{
+		\includegraphics[width=0.5\textwidth]{OT_traj_PE_r0_JAS_20}}
+}
+\centerline{
 	\subfigure[Estimation errors]{
-		\includegraphics[width=0.45\textwidth]{EKF_PE_r0_20_e.pdf}}
-	\hspace*{0.05\textwidth}
+		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_20_e.pdf}}
+%	\hspace*{0.15\textwidth}
 		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.45\textwidth]{EKF_PE_r0_20_xx_norm.pdf}}
+		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_20_xx_norm.pdf}}
 }
@@
 \centerline{
+	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
+		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_r0_JAS_50}}
+%	\hspace*{0.15\textwidth}
+		\subfigure[Orbital trajectory in the LVLH frame]{
+		\includegraphics[width=0.5\textwidth]{OT_traj_PE_r0_JAS_50}}
+}
+\centerline{
 	\subfigure[Estimation errors]{
-		\includegraphics[width=0.45\textwidth]{EKF_PE_r0_50_e.pdf}}
-	\hspace*{0.05\textwidth}
+		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_50_e.pdf}}
+%	\hspace*{0.15\textwidth}
 		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.45\textwidth]{EKF_PE_r0_50_xx_norm.pdf}}
+		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_50_xx_norm.pdf}}
 }
@@
 \centerline{
+	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
+		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_r0_JAS_100}}
+%	\hspace*{0.15\textwidth}
+		\subfigure[Orbital trajectory in the LVLH frame]{
+		\includegraphics[width=0.5\textwidth]{OT_traj_PE_r0_JAS_100}}
+}
+\centerline{
 	\subfigure[Estimation errors]{
-		\includegraphics[width=0.45\textwidth]{EKF_PE_r0_100_e.pdf}}
-	\hspace*{0.05\textwidth}
+		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_100_e.pdf}}
+%	\hspace*{0.15\textwidth}
 		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.45\textwidth]{EKF_PE_r0_100_xx_norm.pdf}}
+		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_100_xx_norm.pdf}}
 }
@@
 \end{figure}
+
 %\begin{figure}
@@
 
-It is well known that the linearized relative orbital dynamics are not observable with angles-only
+It is well known that linearized relative orbital dynamics are not observable with angles-only
 measurements.
-This paper studies the observability for nonlinear relative orbital dynamics, and it shows that the relative orbit is indeed observable under certain geometric conditions.
+This paper studies the observability for nonlinear relative orbital dynamics, and it derives a set of sufficiency conditions to show that the relative orbit is indeed observable under certain geometric conditions.
 Also, a quantitative observability measure that represents the degree of observability is proposed.
-They are illustrated by several numerical examples.
-Future works include constructing observability criteria based on higher-order Lie derivatives of the measurement, and developing nonlinear filtering techniques that can capture the weak coupling between the dynamics of relative distance and the line-of-sight measurements effectively.
+It is illustrated by several numerical examples.
+Future work objectives include constructing observability criteria based on higher-order Lie derivatives of the measurement, and developing necessary conditions for observability to classify the type of relative orbits that are not observable completely.
+We also plan to develop nonlinear filtering techniques that can capture the weak coupling between the dynamics of relative motion and the line-of-sight measurements effectively, as well as studying the effects of higher-order gravitational perturbations on the observability.
 
@@
 %\bibliographystyle{spphys}       % APS-like style for physics
+
 \bibliography{BibSources}   % name your BibTeX data base

### 2018-06-09 Journal/JAS OD EKF/JAS_OD_EKF.tex
journal added

--- 2018-06-09 Journal/JAS OD EKF/JAS OD EKF Old/JAS_OD_EKF_rev1.tex
@@
 In scenarios where the spacecraft are in close-proximity, this is commonly referred to as relative navigation.
-When baselines between the spacecraft are longer, a more appropriate term would be space-based space surveillance or space-based orbit determination.
+When baselines between spacecraft are longer, a more appropriate term would be space-based space surveillance or space-based orbit determination.
 In particular, vision-based ROD has received attention recently, since optical sensors have the desirable properties of being low cost and minimal maintenance, while providing accurate line-of-sight measurements, or angles, from a chief to a deputy.
@@
 Next, we present a new measure of observability for relative orbit determination with angles-only measurements.
-The objectives are to understand the degree of observability for the relative orbital dynamics, and to identify the class of relative orbits that have stronger or weaker observability.
+This is motivated to understand the degree of observability for the relative orbital dynamics, and to identify the class of relative orbits that have stronger or weaker observability.
 For a given initial condition of relative orbits, the proposed observability measure determines how much the corresponding relative orbits are easy or difficult to estimate, thereby providing a quantitative measure for relative orbit determination.
@@
 
-In short, the main contributions of this paper are constructing sufficient conditions for the observability of nonlinear relative orbits with angles-only measurements, and proposing a new quantitative measure of observability.
-The main objective is to provide assurance that relative orbits are observable with angles-only measurements under certain conditions, if the nonlinearity of relative orbital dynamics are properly considered.
-More comprehensive study for observability of nonlinear relative orbits, such as obtaining necessary conditions for observability or sufficient conditions for unobservability, is deferred for future study.
+In short, the main contributions of this paper are constructing sufficient conditions for the observability of nonlinear relative orbits with angles-only measurements that have not been studied before, and proposing a new quantitative measure of observability.
 
@@
 The system is \textit{locally weakly observable} at $\x_0$, if there exists an open neighborhood $V$ of $\x_0$ such that for every open neighborhood $V_0$ of $\x_0$ contained in $V$, the only indistinguishable point to $\x_0$ is the point $\x_0$ itself.
-This implies that $\x_0$ is distinguishable from any other points in a certain neighborhood of it.
+This implies that $\x_0$ is distinguishable from any other points in a neighborhood of it.
 
@@
 \begin{bmatrix}
-\deriv{\y}{\r} & \deriv{\y}{\dot\r}\\[0.4em]
-\deriv{\dot\y}{\r} & \deriv{\dot\y}{\dot\r}\\[0.4em]
+\deriv{\y}{\r} & \deriv{\y}{\dot\r}\\
+\deriv{\dot\y}{\r} & \deriv{\dot\y}{\dot\r}\\
 \deriv{\ddot\y}{\r} & \deriv{\ddot\y}{\dot\r}
@@
 But, this still provides sufficient conditions for observability under certain geometric conditions described later.
+%asdf
 
@@
 This may happen when there is no out-of-plane motion of the deputy, i.e., the orbital plane of the deputy is identical to the chief, and $z(t)\equiv 0$ for all $t$.
-As such, unlike the prior two conditions, the third condition is violated always for any coplanar relative motions.
+As such, unlike the prior two conditions, the third condition is violated always for any planar relative motions.
 The third condition of \refeqn{cond2} has a similar structure as well.
@@
 However, \refeqn{cond1} and \refeqn{cond2} are sufficient conditions for observability, and the fact that any of these conditions is not satisfied does not necessarily imply that the considered point is not observable.
-In such cases, the third or higher order Lie derivatives should be checked to confirm observability.
-Furthermore, the presented differential geometric approaches to determine observability do not yield necessary conditions for observability.
-As such, these conditions cannot be used to determine the certain types of relative motions that are not observable with angles-only measurements.
-This paper is focused on verifying the observability of nonlinear relative orbits via angles-only measurements under certain conditions, and determining observability with higher-order Lie derivatives or necessary conditions for observability is deferred for future investigation.
+In such cases, the third or higher order Lie derivatives should be checked to determine observability.
+This paper is focused on verifying the observability of nonlinear relative orbits via angles-only measurements under certain conditions, and determining observability with higher-order Lie derivatives is referred to future investigation.
 
@@
 The observability criteria derived in the previous section guarantee that the nonlinear relative orbital dynamics are indeed observable under certain geometric conditions.
-However, they are limited in the sense that they only provide binary information about observability.
-Therefore, they cannot be used to determine whether a certain type of orbit is more easily estimated compared with other orbits based on angles-only measurements.
-Furthermore, as they are defined in terms of the state vector at a specific time, they do not capture the effects of the measurement trajectory over a certain time period on the observability, which often predicts the performance of estimation.
-
-Motivated by the above issues, we propose a quantitative measure that represents the degree of observability for relative orbits in this section.
-More explicitly, this measure evaluates the observability in a quantitative manner as it determines how effectively the relative orbit can be estimated by line-of-sight measurements.
-
-The problem of assigning physically meaningful measures of the quality of observability has been introduced in~\cite{BroPNEC66,MonITAC67,MulWebA72} for linear systems, and in particular, in \cite{MulWebA72} the determinant, trace, and maximal eigenvalue of the inverse characteristic observability matrix are maximized with respect to certain parameters of a dynamic system.
+However, it does not provide any information about how strong the observability is.
+
 Several approaches have been proposed for an observability measure of nonlinear systems. 
@@
 Therefore, the observability Gramian defined in \refeqn{Wo_NL} measures how much the sensitivities of the output with respect to the initial condition are linearly independent of each other.
-This represents the degree of observability, since for example, if $\deriv{\y}{\x_{0_i}}$ is linearly dependent on $\deriv{\y}{\x_{0_j}}$, then it is difficult to distinguish the effects on $\delta \x_{0_i}$ to the output $\y$ from the effects of $\delta \x_{0_j}$ on $\y$.
+This represents the degree of observability, since for example, if $\deriv{\y}{\x_{0_i}}$ is linearly dependent to $\deriv{\y}{\x_{0_j}}$, then it is difficult to distinguish the effects on $\delta \x_{0_i}$ to the output $\y$ from the effects of $\delta \x_{0_j}$ on $\y$.
 Therefore, the condition number of $\mathcal{W} (\x_0,t_0,t_f)$ can be used as a measure of observability.
@@
 The positive integer $N_f$ is chosen such that $t_{N_f}=t_f$.
-
-In contrast to the observability criteria presented in Section 3, the proposed observability measure is obtained numerically.
-As such, it is not useful to identify the particular class of the relative orbits that yield stronger, or weaker observability.
-However, it can be used to analyze the degree of observability in a complementary manner when the sufficient conditions of the observability criteria are violated, as illustrated by the following numerical examples.
 
@@
 \end{align*}
-Therefore, the relative orbits are periodic.
-The inclination of the deputy is varied as $i_{deputy}=0,\,10,\,\cdots,\,50^\circ$, where the $i_{deputy}=0$ case corresponds to coplanar orbits such that the third condition of \refeqn{cond1} and \refeqn{cond2} is violated.
+Therefore, the relative orbits are periodic. The inclination of the deputy is varied as $i_{deputy}=0,\,10,\,\cdots,\,50^\circ$.
 
@@
 The initial covariance of the state is chosen as $P_0=\mathrm{diag}[(50\,\mathrm{km})^2I_{3\times 3}, (50n\,\mathrm{km/s})^2I_{3\times 3}]$.
-The covariance matrices for the process noise and the measurement noise are $Q_k=\mathrm{diag}[10^{-6}I_{3\times 3}, 10^{-8}I_{3\times 3}]$, and \\$R_k=1.306^2 I_{3\times 3}\,\mathrm{deg}^2$.
+The covariance matrices for the process noise and the measurement noise are $Q_k=\mathrm{diag}[10^{-6}I_{3\times 3}, 10^{-8}I_{3\times 3}]$, and $R_k=1.306^2 I_{3\times 3}\,\mathrm{deg}^2$.
 It is assumed that the line-of-sight is measured at every $\Delta t = 0.47$ seconds.
@@
 
-%\clearpage\newpage%asdf
+\clearpage\newpage
 \begin{figure}
@@
 	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
-		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_i_0.pdf}}
-%	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.3\textwidth]{OT_itraj_PE_i_0.pdf}}
+	\hspace*{0.15\textwidth}
 		\subfigure[Orbital trajectory in the LVLH frame]{
-		\includegraphics[width=0.5\textwidth]{OT_traj_PE_i_0.pdf}}
+		\includegraphics[width=0.3\textwidth]{OT_traj_PE_i_0}}
 }
@@
 	\subfigure[Estimation errors]{
-		\includegraphics[width=0.5\textwidth]{PE_i_0_e.pdf}}
-%	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.3\textwidth]{PE_i_0_e.pdf}}
+	\hspace*{0.15\textwidth}
 		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.5\textwidth]{PE_i_0_xx_norm.pdf}}
+		\includegraphics[width=0.3\textwidth]{PE_i_0_xx_norm.pdf}}
 }
@@
 	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
-		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_i_30.pdf}}
-%	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.3\textwidth]{OT_itraj_PE_i_30.pdf}}
+	\hspace*{0.15\textwidth}
 		\subfigure[Orbital trajectory in the LVLH frame]{
-		\includegraphics[width=0.5\textwidth]{OT_traj_PE_i_30.pdf}}
+		\includegraphics[width=0.25\textwidth]{OT_traj_PE_i_30.pdf}}
 }
@@
 	\subfigure[Estimation errors]{
-		\includegraphics[width=0.5\textwidth]{PE_i_30_e.pdf}}
-%	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.3\textwidth]{PE_i_30_e.pdf}}
+	\hspace*{0.15\textwidth}
 		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.5\textwidth]{PE_i_30_xx_norm.pdf}}
+		\includegraphics[width=0.3\textwidth]{PE_i_30_xx_norm.pdf}}
 }
@@
 \end{figure}
-
 \begin{figure}
@@
 	\subfigure[Estimation errors]{
-		\includegraphics[width=0.5\textwidth]{PE_i_30_e_J2.pdf}}
-%	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.3\textwidth]{PE_i_30_e_J2.pdf}}
+	\hspace*{0.15\textwidth}
 		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.5\textwidth]{PE_i_30_xx_norm_J2.pdf}}
+		\includegraphics[width=0.3\textwidth]{PE_i_30_xx_norm_J2.pdf}}
 }
@@
 	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
-		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_i_50.pdf}}
-%	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.3\textwidth]{OT_itraj_PE_i_50.pdf}}
+	\hspace*{0.15\textwidth}
 		\subfigure[Orbital trajectory in the LVLH frame]{
-		\includegraphics[width=0.5\textwidth]{OT_traj_PE_i_50.pdf}}
+		\includegraphics[width=0.25\textwidth]{OT_traj_PE_i_50.pdf}}
 }
@@
 	\subfigure[Estimation errors]{
-		\includegraphics[width=0.5\textwidth]{PE_i_50_e.pdf}}
-%	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.3\textwidth]{PE_i_50_e.pdf}}
+	\hspace*{0.15\textwidth}
 		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.5\textwidth]{PE_i_50_xx_norm.pdf}}
+		\includegraphics[width=0.3\textwidth]{PE_i_50_xx_norm.pdf}}
 }
@@
 
-%\clearpage\newpage%asdf
+\clearpage\newpage
 
@@
 For the linearized relative orbital dynamics of the two-body problem, these initial conditions yield elliptic orbits around the chief, but for the nonlinear relative orbital dynamics, there is drift proportional to the size of the orbit parameterized by $r_{x_0}$.
-While the resulting orbits are coplanar such that the third condition of \refeqn{cond1} and \refeqn{cond2} is violated, it is still desirable to understand how the orbital size or drift affects nonlinear observability.
+While the resulting orbits are planar such that the third condition of \refeqn{cond1} and \refeqn{cond2} is violated, it is still desirable to understand how the orbital size or drift affects nonlinear observability.
 For the extended Kalman filter, the initial estimate of the state is chosen as $\hat\x_0=1.5\x_0$, and the other parameters are identical to the prior case.
-
-As discussed in Section 3, these relative orbits violate the observability criteria summarized in Proposition 1 since they are coplanar orbits.
-Therefore, Proposition 1 cannot be applied to investigate the observability of these orbits.
-Instead, the observability is analyzed by the observability measure proposed in Section 4 as follows.
 
@@
 	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
-		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_r0_JAS_20}}
-%	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.3\textwidth]{OT_itraj_PE_r0_JAS_20}}
+	\hspace*{0.15\textwidth}
 		\subfigure[Orbital trajectory in the LVLH frame]{
-		\includegraphics[width=0.5\textwidth]{OT_traj_PE_r0_JAS_20}}
+		\includegraphics[width=0.25\textwidth]{OT_traj_PE_r0_JAS_20}}
 }
@@
 	\subfigure[Estimation errors]{
-		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_20_e.pdf}}
-%	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_20_e.pdf}}
+	\hspace*{0.15\textwidth}
 		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_20_xx_norm.pdf}}
+		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_20_xx_norm.pdf}}
 }
@@
 	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
-		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_r0_JAS_50}}
-%	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.3\textwidth]{OT_itraj_PE_r0_JAS_50}}
+	\hspace*{0.15\textwidth}
 		\subfigure[Orbital trajectory in the LVLH frame]{
-		\includegraphics[width=0.5\textwidth]{OT_traj_PE_r0_JAS_50}}
+		\includegraphics[width=0.25\textwidth]{OT_traj_PE_r0_JAS_50}}
 }
@@
 	\subfigure[Estimation errors]{
-		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_50_e.pdf}}
-%	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_50_e.pdf}}
+	\hspace*{0.15\textwidth}
 		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_50_xx_norm.pdf}}
+		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_50_xx_norm.pdf}}
 }
@@
 	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
-		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_r0_JAS_100}}
-%	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.3\textwidth]{OT_itraj_PE_r0_JAS_100}}
+	\hspace*{0.15\textwidth}
 		\subfigure[Orbital trajectory in the LVLH frame]{
-		\includegraphics[width=0.5\textwidth]{OT_traj_PE_r0_JAS_100}}
+		\includegraphics[width=0.25\textwidth]{OT_traj_PE_r0_JAS_100}}
 }
@@
 	\subfigure[Estimation errors]{
-		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_100_e.pdf}}
-%	\hspace*{0.15\textwidth}
+		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_100_e.pdf}}
+	\hspace*{0.15\textwidth}
 		\subfigure[Magnitude of state (red:true, blue:estimate)]{
-		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_100_xx_norm.pdf}}
+		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_100_xx_norm.pdf}}
 }
@@
 It is illustrated by several numerical examples.
-Future work objectives include constructing observability criteria based on higher-order Lie derivatives of the measurement, and developing necessary conditions for observability to classify the type of relative orbits that are not observable completely.
-We also plan to develop nonlinear filtering techniques that can capture the weak coupling between the dynamics of relative motion and the line-of-sight measurements effectively, as well as studying the effects of higher-order gravitational perturbations on the observability.
+Future works include constructing observability criteria based on higher-order Lie derivatives of the measurement, and developing nonlinear filtering techniques that can capture the weak coupling between the dynamics of relative motion and the line-of-sight measurements effectively, as well as studying the effects of higher-order gravitational perturbations on the observability.
 
@@
 %\bibliographystyle{spbasic}      % basic style, author-year citations
-\bibliographystyle{spmpsci}      % mathematics and physical sciences
+%\bibliographystyle{spmpsci}      % mathematics and physical sciences
 %\bibliographystyle{spphys}       % APS-like style for physics
-
 \bibliography{BibSources}   % name your BibTeX data base

### 2018-06-09 Journal/JINT 2017/JIRS16.tex
journal added

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{amsmath,amssymb,url,times}%,subfigure}% amsthm is the one!
\usepackage{caption,subcaption,hyperref}
\usepackage{color,comment}
\usepackage{curves,pgfgantt}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
 \journalname{Journal of Intelligent \& Robotic Systems}

\newtheorem{prop}{Proposition}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\abs}[1]{\ensuremath{\left| #1 \right|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\ip}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Figure \ref{fig:#1}}
\newcommand{\tr}[1]{\mbox{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mbox{tr}\ensuremath{\!\bracket{#1}}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathbf{d}}}
\newcommand{\pair}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\!\langle #1 \rangle\!\rangle}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\ad}{\ensuremath{\mathrm{ad}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\graphicspath{{./Fig/}}

\begin{document}

\title{Autonomous Exploration with Exact Inverse Sensor Models%Exact Occupancy Grid Mapping and Autonomous Exploration%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}

%\titlerunning{Short form of title}        % if too long for running head

\author{Evan Kaufman \and Kuya Takami \and \\ Taeyoung Lee \and Zhuming Ai}%
%\thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 {\tt \{evankaufman,tylee\}@gwu.edu}}
%\thanks{Zhuming Ai, Ira S. Moskowitz, Information Management \& Decision Architectures, U.S. Naval Research Laboratory,  Washington, DC 20375}
%}

%\authorrunning{Short form of author list} % if too long for running head

% Expanded:
\institute{E. Kaufman \at
              Dept. of Mechanical and Aerospace Engineering \\
		 The George Washington University \\
		 Washington DC 20052 \\
              Tel.: 216-346-0946\\
              \email{evankaufman@gwu.edu} 
           \and
           K. Takami \at
              Dept. of Mechanical and Aerospace Engineering \\
		 The George Washington University \\
		 Washington DC 20052 \\
              Tel.: 414-212-5155\\
              \email{kuya@gwu.edu} 
           \and
           T. Lee \at
              Dept. of Mechanical and Aerospace Engineering \\
		 The George Washington University \\
		 Washington DC 20052 \\
              Tel.: 202-994-8710\\
		 Fax: 202-994-0238\\
              \email{tylee@gwu.edu}
           \and
           Z. Ai \at
              Information Management \& Decision Architectures
              \\ U.S. Naval Research Laboratory \\
              Washington, DC 20375
}

% Condensed:
\institute{E. Kaufman, K. Takami, and T. Lee \at
              Dept. of Mechanical and Aerospace Engineering \\
		 The George Washington University \\
		 Washington DC 20052 \\
              Tel.: 202-994-8710\\
              \email{\{evankaufman,kuya,tylee\}@gwu.edu} 
           \and
           Z. Ai \at
              Information Management \& Decision Architectures
              \\ U.S. Naval Research Laboratory \\
              Washington, DC 20375
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
This paper is focused on probabilistic occupancy grid mapping and motion planning such that a robot may build a map and explore a target area autonomously in real time. The desired path of the robot is developed in an optimal fashion to maximize the information gain from the sensor measurements on its path, thereby increasing the accuracy and efficiency of mapping, while explicitly considering the sensor limitations such as the maximum sensing range and viewing angle. 
Most current exploration techniques require frequent human intervention, often developed for omnidirectional sensors with infinite range.
The proposed research is based on realistic assumptions on sensor capabilities. The unique contribution is that the mapping and autonomous exploration techniques are systematically developed in a rigorous, probabilistic formulation. The mapping approach exploits the probabilistic properties of the sensor and map explicitly, and the autonomous exploration is designed to maximize the expected map information gain, thereby improving the efficiency of the mapping procedure and the quality of the map substantially. The efficacy of the proposed optimal approach is illustrated by both numerical simulations and experimental results. 
%%%% Before minor modifications restructuring the abstract by Evan Kaufman on October 14, 2016:
%This paper is focused on probabilistic occupancy grid mapping and motion planning such that a robot may build a map and explore a target area autonomously in real time. The desired path of the robot is developed in an optimal fashion to maximize the information gain from the sensor measurements on its path, thereby increasing the accuracy and efficiency of mapping, while explicitly considering the sensor limitations such as the maximum sensing range and viewing angle. Most current exploration techniques require frequent human intervention, and they are frequently developed for omnidirectional sensors with infinite range. 
%%For example, when applying SLAM techniques to explore an unknown area, a human operator must provide the path of the vehicle while continuously monitoring the updated map. 
%The proposed research will greatly improve autonomy of robotic exploration in a systematic optimal manner, under realistic assumptions on sensor capabilities. The unique contribution is that the mapping and autonomous exploration techniques are systematically developed in a rigorous, probabilistic formulation. The mapping approach exploits the probabilistic properties of the sensor and map explicitly, and the autonomous exploration is designed to maximize the expected map information gain, thereby improving the efficiency of the mapping procedure and the quality of the map substantially. The efficacy of the proposed approach is illustrated by both numerical simulations and experimental results. 
\keywords{Occupancy Grid Mapping \and Inverse Sensor Model \and Autonomous Exploration}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
Robotic mapping is the process of generating maps, which represent the environment in close proximity of a robot. 
%Mapping in the context of robotics is to generate maps, which represent the environment in close proximity of a robot. 
This problem is crucial to understanding the surrounding regions when a robot enters an uncertain environment.
Various map representations include occupancy grids~\cite{WolSuk05}, octomaps~\cite{WurHorBenStaBur10}, or feature-based maps%in a landmark representation
~\cite{MonThrKolWeg02}.
%To build maps in real time, a computationally-efficient algorithm is required.% \EditTL{rewrite this paragraph for mapping, without introducing SLAM here}

For mobile robots to operate in an unknown environment autonomously, they should be able to construct a map of the environment by using onboard sensors, and estimate their position in the constructed map. Simultaneous localization and mapping problem (SLAM) serves to estimate the map of the environment and the pose of robots concurrently, and there have been various approaches developed for SLAM with successful demonstrations.

However, there are certain issues with the current SLAM techniques. First, the vast majority of work deals only with the aspect of estimating the environment and the poses of vehicles. They are passive in the sense that SLAM is performed on incoming sensor measurements from vehicles following an arbitrary path. As such, human teleoperation and monitoring are often essential to guide the vehicles safely through unknown surroundings. Therefore, it is desired that an active motion planning scheme is developed and integrated with SLAM such that vehicles are able to determine their path without human intervention, and explore unknown areas autonomously.

To achieve these, it is proposed that the problems of mapping and motion planning are integrated in a probabilistic formulation. More explicitly, the map is constructed in a probabilistic sense, in contrast to the common approaches where the map is defined by deterministic objects that are definitely occupied or free. One of the desirable features for the proposed stochastic formulation is that we can extract the information on the degree of uncertainty for the each part of the map, such that the motion of the vehicle is planned to minimize the map uncertainty in an optimal fashion.

This is addressed in two steps. The first part of this paper focuses on probabilistic occupancy grid mapping given a stochastic model of a range sensor, and the second part deals with stochastic motion planning to construct optimal exploration strategies.

\paragraph{Probabilistic Occupancy Grid Mapping}

The proposed approach is based  on the occupancy grid representation, where the environment is decomposed into evenly-spaced grid cells, considered as either occupied or free. The probabilistic model of a range sensor is referred to as the \textit{forward sensor model}, which provides a probability of range measurements for the given pose of the vehicle and the exact map~\cite{ThrBurFox05}. One can characterize the performance of a range sensor, such as its accuracy or maximum range, completely with its forward sensor model.

The mapping problem can be defined as finding the \emph{inverse sensor model}, which is the probability of occupancy for a grid cell based on range measurements and the configuration of the mobile robot, which is of fundamental importance with occupancy grid mapping. However, the exact solution to the inverse sensor model has not been utilized in occupancy grid mapping, because the computational requirements are too cumbersome to compute the exact model in real time~\cite{ThrBurFox05}.

Instead, several techniques have been proposed to approximate the exact inverse sensor model. %Due to the computational limitations, two common approaches are applied to obtain a function close to the inverse sensor model.
In the early occupancy grid mapping applications based on sonar sensors, \emph{ad hoc} techniques to obtain an inverse sensor model were developed based off highly-innacurate bouncing sound waves~\cite{MorElf85,Elf89}.
%These occupancy grid mapping schemes employ \emph{ad hoc} techniques to obtain the inverse sensor model.
These techniques rely on approximate functions to represent the inverse sensor model with questionable accuracy~\cite{ChoLynHutKanBurKavThr05} and applied to more modern sensors in~\cite{And09},~\cite{PirRutBisSch11}.

The other popular approach to obtain an approximate inverse sensor model is to simulate maps, robot poses, and measurements and use a learning algorithm to approximate the inverse sensor model~\cite{ThrBurFox05,Thr01}. These approaches are undesirable in practice due to complexities associated with implementing a learning algorithm. For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples, or how many samples are required to obtain a reasonable approximation. Furthermore, it is challenging to apply any learning algorithm over the large dimensional space composed of maps, poses, and measurements.

%and they only yield probabilistic approximations without strong mathematical justification.

%These approaches simply treat occlusions with discrete probabilities of detection, regardless of whether one cell may completely, partially, or not block another cell in the sensor field of view.

%Furthermore, both approaches follow a log-odds ratio Bayesian framework, imposing an additional Markov assumption.

%These techniques are not also suitable for modern range sensor with the highly accurate bearing~\cite{Thr03}. Based on recent improvements in sensor technology, the occupancy grid mapping problem has been restructured using the fact that depth measurement bearings are known. \EditTL{do we need this paragraph?}

This paper proposes a computationally efficient algorithm to construct the exact inverse sensor model. More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, this algorithm yields the a posteriori probabilities of occupancy of all the cells within the area covered by the range sensor from the range measurements. The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancies of the other cells blocked by it is irrelevant to the forward sensor model, and this property is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model. Furthermore, the proposed approach integrates a priori probabilities of occupancy and multiple range measurements according to the Bayesian framework to obtain more accurate maps. As such, it contrasts from the existing framework based on log-odds ratios that impose an additional Markov assumption.

\paragraph{Optimal Autonomous Exploration}

Next, this paper addresses autonomous exploration to plan the trajectories of the vehicle to generate a complete, accurate map~\cite{Yam97}. We present an approach where the robot chooses a trajectory designed to maximize information gain of the occupancy grid map.

%Another major problem that this paper addresses is to determine an effective trajectory given little knowledge about the environment.  Autonomous exploration solves this problem, by choosing robotic motion through uncertain space, while building a map that is used for subsequent navigation~\cite{Yam97}.

A common approach to solve the autonomous exploration problem is known as frontier-based exploration, such as ~\cite{Yam97,Yam98}. This is the process of executing actions that move the robot toward the closest boundary between visited and unvisited space, known as a frontier.
Then the robot takes measurements at this location such that the mapped territory expands, and thus the new frontiers are pushed back. This process is repeated until the map is well-known.
Frontier-based exploration assumes that repeatedly moving toward the closest frontier and taking measurements are the best actions to gain new information about the map, though these systematic actions are not based on any consideration of the future uncertainty of the probabilistic map or optimality.
%there is no guarantee that these actions satisfy any optimality criteria.

%The uncertainty of an occupancy grid map is commonly measured with Shannon's entropy, which is a metric that measures map uncertainty.

%Approaches aimed at determining the future uncertainty of an occupancy grid map commonly suffer from several approximations.
%%Determining future occupancy grid map uncertainty requires predictive estimations, which suffer from several approximations. 
%Most evident is that all approaches rely on inaccurate probabilistic occupancy grids. This mapping representation is typically based on heuristic approximations (e.g.~\cite{MorElf85,Elf89,ChoLynHutKanBurKavThr05,And09,PirRutBisSch11}) or simulated learning techniques \cite{ThrBurFox05,Thr01}. Map-based information gain is commonly determined with Shannon's entropy~\cite{CarDamKumCas15}, an uncertainty metric based on grid cell occupancy probability. Since this probability is heuristic or learned, any measure of entropy is subject to errors from approximated probability.
%Furthermore, the expected entropy of an action is not directly calculated~\cite{CarDamKumCas15,StaGriBur05}. Instead, these particle filter-based approaches assume that expected entropy is equivalent to entropy based on expected measurement scans~\cite{JohStaPfaBur07}. This assumption is subject to inaccuracies due to the nonlinear nature of Shannon's entropy. %Furthermore, these approaches typically generate maps with a particle filter, which commonly require some tuning and may consume large amounts of onboard memory.
%
%In this paper, we propose an autonomous exploration technique based on expected entropy change that avoids the aforementioned issues. 
%%The occupancy grid mapping scheme is based on~\cite{KauLeeAiMos16}, which provides the exact solution of occupancy grid mapping of a single measurement ray. 
%We develop a solution for expected entropy gain based on exact occupancy grid mapping~\cite{KauLeeAiMos16}. This solution directly calculates expected entropy from a probabilistic map based on similar assumptions of those applied to occupancy grid mapping.
%Though more accurate than several other approaches, the proposed solution requires large computational resources, making real-time implementation difficult in certain scenarios.
%Thus, we provide the tools to approximate the exact solution that reduce the computational requirements with realistic assumptions.
%The motion of the robot is formulated as an optimization problem, where the expected entropy change is minimized, or equivalently the expected information gain is maximized.

Alternatively, there have been exploration techniques that are based on a quantitative measure of uncertainty. However, these approaches commonly suffer from several approximations. Most evident is that all approaches rely on inaccurate probabilistic occupancy grids. 
This mapping representation is typically based on heuristic approximations (e.g.~\cite{MorElf85,Elf89,ChoLynHutKanBurKavThr05,And09,PirRutBisSch11}) or simulated learning techniques \cite{ThrBurFox05,Thr01}. Map-based information gain is commonly determined with Shannon's entropy~\cite{StaGriBur05}, an uncertainty metric based on grid cell occupancy probability. Since this probability is heuristic or learned, any measure of entropy is subject to errors from approximated probability or ad hoc learning.
Furthermore, the expected entropy of an action is not directly calculated. Instead, these particle filter-based approaches assume that expected entropy is equivalent to entropy based on expected measurement scans~\cite{JohStaPfaBur07}. This assumption is subject to inaccuracies due to the nonlinear nature of Shannon's entropy. For these reasons, more careful computation of map uncertainty is warranted. %Furthermore, these approaches typically generate maps with a particle filter, which commonly require some tuning and may consume large amounts of onboard memory.

In this paper, we propose an autonomous exploration technique based on expected entropy change that avoids the aforementioned issues. 
%The occupancy grid mapping scheme is based on~\cite{KauLeeAiMos16}, which provides the exact solution of occupancy grid mapping of a single measurement ray. 
We develop a solution for expected entropy gain based on exact occupancy grid mapping~\cite{KauLeeAiMos16}. This solution directly calculates expected entropy from a probabilistic map based on similar assumptions of those applied to occupancy grid mapping.
Though more accurate than several other approaches, the proposed solution requires large computational resources, making real-time implementation difficult in certain scenarios.
Thus, we provide the tools to approximate the exact solution that reduces the computational requirements with realistic assumptions.
The motion of the robot is formulated as an optimization problem, where the expected entropy change is minimized, or equivalently the expected information gain is maximized.
Compared with the recent results (e.g. [12], [13], [16], [17]) that employ Rao-Blackwellized particle filters to deal with localization and autonomous exploration simultaneously with an approximate entropy change, this paper is based on an accurate probabilistic solution to map information gain, under the assumption that the robot pose is known.

In short, the main contribution of this paper is proposing the exact inverse sensor model and constructing an exact Bayesian occupancy grid mapping based on that. Compared with the current mapping algorithms based on an approximate inverse sensor model, this approach yields more accurate occupancy probabilities. This constructs substantially more accurate occupancy grid maps with less uncertainty for the same set of measurements, and these are directly illustrated by numerical examples and experimental results. Additionally, we derive a computationally-efficient approach to predict map information gain for autonomous exploration with a novel structure that avoids several common assumptions and approximations. By computing the expected information precisely with the exact inverse sensor model, the proposed autonomous exploration approach constructs a more accurate map with limited sensor capabilities in an optimal fashion. While this paper is focused on mapping and exploration in two-dimensional environments, the presented results can be certainly utilized in SLAM or stochastic motion planning in three dimensions.

\section{Problem Formulation}
\label{sec:ProbForm}

\subsection{Probabilistic Occupancy Grid Mapping}
\label{subsec:POGM}
Let a map $m$ be decomposed into $n_m$ evenly-spaced grid cells, where the $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,\ldots,n_m}$, that is defined as $\mathbf{m}_i=1$ when occupied, and $\mathbf{m}_i=0$ when free. The location and size of each grid cell is assumed known. Therefore, a map $m$ is defined by $\{\mathbf{m}_1,\ldots, \mathbf{m}_{n_m}\}$ ($2^{n_{m}}$ possible maps).

Another random variable is defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience. The probability that the $i$-th cell is occupied is $P(\mathbf{m}_i)$, and the probability that it is free is $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$. The random variables $\mathbf{m}_i$ are mutually independent, 
\begin{align}
P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i).
\end{align}

Consider a range sensor that provides scans of the surrounding environment in order to identify the closest occupied space. The location and the direction of the sensor, referred to as the \emph{pose} at time $t$, is denoted by $X_t=(x_t,R_t)$, where the planar position and direction of the robot at $t$ are denoted by $x_t\in\Re^2$ and $R_t\in\Sph^1=\braces{q\in\Re^2|\norm{q}=1}$, i.e., the attitude corresponds to a direction on a 2D plane.  Let $X_{1:t}$ denote the history of poses from the initial time to the current time, i.e., $X_{1:t}=\{X_1,X_2,\ldots, X_t\}$. At each pose, the robot receives a 2D measurement \emph{scan}, composed of $n_z$ measurement \emph{rays}. These rays are 1D depth measurements of known direction from the current pose to the closest occupied space, subject to a known \emph{forward sensor model} $p(z_{t,l}|m,X_t)$, where $z_{t,l}$ is the $l$-th measurement ray of the $t$-th scan $Z_t=\braces{z_{t,1},z_{t,2},\ldots,z_{t,n_z}}$. The measurement history is denoted by $Z_{1:t}=\braces{Z_1,Z_2,\ldots,Z_t}$.

The probability density function, namely $p(z_{t,l}|m,X_t)$ with respect to the depth of the $l$-th measurement ray conditioned on the map $m$ and the pose $X_t$ is commonly referred to as the \emph{forward sensor model}, which characterizes the corresponding depth sensor, such as the maximum range or accuracy. The forward sensor model satisfies  (i) the ranges of all depth measurements are positive and finite, and (ii) measurement rays cannot pass through occupied regions. Throughout this paper, we assume that the forward sensor model of the selected sensor is given. This can be determined empirically or analytically. For example, the \emph{beam model for range finders} satisfying  the above criteria is described in~\cite{ThrBurFox05}.

Occupancy grid mapping provides occupancy probabilities based on robot poses and measurement scans. The goal is to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, commonly referred to as the \emph{inverse sensor model} for given forward sensor model $p(z_{t,l}|m,X_t)$ and the initial estimate of the map $P(m)$ with $X_{1:t}$ and $Z_{1:t}$.

\subsection{Autonomous Exploration}
\label{subsec:AE}
Since occupancy grid mapping provides a probabilistic representation of surrounding space, this mapping scheme holds probabilistic information about the uncertainty of the map. Shannon's entropy is commonly used as a measure of uncertainty, such as in~\cite{StaGriBur05}. Given the probabilities of each grid cell of map $m$, Shannon's entropy is defined as the sum of the entropy of each cell,
\begin{align}
\label{eqn:ShannonsEntropyDef}
H(&P(m))=%\nonumber\\&
-\sum_{i=1}^{n_m}\braces{
P(\mathbf{m}_i)\log{P(\mathbf{m}_i})+P(\bar{\mathbf{m}}_i)\log{(P(\bar{\mathbf{m}}_i}))}.
\end{align}
The entropy of the $i$-th grid cell is maximized when $P(\mathbf{m}_i)=0.5$ (greatest uncertainty) and minimized when $P(\mathbf{m}_i)\in\braces{0,1}$ (smallest uncertainty). %; thus Shannon's entropy is a measure of uncertainty.

The goal of autonomous exploration is to determine future poses that maximize the map information gain, or equivalently minimize the map entropy. 
Since autonomous exploration is conducted in uncertain environments, a complete trajectory cannot be known before the robot takes measurements of the entire reachable space.
Instead, an objective function governs the motion between intermediate poses, building an occupancy grid map along the way.

Suppose at time $t$, $P(m|X_{1:t},Z_{1:t})$ is given. Let $X_c$ be a candidate pose under consideration for some future time step.
The expected information gain $\mathcal I(X_c)$ is defined as the negative change of entropy
\begin{align}
\label{eqn:ObjFun}
\mathcal I(X_c)&=H(P(m|X_{1:t},Z_{1:t}))%\nonumber\\&\quad
-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right],
\end{align}
where $H(P(m|X_{1:t},Z_{1:t}))$ may be computed with \refeqn{ShannonsEntropyDef} and \\$\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right]$ is the expected entropy when the robot is located at $X_c$, where the expectation is taken over $Z_c$.
Since autonomous exploration is formulated as a trajectory optimization, the goal is to repeatedly choose optimal $X_c^*$ that maximizes the objective function
\begin{align}
\label{eqn:Objective}
X_c^*=\argmax_{X_c}{\mathcal I(X_c)},
\end{align}
where $X_c^*$ satisfies the inequality constraint to avoid collisions,
\begin{align}
\label{eqn:CollisionInequalityConstraint}
P_\text{collision}(X)=1-\prod_{i\in\mathcal C_X}P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})\leq\beta,
\end{align}
where $\mathcal C_X$ is the set of grid cells falling inside a volume of preselected size around $X$ that may cause collision and $\beta>0$ is a threshold of the probability of collision.
Once the robot has translated to $X_c^*$, the above process is repeated.

\section{Occupancy Grid Mapping with an Exact Inverse Sensor Model}
\label{sec:ISM}

This section outlines issues with the common log-odds ratio formulation and perceived computational burdens with the exact solution to the inverse sensor model for occupancy grid mapping. Then, a method that exploits several mapping and probabilistic properties provides a computationally efficient exact solution to probabilistic occupancy grid mapping.

\subsection{Mapping via Log-Odds Ratio}

One of the popular frameworks of updating binary random variables with static state is with a binary Bayesian filter using the log-odds ratio formulation.
The main idea is that instead of multiplying terms from prior and current time steps, the properties of logarithms allow these terms to be simply added, while avoiding truncation issues associated with probabilities close to $0$ or $1$~\cite{ThrBurFox05}.
The log-odds ratio is also popular because learning techniques with forward models~\cite{Thr01},~\cite{Thr03} or ad hoc techniques~\cite{MorElf85},~\cite{Elf89} become simplified.%are simpler to derive when they need not consider $P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})$ as part of the inverse sensor model, but rather only $P(\mathbf{m}_i)$, which is typically fixed and uniform for all cells.

However, the log-odds ratio formulation makes an assumption that is not consistent with the occupancy grid mapping problem. It is assumed that if $X_t$ and $\mathbf{m}_i$ are known, then $Z_t$ is independent of $X_{1:t-1}$ and $Z_{1:t-1}$, i.e.,
%\begin{align}
%\label{eqn:FirstBayesRule}
%P(\mathbf{m}_i|X_{1:t},Z_{1:t})
%&
%=\frac{P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})}{P(Z_t|X_{1:t},Z_{1:t-1})}.
%\end{align}
\begin{align}
\label{eqn:AssumptionEarly}
P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})\approx P(Z_t|\mathbf{m}_i,X_t).
\end{align}
In other words, the past history of measurements and poses are irrelevant to the current measurement. However, the prior measurements and poses provide a certain information to the occupancy map, which affects the probabilistic model of the current measurement.

More specifically, forward models are based on the entire map $m$. Since \refeqn{AssumptionEarly} conditioned only by the $i$-th cell, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may contain important information about the occupancy of the other cells, which should be considered in calculating the probability distribution of the measurement scan $Z_t$. For example, $X_{1:t-1}$ and $Z_{1:t-1}$ might suggest that the $i$-th cell is occluded, so the approximation of \refeqn{AssumptionEarly} would not be justified. In short, valuable information about the map that can be obtained by the past measurements is ignored in the common log-odds ratio formulation. The probabilistic mapping proposed in this paper avoids this assumption of the log-odds ratio framework to construct a more accurate map by utilizing all of the past and the current measurements.

%In short, this log-odds ratio framework is based on the assumption that neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$.

%However, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may suggest that there exist occupied regions that occlude some or all of $Z_t$ from detecting $\mathbf{m}_i$ from the current pose $X_t$.
%Since this assumption neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$, this framework is subject to poor results.
%Hence, the log-odds ratio formulation is completely avoided in further analysis.

	% approximated inverse sensor model
	% markov assumption
	% shortcoming, reason this is avoided

%
%
%
%

\subsection{Inverse Sensor Model}

%Due to computational complexity, approximate inverse sensor models have been used for occupancy grid mapping with simplifying assumptions.

In this section, we propose an algorithm to compute the exact inverse sensor model efficiently. 
%Much like the forward sensor model that reasons measurements from a known map, he inverse sensor model reasons that map from known measurements~\cite{ThrBurFox05}.
%While the inverse sensor model varies slightly in meaning in numerous publications, two versions are derived in this paper, both based on a Bayesian framework.
%We present two versions according to a Bayesian framework. The first type is referred to as the \emph{ray inverse sensor model} $P(\mathbf{m}_i|X_{1:t},Z_{1:t-1},z_{t,l})$, which only considers $l$-th ray at time $t$, i.e., one-dimensional inverse sensor model. The second type is referred to as the \emph{scan inverse sensor model} $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, which combines the information from \emph{all} measurement rays \emph{simultaneously} at the $t$-th time step for two-dimensional cases. These are developed in a sequential form to be repeated at each time step.
%These are contrary to inverse sensor models commonly used with the log-odds ratio formulations, and the proposed algorithm integrates all of past poses and measurement scans with the current ones to obtain more accurate estimate of the map.
Suppose the probability of the map conditioned on the past poses and measurements, namely $P(m|X_{1:t-1},Z_{1:t-1})$, is known. Here we construct a posteriori probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$, based on the current pose $X_t$, the measurement from the $l$-th ray $z_{t,l}$, and the given forward sensor model $p(z_{t_l}|m,X_t)$.

\paragraph{Bayesian Framework}
%Assuming that the probability density of thxe sensor measurement is fixed over a small measurement region $\Delta z$, 
Bayes' rule yields
%First, Bayes' rule is applied for a particular map $m$ and arbitrarily small measurement region $\Delta z$, yielding
\begin{align}
\label{eqn:BayesRuleRayISM}
P(&m|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
=\frac{p(z_{t,l}|m,X_{1:t},Z_{1:t-1})P(m|X_{1:t-1},Z_{1:t-1})}{p(z_{t,l}|X_{1:t},Z_{1:t-1})},
\end{align}
considering that $X_t$ carries no information about $m$ without $Z_t$.
If the current pose $X_t$ and map $m$ are known, then the measurement ray $z_{t,l}$ is independent of past poses $X_{1:t-1}$ and past measurements $Z_{1:t-1}$ to obtain
%Note that this is contrary to approaches that consider $X_t$ and $\mathbf{m}_i$ (not the full map $m$) as enough information, ignoring the occupancy of other cells in this term.
%Then, \refeqn{BayesRuleRayISM} can be rewritten as
\begin{align}
P(m|z_{t,l},&X_{1:t},Z_{1:t-1})%\nonumber\\&
=\eta_{t,l}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}),
\end{align}
where the normalizing constant $\eta_{t,l}\in\Re$ absorbs all terms that do not depend on the map $m$.

%Assuming $z_{t,l}$ measures an occupied cell, then $z_{t,l}$ may be a result of the occupancy of $\mathbf{m}_i$, or $z_{t,l}$ may measure a different occupied cell.

Next, we compute the occupancy probability of each cell. Let $\mathcal{M}_i$ be the set of maps where the $i$-th cell is occupied, i.e., $\mathcal{M}_i =\{m\in\{0,1\}^{{n_m}}\,|\ \mathbf{m}_i=1\}$. To compute the probability of occupancy of the $i$-th cell, all possible combinations of map in $\mathcal{M}_i$ should be considered, i.e., 
%All possible combinations of map $m$ must be considered in which $\mathbf{m}_i$ is occupied, denoted as $m:m(i)=\mathbf{m}_i$,
\begin{align}
\label{eqn:InvSenModWithProbDens}
&P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
=\eta_{t,l}\sum_{m\in\mathcal{M}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
\end{align}
Furthermore, to determine the normalizing constant $\eta_{t,l}$, the complement $\\P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in the similar manner. Since there are $2^{n_{m}-1}$ maps in $\mathcal{M}_i$, these equations require $2^{n_m}$ terms to compute the summation for the $i$-th grid cell and normalizer, and this process should be repeated for all other cells along the measurement ray. This is the main reason why the existing results are based on certain approximations of the above expression.

\paragraph{Computationally Efficient Approach}

We propose a computational algorithm to evaluate \refeqn{InvSenModWithProbDens} efficiently. 
Since the cells outside of the sensor field of view (FOV) are not affected, we focus on a reduced map $r_l$ in the FOV of the $l$-th ray. This reduced map is chosen such that each cell of $r_l$ corresponds to a grid cell of $m$ that the $l$-th ray intersects, ordered by increasing distance, which is easily determined from geometry. Let $\mathbf{r}_{l,k}$ be the binary random variable representing the occupancy of the $k$-th cell of the $l$-th ray. The number of cells in the reduced map is denoted by $n_{r,l}\leq n_m$.

Next, let $\mathbf{r}_{l,k+}$ correspond to the event that the $k$-th cell of the $l$-th ray is occupied, cells with lower index are free, and cells with greater index may or may not be occupied, i.e., event $\mathbf{r}_{l,k+}$ occurs when \\$r\in\braces{r\in\{0,1\}^{{n_{r,l}}}\,|\mathbf{r}_{l,1}=0,\mathbf{r}_{l,2}=0,\ldots,\mathbf{r}_{l,k-1}=0,\mathbf{r}_{l,k}=1}$. In other words, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray, and this concept is illustrated with a simple 1D example in Figure \ref{fig:show_rkplus}.
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{rkplus_1.png}
    \centering
  \includegraphics[width=0.4\textwidth]{rkplus_2.png}  
  \centering
  \includegraphics[width=0.4\textwidth]{rkplus_3.png}
    \centering
  \includegraphics[width=0.4\textwidth]{rkplus_4.png}
  \caption{A robot located to the left of the 1D occupancy grid map $r_l$ composed of $n_{r,l}=6$ grid cells in $4$ cases. In each case, cells $1$--$3$ are free, cell $4$ is occupied, and cells $5$ and $6$ may or may not be occupied. Each case corresponds to the event $\mathbf{r}_{4+}$.}
  \label{fig:show_rkplus}
\end{figure}
Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell. Based on this, we present the following computational algorithm to compute the inverse sensor model.

\begin{prop}
For the $l$-th measurement ray, the a posteriori probability of the occupancy of the $k$-th cell, namely the ray inverse sensor model, is given by
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}),
\end{align}
where the unnormalized probability of the inverse sensor model is defined as
\begin{align}
\label{eqn:Unnormalized}
& \tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
=P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\
&\quad\times 
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
%&\quad\times 
p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]\nonumber\\
&\quad + \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
%&\quad\times 
p(z_{t,l}|\mathbf{r}_{l,k+},X_t)P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1}),
\end{align}
where $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ is chosen for convenience and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading. The normalizer $\eta_{t,l}$ is given by
\begin{align}
\label{eqn:allEta}
\eta_{t,l}
&=
\bigg[\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
\end{align}
and it is independent of the cell index $k$.
\end{prop}
\begin{proof}
See Appendix.
\end{proof}

%
%%respectively.
%These have the follow solutions that may be obtained in a computationally-efficient manner:
%\begin{align}
%\label{eqn:tildePresult}
%\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
%\nonumber
%\\
%&\quad
%+
%\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{k+},X_t)P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1}),
%\\
%\label{eqn:tildePcomplementresult}
%\tilde P(\bar{\mathbf{r}}_k|z_{t,l},&X_{1:t},Z_{1:t-1})
%=P(\bar{\mathbf{r}}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
%\nonumber
%\\
%&\quad
%+
%\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}).
%\end{align}

%
%
%Because $\mathbf{r}_k$ is binary, the following holds:
%\begin{align}
%\label{eqn:allEta}
%\eta_{t,l}&=\frac1{\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})+\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})}\nonumber
%\\
%&=
%\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
%\end{align}
%where $\eta_{t,l}$ is independent of the index $k$ of the reduced map.
%Rearranging \refeqn{Unnormalized} and \refeqn{UnnormalizedCompl}, the ray inverse sensor model is 
%\begin{align}
%\label{eqn:RayISMAnswer}
%P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
%\\
%P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
%\end{align}
%respectively, where $\eta_{t,l}$ comes from \refeqn{allEta} and the unnormalized probabilities come from \refeqn{tildePresult} and \refeqn{tildePcomplementresult}.

Note that the a priori estimate, $P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ and its compliment $\\P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})=1-P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ are available at the $t$-th step. Then, \refeqn{RayISMAnswer} yields a sequential occupancy grid mapping that can be applied whenever new measurements are available.

Compared with \refeqn{InvSenModWithProbDens}, where the terms of the summation should be repeated $2^{n_{r,l}}$ times \emph{per each cell of the reduced map}, the proposed expressions \refeqn{RayISMAnswer} and \refeqn{allEta} are \textit{substantially} simpler. In fact, if \refeqn{Unnormalized} and \refeqn{allEta} are obtained recursively, requiring the summation of $O(n_{r,l}+1)$ rather than $O(n_{r,l}\times2^{n_{r,l}})$ as previously thought for \emph{all cells of the reduced map}, this yields an algorithm that is $\mathbf{n_{r,l}\times2^{n_{r,l}}/(n_{r,l}+1)}$ \textbf{times faster}.

\paragraph{Scan Inverse Sensor Model}
The above one-dimensional ray inverse sensor model may be generalized into a two-dimensional scan inverse sensor model by updating the map sequentially for each ray, according to
%The most natural way to generalize the ray inverse sensor model into a scan inverse sensor model is applying the conditions in a recursive way,
\begin{align}
\label{eqn:RayByRayScanISM}
P&(\mathbf{m}_i|X_{1:t},Z_{1:t})%\nonumber\\&%=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1},z_{t,2},\ldots,z_{t,n_z})\nonumber\\&
=P((\dots(((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z}).
\end{align}
In short, this approach updates the map occupancies along the first measurement ray, then repeats this process for each subsequent ray until all rays composing the measurement scan have updated the map.

\subsection{Summary of Algorithm}

In short, the proposed scan inverse sensor model is computed by \refeqn{RayISMAnswer}-\refeqn{RayByRayScanISM}. These are summarized in Table \ref{tab:RayByRayISM} as a computationally efficient, recursive algorithm that avoids several repeated calculations of the same quantity. This algorithm utilizes temporary variables, where the required computational resources are minimized by computing the updated occupancy probabilities of all grid cells along a measurement ray together.

%% More detailed (completed subscripts)
%\begin{table}[p]
%\begin{tabular}{ l }
%  Function: $P(\mathbf{r}_k|X_{1:t},Z_{1:t-1},z_{t,1:l})=\text{InverseSensorModel}(P(\mathbf{r}_k|X_{1:t},Z_{1:t-1},z_{t,1:l-1}),z_{t,l})$\\
%  for all $k\in\braces{1,2,\ldots,n_{r,l}}$\\
%  Note: all probabilities are conditioned on $X_{1:t},Z_{1:t-1},z_{t,1:l-1}$, so these are removed in probabilities.\\
%  Initialize $\eta^{-1}_{t,l}=0$ and $P(\bar{\mathbf{r}}_{0})=P(\bar{\mathbf{r}}_{n_{r,l}+1})=1$;\\
%  for all $k = 1,2,\ldots,n_{r,l}$\\
%   \ \ $P(\mathbf{r}_{k+}) = P(\bar{\mathbf{r}}_{0:k-1})P(\mathbf{r}_{k})$;\\
%   \ \ $P(\bar{\mathbf{r}}_{0:k})=P(\bar{\mathbf{r}}_{0:k-1})P(\bar{\mathbf{r}}_{k})$;\\
%   \ \ $a_\text{temp}=P(\mathbf{r}_{k+})p(z_l|\mathbf{r}_{k+},X_t)$;\\
%   \ \ $\tilde P(\mathbf{r}_{k}|z_{t,l})=P(\mathbf{r}_{k})\eta^{-1}_{t,l}+a_\text{temp}$;\\
%   \ \ $\eta^{-1}_{t,l}=\eta^{-1}_{t,l}+a_\text{temp}$;\\
%  end for\\
%  $\eta^{-1}_{t,l}=\eta^{-1}_{t,l}+P(\bar{\mathbf{r}}_{0:k})p(z_\text{max})$;\\
%  $P(\mathbf{r}_{k}|z_{t,l})=\eta_{t,l}\tilde P(\mathbf{r}_{k}|z_{t,l})$ for all $k\in\braces{1,2,\ldots,n_{r,l}}$;\\
%  \end{tabular}
%\caption{Inverse Sensor Model of an Individual Measurement Ray}
%\label{tab:RayByRayISM}
%\end{table}

% More simple (subscripts omitted)
\begin{table}[p]
\begin{tabular}{ l }
  Function: $P(\mathbf{r}_k|z)=\text{InverseSensorModel}(P(\mathbf{r}_k),z)$\\
  for all $k\in\braces{1,2,\ldots,n_{r}}$\\
  Initialize $\eta^{-1}=0$ and $P(\bar{\mathbf{r}}_{0})=1$;\\
  for all $k = 1,2,\ldots,n_{r}$\\
   \ \ $P(\mathbf{r}_{k+}) = P(\bar{\mathbf{r}}_{0:k-1})P(\mathbf{r}_{k})$;\\
   \ \ $P(\bar{\mathbf{r}}_{0:k})=P(\bar{\mathbf{r}}_{0:k-1})P(\bar{\mathbf{r}}_{k})$;\\
   \ \ $a_\text{temp}=P(\mathbf{r}_{k+})p(z_l|\mathbf{r}_{k+},X)$;\\
   \ \ $\tilde P(\mathbf{r}_{k}|z)=P(\mathbf{r}_{k})\eta^{-1}_{t,l}+a_\text{temp}$;\\
   \ \ $\eta^{-1}=\eta^{-1}+a_\text{temp}$;\\
  end for\\
  $\eta^{-1}=\eta^{-1}+P(\bar{\mathbf{r}}_{0:k})p(z_\text{max})$;\\
  $P(\mathbf{r}_{k}|z)=\eta\tilde P(\mathbf{r}_{k}|z)$ for all $k\in\braces{1,2,\ldots,n_{r}}$;\\
  \end{tabular}
\caption{Inverse Sensor Model of an Individual Measurement Ray}
\raggedright{Note: this function is only considering $l$-th ray at the $t$-th time step, where probabilities are subject to conditions on $X_{1:t},Z_{1:t-1},z_{t,1:l-1}$, so these are removed from the algorithm for simplicity.}
\label{tab:RayByRayISM}
\end{table}

In contrast to the current approximate inverse sensor models, the proposed algorithm evaluates the exact inverse sensor model \refeqn{InvSenModWithProbDens} efficiently without relying on assumptions like \refeqn{AssumptionEarly}. Therefore, it yields substantially more accurate maps for the same set of measurements. These are illustrated by numerical examples and experiments in Sections \ref{sec:NumRes} and \ref{sec:ExpRes}.

\section{Computation of Expected Information Gain}
\label{sec:ExpectedInfoGain}

In this section, we present a method to determine the expected entropy from an individual measurement ray from a known pose. Based on this, an autonomous exploration algorithm will be developed in the next section.

%This method is based on the exact solution to occupancy grid mapping from~\cite{KauLeeAiMos16}.

\subsection{Single Ray Expected Value of Entropy}
Suppose there is a probabilistic map constructed by past measurements, namely \\$P(m|X_{1:t},Z_{1:t})$. We wish to compute the expected entropy when the next measurement is taken at the candidate pose $X_c=(x_c,R_c)$, consisting of location $x_c\in\Re^2$ and attitude $R_c\in\Sph^1$. Assume that the direction of the $l$-th ray is determined by the attitude $R_c$.
The expected entropy due to the $l$-th ray $z_{c,l}$ at position $x_c$ is
\begin{align}
\label{eqn:HRayInt}
\text{E}[H(P(m|&X_{1:t},Z_{1:t},x_c,z_{c,l}))]
\nonumber\\&
=\int_{z_\text{min}}^{z_\text{max}}
H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))%\nonumber\\&\qquad\times 
p(z_{c,l}|X_{1:t},Z_{1:t},x_c)
dz_{c,l}.
\end{align}
%where $z_\text{min}$ and $z_\text{max}$ are sensor limits. Considering the continuous space of possible measurements of $z_{c,l}$ would make \refeqn{HRayInt} intractable to calculate. Thus, w
We discretize the measurement ray space such that $z_{c,l}$ falls on points along the measurement ray intersecting with grid cell edges. %, similar with how occupancy grid mapping decomposes continuous space into grid cells of length $\alpha$.
The discretized expected value of \refeqn{HRayInt} is
\begin{align}
\label{eqn:DiscExpEntropyRay}
\text{E}[H(P(&m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]
\nonumber\\&=\sum_{k=1}^{n_{r,l}+1}\bigg\{H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l,k}))%\nonumber\\&\qquad\times 
P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)\bigg\},
\end{align}
where index $z_{c,l,k}$ is the distance from position $x_c$ to the $k$-th grid cell of the $l$-th reduced map $r_l$, known from geometry.

Standing alone, the term $P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)$ from \refeqn{DiscExpEntropyRay} has a convoluted meaning because the depth $z_{c,l,k}$ does not directly depend on the map. However, we present a method to obtain this discretized probability. %Despite how this term is replaced with a normalizer in ~\cite{ThrBurFox05} replaces
Following the assumption that $z_{c,l}$ is discretized to known distances, the probabilities are proportional to their densities, as the area under the density curve is infinitesimal and fixed. Accounting for all cases, the probability is
\begin{align}
\label{eqn:ProbWithDelta}
P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)=\frac{p(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)}{\sum_{i=1}^{n_{r,l}+1}p(z_{c,l,i}|X_{1:t},Z_{1:t},x_c)}.
\end{align}
Conveniently, this density corresponds to the inverse normalizer according to \refeqn{BayesRuleRayISM},
\begin{align}
\label{eqn:DiscretizedProb}
p(z_{c,l,k}|X_{1:t},X_c,Z_{1:t})=\eta_{c,l,k}^{-1},
\end{align}
where $\eta_{c,l,k}^{-1}$ is defined in \refeqn{allEta}. %However, instead of evaluating the $t$-th pose $X_t$ and measured ray $z_{t,l}$, we simply consider $X_c$ and ray depth $z_{c,l,k}$ based on histories $X_{1:t}$ and $Z_{1:t}$ to obtain the inverse normalizer,
Updating variable subscripts, the normalizer becomes
\begin{align}
\label{eqn:allEtaAtC}
\eta_{c,l,k}^{-1}
&=
\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t},Z_{1:t})\bigg\}%\nonumber\\&\qquad\times 
p(z_{c,l,k}|\mathbf{r}_{l,i+},X_c)P(\mathbf{r}_{l,i}|X_{1:t},Z_{1:t}).
\end{align}
By substituting \refeqn{ProbWithDelta} and \refeqn{DiscretizedProb} into \refeqn{DiscExpEntropyRay}, the expected entropy at $X_c$ from individual measurement ray $z_{c,l}$ is
\begin{align}
\label{eqn:HRayComplete}
\text{E}[H(P&(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]\nonumber\\&=
\left(\sum_{i=1}^{n_{r,l}+1}\eta_{c,l,i}^{-1}\right)^{-1}%\nonumber\\&\quad\quad\times
\sum_{k=1}^{n_{r,l}+1}\bigg\{H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l,k}))\eta_{c,l,k}^{-1}\bigg\},
\end{align}
where $\eta_{c,l,k}^{-1}$ is taken from \refeqn{allEtaAtC}.
Here, \refeqn{HRayComplete} provides the expected entropies for those cells inside the field of view of $r_l$. %; cells outside this reduced map are trivially changed by $0$ due to $z_{c,l}$.
The entropies of cells outside the field of view remain unchanged.

\subsection{Approximation of Expected Ray Entropy}

The order of computation for each measurement ray is $\mathcal O((n_{r,l}+1)^2)$ since the summations of \refeqn{allEtaAtC} are embedded in \refeqn{HRayComplete}. However, several of those intersections provide negligible information since the probability of the $l$-th measurement ray capturing certain cell depths is close to zero.

The approximation of expected ray entropy provides a method to reduce the computation of \refeqn{allEtaAtC} substantially. This goal is achieved by systematically selecting a smaller set of grid cells to consider over the summations of \refeqn{allEtaAtC} and \refeqn{HRayComplete}.
The smaller set is determined by the probability that each cell is captured by the measurement ray, known as the detection probability. This can be found recursively as
\begin{align}
\label{eqn:ProbOfFirstCell}
&P(\mathbf{r}_{l,k+}|X_{1:t},Z_{1:t})%\nonumber\\&
=\bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t},Z_{1:t})\bigg\}P(\mathbf{r}_{l,k}|X_{1:t},Z_{1:t}),
\end{align}
which is the probability that $\mathbf{r}_{l,k}$ is the closest occupied grid cell based on past poses and measurement scans, independent of cells beyond the $k$-th cell from $x_c$.
Let $\hat n>0$ be a fixed number of grid cells for all rays such that $\hat n\leq n_{r,l}+1$ for all $l$.
Let $\hat{r}_{l}$ correspond to the grid cells that yield the $\hat{n}$ maximum values of \refeqn{ProbOfFirstCell} (the $\hat n$ most likely ray detections), indexed by increasing distance from candidate location $x_c$.
By replacing the reduced map $r_l$ with $\hat{r}_l$ and changing the summation limits to $\braces{1,2,...,\hat n}$ in \refeqn{allEtaAtC} and \refeqn{HRayComplete}, the order of computation is reduced to $\mathcal O({\hat{n}}^2)$.
Even though the value of $n_{r,l}$ is different among various rays in general, $\hat n$ is fixed, so the computational order is fixed as well.
In short, this method reduces the required computation substantially by systematically neglecting those grid cells with little effect. It can be noted that if $\hat n=n_{r,l}+1$, the ray objective function is computed without approximation.

\subsection{1D Ray Expected Information Gain Algorithm}

We present an algorithm pseudocode providing the necessary steps to obtain the objective function for a single measurement ray (Table \ref{tab:RayExpectedEntropyGain}). Much like the algorithm pseudocode for the inverse sensor model (Table \ref{tab:RayByRayISM}), the variable $a_\text{temp}$ serves as an intermediate variable designed to avoid repeated calculations.
Since this algorithm operates as a function, fixed indices and condition variables are removed for simplification.

\begin{table}[p]
\begin{tabular}{ l }
  Function: $\mathcal I_\text{ray}=\text{RayExpInfoGain}(x,P(r),z_{1:n_{r}})$\\
  Define $P(\bar{\mathbf{r}}_{0})=P(\bar{\mathbf{r}}_{n_{r}+1})=1$;\\
  for $k = 1,2,\ldots,n_{r}+1$\\
   \ \ $P(\mathbf{r}_{k+}) = P(\bar{\mathbf{r}}_{0:k-1})P(\mathbf{r}_{k})$;\\
   \ \ $P(\bar{\mathbf{r}}_{0:k})=P(\bar{\mathbf{r}}_{0:k-1})P(\bar{\mathbf{r}}_{k})$;\\
  end for\\
%  $P(\mathbf{r}_{(n_{r}+1)+}) = P(\bar{\mathbf{r}}_{0:n_{r}})$;\\
  Find $\hat r\subset r$ corresponding to the $\hat n$ greatest values of\\
  $\braces{P(\mathbf{r}_{1+}),P(\mathbf{r}_{2+}),\ldots,P(\mathbf{r}_{(n_{r}+1)+})}$;\\
  Define $P(\hat{\bar{\mathbf{r}}}_{0})=1$;\\
  for $k = 1,2,\ldots,\hat{n}$\\
   \ \ $P(\hat{\mathbf{r}}_{k+}) = P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\mathbf{r}}_{k})$;\\
   \ \ $P(\hat{\bar{\mathbf{r}}}_{0:k})=P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\bar{\mathbf{r}}}_{k})$;\\
  end for\\
  for $k_\text{m}=1,2,\ldots,\hat n$\\
   \ \ Initialize $\eta^{-1}_{k_\text{m}}=0$;\\
   \ \ for $k_\text{c}=1,2,\ldots,\hat n$\\
   \ \ \ \ $a_\text{temp}=P(\hat{\mathbf{r}}_{k_\text{c}+})p(z_{k_\text{m}}|\hat{\mathbf{r}}_{k_\text{c}+},x)$;\\
   \ \ \ \ $\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=P(\hat{\mathbf{r}}_{k_\text{c}})\eta^{-1}_{k_\text{m}}+a_\text{temp}$;\\
   \ \ \ \ $\eta^{-1}_{k_\text{m}}=\eta^{-1}_{k_\text{m}}+a_\text{temp}$;\\
   \ \ end for\\
   \ \ $P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=\eta_{k_\text{m}}\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})$ for all $k_\text{c}=1,2,\ldots,\hat n$;\\
   end for\\ 
   $P(z_{k_\text{m}}|x)=\frac{\eta^{-1}_{k_\text{m}}}{\sum_{i=1}^{\hat n}\eta^{-1}_{i}}$ for $k_\text{m}=1,2,\ldots,\hat n$;\\

   Initialize $\mathcal I_\text{ray}=0$;\\
   for $k_\text{c}=1,2,\ldots,\hat n$\\
   \ \ $\mathcal I_\text{ray}=\mathcal I_\text{ray}+H(P(\hat{\mathbf{r}}_{k_\text{c}}))$;\\
   \ \ for $k_\text{m}=1,2,\ldots,\hat n$\\
   \ \ \ \ $\mathcal I_\text{ray} = \mathcal I_\text{ray}-H(P(\hat{\mathbf{r}}_{k_\text{c}}|x_c,z_{k_\text{m}}))P(z_{k_\text{m}}|x_{c})$;\\
   \ \ end for\\
  end for\\
  Return: $\mathcal I_\text{ray}$\\
\end{tabular}
\caption{Expected Information Gain from a Measurement Ray}
\raggedright{Note: this function is only considering a single ray for a future time step, where probabilities are subject to conditions on $X_{1:t},Z_{1:t}$, so these are removed from the algorithm for simplicity.}
\label{tab:RayExpectedEntropyGain}
\end{table}

\subsection{Numerical Justification for Approximation}

The purpose of this numerical example is to provide evidence that the approximations are reasonable and increase the algorithm speed substantially.
Since a measurement ray produces a depth measurement in a single direction, we only consider a 1D map where the grid cells have spacing $\alpha=0.2$ meters, and the properties of the range sensor are based on the Microsoft Kinect~\cite{PirRutBisSch11,KhoElb12} with maximum reading depth $z_\text{max}=4$m ($20$ grid cells inside the sensor FOV). The goal is to compare the expected entropy $\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$ from \refeqn{HRayComplete} and with an approximation $\text{E}[H_\text{approx}(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$, which only considers $\hat n$ grid cells with highest detection probability \refeqn{ProbOfFirstCell}.

We consider $100$ probabilistic maps to obtain Monte Carlo results to evaluate the approximate entropy. In every Monte Carlo trial, each grid cell has an $80\%$ chance of being free and a $20\%$ chance of receiving an a priori probability uniformly distributed between $0$ and $1$. 
%We consider $100$ probabilistic maps to obtain Monte Carlo results to evaluate the approximate entropy. In every Monte Carlo trial, each grid cell has an $80\%$ chance of being free, and the remaining $20\%$ of grid cells are given an a priori probability uniformly distributed between $0$ and $1$. 
Several metrics serve to evaluate $\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$ with $\text{E}[H_\text{approx}(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$. The median expected entropy change is $\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]-H(P(m|X_{1:t},Z_{1:t}))=-0.83792$. The error for the $100$ Monte Carlo cases is defined simply as
\begin{align}
e_{H}&=\frac1{100}\sum_{i=1}^{100}\text{abs}\bigg(\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]\nonumber\\&\qquad\quad-\text{E}[H_\text{approx}(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]\bigg).
\end{align}
The Monte Carlo trials are repeated for $\hat n=\braces{1,2,\ldots,10}$ and the results are plotted in Figure \ref{fig:ApproxJust}.
This example shows a typical case when the summation limits generated from \refeqn{ProbOfFirstCell} have only small effects on \refeqn{HRayComplete}, while providing very large improvements in reducing computation.

\begin{figure}
	\centering
    	\begin{subfigure}[b]{0.35\textwidth}
        		\includegraphics[width=\textwidth]{JustifyApprox_eH.pdf}
        		\caption{Entropy Error}
        		\label{fig:H_err}
    	\end{subfigure}
%	\hspace*{0.1\columnwidth}
	\begin{subfigure}[b]{0.35\textwidth}
        		\includegraphics[width=\textwidth]{JustifyApprox_t.pdf}
        		\caption{Computation Ratio}
        		\label{fig:H_comp_ratio}
    	\end{subfigure}
%	\begin{subfigure}[Computation Ratio]
%		{\includegraphics[height=0.35\columnwidth]{JustifyApprox_tRatio.pdf}
%	}
%	\end{subfigure}

\caption{The entropy error is decreased at the cost of increasing computation time in this Monte Carlo 1D measurement ray expected entropy case.}
\label{fig:ApproxJust}
\end{figure}

\section{Autonomous Exploration}

In this section, we develop an autonomous exploration scheme inside a 2D environment based on choosing movements that minimize map entropy and avoid collisions.

\subsection{Pose Selection Optimization}

The objective is to choose a future pose that minimizes the map uncertainty, such that the robot can autonomously move toward this optimal goal. The robot pose is selected among various positions and attitudes. Since searching over all locations and attitudes would require infinite computational resources, these search spaces are discretized into a limited number of robot attitudes and positions.
The procedure involves three steps. First, the attitude that maximizes the information gain objective function is selected at each candidate future pose location. Second, the location with the maximum objective function of all candidates is selected. Thus, the optimal attitude and the optimal location compose the optimal future pose. Finally, the robot determines and follows a collision-free path to the optimal future pose. This process is repeated, resulting in autonomous exploration.

%First, the optimal attitude for an arbitrary location is presented. Then, two methods to select pose candidates, namely \emph{expanding ring} and \emph{complete Cartesian}, are described below, with a brief discussion of the benefits and drawbacks of choosing each search space. Finally, the equations and conditions to determine an optimal future pose are presented.

\paragraph{Attitude Optimization}
Consider a collision-free pose candidate at an arbitrary location $x_c$. At this pose, consider $n_d$ evenly-spaced measurement rays oriented radially outward from the robot in a circular pattern (see red lines in Figure \ref{fig:OptProcess}). These rays are denoted $\braces{z_{c,1},z_{c,2},\ldots,z_{c,n_d}}$. Attitudes correspond to the rays such that the robot sensor is aligned with the ray direction, and these are denoted $\braces{R_{c,1},R_{c,2},\ldots,R_{c,n_d}}$, where the scan with attitude $R_{c,d}$ might cover several ray directions depending on the sensor field of view (FOV). We choose optimal attitude $R_c^*$ as the summation of the expected entropy changes covered by the scan,
\begin{align}
\label{eqn:FindRc}
&R_c^*=\argmax_{R_{c,d}}\sum_{z_{c,i}\in R_{c,d}\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\qquad\quad\quad\quad\quad
-\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,i}))]\bigg).
%\mathcal I_\text{ray}(x_c,z_{c,i})\right).
\end{align}

This method provides the attitude that maximizes the information gain at an arbitrary location in 2D space. The set of candidate positions that warrant consideration is determined with one of two methods, namely \emph{expanding ring} and \emph{complete Cartesian}, described next.

\paragraph{Expanding Ring}
The expanding ring technique is advantageous for searching local solutions quickly, only considering distant future poses when necessary. The key idea is that future candidate locations lie on a circular ``ring'' centered around the robot, evenly spaced around the ring (see red circles in Figure \ref{fig:OptProcess}), and the ring is expanded if certain criteria are not met. More explicitly, consider $n_c$ candidate locations denoted by $x_c\in\braces{x_1,x_2,\ldots,x_{n_c}}$ located with the distance $\delta$ away from the robot. Any location that violates inequality constraint \refeqn{CollisionInequalityConstraint} is excluded to avoid collisions. The set of optimal attitudes at each candidate location is $\braces{R^*_{1},R^*_{2},\ldots,R^*_{n_c}}$, obtained from \refeqn{FindRc}. The objective function \refeqn{ObjFun} is computed by a summation about the $n_d$ rays as %\refeqn{Objective},
\begin{align}
\label{eqn:ObjFunApprox}
\mathcal I(x_c,R_c^*)&\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\qquad-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*,z_{c,i}))\right]\bigg),
\\
x_c^*&=\argmax_{x_c}\mathcal I(x_c,R_c^*).
\end{align}
At the optimal pose $X^*_c=(x^*_c,R^*_c)$, the resulting information gain must satisfy $\mathcal I(X_c^*)\geq\mathcal I_\text{min}$, where $\mathcal I_\text{min}$ is some minimum threshold for expected information gain; robot motion is only justified when the expected information gain is significantly large. If the minimum threshold is not met, the ring of candidate pose locations is increased such that $n_c$ and $\delta$ are multiplied by a scale function $\lambda>1$.
This process is repeated until the expected information gain of the optimal pose is at least $\mathcal I_\text{min}$, or all candidates lie in collision zones or outside map limits. %A pseudocode of the process is found in Table \ref{tab:Alg_AutomousExploration}.

The primary advantage of the expanding ring approach to search the 2D space is that the number of pose candidates need not be proportional to the map area, and that local maxima tend to fall within short distances of the current robot pose. The main drawback is that poses outside a small local region of the robot are frequently neglected, and the robot frequently executes repeated short motions, often without large information gains.

\paragraph{Complete Cartesian}
The complete Cartesian method to search the 2D space provides candidates located fixed distance $d$ apart in each Cartesian direction. This approach is advantageous for capturing expected information gains in regions outside of the immediate vicinity of the robot, but must be applied carefully to avoid issues with computational bottlenecks due to the potentially-large search space.

Similar to the expanding ring approach, $n_c$ candidate pose locations are considered, where those violating \refeqn{CollisionInequalityConstraint} are neglected from further consideration. However, unlike the ring expanding approach, the candidates have different distances from the robot in general, so the objective function is modified to enforce a cost on the squared distance from the current pose location $x_t$ to the candidate location $x_c$, i.e.,
\begin{align}
\label{eqn:ObjFunApproxCompleteCartesian}
\mathcal I&(x_c,R_c^*,x_t)\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*(x_c),z_{c,i}))\right]\bigg)-k_\text{dist}(x_t-x_c)^\T(x_t-x_c),
\end{align}
where weighting function $k_\text{dist}$ represents the sensitivity to avoiding large motions across the map. Furthermore, the same candidate locations are considered throughout the repeated processes of exploration, and the map cell occupancies are assumed static. Therefore, if candidate location $x_c$ fails to satisfy $\mathcal I(x_c,R_c^*,x_t)\geq\mathcal I_\text{min}$, then $x_c$ need \emph{never} be considered again. Avoiding these unnecessary calculations yields a greatly lowered computational burden, particularly late in exploration where several regions are collision-free but disadvantageous to revisit.

\begin{figure}
\vspace*{0.1\columnwidth}
\centerline{
	\includegraphics[height=0.35\columnwidth]{ExampleOptimalPose.png}
}
\begin{picture}(0,0)(0,0)
\setlength{\unitlength}{0.1\columnwidth}\scriptsize
\put(4.4,2.1){\color{blue}$X_t$}
\put(5.9,1.9){\color{red}$x_1$}
\put(2.9,2.0){\color{red}$x_5$}
\put(4.4,3.6){\color{red}$x_3$}
\put(4.4,0.6){\color{red}$x_7$}
\put(5.4,3.0){\color{red}$x_2$}
\put(3.3,3.1){\color{red}$x_4$}
\put(3.3,1.0){\color{red}$x_6$}
\put(5.4,1.0){\color{red}$x_8$}
\put(6.8,3.1){\color{red}$z_{2,1}$}
\put(4.5,3.1){\color{red}$z_{2,5}$}
\put(5.6,4.1){\color{red}$z_{2,3}$}
\put(5.6,2.2){\color{red}$z_{2,7}$}
\put(6.5,3.7){\color{red}$z_{2,2}$}
\put(4.9,3.8){\color{red}$z_{2,4}$}
\put(5.0,2.5){\color{red}$z_{2,6}$}
\put(6.5,2.5){\color{red}$z_{2,8}$}
\put(6.1,2.9){$X_c^*$}
\end{picture}
\caption{This figure illustrates the expanding ring method. Initially, a robot (blue circle) views down and left (blue sector). Then, the robot considers $n_c=8$ poses (red circles) as possible candidates for where to move next. For each candidate, $n_d=8$ directions are considered (red lines, only displayed on candidate location $c=2$). Then, the expected entropy of each ray is calculated. The scan (red sector) covering those rays with the largest expected entropy decreases is chosen for each candidate, and the best candidate $X_c^*$ (black circle: location, black sector: scan) is chosen to maximize information gain. Finally, Dijkstra's algorithm provides the collision-free motion between $X_t$ and $X_c^*$, and the process is repeated.
}
\label{fig:OptProcess}
\end{figure}

%avoiding collisions subject to the inequality constraint \refeqn{CollisionInequalityConstraint}.

\paragraph{Motion Planning}
Once the optimal future pose $X_c^*$ is selected, the robot must traverse its environment to reach $X_c^*$ without collisions.
The locations of cells on a grid-based map provide nodes for Dijkstra's algorithm~\cite{Dij59}. At each node, \refeqn{CollisionInequalityConstraint} must be satisfied for the node to be considered reachable, and edges between nodes correspond to neighboring grid cells.
Dijkstra's algorithm provides a relatively simple motion planning strategy to move from $x_t$ to $x_c^*$ that avoids collisions and local minima.

The trajectory to follow the path outlined by Dijkstra's algorithm is determined with a linear constrained least squares optimization, assuming the robot follows a polynomial trajectory with fixed speed. The starting and ending positions and attitudes may be constrained, and polynomials patched together for long trajectories share a common position and velocity with respect to time. Then, a controller on the robot tracks this trajectory until the robot falls within acceptable thresholds of the final optimal pose.
Once the robot completes this motion, the entire process is repeated. %The complete autonomous exploration process is summarized in Table \ref{tab:Alg_AutomousExploration}.

%The first pseudocode in table \ref{tab:Alg_AutomousExploration} outlines how a candidate pose is selected given an objective function. The second pseudocode in table \ref{tab:Alg_ObjFun}

%\begin{table}[p]
%\begin{tabular}{ l }
%%  Define $n_c$: number of candidate locations\\
%%  Define $n_d$: number of evenly-spaced measurement rays per candidate\\
%  Given: $t=1$, $X_{1}$, $Z_{1}$, $P(m|X_{1},Z_{1})$ from the inverse sensor model\\
%  TrajectoryComplete = false;\\
%  while TrajectoryComplete = false\\
%  \ \ if $X_{t+1}$ is unknown\\
%  \ \ \ \ MotionPlanning = false;\\
%  \ \ \ \ while MotionPlanning = false\\
%  \ \ \ \ \ \ for $c = 1,2,\ldots,n_c$\\
%  \ \ \ \ \ \ \ \ Candidate location: $x_{c}$ from evenly-spaced circle around\\
%  \ \ \ \ \ \ \ \ $X_t$ with radius $\delta$;\\
%  \ \ \ \ \ \ \ \ for $d=1,2,\ldots,n_d$\\
%  \ \ \ \ \ \ \ \ \ \ Obtain reduced map $r_{c,d}$ with $n_{r_{c,d}}$ intersections\\
%  \ \ \ \ \ \ \ \ \ \ through grid cells with depths $z_{c,d,1:n_{r,d}}$ from geometry;\\
%  \ \ \ \ \ \ \ \ \ \ $\mathcal I_\text{ray}(x_c,z_{c,d})=\text{RayExpInfoGain}(x_c,P(r_{c,d}),z_{c,d,1:n_{r,d}})$;\\
%  \ \ \ \ \ \ \ \ end for\\
%  \ \ \ \ \ \ \ \ $R_{c}^*=\argmax_{R_c}\left(\sum_{z_{c,i}\in R_c \text{ FOV}}\mathcal I_\text{ray}(x_c,z_{c,i})\right)$;\\%\argmax_{R}{\sum_{d=1}^{n_d}\mathcal I_{\text{ray},c,d}|d\in\text{FOV}(R))}$;\\
%  \ \ \ \ \ \ \ \ $\mathcal I(x_c,R_c^*)=\sum_{z_{c,i}\in R_c^* \text{ FOV}}\mathcal I_\text{ray}(x_c,z_{c,i})$;\\%\sum_{d\in \text{FOV}(R_c^*)}\mathcal I_{\text{ray}_{c,d}}$;\\
%  \ \ \ \ \ \ end for\\
%  \ \ \ \ \ \ $x_c^*=\argmax_{x_c}\mathcal I(x_c,R_c^*)$;\\
%  \ \ \ \ \ \ $X_c^*=\braces{x_c^*,R_c^*}$;\\
%  \ \ \ \ \ \ if $\mathcal I(X_c^*)<\mathcal I_\text{min}$\\
%  \ \ \ \ \ \ \ \ $n_c=\text{round}(\lambda n_c)$;\\
%  \ \ \ \ \ \ \ \ $\delta=\lambda \delta$;\\
%  \ \ \ \ \ \ else\\
%  \ \ \ \ \ \ \ \ Obtain $X_{t+1},X_{t+2},\ldots,X_{c}^*$ using Dijkstra's algorithm and minimum rotation;\\
%  \ \ \ \ \ \ \ \ Reset $n_c$ and $\delta$ to initial values;\\
%  \ \ \ \ \ \ \ \ MotionPlanning = true;\\  
%  \ \ \ \ \ \ end if\\
%  \ \ \ \ end while\\
%%  $X_c^*=\argmax_{X}$\mathcal I_{\text{ray},c
%%  \ \ $\mathcal D\subset\braces{1,2,\ldots,n_d}$ represents scan coverage of candidate rays\\
%%  \ \ The $c$-th pose candidate: $X_{c}=\braces{x_{c},R_{c}}$;\\
%%   end for\\
%%  $X_c^*=\argmin_{X_c}({\Delta H(X_c)}|X_c\in\braces{X_1,X_2,\ldots,X_{n_c}})$;\\
%%  Motion planner generates a collision-free path between $X_t$ to $X_c^*$\\
%%  Move to $X_{t+1}$, obtain $Z_{t+1}$, and find $P(m|X_{1:t+1},Z_{1:t+1})$;
%  \ \ end if\\
%  \ \ t = t+1;\\
%  \ \ Obtain $P(m|X_{1:t},Z_{1:t})$ from the inverse sensor model;\\
%  \ \ if map $m$ is completely explored\\
%  \ \ \ \ TrajectoryComplete = true;\\
%  \ \ end if\\
%  end while
%
%
%\end{tabular}
%\caption{Autonomous Exploration via Expected Information Gain from Expanding Ring Searches}
%\label{tab:Alg_AutomousExploration}
%\end{table}

\section{Numerical Examples}
\label{sec:NumRes}

The efficacy of the proposed algorithms are verified by numerical examples. First, the occupancy grid mapping strategies are simulated in Section \ref{sec:NEOGC}, followed by numerical examples for autonomous exploration in Section \ref{sec:NEAE}.

% It is common practice to approximate\refeqn{InvSenModWithProbDens} with a simplified function, frequently following the structure of the approximate inverse sensor model proposed in~\cite{And09}. The main idea of this approach is that the probability of a cell being occupied ($i$) near a measurement is high (measurement likely hits this cell), ($ii$) between the robot and the measurement is low (measurement passes through this cell), and ($iii$) beyond the measurement is unchanged (the robot cannot measure through a wall/object). The function is based on intuition, not thorough mathematical derivation. In short, this approach simplifies the inverse sensor model at the cost of accuracy.

%The Kinect depth sensor is used as an example, with specifications from~\cite{KhoElb12}. With these parameters and the identical set of measurements, the exact solution is compared with a benchmark algorithm for a Kinect sensor taken from summarized as follows.

\subsection{Occupancy Grid Mapping}\label{sec:NEOGC}

First, we compare the proposed exact solution to the inverse sensor model summarized in Table \ref{tab:RayByRayISM} with an approximate algorithm presented in~\cite{PirRutBisSch11,KhoElb12}, summarized as follows. The probability that the $i$-th grid cell $\mathbf{m}_i$ is occupied conditioned on the measurement ray $z_{t,l}$ at the pose $X_t$ is the continuous function
\begin{align*}
%\label{eqn:ISM_Approx_1}
&P(\mathbf{m}_i|z_{t,l},X_t)%\nonumber\\&
=\begin{cases}
0.3+(\frac{k}{\sigma\sqrt{2\pi}}+0.2)e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ \hat z_{l,i} \leq z_{t,l},%z_{t,l}\leq \hat z_{l,i},
\\
0.5+\frac{k}{\sigma\sqrt{2\pi}}e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{otherwise},
\end{cases}
\end{align*}
which is based on the expected distance to the cell $\hat z_{l,i}$ with parameters $k=\sigma=0.6$. This follows the structure of the approximate inverse sensor model proposed in~\cite{And09}. The main idea of this approach is that the probability of a cell being occupied (i) near a measurement is high (measurement likely hits this cell), (ii) between the robot and the measurement is low (measurement passes through this cell), and (iii) beyond the measurement is unchanged (the robot cannot measure through a wall/object). 
Then, these probabilities are combined in a weighted fashion such that all measurements rays of scan $Z_t$ simultaneously update the same grid cell in a log-odds format,
\begin{align*}
%\label{eqn:ISM_Approx_2}
&\log\left(\frac{P(\mathbf{m}_i|Z_{t},X_t)}{1-P(\mathbf{m}_i|Z_{t},X_t)}\right)
%\nonumber\\&
=
\frac1{\sum_{z_{t,l}\in\mathbf{m}_i}\hat z_{l,i}}\sum_{z_{t,l}\in\mathbf{m}_i}\log\left(\frac{P(\mathbf{m}_i|z_{t,l},X_t)}{1-P(\mathbf{m}_i|z_{t,l},X_t)}\hat z_{l,i}\right).
\end{align*}

%Given the same poses and measurement sets, the proposed algorithm from \refeqn{RayISMAnswer}-\refeqn{RayByRayScanISM} are compared with the approximate algorithm described above. A probabilistic map consisting of NUMX by NUMY square grid cells with edge length ALPHA is assumed to accurately represent the occupancy of the environment surrounding a robot. The robot follows (SOMETHING ABOUT THE TRAJECTORY). With both algorithms, the probabilistic maps are updated in real time.
%
%(PARAGRAPH ABOUT MAP CLARITY, ENTROPY, AND COMPUATION TIME)

Table~\ref{tab:penn} shows the mapping parameters and data specifications for the actual odometry and lidar measurements from University of Pennsylvania through Coursera open course~\cite{coursera}.
Figure~\ref{fig:penn_map} shows the direct comparison of the resulting map based on the approximate and proposed inverse sensor model.
As shown in the figure, both approaches capture the structure of the environment. However, the significant differences can be seen in the certain area of the map such as in Figure~\ref{fig:penn_map_zoom}.
The differences can be clearly depicted in Figure~\ref{fig:entropy_comp} where entropy of maps are compared at frame 2280.
The red part of the map shows the maximum  entropy whereas blue shows lowest entropy for the cell.
Finally, overall improvement was demonstrated in Figure~\ref{fig:entropy} resulting improved entropy value per grid cell throughout the mapping process.
In short, the proposed exact occupancy grid mapping produced a cleaner map with lower entropy from the same set of measurements.

\begin{center}
\captionof{table}{Experimental parameters provided from dataset}
\label{tab:penn}
    \begin{tabular}{r | c}
        Parameters & Value\\ \hline\hline
        Map dimension & 576x448 [pixel]\\
        Resolution & 1/16 [m]\\
        Scan angle & [-2.36, 2.36]\\
        Ray number & 1081\\
        Total frame & Every 120 scans out of 3701\\
    \end{tabular}
\end{center}

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[t]{0.8\columnwidth}
        \centering
        \includegraphics[trim={1.5cm 3.5cm 1.5cm 1.5cm},clip,width=\textwidth]{map_comparison.png}
        \caption{Comparison of the total mapped environment based on the approximate (left) and proposed (right) inverse sensor model.}
        \label{fig:penn_map_total}
    \end{subfigure}
    \begin{subfigure}[t]{0.8\columnwidth}
        \centering
        \includegraphics[trim={1.5cm 3.5cm 1.5cm 1.5cm},clip,width=\textwidth]{map_zoom.png}
        \caption{Evidence of significant improvement at certain area of the map.}
        \label{fig:penn_map_zoom}
    \end{subfigure}
    \caption{Using experimental sensor data, the map was generated using an approximate and the proposed exact inverse sensor models. The laser sensor and odometry data used for the experiment was provided by University of Pennsylvania open course robotics estimation and learning on Coursera~\cite{coursera}.}
    \label{fig:penn_map}
\end{figure}

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[t]{0.35\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{AISM_Image_inf_19.pdf}
        \caption{Approximate inverse sensor model}
        \label{fig:AISM}
    \end{subfigure}
    \begin{subfigure}[t]{0.35\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{EISM_Image_inf_19.pdf}
        \caption{Exact inverse sensor model}
        \label{fig:EISM}
    \end{subfigure}
    \caption{The entropy is compared between the approximate and proposed exact inverse sensor models at step 2280. The blue areas show the lowest entropy while red shows highest.}
    \label{fig:entropy_comp}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{entropy_frame.png}
  \caption{The map entropies of the approximate inverse sensor model (blue) and the proposed exact inverse sensor model (red) show that the exact version yields a more certain occupancy grid map.}
  \label{fig:entropy}
\end{figure}

%\paragraph{Computation Time}
%The proposed ray inverse sensor model is designed to maintain the accuracy of the exact solution with substantial computational improvements.
%A $1$-dimensional example is simulated on a Mac desktop computer with $11$ grid cells, $10$ of which are visible to the robot, where the same depth measurement is used for all occupancy grid algorithms.
%The computational time of the exact solution is computed in the conventional manner according to \refeqn{InvSenModWithProbDens} in $2.3314$ seconds to obtain the same solution of \refeqn{RayISMAnswer} from \refeqn{Unnormalized} and \refeqn{allEta} in just $0.0027$ seconds, which is $874.77$ times faster.
%The approximate ray inverse sensor model completes in $0.0011$ seconds, but this algorithm does not follow a forward sensor model, so the probabilities of the map may be inaccurate.

%\subsection{Trajectory Map Building}
%In this numerical example, we consider a robot in a two-dimensional environment composed of ten wall edges, and the robot follows a figure-eight curve, then turns around and completes the same curve in the reverse direction.
%The map is composed of $100\times100$ square grid cells with edge length of $1$ cm.
%A forward sensor model $p(z_t|m,X_t)$~\cite{ThrBurFox05} is constructed according to the specifications of the Kinect sensor~\cite{KhoElb12}, and a pseudo-random measurement scan is sampled at each time step via the inverse transform sampling~\cite{DevBK86}. The same set of measurements updated each second are used with both occupancy grid mapping algorithms to construct the map.
%
%The resulting maps are illustrated in Figure \ref{fig:NumResOccProbs} for both algorithms, where it is shown that the proposed algorithm yields a substantially more accurate and clear map with less uncertainty. To quantify the degree of map uncertainty, we define the entropy of the map as 
%\begin{align*}
%&H(P(m|X_{1:t},Z_{1:t}))=\nonumber\\
%&-\sum_{i=1}^n\big\{P(\mathbf{m}_i|X_{1:t},Z_{1:t})\log P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\
%&+(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\log(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\big\},
%\end{align*}
%which is maximized when the probability of occupancy is $0.5$ for all cells (more uncertain), and it is minimized when they are either $0$ or $1$ (less uncertain). 
%
%The change of the map entropy over time, and the entropy of the completed maps, for both methods are depicted in Figure \ref{fig:NumResOccH}. The subfigure (a) illustrates that the proposed exact inverse sensor model exhibits rapid decreases of entropies, and smaller entropies always. The resulting terminal map obtained from the proposed approach, shown in the subfigure (b), has less uncertainty than (c) constructed by the approximate model. In short, the proposed approach is more efficient at extracting information about the environment from the same set of set measurements.

\subsection{Autonomous Exploration}\label{sec:NEAE}

%Maximizing the proposed expected information gain serves as the objective for autonomous exploration. A robot begins in a completely uncertain environment, except for the grid cells inside the immediate vicinity the robot location.

%At each time step, the robot receives a measurement scan, where the probabilistic properties of the sensor are taken from~\cite{PirRutBisSch11,KhoElb12}. Then, the $n_c=8$ evenly-spaced candidate locations about a circle of radius $\delta=0.5$m around the current pose location are considered, where $n_d=16$ measurement rays are evenly-spaced about the candidate location.  When no current candidates yield expected information gains above $\mathcal I_\text{min}=1.25$, $\lambda=1.25$ is multiplied to the current $n_c$ and $\delta$ values. The motion of the robot is restricted to movement on grid cells satisfying \refeqn{CollisionInequalityConstraint} with $\beta=0.01$. Dijkstra's algorithm guides the robot through the environment without collision.

Next, the proposed exploration approach is applied to the floor plan of Intel Research Lab, illustrated in Figure~\ref{fig:intel}.
The robot explores its surroundings with an occupancy grid with $90,000$ cells where grid cell edges are $\alpha=0.1$m, composing a map with dimensions $30\text{m}\times30\text{m}$.
The initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) for grid cells covered by the circular robot of radius $0.3$m and $P(\mathbf{m}_i)=0.5$ for all other cells.

The robot follows the complete Cartesian candidate search method to determine future poses, with $k_\text{dist}=5$ to avoid large motions with little added expected information gains. Due to the large size of the map, candidates below the expected information gain of $1.25$ are neglected from future consideration.

The time sequence of the exploration for every 5~minutes are simulated in ROS Stage 2D environment. The corresponding simulation results are illustrated in Figure \ref{fig:IRL}.
Lighter red dots and darker blue dots in the explored area correlate to the level of information gain at those positions with low and high value, respectively. It is shown that the proposed autonomous exploration algorithm successfully mapped the complicated environment composed of a number of rooms, narrow hallways, and open spaces that are irregularly shaped.

%This approach not only achieved exploration of the entire environment but also improved computational efficiency.

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{intel_clean.png}
    \caption{Modified Intel Research Lab floor plan form the SLAM benchmark~\cite{kummerle2009measuring}.}
\label{fig:intel}
\end{figure}

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[t]{0.3\columnwidth}
        \centering
        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{0min.png}
        \caption{$t=0$ min}
        \label{fig:IRL0min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
        \centering
        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{5min.png}
        \caption{$t=5$ min}
        \label{fig:IRL5min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{10min.png}
        \caption{$t=10$ min}
        \label{fig:IRL10min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{15min.png}
        \caption{$t=15$ min}
        \label{fig:IRL15min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
         \centering
         \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{20min.png}
        \caption{$t=20$ min}
        \label{fig:IRL20min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{25min.png}
        \caption{$t=25$ min}
        \label{fig:IRL25min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{30min.png}
        \caption{$t=30$ min}
        \label{fig:IRL30min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{35min.png}
        \caption{$t=35$ min}
        \label{fig:IRL35min}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\columnwidth}
           \centering
           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{40min.png}
        \caption{$t=40$ min}
        \label{fig:IRL40min}
    \end{subfigure}
    \caption{A robot (green) measures a room in ROS Stage simulator using modified Intel Research Lab floor plan. The exact inverse sensor model is used for the autonomous exploration algorithm.}
    \label{fig:IRL}
\end{figure}

%\section{Numerical Examples}
%\label{sec:NumRes}
%
%The efficacy of the proposed algorithms are verified by numerical examples. First, the occupancy grid mapping strategies are simulated in Section \ref{sec:NEOGC}, followed by numerical examples for autonomous exploration in Section \ref{sec:NEAE}. 
%
%% It is common practice to approximate\refeqn{InvSenModWithProbDens} with a simplified function, frequently following the structure of the approximate inverse sensor model proposed in~\cite{And09}. The main idea of this approach is that the probability of a cell being occupied ($i$) near a measurement is high (measurement likely hits this cell), ($ii$) between the robot and the measurement is low (measurement passes through this cell), and ($iii$) beyond the measurement is unchanged (the robot cannot measure through a wall/object). The function is based on intuition, not thorough mathematical derivation. In short, this approach simplifies the inverse sensor model at the cost of accuracy.
%
%%The Kinect depth sensor is used as an example, with specifications from~\cite{KhoElb12}. With these parameters and the identical set of measurements, the exact solution is compared with a benchmark algorithm for a Kinect sensor taken from summarized as follows.
%
%
%
%
%\subsection{Occupancy Grid Mapping}\label{sec:NEOGC}
%
%First, we compare the proposed exact solution to the inverse sensor model summarized in Table \ref{tab:RayByRayISM} with an approximate algorithm presented in~\cite{PirRutBisSch11,KhoElb12}, summarized as follows. The probability that the $i$-th grid cell $\mathbf{m}_i$ is occupied conditioned on the measurement ray $z_{t,l}$ at the pose $X_t$ is the continuous function
%\begin{align*}
%%\label{eqn:ISM_Approx_1}
%&P(\mathbf{m}_i|z_{t,l},X_t)%\nonumber\\&
%=\begin{cases}
%0.3+(\frac{k}{\sigma\sqrt{2\pi}}+0.2)e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ \hat z_{l,i} \leq z_{t,l},%z_{t,l}\leq \hat z_{l,i},
%\\
%0.5+\frac{k}{\sigma\sqrt{2\pi}}e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{otherwise},
%\end{cases}
%\end{align*}
%which is based on the expected distance to the cell $\hat z_{l,i}$ with parameters $k=\sigma=0.6$. This follows the structure of the approximate inverse sensor model proposed in~\cite{And09}. The main idea of this approach is that the probability of a cell being occupied (i) near a measurement is high (measurement likely hits this cell), (ii) between the robot and the measurement is low (measurement passes through this cell), and (iii) beyond the measurement is unchanged (the robot cannot measure through a wall/object). 
%Then, these probabilities are combined in a weighted fashion such that all measurements rays of scan $Z_t$ simultaneously update the same grid cell in a log-odds format,
%\begin{align*}
%%\label{eqn:ISM_Approx_2}
%&\log\left(\frac{P(\mathbf{m}_i|Z_{t},X_t)}{1-P(\mathbf{m}_i|Z_{t},X_t)}\right)
%%\nonumber\\&
%=
%\frac1{\sum_{z_{t,l}\in\mathbf{m}_i}\hat z_{l,i}}\sum_{z_{t,l}\in\mathbf{m}_i}\log\left(\frac{P(\mathbf{m}_i|z_{t,l},X_t)}{1-P(\mathbf{m}_i|z_{t,l},X_t)}\hat z_{l,i}\right).
%\end{align*}
%
%Given the same poses and measurement sets, the proposed algorithm from \refeqn{RayISMAnswer}-\refeqn{RayByRayScanISM} are compared with the approximate algorithm described above. A probabilistic map consisting of NUMX by NUMY square grid cells with edge length ALPHA is assumed to accurately represent the occupancy of the environment surrounding a robot. The robot follows (SOMETHING ABOUT THE TRAJECTORY). With both algorithms, the probabilistic maps are updated in real time.
%
%(PARAGRAPH ABOUT MAP CLARITY, ENTROPY, AND COMPUATION TIME)
%
%%\paragraph{Computation Time}
%%The proposed ray inverse sensor model is designed to maintain the accuracy of the exact solution with substantial computational improvements.
%%A $1$-dimensional example is simulated on a Mac desktop computer with $11$ grid cells, $10$ of which are visible to the robot, where the same depth measurement is used for all occupancy grid algorithms.
%%The computational time of the exact solution is computed in the conventional manner according to \refeqn{InvSenModWithProbDens} in $2.3314$ seconds to obtain the same solution of \refeqn{RayISMAnswer} from \refeqn{Unnormalized} and \refeqn{allEta} in just $0.0027$ seconds, which is $874.77$ times faster.
%%The approximate ray inverse sensor model completes in $0.0011$ seconds, but this algorithm does not follow a forward sensor model, so the probabilities of the map may be inaccurate.
%
%
%
%
%
%
%
%
%%\subsection{Trajectory Map Building}
%%In this numerical example, we consider a robot in a two-dimensional environment composed of ten wall edges, and the robot follows a figure-eight curve, then turns around and completes the same curve in the reverse direction.
%%The map is composed of $100\times100$ square grid cells with edge length of $1$ cm.
%%A forward sensor model $p(z_t|m,X_t)$~\cite{ThrBurFox05} is constructed according to the specifications of the Kinect sensor~\cite{KhoElb12}, and a pseudo-random measurement scan is sampled at each time step via the inverse transform sampling~\cite{DevBK86}. The same set of measurements updated each second are used with both occupancy grid mapping algorithms to construct the map.
%%
%%The resulting maps are illustrated in Figure \ref{fig:NumResOccProbs} for both algorithms, where it is shown that the proposed algorithm yields a substantially more accurate and clear map with less uncertainty. To quantify the degree of map uncertainty, we define the entropy of the map as 
%%\begin{align*}
%%&H(P(m|X_{1:t},Z_{1:t}))=\nonumber\\
%%&-\sum_{i=1}^n\big\{P(\mathbf{m}_i|X_{1:t},Z_{1:t})\log P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\
%%&+(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\log(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\big\},
%%\end{align*}
%%which is maximized when the probability of occupancy is $0.5$ for all cells (more uncertain), and it is minimized when they are either $0$ or $1$ (less uncertain). 
%%
%%The change of the map entropy over time, and the entropy of the completed maps, for both methods are depicted in Figure \ref{fig:NumResOccH}. The subfigure (a) illustrates that the proposed exact inverse sensor model exhibits rapid decreases of entropies, and smaller entropies always. The resulting terminal map obtained from the proposed approach, shown in the subfigure (b), has less uncertainty than (c) constructed by the approximate model. In short, the proposed approach is more efficient at extracting information about the environment from the same set of set measurements. 
%
%
%
%
%
%
%\subsection{Autonomous Exploration}\label{sec:NEAE}
%
%%Maximizing the proposed expected information gain serves as the objective for autonomous exploration. A robot begins in a completely uncertain environment, except for the grid cells inside the immediate vicinity the robot location. 
%
%WILL UPDATE ONCE RESULTS ARE BETTER
%The robot explores its surroundings with an occupancy grid with $15000$ cells where grid cell edges are $\alpha=0.2$m, composing a map with dimensions $30\text{m}\times20\text{m}$. The initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) for grid cells covered by the circular robot of radius $0.1$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
%At each time step, the robot receives a measurement scan, where the probabilistic properties of the sensor are taken from~\cite{PirRutBisSch11,KhoElb12}. Then, the $n_c=8$ evenly-spaced candidate locations about a circle of radius $\delta=0.5$m around the current pose location are considered, where $n_d=32$ measurement rays are evenly-spaced about the candidate location.  When no current candidates yield expected information gains above $\mathcal I_\text{min}=2$, $\lambda=1.25$ is multiplied to the current $n_c$ and $\delta$ values. The motion of the robot is restricted to movement on grid cells satisfying \refeqn{CollisionInequalityConstraint} with $\beta=0.01$. Dijkstra's algorithm guides the robot through the environment without collision. The results are illustrated in Figure \ref{fig:AutonomousExploration}.

\section{Experimental Results}
\label{sec:ExpRes}
	% Kinect result

%A preliminary experiment shows how the exact inverse sensor model is successfully implemented with the Kinect sensor.

The efficacy of the proposed mapping and autonomous exploration algorithms are also verified with experiments. The experiment involves a ground vehicle producing a probabilistic occupancy grid map in real time. The robot motion is governed by the information gain-maximizing policy described in this paper.

\paragraph{Hardware Configuration}
The Pioneer 3 ground robot was chosen. The vehicle accepts two inputs, namely linear velocity (aligned with the robot wheels) and angular velocity (about a central axis passing vertically through the robot). The robot pose was estimated via Vicon Tracker, which provides the location and attitude of a rigid body. The depth readings were obtained by a Kinect depth sensor. Since the experiment was to map and explore a 2D environment, only a single central row of the 3D Kinect depth scan provided the necessary measurements.
An environment with features of walls and obstacles was constructed by Styrofoam.

\paragraph{Software Configuration}
The proposed algorithm was implemented with ROS. The ROS framework provided seamless communication betweens sensors and actuators, and the executable programs (referred to as ``nodes'') operate quickly when compiled from C++ code. The robot and map are plotted using ``OpenGL'' libraries.

Synchronization between the sensors is of high importance to the proposed occupancy grid mapping approach. Any time delay between the robot pose estimation and the sensor depth readings can cause conflicting information, which may be harmful to the probabilistic map. Thus, ROS approximate message filters were applied such that the pose from Vicon Tracker ($100$Hz) and the Microsoft Kinect ($30$Hz) were nearly synchronized, which provided validity of the assumption that the measurement ray positions and directions are known deterministically.

\paragraph{Exploring and Mapping a 2D Environment}
The environment consisted of Styrofoam walls on a flat floor. The parameter was oriented in a rectangle, with several obstacles and angled walls (see Figure \ref{fig:ExpSetupPhoto}). The robot autonomously explored the space using the complete Cartesian searching approach, and Dijkstra's algorithm provided collision-free waypoints to the optimal future poses. Based on these waypoints, a constrained least squares polynomial fitting method provided a smooth trajectory to follow the waypoints with fixed velocity. Finally, a simple controller was implemented such that the Pioneer moved along the trajectory by moving and orienting toward the desired robot location $1$sec in the future.

\begin{figure}
	\centering
    	\begin{subfigure}[b]{0.35\textwidth}
        		\includegraphics[width=\textwidth]{test_setup_1.jpg}
        		\caption{Bottom-left view}
        		\label{fig:Experiment_blv}
    	\end{subfigure}
	\hspace*{0.05\columnwidth}
	\begin{subfigure}[b]{0.35\textwidth}
        		\includegraphics[width=\textwidth]{test_setup_2.jpg}
        		\caption{Bottom-right view}
        		\label{fig:Experiment_brv}
    	\end{subfigure}
\caption{Images from two perspectives show the walls and obstacles of the experimental environment.}
\label{fig:ExpSetupPhoto}
\end{figure}

The resulting occupancy grid maps and trajectories are depicted in Figure \ref{fig:ExperimentOGM}. Most importantly, the robot builds a clear occupancy grid map despite sensor noises and imperfect sensor synchronization. The autonomous exploration successfully guides the robot such that it builds a map of the reachable space without colliding with any obstacles.

\begin{figure}
	\centering{
    	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_start.jpg}
        		\caption{$t=0.0$min}
        		\label{fig:Experiment_ogm_t0}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_0p5min.jpg}
        		\caption{$t=0.5$min}
        		\label{fig:Experiment_ogm_t0p5}
    	\end{subfigure}    
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1min.jpg}
        		\caption{$t=1.0$min}
        		\label{fig:Experiment_ogm_t1}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1p5min.jpg}
        		\caption{$t=1.5$min}
        		\label{fig:Experiment_ogm_t1p5}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2min.jpg}
        		\caption{$t=2.0$min}
        		\label{fig:Experiment_ogm_t2}
    	\end{subfigure}
		\vspace*{0.05\textwidth}

	}
	\centering{
    	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2p5min.jpg}
        		\caption{$t=2.5$min}
        		\label{fig:Experiment_ogm_t2p5}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3min.jpg}
        		\caption{$t=3.0$min}
        		\label{fig:Experiment_ogm_t3}
    	\end{subfigure}    
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3p5min.jpg}
        		\caption{$t=3.5$min}
        		\label{fig:Experiment_ogm_t3p5}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4min.jpg}
        		\caption{$t=4.0$min}
        		\label{fig:Experiment_ogm_t4}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4p5min.jpg}
        		\caption{$t=4.5$min}
        		\label{fig:Experiment_ogm_t4p5}
    	\end{subfigure}
	}
\caption{A robot autonomously explores the experimental environment and produces an occupancy grid map along the way. The robot (shaded green circle body with shaded blue circle sensor) is controlled toward the desired pose $1$sec in the future (unshaded grey robot), which is following a constrained least-squares trajectory (cyan curve). This trajectory is based on waypoints (black circles) from Dijkstra's algorithm to arrive at the optimal pose (unshaded green circle body with unshaded blue circle sensor) as proposed in this paper.}
\label{fig:ExperimentOGM}
\end{figure}

The autonomous exploration is governed by a policy to maximize expected entropy decreases, so entropy and entropy change with time (Figure \ref{fig:ExperimentH}) are important metrics. In this example, the a priori occupancy probabilities of all cells are $0.1$, so the entropies could temporarily rise for likely occupied cells. The total map entropy begins at roughly $2613$ and decreases by roughly half to $1338$. Many grid cells within the map limits are occluded by walls, preventing the blocked cells from occupancy probability changes. In some cases, the robot is able to update the cell occupancy probabilities; in other cases, cells fall inside regions occluded from all reachable space, and the robot cannot update these cells. Among those grid cells visible from collision-free locations, the vast majority of cells were updated in less than $5$min while following a very slow ($5$cm/sec) trajectory.

%Additionally, even though the walls appear fairly well-defined, they are not pefectly straight boundaries. The occupancy grid assumptions do not perfectly align with the true environment; an edge of a wall might occupy only part of a grid cell. Therefore, some measurement rays passing through this cell provide evidence that the cell is free, whereas other rays might provide conflicting evidence that the cell is occupied. Furthermore, any error in robot attitude estimation can degrade the map; an attitude error of $1^{\circ}$ corresponds to an offset of $11$ measurement rays propagating across the $640$ measurement rays composing the Kinect depth scan.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{experiment_H_per_min.pdf}
	\caption{The map entropy $H$ averaged over all $90601$ grid cells is reduced during experimentation. The quick changes $\Delta H$ per cell correspond to map updates when the robot captures new territory.}
\label{fig:ExperimentH}
\end{figure}

\section{Conclusions}
\label{sec:Conclusion}

This paper serves to improve the accuracy occupancy grid mapping and the policy governing robotic motion during autonomous exploration.
With other approaches, each probability of an occupancy grid map is approximated in practice due to the perceived complications of the exact solution.
This paper provides an exact solution to this complicated probability problem using cell occlusion in the forward sensor model. This yields a substantially simpler inverse sensor model, which avoids a potentially harmful Markov assumption that commonly appears in the log-odd ratio formulations. The main contribution of this paper is that all of the information available in the measurements and the a priori estimate is extracted and integrated to generate a more accurate a posteriori map. The numerical and experimental results show that maps using this exact approach are substantially improved.

The exact probabilistic occupancy grid mapping framework is utilized in autonomous exploration. The robot pose is chosen to minimize the expected map information gain, which is now directly solved more precisely in an unprecedented way. Unlike common techniques based on finding frontiers or predicting measurement scans with particle filters, the proposed approach is based on an expected value of the entropy directly. The approach is approximated with large improvements in computation and small losses in accuracy. Numerical examples and an experiment show how robots navigate uncertain maps governed by the proposed autonomous exploration algorithm while building accurate probabilistic maps.

\section*{Appendix}\label{append}

\subsection*{Proof of Proposition 1}

\paragraph{Unnormalized Reduced Map Inverse Sensor Model}
For the reduced map of the $l$-th ray from \refeqn{InvSenModWithProbDens}, the unnormalized part corresponding to terms without $\eta_{t,l}$ in \refeqn{RayISMAnswer} is %given by
\begin{align}
\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\
= \sum_{r\in\mathcal{M}_{k}} p(z_{t,l}|r,X_t) P(r|X_{1:t-1},Z_{1:t-1}).\label{eqn:tildePlk}
\end{align}
Recall $\mathcal{M}_k$ corresponds to the set of maps where the $k$-th cell is occupied. Define a subset $\mathcal{N}_{i,k}\subset \mathcal{M}_k$ for $1\leq i\leq k$ be the set of maps where the $i$-th cell is the first occupied cell. More explicitly, 
\begin{align*}
\mathcal{N}_{i,k}  = \{ r\in\mathcal{M}_k| \mathbf{r}_{l,i+}\}
= \{ r\in\mathcal{M}_k| \mathbf{r}_{l,1}=0,\ldots, \mathbf{r}_{l,i-1}=0, \mathbf{r}_{l,i}=1,
\mathbf{r}_{l,k}=1\}.
\end{align*}
Then, $\mathcal{M}_k$ can be written as $\mathcal{M}_k =\bigcup_{i=1}^{k} \mathcal{N}_{i,k}$. Using this, the summation over $\mathcal{M}_k$ at \refeqn{tildePlk} can be decomposed of the summation over each $\mathcal{N}_{i,k}$ to obtain
\begin{align*}
\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}) = \sum_{i=1}^k\braces{\sum_{r\in\mathcal{N}_{i,k}} p(z_{t,l}|r,X_t) P(r|X_{1:t-1},Z_{1:t-1})}.
\end{align*}
This is motivated by the fact that the forward sensor model $p(z_{t,l}|r,X_t)$ is identical for all maps in $\mathcal{N}_{i,k}$, such that $p(z_{t,l}|r,X_t)=p(z_{t,l}|\mathbf{r}_{l,i+},X_t)$ and it is moved left of the summation to obtain
\begin{align}
\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})= \sum_{i=1}^k \braces{p(z_{t,l}|\mathbf{r}_{l,i+},X_t) \sum_{r\in\mathcal{N}_{i,k}} P(r|X_{1:t-1},Z_{1:t-1})}.\label{eqn:tildePlk0}
\end{align}
The last term of the above expression corresponds to the a priori probability of $\mathcal{N}_{i,k}$. When $i<k$, it is given by
\begin{align}
\sum_{r\in\mathcal{N}_{i,k}}  P(r|X_{1:t-1},Z_{1:t-1}) &= \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
 \times &P({\mathbf{r}}_{l,i}|X_{1:t-1},Z_{1:t-1})P({\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1}),\label{eqn:PNik}
\end{align}
and when $i=k$, 
\begin{align}
\sum_{r\in\mathcal{N}_{k,k}} P(r|&X_{1:t-1},Z_{1:t-1})\nonumber\\&= \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
%&\quad \times 
P({\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1}).\label{eqn:PNkk}
\end{align}
Substituting \refeqn{PNik} and \refeqn{PNkk} into \refeqn{tildePlk0}, we obtain \refeqn{Unnormalized}.

\paragraph{Complement of the Unnormalized Reduced Map Inverse Sensor Model}
An analytic expression for the complement of the unnormalized inverse sensor model is also required  for obtaining the normalizer of the $l$-th ray at time $t$, namely $\eta_{t,l}$. Let $\bar{\mathcal{M}}_k$ be the set of maps where the $k$-th cell is unoccupied, i.e., $\bar{\mathcal{M}}_k = \{ r\in\{0,1\}^{n_{r,l}}\,|\, \mathbf{r}_{l,k}=0\}$. Similar to \refeqn{tildePlk},
\begin{align}
\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})= \sum_{r\in\bar{\mathcal{M}}_{k}} p(z_{t,l}|r,X_t) P(r|X_{1:t-1},Z_{1:t-1}).\label{eqn:tildePbarlk}
\end{align}

Let $\bar{\mathcal{N}}_{i,k}\subset \bar{\mathcal{M}}_k $ for $1\leq i\leq n_{r,l}$ and $i\neq k$ be the set of maps where the $i$-th cell is the first occupied cell. More explicitly, 
\begin{align*}
\bar{\mathcal{N}}_{i,k}=\{r\in\bar{\mathcal{M}}_k\,|\, \mathbf{r}_{l,1}=0,\ldots,\mathbf{r}_{l,i-1}=0,
\mathbf{r}_{l,i}=1,\mathbf{r}_{l,k}=0\}.
\end{align*}
Then, we have $\bar{\mathcal{M}}_k=\bigcup_{\substack{i=1\\i\neq k}}^{n_{r,l}} \bar{\mathcal{N}}_{i,k}$. Note that the forward sensor model $p(z_{t,l}|r,X_t)$ is identical for any maps in $\bar{\mathcal{N}}_{i,k}$, such that $p(z_{t,l}|r,X_t)=p(z_{t,l}|\mathbf{r}_{l,i+},X_t)$. Similar to \refeqn{tildePlk0},
\begin{align}
\tilde P&(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=\sum_{\substack{i=1\\i\neq k}}^{n_{r,l}} \braces{p(z_{t,l}|\mathbf{r}_{l,i+},X_t) \sum_{r\in\bar{\mathcal{N}}_{i,k}} P(r|X_{1:t-1},Z_{1:t-1})},\label{eqn:tildePbarlk0}
\end{align}
where the last term corresponds to the a priori probability of $\bar{\mathcal{N}}_{i,k}$. When $i<k$, it is given by
\begin{align}
\sum_{r\in\bar{\mathcal{N}}_{i,k}} P(r|X_{1:t-1},&Z_{1:t-1}) = \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
&\quad \times P({\mathbf{r}}_{l,i}|X_{1:t-1},Z_{1:t-1})P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1}),\label{eqn:PbarNik1}
\end{align}
and when $k<i$,
\begin{align}
\sum_{r\in\bar{\mathcal{N}}_{i,k}} & P(r|X_{1:t-1},Z_{1:t-1}) = \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\} P({\mathbf{r}}_{l,i}|X_{1:t-1},Z_{1:t-1}).\label{eqn:PbarNik2}
\end{align}
Substituting \refeqn{PbarNik1} and \refeqn{PbarNik2} into \refeqn{tildePbarlk0}, we obtain
\begin{align}
&\tilde P (\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})
=P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\
&\quad\times \bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]
\nonumber
\\
&\quad
+
\bigg[\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg],\label{eqn:tildePbar}
\end{align}
where $p(z_{t,l}|\mathbf{r}_{(n_r+1)+},X_t)$ corresponds to the forward sensor model of maximum reading and $P(\mathbf{r}_{l,n_{r,l}+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience for the empty map case.

\paragraph{Normalizer}
We have 
\begin{align*}
P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})+
P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=1.
\end{align*}
Since they share the same normalizer $\eta_{t,l}$, this implies 
\begin{align*}
&\eta_{t,l}=\frac1{\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})+
\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}.
\end{align*}
Substituting \refeqn{Unnormalized} and \refeqn{tildePbar}, and rearranging, we obtain \refeqn{allEta}.

\bibliography{BibSources}
\bibliographystyle{IEEEtran}

%% For one-column wide figures use
%\begin{figure}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:1}       % Give a unique label
%\end{figure}
%%
%% For two-column wide figures use
%\begin{figure*}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics[width=0.75\textwidth]{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:2}       % Give a unique label
%\end{figure*}
%%
%% For tables use
%\begin{table}
%% table caption is above the table
%\caption{Please write your table caption here}
%\label{tab:1}       % Give a unique label
%% For LaTeX tables use
%\begin{tabular}{lll}
%\hline\noalign{\smallskip}
%first & second & third  \\
%\noalign{\smallskip}\hline\noalign{\smallskip}
%number & number & number \\
%number & number & number \\
%\noalign{\smallskip}\hline
%\end{tabular}
%\end{table}

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

\end{document}
% end of file template.tex

### 2018-06-09 Journal/JINT 2017/JIRS16_ans1.tex
journal added

\documentclass[11pt]{article}
\usepackage[letterpaper,text={6.5in,8.6in},centering]{geometry}
\usepackage{amssymb,amsmath,times,url}
\usepackage{xr,color}
\usepackage{hyperref}

\externaldocument[]{JIRS16_rev1}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand{\refprop}[1]{Proposition \ref{prop:#1}}

\newcommand{\RNum}[1]{\uppercase\expandafter{\romannumeral #1\relax}}
\newcommand{\RI}{\text{\RNum{1}}}
\newcommand{\RII}{\text{\RNum{2}}}
\newcommand{\RIII}{\text{\RNum{3}}}

\newenvironment{correction}{\begin{list}{}{\setlength{\leftmargin}{1cm}\setlength{\rightmargin}{1cm}}\vspace{\parsep}\item[]``}{''\end{list}}

\newcommand{\EditTL}[1]{{\color{red}\protect #1}}

\begin{document}

%\pagestyle{empty}

\section*{Response to the Reviewers' Comments for JINT-D-16-00659}

The authors would like to thank the reviewers for their thoughtful comments, which are aimed toward improving the quality of the paper and the clarity of the results. In accordance with the comments and suggestions, the paper has been revised as follows.

\subsection*{Reviewer 1}

\setlength{\leftmargini}{0pt}
\begin{itemize}\setlength{\itemsep}{2\parsep}

\item {\itshape Reviewer \#1: The paper presents `probabilistic occupancy grid mapping and motion planning approaches such that a robot may build autonomously a map and explore a target area  in real time'. The sensor characteristics are explored probabilistically and the approaches are presented and summarized using algorithms. An Analysis of  computational efficient  is also presented together with numerical and experimental results. The news are described by authors and the experimental results help the readers to understanding the contributions.

However, I have some questions for the authors;}
\end{itemize}

%\renewcommand{\labelenumi}{\theenumi}
%\renewcommand{\theenumi}{(\roman{enumi})}%

\begin{enumerate}
\item {\itshape Reviewer \#1: Figure 1 and the explanation in subsection ``Computationally Efficient Approach'' is not so clear for readers. The authors work with bi-dimensional space and present in figure an example in one-dimensional space. Also, the explanation in subsection is not clear and the second paragraph of the subsection is concluded with ``we present the following computational algorithm to compute the inverse sensor model'', but only the ``Proposition 1'' is seen in sequence.}

The proposed algorithm to compute the exact inverse sensor model is applied to any of one, two, or three-dimensional depth measurements and mappings. The authors decided to use one-dimensional example in Figure 1, only to illustrate the proposed approach based on cell occlusion more explicitly through simpler cases.

To clarify these, a few sentences are added to explain how the 1D concept is extended to 2D maps. Additionally, several other minor changes are added for clarity. Also, the last statement refers to Section \ref{sec:SA} for a summary of the proposed computational algorithm.

\begin{correction}
%
We propose a computational algorithm to evaluate \refeqn{InvSenModWithProbDens} efficiently. 
Since the cells outside of the sensor field of view (FOV) are not affected, we focus on a reduced map $r_l$ in the FOV of the $l$-th ray. This reduced map is chosen such that each cell of $r_l$ corresponds to a grid cell of map $m$ that the $l$-th ray intersects, ordered by increasing distance, which is easily determined from geometry. Let $\mathbf{r}_{l,k}$ be the binary random variable representing the occupancy of the $k$-th cell of the $l$-th ray. The number of cells in the reduced map is denoted by $n_{r,l}\leq n_m$.

Next, let $\mathbf{r}_{l,k+}$ correspond to the event that the $k$-th cell of the $l$-th ray is occupied, cells with lower index (closer cells) are free, and cells with greater index (farther cells) may or may not be occupied, i.e., event $\mathbf{r}_{l,k+}$ occurs when \\$r\in\braces{r\in\{0,1\}^{{n_{r,l}}}\,|\mathbf{r}_{l,1}=0,\mathbf{r}_{l,2}=0,\ldots,\mathbf{r}_{l,k-1}=0,\mathbf{r}_{l,k}=1}$. In other words, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray. For example, see Figure \ref{fig:show_rkplus} that illustrates the event of $\mathbf{r}_{l,k+}$ when $l=1$ and $k=4$ for an one-dimensional cell array. This concept of grouping map outcomes can be easily extended to 2D using ray casting, where a 1D ray is spanning 2D space. The cells considered for the measurement ray are determined with the geometric intersection between the ray and map cell edges.
Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell. Based on this, we present the mathematical expressions for the exact inverse sensor model, as summarized by \refeqn{RayISMAnswer} in Proposition 1. Later in Section \ref{sec:SA}, these are also rearranged as a computational algorithm. 
\end{correction}

%The algorithm was written sequentially with LaTeX, but placed awkwardly in the PDF, which is now resolved.

\item {\itshape Reviewer \#1: In page 13, the authors present two algorithms as table. Well, algorithms can not be table and seemly the authors are using latex. So, the authors must use the  algorithm2e package to present all algorithms into the paper and also to citing them suitably.}

Both algorithms are rewritten using the suggested \texttt{algorithm2e} package and they are referenced properly in the revised manuscript.

\item {\itshape Reviewer \#1: About the computational analysis, the authors talk in introduction of the paper that the approaches are to be used in real time. However, discussion about real time questions are not done into paper. For example: what information about each cell are stored in the memory in real time? For the experiment, what the computational load of the algorithms were? what time slice (tick) was? how much memories were used in the examples presented in figures 8 and 10? Well, I think that the discussion about real time could be improved.}\\

The following discussion on real-time implementation and memory is added to Section 7 of the experimental results.

%\EditTL{It is common to specify the processor and the programming language when presenting computation time. We may also add a remark that those were suitable for our experiments.}

\begin{correction}The resulting map and exploration commands are generated in real time. The grid cells are stored as double-float variables in a vector, where the vector index is mapped to a location on the occupancy grid. Thus, the memory requirements are proportional to the number of grid cells, which is roughly the mapping area divided by the area of a grid cell. Since cell locations need not be saved, memory requirements are reduced. On a Lenovo T540p laptop with an Intel Core i7-4900MQ processor (quad-core, 2.8GHz per core) and 16GB of RAM, the mean time for updating the occupancy grid is $0.0115$ seconds, and the mean time to determine an exploration strategy is $0.6820$ seconds. Time requirements for mapping and exploration are easily modified by changing grid cell resolution or the number of exploration pose candidates, respectively. In short, there is a tradeoff between computational speed and accuracy of mapping or exploration.\end{correction}

\end{enumerate}

%\vspace*{0.1\textwidth}

\clearpage\newpage

\subsection*{Reviewer 2}

\setlength{\leftmargini}{0pt}
\begin{itemize}\setlength{\itemsep}{2\parsep}

\item {\itshape Reviewer \#2: This paper serves to improve the accuracy occupancy grid mapping and the policy governing robotic motion during autonomous exploration.This paper provides an exact solution to this complicated probability problem using cell occlusion in the forward sensor model, which yields a substantially simpler inverse sensor model, which avoids a potentially harmful Markov assumption that commonly appears in the log-odd ratio formulations. There are some minor comments:}
\end{itemize}

\begin{enumerate}\setlength{\itemsep}{2\parsep}

\item {\itshape Reviewer \#2: ``The robot motion is governed by the information gain-maximizing policy described in this paper.'' What is the advantage to maximize the information gain? In the field of motion control, the big control gain is unwanted when designing an algorithm.}

The objective of exploration pursued in this paper is to build an accurate probabilistic map in an optimal fashion. As described in \refeqn{ShannonsEntropyDef}, the entropy of a probabilistic map has been used as a measure of uncertainties. Therefore, the decrease of the entropy due to measurements is regarded as the information gain to obtain a more accurate map. By planning the motion of robots to maximize the information gain, we can construct the complete map in an optimal fashion.

The proposed optimal motion planning scheme to generate the desired path and the pose of the robot should be distinguished from the motion control problem that is applied in the inner loop to physically direct the robot toward the desired path.

In other words, the information gain is separate from control gain. This paper serves to determine where the robot should go (optimal pose) based on its available information of a probabilistic map, not the tracking control inputs along a trajectory to the optimal pose, or which type of controller to use. To clarify this point, the following is added to the last paragraph of Section 2.2.
\begin{correction}It is considered that once a collision-free trajectory toward $X_c^*$ is obtained, the robot moves along the trajectory controlled by a motion control system in the inner loop. Once the robot has translated to $X_c^*$, the above process is repeated.\end{correction}

\item {\itshape Reviewer \#2: Does the methods proposed in this paper can be applied in controlling mobile robots with visual servoing? for example

[1] Chen H, Wang C, Liang Z, et al. Robust practical stabilization of nonholonomic mobile robots based on visual servoing feedback with inputs saturation[J]. Asian Journal of Control, 2014, 16(3): 692-702.

[2] Chen H, Ding S, Chen X, et al. Global finite-time stabilization for nonholonomic mobile robots based on visual servoing[J]. International Journal of Advanced Robotic Systems, 2014, 11(11): 180.

[3] Chen H, Zhang J, Chen B, et al. Global practical stabilization for non-holonomic mobile robots with uncalibrated visual parameters by using a switching controller[J]. IMA Journal of Mathematical Control and Information, 2013: dns044.}

%In the context of this paper, maximizing information gain is based on how expected measurement ray depths determine some action.

As discussed above, the proposed algorithm generates the collision-free paths to construct a probabilistic map in an optimal fashion. Therefore, it can be integrated with any control algorithm, including methods based on visual servoing. To clarify these, the last part of Section 5 has been revised as follows.

%In this sense, a geometric interpretation of visual servoing relates to the proposed research, i.e., an image does not directly control the robot degrees of freedom. However, the main difference is that this paper provides a desired robot trajectory, whereas visual servoing focusses heavily on robotic control.
%So, if the proposed research in this paper were combined with control, the research could be classified under visual servoing with a primary focus on map information.
%The authors chose not to include a particular controller in this paper to maintain generality to the resulting method. The following sentence is added to the motion planning description of Section 5.1.

\begin{correction}Then, a controller on the robot tracks this trajectory until the robot falls within acceptable thresholds of the final optimal pose. Any control scheme for the motion of the robot can be integrated with the proposed exploration algorithm. Once the robot completes this motion, the entire process is repeated.\end{correction}

\end{enumerate}

%\vspace*{0.1\textwidth}

\clearpage\newpage
\subsection*{Reviewer 3}

\begin{itemize}\setlength{\itemsep}{2\parsep}

\item {\itshape Reviewer \#3: Equation (5), determination of beta, collision threshold, should be clarified through numeric/experimental section as well...}

The collision threshold $\beta$ corresponds to the probability of collision that may occur when the robot follows the desired trajectory. As $\beta$ decreases, the chances for collision decreases, but at the time, the planned trajectory may become longer and complicated while keeping a distance from potential obstacles.

%Setting the particular value of $\beta$ depend

%is an important parameter selected to avoid collisions. 
%It is sensitive to grid cell size, since cell resolution determines the number of grid cells falling inside the possible collision region $\mathcal{C}_X$. One practical implication of (5) is that cells with greater occupancy probabilities inside $\mathcal{C}_X$ tend to fall near the perimeter of the region as the robot moves because internal cells are typically well-known to be free assuming the robot covered open space beforehand. This point, and 
The following description about collision probability is added to Section 6 (numerical exploration example), and the same properties are held for Section 7 (experiment).
\begin{correction}For added safety, cells falling within $0.6$m of the robot are considered inside the possible collision zone $\mathcal{C}_X$, from \refeqn{CollisionInequalityConstraint}, where $\beta=0.1$ is selected.\end{correction}

\item {\itshape Reviewer \#3: Just before equation (7) forward sensor model p(...) should be explained more.}

The following is added to the ``Bayesian Framework'' paragraph of Section 3.2.

\begin{correction}The occupancy probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$ is based on the forward sensor model $p(z_{t,l}|m,X_{1:t},Z_{1:t-1})$ that describe the distribution of the measurements for the given robot pose and the map. This specifies the stochastic characteristics of the sensor, such as the mean squared error, and it is assumed as a known distribution, specific to a particular depth sensor. One could certainly obtain a stochastic model by fitting a probability density to a controlled set of measurements.
\end{correction}

\item {\itshape Reviewer \#3: Likewise, in the Proposition 1, ``$p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading'' should be explained more rigorously through the Appendix 1 as well.}

The statement in Proposition 1 is revised as follows.

\begin{correction}$p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the probability density of the measurement when all of cells in the field of view of the $l$-th ray is not occupied.\end{correction}

Referring to the maximum reading, the following explanation is added after (34).

\begin{correction}For this special case, probability density can be easily represented with a uniform distribution over a short distance. The properties of this distribution depend on the depth sensor fidelity, because a maximum reading may indicate that all cells within the sensor range are empty, but it might also represent a failed reading.\end{correction}

\item {\itshape Reviewer \#3: Uncertainty on position and attitude, cell size adjustment can be discussed at the beginning.}

Accounting for localization is very important, particularly in conjunction with mapping and autonomous exploration, though this is beyond the paper scope. The following is added to the Section 1 (introduction).

\begin{correction}Accounting for the stochastic properties of position and attitude is important to the field of robotics, and accounting for localization in conjunction with mapping and exploration are considered as a future step to this paper.\end{correction}

A brief explanation of grid cell size is added to the first paragraph of Section 2.1.

\begin{correction}The location and size of each grid cell is assumed known, where a smaller cell size (greater grid resolution) better represents a space, but increases computation and memory. \end{correction}

Further discussion on cell resolution is addressed two comments below.

\item {\itshape Reviewer \#3: Section 5, motion planning should cite the following paper: (grid-based, Dijkstra)}

N. Gasilov, M. Dogan, V. Arici, `Two-stage shortest path algorithm for solving optimal obstacle avoidance problem', IETE Journal of Research, 57(3), 278-285, 2011, doi:10.4103/0377-2063.83650.

The reference is added.

%\EditTL{Do we need a brief comparison with this?}

\item {\itshape Reviewer \#3: Choosing the optimal parameters for each case, e.g. ray numbers, resolution in Table 1, can be explained more to understand the proposed algorithm.}

The following is added in reference to Table 1 in Section 6.1.

\begin{correction}Among these parameters, cell resolution and the number of rays have the greatest impact on computation. In general, the finer the grid, the greater the memory and computational requirements are because the computational order of (10) grows linearly with the number of grid cells inside the sensor range limits. Since the inverse sensor models are combined sequentially as shown in (13), the number of rays considered are proportional to the computation order as well.\end{correction}

\item {\itshape Reviewer \#3: Convergence rate of algorithm, or speed of robot while achieving its goal, should be discussed, e.g. possible extensions or improvements to overcome some unnecessary repetitions and back-and-forth small step size bottlenecks.}

Thanks for the comments. This is an important point, especially as it relates to future goals of the research. The following is added as the last paragraph of Section 7 (experiment).

\begin{correction}Both the numerical and experimental results demonstrated successful exploration strategies, though the algorithm can be improved. In particular, the robot sometimes revisits spaces that it has explored before, most frequently when the robot did not take measurements toward a particular direction. A few methods to improve this issue are treated as future work. One options is adding cost functions (e.g. distance based, number of past measurements inside an area) to prevent the robot from leaving an area before it is well-known. A second option is partitioning the space into sections, where the robot moves toward sections with large numbers of pose candidates with high expected information gains. Another option is reevaluating the future pose optimization while the robot is translating between poses because the occupancy grid map is changing, and so is expected information gain. The efficacy of these proposed changes will be evaluated, particularly as they relate to more useful systems, such as 3D map generation and multi-robot exploration.\end{correction}

\end{itemize}

\end{document}

### 2018-06-09 Journal/JINT 2017/JIRS16_ans2.tex
journal added

--- 2018-06-09 Journal/JINT 2017/JIRS16_ans1.tex
@@
 \subsection*{Reviewer 1}
-
-\setlength{\leftmargini}{0pt}
-\begin{itemize}\setlength{\itemsep}{2\parsep}
-
-\item {\itshape Reviewer \#1: The paper presents `probabilistic occupancy grid mapping and motion planning approaches such that a robot may build autonomously a map and explore a target area  in real time'.
-The sensor characteristics are explored probabilistically and the approaches are presented and summarized using algorithms.
-An Analysis of  computational efficient  is also presented together with numerical and experimental results.
-The news are described by authors and the experimental results help the readers to understanding the contributions.
-
-However, I have some questions for the authors;}
+\begin{itemize}
+\item The authors have payed attention to suggestions in my previous review. I do not have new questions to him.
 \end{itemize}
 
-%\renewcommand{\labelenumi}{\theenumi}
-%\renewcommand{\theenumi}{(\roman{enumi})}%
-
-\begin{enumerate}
-\item {\itshape Reviewer \#1: Figure 1 and the explanation in subsection ``Computationally Efficient Approach'' is not so clear for readers.
-The authors work with bi-dimensional space and present in figure an example in one-dimensional space.
-Also, the explanation in subsection is not clear and the second paragraph of the subsection is concluded with ``we present the following computational algorithm to compute the inverse sensor model'', but only the ``Proposition 1'' is seen in sequence.}
-
-The proposed algorithm to compute the exact inverse sensor model is applied to any of one, two, or three-dimensional depth measurements and mappings.
-The authors decided to use one-dimensional example in Figure 1, only to illustrate the proposed approach based on cell occlusion more explicitly through simpler cases.
-
-To clarify these, a few sentences are added to explain how the 1D concept is extended to 2D maps.
-Additionally, several other minor changes are added for clarity.
-Also, the last statement refers to Section \ref{sec:SA} for a summary of the proposed computational algorithm.
-
-\begin{correction}
-%
-We propose a computational algorithm to evaluate \refeqn{InvSenModWithProbDens} efficiently. 
-Since the cells outside of the sensor field of view (FOV) are not affected, we focus on a reduced map $r_l$ in the FOV of the $l$-th ray.
-This reduced map is chosen such that each cell of $r_l$ corresponds to a grid cell of map $m$ that the $l$-th ray intersects, ordered by increasing distance, which is easily determined from geometry.
-Let $\mathbf{r}_{l,k}$ be the binary random variable representing the occupancy of the $k$-th cell of the $l$-th ray.
-The number of cells in the reduced map is denoted by $n_{r,l}\leq n_m$.
-
-Next, let $\mathbf{r}_{l,k+}$ correspond to the event that the $k$-th cell of the $l$-th ray is occupied, cells with lower index (closer cells) are free, and cells with greater index (farther cells) may or may not be occupied, i.e., event $\mathbf{r}_{l,k+}$ occurs when \\$r\in\braces{r\in\{0,1\}^{{n_{r,l}}}\,|\mathbf{r}_{l,1}=0,\mathbf{r}_{l,2}=0,\ldots,\mathbf{r}_{l,k-1}=0,\mathbf{r}_{l,k}=1}$.
-In other words, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray.
-For example, see Figure \ref{fig:show_rkplus} that illustrates the event of $\mathbf{r}_{l,k+}$ when $l=1$ and $k=4$ for an one-dimensional cell array.
-This concept of grouping map outcomes can be easily extended to 2D using ray casting, where a 1D ray is spanning 2D space.
-The cells considered for the measurement ray are determined with the geometric intersection between the ray and map cell edges.
-Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell.
-Based on this, we present the mathematical expressions for the exact inverse sensor model, as summarized by \refeqn{RayISMAnswer} in Proposition 1.
-Later in Section \ref{sec:SA}, these are also rearranged as a computational algorithm.
-\end{correction}
-
-%The algorithm was written sequentially with LaTeX, but placed awkwardly in the PDF, which is now resolved.
-
-\item {\itshape Reviewer \#1: In page 13, the authors present two algorithms as table.
-Well, algorithms can not be table and seemly the authors are using latex.
-So, the authors must use the  algorithm2e package to present all algorithms into the paper and also to citing them suitably.}
-
-Both algorithms are rewritten using the suggested \texttt{algorithm2e} package and they are referenced properly in the revised manuscript.
-
-\item {\itshape Reviewer \#1: About the computational analysis, the authors talk in introduction of the paper that the approaches are to be used in real time.
-However, discussion about real time questions are not done into paper.
-For example: what information about each cell are stored in the memory in real time?
-For the experiment, what the computational load of the algorithms were?
-what time slice (tick) was?
-how much memories were used in the examples presented in figures 8 and 10?
-Well, I think that the discussion about real time could be improved.}\\
-
-The following discussion on real-time implementation and memory is added to Section 7 of the experimental results.
-
-%\EditTL{It is common to specify the processor and the programming language when presenting computation time. We may also add a remark that those were suitable for our experiments.}
-
-\begin{correction}The resulting map and exploration commands are generated in real time.
-The grid cells are stored as double-float variables in a vector, where the vector index is mapped to a location on the occupancy grid.
-Thus, the memory requirements are proportional to the number of grid cells, which is roughly the mapping area divided by the area of a grid cell.
-Since cell locations need not be saved, memory requirements are reduced.
-On a Lenovo T540p laptop with an Intel Core i7-4900MQ processor (quad-core, 2.8GHz per core) and 16GB of RAM, the mean time for updating the occupancy grid is $0.0115$ seconds, and the mean time to determine an exploration strategy is $0.6820$ seconds.
-Time requirements for mapping and exploration are easily modified by changing grid cell resolution or the number of exploration pose candidates, respectively.
-In short, there is a tradeoff between computational speed and accuracy of mapping or exploration.\end{correction}
-
-\end{enumerate}
-
-%\vspace*{0.1\textwidth}
-
-\clearpage\newpage
+The authors appreciate your questions and suggestions, which have improved the paper quality.
 
@@
 \setlength{\leftmargini}{0pt}
-\begin{itemize}\setlength{\itemsep}{2\parsep}
-
-\item {\itshape Reviewer \#2: This paper serves to improve the accuracy occupancy grid mapping and the policy governing robotic motion during autonomous exploration.This paper provides an exact solution to this complicated probability problem using cell occlusion in the forward sensor model, which yields a substantially simpler inverse sensor model, which avoids a potentially harmful Markov assumption that commonly appears in the log-odd ratio formulations.
-There are some minor comments:}
-\end{itemize}
 
@@
 
-\item {\itshape Reviewer \#2: ``The robot motion is governed by the information gain-maximizing policy described in this paper.'' What is the advantage to maximize the information gain?
-In the field of motion control, the big control gain is unwanted when designing an algorithm.}
+\item {\itshape Reviewer \#2: Under the exact probabilistic occupancy grid mapping framework, how to propose a simple controller to complete the control task such as stabilization or tracking problem?}
 
-The objective of exploration pursued in this paper is to build an accurate probabilistic map in an optimal fashion.
-As described in \refeqn{ShannonsEntropyDef}, the entropy of a probabilistic map has been used as a measure of uncertainties.
-Therefore, the decrease of the entropy due to measurements is regarded as the information gain to obtain a more accurate map.
-By planning the motion of robots to maximize the information gain, we can construct the complete map in an optimal fashion.
+The focus of this paper is probabilistic mapping and autonomous exploration, not control system design for stabilization or tracking.
+In fact, the exploration algorithm serves to determine an optimal trajectory for the robot as it gathers knowledge about its environment, and as such, it deals with a guidance problem.
+Designing a specific control system to follow this trajectory may be achieved by a wide range of controllers, and it is out of scope of this paper.
 
-The proposed optimal motion planning scheme to generate the desired path and the pose of the robot should be distinguished from the motion control problem that is applied in the inner loop to physically direct the robot toward the desired path.
+To clarify this, the end of Section 5 is modified as follows.
 
-In other words, the information gain is separate from control gain.
-This paper serves to determine where the robot should go (optimal pose) based on its available information of a probabilistic map, not the tracking control inputs along a trajectory to the optimal pose, or which type of controller to use.
-To clarify this point, the following is added to the last paragraph of Section 2.2.
-\begin{correction}It is considered that once a collision-free trajectory toward $X_c^*$ is obtained, the robot moves along the trajectory controlled by a motion control system in the inner loop.
-Once the robot has translated to $X_c^*$, the above process is repeated.\end{correction}
+\begin{correction}The starting and ending positions and attitudes may be constrained, and polynomials patched together for long trajectories share a common position and velocity with respect to time.
+Then, the robot tracks this trajectory until the robot falls within acceptable thresholds of the final optimal pose.
+For generality of the proposed exploration algorithm, any position controller yielding robotic motion that follows this desired trajectory may be selected.
+Once the robot completes this motion, the entire process is repeated.\end{correction}
 
-\item {\itshape Reviewer \#2: Does the methods proposed in this paper can be applied in controlling mobile robots with visual servoing? for example
+\item {\itshape Reviewer \#2: I suggest add a simple application to the visual servoing control design problem of robotic system by using the grid mapping inverse sensor models as an additional experiment to highlight its utility.}
 
-[1] Chen H, Wang C, Liang Z, et al.
+Visual servoing involves using visual information to control the pose of a robot precisely relative to a target object, such as controlling the pose of a vehicle relative to a landmark.
+While this important field supports numerous applications in estimation and control of robotic systems, it is distant from the submitted manuscript, focused on probabilistic mapping and motion planning.
+
+Certainly, one may apply both visual servoing and the proposed mapping algorithm together, as a control scheme and a motion planning scheme, respectively.
+However, as discussed above, control system design is beyond the focus of the manuscript, and illustrating the efficacy of the proposed exploration algorithm with another type of control system will not support the main contribution of the submitted manuscript.
+
+Instead, the following is added to the end of Section 3.3 of the revised manuscript.
+
+\begin{correction}This proposed mapping approach follows a Bayesian framework that focusses on occupied and free space, not tracking particular systems or mapping features, such as visual servoing [17].
+The proposed algorithm may be combined with feature-based approaches for object tracking or stabilization purposes, with the added challenge of associating occupied spaces with features.\end{correction}
+
+%[1] Chen H, Wang C, Liang Z, et al.
 Robust practical stabilization of nonholonomic mobile robots based on visual servoing feedback with inputs saturation[J].
 Asian Journal of Control, 2014, 16(3): 692-702.
-
-[2] Chen H, Ding S, Chen X, et al. Global finite-time stabilization for nonholonomic mobile robots based on visual servoing[J]. International Journal of Advanced Robotic Systems, 2014, 11(11): 180.
-
-[3] Chen H, Zhang J, Chen B, et al.
+%
+%[2] Chen H, Ding S, Chen X, et al. Global finite-time stabilization for nonholonomic mobile robots based on visual servoing[J]. International Journal of Advanced Robotic Systems, 2014, 11(11): 180.
+%
+%[3] Chen H, Zhang J, Chen B, et al.
 Global practical stabilization for non-holonomic mobile robots with uncalibrated visual parameters by using a switching controller[J].
-IMA Journal of Mathematical Control and Information, 2013: dns044.}
-
-%In the context of this paper, maximizing information gain is based on how expected measurement ray depths determine some action.
-
-As discussed above, the proposed algorithm generates the collision-free paths to construct a probabilistic map in an optimal fashion.
-Therefore, it can be integrated with any control algorithm, including methods based on visual servoing.
-To clarify these, the last part of Section 5 has been revised as follows.
-
-%In this sense, a geometric interpretation of visual servoing relates to the proposed research, i.e., an image does not directly control the robot degrees of freedom.
-However, the main difference is that this paper provides a desired robot trajectory, whereas visual servoing focusses heavily on robotic control.
-%So, if the proposed research in this paper were combined with control, the research could be classified under visual servoing with a primary focus on map information.
-%The authors chose not to include a particular controller in this paper to maintain generality to the resulting method.
-The following sentence is added to the motion planning description of Section 5.1.
-
-\begin{correction}Then, a controller on the robot tracks this trajectory until the robot falls within acceptable thresholds of the final optimal pose.
-Any control scheme for the motion of the robot can be integrated with the proposed exploration algorithm.
-Once the robot completes this motion, the entire process is repeated.\end{correction}
+IMA Journal of Mathematical Control and Information, 2013: dns044.
 
@@
 
-%\vspace*{0.1\textwidth}
-
-\clearpage\newpage
-\subsection*{Reviewer 3}
-
-\begin{itemize}\setlength{\itemsep}{2\parsep}
-
-\item {\itshape Reviewer \#3: Equation (5), determination of beta, collision threshold, should be clarified through numeric/experimental section as well...}
-
-The collision threshold $\beta$ corresponds to the probability of collision that may occur when the robot follows the desired trajectory.
-As $\beta$ decreases, the chances for collision decreases, but at the time, the planned trajectory may become longer and complicated while keeping a distance from potential obstacles.
-
-%Setting the particular value of $\beta$ depend
-
-%is an important parameter selected to avoid collisions. 
-%It is sensitive to grid cell size, since cell resolution determines the number of grid cells falling inside the possible collision region $\mathcal{C}_X$.
-One practical implication of (5) is that cells with greater occupancy probabilities inside $\mathcal{C}_X$ tend to fall near the perimeter of the region as the robot moves because internal cells are typically well-known to be free assuming the robot covered open space beforehand.
-This point, and
-The following description about collision probability is added to Section 6 (numerical exploration example), and the same properties are held for Section 7 (experiment).
-\begin{correction}For added safety, cells falling within $0.6$m of the robot are considered inside the possible collision zone $\mathcal{C}_X$, from \refeqn{CollisionInequalityConstraint}, where $\beta=0.1$ is selected.\end{correction}
-
-\item {\itshape Reviewer \#3: Just before equation (7) forward sensor model p(...) should be explained more.}
-
-The following is added to the ``Bayesian Framework'' paragraph of Section 3.2.
-
-\begin{correction}The occupancy probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$ is based on the forward sensor model $p(z_{t,l}|m,X_{1:t},Z_{1:t-1})$ that describe the distribution of the measurements for the given robot pose and the map.
-This specifies the stochastic characteristics of the sensor, such as the mean squared error, and it is assumed as a known distribution, specific to a particular depth sensor.
-One could certainly obtain a stochastic model by fitting a probability density to a controlled set of measurements.
-\end{correction}
-
-\item {\itshape Reviewer \#3: Likewise, in the Proposition 1, ``$p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading'' should be explained more rigorously through the Appendix 1 as well.}
-
-The statement in Proposition 1 is revised as follows.
-
-\begin{correction}$p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the probability density of the measurement when all of cells in the field of view of the $l$-th ray is not occupied.\end{correction}
-
-Referring to the maximum reading, the following explanation is added after (34).
-
-\begin{correction}For this special case, probability density can be easily represented with a uniform distribution over a short distance.
-The properties of this distribution depend on the depth sensor fidelity, because a maximum reading may indicate that all cells within the sensor range are empty, but it might also represent a failed reading.\end{correction}
-
-\item {\itshape Reviewer \#3: Uncertainty on position and attitude, cell size adjustment can be discussed at the beginning.}
-
-Accounting for localization is very important, particularly in conjunction with mapping and autonomous exploration, though this is beyond the paper scope.
-The following is added to the Section 1 (introduction).
-
-\begin{correction}Accounting for the stochastic properties of position and attitude is important to the field of robotics, and accounting for localization in conjunction with mapping and exploration are considered as a future step to this paper.\end{correction}
-
-A brief explanation of grid cell size is added to the first paragraph of Section 2.1.
-
-\begin{correction}The location and size of each grid cell is assumed known, where a smaller cell size (greater grid resolution) better represents a space, but increases computation and memory.
-\end{correction}
-
-Further discussion on cell resolution is addressed two comments below.
-
-\item {\itshape Reviewer \#3: Section 5, motion planning should cite the following paper: (grid-based, Dijkstra)}
-
-N. Gasilov, M. Dogan, V. Arici, `Two-stage shortest path algorithm for solving optimal obstacle avoidance problem', IETE Journal of Research, 57(3), 278-285, 2011, doi:10.4103/0377-2063.83650.
-
-The reference is added.
-
-%\EditTL{Do we need a brief comparison with this?}
-
-\item {\itshape Reviewer \#3: Choosing the optimal parameters for each case, e.g. ray numbers, resolution in Table 1, can be explained more to understand the proposed algorithm.}
-
-The following is added in reference to Table 1 in Section 6.1.
-
-\begin{correction}Among these parameters, cell resolution and the number of rays have the greatest impact on computation.
-In general, the finer the grid, the greater the memory and computational requirements are because the computational order of (10) grows linearly with the number of grid cells inside the sensor range limits.
-Since the inverse sensor models are combined sequentially as shown in (13), the number of rays considered are proportional to the computation order as well.\end{correction}
-
-\item {\itshape Reviewer \#3: Convergence rate of algorithm, or speed of robot while achieving its goal, should be discussed, e.g.
-possible extensions or improvements to overcome some unnecessary repetitions and back-and-forth small step size bottlenecks.}
-
-Thanks for the comments. This is an important point, especially as it relates to future goals of the research. The following is added as the last paragraph of Section 7 (experiment).
-
-\begin{correction}Both the numerical and experimental results demonstrated successful exploration strategies, though the algorithm can be improved.
-In particular, the robot sometimes revisits spaces that it has explored before, most frequently when the robot did not take measurements toward a particular direction.
-A few methods to improve this issue are treated as future work.
-One options is adding cost functions (e.g.
-distance based, number of past measurements inside an area) to prevent the robot from leaving an area before it is well-known.
-A second option is partitioning the space into sections, where the robot moves toward sections with large numbers of pose candidates with high expected information gains.
-Another option is reevaluating the future pose optimization while the robot is translating between poses because the occupancy grid map is changing, and so is expected information gain.
-The efficacy of these proposed changes will be evaluated, particularly as they relate to more useful systems, such as 3D map generation and multi-robot exploration.\end{correction}
-
-\end{itemize}
-
 \end{document}

### 2018-06-09 Journal/JINT 2017/JIRS16_rev1.tex
journal added

--- 2018-06-09 Journal/JINT 2017/JIRS16.tex
@@
 \usepackage{curves,pgfgantt}
-%
+\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
+
 % \usepackage{mathptmx}      % use Times fonts if available on your TeX system
@@
 \graphicspath{{./Fig/}}
+\newcommand{\WriteBlue}[1]{{\color{blue}\protect #1}}
 
@@
 [12], [13], [16], [17]) that employ Rao-Blackwellized particle filters to deal with localization and autonomous exploration simultaneously with an approximate entropy change, this paper is based on an accurate probabilistic solution to map information gain, under the assumption that the robot pose is known.
+Accounting for the stochastic properties of position and attitude is important to the field of robotics, and accounting for localization in conjunction with mapping and exploration are considered as a future step to this paper.
 
@@
 Let a map $m$ be decomposed into $n_m$ evenly-spaced grid cells, where the $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,\ldots,n_m}$, that is defined as $\mathbf{m}_i=1$ when occupied, and $\mathbf{m}_i=0$ when free.
-The location and size of each grid cell is assumed known.
+The location and size of each grid cell is assumed known, where a smaller cell size (greater grid resolution) better represents a space, but increases computation and memory.
 Therefore, a map $m$ is defined by $\{\mathbf{m}_1,\ldots, \mathbf{m}_{n_m}\}$ ($2^{n_{m}}$ possible maps).
@@
 where $\mathcal C_X$ is the set of grid cells falling inside a volume of preselected size around $X$ that may cause collision and $\beta>0$ is a threshold of the probability of collision.
+It is considered that once a collision-free trajectory toward $X_c^*$ is obtained, the robot moves along the trajectory controlled by a motion control system in the inner loop.
 Once the robot has translated to $X_c^*$, the above process is repeated.
@@
 \paragraph{Bayesian Framework}
-%Assuming that the probability density of thxe sensor measurement is fixed over a small measurement region $\Delta z$, 
+%Assuming that the probability density of the sensor measurement is fixed over a small measurement region $\Delta z$, 
+The occupancy probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$ is based on the forward sensor model $p(z_{t,l}|m,X_{1:t},Z_{1:t-1})$ that describe the distribution of the measurements for the given robot pose and the map.
+This specifies the stochastic characteristics of the sensor, such as the mean squared error, and it is assumed as a known distribution, specific to a particular depth sensor.
+One could certainly obtain a stochastic model by fitting a probability density to a controlled set of measurements.
 Bayes' rule yields
@@
 Since the cells outside of the sensor field of view (FOV) are not affected, we focus on a reduced map $r_l$ in the FOV of the $l$-th ray.
-This reduced map is chosen such that each cell of $r_l$ corresponds to a grid cell of $m$ that the $l$-th ray intersects, ordered by increasing distance, which is easily determined from geometry.
+This reduced map is chosen such that each cell of $r_l$ corresponds to a grid cell of map $m$ that the $l$-th ray intersects, ordered by increasing distance, which is easily determined from geometry.
 Let $\mathbf{r}_{l,k}$ be the binary random variable representing the occupancy of the $k$-th cell of the $l$-th ray.
@@
 
-Next, let $\mathbf{r}_{l,k+}$ correspond to the event that the $k$-th cell of the $l$-th ray is occupied, cells with lower index are free, and cells with greater index may or may not be occupied, i.e., event $\mathbf{r}_{l,k+}$ occurs when \\$r\in\braces{r\in\{0,1\}^{{n_{r,l}}}\,|\mathbf{r}_{l,1}=0,\mathbf{r}_{l,2}=0,\ldots,\mathbf{r}_{l,k-1}=0,\mathbf{r}_{l,k}=1}$.
-In other words, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray, and this concept is illustrated with a simple 1D example in Figure \ref{fig:show_rkplus}.
+Next, let $\mathbf{r}_{l,k+}$ correspond to the event that the $k$-th cell of the $l$-th ray is occupied, cells with lower index (closer cells) are free, and cells with greater index (farther cells) may or may not be occupied, i.e., event $\mathbf{r}_{l,k+}$ occurs when \\$r\in\braces{r\in\{0,1\}^{{n_{r,l}}}\,|\mathbf{r}_{l,1}=0,\mathbf{r}_{l,2}=0,\ldots,\mathbf{r}_{l,k-1}=0,\mathbf{r}_{l,k}=1}$.
+In other words, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray.
+For example, see Figure \ref{fig:show_rkplus} that illustrates the event of $\mathbf{r}_{l,k+}$ when $l=1$ and $k=4$ for an one-dimensional cell array.
+This concept of grouping map outcomes can be easily extended to 2D using ray casting, where a 1D ray is spanning 2D space.
+The cells considered for the measurement ray are determined with the geometric intersection between the ray and map cell edges.
+Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell.
+Based on this, we present the mathematical expressions for the exact inverse sensor model, as summarized by \refeqn{RayISMAnswer} in Proposition 1.
+Later in Section \ref{sec:SA}, these are also rearranged as a computational algorithm.
+
 \begin{figure}
@@
   \includegraphics[width=0.4\textwidth]{rkplus_4.png}
-  \caption{A robot located to the left of the 1D occupancy grid map $r_l$ composed of $n_{r,l}=6$ grid cells in $4$ cases.
+  \caption{A robot located to the left of a 1D occupancy grid map $r_l$ composed of $n_{r,l}=6$ grid cells in $4$ cases.
 In each case, cells $1$--$3$ are free, cell $4$ is occupied, and cells $5$ and $6$ may or may not be occupied.
-Each case corresponds to the event $\mathbf{r}_{4+}$.}
+All above outcomes correspond to the event $\mathbf{r}_{l,4+}$.}
   \label{fig:show_rkplus}
 \end{figure}
-Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell.
-Based on this, we present the following computational algorithm to compute the inverse sensor model.
 
@@
 \end{align}
-where $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ is chosen for convenience and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading.
+where $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ is chosen for convenience and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the probability density of the measurement when all of cells in the field of view of the $l$-th ray is not occupied.
 The normalizer $\eta_{t,l}$ is given by
@@
 
-\subsection{Summary of Algorithm}
+\subsection{Summary of Algorithm}\label{sec:SA}
 
 In short, the proposed scan inverse sensor model is computed by \refeqn{RayISMAnswer}-\refeqn{RayByRayScanISM}.
-These are summarized in Table \ref{tab:RayByRayISM} as a computationally efficient, recursive algorithm that avoids several repeated calculations of the same quantity.
+These are summarized in Algorithm \ref{alg:RayByRayISM} as a computationally efficient, recursive algorithm that avoids several repeated calculations of the same quantity.
 This algorithm utilizes temporary variables, where the required computational resources are minimized by computing the updated occupancy probabilities of all grid cells along a measurement ray together.
+Note that this function is only considering $l$-th ray at the $t$-th time step, where probabilities are subject to conditions on $X_{1:t},Z_{1:t-1},z_{t,1:l-1}$, so these are removed from the algorithm for simplicity.
 
@@
 % More simple (subscripts omitted)
-\begin{table}[p]
-\begin{tabular}{ l }
-  Function: $P(\mathbf{r}_k|z)=\text{InverseSensorModel}(P(\mathbf{r}_k),z)$\\
-  for all $k\in\braces{1,2,\ldots,n_{r}}$\\
-  Initialize $\eta^{-1}=0$ and $P(\bar{\mathbf{r}}_{0})=1$;\\
-  for all $k = 1,2,\ldots,n_{r}$\\
-   \ \ $P(\mathbf{r}_{k+}) = P(\bar{\mathbf{r}}_{0:k-1})P(\mathbf{r}_{k})$;\\
-   \ \ $P(\bar{\mathbf{r}}_{0:k})=P(\bar{\mathbf{r}}_{0:k-1})P(\bar{\mathbf{r}}_{k})$;\\
-   \ \ $a_\text{temp}=P(\mathbf{r}_{k+})p(z_l|\mathbf{r}_{k+},X)$;\\
-   \ \ $\tilde P(\mathbf{r}_{k}|z)=P(\mathbf{r}_{k})\eta^{-1}_{t,l}+a_\text{temp}$;\\
-   \ \ $\eta^{-1}=\eta^{-1}+a_\text{temp}$;\\
-  end for\\
-  $\eta^{-1}=\eta^{-1}+P(\bar{\mathbf{r}}_{0:k})p(z_\text{max})$;\\
-  $P(\mathbf{r}_{k}|z)=\eta\tilde P(\mathbf{r}_{k}|z)$ for all $k\in\braces{1,2,\ldots,n_{r}}$;\\
-  \end{tabular}
+
+\begin{algorithm}[H]
+	Function: $P(\mathbf{r}_k|z)=\text{InverseSensorModel}(P(\mathbf{r}_k),z)$ $\forall$ $k\in\braces{1,2,\ldots,n_{r}}$\;
+	Initialize $\eta^{-1}=0$ and $P(\bar{\mathbf{r}}_{0})=1$\;
+	\For{$k=1,2,\ldots,n_{r}$}{
+		$P(\mathbf{r}_{k+}) = P(\bar{\mathbf{r}}_{0:k-1})P(\mathbf{r}_{k})$\;
+		$P(\bar{\mathbf{r}}_{0:k})=P(\bar{\mathbf{r}}_{0:k-1})(1-P(\mathbf{r}_{k}))$\;
+    		$a_\text{temp}=P(\mathbf{r}_{k+})p(z_l|\mathbf{r}_{k+},X)$\;
+   		$\tilde P(\mathbf{r}_{k}|z)=P(\mathbf{r}_{k})\eta^{-1}_{t,l}+a_\text{temp}$\;
+   		$\eta^{-1}=\eta^{-1}+a_\text{temp}$\;
+	}
+	$\eta^{-1}=\eta^{-1}+P(\bar{\mathbf{r}}_{0:k})p(z_\text{max})$\;
+	Return $P(\mathbf{r}_{k}|z)=\eta\tilde P(\mathbf{r}_{k}|z)$ for all $k\in\braces{1,2,\ldots,n_{r}}$\;
 \caption{Inverse Sensor Model of an Individual Measurement Ray}
-\raggedright{Note: this function is only considering $l$-th ray at the $t$-th time step, where probabilities are subject to conditions on $X_{1:t},Z_{1:t-1},z_{t,1:l-1}$, so these are removed from the algorithm for simplicity.}
-\label{tab:RayByRayISM}
-\end{table}
+\label{alg:RayByRayISM}
+\end{algorithm}
 
@@
 
-We present an algorithm pseudocode providing the necessary steps to obtain the objective function for a single measurement ray (Table \ref{tab:RayExpectedEntropyGain}).
-Much like the algorithm pseudocode for the inverse sensor model (Table \ref{tab:RayByRayISM}), the variable $a_\text{temp}$ serves as an intermediate variable designed to avoid repeated calculations.
+We present an algorithm pseudocode providing the necessary steps to obtain the objective function for a single measurement ray (Algorithm \ref{alg:RayExpectedEntropyGain}).
+Much like the algorithm pseudocode for the inverse sensor model (Algorithm \ref{alg:RayByRayISM}), the variable $a_\text{temp}$ serves as an intermediate variable designed to avoid repeated calculations.
 Since this algorithm operates as a function, fixed indices and condition variables are removed for simplification.
 
-\begin{table}[p]
-\begin{tabular}{ l }
-  Function: $\mathcal I_\text{ray}=\text{RayExpInfoGain}(x,P(r),z_{1:n_{r}})$\\
-  Define $P(\bar{\mathbf{r}}_{0})=P(\bar{\mathbf{r}}_{n_{r}+1})=1$;\\
-  for $k = 1,2,\ldots,n_{r}+1$\\
-   \ \ $P(\mathbf{r}_{k+}) = P(\bar{\mathbf{r}}_{0:k-1})P(\mathbf{r}_{k})$;\\
-   \ \ $P(\bar{\mathbf{r}}_{0:k})=P(\bar{\mathbf{r}}_{0:k-1})P(\bar{\mathbf{r}}_{k})$;\\
-  end for\\
-%  $P(\mathbf{r}_{(n_{r}+1)+}) = P(\bar{\mathbf{r}}_{0:n_{r}})$;\\
-  Find $\hat r\subset r$ corresponding to the $\hat n$ greatest values of\\
-  $\braces{P(\mathbf{r}_{1+}),P(\mathbf{r}_{2+}),\ldots,P(\mathbf{r}_{(n_{r}+1)+})}$;\\
-  Define $P(\hat{\bar{\mathbf{r}}}_{0})=1$;\\
-  for $k = 1,2,\ldots,\hat{n}$\\
-   \ \ $P(\hat{\mathbf{r}}_{k+}) = P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\mathbf{r}}_{k})$;\\
-   \ \ $P(\hat{\bar{\mathbf{r}}}_{0:k})=P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\bar{\mathbf{r}}}_{k})$;\\
-  end for\\
-  for $k_\text{m}=1,2,\ldots,\hat n$\\
-   \ \ Initialize $\eta^{-1}_{k_\text{m}}=0$;\\
-   \ \ for $k_\text{c}=1,2,\ldots,\hat n$\\
-   \ \ \ \ $a_\text{temp}=P(\hat{\mathbf{r}}_{k_\text{c}+})p(z_{k_\text{m}}|\hat{\mathbf{r}}_{k_\text{c}+},x)$;\\
-   \ \ \ \ $\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=P(\hat{\mathbf{r}}_{k_\text{c}})\eta^{-1}_{k_\text{m}}+a_\text{temp}$;\\
-   \ \ \ \ $\eta^{-1}_{k_\text{m}}=\eta^{-1}_{k_\text{m}}+a_\text{temp}$;\\
-   \ \ end for\\
-   \ \ $P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=\eta_{k_\text{m}}\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})$ for all $k_\text{c}=1,2,\ldots,\hat n$;\\
-   end for\\ 
-   $P(z_{k_\text{m}}|x)=\frac{\eta^{-1}_{k_\text{m}}}{\sum_{i=1}^{\hat n}\eta^{-1}_{i}}$ for $k_\text{m}=1,2,\ldots,\hat n$;\\
-
-   Initialize $\mathcal I_\text{ray}=0$;\\
-   for $k_\text{c}=1,2,\ldots,\hat n$\\
-   \ \ $\mathcal I_\text{ray}=\mathcal I_\text{ray}+H(P(\hat{\mathbf{r}}_{k_\text{c}}))$;\\
-   \ \ for $k_\text{m}=1,2,\ldots,\hat n$\\
-   \ \ \ \ $\mathcal I_\text{ray} = \mathcal I_\text{ray}-H(P(\hat{\mathbf{r}}_{k_\text{c}}|x_c,z_{k_\text{m}}))P(z_{k_\text{m}}|x_{c})$;\\
-   \ \ end for\\
-  end for\\
+\begin{algorithm}[H]
+	Function: $\mathcal I_\text{ray}=\text{RayExpInfoGain}(x,P(r),z_{1:n_{r}})$\;
+	Initialize $P(\bar{\mathbf{r}}_{0})=P(\hat{\bar{\mathbf{r}}}_{0})=P(\bar{\mathbf{r}}_{n_{r}+1})=1$\;
+	\For{$k = 1,2,\ldots,n_{r}+1$}{
+		$P(\mathbf{r}_{k+}) = P(\bar{\mathbf{r}}_{0:k-1})P(\mathbf{r}_{k})$\;
+		$P(\bar{\mathbf{r}}_{0:k})=P(\bar{\mathbf{r}}_{0:k-1})(1-P(\mathbf{r}_{k}))$\;
+	}
+	Find $\hat r\subset r$ of the $\hat n$ greatest values of 
+	$\braces{P(\mathbf{r}_{1+}),P(\mathbf{r}_{2+}),\ldots,P(\mathbf{r}_{(n_{r}+1)+})}$\;
+	\For{$k = 1,2,\ldots,\hat{n}$}{
+		$P(\hat{\mathbf{r}}_{k+}) = P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\mathbf{r}}_{k})$\;
+		$P(\hat{\bar{\mathbf{r}}}_{0:k})=P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\bar{\mathbf{r}}}_{k})$\;
+	}
+	\For{$k_\text{m}=1,2,\ldots,\hat n$}{
+		Initialize $\eta^{-1}_{k_\text{m}}=0$\;
+		\For{$k_\text{c}=1,2,\ldots,\hat n$}{
+			$a_\text{temp}=P(\hat{\mathbf{r}}_{k_\text{c}+})p(z_{k_\text{m}}|\hat{\mathbf{r}}_{k_\text{c}+},x)$\;
+			$\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=P(\hat{\mathbf{r}}_{k_\text{c}})\eta^{-1}_{k_\text{m}}+a_\text{temp}$\;
+			$\eta^{-1}_{k_\text{m}}=\eta^{-1}_{k_\text{m}}+a_\text{temp}$\;
+		}
+		$P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=\eta_{k_\text{m}}\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})$ for all $k_\text{c}=1,2,\ldots,\hat n$\;
+	}
+	$P(z_{k_\text{m}}|x)=\frac{\eta^{-1}_{k_\text{m}}}{\sum_{i=1}^{\hat n}\eta^{-1}_{i}}$ for $k_\text{m}=1,2,\ldots,\hat n$\;
+	Initialize $\mathcal I_\text{ray}=0$\;
+	\For{$k_\text{c}=1,2,\ldots,\hat n$}{
+		$\mathcal I_\text{ray}=\mathcal I_\text{ray}+H(P(\hat{\mathbf{r}}_{k_\text{c}}))$\;
+		\For{$k_\text{m}=1,2,\ldots,\hat n$}{
+			$\mathcal I_\text{ray} = \mathcal I_\text{ray}-H(P(\hat{\mathbf{r}}_{k_\text{c}}|x_c,z_{k_\text{m}}))P(z_{k_\text{m}}|x_{c})$\;
+		}
+	}
 	Return: $\mathcal I_\text{ray}$\\
-\end{tabular}
 \caption{Expected Information Gain from a Measurement Ray}
-\raggedright{Note: this function is only considering a single ray for a future time step, where probabilities are subject to conditions on $X_{1:t},Z_{1:t}$, so these are removed from the algorithm for simplicity.}
-\label{tab:RayExpectedEntropyGain}
-\end{table}
+\label{alg:RayExpectedEntropyGain}
+\end{algorithm}
 
@@
 Once the optimal future pose $X_c^*$ is selected, the robot must traverse its environment to reach $X_c^*$ without collisions.
-The locations of cells on a grid-based map provide nodes for Dijkstra's algorithm~\cite{Dij59}.
+The locations of cells on a grid-based map provide nodes for Dijkstra's algorithm~\cite{Dij59},~\cite{Dij11}.
 At each node, \refeqn{CollisionInequalityConstraint} must be satisfied for the node to be considered reachable, and edges between nodes correspond to neighboring grid cells.
@@
 Then, a controller on the robot tracks this trajectory until the robot falls within acceptable thresholds of the final optimal pose.
-Once the robot completes this motion, the entire process is repeated. %The complete autonomous exploration process is summarized in Table \ref{tab:Alg_AutomousExploration}.
+Any control scheme for the motion of the robot can be integrated with the proposed exploration algorithm.
+Once the robot completes this motion, the entire process is repeated.
+
+%The complete autonomous exploration process is summarized in Table \ref{tab:Alg_AutomousExploration}.
 
@@
 
-First, we compare the proposed exact solution to the inverse sensor model summarized in Table \ref{tab:RayByRayISM} with an approximate algorithm presented in~\cite{PirRutBisSch11,KhoElb12}, summarized as follows.
+First, we compare the proposed exact solution to the inverse sensor model summarized in Algorithm \ref{alg:RayByRayISM} with an approximate algorithm presented in~\cite{PirRutBisSch11,KhoElb12}, summarized as follows.
 The probability that the $i$-th grid cell $\mathbf{m}_i$ is occupied conditioned on the measurement ray $z_{t,l}$ at the pose $X_t$ is the continuous function
@@
 Table~\ref{tab:penn} shows the mapping parameters and data specifications for the actual odometry and lidar measurements from University of Pennsylvania through Coursera open course~\cite{coursera}.
+Among these parameters, cell resolution and the number of rays have the greatest impact on computation.
+In general, the finer the grid, the greater the memory and computational requirements are because the computational order of \refeqn{RayISMAnswer} grows linearly with the number of grid cells inside the sensor range limits.
+Since the inverse sensor models are combined sequentially as shown in \refeqn{RayByRayScanISM}, the number of rays considered are proportional to the computation order as well.
 Figure~\ref{fig:penn_map} shows the direct comparison of the resulting map based on the approximate and proposed inverse sensor model.
@@
 The initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) for grid cells covered by the circular robot of radius $0.3$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
+For added safety, cells falling within $0.6$m of the robot are considered inside the possible collision zone $\mathcal{C}_X$, from \refeqn{CollisionInequalityConstraint}, where $\beta=0.1$ is selected.
+% One practical implication of \refeqn{CollisionInequalityConstraint} is that cells with greater occupancy probabilities inside $\mathcal{C}_X$ tend to fall near the perimeter of the region because internal cells are typically well-known to be free assuming the robot covered open space beforehand.
+Thus, a fairly large $\beta=0.1$ selected.
 
@@
 
-The time sequence of the exploration for every 5~minutes are simulated in ROS Stage 2D environment. The corresponding simulation results are illustrated in Figure \ref{fig:IRL}.
+The time sequence of the exploration for every 5~minutes are simulated in ROS Stage 2D environment.
+The corresponding simulation results are illustrated in Figure \ref{fig:IRL} and a video is available at \href{https://www.youtube.com/watch?v=5VdzKHreB_s}{\WriteBlue{https://www.youtube.com/watch?v=5VdzKHreB\_s}}.
 Lighter red dots and darker blue dots in the explored area correlate to the level of information gain at those positions with low and high value, respectively.
@@
     	\begin{subfigure}[b]{0.35\textwidth}
-        		\includegraphics[width=\textwidth]{test_setup_1.jpg}
+        		\includegraphics[width=\textwidth]{test_setup_bottom_left_cropped.jpg}
         		\caption{Bottom-left view}
@@
 	\begin{subfigure}[b]{0.35\textwidth}
-        		\includegraphics[width=\textwidth]{test_setup_2.jpg}
+        		\includegraphics[width=\textwidth]{test_setup_bottom_right_cropped.jpg}
         		\caption{Bottom-right view}
@@
 
-The resulting occupancy grid maps and trajectories are depicted in Figure \ref{fig:ExperimentOGM}.
-Most importantly, the robot builds a clear occupancy grid map despite sensor noises and imperfect sensor synchronization.
+The resulting occupancy grid maps and trajectories are depicted in Figure \ref{fig:ExperimentOGM} and a video is available at \href{https://www.youtube.com/watch?v=CRQfhhICSj0&feature=youtu.be}{\WriteBlue{www.youtube.com/watch?v=CRQfhhICSj0\&feature=youtu.be}}.
+Most importantly, the robot builds a clear occupancy grid map despite sensor noise and imperfect sensor synchronization.
 The autonomous exploration successfully guides the robot such that it builds a map of the reachable space without colliding with any obstacles.
@@
     	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_start.jpg}
-        		\caption{$t=0.0$min}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t0sec.jpg}
+        		\caption{$t=0$sec}
         		\label{fig:Experiment_ogm_t0}
@@
 	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_0p5min.jpg}
-        		\caption{$t=0.5$min}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t20sec.jpg}
+        		\caption{$t=20$sec}
         		\label{fig:Experiment_ogm_t0p5}
@@
 	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1min.jpg}
-        		\caption{$t=1.0$min}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t40sec.jpg}
+        		\caption{$t=40$sec}
         		\label{fig:Experiment_ogm_t1}
@@
 	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1p5min.jpg}
-        		\caption{$t=1.5$min}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t60sec.jpg}
+        		\caption{$t=60$sec}
         		\label{fig:Experiment_ogm_t1p5}
@@
 	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2min.jpg}
-        		\caption{$t=2.0$min}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t80sec.jpg}
+        		\caption{$t=80$sec}
         		\label{fig:Experiment_ogm_t2}
@@
     	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2p5min.jpg}
-        		\caption{$t=2.5$min}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t100sec.jpg}
+        		\caption{$t=100$sec}
         		\label{fig:Experiment_ogm_t2p5}
@@
 	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3min.jpg}
-        		\caption{$t=3.0$min}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t120sec.jpg}
+        		\caption{$t=120$sec}
         		\label{fig:Experiment_ogm_t3}
@@
 	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3p5min.jpg}
-        		\caption{$t=3.5$min}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t140sec.jpg}
+        		\caption{$t=140$sec}
         		\label{fig:Experiment_ogm_t3p5}
@@
 	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4min.jpg}
-        		\caption{$t=4.0$min}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t160sec.jpg}
+        		\caption{$t=160$sec}
         		\label{fig:Experiment_ogm_t4}
@@
 	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4p5min.jpg}
-        		\caption{$t=4.5$min}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t180sec.jpg}
+        		\caption{$t=180$sec}
         		\label{fig:Experiment_ogm_t4p5}
@@
 
+%The resulting map and exploration commands are generated in real time.
+The grid cells are stored as double-float variables in a vector, where the vector index is mapped to a location on the occupancy grid.
+Thus, the memory requirements are proportional to the number of grid cells, which is roughly the mapping area divided by the area of a grid cell.
+Since cell locations need not be saved, memory requirements are reduced.
+The mean time for updating the occupancy grid is $0.0115$ seconds, and the mean time to determine an exploration strategy is $0.6820$ seconds.
+Time requirements for mapping and exploration are easily modified by changing grid cell resolution or the number of exploration pose candidates, respectively.
+In short, there is a tradeoff between computational speed and accuracy of mapping or exploration.
+The resulting map and exploration commands are generated in real time.
+The grid cells are stored as double-float variables in a vector, where the vector index is mapped to a location on the occupancy grid.
+Thus, the memory requirements are proportional to the number of grid cells, which is roughly the mapping area divided by the area of a grid cell.
+Since cell locations need not be saved, memory requirements are reduced.
+On a Lenovo T540p laptop with an Intel Core i7-4900MQ processor (quad-core, 2.8GHz per core) and 16GB of RAM, the mean time for updating the occupancy grid is $0.0115$ seconds, and the mean time to determine an exploration strategy is $0.6820$ seconds.
+Time requirements for mapping and exploration are easily modified by changing grid cell resolution or the number of exploration pose candidates, respectively.
+In short, there is a tradeoff between computational speed and accuracy of mapping or exploration.
+
 The autonomous exploration is governed by a policy to maximize expected entropy decreases, so entropy and entropy change with time (Figure \ref{fig:ExperimentH}) are important metrics.
@@
 In some cases, the robot is able to update the cell occupancy probabilities; in other cases, cells fall inside regions occluded from all reachable space, and the robot cannot update these cells.
-Among those grid cells visible from collision-free locations, the vast majority of cells were updated in less than $5$min while following a very slow ($5$cm/sec) trajectory.
-
-%Additionally, even though the walls appear fairly well-defined, they are not pefectly straight boundaries.
+Among those grid cells visible from collision-free locations, the vast majority of cells were updated in less than $3$min while following a very slow ($5$cm/sec) trajectory.
+
+Both the numerical and experimental results demonstrated successful exploration strategies, though the algorithm can be improved.
+In particular, the robot sometimes revisits spaces that it has explored before, most frequently when the robot did not take measurements toward a particular direction.
+A few methods to improve this issue are treated as future work.
+One options is adding cost functions (e.g.
+distance based, number of past measurements inside an area) to prevent the robot from leaving an area before it is well-known.
+A second option is partitioning the space into sections, where the robot moves toward sections with large numbers of pose candidates with high expected information gains.
+Another option is reevaluating the future pose optimization while the robot is translating between poses because the occupancy grid map is changing, and so is expected information gain.
+The efficacy of these proposed changes will be evaluated, particularly as they relate to more useful systems, such as 3D map generation and multi-robot exploration.
+
+%Additionally, even though the walls appear fairly well-defined, they are not perfectly straight boundaries.
 The occupancy grid assumptions do not perfectly align with the true environment; an edge of a wall might occupy only part of a grid cell.
@@
 \end{align}
-where $p(z_{t,l}|\mathbf{r}_{(n_r+1)+},X_t)$ corresponds to the forward sensor model of maximum reading and $P(\mathbf{r}_{l,n_{r,l}+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience for the empty map case.
+where $P(\mathbf{r}_{l,n_{r,l}+1}|X_{1:t-1},Z_{1:t-1})=1$ is selected for convenience and \\$p(z_{t,l}|\mathbf{r}_{(n_r+1)+},X_t)$ corresponds to the forward sensor model of maximum reading.
+For this special case, probability density can be easily represented with a uniform distribution over a short distance.
+The properties of this distribution depend on the depth sensor fidelity, because a maximum reading may indicate that all cells within the sensor range are empty, but it might also represent a failed reading.
 

### 2018-06-09 Journal/JINT 2017/JIRS16_rev2.tex
journal added

--- 2018-06-09 Journal/JINT 2017/JIRS16_rev1.tex
@@
 These are illustrated by numerical examples and experiments in Sections \ref{sec:NumRes} and \ref{sec:ExpRes}.
+
+This proposed mapping approach follows a Bayesian framework that focusses on occupied and free space, not tracking particular systems or mapping features, such as visual servoing~\cite{ChiDinChe14}.
+The proposed algorithm may be combined with feature-based approaches for object tracking or stabilization purposes, with the added challenge of associating occupied spaces with features.
 
@@
 The starting and ending positions and attitudes may be constrained, and polynomials patched together for long trajectories share a common position and velocity with respect to time.
-Then, a controller on the robot tracks this trajectory until the robot falls within acceptable thresholds of the final optimal pose.
-Any control scheme for the motion of the robot can be integrated with the proposed exploration algorithm.
+Then, the robot tracks this trajectory until the robot falls within acceptable thresholds of the final optimal pose.
+For generality of the proposed exploration algorithm, any position controller yielding robotic motion that follows this desired trajectory may be selected.
 Once the robot completes this motion, the entire process is repeated.
-
 %The complete autonomous exploration process is summarized in Table \ref{tab:Alg_AutomousExploration}.
@@
 	\centering
-	\includegraphics[width=\textwidth]{experiment_H_per_min.pdf}
+	\includegraphics[width=\textwidth]{JIRS_fig11.eps}
 	\caption{The map entropy $H$ averaged over all $90601$ grid cells is reduced during experimentation.

### 2018-06-09 Journal/JINT 2018/JINT18.tex
journal added

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{amsmath,amssymb,url,times}%,subfigure}% amsthm is the one!
\usepackage{caption,subcaption,hyperref}
\usepackage{color,comment}
\usepackage{curves,pgfgantt}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
\usepackage{hyperref}
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
 \journalname{Journal of Intelligent \& Robotic Systems}

\newtheorem{prop}{Proposition}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\abs}[1]{\ensuremath{\left| #1 \right|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\ip}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Figure \ref{fig:#1}}
\newcommand{\tr}[1]{\mbox{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mbox{tr}\ensuremath{\!\bracket{#1}}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathbf{d}}}
\newcommand{\pair}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\!\langle #1 \rangle\!\rangle}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\ad}{\ensuremath{\mathrm{ad}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\graphicspath{{../../Fig/}}
\newcommand{\WriteBlue}[1]{{\color{blue}\protect #1}}

\begin{document}

\title{Multi-Vehicle Cooperative Exploration and Patrol of Uncertain 3D Environments}

%\titlerunning{Short form of title}        % if too long for running head

\author{Evan Kaufman \and Kuya Takami \and \\ Zhuming Ai \and Taeyoung Lee}%
%\thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 {\tt \{evankaufman,tylee\}@gwu.edu}}
%\thanks{Zhuming Ai, Ira S. Moskowitz, Information Management \& Decision Architectures, U.S. Naval Research Laboratory,  Washington, DC 20375}
%}

%\authorrunning{Short form of author list} % if too long for running head

% Expanded:
\institute{E. Kaufman \at
              Dept. of Mechanical and Aerospace Engineering \\
		 The George Washington University \\
		 Washington DC 20052 \\
              Tel.: 216-346-0946\\
              \email{evankaufman@gwu.edu} 
           \and
           K. Takami \at
              Dept. of Mechanical and Aerospace Engineering \\
		 The George Washington University \\
		 Washington DC 20052 \\
              Tel.: 414-212-5155\\
              \email{kuya@gwu.edu} 
           \and
           T. Lee \at
              Dept. of Mechanical and Aerospace Engineering \\
		 The George Washington University \\
		 Washington DC 20052 \\
              Tel.: 202-994-8710\\
		 Fax: 202-994-0238\\
              \email{tylee@gwu.edu}
           \and
           Z. Ai \at
              Information Management \& Decision Architectures
              \\ U.S. Naval Research Laboratory \\
              Washington, DC 20375
}

% Condensed:
\institute{E. Kaufman, K. Takami, and T. Lee \at
              Dept. of Mechanical and Aerospace Engineering \\
		 The George Washington University \\
		 Washington DC 20052 \\
              Tel.: 202-994-8710\\
              \email{\{evankaufman,kuya,tylee\}@gwu.edu} 
           \and
           Z. Ai \at
              Information Management \& Decision Architectures
              \\ U.S. Naval Research Laboratory \\
              Washington, DC 20375
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
This paper presents a stochastic framework for autonomous exploration and patrol with multiple cooperating robots. The first contribution extends the authors' prior work in single-robot exact occupancy grid mapping and autonomous exploration in a 2D environment to mapping and exploring in a 3D environment. The proposed 3D occupancy grid map is computed efficiently and is scalable to larger and more complex scenarios for real-time mapping. Important aspects of this 3D map are projected onto a 2D occupancy grid map, where an autonomous exploration algorithm is designed to select robotic motions that maximize map information gain. The mapping and exploration algorithms are demonstrated with an experiment where a quadrotor autonomously maps and explores an initially-uncertain environment. The second contribution is a novel approach to multi-vehicle cooperative patrol of environments based on map uncertainty. A cooperative autonomous exploration algorithm applies a bidding-based framework to coordinate robotic efforts for improving occupancy grid map information gain. Since these exploration approaches are based on probabilistic knowledge about the map, the 3D occupancy grid map is systematically degraded over time to encourage the robots to revisit regions as time passes, thereby patrolling the environment. Furthermore, using a Bayesian framework and receding horizons, the algorithm is robust to dynamic obstacles within the mapping space. The efficacy of the proposed multi-vehicle cooperative patrol is illustrated with a simulation involving three robots patrolling a large floor plan with a non-cooperative person walking around the space.
\keywords{Cooperative Patrol \and Multi-Vehicle Autonomous Exploration \and 3D Occupancy Grid Mapping}
\end{abstract}

\section{Introduction}
\label{intro}

Occupancy grid mapping is a popular choice for representing the environment surrounding robots. Accurate map generation provides important information for applications such as search-and-rescue, surveillance, and robotic cleaning. Occupancy grid maps can further assist in autonomous exploration, where robots safely explore new territories while avoiding collisions. Additionally, several robotic tasks, particularly surveillance and cleaning, require periodic visitation to various regions of the mapped environment, motivating a means for autonomous patrol.

This paper addresses three key aspects to multi-vehicle mapping and patrol of uncertain environments. First, a novel exact solution to occupancy grid mapping previously-proposed by the authors is extended to a 3D representation. Important occupancy information from this 3D map is projected onto a 2D occupancy grid map, where an autonomous exploration algorithm is designed to maximize map information gain based on a measure known as Shannon's entropy. The second key aspect addresses multiple vehicles cooperating together in mapping and exploration. A bidding-based cooperative autonomous exploration strategy is proposed to coordinate robotic efforts to maximize global map information. Finally, we propose a novel method of multi-vehicle coordinated patrol based on the multi-vehicle autonomous exploration and map degradation.

\subsection{Probabilistic Occupancy Grid Mapping and Autonomous Exploration in 3D}

%Mapping is of fundamental importance for robots to understand their environment, avoid collisions, and efficiently maneuver through their surroundings.
%Several grid-based map representations have been studied, e.g.,~\cite{WolSuk05,MeyBeiBur12,TanThoWolBus14}, because they provide abundant occupancy knowledge about the environment for collision-free maneuvers and are computationally tractable.

Mapping is an important and well-studied topic in robotics for understanding an environment, avoiding collisions, and efficiently planning robotic motions.
A popular approach to mapping is defining a grid-based map representation, e.g.,~\cite{WolSuk05,MeyBeiBur12,TanThoWolBus14}, because these representations are computationally tractable and simple to understand.

However, these conventional grid-based mapping representations commonly approximate occupancy probabilities of the grid cells. The \emph{inverse sensor model}, which determines the map occupancy probabilities based on depth sensor measurements, is approximated with a simplified function, providing an inexact solution. As such, the prior belief of the environment and the stochastic properties of the sensor are not properly accounted. Another popular approach is based on simulated learning trials~\cite{Thr01,ThrBurFox05,SouMaiGon12} to determine an inverse sensor model. This approach is sensitive to simulation errors, and is not suitable for the higher dimensional configurations composed of maps, poses, and measurements due to the large number of required samples. In short, none of the existing approaches use the sensor properties to directly obtain a Bayesian solution to occupancy probability.

%        1.      sensor fusion in mapping
The approximate nature of the occupancy probability, either by a simplified function or a learned solution, motivates an exact Bayesian solution to the inverse sensor model. In this paper, we propose a computationally-efficient algorithm to obtain the exact solution to the inverse sensor model. This serves to generate probabilities in a 3D environment following the framework of~\cite{KauLeeAiMos16,KauTakAiLee17}. The proposed approach integrates stochastic properties of all sensors and the prior map estimate, following a Bayesian framework, to obtain the exact 3D occupancy probabilities~\cite{KauTakAiLee18}. As such, it provides a unified framework for sensor fusion between multiple sensors with various properties, such as sensor accuracy and fields of view.

%        2.      proposing a method to combine 3D mapping and 2D exploration: Also, we need to motivate why we are not doing full 3D exploration. Stress that we are proposing a computationally friendly approach to perform 3D mapping with projection using the fact there is less variation in the floor plan along the vertical direction.
The 3D occupancy grid map provides key information to produce the policies that govern robotic motion, referred to as autonomous exploration. Among several approaches to determine these map-based motions in real-time, the most popular approach is known as frontier-based exploration. Using frontiers for autonomous exploration was proposed for 2D environments in~\cite{Yam97,Yam98} and extended to 3D with a visibility metric in~\cite{SawKriSri09} and~\cite{ZhuDinLinWu15,SenWan16,KleDor13} using the Octomap representation. The main idea of this approach is that robots move toward the boarder between certain and uncertain spaces to push back the boundary of known space. However, these systematic actions are not based on any consideration of the future probabilistic map uncertainty.

An alternative approach that addresses the probabilistic map properties is based on Shannon's entropy~\cite{StaGriBur05}. This metric provides a measure of uncertainty, which is used to acquire expected map information gain in~\cite{KauAiLee16,KauTakAiLee17}. Furthermore, entropy-based autonomous exploration approaches can be extended to 3D in a numerically efficient manner~\cite{KauTakAiLee18}. Due to several challenges with obtaining expected information gain of 3D measurement scans, particularly with the extended computational load, we propose a numerically efficient scheme that projects stochastic properties of the 3D probabilistic occupancy grid to 2D spaces for optimal information gain and collision avoidance. This approach is appropriate for floor plans with typically low variation along the vertical direction by exploiting just the important characteristics of 3D maps within a 2D plane. This process is designed for real-time implementation by combining 3D mapping and 2D exploration in a computationally-efficient manner.

%        3.      hardware/software for aerial exploration experiment.
In this paper, we present a flight experiment for the proposed single-vehicle exploration scheme. Multiple tasks, such as flight control, sensor data acquisition, mapping, exploration, and communications are executed simultaneously in a ROS environment using Python/C++. These tasks are verified with a quadrotor unmanned aerial vehicle holding an ASUS Xtion infrared depth sensor and Hokuyo laser scanner operated by the NVidia Jetson computing module.

\subsection{Multi-Vehicle Cooperative Autonomous Exploration}

Exploration with multiple robots can be beneficial in large or complicated environments to complete the task quickly and efficiently. Here, we propose a bidding-based autonomous exploration scheme to maximize global map information. The main idea is a central executive predicts how robotic actions are expected to improve map information gain cooperatively in a probabilistic fashion. A series of auctions determines where robots should travel. During each auction, robots submit bids based on how their sensors are expected to improve the quality of the occupancy grid map. Between auctions, bids are modified to reflect the impact of other robot members on the map. Because this approach is directly based on cooperative improvement of map information, considering coverage overlap among the robots, such as the work in~\cite{SimApfBurFoxMooThrYou00}, becomes unnecessary. This multi-vehicle exploration optimization is further augmented with travel time considerations and collision avoidance among robots. In short, the proposed multi-vehicle exploration scheme employs a bidding-based framework that maximizes map information while considering multiple vehicles cooperating together, and cell degradation promotes repeatedly visiting regions for patrol.

Furthermore, this bidding-based exploration approach is integrated with a receding horizon framework to optimize map information and collision-avoidance in a dynamic environment. The receding horizon framework serves to update the exploration strategy with a possibly-changing occupancy grid map, which directly impacts the viability of pose candidates. Additionally, a receding horizon provides an optimal feedback policy that can handle dynamic obstacles, improving collision-avoidance with cooperative robots and other moving obstacles, which is essential for patrolling dynamic environments.

\subsection{Multi-Vehicle Cooperative Autonomous Patrol}

Robotic surveillance and cleaning are examples of tasks that require periodic observations of the same regions. Several patrolling approaches have been proposed in recent years to address this need. A popular approach is decomposing the reachable space into Voronoi diagrams~\cite{KolCar08,PorRoc10,PipChrWei13}. In~\cite{KolCar08}, edges between nodes of the Voronoi graph are heavily sensed to track intruders. In~\cite{PorRoc10}, regions of the Voronoi diagram are assigned to different robots, and issues with robots having underperforming surveillance capabilities are addressed in~\cite{PipChrWei13}. These approaches assume the map is well-known in advance, and require a Voronoi graph be generated from the map.

Other patrol approaches create new metrics about the environment. Cell periodicity is estimated in~\cite{KraFenCieDonDuc14,KraFenHanDuc16} by predicting and observing the long-term dynamics of objects at the same locations on an occupancy grid. This work is extended to exploring and navigating a human-populated environment for long-term autonomy. A bio-inspired approach is proposed in~\cite{ZhaXia11}, where robots cooperate by dropping digital pheromones based on the importance of events, where the pheromones weaken over time to prioritize future tasks. Another approach is formulating patrol as an optimization problem with respect to robot idleness in~\cite{YanZha16}.

In this paper, the proposed approach to coordinated patrol is based on cooperative multi-vehicle autonomous exploration and map degradation, which is different from prior work in two key ways. First, the patrol is based on map uncertainty, where the environment can be initially-uncertain. Second, the proposed approach is independent of external variables or structures such as Voronoi diagrams, cell periodicity estimates, digital pheromones, or idleness estimates. Instead, the proposed patrol approach is based on predicting future occupancy grid cell uncertainties. All grid cells are subjected to a continuous-time Markov degradation process. Highly certain cells subject to this Bayesian process slowly increase in uncertainty over time. Then, the multi-vehicle exploration scheme assigns cooperating robots to revisit these cells. Thus, degradation promotes periodic observations of different regions of the map.

We present a numerical simulation of three quadrotors mapping and patrolling a floor plan of a building, which is initially uncertain. The robots must safely patrol this space while a human walks independently around the environment.

In short, the contributions of this paper are as follows. First, we propose a computation technique to extend 2D mapping and autonomous exploration to a 3D environment in real-time. Then, we propose a bidding-based cooperative exploration with multiple vehicles. The exploration is developed in a receding horizon framework for dynamic environments. Finally, we develop a stochastic framework for patrol. These are demonstrated with an experiment and a simulation.

%In short, there are two contributions of this paper. First, single-vehicle 2D mapping and autonomous exploration approaches are extended to realistic 3D environments. Second, multi-vehicle cooperative patrol is proposed using bidding-based autonomous exploration and map grid cell degradation. These two concepts are demonstrated with an experiment and a simulation.

\section{Problem Formulation}
\label{sec:ProbDef}

This sections summarizes the key equations providing the exact solutions to occupancy grid mapping and expected entropy change proposed in~\cite{KauLeeAiMos16,KauTakAiLee17,KauTakAiLee18,KauAiLee16}.

\subsection{Probabilistic Occupancy Grid Mapping}
\label{subsec:POGM}
Let map $m$ be an occupancy grid map decomposed into $n_m$ evenly-spaced cubic grid cells with fixed edge length $\alpha$. The $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,\ldots,n_m}$, that is defined as $\mathbf{m}_i=1$ when occupied, and $\mathbf{m}_i=0$ when free. Let another random variable be defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience, which is the event that the $i$-th cell is free, i.e., $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$. The location and size of each grid cell is known, and the random variables composing $m$ are assumed mutually independent, i.e., $P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i)$, so $2^{n_m}$ maps are possible.

Consider a range sensor that repeatedly scans the environment surrounding the robot and returns the distances to the closest occupied spaces. Let the measurement scan at the $t$-th time step be $Z_t$, where $z\in Z_t$ is a single measurement ray. Let the $t$-th pose be denoted by $X_t=\braces{x_t,R_t}$ with location $x_t\in\Re^3$ and attitude $R_t\in\SO=\braces{R\in\Re^{3\times3}|R^\T R=I,\det{R}=1}$, such that $X_t$ is assumed to be known. Given the sensor probabilistic properties $p(z|X_t,m)$, commonly referred to as the \emph{forward sensor model}, the goal is finding the \emph{inverse sensor model}, $P(m|z,X_{1:t},Z_{1:t-1})$. The authors of~\cite{KauLeeAiMos16,KauTakAiLee17} proposed a reduced map $r\subset m$ composed of $n_r$ cells along $z$, where cells are indexed by increasing distance from $x_t$. Defining $\mathbf{r}_{k+}$ as the event that the $k$-th cell of map $r$ is the closest occupied space, the inverse sensor model for the $k$-th cell is
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})&=\eta\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1}),
\end{align}
where $\eta$ is the normalizer,
\begin{align}
\label{eqn:allEta}
\eta
&=
\bigg[\sum_{i=1}^{n_{r}+1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\} p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_i^-\bigg]^{-1},
\end{align}
such that the probability summation of the entire space is $1$, and the unnormalized probability of the inverse sensor model is defined as
\begin{align}
\label{eqn:Unnormalized}
\tilde P(\mathbf{r}_{k}|z,X_{1:t},&Z_{1:t-1})\nonumber\\
&=\mathbf{P}_k^-
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]\nonumber\\
&\quad + \bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{k+},X_t)\mathbf{P}_k^-,
\end{align}
where a priori probability is $\mathbf{P}_k^-=P(\mathbf{r}_{k}|X_{1:t-1},Z_{1:t-1})$, its complement is $\bar{\mathbf{P}}_k^-=1-\mathbf{P}_k^-$, $P(\bar{\mathbf{r}}_{0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience, and $p(z|\mathbf{r}_{(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading. The proofs of \refeqn{RayISMAnswer}--\refeqn{Unnormalized} are given in~\cite{KauLeeAiMos16}. Since $\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})$ from \refeqn{Unnormalized} uses several repeated terms from the prior $\tilde P(\mathbf{r}_{k-1}|z,X_{1:t},Z_{1:t-1})$, and $\eta$ is easily obtained from these as well, the computational cost of \refeqn{RayISMAnswer} is linear with respect to the number of cells along a measurement ray, amortized to $\mathcal{O}(1)$ for each cell, so the exact inverse sensor model can be applied in real-time. The process of updating each cell along a measurement ray is repeated for all rays composing scan $Z_t$.

\subsection{Predictive Entropy-Based Autonomous Exploration}

Here we present how the probabilistic properties of an occupancy grid map provide key information about the uncertainty of the space for motion planning. Shannon's entropy commonly serves as an uncertainty measure~\cite{StaGriBur05}. Provided grid cell probabilities of map $m$, Shannon's entropy is defined as
\begin{align}
\label{eqn:ShannonsEntropyCell}
H(P(\mathbf{m}_i))&=-P(\mathbf{m}_i)\log{P(\mathbf{m}_i})-P(\bar{\mathbf{m}}_i)\log{P(\bar{\mathbf{m}}_i}),
\\
\label{eqn:ShannonsEntropyMap}
H(P(m))&=\sum_{i=1}^{n_m}H(P(\mathbf{m}_i)),
\end{align}
for an individual cell and the entire map, respectively.
Thus, entropy is maximized when $P(\mathbf{m}_i)=0.5$, which corresponds to the largest uncertainty; similarly, entropy is minimized as $P(\mathbf{m}_i)$ approaches $0$ or $1$, which corresponds to the smallest uncertainty. %; thus Shannon's entropy is a measure of uncertainty.

Suppose a future pose candidate and its associated future measurement scan is $X_c=\braces{x_c,R_c}$ and $Z_c$, respectively, where $c\in\mathcal C$ such that $\mathcal C=\braces{1,2,...,n_c}$ accounts for all $n_c$ candidates under consideration for the next robot pose. The current pose of the robot is \emph{not} $X_c$ in general, so any change to the probabilistic map from $X_c$ must be predicted. Much like the probabilistic mapping, this is achieved ray-by-ray~\cite{KauAiLee16,KauTakAiLee17}. Considering that all grid cell probabilities are conditioned on the history of poses $X_{1:t}$ and measurement scans $Z_{1:t}$, these terms are removed from the remaining equations for simplicity. For candidate ray $z_c$, the expected entropy is
\begin{align}
\label{eqn:DiscExpEntropyRay}
&\text{E}[H(P(m|x_c,z_{c}))]=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(m|x_c,z_{c,k}))P(z_{c,k}|x_c)\bigg\},
\end{align}
where $z_{c,k}$ refers to the distance from $x_c$ to the $k$-th grid cell along the measurement ray. The first term of the summation of \refeqn{DiscExpEntropyRay}, namely $H(P(m|x_c,z_{c,k}))$, is obtained with entropy definitions \refeqn{ShannonsEntropyCell}, \refeqn{ShannonsEntropyMap} and the inverse sensor model \refeqn{RayISMAnswer}--\refeqn{Unnormalized}. The second term is derived from \refeqn{allEta} with
\begin{align}
\label{eqn:ProbMeas}
P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}},
\end{align}
where $\eta_{c,k}$ refers to the normalizer based on the measurement $z_{c,k}$.
The expected negative entropy change for candidate pose $X_c$ is equivalently the expected information gain,
\begin{align}
%\label{eqn:objectiveFunOptimized}
%X_c^*&=\argmax_{X_c}{\ \mathcal I(X_c)},
%\\
\label{eqn:expectedInfoGainRay}
\mathcal I(X_c)&=H(P(m|X_{1:t},Z_{1:t}))-\text{E}\left[H(P(m|X_c,Z_c))\right].
\end{align}
Then, \refeqn{DiscExpEntropyRay} is applied repeatedly over several sample measurement rays that compose scan $Z_c$ as shown in~\cite{KauAiLee16}.

\section{Mapping and Exploring in 3D}

In this section, we present computational techniques to implement the probabilistic mapping and exploration schemes discussed in the prior section to a 3D environment in real-time with explicit consideration of the computation time.

\subsection{Mapping in 3D Space}

Grid cells along a measurement ray are updated using the inverse sensor model. Since a measurement ray is assumed to follow a straight line, the inverse sensor model is evaluated only along that vector. Prior research~\cite{KauLeeAiMos16,KauTakAiLee17} is based on 2D distances from the robot to grid cells, which are obtained geometrically through ray casting. This approach is easily extended to 3D space, where the inverse sensor model is applied identically after ray casting. Each measurement ray is considered individually, which simplifies measurement fusion from any number of sensors with differing stochastic properties. Furthermore, multiple sensor scans may update the map in parallel threads, thereby increasing the number of measurements to update the 3D map. In short, extending the exact inverse sensor model from 2D to 3D with multiple sensors involves straightforward geometry for ray casting and processing the sensor readings in parallel.

A robot might be equipped with multiple sensors for several reasons, e.g., different depth accuracy or fields-of-view. For example, consider a robot carying a 2D LIDAR providing highly accurate measurements over a horizontal plane. This sensor might rapidly update the grid cells at the robot height, but neglect spaces above or below the robot. The robot can compensate for these limitations with an IR depth sensor providing measurements with nonzero components in the vertical direction, even if the IR depth sensor accuracy is lower than the LIDAR accuracy. Measurements from both sensors may update the same probabilistic map because each sensor model is explicitly considered in the cell probability calculations. In short, the proposed 3D mapping approach can fuse the impact of measurements on the probabilistic map from different sources because the stochastic properties of measurements are considered according to a Bayesian framework explicitly.

% EditEK 1: example of 2 sensor measurements fused, starting with "For example, a robot may be equipped..." until the end of the paragraph.

However, several careful considerations are required for 3D implementation. First, increasing the map dimensions from 2D to 3D requires memory reserves to multiply by a factor of the number of cells in the added dimension. Second, rays spanning 3D space tend to intersect more grid cells than rays spanning 2D space, which increases the number of grid cells updated per ray. Computation time for the inverse sensor model is linear with respect to the number of cells considered, so careful considerations on sensor limits and number of rays updated per scan are important.
It is common that computers onboard flying robots lack the processing capabilities to update massive 3D maps quickly due to these limitations. Therefore, a simple solution is to stream the sensor data from an onboard computer to a more powerful computer that runs the mapping.

Additionally, mobile robots, particularly aerial vehicles, frequently make fast movements, which may cause time synchronization issues. The most problematic cases correspond to pose estimates and depth sensor readings not synchronizing. For example, the time stamps from the most recent pose estimate and measurement scan may be significantly different while the robot is rapidly rotating. Combining this pose and scan together violates the assumption that the robot pose is well-known for every measurement scan, i.e., $X_t$ and $Z_t$ are given. Therefore, pose estimates and sensor readings must be paired according to their time stamps. A useful tool to accomplish this requirement is the Message Filters package with the Robot Operating System (ROS), where time stamps are synchronized using recently received data of poses and measurement scans.

\subsection{Autonomous Exploration in 3D Space}

We propose an autonomous exploration strategy that selects motions that move the robot to poses that are expected to maximize information gain. The probabilistic map is used for acquiring entropy as a measure of uncertainty, and for collision avoidance. For these processes, careful consideration must be placed on computation. In particular, the computational complexity for expected entropy of a single measurement ray intersecting $n_r$ cells is $\mathcal O(n_r^2)$. To address this issue, we propose projecting the 3D map onto a 2D occupancy grid map, which is located at the desired height of exploration. This approach corresponds to vertically-uniform environments such as rooms and hallways without complex geometries. Projecting the occupancy grid map simplifies the exploration through 3D space, and decreases $n_r$, thereby increasing the algorithm speed.

Next we describe how the 2D map projection is defined such that important information about collision and entropy is extracted from the 3D map. Let a 2D projected map be denoted $m_\text{2D}$. The two horizontal dimensions, namely $x$ and $y$, have edge length $\alpha_{xy}=k_{xy}\alpha$, where $k_{x,y}\geq1$ is an integer representing the relative edge length between the 3D map $m$ and the projected map $m_\text{2D}$. The value of $k_{x,y}$ is selected such that the area of each cell in $m_\text{2D}$ is at minimum the footprint size of the robot for simplified collision analysis. The number of cells in the vertical $z$-direction, namely $k_z\geq1$, is selected such that $k_z\alpha$ is at minimum the robot size in the vertical direction, and at maximum $k_z\alpha$ is height of the environment excluding floors and ceilings.

Here, we show how multiple the 3D cells from map $m$ are used to find the probability of a single cell in $m_\text{2D}$. 
Let the $k$-th cell of the projected map be denoted $\mathbf{m}_{\text{2D},k}$. Let the set $a_k\subset m$ correspond to 3D cells falling within the dimensions $k_{x,y}\times k_{x,y}\times k_z$ centered at the location of $\mathbf{m}_{\text{2D},k}$. Considering that the cells begin with uniform probability $P_\text{init}$ where $0<P_\text{init}<1$, define a threshold $P_\text{thresh}$ such that $P_\text{init}<P_\text{thresh}<1$. Then, the probability of $\mathbf{m}_{\text{2D},k}$ is
\begin{align}
\label{eqn:Proj2DMapComb}
P(\mathbf{m}_{\text{2D},k})= 
\begin{cases}
    \max_{i\in a_k}{P(\mathbf{m}_i)},			&\text{if} \ \max_{i\in a_k}{P(\mathbf{m}_i)}\geq P_\text{thresh}\\
    \min_{i\in a_k}{P(\mathbf{m}_i)},              & \text{otherwise}
\end{cases}.
\end{align}
For example, if the $i$-th cell belonging to $a_k$ is considered occupied, i.e., $P(\mathbf{m}_i)>P_\text{thresh}$, then the projected 2D cell takes on the maximum probability of cells in $a_k$, and therefore is considered occupied. This conservative choice is beneficial for collision-avoidance. Alternatively, if all cell probabilities in $a_k$ fall below $P_\text{thresh}$, and the $i$-th cell probability is $P(\mathbf{m}_i)\approx0$ (typically from measurement rays passing through), then the projected cell takes on the minimum probability of cells in $a_k$, and therefore is considered free.

There are several advantages to this approach. First, the $k$-th cell probability $P(\mathbf{m}_{\text{2D},k})$ remains at $P_\text{init}$ unless the cell is measured at least once. This corresponds to a large expected information gain \refeqn{expectedInfoGainRay} for capturing this unmeasured space. Second, the collision risk is simply calculated, and can be easily modified by changing $P_\text{thresh}$. Additionally, the case when $P(\mathbf{m}_{\text{2D},k})<P_\text{thresh}$ typically implies that this cell is free, thereby increasing the impact cells further away on exploration. Furthermore, this projected map is much simpler to analyze for exploration.

\section{Multi-Vehicle Bidding-Based Autonomous Exploration}
\label{sec:BiddingExploration}

When exploring uncertain environments, robots can generate large maps much faster if they coordinate their efforts. However, the problem of determining a multi-vehicle exploration strategy is complicated and computationally expensive, but must be performed in real-time.

In this section, we formulate a cooperative and autonomous exploration scheme based on sequential auctions for maximizing map information gain. The first auction is based on the expected information gain presented in the prior section, scaled by travel distance. Once the winner of the first auction is determined, the map for bidding is updated based on the expected measurements of the first robot. The second auction for the remaining robots is determined by the information gain from the revised map that is further scaled by the travel distance and a penalty for collision-avoidance. This process is repeated until all robots are assigned tasks.

%While auctions proceed, the occupancy grid cell probabilities are degrading according to a continuous-time Markov process. Cell degradation increases the expected information gain, thereby incentivizing robots to revisit regions after some time passes.

%Here, we formulate a cooperative patrol scheme based on a coordinated autonomous exploration with multiple robots using sequential auctions. The first auction is based on the expected information gain presented in the prior section, scaled by travel distance. Once the winner of the first auction is determined, the map for bidding is updated based on the expected measurements of the first robot. The second auction for the remaining robots is determined by the information gain from the revised map that is further scaled by the travel distance and a penalty for collision-avoidance. This process is repeated until all robots are assigned tasks. While auctions proceed, the occupancy grid cell probabilities are degrading according to a continuous-time Markov process. Cell degradation increases the expected information gain, thereby incentivizing robots to revisit regions after some time passes.

\subsection{Objective Function for the First Auction}

Here, we describe how robots participate in the first auction for maximizing map information gain while accounting for travel time. The information gain \refeqn{expectedInfoGainRay} is computed for each candidate pose. Expected information gains may vary among the robots only if they have different sensor configurations. All expected information gains must be computed prior to the first auction.

Next, we describe how travel time is integrated into exploration. Considering that distances from current robot poses to candidate poses may differ greatly, accounting for these varying travel times properly is essential for exploration time efficiency. Travel distances are computed using Dijkstra's search to provide collision-free waypoints for each robot. There are two steps: first, generate a cost map from the robot location to each location on the 2D projected map. This provides the collision-free distances to all reachable locations on the map. Second, the waypoints from a candidate pose to the current robot pose are easily obtained along the cost map using steepest descent.

%    if x(k) < d
%        f(k) = 0.5*fMax*(1-cos(pi*x(k)/d));
%    else
%        f(k) = fDiff*exp(-1e-2*(x(k)-d)^2)+fFar;
%    end

%Dijkstra's search demonstrates that the distances from each robot to $X_c$ are different in general, so travel times to $X_c$ vary. This motivates a unified approach to account for travel time, which promotes local travel and discourages robots from traversing across the map when possible.

Here, we define a bump function to account for varying travel distances from Dijkstra's search. Let the distance along the collision-free path from the $k$-th robot pose, namely $X_k$, to the $c$-th candidate pose, namely $X_c$, be denoted $d(X_k,X_c,m_\text{2D})\geq0$, which is taken from the cost map belonging to $X_k$. A continuous bump function is defined to account for traveling costs, and is composed of two parts.

The first part of the bump function corresponds to short trajectories. Consider $d_\text{opt}$, the time-optimal distance that a robot may travel at full speed during the time between exploration updates. Here, we consider the case when $d(X_k,X_c,m_\text{2D})\leq d_\text{opt}$, i.e., the robot can completely traverse this distance during the allotted travel time. If $d(X_k,X_c,m_\text{2D})\ll d_\text{opt}$, then the robot is not moving at its full potential. This can be wasteful, because the robot tends to capture larger regions while moving. Therefore, it is desirable for $d(X_k,X_c,m_\text{2D})\rightarrow d_\text{opt}$, which corresponds to maximizing map coverage without time cost. The first part of the bump function is sinusoidal and is optimized at $d_\text{opt}$ to maximize robotic movement as
\begin{align}
\label{eqn:BumpFunIncreasing}
\mathcal B_1(d)=\frac12 f_\text{max}\left(1-\cos{\frac{d\pi}{d_\text{opt}}}\right),
\end{align}
where $f_\text{max}>0$ is the maximum value when $d(X_k,X_c,m_\text{2D})=d_\text{opt}$.

The second part of the bump function is defined over the domain where \\$d(X_k,X_c,m_\text{2D})>d_\text{opt}$ to minimize time required for $X_k$ to arrive at $X_c$, putting a penalty on traversing across the environment. This choice is beneficial to generating accurate local maps before exploring new regions. The second half is defined to be nonzero everywhere and is strictly decreasing as 
\begin{align}
\label{eqn:BumpFunDecreasing}
\mathcal B_2(d)=(f_\text{max}-f_\text{far})\exp\braces{-\beta(d_\text{opt}-d)^2}+f_\text{far},
\end{align}
where $f_\text{max}>f_\text{far}>0$ guarantees $\mathcal B_2>0$ such that $\mathcal B_2\rightarrow f_\text{far}$ as $d(X_k,X_c,m_\text{2D})\rightarrow\infty$ and $\beta>0$ assigns the rate of functional decrease relative to $f_\text{max}$ and $f_\text{far}$. Then, the complete bump function is defined as
\begin{align}
\label{eqn:BumpFun}
\mathcal B(d)=
\begin{cases}
    \mathcal B_1(d),		& \text{if }d\leq d_\text{opt},\\
    \mathcal B_2(d),         & \text{otherwise},
\end{cases}
\end{align}
which is illustrated in Fig. \ref{fig:nonzeroBumpFun}. % In short, the bump function prioritizes short movements to capture the local environment over long traversals for exploration time efficiency.

	\begin{figure}
		\centerline{
			\includegraphics[width=0.9\columnwidth]{NonZeroBump.pdf}
		}
		\caption{The bump function is designed to promote travel distances close to $d_\text{opt}$ by multiplying this function by the expected information gain objective. Travel distances less than $d_\text{opt}$ are below the robot capabilities. Travel distances beyond $d_\text{opt}$ are beyond what the robot can reach within the minimum exploration computation time.}
		\label{fig:nonzeroBumpFun}
	\end{figure}

Then, using the information gain from \refeqn{expectedInfoGainRay} and considering the impact from travel distance from \refeqn{BumpFun}, the objective function for a single agent with respect to $X_{c}$ is
\begin{align}
\label{eqn:CandidateBidSingle}
\text{Obj}_\text{single}(X_k,X_c,m_\text{2D})&=\mathcal B(d(X_k,X_c,m_\text{2D}))\mathcal I(X_{c},m_\text{2D}).
%X_{k,c}^*&=\argmax_{X_c}{\ \mathcal B_{k,c}\mathcal I(X_c)},
\end{align}
Then, the optimal pose selection for the $k$-th agent is
\begin{align}
\label{eqn:OptPoseSingle}
X^*_{k}=\argmax_{X_c}\text{Obj}_\text{single}(X_k,X_c,m_\text{2D}).
\end{align}
Therefore, the bump function from \refeqn{BumpFun} prioritizes robotic movement within the local surrounding space before the robot moves across the map to distant regions, while still considering faraway pose candidates.

The first auction includes bids from all robots, and proceeds as follows. Let the set of all $n_R$ robots be denoted $\mathcal R=\braces{R_1,R_2,\dots,R_{n_R}}$ such that the $k$-th robot, namely $R_k$, bids $\text{Obj}_\text{single}(X^*_k,X_c,m_\text{2D})$ from \refeqn{CandidateBidSingle} and \refeqn{OptPoseSingle}. This is repeated for all $k$ such that $R_k\in\mathcal R$. The robot with the largest bid wins the first auction at index $k^*$, and is assigned to travel to $X^*_{k^*}$. Then, the remaining robots must account for the planned trajectory of the robot that won the first auction in subsequent auctions, described next.

\subsection{Subsequent Auctions}

The robots unable to win the first auction are updated to reflect the expected impact of $R_{k^*}$. This robot is expected to modify the probabilistic occupancy grid map, which may significantly change the expected information gains and collision properties of candidate poses nearby $X^*_{k^*}$. The process of auctioning and modifying bids is repeated, removing the winning robot from consideration after each auction, until no robots remain. Between auctions, updating the map with expected measurements from auction-winning robots discourages the remaining robots from entering the same regions and capturing the same cells. As such, there is no need to consider the coverage overlap explicitly: it is avoided in a systematic way as increased coverage overlap would reduce the overall information gain.

% Furthermore, an additional term serves to avoid collisions between robots.

%When agents submit their first candidate bid, namely $\text{Obj}_\text{single}(X^*_k,X_c,m_\text{2D})$ from \refeqn{CandidateBidSingle} and \refeqn{OptPoseSingle}, this serves to maximize the expected information gain of the map while accounting for distance. However, this bid alone disregards the impact of other robots. Hence, using bids generated this way repeatedly would lead to uncoordinated robot efforts.

More specifically, we coordinate robotic efforts by modifying bids to prevent robots from updating the same grid cells and to avoid collisions among robots. The goal is to find the winning bids $\mathcal X^*=\braces{X^*_{1},X^*_{2},\dots,X^*_{n_R}}$ corresponding to each robot from $\mathcal R$. During the auctioning process, let $\mathcal W\subset\mathcal R$ be the set of robots that have already won auctions by producing the largest objective function bid. After the first auction described above, $\mathcal W=\braces{R_{k^*}}$; after all auctions are complete, $\mathcal W=\mathcal R$. Between auctions, the candidate poses are modified for information gain and collision-avoidance, coordinating the multi-vehicle exploration.

% The main idea is that accounting for the expected change in the occupancy grid map from an auction-winning robot provides a multi-vehicle exploration policy that avoids excessive coverage of the same grid cells.

The first step between auctions is to modify the expected map information gain for efficient map coverage.  Once a robot wins a bid, the measurement ray expected values serve to update a temporary copy of $m_\text{2D}$, namely $m_\text{2D,copy}$, and those candidates located in a local neighborhood (twice the radius of a maximum sensor reading) of the winning candidate from the prior auction are recalculated, thereby coordinating total map information gain with auction-winning robots. Using the same local map notation from \refeqn{allEta} and \refeqn{Unnormalized} along a measurement ray, the expected measurement value is
\begin{align}
\label{eqn:ExpectedMeasRay}
\text{E}[z]=\sum_{k=1}^{n_{r}}\bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}\mathbf{P}_k^-z_k,
\end{align}
where $z_k$ denotes the distance from the robot sensor to the $k$-th cell along the measurement ray. Then $\text{E}[z]$ is substituted into \refeqn{RayISMAnswer}--\refeqn{Unnormalized} to modify $m_\text{2D,copy}$, and expected map information gains are recomputed with \refeqn{ProbMeas} and \refeqn{expectedInfoGainRay}, while the bump function \refeqn{BumpFun} remains the same. In short, the expected changes of $m_\text{2D}$ are integrated into $m_\text{2D,copy}$ for expected information gain estimation between multiple vehicles.

The second step focusses on collision-avoidance between robots based on their proximity. Let $\rho_\text{max}>0$ be a fixed maximum radius to consider collision-avoidance and $\rho_{i,c}\geq0$ be the Euclidean distance from the $i$-th already-assigned pose $X^*_i$ to the $c$-th candidate pose $X_c$, i.e.,
\begin{align}
\rho_{i,c}&=\norm{x^*_i-x_c},
\end{align}
where $x^*_i$ and $x_c$ are the pose locations of $X^*_i$ and $X_c$, respectively. Then the collision-avoidance factor for the $k$-th robot such that $R_k\notin\mathcal W$ is,
\begin{align}
\label{eqn:CollisionAvoidanceAmongRobots}
\mathbf C(X_c)&=
\begin{cases}
    \prod_{i|\mathcal{R}_i\in\mathcal W} \left(\frac{\rho_{i,c}}{\rho_\text{max}}\right)^2,		& \text{if }\rho_{i,c}<\rho_\text{max},\\
    1,              				& \text{otherwise},
\end{cases}
\end{align}
where this product serves to decrease the value of bids for candidate poses in close proximity with already-assigned poses to avoid collisions between robots. The multi-vehicle objective function for the $k$-th robot accounting for collision-avoidance is
\begin{align}
\label{eqn:CandidateBidMulti}
\text{Obj}_\text{multi}&(X_k,X_c,m_\text{2D,copy})
\nonumber\\&=\mathbf C(X_c)\mathcal B(d(X_k,X_c,m_\text{2D}))\mathcal I(X_{c},m_\text{2D,copy}),
\end{align}
where its optimal pose during this auction is
\begin{align}
\label{eqn:OptPoseMulti}
X^*_{k}=\argmax_{X_c}\text{Obj}_\text{multi}&(X_k,X_c,m_\text{2D,copy}),
\end{align}
where $\text{Obj}_\text{multi}(X^*_k,X_c,m_\text{2D,copy})$ is the bid for the $k$-th robot.
Optimal pose selection and bidding is repeated for all robots not belonging to $\mathcal W$. The largest bid among these wins the auction and is tasked with moving to the associated pose candidate, and then this robot is included with $\mathcal W$ to avoid further consideration. Completing $n_R$ auctions produces $\mathcal X^*$.

%In short, the robot with the largest bid wins the first auction, then discounts all candidate poses within a small neighborhood of the auction-winning robot, and this robot is removed from further auctions. This process is repeated, removing a robot from consideration each auction, until just a single robot remains.

The proposed approach uses auctions and bid modifications based on total map expected information gain and collision-avoidance, promoting exploration of different spaces without explicit consideration of coverage overlaps. The pseudocode for the bidding process is shown with Algorithm \ref{alg:bidding}.

\begin{algorithm}
	Function: $Bidding(\mathcal X_r,\mathcal I(X_c)\forall c\in\mathcal C,\rho_\text{max})$\;
	Initialize $k=0$, $\mathcal W=0_{n_R\times1}$, $B=0_{n_R\times1}$\;
	Update candidate expected information gains using \refeqn{DiscExpEntropyRay}\;
	\For{$i=n_R,n_R-1,\ldots,1$}{
		\For{$j=1,2,\ldots,n_R$}{
			\If{$\mathcal W(j)==0$}{
				\If{$i==n_R$}{
					Maximize bid \refeqn{CandidateBidSingle} for $X^*_{j}$ from \refeqn{OptPoseSingle}\;
				}
				\Else{
					Update $\mathcal I(X_c)$ close to $\mathcal X^*(k)$\;
					\For{$c=1,2,\ldots,n_c$}{
						Find Euclidean distance $\rho_{k,c}$\;
					Get $\mathbf C(X_c)$ with \refeqn{CollisionAvoidanceAmongRobots}\;
					}
					Maximize bid \refeqn{CandidateBidMulti} for $X^*_{j}$ from \refeqn{OptPoseMulti}\;
				}
				Insert the $j$-th maximum bid into $B(j)$\;
			}
		}
		$k$: index maximizing $B$ with corresponding $X^*_{k}$\;
		$\mathcal X^*(k)=X^*_{k}$\;
		$\mathcal W(k)=1$\;
		$B(k)=0$\;
	}

	Return $\mathcal X^*$\;
\caption{Robot Task Bidding}
\label{alg:bidding}
\end{algorithm}

%              C++
%              if(distToOptimalPose < explore.discountRadius)
%                proximityDiscount = pow(distToOptimalPose/explore.discountRadius, 2);
%              else
%                proximityDiscount = 1.0;

%\begin{align}
%\label{eqn:BumpFun}
%\text{bump}(d_\text{cell})= 
%\begin{cases}
%    e^{1-\frac1{1-(d_\text{cell}/d_\text{max})^2}},			& \text{if }-d_\text{max}<d_\text{cell}<d_\text{max}\\
%    0,              				& \text{otherwise}
%\end{cases},
%\end{align}

\subsection{Receding Horizon Framework for a Dynamic Environment}
This algorithm is further aided by following a receding horizon framework for improved information gain maximizations and collision-avoidance with dynamic obstacles and other robots. A receding horizon simply repeats the autonomous exploration steps as quickly as possible over a finite time period, frequently before a robot reaches its desired pose. Since exploration optimizations can only occur during exploration updates, a receding horizon maximizes the rate at which these updates occur. These updates require time for computation, over which the robot can move $d_\text{opt}$, used to optimize the robot travel distances with \refeqn{BumpFun}. Hence, the receding horizon framework serves to repeat optimizations as quickly as possible with maximal robotic movement while exploring a changing occupancy grid.

Furthermore, a receding horizon framework enhances autonomous exploration in two other ways. First, the occupancy grid map is constantly updated while a robot traverses a trajectory, so the expected information gains change as well. Since the bids depend on the occupancy grid, $\mathcal X^*$ is updated accordingly. This prevents robots from completing trajectories that have become unnecessary as new terrain becomes captured. Second, the receding horizon prevents robots from colliding with each other when they become too close together (e.g., crossing paths, traversing a tight passage) or other moving obstacles (e.g., a human). The cost map for each robot is updated, and the location of other robots are considered as collision-prone space. Hence, the robot trajectories become better separated due to rapid updates of the cost maps with the receding horizon.

\section{Multi-Vehicle Cooperative Patrol}
\label{sec:MultiVehicleCooperativePatrol}

Patrol is the continuous and repetitive observation of different regions within a larger environment. An effective patrol algorithm is particularly valuable for surveillance operations, especially when the robot members are incapable of measuring all regions of an environment from fixed poses. Here, we formulate cooperative patrol as an optimization problem, where expected map information gains are maximized cooperatively according to the methods outlined in Section \ref{sec:BiddingExploration}. However, the grid cells experience slow degradation over time, thereby modifying the expected map information gains that govern exploration. This way, different regions are revisited after some time passes. This method of cell degradation for cooperative robotic patrol is described next.

\subsection{Continuous-Time Markov Process for Cell Degradation}

The key idea to cell degradation is that as time progresses forward, cells become more uncertain, which causes the exploration algorithm to revisit these cells. All cells are degraded through a continuous-time Markov process at an equal rate, thereby creating a fair policy for revisiting regions that exploits the benefits of multi-vehicle exploration, such as following a receding horizon and optimizing information gain and travel time. Cell degradation is applied simply and uniformly to the entire 3D map for autonomous cooperative patrol. Since this analysis is valid for all grid cells, cell index notation is neglected for this analysis.

Consider that a single grid cell begins with probability $P(\mathbf{m}_{t_0})$ at time $t_0$ and degrades to $P(\mathbf{m}_{t_f})$ at time $t_f>t_0$. Suppose that as degradation time progresses forward after $t_0$, the cell probability approaches a terminal value $P_\infty\in[0,1]$, i.e.,
\begin{align}
\label{eqn:DegradationQualification}
\lim_{t\rightarrow\infty}\begin{bmatrix}
P(\mathbf{m}_{t})
\\
P(\bar{\mathbf{m}}_{t})
\end{bmatrix}
=
\begin{bmatrix}
P_\infty
\\
1-P_\infty
\end{bmatrix}.
\end{align}
Let the degradation rate be $\lambda>0$ and matrix $A\in\Re^{2\times2}$ be defined as
\begin{align}
A=\lambda
\begin{bmatrix}
-(1-P_\infty) & P_\infty
\\
(1-P_\infty) & -P_\infty
\end{bmatrix}.
\end{align}
Then the first-order ordinary differential equation is expressed as
\begin{align}
\begin{bmatrix}
\dot{P}(\mathbf{m}_{t})
\\
\dot{P}(\bar{\mathbf{m}}_{t})
\end{bmatrix}
=
A
\begin{bmatrix}
P(\mathbf{m}_{t})
\\
P(\bar{\mathbf{m}}_{t})
\end{bmatrix}.
\end{align}
Hence, the state transition from $t_0$ to $t_f$ is
\begin{align}
\label{eqn:DegradeStateTransition}
\begin{bmatrix}
P(\mathbf{m}_{t_f})
\\
P(\bar{\mathbf{m}}_{t_f})
\end{bmatrix}
&=
\exp\braces{A(t_f-t_0)}
\begin{bmatrix}
P(\mathbf{m}_{t_0})
\\
P(\bar{\mathbf{m}}_{t_0})
\end{bmatrix},
\end{align}
which is illustrated with Fig. \ref{fig:MarkovDegradeContinuous}. Most importantly, \refeqn{DegradeStateTransition} satisfies \refeqn{DegradationQualification} as $t_f\rightarrow\infty$. Solving the top row of \refeqn{DegradeStateTransition} yields
\begin{align}
\label{eqn:CellDegradationScalar}
P(\mathbf{m}_{t_f})=P(\mathbf{m}_{t_0})\exp\braces{-\lambda (t_f-t_0)}+P_\infty(1-\exp\braces{-\lambda (t_f-t_0)}).
\end{align}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{markov_diagram_continuous.pdf}
\caption{The continuous-time Markov process for a single binary cell occupancy is illustrated. As time progresses, this process degrades the cell slowly toward $P_\infty.$}
\label{fig:MarkovDegradeContinuous}
\end{figure}

Applying degradation is computationally-inexpensive due to the simplicity of \refeqn{CellDegradationScalar}. For an arbitrary positive fixed time step, $\exp\braces{-\lambda (t_f-t_0)}$ needs only be calculated once and is multiplied to every map cell. Similarly, the second term of \refeqn{CellDegradationScalar}, namely $P_\infty(1-\exp\braces{-\lambda (t_f-t_0)})$, is simply added to every cell. For grid cells with probabilities far from $P_\infty$, the times to converge to this final degraded value increase, shown in Fig. \ref{fig:DegradeExamples}. Slow convergence of cells with probabilities close to $0$ or $1$ allows the robots to stop viewing these cells, giving the robots opportunities to visit different regions before returning.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{DegradeExamples.pdf}
\caption{The rate at which probabilities converge toward $P_\infty=0.4$ increases based on the difference magnitude of their initial values.}
\label{fig:DegradeExamples}
\end{figure}

In short, each grid cell is continuously degrading according to \refeqn{CellDegradationScalar} toward $P_\infty$. Only in the presence of measurements does the cell occupancy become well-known according to \refeqn{RayISMAnswer}--\refeqn{Unnormalized}. Cell degradation encourages the robot to revisit cells from the distant past using bidding-based multi-vehicle autonomous exploration to patrol the environment continuously. Adopting the receding horizon framework allows patrol to account for dynamic obstacles by quickly updating patrol strategies as map information changes.

\section{Single-Vehicle Exploration Experimental Results}
\label{sec:ExploreExperiment}

In this section, we present results from a single-vehicle autonomous exploration experiment. A robots builds a 3D map of its environment, which is projected onto a 2D map for exploration. Grid cell probabilities are determined from the onboard sensors without degradation, so this experiment focusses on exploration rather than patrol.

\subsection{Exploration Environment}

The U.S. Naval Research Laboratory (NRL) has a large experimental space for testing known as the Laboratory for Autonomous Systems Research (LASR). Here, we designed an experimental setup containing walls and objects to resemble a building floor plan. The perimeter walls were constructed with tan cardboard, which were suspended from $1$m tall stanchions, and an internal wall was built with gray metal sheeting mounted to $80/20$ supports~\cite{url_8020}. Additionally, a hard flooring reflected some florescent lighting above. The objects inside the experimental environment included various trash cans, desks, and chairs in addition to a few miscellaneous objects. The map size was selected to slightly exceed the volume of the experiment such that the the x-direction (positive east) spanned $0.5$m to $10.3$m, the y-direction (positive north) spanned $-8.3$m to $3.5$m, and the z-direction (positive vertically up) spanned $-0.15$m to $1.5$m. Two images of the experimental environment are displayed in Fig. \ref{fig:exp3DEnvironment}.

\begin{figure}[!t]
\centering
    	\begin{subfigure}[t]{0.95\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_north.jpg}
        		\caption{North Region}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.95\columnwidth}
	\vspace*{0.03\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_south.jpg}
        		\caption{South Region}
    	\end{subfigure}
	\caption{The 3D experimental environment is designed to resemble a large room within an office building. The walls are build from from cardboard suspended from stanchions and gray sheeting with $80/20$ supports, and miscellaneous objects are placed around the room as well.}
	\label{fig:exp3DEnvironment}
\end{figure}

Next we describe the key parameters used for mapping and exploration. The inverse sensor model \refeqn{RayISMAnswer}--\refeqn{Unnormalized} relies on the stochastic properties of two sensors, namely an Asus Xtion color IR depth sensor and a Hokuyo LIDAR. The Xtion and Hokuyo sensors are modeled using a modified version of the beam model for range finders~\cite{ThrBurFox05}. This includes uniform distribution for a phantom reading with probability $P_\text{rand}=0.1$ and a Gaussian distribution with probability $P_\text{hit}=0.9$. The Gaussian distribution corresponds to the desired case when the measurement $z$ correctly captures a particular cell at expected distance $\hat{z}$ with standard deviation $\sigma$. The total forward sensor model for the Xtion and Hokuyo is
\begin{align}
%p_\text{total}(z,z_\text{min},z_\text{max},\hat{z},\sigma)&=P_\text{rand}p_\text{rand}(z,z_\text{min},z_\text{max})+P_\text{hit}p_\text{hit}(z,\hat{z},\sigma),
&p_\text{total}=P_\text{rand}p_\text{rand}+P_\text{hit}p_\text{hit},
\\
&p_\text{rand}(z,z_\text{min},z_\text{max})=\frac1{z_\text{max}-z_\text{min}},
\\
&p_\text{hit}(z,\hat{z},\sigma)=\frac1{\sqrt{2\pi\sigma^2}}\exp\braces{\frac{(\hat{z}-z)^2}{{2\sigma^2}}},
\end{align}
such that $z_\text{min}=0.5$m and $z_\text{max}=4.0$m for both the Xtion and Hokuyo, and the standard deviations are $\sigma_\text{Xtion}=0.25$m and $\sigma_\text{Hokuyo}=0.1$m, respectively, which exceed their specified values to account for localization uncertainty and sensor distortion of the Xtion. These sensors update the probabilities of cubic 3D cells with edge length $\alpha=0.075$m and initial probability $P_\text{init}=0.1$.

The grid cells for the projected 2D map used during exploration are $3$ times the size, $0.225$m. Since this experiment does not require multi-vehicle patrol with a receding horizon framework, the bump function is simplified to decrease with increasing distance, similar to \refeqn{BumpFunDecreasing} following the structure of~\cite{Joh06} with $d_\text{max}>0$,
\begin{align}
\label{eqn:BumpFunRef}
\text{bump}(d)= 
\begin{cases}
    \exp\braces{{1-\frac1{1-(d/d_\text{max})^2}}},			& \text{if }0<d<d_\text{max}\\
    0,              				& \text{otherwise}
\end{cases},
\end{align}
such that $d_\text{max}$ is selected as the maximum cost map value. Then, $\text{bump}(d)$ is multiplied to \refeqn{expectedInfoGainRay} when finding optimal poses. The bump function prioritizes actions that improve the knowledge of the local surrounding space before distant traversals across the map.

\subsection{Hardware Structure}

Several components are used for localization, sensing, and actuation. A Vicon motion capture system provides the transformation between a reference frame fixed to the world and the moving vehicle. This transformation, along with fixed transformations between the robot body and the sensors, provides sufficient information for the transformation from the world to each sensor frame. The sensor readings from the Xtion and Hokuyo are transmitted from an NVidia Jetson TX2 onboard the vehicle (Fig. \ref{fig:QuadrotorHardware}). The sensor readings are received via WIFI on a host computer with an Intel Core i7-6800K CPU (12$\times$3.40GHz), which runs the mapping and exploration nodes. The host computer returns the exploration trajectories to the Jetson TX2 over WIFI. Then, the Jetson executes a nonlinear geometric flight controller~\cite{GooDaeLee13} onboard. With this configuration, large processing tasks are avoided onboard the robot.

\begin{figure}[!t]
\centering
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{quad_top.png}
        		\caption{Quadrotor Top}
    	\end{subfigure}
	\hspace*{0.05\columnwidth}
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{quad_bottom.png}
        		\caption{Quadrotor Bottom}
    	\end{subfigure}
	\caption{A Jetson TX2 (quadrotor top) streams depth measurements via WIFI from an Asus Xtion and Hokuyo LIDAR (quadrotor bottom). The host computer builds a 3D probabilistic occupancy grid map from these measurements. For autonomous exploration, the host computer designs exploration trajectories, which an onboard flight controller can track in real-time.}
	\label{fig:QuadrotorHardware}
\end{figure}

\subsection{Software Structure}

The software is developed using the Robot Operating System (ROS) framework for simple communication among the processes. The programs, referred to as \emph{nodes}, are primarily coded in C++ for efficiency, except a GUI and trivial message conversions, which use Python scripts. Additionally, a few launch, yaml, cfg scripts were used for running nodes and setting parameters.

The first node creates a 3D probabilistic map. This node uses transform message filters (a variation of ROS message filters) to synchronize sensor transformations and associated sensor readings. This way, each measurement origin and direction is well-known, and is applied to ray casting in 3D, where all possible intersections between the measurement and grid cells are computed through geometry. Next, the cell probabilities along this ray are updated with the inverse sensor model \refeqn{RayISMAnswer}--\refeqn{Unnormalized}. The measurements are analyzed ray-by-ray, so any number of depth sensors with known stochastic properties may be used. For this experiment, the stochastic properties of the Xtion and Hokuyo are listed inside a yaml configuration file.

There are three nodes for exploration. The first node projects 3D grid cells onto a 2D map according to \refeqn{Proj2DMapComb}, then publishes this 2D map for visualization and further analysis. The second node subscribes to the 2D map for Dijkstra's search and information gain optimization. This node produces path messages, which consist of desired poses $0.01$sec apart for visualization using the ROS package, Rviz. Finally, an additional node sends the desired pose at the current time to the geometric nonlinear controller~\cite{GooDaeLee13} onboard the robot.

\subsection{Experimental Results}

The quadrotor takes off, completes a rotation to understand its immediate surroundings, then explores the space over $2$ minutes and $47$ seconds. Fig. \ref{fig:exp3DMap} shows the probabilistic 3D map generated in real-time at four selected times, and the corresponding 2D projected maps (x-direction upward) are illustrated in Fig. \ref{fig:exp2DMap}, where arrows represent pose candidates and a trajectory is shown to the optimal pose. Fig. \ref{fig:expH} shows the total map entropy decreasing over the experiment duration. A video of the experiment can be found at \href{https://www.youtube.com/watch?v=I_1rXV2XRqk}{https://www.youtube.com/watch?v=I\_1rXV2XRqk}.

% trim={<left> <lower> <right> <upper>}

\begin{figure}[!t]
\centering
    	\begin{subfigure}[t]{0.95\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_ogm3D_0min.jpg}
        		\caption{$0$ min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.95\columnwidth}
		\vspace*{0.03\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_ogm3D_1min.jpg}
        		\caption{$1$ min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.95\columnwidth}
	\vspace*{0.03\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_ogm3D_2min.jpg}
        		\caption{$2$ min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.95\columnwidth}
	\vspace*{0.03\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_ogm3D_2min47sec.jpg}
        		\caption{$2$ min $47$ sec (end)}
    	\end{subfigure}
	\caption{The 3D occupancy grid map is overlaid with point cloud measurements. The map probabilities directly rely on the Xtion and Hokuyo measurements and the sensor stochastic properties.}
	\label{fig:exp3DMap}
\end{figure}

\begin{figure}[!t]
\centering
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_0min_2D.jpg}
        		\caption{$0$ min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_1min_2D.jpg}
        		\caption{$1$ min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_2min_2D.jpg}
        		\caption{$2$ min}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.44\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{experiment_2min47sec_2D_labeled.jpeg}
        		\caption{$2$ min $47$ sec (end)}
    	\end{subfigure}
	\caption{The 2D projected map is easily produced in real-time. Candidate future poses, shown by arrows, have more blue color for larger objective functions, found with expected information gains and travel distances. Collision-free waypoints (blue line) serve as input to a constrained polynomial least-squares path (light green curve) for the robot controller to follow, where the desired pose (large axes) is tracked by the onboard controller from the current robot pose (small axes).}
	\label{fig:exp2DMap}
\end{figure}

\begin{figure}[!t]
	\centering
	\includegraphics[width=0.85\textwidth]{entropy_experiment_aspect3by1.png}
	\caption{Map entropy decreases as the robot explores and captures more of the experimental environment.}
	\label{fig:expH}
\end{figure}

The robot is able to produce a 3D occupancy grid in under $3$ minutes without a human providing a trajectory. The 3D map can be easily interpreted by a human, and provides sufficient information in real-time for autonomous exploration. Optimal poses and collision-free exploration trajectories are calculated by the exploration node in under $1$ second. These quick decisions are vital to the autonomous exploration performance since the map is initially uncertain.

The projected map used for both collision and exploration has several benefits. The robot measures walls and objects with the 3D map, and nicely represents these as occupied spaces with the 2D projected map. Additionally, the exploration computation is low, allowing only for brief hover periods (less than $1$ second) between trajectories. However, the robot exploration does not consider those cells far below the robot, particularly those near the ground. As a result, some floor regions are not captured by the 3D occupancy grid even though cells above the floor are well-known.

\section{Multi-Vehicle Patrol Simulation Results}
\label{sec:PatrolSimulation}

In this section, we present a multi-vehicle exploration and patrol simulation. The environment is much larger, and the rate of measurements is increased with six active sensors, presenting several computational challenges. Furthermore, a human is walking around the space, unknown to the cooperating robots, requiring a receding horizon framework to avoid collisions.

\subsection{Simulated Environment}

Several parameters used during the multi-vehicle patrol simulation are identical to those of the prior section, with a few key differences relating specifically to the changed simulated environment and the bump function. There are three quadrotors with identical sensors as the last section cooperatively bidding on future poses in a much larger space. The environment is based on The George Washington University's (GWU) Science and Engineering Hall (SEH) floor plan, shown in Fig. \ref{fig:SEHEnvironment}. The map limits are much larger to accommodate the space, modified to $-55.0$m to $55.0$m in the x- and y-directions, and $0.0$m to $1.5$m in the z-direction, producing a total of $45,255,504$ grid cells with edge length $0.075$m (dimensioned $1468\times1468\times21$). For the bump function, $f_\text{max}=1$, $f_\text{far}=0.1$, and $\beta=0.01$ are chosen. For bidding $\rho_\text{max}=3.0$m is chosen.

\subsection{Software Structure}

The same ROS structure is largely used from the prior section, with key differences relating to process scalability and increased functionality with three vehicles patrolling the space. The mapping node, which computes 3D mapping probabilities, is modified in two ways: first, a thread is allocated to each sensor to update the map due to the much greater number of measurements, now originating from an Asus Xtion and Hokuyo LIDAR from each of three robots (six sensor scans total). An additional thread handles cell degradation with $\lambda=1\times10^{-3}$, which corresponds to a cell half life of roughly $11.5$ minutes; if this time passes since a measurement captures a cell, the occupancy probability will be halfway between its value $11.5$ minutes ago and $P_\infty=P_\text{init}=0.1$. While degrading cells with \refeqn{CellDegradationScalar} is inexpensive, degradation must cycle through all cells, which can be costly with a map containing over $45$ million cells.

The exploration node is modified as well to handle a larger map with degrading cell probabilities. This node is composed of four threads. The first thread updates a local copy of the 3D occupancy grid map by subscribing to map changes from each sensor scan, which updates a small portion of the 2D projected map each time. This is accomplished by finding a rectangular prism of the minimum and maximum limits of cells that may have changed during a particular measurement scan update. Publishing and subscribing to map changes provides an efficient means to transfer mapping information, independent of map size. This way, robots understand their immediate surroundings quickly. The second node subscribes to the full 3D map, acquiring just the memory necessary to update the entire 2D projected occupancy grid map. This process is slower than the first thread due to the large 3D map size, but vital for tracking degradation of grid cells far from the current robot positions. The third thread runs the bidding-based information gain and travel distance optimizations following the receding horizon framework. During this thread, a local copy of the projected 2D map is locked to ensure a fair bidding process. The final thread simply broadcasts the projected map for visualization. The node for interpreting trajectory commands is the same as that of the prior section, and is run three times (once for each robot) in parallel. In short, this node structure is capable of multi-vehicle exploration and patrol with a receding horizon, and updates all map cell probabilities, prioritizing those close to the current robot positions.

\begin{figure}[!t]
\centering
    	\begin{subfigure}[t]{0.4\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{SEH_cropped.png}
        		\caption{Floor Plan}
    	\end{subfigure}
	\hspace*{0.1\textwidth}
    	\begin{subfigure}[t]{0.4\columnwidth}
           	\centering
          	\includegraphics[width=\textwidth]{gazeboSEH.png}
        		\caption{Gazebo Model}
    	\end{subfigure}
	\caption{The SEH second floor plan (about $15,000$ sqft) is reconstructed with a Gazebo simulation environment.}
	\label{fig:SEHEnvironment}
\end{figure}

\subsection{Simulated Results}

In the simulation, three quadrotors explore and patrol the simulated environment of The George Washington University's (GWU) Science and Engineering Hall (SEH) for $15$ minutes while taking measurements. Initially, the robots are given no information about the environment except that their immediate surroundings are free. Unknown to the robot team, a simulated human walks in a rectangular repeated pattern at $0.5$ m/s. The 3D occupancy grid maps are shown in Fig. \ref{fig:sim3DMap}, and the corresponding 2D projected maps for exploration are shown in Fig. \ref{fig:sim2Dmaps}. The full video of both maps is available at \href{https://youtu.be/bsLG2romP_8}{https://youtu.be/bsLG2romP\_8}.

\begin{figure}[!t]
\centering
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_1min.jpg}
        		\caption{$1$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
	\hspace*{0.05\textwidth}
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_3min.jpg}
        		\caption{$3$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_5min.jpg}
        		\caption{$5$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
	\hspace*{0.05\textwidth}
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_7min.jpg}
        		\caption{$7$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_9min.jpg}
        		\caption{$9$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
	\hspace*{0.05\textwidth}
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_11min.jpg}
        		\caption{$11$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_13min.jpg}
        		\caption{$13$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
	\hspace*{0.05\textwidth}
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_15min.jpg}
        		\caption{$15$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
	\caption{The 3D occupancy grid map representation of the SEH second floor is shown during patrol. The robots generate these maps while avoiding collisions with each other, the environment, and a non-cooperative human.}
	\label{fig:sim3DMap}
\end{figure}

\begin{figure}[!t]
\centering
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_1min.jpg}
        		\caption{$1$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
	\hspace*{0.05\textwidth}
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_3min.jpg}
        		\caption{$3$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_5min.jpg}
        		\caption{$5$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
	\hspace*{0.05\textwidth}
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_7min.jpg}
        		\caption{$7$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_9min.jpg}
        		\caption{$9$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
	\hspace*{0.05\textwidth}
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_11min.jpg}
        		\caption{$11$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_13min.jpg}
        		\caption{$13$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
	\hspace*{0.05\textwidth}
    	\begin{subfigure}[t]{0.38\columnwidth}
           	\centering
          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\textwidth]{Patrol_Split_Screen_15min.jpg}
        		\caption{$15$ min}
		\vspace*{0.05\textwidth}
    	\end{subfigure}
	\caption{The 2D projected map from of the SEH second floor shows degradation while robots patrol the space. The robots evade the non-cooperative human walking in a rectangular pattern, occasionally capturing the temporarily-occupied space and identifying these cells as occupied on the grid.}
	\label{fig:sim2Dmaps}
\end{figure}
%	% trim={<left> <lower> <right> <upper>}

The robot generated a fairly-complete occupancy grid within $9$ minutes, and proceeded to patrol the space during the remaining time. After $9$ minutes, the total map entropy no longer increases or decreases significantly; the rate at which cell degradation increases map entropy roughly cancels the rate at which new measurements decrease map entropy. In contrast, if the robots simply hover instead of patrol after $9$ minutes, the map uncertainty quickly increases, shown in Fig. \ref{fig:DegradeExamples}.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{entropy_patrol_9minSwitch.pdf}
\caption{Total map entropies are tracked for separate trials: patrol of the building for $15$ minutes (blue) and patrol of the building for $9$ minutes, then hover (red). Even though the map is fairly well-known after $9$ minutes in either case, the continuous effort of patrolling over the complete time period shows lower map uncertainty beyond this time as cells are constantly degrading.}
\label{fig:DegradeExamples}
\end{figure}

Special attention had to be placed on parameter selection for the bump function and collision-avoidance. For the bump function, increasing the parameter $\beta$ narrows the bump, placing a stronger influence on nearby candidates. However, this choice corresponds to the bump function decreasing at a negligible rate beyond a close neighborhood of the robot, i.e., a large range of distances yield $\mathcal B\approx f_\text{far}$. Hence, there is a tradeoff between prioritization of local movements and being able to differentiate the effects of distance on large trajectories.

When robots competed for future poses in a centralized auction framework, the near-optimal solution provided an effective multi-vehicle exploration policy. Between auctions, only a small set of candidates within the neighborhood of the last auction-winning candidate require consideration when updating expected information gains with \refeqn{ExpectedMeasRay} and avoiding collisions between robots using \refeqn{CollisionAvoidanceAmongRobots}, both of which modify the optimization of \refeqn{OptPoseMulti}. Since bidding is applied efficiently, its computation time is negligible compared with other parts of the exploration update such as computing the initial candidate information gains and robot cost maps. Furthermore, the receding horizon framework minimizes time between exploration updates, which allows the robots quickly react to dynamic obstacles like the non-cooperative human, shown in Fig. \ref{fig:EvadeHuman}.

\begin{figure}[!t]
\centering
    	\begin{subfigure}[t]{0.3\columnwidth}
           	\centering
          	\includegraphics[trim={22cm 25cm 20cm 0cm}, clip, width=\textwidth]{evade_15sec.jpg}
        		\caption{$15$ sec}
    	\end{subfigure}
	\hspace*{0.02\textwidth}
    	\begin{subfigure}[t]{0.3\columnwidth}
           	\centering
          	\includegraphics[trim={22cm 25cm 20cm 0cm}, clip, width=\textwidth]{evade_16sec.jpg}
        		\caption{$16$ sec}
    	\end{subfigure}
	\hspace*{0.02\textwidth}
    	\begin{subfigure}[t]{0.3\columnwidth}
           	\centering
          	\includegraphics[trim={22cm 25cm 20cm 0cm}, clip, width=\textwidth]{evade_17sec.jpg}
        		\caption{$17$ sec}
    	\end{subfigure}
	\caption{A cooperating robot must evade a non-cooperative human. The receding horizon framework allows a quickly-updating map to modify the exploration commands to avoid collisions with a dynamic obstacle.}
	\label{fig:EvadeHuman}
\end{figure}

The results also demonstrate an interesting tradeoff between exploring new terrain and patrolling previously-visited spaces. When the degradation rate $\lambda$ is increased, grid cells are degrading faster, incentivizing the exploration strategy to revisit spaces faster. This behavior is beneficial for frequent patrol, but certain areas, particularly those difficult to reach, are not visited as quickly or as frequently. Should an area receive extra attention, a greater value of $\lambda$ may be applied to this region.

\section{Conclusions}
\label{sec:Conclusions}

This paper introduces a near-optimal bidding-based solution to multi-vehicle exploration and patrol to map a 3D space. The proposed approach integrates recent developments in obtaining occupancy grid probabilities and entropies into several auctions to assign tasks to robots. The bidding process relies on expected map information gains and travel distances, and modifies bids between auctions to promote robotic cooperation. Information gains are modified by updating a temporary expected map, and an additional term prevents collisions among robots. Grid cells are continuously degraded using a continuous-time Markov process, which incentivizes the autonomous exploration-based patrol scheme to revisit regions after time passes. Furthermore, a receding horizon constantly updates the bidding process and improves collision avoidance within a dynamic environment. The efficacy of the approach is shown with a numerical simulation patrolling a large floor plan with a walking human, with several computational enhancements for scalability of the algorithms.

%\section*{Appendix}\label{append}
%
%
%\subsection*{Proof of Proposition 1}
%
%Continuous Markov process

\bibliography{../../BibSources}
\bibliographystyle{IEEEtran}

\vspace*{0.1\textwidth}

\noindent{\bf Evan Kaufman} received his B.S. degree in Mechanical Engineering from Bucknell University in 2012, and is currently a Ph.D. candidate at The George Washington University in the Mechanical and Aerospace department. His research focusses are in robotic mapping and exploration with range-based sensors, estimation, data association, and flight controls. He has worked on several research projects through the U.S. Air Force Research Laboratory and Naval Research Laboratory.

\vspace*{0.1\textwidth}

\noindent{\bf Kuya Takami} is a postdoctoral research fellow at the Department of Mechanical and Aerospace Engineering at the George Washington University. He received his doctoral degree in Mechanical Engineering at Virginia Polytechnic Institute and State University in 2015 specializing in robot perception. His main interests include mobile robot perception and autonomous systems.

\vspace*{0.1\textwidth}

\noindent{\bf Zhuming Ai} is a Computer Engineer in the Information Technology Division at the U.S. Naval Research Laboratory (NRL). Before he joined NRL, he served as the Director of the Virtual Reality in Medicine Lab at the University of Illinois at Chicago. He is interested in virtual reality, augmented reality, computer graphics, image processing, and robotics.

\vspace*{0.1\textwidth}

\noindent{\bf Taeyoung Lee} is an associate professor of the Department of Mechanical and Aerospace Engineering at The George Washington University. He received his doctoral degree in Aerospace Engineering and his master?s degree in Mathematics at the University of Michigan in 2008. His research interests include geometric mechanics and control with applications to complex aerospace systems.

\end{document}

### 2018-06-09 Journal/JINT 2018/Related Work/BackgroundResearch.tex
journal added

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Journal of Robotics and Autonomous Systems}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Background Research on Patrol}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author[GWU]{Evan Kaufman\corref{mycorrespondingauthor}}

\author[GWU]{Kuya Takami}

\author[NRL]{Zhuming Ai}

\author[GWU]{Taeyoung Lee}
%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}

\fntext[GWU]{Department of Mechanical and Aerospace Engineering\\
The George Washington University, Washington, DC 20052\\
Email: \{evankaufman, kuya, tylee\}@gwu.edu}
\fntext[NRL]{Information Management \& Decision Architectures
\\ U.S. Naval Research Laboratory, Washington, DC 20375}

%\author[mysecondaryaddress]{Global Customer Service}
\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{support@elsevier.com}

%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

\end{frontmatter}

\begin{itemize}
	\item Survey
	\begin{itemize}
		\item A Survey on Multi-robot Patrolling Algorithms~\cite{PorRoc11}
		\begin{itemize}
			\item Refs 2,3,4 use unweighted or weighted graphs to visit various environment nodes using local or global information
			\item Ref. 5 uses minimal-cost cyclic patrol paths (Hamilton cycles) at an optimal frequency, and independent of robot number (robust, best for largely-connected closed-loop graphs)
			\item Ref. 6 is like 5 except it uses a TSP cycle and segmenting the environment into different partitioned regions for patrolling (best for long corridors)
			\item Ref. 7 uses reinforcement learning to select actions to maximize a long-term performance criterion
			\item Ref. 8 has auctions between negotiating agents to visit random vertices to minimize visits to the same vertex
			\item Ref. 9 considers the length of patrolling roots for negotiations, but is only tested in open space
			\item Ref. 12 uses a neural net for patrol patterns, where a threat mode is activated when an intruder is detected
			\item Ref. 13 has robots drop pheromones that evaporate to indicate time since a robot has visited
			\item Ref. 14 focusses on catching intruders with various algorithms
			\item Refs 15,16 present a centralized and efficient algorithm for multi-level partitioning of the space to assign different regions to different agents
		\end{itemize}
	\end{itemize}
	\item Cell periodicity
	\begin{itemize}
		\item Spectral Analysis for Long-Term Robotic Mapping~\cite{KraFenCieDonDuc14}, Persistent Localization and Life-long Mapping in Changing Environments using the Frequency Map Enhancement~\cite{KraFenHanDuc16}
		\begin{itemize}
			\item Long-term dynamics of objects are predicted by re-observing the same locations on an occupancy grid using the periodicity of individual cells
			\item Experiment with exploring and navigating a human-populated environment for long-term autonomy
			\item Uses the \emph{FreMEn map} which provided probabilities as a function of time based on re-observations of each cell
		\end{itemize}
	\end{itemize}
	\item Voronoi Diagrams
	\begin{itemize}
		\item Extracting Surveillance Graphs from Robot Maps~\cite{KolCar08}
		\begin{itemize}
			\item Create surveillance graphs from occupancy grids: breaking down environment into partitions and tracking intruders through regions
			\item Voronoi Diagram extracted from 2D maps
			\item Use \emph{blocking}: detect intruders passing through a heavily-sensed edge
			\item Use \emph{contractions}: graph modifiers to use less robots for the same surveillance tasks
		\end{itemize}
		\item MSP Algorithm Multi-Robot Patrolling based on Territory Allocation using Balanced Graph Partitioning~\cite{PorRoc10}
		\begin{itemize}
			\item Ref. 16 of survey paper
			\item Known environment
			\item Assigns different regions (subgraphs) to different robots
			\item Deterministic
		\end{itemize}
		\item Performance Based Task Assignment in Multi-Robot Patrolling~\cite{PipChrWei13}
		\begin{itemize}
			\item Distribute patrol areas among robot team members
			\item Addresses issue when some team members patrol inefficiently
			\item Monitor robot performance and dynamically reassigns tasks from poorly-performing members
			\item Assume map provided in advance
			\item Partition using same Voronoi approach as Portugal and Rocha
		\end{itemize}	
	\end{itemize}
	\item Other Metrics
	\begin{itemize}
		\item A Patrolling Scheme in Wireless Sensor and Robot Networks~\cite{ZhaXia11}
		\begin{itemize}
			\item Use pheromones to show where robots were last, but does not reference Ref. 13 of survey paper
		\end{itemize}
		\item Multi-robot patrol: A distributed algorithm based on expected idleness~\cite{YanZha16}
		\begin{itemize}
			\item Decentralized patrol algorithm along a graph to minimize idleness
			\item Assumes a graph is already known, possibly from a Voronoi Diagram
		\end{itemize}
	\end{itemize}
\end{itemize}
% NOTE: at GW, look up "degradation" and "patrol" together (maybe "occupancy grid map")

%Example reference: \cite{KauTakAiLee17}.

\section*{References}

\bibliography{../../../BibSources}% master source for all publications

\end{document}

### 2018-06-09 Conference/SCITECH19/SCITECH19.tex
started numerical results txt

@@
 
+In this section, we demonstrate the efficacy of the proposed 3D mapping and autonomous exploration with a simulated Mars environment.
+First we cover the software structure and parameters, then compare results using two different maps for entropy prediction.
+
+\subsection{Software}
+
+The mapping, exploration, and visualization processes communicate over the Robot Operating System (ROS) framework, and Gazebo 8 serves as the simulator for the physical objects and measurements.
+
+\subsection{Results}
+
 \begin{itemize}

### 2018-06-09 Dissertation/Proposal/DissertationProposalEvanKaufmanNov16.tex
gitignore & dissertation updated

--- 2018-06-09 Conference/SIMPAR 2016/SIMPAR16 Presentation.tex
@@
   \leavevmode%
-  \hbox{\begin{beamercolorbox}[wd=.9\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{author in head/foot}%
+  \hbox{\begin{beamercolorbox}[wd=.7\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{author in head/foot}%
     \usebeamerfont{author in head/foot}\insertshorttitle
   \end{beamercolorbox}%
-  \begin{beamercolorbox}[wd=.1\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{title in head/foot}
+  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm]{title in head/foot}
     \usebeamerfont{title in head/foot}\mypaper\hfill \insertframenumber/\inserttotalframenumber
+    \usebeamerfont{title in head/foot}\hfill \insertframenumber/\inserttotalframenumber
   \end{beamercolorbox}}%
@@
 
-\title[Autonomous Exploration by Expected Information Gain from Prob. Occupancy Grid Mapping]
-{\large Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping}
+\title[Multi-Robot Probabilistic Mapping and Exploration]{\large Multi-Robot Probabilistic Mapping and Exploration}
 
@@
 \institute{\footnotesize
-{\normalsize Evan Kaufman, Taeyoung Lee, and Zhuming Ai}\vspace*{0.2cm}\\
-  Mechanical and Aerospace Engineering\\ George Washington University \vspace*{0.6cm}\\ {\normalsize Special Thanks to Ira.
+{\normalsize Evan Kaufman\\Research Advisor: Taeyoung Lee}\vspace*{0.2cm}\\
+  Mechanical and Aerospace Engineering\\ George Washington University \vspace*{0.6cm}\\ {\normalsize Special Thanks to Zhuming Ai, \\Ira.
 S.
-Moskowitz and Mark Livingston} \vspace*{0.2cm}\\Information Management \& Decision Architectures\\U.S.
+Moskowitz, and Mark Livingston} \vspace*{0.2cm}\\Information Management \& Decision Architectures\\U.S.
 Naval Research Laboratory}
@@
 \section*{}
-\subsection*{Background and Motivation}
+\subsection*{Research Description}
 
 \begin{frame}
-\frametitle{Motivation}
+\frametitle{Intelligent Microflyer}
 \begin{itemize}
-	\item Building Maps and Exploration
+	\item Description and Motivations
 	\begin{itemize}
-		\item Search and rescue: quickly building a map and finding targets in places with physical or health hazards
-		\item Surveillance: gather information of enemy locations
-    		\item Convenience: autonomous robotic cleaning
+    		\item Description: determine a method for multiple flying vehicles to autonomously explore an uncertain space while generating a map
+		\item Surveillance or search-and-rescue
+		\item Particularly useful for potentially dangerous environments
+%		\item Research problems within these goals are approached with several unrealistic assumptions or computational limitations
+		\item Collaborative effort with the NRL and GWU, focussing on localization, \emph{mapping}, and \emph{autonomous exploration}
 	\end{itemize}
@@
 	\begin{itemize}
-		\item Build \emph{accurate} probabilistic maps
-		\item Use map properties to \emph{predict} how the \emph{map information} will change with various actions
-		\item Motion planning for \emph{collision-free} trajectories
+		\item Determine \emph{accurate} probabilistic maps
+    		\item Avoid inexact solutions using heuristic or learned solutions
+		\item Consider map attributes explicitly to determine an exploration strategy
+		\item Use \emph{map expectations} to plan multiple vehicle paths simultaneously
 		\item Generate algorithms capable of \emph{real-time} implementation
+	\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Outline}
+\begin{itemize}
+	\item Probabilistic Mapping
+	\begin{itemize}
+    		\item Exact solution to a probabilistic map
+		\item Computationally-efficient solution for real-time implementation
+	\end{itemize}
+	\item Single-Robot Autonomous Exploration
+	\begin{itemize}
+    		\item Entropy-based solution relies on map uncertainty
+		\item Direct computationally-efficient solution
+	\end{itemize}
+	\item Future Direction: Multi-Robot Autonomous Exploration
+	\begin{itemize}
+		\item Centralized mapping from several measurement sources
+		\item Robots assigned actions from a bidding-based approach
+	\end{itemize}
+\end{itemize}
+\end{frame}
+
+\section*{}
+
+\begin{frame}
+\frametitle{}
+%\center
+\center{\bf \color{blue} Probabilistic Mapping}
+\end{frame}
+
+\subsection*{Probabilistic Mapping}
+\begin{frame}
+\frametitle{Background}
+
+\begin{itemize}
+    \item Robotic Mapping
+    \begin{itemize}
+    	\item Goal: generate a map representing surrounding regions
+	\item Crucial for simultaneous localization and mapping (SLAM) and autonomous exploration of uncertain environments
+    	\item Popular mapping representations include \emph{occupancy grids}, octomaps, and feature-based maps
+    \end{itemize}
+\item Occupancy Grid Mapping
+\begin{itemize}
+	\item The environment may be decomposed into evenly spaced grid cells that are either \emph{occupied} or \emph{free}
+	\item Probabilistic map: the goal is to obtain the \emph{probability} of each grid cell being occupied
 \end{itemize}
@@
 \end{figure}}
+
 \end{frame}
@@
 \begin{frame}
-\frametitle{Background: Occupancy Grid Mapping}
+\frametitle{Problem Definition}
+
 \begin{itemize}
-		\item Environment decomposed into evenly-space grid cells
-		\item \emph{Binary} random variable: cells are either \emph{occupied} or \emph{free}
-		\item Sensor reading $z$, the \emph{forward sensor model} is $p(z|m,X)$ is the measurement probability density, assuming the stochastic properties are known
+	\item The Map and the Robot
+	\begin{itemize}
+	\item Map $m$ is composed of $n_m$ grid cells with known location and size
+	\item The $i$-th grid cell $\mathbf{m}_i$ is a \emph{static binary} random variable, independent of other grid cells: $P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i)$
+	\item \emph{Pose $X_t$} is known, containing robot \emph{position} and \emph{attitude}
+	\end{itemize}
+\vspace*{0.0cm}\pause
+\end{itemize}
+\begin{minipage}[t]{7.0cm}
+\begin{itemize}
+	\item Depth Measurements
+	\begin{itemize}
+	\item Each measurement origin and direction is known \emph{deterministically}
+	\item A measurement \emph{scan} $Z_t=\braces{z_{t,1},z_{t,2},\ldots,z_{t,n_z}}$ contains $n_z$ measurement \emph{rays} (depths)% at the $t$-th time step, and the history of measurement scans $Z_{1:t}$ is known
+\item The \emph{forward sensor model} is known from the sensor properties
+\end{itemize}
+\end{itemize}
+\end{minipage}
+\begin{minipage}[t]{3.0cm}
+%The \emph{forward sensor model} is the probability density distribution $p(z_{t,l}|m,X_{t})$  known from the sensor properties
+\hspace*{0.25cm}
 \begin{figure}[!htbp]
-		\vspace*{-0.0cm}
+\vspace*{-0.25cm}
 %\vspace*{0.25cm}
 \centerline{
-	  	  \includegraphics[width=7cm]{BeamModel.png}\hspace*{0.1cm}
+    \includegraphics[width=3.5cm]{BeamModel.png}\hspace*{0.1cm}
     }
@@
 \centerline{
-	    	\includegraphics[width=5cm]{1D_True_Grid.png}\hspace*{0.1cm}
+    \includegraphics[width=2.5cm]{1D_True_Grid.png}\hspace*{0.1cm}
 %    \hspace*{0.75cm}
@@
 \end{figure}
-	    	\item Goal: generate a \emph{probabilistic} map $m$ given robot pose $X$
-	\end{itemize}
+
+\end{minipage}
+
 \end{frame}
@@
 \begin{frame}
-\frametitle{Background: Occupancy Grid Mapping}
+\frametitle{Problem Definition}
+
 \begin{itemize}
-		\item Cell occupancy probabilities obtained using the \emph{inverse sensor model}, $P(m|z,X)$, which gives grid cell probabilities based on the latest measurement $z$
-		\item The exact inverse sensor model was believed to require $\mathcal O(2^n)$ for \emph{each} grid cell within the sensor field of view, so popular substitutes were \emph{heuristic} or \emph{learned} solutions
-		\item Recently, the exact inverse sensor model was solved in real-time by grouping map outcomes
+	\item Bayesian Framework
+	\begin{itemize}
+	\item Markov Assumption: latest a priori cell occupancy probabilities capture the information from all prior observations
+	\item Log-Odds Ratio: popular representation due to simple additive update structure and probability truncation avoidance, but requires the \emph{assumption}
+	\begin{align*}
+		P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})\approx P(Z_t|\mathbf{m}_i,X_t)
+	\end{align*}
+	\end{itemize}
+\vspace*{0.0cm}\pause
+	\item Inverse Sensor Model
+	\begin{align*}
+P(\mathbf{m}_i|z_{t,l},&X_{1:t},Z_{1:t-1})\nonumber
+\\
+&=\eta_{t,l} \textcolor{blue}{\sum_{m\in\mathcal{M}_i}}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
+\end{align*}
+	\begin{itemize}
+	\item Given $n$ grid cells: $\textcolor{blue}{\mathcal O(2^n)}$ is \emph{computationally intractable}, motivating a different solution
+	\end{itemize}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Related Work }
+
+\begin{minipage}[t]{7.0cm}
+\begin{itemize}
+	\item Approximate a function for the inverse sensor model based on intuition
+	\begin{itemize}
+		\item Simple to implement, but mathematically inaccurate
+	\end{itemize}
+\end{itemize}
+\end{minipage}
+\begin{minipage}[t]{3.0cm}
 \begin{figure}[!htbp]
-		\vspace*{0.5cm}
+\centerline{
+    \vspace*{-0.5cm}
+%	\hspace*{0.25cm}
+    \includegraphics[width=4.0cm]{Approx_ISM_shortened.png}\hspace*{0.1cm}
+    }
+\centerline{
+    \includegraphics[width=3.5cm]{1D_True_Grid.png}
+%    \hspace*{0.75cm}
+    }
+		\end{figure}
+\end{minipage}
+\vspace*{-0.5cm}
+\begin{itemize}
+	\item Find a solution through learning
+	\begin{itemize}
+		\item Simulate maps, poses, and measurements and use learning to obtain an inverse sensor model
+		\item Complicated, unclear how these parameters are chosen
+	\end{itemize}
+	\item Goal: design a \emph{simple} and \emph{accurate} occupancy grid mapping method avoiding log-odds ratio assumptions
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Single Measurement Ray}
+
+\begin{itemize}
+    \item Main idea: make use of occupancy grid mapping \emph{assumptions} and extract \emph{patterns} from probabilistic properties to find a \emph{computationally-efficient} solution
+	\begin{itemize}
+		\item Since the origin and direction of each measurement ray is known deterministically, the set of grid cells that the ray intersects is \emph{known through geometry}
+		\item A depth reading follows the forward sensor model, which \emph{only depends} on the first occupied grid cell along the measurement ray
+	\end{itemize}
+\end{itemize}
+
+\begin{figure}[!htbp]
+\vspace*{-0.25cm}
 \centerline{
@@
 \end{figure}
-		\item By grouping, the inverse sensor model requires $\mathcal O(n+1)$, compared with $\mathcal O(2^n)$: applied in \emph{real-time}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Single Measurement Ray}
+\begin{itemize}
+    \item Reduced Map
+	\begin{itemize}
+		\item Consider a reduced map of the $l$-th measurement ray, namely $r_l=\braces{\mathbf{r}_{l,1},\mathbf{r}_{l,2},\ldots,\mathbf{r}_{l,n_{r,l}}}$, consisting of grid cells \emph{intersected} by the measurement ray within the sensor limits, \emph{indexed by increasing distance}
+		\item Since reduced map $r_l$ is composed of binary random variables, it has \emph{$2^{n_{r,l}}$} possible solutions
+		\item However, the combinations of possible maps are grouped based on the \emph{closest occupied cells}
+		\item Probability calculations have several repeated terms, and may be determined \emph{recursively}
+		\item The number of possible map combinations is just \emph{$n_{r,l}+1$}
+%		\item If $\mathbf{r}_{l,1}$ is occupied (a priori probability $P(\mathbf{r}_{l,1}|X_{1:t-1},Z_{1:t-1})$), then each occupancy $\mathbf{r}_{l,2:n_{r,l}}$ does \emph{not} affect forward sensor model $P(z_{t,l}|\mathbf{r}_{l,1},X_t)$
+%		\item Similarly, if $\mathbf{r}_{l,2}$ is occupied (a priori probability $P(\bar{\mathbf{r}}_{l,1}|X_{1:t-1},Z_{1:t-1})P(\mathbf{r}_{l,2}|X_{1:t-1},Z_{1:t-1})$), then forward sensor model $P(z_{t,l}|\mathbf{r}_{l,2},X_t)$ is independent of grid cells of higher index
+%		\item This concept is extended for all grid cells in $r_l$
+	\end{itemize}
 \end{itemize}
@@
 
-% check here
 %\begin{frame}
-%\frametitle{Exact Occupancy Grid Mapping: Hallway Comparison}
-%	\begin{itemize}
-%		\item Comparison between approximate and exact results
-%		\begin{figure}[!htbp]
-%		    \centering
-%		    \begin{subfigure}{.45\textwidth}
-%		        \centering
-%		        \includegraphics[width=\textwidth]{map_comparison.png}
-%		        \caption*{Occupancy Grid Map}
-%		    \end{subfigure}
-%		    \hspace*{-.05\textwidth}
-%		    \begin{subfigure}{.45\textwidth}
-%		        \centering
-%		        \includegraphics[width=\textwidth]{map_zoom.png}
-%		        \caption*{Zoomed In Map}
-%		    \end{subfigure}
-%		\end{figure}
-%	\end{itemize}
+%\frametitle{Single Measurement Ray}
+%\begin{itemize}
+%	\item Let $P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$
+%	\item The $k$-th unnormalized probability is written generally as
+%	{\tiny
+%	\begin{align*}
+%		& \tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})\\\nonumber&\quad=P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\&\quad\qquad\times 
+%		\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]\nonumber\\
+%		&\quad\quad +\qquad\quad \ \ \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\} p(z_{t,l}|\mathbf{r}_{l,k+},X_t)P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})
+%	\end{align*}
+%	}
+%	\pause
+%	\item The \emph{normalizer} $\eta_{t,l}$ is
+%	{\tiny
+%	\begin{align*}
+%		\eta_{t,l}=
+%		\bigg[\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1}
+%	\end{align*}
+%	}
+%\end{itemize}
 %\end{frame}
@@
 \begin{frame}
-\frametitle{Exact Occupancy Grid Mapping: Hallway Comparison}
-\begin{minipage}[t]{0.4\textwidth}
+\frametitle{Practical Implications}
+
 \begin{itemize}
-	\item Occupancy grid mapping using an approximate and the exact solution are compared
-	\item Both algorithms subject to the same pose and measurement sets 
-	\item The exact algorithm yields a more \emph{certain} map shown with entropy $H$ (blue: certain, red: uncertain)
+	\item Computational Efficiency
+	\begin{itemize}
+%		\item Several of the terms required for $\eta_{t,l}$ and $\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ for each $k=1,2,\ldots,n_{r,l}$ are repeated and need not be computed more than once
+		\item \emph{Computational order $\mathcal O(n_{r,l}+1)$} for all grid cells in reduced map $r$: each grid cell once and the empty map case
+		\item It is commonly believed that the computational order is $\mathcal O(2^{n_{r,l}})$; the proposed method is \emph{$2^{n_{r,l}}\left(\frac{n_{r,l}}{n_{r,l}+1}\right)$ times faster} for the \emph{same solution}
+		\item Proposed method completely avoids inaccuracies associated with heuristic solutions, learned methods, and log-odds ratio update assumptions
+%		\item Analytic solution is valid for any forward sensor model dependent on the closest occupied space
+	\end{itemize}
+	\item A \emph{ray-by-ray} approach accounts for all measurements rays composing a measurement \emph{scan}
+		\begin{align*}
+			P(m|X_{1:t}&,Z_{1:t})=P((\dots(((m|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z})
+		\end{align*}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Numerical Example}
+
+\begin{minipage}[t]{5.0cm}
+\vspace*{0.5cm}
+\begin{itemize}
+	\item A standard \emph{approximate} inverse sensor model technique is compared with the proposed \emph{exact} method
 \end{itemize}
 \end{minipage}
-	\begin{minipage}[t]{0.5\textwidth}
+\begin{minipage}[t]{5.0cm}
+%\vspace*{0.5cm}
 \begin{figure}[!htbp]
@@
         \centering
-		        \includegraphics[width=.8\textwidth]{AISM_Image_19.pdf}
-		        \caption*{Approx. $P(m)$}
+        \includegraphics[width=\textwidth]{AISM_Image_inf_19.pdf}
+        \caption*{Approximate}
     \end{subfigure}
@@
         \centering
-		        \includegraphics[width=.8\textwidth]{EISM_Image_19.pdf}
-		        \caption*{Exact $P(m)$}
-		    \end{subfigure}
-		\end{figure}
-		\vspace*{-.1\textwidth}
-		\begin{figure}[!htbp]
-		    \centering
-		    \begin{subfigure}{0.5\textwidth}
-		        \centering
-		        \includegraphics[width=.8\textwidth]{AISM_Image_inf_19.pdf}
-		        \caption*{Approx. $H$}
-		    \end{subfigure}
-		    \hspace*{-0.1\textwidth}
-		    \begin{subfigure}{0.5\textwidth}
-		        \centering
-		        \includegraphics[width=.8\textwidth]{EISM_Image_inf_19.pdf}
-		        \caption*{Exact $H$}
+        \includegraphics[width=\textwidth]{EISM_Image_inf_19.pdf}
+        \caption*{Exact}
     \end{subfigure}
@@
 
-%	\begin{itemize}
-%		\item The exact algorithm yields a more \emph{certain} map from the same measurement set
-%		\begin{figure}[!htbp]
+%\begin{figure}[!ht]
 %    \centering
-%		    \begin{subfigure}{0.5\textwidth}
+%    \begin{subfigure}[t]{0.35\columnwidth}
 %        \centering
-%		        \includegraphics[width=.5\textwidth]{AISM_Image_inf_19.pdf}
+%        \includegraphics[width=0.5\textwidth]{AISM_Image_inf_19.pdf}
 %        \caption*{Approximate}
 %    \end{subfigure}
-%		    \hspace*{-0.1\textwidth}
-%		    \begin{subfigure}{0.5\textwidth}
+%    \hspace*{0.5cm}
+%    \begin{subfigure}[t]{0.35\columnwidth}
 %        \centering
@@
 %    \end{subfigure}
-%		\end{figure}
-%	\end{itemize}
+%\end{figure}
+
+\begin{itemize}
+\item The exact algorithm yields a more \emph{certain} map from the same measurement set
+\end{itemize}
+
 \end{frame}
@@
 \section*{}
-\subsection*{Autonomous Exploration}
 
 \begin{frame}
-\frametitle{Autonomous Exploration Description}
+\frametitle{}
+%\center
+\center{\bf \color{blue} Single-Robot Autonomous Exploration}
+\end{frame}
+
+\subsection*{Single-Robot Autonomous Exploration}
+
+\begin{frame}
+\frametitle{Introduction}
 \begin{itemize}
-        \item The environment is initially unknown, so the robot path is initially uncertain
+        	\item The environment is initially uncertain, so the robot path is initially unknown
 	\item Simultaneous localization and mapping assumes robotic actions are provided
 	\item Human interaction is common to guide the robot
-	\item Goal: develop effective policies to govern robotic motion that generates a map accurately and efficiently
+	\item Goal: develop effective policies to govern robotic motion 
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Background: Autonomous Exploration}
+\frametitle{Problem Definition}
 \begin{itemize}
@@
 		\item Main idea: robot moves toward boundaries between certain and uncertain space, takes measurements, then pushes back the boundaries
-		\item Typically not based directly on grid cell entropy
+		\item Non-entropy-based exploration
 		\item Heuristic
@@
 \begin{itemize}
-        \item Instead of relying on frontiers or making assumptions about measurements, the expected map entropy is \emph{calculated directly} by the expected value:
+        	\item Instead of relying on frontiers or making assumptions about measurements, the expected map entropy is \emph{calculated directly} by the law of total expectation:
 	\begin{align*}
@@
 	\end{align*}
-	\pause
-	\item The integral is discretized by intersections between the measurement and grid cells:
-	\begin{align*}
-		\text{E}[H(P(m|X,z))]=\sum_{k=1}^{n+1}\bigg\{H(P(m|X,z_k))P(z_{k}|X)\bigg\}
-	\end{align*}
-	\pause
-	\item Entropy $H(P(m|X,z_k))$ is obtained with the exact inverse sensor model and definition of entropy
-	\item Probability $P(z_{k}|X)$ is obtained with normalizing values from the exact inverse sensor model
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Direct Entropy Calculation: Computational Challenges}
-\begin{itemize}
-        \item Problem: $\text{E}[H(P(m|X,z))]$ requires $\mathcal O((n+1)^2)$ for $n+1$ possible measurement outcomes: measurement detects any of the $n$ cells or returns a maximum reading
-        \pause
-        \item Solution: select the $\hat n$ most probable outcomes for $\mathcal O(\hat n^2)$
-        \pause
-        \item Monte Carlo simulations ($100$ trials) to determine the expected entropy of a measurement ray compare the exact and approximate resulting accuracies and times
-\end{itemize}
-
-		\begin{figure}[!htbp]
-		    \centering
-		    \begin{subfigure}{0.5\textwidth}
-		        \centering
-		        \includegraphics[width=.8\textwidth]{simpar_approx_err.pdf}
-		        \caption*{Error}
-		    \end{subfigure}
-		    \hspace*{-0.1\textwidth}
-		    \begin{subfigure}{0.5\textwidth}
-		        \centering
-		        \includegraphics[width=.8\textwidth]{simpar_approx_t.pdf}
-		        \caption*{Time Ratio}
-		    \end{subfigure}
-		\end{figure}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Pose Selection}
-
-\begin{minipage}[t]{7.0cm}
-\begin{itemize}
-	\item Several locations serve as candidates for the robot to move toward
-	\item The optimal attitude at each location maximized expected information gain
-	\item The pose (location and attitude) with the largest overall expected information gain is selected
-	\item Other costs, such as distance or travel time, can be included in the optimization
-\end{itemize}
-\end{minipage}
-\begin{minipage}[t]{3.0cm}
-\vspace*{0.5cm}
-\begin{figure}[!htbp]
-	\centerline{
-		\hspace*{1.0cm}
-   		\includegraphics[width=5.0cm]{pose_selection}%\hspace*{0.1cm}
-	}
-\end{figure}
-\end{minipage}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Collision-Free Path}
-
-\begin{itemize}
-	\item \emph{Waypoints}
-	\begin{itemize}
-		\item Using the grid map as the graph, \emph{Dijkstra's algorithm} provides collision-free waypoints for the robot path
-		\item Probability of collision at pose $X$ must be within threshold $\beta>0$,
-		\begin{align*}
-			P_\text{collision}(X)=1-\prod_{i\in\mathcal C_X}P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})\leq\beta
-		\end{align*}
-	\end{itemize}
-	\pause
-	\item \emph{Constrained Polynomial Least Squares}
-	\begin{itemize}
-		\item Smooth trajectory as a function of time minimizes the squared error between the curve and the waypoints
-		\item Constrained to beginning and terminal poses
-		\item For complicated trajectories, splines are patched together by constraining position and velocity at the connection
-	\end{itemize}
+	\item The robot pose is chosen to maximize map information gain from this expectation
+	\item Dijkstra's algorithm provides collision-free waypoints for the robot path
+	\item Smooth trajectory: constrained polynomial least squares 
 \end{itemize}
@@
 
-%% VIDEO 1: Simulation %%
 \begin{frame}
-%\href{run:JIRS16_simulation_speedup10x_then_speedup100x.mov}{Simulation Video Link}
 \frametitle{Numerical Results}
 	\centering{
-	\includemovie[poster,autoplay]{\textwidth}{0.66667\textwidth}{JIRS16_simulation_speedup10x_then_speedup100x.mov}}
+	\includemovie[poster,autoplay]{0.720\textwidth}{0.576\textwidth}{JIRS16_simulation_speedup10x_then_speedup100x.mov}}
 \end{frame}
@@
 %    \end{subfigure}
-\vspace*{-0.25cm}
-    \caption*{A robot (green) measures a room in ROS Stage simulator using modified Intel Research Lab floor plan using the proposed mapping and exploration algorithms.
-\href{run:JIRS16_simulation_speedup10x_then_speedup100x.mov}{(Simulation Video Link)}
-}
+    \caption*{A robot (green) measures a room in ROS Stage simulator using modified Intel Research Lab floor plan using the proposed mapping and exploration algorithms.}
     \label{fig:IRL}
 \end{figure}
-\href{run:JIRS16_simulation_speedup10x_then_speedup100x.mov}{Simulation Video Link}
 
@@
 
-%% VIDEO 2: Experiment %%
 \begin{frame}
-%\href{run:JIRS16_experiment_SideBySide_speedup8x.mov}{Experimental Video Link}
 \frametitle{Experimental Results}
 	\centering{
-	\includemovie[poster,autoplay]{\textwidth}{0.5625\textwidth}{JIRS16_experiment_SideBySide_speedup8x.mov}}
+	\includemovie[poster,autoplay]{0.720\textwidth}{0.576\textwidth}{JIRS16_experiment_SideBySide_speedup8x.mov}}
 \end{frame}
@@
 \end{figure}
-\vspace{-0.5cm}
-
-\href{run:JIRS16_experiment_SideBySide_speedup8x.mov}{\tiny Experimental Video Link}
 
@@
 
-% Conclusions and Future directions
-%
-%\section*{}
-%
-%\begin{frame}
-%\frametitle{}
-%%\center
-%\center{\bf \color{blue} Future Direction:\\Multi-Robot Autonomous Exploration}
-%\end{frame}
-%
-%\subsection*{Future Direction: Multi-Robot Autonomous Exploration}
-%
-%\begin{frame}
-%\frametitle{Proposed Topics}
-%\begin{itemize}
-%        \item The ray-by-ray exact occupancy grid mapping approach in centralized probabilistic mapping
-%	\item Multiple robots choosing optimal motions considering each other
-%	\item Robots avoiding collisions with each other
-%	\item Coverage optimization of robots from different poses
-%\end{itemize}
-%
-%\end{frame}
-%
-%\begin{frame}
-%\frametitle{Proposed Future Steps}
-%\begin{itemize}
-%        	\item Mapping:
-%	\begin{itemize}
-%		\item Restructure algorithms for centralized map generation
-%		\item Communicate with multiple measurement sources
-%	\end{itemize}
-%	\item Exploration:
-%	\begin{itemize}
-%		\item Simplify enormous computation: bidding-based approach
-%		\item Temporary theoretical map building
-%		\item Collision avoidance between robots
-%	\end{itemize}
-%\end{itemize}
-%
-%\end{frame}
+\section*{}
+
+\begin{frame}
+\frametitle{}
+%\center
+\center{\bf \color{blue} Future Direction:\\Multi-Robot Autonomous Exploration}
+\end{frame}
+
+\subsection*{Future Direction: Multi-Robot Autonomous Exploration}
+
+\begin{frame}
+\frametitle{Proposed Topics}
+\begin{itemize}
+        	\item The ray-by-ray exact occupancy grid mapping approach in centralized probabilistic mapping
+	\item Multiple robots choosing optimal motions considering each other
+	\item Robots avoiding collisions with each other
+	\item Coverage optimization of robots from different poses
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Proposed Future Steps}
+\begin{itemize}
+        	\item Mapping:
+	\begin{itemize}
+		\item Restructure algorithms for centralized map generation
+		\item Communicate with multiple measurement sources
+	\end{itemize}
+	\item Exploration:
+	\begin{itemize}
+		\item Simplify enormous computation: bidding-based approach
+		\item Temporary theoretical map building
+		\item Collision avoidance between robots
+	\end{itemize}
+\end{itemize}
+
+\end{frame}
 
 \section*{}
-\subsection*{Conclusions}
+
+\begin{frame}
+\frametitle{}
+%\center
+\center{\bf \color{blue} Conclusions and Publications}
+\end{frame}
+
+\subsection*{Conclusions and Publications}
 
@@
 \begin{itemize}
-	\item Map Information
-	\begin{itemize}
-		\item Exact occupancy grid mapping concepts are extended to map uncertainty to accomplish \emph{autonomous exploration}
-		\item Shannon's entropy is calculated \emph{directly} by discretization and using terms from mapping
-	\end{itemize}
-	\pause
-	\item Exploration
-	\begin{itemize}
-		\item Optimal pose selection considers both robot location and orientation
-		\item Dijkstra's algorithm and constrained polynomial least squares provide a \emph{collision-free trajectory}
-		\item The efficacy of the approach is verified with numerical and experimental results
-	\end{itemize}
-	\pause
-	\item Future Steps
-	\begin{itemize}
-		\item 3D mapping and exploration
-		\item Multi-vehicle exploration
-	\end{itemize}
+        	\item Proposed an occupancy grid mapping technique that uses the \emph{exact probabilistic solution}
+	\item Computational cost is reduced \emph{substantially} for \emph{real-time implementation} using probabilisitic properties and exploiting mathematical patterns
+	\item Exact mapping concepts are extended to map uncertainty to accomplish \emph{autonomous exploration}
+	\item Proposing multiple vehicles can explore an uncertain environment using a \emph{bidding}-based approach
 \end{itemize}
@@
 
+\begin{frame}
+\frametitle{Publications}
+{\tiny 
+\begin{itemize}
+	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Autonomous Exploration with Exact Inverse Sensor Models,'' \textit{Journal of Intelligent }\&\textit{ Robotic Systems}, 2016, submitted.
+	\item E.
+Kaufman, T.
+Lee, and Z.
+Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, San Francisco, December 2016, accepted.
+	\item E.
+Kaufman, T.
+Lee, Z.
+Ai, and I.
+S.
+Moskowitz, ``Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model,'' \textit{Proceedings of the American Control Conference}, Boston, July 2016, pp.
+5709-5715.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{The Journal of the Astronautical Sciences}, 63(1): pp.
+60-80, 2016, doi: 10.1007/s40295-015-0082-9.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Minimum Uncertainty JPDA Filters and Coalescence Avoidance for Multiple Object Tracking,'' \textit{The Journal of the Astronautical Sciences}, 63(4): pp.
+308-334, 2016, doi: 10.1007/s40295-016-0092-2.
+	\item T.
+Wu, E.
+Kaufman, and T.
+Lee, ``Globally Asymptotically Stable Attitude Observer on SO(3),'' \textit{Proceedings of the 54th IEEE Conference on Decision and Control}, pp.
+2164-2168, Osaka, Japan, December 2015.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting}, Williamsburg, VA, Jan.
+2015, AAS 15-451.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Minimum Uncertainty JPDA Filter and Coalescence Avoidance Performance Evaluations,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting, Williamsburg}, VA, Jan.
+2015, AAS 15-432.
+	\item E.
+Kaufman, K.
+Caldwell, D.
+Lee, and T.
+Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, Mar.
+2014, ASC 14-2527.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Optimal Joint Probabilistic Data Association Filter Avoiding Coalescence in Close Proximity,'' \textit{Proceedings of the European Control Conference}, pp.
+2709-2714, June 2014.
+	\vspace*{0.5cm}
+\end{itemize}
+}
+\end{frame}
+
 \end{document}

### 2018-06-09 Dissertation/Thesis/tex/ae2D.tex
gitignore & dissertation updated

% !TEX root = ../thesis-sample.tex

\chapter{Autonomous Exploration in 2D Space} \label{chap:ae2D}

\section{Entropy-Based Exploration}

\subsection{Shannon's Entropy}
% Description, figure

\subsection{Expected Information Gain}
% Derivation and link to Appendix

\subsection{Computational Limitations and Approximations}
% SIMPAR figure on entropy/computation time relationship

\section{Future Pose Optimization}

\subsection{Optimal Attitude from Sample Rays}
% SIMPAR figure on optimal pose

\subsection{Travel Cost with Obstacles}
% Dijkstra's cost map

\subsection{Optimal Pose Selection}
% Single-Vehicle argmax

\subsection{Collision-Free Motion Planning}
% Dijkstra's algorithm (after cost map), no least squares yet

\section{Numerical Examples}

\subsection{Exploring a Simple Environment}
% SIMPAR numerical example (2 rooms, 1 hallway)

\subsection{Exploring a Complicated Benchmark Environment}
% JINT17 Fig. 7 & 8

\section{Experimental Result}
% JINT17 9, 10, 11 (ground experiment at the NRL)

\section{Conclusions}

### 2018-06-09 Dissertation/Thesis/tex/ae3Dcomplex.tex
gitignore & dissertation updated

% !TEX root = ../thesis-sample.tex

\chapter{Autonomous Exploration in Complex 3D Space} \label{chap:ae3Dcomplex}

\section{Exploration with Full 3D Maps}

\subsection{Collision Mapping in 3D}

\subsection{Future Pose Optimization in 3D}

\section{Numerical Simulation}
% Mars simulation

\section{Experiment}
% SEH indoor simulation (needs completing)

\section{Conclusions}

### 2018-06-09 Dissertation/Thesis/tex/ae3Dsimple.tex
gitignore & dissertation updated

% !TEX root = ../thesis-sample.tex

\chapter{Autonomous Exploration in Vertically-Uniform 3D Space} \label{chap:ae3Dsimple}

\section{Occupancy Grid Mapping in 3D}

\section{Exploration in Simple Environments }

\subsection{Map Projections}

\section{Exploring a Large Room}
% IRC18 result: Fig. 4, 5, 6, 7

\section{Conclusions}

### 2018-06-09 Dissertation/Thesis/tex/intro.tex
gitignore & dissertation updated

% !TEX root = ../thesis-sample.tex
% COMMENTED OUT DURING EDITS
%\chapter{Introduction} \label{chap:intro}
%
%% Add here
%
%\section{Motivation and Goals}
%
%\subsection{Building Probabilistic Maps}
%
%\subsection{Exploration in Uncertain Spaces}
%
%\subsection{Cooperation Among Autonomous Systems}
%
%\section{Literature Review}
%
%\subsection{Occupancy Grid Mapping}
%
%\subsection{Autonomous Exploration}
%
%\subsection{Multi-Vehicle Cooperation}
%
%\section{Outline of Dissertation}
%
%\section{Contributions}
%\subsection{Summary of Contributions}
%\subsection{List of Publications}

### 2018-06-09 Dissertation/Thesis/tex/multivehicle.tex
gitignore & dissertation updated

% !TEX root = ../thesis-sample.tex

\chapter{Multi-Vehicle Autonomous Exploration and Patrol} \label{chap:multivehicle}

\section{Bidding-Based Autonomous Exploration}

\subsection{Objective Function for the First Auction}

\subsection{Objective Function for Subsequent Auctions}

\subsection{Receding Horizon Framework}

\subsection{Multi-Vehicle Exploration Numerical Simulation}
% Multi-vehicle ICRA figs 3-6 (remake fig. 5 with larger contrast)

\section{Multi-Vehicle Cooperative Patrol}

\subsection{Continuous-Time Markov Process for Cell Degradation}
% JINT18 fig. 3

\subsection{Multi-Vehicle Patrol Numerical Simulation}
% JINT18 fig. 10-13 (ref. ICRA fig 3 from above in this chapter)

\section{Conclusions}

### 2018-06-09 Dissertation/Thesis/dissertation.tex
merged evan: mostly tracking changes in .gitignore

\documentclass[thesis]{thesis-gwu}

% this package is only used to generate some random text. 
% it is not needed in a true document
\usepackage{lipsum}
\graphicspath{{../../Fig/}}

\input{tex/frontmatter}

%% DOCUMENT AREA
\begin{document}

%\include{tex/intro}

\chapter{Introduction} \label{chap:intro}

This dissertation explores three key aspects of mobile robotics: occupancy grid mapping, autonomous exploration, and cooperation among robot teams. Several novel contributions are presented on both related topics for complex and dynamic situations.

\section{Motivation and Goals}

The study of robotics has expanded tremendously in recent years as robots have taken on dangerous, difficult, and repetitive tasks instead of humans. Certain tasks, such as search-and-rescue, surveillance, and cleaning, require some level of autonomy, particularly with mobile robots exploring new and uncertain environments. While these tasks provide significant value to our society, they present nontrivial challenges for the robots, particularly in mapping, exploration, and cooperation.

\subsection{Building Probabilistic Maps}

The first topic this dissertation addresses is building accurate probabilistic maps, known as \emph{mapping}. Map generation is crucial for simultaneous localization and mapping (SLAM) because maps provide important information for the robot to determine its pose (localization), which includes its position and attitude. Mapping also provides knowledge for trajectory planning through an environment and collision avoidance with mapped obstacles. In short, mapping is of fundamental importance in mobile robotics.

Therefore, mapping quality deserves much attention. Robotics is inherently \emph{probabilistic}; no sensor reading or action from an actuator is deterministic. All robotic processes have some degree of uncertainty, and the mapping must account for this. Onboard sensors serve to improve the understanding of surrounding spaces, but the stochastic properties of those sensors must be reflected in the stochastic properties of the map. Therefore, we propose to develop probabilistic map that accounts for the history of measurements and poses, and their associated stochastic properties. Here, each element of the map holds a probability based on measured data. In turn, this probability can be applied to tasks such as collision-avoidance and predicting future map outcomes.

\subsection{Exploration in Uncertain Spaces}

The second topic of this dissertation focusses on the motions robots must choose to maximize their knowledge about the map while avoiding collisions, known as \emph{autonomous exploration}.

\subsection{Cooperation Among Autonomous Systems}

The final topic of this dissertation is cooperation among members of a robot team.

\section{Literature Review}

In this section, we discuss the existing approaches to occupancy grid mapping, autonomous exploration, and multi-vehicle cooperation. Several shortcomings with existing approaches are highlighted, which are addressed in the dissertation contributions.

\subsection{Occupancy Grid Mapping}

\subsection{Autonomous Exploration}

\subsection{Multi-Vehicle Cooperation}

\section{Outline of Dissertation}

ADD ONCE COMPLETE

\section{Contributions}

\subsection{Summary of Contributions}
\subsection{List of Publications}

\include{tex/pogm}
\include{tex/ae2D}
\include{tex/ae3Dsimple}
\include{tex/ae3Dcomplex}
\include{tex/multivehicle}

\bibliographystyle{plain}
\bibliography{thesis-bib}

% appendices must appear after
\include{tex/appendix}
\end{document}

### 2018-06-09 Dissertation/Thesis/tex/frontmatter.tex
merged evan: mostly tracking changes in .gitignore

% !TEX root = ../thesis-sample.tex

% --------- FRONT MATTER PAGES ---------------------
% Title of the thesis
\title{Multi-Robot Probabilistic Mapping and Exploration}

% Author name
\author{Evan Kaufman}

% Previous degrees
\bsdepartment{Mechanical Engineering}
\bsschool{Bucknell University}
\bsgrad{May 2012}

\hidemsdegree

\showcommitteepage % hide this page if you're doing a MS thesis
%\hidecommitteepage 
\committee{ %
Taeyoung Lee\\
Department of Mechanical and Aerospace Engineering\\
The George Washington University\\ 
Dissertation Director\\ % remember to add a space between committee members

James Hahn\\
Department of Computer Science\\
The George Washington University\\
Committee Member\\

Yin-Lin Shen\\
Department of Mechanical and Aerospace Engineering\\
The George Washington University\\ 
Committee Member
}

% Chair must be entered separately for formatting reasons.
\chair{Tayeoung Lee}
\chairtitle{Associate Professor of Mechanical and Aerospace Engineering}
% Department
\department{Mechanical and Aerospace Engineering}

\phdgrad{August 31, 2018}
\defensedate{July 11, 2018}
% Year of completion for copyright page and perhaps other places
\year=2018

% Copyright page
%\copyrightholder{Someone else}

%% Dedication
%\dedication{ 
%Add here...
%%
%%This dissertation is dedicated to my parents, Elizabeth Kaufman and Stephen Kaufman, who have consistently supported me throughout my many years of education. And others...
%}

% Acknowledgments
\acknowledgments{
My academic career has certainly been a long and bumpy road, but I was never alone. I would like thank many people for their support throughout my life for pursuing difficult and interesting work.

First and foremost, I would like to thank my parents. In my younger years, I was not always a great student, but they believed in me. When I struggled, they helped every time. I figured that this is what all parents do, but later I realized how lucky I was. Looking back, my academic career not ending prematurely is largely thanks to my mother and father. They have loved me and supported my academic pursuits, and I am extremely grateful.

I would also like to thank all other parts of my family, including my fianc\'ee, brother, grandparents, aunts, uncles, and cousins. We make it a priority to come together and support each other, even when our careers differ greatly and our lives change unpredictably. Despite the long academic road, and they provide support and love.

Finally, I would like to thank my research advisor, Taeyoung Lee, for the excellent guidance over the last six years. Unlike many relationships between advisors and students, we treats his students like colleagues rather than subordinates. This way, I could focus on completing my research objectives properly rather than trying to impress superior figure. Taeyoung Lee knew just what to say to push my research forward, even when my mind was going another direction. His positive influence over my research has been unparalleled, and I will continue using his valuable lessons.% in the years to come.

 I think that following a passion is something truly special. Even though I find great satisfaction in doing anything well, regardless of how simple or complicated the task, I find myself extremely fortunate to pursue endeavors that I actually find exciting and stimulating. With tremendous support around me, I begin a career in an amazing field and ready to take on future challenges.
}

% -----------------------------------------------------------------
% Typically only one of Preface/Foreward/Prologue would be in your thesis.
% To choose one simply delete the others and they will automatically dissappear

%% Preface
%\preface{
%Add here...
%%    This is the preface. 
%%    It's another front matter page that offers additional detail into your work.
%%    Typically, only one (preface OR prologue OR foreword) is used. 
%%    You can remove the other sections by deleting them inside \texttt{tex/frontmatter.tex} or using the appropriate show or hide commands.
%}

%\prologue{
%    This is the prologue. 
%    It's another front matter page that offers additional detail into your work.
%    Typically, only one (preface OR prologue OR foreword) is used. 
%    You can remove the other sections by deleting them inside \texttt{tex/frontmatter.tex} or using the appropriate show or hide commands.
%}
%
%\foreword[2]{
%    This is the foreword. 
%    It's another front matter page that offers additional detail into your work.
%    Typically, only one (preface OR prologue OR foreword) is used. 
%    You can remove the other sections by deleting them inside \texttt{tex/frontmatter.tex} or using the appropriate show or hide commands.
%}
% ----------------------------------------------------------------------

% commands to show or hide front matter pages

\showcopyright
\showabstract
\showcommitteepage
\hidededication
\showacknowledgments
\hidepreface
\hideprologue
\hideforeword

% ------------ TABLE OF CONTENTS ----------------------
% Commands to hide or show lists of figures, tables, etc.
\showlistoffigures
\showlistoftables
\hidenomenclature

% --------- ACRONYMS and SYMBOLS ------------------------------
% TODO Deprecate the entire acronym package and switch to glossaries

% You can either use the acronymn or glossaries package (both work)
% Definition of any abbreviations used.
\abbreviations{
    \acro{CRTBP}{Circular Restricted Three Body Problem}
    \acro{NSA}{National Security Agency}
    \acro{SSME}{Space Shuttle Main Engine}
}
% call an abbreviation using \ac{abbrev}

% symbols and acronyms only show up when used in the text
\symbols{
    \acro{J}{Moment of Inertia}
}

% if you want acronymn (simpler) then change these to show
\hidelistofabbreviations
\hidelistofsymbols

% if you want glossaries (more powerful) then leave above as hide
% GLOSSARIES package options - automatically turns off front pages from acronym package

% acronymns and symbols are basically the same, but there are two provided 
% locations where they can show up
\setabbreviationstyle[acronym]{long-short}
\setabbreviationstyle[abbreviation]{long-short}
\makeglossaries
% you can hide/show the glossaries page
\showglossarieslistofabbreviations
\showglossarieslistofsymbols
\showglossariesglossaryofterms

% acronyms defined in glossaries
\newabbreviation{lidar}{LIDAR}{Light Detection and Ranging}
% defining abbreviations like this allows for autocompletion
\newglossaryentry{filo}{
    name={FILO},
    type=\glsxtrabbrvtype,
    description={first in last out},
    first={first in last out (FILO)}
}

% glossary entries
\newglossaryentry{linux}{
    name=Linux,
    description={is a generic term referring to the family of Unix-like computer operating systems that use the Linux kernel},
    plural=Linuces
}

\newglossaryentry{matrix}{
    name={matrix},
    plural={matrices},
    description={rectangular array of quantiities}
}

% symbols
\newglossaryentry{M}{
    type=symbols,
    name={\ensuremath{M}},
    sort=M,
    description={a \gls{matrix}}
}

\newglossaryentry{F}{
    type=symbols,
    name={\ensuremath{F}},
    sort=F,
    description={External Force}
}
% Some abstract text
\abstract{
This dissertation focusses on robotic mapping and exploration of uncertain environments. Computational algorithms are developed to provide important probability and uncertainty properties. These algorithms are designed for real-time implementation for robotic autonomy.

Probabilistic occupancy grid mapping is developed using Bayesian probability. A novel approach to this problem is explored, which uses important physical properties of the environment and stochastic properties of depth sensors. We develop an exact solution to occupancy grid probability that can be achieved in real-time using sensor properties and conventional assumptions of an occupancy grid. The rapid computation allows the algorithm to consider large scans of measurements in 2D and 3D environments. The mapping algorithm is demonstrated with several numerical examples and experiments.

The next topic this dissertation addresses is autonomous exploration, where a robot selects actions to maximize its knowledge about the probabilistic map. We select Shannon's entropy as a metric that represents grid cell uncertainty. Using the earlier contributions on probabilistic occupancy grid mapping, we determine the expected value of entropy change from possible future measurements. This entropy change provides important insights for where a robot should move to maximize its mapping information gain. Dijkstra's search is integral to the algorithm to account for collision-free distances during motion planning. This algorithm is designed for 2D and 3D, where computation time is carefully considered to ensure real-time algorithm performance. Several versions of the exploration algorithm are applied to simulations and experiments.

The final topic of this dissertation relates to multi-vehicle cooperative scenarios. The mapping algorithm is multi-threaded to accept measurements from multiple sources with differing sensor properties. More importantly, the exploration algorithm structure is modified with a bidding-based framework. A series of auctions determines where robots should travel such that the members act together as a team. This solution is further extended to multi-vehicle patrol, where robots begin with an uncertain map, autonomously explore the space, and periodically revisit regions. Autonomous multi-vehicle patrol is accomplished through map degradation, where the probabilistic map becomes more uncertain over time, and the robots must revisit these spaces. These complex algorithms are demonstrated with numerical simulations.

In short, this dissertation proposes novel solutions to probabilistic occupancy grid mapping, autonomous exploration, and patrol in single-vehicle and multi-vehicle scenarios. Real-time implementation is paramount to ensure autonomy during a task. The efficacy of the approaches are shown with several experiments and numerical examples.

}

### 2018-06-09 Dissertation/Thesis/tex/manual.tex
merged evan: mostly tracking changes in .gitignore

% !TEX root = ../thesis-sample.tex

\chapter{Introduction} \label{chap:intro}

%This chapter is stuck among the others as a brief users' manual for this template.  
%The approach to this template is to result in \LaTeX~source code files (\texttt{.tex} files) that are as simple as possible.  
%It also tries to do as much as possible automatically so that the user does not have to spend a lot of effort trying to match the poorly documented ``guidelines'' provided by GWU or SEAS. 
%This is particularly useful for the first few pages, for example the title page, dedication, and abstract page, which are difficult to make in \LaTeX~and are supposed to go in a certain order.
%
%In addition to the description in this chapter, anyone may, of course, also look at the source code for this file, \texttt{thesis-sample.tex}.
%That file contains all of the source for this \texttt{.pdf} in a single file, but it will work just as well with multiple input files combined with the \texttt{input} or \texttt{include} commands.
%Generally, it is a better idea to use \texttt{include} as this will allow for selective compiling of a specific portion of your thesis, i.e. only a single chapter at a time.
%
%The final generic comment about this template is that it has been updated to take advantage of \LaTeX's capabilities to create documents with links in them.  
%Provided you are using a modern PDF viewer to view this document, you may have already noticed this.  
%It creates a list of bookmarks, which can be used to quickly navigate what may be a long document.  
%It also turns references within the text into links.  
%The best examples in this file are the entries in the Table of Contents.
%Although the chapter and section names are shown in black (in accordance with the GWU guidelines), clicking on them does navigate to the start of the chapter, section, etc.

\section{Robotic Mapping}
The way to invoke usage of this template is to put
\begin{verbatim}
\documentclass{thesis-gwu.cls}
\end{verbatim}
at the beginning of your preamble.  
This can also work if the \texttt{thesis-gwu.cls} file is not in the same directory as your \texttt{.tex} file.  
To do so, just give the relative path.
\begin{verbatim}
\documentclass{./tex/thesis-gwu.cls}
\end{verbatim}

Much like a usual article or report in \LaTeX, the user specifies the primary information about the document in the preamble with commands like
\begin{verbatim}
\author{Shankar Kulumani}
\chair{Taeyoung Lee}
\end{verbatim}
At the beginning of the document, the title page will automatically be created and inserted at the beginning of the document.  
If you forget to declare any of the required fields, it will generate a title page with a message such as ``Insert an author!''

However, the template does a lot more in the preamble than just create a title page.  
The preamble (that is, whatever comes before \texttt{begin\{document\}} in the primary \texttt{.tex} file) is also the place for the user to specify a dedication, any acknowledgments, a foreword, \textit{etc.}  
This is done in a manner very similar manner to declaring the author, title, and so on. Suppose that someone wants to have a simple dedication ``To Mom'', the following command is all that is needed.
\begin{verbatim}
\dedication{To Mom}
\end{verbatim}
This will cause the document to have a dedication page with the
corresponding text.  
If the \texttt{dedication} command is not present, there will not be a dedication page.  
All the work of either having or not having a dedication has been compressed into a single command!
Things other than simple text \emph{are} allowed in the dedication, so feel free to put equations or whatever inside there.  
There are a few more commands that can be used to customize the appearance of the dedication page, and also for the other preamble text pages, but that is left to~\cref{ssec:dedication}.

\section{Front Matter}
The \LaTeX term ``frontmatter'' refers to all of the pages that occur before the beginning of the first chapter.  
It is usually made clear to the reader because the pages in the front matter are numbered with lower-case Roman numerals instead of Arabic numerals.

The present template, \texttt{thesis-gwu.cls} attempts to remove as much work associated with the front matter as possible.  
The template inserts all of the front matter pages automatically, so that there is
not even a need to use a command like \verb+\maketitle+.  
The first thing after \verb+\begin{document}+ should be the start of the first
chapter.

\subsection{Identifiers}
The template is not able to read minds, of course, so there needs to be some way of inputting the relevant information.  
This section covers how to specify the author, title, and so on.  
For the most part, this works just like any other \LaTeX~document, but a dissertation has a few more identifiers than most documents (How many books or reports have a
committee?).  
So there are a few extra commands provided by this template, and they work \emph{almost} exactly like the standard commands.

\begin{table}
 \caption{ \label{tab:identifiers}
  List of all identifier commands}
 \centering
 \small
 \begin{tabular}{l @{\hspace{16pt}} l @{\hspace{16pt}} p{6cm}}
  \hline \hline
  \textsc{Item} & \textsc{Usage} & \textsc{Comment} \\
  \hline
  Author      & \verb|\author{...}|
   & Works as in standard \LaTeX \\
  Chair       & \verb|\chair{...}|
   & Name of chair \emph{without} any title or affiliation.  This
     appears only on the abstract page, and only if there is no
     co-chair. \\
  Co-chair    & \verb|\cochair{...}|
   & Names of all co-chairs \emph{without} any titles or
     affiliations.  This appears only on the abstract page.  Note
     that by convention, it is not chair \emph{and} co-chair, but
     just two co-chairs. \\
  Committee   & \verb|\committee{...}|
   & Formatted names of committee members \emph{with} the
     appropriate titles and university names.  This will appear
     only on the title page. \\
  Department  & \verb|\department{...}|
   & Title of department of student \\
  Title       & \verb|\title{...}|
   & Works as in standard \LaTeX \\
  Year        & \verb|\year=2012|
   & Year that dissertation will be \emph{completed} \\
  \hline \hline
 \end{tabular}
\end{table}

A full list of the identifiers is given in Table \cref{tab:identifiers}. 
You're free to adjust/delete these commands as desired.

\subsection{Frontispiece and Copyright}

By default, the template assumes that there should be a copyright page, and the copyright holder is the author.  
To prevent the copyright page from appearing, use the command \verb|\hidecopyright|.  
To assign the copyright to someone other than the author, use the following command.
\begin{verbatim}
\copyrightholder{Someone Else}
\end{verbatim}

\subsection{Text Pages}  \label{ssec:dedication}
The handling of the first few pages after the title page is one of the best features of this template.  
The pages that occur between the copyright page and the abstract page all consist of short pieces of text that are usually a single paragraph.  
The text for each of these pages is set up using a command of the same name.
Generally, only one of the front matter pages is used. 
\begin{verbatim}
\foreword{This is going to be the best dissertation ever.}
\end{verbatim}
Usually the contents of each of these pages will be longer than a single sentence, and thus it should be noted that each of these commands allows most types of \LaTeX~input.  
For example, the following is perfectly acceptable input--at least as far as the template is concerned.
\begin{verbatim}
\foreword{This is going to be the \emph{best}. \\
 \begin{center} Really, really. \end{center}}
\end{verbatim}

As I mentioned before, a given page will appear in the document if and only if the corresponding command is used.  
The order in which the pages appear does not depend on the order the commands are used in the preamble.  
You can also prevent the pages from appearing by using commands like \verb|\hideforeword|.

The style of each page can also be set by the user.  
By default, each page will appear with a bold, italic heading corresponding to the name
of the page.  
However, there are five other formats, which can be controlled using an optional argument.  
For example, the following command creates a dedication page with no heading (\textit{i.e.}~it does not say ``Dedication'' on the page) but with lines above and below the dedication text.
\begin{verbatim}
\dedication[4]{To Mom}
\end{verbatim}
A complete list of the available styles is given in \cref{tab:fronstyle}.  
The style of each page can be set independently, but it is also possible to change which style is used by default.
\begin{verbatim}
\frontpagestyle{6}
\end{verbatim}
This would make all of the commands that were called without optional
inputs to create pages using style 6.

\begin{table}
 \caption{ \label{tab:fronstyle}
  List of styles for frontmatter text pages}
 \centering
 \begin{tabular}{c @{\hspace{16pt}} p{8cm}}
  \hline \hline
  \textsc{Style} & \textsc{Description} \\
  \hline
  1 & Justified text with no header or lines \\
  2 & Justified text with bold italic header and no lines \\
  3 & Justified text, capitalized header, no lines \\
  4 & Justified text with lines and no header \\
  5 & Justified text with bold italic header and lines \\
  6 & Justified text with capitalized header and lines \\
  other & Centered text with no header or lines \\
  \hline \hline
 \end{tabular}
\end{table}

\subsection{Lists of Things}   \label{ssec:lists}
Suffice it to say that this template handles the Table of Contents appropriately, but this section is also meant to address the List of Figures, List of Tables, \textit{etc}.  According to the guidelines, a corresponding list must appear if there is more than one figure, table, map, program, illustration, or appendix.  
The template assumes that the dissertation will contain at least two figures and tables.
If, for example, there is only one figure,
\begin{verbatim}
\hidelistoffigures
\end{verbatim}
must be put in the preamble.
There are similar commands to hide/show all of the front matter pages.
These can be used at will as desired.

\subsection{Glossary of Terms}\label{ssec:glossary}
Here is an example.
\Gls{linux} is a computer operating system, and its completely free and open.

A \gls{matrix}, denoted \gls{M}.
Lots of \glspl{matrix}.
\Glspl{matrix} are a rectangular array of quantities.
No indexing, linking, or formatting: \glsentrytext{matrix}.
The identity \gls{matrix}['s] diagonal consists of ones.

First use: \gls{filo}.
Next use: \gls{filo}.
Full form: \acrfull{filo}.

\subsection{Acronymns and Symbols}\label{ssec:acronymns}
The other type of list that can occur is for abbreviations of various types.  
This is a somewhat convenient feature, particularly there are a lot of acronyms in the dissertation.  
This template utilizes the \texttt{acronym} or \texttt{glossaries} packages, but eventually I would like to migrate to the use of \texttt{glossaries} which is continuing to be supported and deprecate the use of \texttt{acronym}, for now both are supported.

\subsubsection{\texttt{acronymn} usage - try to avoid using this as it's not the best}
In the preamble put a command like the following.
\begin{verbatim}
\abbreviations{
 \acro{CFD}{Computational Fluid Dynamics}
 \acro{LOA}{List of Abbreviations}
 \acro{H2O}[$\mathrm{H_2O}$]{water}}
\end{verbatim}
This will define a bunch of abbreviations that can be used.  
When you want to use one of the acronyms within the text, simply use the \verb|\ac| command to refer to the abbreviation you want.  
This will automatically spell out what the abbreviation stands for on the first use and only print out the abbreviation on subsequent uses.

\subsubsection{\texttt{glossaries} usage - much improved and more powerful}
In addition, you can utilize the much more powerful \texttt{glossaries} package.
You can define symbols, acronymns, or full glossary entries as desired.
Each acronymn is defined using the following:
\begin{verbatim}
\newacronym{crtb}{CRTBP}{Circular Restricted Three Body Problem}
\end{verbatim}
Here we can  use an acronym, such as \gls{crtbp}.
Or display the full name, \gls{crtbp}.

\section{Float environments}
There are many possible float enviornments, and this section will serve as an introduction and demonstration of some of them.
In addition, it offers the ability to ensure that this template actually follows the guidelines.

\subsection{Figures}\label{ssec:figures}

Here is a figure as shown in~\cref{fig:picard}.
Notice how we're using the fancy referencing offered by the \verb+cleveref+ package. 
Instead of using the normal~\verb+\ref+ command we instead use~\verb+\cref+. 
The magic of \LaTeX automatically figures out that the previous reference points to a figure while~\cref{ssec:figures} points to a section.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/picard_yes.jpg}
    \caption[Damage report!]{Glad to have a thesis class\label{fig:picard}}
\end{figure}

\subsection{Tables}\label{ssec:tables}

Here's a table in~\cref{tab:table}

\begin{table}
\begin{center}
    \begin{tabular}{ | l | l | l | p{5cm} |}
    \hline
    Day & Min Temp & Max Temp & Summary \\ \hline
    Monday & 11C & 22C & A clear day with lots of sunshine.  
    However, the strong breeze will bring down the temperatures. \\ \hline
    Tuesday & 9C & 19C & Cloudy with rain, across many northern regions. Clear spells 
    across most of Scotland and Northern Ireland, 
    but rain reaching the far northwest. \\ \hline
    Wednesday & 10C & 21C & Rain will still linger for the morning. 
    Conditions will improve by early afternoon and continue 
    throughout the evening. \\
    \hline
    \end{tabular}
    \caption[Short caption for table]{Long caption for text \label{tab:table}}
    \end{center}
\end{table}

\section{References and Citation}
Finally, we'll add a subfigure to demonstrate it's proper use. 
Many people use the package~\verb+subfigure+ but this is in fact, quite wrong. 
To begin, the~\verb+subfigure+ package has been deprecated, which one can check by going to \url{https://www.ctan.org/pkg/subfigure}{CTAN}.
Instead, everyone should be using~\verb+subcaption+, just as this class file is already doing.
Here, in~\cref{fig:xkcd}, we see two subfigures encapsulated in a larger figure environment.
Luckily, with our fancy referencing we have access to both~\cref{fig:ext,fig:ksp} using the same commands.
The key thing to note from~\cref{fig:ext} is that trustworthiness reaches a maximum for those using~\verb+.tex+.
\begin{figure}[htbp] 
    \centering 
    \begin{subfigure}[htbp]{0.5\textwidth} 
        \includegraphics[width=\textwidth]{figures/file_extensions.png} 
        \caption{File Extensions} \label{fig:ext} 
    \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
    \begin{subfigure}[htbp]{0.5\textwidth} 
        \includegraphics[width=\textwidth]{figures/orbital_mechanics.png} 
        \caption{Kerbal Space Program} \label{fig:ksp} 
    \end{subfigure}
    \caption[XKCD]{Some words of wisdom from Randall Munroe}
    \label{fig:xkcd} 
\end{figure}

\subsection{References}

Lots of famous people tend to write famous papers~\cite{newton1999}. 
Were they famous because or in-spite of their papers?
Regardless, they're famous now and we all should read them.
Certain people are so famous and do such great work that they invent a whole new field of study with a single paper~\cite{kalman1960,shannon1949}

\section{Math}

Here are some nice equations~\cref{prob_def,prob_def_constrained}
\begin{align}
\label{prob_def}
&\min_{s\subset W}\ J(s) = \sum_{i=1}^{l-1} H(s_j, s_{j+1}) \\
&\max_{s\subset W}\ P_{tr}(s) = \prod_{i=1}^{l-1} P_{tr}(s_j, s_{j+1}) \nonumber
\end{align}

Here's another equation.
\begin{align}
\label{prob_def_constrained}
&\min_{s\subset W}\ J(s) = \sum_{i=1}^{l-1} H(s_j, s_{j+1}) \\
&\text{subject to} \ P_{tr}(s)>\epsilon_{tr} \nonumber
\end{align}

### 2018-06-09 Conference/SCITECH19/SCITECH19.tex
merged evan: mostly tracking changes in .gitignore
first part of rS Sf SCITECH esults

@@
 \begin{align}
+\label{eqn:CandidateLocationOptimization}
 x_c^*=\argmax_{x_c}\mathcal{I}_\text{scan}(x_c,R_c^*)\mathcal B(d(x_c)),
@@
 The mapping, exploration, and visualization processes communicate over the Robot Operating System (ROS) framework, and Gazebo 8 serves as the simulator for the physical objects and measurements.
+There are two major programs, referred to as \emph{nodes} in ROS, used during the simulation: mapping and exploration.
+The mapping node takes measurements from a 3D laser scanner and pairs those with the sensor transformation relative to Mars using the ROS Transform Message Filters package.
+Then, it builds a 3D probabilistic map using \refeqn{RayISMAnswer}-\refeqn{Unnormalized}.
+The mapping node sends updates from the probabilistic map to the exploration node, which decomposes the map into $m_\text{reduced}$ according to \refeqn{Proj3DMapComb}, and builds a cost map based on Dijkstra's search to determine collision-free trajectory costs for all pose candidates.
+Finally, it applies \refeqn{RcStar}, \refeqn{BumpFunRecedingHorizon}, and \refeqn{CandidateLocationOptimization} to update the desired robot trajectory on a receding-horizon.
+An overview of the process is illustrated in Fig.
+\ref{fig:NodesDiagram}.
+
+	\begin{figure}
+		\centerline{
+			\includegraphics[width=0.6\columnwidth]{node_diagram_3D_mars.png}
+		}
+		\caption{The diagram of the 3D probabilistic mapping and autonomous exploration nodes shows the primary processes and communication lines.}
+		\label{fig:NodesDiagram}
+	\end{figure}
 

### 2018-06-09 Conference/SCITECH19/SCITECH19.tex
mars pictures, and some text added to results

@@
 
+A 3D environment of Mars is simulated in Gazebo.
+A height map~\cite{THATwebsite} is used to generate a contoured surface, and the corresponding picture of Mars is draped over this contour, shown in Fig.
+\ref{fig:MarsGazebo}.
+A laser scanner was also simulated in Gazebo.
+In the horizontal direction, the sensor has limits $\Psi=$ with a total of $640$ measurement rays inside, the sensor is fixed at angle $\tilde{\psi}=45$, and $n_\psi=16$ sample measurements are used.
+In the vertical direction, the sensor has limits $\Theta=$ with a total of $480$ measurement rays inside, the sensor is fixed at angle $\tilde{\theta}=15$, and $n_\psi=7$ sample measurements are used.
+These ray samples are taken from candidate locations separated $2$m apart.
+
+% trim={<left> <lower> <right> <upper>}
+
+\begin{figure}[!t]
+	\centering
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[height=0.9\textwidth]{mars_heightmap.png}
+        		\caption{Height Map}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[height=0.9\textwidth]{mars_colormap.png}
+        		\caption{Color Map}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering{
+          	\includegraphics[height=0.9\textwidth]{mars_gazebo_sqr.png}}
+        		\caption{Gazebo Environment}
+    	\end{subfigure}
+%	\begin{subfigure}[t]{0.3\columnwidth}
+%           	\centering
+%          	\includegraphics[trim = {23cm 4cm 17cm 6cm}, clip, height=1.0\textwidth]{mars_theta.png}
+%        		\caption{$\theta$ Rotation}
+%    	\end{subfigure}
+%\includegraphics[width=2.5in]{myfigure}
+% where an .eps filename suffix will be assumed under latex, 
+% and a .pdf suffix will be assumed for pdflatex; or what has been declared
+% via \DeclareGraphicsExtensions.
+\caption{The height map (a) and color map (b) are used in Gazebo (c) to simulate a Mars environment.}
+\label{fig:MarsGazebo}
+\end{figure}
+
+The map parameters are also important to the success of the exploration.
+The full map has dimensions $100$m$\times$$100$m$10$m in the Mars-fixed frame, with cell edge length $\alpha=0.075$m.
+The reduced map for collision-avoidance and motion planning has cells $k=3$ times the size ($0.225$m), so $k^3=27$ cells are considered in \refeqn{Proj3DMapComb}.
+The pose candidates are spaced $2$m apart in each dimension, and the bump functions use $f_\text{max}=$, $f_\text{min}=$, and $\beta=$ to account for travel costs with \refeqn{BumpFunRecedingHorizon}.
+The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.5$m/sec and the expected computation time varies from $0.25$sec to $1.0$sec.
+
+The simulation is run twice.
+Case 1 is as described in this paper.
+Case 2 is identical, except the reduced map $m_\text{reduced}$ is used when computing \refeqn{Iscan}.
+Case 2 is largely inspired by the promising results of~\cite{KauTakAiLee18}, where a projected map based on the same criteria as \refeqn{Proj3DMapComb} provided useful effective information during level flight.
+The resulting occupancy grid maps and complete map entropy are shown in Figs.
+\refeqn{mars3Dogm} and \ref{fig:mars3Dentropy}, respectively.
+
+In both cases, the ...
+
+However, the choice of occupancy grid map introduced an interesting tradeoff.
+
+In short,
+
 \begin{itemize}
 	\item Mars Environment: simulation environment generation, depth sensor facing slightly downward
-	\item Software structure: ROS nodes, full map/reduced map for entropy options
+	\item Full map/reduced map for entropy options
 	\item Numerical Results: reduced map used for entropy, full map used for entropy, Figs: RGB/heightmap/Gazebo, 3D colored grid cells with candidates (both full and reduced maps), entropy over time (full and reduced on same plot)
 \end{itemize}
+
 % software structure

### 2018-06-09 Conference/SCITECH19/SCITECH19.tex
updated paper with true numbers

@@
 
-\usepackage{amssymb,url,times}%,subfigure}% amsthm is the one!
+\usepackage{amssymb,url,times,gensymb}%,subfigure}% amsthm is the one!
 \usepackage{caption,subcaption,hyperref}
@@
 \end{align}
-where the mounting yaw $\tilde{\psi}$ and pitch $\tilde{\theta}$ of the sensor is a fixed rotation in the horizontal and vertical directions, respectively in that order, such that $-\pi<\tilde{\psi}\leq\pi$ and $-\frac{\pi}{2}+\frac\Theta2\leq\tilde{\theta}\leq\frac{\pi}{2}-\frac\Theta2$.
+where the mounting yaw $\tilde{\psi}$ and pitch $\tilde{\theta}$ of the sensor is a fixed rotation in the horizontal and vertical directions, respectively in that order, such that $-\pi<\tilde{\psi}\leq\pi$ and $-\frac{\pi}{2}\leq\tilde{\theta}\leq\frac{\pi}{2}$.
 A positive $\tilde{\theta}$ corresponds to rotating the depth sensor from forward to downward toward the surface of Mars.
@@
 A laser scanner was also simulated in Gazebo.
-In the horizontal direction, the sensor has limits $\Psi=$ with a total of $640$ measurement rays inside, the sensor is fixed at angle $\tilde{\psi}=45$, and $n_\psi=16$ sample measurements are used.
-In the vertical direction, the sensor has limits $\Theta=$ with a total of $480$ measurement rays inside, the sensor is fixed at angle $\tilde{\theta}=15$, and $n_\psi=7$ sample measurements are used.
-These ray samples are taken from candidate locations separated $2$m apart.
+In the horizontal direction, the sensor has limits $\Psi=120\degree$ with a total of $1000$ measurement rays inside, the sensor is fixed at angle $\tilde{\psi}=45\degree$, and $n_\psi=16$ sample measurements are used.
+In the vertical direction, the sensor also has limits $\Theta=120\degree$ with a total of $1000$ measurement rays inside, but the sensor is fixed at angle $\tilde{\theta}=30\degree$, and $n_\psi=7$ sample measurements are used.
+These ray samples are taken from candidate locations separated $1$m apart in each dimension.
 
@@
 The map parameters are also important to the success of the exploration.
-The full map has dimensions $100$m$\times$$100$m$10$m in the Mars-fixed frame, with cell edge length $\alpha=0.075$m.
+The full map has dimensions $20$m$\times20$m$\times5$m in the Mars-fixed frame, with cell edge length $\alpha=0.075$m.
 The reduced map for collision-avoidance and motion planning has cells $k=3$ times the size ($0.225$m), so $k^3=27$ cells are considered in \refeqn{Proj3DMapComb}.
-The pose candidates are spaced $2$m apart in each dimension, and the bump functions use $f_\text{max}=$, $f_\text{min}=$, and $\beta=$ to account for travel costs with \refeqn{BumpFunRecedingHorizon}.
-The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.5$m/sec and the expected computation time varies from $0.25$sec to $1.0$sec.
+The bump functions use $f_\text{max}=1.0$, $f_\text{far}=0.1$, and $\beta=0.1$ to account for travel costs with \refeqn{BumpFunRecedingHorizon}.
+The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.5$m/sec and the computation time varies from $0.6$sec to $1.0$sec.
 
@@
 The resulting occupancy grid maps and complete map entropy are shown in Figs.
-\refeqn{mars3Dogm} and \ref{fig:mars3Dentropy}, respectively.
+\ref{mars3Dogm} and \ref{fig:mars3Dentropy}, respectively.
 

### 2018-06-09 Conference/SCITECH19/SCITECH19.tex
added reference to Mars data

@@
 A 3D environment of Mars is simulated in Gazebo.
-A height map~\cite{THATwebsite} is used to generate a contoured surface, and the corresponding picture of Mars is draped over this contour, shown in Fig.
+A height map~\cite{mapaplanet} is used to generate a contoured surface, and the corresponding picture of Mars is draped over this contour, shown in Fig.
 \ref{fig:MarsGazebo}.

### 2018-06-09 Conference/SCITECH19/SCITECH19.tex
added resulting figs, still 3 paragraphs for results and 1-2 for conclusions

@@
 First we define the map, measurements, and poses as follows.
-Let 3D map $m$ be composed of $n_m$ evenly-spaced cubic grid cells with edge length $\alpha$.
+Let 3D map $ m$ be composed of $n_m$ evenly-spaced cubic grid cells with edge length $\alpha$.
 For index $i\in\braces{1,2,\ldots,n_m}$, let $\mathbf{m}_i$ and $\bar{\mathbf{m}}_i$ be the events that the $i$-th grid cell are occupied and free, respectively, where each cell occupancy is assumed static, binary, and independent of other cells.
@@
 Let $k\geq1$ be an integer such that $k\alpha\geq\rho$ and $(k-1)\alpha<\rho$, i.e., $k$ is the minimum number of grid cells in each dimension capable of fully-enclosing the robot inside this cube.
-Then, we decompose the complete probabilistic 3D into map $m_\text{reduced}$, which is composed of larger cubic cells that encompass $k^3$ cells from map $m$.
-Let $\mathbf{m}_{\text{reduced},k}$ denote the $k$-th cell of $m_\text{reduced}$ being occupied, and let $a_k\subset m$ be those $k^3$ cells from $m$ composing this larger cell.
+Then, we decompose the complete probabilistic 3D into map $ m_\text{reduced}$, which is composed of larger cubic cells that encompass $k^3$ cells from map $ m$.
+Let $\mathbf{m}_{\text{reduced},k}$ denote the $k$-th cell of $ m_\text{reduced}$ being occupied, and let $a_k\subset m$ be those $k^3$ cells from $ m$ composing this larger cell.
 Then, we apply an effective rule for collision-avoidance from~\cite{KauTakAiLee18} to obtain the probability of $\mathbf{m}_{\text{reduced},k}$,
@@
 Next, we show how $\mathbf{m}_{\text{reduced},k}$ is used in Dijkstra's search~\cite{Dij59}.
-Defining $0<P_\text{coll}<1$ as the acceptable probability of collision, every grid cell of $m_\text{reduced}$ is considered a safe and unvisited node if its occupancy probability, and the occupancy probabilities of its neighbors (sharing a face, edge, or corner), are below $P_\text{coll}$.
+Defining $0<P_\text{coll}<1$ as the acceptable probability of collision, every grid cell of $ m_\text{reduced}$ is considered a safe and unvisited node if its occupancy probability, and the occupancy probabilities of its neighbors (sharing a face, edge, or corner), are below $P_\text{coll}$.
 A cost map is built from the starting robot location by neighboring nodes, where the cost to travel to another node is based on Euclidian distance: a cell sharing a face is $\alpha$ away, a cell sharing an edge is $\sqrt2\alpha$ away, and a cell sharing a corner is $\sqrt3\alpha$ away.
@@
 Then, it builds a 3D probabilistic map using \refeqn{RayISMAnswer}-\refeqn{Unnormalized}.
-The mapping node sends updates from the probabilistic map to the exploration node, which decomposes the map into $m_\text{reduced}$ according to \refeqn{Proj3DMapComb}, and builds a cost map based on Dijkstra's search to determine collision-free trajectory costs for all pose candidates.
+The mapping node sends updates from the probabilistic map to the exploration node, which decomposes the map into $ m_\text{reduced}$ according to \refeqn{Proj3DMapComb}, and builds a cost map based on Dijkstra's search to determine collision-free trajectory costs for all pose candidates.
 Finally, it applies \refeqn{RcStar}, \refeqn{BumpFunRecedingHorizon}, and \refeqn{CandidateLocationOptimization} to update the desired robot trajectory on a receding-horizon.
@@
 In the vertical direction, the sensor also has limits $\Theta=120\degree$ with a total of $1000$ measurement rays inside, but the sensor is fixed at angle $\tilde{\theta}=30\degree$, and $n_\psi=7$ sample measurements are used.
-These ray samples are taken from candidate locations separated $1$m apart in each dimension.
+These ray samples are taken from candidate locations separated $1$ m apart in each dimension.
 
@@
     	\end{subfigure}
-%	\begin{subfigure}[t]{0.3\columnwidth}
-%           	\centering
-%          	\includegraphics[trim = {23cm 4cm 17cm 6cm}, clip, height=1.0\textwidth]{mars_theta.png}
-%        		\caption{$\theta$ Rotation}
-%    	\end{subfigure}
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
 \caption{The height map (a) and color map (b) are used in Gazebo (c) to simulate a Mars environment.}
@@
 The map parameters are also important to the success of the exploration.
-The full map has dimensions $20$m$\times20$m$\times5$m in the Mars-fixed frame, with cell edge length $\alpha=0.075$m.
-The reduced map for collision-avoidance and motion planning has cells $k=3$ times the size ($0.225$m), so $k^3=27$ cells are considered in \refeqn{Proj3DMapComb}.
+The full map has dimensions $20$ m$\times20$ m$\times5$ m in the Mars-fixed frame, with cell edge length $\alpha=0.075$ m.
+The reduced map for collision-avoidance and motion planning has cells $k=3$ times the size ($0.225$ m), so $k^3=27$ cells are considered in \refeqn{Proj3DMapComb}.
 The bump functions use $f_\text{max}=1.0$, $f_\text{far}=0.1$, and $\beta=0.1$ to account for travel costs with \refeqn{BumpFunRecedingHorizon}.
-The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.5$m/sec and the computation time varies from $0.6$sec to $1.0$sec.
+The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.5$ m/sec and the computation time varies from $0.6$ sec to $1.0$ sec.
 
@@
 Case 1 is as described in this paper.
-Case 2 is identical, except the reduced map $m_\text{reduced}$ is used when computing \refeqn{Iscan}.
+Case 2 is identical, except the reduced map $ m_\text{reduced}$ is used when computing \refeqn{Iscan}.
 Case 2 is largely inspired by the promising results of~\cite{KauTakAiLee18}, where a projected map based on the same criteria as \refeqn{Proj3DMapComb} provided useful effective information during level flight.
 The resulting occupancy grid maps and complete map entropy are shown in Figs.
-\ref{mars3Dogm} and \ref{fig:mars3Dentropy}, respectively.
+\ref{fig:mars3Dogm} and \ref{fig:mars3Dentropy}, respectively.
+
+\begin{figure}[!t]
+	\centering
+	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsFullMap1min.jpg}
+        		\caption{Case 1: $1$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap1min.jpg}
+        		\caption{Case 2: $1$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+	\centering
+	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsFullMap2min.jpg}
+        		\caption{Case 1: $2$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap2min.jpg}
+        		\caption{Case 2: $2$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+	\centering
+	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsFullMap5min.jpg}
+        		\caption{Case 1: $5$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap5min.jpg}
+        		\caption{Case 2: $5$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+	\centering
+	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsFullMap10min.jpg}
+        		\caption{Case 1: $10$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap10min.jpg}
+        		\caption{Case 2: $10$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+	\centering
+	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsFullMap15min.jpg}
+        		\caption{Case 1: $15$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsFullMap20min.jpg}
+        		\caption{Case 1: $20$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+\caption{The robot (red disk with arrow indicating laser direction) moves toward candidate poses (red arrows, more opaque for greater reward) while generating a 3D probabilistic occupancy grid map (cubes: greater opacity for greater occupancy probability) of the surface of Mars.}
+\label{fig:mars3Dogm}
+\end{figure}
+
+	\begin{figure}
+		\centerline{
+			\includegraphics[width=0.6\columnwidth]{mars_H_comparison_flat.pdf}
+		}
+		\caption{The complete map entropy for Case 1 (red solid line) decreases at roughly the same rate as the reduced map entropy for Case 2 (dashed blue line), but the complete map entropy highlights a few locations where entropy can decrease further, shown by lower entropy over the subsequent $10$ min.}
+		\label{fig:mars3Dentropy}
+	\end{figure}
+
+NEXT: ADD 3 PARAGRAPHS
 
@@
 In short,
-
-\begin{itemize}
-	\item Mars Environment: simulation environment generation, depth sensor facing slightly downward
-	\item Full map/reduced map for entropy options
-	\item Numerical Results: reduced map used for entropy, full map used for entropy, Figs: RGB/heightmap/Gazebo, 3D colored grid cells with candidates (both full and reduced maps), entropy over time (full and reduced on same plot)
-\end{itemize}
 

### 2018-06-10 Conference/SCITECH19/SCITECH19.tex
completed paper text, needs editing

@@
 Then, it builds a 3D probabilistic map using \refeqn{RayISMAnswer}-\refeqn{Unnormalized}.
-The mapping node sends updates from the probabilistic map to the exploration node, which decomposes the map into $ m_\text{reduced}$ according to \refeqn{Proj3DMapComb}, and builds a cost map based on Dijkstra's search to determine collision-free trajectory costs for all pose candidates.
+The mapping node sends updates from the probabilistic map to the exploration node, which decomposes the full map $m$ into $ m_\text{reduced}$ according to \refeqn{Proj3DMapComb}, and builds a cost map based on Dijkstra's search to determine collision-free trajectory costs for all pose candidates.
 Finally, it applies \refeqn{RcStar}, \refeqn{BumpFunRecedingHorizon}, and \refeqn{CandidateLocationOptimization} to update the desired robot trajectory on a receding-horizon.
@@
 
-NEXT: ADD 3 PARAGRAPHS
+In both cases, the robot built the 3D probabilistic map of the $2000$ cubic meter space composed of $4.74\times10^6$ grid cells.
+The maps were mostly complete within $10$ minutes.
+The $\tilde{\theta}=30\degree$ downward viewing angle proved effective, as candidate attitudes tended to direct the onboard sensor toward uncertain regions on the surface of Mars.
+The proposed approach provided collision-free mapping and autonomous exploration in real-time.
 
-In both cases, the ...
+However, the choice of occupancy grid map used for entropy predictions introduced an interesting tradeoff.
+In Case 1, when the full probabilistic map $m$ is used in \refeqn{Iscan}, regions where the grid cells are mostly-known are frequently reconsidered until the space is well-known.
+Conversely in Case 2, when $m_\text{reduced}$ is used instead, the robot will leave a space that is missing a few grid cells to visit new terrain.
+This is because cells of $m_\text{reduced}$ with large occupancy probabilities are based on \refeqn{Proj3DMapComb}, which can allow some grid cells of $m$ enclosed within a cell of $m_\text{reduced}$ to be uncertain.
+Ironically, this false assumption can actually lead to greater information gains at certain times if the vehicle leaves this mostly-known space.
+Here, the expected information gains after about $10$ minutes were insignificant, so exploration was complete.
+However with Case $1$, which used $m$ for entropy, the exploration continued for an additional $10$ minutes, overall decreasing total map entropy further, and getting a more complete occupancy grid of the surface of Mars.
 
-However, the choice of occupancy grid map introduced an interesting tradeoff.
-
-In short,
-
-% software structure
-% Numerical Result
-  % reduced map used for entropy
-  % full map used for entropy
+In short, the proposed 3D probabilistic occupancy grid mapping and autonomous exploration were simulated successfully in real-time over the surface of Mars.
+Using $m$ for entropy predictions produced a more-complete map, but using $m_\text{reduced}$ was sometimes beneficial for exploring new terrain faster, while forgoing map completeness.
 
 \section{Conclusions}
+
+In this paper, we proposed an approach to autonomously explore the complex and uncertain terrain of Mars.
+The exploration strategy is based on flying vehicles near the surface of Mars, where their exploration policy is based on expected information gain from Shannon's entropy of a probabilistic map, and travel costs from Dijkstra's algorithm.
+The autonomous exploration scheme was applied in real-time, following a receding-horizon framework, so the map could be generated as quickly as possible.
+Numerical simulations demonstrated the efficacy of the approach, and two possible versions of the 3D exploration, depending on the accuracy requirements of the probabilistic map.
 

### 2018-06-10 Conference/SCITECH19/SCITECH19.tex
edit done for results and conclusion

--- 2018-06-09 Conference/SCITECH19/SCITECH19.tex
@@
 
-\author{Evan Kaufman\footnote{Graduate Research Assistant} and Taeyoung Lee\footnote{Associate Professor, AIAA Member Grade (if any)}}
+\author{Evan Kaufman\footnote{Graduate Research Assistant} and Taeyoung Lee\footnote{Associate Professor, AIAA Member Grade (if any)}}% TODO: update if AIAA member
 \affil{Department of Mechanical and Aerospace Engineering, The George Washington University, Washington, DC, 20052}
@@
 
-The mapping, exploration, and visualization processes communicate over the Robot Operating System (ROS) framework, and Gazebo 8 serves as the simulator for the physical objects and measurements.
+The mapping, exploration, and visualization processes communicate over the Robot Operating System (ROS) framework (Kinetic distribution), and Gazebo 8 serves as the simulator for the physical objects and measurements.
 There are two major programs, referred to as \emph{nodes} in ROS, used during the simulation: mapping and exploration.
@@
 \ref{fig:MarsGazebo}.
-A laser scanner was also simulated in Gazebo.
+A 3D laser scanner is also simulated in Gazebo.
 In the horizontal direction, the sensor has limits $\Psi=120\degree$ with a total of $1000$ measurement rays inside, the sensor is fixed at angle $\tilde{\psi}=45\degree$, and $n_\psi=16$ sample measurements are used.
 In the vertical direction, the sensor also has limits $\Theta=120\degree$ with a total of $1000$ measurement rays inside, but the sensor is fixed at angle $\tilde{\theta}=30\degree$, and $n_\psi=7$ sample measurements are used.
-These ray samples are taken from candidate locations separated $1$ m apart in each dimension.
+These ray samples are taken from each candidate location, which are separated $1$ m apart in each of the $3$ dimensions.
 
@@
 The map parameters are also important to the success of the exploration.
-The full map has dimensions $20$ m$\times20$ m$\times5$ m in the Mars-fixed frame, with cell edge length $\alpha=0.075$ m.
+The full map has dimensions $20$ m $\times$ $20$ m $\times$ $5$ m in the Mars-fixed frame, with cell edge length $\alpha=0.075$ m.
 The reduced map for collision-avoidance and motion planning has cells $k=3$ times the size ($0.225$ m), so $k^3=27$ cells are considered in \refeqn{Proj3DMapComb}.
 The bump functions use $f_\text{max}=1.0$, $f_\text{far}=0.1$, and $\beta=0.1$ to account for travel costs with \refeqn{BumpFunRecedingHorizon}.
-The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.5$ m/sec and the computation time varies from $0.6$ sec to $1.0$ sec.
+The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.25$ m/sec and the computation time varies from $0.6$ sec to $1.0$ sec.
 
-The simulation is run twice.
-Case 1 is as described in this paper.
-Case 2 is identical, except the reduced map $ m_\text{reduced}$ is used when computing \refeqn{Iscan}.
-Case 2 is largely inspired by the promising results of~\cite{KauTakAiLee18}, where a projected map based on the same criteria as \refeqn{Proj3DMapComb} provided useful effective information during level flight.
+The simulation was run twice.
+Case 1 was as described in this paper.
+Case 2 was identical to Case 1, except the reduced map $ m_\text{reduced}$ is used when computing \refeqn{Iscan}.
+Case 2 is largely inspired by the promising results of~\cite{KauTakAiLee18}, where a projected map based on the same criteria as \refeqn{Proj3DMapComb} proved effective for level flight.
 The resulting occupancy grid maps and complete map entropy are shown in Figs.
@@
 
-NEXT: ADD 3 PARAGRAPHS
+In both cases, the robot built the 3D probabilistic map of the $2000$ cubic meter space composed of $4.74\times10^6$ grid cells.
+The maps were mostly complete within $10$ minutes.
+The $\tilde{\theta}=30\degree$ downward viewing angle captured the ground nicely, as candidate attitudes tended to direct the onboard sensor toward uncertain regions on the surface of Mars.
+The proposed approach provided collision-free mapping and autonomous exploration in real-time.
 
-In both cases, the ...
+However, the choice of occupancy grid map used for entropy predictions introduced an interesting tradeoff.
+In Case 1, when the full probabilistic map $m$ was used in \refeqn{Iscan}, regions where the grid cells were partially-known were frequently reconsidered until the space was well-known.
+Conversely in Case 2, when $m_\text{reduced}$ was used instead, the robot repeatedly left spaces that were missing a few grid cells to visit new terrain.
+This is because $m_\text{reduced}$ had some cells with large occupancy probabilities based on \refeqn{Proj3DMapComb}, which allowed some grid cells of $m$ enclosed within a cell of $m_\text{reduced}$ to be uncertain.
+The exploration policy of Case 2 incorrectly assumed these regions were well-known.
+Ironically, this false assumption sometimes actually led to greater information gains when the vehicle moved on to unvisited terrain.
+Here, the exploration policy of Case 2 suspended operations once expected information gains were insignificant, which happened after about $10$ minutes.
+However with Case $1$, which used $m$ for computing expected entropy, the exploration continued for an additional $10$ minutes, overall decreasing total map entropy further, generating a more complete occupancy grid of the surface of Mars.
 
-However, the choice of occupancy grid map introduced an interesting tradeoff.
-
-In short,
-
-% software structure
-% Numerical Result
-  % reduced map used for entropy
-  % full map used for entropy
+In short, the proposed 3D probabilistic occupancy grid mapping and autonomous exploration were simulated successfully in real-time over the surface of Mars using ROS and Gazebo.
+Choosing $m$ for entropy predictions produced a more-complete map, but choosing $m_\text{reduced}$ was sometimes beneficial for exploring new terrain faster, while forgoing map completeness.
 
 \section{Conclusions}
+
+In this paper, we proposed an approach to autonomously explore the complex and uncertain terrain of Mars.
+The exploration strategy is based on flying vehicles near the surface of Mars, where their exploration policy is based on expected information gain from Shannon's entropy of a probabilistic map, and travel costs from Dijkstra's algorithm.
+The autonomous exploration scheme was applied in real-time, following a receding-horizon framework, so the map could be generated as quickly as possible.
+Numerical simulations demonstrated the efficacy of the approach, showing two possible versions of the 3D exploration with differing behavior in probabilistic map accuracy and exploration speed.
 

### 2018-06-10 Conference/SCITECH19/SCITECH19.tex
tylee edit

@@
 
-\title{Autonomous Aerial for Topological Mapping of Mars Environments}
+\newcommand{\todo}[1]{\vspace{5 mm}\par \noindent
+\marginpar{\textsc{ToDo}} \framebox{\begin{minipage}[c]{0.95
+\textwidth} \tt #1 \end{minipage}}\vspace{5 mm}\par}
+%\newcommand{\todo}[1]{}
 
-\author{Evan Kaufman\footnote{Graduate Research Assistant} and Taeyoung Lee\footnote{Associate Professor, AIAA Member Grade (if any)}}% TODO: update if AIAA member
-\affil{Department of Mechanical and Aerospace Engineering, The George Washington University, Washington, DC, 20052}
+\title{Autonomous Aerial Exploration for Topological Mapping of Mars Environments}
+
+\author{Evan Kaufman\footnote{Graduate Research Assistant, Department of Mechanical and Aerospace Engineering, The George Washington University, 800 22nd St NW, Washington, DC, 20052} and Taeyoung Lee\footnote{Associate Professor, Department of Mechanical and Aerospace Engineering, The George Washington University, 800 22nd St NW, Washington, DC, 20052}}% TODO: update if AIAA member
+\affil{The George Washington University, Washington, DC, 20052}
 % TODO: funding source
@@
 \begin{abstract}
-This paper proposes a 3D autonomous exploration approach designed for a flying vehicle while mapping the complex environment of the surface of Mars.
-Shannon's entropy serves as a measure of map uncertainty, where the robot goal is to minimize entropy, thereby maximizing map information.
-We propose a novel technique to predict future map entropy from depth sensors capable of scanning 3D spaces, and formulate autonomous exploration as an optimization problem based on map information gain and collision-free travel costs.
+This paper proposes a 3D autonomous exploration scheme designed for a flying vehicle to build a topological mapping the complex environment of the surface of Mars. 
+The environment is modeled as a uniform grid, and the measurements from depth scans are utilized to determine the probability of occupancy at every cell within the scanned area. 
+Shannon's entropy serves as a measure of map uncertainty, where the motion of the aerial vehicle is guided to minimize entropy, thereby maximizing map information.
+We propose a novel technique to predict future map entropy from depth sensors capable of scanning 3D spaces, and formulate autonomous exploration as an optimization problem based on map information gain subject to inequality constraints to void collision.
 The proposed approach is demonstrated with numerical simulations using contour data from the surface of Mars.
@@
 The predicted negative entropy changes from measurements passing through 3D space are maximized while also considering travel costs.
-The robot then uses Dijkstra's search in 3D for collision-free motion planning, and the complete process follows a receding-horizon framework for consistently using updated information.
+The aerial vehicle then uses Dijkstra's search in 3D for collision-free motion planning, and the complete process follows a receding-horizon framework for consistently using updated information.
 The proposed optimal approach is designed specifically for real-time exploration of the surface of Mars, where the flying robot mostly maps occupied space below the vehicle, but is also capable of looking ahead to capture steep and quickly-changing terrain.
@@
 		\centerline{
-			\includegraphics[width=0.6\columnwidth]{node_diagram_3D_mars.png}
+			\includegraphics[width=0.8\columnwidth]{node_diagram_3D_mars.png}
 		}
@@
 
+\todo{I wish we have a little more pages. Let's split Figure 5 into two, one for Case 1 and another for Case 2, where each case has more time instances. (Also, check the caption for (j)).
+We can have another zoomed-in figure hat shows the aerial vehicle with the simulated measurements of the point clouds and the candidate poses more clearly}
+
 \section{Conclusions}
@@
 
+\section*{Acknowledgments}
+This research has been supported in parts by the 2018 NASA Innovative Advanced Concepts (NIAC) award, and NSF under the grants CMMI-1243000, CMMI-1335008, CNS-1337722, and CMMI-1760928.
+
 \bibliography{../../BibSources}% master source for all publications

### 2018-06-10 Conference/SCITECH19/SCITECH19.tex
merged TLee edits, updated gitignore

@@
 \begin{abstract}
-This paper proposes a 3D autonomous exploration scheme designed for a flying vehicle to build a topological mapping the complex environment of the surface of Mars. 
+This paper proposes a 3D autonomous exploration scheme designed for a flying vehicle to build a topological map representing the complex environment of the surface of Mars. 
 The environment is modeled as a uniform grid, and the measurements from depth scans are utilized to determine the probability of occupancy at every cell within the scanned area. 
 Shannon's entropy serves as a measure of map uncertainty, where the motion of the aerial vehicle is guided to minimize entropy, thereby maximizing map information.
-We propose a novel technique to predict future map entropy from depth sensors capable of scanning 3D spaces, and formulate autonomous exploration as an optimization problem based on map information gain subject to inequality constraints to void collision.
+We propose a novel technique to predict future map entropy from depth sensors capable of scanning 3D spaces, and formulate autonomous exploration as an optimization problem based on map information gain subject to inequality constraints to avoid collision.
 The proposed approach is demonstrated with numerical simulations using contour data from the surface of Mars.
 \end{abstract}
-
-%\section{Nomenclature}
-%
-%{\renewcommand\arraystretch{1.0}
-%\noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
-%$A$  & amplitude of oscillation \\
-%$a$ &    cylinder diameter \\
-%$C_p$& pressure coefficient \\
-%$Cx$ & force coefficient in the \textit{x} direction \\
-%$Cy$ & force coefficient in the \textit{y} direction \\
-%c   & chord \\
-%d$t$ & time step \\
-%$Fx$ & $X$ component of the resultant pressure force acting on the vehicle \\
-%$Fy$ & $Y$ component of the resultant pressure force acting on the vehicle \\
-%$f, g$   & generic functions \\
-%$h$  & height \\
-%$i$  & time index during navigation \\
-%$j$  & waypoint index \\
-%$K$  & trailing-edge (TE) nondimensional angular deflection rate
-%\end{longtable*}}
 
@@
 \section*{Acknowledgments}
-This research has been supported in parts by the 2018 NASA Innovative Advanced Concepts (NIAC) award, and NSF under the grants CMMI-1243000, CMMI-1335008, CNS-1337722, and CMMI-1760928.
+This research has been supported by the 2018 NASA Innovative Advanced Concepts (NIAC) award and NSF under the grants CMMI-1243000, CMMI-1335008, CNS-1337722, and CMMI-1760928.
 

### 2018-06-11 Conference/SCITECH19/SCITECH19.tex
Fig. 5 split in 2 for each case

@@
 Case 2 is largely inspired by the promising results of~\cite{KauTakAiLee18}, where a projected map based on the same criteria as \refeqn{Proj3DMapComb} proved effective for level flight.
-The resulting occupancy grid maps and complete map entropy are shown in Figs.
-\ref{fig:mars3Dogm} and \ref{fig:mars3Dentropy}, respectively.
+The resulting occupancy grid maps for Case 1 and 2 are shown in Figs.
+\ref{fig:mars3DogmCase1} and \ref{fig:mars3DogmCase2}, respectively, and complete map entropy is illustrated in Fig.
+\ref{fig:mars3Dentropy}.
+
+%\begin{figure}[!t]
+%	\centering
+%	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[height=0.5\textwidth]{MarsFullMap1min.jpg}
+%        		\caption{Case 1: $1$ min}
+%		\vspace*{0.025\textwidth}
+%    	\end{subfigure}
+%    	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap1min.jpg}
+%        		\caption{Case 2: $1$ min}
+%		\vspace*{0.025\textwidth}
+%    	\end{subfigure}
+%	\centering
+%	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[height=0.5\textwidth]{MarsFullMap2min.jpg}
+%        		\caption{Case 1: $2$ min}
+%		\vspace*{0.025\textwidth}
+%    	\end{subfigure}
+%    	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap2min.jpg}
+%        		\caption{Case 2: $2$ min}
+%		\vspace*{0.025\textwidth}
+%    	\end{subfigure}
+%	\centering
+%	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[height=0.5\textwidth]{MarsFullMap5min.jpg}
+%        		\caption{Case 1: $5$ min}
+%		\vspace*{0.025\textwidth}
+%    	\end{subfigure}
+%    	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap5min.jpg}
+%        		\caption{Case 2: $5$ min}
+%		\vspace*{0.025\textwidth}
+%    	\end{subfigure}
+%	\centering
+%	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[height=0.5\textwidth]{MarsFullMap10min.jpg}
+%        		\caption{Case 1: $10$ min}
+%		\vspace*{0.025\textwidth}
+%    	\end{subfigure}
+%    	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap10min.jpg}
+%        		\caption{Case 2: $10$ min}
+%		\vspace*{0.025\textwidth}
+%    	\end{subfigure}
+%	\centering
+%	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[height=0.5\textwidth]{MarsFullMap15min.jpg}
+%        		\caption{Case 1: $15$ min}
+%		\vspace*{0.025\textwidth}
+%    	\end{subfigure}
+%    	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[height=0.5\textwidth]{MarsFullMap20min.jpg}
+%        		\caption{Case 1: $20$ min}
+%		\vspace*{0.025\textwidth}
+%    	\end{subfigure}
+%\caption{The robot (red disk with arrow indicating laser direction) moves toward candidate poses (red arrows, more opaque for greater reward) while generating a 3D probabilistic occupancy grid map (cubes: greater opacity for greater occupancy probability) of the surface of Mars.}
+%\label{fig:mars3Dogm}
+%\end{figure}
 
@@
 	\centering
-	\begin{subfigure}[t]{0.4\columnwidth}
+	\begin{subfigure}[t]{0.49\columnwidth}
            	\centering
           	\includegraphics[height=0.5\textwidth]{MarsFullMap1min.jpg}
-        		\caption{Case 1: $1$ min}
+        		\caption{$1$ min}
 		\vspace*{0.025\textwidth}
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.4\columnwidth}
+    	\begin{subfigure}[t]{0.49\columnwidth}
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap1min.jpg}
-        		\caption{Case 2: $1$ min}
+          	\includegraphics[height=0.5\textwidth]{MarsFullMap2min.jpg}
+        		\caption{$2$ min}
 		\vspace*{0.025\textwidth}
@@
 	\centering
-	\begin{subfigure}[t]{0.4\columnwidth}
+	\begin{subfigure}[t]{0.49\columnwidth}
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsFullMap2min.jpg}
-        		\caption{Case 1: $2$ min}
+          	\includegraphics[height=0.5\textwidth]{MarsFullMap3min.jpg}
+        		\caption{$3$ min}
 		\vspace*{0.025\textwidth}
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.4\columnwidth}
+    	\begin{subfigure}[t]{0.49\columnwidth}
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap2min.jpg}
-        		\caption{Case 2: $2$ min}
+          	\includegraphics[height=0.5\textwidth]{MarsFullMap4min.jpg}
+        		\caption{$4$ min}
 		\vspace*{0.025\textwidth}
@@
 	\centering
-	\begin{subfigure}[t]{0.4\columnwidth}
+	\begin{subfigure}[t]{0.49\columnwidth}
            	\centering
           	\includegraphics[height=0.5\textwidth]{MarsFullMap5min.jpg}
-        		\caption{Case 1: $5$ min}
+        		\caption{$5$ min}
 		\vspace*{0.025\textwidth}
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.4\columnwidth}
+    	\begin{subfigure}[t]{0.49\columnwidth}
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap5min.jpg}
-        		\caption{Case 2: $5$ min}
+          	\includegraphics[height=0.5\textwidth]{MarsFullMap10min.jpg}
+        		\caption{$10$ min}
 		\vspace*{0.025\textwidth}
@@
 	\centering
-	\begin{subfigure}[t]{0.4\columnwidth}
+	\begin{subfigure}[t]{0.49\columnwidth}
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsFullMap10min.jpg}
-        		\caption{Case 1: $10$ min}
+          	\includegraphics[height=0.5\textwidth]{MarsFullMap15min.jpg}
+        		\caption{$15$ min}
 		\vspace*{0.025\textwidth}
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.4\columnwidth}
+    	\begin{subfigure}[t]{0.49\columnwidth}
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap10min.jpg}
-        		\caption{Case 2: $10$ min}
+          	\includegraphics[height=0.5\textwidth]{MarsFullMap20min.jpg}
+        		\caption{$20$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+\caption{In Case 1, the robot (red disk with arrow indicating laser direction) moves toward candidate poses (red arrows, more opaque for greater reward) based on expected entropy change of the 3D probabilistic occupancy grid map $m$ (cubes: greater opacity for greater occupancy probability) of the surface of Mars.}
+\label{fig:mars3DogmCase1}
+\end{figure}
+
+\begin{figure}[!t]
+	\centering
+	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap1min.jpg}
+        		\caption{$1$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap2min.jpg}
+        		\caption{$2$ min}
 		\vspace*{0.025\textwidth}
@@
 	\centering
-	\begin{subfigure}[t]{0.4\columnwidth}
+	\begin{subfigure}[t]{0.49\columnwidth}
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsFullMap15min.jpg}
-        		\caption{Case 1: $15$ min}
+          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap3min.jpg}
+        		\caption{$3$ min}
 		\vspace*{0.025\textwidth}
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.4\columnwidth}
+    	\begin{subfigure}[t]{0.49\columnwidth}
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsFullMap20min.jpg}
-        		\caption{Case 1: $20$ min}
+          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap4min.jpg}
+        		\caption{$4$ min}
 		\vspace*{0.025\textwidth}
     	\end{subfigure}
-\caption{The robot (red disk with arrow indicating laser direction) moves toward candidate poses (red arrows, more opaque for greater reward) while generating a 3D probabilistic occupancy grid map (cubes: greater opacity for greater occupancy probability) of the surface of Mars.}
-\label{fig:mars3Dogm}
+	\centering
+	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap5min.jpg}
+        		\caption{$5$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap10min.jpg}
+        		\caption{$10$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+\caption{In Case 2, the robot (red disk with arrow indicating laser direction) generates a 3D probabilistic occupancy grid map $m$ (cubes: greater opacity for greater occupancy probability), which is used to generate a reduced map $m_\text{reduced}$ based on \refeqn{Proj3DMapComb}.
+The robot moves toward candidate poses (red arrows, more opaque for greater reward) based on expected entropy change of $m_\text{reduced}$.}
+\label{fig:mars3DogmCase2}
 \end{figure}
@@
 
-\todo{I wish we have a little more pages. Let's split Figure 5 into two, one for Case 1 and another for Case 2, where each case has more time instances. (Also, check the caption for (j)).
-We can have another zoomed-in figure hat shows the aerial vehicle with the simulated measurements of the point clouds and the candidate poses more clearly}
+\todo{I wish we have a little more pages. We can have another zoomed-in figure hat shows the aerial vehicle with the simulated measurements of the point clouds and the candidate poses more clearly}
 

### 2018-06-11 Conference/SCITECH19/SCITECH19.tex
added new Mars results with butterscotch color and 20k costmap size

@@
            	\centering{
-          	\includegraphics[height=0.9\textwidth]{mars_gazebo_sqr.png}}
+          	\includegraphics[height=0.9\textwidth]{mars_gazebo_full_lighting.png}}
         		\caption{Gazebo Environment}
@@
 The bump functions use $f_\text{max}=1.0$, $f_\text{far}=0.1$, and $\beta=0.1$ to account for travel costs with \refeqn{BumpFunRecedingHorizon}.
-The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.25$ m/sec and the computation time varies from $0.6$ sec to $1.0$ sec.
+The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.25$ m/sec and the computation time varies from $0.6$ sec to $1.9$ sec.
 
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsFullMap1min.jpg}
-        		\caption{$1$ min}
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap30sec.jpg}
+        		\caption{$30$ sec}
 		\vspace*{0.025\textwidth}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsFullMap2min.jpg}
-        		\caption{$2$ min}
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap1min.jpg}
+        		\caption{$1$ min}
 		\vspace*{0.025\textwidth}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsFullMap3min.jpg}
-        		\caption{$3$ min}
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap2min.jpg}
+        		\caption{$2$ min}
 		\vspace*{0.025\textwidth}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsFullMap4min.jpg}
-        		\caption{$4$ min}
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap3min.jpg}
+        		\caption{$3$ min}
 		\vspace*{0.025\textwidth}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsFullMap5min.jpg}
-        		\caption{$5$ min}
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap4min.jpg}
+        		\caption{$4$ min}
 		\vspace*{0.025\textwidth}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsFullMap10min.jpg}
-        		\caption{$10$ min}
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap5min.jpg}
+        		\caption{$5$ min}
 		\vspace*{0.025\textwidth}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsFullMap15min.jpg}
-        		\caption{$15$ min}
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap10min.jpg}
+        		\caption{$10$ min}
 		\vspace*{0.025\textwidth}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsFullMap20min.jpg}
-        		\caption{$20$ min}
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap15min.jpg}
+        		\caption{$15$ min}
 		\vspace*{0.025\textwidth}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap1min.jpg}
-        		\caption{$1$ min}
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap30sec.jpg}
+        		\caption{$30$ sec}
 		\vspace*{0.025\textwidth}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap2min.jpg}
-        		\caption{$2$ min}
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap1min.jpg}
+        		\caption{$1$ min}
 		\vspace*{0.025\textwidth}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap3min.jpg}
-        		\caption{$3$ min}
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap2min.jpg}
+        		\caption{$2$ min}
 		\vspace*{0.025\textwidth}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap4min.jpg}
-        		\caption{$4$ min}
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap3min.jpg}
+        		\caption{$3$ min}
 		\vspace*{0.025\textwidth}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap5min.jpg}
-        		\caption{$5$ min}
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap4min.jpg}
+        		\caption{$4$ min}
 		\vspace*{0.025\textwidth}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{MarsRdcdMap10min.jpg}
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap5min.jpg}
+        		\caption{$5$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+	\centering
+	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap10min.jpg}
         		\caption{$10$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap15min.jpg}
+        		\caption{$15$ min}
 		\vspace*{0.025\textwidth}
@@
 		\centerline{
-			\includegraphics[width=0.6\columnwidth]{mars_H_comparison_flat.pdf}
+			\includegraphics[width=0.6\columnwidth]{scitech_entropy_comparison_flat.pdf}
 		}
-		\caption{The complete map entropy for Case 1 (red solid line) decreases at roughly the same rate as the reduced map entropy for Case 2 (dashed blue line), but the complete map entropy highlights a few locations where entropy can decrease further, shown by lower entropy over the subsequent $10$ min.}
+		\caption{The complete map entropy for Case 1 (red solid line) decreases at roughly the same rate as the reduced map entropy for Case 2 (dashed blue line) toward the beginning.
+However, the expected entropy in Case 2 promotes actions toward unexplored territory, while the policy of Case 1 yields a more complete map before moving forward.}
 		\label{fig:mars3Dentropy}
@@
 The exploration policy of Case 2 incorrectly assumed these regions were well-known.
-Ironically, this false assumption sometimes actually led to greater information gains when the vehicle moved on to unvisited terrain.
-Here, the exploration policy of Case 2 suspended operations once expected information gains were insignificant, which happened after about $10$ minutes.
-However with Case $1$, which used $m$ for computing expected entropy, the exploration continued for an additional $10$ minutes, overall decreasing total map entropy further, generating a more complete occupancy grid of the surface of Mars.
+Ironically, this false assumption actually led to greater information gains when the vehicle moved on to unvisited terrain.
+Conversely with Case $1$, which used $m$ for computing expected entropy, the total map entropy decreased more steadily and generated a more-complete 3D occupancy grid of the surface of Mars.
 

### 2018-06-11 Conference/SCITECH19/SCITECH19.tex
added close-up figures

@@
 The resulting occupancy grid maps for Case 1 and 2 are shown in Figs.
-\ref{fig:mars3DogmCase1} and \ref{fig:mars3DogmCase2}, respectively, and complete map entropy is illustrated in Fig.
+\ref{fig:mars3DogmCase1} and \ref{fig:mars3DogmCase2}, respectively, and close-up pictures from Case 2 are shown in \ref{fig:marsZoomedIn}.
+The complete map entropy is illustrated in Fig.
 \ref{fig:mars3Dentropy}.
@@
 
+\begin{figure}[!t]
+	\centering
+	\begin{subfigure}[t]{0.95\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{mars_closeup.png}
+        		\caption{Full Map Cells and Sensor Scan}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+	\centering
+	\begin{subfigure}[t]{0.95\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{scitech_reducedCells.png}
+        		\caption{Full Map and Reduced Map Cells}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+\caption{The close-up images from the Case 2 trial show the candidate future poses (red arrows), where greater opacity represents a larger objective function of \refeqn{CandidateLocationOptimization}.
+In (a), we show the 3D scan with color corresponding to the surface of Mars.
+In (b), we overlay the full map $m$ (colored) with the reduced map $m_\text{reduced}$ (gray).}
+\label{fig:marsZoomedIn}
+\end{figure}
+
 	\begin{figure}
@@
 
-\todo{I wish we have a little more pages. We can have another zoomed-in figure hat shows the aerial vehicle with the simulated measurements of the point clouds and the candidate poses more clearly}
-
 \section{Conclusions}

### 2018-06-11 Conference/SCITECH19/SCITECH19.tex
edited abstract, intro, and problem def

@@
 
-The topography of Mars has several benefits in future mission planning for landing spacecraft and  image distortion correction among other research objectives.
+The topography of Mars has several benefits in future mission planning for landing spacecraft and  image distortion correction, among other research objectives.
 The Mars Orbiting Laser Altimeter (MOLA) mission greatly improved the topography map of the surface, but was limited to an uncertainty of roughly eight meters~\cite{MOLA99}.
@@
 At the $t$-th time step, we assume a known current pose $X_t=\braces{x_t,R_t}$ such that $x_t\in\Re^3$ is a 3D location and $R_t\in\SO$ is the 3D attitude, and history of poses $X_{1:t-1}$ is assumed known as well.
-Considering that most modern sensors return scans with multiple measurement rays, let measurement ray $z\in\Sph^2$ belong the the most recent scan $Z_t$, and let $Z_{1:t-1}$ be the history of measurement scans.
+Considering that most modern sensors return scans with multiple measurement rays, let measurement ray $z$ be a single depth measurement belonging the the most recent scan $Z_t$, and let $Z_{1:t-1}$ be the history of measurement scans.
 
@@
 respectively.
-This serves as an effective measure because \refeqn{ShannonsEntropyCell} is minimized for $P(\mathbf{m}_i)=0$ and $P(\mathbf{m}_i)=1$ (smallest uncertainty), but is maximized when $P(\mathbf{m}_i)=0.5$ (largest uncertainty).
+This serves as an effective measure because \refeqn{ShannonsEntropyCell} is minimized as $P(\mathbf{m}_i)$ approaches $0$ or $1$ (smallest uncertainty), but is maximized when $P(\mathbf{m}_i)=0.5$ (largest uncertainty).
 
@@
 The bump functions use $f_\text{max}=1.0$, $f_\text{far}=0.1$, and $\beta=0.1$ to account for travel costs with \refeqn{BumpFunRecedingHorizon}.
-The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.25$ m/sec and the computation time varies from $0.6$ sec to $1.9$ sec.
+The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.25$ m/sec and the computation time varies from $1.8$ sec to $2.5$ sec.
 

### 2018-06-11 Conference/SCITECH19/SCITECH19.tex
edited approach section

@@
 -\sin\theta & 0 & \cos\theta
-\end{bmatrix}, \quad -\frac{\pi}{2}\leq\theta\leq\frac{\pi}{2},
+\end{bmatrix}, \quad -\frac{\pi}{2}\leq\theta\leq\frac{\pi}{2}.
 \end{align*}
-where $\Theta$ is the angular range for sensor field-of-view in the vertical direction.
 Combining these two rotation matrices, the measurement $z$ with depth $\norm{z}$ is expressed with respect to a frame fixed to Mars as
@@
 \end{align}
-where the mounting yaw $\tilde{\psi}$ and pitch $\tilde{\theta}$ of the sensor is a fixed rotation in the horizontal and vertical directions, respectively in that order, such that $-\pi<\tilde{\psi}\leq\pi$ and $-\frac{\pi}{2}\leq\tilde{\theta}\leq\frac{\pi}{2}$.
-A positive $\tilde{\theta}$ corresponds to rotating the depth sensor from forward to downward toward the surface of Mars.
+where the mounting yaw $\tilde{\psi}$ and pitch $\tilde{\theta}$ of the sensor is a fixed rotation in the horizontal and vertical directions, respectively in that order, such that $-\pi<\tilde{\psi}\leq\pi$ and $-\frac{\pi}{2}+\frac\Theta2\leq\tilde{\theta}\leq\frac{\pi}{2}-\frac\Theta2$.
+A positive $\tilde{\theta}$ corresponds to rotating the depth sensor from a forward direction to a downward direction toward the surface of Mars.
 The expected information gain for the full scan is the summation of expected information gains for individual rays
@@
 \end{align}
-where initial probability is $0<P_\text{init}<1$ and threshold the threshold probability is constrained to $P_\text{init}<P_\text{thresh}<1$.
+where initial probability is $0<P_\text{init}<1$ and the threshold probability is constrained to $P_\text{init}<P_\text{thresh}<1$.
 This approach uses probabilities that have changed from $P_\text{init}$, and favors occupied cells to avoid risking collisions.
 
-Next, we show how $\mathbf{m}_{\text{reduced},k}$ is used in Dijkstra's search~\cite{Dij59}.
+Next, we show how $\mathbf{m}_{\text{reduced},k}$ is used in Dijkstra's search~\cite{Dij59} for collision avoidance and motion planning.
 Defining $0<P_\text{coll}<1$ as the acceptable probability of collision, every grid cell of $ m_\text{reduced}$ is considered a safe and unvisited node if its occupancy probability, and the occupancy probabilities of its neighbors (sharing a face, edge, or corner), are below $P_\text{coll}$.

### 2018-06-11 Conference/SCITECH19/SCITECH19.tex
edited the remaining text & figure captions

@@
 The Mars frame (short and thick) is fixed to the planet.
-The body-fixed frame (medium thickness and length) is fixed to the camera frame (long and narrow).}
+The body-fixed frame (medium thickness and length) is fixed to the camera frame (long and narrow).
+The rotation of $\psi$ (b) represents the fixed camera yaw rotation and $\theta$ (c) represents the downward angle of the camera to capture the surface of Mars below the flying robot.}
 \label{fig:transforms}
@@
 	\vspace*{-0.25\columnwidth}
-		\caption{The bump function is designed to prioritize local trajectories to avoid traversals across the map.
+		\caption{The bump function is multiplied to the objective function of \refeqn{CandidateLocationOptimization} to prioritize local trajectories to avoid traversals across the map, where travel costs are determined using Dijkstra's algorithm.
 However, there is no time wasted up to $d_\text{opt}$ because this distance may be achieved in the minimum exploration computation time.}
@@
 The mapping node takes measurements from a 3D laser scanner and pairs those with the sensor transformation relative to Mars using the ROS Transform Message Filters package.
-Then, it builds a 3D probabilistic map using \refeqn{RayISMAnswer}-\refeqn{Unnormalized}.
+Then, it builds a 3D probabilistic map using ray casting and the inverse sensor model \refeqn{RayISMAnswer}-\refeqn{Unnormalized}.
 The mapping node sends updates from the probabilistic map to the exploration node, which decomposes the full map $m$ into $ m_\text{reduced}$ according to \refeqn{Proj3DMapComb}, and builds a cost map based on Dijkstra's search to determine collision-free trajectory costs for all pose candidates.
-Finally, it applies \refeqn{RcStar}, \refeqn{BumpFunRecedingHorizon}, and \refeqn{CandidateLocationOptimization} to update the desired robot trajectory on a receding-horizon.
+Finally, the exploration node applies \refeqn{RcStar}, \refeqn{BumpFunRecedingHorizon}, and \refeqn{CandidateLocationOptimization} to update the desired robot trajectory on a receding-horizon.
 An overview of the process is illustrated in Fig.
@@
 		}
-		\caption{The diagram of the 3D probabilistic mapping and autonomous exploration nodes shows the primary processes and communication lines.}
+		\caption{The diagram of the 3D probabilistic mapping and autonomous exploration nodes shows the primary processes and communication lines. These ROS nodes (boxes) operate together in real-time.}
 		\label{fig:NodesDiagram}
@@
     	\end{subfigure}
-\caption{The height map (a) and color map (b) are used in Gazebo (c) to simulate a Mars environment.}
+\caption{The height map (a) and color map (b) are combined in the Gazebo simulator (c) to model a Mars environment. The robot captures the simulated environment to generate an occupancy grid map.}
 \label{fig:MarsGazebo}
@@
 Case 2 is largely inspired by the promising results of~\cite{KauTakAiLee18}, where a projected map based on the same criteria as \refeqn{Proj3DMapComb} proved effective for level flight.
-The resulting occupancy grid maps for Case 1 and 2 are shown in Figs.
-\ref{fig:mars3DogmCase1} and \ref{fig:mars3DogmCase2}, respectively, and close-up pictures from Case 2 are shown in \ref{fig:marsZoomedIn}.
+The resulting occupancy grid maps for Cases 1 and 2 are shown in Figs.
+\ref{fig:mars3DogmCase1} and \ref{fig:mars3DogmCase2}, respectively, and close-up pictures from Case 2 are shown in Fig.
+\ref{fig:marsZoomedIn}.
 The complete map entropy is illustrated in Fig.
@@
 \caption{In Case 2, the robot (red disk with arrow indicating laser direction) generates a 3D probabilistic occupancy grid map $m$ (cubes: greater opacity for greater occupancy probability), which is used to generate a reduced map $m_\text{reduced}$ based on \refeqn{Proj3DMapComb}.
-The robot moves toward candidate poses (red arrows, more opaque for greater reward) based on expected entropy change of $m_\text{reduced}$.}
+The robot moves toward candidate poses (red arrows, more opaque for greater reward) based on expected entropy change of $m_\text{reduced}$.
+Using $m_\text{reduced}$ instead of $m$ for entropy calculation leads to faster exploration of new terrain, but this leaves some grid cells missing.}
 \label{fig:mars3DogmCase2}
@@
 In (a), we show the 3D scan with color corresponding to the surface of Mars.
-In (b), we overlay the full map $m$ (colored) with the reduced map $m_\text{reduced}$ (gray).}
+In (b), we overlay the full map $m$ (colored) with the reduced map $m_\text{reduced}$ (gray) from \refeqn{Proj3DMapComb}.}
 \label{fig:marsZoomedIn}
@@
 		\caption{The complete map entropy for Case 1 (red solid line) decreases at roughly the same rate as the reduced map entropy for Case 2 (dashed blue line) toward the beginning.
-However, the expected entropy in Case 2 promotes actions toward unexplored territory, while the policy of Case 1 yields a more complete map before moving forward.}
+However, the expected entropy in Case 2 promotes actions toward unexplored territory, while the policy of Case 1 yields a more complete map before moving on to unexplored spaces.}
 		\label{fig:mars3Dentropy}
@@
 Conversely in Case 2, when $m_\text{reduced}$ was used instead, the robot repeatedly left spaces that were missing a few grid cells to visit new terrain.
-This is because $m_\text{reduced}$ had some cells with large occupancy probabilities based on \refeqn{Proj3DMapComb}, which allowed some grid cells of $m$ enclosed within a cell of $m_\text{reduced}$ to be uncertain.
+This is because $m_\text{reduced}$ contained some cells with large occupancy probabilities based on \refeqn{Proj3DMapComb}, which allowed some grid cells of $m$ enclosed within a cell of $m_\text{reduced}$ to be uncertain.
 The exploration policy of Case 2 incorrectly assumed these regions were well-known.
@@
 In this paper, we proposed an approach to autonomously explore the complex and uncertain terrain of Mars.
-The exploration strategy is based on flying vehicles near the surface of Mars, where their exploration policy is based on expected information gain from Shannon's entropy of a probabilistic map, and travel costs from Dijkstra's algorithm.
-The autonomous exploration scheme was applied in real-time, following a receding-horizon framework, so the map could be generated as quickly as possible.
+The exploration strategy is designed for flying vehicles near the surface of Mars, where the exploration policy is based on expected information gain from Shannon's entropy of a probabilistic map, and travel costs from Dijkstra's algorithm.
+The autonomous exploration scheme is applied in real-time, following a receding-horizon framework, so the map could be generated as quickly as possible.
 Numerical simulations demonstrated the efficacy of the approach, showing two possible versions of the 3D exploration with differing behavior in probabilistic map accuracy and exploration speed.

### 2018-06-11 Conference/SCITECH19/SCITECH19.tex
submitted draft

@@
 
-\author{Evan Kaufman\footnote{Graduate Research Assistant, Department of Mechanical and Aerospace Engineering, The George Washington University, 800 22nd St NW, Washington, DC, 20052} and Taeyoung Lee\footnote{Associate Professor, Department of Mechanical and Aerospace Engineering, The George Washington University, 800 22nd St NW, Washington, DC, 20052}}% TODO: update if AIAA member
+\author{Evan Kaufman\footnote{Graduate Research Assistant, Department of Mechanical and Aerospace Engineering, The George Washington University, 800 22nd St NW, Washington, DC, 20052} and Taeyoung Lee\footnote{Associate Professor, Department of Mechanical and Aerospace Engineering, The George Washington University, 800 22nd St NW, Washington, DC, 20052}}
 \affil{The George Washington University, Washington, DC, 20052}
-% TODO: funding source
 
@@
 The resulting occupancy grid maps for Cases 1 and 2 are shown in Figs.
-\ref{fig:mars3DogmCase1} and \ref{fig:mars3DogmCase2}, respectively, and close-up pictures from Case 2 are shown in Fig.
+\ref{fig:mars3DogmCase1} and \ref{fig:mars3DogmCase2}, respectively.
+A video of Case 1 can be found at \href{https://youtu.be/FrBcL2UMW9w}{https://youtu.be/FrBcL2UMW9w} and close-up pictures from Case 2 are shown in Fig.
 \ref{fig:marsZoomedIn}.

### 2018-06-14 Dissertation/Thesis/dissertation.tex
added some paragraphs to intro up to list of pub

@@
 
+Here we describe the existing approaches to building probabilistic occupancy grid maps.
+Though several variations have been proposed, no approach uses the stochastic properties of depth sensors directly to obtain an exact Bayesian solution that can be computed in real-time.
+
+% First proposed in 80s 
+% Log-odds ratio solution
+% simplified function - lines, patched Gaussians
+
+% Learning
+% Hill-climbing
+
+In short, other existing solutions to occupancy probability are inexact, and involve a potentially-harmful log-odds ratio assumption.
+These shortcomings motivate an exact solution to occupancy grid mapping that can be computed in real-time.
+
+%It was previously-believed that the exact solution to occupancy grid mapping was impossible in real-time.
+
 \subsection{Autonomous Exploration}
@@
 
+There are three key contributions in this dissertation.
+The first two are novel solutions that find exact occupancy probabilities and predicted entropies, respectively.
+These contributions improve the quality and efficiency of autonomous mapping and exploration of uncertain environments for single-vehicle missions.
+The third contribution extends these concepts to multi-vehicle scenarios, where multiple members of a team work together to autonomously explore and patrol large environments.
+
 \subsection{Summary of Contributions}
+
+The first contribution is an exact Bayesian solution to occupancy grid mapping.
+The proposed approach considers that occupied spaces occlude measurement behind them, which allows certain mapping outcomes to be systematically grouped together.
+The exact solution to occupancy grid cell probability, which was previously-believed to have exponential complexity, is now possible to solve in real-time.
+Numerous numerical simulations and experiments demonstrate that the proposed mapping approach in effective and inexpensive in 2D and 3D scenarios.
+
+The second contribution is a new approach to autonomous exploration, where the robot predicts future occupancy grid map uncertainty, using a measure known as Shannon's entropy.
+This approach computes the expected values of grid cell entropies from potential future measurement rays with a novel approach to predict future measurement probabilities.
+These predictions are used to evaluate the benefit of possible robot actions to maximize map information gain.
+Then, we formulate autonomous exploration as an optimization problem to maximize an objective function that includes map uncertainty and travel costs.
+The probabilistic map is used for collision-avoidance and motion planning.
+These processes are designed for real-time applications, where the robot makes decisions as it learns information about the surrounding environment.
+
+The final contribution coordinates the mapping and exploration efforts of multiple robots together.
+The mapping process is multi-threaded such that multiple members update the same occupancy grid.
+The exploration algorithm follows a bidding-based structure.
+Robots compete for tasks in a series of auctions, where the winning bid of one auction affects the bids of subsequent auctions, thereby coordinating the robot efforts.
+This cooperative approach is computationally-efficient in large environments, and follows a receding-horizon framework, such that map expected information gains are updated as quickly as possible.
+The proposed coordination is extended to autonomous patrol, where occupancy grid map probabilities are degraded over time, which promotes the robots to revisit areas periodically.
+
 \subsection{List of Publications}
+
+The following is a list of publications. These include research included in this dissertation, as well as contributions in control, data association, and estimation.
+
+ADD LIST
 

### 2018-06-15 Dissertation/Thesis/dissertation.tex
added list of pub with \item cmd, bib versions for this list not working

@@
 \usepackage{lipsum}
+
 \graphicspath{{../../Fig/}}
@@
 
-ADD LIST
+% TODO: update this list
+
+\begin{itemize}
+	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Autonomous Exploration with Exact Inverse Sensor Models,'' \textit{Journal of Intelligent }\&\textit{ Robotic Systems}, 2016, submitted.
+	\item E.
+Kaufman, T.
+Lee, and Z.
+Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, San Francisco, December 2016, accepted.
+	\item E.
+Kaufman, T.
+Lee, Z.
+Ai, and I.
+S.
+Moskowitz, ``Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model,'' \textit{Proceedings of the American Control Conference}, Boston, July 2016, pp.
+5709-5715.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{The Journal of the Astronautical Sciences}, 63(1): pp.
+60-80, 2016, doi: 10.1007/s40295-015-0082-9.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Minimum Uncertainty JPDA Filters and Coalescence Avoidance for Multiple Object Tracking,'' \textit{The Journal of the Astronautical Sciences}, 63(4): pp.
+308-334, 2016, doi: 10.1007/s40295-016-0092-2.
+	\item T.
+Wu, E.
+Kaufman, and T.
+Lee, ``Globally Asymptotically Stable Attitude Observer on SO(3),'' \textit{Proceedings of the 54th IEEE Conference on Decision and Control}, pp.
+2164-2168, Osaka, Japan, December 2015.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting}, Williamsburg, VA, Jan.
+2015, AAS 15-451.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Minimum Uncertainty JPDA Filter and Coalescence Avoidance Performance Evaluations,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting, Williamsburg}, VA, Jan.
+2015, AAS 15-432.
+	\item E.
+Kaufman, K.
+Caldwell, D.
+Lee, and T.
+Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, Mar.
+2014, ASC 14-2527.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Optimal Joint Probabilistic Data Association Filter Avoiding Coalescence in Close Proximity,'' \textit{Proceedings of the European Control Conference}, pp.
+2709-2714, June 2014.
+	\vspace*{0.5cm}
+\end{itemize}
 
@@
 
-\bibliographystyle{plain}
-\bibliography{thesis-bib}
+%\bibliographystyle{plain}
+%\bibliography{../../BibSources.bib}
 

### 2018-06-20 Dissertation/Thesis/dissertation.tex
eliminated some unnecessary files from dissertation
beginnings of mapping lit review

@@
 Though several variations have been proposed, no approach uses the stochastic properties of depth sensors directly to obtain an exact Bayesian solution that can be computed in real-time.
+
+\paragraph{Inverse Sensor Model} The key to generating a probabilistic occupancy grid map is known as the \emph{inverse sensor model}~\cite{ThrBurFox05}.
+The main idea is that a depth sensor has a probability density (e.g.
+Gaussian) that describes the stochastic properties of a sensor reading (e.g.
+a LIDAR range) with respect to the expected value of a reading (e.g.
+distance to a wall).
+This probability density is referred to as the \emph{forward sensor model}, as it relates a depth measurement conditioned on the occupancy of the map.
+
+In contrast, the inverse sensor model relates the occupancy of the map conditioned on the depth measurement.
+This problem is solved using Bayes' theorem.
+However, the occupancy grid map properties yield a complicated Bayesian probability; the computational cost has exponential complexity with respect to the number of grid cells captured within the depth sensor limits.
+It was previously-believed that this restriction makes solving the inverse sensor model in real-time impossible.
+This motivated approximate or estimated solutions, described next.
+
+\paragraph{Approximated Function for the Inverse Sensor Model} Elfs et al originally proposed occupancy grid mapping in 1988...
 
@@
 
-%\bibliographystyle{plain}
-%\bibliography{../../BibSources.bib}
+\bibliographystyle{plain}
+\bibliography{../../BibSources.bib}
 

### 2018-06-21 Dissertation/Thesis/dissertation.tex
added mapping background & beginning of autonomous exploration lit review

@@
 
-\paragraph{Approximated Function for the Inverse Sensor Model} Elfs et al originally proposed occupancy grid mapping in 1988...
+\paragraph{Approximated Function for the Inverse Sensor Model} Moravec and Elfes originally proposed a grid cell-based mapping for sonar sensors in~\cite{MorElf85}, and the probabilistic properties of the occupancy grid were formalized in~\cite{MorElf85,Elf89}.
+This approach models measurements with Gaussian probability density, although the occupancy of grid cells is not exactly known.
+Instead, the probability of grid cells are estimated with a Gaussian-like \emph{approximate function}.
+The authors show that occupancy probability converges to a final value following a Bayesian framework, though there is no conclusive proof that this probability is accurate.
+Nevertheless, this approach garnered much attention because it could efficiently produce robotic maps that nicely-resembled the occupancy of spaces around a robot.
 
-% First proposed in 80s 
-% Log-odds ratio solution
-% simplified function - lines, patched Gaussians
+Several other research studies have applied approximate functions in place of the inverse sensor model.
+For example, \cite{ChoLynHutKanBurKavThr05} proposed a simplified continuous function composed of lines patched together near a measurement reading.
+Furthermore, variations of approximate smoother functions in~\cite{and09,PirRutBisSch11,KhoElb12} contain Gaussian-like terms to apply occupancy grid mapping to 3D environments with more modern sensors.
+Numerous research studies have used the proposed approximate functions with questionable accuracy.
 
-% Learning
-% Hill-climbing
+\paragraph{Learning the Inverse Sensor Model} The approximate nature of the inverse sensor model motivated a solution using machine learning.
+The main idea is to simulate maps, robot poses, and measurements with simulated noised based on the sensor stochastic properties.
+Then, an approximation the inverse sensor model is learned using an expectation maximization algorithm~\cite{Thr01,ThrBurFox05}.
+Unlike approximate functions based on intuition, the machine learning approach is based on the physical properties of sensors.
+However, this approach is undesirable in practice due to complexities associated to implementing a learning algorithm.
+For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples, or how many samples are required to obtain a reasonable approximation.
+Furthermore, it is challenging to apply any learning algorithm over the large dimensional space composed of maps, poses, and measurements.
+
+\paragraph{Log-Odds Ratio} Both approaches described above use the log-odds ratio to update grid cell probabilities.
+The log-odds ratio formulation is a popular framework for updating binary random variables with static state within a Bayesian filter.
+The main idea is that instead of multiplying terms from prior and current time steps, the properties of logarithms allow these terms to be simply added, while avoiding some truncation issues associated with probabilities close to $0$ or $1$~\cite{ThrBurFox05}.
+This approach is also popular because the inverse sensor model needs not consider prior probabilities.
+Instead, these approaches typically consider fixed uniform initial probabilities of all grid cells, which makes the formulation simpler.
+
+However, the log-odds ratio formulation makes an assumption that is not consistent with the occupancy grid mapping problem.
+The approach yields a simplified solution by assuming that the probability of a measurement is independent of past measurements and robot poses.
+This assumption is frequently violated when past measurements and poses indicate the occupancy of other cells between the robot and the cell in question.
+As such, this approach neglects potentially important information when considering grid cell occupancy probability.
 
@@
 \subsection{Autonomous Exploration}
+
+The next topic this dissertation addresses is autonomous exploration.
+This topic is important for planning the trajectories of a robot to generate a complete and accurate map.
+Frontier-based and entropy-based approaches are commonly used to achieve this goal, described next.
+
+\paragraph{Frontier-Based Autonomous Exploration}
+
+\paragraph{Entropy-Based Autonomous Exploration}
+
+\
+
+In short, existing autonomous exploration approaches are effective for eliminating the need for humans to control the robot.
+However, optimality remains quite difficult, so these approaches apply suboptimal solutions to achieve autonomous motions.
+
+% JINT17: However, there are certain issues with the current SLAM techniques.
+First, the vast majority of work deals only with the aspect of estimating the environment and the poses of vehicles.
+They are passive in the sense that SLAM is per- formed on incoming sensor measurements from vehicles following an arbitrary path.
+As such, human teleopera- tion and monitoring are often essential to guide the vehi- cles safely through unknown surroundings.
+Therefore, it is desired that an active motion planning scheme is devel- oped and integrated with SLAM such that vehicles are able to determine their path without human intervention, and explore unknown areas autonomously.
 
@@
 The exact solution to occupancy grid cell probability, which was previously-believed to have exponential complexity, is now possible to solve in real-time.
-Numerous numerical simulations and experiments demonstrate that the proposed mapping approach in effective and inexpensive in 2D and 3D scenarios.
+Numerous numerical simulations and experiments demonstrate that the proposed mapping approach is effective and inexpensive in 2D and 3D scenarios.
+
+% JINT17: This paper proposes a computationally efficient algo- rithm to construct the exact inverse sensor model.
+More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, this algo- rithm yields the a posteriori probabilities of occupancy of all the cells within the area covered by the range sensor from the range measurements.
+The key idea is reducing the computational load by using the fact that if a cell is occu- pied, the occupancies of the other cells blocked by it is irrelevant to the forward sensor model, and this property is systematically utilized with various probabilistic prop- erties to derive a computationally-efficient solution to the inverse sensor model.
+Furthermore, the proposed approach integrates a priori probabilities of occupancy and multiple range measurements according to the Bayesian framework to obtain more accurate maps.
+As such, it contrasts from the existing framework based on log-odds ratios that impose an additional Markov assumption.
 

### 2018-06-21 Dissertation/Thesis/dissertation.tex
exploration lit review, start of patrol

@@
 The next topic this dissertation addresses is autonomous exploration.
-This topic is important for planning the trajectories of a robot to generate a complete and accurate map.
-Frontier-based and entropy-based approaches are commonly used to achieve this goal, described next.
+The vast majority of work in simultaneous localization and mapping (SLAM) deals only with the aspect of estimating the environment and the poses of vehicles.
+These approaches are passive in the sense that SLAM is performed on incoming sensor measurements from vehicles following an arbitrary path.
+As such, human teleoperation and monitoring are often essential to guide the vehicles safely through unknown surroundings.
+Therefore, it is desirable that an active motion planning scheme is developed and integrated with SLAM such that vehicles are able to determine their path without human intervention, and explore unknown areas autonomously.
+The two approaches used to solve this problem are described next.
 
@@
 
+The most popular approach to solve the autonomous exploration problem is known as frontier-based exploration, originally-proposed in~\cite{Yam97,Yam98} for 2D applications, and was later extended to 3D with a visibility metric in~\cite{SawKriSri09} and~\cite{ZhuDinLinWu15,SenWan16,KleDor13} using the popular Octomap representation (a variation on occupancy grid mapping).
+The main idea is that robots toward the border between certain and uncertain space, referred to as a frontier.
+Then the robot takes depth measurement at this frontier, thus pushing back the boundary.
+This process is repeated until the map is well-known.
+Frontier-based exploration assumes that repeatedly moving toward the closest frontier and taking measurements are the best actions to gain new information about the map.
+However, these systematic actions, based on ad-hoc rules for what constitutes a frontier, do not consider the future uncertainty of the probabilistic map.
+Frontier-based approaches provide an intuitive solution, but lack optimality in map uncertainty or exploration time.
+
 \paragraph{Entropy-Based Autonomous Exploration}
+Entropy-based approaches address the suboptimal nature of frontier-based approaches by selecting robotic actions designed to decrease minimize map uncertainty~\cite{StaGriBur05,BurMooStaSch05}.
+These approaches use a measure of uncertainty known as Shannon's entropy, which becomes smaller as cell probabilities approach $0$ or $1$ (becoming more certain).
+Existing entropy-based approaches simulate possible future measurements from various locations with so-called ``hallucination measurements,'' which typically corresponds to expected depth measurements from the robot to the closest grid cell that is possibly occupied.
+Then, the robot analyzes how this measurement would affect the map uncertainty.
+However, this approach assumes that expected map entropy is equivalent to map entropy from expected measurements.
+This is not the case in general, however, as there is typically a nonzero probability that the hallucination measurement and true future measurement are not the same.
+Entropy-based approaches seek to optimize map uncertainty, but require large approximations which can decrease the accuracy of predictions.
+
+In conclusion, existing autonomous exploration approaches are effective for eliminating the need for humans to control the robot.
+However, optimality remains quite difficult, so these approaches apply suboptimal or approximated solutions to achieve autonomous motions.
+
+\subsection{Multi-Vehicle Cooperation}
+
+The final topic relates to coordinating robotic efforts to build occupancy grid maps together.
+These concepts are further extended to autonomous cooperative patrol, where robots periodically revisit regions to monitor a large environment.
+
+\paragraph{Cooperative Autonomous Exploration}
+
+\paragraph{Autonomous Patrol}
 
@@
 
-In short, existing autonomous exploration approaches are effective for eliminating the need for humans to control the robot.
-However, optimality remains quite difficult, so these approaches apply suboptimal solutions to achieve autonomous motions.
-
-% JINT17: However, there are certain issues with the current SLAM techniques.
-First, the vast majority of work deals only with the aspect of estimating the environment and the poses of vehicles.
-They are passive in the sense that SLAM is per- formed on incoming sensor measurements from vehicles following an arbitrary path.
-As such, human teleopera- tion and monitoring are often essential to guide the vehi- cles safely through unknown surroundings.
-Therefore, it is desired that an active motion planning scheme is devel- oped and integrated with SLAM such that vehicles are able to determine their path without human intervention, and explore unknown areas autonomously.
-
-\subsection{Multi-Vehicle Cooperation}
+In short, ...
 

### 2018-06-22 Dissertation/Thesis/dissertation.tex
finished intro w/o outline of paper and editing

@@
 
+Multi-vehicle cooperation is essential for robots to explore large spaces autonomously so that robots coordinate their efforts to improve map information and avoid collisions.
+However, optimizing map coverage simultaneously from multiple vehicles becomes very complicated and computationally intractable.
+Instead, an auction-based approach addresses the computational restrictions effectively.
+Bidding-based approaches were introduced with~\cite{SimApfBurFoxMooThrYou00,DiaSte00}, and extended to several other robotic problems~\cite{GerMaj02,ZloSteDiaTha02,SarBal05,ChoBruHow09} exploring decentralization and task allocation with varying robot capabilities.
+In~\cite{SimApfBurFoxMooThrYou00}, a central executive assigns tasks to robots for maximizing overall utility of the group.
+This is accomplished by minimizing a coverage overlap among the members.
+Robots submit bids based on expected coverage of uncertain spaces and travel distance.
+The robot winning the auction is awarded the task, and the remaining bids are discounted to prevent robots from covering the same area, thereby coordinating the mapping efforts.
+Bidding-based approaches simplify multi-vehicle cooperation in a computationally-efficient manner, and integrate effectively with the autonomous mapping and exploration approaches proposed in this dissertation.
+
 \paragraph{Autonomous Patrol}
 
-\
+Robotic surveillance and cleaning are examples of tasks that require periodic observations of the same regions.
+Several patrolling approaches have been proposed in recent years to address this need.
+A popular approach is decomposing the reachable space into Voronoi diagrams~\cite{KolCar08,PorRoc10,PipChrWei13}.
+In~\cite{KolCar08}, edges between nodes of the Voronoi graph are heavily sensed to track intruders.
+In~\cite{PorRoc10}, regions of the Voronoi diagram are assigned to different robots, and issues with robots having underperforming surveillance capabilities are addressed in~\cite{PipChrWei13}.
+These approaches assume the map is well-known in advance, and require a Voronoi graph be generated from the map.
 
-In short, ...
+Other patrol approaches create new metrics about the environment.
+Cell periodicity is estimated in~\cite{KraFenCieDonDuc14,KraFenHanDuc16} by predicting and observing the long-term dynamics of objects at the same locations on an occupancy grid.
+This work is extended to exploring and navigating a human-populated environment for long-term autonomy.
+A bio-inspired approach is proposed in~\cite{ZhaXia11}, where robots cooperate by dropping digital pheromones based on the importance of events, where the pheromones weaken over time to prioritize future tasks.
+Another approach is formulating patrol as an optimization problem with respect to robot idleness in~\cite{YanZha16}.
+The prior work in autonomous patrol demonstrates that periodic actions can be incentivized with a metric to be optimized.
 

### 2018-06-23 Dissertation/Thesis/dissertation.tex
I think I deleted everyting

--- 2018-06-21 Dissertation/Thesis/dissertation.tex
@@
 
-\chapter{Introduction} \label{chap:intro}
-
-This dissertation explores three key aspects of mobile robotics: occupancy grid mapping, autonomous exploration, and cooperation among robot teams.
-Several novel contributions are presented on both related topics for complex and dynamic situations.
-
-\section{Motivation and Goals}
-
-The study of robotics has expanded tremendously in recent years as robots have taken on dangerous, difficult, and repetitive tasks instead of humans.
-Certain tasks, such as search-and-rescue, surveillance, and cleaning, require some level of autonomy, particularly with mobile robots exploring new and uncertain environments.
-While these tasks provide significant value to our society, they present nontrivial challenges for the robots, particularly in mapping, exploration, and cooperation.
-
-\subsection{Building Probabilistic Maps}
-
-The first topic this dissertation addresses is building accurate probabilistic maps, known as \emph{mapping}.
-Map generation is crucial for simultaneous localization and mapping (SLAM) because maps provide important information for the robot to determine its pose (localization), which includes its position and attitude.
-Mapping also provides knowledge for trajectory planning through an environment and collision avoidance with mapped obstacles.
-In short, mapping is of fundamental importance in mobile robotics.
-
-Therefore, mapping quality deserves much attention.
-Robotics is inherently \emph{probabilistic}; no sensor reading or action from an actuator is deterministic.
-All robotic processes have some degree of uncertainty, and the mapping must account for this.
-Onboard sensors serve to improve the understanding of surrounding spaces, but the stochastic properties of those sensors must be reflected in the stochastic properties of the map.
-Therefore, we propose to develop probabilistic map that accounts for the history of measurements and poses, and their associated stochastic properties.
-Here, each element of the map holds a probability based on measured data.
-In turn, this probability can be applied to tasks such as collision-avoidance and predicting future map outcomes.
-
-\subsection{Exploration in Uncertain Spaces}
-
-The second topic of this dissertation focusses on the motions robots must choose to maximize their knowledge about the map while avoiding collisions, known as \emph{autonomous exploration}.
-
-\subsection{Cooperation Among Autonomous Systems}
-
-The final topic of this dissertation is cooperation among members of a robot team.
-
-\section{Literature Review}
-
-In this section, we discuss the existing approaches to occupancy grid mapping, autonomous exploration, and multi-vehicle cooperation.
-Several shortcomings with existing approaches are highlighted, which are addressed in the dissertation contributions.
-
-\subsection{Occupancy Grid Mapping}
-
-Here we describe the existing approaches to building probabilistic occupancy grid maps.
-Though several variations have been proposed, no approach uses the stochastic properties of depth sensors directly to obtain an exact Bayesian solution that can be computed in real-time.
-
-\paragraph{Inverse Sensor Model} The key to generating a probabilistic occupancy grid map is known as the \emph{inverse sensor model}~\cite{ThrBurFox05}.
-The main idea is that a depth sensor has a probability density (e.g.
-Gaussian) that describes the stochastic properties of a sensor reading (e.g.
-a LIDAR range) with respect to the expected value of a reading (e.g.
-distance to a wall).
-This probability density is referred to as the \emph{forward sensor model}, as it relates a depth measurement conditioned on the occupancy of the map.
-
-In contrast, the inverse sensor model relates the occupancy of the map conditioned on the depth measurement.
-This problem is solved using Bayes' theorem.
-However, the occupancy grid map properties yield a complicated Bayesian probability; the computational cost has exponential complexity with respect to the number of grid cells captured within the depth sensor limits.
-It was previously-believed that this restriction makes solving the inverse sensor model in real-time impossible.
-This motivated approximate or estimated solutions, described next.
-
-\paragraph{Approximated Function for the Inverse Sensor Model} Moravec and Elfes originally proposed a grid cell-based mapping for sonar sensors in~\cite{MorElf85}, and the probabilistic properties of the occupancy grid were formalized in~\cite{MorElf85,Elf89}.
-This approach models measurements with Gaussian probability density, although the occupancy of grid cells is not exactly known.
-Instead, the probability of grid cells are estimated with a Gaussian-like \emph{approximate function}.
-The authors show that occupancy probability converges to a final value following a Bayesian framework, though there is no conclusive proof that this probability is accurate.
-Nevertheless, this approach garnered much attention because it could efficiently produce robotic maps that nicely-resembled the occupancy of spaces around a robot.
-
-Several other research studies have applied approximate functions in place of the inverse sensor model.
-For example, \cite{ChoLynHutKanBurKavThr05} proposed a simplified continuous function composed of lines patched together near a measurement reading.
-Furthermore, variations of approximate smoother functions in~\cite{and09,PirRutBisSch11,KhoElb12} contain Gaussian-like terms to apply occupancy grid mapping to 3D environments with more modern sensors.
-Numerous research studies have used the proposed approximate functions with questionable accuracy.
-
-\paragraph{Learning the Inverse Sensor Model} The approximate nature of the inverse sensor model motivated a solution using machine learning.
-The main idea is to simulate maps, robot poses, and measurements with simulated noised based on the sensor stochastic properties.
-Then, an approximation the inverse sensor model is learned using an expectation maximization algorithm~\cite{Thr01,ThrBurFox05}.
-Unlike approximate functions based on intuition, the machine learning approach is based on the physical properties of sensors.
-However, this approach is undesirable in practice due to complexities associated to implementing a learning algorithm.
-For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples, or how many samples are required to obtain a reasonable approximation.
-Furthermore, it is challenging to apply any learning algorithm over the large dimensional space composed of maps, poses, and measurements.
-
-\paragraph{Log-Odds Ratio} Both approaches described above use the log-odds ratio to update grid cell probabilities.
-The log-odds ratio formulation is a popular framework for updating binary random variables with static state within a Bayesian filter.
-The main idea is that instead of multiplying terms from prior and current time steps, the properties of logarithms allow these terms to be simply added, while avoiding some truncation issues associated with probabilities close to $0$ or $1$~\cite{ThrBurFox05}.
-This approach is also popular because the inverse sensor model needs not consider prior probabilities.
-Instead, these approaches typically consider fixed uniform initial probabilities of all grid cells, which makes the formulation simpler.
-
-However, the log-odds ratio formulation makes an assumption that is not consistent with the occupancy grid mapping problem.
-The approach yields a simplified solution by assuming that the probability of a measurement is independent of past measurements and robot poses.
-This assumption is frequently violated when past measurements and poses indicate the occupancy of other cells between the robot and the cell in question.
-As such, this approach neglects potentially important information when considering grid cell occupancy probability.
-
-In short, other existing solutions to occupancy probability are inexact, and involve a potentially-harmful log-odds ratio assumption.
-These shortcomings motivate an exact solution to occupancy grid mapping that can be computed in real-time.
-
-%It was previously-believed that the exact solution to occupancy grid mapping was impossible in real-time.
-
-\subsection{Autonomous Exploration}
-
-The next topic this dissertation addresses is autonomous exploration.
-This topic is important for planning the trajectories of a robot to generate a complete and accurate map.
-Frontier-based and entropy-based approaches are commonly used to achieve this goal, described next.
-
-\paragraph{Frontier-Based Autonomous Exploration}
-
-\paragraph{Entropy-Based Autonomous Exploration}
-
-\
-
-In short, existing autonomous exploration approaches are effective for eliminating the need for humans to control the robot.
-However, optimality remains quite difficult, so these approaches apply suboptimal solutions to achieve autonomous motions.
-
-% JINT17: However, there are certain issues with the current SLAM techniques.
-First, the vast majority of work deals only with the aspect of estimating the environment and the poses of vehicles.
-They are passive in the sense that SLAM is per- formed on incoming sensor measurements from vehicles following an arbitrary path.
-As such, human teleopera- tion and monitoring are often essential to guide the vehi- cles safely through unknown surroundings.
-Therefore, it is desired that an active motion planning scheme is devel- oped and integrated with SLAM such that vehicles are able to determine their path without human intervention, and explore unknown areas autonomously.
-
-\subsection{Multi-Vehicle Cooperation}
-
-\section{Outline of Dissertation}
-
-ADD ONCE COMPLETE
-
-\section{Contributions}
-
-There are three key contributions in this dissertation.
-The first two are novel solutions that find exact occupancy probabilities and predicted entropies, respectively.
-These contributions improve the quality and efficiency of autonomous mapping and exploration of uncertain environments for single-vehicle missions.
-The third contribution extends these concepts to multi-vehicle scenarios, where multiple members of a team work together to autonomously explore and patrol large environments.
-
-\subsection{Summary of Contributions}
-
-The first contribution is an exact Bayesian solution to occupancy grid mapping.
-The proposed approach considers that occupied spaces occlude measurement behind them, which allows certain mapping outcomes to be systematically grouped together.
-The exact solution to occupancy grid cell probability, which was previously-believed to have exponential complexity, is now possible to solve in real-time.
-Numerous numerical simulations and experiments demonstrate that the proposed mapping approach is effective and inexpensive in 2D and 3D scenarios.
-
-% JINT17: This paper proposes a computationally efficient algo- rithm to construct the exact inverse sensor model.
-More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, this algo- rithm yields the a posteriori probabilities of occupancy of all the cells within the area covered by the range sensor from the range measurements.
-The key idea is reducing the computational load by using the fact that if a cell is occu- pied, the occupancies of the other cells blocked by it is irrelevant to the forward sensor model, and this property is systematically utilized with various probabilistic prop- erties to derive a computationally-efficient solution to the inverse sensor model.
-Furthermore, the proposed approach integrates a priori probabilities of occupancy and multiple range measurements according to the Bayesian framework to obtain more accurate maps.
-As such, it contrasts from the existing framework based on log-odds ratios that impose an additional Markov assumption.
-
-The second contribution is a new approach to autonomous exploration, where the robot predicts future occupancy grid map uncertainty, using a measure known as Shannon's entropy.
-This approach computes the expected values of grid cell entropies from potential future measurement rays with a novel approach to predict future measurement probabilities.
-These predictions are used to evaluate the benefit of possible robot actions to maximize map information gain.
-Then, we formulate autonomous exploration as an optimization problem to maximize an objective function that includes map uncertainty and travel costs.
-The probabilistic map is used for collision-avoidance and motion planning.
-These processes are designed for real-time applications, where the robot makes decisions as it learns information about the surrounding environment.
-
-The final contribution coordinates the mapping and exploration efforts of multiple robots together.
-The mapping process is multi-threaded such that multiple members update the same occupancy grid.
-The exploration algorithm follows a bidding-based structure.
-Robots compete for tasks in a series of auctions, where the winning bid of one auction affects the bids of subsequent auctions, thereby coordinating the robot efforts.
-This cooperative approach is computationally-efficient in large environments, and follows a receding-horizon framework, such that map expected information gains are updated as quickly as possible.
-The proposed coordination is extended to autonomous patrol, where occupancy grid map probabilities are degraded over time, which promotes the robots to revisit areas periodically.
-
-\subsection{List of Publications}
-
-The following is a list of publications. These include research included in this dissertation, as well as contributions in control, data association, and estimation.
-
-% TODO: update this list
-
-\begin{itemize}
-	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Autonomous Exploration with Exact Inverse Sensor Models,'' \textit{Journal of Intelligent }\&\textit{ Robotic Systems}, 2016, submitted.
-	\item E.
-Kaufman, T.
-Lee, and Z.
-Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, San Francisco, December 2016, accepted.
-	\item E.
-Kaufman, T.
-Lee, Z.
-Ai, and I.
-S.
-Moskowitz, ``Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model,'' \textit{Proceedings of the American Control Conference}, Boston, July 2016, pp.
-5709-5715.
-	\item E.
-Kaufman, T.
-A.
-Lovell, and T.
-Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{The Journal of the Astronautical Sciences}, 63(1): pp.
-60-80, 2016, doi: 10.1007/s40295-015-0082-9.
-	\item E.
-Kaufman, T.
-A.
-Lovell, and T.
-Lee, ``Minimum Uncertainty JPDA Filters and Coalescence Avoidance for Multiple Object Tracking,'' \textit{The Journal of the Astronautical Sciences}, 63(4): pp.
-308-334, 2016, doi: 10.1007/s40295-016-0092-2.
-	\item T.
-Wu, E.
-Kaufman, and T.
-Lee, ``Globally Asymptotically Stable Attitude Observer on SO(3),'' \textit{Proceedings of the 54th IEEE Conference on Decision and Control}, pp.
-2164-2168, Osaka, Japan, December 2015.
-	\item E.
-Kaufman, T.
-A.
-Lovell, and T.
-Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting}, Williamsburg, VA, Jan.
-2015, AAS 15-451.
-	\item E.
-Kaufman, T.
-A.
-Lovell, and T.
-Lee, ``Minimum Uncertainty JPDA Filter and Coalescence Avoidance Performance Evaluations,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting, Williamsburg}, VA, Jan.
-2015, AAS 15-432.
-	\item E.
-Kaufman, K.
-Caldwell, D.
-Lee, and T.
-Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, Mar.
-2014, ASC 14-2527.
-	\item E.
-Kaufman, T.
-A.
-Lovell, and T.
-Lee, ``Optimal Joint Probabilistic Data Association Filter Avoiding Coalescence in Close Proximity,'' \textit{Proceedings of the European Control Conference}, pp.
-2709-2714, June 2014.
-	\vspace*{0.5cm}
-\end{itemize}
-
+\include{tex/intro}
 \include{tex/pogm}

### 2018-06-23 Dissertation/Thesis/tex/intro.tex
I think I deleted everyting
fixed the delete issue! it works

@@
 % !TEX root = ../thesis-sample.tex
-% COMMENTED OUT DURING EDITS
-%\chapter{Introduction} \label{chap:intro}
-%
-%% Add here
-%
-%\section{Motivation and Goals}
-%
-%\subsection{Building Probabilistic Maps}
-%
-%\subsection{Exploration in Uncertain Spaces}
-%
-%\subsection{Cooperation Among Autonomous Systems}
-%
-%\section{Literature Review}
-%
-%\subsection{Occupancy Grid Mapping}
-%
-%\subsection{Autonomous Exploration}
-%
-%\subsection{Multi-Vehicle Cooperation}
-%
-%\section{Outline of Dissertation}
-%
-%\section{Contributions}
-%\subsection{Summary of Contributions}
-%\subsection{List of Publications}
+
+\chapter{Introduction} \label{chap:intro}
+
+This dissertation explores three key aspects of mobile robotics: occupancy grid mapping, autonomous exploration, and cooperation among robot teams.
+Several novel contributions are presented on these topics for complex and dynamic situations.
+
+\section{Motivation and Goals}
+
+The study of robotics has expanded tremendously in recent years as robots have replaced humans for dangerous, difficult, and repetitive tasks.
+Certain objectives, such as search-and-rescue, surveillance, and cleaning, require some level of autonomy, particularly with mobile robots exploring new and uncertain environments.
+While these tasks provide significant value to our society, they present nontrivial challenges for the robots, particularly in mapping, exploration, and cooperation.
+
+\subsection{Building Probabilistic Maps}
+
+The first topic this dissertation addresses is building accurate probabilistic maps, known as \emph{mapping}.
+Map generation is crucial for simultaneous localization and mapping (SLAM) because maps provide important information for the robot to determine its pose (localization), which includes its position and attitude.
+Mapping is also an integral part of vital mobile robot objectives, such as collision avoidance and trajectory planning.
+In short, mapping is of fundamental importance in mobile robotics.
+
+Therefore, the map \emph{quality} deserves much attention.
+Robotics is inherently \emph{probabilistic}; no sensor reading or action from an actuator is deterministic.
+All robotic processes have some degree of uncertainty, and the mapping must account for this.
+Onboard sensors serve to improve the understanding of surrounding spaces, but the stochastic properties of those sensors must be reflected in the stochastic properties of the map.
+Therefore, we propose to develop a probabilistic map that accounts for the history of measurements and poses, and their associated stochastic properties.
+Here, each element of the map holds a probability based on measured data.
+In turn, this probability can be applied to tasks such as collision avoidance and predicting future map outcomes.
+
+\subsection{Exploration in Uncertain Spaces}
+
+The second topic of this dissertation focusses on the motions robots must choose to maximize their knowledge about the map while avoiding collisions, known as \emph{autonomous exploration}.
+Unlike conventional motion planning, where the map of the environment is assumed known, the robot must select motions with only the current knowledge of the environment.
+Therefore, real-time implementation of an active motion planning scheme is necessary while the probabilistic map is changing.
+
+There are several important aspects to consider in autonomous exploration.
+Most importantly, the robot motion must be constrained to collision-free paths.
+The probabilistic map provides collision-free regions for this objective, and it also serves as a graph to determine an ordered set of collision-free waypoints.
+Additionally, the uncertainty of the map must be considered in the policy governing robotic motion.
+The method of decreasing map uncertainty can be computationally-expensive, so careful consideration with respect to algorithm complexity is vital.
+Furthermore, the efficiency of robot motions should be considered to avoid unnecessary travel distances.
+These objectives must all be achieved for an effective autonomous exploration policy.
+
+%We proposed an autonomous exploration approach based on maximizing map information gain.
+Shannon's entropy serves as measure of map uncertainty to be minimized.
+This exploration scheme is formulated as a constrained optimization problem, where the robot predicts how much information will be acquired from various possible future actions, while also considering travel distance.
+
+\subsection{Cooperation Among Autonomous Systems}
+
+The final topic of this dissertation is cooperation among members of a robot team.
+All robots share the mutual objective of learning the map as quickly as possible.
+Each robot must select actions that consider the other team members.
+These considerations involve avoiding collisions among members, and avoiding actions that provide too much coverage of one region of the map while another region remains uncertain.
+Much like single-vehicle autonomous exploration, the multi-vehicle problem requires a computationally-inexpensive algorithm to handle numerous possible robot motions.
+These nontrivial goals can be handled by extending single-vehicle autonomous exploration with a bidding-based approach.
+
+Multi-vehicle autonomous exploration can be extended further for surveillance purposes with autonomous patrol.
+The goal is to periodically capture many regions within a large space.
+Frequently, the robots are subject to dynamic obstacles (e.g.
+a walking person) or a changing environment (e.g.
+moving a chair), so the robots must account for these as well.
+Furthermore, an effective patrol policy can decrease the number of robots required to patrol an environment.
+The multi-vehicle autonomous exploration can be integrated with probabilistic map degradation to achieve autonomous patrol.
+
+\section{Literature Review}
+
+In this section, we discuss the existing approaches to occupancy grid mapping, autonomous exploration, and multi-vehicle cooperation.
+Several shortcomings with existing approaches are highlighted, which are addressed in the dissertation contributions.
+
+\subsection{Occupancy Grid Mapping}
+
+Here we describe the existing approaches to building probabilistic occupancy grid maps.
+Though several variations have been proposed, no approach uses the stochastic properties of depth sensors directly to obtain an exact Bayesian solution that can be computed in real-time.
+
+\paragraph{Inverse Sensor Model} The key to generating a probabilistic occupancy grid map is known as the \emph{inverse sensor model}~\cite{ThrBurFox05}.
+The main idea is that a depth sensor has a probability density (e.g.
+Gaussian) that describes the stochastic properties of a sensor reading (e.g.
+a LIDAR range) with respect to the expected value of a reading (e.g.
+distance to a wall).
+This probability density is referred to as the \emph{forward sensor model}, as it relates a depth measurement conditioned on the occupancy of the map.
+
+In contrast, the inverse sensor model relates the occupancy of the map conditioned on the depth measurement.
+This problem is solved using Bayes' theorem.
+However, the occupancy grid map properties yield a complicated Bayesian probability; the computational cost has exponential complexity with respect to the number of grid cells captured within the depth sensor limits.
+It was previously-believed that this restriction makes solving the inverse sensor model in real-time impossible.
+This motivated approximate or estimated solutions, described next.
+
+\paragraph{Approximated Function for the Inverse Sensor Model} Moravec and Elfes originally proposed a grid cell-based mapping for sonar sensors in~\cite{MorElf85}, and the probabilistic properties of the occupancy grid were formalized in~\cite{MorElf85,Elf89}.
+This approach models measurements with Gaussian probability density, although the occupancy of grid cells is not exactly known.
+Instead, the probability of grid cells are estimated with a Gaussian-like \emph{approximate function}.
+The authors show that occupancy probability converges to a final value following a Bayesian framework, though there is no conclusive proof that this probability is accurate.
+Nevertheless, this approach garnered much attention because it could efficiently produce robotic maps that nicely-resembled the occupancy of spaces around a robot.
+
+Several other research studies have applied approximate functions in place of the inverse sensor model.
+For example, \cite{ChoLynHutKanBurKavThr05} proposed a simplified continuous function composed of lines patched together near a measurement reading.
+Furthermore, variations of approximate smoother functions in~\cite{and09,PirRutBisSch11,KhoElb12} contain Gaussian-like terms to apply occupancy grid mapping to 3D environments with more modern sensors.
+Numerous research studies have used the proposed approximate functions with questionable accuracy.
+
+\paragraph{Learning the Inverse Sensor Model} The approximate nature of the inverse sensor model motivated a solution using machine learning.
+The main idea is to simulate maps, robot poses, and measurements with simulated noised based on the sensor stochastic properties.
+Then, an approximation the inverse sensor model is learned using an expectation maximization algorithm~\cite{Thr01,ThrBurFox05}.
+Unlike approximate functions based on intuition, the machine learning approach is based on the physical properties of sensors.
+However, this approach is undesirable in practice due to complexities associated to implementing a learning algorithm.
+For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples, or how many samples are required to obtain a reasonable approximation.
+Furthermore, it is challenging to apply any learning algorithm over the large dimensional space composed of maps, poses, and measurements.
+
+\paragraph{Log-Odds Ratio} Both approaches described above use the log-odds ratio to update grid cell probabilities.
+The log-odds ratio formulation is a popular framework for updating binary random variables with static state within a Bayesian filter.
+The main idea is that instead of multiplying terms from prior and current time steps, the properties of logarithms allow these terms to be simply added, while avoiding some truncation issues associated with probabilities close to $0$ or $1$~\cite{ThrBurFox05}.
+This approach is also popular because the inverse sensor model needs not consider prior probabilities.
+Instead, these approaches typically consider fixed uniform initial probabilities of all grid cells, which makes the formulation simpler.
+
+However, the log-odds ratio formulation makes an assumption that is not consistent with the occupancy grid mapping problem.
+The approach yields a simplified solution by assuming that the probability of a measurement is independent of past measurements and robot poses.
+This assumption is frequently violated when past measurements and poses indicate the occupancy of other cells between the robot and the cell in question.
+As such, this approach neglects potentially important information when considering grid cell occupancy probability.
+
+In short, other existing solutions to occupancy probability are inexact, and involve a potentially-harmful log-odds ratio assumption.
+These shortcomings motivate an exact solution to occupancy grid mapping that can be computed in real-time.
+
+%It was previously-believed that the exact solution to occupancy grid mapping was impossible in real-time.
+
+\subsection{Autonomous Exploration}
+
+The next topic this dissertation addresses is autonomous exploration.
+The vast majority of work in simultaneous localization and mapping (SLAM) deals only with the aspect of estimating the environment and the poses of vehicles.
+These approaches are passive in the sense that SLAM is performed on incoming sensor measurements from vehicles following an arbitrary path.
+As such, human teleoperation and monitoring are often essential to guide the vehicles safely through unknown surroundings.
+Therefore, it is desirable that an active motion planning scheme is developed and integrated with SLAM such that vehicles are able to determine their path without human intervention, and explore unknown areas autonomously.
+The two approaches used to solve this problem are described next.
+
+\paragraph{Frontier-Based Autonomous Exploration}
+
+The most popular approach to solve the autonomous exploration problem is known as frontier-based exploration, originally-proposed in~\cite{Yam97,Yam98} for 2D applications, and was later extended to 3D with a visibility metric in~\cite{SawKriSri09} and~\cite{ZhuDinLinWu15,SenWan16,KleDor13} using the popular Octomap representation (a variation on occupancy grid mapping).
+The main idea is that robots toward the border between certain and uncertain space, referred to as a frontier.
+Then the robot takes depth measurement at this frontier, thus pushing back the boundary.
+This process is repeated until the map is well-known.
+Frontier-based exploration assumes that repeatedly moving toward the closest frontier and taking measurements are the best actions to gain new information about the map.
+However, these systematic actions, based on ad-hoc rules for what constitutes a frontier, do not consider the future uncertainty of the probabilistic map.
+Frontier-based approaches provide an intuitive solution, but lack optimality in map uncertainty or exploration time.
+
+\paragraph{Entropy-Based Autonomous Exploration}
+Entropy-based approaches address the suboptimal nature of frontier-based approaches by selecting robotic actions designed to decrease minimize map uncertainty~\cite{StaGriBur05,BurMooStaSch05}.
+These approaches use a measure of uncertainty known as Shannon's entropy, which becomes smaller as cell probabilities approach $0$ or $1$ (becoming more certain).
+Existing entropy-based approaches simulate possible future measurements from various locations with so-called ``hallucination measurements,'' which typically corresponds to expected depth measurements from the robot to the closest grid cell that is possibly occupied.
+Then, the robot analyzes how this measurement would affect the map uncertainty.
+However, this approach assumes that expected map entropy is equivalent to map entropy from expected measurements.
+This is not the case in general, however, as there is typically a nonzero probability that the hallucination measurement and true future measurement are not the same.
+Entropy-based approaches seek to optimize map uncertainty, but require large approximations which can decrease the accuracy of predictions.
+
+In conclusion, existing autonomous exploration approaches are effective for eliminating the need for humans to control the robot.
+However, optimality remains quite difficult, so these approaches apply suboptimal or approximated solutions to achieve autonomous motions.
+
+\subsection{Multi-Vehicle Cooperation}
+
+The final topic relates to coordinating robotic efforts to build occupancy grid maps together.
+These concepts are further extended to autonomous cooperative patrol, where robots periodically revisit regions to monitor a large environment.
+
+\paragraph{Cooperative Autonomous Exploration}
+
+Multi-vehicle cooperation is essential for robots to explore large spaces autonomously so that robots coordinate their efforts to improve map information and avoid collisions.
+However, optimizing map coverage simultaneously from multiple vehicles becomes very complicated and computationally intractable.
+Instead, an auction-based approach addresses the computational restrictions effectively.
+Bidding-based approaches were introduced with~\cite{SimApfBurFoxMooThrYou00,DiaSte00}, and extended to several other robotic problems~\cite{GerMaj02,ZloSteDiaTha02,SarBal05,ChoBruHow09} exploring decentralization and task allocation with varying robot capabilities.
+In~\cite{SimApfBurFoxMooThrYou00}, a central executive assigns tasks to robots for maximizing overall utility of the group.
+This is accomplished by minimizing a coverage overlap among the members.
+Robots submit bids based on expected coverage of uncertain spaces and travel distance.
+The robot winning the auction is awarded the task, and the remaining bids are discounted to prevent robots from covering the same area, thereby coordinating the mapping efforts.
+Bidding-based approaches simplify multi-vehicle cooperation in a computationally-efficient manner, and integrate effectively with the autonomous mapping and exploration approaches proposed in this dissertation.
+
+\paragraph{Autonomous Patrol}
+
+Robotic surveillance and cleaning are examples of tasks that require periodic observations of the same regions.
+Several patrolling approaches have been proposed in recent years to address this need.
+A popular approach is decomposing the reachable space into Voronoi diagrams~\cite{KolCar08,PorRoc10,PipChrWei13}.
+In~\cite{KolCar08}, edges between nodes of the Voronoi graph are heavily sensed to track intruders.
+In~\cite{PorRoc10}, regions of the Voronoi diagram are assigned to different robots, and issues with robots having underperforming surveillance capabilities are addressed in~\cite{PipChrWei13}.
+These approaches assume the map is well-known in advance, and require a Voronoi graph be generated from the map.
+
+Other patrol approaches create new metrics about the environment.
+Cell periodicity is estimated in~\cite{KraFenCieDonDuc14,KraFenHanDuc16} by predicting and observing the long-term dynamics of objects at the same locations on an occupancy grid.
+This work is extended to exploring and navigating a human-populated environment for long-term autonomy.
+A bio-inspired approach is proposed in~\cite{ZhaXia11}, where robots cooperate by dropping digital pheromones based on the importance of events, where the pheromones weaken over time to prioritize future tasks.
+Another approach is formulating patrol as an optimization problem with respect to robot idleness in~\cite{YanZha16}.
+The prior work in autonomous patrol demonstrates that periodic actions can be incentivized with a metric to be optimized.
+
+\section{Outline of Dissertation}
+
+ADD ONCE COMPLETE
+
+\section{Contributions}
+
+There are three key contributions in this dissertation.
+The first two are novel solutions that find exact occupancy probabilities and predicted entropies, respectively.
+These contributions improve the quality and efficiency of autonomous mapping and exploration of uncertain environments for single-vehicle missions.
+The third contribution extends these concepts to multi-vehicle scenarios, where multiple members of a team work together to autonomously explore and patrol large environments.
+
+\subsection{Summary of Contributions}
+
+The first contribution is an exact Bayesian solution to occupancy grid mapping.
+The proposed approach considers that occupied spaces occlude measurement behind them, which allows certain mapping outcomes to be systematically grouped together.
+The exact solution to occupancy grid cell probability, which was previously-believed to have exponential complexity, is now possible to solve in real-time.
+Numerous numerical simulations and experiments demonstrate that the proposed mapping approach is effective and inexpensive in 2D and 3D scenarios.
+
+% JINT17: This paper proposes a computationally efficient algo- rithm to construct the exact inverse sensor model.
+More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, this algo- rithm yields the a posteriori probabilities of occupancy of all the cells within the area covered by the range sensor from the range measurements.
+The key idea is reducing the computational load by using the fact that if a cell is occu- pied, the occupancies of the other cells blocked by it is irrelevant to the forward sensor model, and this property is systematically utilized with various probabilistic prop- erties to derive a computationally-efficient solution to the inverse sensor model.
+Furthermore, the proposed approach integrates a priori probabilities of occupancy and multiple range measurements according to the Bayesian framework to obtain more accurate maps.
+As such, it contrasts from the existing framework based on log-odds ratios that impose an additional Markov assumption.
+
+The second contribution is a new approach to autonomous exploration, where the robot predicts future occupancy grid map uncertainty, using a measure known as Shannon's entropy.
+This approach computes the expected values of grid cell entropies from potential future measurement rays with a novel approach to predict future measurement probabilities.
+These predictions are used to evaluate the benefit of possible robot actions to maximize map information gain.
+Then, we formulate autonomous exploration as an optimization problem to maximize an objective function that includes map uncertainty and travel costs.
+The probabilistic map is used for collision-avoidance and motion planning.
+These processes are designed for real-time applications, where the robot makes decisions as it learns information about the surrounding environment.
+
+The final contribution coordinates the mapping and exploration efforts of multiple robots together.
+The mapping process is multi-threaded such that multiple members update the same occupancy grid.
+The exploration algorithm follows a bidding-based structure.
+Robots compete for tasks in a series of auctions, where the winning bid of one auction affects the bids of subsequent auctions, thereby coordinating the robot efforts.
+This cooperative approach is computationally-efficient in large environments, and follows a receding-horizon framework, such that map expected information gains are updated as quickly as possible.
+The proposed coordination is extended to autonomous patrol, where occupancy grid map probabilities are degraded over time, which promotes the robots to revisit areas periodically.
+
+\subsection{List of Publications}
+
+The following is a list of publications. These include research included in this dissertation, as well as contributions in control, data association, and estimation.
+
+% TODO: update this list
+
+\begin{itemize}
+	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Autonomous Exploration with Exact Inverse Sensor Models,'' \textit{Journal of Intelligent }\&\textit{ Robotic Systems}, 2016, submitted.
+	\item E.
+Kaufman, T.
+Lee, and Z.
+Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, San Francisco, December 2016, accepted.
+	\item E.
+Kaufman, T.
+Lee, Z.
+Ai, and I.
+S.
+Moskowitz, ``Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model,'' \textit{Proceedings of the American Control Conference}, Boston, July 2016, pp.
+5709-5715.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{The Journal of the Astronautical Sciences}, 63(1): pp.
+60-80, 2016, doi: 10.1007/s40295-015-0082-9.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Minimum Uncertainty JPDA Filters and Coalescence Avoidance for Multiple Object Tracking,'' \textit{The Journal of the Astronautical Sciences}, 63(4): pp.
+308-334, 2016, doi: 10.1007/s40295-016-0092-2.
+	\item T.
+Wu, E.
+Kaufman, and T.
+Lee, ``Globally Asymptotically Stable Attitude Observer on SO(3),'' \textit{Proceedings of the 54th IEEE Conference on Decision and Control}, pp.
+2164-2168, Osaka, Japan, December 2015.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting}, Williamsburg, VA, Jan.
+2015, AAS 15-451.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Minimum Uncertainty JPDA Filter and Coalescence Avoidance Performance Evaluations,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting, Williamsburg}, VA, Jan.
+2015, AAS 15-432.
+	\item E.
+Kaufman, K.
+Caldwell, D.
+Lee, and T.
+Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, Mar.
+2014, ASC 14-2527.
+	\item E.
+Kaufman, T.
+A.
+Lovell, and T.
+Lee, ``Optimal Joint Probabilistic Data Association Filter Avoiding Coalescence in Close Proximity,'' \textit{Proceedings of the European Control Conference}, pp.
+2709-2714, June 2014.
+	\vspace*{0.5cm}
+\end{itemize}

### 2018-06-23 Dissertation/Thesis/tex/pogm.tex
started chapter 2: mapping

% !TEX root = ../thesis-sample.tex

\chapter{Probabilistic Occupancy Grid Mapping} \label{chap:pogm}

In this chapter, we present the exact solution to occupancy grid map probability. The Bayesian approach uses the stochastic properties of sensors, and exploits important patterns in conditional probabilities. This solution is simulated in several examples to show the efficacy of the approach.

\section{Mapping Problem Definition}

\section{Inverse Sensor Model}

\subsection{Bayesian Probability}

\subsection{Computationally-Efficient Solution}
% Include Fig 1 from JINT 2017

\section{Multi-Dimensional Maps}

\subsection{Ray Casting}

\subsection{Combining Measurements from a Single Scan}

Next, we describe two approaches to combine the solutions to the inverse sensor model for a single measurement ray...

\paragraph{Ray-By-Ray Approach}

\paragraph{Synergistic Update Approach}

\subsection{Multi-Sensor Fusion}

\subsection{Practical Implications}

\section{Numerical and Experimental Examples}

\subsection{Mapping a Room}
% ACC16 comparison of exact/approx. inverse sensor models

\subsection{Kinect Measurement Scan Analysis}
% ACC16 Single scan exact/approx. comparison

\section{Conclusions}

### 2018-06-23 Dissertation/Thesis/tex/pogm.tex
added some materials from JINT18

@@
 
+Let a map $m$ be decomposed into $n_m$ evenly-spaced grid cells, where the $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,\ldots,n_m}$, that is defined as $\mathbf{m}_i=1$ when occupied, and $\mathbf{m}_i=0$ when free.
+The location and size of each grid cell is assumed known, where a smaller cell size (greater grid resolution) better represents a space, but increases computation and memory.
+Therefore, a map $m$ is defined by $\{\mathbf{m}_1,\ldots, \mathbf{m}_{n_m}\}$ ($2^{n_{m}}$ possible maps).
+
+Another random variable is defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience.
+The probability that the $i$-th cell is occupied is $P(\mathbf{m}_i)$, and the probability that it is free is $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$.
+The random variables $\mathbf{m}_i$ are mutually independent,
+\begin{align}
+P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i).
+\end{align}
+
+Consider a range sensor that provides scans of the surrounding environment in order to identify the closest occupied space.
+The location and the direction of the sensor, referred to as the \emph{pose} at time $t$, is denoted by $X_t=(x_t,R_t)$, where the planar position and direction of the robot at $t$ are denoted by $x_t\in\Re^2$ and $R_t\in\Sph^1=\braces{q\in\Re^2|\norm{q}=1}$, i.e., the attitude corresponds to a direction on a 2D plane.
+Let $X_{1:t}$ denote the history of poses from the initial time to the current time, i.e., $X_{1:t}=\{X_1,X_2,\ldots, X_t\}$.
+At each pose, the robot receives a 2D measurement \emph{scan}, composed of $n_z$ measurement \emph{rays}.
+These rays are 1D depth measurements of known direction from the current pose to the closest occupied space, subject to a known \emph{forward sensor model} $p(z_{t,l}|m,X_t)$, where $z_{t,l}$ is the $l$-th measurement ray of the $t$-th scan $Z_t=\braces{z_{t,1},z_{t,2},\ldots,z_{t,n_z}}$.
+The measurement history is denoted by $Z_{1:t}=\braces{Z_1,Z_2,\ldots,Z_t}$.
+
+The probability density function, namely $p(z_{t,l}|m,X_t)$ with respect to the depth of the $l$-th measurement ray conditioned on the map $m$ and the pose $X_t$ is commonly referred to as the \emph{forward sensor model}, which characterizes the corresponding depth sensor, such as the maximum range or accuracy.
+The forward sensor model satisfies (i) the ranges of all depth measurements are positive and finite, and (ii) measurement rays cannot pass through occupied regions.
+Throughout this paper, we assume that the forward sensor model of the selected sensor is given.
+This can be determined empirically or analytically.
+For example, the \emph{beam model for range finders} satisfying  the above criteria is described in~\cite{ThrBurFox05}.
+
+Occupancy grid mapping provides occupancy probabilities based on robot poses and measurement scans.
+The goal is to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, commonly referred to as the \emph{inverse sensor model} for given forward sensor model $p(z_{t,l}|m,X_t)$ and the initial estimate of the map $P(m)$ with $X_{1:t}$ and $Z_{1:t}$.
+
 \section{Inverse Sensor Model}
 
-\subsection{Bayesian Probability}
+In this section, we propose an algorithm to compute the exact inverse sensor model efficiently. 
+Suppose the probability of the map conditioned on the past poses and measurements, namely $P(m|X_{1:t-1},Z_{1:t-1})$, is known.
+Here we construct a posteriori probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$, based on the current pose $X_t$, the measurement from the $l$-th ray $z_{t,l}$, and the given forward sensor model $p(z_{t_l}|m,X_t)$.
+
+\subsection{Bayesian Framework}
+
+The occupancy probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$ is based on the forward sensor model $p(z_{t,l}|m,X_{1:t},Z_{1:t-1})$ that describe the distribution of the measurements for the given robot pose and the map.
+This specifies the stochastic characteristics of the sensor, such as the mean squared error, and it is assumed as a known distribution, specific to a particular depth sensor.
+One could certainly obtain a stochastic model by fitting a probability density to a controlled set of measurements.
+Bayes' rule yields
+\begin{align}
+\label{eqn:BayesRuleRayISM}
+P(&m|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
+=\frac{p(z_{t,l}|m,X_{1:t},Z_{1:t-1})P(m|X_{1:t-1},Z_{1:t-1})}{p(z_{t,l}|X_{1:t},Z_{1:t-1})},
+\end{align}
+considering that $X_t$ carries no information about $m$ without $Z_t$.
+If the current pose $X_t$ and map $m$ are known, then the measurement ray $z_{t,l}$ is independent of past poses $X_{1:t-1}$ and past measurements $Z_{1:t-1}$ to obtain
+\begin{align}
+P(m|z_{t,l},&X_{1:t},Z_{1:t-1})%\nonumber\\&
+=\eta_{t,l}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}),
+\end{align}
+where the normalizing constant $\eta_{t,l}\in\Re$ absorbs all terms that do not depend on the map $m$.
+Next, we compute the occupancy probability of each cell.
+Let $\mathcal{M}_i$ be the set of maps where the $i$-th cell is occupied, i.e., $\mathcal{M}_i =\{m\in\{0,1\}^{{n_m}}\,|\ \mathbf{m}_i=1\}$.
+To compute the probability of occupancy of the $i$-th cell, all possible combinations of map in $\mathcal{M}_i$ should be considered, i.e.,
+\begin{align}
+\label{eqn:InvSenModWithProbDens}
+&P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
+=\eta_{t,l}\sum_{m\in\mathcal{M}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
+\end{align}
+Furthermore, to determine the normalizing constant $\eta_{t,l}$, the complement $\\P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in the similar manner.
+Since there are $2^{n_{m}-1}$ maps in $\mathcal{M}_i$, these equations require $2^{n_m}$ terms to compute the summation for the $i$-th grid cell and normalizer, and this process should be repeated for all other cells along the measurement ray.
+This is the main reason why the existing results are based on certain approximations of the above expression.
 
@@
 
+We propose a computational algorithm to evaluate \refeqn{InvSenModWithProbDens} efficiently. 
+Since the cells outside of the sensor field of view (FOV) are not affected, we focus on a reduced map $r_l$ in the FOV of the $l$-th ray.
+This reduced map is chosen such that each cell of $r_l$ corresponds to a grid cell of map $m$ that the $l$-th ray intersects, ordered by increasing distance, which is easily determined from geometry.
+Let $\mathbf{r}_{l,k}$ be the binary random variable representing the occupancy of the $k$-th cell of the $l$-th ray.
+The number of cells in the reduced map is denoted by $n_{r,l}\leq n_m$.
+
+Next, let $\mathbf{r}_{l,k+}$ correspond to the event that the $k$-th cell of the $l$-th ray is occupied, cells with lower index (closer cells) are free, and cells with greater index (farther cells) may or may not be occupied, i.e., event $\mathbf{r}_{l,k+}$ occurs when \\$r\in\braces{r\in\{0,1\}^{{n_{r,l}}}\,|\mathbf{r}_{l,1}=0,\mathbf{r}_{l,2}=0,\ldots,\mathbf{r}_{l,k-1}=0,\mathbf{r}_{l,k}=1}$.
+In other words, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray.
+For example, see Figure \ref{fig:show_rkplus} that illustrates the event of $\mathbf{r}_{l,k+}$ when $l=1$ and $k=4$ for an one-dimensional cell array.
+This concept of grouping map outcomes can be easily extended to 2D using ray casting, where a 1D ray is spanning 2D space.
+The cells considered for the measurement ray are determined with the geometric intersection between the ray and map cell edges.
+Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell.
+Based on this, we present the mathematical expressions for the exact inverse sensor model, as summarized by \refeqn{RayISMAnswer} in Proposition 1.
+Later in Section \ref{sec:SA}, these are also rearranged as a computational algorithm.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.4\textwidth]{rkplus_1.png}
+    \centering
+  \includegraphics[width=0.4\textwidth]{rkplus_2.png}  
+  \centering
+  \includegraphics[width=0.4\textwidth]{rkplus_3.png}
+    \centering
+  \includegraphics[width=0.4\textwidth]{rkplus_4.png}
+  \caption{A robot located to the left of a 1D occupancy grid map $r_l$ composed of $n_{r,l}=6$ grid cells in $4$ cases.
+In each case, cells $1$--$3$ are free, cell $4$ is occupied, and cells $5$ and $6$ may or may not be occupied.
+All above outcomes correspond to the event $\mathbf{r}_{l,4+}$.}
+  \label{fig:show_rkplus}
+\end{figure}
+
+\begin{prop}
+For the $l$-th measurement ray, the a posteriori probability of the occupancy of the $k$-th cell, namely the ray inverse sensor model, is given by
+\begin{align}
+\label{eqn:RayISMAnswer}
+P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}),
+\end{align}
+where the unnormalized probability of the inverse sensor model is defined as
+\begin{align}
+\label{eqn:Unnormalized}
+& \tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
+=P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\
+&\quad\times 
+\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
+%&\quad\times 
+p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]\nonumber\\
+&\quad + \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
+%&\quad\times 
+p(z_{t,l}|\mathbf{r}_{l,k+},X_t)P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1}),
+\end{align}
+where $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ is chosen for convenience and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the probability density of the measurement when all of cells in the field of view of the $l$-th ray is not occupied.
+The normalizer $\eta_{t,l}$ is given by
+\begin{align}
+\label{eqn:allEta}
+\eta_{t,l}
+&=
+\bigg[\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
+\end{align}
+and it is independent of the cell index $k$.
+\end{prop}
+\begin{proof}
+See Appendix.
+\end{proof}
+
+Note that the a priori estimate, $P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ and its compliment $\\P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})=1-P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ are available at the $t$-th step.
+Then, \refeqn{RayISMAnswer} yields a sequential occupancy grid mapping that can be applied whenever new measurements are available.
+
+% TODO: add update paragraph from JINT18
+
+%Compared with \refeqn{InvSenModWithProbDens}, where the terms of the summation should be repeated $2^{n_{r,l}}$ times \emph{per each cell of the reduced map}, the proposed expressions \refeqn{RayISMAnswer} and \refeqn{allEta} are \textit{substantially} simpler.
+In fact, if \refeqn{Unnormalized} and \refeqn{allEta} are obtained recursively, requiring the summation of $O(n_{r,l}+1)$ rather than $O(n_{r,l}\times2^{n_{r,l}})$ as previously thought for \emph{all cells of the reduced map}, this yields an algorithm that is $\mathbf{n_{r,l}\times2^{n_{r,l}}/(n_{r,l}+1)}$ \textbf{times faster}.
+
 \section{Multi-Dimensional Maps}
+
+The above formulations describe how a single ray updates grid cells along its 1D path.
+Next we describe how measurement rays can update 2D maps, and how large scans of measurement rays are handled together.
 

### 2018-06-23 Dissertation/Thesis/tex/pogm.tex
imported an explanation on map comp time from JINT18

@@
 
-Note that the a priori estimate, $P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ and its compliment $\\P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})=1-P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ are available at the $t$-th step.
+%Note that the a priori estimate, $P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ and its compliment $\\P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})=1-P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ are available at the $t$-th step.
 Then, \refeqn{RayISMAnswer} yields a sequential occupancy grid mapping that can be applied whenever new measurements are available.
+
+%where a priori probability is $\mathbf{P}_k^-=P(\mathbf{r}_{k}|X_{1:t-1},Z_{1:t-1})$, its complement is $\bar{\mathbf{P}}_k^-=1-\mathbf{P}_k^-$, $P(\bar{\mathbf{r}}_{0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{n_{r,l}+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience, and $p(z_{t,l}|\mathbf{r}_{(n_{r,l}+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading.
+%The proofs of \refeqn{RayISMAnswer}--\refeqn{Unnormalized} are given in~\cite{KauLeeAiMos16}.
+
+Since $\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ from \refeqn{Unnormalized} uses several repeated terms from the prior $\tilde P(\mathbf{r}_{l,k-1}|z_{t,l},X_{1:t},Z_{1:t-1})$, and $\eta_{t,l}$ is easily obtained from these as well, the computational cost of \refeqn{RayISMAnswer} is linear with respect to the number of cells along a measurement ray, amortized to $\mathcal{O}(1)$ for each cell, so the exact inverse sensor model can be applied in real-time.
+The process of updating each cell along a measurement ray is repeated for all rays composing scan $Z_t$.
 
@@
 
-\section{Multi-Dimensional Maps}
+\section{Mapping in 2D Space}
 
@@
 \subsection{Ray Casting}
+
+Ray casting is the process of determining which cells a measurement ray might intersect, and the distances to these cells.
 

### 2018-06-23 Dissertation/Thesis/tex/intro.tex
edited up to Ray Casting

@@
 \begin{itemize}
-	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Autonomous Exploration with Exact Inverse Sensor Models,'' \textit{Journal of Intelligent }\&\textit{ Robotic Systems}, 2016, submitted.
+	\item E.
+Kaufman, K.
+Takami, Z.
+Ai, and T.
+Lee, ``Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities,'' \textit{The Second IEEE International Conference on Robotic Computing}, Laguna Hills, January 2018, pp.
+49--55.
+	\item E.
+Kaufman, K.
+Takami, T.
+Lee, and Z.
+Ai, ``Autonomous Exploration with Exact Inverse Sensor Models,'' \textit{Journal of Intelligent }\&\textit{ Robotic Systems}, 2017, doi: 10.1007/s10846-017-0710-7.
 	\item E.
@@
 Lee, and Z.
-Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, San Francisco, December 2016, accepted.
+Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, San Francisco, December 2016, pp.
+246--251.
 	\item E.

### 2018-06-23 Dissertation/Thesis/tex/pogm.tex
edited up to Ray Casting

@@
 In this chapter, we present the exact solution to occupancy grid map probability.
-The Bayesian approach uses the stochastic properties of sensors, and exploits important patterns in conditional probabilities.
+This Bayesian approach uses the stochastic properties of sensors, and exploits important patterns in conditional probabilities.
 This solution is simulated in several examples to show the efficacy of the approach.
@@
 The probability that the $i$-th cell is occupied is $P(\mathbf{m}_i)$, and the probability that it is free is $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$.
-The random variables $\mathbf{m}_i$ are mutually independent,
+The random variables $\mathbf{m}_i$ are mutually independent, i.e.,
 \begin{align}
@@
 Consider a range sensor that provides scans of the surrounding environment in order to identify the closest occupied space.
-The location and the direction of the sensor, referred to as the \emph{pose} at time $t$, is denoted by $X_t=(x_t,R_t)$, where the planar position and direction of the robot at $t$ are denoted by $x_t\in\Re^2$ and $R_t\in\Sph^1=\braces{q\in\Re^2|\norm{q}=1}$, i.e., the attitude corresponds to a direction on a 2D plane.
+The location and the direction of the sensor, referred to as the \emph{pose} at time $t$, is denoted by $X_t=(x_t,R_t)$.
+In a 2D environment, the planar position and direction of the robot at $t$ are denoted by $x_t\in\Re^2$ and $R_t\in\Sph^1=\braces{q\in\Re^2|\norm{q}=1}$, i.e., the attitude corresponds to a direction on a 2D plane.
+Similarly in a 3D environment, the spacial position and attitude are denoted by $x_t\in\Re^3$ and $R_t\in\SO=\braces{R\in\Re^{3\times3}|R^\T R=I,\det{R}=1}$.
 Let $X_{1:t}$ denote the history of poses from the initial time to the current time, i.e., $X_{1:t}=\{X_1,X_2,\ldots, X_t\}$.
@@
 The probability density function, namely $p(z_{t,l}|m,X_t)$ with respect to the depth of the $l$-th measurement ray conditioned on the map $m$ and the pose $X_t$ is commonly referred to as the \emph{forward sensor model}, which characterizes the corresponding depth sensor, such as the maximum range or accuracy.
-The forward sensor model satisfies (i) the ranges of all depth measurements are positive and finite, and (ii) measurement rays cannot pass through occupied regions.
+The forward sensor model satisfies (i) the ranges of all depth measurements are positive and finite, and (ii) a measurement cannot pass through occupied regions.
 Throughout this paper, we assume that the forward sensor model of the selected sensor is given.
@@
 Occupancy grid mapping provides occupancy probabilities based on robot poses and measurement scans.
-The goal is to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, commonly referred to as the \emph{inverse sensor model} for given forward sensor model $p(z_{t,l}|m,X_t)$ and the initial estimate of the map $P(m)$ with $X_{1:t}$ and $Z_{1:t}$.
+The goal is to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, commonly referred to as the \emph{inverse sensor model} for a given forward sensor model $p(z_{t,l}|m,X_t)$ and the initial estimate of the map $P(m)$ with $X_{1:t}$ and $Z_{1:t}$.
 
@@
 The occupancy probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$ is based on the forward sensor model $p(z_{t,l}|m,X_{1:t},Z_{1:t-1})$ that describe the distribution of the measurements for the given robot pose and the map.
-This specifies the stochastic characteristics of the sensor, such as the mean squared error, and it is assumed as a known distribution, specific to a particular depth sensor.
+This specifies the stochastic characteristics of the sensor are a known distribution, specific to a particular depth sensor.
 One could certainly obtain a stochastic model by fitting a probability density to a controlled set of measurements.
@@
 \end{align}
-considering that $X_t$ carries no information about $m$ without $Z_t$.
+where the second term in the numerator considers that $X_t$ carries no information about $m$ without $Z_t$.
 If the current pose $X_t$ and map $m$ are known, then the measurement ray $z_{t,l}$ is independent of past poses $X_{1:t-1}$ and past measurements $Z_{1:t-1}$ to obtain
@@
 Since there are $2^{n_{m}-1}$ maps in $\mathcal{M}_i$, these equations require $2^{n_m}$ terms to compute the summation for the $i$-th grid cell and normalizer, and this process should be repeated for all other cells along the measurement ray.
-This is the main reason why the existing results are based on certain approximations of the above expression.
+This is the main reason why the existing results are based on approximations or learned solutions of the above expression.
 
@@
 Since the cells outside of the sensor field of view (FOV) are not affected, we focus on a reduced map $r_l$ in the FOV of the $l$-th ray.
-This reduced map is chosen such that each cell of $r_l$ corresponds to a grid cell of map $m$ that the $l$-th ray intersects, ordered by increasing distance, which is easily determined from geometry.
+This reduced map is chosen such that each cell of $r_l$ corresponds to a grid cell of map $m$ that the $l$-th ray intersects, ordered by increasing distance.
 Let $\mathbf{r}_{l,k}$ be the binary random variable representing the occupancy of the $k$-th cell of the $l$-th ray.
@@
 Next, let $\mathbf{r}_{l,k+}$ correspond to the event that the $k$-th cell of the $l$-th ray is occupied, cells with lower index (closer cells) are free, and cells with greater index (farther cells) may or may not be occupied, i.e., event $\mathbf{r}_{l,k+}$ occurs when \\$r\in\braces{r\in\{0,1\}^{{n_{r,l}}}\,|\mathbf{r}_{l,1}=0,\mathbf{r}_{l,2}=0,\ldots,\mathbf{r}_{l,k-1}=0,\mathbf{r}_{l,k}=1}$.
-In other words, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray.
+Put differently, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray.
 For example, see Figure \ref{fig:show_rkplus} that illustrates the event of $\mathbf{r}_{l,k+}$ when $l=1$ and $k=4$ for an one-dimensional cell array.
-This concept of grouping map outcomes can be easily extended to 2D using ray casting, where a 1D ray is spanning 2D space.
-The cells considered for the measurement ray are determined with the geometric intersection between the ray and map cell edges.
+This concept of grouping map outcomes can be easily extended to 2D using ray casting, where a 1D ray is spanning 2D space, which is easily determined from geometry, described in further detail in Section \ref{sec:RayCasting}.
 Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell.
 Based on this, we present the mathematical expressions for the exact inverse sensor model, as summarized by \refeqn{RayISMAnswer} in Proposition 1.
-Later in Section \ref{sec:SA}, these are also rearranged as a computational algorithm.
+Later in Sections \ref{sec:RayByRay} and \ref{sec:Synergistic}, these are also rearranged in computational algorithms.
 
@@
   \includegraphics[width=0.4\textwidth]{rkplus_4.png}
-  \caption{A robot located to the left of a 1D occupancy grid map $r_l$ composed of $n_{r,l}=6$ grid cells in $4$ cases.
-In each case, cells $1$--$3$ are free, cell $4$ is occupied, and cells $5$ and $6$ may or may not be occupied.
+  \caption{A robot located to the left of a 1D occupancy grid map $r_l$ composed of $n_{r,l}=6$ grid cells in four cases.
+In each case, cells the first three cells are free, the fourth cell is occupied, and the fifth and sixth cells may or may not be occupied.
 All above outcomes correspond to the event $\mathbf{r}_{l,4+}$.}
@@
 &=
-\bigg[\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
+\bigg[\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
 \end{align}
@@
 \end{prop}
-\begin{proof}
+\begin{proof}% TODO: add ref
 See Appendix.
@@
 
-Since $\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ from \refeqn{Unnormalized} uses several repeated terms from the prior $\tilde P(\mathbf{r}_{l,k-1}|z_{t,l},X_{1:t},Z_{1:t-1})$, and $\eta_{t,l}$ is easily obtained from these as well, the computational cost of \refeqn{RayISMAnswer} is linear with respect to the number of cells along a measurement ray, amortized to $\mathcal{O}(1)$ for each cell, so the exact inverse sensor model can be applied in real-time.
+Since $\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ from \refeqn{Unnormalized} uses several repeated terms from the prior $\tilde P(\mathbf{r}_{l,k-1}|z_{t,l},X_{1:t},Z_{1:t-1})$, and $\eta_{t,l}$ is easily obtained from these as well, the computational cost of \refeqn{RayISMAnswer} is linear with respect to the number of cells along a measurement ray, amortized to $\mathcal{O}(1)$ for each cell.
+Because of this substantial computational improvement, the exact inverse sensor model can be applied in real-time.
 The process of updating each cell along a measurement ray is repeated for all rays composing scan $Z_t$.
@@
 
+% show RayCastingIllustration.png
 \subsection{Ray Casting}
+\label{sec:RayCasting}
 
@@
 \paragraph{Ray-By-Ray Approach}
+\label{sec:RayByRay}
 
 \paragraph{Synergistic Update Approach}
+\label{sec:Synergistic}
 

### 2018-06-23 Dissertation/Thesis/tex/pogm.tex
added new ray casting figs

@@
 Based on this, we present the mathematical expressions for the exact inverse sensor model, as summarized by \refeqn{RayISMAnswer} in Proposition 1.
-Later in Sections \ref{sec:RayByRay} and \ref{sec:Synergistic}, these are also rearranged in computational algorithms.
+Later in Section \ref{sec:RayScanComb}, these are also rearranged in computational algorithms.
 
@@
 Ray casting is the process of determining which cells a measurement ray might intersect, and the distances to these cells.
+The process is straightforward: follow a measurement unit vector from its minimum range $z_\text{min}$ to maximum range $z_\text{max}$, and identify the edges of all grid cells that it intersects.
+Save the cell index and Cartesian distance from the robot, and re-order the cells by increasing distance.
+An illustration of a simple example is shown in Figure \ref{fig:RayCastingIllustration}.
+
+\begin{figure}[!ht]
+    \centering
+    \begin{subfigure}[t]{0.4\columnwidth}
+        \centering
+        \includegraphics[width=\textwidth]{RayCastingIllustration.png}
+        \caption{Ray Casting on a 2D Grid}
+%        \label{fig:penn_map_total}
+    \end{subfigure}
+    \hspace*{0.05\columnwidth}
+    \begin{subfigure}[t]{0.4\columnwidth}
+        \centering
+        \includegraphics[width=\textwidth]{RayCastIllustrationReducedMapOnly.png}
+        \caption{Reduced Map Only}
+%        \label{fig:penn_map_zoom}
+    \end{subfigure}
+    \caption{This simple 2D example illustrates how ray casting determines the cells that compose the reduced map for the inverse sensor model.
+The closest edges are used to determine which cells should be considered, and in what order.
+The complete map cell indices are temporarily saved as well, which are used to associate probabilities to cells of the reduced map.}
+    \label{fig:RayCastingIllustration}
+\end{figure}
 
 \subsection{Combining Measurements from a Single Scan}
-
-Next, we describe two approaches to combine the solutions to the inverse sensor model for a single measurement ray...
+\label{sec:RayScanComb}
+Typically, many measurement rays are provided from a single scan from modern sensors.
+Next, we describe two approaches to combine the inverse sensor models of numerous measurement rays.
 
 \paragraph{Ray-By-Ray Approach}
-\label{sec:RayByRay}
 
 \paragraph{Synergistic Update Approach}
-\label{sec:Synergistic}
 

### 2018-06-23 Dissertation/Thesis/tex/pogm.tex
added ray-by-ray and partway done with pseudocodes

@@
 
+At the $t$-th time step, consider the measurement scan $Z_t=\{z_{t,1},z_{t,2},...,z_{t,n_z}\}$.
+One may apply the results of Proposition 1 repeatedly for each ray to obtain a two-dimensional inverse sensor model because
+\begin{align}
+\label{eqn:RayByRayScanISM}
+P(\mathbf{m}_i|X_{1:t},Z_{1:t})&%=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1},z_{t,2},...,z_{t,n_z})\nonumber\\&
+=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1}|z_{t,2}|...|z_{t,n_z}).
+\end{align}
+In this approach, each ray is considered individually such that the inverse sensor model updates the probabilities of grid cells from one ray before subsequent rays of $Z_t$ are considered.
+Therefore, we assume $z_{t,1}$ is independent of $z_{t,2:n_z}$.
+For the second ray, namely $z_{t,2}$, this may depend on $z_{t,1}$, but not subsequent rays $z_{t,3:n_z}$.
+By the time $z_{t,n_z}$ is considered, it may depend on all prior rays $z_{t,1:n_z-1}$.
+Therefore, the ray-by-ray approach considers rays within a single scan as partially dependent on each other.
+
+In short, the proposed scan inverse sensor model is computed by \refeqn{RayISMAnswer}-\refeqn{RayByRayScanISM}.
+These are summarized in Algorithm \ref{alg:RayByRayISM} as a computationally efficient, recursive algorithm that avoids several repeated calculations of the same quantity.
+This algorithm utilizes temporary variables, where the required computational resources are minimized by computing the updated occupancy probabilities of all grid cells along a measurement ray together.
+Note that this function is only considering $l$-th ray at the $t$-th time step, where probabilities are subject to conditions on $X_{1:t},Z_{1:t-1},z_{t,1:l-1}$, so these are removed from the algorithm for simplicity.
+
+\vspace*{0.05\columnwidth}
+\begin{algorithm}[H]
+	Function: $P(\mathbf{r}_k|Z)=\text{InverseSensorModelRayByRay}(P(\mathbf{r}_k),Z)$\;
+	% $\forall$ $l\in\braces{1,2,\ldots,n_{l}}$\;
+
+	\For{$l=1,2,\ldots,n_{l}$}{
+		Find $n_{r,l}$ cells from $m$ corresponding to $z_l$\;
+		Initialize $\eta_l^{-1}=0$ and $P(\bar{\mathbf{r}}_{1:l,0})=1$\;
+		\For{$k=1,2,\ldots,n_{r,l}$}{
+			$P(\mathbf{r}_{1:l,k+}) = P(\bar{\mathbf{r}}_{1:l,0:k-1})P(\mathbf{r}_{1:l,k})$\;
+			$P(\bar{\mathbf{r}}_{1:l,0:k})=P(\bar{\mathbf{r}}_{1:l,0:k-1})(1-P(\mathbf{r}_{1:l,k}))$\;
+    			$a_\text{temp}=P(\mathbf{r}_{1:l,k+})p(z_l|\mathbf{r}_{1:l,k+},X)$\;
+   			$\tilde P(\mathbf{r}_{l,k}|z_{1:l})=P(\mathbf{r}_{1:l,k})\eta^{-1}_{l}+a_\text{temp}$\;
+   			$\eta_l^{-1}=\eta_l^{-1}+a_\text{temp}$\;
+		}
+	$\eta_l^{-1}=\eta_l^{-1}+P(\bar{\mathbf{r}}_{1:l,0:k})p(z_\text{max})$\;
+	$P(\mathbf{r}_{k}|z_{1:l})=\eta\tilde P(\mathbf{r}_{k}|z_{1:l})$ for all $k\in\braces{1,2,\ldots,n_{r,l}}$\;
+	Substitute $P(\mathbf{r}_{k}|z_{1:l})$ back into $P(m|z_{1:l})$\;
+	}
+	Return $P(m|Z)$\;
+\caption{Inverse Sensor Model of an Individual Measurement Ray}
+\label{alg:RayByRayISM}
+\end{algorithm}
+\vspace*{0.05\columnwidth}
+
 \paragraph{Synergistic Update Approach}
 
-\subsection{Multi-Sensor Fusion}
+In this approach, all rays from a single scan are considered simultaneously for a synergistic update of all grid cells falling inside the scan FOV.
+However, the subsequent formulations require the assumption that all measurement rays within the scan are mutually independent.
+Here, we construct such two-dimensional inverse sensor model, referred to as the synergistic scan inverse sensor model.
+Let $\mathcal L_i\subset\{1,\ldots, n_z\}$ be the set of rays that pass through the $i$-th cell, $\mathbf{m}_i$. Applying Bayes' rule repeatedly, we obtain
+\begin{align}
+&P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\
+&=\tilde\zeta_i\braces{\prod_{l\in\mathcal{L}_i}p(z_{t,l}|\mathbf{m}_i,X_{1:t},Z_{1:t-1})}P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})\nonumber\\
+&=
+\zeta_i P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})%\nonumber\\&\quad\times
+\prod_{l\in\mathcal{L}_i}\frac{P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})},
+\label{eqn:ThirdBayesRule}
+\end{align}
+where $\tilde\zeta_i,\zeta_i\in\Re$ correspond to normalizing constants independent of $\mathbf{m}_i$.
 
-\subsection{Practical Implications}
+Suppose the $l$-th measurement ray intersects with the cell $\mathbf{m}_i$, and it is the $k$-th cell of the corresponding reduced map, i.e., $\mathbf{r}_{l,k}$ corresponds to the cell $\mathbf{m}_i$ in the $l$-th reduced map, and $\mathbf{r}_{l,k}=\mathbf{m}_i$ since they represent the same cell.
+Then, we have $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})=P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$.
 
-\section{Numerical and Experimental Examples}
+Using this, \refeqn{ThirdBayesRule} is rewritten as
+\begin{align}
+P(\mathbf{m}_i|X_{1:t},Z_{1:t})
+&=\xi_i P(\mathbf{m}_i|{X_{1:t-1}},Z_{1:t-1})
+\prod_{l\in\mathcal L_i}
+\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
+,
+\label{eqn:ISM_Fusion}
+\end{align}
+where the normalizer $\xi_i$ is composed of the product of normalizers $\zeta_i$, and $\eta_i$.
+Here, we introduce a new term $\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ for computational efficiency as
+\begin{align}
+&\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
+\triangleq \frac{\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})}
+\nonumber\\&=
+\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})
+\nonumber\\&\quad
++
+\bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,k+},X_t),
+\end{align}
+where we have used \refeqn{Unnormalized}. Similarly, its complement is
+\begin{align}
+P(\bar{\mathbf{m}}_i|{{X_{1:t}}},Z_{1:t})
+&=\xi_i P(\bar{\mathbf{m}}_i|{X_{1:t-1}},Z_{1:t-1})
+\prod_{l\in\mathcal L_i}
+\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
+,
+\label{eqn:ISM_Bar_Fusion}
+\end{align}
+where $\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ is defined as
+\begin{align}
+&\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
+\triangleq\frac{\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t-1})}
+\nonumber\\
+&=\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\} p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})
+\nonumber
+\\
+&\quad
++\frac{
+\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})}{P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})},
+\end{align}
+which is obtained from \refeqn{tildePbar}. Since % TODO: ref in appendix
+\begin{align*}
+P(\mathbf{m}_i|X_{1:t},Z_{1:t})+P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})=1,
+\end{align*}
+the normalizer $\xi_i$ is obtained using \refeqn{ISM_Fusion} and \refeqn{ISM_Bar_Fusion} as,
+\begin{align}
+&\xi_i=
+\bigg[
+P(\mathbf{m}_i|{X_{1:t-1}},Z_{1:t-1})
+\prod_{\mathcal L_i}
+\hat P(\mathbf{r}_{l,k}|z_{t,l},{X_{1:t}},Z_{1:t-1})
+\nonumber\\&\quad
++
+P(\bar{\mathbf{m}}_i|{X_{1:t-1}},Z_{1:t-1})
+\prod_{\mathcal L_i}
+\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
+\bigg]^{-1},\label{eqn:xi}
+\end{align}
+which is substituted into \refeqn{ISM_Fusion} to obtain the complete scan inverse sensor model $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$.
+
+The synergistic scan inverse sensor model algorithm is summarized with the pseudocode of Algorithm \ref{alg:SynergisticScanISM} as a computationally efficient, recursive algorithm that avoids repeated calculations of the same quantity.
+This algorithm utilizes the following temporary variables to develop the algorithm in an efficient recursive form, defined as
+\begin{align*}
+a_k&=\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
+\\
+b_k&=\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1}),
+\nonumber\\
+c_k&=P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})^{-1}
+\nonumber\\&\quad\times 
+\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
+\end{align*}
+to compute \refeqn{ISM_Fusion}-\refeqn{xi}.
+
+\begin{table}
+\begin{tabular}{ l }
+  for $l = 1,2,...,n_z$\\
+   \ \ Obtain reduced map $r_{t,l}$\\
+   \ \ Define $P(\mathbf{r}_{l,0}|X_{1:t-1},Z_{1:t-1})=0$, $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=1$, \\  \ \ \ \ \ \ \ \ \ \   $P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$;\\
+   \ \ for $k = 1,2,...,n_r$\\
+   \ \ \ \ Obtain $p(z_{t,l}|X_t,\mathbf{r}_{l,k})$ from a forward sensor model\\
+   \ \ \ \ if $k=1$\\
+   \ \ \ \ \ \ $a_1=0$, $b_1=1$;\\
+   \ \ \ \ else\\
+   \ \ \ \ \ \ $a_k=a_{k-1}+b_{k-1}p(z_{t,l}|\mathbf{r}_{l,k-1},X_t)P(\mathbf{r}_{l,k-1}|X_{1:t-1},Z_{1:t-1})$;\\
+   \ \ \ \ \ \ $b_k=b_{k-1}P(\bar{\mathbf{r}}_{l,k-1}|X_{1:t-1},Z_{1:t-1})$;\\
+   \ \ \ \ end if\\
+   \ \ end for\\
+   \ \ $c_{n_{r,l}+1}=0$;\\
+   \ \ for $k = n_{r,l},n_{r,l}-1,...,1$\\
+   \ \ \ \ $c_k=\frac{P(\mathbf{r}_{l,k+1}|X_{1:t-1},Z_{1:t-1})}{P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})}c_{k+1}$\\
+   \ \ \ \ \ \ \ \ \ \ \ \ $+b_{k}p(z_{t,l}|\mathbf{r}_{l,k+1},X_t)P(\mathbf{r}_{l,k+1}|X_{1:t-1},Z_{1:t-1})$;\\
+   \ \ end for\\
+   \ \ for $k = 1,2,...,n_r$\\
+   \ \ \ \ $\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=a_k+b_kp(z_{t,l}|\mathbf{r}_{l,k},X_t)$;\\
+   \ \ \ \ $\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=a_k+c_k$;\\
+   \ \ end for\\
+   end for\\
+   for all $i$ map cells inside the FOV\\
+   \ \ Obtain the set $\mathcal L_i$ of $l_i$ measurement rays intersecting this cell\\
+   \ \ if $l_i>0$\\
+   \ \ \ \ $d=P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})\prod_{\mathcal L_i}\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$;\\
+   \ \ \ \ $e=P(\bar{\mathbf{m}}_i|X_{1:t-1},Z_{1:t-1})\prod_{\mathcal L_i}\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$;\\
+   \ \ \ \ $P(\mathbf{m}_i|X_{1:t},Z_{1:t})=\frac{d}{d+e}$;\\
+   \ \ else\\
+   \ \ \ \ $P(\mathbf{m}_i|X_{1:t},Z_{1:t})=P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1});$\\
+   \ \ end if\\
+   \ \ Return: $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$\\
+   end for\\
+
+\end{tabular}
+\caption{Fast Algorithm of the Exact Solution of the Inverse Sensor Model for $l$ Simultaneous Measurement Rays}
+\label{tab:Alg_ISM_2D}
+\end{table}
+
+In short, we propose two alternatives for combining inverse sensor models for multiple rays within a scan.
+The ray-by-ray approach updates the entire map one ray at a time with some dependancies among the rays, while the synergistic approach considers all rays together where rays are assumed mutually independent.
+
+% From ACC16: In contrast to the current approximate inverse sensor models, the proposed algorithm evaluates the exact inverse sensor model \refeqn{InvSenModWithProbDens} efficiently without relying on assumptions like \refeqn{AssumptionEarly}, while integrating multiple measurements synergistically.
+Therefore, it yields substantially more accurate maps for the same set of measurements.
+These are illustrated by numerical examples and experiments as follows.
+
+% TODO: move 2 below to 3D mapping part
+%\subsection{Multi-Sensor Fusion}
+%
+%\subsection{Practical Implications}
 
@@
 
+% Upon further study, the ray-by-ray approach has proven more robust against localization errors and changing environments because the rays are not assumed completely independent.
+
 \section{Conclusions}

### 2018-06-23 Dissertation/Thesis/tex/pogm.tex
fixed algorithms

@@
 \begin{algorithm}[H]
-	Function: $P(\mathbf{r}_k|Z)=\text{InverseSensorModelRayByRay}(P(\mathbf{r}_k),Z)$\;
+	Function: $P(m|X,Z)=\text{InverseSensorModelRayByRay}(P(m),X,Z)$\;
 	% $\forall$ $l\in\braces{1,2,\ldots,n_{l}}$\;
@@
 		Find $n_{r,l}$ cells from $m$ corresponding to $z_l$\;
-		Initialize $\eta_l^{-1}=0$ and $P(\bar{\mathbf{r}}_{1:l,0})=1$\;
+		Initialize $\eta_l^{-1}=0$ and $P(\bar{\mathbf{r}}_{l,0}|X,z_{1:l-1})=1$ ($z_{1:0}$ is no condition)\;
 		\For{$k=1,2,\ldots,n_{r,l}$}{
-			$P(\mathbf{r}_{1:l,k+}) = P(\bar{\mathbf{r}}_{1:l,0:k-1})P(\mathbf{r}_{1:l,k})$\;
-			$P(\bar{\mathbf{r}}_{1:l,0:k})=P(\bar{\mathbf{r}}_{1:l,0:k-1})(1-P(\mathbf{r}_{1:l,k}))$\;
-    			$a_\text{temp}=P(\mathbf{r}_{1:l,k+})p(z_l|\mathbf{r}_{1:l,k+},X)$\;
-   			$\tilde P(\mathbf{r}_{l,k}|z_{1:l})=P(\mathbf{r}_{1:l,k})\eta^{-1}_{l}+a_\text{temp}$\;
+			$P(\mathbf{r}_{l,k+}|X,z_{1:l-1}) = P(\bar{\mathbf{r}}_{l,0:k-1}|X,z_{1:l-1})P(\mathbf{r}_{l,k}|X,z_{1:l-1})$\;
+			$P(\bar{\mathbf{r}}_{l,0:k}|X,z_{1:l-1})=P(\bar{\mathbf{r}}_{l,0:k-1}|X,z_{1:l-1})(1-P(\mathbf{r}_{l,k}|X,z_{1:l-1}))$\;
+    			$a_\text{temp}=P(\mathbf{r}_{l,k+}|X,z_{1:l-1})p(z_l|\mathbf{r}_{1:l,k+},X)$\;
+   			$\tilde P(\mathbf{r}_{l,k}|X,z_{1:l})=P(\mathbf{r}_{l,k}|X,z_{1:l-1})\eta^{-1}_{l}+a_\text{temp}$\;
    			$\eta_l^{-1}=\eta_l^{-1}+a_\text{temp}$\;
 		}
-	$\eta_l^{-1}=\eta_l^{-1}+P(\bar{\mathbf{r}}_{1:l,0:k})p(z_\text{max})$\;
-	$P(\mathbf{r}_{k}|z_{1:l})=\eta\tilde P(\mathbf{r}_{k}|z_{1:l})$ for all $k\in\braces{1,2,\ldots,n_{r,l}}$\;
-	Substitute $P(\mathbf{r}_{k}|z_{1:l})$ back into $P(m|z_{1:l})$\;
+	$\eta_l^{-1}=\eta_l^{-1}+P(\bar{\mathbf{r}}_{l,0:k}|X,z_{1:l})p(z_\text{max})$\;
+	$P(\mathbf{r}_{k}|z_{1:l})=\eta\tilde P(\mathbf{r}_{k}|X,z_{1:l})$ for all $k\in\braces{1,2,\ldots,n_{r,l}}$\;
+	Substitute $P(\mathbf{r}_{k}|X,z_{1:l})$ back into $P(m|X,z_{1:l})$\;
 	}
-	Return $P(m|Z)$\;
-\caption{Inverse Sensor Model of an Individual Measurement Ray}
+	Return $P(m|X,z_{1:n_{l}})=P(m|X,Z)$\;
+\caption{Inverse Sensor Model of a Scan Updated Ray-By-Ray}
 \label{alg:RayByRayISM}
@@
 \vspace*{0.05\columnwidth}
+\newpage
 
@@
 \end{align*}
-to compute \refeqn{ISM_Fusion}-\refeqn{xi}.
+to compute \refeqn{ISM_Fusion}-\refeqn{xi}. Once again, the history of poses and measurement scans are removed from the pseudocode for simplicity.
 
-\begin{table}
-\begin{tabular}{ l }
-  for $l = 1,2,...,n_z$\\
-   \ \ Obtain reduced map $r_{t,l}$\\
-   \ \ Define $P(\mathbf{r}_{l,0}|X_{1:t-1},Z_{1:t-1})=0$, $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=1$, \\  \ \ \ \ \ \ \ \ \ \   $P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$;\\
-   \ \ for $k = 1,2,...,n_r$\\
-   \ \ \ \ Obtain $p(z_{t,l}|X_t,\mathbf{r}_{l,k})$ from a forward sensor model\\
-   \ \ \ \ if $k=1$\\
-   \ \ \ \ \ \ $a_1=0$, $b_1=1$;\\
-   \ \ \ \ else\\
-   \ \ \ \ \ \ $a_k=a_{k-1}+b_{k-1}p(z_{t,l}|\mathbf{r}_{l,k-1},X_t)P(\mathbf{r}_{l,k-1}|X_{1:t-1},Z_{1:t-1})$;\\
-   \ \ \ \ \ \ $b_k=b_{k-1}P(\bar{\mathbf{r}}_{l,k-1}|X_{1:t-1},Z_{1:t-1})$;\\
-   \ \ \ \ end if\\
-   \ \ end for\\
-   \ \ $c_{n_{r,l}+1}=0$;\\
-   \ \ for $k = n_{r,l},n_{r,l}-1,...,1$\\
-   \ \ \ \ $c_k=\frac{P(\mathbf{r}_{l,k+1}|X_{1:t-1},Z_{1:t-1})}{P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})}c_{k+1}$\\
-   \ \ \ \ \ \ \ \ \ \ \ \ $+b_{k}p(z_{t,l}|\mathbf{r}_{l,k+1},X_t)P(\mathbf{r}_{l,k+1}|X_{1:t-1},Z_{1:t-1})$;\\
-   \ \ end for\\
-   \ \ for $k = 1,2,...,n_r$\\
-   \ \ \ \ $\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=a_k+b_kp(z_{t,l}|\mathbf{r}_{l,k},X_t)$;\\
-   \ \ \ \ $\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=a_k+c_k$;\\
-   \ \ end for\\
-   end for\\
-   for all $i$ map cells inside the FOV\\
-   \ \ Obtain the set $\mathcal L_i$ of $l_i$ measurement rays intersecting this cell\\
-   \ \ if $l_i>0$\\
-   \ \ \ \ $d=P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})\prod_{\mathcal L_i}\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$;\\
-   \ \ \ \ $e=P(\bar{\mathbf{m}}_i|X_{1:t-1},Z_{1:t-1})\prod_{\mathcal L_i}\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$;\\
-   \ \ \ \ $P(\mathbf{m}_i|X_{1:t},Z_{1:t})=\frac{d}{d+e}$;\\
-   \ \ else\\
-   \ \ \ \ $P(\mathbf{m}_i|X_{1:t},Z_{1:t})=P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1});$\\
-   \ \ end if\\
-   \ \ Return: $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$\\
-   end for\\
-
-\end{tabular}
-\caption{Fast Algorithm of the Exact Solution of the Inverse Sensor Model for $l$ Simultaneous Measurement Rays}
-\label{tab:Alg_ISM_2D}
-\end{table}
+% TODO: check spacing of algorithms once complete
+\vspace*{0.05\columnwidth}
+\begin{algorithm}[H]
+	Function: $P(m|X,Z)=\text{InverseSensorModelSynergistic}(P(m),X,Z)$\;
+	\For{$l=1,2,\ldots,n_{l}$}{
+		Find $n_{r,l}$ cells from $m$ corresponding to $z_l$\;
+		Define $P(\mathbf{r}_{l,0})=0$, $P(\bar{\mathbf{r}}_{l,0})=1$, $P(\mathbf{r}_{l,n_{r,l}+1})=1$, and $c_{n_{r,l}+1}=0$\;
+		\For{$k=1,2,\ldots,n_{r,l}$}{
+			\If{$k=1$}{
+				$a_1=0$, $b_1=1$\;
+			}
+			\Else{
+				$a_k=a_{k-1}+b_{k-1}p(z_{l}|\mathbf{r}_{l,k-1},X)P(\mathbf{r}_{l,k-1})$, $b_k=b_{k-1}P(\bar{\mathbf{r}}_{l,k-1})$\;
+			}
+		}
+		\For{$k = n_{r,l},n_{r,l}-1,...,1$}{
+			$c_k=\frac{P(\mathbf{r}_{l,k+1})}{P(\bar{\mathbf{r}}_{l,k})}c_{k+1}+b_{k}p(z_{t,l}|\mathbf{r}_{l,k+1},X)P(\mathbf{r}_{l,k+1})$\;
+		}
+		\For{$k = 1,2,...,n_{r,l}$}{
+			$\hat P(\mathbf{r}_{l,k}|z_{t,l},X)=a_k+b_kp(z_{t,l}|\mathbf{r}_{l,k},X)$, $\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X,)=a_k+c_k$\;
+		}
+	}
+	\For{$i\in m_\text{FOV}$ \text{(map cells inside FOV)}}{
+		Obtain the set $\mathcal L_i$ of $l_i$ measurement rays intersecting this cell\;
+		\If{$l_i>0$}{
+			$d=P(\mathbf{m}_i)\prod_{\mathcal L_i}\hat P(\mathbf{r}_{l,k}|z_{t,l},X)$\;
+			$e=P(\bar{\mathbf{m}}_i)\prod_{\mathcal L_i}\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X)$\;
+			$P(\mathbf{m}_i|X,Z)=\frac{d}{d+e}$\;
+		}
+		\Else{
+			$P(\mathbf{m}_i|X,Z)=P(\mathbf{m}_i)$\;
+		}
+	}
+	Return $P(m|X,Z)$\;
+\caption{Inverse Sensor Model of a Scan Updated Synergistically}
+\label{alg:SynergisticScanISM}
+\end{algorithm}
+\vspace*{0.05\columnwidth}
 

### 2018-06-24 Dissertation/Thesis/tex/pogm.tex
finished chap. 2, sec 2.3.3 forward needs editing

@@
 \begin{align*}
-a_k&=\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
+a_k&=\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
 \\
@@
 \nonumber\\
-c_k&=P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})^{-1}
-\nonumber\\&\quad\times 
-\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
+c_k&=\frac{
+\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})}{P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})},
 \end{align*}
-to compute \refeqn{ISM_Fusion}-\refeqn{xi}. Once again, the history of poses and measurement scans are removed from the pseudocode for simplicity.
+as well as $d$ and $e$ as the two terms composing \refeqn{xi}. Once again, the history of poses and measurement scans are removed from the pseudocode for simplicity.
 
@@
 
-% From ACC16: In contrast to the current approximate inverse sensor models, the proposed algorithm evaluates the exact inverse sensor model \refeqn{InvSenModWithProbDens} efficiently without relying on assumptions like \refeqn{AssumptionEarly}, while integrating multiple measurements synergistically.
-Therefore, it yields substantially more accurate maps for the same set of measurements.
-These are illustrated by numerical examples and experiments as follows.
+% TODO: move to 3D mapping part
+%\subsection{Multi-Sensor Fusion}
 
-% TODO: move 2 below to 3D mapping part
-%\subsection{Multi-Sensor Fusion}
-%
+% TODO: move to experiment part
 %\subsection{Practical Implications}
 
-\subsection{Mapping a Room}
-% ACC16 comparison of exact/approx. inverse sensor models
+\subsection{Numerical Examples}
 
-\subsection{Kinect Measurement Scan Analysis}
-% ACC16 Single scan exact/approx. comparison
+In contrast to the current approximate inverse sensor models, the proposed algorithms evaluates the exact inverse sensor model efficiently without relying on approximations, learned solutions, or log-odds ratio assumptions.
+The following simulations show two examples comparing an approximate inverse sensor model with the proposed exact inverse sensor model, either with integrating multiple measurements ray-by-ray or synergistically.
+The proposed algorithms yield substantially more accurate maps for the same set of measurements.
 
-% Upon further study, the ray-by-ray approach has proven more robust against localization errors and changing environments because the rays are not assumed completely independent.
+\paragraph{Approximate Inverse Sensor Model}
+
+We compare the proposed exact solution to the inverse sensor model with an approximate algorithm developed for a Microsoft Kinect sensor~\cite{PirRutBisSch11,KhoElb12}, summarized as follows.
+The probability that the $i$-th grid cell $\mathbf{m}_i$ is occupied conditioned on the measurement ray $z_{t,l}$ (the $l$-th ray at the $t$-th time step) at the pose $X_t$ is the continuous function
+\begin{align*}
+%\label{eqn:ISM_Approx_1}
+&P(\mathbf{m}_i|z_{t,l},X_t)\nonumber\\&=\begin{cases}
+0.3+(\frac{k}{\sigma\sqrt{2\pi}}+0.2)e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ z_{t,l}\leq \hat z_{l,i},
+\\
+0.5+\frac{k}{\sigma\sqrt{2\pi}}e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ z_{t,l}>\hat z_{l,i},
+\end{cases}
+\end{align*}
+which is based on the expected distance to the cell $\hat z_{l,i}$ with parameters $k=\sigma=0.6$.
+This follows the structure of the approximate inverse sensor model proposed in~\cite{And09}.
+The main idea of this approach is that the probability of a cell being occupied (i) near a measurement is high (measurement likely hits this cell), (ii) between the robot and the measurement is low (measurement passes through this cell), and (iii) beyond the measurement is unchanged (the robot cannot measure through a wall/object).
+
+Then, these probabilities are combined in a weighted fashion such that all measurements rays of scan $Z_t$ simultaneously update the same grid cell in a log-odds format,
+\begin{align*}
+%\label{eqn:ISM_Approx_2}
+&\log\left(\frac{P(\mathbf{m}_i|Z_{t},X_t)}{1-P(\mathbf{m}_i|Z_{t},X_t)}\right)
+\nonumber\\&=
+\frac1{\sum_{z_{t,l}\in\mathbf{m}_i}\hat z_{l,i}}\sum_{z_{t,l}\in\mathbf{m}_i}\log\left(\frac{P(\mathbf{m}_i|z_{t,l},X_t)}{1-P(\mathbf{m}_i|z_{t,l},X_t)}\hat z_{l,i}\right).
+\end{align*}
+
+\paragraph{Mapping a Hallway with the Ray-By-Ray Approach}
+
+First, we compare the proposed exact solution to the inverse sensor model summarized in Algorithm \ref{alg:RayByRayISM} with an approximate algorithm presented in~\cite{PirRutBisSch11,KhoElb12}, summarized as follows.
+The probability that the $i$-th grid cell $\mathbf{m}_i$ is occupied conditioned on the measurement ray $z_{t,l}$ at the pose $X_t$ is the continuous function
+\begin{align*}
+%\label{eqn:ISM_Approx_1}
+&P(\mathbf{m}_i|z_{t,l},X_t)%\nonumber\\&
+=\begin{cases}
+0.3+(\frac{k}{\sigma\sqrt{2\pi}}+0.2)e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ \hat z_{l,i} \leq z_{t,l},%z_{t,l}\leq \hat z_{l,i},
+\\
+0.5+\frac{k}{\sigma\sqrt{2\pi}}e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{otherwise},
+\end{cases}
+\end{align*}
+which is based on the expected distance to the cell $\hat z_{l,i}$ with parameters $k=\sigma=0.6$.
+This follows the structure of the approximate inverse sensor model proposed in~\cite{And09}.
+The main idea of this approach is that the probability of a cell being occupied (i) near a measurement is high (measurement likely hits this cell), (ii) between the robot and the measurement is low (measurement passes through this cell), and (iii) beyond the measurement is unchanged (the robot cannot measure through a wall/object).
+Then, these probabilities are combined in a weighted fashion such that all measurements rays of scan $Z_t$ simultaneously update the same grid cell in a log-odds format,
+\begin{align*}
+%\label{eqn:ISM_Approx_2}
+&\log\left(\frac{P(\mathbf{m}_i|Z_{t},X_t)}{1-P(\mathbf{m}_i|Z_{t},X_t)}\right)
+%\nonumber\\&
+=
+\frac1{\sum_{z_{t,l}\in\mathbf{m}_i}\hat z_{l,i}}\sum_{z_{t,l}\in\mathbf{m}_i}\log\left(\frac{P(\mathbf{m}_i|z_{t,l},X_t)}{1-P(\mathbf{m}_i|z_{t,l},X_t)}\hat z_{l,i}\right).
+\end{align*}
+
+Table~\ref{tab:penn} shows the mapping parameters and data specifications for the actual odometry and lidar measurements from University of Pennsylvania through Coursera open course~\cite{coursera}.
+Among these parameters, cell resolution and the number of rays have the greatest impact on computation.
+In general, the finer the grid, the greater the memory and computational requirements are because the computational order of \refeqn{RayISMAnswer} grows linearly with the number of grid cells inside the sensor range limits.
+Since the inverse sensor models are combined sequentially as shown in \refeqn{RayByRayScanISM}, the number of rays considered are proportional to the computation order as well.
+Figure~\ref{fig:penn_map} shows the direct comparison of the resulting map based on the approximate and proposed inverse sensor model.
+As shown in the figure, both approaches capture the structure of the environment.
+However, the significant differences can be seen in the certain area of the map such as in Figure~\ref{fig:penn_map_zoom}.
+The differences can be clearly depicted in Figure~\ref{fig:entropy_comp} where entropy of maps are compared at frame 2280.
+The red part of the map shows the maximum  entropy whereas blue shows lowest entropy for the cell.
+Finally, overall improvement was demonstrated in Figure~\ref{fig:entropy} resulting improved entropy value per grid cell throughout the mapping process.
+In short, the proposed exact occupancy grid mapping produced a cleaner map with lower entropy from the same set of measurements.
+
+\begin{center}
+\captionof{table}{Experimental parameters provided from dataset}
+\label{tab:penn}
+    \begin{tabular}{r | c}
+        Parameters & Value\\ \hline\hline
+        Map dimension & 576x448 [pixel]\\
+        Resolution & 1/16 [m]\\
+        Scan angle & [-2.36, 2.36]\\
+        Ray number & 1081\\
+        Total frame & Every 120 scans out of 3701\\
+    \end{tabular}
+\end{center}
+
+\begin{figure}[!ht]
+    \centering
+    \begin{subfigure}[t]{0.8\columnwidth}
+        \centering
+        \includegraphics[trim={1.5cm 3.5cm 1.5cm 1.5cm},clip,width=\textwidth]{map_comparison.png}
+        \caption{Comparison of the total mapped environment based on the approximate (left) and proposed (right) inverse sensor model.}
+        \label{fig:penn_map_total}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.8\columnwidth}
+        \centering
+        \includegraphics[trim={1.5cm 3.5cm 1.5cm 1.5cm},clip,width=\textwidth]{map_zoom.png}
+        \caption{Evidence of significant improvement at certain area of the map.}
+        \label{fig:penn_map_zoom}
+    \end{subfigure}
+    \caption{Using experimental sensor data, the map was generated using an approximate and the proposed exact inverse sensor models.
+The laser sensor and odometry data used for the experiment was provided by University of Pennsylvania open course robotics estimation and learning on Coursera~\cite{coursera}.}
+    \label{fig:penn_map}
+\end{figure}
+
+\begin{figure}[!ht]
+    \centering
+    \begin{subfigure}[t]{0.35\columnwidth}
+        \centering
+        \includegraphics[width=\textwidth]{AISM_Image_inf_19.pdf}
+        \caption{Approximate inverse sensor model}
+        \label{fig:AISM}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.35\columnwidth}
+        \centering
+        \includegraphics[width=\textwidth]{EISM_Image_inf_19.pdf}
+        \caption{Exact inverse sensor model}
+        \label{fig:EISM}
+    \end{subfigure}
+    \caption{The entropy is compared between the approximate and proposed exact inverse sensor models at step 2280. The blue areas show the lowest entropy while red shows highest.}
+    \label{fig:entropy_comp}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.7\textwidth]{entropy_frame.png}
+  \caption{The map entropies of the approximate inverse sensor model (blue) and the proposed exact inverse sensor model (red) show that the exact version yields a more certain occupancy grid map.}
+  \label{fig:entropy}
+\end{figure}
+
+\paragraph{Mapping a Room with the Synergistic Approach}
+
+Next, we compare the same approximate inverse sensor model against the synergistic exact inverse sensor model.
+Here, the robot maps in a two-dimensional environment composed of ten wall edges, and the robot follows a figure-eight curve, then turns around and completes the same curve in the reverse direction.
+Like the prior example, the same set of measurements updated each second are used with both occupancy grid mapping algorithms to construct the map.
+
+The resulting maps are illustrated in Figure \ref{fig:NumResOccProbs} for both algorithms, where it is shown that the proposed algorithm yields a substantially more accurate and clear map with less uncertainty.
+To quantify the degree of map uncertainty, we define the entropy of the map as
+\begin{align*}
+H(P(m|X_{1:t},Z_{1:t}))&=-\sum_{i=1}^n\big\{P(\mathbf{m}_i|X_{1:t},Z_{1:t})\log P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\
+&\qquad+(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\log(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\big\},
+\end{align*}
+which is maximized when the probability of occupancy is $0.5$ for all cells (more uncertain), and it is minimized when they are either $0$ or $1$ (less uncertain).
+
+\begin{figure}
+%    \centering{
+%    \begin{subfigure}[t]{0.4\columnwidth}
+%        \centering
+%        \includegraphics[width=0.7\columnwidth]{Compare_ISM_1.png}
+%        \caption{Map at $t=0$ sec (exact)}
+%    \end{subfigure}
+%    \begin{subfigure}[t]{0.4\columnwidth}
+%        \centering
+%        \includegraphics[width=0.7\columnwidth]{Compare_Approx_1.png}
+%        \caption{Map at $t=0$ sec (approx.)}
+%    \end{subfigure}
+%    }
+    \centering{
+    \begin{subfigure}[t]{0.4\columnwidth}
+        \centering
+        \includegraphics[width=0.7\columnwidth]{Compare_ISM_2.png}
+        \caption{Map at $t=13$ sec (exact)}
+    \end{subfigure}
+    \hspace*{-0.05\columnwidth}
+    \begin{subfigure}[t]{0.4\columnwidth}
+        \centering
+        \includegraphics[width=0.7\columnwidth]{Compare_Approx_2.png}
+        \caption{Map at $t=13$ sec (approx.)}
+    \end{subfigure}
+    }
+    \centering{
+    \begin{subfigure}[t]{0.4\columnwidth}
+        \centering
+        \includegraphics[width=0.7\columnwidth]{Compare_ISM_3.png}
+        \caption{Map at $t=25$ sec (exact)}
+    \end{subfigure}
+    \hspace*{-0.05\columnwidth}
+    \begin{subfigure}[t]{0.4\columnwidth}
+        \centering
+        \includegraphics[width=0.7\columnwidth]{Compare_Approx_3.png}
+        \caption{Map at $t=25$ sec (approx.)}
+    \end{subfigure}
+    }
+    \centering{
+    \begin{subfigure}[t]{0.4\columnwidth}
+        \centering
+        \includegraphics[width=0.7\columnwidth]{Compare_ISM_4.png}
+        \caption{Map at $t=38$ sec (exact)}
+    \end{subfigure}
+    \hspace*{-0.05\columnwidth}
+    \begin{subfigure}[t]{0.4\columnwidth}
+        \centering
+        \includegraphics[width=0.7\columnwidth]{Compare_Approx_4.png}
+        \caption{Map at $t=38$ sec (approx.)}
+    \end{subfigure}
+    }
+    \centering{
+    \begin{subfigure}[t]{0.4\columnwidth}
+        \centering
+        \includegraphics[width=0.7\columnwidth]{Compare_ISM_5.png}
+        \caption{Map at $t=50$ sec (exact)}
+    \end{subfigure}
+    \hspace*{-0.05\columnwidth}
+    \begin{subfigure}[t]{0.4\columnwidth}
+        \centering
+        \includegraphics[width=0.7\columnwidth]{Compare_Approx_5.png}
+        \caption{Map at $t=50$ sec (approx.)}
+    \end{subfigure}
+    }
+\caption{A robot (green crossed circle, green curve of its previous $13$ seconds) measures a room with a Kinect depth sensor.
+The exact inverse sensor model is compared with an approximate algorithm designed for the Kinect depth sensor.
+Grid cells are either known as free (white) or occupied (black), or uncertain with a grayscale color.
+%The exact inverse sensor model approach generates a more accurate result.
+}
+\label{fig:NumResOccProbs}
+\end{figure}
+
+The change of the map entropy over time, and the entropy of the completed maps, for both methods are depicted in Figure \ref{fig:NumResOccH}.
+The subfigure (a) illustrates that the proposed exact inverse sensor model exhibits rapid decreases of entropies, and smaller entropies always.
+The resulting terminal map obtained from the proposed approach, shown in the subfigure (b), has less uncertainty than (c) constructed by the approximate model.
+In short, the proposed approach is more efficient at extracting information about the environment from the same set of set measurements.
+
+\begin{figure}
+    \centering{
+    \begin{subfigure}[t]{0.8\columnwidth}
+        \centering
+        \includegraphics[width=0.7\columnwidth]{EntropyEvolution.png}
+        \caption{Entropy (black: exact model, green: approx.)}
+    \end{subfigure}
+}
+    \centering{
+    \begin{subfigure}[t]{0.4\columnwidth}
+        \centering
+        \includegraphics[width=0.7\columnwidth]{Compare_ISM_H.png}
+        \caption{Entropy map at $t=50$ (exact)}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.4\columnwidth}
+        \centering
+        \includegraphics[width=0.7\columnwidth]{Compare_Approx_H.png}
+        \caption{Entropy map at $t=50$  (approx.)}
+    \end{subfigure}
+    }
+\caption{Entropy serves as a measure of uncertainty of the occupancy grid, where more blue regions are more certain, and red regions are more uncertain.
+The uncertainty is always less when the exact solution is applied.}\label{fig:NumResOccH}
+\end{figure}
+
+% I think I'll skip this trivial result:
+%\subsection{Kinect Measurement Scan Analysis}
+%% ACC16 Single scan exact/approx. comparison
 
 \section{Conclusions}
+
+In this chapter, we presented the exact solution to the inverse sensor model for a single measurement ray, then extended this result to measurement scans with two techniques, namely ray-by-ray and synergistic combinations.
+Then, numerical results show how both proposed solutions outperform existing approximate solutions.
+Upon further study, the ray-by-ray approach has proven more robust against localization errors and changing environments because the rays are not assumed completely independent.
+Therefore, the ray-by-ray approach is applied to the subsequent proposals, numerical simulations, and experimental results.

### 2018-06-24 Dissertation/Thesis/tex/pogm.tex
edit chap 2

@@
 In this approach, each ray is considered individually such that the inverse sensor model updates the probabilities of grid cells from one ray before subsequent rays of $Z_t$ are considered.
-Therefore, we assume $z_{t,1}$ is independent of $z_{t,2:n_z}$.
+Therefore, we assume the first measurement ray $z_{t,1}$ is independent of the remaining rays $z_{t,2:n_z}$.
 For the second ray, namely $z_{t,2}$, this may depend on $z_{t,1}$, but not subsequent rays $z_{t,3:n_z}$.
@@
 This algorithm utilizes temporary variables, where the required computational resources are minimized by computing the updated occupancy probabilities of all grid cells along a measurement ray together.
-Note that this function is only considering $l$-th ray at the $t$-th time step, where probabilities are subject to conditions on $X_{1:t},Z_{1:t-1},z_{t,1:l-1}$, so these are removed from the algorithm for simplicity.
+Note that this function is only considering $l$-th ray at the $t$-th time step, where probabilities are subject to conditions on the history of poses $X_{1:t-1}$ and measurement scans $Z_{1:t-1}$, so these are removed from the algorithm for simplicity.
 
@@
 		}
-	$\eta_l^{-1}=\eta_l^{-1}+P(\bar{\mathbf{r}}_{l,0:k}|X,z_{1:l})p(z_\text{max})$\;
-	$P(\mathbf{r}_{k}|z_{1:l})=\eta\tilde P(\mathbf{r}_{k}|X,z_{1:l})$ for all $k\in\braces{1,2,\ldots,n_{r,l}}$\;
+	$\eta_l^{-1}=\eta_l^{-1}+P(\bar{\mathbf{r}}_{l,0:n_{r,l}}|X,z_{1:l-1})p(z_\text{max})$\;
+	$P(\mathbf{r}_{k}|z_{1:l})=\eta_l\tilde P(\mathbf{r}_{k}|X,z_{1:l})$ for all $k\in\braces{1,2,\ldots,n_{r,l}}$\;
 	Substitute $P(\mathbf{r}_{k}|X,z_{1:l})$ back into $P(m|X,z_{1:l})$\;
@@
 In this approach, all rays from a single scan are considered simultaneously for a synergistic update of all grid cells falling inside the scan FOV.
-However, the subsequent formulations require the assumption that all measurement rays within the scan are mutually independent.
+However, the subsequent formulations require the assumption that all measurement rays within the scan are mutually independent, i.e.,
+\begin{align*}
+p(z_{t,1},z_{t,2},\ldots,z_{t,n_l}|\mathbf{m}_i,X_{1:t},Z_{1:t-1})=\prod_{l=1}^{n_l}p(z_{t,l}|\mathbf{m}_i,X_{1:t},Z_{1:t-1}).
+\end{align*}
 Here, we construct such two-dimensional inverse sensor model, referred to as the synergistic scan inverse sensor model.
-Let $\mathcal L_i\subset\{1,\ldots, n_z\}$ be the set of rays that pass through the $i$-th cell, $\mathbf{m}_i$. Applying Bayes' rule repeatedly, we obtain
+Let $\mathcal L_i\subset\{1,\ldots, n_l\}$ be the set of rays that pass through the $i$-th cell, $\mathbf{m}_i$. Applying Bayes' rule repeatedly, we obtain
 \begin{align}
-&P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\
+P(\mathbf{m}_i|X_{1:t},Z_{1:t})
 &=\tilde\zeta_i\braces{\prod_{l\in\mathcal{L}_i}p(z_{t,l}|\mathbf{m}_i,X_{1:t},Z_{1:t-1})}P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})\nonumber\\
@@
 \end{align*}
-as well as $d$ and $e$ as the two terms composing \refeqn{xi}. Once again, the history of poses and measurement scans are removed from the pseudocode for simplicity.
+as well as $d$ and $e$ for the two terms composing \refeqn{xi}. Once again, the history of poses and measurement scans are removed from the pseudocode for simplicity.
 
@@
 
-In contrast to the current approximate inverse sensor models, the proposed algorithms evaluates the exact inverse sensor model efficiently without relying on approximations, learned solutions, or log-odds ratio assumptions.
+In contrast to the current approximate inverse sensor models, the proposed ray-by-ray and synergistic algorithms evaluate the exact inverse sensor model efficiently without relying on approximations, learned solutions, or log-odds ratio assumptions.
 The following simulations show two examples comparing an approximate inverse sensor model with the proposed exact inverse sensor model, either with integrating multiple measurements ray-by-ray or synergistically.
@@
 The probability that the $i$-th grid cell $\mathbf{m}_i$ is occupied conditioned on the measurement ray $z_{t,l}$ (the $l$-th ray at the $t$-th time step) at the pose $X_t$ is the continuous function
-\begin{align*}
-%\label{eqn:ISM_Approx_1}
-&P(\mathbf{m}_i|z_{t,l},X_t)\nonumber\\&=\begin{cases}
+\begin{align}
+\label{eqn:ISM_Approx_1}
+P(\mathbf{m}_i|z_{t,l},X_t)=\begin{cases}
 0.3+(\frac{k}{\sigma\sqrt{2\pi}}+0.2)e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ z_{t,l}\leq \hat z_{l,i},
@@
 \end{cases}
-\end{align*}
+\end{align}
 which is based on the expected distance to the cell $\hat z_{l,i}$ with parameters $k=\sigma=0.6$.
-This follows the structure of the approximate inverse sensor model proposed in~\cite{And09}.
-The main idea of this approach is that the probability of a cell being occupied (i) near a measurement is high (measurement likely hits this cell), (ii) between the robot and the measurement is low (measurement passes through this cell), and (iii) beyond the measurement is unchanged (the robot cannot measure through a wall/object).
+This follows the structure of the approximate inverse sensor model proposed in~\cite{Andert09}.
+The main idea of this approach is that the probability of a cell being occupied (i) near a measurement is high (measurement likely hits this cell), (ii) between the robot and the measurement is low (measurement passes through these cells), and (iii) beyond the measurement is unchanged (the robot cannot measure through a wall/object).
 
 Then, these probabilities are combined in a weighted fashion such that all measurements rays of scan $Z_t$ simultaneously update the same grid cell in a log-odds format,
-\begin{align*}
-%\label{eqn:ISM_Approx_2}
-&\log\left(\frac{P(\mathbf{m}_i|Z_{t},X_t)}{1-P(\mathbf{m}_i|Z_{t},X_t)}\right)
-\nonumber\\&=
+\begin{align}
+\label{eqn:ISM_Approx_2}
+\log\left(\frac{P(\mathbf{m}_i|Z_{t},X_t)}{1-P(\mathbf{m}_i|Z_{t},X_t)}\right)
+=
 \frac1{\sum_{z_{t,l}\in\mathbf{m}_i}\hat z_{l,i}}\sum_{z_{t,l}\in\mathbf{m}_i}\log\left(\frac{P(\mathbf{m}_i|z_{t,l},X_t)}{1-P(\mathbf{m}_i|z_{t,l},X_t)}\hat z_{l,i}\right).
-\end{align*}
+\end{align}
 
@@
 
-First, we compare the proposed exact solution to the inverse sensor model summarized in Algorithm \ref{alg:RayByRayISM} with an approximate algorithm presented in~\cite{PirRutBisSch11,KhoElb12}, summarized as follows.
-The probability that the $i$-th grid cell $\mathbf{m}_i$ is occupied conditioned on the measurement ray $z_{t,l}$ at the pose $X_t$ is the continuous function
-\begin{align*}
-%\label{eqn:ISM_Approx_1}
-&P(\mathbf{m}_i|z_{t,l},X_t)%\nonumber\\&
-=\begin{cases}
-0.3+(\frac{k}{\sigma\sqrt{2\pi}}+0.2)e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ \hat z_{l,i} \leq z_{t,l},%z_{t,l}\leq \hat z_{l,i},
-\\
-0.5+\frac{k}{\sigma\sqrt{2\pi}}e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{otherwise},
-\end{cases}
-\end{align*}
-which is based on the expected distance to the cell $\hat z_{l,i}$ with parameters $k=\sigma=0.6$.
-This follows the structure of the approximate inverse sensor model proposed in~\cite{And09}.
-The main idea of this approach is that the probability of a cell being occupied (i) near a measurement is high (measurement likely hits this cell), (ii) between the robot and the measurement is low (measurement passes through this cell), and (iii) beyond the measurement is unchanged (the robot cannot measure through a wall/object).
-Then, these probabilities are combined in a weighted fashion such that all measurements rays of scan $Z_t$ simultaneously update the same grid cell in a log-odds format,
-\begin{align*}
-%\label{eqn:ISM_Approx_2}
-&\log\left(\frac{P(\mathbf{m}_i|Z_{t},X_t)}{1-P(\mathbf{m}_i|Z_{t},X_t)}\right)
-%\nonumber\\&
-=
-\frac1{\sum_{z_{t,l}\in\mathbf{m}_i}\hat z_{l,i}}\sum_{z_{t,l}\in\mathbf{m}_i}\log\left(\frac{P(\mathbf{m}_i|z_{t,l},X_t)}{1-P(\mathbf{m}_i|z_{t,l},X_t)}\hat z_{l,i}\right).
-\end{align*}
-
+First, we compare the proposed exact solution to the ray-by-ray inverse sensor model summarized in Algorithm \ref{alg:RayByRayISM} with an approximations of \refeqn{ISM_Approx_1}--\refeqn{ISM_Approx_2}.
 Table~\ref{tab:penn} shows the mapping parameters and data specifications for the actual odometry and lidar measurements from University of Pennsylvania through Coursera open course~\cite{coursera}.
@@
 As shown in the figure, both approaches capture the structure of the environment.
-However, the significant differences can be seen in the certain area of the map such as in Figure~\ref{fig:penn_map_zoom}.
-The differences can be clearly depicted in Figure~\ref{fig:entropy_comp} where entropy of maps are compared at frame 2280.
+However, the significant differences can be seen, particularly with an enlarged mapping section in Figure~\ref{fig:penn_map_zoom}.
+
+To quantify the degree of map uncertainty, we define the entropy of the map as 
+\begin{align*}
+H(P(m|X_{1:t},Z_{1:t}))&=-\sum_{i=1}^n\big\{P(\mathbf{m}_i|X_{1:t},Z_{1:t})\log P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\
+&\qquad+(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\log(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\big\},
+\end{align*}
+which is maximized when the probability of occupancy is $0.5$ for all cells (more uncertain), and it is minimized when they are either $0$ or $1$ (less uncertain). 
+The entropy differences among the algorithms can be clearly depicted in Figure~\ref{fig:entropy_comp} at frame 2280.
 The red part of the map shows the maximum  entropy whereas blue shows lowest entropy for the cell.
-Finally, overall improvement was demonstrated in Figure~\ref{fig:entropy} resulting improved entropy value per grid cell throughout the mapping process.
+Finally, overall improvement was demonstrated in Figure~\ref{fig:entropy} resulting in an improved entropy value per grid cell throughout the mapping process.
 In short, the proposed exact occupancy grid mapping produced a cleaner map with lower entropy from the same set of measurements.
@@
 \begin{center}
-\captionof{table}{Experimental parameters provided from dataset}
+\captionof{table}{Experimental Parameters Provided from the Dataset}
 \label{tab:penn}
@@
 Like the prior example, the same set of measurements updated each second are used with both occupancy grid mapping algorithms to construct the map.
-
 The resulting maps are illustrated in Figure \ref{fig:NumResOccProbs} for both algorithms, where it is shown that the proposed algorithm yields a substantially more accurate and clear map with less uncertainty.
-To quantify the degree of map uncertainty, we define the entropy of the map as
-\begin{align*}
-H(P(m|X_{1:t},Z_{1:t}))&=-\sum_{i=1}^n\big\{P(\mathbf{m}_i|X_{1:t},Z_{1:t})\log P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\
-&\qquad+(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\log(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\big\},
-\end{align*}
-which is maximized when the probability of occupancy is $0.5$ for all cells (more uncertain), and it is minimized when they are either $0$ or $1$ (less uncertain).
 
@@
 The change of the map entropy over time, and the entropy of the completed maps, for both methods are depicted in Figure \ref{fig:NumResOccH}.
-The subfigure (a) illustrates that the proposed exact inverse sensor model exhibits rapid decreases of entropies, and smaller entropies always.
-The resulting terminal map obtained from the proposed approach, shown in the subfigure (b), has less uncertainty than (c) constructed by the approximate model.
+The subfigure \ref{fig:NumResOccHa} illustrates that the proposed exact inverse sensor model exhibits rapid decreases of entropies, and smaller entropies always.
+The resulting terminal map obtained from the proposed approach, shown in the subfigure \ref{fig:NumResOccHb}, has less uncertainty than \ref{fig:NumResOccHc} constructed by the approximate model.
 In short, the proposed approach is more efficient at extracting information about the environment from the same set of set measurements.
@@
         \caption{Entropy (black: exact model, green: approx.)}
+        \label{fig:NumResOccHa}
     \end{subfigure}
@@
         \caption{Entropy map at $t=50$ (exact)}
+        \label{fig:NumResOccHb}
     \end{subfigure}
@@
         \caption{Entropy map at $t=50$  (approx.)}
+        \label{fig:NumResOccHc}
     \end{subfigure}
@@
 
-In this chapter, we presented the exact solution to the inverse sensor model for a single measurement ray, then extended this result to measurement scans with two techniques, namely ray-by-ray and synergistic combinations.
+In this chapter, we presented the exact solution to the inverse sensor model for a single measurement ray, then extended this result to 2D occupancy grids.
+Measurement scans were analyzed with two techniques, namely ray-by-ray and synergistic combinations.
 Then, numerical results show how both proposed solutions outperform existing approximate solutions.

### 2018-06-24 Dissertation/Thesis/tex/appendix.tex
added Appendix A: formal ISM proof

--- 2018-06-09 Journal/JINT 2017/JIRS16.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% This is a general template file for the LaTeX package SVJour3
-% for Springer journals.          Springer Heidelberg 2010/09/16
-%
-% Copy it to a new file with a new name and use it as the basis
-% for your article. Delete % signs as needed.
-%
-% This template includes a few options for different layouts and
-% content for various journals. Please consult a previous issue of
-% your journal as needed.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\RequirePackage{fix-cm}
-%
-%\documentclass{svjour3}                     % onecolumn (standard format)
-%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
-\documentclass[smallextended]{svjour3}       % onecolumn (second format)
-%\documentclass[twocolumn]{svjour3}          % twocolumn
-%
-\smartqed  % flush right qed marks, e.g. at end of proof
-%
-\usepackage{graphicx}
-\usepackage{amsmath,amssymb,url,times}%,subfigure}% amsthm is the one!
-\usepackage{caption,subcaption,hyperref}
-\usepackage{color,comment}
-\usepackage{curves,pgfgantt}
-%
-% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
-%
-% insert here the call for the packages your document requires
-%\usepackage{latexsym}
-% etc.
-%
-% please place your own definitions here and don't use \def but
-% \newcommand{}{}
-%
-% Insert the name of "your journal" with
- \journalname{Journal of Intelligent \& Robotic Systems}
-
-\newtheorem{prop}{Proposition}
-\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
-\newcommand{\abs}[1]{\ensuremath{\left| #1 \right|}}
-\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
-\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
-\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
-\newcommand{\ip}[1]{\ensuremath{\langle #1 \rangle}}
-\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
-\newcommand{\reffig}[1]{Figure \ref{fig:#1}}
-\newcommand{\tr}[1]{\mbox{tr}\ensuremath{\negthickspace\bracket{#1}}}
-\newcommand{\trs}[1]{\mbox{tr}\ensuremath{\!\bracket{#1}}}
-\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
-\newcommand{\G}{\ensuremath{\mathsf{G}}}
-\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
-\newcommand{\T}{\ensuremath{\mathsf{T}}}
-\renewcommand{\L}{\ensuremath{\mathsf{L}}}
-\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
-\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
-\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
-\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
-\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
-\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
-\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
-\newcommand{\D}{\ensuremath{\mathbf{D}}}
-\renewcommand{\d}{\ensuremath{\mathbf{d}}}
-\newcommand{\pair}[1]{\ensuremath{\left\langle #1 \right\rangle}}
-\newcommand{\met}[1]{\ensuremath{\langle\!\langle #1 \rangle\!\rangle}}
-\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
-\newcommand{\ad}{\ensuremath{\mathrm{ad}}}
-\newcommand{\g}{\ensuremath{\mathfrak{g}}}
-\newcommand{\argmin}{\operatornamewithlimits{argmin}}
-\newcommand{\argmax}{\operatornamewithlimits{argmax}}
-\graphicspath{{./Fig/}}
-
-\begin{document}
-
-\title{Autonomous Exploration with Exact Inverse Sensor Models%Exact Occupancy Grid Mapping and Autonomous Exploration%\thanks{Grants or other notes
-%about the article that should go on the front page should be
-%placed here. General acknowledgments should be placed at the end of the article.}
-}
-
-%\titlerunning{Short form of title}        % if too long for running head
-
-\author{Evan Kaufman \and Kuya Takami \and \\ Taeyoung Lee \and Zhuming Ai}%
-%\thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 {\tt \{evankaufman,tylee\}@gwu.edu}}
-%\thanks{Zhuming Ai, Ira S. Moskowitz, Information Management \& Decision Architectures, U.S. Naval Research Laboratory,  Washington, DC 20375}
-%}
-
-%\authorrunning{Short form of author list} % if too long for running head
-
-% Expanded:
-\institute{E. Kaufman \at
-              Dept. of Mechanical and Aerospace Engineering \\
-		 The George Washington University \\
-		 Washington DC 20052 \\
-              Tel.: 216-346-0946\\
-              \email{evankaufman@gwu.edu} 
-           \and
-           K. Takami \at
-              Dept. of Mechanical and Aerospace Engineering \\
-		 The George Washington University \\
-		 Washington DC 20052 \\
-              Tel.: 414-212-5155\\
-              \email{kuya@gwu.edu} 
-           \and
-           T. Lee \at
-              Dept. of Mechanical and Aerospace Engineering \\
-		 The George Washington University \\
-		 Washington DC 20052 \\
-              Tel.: 202-994-8710\\
-		 Fax: 202-994-0238\\
-              \email{tylee@gwu.edu}
-           \and
-           Z. Ai \at
-              Information Management \& Decision Architectures
-              \\ U.S. Naval Research Laboratory \\
-              Washington, DC 20375
-}
-
-% Condensed:
-\institute{E. Kaufman, K. Takami, and T. Lee \at
-              Dept. of Mechanical and Aerospace Engineering \\
-		 The George Washington University \\
-		 Washington DC 20052 \\
-              Tel.: 202-994-8710\\
-              \email{\{evankaufman,kuya,tylee\}@gwu.edu} 
-           \and
-           Z. Ai \at
-              Information Management \& Decision Architectures
-              \\ U.S. Naval Research Laboratory \\
-              Washington, DC 20375
-}
-
-\date{Received: date / Accepted: date}
-% The correct dates will be entered by the editor
-
-\maketitle
-
-\begin{abstract}
-This paper is focused on probabilistic occupancy grid mapping and motion planning such that a robot may build a map and explore a target area autonomously in real time.
-The desired path of the robot is developed in an optimal fashion to maximize the information gain from the sensor measurements on its path, thereby increasing the accuracy and efficiency of mapping, while explicitly considering the sensor limitations such as the maximum sensing range and viewing angle.
-Most current exploration techniques require frequent human intervention, often developed for omnidirectional sensors with infinite range.
-The proposed research is based on realistic assumptions on sensor capabilities.
-The unique contribution is that the mapping and autonomous exploration techniques are systematically developed in a rigorous, probabilistic formulation.
-The mapping approach exploits the probabilistic properties of the sensor and map explicitly, and the autonomous exploration is designed to maximize the expected map information gain, thereby improving the efficiency of the mapping procedure and the quality of the map substantially.
-The efficacy of the proposed optimal approach is illustrated by both numerical simulations and experimental results.
-%%%% Before minor modifications restructuring the abstract by Evan Kaufman on October 14, 2016:
-%This paper is focused on probabilistic occupancy grid mapping and motion planning such that a robot may build a map and explore a target area autonomously in real time.
-The desired path of the robot is developed in an optimal fashion to maximize the information gain from the sensor measurements on its path, thereby increasing the accuracy and efficiency of mapping, while explicitly considering the sensor limitations such as the maximum sensing range and viewing angle.
-Most current exploration techniques require frequent human intervention, and they are frequently developed for omnidirectional sensors with infinite range.
-%%For example, when applying SLAM techniques to explore an unknown area, a human operator must provide the path of the vehicle while continuously monitoring the updated map. 
-%The proposed research will greatly improve autonomy of robotic exploration in a systematic optimal manner, under realistic assumptions on sensor capabilities.
-The unique contribution is that the mapping and autonomous exploration techniques are systematically developed in a rigorous, probabilistic formulation.
-The mapping approach exploits the probabilistic properties of the sensor and map explicitly, and the autonomous exploration is designed to maximize the expected map information gain, thereby improving the efficiency of the mapping procedure and the quality of the map substantially.
-The efficacy of the proposed approach is illustrated by both numerical simulations and experimental results.
-\keywords{Occupancy Grid Mapping \and Inverse Sensor Model \and Autonomous Exploration}
-% \PACS{PACS code1 \and PACS code2 \and more}
-% \subclass{MSC code1 \and MSC code2 \and more}
-\end{abstract}
-
-\section{Introduction}
-\label{intro}
-Robotic mapping is the process of generating maps, which represent the environment in close proximity of a robot. 
-%Mapping in the context of robotics is to generate maps, which represent the environment in close proximity of a robot. 
-This problem is crucial to understanding the surrounding regions when a robot enters an uncertain environment.
-Various map representations include occupancy grids~\cite{WolSuk05}, octomaps~\cite{WurHorBenStaBur10}, or feature-based maps%in a landmark representation
-~\cite{MonThrKolWeg02}.
-%To build maps in real time, a computationally-efficient algorithm is required.% \EditTL{rewrite this paragraph for mapping, without introducing SLAM here}
-
-For mobile robots to operate in an unknown environment autonomously, they should be able to construct a map of the environment by using onboard sensors, and estimate their position in the constructed map.
-Simultaneous localization and mapping problem (SLAM) serves to estimate the map of the environment and the pose of robots concurrently, and there have been various approaches developed for SLAM with successful demonstrations.
-
-However, there are certain issues with the current SLAM techniques.
-First, the vast majority of work deals only with the aspect of estimating the environment and the poses of vehicles.
-They are passive in the sense that SLAM is performed on incoming sensor measurements from vehicles following an arbitrary path.
-As such, human teleoperation and monitoring are often essential to guide the vehicles safely through unknown surroundings.
-Therefore, it is desired that an active motion planning scheme is developed and integrated with SLAM such that vehicles are able to determine their path without human intervention, and explore unknown areas autonomously.
-
-To achieve these, it is proposed that the problems of mapping and motion planning are integrated in a probabilistic formulation.
-More explicitly, the map is constructed in a probabilistic sense, in contrast to the common approaches where the map is defined by deterministic objects that are definitely occupied or free.
-One of the desirable features for the proposed stochastic formulation is that we can extract the information on the degree of uncertainty for the each part of the map, such that the motion of the vehicle is planned to minimize the map uncertainty in an optimal fashion.
-
-This is addressed in two steps.
-The first part of this paper focuses on probabilistic occupancy grid mapping given a stochastic model of a range sensor, and the second part deals with stochastic motion planning to construct optimal exploration strategies.
-
-\paragraph{Probabilistic Occupancy Grid Mapping}
-
-The proposed approach is based  on the occupancy grid representation, where the environment is decomposed into evenly-spaced grid cells, considered as either occupied or free.
-The probabilistic model of a range sensor is referred to as the \textit{forward sensor model}, which provides a probability of range measurements for the given pose of the vehicle and the exact map~\cite{ThrBurFox05}.
-One can characterize the performance of a range sensor, such as its accuracy or maximum range, completely with its forward sensor model.
-
-The mapping problem can be defined as finding the \emph{inverse sensor model}, which is the probability of occupancy for a grid cell based on range measurements and the configuration of the mobile robot, which is of fundamental importance with occupancy grid mapping.
-However, the exact solution to the inverse sensor model has not been utilized in occupancy grid mapping, because the computational requirements are too cumbersome to compute the exact model in real time~\cite{ThrBurFox05}.
-
-Instead, several techniques have been proposed to approximate the exact inverse sensor model.
-%Due to the computational limitations, two common approaches are applied to obtain a function close to the inverse sensor model.
-In the early occupancy grid mapping applications based on sonar sensors, \emph{ad hoc} techniques to obtain an inverse sensor model were developed based off highly-innacurate bouncing sound waves~\cite{MorElf85,Elf89}.
-%These occupancy grid mapping schemes employ \emph{ad hoc} techniques to obtain the inverse sensor model.
-These techniques rely on approximate functions to represent the inverse sensor model with questionable accuracy~\cite{ChoLynHutKanBurKavThr05} and applied to more modern sensors in~\cite{And09},~\cite{PirRutBisSch11}.
-
-The other popular approach to obtain an approximate inverse sensor model is to simulate maps, robot poses, and measurements and use a learning algorithm to approximate the inverse sensor model~\cite{ThrBurFox05,Thr01}.
-These approaches are undesirable in practice due to complexities associated with implementing a learning algorithm.
-For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples, or how many samples are required to obtain a reasonable approximation.
-Furthermore, it is challenging to apply any learning algorithm over the large dimensional space composed of maps, poses, and measurements.
-
-%and they only yield probabilistic approximations without strong mathematical justification.
-
-%These approaches simply treat occlusions with discrete probabilities of detection, regardless of whether one cell may completely, partially, or not block another cell in the sensor field of view.
-
-%Furthermore, both approaches follow a log-odds ratio Bayesian framework, imposing an additional Markov assumption.
-
-%These techniques are not also suitable for modern range sensor with the highly accurate bearing~\cite{Thr03}.
-Based on recent improvements in sensor technology, the occupancy grid mapping problem has been restructured using the fact that depth measurement bearings are known.
-\EditTL{do we need this paragraph?}
-
-This paper proposes a computationally efficient algorithm to construct the exact inverse sensor model.
-More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, this algorithm yields the a posteriori probabilities of occupancy of all the cells within the area covered by the range sensor from the range measurements.
-The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancies of the other cells blocked by it is irrelevant to the forward sensor model, and this property is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model.
-Furthermore, the proposed approach integrates a priori probabilities of occupancy and multiple range measurements according to the Bayesian framework to obtain more accurate maps.
-As such, it contrasts from the existing framework based on log-odds ratios that impose an additional Markov assumption.
-
-\paragraph{Optimal Autonomous Exploration}
-
-Next, this paper addresses autonomous exploration to plan the trajectories of the vehicle to generate a complete, accurate map~\cite{Yam97}.
-We present an approach where the robot chooses a trajectory designed to maximize information gain of the occupancy grid map.
-
-%Another major problem that this paper addresses is to determine an effective trajectory given little knowledge about the environment.
-Autonomous exploration solves this problem, by choosing robotic motion through uncertain space, while building a map that is used for subsequent navigation~\cite{Yam97}.
-
-A common approach to solve the autonomous exploration problem is known as frontier-based exploration, such as ~\cite{Yam97,Yam98}.
-This is the process of executing actions that move the robot toward the closest boundary between visited and unvisited space, known as a frontier.
-Then the robot takes measurements at this location such that the mapped territory expands, and thus the new frontiers are pushed back. This process is repeated until the map is well-known.
-Frontier-based exploration assumes that repeatedly moving toward the closest frontier and taking measurements are the best actions to gain new information about the map, though these systematic actions are not based on any consideration of the future uncertainty of the probabilistic map or optimality.
-%there is no guarantee that these actions satisfy any optimality criteria.
-
-%The uncertainty of an occupancy grid map is commonly measured with Shannon's entropy, which is a metric that measures map uncertainty.
-
-%Approaches aimed at determining the future uncertainty of an occupancy grid map commonly suffer from several approximations.
-%%Determining future occupancy grid map uncertainty requires predictive estimations, which suffer from several approximations. 
-%Most evident is that all approaches rely on inaccurate probabilistic occupancy grids.
-This mapping representation is typically based on heuristic approximations (e.g.~\cite{MorElf85,Elf89,ChoLynHutKanBurKavThr05,And09,PirRutBisSch11}) or simulated learning techniques \cite{ThrBurFox05,Thr01}.
-Map-based information gain is commonly determined with Shannon's entropy~\cite{CarDamKumCas15}, an uncertainty metric based on grid cell occupancy probability.
-Since this probability is heuristic or learned, any measure of entropy is subject to errors from approximated probability.
-%Furthermore, the expected entropy of an action is not directly calculated~\cite{CarDamKumCas15,StaGriBur05}.
-Instead, these particle filter-based approaches assume that expected entropy is equivalent to entropy based on expected measurement scans~\cite{JohStaPfaBur07}.
-This assumption is subject to inaccuracies due to the nonlinear nature of Shannon's entropy.
-%Furthermore, these approaches typically generate maps with a particle filter, which commonly require some tuning and may consume large amounts of onboard memory.
-%
-%In this paper, we propose an autonomous exploration technique based on expected entropy change that avoids the aforementioned issues. 
-%%The occupancy grid mapping scheme is based on~\cite{KauLeeAiMos16}, which provides the exact solution of occupancy grid mapping of a single measurement ray. 
-%We develop a solution for expected entropy gain based on exact occupancy grid mapping~\cite{KauLeeAiMos16}.
-This solution directly calculates expected entropy from a probabilistic map based on similar assumptions of those applied to occupancy grid mapping.
-%Though more accurate than several other approaches, the proposed solution requires large computational resources, making real-time implementation difficult in certain scenarios.
-%Thus, we provide the tools to approximate the exact solution that reduce the computational requirements with realistic assumptions.
-%The motion of the robot is formulated as an optimization problem, where the expected entropy change is minimized, or equivalently the expected information gain is maximized.
-
-Alternatively, there have been exploration techniques that are based on a quantitative measure of uncertainty.
-However, these approaches commonly suffer from several approximations.
-Most evident is that all approaches rely on inaccurate probabilistic occupancy grids.
-This mapping representation is typically based on heuristic approximations (e.g.~\cite{MorElf85,Elf89,ChoLynHutKanBurKavThr05,And09,PirRutBisSch11}) or simulated learning techniques \cite{ThrBurFox05,Thr01}.
-Map-based information gain is commonly determined with Shannon's entropy~\cite{StaGriBur05}, an uncertainty metric based on grid cell occupancy probability.
-Since this probability is heuristic or learned, any measure of entropy is subject to errors from approximated probability or ad hoc learning.
-Furthermore, the expected entropy of an action is not directly calculated.
-Instead, these particle filter-based approaches assume that expected entropy is equivalent to entropy based on expected measurement scans~\cite{JohStaPfaBur07}.
-This assumption is subject to inaccuracies due to the nonlinear nature of Shannon's entropy.
-For these reasons, more careful computation of map uncertainty is warranted.
-%Furthermore, these approaches typically generate maps with a particle filter, which commonly require some tuning and may consume large amounts of onboard memory.
-
-In this paper, we propose an autonomous exploration technique based on expected entropy change that avoids the aforementioned issues. 
-%The occupancy grid mapping scheme is based on~\cite{KauLeeAiMos16}, which provides the exact solution of occupancy grid mapping of a single measurement ray. 
-We develop a solution for expected entropy gain based on exact occupancy grid mapping~\cite{KauLeeAiMos16}.
-This solution directly calculates expected entropy from a probabilistic map based on similar assumptions of those applied to occupancy grid mapping.
-Though more accurate than several other approaches, the proposed solution requires large computational resources, making real-time implementation difficult in certain scenarios.
-Thus, we provide the tools to approximate the exact solution that reduces the computational requirements with realistic assumptions.
-The motion of the robot is formulated as an optimization problem, where the expected entropy change is minimized, or equivalently the expected information gain is maximized.
-Compared with the recent results (e.g.
-[12], [13], [16], [17]) that employ Rao-Blackwellized particle filters to deal with localization and autonomous exploration simultaneously with an approximate entropy change, this paper is based on an accurate probabilistic solution to map information gain, under the assumption that the robot pose is known.
-
-In short, the main contribution of this paper is proposing the exact inverse sensor model and constructing an exact Bayesian occupancy grid mapping based on that.
-Compared with the current mapping algorithms based on an approximate inverse sensor model, this approach yields more accurate occupancy probabilities.
-This constructs substantially more accurate occupancy grid maps with less uncertainty for the same set of measurements, and these are directly illustrated by numerical examples and experimental results.
-Additionally, we derive a computationally-efficient approach to predict map information gain for autonomous exploration with a novel structure that avoids several common assumptions and approximations.
-By computing the expected information precisely with the exact inverse sensor model, the proposed autonomous exploration approach constructs a more accurate map with limited sensor capabilities in an optimal fashion.
-While this paper is focused on mapping and exploration in two-dimensional environments, the presented results can be certainly utilized in SLAM or stochastic motion planning in three dimensions.
-
-\section{Problem Formulation}
-\label{sec:ProbForm}
-
-\subsection{Probabilistic Occupancy Grid Mapping}
-\label{subsec:POGM}
-Let a map $m$ be decomposed into $n_m$ evenly-spaced grid cells, where the $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,\ldots,n_m}$, that is defined as $\mathbf{m}_i=1$ when occupied, and $\mathbf{m}_i=0$ when free.
-The location and size of each grid cell is assumed known.
-Therefore, a map $m$ is defined by $\{\mathbf{m}_1,\ldots, \mathbf{m}_{n_m}\}$ ($2^{n_{m}}$ possible maps).
-
-Another random variable is defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience.
-The probability that the $i$-th cell is occupied is $P(\mathbf{m}_i)$, and the probability that it is free is $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$.
-The random variables $\mathbf{m}_i$ are mutually independent,
-\begin{align}
-P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i).
-\end{align}
-
-Consider a range sensor that provides scans of the surrounding environment in order to identify the closest occupied space.
-The location and the direction of the sensor, referred to as the \emph{pose} at time $t$, is denoted by $X_t=(x_t,R_t)$, where the planar position and direction of the robot at $t$ are denoted by $x_t\in\Re^2$ and $R_t\in\Sph^1=\braces{q\in\Re^2|\norm{q}=1}$, i.e., the attitude corresponds to a direction on a 2D plane.
-Let $X_{1:t}$ denote the history of poses from the initial time to the current time, i.e., $X_{1:t}=\{X_1,X_2,\ldots, X_t\}$.
-At each pose, the robot receives a 2D measurement \emph{scan}, composed of $n_z$ measurement \emph{rays}.
-These rays are 1D depth measurements of known direction from the current pose to the closest occupied space, subject to a known \emph{forward sensor model} $p(z_{t,l}|m,X_t)$, where $z_{t,l}$ is the $l$-th measurement ray of the $t$-th scan $Z_t=\braces{z_{t,1},z_{t,2},\ldots,z_{t,n_z}}$.
-The measurement history is denoted by $Z_{1:t}=\braces{Z_1,Z_2,\ldots,Z_t}$.
-
-The probability density function, namely $p(z_{t,l}|m,X_t)$ with respect to the depth of the $l$-th measurement ray conditioned on the map $m$ and the pose $X_t$ is commonly referred to as the \emph{forward sensor model}, which characterizes the corresponding depth sensor, such as the maximum range or accuracy.
-The forward sensor model satisfies  (i) the ranges of all depth measurements are positive and finite, and (ii) measurement rays cannot pass through occupied regions.
-Throughout this paper, we assume that the forward sensor model of the selected sensor is given.
-This can be determined empirically or analytically.
-For example, the \emph{beam model for range finders} satisfying  the above criteria is described in~\cite{ThrBurFox05}.
-
-Occupancy grid mapping provides occupancy probabilities based on robot poses and measurement scans.
-The goal is to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, commonly referred to as the \emph{inverse sensor model} for given forward sensor model $p(z_{t,l}|m,X_t)$ and the initial estimate of the map $P(m)$ with $X_{1:t}$ and $Z_{1:t}$.
-
-\subsection{Autonomous Exploration}
-\label{subsec:AE}
-Since occupancy grid mapping provides a probabilistic representation of surrounding space, this mapping scheme holds probabilistic information about the uncertainty of the map.
-Shannon's entropy is commonly used as a measure of uncertainty, such as in~\cite{StaGriBur05}.
-Given the probabilities of each grid cell of map $m$, Shannon's entropy is defined as the sum of the entropy of each cell,
-\begin{align}
-\label{eqn:ShannonsEntropyDef}
-H(&P(m))=%\nonumber\\&
--\sum_{i=1}^{n_m}\braces{
-P(\mathbf{m}_i)\log{P(\mathbf{m}_i})+P(\bar{\mathbf{m}}_i)\log{(P(\bar{\mathbf{m}}_i}))}.
-\end{align}
-The entropy of the $i$-th grid cell is maximized when $P(\mathbf{m}_i)=0.5$ (greatest uncertainty) and minimized when $P(\mathbf{m}_i)\in\braces{0,1}$ (smallest uncertainty).
-%; thus Shannon's entropy is a measure of uncertainty.
-
-The goal of autonomous exploration is to determine future poses that maximize the map information gain, or equivalently minimize the map entropy. 
-Since autonomous exploration is conducted in uncertain environments, a complete trajectory cannot be known before the robot takes measurements of the entire reachable space.
-Instead, an objective function governs the motion between intermediate poses, building an occupancy grid map along the way.
-
-Suppose at time $t$, $P(m|X_{1:t},Z_{1:t})$ is given. Let $X_c$ be a candidate pose under consideration for some future time step.
-The expected information gain $\mathcal I(X_c)$ is defined as the negative change of entropy
-\begin{align}
-\label{eqn:ObjFun}
-\mathcal I(X_c)&=H(P(m|X_{1:t},Z_{1:t}))%\nonumber\\&\quad
--\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right],
-\end{align}
-where $H(P(m|X_{1:t},Z_{1:t}))$ may be computed with \refeqn{ShannonsEntropyDef} and \\$\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right]$ is the expected entropy when the robot is located at $X_c$, where the expectation is taken over $Z_c$.
-Since autonomous exploration is formulated as a trajectory optimization, the goal is to repeatedly choose optimal $X_c^*$ that maximizes the objective function
-\begin{align}
-\label{eqn:Objective}
-X_c^*=\argmax_{X_c}{\mathcal I(X_c)},
-\end{align}
-where $X_c^*$ satisfies the inequality constraint to avoid collisions,
-\begin{align}
-\label{eqn:CollisionInequalityConstraint}
-P_\text{collision}(X)=1-\prod_{i\in\mathcal C_X}P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})\leq\beta,
-\end{align}
-where $\mathcal C_X$ is the set of grid cells falling inside a volume of preselected size around $X$ that may cause collision and $\beta>0$ is a threshold of the probability of collision.
-Once the robot has translated to $X_c^*$, the above process is repeated.
-
-\section{Occupancy Grid Mapping with an Exact Inverse Sensor Model}
-\label{sec:ISM}
-
-This section outlines issues with the common log-odds ratio formulation and perceived computational burdens with the exact solution to the inverse sensor model for occupancy grid mapping.
-Then, a method that exploits several mapping and probabilistic properties provides a computationally efficient exact solution to probabilistic occupancy grid mapping.
-
-\subsection{Mapping via Log-Odds Ratio}
-
-One of the popular frameworks of updating binary random variables with static state is with a binary Bayesian filter using the log-odds ratio formulation.
-The main idea is that instead of multiplying terms from prior and current time steps, the properties of logarithms allow these terms to be simply added, while avoiding truncation issues associated with probabilities close to $0$ or $1$~\cite{ThrBurFox05}.
-The log-odds ratio is also popular because learning techniques with forward models~\cite{Thr01},~\cite{Thr03} or ad hoc techniques~\cite{MorElf85},~\cite{Elf89} become simplified.%are simpler to derive when they need not consider $P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})$ as part of the inverse sensor model, but rather only $P(\mathbf{m}_i)$, which is typically fixed and uniform for all cells.
-
-However, the log-odds ratio formulation makes an assumption that is not consistent with the occupancy grid mapping problem.
-It is assumed that if $X_t$ and $\mathbf{m}_i$ are known, then $Z_t$ is independent of $X_{1:t-1}$ and $Z_{1:t-1}$, i.e.,
-%\begin{align}
-%\label{eqn:FirstBayesRule}
-%P(\mathbf{m}_i|X_{1:t},Z_{1:t})
-%&
-%=\frac{P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})}{P(Z_t|X_{1:t},Z_{1:t-1})}.
-%\end{align}
-\begin{align}
-\label{eqn:AssumptionEarly}
-P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})\approx P(Z_t|\mathbf{m}_i,X_t).
-\end{align}
-In other words, the past history of measurements and poses are irrelevant to the current measurement.
-However, the prior measurements and poses provide a certain information to the occupancy map, which affects the probabilistic model of the current measurement.
-
-More specifically, forward models are based on the entire map $m$.
-Since \refeqn{AssumptionEarly} conditioned only by the $i$-th cell, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may contain important information about the occupancy of the other cells, which should be considered in calculating the probability distribution of the measurement scan $Z_t$.
-For example, $X_{1:t-1}$ and $Z_{1:t-1}$ might suggest that the $i$-th cell is occluded, so the approximation of \refeqn{AssumptionEarly} would not be justified.
-In short, valuable information about the map that can be obtained by the past measurements is ignored in the common log-odds ratio formulation.
-The probabilistic mapping proposed in this paper avoids this assumption of the log-odds ratio framework to construct a more accurate map by utilizing all of the past and the current measurements.
-
-%In short, this log-odds ratio framework is based on the assumption that neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$.
-
-%However, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may suggest that there exist occupied regions that occlude some or all of $Z_t$ from detecting $\mathbf{m}_i$ from the current pose $X_t$.
-%Since this assumption neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$, this framework is subject to poor results.
-%Hence, the log-odds ratio formulation is completely avoided in further analysis.
-
-	% approximated inverse sensor model
-	% markov assumption
-	% shortcoming, reason this is avoided
-
-%
-%
-%
-%
-
-\subsection{Inverse Sensor Model}
-
-%Due to computational complexity, approximate inverse sensor models have been used for occupancy grid mapping with simplifying assumptions.
-
-In this section, we propose an algorithm to compute the exact inverse sensor model efficiently. 
-%Much like the forward sensor model that reasons measurements from a known map, he inverse sensor model reasons that map from known measurements~\cite{ThrBurFox05}.
-%While the inverse sensor model varies slightly in meaning in numerous publications, two versions are derived in this paper, both based on a Bayesian framework.
-%We present two versions according to a Bayesian framework.
-The first type is referred to as the \emph{ray inverse sensor model} $P(\mathbf{m}_i|X_{1:t},Z_{1:t-1},z_{t,l})$, which only considers $l$-th ray at time $t$, i.e., one-dimensional inverse sensor model.
-The second type is referred to as the \emph{scan inverse sensor model} $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, which combines the information from \emph{all} measurement rays \emph{simultaneously} at the $t$-th time step for two-dimensional cases.
-These are developed in a sequential form to be repeated at each time step.
-%These are contrary to inverse sensor models commonly used with the log-odds ratio formulations, and the proposed algorithm integrates all of past poses and measurement scans with the current ones to obtain more accurate estimate of the map.
-Suppose the probability of the map conditioned on the past poses and measurements, namely $P(m|X_{1:t-1},Z_{1:t-1})$, is known.
-Here we construct a posteriori probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$, based on the current pose $X_t$, the measurement from the $l$-th ray $z_{t,l}$, and the given forward sensor model $p(z_{t_l}|m,X_t)$.
-
-\paragraph{Bayesian Framework}
-%Assuming that the probability density of thxe sensor measurement is fixed over a small measurement region $\Delta z$, 
-Bayes' rule yields
-%First, Bayes' rule is applied for a particular map $m$ and arbitrarily small measurement region $\Delta z$, yielding
-\begin{align}
-\label{eqn:BayesRuleRayISM}
-P(&m|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
-=\frac{p(z_{t,l}|m,X_{1:t},Z_{1:t-1})P(m|X_{1:t-1},Z_{1:t-1})}{p(z_{t,l}|X_{1:t},Z_{1:t-1})},
-\end{align}
-considering that $X_t$ carries no information about $m$ without $Z_t$.
-If the current pose $X_t$ and map $m$ are known, then the measurement ray $z_{t,l}$ is independent of past poses $X_{1:t-1}$ and past measurements $Z_{1:t-1}$ to obtain
-%Note that this is contrary to approaches that consider $X_t$ and $\mathbf{m}_i$ (not the full map $m$) as enough information, ignoring the occupancy of other cells in this term.
-%Then, \refeqn{BayesRuleRayISM} can be rewritten as
-\begin{align}
-P(m|z_{t,l},&X_{1:t},Z_{1:t-1})%\nonumber\\&
-=\eta_{t,l}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}),
-\end{align}
-where the normalizing constant $\eta_{t,l}\in\Re$ absorbs all terms that do not depend on the map $m$.
-
-%Assuming $z_{t,l}$ measures an occupied cell, then $z_{t,l}$ may be a result of the occupancy of $\mathbf{m}_i$, or $z_{t,l}$ may measure a different occupied cell.
-
-Next, we compute the occupancy probability of each cell.
-Let $\mathcal{M}_i$ be the set of maps where the $i$-th cell is occupied, i.e., $\mathcal{M}_i =\{m\in\{0,1\}^{{n_m}}\,|\ \mathbf{m}_i=1\}$.
-To compute the probability of occupancy of the $i$-th cell, all possible combinations of map in $\mathcal{M}_i$ should be considered, i.e.,
-%All possible combinations of map $m$ must be considered in which $\mathbf{m}_i$ is occupied, denoted as $m:m(i)=\mathbf{m}_i$,
-\begin{align}
-\label{eqn:InvSenModWithProbDens}
-&P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
-=\eta_{t,l}\sum_{m\in\mathcal{M}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
-\end{align}
-Furthermore, to determine the normalizing constant $\eta_{t,l}$, the complement $\\P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in the similar manner.
-Since there are $2^{n_{m}-1}$ maps in $\mathcal{M}_i$, these equations require $2^{n_m}$ terms to compute the summation for the $i$-th grid cell and normalizer, and this process should be repeated for all other cells along the measurement ray.
-This is the main reason why the existing results are based on certain approximations of the above expression.
-
-\paragraph{Computationally Efficient Approach}
-
-We propose a computational algorithm to evaluate \refeqn{InvSenModWithProbDens} efficiently. 
-Since the cells outside of the sensor field of view (FOV) are not affected, we focus on a reduced map $r_l$ in the FOV of the $l$-th ray.
-This reduced map is chosen such that each cell of $r_l$ corresponds to a grid cell of $m$ that the $l$-th ray intersects, ordered by increasing distance, which is easily determined from geometry.
-Let $\mathbf{r}_{l,k}$ be the binary random variable representing the occupancy of the $k$-th cell of the $l$-th ray.
-The number of cells in the reduced map is denoted by $n_{r,l}\leq n_m$.
-
-Next, let $\mathbf{r}_{l,k+}$ correspond to the event that the $k$-th cell of the $l$-th ray is occupied, cells with lower index are free, and cells with greater index may or may not be occupied, i.e., event $\mathbf{r}_{l,k+}$ occurs when \\$r\in\braces{r\in\{0,1\}^{{n_{r,l}}}\,|\mathbf{r}_{l,1}=0,\mathbf{r}_{l,2}=0,\ldots,\mathbf{r}_{l,k-1}=0,\mathbf{r}_{l,k}=1}$.
-In other words, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray, and this concept is illustrated with a simple 1D example in Figure \ref{fig:show_rkplus}.
-\begin{figure}
-  \centering
-  \includegraphics[width=0.4\textwidth]{rkplus_1.png}
-    \centering
-  \includegraphics[width=0.4\textwidth]{rkplus_2.png}  
-  \centering
-  \includegraphics[width=0.4\textwidth]{rkplus_3.png}
-    \centering
-  \includegraphics[width=0.4\textwidth]{rkplus_4.png}
-  \caption{A robot located to the left of the 1D occupancy grid map $r_l$ composed of $n_{r,l}=6$ grid cells in $4$ cases.
-In each case, cells $1$--$3$ are free, cell $4$ is occupied, and cells $5$ and $6$ may or may not be occupied.
-Each case corresponds to the event $\mathbf{r}_{4+}$.}
-  \label{fig:show_rkplus}
-\end{figure}
-Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell.
-Based on this, we present the following computational algorithm to compute the inverse sensor model.
-
-\begin{prop}
-For the $l$-th measurement ray, the a posteriori probability of the occupancy of the $k$-th cell, namely the ray inverse sensor model, is given by
-\begin{align}
-\label{eqn:RayISMAnswer}
-P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}),
-\end{align}
-where the unnormalized probability of the inverse sensor model is defined as
-\begin{align}
-\label{eqn:Unnormalized}
-& \tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
-=P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\
-&\quad\times 
-\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
-%&\quad\times 
-p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]\nonumber\\
-&\quad + \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
-%&\quad\times 
-p(z_{t,l}|\mathbf{r}_{l,k+},X_t)P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1}),
-\end{align}
-where $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ is chosen for convenience and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading.
-The normalizer $\eta_{t,l}$ is given by
-\begin{align}
-\label{eqn:allEta}
-\eta_{t,l}
-&=
-\bigg[\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
-\end{align}
-and it is independent of the cell index $k$.
-\end{prop}
-\begin{proof}
-See Appendix.
-\end{proof}
-
-%
-%%respectively.
-%These have the follow solutions that may be obtained in a computationally-efficient manner:
-%\begin{align}
-%\label{eqn:tildePresult}
-%\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
-%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
-%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
-%\nonumber
-%\\
-%&\quad
-%+
-%\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
-%&\quad\times p(z_{t,l}|\mathbf{r}_{k+},X_t)P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1}),
-%\\
-%\label{eqn:tildePcomplementresult}
-%\tilde P(\bar{\mathbf{r}}_k|z_{t,l},&X_{1:t},Z_{1:t-1})
-%=P(\bar{\mathbf{r}}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
-%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
-%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
-%\nonumber
-%\\
-%&\quad
-%+
-%\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
-%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}).
-%\end{align}
-
-%
-%
-%Because $\mathbf{r}_k$ is binary, the following holds:
-%\begin{align}
-%\label{eqn:allEta}
-%\eta_{t,l}&=\frac1{\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})+\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})}\nonumber
-%\\
-%&=
-%\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
-%\end{align}
-%where $\eta_{t,l}$ is independent of the index $k$ of the reduced map.
-%Rearranging \refeqn{Unnormalized} and \refeqn{UnnormalizedCompl}, the ray inverse sensor model is 
-%\begin{align}
-%\label{eqn:RayISMAnswer}
-%P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
-%\\
-%P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
-%\end{align}
-%respectively, where $\eta_{t,l}$ comes from \refeqn{allEta} and the unnormalized probabilities come from \refeqn{tildePresult} and \refeqn{tildePcomplementresult}.
-
-Note that the a priori estimate, $P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ and its compliment $\\P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})=1-P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ are available at the $t$-th step.
-Then, \refeqn{RayISMAnswer} yields a sequential occupancy grid mapping that can be applied whenever new measurements are available.
-
-Compared with \refeqn{InvSenModWithProbDens}, where the terms of the summation should be repeated $2^{n_{r,l}}$ times \emph{per each cell of the reduced map}, the proposed expressions \refeqn{RayISMAnswer} and \refeqn{allEta} are \textit{substantially} simpler.
-In fact, if \refeqn{Unnormalized} and \refeqn{allEta} are obtained recursively, requiring the summation of $O(n_{r,l}+1)$ rather than $O(n_{r,l}\times2^{n_{r,l}})$ as previously thought for \emph{all cells of the reduced map}, this yields an algorithm that is $\mathbf{n_{r,l}\times2^{n_{r,l}}/(n_{r,l}+1)}$ \textbf{times faster}.
-
-\paragraph{Scan Inverse Sensor Model}
-The above one-dimensional ray inverse sensor model may be generalized into a two-dimensional scan inverse sensor model by updating the map sequentially for each ray, according to
-%The most natural way to generalize the ray inverse sensor model into a scan inverse sensor model is applying the conditions in a recursive way,
-\begin{align}
-\label{eqn:RayByRayScanISM}
-P&(\mathbf{m}_i|X_{1:t},Z_{1:t})%\nonumber\\&%=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1},z_{t,2},\ldots,z_{t,n_z})\nonumber\\&
-=P((\dots(((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z}).
-\end{align}
-In short, this approach updates the map occupancies along the first measurement ray, then repeats this process for each subsequent ray until all rays composing the measurement scan have updated the map.
-
-\subsection{Summary of Algorithm}
-
-In short, the proposed scan inverse sensor model is computed by \refeqn{RayISMAnswer}-\refeqn{RayByRayScanISM}.
-These are summarized in Table \ref{tab:RayByRayISM} as a computationally efficient, recursive algorithm that avoids several repeated calculations of the same quantity.
-This algorithm utilizes temporary variables, where the required computational resources are minimized by computing the updated occupancy probabilities of all grid cells along a measurement ray together.
-
-%% More detailed (completed subscripts)
-%\begin{table}[p]
-%\begin{tabular}{ l }
-%  Function: $P(\mathbf{r}_k|X_{1:t},Z_{1:t-1},z_{t,1:l})=\text{InverseSensorModel}(P(\mathbf{r}_k|X_{1:t},Z_{1:t-1},z_{t,1:l-1}),z_{t,l})$\\
-%  for all $k\in\braces{1,2,\ldots,n_{r,l}}$\\
-%  Note: all probabilities are conditioned on $X_{1:t},Z_{1:t-1},z_{t,1:l-1}$, so these are removed in probabilities.\\
-%  Initialize $\eta^{-1}_{t,l}=0$ and $P(\bar{\mathbf{r}}_{0})=P(\bar{\mathbf{r}}_{n_{r,l}+1})=1$;\\
-%  for all $k = 1,2,\ldots,n_{r,l}$\\
-%   \ \ $P(\mathbf{r}_{k+}) = P(\bar{\mathbf{r}}_{0:k-1})P(\mathbf{r}_{k})$;\\
-%   \ \ $P(\bar{\mathbf{r}}_{0:k})=P(\bar{\mathbf{r}}_{0:k-1})P(\bar{\mathbf{r}}_{k})$;\\
-%   \ \ $a_\text{temp}=P(\mathbf{r}_{k+})p(z_l|\mathbf{r}_{k+},X_t)$;\\
-%   \ \ $\tilde P(\mathbf{r}_{k}|z_{t,l})=P(\mathbf{r}_{k})\eta^{-1}_{t,l}+a_\text{temp}$;\\
-%   \ \ $\eta^{-1}_{t,l}=\eta^{-1}_{t,l}+a_\text{temp}$;\\
-%  end for\\
-%  $\eta^{-1}_{t,l}=\eta^{-1}_{t,l}+P(\bar{\mathbf{r}}_{0:k})p(z_\text{max})$;\\
-%  $P(\mathbf{r}_{k}|z_{t,l})=\eta_{t,l}\tilde P(\mathbf{r}_{k}|z_{t,l})$ for all $k\in\braces{1,2,\ldots,n_{r,l}}$;\\
-%  \end{tabular}
-%\caption{Inverse Sensor Model of an Individual Measurement Ray}
-%\label{tab:RayByRayISM}
-%\end{table}
-
-% More simple (subscripts omitted)
-\begin{table}[p]
-\begin{tabular}{ l }
-  Function: $P(\mathbf{r}_k|z)=\text{InverseSensorModel}(P(\mathbf{r}_k),z)$\\
-  for all $k\in\braces{1,2,\ldots,n_{r}}$\\
-  Initialize $\eta^{-1}=0$ and $P(\bar{\mathbf{r}}_{0})=1$;\\
-  for all $k = 1,2,\ldots,n_{r}$\\
-   \ \ $P(\mathbf{r}_{k+}) = P(\bar{\mathbf{r}}_{0:k-1})P(\mathbf{r}_{k})$;\\
-   \ \ $P(\bar{\mathbf{r}}_{0:k})=P(\bar{\mathbf{r}}_{0:k-1})P(\bar{\mathbf{r}}_{k})$;\\
-   \ \ $a_\text{temp}=P(\mathbf{r}_{k+})p(z_l|\mathbf{r}_{k+},X)$;\\
-   \ \ $\tilde P(\mathbf{r}_{k}|z)=P(\mathbf{r}_{k})\eta^{-1}_{t,l}+a_\text{temp}$;\\
-   \ \ $\eta^{-1}=\eta^{-1}+a_\text{temp}$;\\
-  end for\\
-  $\eta^{-1}=\eta^{-1}+P(\bar{\mathbf{r}}_{0:k})p(z_\text{max})$;\\
-  $P(\mathbf{r}_{k}|z)=\eta\tilde P(\mathbf{r}_{k}|z)$ for all $k\in\braces{1,2,\ldots,n_{r}}$;\\
-  \end{tabular}
-\caption{Inverse Sensor Model of an Individual Measurement Ray}
-\raggedright{Note: this function is only considering $l$-th ray at the $t$-th time step, where probabilities are subject to conditions on $X_{1:t},Z_{1:t-1},z_{t,1:l-1}$, so these are removed from the algorithm for simplicity.}
-\label{tab:RayByRayISM}
-\end{table}
-
-In contrast to the current approximate inverse sensor models, the proposed algorithm evaluates the exact inverse sensor model \refeqn{InvSenModWithProbDens} efficiently without relying on assumptions like \refeqn{AssumptionEarly}.
-Therefore, it yields substantially more accurate maps for the same set of measurements.
-These are illustrated by numerical examples and experiments in Sections \ref{sec:NumRes} and \ref{sec:ExpRes}.
-
-\section{Computation of Expected Information Gain}
-\label{sec:ExpectedInfoGain}
-
-In this section, we present a method to determine the expected entropy from an individual measurement ray from a known pose.
-Based on this, an autonomous exploration algorithm will be developed in the next section.
-
-%This method is based on the exact solution to occupancy grid mapping from~\cite{KauLeeAiMos16}.
-
-\subsection{Single Ray Expected Value of Entropy}
-Suppose there is a probabilistic map constructed by past measurements, namely \\$P(m|X_{1:t},Z_{1:t})$.
-We wish to compute the expected entropy when the next measurement is taken at the candidate pose $X_c=(x_c,R_c)$, consisting of location $x_c\in\Re^2$ and attitude $R_c\in\Sph^1$.
-Assume that the direction of the $l$-th ray is determined by the attitude $R_c$.
-The expected entropy due to the $l$-th ray $z_{c,l}$ at position $x_c$ is
-\begin{align}
-\label{eqn:HRayInt}
-\text{E}[H(P(m|&X_{1:t},Z_{1:t},x_c,z_{c,l}))]
-\nonumber\\&
-=\int_{z_\text{min}}^{z_\text{max}}
-H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))%\nonumber\\&\qquad\times 
-p(z_{c,l}|X_{1:t},Z_{1:t},x_c)
-dz_{c,l}.
-\end{align}
-%where $z_\text{min}$ and $z_\text{max}$ are sensor limits. Considering the continuous space of possible measurements of $z_{c,l}$ would make \refeqn{HRayInt} intractable to calculate. Thus, w
-We discretize the measurement ray space such that $z_{c,l}$ falls on points along the measurement ray intersecting with grid cell edges.
-%, similar with how occupancy grid mapping decomposes continuous space into grid cells of length $\alpha$.
-The discretized expected value of \refeqn{HRayInt} is
-\begin{align}
-\label{eqn:DiscExpEntropyRay}
-\text{E}[H(P(&m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]
-\nonumber\\&=\sum_{k=1}^{n_{r,l}+1}\bigg\{H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l,k}))%\nonumber\\&\qquad\times 
-P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)\bigg\},
-\end{align}
-where index $z_{c,l,k}$ is the distance from position $x_c$ to the $k$-th grid cell of the $l$-th reduced map $r_l$, known from geometry.
-
-Standing alone, the term $P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)$ from \refeqn{DiscExpEntropyRay} has a convoluted meaning because the depth $z_{c,l,k}$ does not directly depend on the map.
-However, we present a method to obtain this discretized probability.
-%Despite how this term is replaced with a normalizer in ~\cite{ThrBurFox05} replaces
-Following the assumption that $z_{c,l}$ is discretized to known distances, the probabilities are proportional to their densities, as the area under the density curve is infinitesimal and fixed.
-Accounting for all cases, the probability is
-\begin{align}
-\label{eqn:ProbWithDelta}
-P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)=\frac{p(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)}{\sum_{i=1}^{n_{r,l}+1}p(z_{c,l,i}|X_{1:t},Z_{1:t},x_c)}.
-\end{align}
-Conveniently, this density corresponds to the inverse normalizer according to \refeqn{BayesRuleRayISM},
-\begin{align}
-\label{eqn:DiscretizedProb}
-p(z_{c,l,k}|X_{1:t},X_c,Z_{1:t})=\eta_{c,l,k}^{-1},
-\end{align}
-where $\eta_{c,l,k}^{-1}$ is defined in \refeqn{allEta}.
-%However, instead of evaluating the $t$-th pose $X_t$ and measured ray $z_{t,l}$, we simply consider $X_c$ and ray depth $z_{c,l,k}$ based on histories $X_{1:t}$ and $Z_{1:t}$ to obtain the inverse normalizer,
-Updating variable subscripts, the normalizer becomes
-\begin{align}
-\label{eqn:allEtaAtC}
-\eta_{c,l,k}^{-1}
-&=
-\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t},Z_{1:t})\bigg\}%\nonumber\\&\qquad\times 
-p(z_{c,l,k}|\mathbf{r}_{l,i+},X_c)P(\mathbf{r}_{l,i}|X_{1:t},Z_{1:t}).
-\end{align}
-By substituting \refeqn{ProbWithDelta} and \refeqn{DiscretizedProb} into \refeqn{DiscExpEntropyRay}, the expected entropy at $X_c$ from individual measurement ray $z_{c,l}$ is
-\begin{align}
-\label{eqn:HRayComplete}
-\text{E}[H(P&(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]\nonumber\\&=
-\left(\sum_{i=1}^{n_{r,l}+1}\eta_{c,l,i}^{-1}\right)^{-1}%\nonumber\\&\quad\quad\times
-\sum_{k=1}^{n_{r,l}+1}\bigg\{H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l,k}))\eta_{c,l,k}^{-1}\bigg\},
-\end{align}
-where $\eta_{c,l,k}^{-1}$ is taken from \refeqn{allEtaAtC}.
-Here, \refeqn{HRayComplete} provides the expected entropies for those cells inside the field of view of $r_l$. %; cells outside this reduced map are trivially changed by $0$ due to $z_{c,l}$.
-The entropies of cells outside the field of view remain unchanged.
-
-\subsection{Approximation of Expected Ray Entropy}
-
-The order of computation for each measurement ray is $\mathcal O((n_{r,l}+1)^2)$ since the summations of \refeqn{allEtaAtC} are embedded in \refeqn{HRayComplete}.
-However, several of those intersections provide negligible information since the probability of the $l$-th measurement ray capturing certain cell depths is close to zero.
-
-The approximation of expected ray entropy provides a method to reduce the computation of \refeqn{allEtaAtC} substantially.
-This goal is achieved by systematically selecting a smaller set of grid cells to consider over the summations of \refeqn{allEtaAtC} and \refeqn{HRayComplete}.
-The smaller set is determined by the probability that each cell is captured by the measurement ray, known as the detection probability. This can be found recursively as
-\begin{align}
-\label{eqn:ProbOfFirstCell}
-&P(\mathbf{r}_{l,k+}|X_{1:t},Z_{1:t})%\nonumber\\&
-=\bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t},Z_{1:t})\bigg\}P(\mathbf{r}_{l,k}|X_{1:t},Z_{1:t}),
-\end{align}
-which is the probability that $\mathbf{r}_{l,k}$ is the closest occupied grid cell based on past poses and measurement scans, independent of cells beyond the $k$-th cell from $x_c$.
-Let $\hat n>0$ be a fixed number of grid cells for all rays such that $\hat n\leq n_{r,l}+1$ for all $l$.
-Let $\hat{r}_{l}$ correspond to the grid cells that yield the $\hat{n}$ maximum values of \refeqn{ProbOfFirstCell} (the $\hat n$ most likely ray detections), indexed by increasing distance from candidate location $x_c$.
-By replacing the reduced map $r_l$ with $\hat{r}_l$ and changing the summation limits to $\braces{1,2,...,\hat n}$ in \refeqn{allEtaAtC} and \refeqn{HRayComplete}, the order of computation is reduced to $\mathcal O({\hat{n}}^2)$.
-Even though the value of $n_{r,l}$ is different among various rays in general, $\hat n$ is fixed, so the computational order is fixed as well.
-In short, this method reduces the required computation substantially by systematically neglecting those grid cells with little effect.
-It can be noted that if $\hat n=n_{r,l}+1$, the ray objective function is computed without approximation.
-
-\subsection{1D Ray Expected Information Gain Algorithm}
-
-We present an algorithm pseudocode providing the necessary steps to obtain the objective function for a single measurement ray (Table \ref{tab:RayExpectedEntropyGain}).
-Much like the algorithm pseudocode for the inverse sensor model (Table \ref{tab:RayByRayISM}), the variable $a_\text{temp}$ serves as an intermediate variable designed to avoid repeated calculations.
-Since this algorithm operates as a function, fixed indices and condition variables are removed for simplification.
-
-\begin{table}[p]
-\begin{tabular}{ l }
-  Function: $\mathcal I_\text{ray}=\text{RayExpInfoGain}(x,P(r),z_{1:n_{r}})$\\
-  Define $P(\bar{\mathbf{r}}_{0})=P(\bar{\mathbf{r}}_{n_{r}+1})=1$;\\
-  for $k = 1,2,\ldots,n_{r}+1$\\
-   \ \ $P(\mathbf{r}_{k+}) = P(\bar{\mathbf{r}}_{0:k-1})P(\mathbf{r}_{k})$;\\
-   \ \ $P(\bar{\mathbf{r}}_{0:k})=P(\bar{\mathbf{r}}_{0:k-1})P(\bar{\mathbf{r}}_{k})$;\\
-  end for\\
-%  $P(\mathbf{r}_{(n_{r}+1)+}) = P(\bar{\mathbf{r}}_{0:n_{r}})$;\\
-  Find $\hat r\subset r$ corresponding to the $\hat n$ greatest values of\\
-  $\braces{P(\mathbf{r}_{1+}),P(\mathbf{r}_{2+}),\ldots,P(\mathbf{r}_{(n_{r}+1)+})}$;\\
-  Define $P(\hat{\bar{\mathbf{r}}}_{0})=1$;\\
-  for $k = 1,2,\ldots,\hat{n}$\\
-   \ \ $P(\hat{\mathbf{r}}_{k+}) = P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\mathbf{r}}_{k})$;\\
-   \ \ $P(\hat{\bar{\mathbf{r}}}_{0:k})=P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\bar{\mathbf{r}}}_{k})$;\\
-  end for\\
-  for $k_\text{m}=1,2,\ldots,\hat n$\\
-   \ \ Initialize $\eta^{-1}_{k_\text{m}}=0$;\\
-   \ \ for $k_\text{c}=1,2,\ldots,\hat n$\\
-   \ \ \ \ $a_\text{temp}=P(\hat{\mathbf{r}}_{k_\text{c}+})p(z_{k_\text{m}}|\hat{\mathbf{r}}_{k_\text{c}+},x)$;\\
-   \ \ \ \ $\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=P(\hat{\mathbf{r}}_{k_\text{c}})\eta^{-1}_{k_\text{m}}+a_\text{temp}$;\\
-   \ \ \ \ $\eta^{-1}_{k_\text{m}}=\eta^{-1}_{k_\text{m}}+a_\text{temp}$;\\
-   \ \ end for\\
-   \ \ $P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=\eta_{k_\text{m}}\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})$ for all $k_\text{c}=1,2,\ldots,\hat n$;\\
-   end for\\ 
-   $P(z_{k_\text{m}}|x)=\frac{\eta^{-1}_{k_\text{m}}}{\sum_{i=1}^{\hat n}\eta^{-1}_{i}}$ for $k_\text{m}=1,2,\ldots,\hat n$;\\
-
-   Initialize $\mathcal I_\text{ray}=0$;\\
-   for $k_\text{c}=1,2,\ldots,\hat n$\\
-   \ \ $\mathcal I_\text{ray}=\mathcal I_\text{ray}+H(P(\hat{\mathbf{r}}_{k_\text{c}}))$;\\
-   \ \ for $k_\text{m}=1,2,\ldots,\hat n$\\
-   \ \ \ \ $\mathcal I_\text{ray} = \mathcal I_\text{ray}-H(P(\hat{\mathbf{r}}_{k_\text{c}}|x_c,z_{k_\text{m}}))P(z_{k_\text{m}}|x_{c})$;\\
-   \ \ end for\\
-  end for\\
-  Return: $\mathcal I_\text{ray}$\\
-\end{tabular}
-\caption{Expected Information Gain from a Measurement Ray}
-\raggedright{Note: this function is only considering a single ray for a future time step, where probabilities are subject to conditions on $X_{1:t},Z_{1:t}$, so these are removed from the algorithm for simplicity.}
-\label{tab:RayExpectedEntropyGain}
-\end{table}
-
-\subsection{Numerical Justification for Approximation}
-
-The purpose of this numerical example is to provide evidence that the approximations are reasonable and increase the algorithm speed substantially.
-Since a measurement ray produces a depth measurement in a single direction, we only consider a 1D map where the grid cells have spacing $\alpha=0.2$ meters, and the properties of the range sensor are based on the Microsoft Kinect~\cite{PirRutBisSch11,KhoElb12} with maximum reading depth $z_\text{max}=4$m ($20$ grid cells inside the sensor FOV).
-The goal is to compare the expected entropy $\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$ from \refeqn{HRayComplete} and with an approximation $\text{E}[H_\text{approx}(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$, which only considers $\hat n$ grid cells with highest detection probability \refeqn{ProbOfFirstCell}.
-
-We consider $100$ probabilistic maps to obtain Monte Carlo results to evaluate the approximate entropy.
-In every Monte Carlo trial, each grid cell has an $80\%$ chance of being free and a $20\%$ chance of receiving an a priori probability uniformly distributed between $0$ and $1$.
-%We consider $100$ probabilistic maps to obtain Monte Carlo results to evaluate the approximate entropy.
-In every Monte Carlo trial, each grid cell has an $80\%$ chance of being free, and the remaining $20\%$ of grid cells are given an a priori probability uniformly distributed between $0$ and $1$.
-Several metrics serve to evaluate $\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$ with $\text{E}[H_\text{approx}(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$.
-The median expected entropy change is $\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]-H(P(m|X_{1:t},Z_{1:t}))=-0.83792$.
-The error for the $100$ Monte Carlo cases is defined simply as
-\begin{align}
-e_{H}&=\frac1{100}\sum_{i=1}^{100}\text{abs}\bigg(\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]\nonumber\\&\qquad\quad-\text{E}[H_\text{approx}(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]\bigg).
-\end{align}
-The Monte Carlo trials are repeated for $\hat n=\braces{1,2,\ldots,10}$ and the results are plotted in Figure \ref{fig:ApproxJust}.
-This example shows a typical case when the summation limits generated from \refeqn{ProbOfFirstCell} have only small effects on \refeqn{HRayComplete}, while providing very large improvements in reducing computation.
-
-\begin{figure}
-	\centering
-    	\begin{subfigure}[b]{0.35\textwidth}
-        		\includegraphics[width=\textwidth]{JustifyApprox_eH.pdf}
-        		\caption{Entropy Error}
-        		\label{fig:H_err}
-    	\end{subfigure}
-%	\hspace*{0.1\columnwidth}
-	\begin{subfigure}[b]{0.35\textwidth}
-        		\includegraphics[width=\textwidth]{JustifyApprox_t.pdf}
-        		\caption{Computation Ratio}
-        		\label{fig:H_comp_ratio}
-    	\end{subfigure}
-%	\begin{subfigure}[Computation Ratio]
-%		{\includegraphics[height=0.35\columnwidth]{JustifyApprox_tRatio.pdf}
-%	}
-%	\end{subfigure}
-
-\caption{The entropy error is decreased at the cost of increasing computation time in this Monte Carlo 1D measurement ray expected entropy case.}
-\label{fig:ApproxJust}
-\end{figure}
-
-\section{Autonomous Exploration}
-
-In this section, we develop an autonomous exploration scheme inside a 2D environment based on choosing movements that minimize map entropy and avoid collisions.
-
-\subsection{Pose Selection Optimization}
-
-The objective is to choose a future pose that minimizes the map uncertainty, such that the robot can autonomously move toward this optimal goal.
-The robot pose is selected among various positions and attitudes.
-Since searching over all locations and attitudes would require infinite computational resources, these search spaces are discretized into a limited number of robot attitudes and positions.
-The procedure involves three steps.
-First, the attitude that maximizes the information gain objective function is selected at each candidate future pose location.
-Second, the location with the maximum objective function of all candidates is selected.
-Thus, the optimal attitude and the optimal location compose the optimal future pose.
-Finally, the robot determines and follows a collision-free path to the optimal future pose.
-This process is repeated, resulting in autonomous exploration.
-
-%First, the optimal attitude for an arbitrary location is presented.
-Then, two methods to select pose candidates, namely \emph{expanding ring} and \emph{complete Cartesian}, are described below, with a brief discussion of the benefits and drawbacks of choosing each search space.
-Finally, the equations and conditions to determine an optimal future pose are presented.
-
-\paragraph{Attitude Optimization}
-Consider a collision-free pose candidate at an arbitrary location $x_c$.
-At this pose, consider $n_d$ evenly-spaced measurement rays oriented radially outward from the robot in a circular pattern (see red lines in Figure \ref{fig:OptProcess}).
-These rays are denoted $\braces{z_{c,1},z_{c,2},\ldots,z_{c,n_d}}$.
-Attitudes correspond to the rays such that the robot sensor is aligned with the ray direction, and these are denoted $\braces{R_{c,1},R_{c,2},\ldots,R_{c,n_d}}$, where the scan with attitude $R_{c,d}$ might cover several ray directions depending on the sensor field of view (FOV).
-We choose optimal attitude $R_c^*$ as the summation of the expected entropy changes covered by the scan,
-\begin{align}
-\label{eqn:FindRc}
-&R_c^*=\argmax_{R_{c,d}}\sum_{z_{c,i}\in R_{c,d}\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\qquad\quad\quad\quad\quad
--\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,i}))]\bigg).
-%\mathcal I_\text{ray}(x_c,z_{c,i})\right).
-\end{align}
-
-This method provides the attitude that maximizes the information gain at an arbitrary location in 2D space.
-The set of candidate positions that warrant consideration is determined with one of two methods, namely \emph{expanding ring} and \emph{complete Cartesian}, described next.
-
-\paragraph{Expanding Ring}
-The expanding ring technique is advantageous for searching local solutions quickly, only considering distant future poses when necessary.
-The key idea is that future candidate locations lie on a circular ``ring'' centered around the robot, evenly spaced around the ring (see red circles in Figure \ref{fig:OptProcess}), and the ring is expanded if certain criteria are not met.
-More explicitly, consider $n_c$ candidate locations denoted by $x_c\in\braces{x_1,x_2,\ldots,x_{n_c}}$ located with the distance $\delta$ away from the robot.
-Any location that violates inequality constraint \refeqn{CollisionInequalityConstraint} is excluded to avoid collisions.
-The set of optimal attitudes at each candidate location is $\braces{R^*_{1},R^*_{2},\ldots,R^*_{n_c}}$, obtained from \refeqn{FindRc}.
-The objective function \refeqn{ObjFun} is computed by a summation about the $n_d$ rays as %\refeqn{Objective},
-\begin{align}
-\label{eqn:ObjFunApprox}
-\mathcal I(x_c,R_c^*)&\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\qquad-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*,z_{c,i}))\right]\bigg),
-\\
-x_c^*&=\argmax_{x_c}\mathcal I(x_c,R_c^*).
-\end{align}
-At the optimal pose $X^*_c=(x^*_c,R^*_c)$, the resulting information gain must satisfy $\mathcal I(X_c^*)\geq\mathcal I_\text{min}$, where $\mathcal I_\text{min}$ is some minimum threshold for expected information gain; robot motion is only justified when the expected information gain is significantly large.
-If the minimum threshold is not met, the ring of candidate pose locations is increased such that $n_c$ and $\delta$ are multiplied by a scale function $\lambda>1$.
-This process is repeated until the expected information gain of the optimal pose is at least $\mathcal I_\text{min}$, or all candidates lie in collision zones or outside map limits.
-%A pseudocode of the process is found in Table \ref{tab:Alg_AutomousExploration}.
-
-The primary advantage of the expanding ring approach to search the 2D space is that the number of pose candidates need not be proportional to the map area, and that local maxima tend to fall within short distances of the current robot pose.
-The main drawback is that poses outside a small local region of the robot are frequently neglected, and the robot frequently executes repeated short motions, often without large information gains.
-
-\paragraph{Complete Cartesian}
-The complete Cartesian method to search the 2D space provides candidates located fixed distance $d$ apart in each Cartesian direction.
-This approach is advantageous for capturing expected information gains in regions outside of the immediate vicinity of the robot, but must be applied carefully to avoid issues with computational bottlenecks due to the potentially-large search space.
-
-Similar to the expanding ring approach, $n_c$ candidate pose locations are considered, where those violating \refeqn{CollisionInequalityConstraint} are neglected from further consideration.
-However, unlike the ring expanding approach, the candidates have different distances from the robot in general, so the objective function is modified to enforce a cost on the squared distance from the current pose location $x_t$ to the candidate location $x_c$, i.e.,
-\begin{align}
-\label{eqn:ObjFunApproxCompleteCartesian}
-\mathcal I&(x_c,R_c^*,x_t)\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*(x_c),z_{c,i}))\right]\bigg)-k_\text{dist}(x_t-x_c)^\T(x_t-x_c),
-\end{align}
-where weighting function $k_\text{dist}$ represents the sensitivity to avoiding large motions across the map.
-Furthermore, the same candidate locations are considered throughout the repeated processes of exploration, and the map cell occupancies are assumed static.
-Therefore, if candidate location $x_c$ fails to satisfy $\mathcal I(x_c,R_c^*,x_t)\geq\mathcal I_\text{min}$, then $x_c$ need \emph{never} be considered again.
-Avoiding these unnecessary calculations yields a greatly lowered computational burden, particularly late in exploration where several regions are collision-free but disadvantageous to revisit.
-
-\begin{figure}
-\vspace*{0.1\columnwidth}
-\centerline{
-	\includegraphics[height=0.35\columnwidth]{ExampleOptimalPose.png}
-}
-\begin{picture}(0,0)(0,0)
-\setlength{\unitlength}{0.1\columnwidth}\scriptsize
-\put(4.4,2.1){\color{blue}$X_t$}
-\put(5.9,1.9){\color{red}$x_1$}
-\put(2.9,2.0){\color{red}$x_5$}
-\put(4.4,3.6){\color{red}$x_3$}
-\put(4.4,0.6){\color{red}$x_7$}
-\put(5.4,3.0){\color{red}$x_2$}
-\put(3.3,3.1){\color{red}$x_4$}
-\put(3.3,1.0){\color{red}$x_6$}
-\put(5.4,1.0){\color{red}$x_8$}
-\put(6.8,3.1){\color{red}$z_{2,1}$}
-\put(4.5,3.1){\color{red}$z_{2,5}$}
-\put(5.6,4.1){\color{red}$z_{2,3}$}
-\put(5.6,2.2){\color{red}$z_{2,7}$}
-\put(6.5,3.7){\color{red}$z_{2,2}$}
-\put(4.9,3.8){\color{red}$z_{2,4}$}
-\put(5.0,2.5){\color{red}$z_{2,6}$}
-\put(6.5,2.5){\color{red}$z_{2,8}$}
-\put(6.1,2.9){$X_c^*$}
-\end{picture}
-\caption{This figure illustrates the expanding ring method.
-Initially, a robot (blue circle) views down and left (blue sector).
-Then, the robot considers $n_c=8$ poses (red circles) as possible candidates for where to move next.
-For each candidate, $n_d=8$ directions are considered (red lines, only displayed on candidate location $c=2$).
-Then, the expected entropy of each ray is calculated.
-The scan (red sector) covering those rays with the largest expected entropy decreases is chosen for each candidate, and the best candidate $X_c^*$ (black circle: location, black sector: scan) is chosen to maximize information gain.
-Finally, Dijkstra's algorithm provides the collision-free motion between $X_t$ and $X_c^*$, and the process is repeated.
-}
-\label{fig:OptProcess}
-\end{figure}
-
-%avoiding collisions subject to the inequality constraint \refeqn{CollisionInequalityConstraint}.
-
-\paragraph{Motion Planning}
-Once the optimal future pose $X_c^*$ is selected, the robot must traverse its environment to reach $X_c^*$ without collisions.
-The locations of cells on a grid-based map provide nodes for Dijkstra's algorithm~\cite{Dij59}.
-At each node, \refeqn{CollisionInequalityConstraint} must be satisfied for the node to be considered reachable, and edges between nodes correspond to neighboring grid cells.
-Dijkstra's algorithm provides a relatively simple motion planning strategy to move from $x_t$ to $x_c^*$ that avoids collisions and local minima.
-
-The trajectory to follow the path outlined by Dijkstra's algorithm is determined with a linear constrained least squares optimization, assuming the robot follows a polynomial trajectory with fixed speed.
-The starting and ending positions and attitudes may be constrained, and polynomials patched together for long trajectories share a common position and velocity with respect to time.
-Then, a controller on the robot tracks this trajectory until the robot falls within acceptable thresholds of the final optimal pose.
-Once the robot completes this motion, the entire process is repeated. %The complete autonomous exploration process is summarized in Table \ref{tab:Alg_AutomousExploration}.
-
-%The first pseudocode in table \ref{tab:Alg_AutomousExploration} outlines how a candidate pose is selected given an objective function. The second pseudocode in table \ref{tab:Alg_ObjFun}
-
-%\begin{table}[p]
-%\begin{tabular}{ l }
-%%  Define $n_c$: number of candidate locations\\
-%%  Define $n_d$: number of evenly-spaced measurement rays per candidate\\
-%  Given: $t=1$, $X_{1}$, $Z_{1}$, $P(m|X_{1},Z_{1})$ from the inverse sensor model\\
-%  TrajectoryComplete = false;\\
-%  while TrajectoryComplete = false\\
-%  \ \ if $X_{t+1}$ is unknown\\
-%  \ \ \ \ MotionPlanning = false;\\
-%  \ \ \ \ while MotionPlanning = false\\
-%  \ \ \ \ \ \ for $c = 1,2,\ldots,n_c$\\
-%  \ \ \ \ \ \ \ \ Candidate location: $x_{c}$ from evenly-spaced circle around\\
-%  \ \ \ \ \ \ \ \ $X_t$ with radius $\delta$;\\
-%  \ \ \ \ \ \ \ \ for $d=1,2,\ldots,n_d$\\
-%  \ \ \ \ \ \ \ \ \ \ Obtain reduced map $r_{c,d}$ with $n_{r_{c,d}}$ intersections\\
-%  \ \ \ \ \ \ \ \ \ \ through grid cells with depths $z_{c,d,1:n_{r,d}}$ from geometry;\\
-%  \ \ \ \ \ \ \ \ \ \ $\mathcal I_\text{ray}(x_c,z_{c,d})=\text{RayExpInfoGain}(x_c,P(r_{c,d}),z_{c,d,1:n_{r,d}})$;\\
-%  \ \ \ \ \ \ \ \ end for\\
-%  \ \ \ \ \ \ \ \ $R_{c}^*=\argmax_{R_c}\left(\sum_{z_{c,i}\in R_c \text{ FOV}}\mathcal I_\text{ray}(x_c,z_{c,i})\right)$;\\%\argmax_{R}{\sum_{d=1}^{n_d}\mathcal I_{\text{ray},c,d}|d\in\text{FOV}(R))}$;\\
-%  \ \ \ \ \ \ \ \ $\mathcal I(x_c,R_c^*)=\sum_{z_{c,i}\in R_c^* \text{ FOV}}\mathcal I_\text{ray}(x_c,z_{c,i})$;\\%\sum_{d\in \text{FOV}(R_c^*)}\mathcal I_{\text{ray}_{c,d}}$;\\
-%  \ \ \ \ \ \ end for\\
-%  \ \ \ \ \ \ $x_c^*=\argmax_{x_c}\mathcal I(x_c,R_c^*)$;\\
-%  \ \ \ \ \ \ $X_c^*=\braces{x_c^*,R_c^*}$;\\
-%  \ \ \ \ \ \ if $\mathcal I(X_c^*)<\mathcal I_\text{min}$\\
-%  \ \ \ \ \ \ \ \ $n_c=\text{round}(\lambda n_c)$;\\
-%  \ \ \ \ \ \ \ \ $\delta=\lambda \delta$;\\
-%  \ \ \ \ \ \ else\\
-%  \ \ \ \ \ \ \ \ Obtain $X_{t+1},X_{t+2},\ldots,X_{c}^*$ using Dijkstra's algorithm and minimum rotation;\\
-%  \ \ \ \ \ \ \ \ Reset $n_c$ and $\delta$ to initial values;\\
-%  \ \ \ \ \ \ \ \ MotionPlanning = true;\\  
-%  \ \ \ \ \ \ end if\\
-%  \ \ \ \ end while\\
-%%  $X_c^*=\argmax_{X}$\mathcal I_{\text{ray},c
-%%  \ \ $\mathcal D\subset\braces{1,2,\ldots,n_d}$ represents scan coverage of candidate rays\\
-%%  \ \ The $c$-th pose candidate: $X_{c}=\braces{x_{c},R_{c}}$;\\
-%%   end for\\
-%%  $X_c^*=\argmin_{X_c}({\Delta H(X_c)}|X_c\in\braces{X_1,X_2,\ldots,X_{n_c}})$;\\
-%%  Motion planner generates a collision-free path between $X_t$ to $X_c^*$\\
-%%  Move to $X_{t+1}$, obtain $Z_{t+1}$, and find $P(m|X_{1:t+1},Z_{1:t+1})$;
-%  \ \ end if\\
-%  \ \ t = t+1;\\
-%  \ \ Obtain $P(m|X_{1:t},Z_{1:t})$ from the inverse sensor model;\\
-%  \ \ if map $m$ is completely explored\\
-%  \ \ \ \ TrajectoryComplete = true;\\
-%  \ \ end if\\
-%  end while
-%
-%
-%\end{tabular}
-%\caption{Autonomous Exploration via Expected Information Gain from Expanding Ring Searches}
-%\label{tab:Alg_AutomousExploration}
-%\end{table}
-
-\section{Numerical Examples}
-\label{sec:NumRes}
-
-The efficacy of the proposed algorithms are verified by numerical examples.
-First, the occupancy grid mapping strategies are simulated in Section \ref{sec:NEOGC}, followed by numerical examples for autonomous exploration in Section \ref{sec:NEAE}.
-
-% It is common practice to approximate\refeqn{InvSenModWithProbDens} with a simplified function, frequently following the structure of the approximate inverse sensor model proposed in~\cite{And09}.
-The main idea of this approach is that the probability of a cell being occupied ($i$) near a measurement is high (measurement likely hits this cell), ($ii$) between the robot and the measurement is low (measurement passes through this cell), and ($iii$) beyond the measurement is unchanged (the robot cannot measure through a wall/object).
-The function is based on intuition, not thorough mathematical derivation.
-In short, this approach simplifies the inverse sensor model at the cost of accuracy.
-
-%The Kinect depth sensor is used as an example, with specifications from~\cite{KhoElb12}.
-With these parameters and the identical set of measurements, the exact solution is compared with a benchmark algorithm for a Kinect sensor taken from summarized as follows.
-
-\subsection{Occupancy Grid Mapping}\label{sec:NEOGC}
-
-First, we compare the proposed exact solution to the inverse sensor model summarized in Table \ref{tab:RayByRayISM} with an approximate algorithm presented in~\cite{PirRutBisSch11,KhoElb12}, summarized as follows.
-The probability that the $i$-th grid cell $\mathbf{m}_i$ is occupied conditioned on the measurement ray $z_{t,l}$ at the pose $X_t$ is the continuous function
-\begin{align*}
-%\label{eqn:ISM_Approx_1}
-&P(\mathbf{m}_i|z_{t,l},X_t)%\nonumber\\&
-=\begin{cases}
-0.3+(\frac{k}{\sigma\sqrt{2\pi}}+0.2)e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ \hat z_{l,i} \leq z_{t,l},%z_{t,l}\leq \hat z_{l,i},
-\\
-0.5+\frac{k}{\sigma\sqrt{2\pi}}e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{otherwise},
-\end{cases}
-\end{align*}
-which is based on the expected distance to the cell $\hat z_{l,i}$ with parameters $k=\sigma=0.6$.
-This follows the structure of the approximate inverse sensor model proposed in~\cite{And09}.
-The main idea of this approach is that the probability of a cell being occupied (i) near a measurement is high (measurement likely hits this cell), (ii) between the robot and the measurement is low (measurement passes through this cell), and (iii) beyond the measurement is unchanged (the robot cannot measure through a wall/object).
-Then, these probabilities are combined in a weighted fashion such that all measurements rays of scan $Z_t$ simultaneously update the same grid cell in a log-odds format,
-\begin{align*}
-%\label{eqn:ISM_Approx_2}
-&\log\left(\frac{P(\mathbf{m}_i|Z_{t},X_t)}{1-P(\mathbf{m}_i|Z_{t},X_t)}\right)
-%\nonumber\\&
-=
-\frac1{\sum_{z_{t,l}\in\mathbf{m}_i}\hat z_{l,i}}\sum_{z_{t,l}\in\mathbf{m}_i}\log\left(\frac{P(\mathbf{m}_i|z_{t,l},X_t)}{1-P(\mathbf{m}_i|z_{t,l},X_t)}\hat z_{l,i}\right).
-\end{align*}
-
-%Given the same poses and measurement sets, the proposed algorithm from \refeqn{RayISMAnswer}-\refeqn{RayByRayScanISM} are compared with the approximate algorithm described above.
-A probabilistic map consisting of NUMX by NUMY square grid cells with edge length ALPHA is assumed to accurately represent the occupancy of the environment surrounding a robot.
-The robot follows (SOMETHING ABOUT THE TRAJECTORY).
-With both algorithms, the probabilistic maps are updated in real time.
-%
-%(PARAGRAPH ABOUT MAP CLARITY, ENTROPY, AND COMPUATION TIME)
-
-Table~\ref{tab:penn} shows the mapping parameters and data specifications for the actual odometry and lidar measurements from University of Pennsylvania through Coursera open course~\cite{coursera}.
-Figure~\ref{fig:penn_map} shows the direct comparison of the resulting map based on the approximate and proposed inverse sensor model.
-As shown in the figure, both approaches capture the structure of the environment.
-However, the significant differences can be seen in the certain area of the map such as in Figure~\ref{fig:penn_map_zoom}.
-The differences can be clearly depicted in Figure~\ref{fig:entropy_comp} where entropy of maps are compared at frame 2280.
-The red part of the map shows the maximum  entropy whereas blue shows lowest entropy for the cell.
-Finally, overall improvement was demonstrated in Figure~\ref{fig:entropy} resulting improved entropy value per grid cell throughout the mapping process.
-In short, the proposed exact occupancy grid mapping produced a cleaner map with lower entropy from the same set of measurements.
-
-\begin{center}
-\captionof{table}{Experimental parameters provided from dataset}
-\label{tab:penn}
-    \begin{tabular}{r | c}
-        Parameters & Value\\ \hline\hline
-        Map dimension & 576x448 [pixel]\\
-        Resolution & 1/16 [m]\\
-        Scan angle & [-2.36, 2.36]\\
-        Ray number & 1081\\
-        Total frame & Every 120 scans out of 3701\\
-    \end{tabular}
-\end{center}
-
-\begin{figure}[!ht]
-    \centering
-    \begin{subfigure}[t]{0.8\columnwidth}
-        \centering
-        \includegraphics[trim={1.5cm 3.5cm 1.5cm 1.5cm},clip,width=\textwidth]{map_comparison.png}
-        \caption{Comparison of the total mapped environment based on the approximate (left) and proposed (right) inverse sensor model.}
-        \label{fig:penn_map_total}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.8\columnwidth}
-        \centering
-        \includegraphics[trim={1.5cm 3.5cm 1.5cm 1.5cm},clip,width=\textwidth]{map_zoom.png}
-        \caption{Evidence of significant improvement at certain area of the map.}
-        \label{fig:penn_map_zoom}
-    \end{subfigure}
-    \caption{Using experimental sensor data, the map was generated using an approximate and the proposed exact inverse sensor models.
-The laser sensor and odometry data used for the experiment was provided by University of Pennsylvania open course robotics estimation and learning on Coursera~\cite{coursera}.}
-    \label{fig:penn_map}
-\end{figure}
-
-\begin{figure}[!ht]
-    \centering
-    \begin{subfigure}[t]{0.35\columnwidth}
-        \centering
-        \includegraphics[width=\textwidth]{AISM_Image_inf_19.pdf}
-        \caption{Approximate inverse sensor model}
-        \label{fig:AISM}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.35\columnwidth}
-        \centering
-        \includegraphics[width=\textwidth]{EISM_Image_inf_19.pdf}
-        \caption{Exact inverse sensor model}
-        \label{fig:EISM}
-    \end{subfigure}
-    \caption{The entropy is compared between the approximate and proposed exact inverse sensor models at step 2280. The blue areas show the lowest entropy while red shows highest.}
-    \label{fig:entropy_comp}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.7\textwidth]{entropy_frame.png}
-  \caption{The map entropies of the approximate inverse sensor model (blue) and the proposed exact inverse sensor model (red) show that the exact version yields a more certain occupancy grid map.}
-  \label{fig:entropy}
-\end{figure}
-
-%\paragraph{Computation Time}
-%The proposed ray inverse sensor model is designed to maintain the accuracy of the exact solution with substantial computational improvements.
-%A $1$-dimensional example is simulated on a Mac desktop computer with $11$ grid cells, $10$ of which are visible to the robot, where the same depth measurement is used for all occupancy grid algorithms.
-%The computational time of the exact solution is computed in the conventional manner according to \refeqn{InvSenModWithProbDens} in $2.3314$ seconds to obtain the same solution of \refeqn{RayISMAnswer} from \refeqn{Unnormalized} and \refeqn{allEta} in just $0.0027$ seconds, which is $874.77$ times faster.
-%The approximate ray inverse sensor model completes in $0.0011$ seconds, but this algorithm does not follow a forward sensor model, so the probabilities of the map may be inaccurate.
-
-%\subsection{Trajectory Map Building}
-%In this numerical example, we consider a robot in a two-dimensional environment composed of ten wall edges, and the robot follows a figure-eight curve, then turns around and completes the same curve in the reverse direction.
-%The map is composed of $100\times100$ square grid cells with edge length of $1$ cm.
-%A forward sensor model $p(z_t|m,X_t)$~\cite{ThrBurFox05} is constructed according to the specifications of the Kinect sensor~\cite{KhoElb12}, and a pseudo-random measurement scan is sampled at each time step via the inverse transform sampling~\cite{DevBK86}.
-The same set of measurements updated each second are used with both occupancy grid mapping algorithms to construct the map.
-%
-%The resulting maps are illustrated in Figure \ref{fig:NumResOccProbs} for both algorithms, where it is shown that the proposed algorithm yields a substantially more accurate and clear map with less uncertainty.
-To quantify the degree of map uncertainty, we define the entropy of the map as
-%\begin{align*}
-%&H(P(m|X_{1:t},Z_{1:t}))=\nonumber\\
-%&-\sum_{i=1}^n\big\{P(\mathbf{m}_i|X_{1:t},Z_{1:t})\log P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\
-%&+(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\log(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\big\},
-%\end{align*}
-%which is maximized when the probability of occupancy is $0.5$ for all cells (more uncertain), and it is minimized when they are either $0$ or $1$ (less uncertain). 
-%
-%The change of the map entropy over time, and the entropy of the completed maps, for both methods are depicted in Figure \ref{fig:NumResOccH}.
-The subfigure (a) illustrates that the proposed exact inverse sensor model exhibits rapid decreases of entropies, and smaller entropies always.
-The resulting terminal map obtained from the proposed approach, shown in the subfigure (b), has less uncertainty than (c) constructed by the approximate model.
-In short, the proposed approach is more efficient at extracting information about the environment from the same set of set measurements.
-
-\subsection{Autonomous Exploration}\label{sec:NEAE}
-
-%Maximizing the proposed expected information gain serves as the objective for autonomous exploration.
-A robot begins in a completely uncertain environment, except for the grid cells inside the immediate vicinity the robot location.
-
-%At each time step, the robot receives a measurement scan, where the probabilistic properties of the sensor are taken from~\cite{PirRutBisSch11,KhoElb12}.
-Then, the $n_c=8$ evenly-spaced candidate locations about a circle of radius $\delta=0.5$m around the current pose location are considered, where $n_d=16$ measurement rays are evenly-spaced about the candidate location.
-When no current candidates yield expected information gains above $\mathcal I_\text{min}=1.25$, $\lambda=1.25$ is multiplied to the current $n_c$ and $\delta$ values.
-The motion of the robot is restricted to movement on grid cells satisfying \refeqn{CollisionInequalityConstraint} with $\beta=0.01$.
-Dijkstra's algorithm guides the robot through the environment without collision.
-
-Next, the proposed exploration approach is applied to the floor plan of Intel Research Lab, illustrated in Figure~\ref{fig:intel}.
-The robot explores its surroundings with an occupancy grid with $90,000$ cells where grid cell edges are $\alpha=0.1$m, composing a map with dimensions $30\text{m}\times30\text{m}$.
-The initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) for grid cells covered by the circular robot of radius $0.3$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
-
-The robot follows the complete Cartesian candidate search method to determine future poses, with $k_\text{dist}=5$ to avoid large motions with little added expected information gains.
-Due to the large size of the map, candidates below the expected information gain of $1.25$ are neglected from future consideration.
-
-The time sequence of the exploration for every 5~minutes are simulated in ROS Stage 2D environment. The corresponding simulation results are illustrated in Figure \ref{fig:IRL}.
-Lighter red dots and darker blue dots in the explored area correlate to the level of information gain at those positions with low and high value, respectively.
-It is shown that the proposed autonomous exploration algorithm successfully mapped the complicated environment composed of a number of rooms, narrow hallways, and open spaces that are irregularly shaped.
-
-%This approach not only achieved exploration of the entire environment but also improved computational efficiency.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.4\textwidth]{intel_clean.png}
-    \caption{Modified Intel Research Lab floor plan form the SLAM benchmark~\cite{kummerle2009measuring}.}
-\label{fig:intel}
-\end{figure}
-
-\begin{figure}[!ht]
-    \centering
-    \begin{subfigure}[t]{0.3\columnwidth}
-        \centering
-        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{0min.png}
-        \caption{$t=0$ min}
-        \label{fig:IRL0min}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.3\columnwidth}
-        \centering
-        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{5min.png}
-        \caption{$t=5$ min}
-        \label{fig:IRL5min}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.3\columnwidth}
-           \centering
-           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{10min.png}
-        \caption{$t=10$ min}
-        \label{fig:IRL10min}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.3\columnwidth}
-           \centering
-           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{15min.png}
-        \caption{$t=15$ min}
-        \label{fig:IRL15min}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.3\columnwidth}
-         \centering
-         \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{20min.png}
-        \caption{$t=20$ min}
-        \label{fig:IRL20min}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.3\columnwidth}
-           \centering
-           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{25min.png}
-        \caption{$t=25$ min}
-        \label{fig:IRL25min}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.3\columnwidth}
-           \centering
-           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{30min.png}
-        \caption{$t=30$ min}
-        \label{fig:IRL30min}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.3\columnwidth}
-           \centering
-           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{35min.png}
-        \caption{$t=35$ min}
-        \label{fig:IRL35min}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.3\columnwidth}
-           \centering
-           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{40min.png}
-        \caption{$t=40$ min}
-        \label{fig:IRL40min}
-    \end{subfigure}
-    \caption{A robot (green) measures a room in ROS Stage simulator using modified Intel Research Lab floor plan. The exact inverse sensor model is used for the autonomous exploration algorithm.}
-    \label{fig:IRL}
-\end{figure}
-
-%\section{Numerical Examples}
-%\label{sec:NumRes}
-%
-%The efficacy of the proposed algorithms are verified by numerical examples.
-First, the occupancy grid mapping strategies are simulated in Section \ref{sec:NEOGC}, followed by numerical examples for autonomous exploration in Section \ref{sec:NEAE}.
-%
-%% It is common practice to approximate\refeqn{InvSenModWithProbDens} with a simplified function, frequently following the structure of the approximate inverse sensor model proposed in~\cite{And09}.
-The main idea of this approach is that the probability of a cell being occupied ($i$) near a measurement is high (measurement likely hits this cell), ($ii$) between the robot and the measurement is low (measurement passes through this cell), and ($iii$) beyond the measurement is unchanged (the robot cannot measure through a wall/object).
-The function is based on intuition, not thorough mathematical derivation.
-In short, this approach simplifies the inverse sensor model at the cost of accuracy.
-%
-%%The Kinect depth sensor is used as an example, with specifications from~\cite{KhoElb12}.
-With these parameters and the identical set of measurements, the exact solution is compared with a benchmark algorithm for a Kinect sensor taken from summarized as follows.
-%
-%
-%
-%
-%\subsection{Occupancy Grid Mapping}\label{sec:NEOGC}
-%
-%First, we compare the proposed exact solution to the inverse sensor model summarized in Table \ref{tab:RayByRayISM} with an approximate algorithm presented in~\cite{PirRutBisSch11,KhoElb12}, summarized as follows.
-The probability that the $i$-th grid cell $\mathbf{m}_i$ is occupied conditioned on the measurement ray $z_{t,l}$ at the pose $X_t$ is the continuous function
-%\begin{align*}
-%%\label{eqn:ISM_Approx_1}
-%&P(\mathbf{m}_i|z_{t,l},X_t)%\nonumber\\&
-%=\begin{cases}
-%0.3+(\frac{k}{\sigma\sqrt{2\pi}}+0.2)e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{if}\ \hat z_{l,i} \leq z_{t,l},%z_{t,l}\leq \hat z_{l,i},
-%\\
-%0.5+\frac{k}{\sigma\sqrt{2\pi}}e^{-\frac12\left(\frac{\hat z_{l,i}-z_{t,l}}{\sigma}\right)^2}\ &\text{otherwise},
-%\end{cases}
-%\end{align*}
-%which is based on the expected distance to the cell $\hat z_{l,i}$ with parameters $k=\sigma=0.6$.
-This follows the structure of the approximate inverse sensor model proposed in~\cite{And09}.
-The main idea of this approach is that the probability of a cell being occupied (i) near a measurement is high (measurement likely hits this cell), (ii) between the robot and the measurement is low (measurement passes through this cell), and (iii) beyond the measurement is unchanged (the robot cannot measure through a wall/object).
-%Then, these probabilities are combined in a weighted fashion such that all measurements rays of scan $Z_t$ simultaneously update the same grid cell in a log-odds format,
-%\begin{align*}
-%%\label{eqn:ISM_Approx_2}
-%&\log\left(\frac{P(\mathbf{m}_i|Z_{t},X_t)}{1-P(\mathbf{m}_i|Z_{t},X_t)}\right)
-%%\nonumber\\&
-%=
-%\frac1{\sum_{z_{t,l}\in\mathbf{m}_i}\hat z_{l,i}}\sum_{z_{t,l}\in\mathbf{m}_i}\log\left(\frac{P(\mathbf{m}_i|z_{t,l},X_t)}{1-P(\mathbf{m}_i|z_{t,l},X_t)}\hat z_{l,i}\right).
-%\end{align*}
-%
-%Given the same poses and measurement sets, the proposed algorithm from \refeqn{RayISMAnswer}-\refeqn{RayByRayScanISM} are compared with the approximate algorithm described above.
-A probabilistic map consisting of NUMX by NUMY square grid cells with edge length ALPHA is assumed to accurately represent the occupancy of the environment surrounding a robot.
-The robot follows (SOMETHING ABOUT THE TRAJECTORY).
-With both algorithms, the probabilistic maps are updated in real time.
-%
-%(PARAGRAPH ABOUT MAP CLARITY, ENTROPY, AND COMPUATION TIME)
-%
-%%\paragraph{Computation Time}
-%%The proposed ray inverse sensor model is designed to maintain the accuracy of the exact solution with substantial computational improvements.
-%%A $1$-dimensional example is simulated on a Mac desktop computer with $11$ grid cells, $10$ of which are visible to the robot, where the same depth measurement is used for all occupancy grid algorithms.
-%%The computational time of the exact solution is computed in the conventional manner according to \refeqn{InvSenModWithProbDens} in $2.3314$ seconds to obtain the same solution of \refeqn{RayISMAnswer} from \refeqn{Unnormalized} and \refeqn{allEta} in just $0.0027$ seconds, which is $874.77$ times faster.
-%%The approximate ray inverse sensor model completes in $0.0011$ seconds, but this algorithm does not follow a forward sensor model, so the probabilities of the map may be inaccurate.
-%
-%
-%
-%
-%
-%
-%
-%
-%%\subsection{Trajectory Map Building}
-%%In this numerical example, we consider a robot in a two-dimensional environment composed of ten wall edges, and the robot follows a figure-eight curve, then turns around and completes the same curve in the reverse direction.
-%%The map is composed of $100\times100$ square grid cells with edge length of $1$ cm.
-%%A forward sensor model $p(z_t|m,X_t)$~\cite{ThrBurFox05} is constructed according to the specifications of the Kinect sensor~\cite{KhoElb12}, and a pseudo-random measurement scan is sampled at each time step via the inverse transform sampling~\cite{DevBK86}.
-The same set of measurements updated each second are used with both occupancy grid mapping algorithms to construct the map.
-%%
-%%The resulting maps are illustrated in Figure \ref{fig:NumResOccProbs} for both algorithms, where it is shown that the proposed algorithm yields a substantially more accurate and clear map with less uncertainty.
-To quantify the degree of map uncertainty, we define the entropy of the map as
-%%\begin{align*}
-%%&H(P(m|X_{1:t},Z_{1:t}))=\nonumber\\
-%%&-\sum_{i=1}^n\big\{P(\mathbf{m}_i|X_{1:t},Z_{1:t})\log P(\mathbf{m}_i|X_{1:t},Z_{1:t})\nonumber\\
-%%&+(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\log(1-P(\mathbf{m}_i|X_{1:t},Z_{1:t}))\big\},
-%%\end{align*}
-%%which is maximized when the probability of occupancy is $0.5$ for all cells (more uncertain), and it is minimized when they are either $0$ or $1$ (less uncertain). 
-%%
-%%The change of the map entropy over time, and the entropy of the completed maps, for both methods are depicted in Figure \ref{fig:NumResOccH}.
-The subfigure (a) illustrates that the proposed exact inverse sensor model exhibits rapid decreases of entropies, and smaller entropies always.
-The resulting terminal map obtained from the proposed approach, shown in the subfigure (b), has less uncertainty than (c) constructed by the approximate model.
-In short, the proposed approach is more efficient at extracting information about the environment from the same set of set measurements.
-%
-%
-%
-%
-%
-%
-%\subsection{Autonomous Exploration}\label{sec:NEAE}
-%
-%%Maximizing the proposed expected information gain serves as the objective for autonomous exploration.
-A robot begins in a completely uncertain environment, except for the grid cells inside the immediate vicinity the robot location.
-%
-%WILL UPDATE ONCE RESULTS ARE BETTER
-%The robot explores its surroundings with an occupancy grid with $15000$ cells where grid cell edges are $\alpha=0.2$m, composing a map with dimensions $30\text{m}\times20\text{m}$.
-The initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) for grid cells covered by the circular robot of radius $0.1$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
-%At each time step, the robot receives a measurement scan, where the probabilistic properties of the sensor are taken from~\cite{PirRutBisSch11,KhoElb12}.
-Then, the $n_c=8$ evenly-spaced candidate locations about a circle of radius $\delta=0.5$m around the current pose location are considered, where $n_d=32$ measurement rays are evenly-spaced about the candidate location.
-When no current candidates yield expected information gains above $\mathcal I_\text{min}=2$, $\lambda=1.25$ is multiplied to the current $n_c$ and $\delta$ values.
-The motion of the robot is restricted to movement on grid cells satisfying \refeqn{CollisionInequalityConstraint} with $\beta=0.01$.
-Dijkstra's algorithm guides the robot through the environment without collision.
-The results are illustrated in Figure \ref{fig:AutonomousExploration}.
-
-\section{Experimental Results}
-\label{sec:ExpRes}
-	% Kinect result
-
-%A preliminary experiment shows how the exact inverse sensor model is successfully implemented with the Kinect sensor.
-
-The efficacy of the proposed mapping and autonomous exploration algorithms are also verified with experiments.
-The experiment involves a ground vehicle producing a probabilistic occupancy grid map in real time.
-The robot motion is governed by the information gain-maximizing policy described in this paper.
-
-\paragraph{Hardware Configuration}
-The Pioneer 3 ground robot was chosen.
-The vehicle accepts two inputs, namely linear velocity (aligned with the robot wheels) and angular velocity (about a central axis passing vertically through the robot).
-The robot pose was estimated via Vicon Tracker, which provides the location and attitude of a rigid body.
-The depth readings were obtained by a Kinect depth sensor.
-Since the experiment was to map and explore a 2D environment, only a single central row of the 3D Kinect depth scan provided the necessary measurements.
-An environment with features of walls and obstacles was constructed by Styrofoam.
-
-\paragraph{Software Configuration}
-The proposed algorithm was implemented with ROS.
-The ROS framework provided seamless communication betweens sensors and actuators, and the executable programs (referred to as ``nodes'') operate quickly when compiled from C++ code.
-The robot and map are plotted using ``OpenGL'' libraries.
-
-Synchronization between the sensors is of high importance to the proposed occupancy grid mapping approach.
-Any time delay between the robot pose estimation and the sensor depth readings can cause conflicting information, which may be harmful to the probabilistic map.
-Thus, ROS approximate message filters were applied such that the pose from Vicon Tracker ($100$Hz) and the Microsoft Kinect ($30$Hz) were nearly synchronized, which provided validity of the assumption that the measurement ray positions and directions are known deterministically.
-
-\paragraph{Exploring and Mapping a 2D Environment}
-The environment consisted of Styrofoam walls on a flat floor.
-The parameter was oriented in a rectangle, with several obstacles and angled walls (see Figure \ref{fig:ExpSetupPhoto}).
-The robot autonomously explored the space using the complete Cartesian searching approach, and Dijkstra's algorithm provided collision-free waypoints to the optimal future poses.
-Based on these waypoints, a constrained least squares polynomial fitting method provided a smooth trajectory to follow the waypoints with fixed velocity.
-Finally, a simple controller was implemented such that the Pioneer moved along the trajectory by moving and orienting toward the desired robot location $1$sec in the future.
-
-\begin{figure}
-	\centering
-    	\begin{subfigure}[b]{0.35\textwidth}
-        		\includegraphics[width=\textwidth]{test_setup_1.jpg}
-        		\caption{Bottom-left view}
-        		\label{fig:Experiment_blv}
-    	\end{subfigure}
-	\hspace*{0.05\columnwidth}
-	\begin{subfigure}[b]{0.35\textwidth}
-        		\includegraphics[width=\textwidth]{test_setup_2.jpg}
-        		\caption{Bottom-right view}
-        		\label{fig:Experiment_brv}
-    	\end{subfigure}
-\caption{Images from two perspectives show the walls and obstacles of the experimental environment.}
-\label{fig:ExpSetupPhoto}
-\end{figure}
-
-The resulting occupancy grid maps and trajectories are depicted in Figure \ref{fig:ExperimentOGM}.
-Most importantly, the robot builds a clear occupancy grid map despite sensor noises and imperfect sensor synchronization.
-The autonomous exploration successfully guides the robot such that it builds a map of the reachable space without colliding with any obstacles.
-
-\begin{figure}
-	\centering{
-    	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_start.jpg}
-        		\caption{$t=0.0$min}
-        		\label{fig:Experiment_ogm_t0}
-    	\end{subfigure}
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_0p5min.jpg}
-        		\caption{$t=0.5$min}
-        		\label{fig:Experiment_ogm_t0p5}
-    	\end{subfigure}    
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1min.jpg}
-        		\caption{$t=1.0$min}
-        		\label{fig:Experiment_ogm_t1}
-    	\end{subfigure}
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1p5min.jpg}
-        		\caption{$t=1.5$min}
-        		\label{fig:Experiment_ogm_t1p5}
-    	\end{subfigure}
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2min.jpg}
-        		\caption{$t=2.0$min}
-        		\label{fig:Experiment_ogm_t2}
-    	\end{subfigure}
-		\vspace*{0.05\textwidth}
-
-	}
-	\centering{
-    	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2p5min.jpg}
-        		\caption{$t=2.5$min}
-        		\label{fig:Experiment_ogm_t2p5}
-    	\end{subfigure}
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3min.jpg}
-        		\caption{$t=3.0$min}
-        		\label{fig:Experiment_ogm_t3}
-    	\end{subfigure}    
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3p5min.jpg}
-        		\caption{$t=3.5$min}
-        		\label{fig:Experiment_ogm_t3p5}
-    	\end{subfigure}
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4min.jpg}
-        		\caption{$t=4.0$min}
-        		\label{fig:Experiment_ogm_t4}
-    	\end{subfigure}
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4p5min.jpg}
-        		\caption{$t=4.5$min}
-        		\label{fig:Experiment_ogm_t4p5}
-    	\end{subfigure}
-	}
-\caption{A robot autonomously explores the experimental environment and produces an occupancy grid map along the way.
-The robot (shaded green circle body with shaded blue circle sensor) is controlled toward the desired pose $1$sec in the future (unshaded grey robot), which is following a constrained least-squares trajectory (cyan curve).
-This trajectory is based on waypoints (black circles) from Dijkstra's algorithm to arrive at the optimal pose (unshaded green circle body with unshaded blue circle sensor) as proposed in this paper.}
-\label{fig:ExperimentOGM}
-\end{figure}
-
-The autonomous exploration is governed by a policy to maximize expected entropy decreases, so entropy and entropy change with time (Figure \ref{fig:ExperimentH}) are important metrics.
-In this example, the a priori occupancy probabilities of all cells are $0.1$, so the entropies could temporarily rise for likely occupied cells.
-The total map entropy begins at roughly $2613$ and decreases by roughly half to $1338$.
-Many grid cells within the map limits are occluded by walls, preventing the blocked cells from occupancy probability changes.
-In some cases, the robot is able to update the cell occupancy probabilities; in other cases, cells fall inside regions occluded from all reachable space, and the robot cannot update these cells.
-Among those grid cells visible from collision-free locations, the vast majority of cells were updated in less than $5$min while following a very slow ($5$cm/sec) trajectory.
-
-%Additionally, even though the walls appear fairly well-defined, they are not pefectly straight boundaries.
-The occupancy grid assumptions do not perfectly align with the true environment; an edge of a wall might occupy only part of a grid cell.
-Therefore, some measurement rays passing through this cell provide evidence that the cell is free, whereas other rays might provide conflicting evidence that the cell is occupied.
-Furthermore, any error in robot attitude estimation can degrade the map; an attitude error of $1^{\circ}$ corresponds to an offset of $11$ measurement rays propagating across the $640$ measurement rays composing the Kinect depth scan.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{experiment_H_per_min.pdf}
-	\caption{The map entropy $H$ averaged over all $90601$ grid cells is reduced during experimentation.
-The quick changes $\Delta H$ per cell correspond to map updates when the robot captures new territory.}
-\label{fig:ExperimentH}
-\end{figure}
-
-\section{Conclusions}
-\label{sec:Conclusion}
-
-This paper serves to improve the accuracy occupancy grid mapping and the policy governing robotic motion during autonomous exploration.
-With other approaches, each probability of an occupancy grid map is approximated in practice due to the perceived complications of the exact solution.
-This paper provides an exact solution to this complicated probability problem using cell occlusion in the forward sensor model.
-This yields a substantially simpler inverse sensor model, which avoids a potentially harmful Markov assumption that commonly appears in the log-odd ratio formulations.
-The main contribution of this paper is that all of the information available in the measurements and the a priori estimate is extracted and integrated to generate a more accurate a posteriori map.
-The numerical and experimental results show that maps using this exact approach are substantially improved.
-
-The exact probabilistic occupancy grid mapping framework is utilized in autonomous exploration.
-The robot pose is chosen to minimize the expected map information gain, which is now directly solved more precisely in an unprecedented way.
-Unlike common techniques based on finding frontiers or predicting measurement scans with particle filters, the proposed approach is based on an expected value of the entropy directly.
-The approach is approximated with large improvements in computation and small losses in accuracy.
-Numerical examples and an experiment show how robots navigate uncertain maps governed by the proposed autonomous exploration algorithm while building accurate probabilistic maps.
-
-\section*{Appendix}\label{append}
-
-\subsection*{Proof of Proposition 1}
+% !TEX root = ../thesis-sample.tex
+\appendix
+\doublespacing
+\chapter{Proof of Proposition 1}
 
 \paragraph{Unnormalized Reduced Map Inverse Sensor Model}
-For the reduced map of the $l$-th ray from \refeqn{InvSenModWithProbDens}, the unnormalized part corresponding to terms without $\eta_{t,l}$ in \refeqn{RayISMAnswer} is %given by
+For the reduced map of the $l$-th ray, the unnormalized probability that corresponds to the terms without $\eta_{t,l}$ in \refeqn{InvSenModWithProbDens} is given by
 \begin{align}
-\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\
-= \sum_{r\in\mathcal{M}_{k}} p(z_{t,l}|r,X_t) P(r|X_{1:t-1},Z_{1:t-1}).\label{eqn:tildePlk}
+\tilde P&(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})= \sum_{r\in\mathcal{M}_{k}} p(z_{t,l}|r,X_t) P(r|X_{1:t-1},Z_{1:t-1}).\label{eqn:tildePlk}
 \end{align}
@@
 \begin{align*}
-\mathcal{N}_{i,k}  = \{ r\in\mathcal{M}_k| \mathbf{r}_{l,i+}\}
-= \{ r\in\mathcal{M}_k| \mathbf{r}_{l,1}=0,\ldots, \mathbf{r}_{l,i-1}=0, \mathbf{r}_{l,i}=1,
+\mathcal{N}_{i,k} & = \{ r\in\mathcal{M}_k| \mathbf{r}_{l,i+}\}= \{ r\in\mathcal{M}_k| \mathbf{r}_{l,1}=0,\ldots, \mathbf{r}_{l,i-1}=0, \mathbf{r}_{l,i}=1,
 \mathbf{r}_{l,k}=1\}.
@@
 \begin{align*}
-\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}) = \sum_{i=1}^k\braces{\sum_{r\in\mathcal{N}_{i,k}} p(z_{t,l}|r,X_t) P(r|X_{1:t-1},Z_{1:t-1})}.
+\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})= \sum_{i=1}^k\braces{\sum_{r\in\mathcal{N}_{i,k}} p(z_{t,l}|r,X_t) P(r|X_{1:t-1},Z_{1:t-1})}.
 \end{align*}
@@
 \begin{align}
-\sum_{r\in\mathcal{N}_{i,k}}  P(r|X_{1:t-1},Z_{1:t-1}) &= \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
- \times &P({\mathbf{r}}_{l,i}|X_{1:t-1},Z_{1:t-1})P({\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1}),\label{eqn:PNik}
+\sum_{r\in\mathcal{N}_{i,k}}  &P(r|X_{1:t-1},Z_{1:t-1}) 
+\nonumber\\&= \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}
+P({\mathbf{r}}_{l,i}|X_{1:t-1},Z_{1:t-1})P({\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1}),\label{eqn:PNik}
 \end{align}
@@
 \begin{align}
-\sum_{r\in\mathcal{N}_{k,k}} P(r|&X_{1:t-1},Z_{1:t-1})\nonumber\\&= \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
-%&\quad \times 
-P({\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1}).\label{eqn:PNkk}
+\sum_{r\in\mathcal{N}_{k,k}} & P(r|X_{1:t-1},Z_{1:t-1})= \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}P({\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1}).\label{eqn:PNkk}
 \end{align}
@@
 \begin{align}
-\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})= \sum_{r\in\bar{\mathcal{M}}_{k}} p(z_{t,l}|r,X_t) P(r|X_{1:t-1},Z_{1:t-1}).\label{eqn:tildePbarlk}
+\tilde P&(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})= \sum_{r\in\bar{\mathcal{M}}_{k}} p(z_{t,l}|r,X_t) P(r|X_{1:t-1},Z_{1:t-1}).\label{eqn:tildePbarlk}
 \end{align}
@@
 \begin{align}
-\tilde P&(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})=\sum_{\substack{i=1\\i\neq k}}^{n_{r,l}} \braces{p(z_{t,l}|\mathbf{r}_{l,i+},X_t) \sum_{r\in\bar{\mathcal{N}}_{i,k}} P(r|X_{1:t-1},Z_{1:t-1})},\label{eqn:tildePbarlk0}
+\tilde P&(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})= \sum_{\substack{i=1\\i\neq k}}^{n_{r,l}} \braces{p(z_{t,l}|\mathbf{r}_{l,i+},X_t) \sum_{r\in\bar{\mathcal{N}}_{i,k}} P(r|X_{1:t-1},Z_{1:t-1})},\label{eqn:tildePbarlk0}
 \end{align}
@@
 \begin{align}
-\sum_{r\in\bar{\mathcal{N}}_{i,k}} P(r|X_{1:t-1},&Z_{1:t-1}) = \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\
-&\quad \times P({\mathbf{r}}_{l,i}|X_{1:t-1},Z_{1:t-1})P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1}),\label{eqn:PbarNik1}
+\sum_{r\in\bar{\mathcal{N}}_{i,k}} & P(r|X_{1:t-1},Z_{1:t-1}) \nonumber\\&= \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\} P({\mathbf{r}}_{l,i}|X_{1:t-1},Z_{1:t-1})P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1}),\label{eqn:PbarNik1}
 \end{align}
@@
 \begin{align}
-\sum_{r\in\bar{\mathcal{N}}_{i,k}} & P(r|X_{1:t-1},Z_{1:t-1}) = \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\} P({\mathbf{r}}_{l,i}|X_{1:t-1},Z_{1:t-1}).\label{eqn:PbarNik2}
+\sum_{r\in\bar{\mathcal{N}}_{i,k}} & P(r|X_{1:t-1},Z_{1:t-1}) = \bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}
+P({\mathbf{r}}_{l,i}|X_{1:t-1},Z_{1:t-1}).\label{eqn:PbarNik2}
 \end{align}
@@
 \begin{align}
-&\tilde P (\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})
+\tilde P &(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})
 =P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\
-&\quad\times \bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]
+&\quad\times \bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\} p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]
 \nonumber
@@
 +
-\bigg[\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg],\label{eqn:tildePbar}
+\bigg[\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\} p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg],\label{eqn:tildePbar}
 \end{align}
@@
 \begin{align*}
-&\eta_{t,l}=\frac1{\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})+
-\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}.
+\eta_{t,l}=[\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})+
+\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})]^{-1}.
 \end{align*}
 Substituting \refeqn{Unnormalized} and \refeqn{tildePbar}, and rearranging, we obtain \refeqn{allEta}.
-
-\bibliography{BibSources}
-\bibliographystyle{IEEEtran}
-
-%% For one-column wide figures use
-%\begin{figure}
-%% Use the relevant command to insert your figure file.
-%% For example, with the graphicx package use
-%  \includegraphics{example.eps}
-%% figure caption is below the figure
-%\caption{Please write your figure caption here}
-%\label{fig:1}       % Give a unique label
-%\end{figure}
-%%
-%% For two-column wide figures use
-%\begin{figure*}
-%% Use the relevant command to insert your figure file.
-%% For example, with the graphicx package use
-%  \includegraphics[width=0.75\textwidth]{example.eps}
-%% figure caption is below the figure
-%\caption{Please write your figure caption here}
-%\label{fig:2}       % Give a unique label
-%\end{figure*}
-%%
-%% For tables use
-%\begin{table}
-%% table caption is above the table
-%\caption{Please write your table caption here}
-%\label{tab:1}       % Give a unique label
-%% For LaTeX tables use
-%\begin{tabular}{lll}
-%\hline\noalign{\smallskip}
-%first & second & third  \\
-%\noalign{\smallskip}\hline\noalign{\smallskip}
-%number & number & number \\
-%number & number & number \\
-%\noalign{\smallskip}\hline
-%\end{tabular}
-%\end{table}
-
-%\begin{acknowledgements}
-%If you'd like to thank anyone, place your comments here
-%and remove the percent signs.
-%\end{acknowledgements}
-
-% BibTeX users please use one of
-%\bibliographystyle{spbasic}      % basic style, author-year citations
-%\bibliographystyle{spmpsci}      % mathematics and physical sciences
-%\bibliographystyle{spphys}       % APS-like style for physics
-%\bibliography{}   % name your BibTeX data base
-
-\end{document}
-% end of file template.tex

### 2018-06-24 Dissertation/Thesis/tex/intro.tex
added Appendix A: formal ISM proof

@@
 For example, \cite{ChoLynHutKanBurKavThr05} proposed a simplified continuous function composed of lines patched together near a measurement reading.
-Furthermore, variations of approximate smoother functions in~\cite{and09,PirRutBisSch11,KhoElb12} contain Gaussian-like terms to apply occupancy grid mapping to 3D environments with more modern sensors.
+Furthermore, variations of approximate smoother functions in~\cite{Andert09,PirRutBisSch11,KhoElb12} contain Gaussian-like terms to apply occupancy grid mapping to 3D environments with more modern sensors.
 Numerous research studies have used the proposed approximate functions with questionable accuracy.

### 2018-06-24 Dissertation/Thesis/tex/ae2D.tex
new prop for entropy, started another appendix which is empty now

@@
 
+The goal of autonomous exploration is to select robotic actions designed to minimize map uncertainty, thereby maximizing map information gain.
+We formulate autonomous exploration as an optimization problem to determine a policy that accounts for map uncertainty and travel cost.
+
 \section{Entropy-Based Exploration}
 
+In this section, we define Shannon's entropy as a measure of map uncertainty, and present a novel approach to predict future entropy from an arbitrary ray.
+Then we discuss computational modifications for real-time implementation.
+
 \subsection{Shannon's Entropy}
-% Description, figure
+
+Here we present how the probabilistic properties of an occupancy grid map provide key information about the uncertainty of the space for motion planning.
+Shannon's entropy commonly serves as an uncertainty measure~\cite{StaGriBur05}.
+Provided grid cell probabilities of map $m$, Shannon's entropy is defined as
+\begin{align}
+\label{eqn:ShannonsEntropyCell}
+H(P(\mathbf{m}_i))&=-P(\mathbf{m}_i)\log{P(\mathbf{m}_i})-P(\bar{\mathbf{m}}_i)\log{P(\bar{\mathbf{m}}_i}),
+\\
+\label{eqn:ShannonsEntropyMap}
+H(P(m))&=\sum_{i=1}^{n_m}H(P(\mathbf{m}_i)),
+\end{align}
+for an individual cell and the entire map, respectively.
+Thus, entropy is maximized when $P(\mathbf{m}_i)=0.5$, which corresponds to the largest uncertainty; similarly, entropy is minimized as $P(\mathbf{m}_i)$ approaches $0$ or $1$, which corresponds to the smallest uncertainty.
+%; thus Shannon's entropy is a measure of uncertainty.
 
 \subsection{Expected Information Gain}
-% Derivation and link to Appendix
+
+Suppose a future pose candidate and its associated future measurement scan is $X_c=\braces{x_c,R_c}$ and $Z_c$, respectively, where $c\in\mathcal C$ such that $\mathcal C=\braces{1,2,...,n_c}$ accounts for all $n_c$ candidates under consideration for the next robot pose.
+The current pose of the robot is \emph{not} $X_c$ in general, so any change to the probabilistic map from $X_c$ must be predicted.
+Much like the probabilistic mapping, this is achieved ray-by-ray~\cite{KauAiLee16,KauTakAiLee17}.
+Considering that all grid cell probabilities are conditioned on the history of poses $X_{1:t}$ and measurement scans $Z_{1:t}$, these terms are removed from the remaining equations for simplicity.
+
+\begin{prop}
+\label{prop:ExpectedH}
+For candidate ray $z_c$, the expected entropy is
+\begin{align}
+\label{eqn:DiscExpEntropyRay}
+&\text{E}[H(P(m|x_c,z_{c}))]=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(m|x_c,z_{c,k}))P(z_{c,k}|x_c)\bigg\},
+\end{align}
+where $z_{c,k}$ refers to the distance from $x_c$ to the $k$-th grid cell along the measurement ray.
+The first term of the summation of \refeqn{DiscExpEntropyRay}, namely $H(P(m|x_c,z_{c,k}))$, is obtained with entropy definitions \refeqn{ShannonsEntropyCell}, \refeqn{ShannonsEntropyMap} and the inverse sensor model \refeqn{RayISMAnswer}--\refeqn{Unnormalized}.
+The second term is derived from \refeqn{allEta} with
+\begin{align}
+\label{eqn:ProbMeas}
+P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}},
+\end{align}
+where $\eta_{c,k}$ refers to the normalizer based on the measurement $z_{c,k}$.
+The expected negative entropy change for candidate pose $X_c$ is equivalently the expected information gain,
+\begin{align}
+\label{eqn:expectedInfoGainRay}
+\mathcal I(X_c)&=H(P(m))-\text{E}\left[H(P(m|X_c,Z_c))\right].
+\end{align}
+\end{prop}
+\begin{proof}% TODO: add ref?
+See Appendix B.
+\end{proof}
+
+The proposed approach discretizes the measurement according to assumptions of occupancy grid mapping, and exploits the probabilistic properties uncovered by the proposed inverse sensor model.
+These are used to directly calculate the expected value of map entropy for a single measurement ray.
 
@@
 % SIMPAR figure on optimal pose
+
+%Then, \refeqn{DiscExpEntropyRay} is applied repeatedly over several sample measurement rays that compose scan $Z_c$ as shown in~\cite{KauAiLee16}.
 

### 2018-06-24 Dissertation/Thesis/tex/pogm.tex
new prop for entropy, started another appendix which is empty now

@@
 Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell.
-Based on this, we present the mathematical expressions for the exact inverse sensor model, as summarized by \refeqn{RayISMAnswer} in Proposition 1.
+Based on this, we present the mathematical expressions for the exact inverse sensor model, as summarized by \refeqn{RayISMAnswer} in Proposition \ref{prop:ISM}.
 Later in Section \ref{sec:RayScanComb}, these are also rearranged in computational algorithms.
@@
 \begin{prop}
+\label{prop:ISM}
 For the $l$-th measurement ray, the a posteriori probability of the occupancy of the $k$-th cell, namely the ray inverse sensor model, is given by
@@
 \end{prop}
-\begin{proof}% TODO: add ref
-See Appendix.
+\begin{proof}% TODO: add ref?
+See Appendix A.
 \end{proof}
@@
 At the $t$-th time step, consider the measurement scan $Z_t=\{z_{t,1},z_{t,2},...,z_{t,n_z}\}$.
-One may apply the results of Proposition 1 repeatedly for each ray to obtain a two-dimensional inverse sensor model because
+One may apply the results of Proposition \ref{prop:ISM} repeatedly for each ray to obtain a two-dimensional inverse sensor model because
 \begin{align}

### 2018-06-24 Dissertation/Thesis/tex/ae2D.tex
added approx justification and pseudocode to approx H

@@
 \subsection{Computational Limitations and Approximations}
+
+The computational order for each measurement ray is $\mathcal O(n_{r}^2)$ since the summations of \refeqn{ProbMeas} are embedded in \refeqn{DiscExpEntropyRay}.
+However, several of those intersections provide negligible information since the probability of the measurement ray capturing certain cell depths is close to zero.
+
+The approximation of expected ray entropy provides a method to reduce the computation of \refeqn{expectedInfoGainRay} substantially.
+This goal is achieved by systematically selecting a smaller set of grid cells to consider over the summations of \refeqn{DiscExpEntropyRay} and \refeqn{ProbMeas}.
+The smaller set is determined by the probability that each cell is captured by the measurement ray, known as the detection probability. This can be found recursively as
+\begin{align}
+\label{eqn:ProbOfFirstCell}
+P(\mathbf{r}_{k+})%\nonumber\\&
+=\bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{j})\bigg\}P(\mathbf{r}_{k}),
+\end{align}
+which is the probability that $\mathbf{r}_{k}$ is the closest occupied grid cell based on past poses and measurement scans, independent of cells beyond the $k$-th cell from $x_c$.
+Let $\hat n>0$ be a fixed number of grid cells for all rays such that $\hat n\leq n_{r}+1$ for all $l$.
+Let $\hat{r}$ correspond to the grid cells that yield the $\hat{n}$ maximum values of \refeqn{ProbOfFirstCell} (the $\hat n$ most likely ray detections), indexed by increasing distance from candidate location $x_c$.
+By replacing the reduced map $r$ with $\hat{r}$ and changing the summation limits to $\braces{1,2,...,\hat n}$ in \refeqn{DiscExpEntropyRay} and \refeqn{ProbMeas}, the order of computation is reduced to $\mathcal O({\hat{n}}^2)$.
+Even though the value of $n_{r}$ is different among various rays in general, $\hat n$ is fixed, so the computational order is fixed as well.
+In short, this method reduces the required computation substantially by systematically neglecting those grid cells with little effect.
+It can be noted that if $\hat n=n_{r}+1$, the ray objective function is computed without approximation.
+
+\paragraph{Algorithm} We present an algorithm pseudocode providing the necessary steps to obtain the objective function for a single measurement ray (Algorithm \ref{alg:RayExpectedEntropyGain}).
+Much like the algorithm pseudocode for the ray-by-ray inverse sensor model (Algorithm \ref{alg:RayByRayISM}), the variable $a_\text{temp}$ serves as an intermediate variable designed to avoid repeated calculations.
+Since this algorithm operates as a function, fixed indices and condition variables are removed for simplification.
+
+\vspace*{0.05\columnwidth}
+\begin{algorithm}[H]
+	Function: $\mathcal I_\text{ray}=\text{RayExpInfoGain}(x,P(r),z_{1:n_{r}})$\;
+	Initialize $P(\bar{\mathbf{r}}_{0})=P(\hat{\bar{\mathbf{r}}}_{0})=P(\bar{\mathbf{r}}_{n_{r}+1})=1$\;
+	\For{$k = 1,2,\ldots,n_{r}+1$}{
+		$P(\mathbf{r}_{k+}) = P(\bar{\mathbf{r}}_{0:k-1})P(\mathbf{r}_{k})$\;
+		$P(\bar{\mathbf{r}}_{0:k})=P(\bar{\mathbf{r}}_{0:k-1})(1-P(\mathbf{r}_{k}))$\;
+	}
+	Find $\hat r\subset r$ of the $\hat n$ greatest values of 
+	$\braces{P(\mathbf{r}_{1+}),P(\mathbf{r}_{2+}),\ldots,P(\mathbf{r}_{(n_{r}+1)+})}$\;
+	\For{$k = 1,2,\ldots,\hat{n}$}{
+		$P(\hat{\mathbf{r}}_{k+}) = P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\mathbf{r}}_{k})$\;
+		$P(\hat{\bar{\mathbf{r}}}_{0:k})=P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\bar{\mathbf{r}}}_{k})$\;
+	}
+	\For{$k_\text{m}=1,2,\ldots,\hat n$}{
+		Initialize $\eta^{-1}_{k_\text{m}}=0$\;
+		\For{$k_\text{c}=1,2,\ldots,\hat n$}{
+			$a_\text{temp}=P(\hat{\mathbf{r}}_{k_\text{c}+})p(z_{k_\text{m}}|\hat{\mathbf{r}}_{k_\text{c}+},x)$\;
+			$\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=P(\hat{\mathbf{r}}_{k_\text{c}})\eta^{-1}_{k_\text{m}}+a_\text{temp}$\;
+			$\eta^{-1}_{k_\text{m}}=\eta^{-1}_{k_\text{m}}+a_\text{temp}$\;
+		}
+		$P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=\eta_{k_\text{m}}\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})$ for all $k_\text{c}=1,2,\ldots,\hat n$\;
+	}
+	$P(z_{k_\text{m}}|x)=\frac{\eta^{-1}_{k_\text{m}}}{\sum_{i=1}^{\hat n}\eta^{-1}_{i}}$ for $k_\text{m}=1,2,\ldots,\hat n$\;
+	Initialize $\mathcal I_\text{ray}=0$\;
+	\For{$k_\text{c}=1,2,\ldots,\hat n$}{
+		$\mathcal I_\text{ray}=\mathcal I_\text{ray}+H(P(\hat{\mathbf{r}}_{k_\text{c}}))$\;
+		\For{$k_\text{m}=1,2,\ldots,\hat n$}{
+			$\mathcal I_\text{ray} = \mathcal I_\text{ray}-H(P(\hat{\mathbf{r}}_{k_\text{c}}|x_c,z_{k_\text{m}}))P(z_{k_\text{m}}|x_{c})$\;
+		}
+	}
+	Return: $\mathcal I_\text{ray}$\\
+\caption{Expected Information Gain from a Measurement Ray}
+\label{alg:RayExpectedEntropyGain}
+\end{algorithm}
+
 % SIMPAR figure on entropy/computation time relationship

### 2018-06-24 Dissertation/Thesis/tex/ae2D.tex
edited up through 3.1

@@
 where $\eta_{c,k}$ refers to the normalizer based on the measurement $z_{c,k}$.
-The expected negative entropy change for candidate pose $X_c$ is equivalently the expected information gain,
+The expected negative entropy change for candidate pose $X_c$ is equivalently the expected information gain from ray $z_c$,
 \begin{align}
 \label{eqn:expectedInfoGainRay}
-\mathcal I(X_c)&=H(P(m))-\text{E}\left[H(P(m|X_c,Z_c))\right].
+\mathcal I(X_c,z_c)&=H(P(m))-\text{E}\left[H(P(m|X_c,z_c))\right].
 \end{align}
@@
 
-% SIMPAR figure on entropy/computation time relationship
+\paragraph{Numerical Justification for the Approximation}
+
+The purpose of this numerical example is to provide evidence that the approximations are reasonable and increase the algorithm speed substantially.
+Since a measurement ray produces a depth measurement in a single direction, we only consider a 1D map where the grid cells have spacing $\alpha=0.2$ m, and the properties of the range sensor are based on the Microsoft Kinect~\cite{PirRutBisSch11,KhoElb12} with maximum reading depth $z_\text{max}=4$ m ($20$ grid cells inside the sensor FOV).
+The goal is to compare the expected entropy $\text{E}[H(P(m|x_c,z_{c}))]$ from \refeqn{DiscExpEntropyRay} and with an approximation $\text{E}[H_\text{approx}(P(m|x_c,z_{c}))]$, which only considers $\hat n$ grid cells with highest detection probability \refeqn{ProbOfFirstCell}.
+
+We consider $100$ probabilistic maps to obtain Monte Carlo results to evaluate the approximate entropy.
+In every Monte Carlo trial, each grid cell has an $80\%$ chance of being free and a $20\%$ chance of receiving an a priori probability uniformly distributed between $0$ and $1$.
+Several metrics serve to evaluate $\text{E}[H(P(m|x_c,z_{c}))]$ with $\text{E}[H_\text{approx}(P(m|x_c,z_{c}))]$.
+The median expected entropy change is $\text{E}[H(P(m|x_c,z_{c}))]-H(P(m|X_{1:t},Z_{1:t}))=-0.83792$.
+The error for the $100$ Monte Carlo cases is defined simply as
+\begin{align}
+e_{H}&=\frac1{100}\sum_{i=1}^{100}\text{abs}\bigg(\text{E}[H(P(m|x_c,z_{c}))]-\text{E}[H_\text{approx}(P(m|x_c,z_{c}))]\bigg).
+\end{align}
+The Monte Carlo trials are repeated for $\hat n=\braces{1,2,\ldots,10}$ and the results are plotted in Figure \ref{fig:ApproxJust}.
+This example shows a typical case when the summation limits generated from \refeqn{ProbOfFirstCell} have only small effects on \refeqn{DiscExpEntropyRay}, while providing very large improvements in reducing computation.
+
+\begin{figure}
+	\centering
+    	\begin{subfigure}[b]{0.45\textwidth}
+        		\includegraphics[width=\textwidth]{JustifyApprox_eH.pdf}
+        		\caption{Entropy Error}
+        		\label{fig:H_err}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.45\textwidth}
+        		\includegraphics[width=\textwidth]{JustifyApprox_t.pdf}
+        		\caption{Computation Ratio}
+        		\label{fig:H_comp_ratio}
+    	\end{subfigure}
+\caption{The entropy error is decreased at the cost of increasing computation time in this Monte Carlo 1D measurement ray expected entropy case.}
+\label{fig:ApproxJust}
+\end{figure}
 

### 2018-06-24 Dissertation/Thesis/tex/ae2D.tex
added text from JINT17, needs revision

@@
 
+In this section, we show how expected entropy changes from single measurement rays can be integrated into predicting the expected information gain of measurement scans.
+This metric, and a cost associated with travel time, are integrated to determine an optimal future pose and collision-free path in a 2D environment.
+
 \subsection{Optimal Attitude from Sample Rays}
-% SIMPAR figure on optimal pose
+
+The objective is to choose a future pose that minimizes the map uncertainty, such that the robot can autonomously move toward this optimal goal.
+The robot pose is selected among various positions and attitudes.
+Since searching over all locations and attitudes would require infinite computational resources, these search spaces are discretized into a limited number of robot attitudes and positions.
+The procedure involves three steps.
+First, the attitude that maximizes the information gain objective function is selected at each candidate future pose location.
+Second, the location with the maximum objective function of all candidates is selected.
+Thus, the optimal attitude and the optimal location compose the optimal future pose.
+Finally, the robot determines and follows a collision-free path to the optimal future pose.
+This process is repeated, resulting in autonomous exploration.
+
+\paragraph{Attitude Optimization}
+Consider a collision-free pose candidate at an arbitrary location $x_c$.
+At this pose, consider $n_d$ evenly-spaced measurement rays oriented radially outward from the robot in a circular pattern (see red lines in Figure \ref{fig:OptProcess}).
+These rays are denoted $\braces{z_{c,1},z_{c,2},\ldots,z_{c,n_d}}$.
+Attitudes correspond to the rays such that the robot sensor is aligned with the ray direction, and these are denoted $\braces{R_{c,1},R_{c,2},\ldots,R_{c,n_d}}$, where the scan with attitude $R_{c,d}$ might cover several ray directions depending on the sensor field of view (FOV).
+We choose optimal attitude $R_c^*$ as the summation of the expected entropy changes covered by the scan,
+\begin{align}
+\label{eqn:FindRc}
+&R_c^*=\argmax_{R_{c,d}}\sum_{z_{c,i}\in R_{c,d}\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\qquad\quad\quad\quad\quad
+-\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,i}))]\bigg).
+%\mathcal I_\text{ray}(x_c,z_{c,i})\right).
+\end{align}
+
+This method provides the attitude that maximizes the information gain at an arbitrary location in 2D space.
+The set of candidate positions that warrant consideration is determined with one of two methods, namely \emph{expanding ring} and \emph{complete Cartesian}, described next.
+
+\paragraph{Expanding Ring}
+The expanding ring technique is advantageous for searching local solutions quickly, only considering distant future poses when necessary.
+The key idea is that future candidate locations lie on a circular ``ring'' centered around the robot, evenly spaced around the ring (see red circles in Figure \ref{fig:OptProcess}), and the ring is expanded if certain criteria are not met.
+More explicitly, consider $n_c$ candidate locations denoted by $x_c\in\braces{x_1,x_2,\ldots,x_{n_c}}$ located with the distance $\delta$ away from the robot.
+Any location that violates inequality constraint \refeqn{CollisionInequalityConstraint} is excluded to avoid collisions.
+The set of optimal attitudes at each candidate location is $\braces{R^*_{1},R^*_{2},\ldots,R^*_{n_c}}$, obtained from \refeqn{FindRc}.
+The objective function \refeqn{ObjFun} is computed by a summation about the $n_d$ rays as %\refeqn{Objective},
+\begin{align}
+\label{eqn:ObjFunApprox}
+\mathcal I(x_c,R_c^*)&\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\qquad-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*,z_{c,i}))\right]\bigg),
+\\
+x_c^*&=\argmax_{x_c}\mathcal I(x_c,R_c^*).
+\end{align}
+At the optimal pose $X^*_c=(x^*_c,R^*_c)$, the resulting information gain must satisfy $\mathcal I(X_c^*)\geq\mathcal I_\text{min}$, where $\mathcal I_\text{min}$ is some minimum threshold for expected information gain; robot motion is only justified when the expected information gain is significantly large.
+If the minimum threshold is not met, the ring of candidate pose locations is increased such that $n_c$ and $\delta$ are multiplied by a scale function $\lambda>1$.
+This process is repeated until the expected information gain of the optimal pose is at least $\mathcal I_\text{min}$, or all candidates lie in collision zones or outside map limits.
+%A pseudocode of the process is found in Table \ref{tab:Alg_AutomousExploration}.
+
+The primary advantage of the expanding ring approach to search the 2D space is that the number of pose candidates need not be proportional to the map area, and that local maxima tend to fall within short distances of the current robot pose.
+The main drawback is that poses outside a small local region of the robot are frequently neglected, and the robot frequently executes repeated short motions, often without large information gains.
+
+\paragraph{Complete Cartesian}
+The complete Cartesian method to search the 2D space provides candidates located fixed distance $d$ apart in each Cartesian direction.
+This approach is advantageous for capturing expected information gains in regions outside of the immediate vicinity of the robot, but must be applied carefully to avoid issues with computational bottlenecks due to the potentially-large search space.
+
+Similar to the expanding ring approach, $n_c$ candidate pose locations are considered, where those violating \refeqn{CollisionInequalityConstraint} are neglected from further consideration.
+However, unlike the ring expanding approach, the candidates have different distances from the robot in general, so the objective function is modified to enforce a cost on the squared distance from the current pose location $x_t$ to the candidate location $x_c$, i.e.,
+\begin{align}
+\label{eqn:ObjFunApproxCompleteCartesian}
+\mathcal I&(x_c,R_c^*,x_t)\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*(x_c),z_{c,i}))\right]\bigg)-k_\text{dist}(x_t-x_c)^\T(x_t-x_c),
+\end{align}
+where weighting function $k_\text{dist}$ represents the sensitivity to avoiding large motions across the map.
+Furthermore, the same candidate locations are considered throughout the repeated processes of exploration, and the map cell occupancies are assumed static.
+Therefore, if candidate location $x_c$ fails to satisfy $\mathcal I(x_c,R_c^*,x_t)\geq\mathcal I_\text{min}$, then $x_c$ need \emph{never} be considered again.
+Avoiding these unnecessary calculations yields a greatly lowered computational burden, particularly late in exploration where several regions are collision-free but disadvantageous to revisit.
+
+\begin{figure}
+\vspace*{0.1\columnwidth}
+\centerline{
+	\includegraphics[height=0.35\columnwidth]{ExampleOptimalPose.png}
+}
+\begin{picture}(0,0)(0,0)
+\setlength{\unitlength}{0.1\columnwidth}\scriptsize
+\put(4.4,2.1){\color{blue}$X_t$}
+\put(5.9,1.9){\color{red}$x_1$}
+\put(2.9,2.0){\color{red}$x_5$}
+\put(4.4,3.6){\color{red}$x_3$}
+\put(4.4,0.6){\color{red}$x_7$}
+\put(5.4,3.0){\color{red}$x_2$}
+\put(3.3,3.1){\color{red}$x_4$}
+\put(3.3,1.0){\color{red}$x_6$}
+\put(5.4,1.0){\color{red}$x_8$}
+\put(6.8,3.1){\color{red}$z_{2,1}$}
+\put(4.5,3.1){\color{red}$z_{2,5}$}
+\put(5.6,4.1){\color{red}$z_{2,3}$}
+\put(5.6,2.2){\color{red}$z_{2,7}$}
+\put(6.5,3.7){\color{red}$z_{2,2}$}
+\put(4.9,3.8){\color{red}$z_{2,4}$}
+\put(5.0,2.5){\color{red}$z_{2,6}$}
+\put(6.5,2.5){\color{red}$z_{2,8}$}
+\put(6.1,2.9){$X_c^*$}
+\end{picture}
+\caption{This figure illustrates the expanding ring method.
+Initially, a robot (blue circle) views down and left (blue sector).
+Then, the robot considers $n_c=8$ poses (red circles) as possible candidates for where to move next.
+For each candidate, $n_d=8$ directions are considered (red lines, only displayed on candidate location $c=2$).
+Then, the expected entropy of each ray is calculated.
+The scan (red sector) covering those rays with the largest expected entropy decreases is chosen for each candidate, and the best candidate $X_c^*$ (black circle: location, black sector: scan) is chosen to maximize information gain.
+Finally, Dijkstra's algorithm provides the collision-free motion between $X_t$ and $X_c^*$, and the process is repeated.
+}
+\label{fig:OptProcess}
+\end{figure}
 

### 2018-06-24 Dissertation/Thesis/tex/ae2D.tex
completed through numerical examples (3.3), added intro to that

@@
 
-\subsection{Optimal Attitude from Sample Rays}
+\subsection{Optimal Pose from Sample Rays}
+\label{sec:OptimalPose2DMap}
 
@@
 \label{eqn:FindRc}
-&R_c^*=\argmax_{R_{c,d}}\sum_{z_{c,i}\in R_{c,d}\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\qquad\quad\quad\quad\quad
+&R_c^*=\argmax_{R_{c,d}}\sum_{z_{c,i}\in R_{c,d}\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))
 -\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,i}))]\bigg).
@@
 More explicitly, consider $n_c$ candidate locations denoted by $x_c\in\braces{x_1,x_2,\ldots,x_{n_c}}$ located with the distance $\delta$ away from the robot.
-Any location that violates inequality constraint \refeqn{CollisionInequalityConstraint} is excluded to avoid collisions.
+All candidates must satisfy the inequality constraint,
+\begin{align}
+\label{eqn:CollisionInequalityConstraint}
+P_\text{collision}(X)=1-\prod_{i\in\mathcal C_X}(P(\bar{\mathbf{m}}_i)\leq\beta,
+\end{align}
+where $\mathcal C_X$ is the set of grid cells falling inside a volume of preselected size around $X$ that may cause collision and $\beta>0$ is a small acceptable probability of collision.
+Any location that violates an inequality constraint \refeqn{CollisionInequalityConstraint} is excluded to avoid collisions.
 The set of optimal attitudes at each candidate location is $\braces{R^*_{1},R^*_{2},\ldots,R^*_{n_c}}$, obtained from \refeqn{FindRc}.
-The objective function \refeqn{ObjFun} is computed by a summation about the $n_d$ rays as %\refeqn{Objective},
+The information gain objective function,
+\begin{align}
+\label{eqn:ObjFun}
+\mathcal I(X_c)=H(P(m))-\text{E}\left[H(P(m|X_c,Z_c))\right],
+\end{align}
+is computed by a summation about the $n_d$ rays as %\refeqn{Objective},
 \begin{align}
 \label{eqn:ObjFunApprox}
-\mathcal I(x_c,R_c^*)&\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\qquad-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*,z_{c,i}))\right]\bigg),
+\mathcal I(x_c,R_c^*)&\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*,z_{c,i}))\right]\bigg),
 \\
@@
 \end{align}
-At the optimal pose $X^*_c=(x^*_c,R^*_c)$, the resulting information gain must satisfy $\mathcal I(X_c^*)\geq\mathcal I_\text{min}$, where $\mathcal I_\text{min}$ is some minimum threshold for expected information gain; robot motion is only justified when the expected information gain is significantly large.
+At the optimal pose $X^*_c=(x^*_c,R^*_c)$, the resulting information gain must satisfy $\mathcal I(X_c^*)\geq\mathcal I_\text{min}$, where $\mathcal I_\text{min}$ is a minimum threshold for expected information gain; robot motion is only justified when the expected information gain is significantly large.
 If the minimum threshold is not met, the ring of candidate pose locations is increased such that $n_c$ and $\delta$ are multiplied by a scale function $\lambda>1$.
@@
 The primary advantage of the expanding ring approach to search the 2D space is that the number of pose candidates need not be proportional to the map area, and that local maxima tend to fall within short distances of the current robot pose.
+Additionally, the distance to these poses need not be explicitly considered in the pose determination optimization because the distance costs are identical to each candidate, assuming objects do not occlude the trajectories.
 The main drawback is that poses outside a small local region of the robot are frequently neglected, and the robot frequently executes repeated short motions, often without large information gains.
@@
 Then, the expected entropy of each ray is calculated.
-The scan (red sector) covering those rays with the largest expected entropy decreases is chosen for each candidate, and the best candidate $X_c^*$ (black circle: location, black sector: scan) is chosen to maximize information gain.
+The scan (red sector) covering those rays with the largest expected entropy decrease is chosen for each candidate, and the best candidate $X_c^*$ (black circle: location, black sector: scan) is chosen to maximize information gain.
 Finally, Dijkstra's algorithm provides the collision-free motion between $X_t$ and $X_c^*$, and the process is repeated.
@@
 
-%Then, \refeqn{DiscExpEntropyRay} is applied repeatedly over several sample measurement rays that compose scan $Z_c$ as shown in~\cite{KauAiLee16}.
-
-\subsection{Travel Cost with Obstacles}
+% TODO: move these to 3D/multi-vehicle sections
+%\subsection{Travel Cost with Obstacles}
 % Dijkstra's cost map
-
-\subsection{Optimal Pose Selection}
+%\subsection{Optimal Pose Selection}
 % Single-Vehicle argmax
@@
 
+There are two important steps to moving a robot on a collision-free trajectory to its optimal pose.
+First, Dijkstra's algorithm provides the cost map and optimal collision-free waypoints.
+Then, a constrained polynomial least squares trajectory serves as the desired path, which is followed by a controller.
+
+\paragraph{Dijkstra's Search}
+
+Dijkstra's search is naturally applied to this problem because the occupancy grid map serves nicely as a graph, and Dijkstra's algorithm provides an optimal collision-free trajectory along that graph.
+There are two steps to Dijkstra's algorithm.
+first, generate a cost map from the robot location to each grid cell on the 2D occupancy grid map.
+Only cells that satisfy \refeqn{CollisionInequalityConstraint} are considered to avoid possible collisions.
+The cost to travel to neighboring cells that share an edge are the cell edge length $\alpha$, and the cost to travel to cells that share a corner is $\sqrt{2}\alpha$.
+This provides the collision-free distances to all reachable locations on the map.
+Second, the waypoints from a candidate pose to the current robot pose are easily obtained along the cost map using steepest descent.
+
+\paragraph{Constrained Polynomial Least Squares Trajectory} The trajectory to follow the path outlined by Dijkstra's algorithm is determined with a linear constrained least squares optimization, assuming the robot follows a polynomial trajectory with fixed speed.
+The starting and ending positions and attitudes may be constrained, and polynomials patched together for long trajectories share a common position and velocity with respect to time.
+Then, a controller on the robot tracks this trajectory until the robot falls within acceptable thresholds of the final optimal pose.
+Any control scheme for the motion of the robot can be integrated with the proposed exploration algorithm.
+Once the robot completes this motion, the entire process is repeated.
+
 \section{Numerical Examples}
+
+Here we present two numerical examples.
+The first uses the expanding-ring technique in a small environment and the second considers a larger benchmark environment, so the complete Cartesian method is applied, as described in Section \ref{sec:OptimalPose2DMap}.
 

### 2018-06-25 Dissertation/Thesis/tex/ae2D.tex
finished chap 3, could use an edit

@@
 
+Here, a robot autonomously maps and explores a 2D environment composed of two rooms and one hallway.
+The robot models its surroundings with an occupancy grid of $15000$ cells where grid cell edges have length $\alpha=0.2$m, composing a map with dimensions $30\text{m}\times20\text{m}$.
+The initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) for grid cells covered by the circular robot of radius $0.1$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
+At each time step, the robot receives a measurement scan, where the probabilistic properties of the sensor are taken from~\cite{PirRutBisSch11,KhoElb12}.
+Then, the $n_c=8$ evenly-spaced candidate locations about a circle of radius $\delta=0.5$m around the current pose location are considered, where $n_d=32$ measurement rays are evenly-spaced about the candidate location.
+When no current candidates yield expected information gains above $\mathcal I_\text{min}=2$, $\lambda=1.25$ is multiplied to $n_c$ and $\delta$.
+The motion of the robot is restricted to movement along cells satisfying \refeqn{CollisionInequalityConstraint} with $\beta=0.01$.
+Dijkstra's algorithm provides collision-free motion planning.
+The results are illustrated in Figure \ref{fig:AutonomousExploration}.
+
+\begin{figure}
+\centering{
+    	\begin{subfigure}[b]{0.3\textwidth}
+        		\includegraphics[width=\textwidth]{CDC16_t50.png}
+        		\caption{Map at $t=50$ sec}
+%        		\label{fig:}
+    	\end{subfigure}
+	\hspace*{0.01\columnwidth}
+	\begin{subfigure}[b]{0.3\textwidth}
+        		\includegraphics[width=\textwidth]{CDC16_t100.png}
+        		\caption{Map at $t=100$ sec}
+%        		\label{fig:}
+    	\end{subfigure}
+	\hspace*{0.01\columnwidth}
+	\begin{subfigure}[b]{0.3\textwidth}
+        		\includegraphics[width=\textwidth]{CDC16_t150.png}
+        		\caption{Map at $t=150$ sec}
+%        		\label{fig:}
+    	\end{subfigure}
+}
+\centering{
+    	\begin{subfigure}[b]{0.3\textwidth}
+        		\includegraphics[width=\textwidth]{CDC16_t200.png}
+        		\caption{Map at $t=100$ sec}
+%        		\label{fig:}
+    	\end{subfigure}
+	\hspace*{0.01\columnwidth}
+	\begin{subfigure}[b]{0.3\textwidth}
+        		\includegraphics[width=\textwidth]{CDC16_t250.png}
+        		\caption{Map at $t=250$ sec}
+%        		\label{fig:}
+    	\end{subfigure}
+	\hspace*{0.01\columnwidth}
+	\begin{subfigure}[b]{0.3\textwidth}
+        		\includegraphics[width=\textwidth]{CDC16_t300.png}
+        		\caption{Map at $t=300$ sec}
+%        		\label{fig:}
+    	\end{subfigure}
+}
+\centering{
+    	\begin{subfigure}[b]{0.3\textwidth}
+        		\includegraphics[width=\textwidth]{CDC16_t350.png}
+        		\caption{Map at $t=350$ sec}
+%        		\label{fig:}
+    	\end{subfigure}
+	\hspace*{0.01\columnwidth}
+	\begin{subfigure}[b]{0.3\textwidth}
+        		\includegraphics[width=\textwidth]{CDC16_t400.png}
+        		\caption{Map at $t=400$ sec}
+%        		\label{fig:}
+    	\end{subfigure}
+	\hspace*{0.01\columnwidth}
+	\begin{subfigure}[b]{0.3\textwidth}
+        		\includegraphics[width=\textwidth]{CDC16_t450.png}
+        		\caption{Map at $t=450$ sec}
+%        		\label{fig:}
+    	\end{subfigure}
+}
+\caption{A robot (green circle, green marker of previous path) measures a room with a Kinect depth sensor to maximize map information gain.}
+\label{fig:AutonomousExploration}
+\end{figure}
+
+Knowing only that the robot is inside free space at the beginning, the robot carefully navigates the environment while avoiding collisions.
+The robot motion is governed by a policy that maximizes the map information gain within its set of pose candidates, where the $\hat n=6$ is chosen to approximate the expected entropy of each ray.
+When running the exploration algorithm in Robot Operating System (ROS), the mean computation time is $0.0194$ seconds to determine the optimal future pose and complete Dijkstra's algorithm on the occupancy grid.
+The computation times for this map reached roughly $1$ second at maximum, corresponding to rare cases when the robot is locally surrounded by a highly-certain environment; this task requires evaluating many more pose candidates and motion planning over a larger terrain.
+
+Throughout the numerical example, the robot chooses numerous actions, based directly on their expected information gains, not frontiers or predicted measurement scans.
+If the obstacles were known a priori, the motion planning problem would yield a simpler path; however, the autonomous exploration is based on only the information of the map that the robot generates, so the motion planning must be reevaluated repeatedly.
+Even with these limitations, the robot explores the vast majority of reachable space in the $450$ second period.
+
 \subsection{Exploring a Complicated Benchmark Environment}
-% JINT17 Fig. 7 & 8
 
-\section{Experimental Result}
+Next, the proposed autonomous exploration approach is applied to the floor plan of Intel Research Lab, illustrated in Figure~\ref{fig:intel}.
+The robot explores its surroundings with an occupancy grid with $90,000$ cells where grid cell edges are $\alpha=0.1$m, composing a map with dimensions $30\text{m}\times30\text{m}$.
+Similar to the prior example, the initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) is chosen for grid cells covered by the circular robot of radius of $0.3$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
+For added safety, cells falling within $0.6$m of the robot are considered inside the possible collision zone $\mathcal{C}_X$, from \refeqn{CollisionInequalityConstraint}, where $\beta=0.1$ is selected.
+
+The robot follows the complete Cartesian candidate search method to determine future poses, with $k_\text{dist}=5$ to avoid large motions with little added expected information gains.
+Due to the large size of the map, candidates below the expected information gain of $1.25$ are neglected from future consideration.
+
+The time sequence of the exploration for every 5~minutes are simulated in ROS Stage 2D environment.
+The corresponding simulation results are illustrated in Figure \ref{fig:IRL} and a video is available at \href{https://www.youtube.com/watch?v=5VdzKHreB_s}{\WriteBlue{https://www.youtube.com/watch?v=5VdzKHreB\_s}}.
+Lighter red dots and darker blue dots in the explored area correlate to the level of information gain at those positions with low and high value, respectively.
+It is shown that the proposed autonomous exploration algorithm successfully mapped the complicated environment composed of a number of rooms, narrow hallways, and open spaces that are irregularly shaped.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.4\textwidth]{intel_clean.png}
+    \caption{Modified Intel Research Lab floor plan form the SLAM benchmark~\cite{kummerle2009measuring}.}
+\label{fig:intel}
+\end{figure}
+
+\begin{figure}[!ht]
+    \centering
+    \begin{subfigure}[t]{0.3\columnwidth}
+        \centering
+        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{0min.png}
+        \caption{$t=0$ min}
+        \label{fig:IRL0min}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.3\columnwidth}
+        \centering
+        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{5min.png}
+        \caption{$t=5$ min}
+        \label{fig:IRL5min}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.3\columnwidth}
+           \centering
+           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{10min.png}
+        \caption{$t=10$ min}
+        \label{fig:IRL10min}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.3\columnwidth}
+           \centering
+           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{15min.png}
+        \caption{$t=15$ min}
+        \label{fig:IRL15min}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.3\columnwidth}
+         \centering
+         \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{20min.png}
+        \caption{$t=20$ min}
+        \label{fig:IRL20min}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.3\columnwidth}
+           \centering
+           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{25min.png}
+        \caption{$t=25$ min}
+        \label{fig:IRL25min}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.3\columnwidth}
+           \centering
+           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{30min.png}
+        \caption{$t=30$ min}
+        \label{fig:IRL30min}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.3\columnwidth}
+           \centering
+           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{35min.png}
+        \caption{$t=35$ min}
+        \label{fig:IRL35min}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.3\columnwidth}
+           \centering
+           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{40min.png}
+        \caption{$t=40$ min}
+        \label{fig:IRL40min}
+    \end{subfigure}
+    \caption{A robot (green) measures a room in ROS Stage simulator using the Intel Research Lab floor plan. The exact inverse sensor model is used for the autonomous exploration algorithm.}
+    \label{fig:IRL}
+\end{figure}
+
+% TODO: move these to final chapter
+%\section{Experimental Result}
 % JINT17 9, 10, 11 (ground experiment at the NRL)
@@
 \section{Conclusions}
+
+The proposed autonomous exploration scheme determines the expected value of entropy for various potential robotic actions.
+The optimal policy maximizes an objective function that includes expected information gain and travel distance.
+Collision-free paths to optimal poses are determined with Dijkstra's search and a constrained polynomial least squares trajectory.
+These are demonstrated with two numerical examples, showing the expanding ring and complete Cartesian approaches.
+Due to superior scalability, the complete Cartesian approach is applied to subsequent exploration algorithms in this dissertation.

### 2018-06-25 Dissertation/Thesis/tex/ae3Dsimple.tex
wrote chapter 4, needs editing (3 & 4 still)

@@
 
+In this chapter, we consider mapping and exploring in 3D environments.
+We consider common spaces like rooms and hallways where objects are largely vertically-uniform.
+A complete 3D probabilistic map is projected onto a simpler 2D map at a fixed exploration height, where the projected map can be used for collision-avoidance and entropy predictions.
+
 \section{Occupancy Grid Mapping in 3D}
 
-\section{Exploration in Simple Environments }
+The inverse sensor model from Chapter \ref{chap:pogm} is based on an arbitrary vector spanning 2D space.
+The distances from the robot to grid cells are obtained geometrically through ray casting.
+This method is easily extended to 3D space, and the inverse sensor model is applied identically.
+Since the map is updated ray-by-ray (each measurement ray is considered once by itself), this approach is extended for measurement fusion of any number of sensors; the map is updated by each ray of a scan sequentially, then the map is updated by the rays of the next scan in queue, which might have different characteristics.
+There are several reasons why a robot might be equipped with multiple sensors, e.g., different fields-of-view or depth sensor properties.
+For example, a robot may be equipped with a 2D LIDAR with highly accurate measurements, but a limited single-plane field-of-view.
+Then, the robot can compensate for the LIDAR limitations with an IR depth sensor that provides many more measurement rays with a 3D scan, where each ray has lower accuracy than the LIDAR rays.
+Since each sensor model is explicitly considered in the cell probability calculations, measurements from both sensors may serve to update the same probabilistic map.
+However, the impact from a LIDAR measurement with high accuracy on the probabilistic map will be greater than that of the IR depth sensor with lower accuracy.
+Since sensor properties are considered explicitly, the proposed 3D mapping approach can fuse the impact of measurements on the probabilistic map from different sources with different properties.
 
-\subsection{Map Projections}
+However, implementation in 3D requires several careful considerations.
+First, the number of cell occupancy probabilities tracked with computer memory increases by a factor of the number of cells in the added dimension.
+In its current form, the proposed method assumes fixed map limits, so these should be carefully selected with grid cell resolution based on available memory.
+Second, measurement rays spanning 3D space typically involve more intersections with grid cells.
+The computational order of ray casting and the inverse sensor model are proportional to the number of grid cells along the ray $n_r$, so careful consideration must be placed on sensor limits and rays per scan that can be considered.
+Typically, computers small enough to be carried onboard flying robots lack the processing capabilities to update massive maps quickly, particularly because the ray-by-ray updating method requires many computations in series with large measurement scans.
+Therefore, an onboard computer can simply stream the sensor data, and a more powerful computer can run the mapping.
 
-\section{Exploring a Large Room}
-% IRC18 result: Fig. 4, 5, 6, 7
+Additionally, mobile robots, aerial vehicles in particular, are frequently subject to fast movements, and the time stamps for pose estimates and depth sensor readings are not guaranteed to align.
+For example, the most recent pose estimate and measurement scan may have significantly different time stamps while the robot is rapidly rotating.
+Using these together violates the assumption that the robot pose is well-known for every measurement scan, i.e., $X_t$ and $Z_t$ are given.
+Therefore, pose estimates and sensor readings must be paired according to their time stamps.
+A useful tool that easily satisfies this goal is the Message Filters package with the Robot Operating System (ROS), where stamped poses and measurement scans are synchronized using recently received data.
+
+\section{Exploration in Vertically-Uniform Environments}
+
+Next, we describe two approaches to projecting the 3D probabilistic map onto a 2D map at the exploration height.
+Assuming projections are sufficient is consistent with rooms and hallways with non-complex geometries.
+
+\subsubsection{Collision and Entropy Maps}
+Occupancy grid map cell volume is frequently smaller than the size of the robot taking measurements, particularly with highly-detailed maps.
+This motivates a method to simply represent locations that risk collision, referred to as a \emph{collision map} $C$.
+Let $k_C\geq1$ be an integer representing the minimum number of 3D grid cells that form a cube that can completely encompass the robot, i.e., the collision map cell edge length is
+\begin{align}
+\label{eqn:alphaC}
+\alpha_C=k_C\alpha.
+\end{align}
+Consider the set $m_{C,k}\subset m$ consisting of the 3D grid cells falling inside the limits of the $k$-th cell of collision map $C$, namely $\mathbf{C}_k$. Then the collision probability is
+\begin{align}
+P(\mathbf{C}_k)=\max{(P(\mathbf{m}_i)\ \forall \ i\in m_{C,k})}.
+\end{align}
+Then, the robot is only allowed to consider moving into $k$-th cell of the collision map if
+\begin{align}
+P(\mathbf{C}_k) \leq C_\text{thresh},
+\end{align}
+where $C_\text{thresh}>0$ is a maximum acceptable collision threshold, typically chosen slightly greater than $0$.
+In short, this conservative approach limits robotic motion to regions with a low risk of collision while neglecting objects far above or below the exploration plane.
+
+In a similar method, \emph{entropy map} $E$ determines cells based on the entropy metric of \refeqn{ShannonsEntropyCell} over a different subset of cells.
+Let the $E$ be a 2D map composed of cells with edge length $\alpha$ located at the exploration height.
+Let the set $m_{E,k}\subset m$ be the cells directly above and below the $k$-th cell of $E$, neglecting any 3D cells located lower than the floor height.
+Selecting this set is advantageous because measurements are assumed not to penetrate occupied cells (edge length $\alpha$) and measurement rays are typically not closely-aligned with the third axis of the inertial frame (vertically long set).
+Furthermore, any cells located below the floor, which cannot be reached, should not impact the uncertainty of the space.
+Then, the occupancy probability of the $k$-th entropy map cell is selected as
+\begin{align}
+\label{eqn:ProbEntropyMap}
+P(\mathbf{E}_k)=\argmax_{P(\mathbf{m}_i)}(H(P(\mathbf{m}_i))\ \forall \ i\in m_{E,k}).
+\end{align}
+After several measurements of the cells belonging to $m_{E,k}$, it is common for $H(P(\mathbf{E}_k))\approx0$, which implies that these cells are well-known.
+However, this does not imply that the captured space is \emph{entirely} free or occupied (e.g.
+an object on the ground that does not breach the upper map boundary), so applying \refeqn{ProbEntropyMap} alone could arbitrarily set $P(\mathbf{E}_k)$ close to $0$ or $1$.
+This can have damaging effects on the expected information of cells beyond the $k$-th cell.
+Let $P_\text{min}>0$ be a lower bound such that $H(P_\text{min})=H(1-P_\text{min})\approx0$.
+If $H(P(\mathbf{E}_k))\leq H(P_\text{min})$, then this probability is corrected to
+\begin{align}
+P(\mathbf{E}_k)= 
+\begin{cases}
+    P_\text{min},			& \text{if } \text{mean}(P(\mathbf{m}_i)\ \forall \ i\in m_{E,k})\leq0.5\\
+    1-P_\text{min},              & \text{otherwise}
+\end{cases},
+\end{align}
+where the first case corresponds to mostly open space and the second case corresponds to mostly occupied space (e.g.
+walls and objects).
+In short, the candidates are only considered if they can be reached over collision map $C$ via Dijkstra's search, and the information gains of the reachable candidates are predicted using entropy map $E$.
+
+\subsubsection{Collision and Entropy Combination Map}
+The second proposed technique combines the advantages of both the collision and entropy maps into a 2D projected occupancy grid map, which can be used for both tasks.
+Let the \emph{collision and entropy combination map} $B$ be a 2D projected map located at the exploration height with grid cell edges the same as the collision map, namely $\alpha_C$ from \refeqn{alphaC}, using the same subset of grid cells from the collision map, namely $m_C$.
+Considering that the cells begin with uniform probability $P_\text{init}$ where $0<P_\text{init}<1$, define a threshold $P_\text{thresh}$ such that $P_\text{init}<P_\text{thresh}<1$.
+Then, the probability assigned to the $k$-th grid cell of $B$ is
+\begin{align}
+P(\mathbf{B}_k)= 
+\begin{cases}
+    P(\mathbf{C}_k),			& \text{if }P(\mathbf{C}_k)\geq P_\text{thresh}\\
+    \min{(P(\mathbf{m}_i)\ \forall \ i\in m_{C,k})},              & \text{otherwise}
+\end{cases}.
+\end{align}
+
+This approach has several advantages.
+First, measurements must enter the vicinity of the $k$-th cell for $P(\mathbf{B}_k)$ to change from $P_\text{init}$.
+Without prior measurements, capturing $P(\mathbf{B}_k)$ produces a large reward for the expected information gain in \refeqn{ObjFun}.
+Second, the value of $P_\text{thresh}$ can be modified based on how conservative the collision risk assessment for a robot should be.
+Additionally, when $P(\mathbf{B}_k)<P_\text{thresh}$, this event tends to favor free cells such that expected information gains of cells beyond the $k$-th cell have greater impact on exploration.
+Furthermore, a single map is simpler, and since cell sizes satisfy $\alpha_C\geq\alpha$, it follows that $n_r$ can only be decreased with $\alpha_C$, thereby increasing computational speed.
+
+\section{Numerical Example}
+
+The following numerical example involves a simulated quadrotor taking measurements of a large room and generating a 3D occupancy grid map.
+Both types of map projections are used for autonomous exploration.
+
+\subsection{Software Structure}
+%        A.      Software Structures
+%                list and explain ROS nodes
+
+The Robot Operating System (ROS) provides an excellent framework for several programs, referred to as \emph{nodes}, to communicate easily.
+These are mostly coded in C++ for speed, except for a few Python scripts for a GUI and trivial message conversions, and launch/yaml scripts for running nodes and setting parameters.
+
+Gazebo is a 3D dynamic simulator~\cite{KaeHow04}, with plugins to simulate an Asus Xtion color depth sensor and a Hokuyo LIDAR.
+Gazebo also provides position and attitude transforms from the world to the quadrotor body, and the ability to set robot model states directly.
+
+Using fixed transforms between the quadrotor body and its onboard sensors, an original mapping node uses a variation of ROS message filters, known as transform message filters, to synchronize the sensor poses with sensor readings.
+These serve as inputs for ray casting in 3D, which provides the cell probabilities and depths along the ray through geometry.
+Then, the inverse sensor model serves to update the map ray-by-ray.
+The node is designed to work for any number of sensors, specified by sensor parameters in a yaml configuration file.
+
+There are two processes for exploration.
+The first process is projecting the 3D map onto 2D maps; the same code under different conditional parameters produces two nodes, for collision and entropy maps.
+The second process subscribes to these two 2D maps; the collision map serves as input for candidate pose consideration and Dijkstra's search, and the entropy map serves to predict information gain.
+This node provides path messages, composed of desired poses $0.01$sec apart, displayed with a ROS visualization package, Rviz.
+
+Finally, a path-interpreter node subscribes to the path messages and interpolates the pose at the current ROS time step. This pose is converted into a message that sets the Gazebo quadrotor state.
+
+\subsection{Simulation Results}
+%        B.      Simulation Results
+%                a little bit more interpretations for the results
+
+The 3D mapping and exploration algorithms are simulated with several parameters.
+In particular, the Xtion and Hokuyo are modeled using normalized Gaussian and uniform distributions for the sensor model, with a Gaussian probability of $0.9$ with $\sigma_\text{Xtion}=0.25$m and $\sigma_\text{Hokuyo}=0.1$m, respectively, which are greater than their specified values to account for pose uncertainty and distortion with the Xtion.
+Since the Xtion tends to distort depths significantly beyond $4$m, this maximum reading considered for both the Xtion and Hokuyo, and the uniform distribution (probability $0.1$) covers the sensor ranges from $0.5$m to $4$m.
+A cell size of $\alpha=0.075$m is selected and the map limits are $-4$m to $4$m in the x-direction, $-6$m to $6$m in the y-direction, and $-0.15$m to $1.5$m in the z-direction.
+The initial probability of cells is selected at $P_\text{init}=0.1$.
+For the collision map, the cell edge factor is $k_C=3$.
+Additionally, Dijkstra's algorithm produces a cost map composed of distance costs for each collision map cell, denoted $d_\text{cell}$, from the current robot position to the candidate pose locations over a collision-free path about the occupancy grid.
+A normalized variation on the \emph{bump function}~\cite{Joh06} with maximum distance $d_\text{max}>0$,% TODO: CHECK!
+\begin{align}
+\label{eqn:bumpFun}
+\text{bump}(d_\text{cell})= 
+\begin{cases}
+    e^{1-\frac1{1-(d_\text{cell}/d_\text{max})^2}},			& \text{if }-d_\text{max}<d_\text{cell}<d_\text{max}\\
+    0,              				& \text{otherwise}
+\end{cases},
+\end{align}
+is multiplied to \refeqn{ObjFunApproxCompleteCartesian} in place of the Euclidean cost term, namely $k_\text{dist}(x_t-x_c)^\T(x_t-x_c)$, which is removed.
+For this experiment, $d_\text{max}$ is chosen as the maximum value of $d_\text{cell}$ on the cost map.
+This serves to prioritize understanding the local surrounding space before the robot moves across the map to distant regions, while still considering all of the pose candidates.
+
+In the simulation, the robot explores the simulated environment for $8$min while taking measurements.
+The 3D occupancy grid map and true environment are shown in Figure \ref{fig:sim3DMap}.
+The Gazebo simulated environment and two 2D projected maps for collision and entropy are shown in Figure \ref{fig:sim2Dmaps}.
+The total map entropy as a function of time is shown in Figure \ref{fig:simH}.
+
+\begin{figure}[!t]
+    	\begin{subfigure}[t]{0.24\columnwidth}
+           	\centering
+          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_1min.jpg}
+        		\caption{$t=1$min}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.24\columnwidth}
+           	\centering
+          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_2min.jpg}
+        		\caption{$t=2$min}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.24\columnwidth}
+           	\centering
+          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_3min.jpg}
+        		\caption{$t=3$min}
+   	\end{subfigure}
+    	\begin{subfigure}[t]{0.24\columnwidth}
+           	\centering
+          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_4min.jpg}
+        		\caption{$t=4$min}
+    	\end{subfigure}
+	\begin{subfigure}[t]{0.24\columnwidth}
+	\vspace*{0.15\columnwidth}
+           	\centering
+          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_5min.jpg}
+        		\caption{$t=5$min}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.24\columnwidth}
+	\vspace*{0.15\columnwidth}
+           	\centering
+          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_6min.jpg}
+        		\caption{$t=6$min}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.24\columnwidth}
+	\vspace*{0.15\columnwidth}
+           	\centering
+          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_7min.jpg}
+        		\caption{$t=7$min}
+   	\end{subfigure}
+    	\begin{subfigure}[t]{0.24\columnwidth}
+	\vspace*{0.15\columnwidth}
+           	\centering
+          	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_8min.jpg}
+        		\caption{$t=8$min}
+    	\end{subfigure}
+%\includegraphics[width=2.5in]{myfigure}
+% where an .eps filename suffix will be assumed under latex, 
+% and a .pdf suffix will be assumed for pdflatex; or what has been declared
+% via \DeclareGraphicsExtensions.
+\caption{Simulated 3D Occupancy Grid Map}
+\label{fig:sim3DMap}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[height=1.5\textwidth]{gazebo_view.png}
+        		\caption{Environment}
+    	\end{subfigure}
+		\hspace*{0.05cm}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim = {3.6cm 17cm 52.5cm 4cm}, clip, height=1.5\textwidth]{gazebo_4min.jpg}
+        		\caption{Collision Map}
+    	\end{subfigure}
+	\hspace*{0.1cm}
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim = {18cm 17cm 38.1cm 4cm}, clip, height=1.5\textwidth]{gazebo_4min.jpg}
+        		\caption{Entropy Map}
+		\label{fig:sim2DmapsC}
+    	\end{subfigure}
+%\includegraphics[width=2.5in]{myfigure}
+% where an .eps filename suffix will be assumed under latex, 
+% and a .pdf suffix will be assumed for pdflatex; or what has been declared
+% via \DeclareGraphicsExtensions.
+\caption{Simulated Environment and 2D Projected Maps at $4$min}
+\label{fig:sim2Dmaps}
+\end{figure}
+
+\begin{figure}[!t]
+	\centering
+	\includegraphics[width=0.9\textwidth]{entropy_gazebo_aspect3by1.png}
+	\caption{Simulated Environment Map Entropy}
+	\label{fig:simH}
+\end{figure}
+
+The two-map approach exhibits advantages and disadvantages.
+The exploration algorithm successfully avoided collisions and gave way to a fairly complete 3D occupancy grid.
+The floor, which is not always visible, is properly estimated in most places.
+Conversely, this attention to detail has negative consequences that encourage the robot to repeatedly look at the same regions from different vantage points.
+Since an entropy map cell is composed from many 3D grid cells single-stacked vertically, if just one of these 3D map cells is not captured or hardly captured, the entropy map cell will be recorded as quite uncertain, shown covering numerous locations in Figure \ref{fig:sim2DmapsC}.
+The robot may attempt to measure this space again, even when other regions might more strongly warrant observation.
+These periods of sluggish entropy decrease are shown with flatter sections of Figure \ref{fig:simH}.
+In short, this algorithm successfully explores the space collision-free while producing a complete map, but over-attention to just a few uncertain cells has negative consequences on the exploration speed.
+
+% TODO: move below to experimental results
+%\section{Exploring a Large Room}
+% IRC18 result: Figure 4, 5, 6, 7
 
 \section{Conclusions}
+
+In this chapter, we proposed an extension of exact probabilistic occupancy grid mapping from 2D to 3D, then proposed a simplified autonomous exploration scheme for vertically-uniform environments.
+Two versions of the projected map, namely one for collision and one for entropy, are used in a numerical example.
+The collision and entropy combination map is applied to an experimental result in Chapter \ref{chap:Experiments}.

### 2018-06-25 Dissertation/Thesis/tex/results.tex
wrote chapter 4, needs editing (3 & 4 still)

% !TEX root = ../thesis-sample.tex

\chapter{Experimental Results}\label{chap:Experiments}

\section{A section}

\subsection{A subsection}

### 2018-06-26 Dissertation/Thesis/tex/appendixB.tex
added new tex files and local pub figs

% !TEX root = ../thesis-sample.tex
\appendix
\doublespacing
\chapter{Proof of Proposition \ref{prop:ExpectedH}}

\subsection{Single Ray Expected Value of Entropy}

In this appendix, we prove the expected value of entropy for a single measurement ray. We consider the expected entropy due to measurement ray $z$ at pose $X$,
\begin{align}
\label{eqn:HRayInt}
\text{E}[H(P(m|&X,z))]
=\int_{z_\text{min}}^{z_\text{max}}
H(P(m|X,z))p(z|X)
dz.
\end{align}
%where $z_\text{min}$ and $z_\text{max}$ are sensor limits. Considering the continuous space of possible measurements of $z$ would make \refeqn{HRayInt} intractable to calculate. Thus, w
We discretize the measurement ray space such that $z$ falls on points along the measurement ray intersecting with grid cell edges.
The discretized expected value of \refeqn{HRayInt} is
\begin{align}
\label{eqn:DiscExpEntropyRayAppendix}
\text{E}[H(P(m|&X,z))]=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(m|X,z_{k}))P(z_{k}|X)\bigg\},
\end{align}
where index $z_{k}$ is the distance from the location of pose $X$ to the $k$-th grid cell of the reduced map $r\subset m$ for this ray, known from geometry (see Figure \ref{fig:RayCastingIllustration}).

Standing alone, the term $P(z_{k}|X)$ from \refeqn{DiscExpEntropyRayAppendix} has a convoluted meaning because the depth $z_{k}$ does not directly depend on the map. However, we present a method to obtain this discretized probability. %Despite how this term is replaced with a normalizer in ~\cite{ThrBurFox05} replaces
Following the assumption that $z$ is discretized to known distances, the probabilities are proportional to their densities, as the area under the density curve is infinitesimal and fixed. Accounting for all cases, the probability is
\begin{align}
\label{eqn:ProbWithDelta}
P(z_{k}|X)=\frac{p(z_{k}|X)}{\sum_{i=1}^{n_{r}+1}p(z_{i}|X)}.
\end{align}
Conveniently, this density corresponds to the inverse normalizer defined in \refeqn{BayesRuleRayISM} as the denominator from Bayes' rule,
\begin{align}
\label{eqn:DiscretizedProb}
p(z_{k}|X)=\eta_{k}^{-1},
\end{align}
where $\eta_{k}^{-1}$ is defined in \refeqn{allEta} as the inverse sensor normalizer corresponding to a measurement of the $k$-th cell of reduced map $r$. Thus, the inverse normalizer is
\begin{align}
\label{eqn:allEtaAtCAppendix}
\eta_{k}^{-1}
&=
\sum_{i=1}^{n_{r}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{j})\bigg\} p(z_{k}|\mathbf{r}_{i+},X)P(\mathbf{r}_{i}).
\end{align}
By substituting \refeqn{ProbWithDelta} and \refeqn{DiscretizedProb} into \refeqn{DiscExpEntropyRayAppendix}, the expected entropy from individual measurement ray $z$ at $X$ is
\begin{align}
\label{eqn:HRayCompleteAppendix}
&\text{E}[H(P(m|X,z))]=\left(\sum_{i=1}^{n_{r}+1}\eta_{i}^{-1}\right)^{-1}\sum_{k=1}^{n_{r}+1}\bigg\{H(P(m|X,z_{k}))\eta_{k}^{-1}\bigg\},
\end{align}
where $\eta_{k}^{-1}$ is taken from \refeqn{allEtaAtCAppendix}.
Here, \refeqn{HRayCompleteAppendix} provides the expected entropies for cells inside the field of view of reduced map $r$. %; cells outside this reduced map are trivially changed by $0$ due to $z$.
The entropies of cells outside the field of view remain unchanged.

### 2018-06-26 Dissertation/Thesis/tex/frontmatter.tex
added mapping background & beginning of autonomous exploration lit review
fixed committee members in frontmatter

@@
 
+Robert Pless\\
+Department of Computer Science\\
+The George Washington University\\
+Committee Member\\
+
 James Hahn\\
@@
 
-Yin-Lin Shen\\
-Department of Mechanical and Aerospace Engineering\\
+Chung Hyuk Park\\
+Department of Biomedical Engineering\\
 The George Washington University\\ 
+Committee Member\\
+
+Zhuming Ai\\
+Information Management and Decision Architectures\\
+U.S. Naval Research Laboratory\\ 
 Committee Member
+
 }
@@
 Finally, I would like to thank my research advisor, Taeyoung Lee, for the excellent guidance over the last six years.
-Unlike many relationships between advisors and students, we treats his students like colleagues rather than subordinates.
+Unlike many relationships between advisors and students, he has treated his students like colleagues rather than subordinates.
 This way, I could focus on completing my research objectives properly rather than trying to impress superior figure.

### 2018-06-26 Dissertation/Thesis/tex/ae2D.tex
edited chap 3 through 3.2.2

@@
 Much like the probabilistic mapping, this is achieved ray-by-ray~\cite{KauAiLee16,KauTakAiLee17}.
-Considering that all grid cell probabilities are conditioned on the history of poses $X_{1:t}$ and measurement scans $Z_{1:t}$, these terms are removed from the remaining equations for simplicity.
+Considering that all grid cell probabilities are conditioned on the history of poses $X_{1:t}$ and measurement scans $Z_{1:t}$, these terms are removed from the remaining equations of this dissertation for simplicity (excluding Appendices).
 
@@
 which is the probability that $\mathbf{r}_{k}$ is the closest occupied grid cell based on past poses and measurement scans, independent of cells beyond the $k$-th cell from $x_c$.
-Let $\hat n>0$ be a fixed number of grid cells for all rays such that $\hat n\leq n_{r}+1$ for all $l$.
+Let $\hat n>0$ be a fixed number of grid cells such that $\hat n\leq n_{r}+1$.
 Let $\hat{r}$ correspond to the grid cells that yield the $\hat{n}$ maximum values of \refeqn{ProbOfFirstCell} (the $\hat n$ most likely ray detections), indexed by increasing distance from candidate location $x_c$.
 By replacing the reduced map $r$ with $\hat{r}$ and changing the summation limits to $\braces{1,2,...,\hat n}$ in \refeqn{DiscExpEntropyRay} and \refeqn{ProbMeas}, the order of computation is reduced to $\mathcal O({\hat{n}}^2)$.
-Even though the value of $n_{r}$ is different among various rays in general, $\hat n$ is fixed, so the computational order is fixed as well.
+Even though the value of $n_{r}$ is different among various measurement rays in general, $\hat n$ is fixed for all rays, so the computational order is fixed as well.
 In short, this method reduces the required computation substantially by systematically neglecting those grid cells with little effect.
@@
 The purpose of this numerical example is to provide evidence that the approximations are reasonable and increase the algorithm speed substantially.
-Since a measurement ray produces a depth measurement in a single direction, we only consider a 1D map where the grid cells have spacing $\alpha=0.2$ m, and the properties of the range sensor are based on the Microsoft Kinect~\cite{PirRutBisSch11,KhoElb12} with maximum reading depth $z_\text{max}=4$ m ($20$ grid cells inside the sensor FOV).
+Since a measurement ray returns a range in a single direction, we only consider a 1D map where the grid cells have spacing $\alpha=0.2$ m, and the properties of the range sensor are based on the Microsoft Kinect~\cite{PirRutBisSch11,KhoElb12} with maximum reading depth $z_\text{max}=4$ m ($20$ grid cells inside the sensor FOV).
 The goal is to compare the expected entropy $\text{E}[H(P(m|x_c,z_{c}))]$ from \refeqn{DiscExpEntropyRay} and with an approximation $\text{E}[H_\text{approx}(P(m|x_c,z_{c}))]$, which only considers $\hat n$ grid cells with highest detection probability \refeqn{ProbOfFirstCell}.
@@
 In this section, we show how expected entropy changes from single measurement rays can be integrated into predicting the expected information gain of measurement scans.
-This metric, and a cost associated with travel time, are integrated to determine an optimal future pose and collision-free path in a 2D environment.
+This metric, and a cost associated with travel time, are combined to determine an optimal future pose and collision-free path in a 2D environment.
 
@@
 These rays are denoted $\braces{z_{c,1},z_{c,2},\ldots,z_{c,n_d}}$.
-Attitudes correspond to the rays such that the robot sensor is aligned with the ray direction, and these are denoted $\braces{R_{c,1},R_{c,2},\ldots,R_{c,n_d}}$, where the scan with attitude $R_{c,d}$ might cover several ray directions depending on the sensor field of view (FOV).
+Attitudes correspond to the rays such that the robot sensor is aligned with the ray direction, and these are denoted $\braces{R_{c,1},R_{c,2},\ldots,R_{c,n_d}}$, where the scan with attitude $R_{c,d}$ might cover several ray directions depending on the sensor FOV.
 We choose optimal attitude $R_c^*$ as the summation of the expected entropy changes covered by the scan,
@@
 \label{eqn:FindRc}
-&R_c^*=\argmax_{R_{c,d}}\sum_{z_{c,i}\in R_{c,d}\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))
--\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,i}))]\bigg).
+&R_c^*=\argmax_{R_{c,d}}\sum_{z_{c,i}\in R_{c,d}\text{ FOV}}\bigg(H(P(m))
+-\text{E}[H(P(m|x_c,z_{c,i}))]\bigg).
 %\mathcal I_\text{ray}(x_c,z_{c,i})\right).
@@
 The set of optimal attitudes at each candidate location is $\braces{R^*_{1},R^*_{2},\ldots,R^*_{n_c}}$, obtained from \refeqn{FindRc}.
-The information gain objective function,
+The information gain objective function for a scan $Z_c$ captured from pose $X_c$ is
 \begin{align}
@@
 \label{eqn:ObjFunApprox}
-\mathcal I(x_c,R_c^*)&\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*,z_{c,i}))\right]\bigg),
+\mathcal I(x_c,R_c^*)&\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m))-\text{E}\left[H(P(m|x_c,z_{c,i}))\right]\bigg),
 \\
@@
 Additionally, the distance to these poses need not be explicitly considered in the pose determination optimization because the distance costs are identical to each candidate, assuming objects do not occlude the trajectories.
-The main drawback is that poses outside a small local region of the robot are frequently neglected, and the robot frequently executes repeated short motions, often without large information gains.
+The main drawback is that poses outside a small local region of the robot are frequently neglected, causing the robot to repeatedly execute short motions, often without large information gains.
 
@@
 \label{eqn:ObjFunApproxCompleteCartesian}
-\mathcal I&(x_c,R_c^*,x_t)\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*(x_c),z_{c,i}))\right]\bigg)-k_\text{dist}(x_t-x_c)^\T(x_t-x_c),
+\mathcal I(x_c,R_c^*,x_t)\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m))-\text{E}\left[H(P(m|x_c,z_{c,i}))\right]\bigg)-k_\text{dist}(x_t-x_c)^\T(x_t-x_c),
 \end{align}
@@
 Furthermore, the same candidate locations are considered throughout the repeated processes of exploration, and the map cell occupancies are assumed static.
-Therefore, if candidate location $x_c$ fails to satisfy $\mathcal I(x_c,R_c^*,x_t)\geq\mathcal I_\text{min}$, then $x_c$ need \emph{never} be considered again.
+Therefore, if candidate location $x_c$ fails to satisfy $\mathcal I(x_c,R_c^*,x_t)+k_\text{dist}(x_t-x_c)^\T(x_t-x_c)\geq\mathcal I_\text{min}$, then $x_c$ need \emph{never} be considered again.
 Avoiding these unnecessary calculations yields a greatly lowered computational burden, particularly late in exploration where several regions are collision-free but disadvantageous to revisit.

### 2018-06-26 Dissertation/Thesis/tex/ae2D.tex
edited up to sec. 4.2

@@
 Dijkstra's search is naturally applied to this problem because the occupancy grid map serves nicely as a graph, and Dijkstra's algorithm provides an optimal collision-free trajectory along that graph.
-There are two steps to Dijkstra's algorithm.
-first, generate a cost map from the robot location to each grid cell on the 2D occupancy grid map.
+There are two steps to Dijkstra's algorithm: first, generate a cost map from the robot location to each grid cell on the 2D occupancy grid map.
 Only cells that satisfy \refeqn{CollisionInequalityConstraint} are considered to avoid possible collisions.
-The cost to travel to neighboring cells that share an edge are the cell edge length $\alpha$, and the cost to travel to cells that share a corner is $\sqrt{2}\alpha$.
+Traveling to a neighboring cells that shares an edge costs the edge length distance $\alpha$, and the cost to travel to cells that share a corner is $\sqrt{2}\alpha$.
 This provides the collision-free distances to all reachable locations on the map.
@@
 Here, a robot autonomously maps and explores a 2D environment composed of two rooms and one hallway.
-The robot models its surroundings with an occupancy grid of $15000$ cells where grid cell edges have length $\alpha=0.2$m, composing a map with dimensions $30\text{m}\times20\text{m}$.
+The robot models its surroundings with an occupancy grid of $15,000$ cells where grid cell edges have length $\alpha=0.2$m, composing a map with dimensions $30\text{m}\times20\text{m}$.
 The initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) for grid cells covered by the circular robot of radius $0.1$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
@@
 The robot motion is governed by a policy that maximizes the map information gain within its set of pose candidates, where the $\hat n=6$ is chosen to approximate the expected entropy of each ray.
-When running the exploration algorithm in Robot Operating System (ROS), the mean computation time is $0.0194$ seconds to determine the optimal future pose and complete Dijkstra's algorithm on the occupancy grid.
+When running the exploration algorithm within the framework of the Robot Operating System (ROS), the mean computation time is $0.0194$ seconds to determine the optimal future pose and complete Dijkstra's algorithm on the occupancy grid.
 The computation times for this map reached roughly $1$ second at maximum, corresponding to rare cases when the robot is locally surrounded by a highly-certain environment; this task requires evaluating many more pose candidates and motion planning over a larger terrain.
@@
 
-Next, the proposed autonomous exploration approach is applied to the floor plan of Intel Research Lab, illustrated in Figure~\ref{fig:intel}.
+Next, the proposed autonomous exploration approach is applied to the floor plan of the Intel Research Lab, illustrated in Figure~\ref{fig:intel}.
 The robot explores its surroundings with an occupancy grid with $90,000$ cells where grid cell edges are $\alpha=0.1$m, composing a map with dimensions $30\text{m}\times30\text{m}$.
@@
 
-The time sequence of the exploration for every 5~minutes are simulated in ROS Stage 2D environment.
+The time sequence of the exploration for every 5~minutes are simulated in the ROS Stage 2D environment.
 The corresponding simulation results are illustrated in Figure \ref{fig:IRL} and a video is available at \href{https://www.youtube.com/watch?v=5VdzKHreB_s}{\WriteBlue{https://www.youtube.com/watch?v=5VdzKHreB\_s}}.
@@
     \includegraphics[width=0.4\textwidth]{intel_clean.png}
-    \caption{Modified Intel Research Lab floor plan form the SLAM benchmark~\cite{kummerle2009measuring}.}
+    \caption{Intel Research Lab floor plan benchmark~\cite{kummerle2009measuring}.}
 \label{fig:intel}
@@
     \end{subfigure}
-    \caption{A robot (green) measures a room in ROS Stage simulator using the Intel Research Lab floor plan. The exact inverse sensor model is used for the autonomous exploration algorithm.}
+    \caption{A robot (green) measures a room in the ROS Stage simulator using the Intel Research Lab floor plan. The exact inverse sensor model is used for the autonomous exploration algorithm.}
     \label{fig:IRL}

### 2018-06-26 Dissertation/Thesis/tex/ae3Dsimple.tex
edited up to sec. 4.2

@@
 
-In this chapter, we consider mapping and exploring in 3D environments.
-We consider common spaces like rooms and hallways where objects are largely vertically-uniform.
+In this chapter, we introduce mapping and autonomous exploration in 3D environments.
+Here, we consider common spaces like rooms and hallways where objects are largely vertically-uniform.
 A complete 3D probabilistic map is projected onto a simpler 2D map at a fixed exploration height, where the projected map can be used for collision-avoidance and entropy predictions.
@@
 Second, measurement rays spanning 3D space typically involve more intersections with grid cells.
-The computational order of ray casting and the inverse sensor model are proportional to the number of grid cells along the ray $n_r$, so careful consideration must be placed on sensor limits and rays per scan that can be considered.
+The computational orders of ray casting and the inverse sensor model are proportional to the number of grid cells along the ray $n_r$, so careful consideration must be placed on sensor limits and rays per scan that can be considered.
 Typically, computers small enough to be carried onboard flying robots lack the processing capabilities to update massive maps quickly, particularly because the ray-by-ray updating method requires many computations in series with large measurement scans.
@@
 
-Additionally, mobile robots, aerial vehicles in particular, are frequently subject to fast movements, and the time stamps for pose estimates and depth sensor readings are not guaranteed to align.
+Additionally, mobile robots, particularly aerial vehicles, are frequently subject to fast movements, and the time stamps for pose estimates and depth sensor readings are not guaranteed to align.
 For example, the most recent pose estimate and measurement scan may have significantly different time stamps while the robot is rapidly rotating.
@@
 Therefore, pose estimates and sensor readings must be paired according to their time stamps.
-A useful tool that easily satisfies this goal is the Message Filters package with the Robot Operating System (ROS), where stamped poses and measurement scans are synchronized using recently received data.
+A useful tool that easily satisfies this goal is the Message Filters package with the Robot Operating System (ROS), where stamped poses and measurement scans are synchronized using recently-received data.
 

### 2018-06-26 Dissertation/Thesis/tex/ae2D.tex
finished editing chap. 4

@@
 \label{eqn:ObjFunApproxCompleteCartesian}
-\mathcal I(x_c,R_c^*,x_t)\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m))-\text{E}\left[H(P(m|x_c,z_{c,i}))\right]\bigg)-k_\text{dist}(x_t-x_c)^\T(x_t-x_c),
+\mathcal I(x_c,R_c^*,x_t)\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m))-\text{E}\left[H(P(m|x_c,z_{c,i}))\right]\bigg)-k_\text{dist}\norm{x_t-x_c}^2,
 \end{align}
@@
 Furthermore, the same candidate locations are considered throughout the repeated processes of exploration, and the map cell occupancies are assumed static.
-Therefore, if candidate location $x_c$ fails to satisfy $\mathcal I(x_c,R_c^*,x_t)+k_\text{dist}(x_t-x_c)^\T(x_t-x_c)\geq\mathcal I_\text{min}$, then $x_c$ need \emph{never} be considered again.
+Therefore, if candidate location $x_c$ fails to satisfy $\mathcal I(x_c,R_c^*,x_t)+k_\text{dist}\norm{x_t-x_c}^2\geq\mathcal I_\text{min}$, then $x_c$ need \emph{never} be considered again.
 Avoiding these unnecessary calculations yields a greatly lowered computational burden, particularly late in exploration where several regions are collision-free but disadvantageous to revisit.

### 2018-06-26 Dissertation/Thesis/tex/ae3Dsimple.tex
finished editing chap. 4

@@
 \end{align}
-Consider the set $m_{C,k}\subset m$ consisting of the 3D grid cells falling inside the limits of the $k$-th cell of collision map $C$, namely $\mathbf{C}_k$. Then the collision probability is
+Consider the set $ m_{C,k}\subset m$ consisting of the 3D grid cells falling inside the limits of the $k$-th cell of collision map $C$, namely $\mathbf{C}_k$. Then the collision probability is
 \begin{align}
@@
 Let the $E$ be a 2D map composed of cells with edge length $\alpha$ located at the exploration height.
-Let the set $m_{E,k}\subset m$ be the cells directly above and below the $k$-th cell of $E$, neglecting any 3D cells located lower than the floor height.
+Let the set $ m_{E,k}\subset m$ be the cells directly above and below the $k$-th cell of $E$, neglecting any 3D cells located lower than the floor height.
 Selecting this set is advantageous because measurements are assumed not to penetrate occupied cells (edge length $\alpha$) and measurement rays are typically not closely-aligned with the third axis of the inertial frame (vertically long set).
@@
 \end{align}
-After several measurements of the cells belonging to $m_{E,k}$, it is common for $H(P(\mathbf{E}_k))\approx0$, which implies that these cells are well-known.
+After several measurements of the cells belonging to $ m_{E,k}$, it is common for $H(P(\mathbf{E}_k))\approx0$, which implies that these cells are well-known.
 However, this does not imply that the captured space is \emph{entirely} free or occupied (e.g.
@@
 The second proposed technique combines the advantages of both the collision and entropy maps into a 2D projected occupancy grid map, which can be used for both tasks.
-Let the \emph{collision and entropy combination map} $B$ be a 2D projected map located at the exploration height with grid cell edges the same as the collision map, namely $\alpha_C$ from \refeqn{alphaC}, using the same subset of grid cells from the collision map, namely $m_C$.
+Let the \emph{collision and entropy combination map} $ m_\text{2D}$ be a 2D projected map located at the exploration height with grid cell edges the same as the collision map, namely $\alpha_C$ from \refeqn{alphaC}, using the same subset of grid cells from the collision map, namely $ m_C$.
 Considering that the cells begin with uniform probability $P_\text{init}$ where $0<P_\text{init}<1$, define a threshold $P_\text{thresh}$ such that $P_\text{init}<P_\text{thresh}<1$.
-Then, the probability assigned to the $k$-th grid cell of $B$ is
+Then, the probability assigned to the $k$-th grid cell of $ m_\text{2D}$, namely $\mathbf{m}_{\text{2D},k}$ is
 \begin{align}
-P(\mathbf{B}_k)= 
+P(\mathbf{m}_{\text{2D},k})= 
 \begin{cases}
@@
 This approach has several advantages.
-First, measurements must enter the vicinity of the $k$-th cell for $P(\mathbf{B}_k)$ to change from $P_\text{init}$.
-Without prior measurements, capturing $P(\mathbf{B}_k)$ produces a large reward for the expected information gain in \refeqn{ObjFun}.
+First, measurements must enter the vicinity of the $k$-th cell for $P(\mathbf{m}_{\text{2D},k})$ to change from $P_\text{init}$.
+Without prior measurements, capturing $P(\mathbf{m}_{\text{2D},k})$ produces a large reward for the expected information gain in \refeqn{ObjFun}.
 Second, the value of $P_\text{thresh}$ can be modified based on how conservative the collision risk assessment for a robot should be.
-Additionally, when $P(\mathbf{B}_k)<P_\text{thresh}$, this event tends to favor free cells such that expected information gains of cells beyond the $k$-th cell have greater impact on exploration.
+Additionally, when $P(\mathbf{m}_{\text{2D},k})<P_\text{thresh}$, this event tends to favor free cells such that expected information gains of cells beyond the $k$-th cell have greater impact on exploration.
 Furthermore, a single map is simpler, and since cell sizes satisfy $\alpha_C\geq\alpha$, it follows that $n_r$ can only be decreased with $\alpha_C$, thereby increasing computational speed.
@@
 The following numerical example involves a simulated quadrotor taking measurements of a large room and generating a 3D occupancy grid map.
-Both types of map projections are used for autonomous exploration.
+Separate collision and entropy map projections are used for autonomous exploration.
 
@@
 
-Using fixed transforms between the quadrotor body and its onboard sensors, an original mapping node uses a variation of ROS message filters, known as transform message filters, to synchronize the sensor poses with sensor readings.
+Using fixed transformations between the quadrotor body and its onboard sensors, an original mapping node uses a variation of ROS message filters, known as transform message filters, to synchronize the sensor poses with sensor readings.
 These serve as inputs for ray casting in 3D, which provides the cell probabilities and depths along the ray through geometry.
@@
 The second process subscribes to these two 2D maps; the collision map serves as input for candidate pose consideration and Dijkstra's search, and the entropy map serves to predict information gain.
-This node provides path messages, composed of desired poses $0.01$sec apart, displayed with a ROS visualization package, Rviz.
+This node provides path messages, composed of desired poses $0.01$ sec apart, displayed with a ROS visualization package, Rviz.
 
@@
 
-The 3D mapping and exploration algorithms are simulated with several parameters.
-In particular, the Xtion and Hokuyo are modeled using normalized Gaussian and uniform distributions for the sensor model, with a Gaussian probability of $0.9$ with $\sigma_\text{Xtion}=0.25$m and $\sigma_\text{Hokuyo}=0.1$m, respectively, which are greater than their specified values to account for pose uncertainty and distortion with the Xtion.
-Since the Xtion tends to distort depths significantly beyond $4$m, this maximum reading considered for both the Xtion and Hokuyo, and the uniform distribution (probability $0.1$) covers the sensor ranges from $0.5$m to $4$m.
-A cell size of $\alpha=0.075$m is selected and the map limits are $-4$m to $4$m in the x-direction, $-6$m to $6$m in the y-direction, and $-0.15$m to $1.5$m in the z-direction.
+The 3D mapping and exploration algorithms are simulated with several parameters, described next.
+
+\paragraph{Maps}
+
+A cell size of $\alpha=0.075$ m is selected and the map limits are $-4$ m to $4$ m in the x-direction, $-6$ m to $6$ m in the y-direction, and $-0.15$ m to $1.5$ m in the z-direction.
 The initial probability of cells is selected at $P_\text{init}=0.1$.
@@
 Additionally, Dijkstra's algorithm produces a cost map composed of distance costs for each collision map cell, denoted $d_\text{cell}$, from the current robot position to the candidate pose locations over a collision-free path about the occupancy grid.
-A normalized variation on the \emph{bump function}~\cite{Joh06} with maximum distance $d_\text{max}>0$,% TODO: CHECK!
+
+\paragraph{Sensors}
+
+Next we describe the key parameters used for mapping and exploration.
+The inverse sensor model \refeqn{RayISMAnswer}--\refeqn{Unnormalized} relies on the stochastic properties of two sensors, namely an Asus Xtion color IR depth sensor and a Hokuyo LIDAR.
+The Xtion and Hokuyo sensors are modeled using a modified version of the beam model for range finders~\cite{ThrBurFox05}.
+This includes uniform distribution for a phantom reading with probability $P_\text{rand}=0.1$ and a Gaussian distribution with probability $P_\text{hit}=0.9$.
+The Gaussian distribution corresponds to the desired case when the measurement $z$ correctly captures a particular cell at expected distance $\hat{z}$ with standard deviation $\sigma$.
+The total forward sensor model for the Xtion and Hokuyo is
 \begin{align}
-\label{eqn:bumpFun}
-\text{bump}(d_\text{cell})= 
+%p_\text{total}(z,z_\text{min},z_\text{max},\hat{z},\sigma)&=P_\text{rand}p_\text{rand}(z,z_\text{min},z_\text{max})+P_\text{hit}p_\text{hit}(z,\hat{z},\sigma),
+&p_\text{total}=P_\text{rand}p_\text{rand}+P_\text{hit}p_\text{hit},
+\\
+&p_\text{rand}(z,z_\text{min},z_\text{max})=\frac1{z_\text{max}-z_\text{min}},
+\\
+&p_\text{hit}(z,\hat{z},\sigma)=\frac1{\sqrt{2\pi\sigma^2}}\exp\braces{\frac{(\hat{z}-z)^2}{{2\sigma^2}}},
+\end{align}
+such that $z_\text{min}=0.5$ m and $z_\text{max}=4.0$ m for both the Xtion and Hokuyo, and the standard deviations are $\sigma_\text{Xtion}=0.25$ m and $\sigma_\text{Hokuyo}=0.1$ m, respectively, which exceed their specified values to account for localization uncertainty and sensor distortion of the Xtion.
+These sensors update the probabilities of cubic 3D cells with edge length $\alpha=0.075$ m and initial probability $P_\text{init}=0.1$.
+
+\paragraph{Bump Function}
+Dijkstra's algorithm produces a cost map composed of distance costs for each collision map cell, denoted $d_\text{cell}$, from the current robot position to the candidate pose locations over a collision-free path about the occupancy grid.
+This cost map provides a more accurate travel distance cost than the Euclidean distance used in \refeqn{ObjFunApproxCompleteCartesian} because the cost map can account for obstacles within the environment, which is particularly important to properly analyze long trajectories.
+Therefore, we replace $k_\text{dist}\norm{x_t-x_c}^2$ from \refeqn{ObjFunApproxCompleteCartesian} with a normalized variation on the \emph{bump function}~\cite{Joh06} with maximum distance $d_\text{max}>0$,
+\begin{align}
+\label{eqn:BumpFunRef}
+\text{bump}(d)= 
 \begin{cases}
-    e^{1-\frac1{1-(d_\text{cell}/d_\text{max})^2}},			& \text{if }-d_\text{max}<d_\text{cell}<d_\text{max}\\
+    \exp\braces{{1-\frac1{1-(d/d_\text{max})^2}}},			& \text{if }0<d<d_\text{max}\\
     0,              				& \text{otherwise}
@@
 \end{align}
-is multiplied to \refeqn{ObjFunApproxCompleteCartesian} in place of the Euclidean cost term, namely $k_\text{dist}(x_t-x_c)^\T(x_t-x_c)$, which is removed.
-For this experiment, $d_\text{max}$ is chosen as the maximum value of $d_\text{cell}$ on the cost map.
-This serves to prioritize understanding the local surrounding space before the robot moves across the map to distant regions, while still considering all of the pose candidates.
+such that $d_\text{max}$ is selected as the maximum cost map value.
+Then, $\text{bump}(d)$ is multiplied to $\sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m))-\text{E}\left[H(P(m|x_c,z_{c,i}))\right]\bigg)$ from \refeqn{ObjFunApproxCompleteCartesian} when finding optimal poses.
+The bump function prioritizes actions that improve the knowledge of the local surrounding space before distant traversals across the map.
 
-In the simulation, the robot explores the simulated environment for $8$min while taking measurements.
-The 3D occupancy grid map and true environment are shown in Figure \ref{fig:sim3DMap}.
-The Gazebo simulated environment and two 2D projected maps for collision and entropy are shown in Figure \ref{fig:sim2Dmaps}.
-The total map entropy as a function of time is shown in Figure \ref{fig:simH}.
+\paragraph{Simulation}
+
+In the simulation, the robot explores the Gazebo environment for $8$ min while taking measurements.
+The 3D occupancy grid map and true environment are shown in Figure~\ref{fig:sim3DMap}.
+The Gazebo simulated environment and two 2D projected maps for collision and entropy are shown in Figure~\ref{fig:sim2Dmaps}.
+The total map entropy as a function of time is shown in Figure~\ref{fig:simH}.
 
@@
           	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_1min.jpg}
-        		\caption{$t=1$min}
+        		\caption{$t=1$ min}
     	\end{subfigure}
@@
           	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_2min.jpg}
-        		\caption{$t=2$min}
+        		\caption{$t=2$ min}
     	\end{subfigure}
@@
           	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_3min.jpg}
-        		\caption{$t=3$min}
+        		\caption{$t=3$ min}
    	\end{subfigure}
@@
           	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_4min.jpg}
-        		\caption{$t=4$min}
+        		\caption{$t=4$ min}
     	\end{subfigure}
@@
           	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_5min.jpg}
-        		\caption{$t=5$min}
+        		\caption{$t=5$ min}
     	\end{subfigure}
@@
           	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_6min.jpg}
-        		\caption{$t=6$min}
+        		\caption{$t=6$ min}
     	\end{subfigure}
@@
           	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_7min.jpg}
-        		\caption{$t=7$min}
+        		\caption{$t=7$ min}
    	\end{subfigure}
@@
           	\includegraphics[trim = {41.5cm 17cm 14.3cm 3.8cm}, clip, height=\textwidth]{gazebo_8min.jpg}
-        		\caption{$t=8$min}
+        		\caption{$t=8$ min}
     	\end{subfigure}
@@
 % via \DeclareGraphicsExtensions.
-\caption{Simulated Environment and 2D Projected Maps at $4$min}
+\caption{Simulated Environment and 2D Projected Maps at $4$ min}
 \label{fig:sim2Dmaps}
@@
 
-The two-map approach exhibits advantages and disadvantages.
+The two-map approach both exhibits advantages and disadvantages.
 The exploration algorithm successfully avoided collisions and gave way to a fairly complete 3D occupancy grid.
@@
 Conversely, this attention to detail has negative consequences that encourage the robot to repeatedly look at the same regions from different vantage points.
-Since an entropy map cell is composed from many 3D grid cells single-stacked vertically, if just one of these 3D map cells is not captured or hardly captured, the entropy map cell will be recorded as quite uncertain, shown covering numerous locations in Figure \ref{fig:sim2DmapsC}.
+Since an entropy map cell is composed from many 3D grid cells single-stacked vertically, if just one of these 3D map cells is not captured or hardly captured, the entropy map cell will be recorded as quite uncertain, shown covering numerous locations in Figure~\ref{fig:sim2DmapsC}.
 The robot may attempt to measure this space again, even when other regions might more strongly warrant observation.
-These periods of sluggish entropy decrease are shown with flatter sections of Figure \ref{fig:simH}.
+These periods of sluggish entropy decrease are shown with flatter sections of Figure~\ref{fig:simH}.
 In short, this algorithm successfully explores the space collision-free while producing a complete map, but over-attention to just a few uncertain cells has negative consequences on the exploration speed.

### 2018-06-26 Dissertation/Thesis/tex/multivehicle.tex
added text for the first auction

@@
 
+This chapter covers a bidding-based formulation for multiple vehicles cooperating together in the context of autonomous exploration and patrol.
+We focus on a series of auctions to assign tasks, and consider dynamic obstacles, such as people, in large office-like environments using a receding-horizon framework.
+
 \section{Bidding-Based Autonomous Exploration}
 
+When exploring uncertain environments, robots can generate large maps much faster if they coordinate their efforts.
+However, the problem of determining a multi-vehicle exploration strategy is complicated and computationally expensive, but must be performed in real-time.
+
+In this section, we formulate a cooperative and autonomous exploration scheme based on sequential auctions for maximizing map information gain.
+The first auction is based on the expected information gain presented in the prior section, scaled by travel distance.
+Once the winner of the first auction is determined, the map for bidding is updated based on the expected measurements of the first robot.
+The second auction for the remaining robots is determined by the information gain from the revised map that is further scaled by the travel distance and a penalty for collision-avoidance.
+This process is repeated until all robots are assigned tasks.
+
 \subsection{Objective Function for the First Auction}
+
+Here, we describe how robots participate in the first auction for maximizing map information gain while accounting for travel time.
+The information gain \refeqn{ObjFunApproxCompleteCartesian} is computed for each candidate pose.
+Expected information gains may vary among the robots only if they have different sensor configurations.
+All expected information gains must be computed prior to the first auction.
+
+Next, we describe how travel time is integrated into exploration.
+Considering that distances from current robot poses to candidate poses may differ greatly, accounting for these varying travel times properly is essential for exploration time efficiency.
+Travel distances are computed using Dijkstra's search to provide collision-free waypoints for each robot.
+There are two steps: first, generate a cost map from the robot location to each location on the 2D projected map.
+This provides the collision-free distances to all reachable locations on the map.
+Second, the waypoints from a candidate pose to the current robot pose are easily obtained along the cost map using steepest descent.
+
+Here, we define a bump function to account for varying travel distances from Dijkstra's search.
+Let the distance along the collision-free path from the $k$-th robot pose, namely $X_k$, to the $c$-th candidate pose, namely $X_c$, be denoted $d(X_k,X_c,m_\text{2D})\geq0$, which is taken from the cost map belonging to $X_k$.
+A continuous bump function is defined to account for traveling costs, and is composed of two parts.
+
+The first part of the bump function corresponds to short trajectories.
+Consider $d_\text{opt}$, the time-optimal distance that a robot may travel at full speed during the time between exploration updates.
+Here, we consider the case when $d(X_k,X_c,m_\text{2D})\leq d_\text{opt}$, i.e., the robot can completely traverse this distance during the allotted travel time.
+If $d(X_k,X_c,m_\text{2D})\ll d_\text{opt}$, then the robot is not moving at its full potential.
+This can be wasteful, because the robot tends to capture larger regions while moving.
+Therefore, it is desirable for $d(X_k,X_c,m_\text{2D})\rightarrow d_\text{opt}$, which corresponds to maximizing map coverage without time cost.
+The first part of the bump function is sinusoidal and is optimized at $d_\text{opt}$ to maximize robotic movement as
+\begin{align}
+\label{eqn:BumpFunIncreasing}
+\mathcal B_1(d)=\frac12 f_\text{max}\left(1-\cos{\frac{d\pi}{d_\text{opt}}}\right),
+\end{align}
+where $f_\text{max}>0$ is the maximum value when $d(X_k,X_c,m_\text{2D})=d_\text{opt}$.
+
+The second part of the bump function is defined over the domain where \\$d(X_k,X_c,m_\text{2D})>d_\text{opt}$ to minimize time required for $X_k$ to arrive at $X_c$, putting a penalty on traversing across the environment.
+This choice is beneficial to generating accurate local maps before exploring new regions.
+The second half is defined to be nonzero everywhere and is strictly decreasing as
+\begin{align}
+\label{eqn:BumpFunDecreasing}
+\mathcal B_2(d)=(f_\text{max}-f_\text{far})\exp\braces{-\beta(d_\text{opt}-d)^2}+f_\text{far},
+\end{align}
+where $f_\text{max}>f_\text{far}>0$ guarantees $\mathcal B_2>0$ such that $\mathcal B_2\rightarrow f_\text{far}$ as $d(X_k,X_c,m_\text{2D})\rightarrow\infty$ and $\beta>0$ assigns the rate of functional decrease relative to $f_\text{max}$ and $f_\text{far}$.
+Then, the complete bump function is defined as
+\begin{align}
+\label{eqn:BumpFun}
+\mathcal B(d)=
+\begin{cases}
+    \mathcal B_1(d),		& \text{if }d\leq d_\text{opt},\\
+    \mathcal B_2(d),         & \text{otherwise},
+\end{cases}
+\end{align}
+which is illustrated in Fig. \ref{fig:nonzeroBumpFun}. % In short, the bump function prioritizes short movements to capture the local environment over long traversals for exploration time efficiency.
+
+	\begin{figure}
+		\centerline{
+			\includegraphics[width=0.9\columnwidth]{NonZeroBump.pdf}
+		}
+		\caption{The bump function is designed to promote travel distances close to $d_\text{opt}$ by multiplying this function by the expected information gain objective.
+Travel distances less than $d_\text{opt}$ are below the robot capabilities.
+Travel distances beyond $d_\text{opt}$ are beyond what the robot can reach within the minimum exploration computation time.}
+		\label{fig:nonzeroBumpFun}
+	\end{figure}
+
+Then, using the information gain from \refeqn{expectedInfoGainRay} and considering the impact from travel distance from \refeqn{BumpFun}, the objective function for a single agent with respect to $X_{c}$ is
+\begin{align}
+\label{eqn:CandidateBidSingle}
+\text{Obj}_\text{single}(X_k,X_c,m_\text{2D})&=\mathcal B(d(X_k,X_c,m_\text{2D}))\mathcal I(X_{c},m_\text{2D}).
+%X_{k,c}^*&=\argmax_{X_c}{\ \mathcal B_{k,c}\mathcal I(X_c)},
+\end{align}
+Then, the optimal pose selection for the $k$-th agent is
+\begin{align}
+\label{eqn:OptPoseSingle}
+X^*_{k}=\argmax_{X_c}\text{Obj}_\text{single}(X_k,X_c,m_\text{2D}).
+\end{align}
+Therefore, the bump function from \refeqn{BumpFun} prioritizes robotic movement within the local surrounding space before the robot moves across the map to distant regions, while still considering faraway pose candidates.
+
+The first auction includes bids from all robots, and proceeds as follows.
+Let the set of all $n_R$ robots be denoted $\mathcal R=\braces{R_1,R_2,\dots,R_{n_R}}$ such that the $k$-th robot, namely $R_k$, bids $\text{Obj}_\text{single}(X^*_k,X_c,m_\text{2D})$ from \refeqn{CandidateBidSingle} and \refeqn{OptPoseSingle}.
+This is repeated for all $k$ such that $R_k\in\mathcal R$.
+The robot with the largest bid wins the first auction at index $k^*$, and is assigned to travel to $X^*_{k^*}$.
+Then, the remaining robots must account for the planned trajectory of the robot that won the first auction in subsequent auctions, described next.
 

### 2018-06-26 Dissertation/Thesis/tex/multivehicle.tex
updated sec 5.1.1

@@
 
+During this chapter, we assume a 3D vertically-uniform environment at a fixed exploration height as described in Chapter \ref{chap:ae3Dsimple} using the 2D projected combination map $m_\text{2D}$.
+Then information gain is based on \refeqn{ObjFun} and \refeqn{ObjFunApprox}, and depends on this reduced map only, i.e.,
+\begin{align}
+\label{eqn:expectedInfoGainMap2D}
+\mathcal I(X_c,m_\text{2D})&=H(P(m_\text{2D}))-\text{E}\left[H(P(m|X_c,Z_c))\right],
+\end{align}
+where the expected scan $Z$ is given from \refeqn{expectedInfoGainRay} and \refeqn{FindRc}.
+
 \subsection{Objective Function for the First Auction}
@@
 There are two steps: first, generate a cost map from the robot location to each location on the 2D projected map.
-This provides the collision-free distances to all reachable locations on the map.
+This provides the collision-free distances to all reachable locations on the map, which are different for each robot in general.
 Second, the waypoints from a candidate pose to the current robot pose are easily obtained along the cost map using steepest descent.
 
-Here, we define a bump function to account for varying travel distances from Dijkstra's search.
+The travel time is integrated into the autonomous exploration optimization using a bump function similar to \refeqn{BumpFunRef}.
 Let the distance along the collision-free path from the $k$-th robot pose, namely $X_k$, to the $c$-th candidate pose, namely $X_c$, be denoted $d(X_k,X_c,m_\text{2D})\geq0$, which is taken from the cost map belonging to $X_k$.
@@
 
-Then, using the information gain from \refeqn{expectedInfoGainRay} and considering the impact from travel distance from \refeqn{BumpFun}, the objective function for a single agent with respect to $X_{c}$ is
+Then, using the information gain from \refeqn{expectedInfoGainMap2D} and considering the impact from travel distance from \refeqn{BumpFun}, the objective function for a single agent with respect to $X_{c}$ is
 \begin{align}
@@
 \end{align}
-Therefore, the bump function from \refeqn{BumpFun} prioritizes robotic movement within the local surrounding space before the robot moves across the map to distant regions, while still considering faraway pose candidates.
+Like \refeqn{BumpFunRef}, the bump function from \refeqn{BumpFun} prioritizes robotic movement within the local surrounding space before the robot moves across the map to distant regions, while still considering faraway pose candidates.
+Unlike \refeqn{BumpFunRef}, \refeqn{BumpFun} discourages the robot from staying motionless or near-motionless.
 

### 2018-06-26 Dissertation/Thesis/tex/multivehicle.tex
added 5.1.2: needs editing

@@
 
+The robots unable to win the first auction are updated to reflect the expected impact of $R_{k^*}$.
+This robot is expected to modify the probabilistic occupancy grid map, which may significantly change the expected information gains and collision properties of candidate poses nearby $X^*_{k^*}$.
+The process of auctioning and modifying bids is repeated, removing the winning robot from consideration after each auction, until no robots remain.
+Between auctions, updating the map with expected measurements from auction-winning robots discourages the remaining robots from entering the same regions and capturing the same cells.
+As such, there is no need to consider the coverage overlap explicitly: it is avoided in a systematic way as increased coverage overlap would reduce the overall information gain.
+
+More specifically, we coordinate robotic efforts by modifying bids to prevent robots from updating the same grid cells and to avoid collisions among robots.
+The goal is to find the winning bids $\mathcal X^*=\braces{X^*_{1},X^*_{2},\dots,X^*_{n_R}}$ corresponding to each robot from $\mathcal R$.
+During the auctioning process, let $\mathcal W\subset\mathcal R$ be the set of robots that have already won auctions by producing the largest objective function bid.
+After the first auction described above, $\mathcal W=\braces{R_{k^*}}$; after all auctions are complete, $\mathcal W=\mathcal R$.
+Between auctions, the candidate poses are modified for information gain and collision-avoidance, coordinating the multi-vehicle exploration.
+
+The first step between auctions is to modify the expected map information gain for efficient map coverage.
+Once a robot wins a bid, the measurement ray expected values serve to update a temporary copy of $m_\text{2D}$, namely $m_\text{2D,copy}$, and those candidates located in a local neighborhood (twice the radius of a maximum sensor reading) of the winning candidate from the prior auction are recalculated, thereby coordinating total map information gain with auction-winning robots.
+Using the same local map notation from \refeqn{allEta} and \refeqn{Unnormalized} along a measurement ray, the expected measurement value is
+\begin{align}
+\label{eqn:ExpectedMeasRay}
+\text{E}[z]=\sum_{k=1}^{n_{r}}\bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{m}}_j)\bigg\}P(\mathbf{m}_k)z_k,
+\end{align}
+where $z_k$ denotes the distance from the robot sensor to the $k$-th cell along the measurement ray.
+Then $\text{E}[z]$ is substituted into \refeqn{RayISMAnswer}--\refeqn{Unnormalized} to modify $m_\text{2D,copy}$, and expected map information gains are recomputed with \refeqn{ProbMeas} and \refeqn{expectedInfoGainRay}, while the bump function \refeqn{BumpFun} remains the same.
+In short, the expected changes of $m_\text{2D}$ are integrated into $m_\text{2D,copy}$ for expected information gain estimation between multiple vehicles.
+
+The second step focusses on collision-avoidance between robots based on their proximity.
+Let $\rho_\text{max}>0$ be a fixed maximum radius to consider collision-avoidance and $\rho_{i,c}\geq0$ be the Euclidean distance from the $i$-th already-assigned pose $X^*_i$ to the $c$-th candidate pose $X_c$, i.e.,
+\begin{align}
+\rho_{i,c}&=\norm{x^*_i-x_c},
+\end{align}
+where $x^*_i$ and $x_c$ are the pose locations of $X^*_i$ and $X_c$, respectively. Then the collision-avoidance factor for the $k$-th robot such that $R_k\notin\mathcal W$ is,
+\begin{align}
+\label{eqn:CollisionAvoidanceAmongRobots}
+\mathbf C(X_c)&=
+\begin{cases}
+    \prod_{i|\mathcal{R}_i\in\mathcal W} \left(\frac{\rho_{i,c}}{\rho_\text{max}}\right)^2,		& \text{if }\rho_{i,c}<\rho_\text{max},\\
+    1,              				& \text{otherwise},
+\end{cases}
+\end{align}
+where this product serves to decrease the value of bids for candidate poses in close proximity with already-assigned poses to avoid collisions between robots.
+The multi-vehicle objective function for the $k$-th robot accounting for collision-avoidance is
+\begin{align}
+\label{eqn:CandidateBidMulti}
+\text{Obj}_\text{multi}(X_k,X_c,m_\text{2D,copy})
+=\mathbf C(X_c)\mathcal B(d(X_k,X_c,m_\text{2D}))\mathcal I(X_{c},m_\text{2D,copy}),
+\end{align}
+where its optimal pose during this auction is
+\begin{align}
+\label{eqn:OptPoseMulti}
+X^*_{k}=\argmax_{X_c}\text{Obj}_\text{multi}&(X_k,X_c,m_\text{2D,copy}),
+\end{align}
+where $\text{Obj}_\text{multi}(X^*_k,X_c,m_\text{2D,copy})$ is the bid for the $k$-th robot.
+Optimal pose selection and bidding is repeated for all robots not belonging to $\mathcal W$.
+The largest bid among these wins the auction and is tasked with moving to the associated pose candidate, and then this robot is included with $\mathcal W$ to avoid further consideration.
+Completing $n_R$ auctions produces $\mathcal X^*$.
+
+The proposed approach uses auctions and bid modifications based on total map expected information gain and collision-avoidance, promoting exploration of different spaces without explicit consideration of coverage overlaps.
+The pseudocode for the bidding process is shown with Algorithm \ref{alg:bidding}.
+
+\begin{algorithm}
+	Function: $Bidding(\mathcal X_r,\mathcal I(X_c)\forall c\in\mathcal C,\rho_\text{max})$\;
+	Initialize $k=0$, $\mathcal W=0_{n_R\times1}$, $B=0_{n_R\times1}$\;
+	Update candidate expected information gains using \refeqn{DiscExpEntropyRay}\;
+	\For{$i=n_R,n_R-1,\ldots,1$}{
+		\For{$j=1,2,\ldots,n_R$}{
+			\If{$\mathcal W(j)==0$}{
+				\If{$i==n_R$}{
+					Maximize bid \refeqn{CandidateBidSingle} for $X^*_{j}$ from \refeqn{OptPoseSingle}\;
+				}
+				\Else{
+					Update $\mathcal I(X_c)$ close to $\mathcal X^*(k)$\;
+					\For{$c=1,2,\ldots,n_c$}{
+						Find Euclidean distance $\rho_{k,c}$\;
+					Get $\mathbf C(X_c)$ with \refeqn{CollisionAvoidanceAmongRobots}\;
+					}
+					Maximize bid \refeqn{CandidateBidMulti} for $X^*_{j}$ from \refeqn{OptPoseMulti}\;
+				}
+				Insert the $j$-th maximum bid into $B(j)$\;
+			}
+		}
+		$k$: index maximizing $B$ with corresponding $X^*_{k}$\;
+		$\mathcal X^*(k)=X^*_{k}$\;
+		$\mathcal W(k)=1$\;
+		$B(k)=0$\;
+	}
+
+	Return $\mathcal X^*$\;
+\caption{Robot Task Bidding}
+\label{alg:bidding}
+\end{algorithm}
+
 \subsection{Receding Horizon Framework}

### 2018-06-26 Dissertation/Thesis/tex/multivehicle.tex
edited 5.1.2 and inserted (unedited) 5.1.3

@@
 
-The robots unable to win the first auction are updated to reflect the expected impact of $R_{k^*}$.
-This robot is expected to modify the probabilistic occupancy grid map, which may significantly change the expected information gains and collision properties of candidate poses nearby $X^*_{k^*}$.
+The robots unable to win the first auction are updated to reflect the expected impact from the first robot traveling to optimal pose $X^*_{k^*}$.
+This robot is expected to modify the probabilistic occupancy grid map, which may significantly change the expected information gains and collision properties of candidate poses nearby.
 The process of auctioning and modifying bids is repeated, removing the winning robot from consideration after each auction, until no robots remain.
 Between auctions, updating the map with expected measurements from auction-winning robots discourages the remaining robots from entering the same regions and capturing the same cells.
-As such, there is no need to consider the coverage overlap explicitly: it is avoided in a systematic way as increased coverage overlap would reduce the overall information gain.
+As such, there is no need to consider the coverage overlap explicitly as in~\cite{SimApfBurFoxMooThrYou00}; instead coverage overlap is avoided in a systematic way as increased coverage overlap would reduce the overall information gain.
 
@@
 where $z_k$ denotes the distance from the robot sensor to the $k$-th cell along the measurement ray.
-Then $\text{E}[z]$ is substituted into \refeqn{RayISMAnswer}--\refeqn{Unnormalized} to modify $m_\text{2D,copy}$, and expected map information gains are recomputed with \refeqn{ProbMeas} and \refeqn{expectedInfoGainRay}, while the bump function \refeqn{BumpFun} remains the same.
+Then $\text{E}[z]$ is substituted into \refeqn{RayISMAnswer}--\refeqn{allEta} to modify $m_\text{2D,copy}$, and expected map information gains are recomputed with \refeqn{ProbMeas} and \refeqn{expectedInfoGainRay}, while the bump function \refeqn{BumpFun} remains the same.
 In short, the expected changes of $m_\text{2D}$ are integrated into $m_\text{2D,copy}$ for expected information gain estimation between multiple vehicles.
+
+% TODO: make sure find/replace of:
+% \refeqn{RayISMAnswer}--\refeqn{Unnormalized}
+%  \refeqn{RayISMAnswer}--\refeqn{allEta}
 
@@
 \end{align}
-where $\text{Obj}_\text{multi}(X^*_k,X_c,m_\text{2D,copy})$ is the bid for the $k$-th robot.
+such that $\text{Obj}_\text{multi}(X^*_k,X_c,m_\text{2D,copy})$ is the bid for the $k$-th robot.
 Optimal pose selection and bidding is repeated for all robots not belonging to $\mathcal W$.
 The largest bid among these wins the auction and is tasked with moving to the associated pose candidate, and then this robot is included with $\mathcal W$ to avoid further consideration.
-Completing $n_R$ auctions produces $\mathcal X^*$.
+Completing $n_R$ auctions produces the set of coordinated poses $\mathcal X^*$.
 
@@
 
+This algorithm is further aided by following a receding horizon framework for improved information gain maximizations and collision-avoidance with dynamic obstacles and other robots.
+A receding horizon simply repeats the autonomous exploration steps as quickly as possible over a finite time period, frequently before a robot reaches its desired pose.
+Since exploration optimizations can only occur during exploration updates, a receding horizon maximizes the rate at which these updates occur.
+These updates require time for computation, over which the robot can move $d_\text{opt}$, used to optimize the robot travel distances with \refeqn{BumpFun}.
+Hence, the receding horizon framework serves to repeat optimizations as quickly as possible with maximal robotic movement while exploring a changing occupancy grid.
+
+Furthermore, a receding horizon framework enhances autonomous exploration in two other ways.
+First, the occupancy grid map is constantly updated while a robot traverses a trajectory, so the expected information gains change as well.
+Since the bids depend on the occupancy grid, $\mathcal X^*$ is updated accordingly.
+This prevents robots from completing trajectories that have become unnecessary as new terrain becomes captured.
+Second, the receding horizon prevents robots from colliding with each other when they become too close together (e.g., crossing paths, traversing a tight passage) or other moving obstacles (e.g., a human).
+The cost map for each robot is updated, and the location of other robots are considered as collision-prone space.
+Hence, the robot trajectories become better separated due to rapid updates of the cost maps with the receding horizon.
+
 \subsection{Multi-Vehicle Exploration Numerical Simulation}

### 2018-06-26 Dissertation/Thesis/tex/multivehicle.tex
updated & edited up to 5.2.2 (no patrol simulation yet)

@@
 \section{Bidding-Based Autonomous Exploration}
+\label{sec:BiddingExploration}
 
@@
 
-During this chapter, we assume a 3D vertically-uniform environment at a fixed exploration height as described in Chapter \ref{chap:ae3Dsimple} using the 2D projected combination map $m_\text{2D}$.
+During this chapter, we assume a 3D vertically-uniform environment at a fixed exploration height as described in Chapter \ref{chap:ae3Dsimple} using the 2D projected combination map $ m_\text{2D}$.
 Then information gain is based on \refeqn{ObjFun} and \refeqn{ObjFunApprox}, and depends on this reduced map only, i.e.,
@@
 	\begin{figure}
-		\centerline{
+		\centering{
 			\includegraphics[width=0.9\columnwidth]{NonZeroBump.pdf}
@@
 The first step between auctions is to modify the expected map information gain for efficient map coverage.
-Once a robot wins a bid, the measurement ray expected values serve to update a temporary copy of $m_\text{2D}$, namely $m_\text{2D,copy}$, and those candidates located in a local neighborhood (twice the radius of a maximum sensor reading) of the winning candidate from the prior auction are recalculated, thereby coordinating total map information gain with auction-winning robots.
+Once a robot wins a bid, the measurement ray expected values serve to update a temporary copy of $ m_\text{2D}$, namely $ m_\text{2D,copy}$, and those candidates located in a local neighborhood (twice the radius of a maximum sensor reading) of the winning candidate from the prior auction are recalculated, thereby coordinating total map information gain with auction-winning robots.
 Using the same local map notation from \refeqn{allEta} and \refeqn{Unnormalized} along a measurement ray, the expected measurement value is
@@
 where $z_k$ denotes the distance from the robot sensor to the $k$-th cell along the measurement ray.
-Then $\text{E}[z]$ is substituted into \refeqn{RayISMAnswer}--\refeqn{allEta} to modify $m_\text{2D,copy}$, and expected map information gains are recomputed with \refeqn{ProbMeas} and \refeqn{expectedInfoGainRay}, while the bump function \refeqn{BumpFun} remains the same.
-In short, the expected changes of $m_\text{2D}$ are integrated into $m_\text{2D,copy}$ for expected information gain estimation between multiple vehicles.
+Then $\text{E}[z]$ is substituted into \refeqn{RayISMAnswer}--\refeqn{allEta} to modify $ m_\text{2D,copy}$, and expected map information gains are recomputed with \refeqn{ProbMeas} and \refeqn{expectedInfoGainRay}, while the bump function \refeqn{BumpFun} remains the same.
+In short, the expected changes of $ m_\text{2D}$ are integrated into $ m_\text{2D,copy}$ for expected information gain estimation between multiple vehicles.
 
@@
 
+The algorithms for mapping, exploration, and visualization are designed for the ROS framework.
+This framework allows the node that computes 3D mapping probabilities to easily transfer these to an exploration node and a visualization node.
+However, unlike the simulation in Section \ref{sec:Compare2MapProjections}, scalability to larger environment is necessary.
+For this task, the exploration node is separated into three threads running in parallel.
+The first thread constantly updates the 2D projected map according to \refeqn{CombinationProjection2DMap}.
+The second thread computes the information gains, runs auctions, and generates trajectories.
+The third thread broadcasts a message for visualizing the 2D projected map.
+This structure is chosen because the 2D map cannot afford to miss messages, and the last two may produce time bottlenecks and are independent of each other.
+The final node is for 3D map visualization, and it creates a message to the ROS visualization package, Rviz, to produce 3D cells with varying opacities corresponding to cell occupancy probabilities.
+
+The multi-vehicle exploration problem is typically applied to large environments, so several additional scalability precautions are taken to avoid large communications and computations.
+Most importantly, only the map \emph{changes} are communicated from the mapping node to the exploration and visualization nodes using a custom ROS message type.
+This change drastically decreases the time needed to generate or interpret the message.
+This is accomplished by finding a rectangular prism of the minimum and maximum limits of cells that may have changed during a particular measurement scan update.
+Publishing and subscribing to map changes provides an efficient means to transfer mapping information, independent of map size.
+
+Another computational improvement key to autonomous exploration success is that only important and changing map information gains are updated.
+The expected information gain for a single candidate is updated ray-by-ray, where only the top $\hat{n}=5$ cells with the largest detection probabilities are considered using \refeqn{ProbOfFirstCell}.
+Since \refeqn{ProbMeas} is embedded in \refeqn{DiscExpEntropyRay}, the computational complexity for a ray with $n_{r}$ cells is $\mathcal O(n_{r}^2)$.
+Reducing the considered cells to $\hat{n}$ limits the computation to $\mathcal O(n_{p}^2)$ where $n_{p}\ll n_r$ in general.
+Furthermore, finding these $\hat{n}$ cells is computationally inexpensive; using a selection algorithm of the $n_{r}$ cells and a sorting algorithm of the $n_{p}$ cells, the computational complexity is amortized to $\mathcal O(n_r+\hat{n}\log(\hat{n}))$, which can be further approximated to $\mathcal O(n_r)$ assuming that $\hat{n}\ll n_r$.
+Thus, this entropy updating scheme scales well to varying sensor ranges and cell sizes.
+Furthermore, the information gains of poses far from any robot trajectory need not be updated between receding horizons.
+
+\paragraph{Parameters and Resulting Maps}
+The 3D mapping and exploration algorithms for three robots are simulated with several parameters, which are the same as those in Section \ref{sec:Compare2MapProjections} except the map limits are extended to $-55.0$ m to $55.0$ m in the x- and y-directions, and $0.0$ m to $1.5$ m in the z-direction, producing a total of $45,255,504$ grid cells with edge length $0.075$ m (dimensioned $1468\times1468\times21$).
+For the bump function, $\mathcal B_\text{max}=1$, $\mathcal B_\text{far}=0.1$, and $\beta=0.01$ are chosen.
+For bidding $\rho_\text{max}=3.0$ m is chosen.
+
+In the simulation, three quadrotors explore the simulated environment of The George Washington University's (GWU) Science and Engineering Hall (SEH) for $10$ minutes while taking measurements.
+Each quadrotor is given an Asus Xtion depth sensor and Hokuyo LIDAR, where the sensor properties for the Asus Xtion are used with exploration expectations because they provide the vast majority of measurements.
+The building floor plan and Gazebo simulated environment are shown in Fig.
+\ref{fig:SEHEnvironment}.
+The 3D occupancy grid maps are shown in Fig.
+\ref{fig:sim3DMap}, and the corresponding 2D projected maps for exploration are shown in Fig.
+\ref{fig:sim2Dmaps} using the ROS visualization package Rviz.
+The full video of both maps is available at \href{https://youtu.be/OMn2c453oik}{\WriteBlue{https://youtu.be/OMn2c453oik}}.
+
+\begin{figure}[!t]
+\centering
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{SEH_cropped.png}
+        		\caption{Floor Plan}
+    	\end{subfigure}
+	\hspace*{0.1\textwidth}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{gazeboSEH.png}
+        		\caption{Gazebo Model}
+    	\end{subfigure}
+	\caption{The SEH second floor plan (about $15,000$ sqft) is reconstructed with a Gazebo simulation environment.}
+	\label{fig:SEHEnvironment}
+\end{figure}
+
+%	% trim={<left> <lower> <right> <upper>}
+
+\begin{figure}[!t]
+	\centering{
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_1min_ss.jpg}
+        		\caption{$1$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_2min_ss.jpg}
+        		\caption{$2$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_3min_ss.jpg}
+        		\caption{$3$ min}
+    	\end{subfigure}
+	}
+	\centering{
+	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_4min_ss.jpg}
+        		\caption{$4$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_5min_ss.jpg}
+        		\caption{$5$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_6min_ss.jpg}
+        		\caption{$6$ min}
+    	\end{subfigure}
+	}
+	\centering{
+   	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_7min_ss.jpg}
+        		\caption{$7$ min}
+    	\end{subfigure}
+ 	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_8min_ss.jpg}
+        		\caption{$8$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_9min_ss.jpg}
+        		\caption{$9$ min}
+    	\end{subfigure}
+	}
+	\caption{3D Occupancy Grid Map of SEH Second Floor}
+	\label{fig:sim3DMap}
+\end{figure}
+
+\begin{figure}[!t]
+	\centering{
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, width=\columnwidth]{multi_discount_raw_2D_1min.jpg}
+        		\caption{$1$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, width=\columnwidth]{multi_discount_raw_2D_2min.jpg}
+        		\caption{$2$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, width=\columnwidth]{multi_discount_raw_2D_3min.jpg}
+        		\caption{$3$ min}
+    	\end{subfigure}
+	}
+	\centering{
+	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, width=\columnwidth]{multi_discount_raw_2D_4min.jpg}
+        		\caption{$4$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, width=\columnwidth]{multi_discount_raw_2D_5min.jpg}
+        		\caption{$5$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, width=\columnwidth]{multi_discount_raw_2D_6min.jpg}
+        		\caption{$6$ min}
+    	\end{subfigure}
+	}
+	\centering{
+   	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, width=\columnwidth]{multi_discount_raw_2D_7min.jpg}
+        		\caption{$7$ min}
+    	\end{subfigure}
+ 	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, width=\columnwidth]{multi_discount_raw_2D_8min.jpg}
+        		\caption{$8$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, width=\columnwidth]{multi_discount_raw_2D_9min.jpg}
+        		\caption{$9$ min}
+    	\end{subfigure}
+	}
+	\caption{2D Projected Occupancy Grid Map of SEH Second Floor}
+	\label{fig:sim2Dmaps}
+\end{figure}
+
+%	
+%	\begin{figure}
+%	% trim={<left> <lower> <right> <upper>}
+%		\centering{
+%			\subfigure[$0$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_0min.jpg}}
+%			\hspace*{0.025\columnwidth}
+%			\subfigure[$1$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_1min.jpg}}
+%		}
+%		\vspace*{0.025\columnwidth}
+%		\centering{
+%			\subfigure[$2$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_2min.jpg}}
+%			\hspace*{0.025\columnwidth}
+%			\subfigure[$3$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_3min.jpg}}
+%		}
+%		\vspace*{0.025\columnwidth}
+%		\centering{
+%			\subfigure[$4$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_4min.jpg}}
+%			\hspace*{0.025\columnwidth}
+%			\subfigure[$5$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_5min.jpg}}
+%		}
+%		\vspace*{0.025\columnwidth}
+%		\centering{
+%			\subfigure[$6$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_6min.jpg}}
+%			\hspace*{0.025\columnwidth}
+%			\subfigure[$7$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_7min.jpg}}
+%		}
+%		\vspace*{0.025\columnwidth}
+%		\centering{
+%			\subfigure[$8$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_8min.jpg}}
+%			\hspace*{0.025\columnwidth}
+%			\subfigure[$9$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_9min.jpg}}
+%		}
+%		\vspace*{0.025\columnwidth}
+%		\caption{2D Projected Occupancy Grid Map of SEH Second Floor}
+%			\label{fig:sim2Dmaps}
+%	\end{figure}
+
+The resulting 3D occupancy grid shows a nearly-completed map after three quadrotors cooperatively explored over $10$ minutes, which is quantified with total map entropy from \refeqn{ShannonsEntropyMap}, shown in Fig.
+\ref{fig:simH}.
+An important observation is that the majority of map space was unreachable by any robot, so about $1.2\times10^7$ of the entropy metric could not be improved, regardless of mapping or exploration strategy.
+
+Special attention had to be placed on parameter selection for the bump function and collision-avoidance.
+For the bump function, increasing the parameter $\beta$ narrows the bump, placing a stronger influence on nearby candidates.
+However, this choice corresponds to the bump function decreasing at a negligible rate beyond a close neighborhood of the robot, i.e., a large range of distances yield $\mathcal B\approx\mathcal B_\text{far}$.
+Hence, there is a tradeoff between prioritization of local movements and being able to differentiate the effects of distance on large trajectories.
+
+When robots competed for future poses in a centralized auction framework, the repeated optimizations provided an effective multi-vehicle exploration policy.
+Between auctions, only a small set of candidates within the neighborhood of the last auction-winning candidate require consideration when updating expected information gains with \refeqn{ExpectedMeasRay} and avoiding collisions between robots using \refeqn{CollisionAvoidanceAmongRobots}, both of which modify the optimization of \refeqn{OptPoseMulti}.
+Since bidding is applied efficiently, its computation time is negligible compared with other parts of the exploration update such as computing the initial candidate information gains and robot cost maps.
+
+	\begin{figure}
+		\centering{
+			\includegraphics[width=0.9\columnwidth]{multi_H.pdf}
+		}
+		\caption{Total 3D Map Cell Entropy}
+		\label{fig:simH}
+	\end{figure}
+
 \section{Multi-Vehicle Cooperative Patrol}
 
+Patrol is the continuous and repetitive observation of different regions within a larger environment.
+An effective patrol algorithm is particularly valuable for surveillance operations, especially when the robot members are incapable of measuring all regions of an environment from fixed poses.
+Here, we formulate cooperative patrol as an optimization problem, where expected map information gains are maximized cooperatively according to the methods outlined in Section \ref{sec:BiddingExploration}.
+However, the grid cells experience slow degradation over time, thereby modifying the expected map information gains that govern exploration.
+This way, different regions are revisited after some time passes.
+This method of cell degradation for cooperative robotic patrol is described next.
+
 \subsection{Continuous-Time Markov Process for Cell Degradation}
-% JINT18 fig. 3
+
+The key idea to cell degradation is that as time progresses forward, cells become more uncertain, which causes the exploration algorithm to revisit these cells.
+All cells are degraded through a continuous-time Markov process at an equal rate, thereby creating a fair policy for revisiting regions that exploits the benefits of multi-vehicle exploration, such as following a receding horizon and optimizing information gain and travel time.
+Cell degradation is applied simply and uniformly to the entire 3D map for autonomous cooperative patrol.
+Since this analysis is valid for all grid cells, cell index notation is neglected for this analysis.
+
+Consider that a single grid cell begins with probability $P(\mathbf{m}_{t_0})$ at time $t_0$ and degrades to $P(\mathbf{m}_{t_f})$ at time $t_f>t_0$.
+Suppose that as time progresses forward after $t_0$, the cell probability approaches a terminal value $P_\infty\in(0,1)$, i.e.,
+\begin{align}
+\label{eqn:DegradationQualification}
+\lim_{t\rightarrow\infty}\begin{bmatrix}
+P(\mathbf{m}_{t})
+\\
+P(\bar{\mathbf{m}}_{t})
+\end{bmatrix}
+=
+\begin{bmatrix}
+P_\infty
+\\
+1-P_\infty
+\end{bmatrix}.
+\end{align}
+Let the degradation rate be $\lambda>0$ and matrix $A\in\Re^{2\times2}$ be defined as
+\begin{align}
+A=\lambda
+\begin{bmatrix}
+-(1-P_\infty) & P_\infty
+\\
+(1-P_\infty) & -P_\infty
+\end{bmatrix}.
+\end{align}
+Then the first-order ordinary differential equation is expressed as
+\begin{align}
+\begin{bmatrix}
+\dot{P}(\mathbf{m}_{t})
+\\
+\dot{P}(\bar{\mathbf{m}}_{t})
+\end{bmatrix}
+=
+A
+\begin{bmatrix}
+P(\mathbf{m}_{t})
+\\
+P(\bar{\mathbf{m}}_{t})
+\end{bmatrix}.
+\end{align}
+Hence, the state transition from $t_0$ to $t_f$ is
+\begin{align}
+\label{eqn:DegradeStateTransition}
+\begin{bmatrix}
+P(\mathbf{m}_{t_f})
+\\
+P(\bar{\mathbf{m}}_{t_f})
+\end{bmatrix}
+&=
+\exp\braces{A(t_f-t_0)}
+\begin{bmatrix}
+P(\mathbf{m}_{t_0})
+\\
+P(\bar{\mathbf{m}}_{t_0})
+\end{bmatrix},
+\end{align}
+which is illustrated with Fig.
+\ref{fig:MarkovDegradeContinuous}.
+Most importantly, \refeqn{DegradeStateTransition} satisfies \refeqn{DegradationQualification} as $t_f\rightarrow\infty$.
+Solving the top row of \refeqn{DegradeStateTransition} yields
+\begin{align}
+\label{eqn:CellDegradationScalar}
+P(\mathbf{m}_{t_f})=P(\mathbf{m}_{t_0})\exp\braces{-\lambda (t_f-t_0)}+P_\infty(1-\exp\braces{-\lambda (t_f-t_0)}).
+\end{align}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{markov_diagram_continuous.pdf}
+\caption{The continuous-time Markov process for a single binary cell occupancy is illustrated. As time progresses, this process degrades the cell slowly toward $P_\infty.$}
+\label{fig:MarkovDegradeContinuous}
+\end{figure}
+
+Applying degradation is computationally-inexpensive due to the simplicity of \refeqn{CellDegradationScalar}.
+For an arbitrary positive fixed time step, $\exp\braces{-\lambda (t_f-t_0)}$ needs only be calculated once and is multiplied to every map cell.
+Similarly, the second term of \refeqn{CellDegradationScalar}, namely $P_\infty(1-\exp\braces{-\lambda (t_f-t_0)})$, is simply added to every cell.
+For grid cells with probabilities far from $P_\infty$, the times to converge to this final degraded value increase, shown in Fig.
+\ref{fig:DegradeExamples}.
+Slow convergence of cells with probabilities close to $0$ or $1$ allows the robots to stop viewing these cells, giving the robots opportunities to visit different regions before returning.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{DegradeExamples.pdf}
+\caption{The rate at which probabilities converge toward $P_\infty=0.4$ increases based on the difference magnitude of their initial values.}
+\label{fig:DegradeExamples}
+\end{figure}
+
+In short, each grid cell is continuously degrading according to \refeqn{CellDegradationScalar} toward $P_\infty$.
+Only in the presence of measurements does the cell occupancy become well-known according to \refeqn{RayISMAnswer}--\refeqn{Unnormalized}.
+Cell degradation encourages the robot to revisit cells from the distant past using bidding-based multi-vehicle autonomous exploration to patrol the environment continuously.
+Adopting the receding horizon framework allows patrol to account for dynamic obstacles by quickly updating patrol strategies as map information changes.
 

### 2018-06-26 Dissertation/Thesis/tex/multivehicle.tex
finished chap 5, but I need to generate figs @ 2 min for patrol (3x3 grid of subfigs needs 1 more)

@@
 \subsection{Multi-Vehicle Exploration Numerical Simulation}
+\label{sec:MultivehicleExplorationSEH}
 % Multi-vehicle ICRA figs 3-6 (remake fig. 5 with larger contrast)
@@
 
-%	
-%	\begin{figure}
-%	% trim={<left> <lower> <right> <upper>}
-%		\centering{
-%			\subfigure[$0$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_0min.jpg}}
-%			\hspace*{0.025\columnwidth}
-%			\subfigure[$1$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_1min.jpg}}
-%		}
-%		\vspace*{0.025\columnwidth}
-%		\centering{
-%			\subfigure[$2$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_2min.jpg}}
-%			\hspace*{0.025\columnwidth}
-%			\subfigure[$3$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_3min.jpg}}
-%		}
-%		\vspace*{0.025\columnwidth}
-%		\centering{
-%			\subfigure[$4$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_4min.jpg}}
-%			\hspace*{0.025\columnwidth}
-%			\subfigure[$5$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_5min.jpg}}
-%		}
-%		\vspace*{0.025\columnwidth}
-%		\centering{
-%			\subfigure[$6$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_6min.jpg}}
-%			\hspace*{0.025\columnwidth}
-%			\subfigure[$7$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_7min.jpg}}
-%		}
-%		\vspace*{0.025\columnwidth}
-%		\centering{
-%			\subfigure[$8$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_8min.jpg}}
-%			\hspace*{0.025\columnwidth}
-%			\subfigure[$9$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_9min.jpg}}
-%		}
-%		\vspace*{0.025\columnwidth}
-%		\caption{2D Projected Occupancy Grid Map of SEH Second Floor}
-%			\label{fig:sim2Dmaps}
-%	\end{figure}
-
 The resulting 3D occupancy grid shows a nearly-completed map after three quadrotors cooperatively explored over $10$ minutes, which is quantified with total map entropy from \refeqn{ShannonsEntropyMap}, shown in Fig.
@@
 \subsection{Multi-Vehicle Patrol Numerical Simulation}
+\label{sec:MultivehiclePatrolSEH}
 % JINT18 fig. 10-13 (ref. ICRA fig 3 from above in this chapter)
 
+In this section, we present a multi-vehicle exploration simulation that is identical to the scenario in Section \ref{sec:MultivehicleExplorationSEH}, except for two key differences.
+The first is that patrol is included in the exploration scheme to incentivize the robots to revisit areas after time passes.
+The second is that the robots must handle a human is walking around the space, unknown to the cooperating robots, requiring the receding horizon framework to avoid collisions with this dynamic obstacle.
+
+\paragraph{Degradation Thread}
+An additional thread within the mapping node handles cell degradation with $\lambda=1\times10^{-3}$, which corresponds to a cell half life of roughly $11.5$ minutes; if this time passes since a measurement captures a cell, the occupancy probability will be halfway between its value $11.5$ minutes ago and $P_\infty=P_\text{init}=0.1$.
+While degrading cells with \refeqn{CellDegradationScalar} is inexpensive, degradation must cycle through all cells, which can be costly with a map containing over $45$ million cells.
+Since all cells are affected by degradation, simply sending the map \emph{changes} from measurement scan updates alone is insufficient for the exploration node to track these changing probabilities.
+Thus, the complete map is sent from this mapping thread, and an additional thread in the exploration node subscribes to this message to update the projected map $m_\text{2D}$ slowly.
+The sluggish speed of this thread is acceptable, as cell degradation is a slow process, so fast updates are uncritical.
+Both the mapping and exploration nodes require an additional thread for degradation.
+
+\paragraph{Simulated Results}
+
+In the simulation, three quadrotors explore and patrol the same simulated environment from Section \ref{sec:MultivehicleExplorationSEH} of The George Washington University's (GWU) Science and Engineering Hall (SEH) for $15$ minutes while taking measurements.
+Initially, the robots are given no information about the environment except that their immediate surroundings are free.
+Unknown to the robot team, a simulated human walks in a rectangular repeated pattern at $0.5$ m/s.
+The 3D occupancy grid maps are shown in Fig.
+\ref{fig:Sim3DMapPatrol}, and the corresponding 2D projected maps for exploration are shown in Fig.
+\ref{fig:Sim2DMapPatrol}.
+The full video of both maps is available at \href{https://youtu.be/bsLG2romP_8}{\WriteBlue{https://youtu.be/bsLG2romP\_8}}.
+
+\begin{figure}[!t]
+	\centering{
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_1min.jpg}
+        		\caption{$1$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_1min.jpg}
+        		\caption{$2$ min}% TODO: generate this figure!
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_3min.jpg}
+        		\caption{$3$ min}
+    	\end{subfigure}
+	}
+	\centering{
+	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_5min.jpg}
+        		\caption{$5$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_7min.jpg}
+        		\caption{$7$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_9min.jpg}
+        		\caption{$9$ min}
+    	\end{subfigure}
+	}
+	\centering{
+   	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_11min.jpg}
+        		\caption{$11$ min}
+    	\end{subfigure}
+ 	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_13min.jpg}
+        		\caption{$13$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_15min.jpg}
+        		\caption{$15$ min}
+    	\end{subfigure}
+	}
+	\caption{The 3D occupancy grid map representation of the SEH second floor is shown during patrol.
+The robots generate these maps while avoiding collisions with each other, the environment, and a non-cooperative human.}
+	\label{fig:Sim3DMapPatrol}
+\end{figure}
+
+\begin{figure}[!t]
+	\centering{
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_1min.jpg}
+        		\caption{$1$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_1min.jpg}
+        		\caption{$2$ min}% TODO: generate this figure!
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_3min.jpg}
+        		\caption{$3$ min}
+    	\end{subfigure}
+	}
+	\centering{
+	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_5min.jpg}
+        		\caption{$5$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_7min.jpg}
+        		\caption{$7$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_9min.jpg}
+        		\caption{$9$ min}
+    	\end{subfigure}
+	}
+	\centering{
+   	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_11min.jpg}
+        		\caption{$11$ min}
+    	\end{subfigure}
+ 	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_13min.jpg}
+        		\caption{$13$ min}
+    	\end{subfigure}
+	\hspace*{0.05\textwidth}
+    	\begin{subfigure}[t]{0.25\columnwidth}
+          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_15min.jpg}
+        		\caption{$15$ min}
+    	\end{subfigure}
+	}
+	\caption{The 2D projected map from of the SEH second floor shows degradation while robots patrol the space.
+The robots evade the non-cooperative human walking in a rectangular pattern, occasionally capturing the temporarily-occupied space and identifying these cells as occupied on the grid.}
+	\label{fig:Sim2DMapPatrol}
+\end{figure}
+
+The robot generated a fairly-complete occupancy grid within $9$ minutes, and proceeded to patrol the space during the remaining time.
+After $9$ minutes, the total map entropy no longer increases or decreases significantly; the rate of cell degradation is roughly equal and opposite to improvements to the probabilistic map from the robot team.
+In contrast, if the robots simply hover instead of patrol after $9$ minutes, the map uncertainty quickly increases, shown in Fig.
+\ref{fig:DegradeExamples}.
+Furthermore, the receding horizon framework minimizes time between exploration updates, which allows the robots quickly react to dynamic obstacles like the non-cooperative human, shown in Fig.
+\ref{fig:EvadeHuman}.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{entropy_patrol_9minSwitch.pdf}
+\caption{Total map entropies are tracked for separate trials: patrol of the building for $15$ minutes (blue) and patrol of the building for $9$ minutes, then hover (red).
+Even though the map is fairly well-known after $9$ minutes in either case, the continuous effort of patrolling over the complete time period shows lower map uncertainty beyond this time as cells are constantly degrading.}
+\label{fig:DegradeExamples}
+\end{figure}
+
+\begin{figure}[!t]
+\centering
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim={22cm 25cm 20cm 0cm}, clip, width=\textwidth]{evade_15sec.jpg}
+        		\caption{$15$ sec}
+    	\end{subfigure}
+	\hspace*{0.02\textwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim={22cm 25cm 20cm 0cm}, clip, width=\textwidth]{evade_16sec.jpg}
+        		\caption{$16$ sec}
+    	\end{subfigure}
+	\hspace*{0.02\textwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim={22cm 25cm 20cm 0cm}, clip, width=\textwidth]{evade_17sec.jpg}
+        		\caption{$17$ sec}
+    	\end{subfigure}
+	\caption{A cooperating robot must evade a non-cooperative human.
+The receding horizon framework allows a quickly-updating map to modify the exploration commands to avoid collisions with a dynamic obstacle.}
+	\label{fig:EvadeHuman}
+\end{figure}
+
+The results also demonstrate an interesting tradeoff between exploring new terrain and patrolling previously-visited spaces.
+When the degradation rate $\lambda$ is increased, grid cells are degrading faster, incentivizing the exploration strategy to revisit spaces faster.
+This behavior is beneficial for frequent patrol, but certain areas, particularly those difficult to reach, are not visited as quickly or as frequently.
+Should an area receive extra attention, a greater value of $\lambda$ may be applied to this region.
+
 \section{Conclusions}
+
+This chapter introduced a bidding-based framework for cooperative autonomous exploration and patrol.
+These contributions exploit the values of the objective functions used in autonomous exploration optimizations for a bids over a series of auctions.
+Between auctions, a copy of the map and collision criteria are updated, thereby coordinating the exploration efforts.
+Finally, a patrol scheme is developed based on simple cell degradation, promoting periodic visitations to the same regions over time.

### 2018-06-27 Dissertation/Thesis/tex/ae3Dcomplex.tex
added and edited Secion 6.1

@@
 
-\section{Exploration with Full 3D Maps}
+This chapter extends the autonomous exploration scheme to complex 3D environments.
+Unlike Chapter \ref{chap:ae2D} where exploration was applied in 2D, or Chapters \ref{chap:ae3Dsimple} and \ref{chap:multivehicle} where exploration was confined to a single height in vertically-uniform environments, this chapter proposes a completely 3D exploration approach for complex environments.
+This proposed approach is designed for rooms with complicated features and places with changing and uncertain terrain such as Mars.
 
-\subsection{Collision Mapping in 3D}
+\section{Exploration in 3D}
 
-\subsection{Future Pose Optimization in 3D}
+In this section, we propose how multiple measurement rays in 3D can determine map information gain from a pose candidate.
+Then, we show how a collision-free trajectory is determined with a 3D Dijkstra's search.
+Finally, we combine 3D information gain and travel costs into a single optimization.
+
+\subsection{Map Information Gain in 3D}
+
+The expected entropy for an arbitrary 1D ray spanning 3D space can be calculated with \refeqn{DiscExpEntropyRay} and \refeqn{ProbMeas} in a computationally tractable manner.
+However, computing the expected entropy from multiple rays simultaneously has exponential complexity, and is therefore computationally intractable.
+Additionally, numerous rays of a single scan commonly intersect the same grid cells, making consideration of \emph{every} measurement ray unnecessary.
+
+Instead of considering expected entropy from a complete scan, we propose a real-time solution that selects sample measurements to determine an optimal attitude at each candidate pose and the associated map information gain, similar to how the scan information gain is found in Section \ref{sec:OptimalPose2DMap}.
+We assume the vehicle is capable of level flight, so the third axes of the world and body are aligned (Fig.
+\ref{fig:transforms}), so $R_c$ can be expressed as
+\begin{align*}
+R_c=\begin{bmatrix}
+\cos\psi & -\sin\psi & 0
+\\
+\sin\psi & \cos\psi & 0
+\\
+0 & 0 & 1
+\end{bmatrix}, \quad 0\leq\psi<2\pi,
+\end{align*}
+where $\psi$ represents the angle about the third body-fixed axis.
+The direction of a 3D measurement may also have a nonzero component in the vertical direction.
+This is achieved by rotating the sensor frame angle $\theta$ about the second sensor-fixed axis,
+\begin{align*}
+R_{c,\text{sensor}}=\begin{bmatrix}
+\cos\theta & 0 & \sin\theta
+\\
+0 & 1 & 0
+\\
+-\sin\theta & 0 & \cos\theta
+\end{bmatrix}, \quad -\frac{\pi}{2}\leq\theta\leq\frac{\pi}{2}.
+\end{align*}
+Combining these two rotation matrices, the measurement $z$ with depth $\norm{z}$ is expressed with respect to a frame fixed to the world as
+\begin{align}
+z(\psi,\theta)=\begin{bmatrix}\cos\theta\cos\psi & -\sin\psi & \cos\psi\sin\theta
+\\
+\cos\theta\sin\psi & \cos\psi & \sin\psi\sin\theta
+\\
+-\sin\theta & 0 & \cos\theta
+\end{bmatrix}
+\begin{bmatrix}
+\norm{z} \\ 0 \\ 0
+\end{bmatrix}
+=
+\begin{bmatrix}
+\cos\theta\cos\psi
+\\
+\cos\theta\sin\psi
+\\
+-\sin\theta
+\end{bmatrix}\norm{z}.
+\end{align}
+In short, unit vectors are acquired via Euler angles within certain sensor limits.
+
+% trim={<left> <lower> <right> <upper>}
+
+\begin{figure}[!t]
+	\centering
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim = {21cm 5cm 20cm 5cm}, clip, height=1.0\textwidth]{mars_general.png}
+        		\caption{General View}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim = {25cm 6cm 16cm 4cm}, clip, height=1.0\textwidth]{mars_psi.png}
+        		\caption{$\psi$ Rotation}
+    	\end{subfigure}
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim = {23cm 4cm 17cm 6cm}, clip, height=1.0\textwidth]{mars_theta.png}
+        		\caption{$\theta$ Rotation}
+    	\end{subfigure}
+%\includegraphics[width=2.5in]{myfigure}
+% where an .eps filename suffix will be assumed under latex, 
+% and a .pdf suffix will be assumed for pdflatex; or what has been declared
+% via \DeclareGraphicsExtensions.
+\caption{The three frames are shown with three axes (red: first axis, green: second axis, blue: third axis).
+The Mars frame (short and thick) is fixed to the planet.
+The body-fixed frame (medium thickness and length) is fixed to the camera frame (long and narrow).
+The rotation of $\psi$ (b) represents the fixed camera yaw rotation and $\theta$ (c) represents the downward angle of the camera to capture the surface of Mars below the flying robot.}
+\label{fig:transforms}
+\end{figure}
+
+Next, we show how multiple 3D measurements provide an estimate for information gain.
+Let $\Psi$ and $\Theta$ be the angular ranges for the sensor FOV in the horizontal and vertical directions, respectively.
+Within the sensor FOV, we select sample measurements for evaluating their expected information gains.
+Let the number of sample measurements be $n_\psi$ and $n_\theta$ for horizontal and vertical rotations, respectively.
+Since $R_c$ and the FOV are known, the set of sample measurement rays, namely $\mathcal{Z}(R_c)$, is known as well.
+Then, the expected information gain for a single measurement ray $z$ from candidate position $x_c$ and attitude $R_c$ is the negative expected entropy change,
+\begin{align}
+\label{eqn:Iray}
+\mathcal{I}&_\text{ray}(x_c,R_c,\psi,\theta)\nonumber\\&=
+\begin{cases}
+H(P(r))-\text{E}[H(P(r|x_c,z(\psi+\tilde{\psi},\theta+\tilde{\theta})))], & \mbox{if } z(\psi+\tilde{\psi},\theta+\tilde{\theta})\in\mathcal{Z}(R_c), \\ 
+0,                                                                                    & \mbox{otherwise},
+\end{cases}
+\end{align}
+where the mounting yaw $\tilde{\psi}$ and pitch $\tilde{\theta}$ of the sensor is a fixed rotation in the horizontal and vertical directions, respectively in that order, such that $-\pi<\tilde{\psi}\leq\pi$ and $-\frac{\pi}{2}+\frac\Theta2\leq\tilde{\theta}\leq\frac{\pi}{2}-\frac\Theta2$.
+A positive $\tilde{\theta}$ corresponds to rotating the depth sensor from a forward direction to a downward direction toward the ground below.
+The expected information gain for the full scan is the summation of expected information gains for individual rays
+\begin{align}
+\label{eqn:Iscan}
+\mathcal{I}_\text{scan}(x_c,R_c)=\sum_{i_\psi=1}^{n_\psi}\sum_{i_\theta=1}^{n_\theta}\mathcal{I}_\text{ray}\left(x_c,R_c,\frac{(2i_\psi-n_\psi-1)\Psi}{2n_\psi},\frac{(2i_\theta-n_\theta-1)\Theta}{2n_\theta}\right).
+\end{align}
+
+Finally, \refeqn{Iscan} is computed for $n_\psi$ possible attitudes when the robot is at location $x_c$, one corresponding to each yaw angle $\psi$. Then, the optimal attitude at $x_c$ is
+\begin{align}
+\label{eqn:RcStar}
+R_c^*=\argmax_{R_c} \mathcal{I}_\text{scan}(x_c,R_c).
+\end{align}
+
+In short, the expected information gain from each possible measurement ray is obtained using predicted entropy from \refeqn{DiscExpEntropyRay} and \refeqn{ProbMeas}.
+Then the expected information gain from all possible scans at a candidate location is calculated from \refeqn{Iray} and \refeqn{Iscan}.
+Finally, the optimal attitude at this candidate location is found using \refeqn{RcStar}.
+
+\subsection{Collision-Free Trajectory in 3D}
+
+Next, we present a fairly straightforward application of Dijkstra's search to 3D occupancy grid mapping.
+First we reduced the number of grid cells to cubic blocks large enough for a robot to fit completely inside.
+Based on the occupancy probability of these blocks and their neighbors, we determine which cells are reachable, and what are their travel costs by building a cost map based on Dijkstra's algorithm.
+
+Here we describe how a reduced map is generated for collision avoidance and motion planning.
+If the largest edge-to-edge distance of the robot, namely $\rho$, is not larger than the 3D grid cell size $\alpha$, this step is unnecessary.
+Let $k\geq1$ be an integer such that $k\alpha\geq\rho$ and $(k-1)\alpha<\rho$, i.e., $k$ is the minimum number of grid cells in each dimension capable of fully-enclosing the robot inside this cube.
+Then, we decompose the complete probabilistic 3D into map $ m_\text{reduced}$, which is composed of larger cubic cells that encompass $k^3$ cells from map $ m$.
+Let $\mathbf{m}_{\text{reduced},k}$ denote the $k$-th cell of $ m_\text{reduced}$ being occupied, and let $a_k\subset m$ be those $k^3$ cells from $ m$ composing this larger cell.
+Then, we apply the same criteria as \refeqn{CombinationProjection2DMap} to obtain the probability of $\mathbf{m}_{\text{reduced},k}$ in 3D,
+
+\begin{align}
+\label{eqn:Proj3DMapComb}
+P(\mathbf{m}_{\text{reduced},k})= 
+\begin{cases}
+    \max_{i\in a_k}{P(\mathbf{m}_i)},	&\text{if} \ \max_{i\in a_k}{P(\mathbf{m}_i)}\geq P_\text{thresh},\\
+    \min_{i\in a_k}{P(\mathbf{m}_i)},	& \text{otherwise},
+\end{cases}
+\end{align}
+where initial probability is $0<P_\text{init}<1$ and the threshold probability is constrained to $P_\text{init}<P_\text{thresh}<1$.
+This approach uses probabilities that have changed from $P_\text{init}$, and favors occupied cells to avoid risking collisions.
+
+Next, we show how $\mathbf{m}_{\text{reduced},k}$ is used in Dijkstra's search for collision avoidance and motion planning.
+Defining $0<P_\text{coll}<1$ as the acceptable probability of collision, every grid cell of $ m_\text{reduced}$ is considered a safe and unvisited node if its occupancy probability, and the occupancy probabilities of its neighbors (sharing a face, edge, or corner), are below $P_\text{coll}$.
+A cost map is built from the starting robot location by neighboring nodes, where the cost to travel to another node is based on Euclidian distance: a cell sharing a face is $\alpha$ away, a cell sharing an edge is $\sqrt2\alpha$ away, and a cell sharing a corner is $\sqrt3\alpha$ away.
+Generating a cost map for all reachable cells in $\mathbf{m}_{\text{reduced},k}$ provides a collision-free travel cost for all reachable candidate pose locations.
+Once an optimal pose is selected, the path is easily found from the cost map using steepest descent.
+
+\subsection{Optimal 3D Pose}
+
+Here we present how an optimal pose is selected based on information gains at optimal attitudes and known collision-free travel costs.
+This process is similar to prior the autonomous exploration proposed in earlier chapters, with a few minor differences.
+
+The proposed 3D autonomous exploration follows a receding-horizon framework, and uses a bump function \refeqn{BumpFunIncreasing}--\refeqn{BumpFun} to account for travel distance.
+The only difference is that the start of the bump function $B_1(d)=f_\text{max}$, illustrated in Figure \ref{fig:recedingHorizonBumpFun}).
+This change, though minor, avoids the possible scenario where a robot repeatedly alternates between neighboring candidates with similar expected information gains.
+Then, we incorporate expected information gain and travel costs into a unified optimization. Given $d(x_c)$ from Dijkstra's cost map, the optimal candidate location is
+\begin{align}
+\label{eqn:CandidateLocationOptimization}
+x_c^*=\argmax_{x_c}\mathcal{I}_\text{scan}(x_c,R_c^*)\mathcal B(d(x_c)),
+\end{align}
+and therefore the optimal pose is $X_c^*=\braces{x_c^*,R_c^*}$.
+
+	\begin{figure}
+	\vspace*{-0.3\columnwidth}
+		\centerline{
+			\includegraphics[width=0.6\columnwidth]{recedingHorizonBumpFunFlattened.pdf}
+		}
+	\vspace*{-0.25\columnwidth}
+		\caption{The bump function is multiplied to the objective function of \refeqn{CandidateLocationOptimization} to prioritize local trajectories to avoid traversals across the map, where travel costs are determined using Dijkstra's algorithm.
+However, there is no time wasted up to $d_\text{opt}$ because this distance may be achieved in the minimum exploration computation time.}
+		\label{fig:recedingHorizonBumpFun}
+	\end{figure}
+
+In conclusion, every reachable candidate pose is considered for its optimal attitude, which is found using sample rays spanning 3D space.
+The travel costs, acquired from Dijkstra's search, serve as input for the bump function, which prioritizes local trajectories to avoid costly map traversals.
+Then, the optimal candidate location is selected to maximize a combination of expected information gain and travel cost.
 

### 2018-06-27 Dissertation/Thesis/tex/ae3Dcomplex.tex
added and partly edited Section 6.2 (Mars simulation)

@@
 \section{Numerical Simulation}
-% Mars simulation
+In this section, we demonstrate the efficacy of the proposed 3D mapping and autonomous exploration with a simulated Mars environment.
+First we cover the parameters, then compare results using two different maps for entropy prediction.
 
-\section{Experiment}
-% SEH indoor simulation (needs completing)
+\subsection{Mars Parameters}
+
+A 3D environment of Mars is simulated in Gazebo.
+A height map~\cite{mapaplanet} is used to generate a contoured surface, and the corresponding picture of Mars is draped over this contour, shown in Fig.
+\ref{fig:MarsGazebo}.
+A 3D laser scanner is also simulated in Gazebo.
+In the horizontal direction, the sensor has limits $\Psi=120^{\circ}$ with a total of $1000$ measurement rays inside, the sensor is fixed at angle $\tilde{\psi}=45^{\circ}$, and $n_\psi=16$ sample measurements are used.
+In the vertical direction, the sensor also has limits $\Theta=120^{\circ}$ with a total of $1000$ measurement rays inside, but the sensor is fixed at angle $\tilde{\theta}=30^{\circ}$, and $n_\psi=7$ sample measurements are used.
+These ray samples are taken from each candidate location, which are separated $1$ m apart in each of the $3$ dimensions.
+
+% trim={<left> <lower> <right> <upper>}
+
+\begin{figure}[!t]
+	\centering
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[height=0.9\textwidth]{mars_heightmap.png}
+        		\caption{Height Map}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[height=0.9\textwidth]{mars_colormap.png}
+        		\caption{Color Map}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering{
+          	\includegraphics[height=0.9\textwidth]{mars_gazebo_full_lighting.png}}
+        		\caption{Gazebo Environment}
+    	\end{subfigure}
+\caption{The height map (a) and color map (b) are combined in the Gazebo simulator (c) to model a Mars environment. The robot captures the simulated environment to generate an occupancy grid map.}
+\label{fig:MarsGazebo}
+\end{figure}
+
+The map parameters are also important to the success of the exploration.
+The full map has dimensions $20$ m $\times$ $20$ m $\times$ $5$ m in the Mars-fixed frame, with cell edge length $\alpha=0.075$ m.
+The reduced map for collision-avoidance and motion planning has cells $k=3$ times the size ($0.225$ m), so $k^3=27$ cells are considered in \refeqn{Proj3DMapComb}.
+The bump functions use $f_\text{max}=1.0$, $f_\text{far}=0.1$, and $\beta=0.1$ to account for travel costs with \refeqn{BumpFun}.
+The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.25$ m/sec and the computation time varies from $1.8$ sec to $2.5$ sec.
+
+\subsection{Mars Results}
+
+The simulation was run twice.
+Case 1 was as described in this paper.
+Case 2 was identical to Case 1, except the reduced map $ m_\text{reduced}$ is used when computing \refeqn{Iscan}.
+Case 2 is largely inspired by the promising experimental results covered in Section \ref{sec:QuadrotorNRL}, where a projected map based on the same criteria as \refeqn{Proj3DMapComb} proved effective for level flight.
+The resulting occupancy grid maps for Cases 1 and 2 are shown in Figs.
+\ref{fig:mars3DogmCase1} and \ref{fig:mars3DogmCase2}, respectively.
+A video of Case 1 can be found at \href{https://youtu.be/FrBcL2UMW9w}{\WriteBlue{https://youtu.be/FrBcL2UMW9w}} and close-up pictures from Case 2 are shown in Fig.
+\ref{fig:marsZoomedIn}.
+The complete map entropy is illustrated in Fig.
+\ref{fig:mars3Dentropy}.
+
+\begin{figure}[!t]
+	\centering
+	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap30sec.jpg}
+        		\caption{$30$ sec}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap1min.jpg}
+        		\caption{$1$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+	\centering
+	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap2min.jpg}
+        		\caption{$2$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap3min.jpg}
+        		\caption{$3$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+	\centering
+	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap4min.jpg}
+        		\caption{$4$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap5min.jpg}
+        		\caption{$5$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+	\centering
+	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap10min.jpg}
+        		\caption{$10$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{FullMarsMap15min.jpg}
+        		\caption{$15$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+\caption{In Case 1, the robot (red disk with arrow indicating laser direction) moves toward candidate poses (red arrows, more opaque for greater reward) based on expected entropy change of the 3D probabilistic occupancy grid map $m$ (cubes: greater opacity for greater occupancy probability) of the surface of Mars.}
+\label{fig:mars3DogmCase1}
+\end{figure}
+
+\begin{figure}[!t]
+	\centering
+	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap30sec.jpg}
+        		\caption{$30$ sec}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap1min.jpg}
+        		\caption{$1$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+	\centering
+	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap2min.jpg}
+        		\caption{$2$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap3min.jpg}
+        		\caption{$3$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+	\centering
+	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap4min.jpg}
+        		\caption{$4$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap5min.jpg}
+        		\caption{$5$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+	\centering
+	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap10min.jpg}
+        		\caption{$10$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.49\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{RdcdMarsMap15min.jpg}
+        		\caption{$15$ min}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+\caption{In Case 2, the robot (red disk with arrow indicating laser direction) generates a 3D probabilistic occupancy grid map $m$ (cubes: greater opacity for greater occupancy probability), which is used to generate a reduced map $m_\text{reduced}$ based on \refeqn{Proj3DMapComb}.
+The robot moves toward candidate poses (red arrows, more opaque for greater reward) based on expected entropy change of $m_\text{reduced}$.
+Using $m_\text{reduced}$ instead of $m$ for entropy calculation leads to faster exploration of new terrain, but this leaves some grid cells missing.}
+\label{fig:mars3DogmCase2}
+\end{figure}
+
+\begin{figure}[!t]
+	\centering
+	\begin{subfigure}[t]{0.95\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{mars_closeup.png}
+        		\caption{Full Map Cells and Sensor Scan}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+	\centering
+	\begin{subfigure}[t]{0.95\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{scitech_reducedCells.png}
+        		\caption{Full Map and Reduced Map Cells}
+		\vspace*{0.025\textwidth}
+    	\end{subfigure}
+\caption{The close-up images from the Case 2 trial show the candidate future poses (red arrows), where greater opacity represents a larger objective function of \refeqn{CandidateLocationOptimization}.
+In (a), we show the 3D scan with color corresponding to the surface of Mars.
+In (b), we overlay the full map $m$ (colored) with the reduced map $m_\text{reduced}$ (gray) from \refeqn{Proj3DMapComb}.}
+\label{fig:marsZoomedIn}
+\end{figure}
+
+	\begin{figure}
+		\centerline{
+			\includegraphics[width=0.6\columnwidth]{scitech_entropy_comparison_flat.pdf}
+		}
+		\caption{The complete map entropy for Case 1 (red solid line) decreases at roughly the same rate as the reduced map entropy for Case 2 (dashed blue line) toward the beginning.
+However, the expected entropy in Case 2 promotes actions toward unexplored territory, while the policy of Case 1 yields a more complete map before moving on to unexplored spaces.}
+		\label{fig:mars3Dentropy}
+	\end{figure}
+
+In both cases, the robot built the 3D probabilistic map of the $2000$ cubic meter space composed of $4.74\times10^6$ grid cells.
+The maps were mostly complete within $10$ minutes.
+The $\tilde{\theta}=30^{\circ}$ downward viewing angle captured the ground nicely, as candidate attitudes tended to direct the onboard sensor toward uncertain regions on the surface of Mars.
+The proposed approach provided collision-free mapping and autonomous exploration in real-time.
+
+However, the choice of occupancy grid map used for entropy predictions introduced an interesting tradeoff.
+In Case 1, when the full probabilistic map $m$ was used in \refeqn{Iscan}, regions where the grid cells were partially-known were frequently reconsidered until the space was well-known.
+Conversely in Case 2, when $m_\text{reduced}$ was used instead, the robot repeatedly left spaces that were missing a few grid cells to visit new terrain.
+This is because $m_\text{reduced}$ contained some cells with large occupancy probabilities based on \refeqn{Proj3DMapComb}, which allowed some grid cells of $m$ enclosed within a cell of $m_\text{reduced}$ to be uncertain.
+The exploration policy of Case 2 incorrectly assumed these regions were well-known.
+Ironically, this false assumption actually led to greater information gains when the vehicle moved on to unvisited terrain.
+Conversely with Case $1$, which used $m$ for computing expected entropy, the total map entropy decreased more steadily and generated a more-complete 3D occupancy grid of the surface of Mars.
+
+In short, the proposed 3D probabilistic occupancy grid mapping and autonomous exploration were simulated successfully in real-time over the surface of Mars using ROS and Gazebo.
+Choosing $m$ for entropy predictions produced a more-complete map, but choosing $m_\text{reduced}$ was sometimes beneficial for exploring new terrain faster, while forgoing map completeness.
 

### 2018-06-27 Dissertation/Thesis/tex/ae3Dcomplex.tex
finished writing and editing chap 6

@@
 Instead of considering expected entropy from a complete scan, we propose a real-time solution that selects sample measurements to determine an optimal attitude at each candidate pose and the associated map information gain, similar to how the scan information gain is found in Section \ref{sec:OptimalPose2DMap}.
-We assume the vehicle is capable of level flight, so the third axes of the world and body are aligned (Fig.
-\ref{fig:transforms}), so $R_c$ can be expressed as
+We assume the vehicle is capable of level flight, so the third axes of the world and body are aligned (Figure \ref{fig:transforms}), so $R_c$ can be expressed as
 \begin{align*}
@@
 A 3D environment of Mars is simulated in Gazebo.
-A height map~\cite{mapaplanet} is used to generate a contoured surface, and the corresponding picture of Mars is draped over this contour, shown in Fig.
-\ref{fig:MarsGazebo}.
+A height map~\cite{mapaplanet} is used to generate a contoured surface, and the corresponding picture of Mars is draped over this contour, shown in Figure \ref{fig:MarsGazebo}.
 A 3D laser scanner is also simulated in Gazebo.
@@
 Case 2 is largely inspired by the promising experimental results covered in Section \ref{sec:QuadrotorNRL}, where a projected map based on the same criteria as \refeqn{Proj3DMapComb} proved effective for level flight.
-The resulting occupancy grid maps for Cases 1 and 2 are shown in Figs.
-\ref{fig:mars3DogmCase1} and \ref{fig:mars3DogmCase2}, respectively.
-A video of Case 1 can be found at \href{https://youtu.be/FrBcL2UMW9w}{\WriteBlue{https://youtu.be/FrBcL2UMW9w}} and close-up pictures from Case 2 are shown in Fig.
-\ref{fig:marsZoomedIn}.
-The complete map entropy is illustrated in Fig.
-\ref{fig:mars3Dentropy}.
+The resulting occupancy grid maps for Cases 1 and 2 are shown in Figures \ref{fig:mars3DogmCase1} and \ref{fig:mars3DogmCase2}, respectively.
+A video of Case 1 can be found at \href{https://youtu.be/FrBcL2UMW9w}{\WriteBlue{https://youtu.be/FrBcL2UMW9w}}, and close-up pictures from Case 2 are shown in Figure \ref{fig:marsZoomedIn}.
+The complete map entropies for both cases are illustrated in Figure \ref{fig:mars3Dentropy}.
 
@@
            	\centering
-          	\includegraphics[width=\textwidth]{mars_closeup.png}
+          	\includegraphics[width=0.9\textwidth]{mars_closeup.png}
         		\caption{Full Map Cells and Sensor Scan}
@@
            	\centering
-          	\includegraphics[width=\textwidth]{scitech_reducedCells.png}
+          	\includegraphics[width=0.9\textwidth]{scitech_reducedCells.png}
         		\caption{Full Map and Reduced Map Cells}
@@
 \section{Conclusions}
+
+This chapter covers autonomous exploration through a 3D environment, such that the robot can account for complex geometries in all dimension and explore with upward and downward motions.
+The expected information gains are calculated with predicting the entropy changes from sample rays in 3D space, and the bump function is applied using the cost from Dijkstra's search.
+An example where a robot maps and explores the surface of Mars demonstrates the efficacy of the approach.

### 2018-06-27 Dissertation/Thesis/tex/results.tex
added and partly edited Section 6.2 (Mars simulation)
added & edited sec. 7.1 (Pioneer ground experiment)

@@
 
-\section{A section}
+In this chapter, we present three experiments that demonstrate the efficacy of the proposed exact occupancy grid mapping and autonomous exploration from earlier chapters.
+The first example involves a ground vehicle mapping and exploring a 2D space.
+The second example is a quadrotor unmanned aerial vehicle (UAV) flying around a vertically-uniform space while generating a 3D map from level flight.
+The final example shows autonomous exploration completely in 3D, where a quadrotor must map and explore around complicated objects.
 
-\subsection{A subsection}
+\section{Pioneer Ground Vehicle Tests}
+\label{sec:PioneerNRL}
+
+The experiment involves a ground vehicle producing a probabilistic occupancy grid map in real time.
+The robot motion is governed by the information gain-maximizing policy described in Chapter \ref{chap:ae2D}.
+
+\paragraph{Hardware Configuration}
+The Pioneer 3 ground robot is chosen for this experiment because of its reliability and simple operation.
+The vehicle accepts two inputs, namely linear velocity (aligned with the robot wheels) and angular velocity (about a central axis passing vertically through the robot).
+The robot pose is estimated via Vicon Tracker, which provides the location and attitude of a rigid body.
+The depth readings are obtained by a Kinect depth sensor.
+Since the experiment is to map and explore a 2D environment, only a single central row of the 3D Kinect depth scan provide the necessary measurements.
+The robot captures an experimental environment with walls constructed from Styrofoam.
+
+\paragraph{Software Configuration}
+
+The software structure is similar to ROS nodes Section \ref{sec:Ae2DNumExamples}, which simply involves a mapping node, an exploration node, and a visualization node using ``OpenGL'' libraries.
+The only differences are eliminating the simulator (Stage) and adding synchronization among the sensors, which is of high importance to the proposed occupancy grid mapping approach.
+Any time delay between the robot pose estimation and the sensor depth readings can cause conflicting information, which may be harmful to the probabilistic map.
+Thus, ROS approximate message filters are applied such that the pose from Vicon Tracker ($100$Hz) and the Microsoft Kinect ($30$Hz) are nearly synchronized, which provides validity of the assumption that the measurement ray positions and directions are known deterministically.
+
+\paragraph{Exploring and Mapping a 2D Environment}
+The environment consists of Styrofoam walls on a flat floor.
+The perimeter is rectangular, with several obstacles and angled walls (see Figure \ref{fig:ExpSetupPhoto}).
+The robot autonomously explored the space using the complete Cartesian searching approach, and Dijkstra's algorithm provides collision-free waypoints to the optimal future poses.
+Based on these waypoints, a constrained least squares polynomial fitting generates a smooth trajectory to follow the waypoints with fixed velocity.
+Finally, a simple controller is implemented such that the Pioneer follows the trajectory by moving and orienting toward the desired robot location $1$ sec in the future.
+
+\begin{figure}
+	\centering
+    	\begin{subfigure}[b]{0.45\textwidth}
+        		\includegraphics[width=\textwidth]{test_setup_bottom_left_cropped.jpg}
+        		\caption{Bottom-left view}
+        		\label{fig:Experiment_blv}
+    	\end{subfigure}
+	\hspace*{0.05\columnwidth}
+	\begin{subfigure}[b]{0.45\textwidth}
+        		\includegraphics[width=\textwidth]{test_setup_bottom_right_cropped.jpg}
+        		\caption{Bottom-right view}
+        		\label{fig:Experiment_brv}
+    	\end{subfigure}
+\caption{Images from two perspectives show the walls and obstacles of the experimental environment.}
+\label{fig:ExpSetupPhoto}
+\end{figure}
+
+The resulting occupancy grid maps and trajectories are depicted in Figure \ref{fig:ExperimentOGM} and a video is available at \href{https://www.youtube.com/watch?v=CRQfhhICSj0&feature=youtu.be}{\WriteBlue{www.youtube.com/watch?v=CRQfhhICSj0\&feature=youtu.be}}.
+Most importantly, the robot builds a clear occupancy grid map despite sensor noise and imperfect sensor synchronization.
+The autonomous exploration successfully guides the robot such that it builds a map of the reachable space without colliding with any obstacles.
+
+\begin{figure}
+	\centering{
+    	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t0sec.jpg}
+        		\caption{$t=0$ sec}
+        		\label{fig:Experiment_ogm_t0}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t20sec.jpg}
+        		\caption{$t=20$ sec}
+        		\label{fig:Experiment_ogm_t0p5}
+    	\end{subfigure}    
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t40sec.jpg}
+        		\caption{$t=40$ sec}
+        		\label{fig:Experiment_ogm_t1}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t60sec.jpg}
+        		\caption{$t=60$ sec}
+        		\label{fig:Experiment_ogm_t1p5}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t80sec.jpg}
+        		\caption{$t=80$ sec}
+        		\label{fig:Experiment_ogm_t2}
+    	\end{subfigure}
+		\vspace*{0.05\textwidth}
+
+	}
+	\centering{
+    	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t100sec.jpg}
+        		\caption{$t=100$ sec}
+        		\label{fig:Experiment_ogm_t2p5}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t120sec.jpg}
+        		\caption{$t=120$ sec}
+        		\label{fig:Experiment_ogm_t3}
+    	\end{subfigure}    
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t140sec.jpg}
+        		\caption{$t=140$ sec}
+        		\label{fig:Experiment_ogm_t3p5}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t160sec.jpg}
+        		\caption{$t=160$ sec}
+        		\label{fig:Experiment_ogm_t4}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{feb23_t180sec.jpg}
+        		\caption{$t=180$ sec}
+        		\label{fig:Experiment_ogm_t4p5}
+    	\end{subfigure}
+	}
+\caption{A robot autonomously explores the experimental environment and produces an occupancy grid map along the way.
+The robot (shaded green circle body with shaded blue circle sensor) is controlled toward the desired pose $1$ sec in the future (unshaded grey robot), which is following a constrained least-squares trajectory (cyan curve).
+This trajectory is based on waypoints (black circles) from Dijkstra's algorithm to arrive at the optimal pose (unshaded green circle body with unshaded blue circle sensor) as proposed in this paper.}
+\label{fig:ExperimentOGM}
+\end{figure}
+
+The resulting map and exploration commands are generated in real time.
+The grid cells are stored as double-float variables in a vector, where the vector index is mapped to a location on the occupancy grid.
+Thus, the memory requirements are proportional to the number of grid cells, which is roughly the mapping area divided by the area of a grid cell.
+Since cell locations need not be saved because the mapping between cell index and location is known, memory requirements are reduced.
+On a Lenovo T540p laptop with an Intel Core i7-4900MQ processor (quad-core, 2.8GHz per core) and 16GB of RAM, the mean time for updating the occupancy grid is $0.0115$ seconds, and the mean time to determine an exploration strategy is $0.6820$ seconds.
+Time requirements for mapping and exploration are easily modified by changing grid cell resolution or the number of exploration pose candidates, respectively.
+In short, there is a tradeoff between computational speed and accuracy of mapping or exploration.
+
+The autonomous exploration is governed by a policy to maximize expected entropy decreases, so entropy and entropy change with time (Figure \ref{fig:ExperimentH}) are important metrics.
+In this example, the a priori occupancy probabilities of all cells are $0.1$, so the entropies could temporarily rise for likely occupied cells.
+The total map entropy begins at roughly $2613$ and decreases by roughly half to $1338$.
+Many grid cells within the map limits are occluded by walls, preventing the blocked cells from occupancy probability changes.
+In some cases, the robot is able to update the cell occupancy probabilities; in other cases, cells fall inside regions occluded from all reachable space, and the robot cannot update these cells.
+Among those grid cells visible from collision-free locations, the vast majority of cells were updated in less than $3$ min while following a very slow ($5$ cm/sec) trajectory.
+
+In conclusion, this example shows how occupancy grid mapping and autonomous exploration are applied to a ground vehicle in a 2D environment such as a floor of a house or building.
+Dijkstra's search is applied about the occupancy grid, and a smooth trajectory is easily tracked with a simple controller that tracks a polynomial least squares trajectory.
+These results show the efficacy of exact occupancy grid mapping and autonomous exploration.
+Next, we extend these experiments to 3D.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{JIRS_fig11.pdf}
+	\caption{The map entropy $H$ averaged over all $90601$ grid cells is reduced during experimentation.
+The quick changes $\Delta H$ per cell correspond to map updates when the robot captures new territory.}
+\label{fig:ExperimentH}
+\end{figure}
+
+\section{Quadrotor Exploration of a Large Room}
+\label{sec:QuadrotorNRL}
+
+\section{Quadrotor 3D Exploration}
+\label{sec:QuadrotorSEH}

### 2018-06-27 Dissertation/Thesis/tex/results.tex
added & edited sec. 7.2 (NRL flight tests), added patrol split screen @ 2 min

@@
 
-\section{Quadrotor 3D Exploration}
+In this section, we present results from a quadrotor as it autonomously explores a nearly vertically-uniform room.
+A robots builds a 3D map of its environment, which is projected onto a 2D map for exploration following the process described in Section~\ref{sec:MapProjections}.
+The software structure and sensor parameters are identical to those presented in Section~\ref{sec:Compare2MapProjections}.
+
+\subsection{Exploration Environment}
+
+The U.S.
+Naval Research Laboratory (NRL) has a large experimental space for testing known as the Laboratory for Autonomous Systems Research (LASR).
+Here, the experimental setup contains walls and objects to resemble a building floor plan.
+The perimeter walls are constructed with tan cardboard, which were suspended from $1$ m tall stanchions, and an internal wall is built with gray metal sheeting mounted to $80/20$ supports~\cite{url_8020}.
+Additionally, a hard flooring reflects some florescent lighting above.
+The objects inside the experimental environment include various trash cans, desks, and chairs in addition to a few miscellaneous objects.
+The map size is selected to slightly exceed the volume of the experiment such that the the x-direction (positive east) spans $0.5$ m to $10.3$ m, the y-direction (positive north) spans $-8.3$ m to $3.5$ m, and the z-direction (positive vertically up) spans $-0.15$ m to $1.5$ m.
+Two images of the experimental environment are displayed in Figure \ref{fig:exp3DEnvironment}.
+
+\begin{figure}[!t]
+\centering
+    	\begin{subfigure}[t]{0.95\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{experiment_north.jpg}
+        		\caption{North Region}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.95\columnwidth}
+	\vspace*{0.03\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{experiment_south.jpg}
+        		\caption{South Region}
+    	\end{subfigure}
+	\caption{The 3D experimental environment is designed to resemble a large room within an office building.
+The walls are build from from cardboard suspended from stanchions and gray sheeting with $80/20$ supports, and miscellaneous objects are placed around the room as well.}
+	\label{fig:exp3DEnvironment}
+\end{figure}
+
+\subsection{Hardware Structure}
+
+Several components are used for localization, sensing, and actuation.
+A Vicon motion capture system provides the transformation between a reference frame fixed to the world and the moving vehicle.
+This transformation, along with fixed transformations between the robot body and the sensors, provides sufficient information for the transformation from the world to each sensor frame.
+The sensor readings from the Xtion and Hokuyo are transmitted from an NVidia Jetson TX2 onboard the vehicle (Figure \ref{fig:QuadrotorHardware}).
+The sensor readings are received via WIFI on a host computer with an Intel Core i7-6800K CPU (12$\times$3.40GHz), which runs the mapping and exploration nodes.
+The host computer returns the exploration trajectories to the Jetson TX2 over WIFI.
+Then, the Jetson executes a nonlinear geometric flight controller~\cite{GooDaeLee13} onboard.
+With this configuration, large processing tasks are avoided onboard the robot.
+
+\begin{figure}[!t]
+\centering
+    	\begin{subfigure}[t]{0.44\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{quad_top.png}
+        		\caption{Quadrotor Top}
+    	\end{subfigure}
+	\hspace*{0.05\columnwidth}
+    	\begin{subfigure}[t]{0.44\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{quad_bottom.png}
+        		\caption{Quadrotor Bottom}
+    	\end{subfigure}
+	\caption{A Jetson TX2 (quadrotor top) streams depth measurements via WIFI from an Asus Xtion and Hokuyo LIDAR (quadrotor bottom).
+The host computer builds a 3D probabilistic occupancy grid map from these measurements.
+For autonomous exploration, the host computer designs exploration trajectories, which an onboard flight controller can track in real-time.}
+	\label{fig:QuadrotorHardware}
+\end{figure}
+
+\subsection{Experimental Results}
+
+The quadrotor takes off, completes a rotation to understand its immediate surroundings, then explores the space over $2$ minutes and $47$ seconds.
+Figure \ref{fig:exp3DMap} shows the probabilistic 3D map generated in real-time at four selected times, and the corresponding 2D projected maps (x-direction upward) are illustrated in Figure \ref{fig:exp2DMap}, where arrows represent pose candidates and a trajectory is shown to the optimal pose.
+Figure \ref{fig:expH} shows the total map entropy decreasing over the experiment duration.
+A video of the experiment can be found at \href{https://www.youtube.com/watch?v=I_1rXV2XRqk}{\WriteBlue{https://www.youtube.com/watch?v=I\_1rXV2XRqk}}.
+
+% trim={<left> <lower> <right> <upper>}
+
+\begin{figure}[!t]
+\centering
+    	\begin{subfigure}[t]{0.65\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{experiment_ogm3D_0min.jpg}
+        		\caption{$0$ min}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.65\columnwidth}
+		\vspace*{0.01\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{experiment_ogm3D_1min.jpg}
+        		\caption{$1$ min}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.65\columnwidth}
+	\vspace*{0.01\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{experiment_ogm3D_2min.jpg}
+        		\caption{$2$ min}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.65\columnwidth}
+	\vspace*{0.01\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{experiment_ogm3D_2min47sec.jpg}
+        		\caption{$2$ min $47$ sec (end)}
+    	\end{subfigure}
+	\caption{The 3D occupancy grid map is overlaid with point cloud measurements. The map probabilities directly rely on the Xtion and Hokuyo measurements and the sensor stochastic properties.}
+	\label{fig:exp3DMap}
+\end{figure}
+
+\begin{figure}[!t]
+\centering
+    	\begin{subfigure}[t]{0.44\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{experiment_0min_2D.jpg}
+        		\caption{$0$ min}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.44\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{experiment_1min_2D.jpg}
+        		\caption{$1$ min}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.44\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{experiment_2min_2D.jpg}
+        		\caption{$2$ min}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.44\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{experiment_2min47sec_2D_labeled.jpeg}
+        		\caption{$2$ min $47$ sec (end)}
+    	\end{subfigure}
+	\caption{The 2D projected map is easily produced in real-time.
+Candidate future poses, shown by arrows, have more blue color for larger objective functions, found with expected information gains and travel distances.
+Collision-free waypoints (blue line) serve as input to a constrained polynomial least-squares path (light green curve) for the robot controller to follow, where the desired pose (large axes) is tracked by the onboard controller from the current robot pose (small axes).}
+	\label{fig:exp2DMap}
+\end{figure}
+
+\begin{figure}[!t]
+	\centering
+	\includegraphics[width=0.85\textwidth]{entropy_experiment_aspect3by1.pdf}
+	\caption{Map entropy decreases as the robot explores and captures more of the experimental environment.}
+	\label{fig:expH}
+\end{figure}
+
+The robot is able to produce a 3D occupancy grid in under $3$ minutes without a human providing a trajectory.
+The 3D map can be easily interpreted by a human, and provides sufficient information in real-time for autonomous exploration.
+Optimal poses and collision-free exploration trajectories are calculated by the exploration node in under $1$ second.
+These quick decisions are vital to the autonomous exploration performance since the map is initially uncertain.
+
+The projected map used for both collision and exploration has several benefits.
+The robot measures walls and objects with the 3D map, and nicely represents these as occupied spaces with the 2D projected map.
+Additionally, the exploration computation is low, allowing only for brief hover periods (less than $1$ second) between trajectories.
+However, the robot exploration does not consider those cells far below the robot, particularly those near the ground.
+As a result, some floor regions are not captured by the 3D occupancy grid even though cells above the floor are well-known.
+These neglected cells on the floor motivate the full 3D exploration, shown in the next experiment.
+
+\section{Full 3D Exploration with a Quadrotor}
 \label{sec:QuadrotorSEH}

### 2018-06-27 Dissertation/Thesis/tex/multivehicle.tex
added & edited sec. 7.2 (NRL flight tests), added patrol split screen @ 2 min
got rid of '_ss' to figs for SEH multi-vehicle autonomous exploration (not patrol)

@@
     	\begin{subfigure}[t]{0.25\columnwidth}
-          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_1min_ss.jpg}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_1min.jpg}
         		\caption{$1$ min}
@@
     	\begin{subfigure}[t]{0.25\columnwidth}
-          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_2min_ss.jpg}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_2min.jpg}
         		\caption{$2$ min}
@@
 	\begin{subfigure}[t]{0.25\columnwidth}
-          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_3min_ss.jpg}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_3min.jpg}
         		\caption{$3$ min}
@@
 	\begin{subfigure}[t]{0.25\columnwidth}
-          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_4min_ss.jpg}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_4min.jpg}
         		\caption{$4$ min}
@@
     	\begin{subfigure}[t]{0.25\columnwidth}
-          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_5min_ss.jpg}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_5min.jpg}
         		\caption{$5$ min}
@@
     	\begin{subfigure}[t]{0.25\columnwidth}
-          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_6min_ss.jpg}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_6min.jpg}
         		\caption{$6$ min}
@@
    	\begin{subfigure}[t]{0.25\columnwidth}
-          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_7min_ss.jpg}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_7min.jpg}
         		\caption{$7$ min}
@@
     	\begin{subfigure}[t]{0.25\columnwidth}
-          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_8min_ss.jpg}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_8min.jpg}
         		\caption{$8$ min}
@@
     	\begin{subfigure}[t]{0.25\columnwidth}
-          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_9min_ss.jpg}
+          	\includegraphics[trim={26cm 10cm 24cm 10cm}, clip, width=\columnwidth]{multi_discount_raw_3D_9min.jpg}
         		\caption{$9$ min}
@@
     	\begin{subfigure}[t]{0.25\columnwidth}
-          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_1min.jpg}
-        		\caption{$2$ min}% TODO: generate this figure!
+          	\includegraphics[trim={0cm 0cm 35cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_2min.jpg}
+        		\caption{$2$ min}
     	\end{subfigure}
@@
     	\begin{subfigure}[t]{0.25\columnwidth}
-          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_1min.jpg}
-        		\caption{$2$ min}% TODO: generate this figure!
+          	\includegraphics[trim={35cm 0cm 0cm 0cm}, clip, width=\columnwidth]{Patrol_Split_Screen_2min.jpg}
+        		\caption{$2$ min}
     	\end{subfigure}

### 2018-07-01 Dissertation/Thesis/tex/conclusions.tex
added beginning of 3D exploration results, made template for conclusions

\chapter{Conclusions}\label{chap:Conclusions}

This dissertation proposed several contributions in probabilistic occupancy grid mapping and autonomous exploration in 2D and 3D environments.

### 2018-07-01 Dissertation/Thesis/tex/results.tex
added beginning of 3D exploration results, made template for conclusions

@@
 \label{sec:QuadrotorSEH}
+
+In this final experiment, we introduce the quadrotor to a space with complex 3D objects.
+The robot generates a 3D map of this space, and performs autonomous exploration while considering the expected information gains of sample measurement rays with nonzero vertical components.
+The robot unlike prior examples, the robot also moves vertically during exploration and tracks the ground below the vehicle.
+
+\subsection{3D Exploration Setup and Paramaters}
+The software follows the same structure as in Section \ref{sec:MarsSimulation}, except the Gazebo simulated environment is replaces with experimental equipment.
+This equipment is identical to the prior experimental example (Section \ref{sec:QuadrotorNRL}), except the experiment took place at The George Washington University's (GWU) Science and Engineering Hall (SEH).
+This environment is smaller, where the x-direction (positive east) spans $-2.0$ m to $2.0$ m, the y-direction (positive north) spans $-2.0$ m to $2.0$ m, and the z-direction (positive vertically up) spans $-0.15$ m to $1.5$ m.
+
+\subsection{3D Exploration Results}

### 2018-07-02 Dissertation/Thesis/tex/conclusions.tex
added conclusions chapter (chap 8) and edited it

@@
 
-This dissertation proposed several contributions in probabilistic occupancy grid mapping and autonomous exploration in 2D and 3D environments.
+This concludes the dissertation.
+We propose several contributions in probabilistic occupancy grid mapping and autonomous exploration in 2D and 3D environments in single- and multi-vehicle scenarios.
+Computational efficiency and numerical accuracy are held paramount in derivations and algorithmic structure.
+Several numerical examples and experiments demonstrate the efficacy of each contribution.
+
+The primary contribution is an exact and efficient solution to probabilistic occupancy grid mapping.
+Previously believed too computationally expensive, the exact solution can now be applied in real-time to 2D and 3D environments.
+The proposed novel Bayesian solution, referred to as the inverse sensor model, relies on grouping several possible map combinations with identical forward sensor models, rather than considering each map combination individually.
+This approach exploits properties of occupancy grid maps, and follows a recursive algorithm that avoids repeated calculations efficiently.
+
+The solution to probabilistic occupancy grid mapping naturally extends to predicting future maps.
+The proposed autonomous exploration scheme capitalizes on map predictions using Shannon's entropy as an uncertainty measure.
+With the novel proposed approach, probabilities of future map measurements can be predicted, which can be used in conjunction with the inverse sensor model to predict future map entropy.
+Robots select poses to minimize expected entropy, equivalently maximizing map information gain.
+The algorithm is designed to be applied in real-time such that robots can handle uncertain environments without prior information.
+
+The contributions in probabilistic occupancy grid mapping and autonomous exploration are extended in several ways.
+Probabilistic mapping and autonomous exploration are extended from 2D to 3D with two approaches: projecting a 3D map onto a 2D plane for exploration, or using the full 3D map for entropy predictions.
+This concept is also extended to cooperative multi-vehicle autonomous exploration, where a bidding-based framework coordinates robotic efforts to understand the map.
+The approach follows a receding horizon, which improves collision-avoidance and substantially reduces unnecessary robotic actions.
+Furthermore, this approach is extended to autonomous patrol, where an exponential map degradation incentivizes robot team members to periodically revisit areas.
+These extensions to occupancy grid mapping and autonomous exploration are demonstrated with numerous simulations and experimental results.
+
+There are several possible future directions with this research.
+Most importantly, localization, where the robot determines its own position and attitude, should be included in conjunction with the proposed mapping approach, thereby completing the full simultaneous localization and mapping (SLAM) problem.
+Secondly, the cost map portion of Dijkstra's search can be expensive, particularly with multiple vehicles in a large space, or in 3D environments.
+A more efficient approach that provides collision-free distance costs during autonomous exploration could improve computational speed for autonomous exploration.
+Finally, enhancements in memory allocation and distributed computing are required for these algorithms to be applied in a decentralized manner for systems with more autonomy.

### 2018-07-02 Dissertation/Thesis/tex/results.tex
added conclusions chapter (chap 8) and edited it

@@
 This equipment is identical to the prior experimental example (Section \ref{sec:QuadrotorNRL}), except the experiment took place at The George Washington University's (GWU) Science and Engineering Hall (SEH).
-This environment is smaller, where the x-direction (positive east) spans $-2.0$ m to $2.0$ m, the y-direction (positive north) spans $-2.0$ m to $2.0$ m, and the z-direction (positive vertically up) spans $-0.15$ m to $1.5$ m.
+This environment is smaller with only $6$ cameras, where the x-direction (positive east) spans $-2.0$ m to $2.0$ m, the y-direction (positive north) spans $-2.0$ m to $2.0$ m, and the z-direction (positive vertically up) spans $-0.15$ m to $1.5$ m.
 

### 2018-07-02 Dissertation/Thesis/tex/pogm.tex
figure captions in chap 2 updated for caption, then description, better for frontmatter

@@
 
-We propose a computational algorithm to evaluate \refeqn{InvSenModWithProbDens} efficiently. 
+We propose a computational algorithm to evaluate \refeqn{InvSenModWithProbDens} efficiently~\cite{KauLeeAiMos16,KauTakAiLee17}. 
 Since the cells outside of the sensor field of view (FOV) are not affected, we focus on a reduced map $r_l$ in the FOV of the $l$-th ray.
@@
   \includegraphics[width=0.4\textwidth]{rkplus_4.png}
-  \caption{A robot located to the left of a 1D occupancy grid map $r_l$ composed of $n_{r,l}=6$ grid cells in four cases.
+  \caption{Occupancy Grid Maps with Identical Forward Sensor Models}
+  \medskip
+  \small
+  A robot located to the left of a 1D occupancy grid map $r_l$ composed of $n_{r,l}=6$ grid cells in four cases.
 In each case, cells the first three cells are free, the fourth cell is occupied, and the fifth and sixth cells may or may not be occupied.
-All above outcomes correspond to the event $\mathbf{r}_{l,4+}$.}
+All above outcomes correspond to the event $\mathbf{r}_{l,4+}$.
   \label{fig:show_rkplus}
@@
     \end{subfigure}
-    \caption{This simple 2D example illustrates how ray casting determines the cells that compose the reduced map for the inverse sensor model.
+    \caption{Ray Casting Illustration}
+    \label{fig:RayCastingIllustration}
+    	\medskip
+  	\small
+	This simple 2D example illustrates how ray casting determines the cells that compose the reduced map for the inverse sensor model.
 The closest edges are used to determine which cells should be considered, and in what order.
-The complete map cell indices are temporarily saved as well, which are used to associate probabilities to cells of the reduced map.}
-    \label{fig:RayCastingIllustration}
+The complete map cell indices are temporarily saved as well, which are used to associate probabilities to cells of the reduced map.
 \end{figure}
@@
     \end{subfigure}
-    \caption{Using experimental sensor data, the map was generated using an approximate and the proposed exact inverse sensor models.
-The laser sensor and odometry data used for the experiment was provided by University of Pennsylvania open course robotics estimation and learning on Coursera~\cite{coursera}.}
+    \caption{Comparison of Proposed Ray-By-Ray Approach with Approximate Solution}
+	\medskip
+	\small
+	Using experimental sensor data, the map was generated using an approximate and the proposed exact inverse sensor models.
+The laser sensor and odometry data used for the experiment was provided by University of Pennsylvania open course robotics estimation and learning on Coursera~\cite{coursera}.
     \label{fig:penn_map}
@@
     \end{subfigure}
-    \caption{The entropy is compared between the approximate and proposed exact inverse sensor models at step 2280. The blue areas show the lowest entropy while red shows highest.}
+    \caption{Map Uncertainty Comparison Between Proposed Ray-By-Ray Approach and Approximate Solution}
+	\medskip
+	\small
+	The entropy is compared between the approximate and proposed exact inverse sensor models at step 2280. The blue areas show the lowest entropy while red shows highest.
 \label{fig:entropy_comp}
@@
   \includegraphics[width=0.7\textwidth]{entropy_frame.png}
-  \caption{The map entropies of the approximate inverse sensor model (blue) and the proposed exact inverse sensor model (red) show that the exact version yields a more certain occupancy grid map.}
+  \caption{Entropy Histories of the Proposed Ray-By-Ray Approach and Approximate Solution}
+  	\medskip
+	\small
+	The map entropies of the approximate inverse sensor model (blue) and the proposed exact inverse sensor model (red) show that the exact version yields a more certain occupancy grid map.
 \label{fig:entropy}
@@
     }
-\caption{A robot (green crossed circle, green curve of its previous $13$ seconds) measures a room with a Kinect depth sensor.
-The exact inverse sensor model is compared with an approximate algorithm designed for the Kinect depth sensor.
-Grid cells are either known as free (white) or occupied (black), or uncertain with a grayscale color.
-%The exact inverse sensor model approach generates a more accurate result.
-}
+    \caption{Comparison of Proposed Synergistic Approach with Approximate Solution}
+	\medskip
+	\small
+	A robot (green crossed circle, green curve of its previous $13$ seconds) measures a room with a Kinect depth sensor.
+Grid cells are either known as free (white) or occupied (black), or uncertain (gray).
 \label{fig:NumResOccProbs}
@@
     }
-\caption{Entropy serves as a measure of uncertainty of the occupancy grid, where more blue regions are more certain, and red regions are more uncertain.
-The uncertainty is always less when the exact solution is applied.}\label{fig:NumResOccH}
+    \caption{Map Uncertainty Comparison Between Proposed Synergistic Approach and Approximate Solution}
+	\medskip
+	\small
+	Entropy serves as a measure of uncertainty of the occupancy grid, where more blue regions are more certain, and red regions are more uncertain.
+The uncertainty is always less when the exact solution is applied.
+\label{fig:NumResOccH}
 \end{figure}

### 2018-07-02 Dissertation/Thesis/tex/results.tex
figure captions in chap 2 updated for caption, then description, better for frontmatter

@@
 
-\paragraph{Hardware Configuration}
+\subsection{Hardware Configuration}
 The Pioneer 3 ground robot is chosen for this experiment because of its reliability and simple operation.
@@
 
-\paragraph{Software Configuration}
+\subsection{Software Configuration}
 
@@
 
-\paragraph{Exploring and Mapping a 2D Environment}
+\subsection{Exploring and Mapping a 2D Environment}
 The environment consists of Styrofoam walls on a flat floor.

### 2018-07-02 Dissertation/Thesis/tex/ae2D.tex
added and edited Secion 6.1
added & edited sec. 7.1 (Pioneer ground experiment)
figure captions in chap 3 updated

@@
     	\end{subfigure}
-\caption{The entropy error is decreased at the cost of increasing computation time in this Monte Carlo 1D measurement ray expected entropy case.}
+	\caption{Expected Entropy Computational Improvement of Reduced Ray Cells Considered}
+	\medskip
+	\small
+	The entropy error is decreased at the cost of increasing computation time in this Monte Carlo 1D measurement ray expected entropy case.
 	\label{fig:ApproxJust}
@@
 \end{picture}
-\caption{This figure illustrates the expanding ring method.
+\caption{Optimal Pose Selection Illustration
+}
+	\medskip
+	\small
+	This figure illustrates the expanding ring method.
 Initially, a robot (blue circle) views down and left (blue sector).
@@
 Finally, Dijkstra's algorithm provides the collision-free motion between $X_t$ and $X_c^*$, and the process is repeated.
-}
 \label{fig:OptProcess}
@@
 There are two important steps to moving a robot on a collision-free trajectory to its optimal pose.
-First, Dijkstra's algorithm provides the cost map and optimal collision-free waypoints.
+First, Dijkstra's algorithm~\cite{Dij59} provides the cost map and optimal collision-free waypoints.
 Then, a constrained polynomial least squares trajectory serves as the desired path, which is followed by a controller.
@@
 \section{Numerical Examples}
+\label{sec:Ae2DNumExamples}
 
@@
 }
-\caption{A robot (green circle, green marker of previous path) measures a room with a Kinect depth sensor to maximize map information gain.}
+	\caption{Autonomous Exploration of a 2D Space Using the Expanding Ring Approach}
+	\medskip
+	\small
+	A robot (green circle, green marker of previous path) measures a room with a Kinect depth sensor to maximize map information gain.
 \label{fig:AutonomousExploration}
@@
 	\includegraphics[width=0.4\textwidth]{intel_clean.png}
-    \caption{Intel Research Lab floor plan benchmark~\cite{kummerle2009measuring}.}
+	\caption{Intel Research Lab Floor Plan Benchmark~\cite{kummerle2009measuring}.}
+	\medskip
+	\small
+	The 2D floor plan from the Intel Research Lab is simplified to eliminate small objects. The remaining black pixels are used to produce walls for 2D simulations.
 \label{fig:intel}
@@
     \end{subfigure}
-    \caption{A robot (green) measures a room in the ROS Stage simulator using the Intel Research Lab floor plan. The exact inverse sensor model is used for the autonomous exploration algorithm.}
+	\caption{Autonomous Exploration of the Intel Research Lab Using the Complete Cartesian Approach}
+	\medskip
+	\small
+	A robot (green) measures a room in the ROS Stage simulator using the Intel Research Lab floor plan. The exact inverse sensor model is used for the autonomous exploration algorithm.
     \label{fig:IRL}

### 2018-07-02 Dissertation/Thesis/tex/ae3Dsimple.tex
updated & edited up to 5.2.2 (no patrol simulation yet)
added & edited sec. 7.2 (NRL flight tests), added patrol split screen @ 2 min
updated figs of chap 4

@@
 \section{Exploration in Vertically-Uniform Environments}
+\label{sec:MapProjections}
 
@@
 
-\subsubsection{Collision and Entropy Maps}
+\subsection{Collision and Entropy Maps}
 Occupancy grid map cell volume is frequently smaller than the size of the robot taking measurements, particularly with highly-detailed maps.
@@
 
-\subsubsection{Collision and Entropy Combination Map}
+\subsection{Collision and Entropy Combination Map}
 The second proposed technique combines the advantages of both the collision and entropy maps into a 2D projected occupancy grid map, which can be used for both tasks.
@@
 \begin{align}
+\label{eqn:CombinationProjection2DMap}
 P(\mathbf{m}_{\text{2D},k})= 
@@
 \section{Numerical Example}
+\label{sec:Compare2MapProjections}
 
@@
 	\caption{Simulated 3D Occupancy Grid Map}
+	\medskip
+	\small
+	A robot (RGB axes) explores a simulated 3D Gazebo environment using separate collision and entropy projected maps.
+The robot considers numerous candidates (arrows, blue: greater objective function, red: smaller objective function) and selects the optimal candidate (blue arrow).
 \label{fig:sim3DMap}
@@
         		\caption{Environment}
+		\label{fig:Sim2ProjMapsGazebo}
     	\end{subfigure}
@@
         		\caption{Collision Map}
+		\label{fig:Sim2ProjMapsCollision}
     	\end{subfigure}
@@
         		\caption{Entropy Map}
-		\label{fig:sim2DmapsC}
+		\label{fig:Sim2ProjMapsEntropy}
     	\end{subfigure}
@@
 \caption{Simulated Environment and 2D Projected Maps at $4$ min}
+	\medskip
+	\small
+	The robot hovering near the beginning of the simulation is shown in Figure \ref{fig:Sim2ProjMapsGazebo}.
+The projected maps for collision (Figure \ref{fig:Sim2ProjMapsCollision}) and entropy (Figure \ref{fig:Sim2ProjMapsEntropy}) are shown $4$ minutes into the simulation.
+The robot and candidates are shown the same as Figure \ref{fig:sim3DMap}, where the space is considered mostly collision-free, though still largely uncertain.
 \label{fig:sim2Dmaps}
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{entropy_gazebo_aspect3by1.png}
+	\includegraphics[width=0.9\textwidth]{entropy_experiment_aspect3by1.pdf}
 	\caption{Simulated Environment Map Entropy}
+	\medskip
+	\small
+	The entropy of the full 3D map decreases while the robot follows an autonomous exploration scheme based on projected map cells that consider maximum entropy in the vertical direction.
 	\label{fig:simH}
@@
 Conversely, this attention to detail has negative consequences that encourage the robot to repeatedly look at the same regions from different vantage points.
-Since an entropy map cell is composed from many 3D grid cells single-stacked vertically, if just one of these 3D map cells is not captured or hardly captured, the entropy map cell will be recorded as quite uncertain, shown covering numerous locations in Figure~\ref{fig:sim2DmapsC}.
+Since an entropy map cell is composed from many 3D grid cells single-stacked vertically, if just one of these 3D map cells is not captured or hardly captured, the entropy map cell will be recorded as quite uncertain, shown covering numerous locations in Figure~\ref{fig:Sim2ProjMapsEntropy}.
 The robot may attempt to measure this space again, even when other regions might more strongly warrant observation.

### 2018-07-02 Dissertation/Thesis/tex/ae3Dsimple.tex
updated figs of chap 5

@@
 	\small
-	The robot hovering near the beginning of the simulation is shown in Figure \ref{fig:Sim2ProjMapsGazebo}.
-The projected maps for collision (Figure \ref{fig:Sim2ProjMapsCollision}) and entropy (Figure \ref{fig:Sim2ProjMapsEntropy}) are shown $4$ minutes into the simulation.
+	The robot hovering near the beginning of the simulation is shown in (a).
+The projected maps for collision (b) and entropy (c) are shown $4$ minutes into the simulation.
 The robot and candidates are shown the same as Figure \ref{fig:sim3DMap}, where the space is considered mostly collision-free, though still largely uncertain.

### 2018-07-02 Dissertation/Thesis/tex/multivehicle.tex
updated figs of chap 5

@@
 This chapter covers a bidding-based formulation for multiple vehicles cooperating together in the context of autonomous exploration and patrol.
-We focus on a series of auctions to assign tasks, and consider dynamic obstacles, such as people, in large office-like environments using a receding-horizon framework.
+We focus on a series of auctions to assign tasks, and consider dynamic obstacles, such as people, in large office-like environments using a receding horizon framework.
 
@@
 		}
-		\caption{The bump function is designed to promote travel distances close to $d_\text{opt}$ by multiplying this function by the expected information gain objective.
+		\caption{Bump Function for Multi-Vehicle Autonomous Exploration}
+		\medskip
+		\small
+		The bump function is designed to promote travel distances close to $d_\text{opt}$ by multiplying this function by the expected information gain objective.
 Travel distances less than $d_\text{opt}$ are below the robot capabilities.
-Travel distances beyond $d_\text{opt}$ are beyond what the robot can reach within the minimum exploration computation time.}
+Travel distances beyond $d_\text{opt}$ are beyond what the robot can reach within the minimum exploration computation time.
 		\label{fig:nonzeroBumpFun}
@@
     	\end{subfigure}
-	\caption{The SEH second floor plan (about $15,000$ sqft) is reconstructed with a Gazebo simulation environment.}
+	\caption{SEH Floor Plan}
+	\medskip
+	\small
+	The SEH second floor plan (about $15,000$ sqft) is reconstructed with a Gazebo simulation environment, excluding some small details and features of closed rooms.
 	\label{fig:SEHEnvironment}
@@
 	}
-	\caption{3D Occupancy Grid Map of SEH Second Floor}
+	\caption{3D Occupancy Grid Map of SEH Second Floor During Multi-Vehicle Exploration}
+	\medskip
+	\small
+	The 3D occupancy grid map of SEH is generated from three robots coordinating their exploration strategy with auctions.
 	\label{fig:sim3DMap}
@@
 	}
-	\caption{2D Projected Occupancy Grid Map of SEH Second Floor}
+	\caption{2D Projected Occupancy Grid Map of SEH Second Floor During Multi-Vehicle Exploration}
+	\medskip
+	\small
+	The 2D projected map serves for both entropy and collision information at the exploration height for all three vehicles as they coordinate their mapping efforts with bidding-based multi-vehicle autonomous exploration.
 	\label{fig:sim2Dmaps}
@@
 		}
-		\caption{Total 3D Map Cell Entropy}
+		\caption{Total 3D Map Cell Entropy for Multi-Vehicle Exploration}
+		\medskip
+		\small
+		The total map cell entropy is calculated using the difference in map entropy before and after each measurement scan from each of the three coordinating vehicles.
 		\label{fig:simH}
@@
 	\includegraphics[width=\textwidth]{DegradeExamples.pdf}
-\caption{The rate at which probabilities converge toward $P_\infty=0.4$ increases based on the difference magnitude of their initial values.}
+	\caption{Grid Cell Degradation Illustration}
+	\medskip
+	\small
+	The rate at which probabilities converge toward $P_\infty=0.4$ increases based on the difference magnitude of their initial values.
 	\label{fig:DegradeExamples}
@@
 	}
-	\caption{The 3D occupancy grid map representation of the SEH second floor is shown during patrol.
-The robots generate these maps while avoiding collisions with each other, the environment, and a non-cooperative human.}
+	\caption{3D Occupancy Grid Map of SEH Second Floor During Multi-Vehicle Patrol}
+	\medskip
+	\small
+	The 3D occupancy grid map representation of the SEH second floor is shown during patrol.
+The robots generate these maps while avoiding collisions with each other, the environment, and a non-cooperative human.
 	\label{fig:Sim3DMapPatrol}
@@
 	}
-	\caption{The 2D projected map from of the SEH second floor shows degradation while robots patrol the space.
-The robots evade the non-cooperative human walking in a rectangular pattern, occasionally capturing the temporarily-occupied space and identifying these cells as occupied on the grid.}
+	\caption{2D Projected Occupancy Grid Map of SEH Second Floor During Multi-Vehicle Patrol}
+	\medskip
+	\small
+	The 2D projected map from of the SEH second floor shows degradation while robots patrol the space.
+The robots evade the non-cooperative human walking in a rectangular pattern, occasionally capturing the temporarily-occupied space and identifying these cells as occupied on the grid.
 	\label{fig:Sim2DMapPatrol}
@@
 	\includegraphics[width=\textwidth]{entropy_patrol_9minSwitch.pdf}
-\caption{Total map entropies are tracked for separate trials: patrol of the building for $15$ minutes (blue) and patrol of the building for $9$ minutes, then hover (red).
-Even though the map is fairly well-known after $9$ minutes in either case, the continuous effort of patrolling over the complete time period shows lower map uncertainty beyond this time as cells are constantly degrading.}
+	\caption{Total 3D Map Cell Entropy for Multi-Vehicle Patrol}
+	\medskip
+	\small
+	Total map entropies are tracked for separate trials: patrol of the building for $15$ minutes (blue) and patrol of the building for $9$ minutes, then hover (red).
+Even though the map is fairly well-known after $9$ minutes in either case, the continuous effort of patrolling over the complete time period shows lower map uncertainty beyond this time as cells are constantly degrading.
 	\label{fig:DegradeExamples}
@@
     	\end{subfigure}
-	\caption{A cooperating robot must evade a non-cooperative human.
-The receding horizon framework allows a quickly-updating map to modify the exploration commands to avoid collisions with a dynamic obstacle.}
+	\caption{Human Collision-Avoidance Using a Receding Horizon}
+	\medskip
+	\small
+	A cooperating robot must evade a non-cooperative human. The receding horizon framework allows a quickly-updating map to modify the exploration commands to avoid collisions with a dynamic obstacle.
 	\label{fig:EvadeHuman}

### 2018-07-02 Dissertation/Thesis/tex/results.tex
updated figs of chap 5

@@
 The software follows the same structure as in Section \ref{sec:MarsSimulation}, except the Gazebo simulated environment is replaces with experimental equipment.
-This equipment is identical to the prior experimental example (Section \ref{sec:QuadrotorNRL}), except the experiment took place at The George Washington University's (GWU) Science and Engineering Hall (SEH).
+This equipment is identical to the prior experimental example (Section \ref{sec:QuadrotorNRL}), except the experiment took place at the Flight Dynamics and Control Lab (FDCL) inside the second floor of SEH.
 This environment is smaller with only $6$ cameras, where the x-direction (positive east) spans $-2.0$ m to $2.0$ m, the y-direction (positive north) spans $-2.0$ m to $2.0$ m, and the z-direction (positive vertically up) spans $-0.15$ m to $1.5$ m.

### 2018-07-02 Dissertation/Thesis/tex/ae3Dcomplex.tex
added beginning of 3D exploration results, made template for conclusions
updated figs of chap 5

@@
     	\end{subfigure}
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-\caption{The three frames are shown with three axes (red: first axis, green: second axis, blue: third axis).
-The Mars frame (short and thick) is fixed to the planet.
+	\caption{Illustration of 3D Exploration Reference Frames}
+	\medskip
+	\small
+	The three frames are shown with three axes (red: first axis, green: second axis, blue: third axis).
+The world frame (short and thick) is fixed to the ground.
 The body-fixed frame (medium thickness and length) is fixed to the camera frame (long and narrow).
-The rotation of $\psi$ (b) represents the fixed camera yaw rotation and $\theta$ (c) represents the downward angle of the camera to capture the surface of Mars below the flying robot.}
+The rotation of $\psi$ (b) represents the fixed camera yaw rotation and $\theta$ (c) represents the downward angle of the camera to capture the ground below the flying robot.
 	\label{fig:transforms}
@@
 
-	\begin{figure}
+\begin{figure}[!t]
+	\centering
 	\vspace*{-0.3\columnwidth}
@@
 	\vspace*{-0.25\columnwidth}
-		\caption{The bump function is multiplied to the objective function of \refeqn{CandidateLocationOptimization} to prioritize local trajectories to avoid traversals across the map, where travel costs are determined using Dijkstra's algorithm.
-However, there is no time wasted up to $d_\text{opt}$ because this distance may be achieved in the minimum exploration computation time.}
+	\caption{Bump Function for Full 3D Exploration}
+	\medskip
+	\small
+	The bump function is multiplied to the objective function of \refeqn{CandidateLocationOptimization} to prioritize local trajectories to avoid traversals across the map, where travel costs are determined using Dijkstra's algorithm.
+However, there is no time wasted up to $d_\text{opt}$ because this distance may be achieved in the minimum exploration computation time.
 	\label{fig:recedingHorizonBumpFun}
@@
 \section{Numerical Simulation}
+\label{sec:MarsSimulation}
+
 In this section, we demonstrate the efficacy of the proposed 3D mapping and autonomous exploration with a simulated Mars environment.
@@
     	\end{subfigure}
-\caption{The height map (a) and color map (b) are combined in the Gazebo simulator (c) to model a Mars environment. The robot captures the simulated environment to generate an occupancy grid map.}
+	\caption{Simulated Mars Surface}
+	\medskip
+	\small
+	The height map (a) and color map (b) are combined in the Gazebo simulator (c) to model a Mars environment. The robot captures the simulated environment to generate an occupancy grid map.
 	\label{fig:MarsGazebo}
@@
 The full map has dimensions $20$ m $\times$ $20$ m $\times$ $5$ m in the Mars-fixed frame, with cell edge length $\alpha=0.075$ m.
-The reduced map for collision-avoidance and motion planning has cells $k=3$ times the size ($0.225$ m), so $k^3=27$ cells are considered in \refeqn{Proj3DMapComb}.
+The reduced map for collision avoidance and motion planning has cells $k=3$ times the size ($0.225$ m), so $k^3=27$ cells are considered in \refeqn{Proj3DMapComb}.
 The bump functions use $f_\text{max}=1.0$, $f_\text{far}=0.1$, and $\beta=0.1$ to account for travel costs with \refeqn{BumpFun}.
@@
     	\end{subfigure}
-\caption{In Case 1, the robot (red disk with arrow indicating laser direction) moves toward candidate poses (red arrows, more opaque for greater reward) based on expected entropy change of the 3D probabilistic occupancy grid map $m$ (cubes: greater opacity for greater occupancy probability) of the surface of Mars.}
+\caption{Mars 3D Occupancy Grid Map for Case 1}
+	\medskip
+	\small
+	In Case 1, the robot (red disk with arrow indicating laser direction) moves toward candidate poses (red arrows, more opaque for greater reward) based on expected entropy change of the 3D probabilistic occupancy grid map $m$ (cubes: greater opacity for greater occupancy probability) of the surface of Mars.
 \label{fig:mars3DogmCase1}
@@
     	\end{subfigure}
-\caption{In Case 2, the robot (red disk with arrow indicating laser direction) generates a 3D probabilistic occupancy grid map $m$ (cubes: greater opacity for greater occupancy probability), which is used to generate a reduced map $m_\text{reduced}$ based on \refeqn{Proj3DMapComb}.
+\caption{Mars 3D Occupancy Grid Map for Case 2}
+	\medskip
+	\small
+	In Case 2, the robot (red disk with arrow indicating laser direction) generates a 3D probabilistic occupancy grid map $m$ (cubes: greater opacity for greater occupancy probability), which is used to generate a reduced map $m_\text{reduced}$ based on \refeqn{Proj3DMapComb}.
 The robot moves toward candidate poses (red arrows, more opaque for greater reward) based on expected entropy change of $m_\text{reduced}$.
-Using $m_\text{reduced}$ instead of $m$ for entropy calculation leads to faster exploration of new terrain, but this leaves some grid cells missing.}
+This approach leads to faster exploration of new terrain, but this leaves some grid cells missing.
 \label{fig:mars3DogmCase2}
@@
     	\end{subfigure}
-\caption{The close-up images from the Case 2 trial show the candidate future poses (red arrows), where greater opacity represents a larger objective function of \refeqn{CandidateLocationOptimization}.
+	\caption{Zoomed-In Occupancy Grid Cells of Mars}
+	\medskip
+	\small
+	The close-up images from the Case 2 trial show the candidate future poses (red arrows), where greater opacity represents a larger objective function of \refeqn{CandidateLocationOptimization}.
 In (a), we show the 3D scan with color corresponding to the surface of Mars.
-In (b), we overlay the full map $m$ (colored) with the reduced map $m_\text{reduced}$ (gray) from \refeqn{Proj3DMapComb}.}
+In (b), we overlay the full map $m$ (colored) with the reduced map $m_\text{reduced}$ (gray) from \refeqn{Proj3DMapComb}.
 	\label{fig:marsZoomedIn}
@@
 	\centerline{
-			\includegraphics[width=0.6\columnwidth]{scitech_entropy_comparison_flat.pdf}
+		\includegraphics[width=0.9\columnwidth]{scitech_entropy_comparison_flat.pdf}
 	}
-		\caption{The complete map entropy for Case 1 (red solid line) decreases at roughly the same rate as the reduced map entropy for Case 2 (dashed blue line) toward the beginning.
-However, the expected entropy in Case 2 promotes actions toward unexplored territory, while the policy of Case 1 yields a more complete map before moving on to unexplored spaces.}
+	\caption{Entropy Histories During Mars Exploration}
+	\medskip
+	\small
+	The complete map entropy for Case 1 (red solid line) decreases at roughly the same rate as the reduced map entropy for Case 2 (dashed blue line) toward the beginning.
+However, the expected entropy in Case 2 promotes actions toward unexplored territory, while the policy of Case 1 yields a more complete map before moving on to unexplored spaces.
 	\label{fig:mars3Dentropy}

### 2018-07-02 Dissertation/Thesis/tex/results.tex
updated figs of chap 5
updated figs of chap 6, started 7

@@
     	\end{subfigure}
-\caption{Images from two perspectives show the walls and obstacles of the experimental environment.}
+\caption{2D Experimental Environment}
+	\medskip
+	\small
+	Images from two perspectives show the walls and obstacles of the experimental environment.
 \label{fig:ExpSetupPhoto}
@@
 	}
-\caption{A robot autonomously explores the experimental environment and produces an occupancy grid map along the way.
+\caption{Experimental 2D Occupancy Grid Map Results}
+	\medskip
+	\small
+	A robot autonomously explores the experimental environment and produces an occupancy grid map along the way.
 The robot (shaded green circle body with shaded blue circle sensor) is controlled toward the desired pose $1$ sec in the future (unshaded grey robot), which is following a constrained least-squares trajectory (cyan curve).
-This trajectory is based on waypoints (black circles) from Dijkstra's algorithm to arrive at the optimal pose (unshaded green circle body with unshaded blue circle sensor) as proposed in this paper.}
+This trajectory is based on waypoints (black circles) from Dijkstra's algorithm to arrive at the optimal pose (unshaded green circle body with unshaded blue circle sensor) as proposed in this paper.
 \label{fig:ExperimentOGM}

### 2018-07-02 Dissertation/Thesis/tex/results.tex
updated figs of chap 7 (none in 8)

@@
 	A robot autonomously explores the experimental environment and produces an occupancy grid map along the way.
-The robot (shaded green circle body with shaded blue circle sensor) is controlled toward the desired pose $1$ sec in the future (unshaded grey robot), which is following a constrained least-squares trajectory (cyan curve).
+The robot (shaded green circle body with shaded blue circle sensor) is controlled toward the desired pose $1$ sec in the future (unshaded gray robot), which is following a constrained least-squares trajectory (cyan curve).
 This trajectory is based on waypoints (black circles) from Dijkstra's algorithm to arrive at the optimal pose (unshaded green circle body with unshaded blue circle sensor) as proposed in this paper.
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{JIRS_fig11.pdf}
+	\caption{Entropy Change During 2D Experiment}
+	\medskip
+	\small
+	The map entropy $H$ averaged over all $90,601$ grid cells is reduced during experimentation. The quick changes $\Delta H$ per cell correspond to map updates when the robot captures new territory.
+	\label{fig:ExperimentH}
+\end{figure}
+
 In conclusion, this example shows how occupancy grid mapping and autonomous exploration are applied to a ground vehicle in a 2D environment such as a floor of a house or building.
@@
 Next, we extend these experiments to 3D.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{JIRS_fig11.pdf}
-	\caption{The map entropy $H$ averaged over all $90601$ grid cells is reduced during experimentation.
-The quick changes $\Delta H$ per cell correspond to map updates when the robot captures new territory.}
-\label{fig:ExperimentH}
-\end{figure}
 
@@
     	\end{subfigure}
-	\caption{The 3D experimental environment is designed to resemble a large room within an office building.
-The walls are build from from cardboard suspended from stanchions and gray sheeting with $80/20$ supports, and miscellaneous objects are placed around the room as well.}
+	\caption{3D Experimental Environment for Level Flight}
+	\medskip
+	\small
+	The 3D experimental environment is designed to resemble a large room within an office building.
+The walls are build from from cardboard suspended from stanchions and gray sheeting with $80/20$ supports, and miscellaneous objects are placed around the room as well.
 	\label{fig:exp3DEnvironment}
@@
     	\end{subfigure}
-	\caption{A Jetson TX2 (quadrotor top) streams depth measurements via WIFI from an Asus Xtion and Hokuyo LIDAR (quadrotor bottom).
+	\caption{Quadrotor Platform for 3D Mapping and Exploration}
+	\medskip
+	\small
+	A Jetson TX2 (quadrotor top) streams depth measurements via WIFI from an Asus Xtion and Hokuyo LIDAR (quadrotor bottom).
 The host computer builds a 3D probabilistic occupancy grid map from these measurements.
-For autonomous exploration, the host computer designs exploration trajectories, which an onboard flight controller can track in real-time.}
+For autonomous exploration, the host computer designs exploration trajectories, which an onboard flight controller can track in real-time.
+This platform is used for both level flight (Section \ref{sec:QuadrotorNRL}) and non-level flight (Section \ref{sec:QuadrotorSEH}).
 	\label{fig:QuadrotorHardware}
@@
     	\end{subfigure}
-	\caption{The 3D occupancy grid map is overlaid with point cloud measurements. The map probabilities directly rely on the Xtion and Hokuyo measurements and the sensor stochastic properties.}
+	\caption{3D Occupancy Grid Map During Level Flight Experiment}
+	\medskip
+	\small
+	The 3D occupancy grid map is overlaid with point cloud measurements. The map probabilities directly rely on the Xtion and Hokuyo measurements and the sensor stochastic properties.
 	\label{fig:exp3DMap}
@@
     	\end{subfigure}
-	\caption{The 2D projected map is easily produced in real-time.
+	\caption{2D Projected Occupancy Grid Map During Level Flight Experiment}
+	\medskip
+	\small
+	The 2D projected map is easily produced in real-time.
 Candidate future poses, shown by arrows, have more blue color for larger objective functions, found with expected information gains and travel distances.
-Collision-free waypoints (blue line) serve as input to a constrained polynomial least-squares path (light green curve) for the robot controller to follow, where the desired pose (large axes) is tracked by the onboard controller from the current robot pose (small axes).}
+Collision-free waypoints (blue line) serve as input to a constrained polynomial least-squares path (light green curve) for the robot controller to follow, where the desired pose (large axes) is tracked by the onboard controller from the current robot pose (small axes).
 	\label{fig:exp2DMap}
@@
 	\includegraphics[width=0.85\textwidth]{entropy_experiment_aspect3by1.pdf}
-	\caption{Map entropy decreases as the robot explores and captures more of the experimental environment.}
+	\caption{Entropy History During Level Flight Experiment}
+	\medskip
+	\small
+	Map entropy decreases as the robot explores and captures more of the experimental environment.
 	\label{fig:expH}

### 2018-07-02 Dissertation/Thesis/tex/ae2D.tex
updated figs of chap 5
updated references of myself

@@
 
-In this section, we define Shannon's entropy as a measure of map uncertainty, and present a novel approach to predict future entropy from an arbitrary ray.
+In this section, we define Shannon's entropy as a measure of map uncertainty, and present a novel approach to predict future entropy from an arbitrary ray~\cite{KauAiLee16}.
 Then we discuss computational modifications for real-time implementation.
@@
 }
+\centering{
 	\medskip
@@
 Finally, Dijkstra's algorithm provides the collision-free motion between $X_t$ and $X_c^*$, and the process is repeated.
+}
 \label{fig:OptProcess}

### 2018-07-02 Dissertation/Thesis/tex/intro.tex
added outline of dissertation to chap 1

@@
 
-ADD ONCE COMPLETE
+This dissertation is organized as follows.
+Chapter \ref{chap:pogm} introduces a novel approach to solve for the exact Bayesian probability of grid cells in real-time.
+Then, Chapter \ref{chap:ae2D} extends these contributions to autonomous exploration in 2D space, where a novel approach to predict future map uncertainty is presented.
+Next, Chapter \ref{chap:ae2D} applies the proposed occupancy grid mapping approach to 3D environments, and presents a method to simplify autonomous exploration in 3D environments with map projections.
+Chapter \ref{chap:multivehicle} applies this 3D mapping and exploration to multi-vehicle exploration and patrol scenarios using a bidding-based framework.
+Then, Chapter \ref{chap:ae3Dcomplex} extends exploration to complex 3D environments without 2D map projections.
+In these chapters, several numerical examples demonstrate the efficacies of the approaches.
+Furthermore, Chapter \ref{chap:Experiments} presents three experiments, which show the key contributions in this dissertation.
+Finally, these are followed by conclusions in Chapter \ref{chap:Conclusions}.
 

### 2018-07-02 Conference/ICRA 2019.1/BackgroundResearch.tex

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% originally created by Michael Shell
%%
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/
%%
%% Originally created by Michael Shell
%% Modifed by J. Kim for URAI conference

\documentclass[conference]{IEEEURAI}
\usepackage{graphicx}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{3D Mapping with Bidding-Based Autonomous Exploration of Cooperative Quadrotors}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Evan Kaufman$^{1}$, Taeyoung Lee$^{2}$}
\IEEEauthorblockA{$^{1,2}$Department of Mechanical and Aerospace Engineering\\The George Washington University, Washington, DC, USA\\
Email: \{evankaufman, tylee\}@gwu.edu}}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
These instructions give you the basic guidelines for preparing the paper for URAI. The abstract text should be organized to include the background, method of approach, results, and conclusions. The abstract is to be in 10-point, single-spaced type, and should not exceed 300 words.\\
\end{abstract}

\begin{IEEEkeywords}
Mapping, Exploration, Cooperation, Entropy
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\subsection*{Background Research}
\begin{itemize}
	\item Simmons et. al.~\cite{SimApfBurFoxMooThrYou00}
	\begin{itemize}
		\item Introduced the concept of ``bids''
		\item Main idea: central executive assigns tasks to maximized overall utility and minimize coverage overlap
		\item \emph{bids}: ``describe their estimates of the expected information gain and costs of traveling to various locations''
		\item Keep robots separated to decrease overall time
		\item Each robot has its own local map
		\item Expected information gain: from frontier cells, count the number unknown cells before free/occupied cells are reached within a circular region
		\item Without discounting (percentage overlap), the actions would be uncoordinated
		\item Interested addition: hystereses is used to make robots finish their tasks
	\end{itemize}
	\item Zlot et. al.~\cite{ZloSteDiaTha02}
	\begin{itemize}
		\item Main idea: robots share information with each other and auction off ``tasks'' 
		\item References~\cite{SimApfBurFoxMooThrYou00} giving credit for bidding, but claiming that the centralized manner prevents robustness
		\item Proposed approach stresses robustness: introduction/loss of team members, communication interruptions \& failures
		\item Central interface is only for humans and can disappear; robots communicate directly with each other for completely distributed implementation
	\end{itemize}
	\item Cavalcante et. al.~\cite{CavNorCha13}
	\begin{itemize}
		\item Team of robots must visit a set of target points on a known map
		\item Goal: compute and execute paths to minimize total mission costs with combinatorial auctions
		\item ``Auction-based mechanisms have been used to coordinate multiple robots in exploration missions, due to their simplicity and flexibility''
		\item A different problem: given $m$ targets and $n$ robots, move robots toward targets, then return in some optimal fashion
		\item Not really about generating a map, as this is assumed known
	\end{itemize}
%	\item Elizondo-Leal et. al.~\cite{LesRamPul08}
	\item Choi et. al.~\cite{ChoBruHow09}
	\begin{itemize}
		\item Decentralized with local communications
		\item ``...presenting two decentralized algorithms: the consensus-based auction algorithm (CBAA) and its generalization to the multi-assignment problem, i.e., the consensus-based bundle algorithm (CBBA)''
		\item Both algorithms proven to guarantee convergence to a conflict-free assignment
	\end{itemize}
	\item Sariel et. al.~\cite{SarBal05,SarBal06}
	\begin{itemize}
		\item Decentralized: ``We propose a framework... that require diverse capabilities and collective work without a central planner/decision maker''
		\item Multi-agent multi-target exploration problem in a dynamic environment
		\item Each target is visited by at least 1 robot
		\item Optimal: total cost minimization
	\end{itemize}
	\item Main ideas
	\begin{itemize}
		\item Decentralized approaches are very common, using one agent as an auctioneer that considers its own situational awareness
		\item Inspired by Simmons et. al.~\cite{SimApfBurFoxMooThrYou00} that applies discounting for cooperation, but now we apply these ideas to a different mapping objective
		\item Argument against decentralized approach: Since the occupancy grid represents the entire environment, it is natural that the exploration strategy adopts a global interpretation.
	\end{itemize}
\end{itemize}

\section{Introduction}

\begin{itemize}
	\item Motivations and definitions
	\begin{itemize}
		\item Motivation for mapping and exploration
		\item Motivation for multi-agent systems
		\item Definitions of mapping and exploration
	\end{itemize}
	\item Reasons for Exact Bayesian mapping
	\begin{itemize}
		\item Uses sensor properties explicitly
		\item Stochastic properties may be different for different sensors
		\item Multiple sensors per vehicle, multiple vehicles okay
	\end{itemize}
	\item Entropy-based exploration
	\begin{itemize}
		\item Reasons for entropy, and recent advances for precise solution
		\item Difficulty of considering various robot configuration combinations
		\item How auction has cooperative benefits with computationally-efficient manner
	\end{itemize}
	\item Organization of paper
\end{itemize}

%\subsection{First Part}
%Author names and affiliations are to be centered beneath the paper title and typed in 11-point Times New Roman, non-boldface type. Multiple authors may be shown in a two- or three-column format, with their affiliations below the authors’ names. Include the e-mail addresses. The author information is followed by two single-spaced blank lines before the main text.
%
%\subsection{Format}
%In formatting your page, set the top margin to 25mm, the bottom margin to 25mm, the side margins to 20mm, and heading to 15mm. This gives a column width of 82mm and the space between two columns 6mm. Left - and right - justify the columns.  
% 
%Type the main text in 10-point Times, single-spaced. All paragraphs should be indented 4mm. Please do not place any additional blank lines between paragraphs. 
%
%
%\section{Elements of a Paper}
%\subsection{Figures and Tables}
%\subsubsection{Figures}
%All figures (graphs, line drawings, photographs, etc.) should be numbered consecutively and have a caption consisting of the figure number and a brief title or description of the figure. Figures should be referenced within the text as “Fig. 1.” When the reference to a figure begins a sentence, the abbreviation "Fig." should be spelled out as “Figure 1.” When the reference to a figure begins a sentence, the abbreviation "Fig." should be spelled out as “Figure 1.”
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{kros.png}
%\caption{Caption of a sample figure. It is indented 4mm and placed below the figure.}
%\label{fig_kros}
%\end{figure}
%
%\subsubsection{Tables}
%All tables should be numbered consecutively and have a caption consisting of the table number and a brief title. This number should be used when referring to the table in the text. Tables should be inserted as part of the text as close as possible to its first reference. 
%
%\begin{table}[!t]
%%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
%\caption{Caption of a sample table.}
%\label{table_example}
%\centering
%\begin{tabular}{cc}
%\hline
%Number & Types\\
%\hline
%1 & Industrial robots\\
%\hline
%2 & Service robots\\
%\hline
%\end{tabular}
%\end{table}
%
%\subsection{Equations}
%Equations should be numbered consecutively beginning with (1) to the end of the paper, including any appendices. The number should be enclosed in parentheses and set flush right in the column on the same line as the equation. Equations should be referenced within the text as “Eq. (x).” When the reference to an equation begins a sentence, it should be spelled out as “Equation (x).”
%
%\begin{equation}
%x^2+y^2=r^2
%\end{equation}
%
%\subsection{References}
%Within the text, references should be cited in numerical order according to their order of appearance. The numbered reference citation within text should be enclosed in brackets (e.g., It was shown by Craig [1] that … . ). In the case of two citations, the numbers should be separated by a comma [1, 2]. In the case of more than two references, the numbers should be separated by a dash [1-4].
%
%References should be listed together at the end of the paper in numerical order according to the sequence of citations within the text. Refer to some references at the end of this template for correct format. 
%
%\section*{Acknowledgement}
%Acknowledgments may be made to individuals or institutions not mentioned elsewhere in the work who have made an important contribution.

% On the last page of your paper, adjust the lengths of the columns so that they are approximately equal.

% references section
%\begin{thebibliography}{1}
%\bibitem{IEEEhowto:kopka}
%G. D. Hong, T. Nakamura and K. G. Kim, “Article Title,” Journal Title, Vol. 10, No. 6, pp. 1-10, 2006. 
%\bibitem{IEEEhowto:kopka}
%A. C. Smith, Introduction to Robotics, Publisher, New York, 2006.
%\end{thebibliography}
\bibliography{../../BibSources}% master source for all publications always 2 directories up
\bibliographystyle{IEEEtran}

\end{document}

### 2018-07-02 Dissertation/Thesis/tex/frontmatter.tex
added & edited sec. 7.2 (NRL flight tests), added patrol split screen @ 2 min

@@
 
-\phdgrad{August 31, 2018}
-\defensedate{July 11, 2018}
+\phdgrad{July 13, 2018}
+\defensedate{July 13, 2018}
 % Year of completion for copyright page and perhaps other places
@@
 First and foremost, I would like to thank my parents.
-In my younger years, I was not always a great student, but they believed in me.
+When I was younger, I was not always a great student, but my parents believed in me.
 When I struggled, they helped every time.
@@
 
-I would also like to thank all other parts of my family, including my fianc\'ee, brother, grandparents, aunts, uncles, and cousins.
+I would also like to thank all other parts of my family, including my fianc\'ee, brother, grandparents, aunts, uncles, cousins, and close friends.
 We make it a priority to come together and support each other, even when our careers differ greatly and our lives change unpredictably.
-Despite the long academic road, and they provide support and love.
+Despite the long academic road, they provided consistent support and love.
 
@@
 
- I think that following a passion is something truly special.
+ I think that finding and pursuing a passion is something truly special.
 Even though I find great satisfaction in doing anything well, regardless of how simple or complicated the task, I find myself extremely fortunate to pursue endeavors that I actually find exciting and stimulating.

### 2018-07-02 Dissertation/Thesis/tex/results.tex

@@
 
-\subsection{3D Exploration Setup and Paramaters}
+\subsection{3D Exploration Setup and Parameters}
 The software follows the same structure as in Section \ref{sec:MarsSimulation}, except the Gazebo simulated environment is replaces with experimental equipment.

### 2018-07-05 Dissertation/Thesis/tex/ae2D.tex
tylee edit

@@
 
-\paragraph{Algorithm} We present an algorithm pseudocode providing the necessary steps to obtain the objective function for a single measurement ray (Algorithm \ref{alg:RayExpectedEntropyGain}).
-Much like the algorithm pseudocode for the ray-by-ray inverse sensor model (Algorithm \ref{alg:RayByRayISM}), the variable $a_\text{temp}$ serves as an intermediate variable designed to avoid repeated calculations.
+\paragraph{Algorithm} We present an algorithm pseudo-code providing the necessary steps to obtain the objective function for a single measurement ray (Algorithm \ref{alg:RayExpectedEntropyGain}).
+Much like the algorithm pseudo-code for the ray-by-ray inverse sensor model (Algorithm \ref{alg:RayByRayISM}), the variable $a_\text{temp}$ serves as an intermediate variable designed to avoid repeated calculations.
 Since this algorithm operates as a function, fixed indices and condition variables are removed for simplification.

### 2018-07-05 Dissertation/Thesis/tex/ae3Dcomplex.tex
tylee edit

@@
 Defining $0<P_\text{coll}<1$ as the acceptable probability of collision, every grid cell of $ m_\text{reduced}$ is considered a safe and unvisited node if its occupancy probability, and the occupancy probabilities of its neighbors (sharing a face, edge, or corner), are below $P_\text{coll}$.
-A cost map is built from the starting robot location by neighboring nodes, where the cost to travel to another node is based on Euclidian distance: a cell sharing a face is $\alpha$ away, a cell sharing an edge is $\sqrt2\alpha$ away, and a cell sharing a corner is $\sqrt3\alpha$ away.
+A cost map is built from the starting robot location by neighboring nodes, where the cost to travel to another node is based on Euclidean distance: a cell sharing a face is $\alpha$ away, a cell sharing an edge is $\sqrt2\alpha$ away, and a cell sharing a corner is $\sqrt3\alpha$ away.
 Generating a cost map for all reachable cells in $\mathbf{m}_{\text{reduced},k}$ provides a collision-free travel cost for all reachable candidate pose locations.

### 2018-07-05 Dissertation/Thesis/tex/frontmatter.tex
tylee edit

@@
     plural={matrices},
-    description={rectangular array of quantiities}
+    description={rectangular array of quantities}
 }
@@
 \abstract{
-This dissertation focusses on robotic mapping and exploration of uncertain environments.
-Computational algorithms are developed to provide important probability and uncertainty properties.
+This dissertation focuses on robotic mapping and exploration of uncertain environments.
+Computational algorithms are developed to provide complete stochastic information of the environments.
 These algorithms are designed for real-time implementation for robotic autonomy.
 
-Probabilistic occupancy grid mapping is developed using Bayesian probability.
+First, probabilistic occupancy grid mapping is developed according to Bayesian framework.
 A novel approach to this problem is explored, which uses important physical properties of the environment and stochastic properties of depth sensors.
@@
 
-The next topic this dissertation addresses is autonomous exploration, where a robot selects actions to maximize its knowledge about the probabilistic map.
+The next topic is autonomous exploration, where a robot selects actions to maximize its knowledge about the probabilistic map.
 We select Shannon's entropy as a metric that represents grid cell uncertainty.
@@
 The final topic of this dissertation relates to multi-vehicle cooperative scenarios.
-The mapping algorithm is multi-threaded to accept measurements from multiple sources with differing sensor properties.
+The mapping algorithm is revised to accept measurements from multiple sources with differing sensor properties.
 More importantly, the exploration algorithm structure is modified with a bidding-based framework.

### 2018-07-05 Dissertation/Thesis/tex/intro.tex
tylee edit

@@
 
-The second topic of this dissertation focusses on the motions robots must choose to maximize their knowledge about the map while avoiding collisions, known as \emph{autonomous exploration}.
+The second topic of this dissertation focuses on the motions robots must choose to maximize their knowledge about the map while avoiding collisions, known as \emph{autonomous exploration}.
 Unlike conventional motion planning, where the map of the environment is assumed known, the robot must select motions with only the current knowledge of the environment.
@@
 Chapter \ref{chap:pogm} introduces a novel approach to solve for the exact Bayesian probability of grid cells in real-time.
-Then, Chapter \ref{chap:ae2D} extends these contributions to autonomous exploration in 2D space, where a novel approach to predict future map uncertainty is presented.
+Then, Chapter \ref{chap:ae2D} extends these contributions to autonomous exploration in 2D space, where a computational approach to predict future map uncertainty is presented.
 Next, Chapter \ref{chap:ae2D} applies the proposed occupancy grid mapping approach to 3D environments, and presents a method to simplify autonomous exploration in 3D environments with map projections.
@@
 % TODO: update this list
+% TL: the paper in preparation can be added as well
 

### 2018-07-05 Dissertation/Thesis/tex/multivehicle.tex
tylee edit

@@
 \subsection{Receding Horizon Framework}
-
-This algorithm is further aided by following a receding horizon framework for improved information gain maximizations and collision-avoidance with dynamic obstacles and other robots.
+% TL: can we expand this section a little bit. for example, discussion on dynamic obstacles, or pseudo-code can be added.
+This algorithm is further aided by following a receding horizon framework for improved information gain maximization and collision-avoidance with dynamic obstacles and other robots.
 A receding horizon simply repeats the autonomous exploration steps as quickly as possible over a finite time period, frequently before a robot reaches its desired pose.

### 2018-07-05 Dissertation/Thesis/tex/pogm.tex
figure captions in chap 3 updated
tylee edit

@@
 
-In this chapter, we present the exact solution to occupancy grid map probability.
+In this chapter, we present the exact solution to occupancy probability for a given depth measurement.
 This Bayesian approach uses the stochastic properties of sensors, and exploits important patterns in conditional probabilities.
@@
 
-Using this, \refeqn{ThirdBayesRule} is rewritten as
+\sloppy Using this, \refeqn{ThirdBayesRule} is rewritten as
 \begin{align}
@@
 \end{align*}
-as well as $d$ and $e$ for the two terms composing \refeqn{xi}. Once again, the history of poses and measurement scans are removed from the pseudocode for simplicity.
+as well as $d$ and $e$ for the two terms composing \refeqn{xi}. Once again, the history of poses and measurement scans are removed from the pseudo-code for simplicity.
 
@@
 In short, we propose two alternatives for combining inverse sensor models for multiple rays within a scan.
-The ray-by-ray approach updates the entire map one ray at a time with some dependancies among the rays, while the synergistic approach considers all rays together where rays are assumed mutually independent.
+The ray-by-ray approach updates the entire map one ray at a time with some dependencies among the rays, while the synergistic approach considers all rays together where rays are assumed mutually independent.
 

### 2018-07-05 Dissertation/Thesis/tex/results.tex
tylee edit

@@
 
-\section{Pioneer Ground Vehicle Tests}
+\section{2D Exploration With an Unmanned Ground Vehicle}
 \label{sec:PioneerNRL}
@@
 
+% TL: add a figure of Pioneer
 \subsection{Hardware Configuration}
@@
 
-The software structure is similar to ROS nodes Section \ref{sec:Ae2DNumExamples}, which simply involves a mapping node, an exploration node, and a visualization node using ``OpenGL'' libraries.
+% add citation for OpenGL
+The software structure is similar to ROS nodes Section \ref{sec:Ae2DNumExamples}, which simply involves a mapping node, an exploration node, and a visualization node using OpenGL libraries.
 The only differences are eliminating the simulator (Stage) and adding synchronization among the sensors, which is of high importance to the proposed occupancy grid mapping approach.
@@
 \subsection{Exploring and Mapping a 2D Environment}
+% TL: I wish there are more photos from the experiments 
 The environment consists of Styrofoam walls on a flat floor.
@@
 
-\section{Quadrotor Exploration of a Large Room}
+\section{Aerial Exploration of a Large Space}
 \label{sec:QuadrotorNRL}
@@
 The objects inside the experimental environment include various trash cans, desks, and chairs in addition to a few miscellaneous objects.
-The map size is selected to slightly exceed the volume of the experiment such that the the x-direction (positive east) spans $0.5$ m to $10.3$ m, the y-direction (positive north) spans $-8.3$ m to $3.5$ m, and the z-direction (positive vertically up) spans $-0.15$ m to $1.5$ m.
+The map size is selected to slightly exceed the volume of the experiment such that the x-direction (positive east) spans $0.5$ m to $10.3$ m, the y-direction (positive north) spans $-8.3$ m to $3.5$ m, and the z-direction (positive vertically up) spans $-0.15$ m to $1.5$ m.
 Two images of the experimental environment are displayed in Figure \ref{fig:exp3DEnvironment}.
@@
 
+% TL: add figures for more time instances
 \begin{figure}[!t]

### 2018-07-07 Dissertation/Thesis/tex/ae2D.tex
updated from tlee edits, more figs for NRL flight experiment

@@
 This process is repeated until the expected information gain of the optimal pose is at least $\mathcal I_\text{min}$, or all candidates lie in collision zones or outside map limits.
-%A pseudocode of the process is found in Table \ref{tab:Alg_AutomousExploration}.
+%A pseudo-code of the process is found in Table \ref{tab:Alg_AutomousExploration}.
 
@@
 The robot motion is governed by a policy that maximizes the map information gain within its set of pose candidates, where the $\hat n=6$ is chosen to approximate the expected entropy of each ray.
-When running the exploration algorithm within the framework of the Robot Operating System (ROS), the mean computation time is $0.0194$ seconds to determine the optimal future pose and complete Dijkstra's algorithm on the occupancy grid.
+When running the exploration algorithm within the framework of the Robot Operating System (ROS)~\cite{ROS}, the mean computation time is $0.0194$ seconds to determine the optimal future pose and complete Dijkstra's algorithm on the occupancy grid.
 The computation times for this map reached roughly $1$ second at maximum, corresponding to rare cases when the robot is locally surrounded by a highly-certain environment; this task requires evaluating many more pose candidates and motion planning over a larger terrain.

### 2018-07-07 Dissertation/Thesis/tex/conclusions.tex
updated from tlee edits, more figs for NRL flight experiment

@@
 
-This concludes the dissertation.
-We propose several contributions in probabilistic occupancy grid mapping and autonomous exploration in 2D and 3D environments in single- and multi-vehicle scenarios.
+\section{Summary of Contributions}
+In this dissertation, we propose several contributions in probabilistic occupancy grid mapping and autonomous exploration in 2D and 3D environments in single- and multi-vehicle scenarios.
 Computational efficiency and numerical accuracy are held paramount in derivations and algorithmic structure.
@@
 
+\section{Future Directions}
+
 There are several possible future directions with this research.

### 2018-07-07 Dissertation/Thesis/tex/frontmatter.tex
updated from tlee edits, more figs for NRL flight experiment

@@
 \committee{ %
+
+James Hahn\\
+Department of Computer Science\\
+The George Washington University\\
+Committee Chair\\
+
 Taeyoung Lee\\
@@
 The George Washington University\\ 
-Dissertation Director\\ % remember to add a space between committee members
+Committee Member\\ % remember to add a space between committee members
 
 Robert Pless\\
-Department of Computer Science\\
-The George Washington University\\
-Committee Member\\
-
-James Hahn\\
 Department of Computer Science\\

### 2018-07-07 Dissertation/Thesis/tex/intro.tex
updated from tlee edits, more figs for NRL flight experiment

@@
 
-% JINT17: This paper proposes a computationally efficient algo- rithm to construct the exact inverse sensor model.
-More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, this algo- rithm yields the a posteriori probabilities of occupancy of all the cells within the area covered by the range sensor from the range measurements.
-The key idea is reducing the computational load by using the fact that if a cell is occu- pied, the occupancies of the other cells blocked by it is irrelevant to the forward sensor model, and this property is systematically utilized with various probabilistic prop- erties to derive a computationally-efficient solution to the inverse sensor model.
-Furthermore, the proposed approach integrates a priori probabilities of occupancy and multiple range measurements according to the Bayesian framework to obtain more accurate maps.
-As such, it contrasts from the existing framework based on log-odds ratios that impose an additional Markov assumption.
-
 The second contribution is a new approach to autonomous exploration, where the robot predicts future occupancy grid map uncertainty, using a measure known as Shannon's entropy.
@@
 
-% TODO: update this list
-% TL: the paper in preparation can be added as well
-
 \begin{itemize}
+	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Multi-Vehicle Cooperative Exploration and Patrol of Uncertain 3D Environments'', in preparation.
+	\item E.
+Kaufman and T.
+Lee, ``Autonomous Aerial Exploration for Topological Mapping of Mars Environments'', \textit{2019 AIAA Guidance, Navigation, and Control Conference}, San Diego, CA, USA, January 2019, submitted.
 	\item E.
@@
 Ai, and T.
-Lee, ``Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities,'' \textit{The Second IEEE International Conference on Robotic Computing}, Laguna Hills, January 2018, pp.
-49--55.
+Lee, ``Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities,'' \textit{The Second IEEE International Conference on Robotic Computing}, pp.
+49--55, Laguna Hills, January 2018.
 	\item E.
@@
 Lee, and Z.
-Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, San Francisco, December 2016, pp.
-246--251.
+Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, pp.
+246--251, San Francisco, CA, USA, December 2016.
 	\item E.
@@
 S.
-Moskowitz, ``Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model,'' \textit{Proceedings of the American Control Conference}, Boston, July 2016, pp.
-5709-5715.
+Moskowitz, ``Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model,'' \textit{Proceedings of the American Control Conference}, pp.
+5709-5715, Boston, MA, USA, July 2016.
 	\item E.
@@
 Lovell, and T.
-Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting}, Williamsburg, VA, Jan.
-2015, AAS 15-451.
+Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting}, AAS 15-451, Williamsburg, VA, January 2015.
 	\item E.
@@
 Lovell, and T.
-Lee, ``Minimum Uncertainty JPDA Filter and Coalescence Avoidance Performance Evaluations,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting, Williamsburg}, VA, Jan.
-2015, AAS 15-432.
+Lee, ``Minimum Uncertainty JPDA Filter and Coalescence Avoidance Performance Evaluations,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting, Williamsburg}, AAS 15-432, VA, January 2015.
 	\item E.
@@
 Lee, and T.
-Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, Mar.
-2014, ASC 14-2527.
+Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, ASC 14-2527, March 2014, Big Sky, MT, USA.
 	\item E.
@@
 Lee, ``Optimal Joint Probabilistic Data Association Filter Avoiding Coalescence in Close Proximity,'' \textit{Proceedings of the European Control Conference}, pp.
-2709-2714, June 2014.
+2709-2714, Strasbourg, France, June 2014.
 	\vspace*{0.5cm}

### 2018-07-07 Dissertation/Thesis/tex/pogm.tex
updated from tlee edits, more figs for NRL flight experiment

@@
 
-In this chapter, we present the exact solution to occupancy probability for a given depth measurement.
+In this chapter, we present the exact solution to occupancy grid map probability for a given depth measurement.
 This Bayesian approach uses the stochastic properties of sensors, and exploits important patterns in conditional probabilities.
-This solution is simulated in several examples to show the efficacy of the approach.
+Then we show how several measurements can be considered together in 2D.
+This solution is simulated with numerical examples to show the efficacy of the approach.
 
@@
 
-The synergistic scan inverse sensor model algorithm is summarized with the pseudocode of Algorithm \ref{alg:SynergisticScanISM} as a computationally efficient, recursive algorithm that avoids repeated calculations of the same quantity.
+The synergistic scan inverse sensor model algorithm is summarized with the pseudo-code of Algorithm \ref{alg:SynergisticScanISM} as a computationally efficient, recursive algorithm that avoids repeated calculations of the same quantity.
 This algorithm utilizes the following temporary variables to develop the algorithm in an efficient recursive form, defined as

### 2018-07-07 Dissertation/Thesis/tex/results.tex
updated from tlee edits, more figs for NRL flight experiment

@@
 
-The experiment involves a ground vehicle producing a probabilistic occupancy grid map in real time.
+The experiment involves a Pioneer ground vehicle producing a probabilistic occupancy grid map in real time.
 The robot motion is governed by the information gain-maximizing policy described in Chapter \ref{chap:ae2D}.
 
-% TL: add a figure of Pioneer
 \subsection{Hardware Configuration}
-The Pioneer 3 ground robot is chosen for this experiment because of its reliability and simple operation.
+The Pioneer 3 ground robot (Figure \ref{fig:Pioneer}) is chosen for this experiment because of its reliability and simple operation.
 The vehicle accepts two inputs, namely linear velocity (aligned with the robot wheels) and angular velocity (about a central axis passing vertically through the robot).
@@
 
+\begin{figure}
+	\centering{
+    	\begin{subfigure}[b]{0.3\textwidth}
+	\centering
+        		\includegraphics[height=0.7\textwidth]{pioneer.png}
+        		\caption{Pioneer 3 Vehicle}
+        		\label{fig:PioneerStockImage}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.3\textwidth}
+	\centering
+        		\includegraphics[height=0.7\textwidth]{pioneer_side.png}
+        		\caption{Side View}
+        		\label{fig:PioneerSide}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.3\textwidth}
+	\centering
+        		\includegraphics[height=0.7\textwidth]{pioneer_kinect.png}
+        		\caption{Front View with Kinect}
+        		\label{fig:PioneerSide}
+    	\end{subfigure}
+	}
+\caption{Pioneer Robot}
+	\medskip
+	\small
+	The Pioneer 3 robot serves as a reliable experimental platform for 2D mapping and autonomous exploration using a Microsoft Kinect.
+\label{fig:Pioneer}
+\end{figure}
+
 \subsection{Software Configuration}
 
-% add citation for OpenGL
-The software structure is similar to ROS nodes Section \ref{sec:Ae2DNumExamples}, which simply involves a mapping node, an exploration node, and a visualization node using OpenGL libraries.
+The software structure is similar to ROS nodes Section \ref{sec:Ae2DNumExamples}, which simply involves a mapping node, an exploration node, and a visualization node using OpenGL libraries~\cite{openGL}.
 The only differences are eliminating the simulator (Stage) and adding synchronization among the sensors, which is of high importance to the proposed occupancy grid mapping approach.
@@
 \subsection{Exploring and Mapping a 2D Environment}
-% TL: I wish there are more photos from the experiments 
 The environment consists of Styrofoam walls on a flat floor.
@@
 
-The quadrotor takes off, completes a rotation to understand its immediate surroundings, then explores the space over $2$ minutes and $47$ seconds.
+The quadrotor takes off, completes a rotation to understand its immediate surroundings, then explores the space over $2$ minutes and $47$ seconds, shown in Figure \ref{fig:expFilming}.
 Figure \ref{fig:exp3DMap} shows the probabilistic 3D map generated in real-time at four selected times, and the corresponding 2D projected maps (x-direction upward) are illustrated in Figure \ref{fig:exp2DMap}, where arrows represent pose candidates and a trajectory is shown to the optimal pose.
@@
 
-% trim={<left> <lower> <right> <upper>}
+\begin{figure}[!t]
+\centering{
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 20cm 4cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_0sec.jpg}
+        		\caption{$0$ sec}
+    	\end{subfigure}
+	\hspace*{0.04\columnwidth}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 20cm 4cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_20sec.jpg}
+        		\caption{$20$ sec}
+    	\end{subfigure}
+}
+\centering{
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 20cm 4cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_40sec.jpg}
+        		\caption{$40$ sec}
+    	\end{subfigure}
+	\hspace*{0.04\columnwidth}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 20cm 4cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_60sec.jpg}
+        		\caption{$60$ sec}
+    	\end{subfigure}
+}
+\centering{
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 20cm 4cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_80sec.jpg}
+        		\caption{$80$ sec}
+    	\end{subfigure}
+	\hspace*{0.04\columnwidth}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 20cm 4cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_100sec.jpg}
+        		\caption{$100$ sec}
+    	\end{subfigure}
+}
+\centering{
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 20cm 4cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_120sec.jpg}
+        		\caption{$120$ sec}
+    	\end{subfigure}
+	\hspace*{0.04\columnwidth}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 20cm 4cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_140sec.jpg}
+        		\caption{$140$ sec}
+    	\end{subfigure}
+}
+\centering{
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 20cm 4cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_160sec.jpg}
+        		\caption{$160$ sec}
+    	\end{subfigure}
+	\hspace*{0.04\columnwidth}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 20cm 4cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_167sec.jpg}
+        		\caption{$167$ sec (end)}
+    	\end{subfigure}
+}
+	\caption{Level Flight Experiment}
+	\medskip
+	\small
+	During the first minute, the robot slowly takes off and does an initial rotation, then explores the space in the remaining time.
+	\label{fig:expFilming}
+\end{figure}
 
-% TL: add figures for more time instances
 \begin{figure}[!t]
+\centering{
+    	\begin{subfigure}[t]{0.4\columnwidth}
            	\centering
-    	\begin{subfigure}[t]{0.65\columnwidth}
+          	\includegraphics[trim={25cm 1cm 4cm 19cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_0sec.jpg}
+        		\caption{$0$ sec}
+    	\end{subfigure}
+	\hspace*{0.04\columnwidth}
+    	\begin{subfigure}[t]{0.4\columnwidth}
            	\centering
-          	\includegraphics[width=\textwidth]{experiment_ogm3D_0min.jpg}
-        		\caption{$0$ min}
+          	\includegraphics[trim={25cm 1cm 4cm 19cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_20sec.jpg}
+        		\caption{$20$ sec}
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.65\columnwidth}
-		\vspace*{0.01\columnwidth}
+}
+\centering{
+    	\begin{subfigure}[t]{0.4\columnwidth}
            	\centering
-          	\includegraphics[width=\textwidth]{experiment_ogm3D_1min.jpg}
-        		\caption{$1$ min}
+          	\includegraphics[trim={25cm 1cm 4cm 19cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_40sec.jpg}
+        		\caption{$40$ sec}
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.65\columnwidth}
-	\vspace*{0.01\columnwidth}
+	\hspace*{0.04\columnwidth}
+    	\begin{subfigure}[t]{0.4\columnwidth}
            	\centering
-          	\includegraphics[width=\textwidth]{experiment_ogm3D_2min.jpg}
-        		\caption{$2$ min}
+          	\includegraphics[trim={25cm 1cm 4cm 19cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_60sec.jpg}
+        		\caption{$60$ sec}
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.65\columnwidth}
-	\vspace*{0.01\columnwidth}
+}
+\centering{
+    	\begin{subfigure}[t]{0.4\columnwidth}
            	\centering
-          	\includegraphics[width=\textwidth]{experiment_ogm3D_2min47sec.jpg}
-        		\caption{$2$ min $47$ sec (end)}
+          	\includegraphics[trim={25cm 1cm 4cm 19cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_80sec.jpg}
+        		\caption{$80$ sec}
     	\end{subfigure}
+	\hspace*{0.04\columnwidth}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 1cm 4cm 19cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_100sec.jpg}
+        		\caption{$100$ sec}
+    	\end{subfigure}
+}
+\centering{
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 1cm 4cm 19cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_120sec.jpg}
+        		\caption{$120$ sec}
+    	\end{subfigure}
+	\hspace*{0.04\columnwidth}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 1cm 4cm 19cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_140sec.jpg}
+        		\caption{$140$ sec}
+    	\end{subfigure}
+}
+\centering{
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 1cm 4cm 19cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_160sec.jpg}
+        		\caption{$160$ sec}
+    	\end{subfigure}
+	\hspace*{0.04\columnwidth}
+    	\begin{subfigure}[t]{0.4\columnwidth}
+           	\centering
+          	\includegraphics[trim={25cm 1cm 4cm 19cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_167sec.jpg}
+        		\caption{$167$ sec (end)}
+    	\end{subfigure}
+}
 	\caption{3D Occupancy Grid Map During Level Flight Experiment}
@@
 
-% TL: add figures for more time instances
+% trim={<left> <lower> <right> <upper>}
+
+%\begin{figure}[!t]
+%\centering{
+%    	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[trim={5cm 0cm 6cm 0cm}, clip, width=\textwidth]{experiment_ogm3D_0min.jpg}
+%        		\caption{$0$ min}
+%    	\end{subfigure}
+%	\hspace*{0.04\columnwidth}
+%    	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[trim={5cm 0cm 6cm 0cm}, clip, width=\textwidth]{experiment_ogm3D_1min.jpg}
+%        		\caption{$1$ min}
+%    	\end{subfigure}
+%}
+%\centering{
+%    	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[trim={5cm 0cm 6cm 0cm}, clip, width=\textwidth]{experiment_ogm3D_2min.jpg}
+%        		\caption{$2$ min}
+%    	\end{subfigure}
+%	\hspace*{0.04\columnwidth}
+%    	\begin{subfigure}[t]{0.4\columnwidth}
+%           	\centering
+%          	\includegraphics[trim={5cm 0cm 6cm 0cm}, clip, width=\textwidth]{experiment_ogm3D_2min47sec.jpg}
+%        		\caption{$2$ min $47$ sec (end)}
+%    	\end{subfigure}
+%}
+%	\caption{3D Occupancy Grid Map During Level Flight Experiment}
+%	\medskip
+%	\small
+%	The 3D occupancy grid map is overlaid with point cloud measurements. The map probabilities directly rely on the Xtion and Hokuyo measurements and the sensor stochastic properties.
+%	\label{fig:exp3DMap}
+%\end{figure}
+
 \begin{figure}[!t]
+\centering{
+    	\begin{subfigure}[t]{0.3\columnwidth}
            	\centering
-    	\begin{subfigure}[t]{0.44\columnwidth}
+          	\includegraphics[trim={0cm 19.8cm 45.5cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_0sec.jpg}
+        		\caption{$0$ sec}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
            	\centering
-          	\includegraphics[width=\textwidth]{experiment_0min_2D.jpg}
-        		\caption{$0$ min}
+          	\includegraphics[trim={0cm 19.8cm 45.5cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_20sec.jpg}
+        		\caption{$20$ sec}
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.44\columnwidth}
+	\hspace*{0.02\columnwidth}
+	\begin{subfigure}[t]{0.3\columnwidth}
            	\centering
-          	\includegraphics[width=\textwidth]{experiment_1min_2D.jpg}
-        		\caption{$1$ min}
+          	\includegraphics[trim={0cm 19.8cm 45.5cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_40sec.jpg}
+        		\caption{$40$ sec}
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.44\columnwidth}
+}
+\\
+\centering{
+    	\begin{subfigure}[t]{0.3\columnwidth}
            	\centering
-          	\includegraphics[width=\textwidth]{experiment_2min_2D.jpg}
-        		\caption{$2$ min}
+          	\includegraphics[trim={0cm 19.8cm 45.5cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_60sec.jpg}
+        		\caption{$60$ sec}
     	\end{subfigure}
-    	\begin{subfigure}[t]{0.44\columnwidth}
+	\hspace*{0.02\columnwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim={0cm 19.8cm 45.5cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_80sec.jpg}
+        		\caption{$80$ sec}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim={0cm 19.8cm 45.5cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_100sec.jpg}
+        		\caption{$100$ sec}
+    	\end{subfigure}
+}
+\\
+\centering{
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim={0cm 19.8cm 45.5cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_120sec.jpg}
+        		\caption{$120$ sec}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim={0cm 19.8cm 45.5cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_140sec.jpg}
+        		\caption{$140$ sec}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim={0cm 19.8cm 45.5cm 0cm}, clip, width=\textwidth]{NRL_quadrotor_4panes_160sec.jpg}
+        		\caption{$160$ sec}
+    	\end{subfigure}
+}
+\\
+\centering{
+    	\begin{subfigure}[t]{0.3\columnwidth}
            	\centering
           	\includegraphics[width=\textwidth]{experiment_2min47sec_2D_labeled.jpeg}
-        		\caption{$2$ min $47$ sec (end)}
+        		\caption{$167$ sec (end)}
     	\end{subfigure}
+}
 	\caption{2D Projected Occupancy Grid Map During Level Flight Experiment}

### 2018-07-07 Dissertation/Thesis/tex/results.tex
added figures to 3D full exploration

@@
 
+\begin{figure}[!t]
+\centering
+    	\begin{subfigure}[t]{0.44\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_space_NW.JPG}
+        		\caption{Northwest View}
+    	\end{subfigure}
+	\hspace*{0.05\columnwidth}
+    	\begin{subfigure}[t]{0.44\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_space_NE.JPG}
+        		\caption{Northeast View}
+    	\end{subfigure}
+	\caption{3D Exploration Space}
+	\medskip
+	\small
+	The experimental space for full 3D exploration consists of normal lab objects such as tables and chairs, as well as a mannequin sitting at a desk and his parked bicycle.
+	\label{fig:Full3DExplorationEnvironment}
+\end{figure}
+
+\begin{figure}[!t]
+\centering{
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_vid_0min.jpg}
+        		\caption{$0$ sec}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_vid_0p25min.jpg}
+        		\caption{$15$ sec}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_vid_0p5min.jpg}
+        		\caption{$30$ sec}
+    	\end{subfigure}
+}
+\\
+\centering{
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_vid_1min.jpg}
+        		\caption{$1$ min}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_vid_1p5min.jpg}
+        		\caption{$1.5$ min}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_vid_2min.jpg}
+        		\caption{$2$ min}
+    	\end{subfigure}
+}
+\\
+\centering{
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_vid_2p5min.jpg}
+        		\caption{$2.5$ min}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_vid_3min.jpg}
+        		\caption{$3$ min}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_vid_3p5min.jpg}
+        		\caption{$3.5$ min}
+    	\end{subfigure}
+}
+\\
+\centering{
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_vid_4min.jpg}
+        		\caption{$4$ min}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_vid_4p5min.jpg}
+        		\caption{$4.5$ min}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_vid_4p75min.jpg}
+        		\caption{$4.75$ min (end)}
+    	\end{subfigure}
+}
+	\caption{Full 3D Autonomous Exploration Experiment}
+	\medskip
+	\small
+	The robot completes a full yaw rotation over $20$ seconds, then autonomously explores the space. At $4$ min (j), the robot flies over the mannequin to explore the southeast corner of the room.
+	\label{fig:exp3DVid}
+\end{figure}
+
+\begin{figure}[!t]
+\centering{
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_0min.jpg}
+        		\caption{$0$ sec}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_0p25min.jpg}
+        		\caption{$15$ sec}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_0p5min.jpg}
+        		\caption{$30$ sec}
+    	\end{subfigure}
+}
+\\
+\centering{
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_1min.jpg}
+        		\caption{$1$ min}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_1p5min.jpg}
+        		\caption{$1.5$ min}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_2min.jpg}
+        		\caption{$2$ min}
+    	\end{subfigure}
+}
+\\
+\centering{
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_2p5min.jpg}
+        		\caption{$2.5$ min}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_3min.jpg}
+        		\caption{$3$ min}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_3p5min.jpg}
+        		\caption{$3.5$ min}
+    	\end{subfigure}
+}
+\\
+\centering{
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_4min.jpg}
+        		\caption{$4$ min}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_4p5min.jpg}
+        		\caption{$4.5$ min}
+    	\end{subfigure}
+	\hspace*{0.02\columnwidth}
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{explore3D_4p75min.jpg}
+        		\caption{$4.75$ min (end)}
+    	\end{subfigure}
+}
+	\caption{3D Occupancy Grid Map During Full 3D Autonomous Exploration}
+	\medskip
+	\small
+	The robot (large axes) is controlled to track the desired pose (small axes), which follows a collision-free trajectory using Dijkstra's search toward candidates (top $75\%$ shown with red arrows, with greater opacity for larger expected information gains).
+	\label{fig:exp3DOccMap}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{H_explore3D_test1.pdf}
+	\caption{Entropy Change During Full 3D Experiment}
+	\medskip
+	\small
+	The map entropy $H$ consistently decreases during the full 3D exploration experiment.
+Toward the end of the experiment, most cells that can be captured by the Asus Xtion have been updated, slowing down the rate of entropy decrease.
+	\label{fig:Explore3D_H}
+\end{figure}
+
+\begin{figure}[!t]
+\centering
+    	\begin{subfigure}[t]{0.8\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{connected_costmap.png}
+        		\caption{Connected Cost Map Cross-Section}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.8\columnwidth}
+           	\centering
+          	\includegraphics[width=\textwidth]{disconnected_costmap.png}
+        		\caption{Disconnected Cost Map Cross-Section}
+    	\end{subfigure}
+	\caption{2D Cross-Sections of 3D Cost Maps}
+	\medskip
+	\small
+	Two examples of the 3D cost maps from Dijkstra's search show connected and disconnected regions at the exploration height, where more blue regions have small distance costs, red regions have larger distance costs, and pink regions are unreachable due to collision risk.
+In (a), the robot does not have to change altitude to reach any of the collision-free poses.
+However in (b), the robot can only reach collision-free poses in the disconnected section by changing altitude.
+	\label{fig:Full3DExplorationCostmaps}
+\end{figure}
+
 \subsection{3D Exploration Results}

### 2018-07-08 Dissertation/Thesis/tex/ae3Dcomplex.tex
updated from tlee edits, more figs for NRL flight experiment
increased refs from 42 to 61 excluding geometric control ones, which will be added later

@@
 
-In this section, we propose how multiple measurement rays in 3D can determine map information gain from a pose candidate.
+In this section, we propose how multiple measurement rays in 3D can determine map information gain from a pose candidate~\cite{KauLee19}.
 Then, we show how a collision-free trajectory is determined with a 3D Dijkstra's search.
@@
 First we cover the parameters, then compare results using two different maps for entropy prediction.
+
+\subsection{Mars Exploration Motivation}
+
+The topography of Mars has several benefits in future mission planning for landing spacecraft and  image distortion correction, among other research objectives.
+The Mars Orbiting Laser Altimeter (MOLA) mission greatly improved the topography map of the surface, but was limited to an uncertainty of roughly eight meters~\cite{MOLA99}.
+The only precise topography knowledge about the surface comes from four rovers, two of which are inactive, and have covered only a small portion of the surface area of Mars.
+Approaches for exploring environments similar to offices, such as \cite{MauDakPet14}, are incompatible with the complex terrain presented in this numerical example.
 

### 2018-07-08 Dissertation/Thesis/tex/intro.tex
increased refs from 42 to 61 excluding geometric control ones, which will be added later

@@
 
+There are several existing mapping representations including occupancy grids~\cite{WolSuk05,WolSuk05,MeyBeiBur12,TanThoWolBus14}, Octomaps~\cite{WurHorBenStaBur10,HorWurBenStaBur13}, or feature-based maps
+~\cite{MonThrKolWeg02}.
+In this dissertation, we focus on occupancy grid maps because of their simple structure, efficient memory usage, and natural application to autonomous exploration in later sections.
+Occupancy grid mapping follows a Bayesian framework that focusses on occupied and free space, not tracking particular systems or mapping features, such as visual servoing~\cite{ChiDinChe14}.
+
 Here we describe the existing approaches to building probabilistic occupancy grid maps.
@@
 The main idea is to simulate maps, robot poses, and measurements with simulated noised based on the sensor stochastic properties.
-Then, an approximation the inverse sensor model is learned using an expectation maximization algorithm~\cite{Thr01,ThrBurFox05}.
+Then, an approximation the inverse sensor model is learned using an expectation maximization algorithm~\cite{Thr01,ThrBurFox05,SouMaiGon12}.
 Unlike approximate functions based on intuition, the machine learning approach is based on the physical properties of sensors.
@@
 The log-odds ratio formulation is a popular framework for updating binary random variables with static state within a Bayesian filter.
-The main idea is that instead of multiplying terms from prior and current time steps, the properties of logarithms allow these terms to be simply added, while avoiding some truncation issues associated with probabilities close to $0$ or $1$~\cite{ThrBurFox05}.
+The main idea is that instead of multiplying terms from prior and current time steps, the properties of logarithms allow these terms to be simply added, while avoiding some truncation issues associated with probabilities close to $0$ or $1$~\cite{ThrBurFox05,Thr03}.
 This approach is also popular because the inverse sensor model needs not consider prior probabilities.
@@
 The vast majority of work in simultaneous localization and mapping (SLAM) deals only with the aspect of estimating the environment and the poses of vehicles.
-These approaches are passive in the sense that SLAM is performed on incoming sensor measurements from vehicles following an arbitrary path.
+These approaches are passive in the sense that SLAM is performed on incoming sensor measurements from vehicles following an arbitrary path (e.g.~\cite{ThrBurFox05,DurBai06,CheChe09}).
 As such, human teleoperation and monitoring are often essential to guide the vehicles safely through unknown surroundings.
@@
 \paragraph{Entropy-Based Autonomous Exploration}
-Entropy-based approaches address the suboptimal nature of frontier-based approaches by selecting robotic actions designed to decrease minimize map uncertainty~\cite{StaGriBur05,BurMooStaSch05}.
+Entropy-based approaches address the suboptimal nature of frontier-based approaches by selecting robotic actions designed to decrease minimize map uncertainty~\cite{CarDamKumCas15,StaGriBur05,BurMooStaSch05}.
 These approaches use a measure of uncertainty known as Shannon's entropy, which becomes smaller as cell probabilities approach $0$ or $1$ (becoming more certain).
@@
 Then, the robot analyzes how this measurement would affect the map uncertainty.
-However, this approach assumes that expected map entropy is equivalent to map entropy from expected measurements.
+However, this approach assumes that expected map entropy is equivalent to map entropy from expected measurements~\cite{JohStaPfaBur07}.
 This is not the case in general, however, as there is typically a nonzero probability that the hallucination measurement and true future measurement are not the same.

### 2018-07-08 Dissertation/Thesis/tex/pogm.tex
increased refs from 42 to 61 excluding geometric control ones, which will be added later

@@
 Here, the robot maps in a two-dimensional environment composed of ten wall edges, and the robot follows a figure-eight curve, then turns around and completes the same curve in the reverse direction.
-Like the prior example, the same set of measurements updated each second are used with both occupancy grid mapping algorithms to construct the map.
+A pseudo-random measurement scan is sampled at each time step via the inverse transform sampling~\cite{DevBK86}, and like the prior example, the same set of measurements are used with both occupancy grid mapping algorithms to construct the map.
 The resulting maps are illustrated in Figure \ref{fig:NumResOccProbs} for both algorithms, where it is shown that the proposed algorithm yields a substantially more accurate and clear map with less uncertainty.

### 2018-07-09 Dissertation/Defense/defense.tex
updated section 7.4 of dissertation to bring figures together with section 7.3, started defense presentation

\documentclass[11pt,professionalfonts,hyperref={pdftex,pdfpagemode=none,pdfstartview=FitH}]{beamer}
%\usepackage{times}
%\usefonttheme{serif}
%\usepackage{helvet}
%\usepackage{amsmath,amssymb}
\usepackage{graphicx,multirow}

\usepackage{movie15}

\usepackage{caption}
\usepackage{subcaption}
\captionsetup{compatibility=false}

%\usepackage{warmread}
%\usepackage[all,import]{xy}

%\renewcommand\mathfamilydefault{\rmdefault}

% JIRS includes
%\usepackage{graphicx}
%\usepackage{amsmath,amssymb,url,times}%,subfigure}% amsthm is the one!
%\usepackage{caption,subcaption,hyperref}
%\usepackage{color,comment}
%\usepackage{curves,pgfgantt}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand{\refprop}[1]{Proposition \ref{prop:#1}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

\definecolor{mygray}{gray}{0.9}

\graphicspath{{../../Fig/}}

\mode<presentation> {
  \usetheme{Warsaw}
  \usefonttheme{serif}
  \setbeamercovered{transparent}
}

\newcommand{\mypaper}{}

\setbeamertemplate{footline}%{split theme}
{%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=.7\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm]{title in head/foot}
    \usebeamerfont{title in head/foot}\mypaper\hfill \insertframenumber/\inserttotalframenumber
    \usebeamerfont{title in head/foot}\hfill \insertframenumber/\inserttotalframenumber
  \end{beamercolorbox}}%
  \vskip0pt%
} \setbeamercolor{box}{fg=black,bg=yellow}

\title[Multi-Robot Probabilistic Mapping and Exploration]{\large Multi-Robot Probabilistic Mapping and Exploration}

\author{\vspace*{-0.1cm}}

\institute{\footnotesize
{\normalsize Evan Kaufman\\Research Advisor: Taeyoung Lee}\vspace*{0.1cm}\\
  Mechanical and Aerospace Engineering\\ George Washington University \vspace*{0.3cm}\\ {\normalsize Special Thanks to:\\
 \vspace*{0.2cm}
 Kuya Takami, Mahdis Bisheban, and Kanishke Gamagedara} \\Department of Mechanical Engineering\\George Washington University\\
\vspace*{0.2cm}
{\normalsize Zhuming Ai, Ira. S. Moskowitz, and Mark Livingston}\\Information Management \& Decision Architectures\\U.S. Naval Research Laboratory}

\date{}

\definecolor{tmp}{rgb}{0.804,0.941,1.0}
\setbeamercolor{numerical}{fg=black,bg=tmp}
\setbeamercolor{exact}{fg=black,bg=red}

\newtheorem{prop}{Proposition}

\renewcommand{\emph}[1]{\textit{\textbf{\color{blue}{#1}}}}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section*{}
\subsection*{Motivation}

\begin{frame}
\frametitle{Motivation}
\begin{itemize}
	\item Environments can be dangerous and uncertain
	\item Robots can replace humans for tasks such as surveillance, search-and-rescue, and cleaning
	\item Robot must understand their environment by generating a map, referred to as \emph{mapping}
	\item Selecting and performing robotic actions to maximize map information, referred to as \emph{autonomous exploration}, is necessary to optimally explore initially-uncertain environments
	\item Cooperative robots can share information to achieve a collective goal, referred to as \emph{multi-vehicle coordination}
	\item These task must be performed in 2D and 3D environments, where computational efficiency is paramount to real-time implementation
\end{itemize}
\end{frame}

\subsection*{Research Contributions}

\begin{frame}
\frametitle{High-Level Contributions}
\begin{itemize}
	\item Occupancy Grid Mapping: solving the Bayesian solution to occupancy probability, known as the \emph{inverse sensor model}
	\pause
	\item Autonomous Exploration: novel approach to predicting future map uncertainty, known as \emph{Shannon's entropy}
	\pause
	\item Multi-Vehicle Coordination: \emph{bidding}-based framework for assigning tasks while exploring uncertain environments
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Outline}
\begin{itemize}
	\item Occupancy Grid Mapping
	\begin{itemize}
		\item Background and problem definition
		\item Common approaches and approximations
		\item Inverse sensor model in 1D
		\item Ray casting for 2D and 3D environments
		\item Numerical examples
	\end{itemize}
	\pause
	\item Autonomous Exploration
	\begin{itemize}
		\item Existing solutions: frontiers, expected measurements
		\item Proposed approach: entropy expectation
		\item Collision-free trajectories
		\item Optimal pose selection
		\item Numerical Examples
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Outline}
\begin{itemize}
	\item Multi-Vehicle Coordination
	\begin{itemize}
		\item Cooperative approaches
		\item Bidding-based framework
		\item Receding horizon
		\item Multi-vehicle autonomous exploration numerical example
		\item Autonomous patrol
		\item Multi-vehicle autonomous patrol numerical example
	\end{itemize}
	\pause
	\item Experiments
	\begin{itemize}
		\item Ground vehicle 2D mapping and autonomous exploration
		\item Flying quadrotor in large vertically-uniform space
		\item Flying quadrotor in complex 3D environment
	\end{itemize}
\end{itemize}
\end{frame}

\section*{}
\subsection*{Probabilistic Occupancy Grid Mapping}

\begin{frame}
\frametitle{Mapping Representation}
    \begin{itemize}
    	\item Simple and well-known grid-based representation
	\pause
	\item Decomposes space into evenly-spaced cells that are either \emph{occupied} or \emph{free}
	\pause
    	\item Probabilistic: goal is to find the \emph{probability} that the cells are occupied
	\pause
	\item Useful for collision-avoidance and graph searching
    \end{itemize}

    \only<1->{
\begin{figure}
\centerline{
    \includegraphics[height=2.1cm]{ogm_ex3.jpeg}\hspace*{0.1cm}
\hspace*{0.5cm}
    \includegraphics[height=2.1cm]{ogm_ex1.jpg}\hspace*{0.1cm}
\hspace*{0.5cm}
    \includegraphics[height=2.1cm]{ogm_ex2.png}\hspace*{0.1cm}
}
\end{figure}}

\end{frame}

\begin{frame}
\frametitle{Problem Definition}
%\framesubtitle{Problem Definition}

\begin{itemize}
	\item The Map and the Robot
	\begin{itemize}
	\item Map $m$ is composed of $n_m$ grid cells with known location and size
	\item The $i$-th grid cell $\mathbf{m}_i$ is a \emph{static binary} random variable, independent of other grid cells: $P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i)$
	\item \emph{Pose} $X_t$ is known, containing robot \emph{position} and \emph{attitude}
	\end{itemize}
\vspace*{0.0cm}\pause
\end{itemize}
\begin{minipage}[t]{7.0cm}
\begin{itemize}
	\item Depth Measurements
	\begin{itemize}
	\item Each measurement origin and direction is known \emph{deterministically}
	\item A measurement \emph{scan} $Z_t=\braces{z_{t,1},z_{t,2},\ldots,z_{t,n_z}}$ contains $n_z$ measurement \emph{rays} (depths)% at the $t$-th time step, and the history of measurement scans $Z_{1:t}$ is known
\item The \emph{forward sensor model} is known from the sensor properties
\end{itemize}
\end{itemize}
\end{minipage}
\begin{minipage}[t]{3.0cm}
%The \emph{forward sensor model} is the probability density distribution $p(z_{t,l}|m,X_{t})$  known from the sensor properties
\hspace*{0.25cm}
\begin{figure}[!htbp]
\vspace*{-0.25cm}
%\vspace*{0.25cm}
\centerline{
    \includegraphics[width=3.5cm]{BeamModel.png}\hspace*{0.1cm}
    }
\vspace*{0.25cm}
\centerline{
    \includegraphics[width=2.5cm]{1D_True_Grid.png}\hspace*{0.1cm}
%    \hspace*{0.75cm}
    }
{Beam Model for Range Finders}
\end{figure}

\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Problem Definition}

\begin{itemize}
	\item Bayesian Framework
	\begin{itemize}
	\item Markov Assumption: latest a priori cell occupancy probabilities capture the information from all prior observations
	\item Log-Odds Ratio: popular representation due to simple additive update structure and probability truncation avoidance, but requires the \emph{assumption}
	\begin{align*}
		P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})\approx P(Z_t|\mathbf{m}_i,X_t)
	\end{align*}
	\end{itemize}
\vspace*{0.0cm}\pause
	\item Inverse Sensor Model
	\begin{align*}
&P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\eta_{t,l}\sum_{m\in\mathcal{M}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
\end{align*}
	\begin{itemize}
	\item Given $n$ grid cells: $\mathcal O(2^n)$ is \emph{computationally intractable}, motivating a different solution
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Inexact Solutions}

\begin{minipage}[t]{7.0cm}
\begin{itemize}
	\item Approximate a function for the inverse sensor model based on intuition
	\begin{itemize}
		\item Simple to implement, but mathematically inaccurate
	\end{itemize}
\end{itemize}
\end{minipage}
\begin{minipage}[t]{3.0cm}
\begin{figure}[!htbp]
\centerline{
    \vspace*{-0.5cm}
%	\hspace*{0.25cm}
    \includegraphics[width=4.0cm]{Approx_ISM_shortened.png}\hspace*{0.1cm}
    }
\centerline{
    \includegraphics[width=3.5cm]{1D_True_Grid.png}
%    \hspace*{0.75cm}
    }
		\end{figure}
\end{minipage}
\vspace*{-0.5cm}
\vspace*{0.0cm}\pause
\begin{itemize}
	\item Find a solution through learning
	\begin{itemize}
		\item Simulate maps, poses, and measurements and use learning to obtain an inverse sensor model
		\item Complicated, unclear how these parameters are chosen
	\end{itemize}
	\vspace*{0.0cm}\pause
	\item Goal: design a \emph{simple} and \emph{accurate} occupancy grid mapping method avoiding log-odds ratio assumptions
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{New Approach: Grouping}

\begin{itemize}
    \item Main idea: make use of occupancy grid mapping \emph{assumptions} and extract \emph{patterns} from probabilistic properties to find a \emph{computationally-efficient} solution
	\begin{itemize}
		\item Since the origin and direction of each measurement ray is known deterministically, the set of grid cells that the ray intersects is \emph{known through geometry}
		\item A depth reading follows the forward sensor model, which \emph{only depends} on the first occupied grid cell along the measurement ray
	\end{itemize}
\end{itemize}
\setcounter{subfigure}{0}
% uncomment
\begin{figure}
  \centering
  \begin{subfigure}[t]{.45\linewidth}
    \centering\includegraphics[width=\linewidth]{rkplus_1.png}
  \end{subfigure}
  \begin{subfigure}[t]{.45\linewidth}
    \centering\includegraphics[width=\linewidth]{rkplus_2.png}
  \end{subfigure}
    \begin{subfigure}[t]{.45\linewidth}
    \centering\includegraphics[width=\linewidth]{rkplus_3.png}
  \end{subfigure}
  \begin{subfigure}[t]{.45\linewidth}
    \centering\includegraphics[width=\linewidth]{rkplus_4.png}
  \end{subfigure}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Bayesian Update}

\begin{itemize}
	\item Unnormalized Probability:
	\begin{align*}
\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})
&=\mathbf{P}_k^-
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]\nonumber\\
&\qquad + \bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{k+},X_t)\mathbf{P}_k^-
\end{align*}
	\item Normalizer:
	\begin{align*}
\label{eqn:allEta}
\eta
&=
\bigg[\sum_{i=1}^{n_{r}+1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\} p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]^{-1}
\end{align*}
	\item Solution: $P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})=\eta\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})$
	\item \emph{Linear Complexity} ($\mathcal{O}(n)$ for \emph{all} $n$ cells in FOV)
\end{itemize}
\end{frame}

% Ray Casting

\begin{frame}
\frametitle{Ray Casting}

\begin{itemize}
	\item Required for 2D and 3D occupancy grid maps
	\item Main idea: use simple geometry to find the distance from the robot sensor to each cell intersected by the measurement ray
\end{itemize}

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[t]{0.4\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{RayCastingIllustration.png}
%        \label{fig:penn_map_total}
    \end{subfigure}
    \hspace*{0.05\columnwidth}
    \begin{subfigure}[t]{0.4\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{RayCastIllustrationReducedMapOnly.png}
%        \label{fig:penn_map_zoom}
    \end{subfigure}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Full Measurement Scan}

\begin{itemize}
	\item Two Methods
	\begin{itemize}
		\item \emph{Ray-By-Ray Inverse Sensor Model}: each measurement ray is evaluated individually for updating the map
		\item \emph{Synergistic Scan Inverse Sensor Model}: all measurement rays are evaluated simultaneously
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Ray-By-Ray Inverse Sensor Model}

\begin{itemize}
	\item Main idea: given a scan of measurement rays, update the map based on each measurement ray \emph{individually} and \emph{sequentially}
	\item Conditional probability:
	\begin{align*}
		P(\mathbf{m}_i|X_{1:t}&,Z_{1:t})\nonumber\\&
		=P((\dots(((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z})
	\end{align*}
	\item Benefit: the measurement rays maintain some dependency as they capture the same map
	\item Drawback: the best order for evaluating rays is not necessarily chosen
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Synergistic Scan Inverse Sensor Model}

\begin{itemize}
	\item Main idea: consider every measurement ray inside a scan \emph{simultaneously} to update all grid cells inside the FOV
	\item Probability derived from a Bayesian approach:
	\begin{align*}
		P(\mathbf{m}_i|X_{1:t},Z_{1:t})
		&=\xi_i P(\mathbf{m}_i|{X_{1:t-1}},Z_{1:t-1})\nonumber\\&\quad\times
		\prod_{l\in\mathcal L_i}
		\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}),
		\\
		\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})&
		\triangleq \frac{\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})}
	\end{align*}
	\item Benefit: synergistic update provides a method where all rays are considered simultaneously
	\item Drawback: required assumption that the measurement rays are independent, though they capture the same map
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Numerical Example}

\begin{minipage}[t]{5.0cm}
\vspace*{0.5cm}
\begin{itemize}
	\item A standard \emph{approximate} inverse sensor model technique is compared with the proposed \emph{exact} method
\end{itemize}
\end{minipage}
\begin{minipage}[t]{5.0cm}
%\vspace*{0.5cm}
\begin{figure}[!htbp]
    \centering
    \begin{subfigure}{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{AISM_Image_inf_19.pdf}
        \caption*{Approximate}
    \end{subfigure}
    \hspace*{-0.1\textwidth}
    \begin{subfigure}{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{EISM_Image_inf_19.pdf}
        \caption*{Exact}
    \end{subfigure}
\end{figure}
\end{minipage}

\begin{itemize}
\item The exact algorithm yields a more \emph{certain} map from the same measurement set
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Probabilistic Occupancy Grid Mapping Conclusions}

\begin{itemize}
	\item The proposed novel Bayesian solution to the inverse sensor model provides the exact solution without approximate functions or learned solutions
	\item The computational complexity is linear with respect to the number of cells considered, making real-time implementation possible
	\item 2D and 3D maps are updated with ray casting and ray-by-ray combination
\end{itemize}

\end{frame}

\section*{}

%
%\section*{}
%
%\begin{frame}
%\frametitle{}
%%\center
%\center{\bf \color{blue} Single-Robot Autonomous Exploration}
%\end{frame}
%
%
%\subsection*{Single-Robot Autonomous Exploration}
%
%\begin{frame}
%\frametitle{Introduction}
%\begin{itemize}
%        	\item The environment is initially uncertain, so the robot path is initially unknown
%	\item Simultaneous localization and mapping assumes robotic actions are provided
%	\item Human interaction is common to guide the robot
%	\item Goal: develop effective policies to govern robotic motion 
%\end{itemize}
%
%\end{frame}
%
%
%\begin{frame}
%\frametitle{Problem Definition}
%\begin{itemize}
%        	\item Uncertainty-Based Exploration
%	\begin{itemize}
%		\item Main idea: choose robot motions that minimize the uncertainty of the occupancy grid map
%		\item Equivalently maximize map information gain
%	\end{itemize}
%\end{itemize}
%\begin{minipage}[t]{7.0cm}
%\begin{itemize}
%	\item Entropy
%	\begin{itemize}		
%		\item \emph{Shannon's entropy} serves as an uncertainty measure
%		\begin{align*}
%			H(P)=-P\log P-\bar{P}\log \bar{P}%-(1-P)\log(1-P)
%		\end{align*}
%		with cell occupancy probability $P$ and complement $\bar{P}=1-P$
%		\item Entropy $H$ is maximized when $P=0.5$ and minimized as $P\rightarrow0$ or $P\rightarrow1$
%	\end{itemize}
%\end{itemize}
%\end{minipage}
%\begin{minipage}[t]{3.0cm}
%\vspace*{0.5cm}
%\begin{figure}[!htbp]
%	\centerline{
%		\hspace*{1.25cm}
%   		\includegraphics[width=5.0cm]{H_Plotted_square.pdf}%\hspace*{0.1cm}
%	}
%\end{figure}
%\end{minipage}
%
%\end{frame}
%
%
%
%\begin{frame}
%\frametitle{Problem Definition}
%\begin{itemize}
%        	\item Map Information Gain Maximization
%	\begin{itemize}
%		\item Goal: choose a set of optimal future poses to maximize map information gain
%		\pause
%		\item Challenge: the complete set cannot be calculated accurately with an initially highly-uncertain environment
%		\pause
%		\item Solution: choose short-term optimal pose goals, reevaluating an updated map for the subsequent optimizations
%		\pause
%		\item The information gain considering candidate future pose $X_c$:% is
%		\begin{align*}
%			\mathcal I(X_c)&=H(P(m|X_{1:t},Z_{1:t}))-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right]
%		\end{align*}
%		\pause
%		\item Optimal pose $X^*_c$ satisfies
%		\begin{align*}
%			X_c^*=\argmax_{X_c}{\mathcal I(X_c)}
%		\end{align*}
%	\end{itemize}
%\end{itemize}
%
%
%\end{frame}
%
%
%
%\begin{frame}
%\frametitle{Related Work}
%\begin{itemize}
%        	\item Frontier-Based Exploration
%	\begin{itemize}
%		\item Main idea: robot moves toward boundaries between certain and uncertain space, takes measurements, then pushes back the boundaries
%		\item Non-entropy-based exploration
%		\item Heuristic
%	\end{itemize}
%	\pause
%	\item Entropy-Based Exploration
%	\begin{itemize}
%		\item Cell occupancy probabilities are inaccurate: inverse sensor model is heuristic or learned, and frequently estimated with Rao-Blackwellized particle filters
%		\item Inaccurate probabilities imply inaccurate entropy measures
%		\item Approaches use ``hallucination measurements'' to predict how future measurements \emph{might} affect the map uncertainty,
%		\begin{align*}
%			\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right]\approx H(P(m|X_{1:t},Z_{1:t},X_c,\text{E}\left[Z_c\right]))
%		\end{align*}
%	\end{itemize}
%\end{itemize}
%
%
%\end{frame}
%
%
%\begin{frame}
%\frametitle{Direct Entropy Calculation}
%\begin{itemize}
%        	\item Instead of relying on frontiers or making assumptions about measurements, the expected map entropy is \emph{calculated directly} by the law of total expectation:
%	\begin{align*}
%		\text{E}[H(P(m|X,z))]=\int_{z_\text{min}}^{z_\text{max}}
%		H(P(m|X,z))p(z|X)dz,
%	\end{align*}
%	\item The robot pose is chosen to maximize map information gain from this expectation
%	\item Dijkstra's algorithm provides collision-free waypoints for the robot path
%	\item Smooth trajectory: constrained polynomial least squares 
%\end{itemize}
%
%
%\end{frame}
%
%
%\begin{frame}
%\frametitle{Numerical Example}
%\begin{itemize}
%        	\item A benchmark example floor plan is explored and mapped using the proposed mapping and exploration algorithms
%	\item Simulations in Robot Operating System (ROS)
%\end{itemize}
%\begin{figure}
%    \centering
%    \includegraphics[width=0.4\textwidth]{intel_clean.png}
%    \caption*{Modified Intel Research Lab floor plan form the SLAM benchmark}
%\end{figure}
%
%\end{frame}
%
%%\begin{frame}
%%\frametitle{Numerical Results}
%%	\centering{
%%	\includemovie[poster,autoplay]{0.720\textwidth}{0.576\textwidth}{../../Fig/JIRS16_simulation_speedup10x_then_speedup100x.mov}}
%%\end{frame}
%
%\begin{frame}
%\frametitle{Numerical Results}
%
%\begin{figure}[!ht]
%    \centering
%    \begin{subfigure}[t]{0.2\columnwidth}
%        \centering
%        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{0min.png}
%        \caption*{$t=0$ min}
%        \label{fig:IRL0min}
%    \end{subfigure}
%    \begin{subfigure}[t]{0.2\columnwidth}
%        \centering
%        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{5min.png}
%        \caption*{$t=5$ min}
%        \label{fig:IRL5min}
%    \end{subfigure}
%    \begin{subfigure}[t]{0.2\columnwidth}
%           \centering
%           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{10min.png}
%        \caption*{$t=10$ min}
%        \label{fig:IRL10min}
%    \end{subfigure}
%    \begin{subfigure}[t]{0.2\columnwidth}
%           \centering
%           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{15min.png}
%        \caption*{$t=15$ min}
%        \label{fig:IRL15min}
%    \end{subfigure}
%    \begin{subfigure}[t]{0.2\columnwidth}
%         \centering
%         \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{20min.png}
%        \caption*{$t=20$ min}
%        \label{fig:IRL20min}
%    \end{subfigure}
%    \begin{subfigure}[t]{0.2\columnwidth}
%           \centering
%           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{25min.png}
%        \caption*{$t=25$ min}
%        \label{fig:IRL25min}
%    \end{subfigure}
%    \begin{subfigure}[t]{0.2\columnwidth}
%           \centering
%           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{30min.png}
%        \caption*{$t=30$ min}
%        \label{fig:IRL30min}
%    \end{subfigure}
%    \begin{subfigure}[t]{0.2\columnwidth}
%           \centering
%           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{35min.png}
%        \caption*{$t=35$ min}
%        \label{fig:IRL35min}
%    \end{subfigure}
%%    \begin{subfigure}[t]{0.2\columnwidth}
%%           \centering
%%           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{40min.png}
%%        \caption*{$t=40$ min}
%%        \label{fig:IRL40min}
%%    \end{subfigure}
%    \caption*{A robot (green) measures a room in ROS Stage simulator using modified Intel Research Lab floor plan using the proposed mapping and exploration algorithms.}
%    \label{fig:IRL}
%\end{figure}
%
%
%\end{frame}
%
%
%
%
%
%\begin{frame}
%\frametitle{Experimental Example}
%\begin{itemize}
%        	\item A pioneer ground vehicle explores a room composed of Styrofoam walls
%	\item Experiments use ROS, a Microsoft Kinect depth sensor, and Vicon Tracker for pose information
%\end{itemize}
%\begin{figure}
%	\centering
%    	\begin{subfigure}[b]{0.28\textwidth}
%        		\includegraphics[width=\textwidth]{pioneer.png}
%        		\caption*{Pioneer}
%    	\end{subfigure}    	
%	\hspace*{0.04\textwidth}
%	\begin{subfigure}[b]{0.28\textwidth}
%        		\includegraphics[width=\textwidth]{test_setup_1.jpg}
%        		\caption*{Bottom-left view}
%    	\end{subfigure}
%	\hspace*{0.03\textwidth}
%	\begin{subfigure}[b]{0.28\textwidth}
%        		\includegraphics[width=\textwidth]{test_setup_2.jpg}
%        		\caption*{Bottom-right view}
%    	\end{subfigure}
%\end{figure}
%
%\end{frame}
%
%%\begin{frame}
%%\frametitle{Experimental Results}
%%	\centering{
%%	\includemovie[poster,autoplay]{0.720\textwidth}{0.576\textwidth}{../../Fig/JIRS16_experiment_SideBySide_speedup8x.mov}}
%%\end{frame}
%
%\begin{frame}
%\frametitle{Experimental Results}
%
%\begin{figure}
%	\centering{
%    	\begin{subfigure}[b]{0.19\textwidth}
%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_start.jpg}
%        		\caption*{$t=0.0$min}
%        		\label{fig:Experiment_ogm_t0}
%    	\end{subfigure}
%	\begin{subfigure}[b]{0.19\textwidth}
%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_0p5min.jpg}
%        		\caption*{$t=0.5$min}
%        		\label{fig:Experiment_ogm_t0p5}
%    	\end{subfigure}    
%	\begin{subfigure}[b]{0.19\textwidth}
%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1min.jpg}
%        		\caption*{$t=1.0$min}
%        		\label{fig:Experiment_ogm_t1}
%    	\end{subfigure}
%	\begin{subfigure}[b]{0.19\textwidth}
%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1p5min.jpg}
%        		\caption*{$t=1.5$min}
%        		\label{fig:Experiment_ogm_t1p5}
%    	\end{subfigure}
%	\begin{subfigure}[b]{0.19\textwidth}
%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2min.jpg}
%        		\caption*{$t=2.0$min}
%        		\label{fig:Experiment_ogm_t2}
%    	\end{subfigure}
%		\vspace*{0.05\textwidth}
%
%	}
%	\vspace*{-1cm}
%	\centering{
%    	\begin{subfigure}[b]{0.19\textwidth}
%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2p5min.jpg}
%        		\caption*{$t=2.5$min}
%        		\label{fig:Experiment_ogm_t2p5}
%    	\end{subfigure}
%	\begin{subfigure}[b]{0.19\textwidth}
%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3min.jpg}
%        		\caption*{$t=3.0$min}
%        		\label{fig:Experiment_ogm_t3}
%    	\end{subfigure}    
%	\begin{subfigure}[b]{0.19\textwidth}
%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3p5min.jpg}
%        		\caption*{$t=3.5$min}
%        		\label{fig:Experiment_ogm_t3p5}
%    	\end{subfigure}
%	\begin{subfigure}[b]{0.19\textwidth}
%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4min.jpg}
%        		\caption*{$t=4.0$min}
%        		\label{fig:Experiment_ogm_t4}
%    	\end{subfigure}
%	\begin{subfigure}[b]{0.19\textwidth}
%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4p5min.jpg}
%        		\caption*{$t=4.5$min}
%        		\label{fig:Experiment_ogm_t4p5}
%    	\end{subfigure}
%	}
%\end{figure}
%
%\end{frame}
%
%
%
%\section*{}
%
%\begin{frame}
%\frametitle{}
%%\center
%\center{\bf \color{blue} Future Direction:\\Multi-Robot Autonomous Exploration}
%\end{frame}
%
%\subsection*{Future Direction: Multi-Robot Autonomous Exploration}
%
%\begin{frame}
%\frametitle{Proposed Topics}
%\begin{itemize}
%        	\item The ray-by-ray exact occupancy grid mapping approach in centralized probabilistic mapping
%	\item Multiple robots choosing optimal motions considering each other
%	\item Robots avoiding collisions with each other
%	\item Coverage optimization of robots from different poses
%\end{itemize}
%
%\end{frame}
%
%\begin{frame}
%\frametitle{Proposed Future Steps}
%\begin{itemize}
%        	\item Mapping:
%	\begin{itemize}
%		\item Restructure algorithms for centralized map generation
%		\item Communicate with multiple measurement sources
%	\end{itemize}
%	\item Exploration:
%	\begin{itemize}
%		\item Simplify enormous computation: bidding-based approach
%		\item Temporary theoretical map building
%		\item Collision avoidance between robots
%	\end{itemize}
%\end{itemize}
%
%\end{frame}
%
%
%
%\section*{}
%
%\begin{frame}
%\frametitle{}
%%\center
%\center{\bf \color{blue} Conclusions and Publications}
%\end{frame}
%
%
%\subsection*{Conclusions and Publications}
%
%\begin{frame}
%\frametitle{Conclusions}
%\begin{itemize}
%        	\item Proposed an occupancy grid mapping technique that uses the \emph{exact probabilistic solution}
%	\item Computational cost is reduced \emph{substantially} for \emph{real-time implementation} using probabilisitic properties and exploiting mathematical patterns
%	\item Exact mapping concepts are extended to map uncertainty to accomplish \emph{autonomous exploration}
%	\item Proposing multiple vehicles can explore an uncertain environment using a \emph{bidding}-based approach
%\end{itemize}
%\end{frame}

\begin{frame}
\frametitle{Publications}
{\tiny 
\begin{itemize}
	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Autonomous Exploration with Exact Inverse Sensor Models,'' \textit{Journal of Intelligent }\&\textit{ Robotic Systems}, 2016, submitted.
	\item E. Kaufman, T. Lee, and Z. Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, San Francisco, December 2016, accepted.
	\item E. Kaufman, T. Lee, Z. Ai, and I. S. Moskowitz, ``Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model,'' \textit{Proceedings of the American Control Conference}, Boston, July 2016, pp. 5709-5715.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{The Journal of the Astronautical Sciences}, 63(1): pp. 60-80, 2016, doi: 10.1007/s40295-015-0082-9.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Minimum Uncertainty JPDA Filters and Coalescence Avoidance for Multiple Object Tracking,'' \textit{The Journal of the Astronautical Sciences}, 63(4): pp. 308-334, 2016, doi: 10.1007/s40295-016-0092-2.
	\item T. Wu, E. Kaufman, and T. Lee, ``Globally Asymptotically Stable Attitude Observer on SO(3),'' \textit{Proceedings of the 54th IEEE Conference on Decision and Control}, pp. 2164-2168, Osaka, Japan, December 2015.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting}, Williamsburg, VA, Jan. 2015, AAS 15-451.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Minimum Uncertainty JPDA Filter and Coalescence Avoidance Performance Evaluations,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting, Williamsburg}, VA, Jan. 2015, AAS 15-432.
	\item E. Kaufman, K. Caldwell, D. Lee, and T. Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, Mar. 2014, ASC 14-2527.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Optimal Joint Probabilistic Data Association Filter Avoiding Coalescence in Close Proximity,'' \textit{Proceedings of the European Control Conference}, pp. 2709-2714, June 2014.
	\vspace*{0.5cm}
\end{itemize}
}
\end{frame}

\end{document}

### 2018-07-09 Dissertation/Thesis/tex/intro.tex
updated section 7.4 of dissertation to bring figures together with section 7.3, started defense presentation

@@
 
-The final contribution coordinates the mapping and exploration efforts of multiple robots together.
+\sloppy The final contribution coordinates the mapping and exploration efforts of multiple robots together.
 The mapping process is multi-threaded such that multiple members update the same occupancy grid.
@@
 Kaufman and T.
-Lee, ``Autonomous Aerial Exploration for Topological Mapping of Mars Environments'', \textit{2019 AIAA Guidance, Navigation, and Control Conference}, San Diego, CA, USA, January 2019, submitted.
+Lee, ``Autonomous Aerial Exploration for Topological Mapping of Mars Environments'', \textit{2019 AIAA Guidance, Navigation, and Control Conference}, San Diego, CA, January 2019, submitted.
 	\item E.
@@
 Lee, ``Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities,'' \textit{The Second IEEE International Conference on Robotic Computing}, pp.
-49--55, Laguna Hills, January 2018.
+49--55, Laguna Hills, CA, January 2018.
 	\item E.
@@
 Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, pp.
-246--251, San Francisco, CA, USA, December 2016.
+246--251, San Francisco, CA, December 2016.
 	\item E.
@@
 Moskowitz, ``Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model,'' \textit{Proceedings of the American Control Conference}, pp.
-5709-5715, Boston, MA, USA, July 2016.
+5709-5715, Boston, MA, July 2016.
 	\item E.
@@
 Lee, and T.
-Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, ASC 14-2527, March 2014, Big Sky, MT, USA.
+Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, ASC 14-2527, March 2014, Big Sky, MT.
 	\item E.
@@
 Lee, ``Optimal Joint Probabilistic Data Association Filter Avoiding Coalescence in Close Proximity,'' \textit{Proceedings of the European Control Conference}, pp.
-2709-2714, Strasbourg, France, June 2014.
+2709-2714, Strasbourg, June 2014.
 	\vspace*{0.5cm}

### 2018-07-09 Dissertation/Thesis/tex/multivehicle.tex
updated from tlee edits, more figs for NRL flight experiment
updated section 7.4 of dissertation to bring figures together with section 7.3, started defense presentation

@@
 The proposed approach uses auctions and bid modifications based on total map expected information gain and collision-avoidance, promoting exploration of different spaces without explicit consideration of coverage overlaps.
-The pseudocode for the bidding process is shown with Algorithm \ref{alg:bidding}.
+The pseudo-code for the bidding process is shown with Algorithm \ref{alg:bidding}.
 
@@
 	}
-	\caption{2D Projected Occupancy Grid Map of SEH Second Floor During Multi-Vehicle Exploration}
+	\caption{2D Projected Occupancy Grid Map During Multi-Vehicle Exploration}
 	\medskip
@@
 	}
-	\caption{2D Projected Occupancy Grid Map of SEH Second Floor During Multi-Vehicle Patrol}
+	\caption{2D Projected Occupancy Grid Map During Multi-Vehicle Patrol}
 	\medskip

### 2018-07-09 Dissertation/Thesis/tex/results.tex
updated section 7.4 of dissertation to bring figures together with section 7.3, started defense presentation

@@
 
+\section{Geometric Control}
+\label{sec:GeometricControl}
+
+The next two experiments involve a quadrotor unmanned aerial vehicle (UAV) mapping and exploring 3D environments.
+A flight controller that accounts for the nonlinear quadrotor attitude dynamics is required to track the trajectories provided by autonomous exploration.
+We choose geometric control because of its stability properties and simple form.
+
+Geometric control is designed for rigid-body attitude dynamics, and is developed directly on $\SO$ without local parameterizations such as Euler angles or quaternions.
+Therefore, this approach avoids singularities, ambiguities, and linearization errors.
+It has been developed for fully-actuated vehicles~\cite{KauCalLeeLee14} and for under-actuated quadrotors~\cite{LeeLeoMcc10,LeeLeo_4457,LeeLeo,GooDaeLee13}.
+This approach is further extended to aerial load transportation~\cite{GooLee14,GooLee16} and attitude estimation~\cite{WuKauLee15}.
+
+In the context of quadrotor control, the main idea is that a translational position-integral-derivative (PID) control produces a desired force in 3D, and the robot selects a desired attitude $R_d$ and angular velocity $\Omega_d$ to align its propellers with this force and to correct its first body-fixed axis to a desired direction.
+The attitude and angular velocity errors from the current robot attitude $R$ and angular velocity $\Omega$ are
+\begin{align}
+e_R&=(R_d^\T R-R^\T R_d)^{\mathsf{V}},
+\\
+e_\Omega&=\Omega-R^\T R_d\Omega_d,
+\end{align}
+respectively, where the \emph{vee} map is defined in~\cite{LeeLeo}. The quadrotor is capable of producing a force $f\in\Re$ and moment $M\in\Re^3$,
+\begin{align}
+\label{eqn:geometricf}
+f&=(k_xe_x+k_ve_v+mge_3-m\ddot{x}_d)\cdot Re_3,
+\\
+\label{eqn:geometricM}
+M&=-k_Re_R-k_\Omega e_\Omega+\Omega\times J\Omega-J(\hat{\Omega}R^\T R_d\Omega_d-R^\T R_d\dot\Omega_d),
+\end{align}
+where $e_x$ and $e_v$ are position and velocity errors, respectively, $k_x$, $k_v$, $k_R$, $k_\Omega$ are positive gain constants, $\ddot{x}_d$ is the desired translational acceleration, $\dot{\Omega}_d$ is the desired rotational acceleration, $m$ is the quadrotor mass, $J$ is the quadrotor moment of inertia, $e_3=[0, 0, 1]^\T$, and the \emph{hat} map is defined in~\cite{LeeLeo}.
+Since \refeqn{geometricf} and \refeqn{geometricM} require only trivial calculations, they are easily computed with onboard modules.
+
 \section{Aerial Exploration of a Large Space}
@@
 The host computer returns the exploration trajectories to the Jetson TX2 over WIFI.
-Then, the Jetson executes a nonlinear geometric flight controller~\cite{GooDaeLee13} onboard.
+Then, the Jetson executes the geometric flight controller onboard.
 With this configuration, large processing tasks are avoided onboard the robot.
@@
 
+%\newpage
+
+%\let\oldsection\section
+%\renewcommand\section{\clearpage\oldsection}
+
+\sectionbreak{\clearpage}
 \section{Full 3D Exploration with a Quadrotor}
@@
 The robot generates a 3D map of this space, and performs autonomous exploration while considering the expected information gains of sample measurement rays with nonzero vertical components.
-The robot unlike prior examples, the robot also moves vertically during exploration and tracks the ground below the vehicle.
+Unlike prior examples, the robot also moves vertically during exploration and tracks the ground below the vehicle.
 
 \subsection{3D Exploration Setup and Parameters}
-The software follows the same structure as in Section \ref{sec:MarsSimulation}, except the Gazebo simulated environment is replaces with experimental equipment.
-This equipment is identical to the prior experimental example (Section \ref{sec:QuadrotorNRL}), except the experiment took place at the Flight Dynamics and Control Lab (FDCL) inside the second floor of SEH.
-This environment is smaller with only $6$ cameras, where the x-direction (positive east) spans $-2.0$ m to $2.0$ m, the y-direction (positive north) spans $-2.0$ m to $2.0$ m, and the z-direction (positive vertically up) spans $-0.15$ m to $1.5$ m.
+The software follows the same structure as in Section \ref{sec:MarsSimulation}, except the Gazebo simulated environment is replaced with experimental equipment.
+This equipment is identical to the prior experimental example (Section \ref{sec:QuadrotorNRL}), except the experiment takes place at the Flight Dynamics and Control Lab (FDCL) inside the second floor of SEH, and the Hokuyo is no longer in use.
+This environment is smaller with only $6$ cameras, where the x-direction (positive east) spans $-4.5$ m to $4.5$ m, the y-direction (positive north) spans $-4.0$ m to $4.5$ m, and the z-direction (positive vertically up) spans $-0.25$ m to $2.0$ m, shown in Figure \ref{fig:Full3DExplorationEnvironment}.
 
@@
 \end{figure}
+
+\subsection{3D Exploration Results}
+
+The experimental results show that the robot successfully explores a 3D space, capturing occupied space above and below the robot.
+The quadrotor does an initial rotation over $20$ sec, then explores the space with planned motions in all three dimensions, shown in Figure \ref{fig:exp3DVid}.
+The robot generates a 3D probabilistic map (Figure \ref{fig:exp3DOccMap}) and uses this for entropy prediction and collision-free motion planning with a 3D cost map (Figure \ref{fig:Full3DExplorationCostmaps}).
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{H_explore3D_test1.pdf}
-	\caption{Entropy Change During Full 3D Experiment}
-	\medskip
-	\small
-	The map entropy $H$ consistently decreases during the full 3D exploration experiment.
-Toward the end of the experiment, most cells that can be captured by the Asus Xtion have been updated, slowing down the rate of entropy decrease.
-	\label{fig:Explore3D_H}
-\end{figure}
-
 \begin{figure}[!t]
@@
 	\small
-	Two examples of the 3D cost maps from Dijkstra's search show connected and disconnected regions at the exploration height, where more blue regions have small distance costs, red regions have larger distance costs, and pink regions are unreachable due to collision risk.
-In (a), the robot does not have to change altitude to reach any of the collision-free poses.
-However in (b), the robot can only reach collision-free poses in the disconnected section by changing altitude.
+	Two examples of the 3D cost maps from Dijkstra's search show connected and disconnected regions at the exploration height, where more blue regions have smaller distance costs, red regions have larger distance costs, and pink regions are unreachable due to collision risk.
+In (a), the robot does not have to change altitude to reach any of the collision-free poses at its current altitude.
+However in (b), the robot can only reach collision-free poses in the disconnected section by changing altitude and following the 3D cost map.
 	\label{fig:Full3DExplorationCostmaps}
@@
 
-\subsection{3D Exploration Results}
+The proposed 3D occupancy grid mapping and autonomous exploration approaches complement each other well.
+The floor and objects are well-captured, as this detail-oriented entropy-based exploration scheme tends to generate completely mapped regions before moving to new terrain, shown with Case 1 of the Mars exploration numerical example (Section \ref{sec:MarsSimulation}).
+As a result, the entropy decreases fairly steadily over the experiment, shown in Figure \ref{fig:Explore3D_H}.
+
+A key improvement to prior experiments is the ability to consider the occupancy properties in full 3D.
+This allows the robot to understand that the mannequin occupies some space that the robot must avoid.
+However, the robot also recognizes that the space above the mannequin is free, and it chooses a trajectory above the mannequin to reach new terrain.
+At other times, the robot observes spaces from lower vantage points, which can be advantageous with a short sensor range and the forward direction of the Asus Xtion.
+These added capabilities are attributed to the full 3D mapping and exploration approach.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{H_explore3D_test1.pdf}
+	\caption{Entropy Change During Full 3D Experiment}
+	\medskip
+	\small
+	The map entropy $H$ consistently decreases during the full 3D exploration experiment.
+Toward the end of the experiment, most cells that can be captured by the Asus Xtion have been updated, slowing down the rate of entropy decrease.
+	\label{fig:Explore3D_H}
+\end{figure}

### 2018-07-11 Dissertation/Defense/defense.tex
added recent figs to dissertation, minor edits to defense

@@
 \institute{\footnotesize
-{\normalsize Evan Kaufman\\Research Advisor: Taeyoung Lee}\vspace*{0.1cm}\\
-  Mechanical and Aerospace Engineering\\ George Washington University \vspace*{0.3cm}\\ {\normalsize Special Thanks to:\\
+{\normalsize Evan Kaufman\\
 \vspace*{0.2cm}
- Kuya Takami, Mahdis Bisheban, and Kanishke Gamagedara} \\Department of Mechanical Engineering\\George Washington University\\
+Dissertation Director: Taeyoung Lee}\\
+\vspace*{0.3cm}
+{\normalsize Committee Members:\\
  \vspace*{0.2cm}
-{\normalsize Zhuming Ai, Ira. S. Moskowitz, and Mark Livingston}\\Information Management \& Decision Architectures\\U.S. Naval Research Laboratory}
+Taeyoung Lee, James Hahn, Robert Pless,\\Chung-Hyuk Park, and Zhuming Ai}
+\vspace*{0.4cm}\\
+  Mechanical and Aerospace Engineering\\George Washington University
+}
+
+%\institute{\footnotesize
+%{\normalsize Evan Kaufman\\Research Advisor: Taeyoung Lee}\vspace*{0.1cm}\\
+%  Mechanical and Aerospace Engineering\\ George Washington University \vspace*{0.3cm}\\ {\normalsize Special Thanks to:\\
+% \vspace*{0.2cm}
+% Kuya Takami, Mahdis Bisheban, and Kanishke Gamagedara} \\Department of Mechanical Engineering\\George Washington University\\
+%\vspace*{0.2cm}
+%{\normalsize Zhuming Ai, Ira. S. Moskowitz, and Mark Livingston}\\Information Management \& Decision Architectures\\U.S. Naval Research Laboratory}
 
@@
 		\item Inverse sensor model in 1D
-		\item Ray casting for 2D and 3D environments
+		\item Ray casting and measurement scans for 2D and 3D environments
 		\item Numerical examples

### 2018-07-11 Dissertation/Thesis/tex/results.tex
added recent figs to dissertation, minor edits to defense

@@
 However, the robot also recognizes that the space above the mannequin is free, and it chooses a trajectory above the mannequin to reach new terrain.
-At other times, the robot observes spaces from lower vantage points, which can be advantageous with a short sensor range and the forward direction of the Asus Xtion.
+At other times, the robot observes spaces from lower vantage points, which can be advantageous with a short sensor range and the forward direction of the Asus Xtion relative to the robot.
 These added capabilities are attributed to the full 3D mapping and exploration approach.
@@
 \end{figure}
+
+The controller is also highly important for following autonomous exploration trajectories accurately.
+The geometric controller tracks the desired trajectory to avoid collsions with obstacles in the room.
+The plot of 3D position is shown in Figure \ref{fig:Explore3D_x3D} and the desired commands and measured outputs of the controlled system are shown in Figures \ref{fig:Explore3D_x_xd}, \ref{fig:Explore3D_v_vd}, \ref{fig:Explore3D_R_Rd}, and \ref{fig:Explore3D_W_Wd}, and the total propeller force $f$ from \refeqn{geometricf} and moment $M$ from \refeqn{geometricM} are shown in Figure \ref{fig:Explore3D_fM}.
+An interesting aspect is that at $1.31$ min, the robot temporarily obtains a poor estimate of the robot, but since this is short-lived, the robot quickly recovers and continues exploring.
+In summary, geometric control successfully tracks the desired commands from autonomous exploration through uncertain environments.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{explore3D_x3D.pdf}
+	\caption{Robot Position over Full 3D Experiment}
+	\medskip
+	\small
+	The history of robot position is shown in 3D with respect to a north-east-down world-fixed frame, such that there are all negative $x_3$ values, making the trajectory appear upside-down.
+The low locations (high relative to the ground) on the left of the figure ($x_1\approx-1$ m, $x_2\approx0.5$ m, and $x_3\approx-1.8$ m) correspond to flying over the mannequin.
+	\label{fig:Explore3D_x3D}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{explore3D_x_xd.pdf}
+	\caption{Geometric Control Position Tracking over Full 3D Experiment}
+	\medskip
+	\small
+	The position of the robot (blue solid) tracks the desired position (red dashed) over the experiment.
+	\label{fig:Explore3D_x_xd}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{explore3D_v_vd.pdf}
+	\caption{Geometric Control Velocity Tracking over Full 3D Experiment}
+	\medskip
+	\small
+	The velocity of the robot (blue solid) tracks the desired velocity (red dashed) over the experiment.
+	\label{fig:Explore3D_v_vd}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{explore3D_R_Rd.pdf}
+	\caption{Geometric Control Attitude Tracking over Full 3D Experiment}
+	\medskip
+	\small
+	The attitude of the robot (blue solid), shown with the rotation matrix between the north-east-down frame and the body-fixed frame, tracks the desired attitude (red dashed) over the experiment.
+The bottom-right plot shows the third body-fixed axis is mostly-aligned with the third world-fixed axis, which corresponds to maintaining level flight.
+	\label{fig:Explore3D_R_Rd}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{explore3D_W_Wd.pdf}
+	\caption{Geometric Control Angular Velocity Tracking over Full 3D Experiment}
+	\medskip
+	\small
+	The angular velocity of the robot (blue solid) tracks the desired angular velocity (red dashed) over the experiment.
+The overall larger $\Omega_3$ corresponds to yaw rotation to change the direction of the Xtion depth sensor.
+	\label{fig:Explore3D_W_Wd}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{explore3D_fM.pdf}
+	\caption{Geometric Control Force and Moments over Full 3D Experiment}
+	\medskip
+	\small
+	The forces and moments provided by the geometric control are shown over the experiment.
+The values of $f$ remain largely-steady to cancel gravitational forces.
+Changes in $M_1$ and $M_2$ correspond to roll and pitch movements to move around horizontally.
+Changes in $M_3$ correspond to correcting yaw for the Xtion direction.
+	\label{fig:Explore3D_fM}
+\end{figure}

### 2018-07-11 Dissertation/Defense/defense.tex
updates to defense

@@
 
-\usepackage{movie15}
+%\usepackage{movie15}
+\usepackage{media9}
 
@@
 \section*{}
+\subsection*{Entropy-Based Autonomous Exploration}
+
+\begin{frame}
+\frametitle{Autonomous Exploration}
+
+\begin{itemize}
+	\item Goal: determine robotic actions that maximize map information
+	\item Solution: predict future map uncertainty and travel distance to select an optimal pose
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Existing Exploration Approaches}
+
+\begin{itemize}
+	\item Frontier-Based Exploration
+	\begin{itemize}
+		\item A robot identifies boundaries of free and uncertain cells
+		\item The robot moves toward these frontiers, thereby pushing back the boundaries
+		\item Heuristic, suboptimal
+	\end{itemize}
+	\vspace*{0.0cm}\pause
+	\item Entropy-Based Approaches
+	\begin{itemize}
+		\item Probabilities from the occupancy grid map are approximate
+		\item Usage of ``hallucination measurements'': assume that $\text{E}[H(P(\mathbf{m}_i|z))]\approx H(P(\mathbf{m}_i|\text{E}[z]))$
+	\end{itemize}
+	\vspace*{0.0cm}\pause
+	\item Proposed Approach
+	\begin{itemize}
+		\item Use occupancy grid map probabilities
+		\item Solve $\text{E}[H(P)]$ and select future poses to maximize map information
+	\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Entropy Definition}
+\begin{itemize}
+        	\item Uncertainty-Based Exploration
+	\begin{itemize}
+		\item Main idea: choose robot motions that minimize the uncertainty of the occupancy grid map
+		\item Equivalently maximize map information gain
+	\end{itemize}
+\end{itemize}
+\begin{minipage}[t]{7.0cm}
+\begin{itemize}
+	\item Entropy
+	\begin{itemize}		
+		\item \emph{Shannon's entropy} serves as an uncertainty measure
+		\begin{align*}
+			H(P)=-P\log P-\bar{P}\log \bar{P}%-(1-P)\log(1-P)
+		\end{align*}
+		with cell occupancy probability $P$ and complement $\bar{P}=1-P$
+		\item Entropy $H$ is maximized when $P=0.5$ and minimized as $P\rightarrow0$ or $P\rightarrow1$
+	\end{itemize}
+\end{itemize}
+\end{minipage}
+\begin{minipage}[t]{3.0cm}
+\vspace*{0.5cm}
+\begin{figure}[!htbp]
+	\centerline{
+		\hspace*{1.25cm}
+   		\includegraphics[width=5.0cm]{H_Plotted_square.pdf}%\hspace*{0.1cm}
+	}
+\end{figure}
+\end{minipage}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Expected Information Gain}
+
+\begin{itemize}
+	\item Main Idea: determine the benefit of future poses based on how their measurements are expected to improve the map
+	\vspace*{0.0cm}\pause
+	\item Goal: decrease total map entropy:
+	\begin{align*}
+		H(P(m))&=\sum_{i=1}^{n_m}H(P(\mathbf{m}_i))
+	\end{align*}
+	\vspace*{0.0cm}\pause
+	\item Measurement Ray Expected Information Gain:
+	\begin{align*}
+		\text{E}[H(P(m|x_c,z_{c}))]&=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(m|x_c,z_{c,k}))P(z_{c,k}|x_c)\bigg\}
+		\\
+		P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}}
+		\\
+		\mathcal I(x_c,z_{c})&=H(P(m))-\text{E}[H(P(m|x_c,z_{c}))]
+	\end{align*}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Expected Information Gain Computation}
+
+\begin{itemize}
+	\item \emph{Squared Complexity}: $\mathcal O(n_r^2)$ for $n_r$ cells along the ray
+	\vspace*{0.0cm}\pause
+	\item Two Approaches:
+	\begin{itemize}
+		\vspace*{0.0cm}\pause
+		\item Decrease the cells considered by selecting $n_p$ most probable cells: $\mathcal O(n_r^2)$ reduces to $\mathcal O(n_p^2)$
+		\vspace*{0.0cm}\pause
+		\item Use 3D cells in a 2D \emph{projection}
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{.3\linewidth}
+    \centering\includegraphics[height=.75\linewidth]{experiment_ogm3D_2min47sec.jpg}
+    \caption*{\ \ \ \ \ \ \ \ \ \ \ \ \ \ 3D Map}
+  \end{subfigure}
+  \hspace*{0.25\linewidth}
+  \begin{subfigure}[t]{.3\linewidth}
+    \centering\includegraphics[height=.75\linewidth]{experiment_2min47sec.jpg}
+    \caption*{2D Projection}
+  \end{subfigure}
+\end{figure}
+	\end{itemize}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Pose Selection}
+
+\begin{itemize}
+	\item Main Idea: select the future pose that maximizes map information gain while accounting for travel distance
+	\vspace*{0.0cm}\pause
+	\item Attitude Selection: choose the attitude that covers the candidate measurement rays with largest objective function summation
+\begin{figure}
+\centerline{
+	\includegraphics[height=0.35\linewidth]{ExampleOptimalPose.png}
+%\begin{picture}(0,0)(0,0)
+%\setlength{\unitlength}{0.1\linewidth}\scriptsize
+%\put(4.4,2.1){\color{blue}$X_t$}
+%\put(5.9,1.9){\color{red}$x_1$}
+%\put(2.9,2.0){\color{red}$x_5$}
+%\put(4.4,3.6){\color{red}$x_3$}
+%\put(4.4,0.6){\color{red}$x_7$}
+%\put(5.4,3.0){\color{red}$x_2$}
+%\put(3.3,3.1){\color{red}$x_4$}
+%\put(3.3,1.0){\color{red}$x_6$}
+%\put(5.4,1.0){\color{red}$x_8$}
+%\put(6.8,3.1){\color{red}$z_{2,1}$}
+%\put(4.5,3.1){\color{red}$z_{2,5}$}
+%\put(5.6,4.1){\color{red}$z_{2,3}$}
+%\put(5.6,2.2){\color{red}$z_{2,7}$}
+%\put(6.5,3.7){\color{red}$z_{2,2}$}
+%\put(4.9,3.8){\color{red}$z_{2,4}$}
+%\put(5.0,2.5){\color{red}$z_{2,6}$}
+%\put(6.5,2.5){\color{red}$z_{2,8}$}
+%\put(6.1,2.9){$X_c^*$}
+%\end{picture}
+}
+\vspace*{-0.05\linewidth}
+\end{figure}
+	\item The attitude $R_c$ is selected for all candidates: the expected information gain becomes $\mathcal I(X_c)$ where $X_c=\braces{x_c,R_c}$
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Pose Selection}
+
+\begin{itemize}
+	\item Travel Distance: generate a cost map from the occupancy grid to obtain distances to each candidate pose $d(x_c)$
+	\vspace*{0.0cm}\pause
+	\item Bump Function: use a bump function \begin{align*}f(d)=\exp\braces{-\frac{d^2}{2\sigma^2}}+\gamma, \quad \sigma>0, \quad \gamma>0\end{align*} to incentivize the robot to explore more locally until nearby spaces are well-known
+	\begin{figure}
+	\vspace*{-0.015\linewidth}
+\centering
+	\includegraphics[width=0.8\linewidth]{GaussianBumpFunFlat.png}
+		\vspace*{-0.015\linewidth}
+\end{figure}
+	% C++: exp(-0.5*distAway*distAway/sigSqrd)+minFunVal GaussianBumpFunFlat.png
+	\vspace*{0.0cm}\pause
+	\item Select the pose that maximizes the product \begin{align*}X^*=\argmax_{X_c}{\mathcal I(X_c)f(d(x_c))}\end{align*}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Collision-Free Path}
+
+\begin{itemize}
+	\item Using the cost map from before, \emph{Dijkstra's algorithm} is completed by finding the waypoints from the desired candidate back to the current location
+	\vspace*{0.0cm}\pause
+	\item These waypoints serve as input to a \emph{constrained polynomial least squares trajectory}
+	\begin{itemize}
+		\item Each trajectory is solved independently as a function of time, i.e. $x(t)$, $y(t)$
+		\item Starting and terminal locations are fixed
+		\item Segments patched together have the same position and velocity at the connection
+	\end{itemize}
+	\vspace*{0.0cm}\pause
+	\item A \emph{geometric controller} tracks the trajectory without \emph{singularities} or \emph{ambiguities}
+	\vspace*{0.0cm}\pause
+	\item The mapping, exploration, and control run in \emph{real-time}
+\end{itemize}
+
+\end{frame}
+
+% TODO: uncomment!
+%\begin{frame}
+%	\frametitle{Numerical Results}
+%	\centering{
+%		\includemedia[
+%  		width=0.720\textwidth,
+%  		height=0.576\textwidth,
+%  		activate=pageopen,
+%  		addresource=IRL.mp4,
+%  		flashvars={source=IRL.mp4}
+%		]{}{VPlayer.swf}
+%		}
+%\end{frame}
+
+%\frametitle{Experimental Results}
+%	\centering{
+%	\includemovie[poster,autoplay]{0.720\textwidth}{0.576\textwidth}{JIRS16_experiment_SideBySide_speedup8x.mov}}
+%\end{frame}
+
+\begin{frame}
+\frametitle{Autonomous Exploration Conclusions}
+
+\begin{itemize}
+	\item Other approaches use heuristic frontiers or entropy of expected measurements
+	\item The proposed approach solves for the expected value of future map entropy directly
+	\item Several sample future measurement rays and Dijkstra's cost map combine in an objective function for optimality
+	\item Numerical simulations show a large 2D environment, and this approach can be extended to 3D environments
+\end{itemize}
+
+\end{frame}
+
+\section*{}
+\subsection*{Multi-Vehicle Coordination}
+
+\section*{}
+\subsection*{Experiments}
 
@@
 %\end{frame}
+
+% TODO: update with dissertation version!
 

### 2018-07-12 Dissertation/Defense/defense.tex
added exploration parts, still need patrol, 2D & 3D NRL tests (mostly copy), and 3D exploration (new)

@@
 
+\begin{frame}
+\frametitle{Motivation and Goal}
+
+\begin{itemize}
+	\item Multiple agents can accomplish tasks faster and more reliably than a single agent
+	\item Large groups are capable of coverage in large environments
+	\item Problem: multi-vehicle map information gain optimality can be computationally intractable
+	\item Solution: bidding-based solution simplifies a complicated optimization into a series of simplified optimizations
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Related Work}
+
+\begin{itemize}
+	\item Most existing bidding-based techniques deal with auctioning tasks among a few agents locally in decentralized frameworks
+	\item The other approach is for a central executive to run a series of auctions
+	\item Bids are discounted between auctions to coordinate the efforts based on coverage overlap
+	\item Heuristic, ad hoc
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Proposed Approach}
+
+\begin{itemize}
+\item Central executive completes $n$ auctions for $n$ robots
+\begin{enumerate}
+	\item Each robot determines its optimal future pose, and bids the objective function
+	\item The winner is awarded this future pose, and the remaining robots update their bids according to
+	\begin{itemize}
+		\item New expected map changes from the winner
+		\item Collision-avoidance with the winner
+	\end{itemize}
+	\item Repeat these steps with the remaining robots
+\end{enumerate}
+\item The efforts are coordinated by updating bids between auctions
+\end{itemize}
+\end{frame}
+
+% TODO: uncomment!
+%\begin{frame}
+%\frametitle{Numerical Example: SEH Second Floor}
+%
+%
+%	\centering{
+%		\includemedia[
+%  		width=0.720\textwidth,
+%  		height=0.576\textwidth,
+%  		activate=pageopen,
+%  		addresource=Multi-Vehicle Mapping and Exploration.mp4,
+%  		flashvars={source=Multi-Vehicle Mapping and Exploration.mp4}
+%		]{}{VPlayer.swf}
+%		}
+%
+%
+%\end{frame}
+
 \section*{}
@@
 \begin{frame}
-\frametitle{Publications}
+\frametitle{Publications on Dissertation Topics}
 {\tiny 
 \begin{itemize}
-	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Autonomous Exploration with Exact Inverse Sensor Models,'' \textit{Journal of Intelligent }\&\textit{ Robotic Systems}, 2016, submitted.
+	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Multi-Vehicle Cooperative Exploration and Patrol of Uncertain 3D Environments'', in preparation.
+	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Bidding-Based Autonomous Exploration of Cooperative Robots Building a 3D Probabilistic Occupancy Grid Map'', in preparation.
+	\item E.
+Kaufman and T.
+Lee, ``Autonomous Aerial Exploration for Topological Mapping of Mars Environments'', \textit{2019 AIAA Guidance, Navigation, and Control Conference}, San Diego, CA, January 2019, submitted.
+	\item E.
+Kaufman, K.
+Takami, Z.
+Ai, and T.
+Lee, ``Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities,'' \textit{The Second IEEE International Conference on Robotic Computing}, pp.
+49--55, Laguna Hills, CA, January 2018.
+	\item E.
+Kaufman, K.
+Takami, T.
+Lee, and Z.
+Ai, ``Autonomous Exploration with Exact Inverse Sensor Models,'' \textit{Journal of Intelligent }\&\textit{ Robotic Systems}, 2017, doi: 10.1007/s10846-017-0710-7.
 	\item E.
@@
 Lee, and Z.
-Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, San Francisco, December 2016, accepted.
+Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, pp.
+246--251, San Francisco, CA, December 2016.
 	\item E.
@@
 S.
-Moskowitz, ``Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model,'' \textit{Proceedings of the American Control Conference}, Boston, July 2016, pp.
-5709-5715.
+Moskowitz, ``Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model,'' \textit{Proceedings of the American Control Conference}, pp.
+5709-5715, Boston, MA, July 2016.
+\end{itemize}
+}
+\end{frame}
+
+\begin{frame}
+\frametitle{Publications on Data Association, Estimation, and Control}
+{\tiny 
+\begin{itemize}
 	\item E.
@@
 Lee, and T.
-Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, Mar.
-2014, ASC 14-2527.
+Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, ASC 14-2527, March 2014, Big Sky, MT.
 	\item E.
@@
 Lee, ``Optimal Joint Probabilistic Data Association Filter Avoiding Coalescence in Close Proximity,'' \textit{Proceedings of the European Control Conference}, pp.
-2709-2714, June 2014.
-	\vspace*{0.5cm}
+2709-2714, Strasbourg, June 2014.
 \end{itemize}

### 2018-07-12 Dissertation/Defense/defense.tex
did patrol, started experiments, need to go back and add ogm of hallway & Mars figs

@@
 	\begin{itemize}
-	\item Given $n$ grid cells: $\mathcal O(2^n)$ is \emph{computationally intractable}, motivating a different solution
+	\item Given $n$ grid cells: $\mathcal O(2^n)$ is \emph{computationally-intractable}, motivating a different solution
 	\end{itemize}
@@
 	\item Large groups are capable of coverage in large environments
-	\item Problem: multi-vehicle map information gain optimality can be computationally intractable
+	\item Problem: multi-vehicle map information gain optimality can be computationally-intractable
 	\item Solution: bidding-based solution simplifies a complicated optimization into a series of simplified optimizations
@@
 \item The efforts are coordinated by updating bids between auctions
+\item The algorithm follows a receding-horizon framework: future pose optimization repeats as quickly as possible
 \end{itemize}
@@
 
+\begin{frame}
+\frametitle{Cooperative Autonomous Patrol}
+
+\begin{itemize}
+	\item Existing Approaches
+	\begin{itemize}
+		\item Rely on a known map and use Voronoi diagrams to disperse
+		\item Another approach introduces cell periodicity: observe long-term dynamics of objects at the same location on the occupancy grid
+		\item Bio-inspired approach uses digital pheromones 
+		\item Minimizing robot idleness
+	\end{itemize}
+	\item Proposed Approach
+	\begin{itemize}
+		\item Degrade occupancy grid cells toward their initial values with exponential decay
+		\item Only affects existing probability values without new variables to track
+		\item Promotes exploration first, then patrol
+	\end{itemize}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Continuous-Time Markov Process}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{markov_diagram_continuous.pdf}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{DegradeExamples.pdf}
+\end{figure}
+
+\end{frame}
+
+% TODO: uncomment!
+%\begin{frame}
+%\frametitle{Numerical Example: SEH Second Floor}
+%
+%
+%	\centering{
+%		\includemedia[
+%  		width=0.720\textwidth,
+%  		height=0.576\textwidth,
+%  		activate=pageopen,
+%  		addresource=Multi-Vehicle Cooperative Patrol.mp4,
+%  		flashvars={source=Multi-Vehicle Cooperative Patrol.mp4}
+%		]{}{VPlayer.swf}
+%		}
+%
+%
+%\end{frame}
+
+\begin{frame}
+\frametitle{Multi-Vehicle Autonomous Exploration and Patrol Conclusions}
+
+\begin{itemize}
+	\item Robotic tasks are assigned following a bidding-based framework to avoid a computationally-intractable solution
+	\item Only one robot wins an auction, and bids are updated for the remaining robots in subsequent auctions
+	\item A receding-horizon framework optimizes the time to update the optimal future poses
+	\item Cell degradation provides a simple means for autonomous patrol via a continuous-time Markov process
+	\item Numerical examples of a large experimental space demonstrate both algorithms
+\end{itemize}
+\end{frame}
+
 \section*{}
 \subsection*{Experiments}
+
+\begin{frame}
+\frametitle{Experimental Overview}
+
+\begin{enumerate}
+	\item Ground vehicle experiment for 2D occupancy grid mapping and autonomous exploration
+	\item Flying vehicle experiment in large 3D space resembling an office
+	\item Flying vehicle experiment in a complex 3D space
+\end{enumerate}
+\end{frame}
 

### 2018-07-12 Dissertation/Thesis/tex/ae2D.tex
increased refs from 42 to 61 excluding geometric control ones, which will be added later
did patrol, started experiments, need to go back and add ogm of hallway & Mars figs

@@
 There are two important steps to moving a robot on a collision-free trajectory to its optimal pose.
-First, Dijkstra's algorithm~\cite{Dij59} provides the cost map and optimal collision-free waypoints.
+First, Dijkstra's algorithm~\cite{Dij59,Dij11} provides the cost map and optimal collision-free waypoints.
 Then, a constrained polynomial least squares trajectory serves as the desired path, which is followed by a controller.
@@
 	\includegraphics[width=0.4\textwidth]{intel_clean.png}
-	\caption{Intel Research Lab Floor Plan Benchmark~\cite{kummerle2009measuring}.}
+	\caption{Intel Research Lab Floor Plan Benchmark}
 	\medskip
 	\small
-	The 2D floor plan from the Intel Research Lab is simplified to eliminate small objects. The remaining black pixels are used to produce walls for 2D simulations.
+	The 2D floor plan from the Intel Research Lab~\cite{kummerle2009measuring} is simplified to eliminate small objects. The remaining black pixels are used to produce walls for 2D simulations.
 \label{fig:intel}

### 2018-07-12 Dissertation/Thesis/tex/multivehicle.tex
did patrol, started experiments, need to go back and add ogm of hallway & Mars figs

@@
 \includegraphics[width=\textwidth]{markov_diagram_continuous.pdf}
-\caption{The continuous-time Markov process for a single binary cell occupancy is illustrated. As time progresses, this process degrades the cell slowly toward $P_\infty.$}
+\caption{Continuous-Time Markov Process Illustration}
+\medskip
+\small
+The continuous-time Markov process for a single binary cell occupancy is illustrated. As time progresses, this process degrades the cell slowly toward $P_\infty.$
 \label{fig:MarkovDegradeContinuous}

### 2018-07-12 Dissertation/Defense/defense.tex
finished defense exxcept for minor issues

@@
 %\begin{frame}
-%\frametitle{Numerical Example: SEH Second Floor}
+%\frametitle{Numerical Example: SEH Second Floor Multi-Vehicle Autonomous Exploration}
 %
@@
 %\begin{frame}
-%\frametitle{Numerical Example: SEH Second Floor}
+%\frametitle{Numerical Example: SEH Second Floor Multi-Vehicle Autonomous Patrol}
 %
@@
 
-%
-%\section*{}
-%
+\begin{frame}
+\frametitle{Pioneer Ground Vehicle 2D Example}
+\begin{itemize}
+        	\item A pioneer ground vehicle explores a room composed of Styrofoam walls
+	\item Experiments use ROS, a Microsoft Kinect depth sensor, and Vicon Tracker for pose information
+\end{itemize}
+\begin{figure}
+	\centering
+    	\begin{subfigure}[b]{0.28\textwidth}
+        		\includegraphics[width=\textwidth]{pioneer.png}
+        		\caption*{Pioneer}
+    	\end{subfigure}    	
+	\hspace*{0.04\textwidth}
+	\begin{subfigure}[b]{0.28\textwidth}
+        		\includegraphics[width=\textwidth]{test_setup_1.jpg}
+        		\caption*{Bottom-left view}
+    	\end{subfigure}
+	\hspace*{0.03\textwidth}
+	\begin{subfigure}[b]{0.28\textwidth}
+        		\includegraphics[width=\textwidth]{test_setup_2.jpg}
+        		\caption*{Bottom-right view}
+    	\end{subfigure}
+\end{figure}
+
+\end{frame}
+
+% TODO: uncomment!
 %\begin{frame}
-%\frametitle{}
-%%\center
-%\center{\bf \color{blue} Single-Robot Autonomous Exploration}
+%\frametitle{Ground Experiment Video}
+%
+%
+%	\centering{
+%		\includemedia[
+%  		width=0.720\textwidth,
+%  		height=0.576\textwidth,
+%  		activate=pageopen,
+%  		addresource=NRL_ground_test.mp4,
+%  		flashvars={source=NRL_ground_test.mp4}
+%		]{}{VPlayer.swf}
+%		}
+%
+%
 %\end{frame}
-%
-%
-%\subsection*{Single-Robot Autonomous Exploration}
-%
+
+\begin{frame}
+\frametitle{Ground Vehicle Experimental Results}
+
+\begin{figure}
+	\centering{
+    	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_start.jpg}
+        		\caption*{$t=0.0$min}
+        		\label{fig:Experiment_ogm_t0}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_0p5min.jpg}
+        		\caption*{$t=0.5$min}
+        		\label{fig:Experiment_ogm_t0p5}
+    	\end{subfigure}    
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1min.jpg}
+        		\caption*{$t=1.0$min}
+        		\label{fig:Experiment_ogm_t1}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1p5min.jpg}
+        		\caption*{$t=1.5$min}
+        		\label{fig:Experiment_ogm_t1p5}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2min.jpg}
+        		\caption*{$t=2.0$min}
+        		\label{fig:Experiment_ogm_t2}
+    	\end{subfigure}
+		\vspace*{0.05\textwidth}
+
+	}
+	\vspace*{-1cm}
+	\centering{
+    	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2p5min.jpg}
+        		\caption*{$t=2.5$min}
+        		\label{fig:Experiment_ogm_t2p5}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3min.jpg}
+        		\caption*{$t=3.0$min}
+        		\label{fig:Experiment_ogm_t3}
+    	\end{subfigure}    
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3p5min.jpg}
+        		\caption*{$t=3.5$min}
+        		\label{fig:Experiment_ogm_t3p5}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4min.jpg}
+        		\caption*{$t=4.0$min}
+        		\label{fig:Experiment_ogm_t4}
+    	\end{subfigure}
+	\begin{subfigure}[b]{0.19\textwidth}
+        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4p5min.jpg}
+        		\caption*{$t=4.5$min}
+        		\label{fig:Experiment_ogm_t4p5}
+    	\end{subfigure}
+	}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Flying Vehicle in a Large Office-Like Environment}
+\begin{itemize}
+        	\item Hardware
+	\begin{itemize}
+		\item NVIDIA Jetson Ubuntu computer-on-module
+		\item Asus Xtion IR color depth sensor
+		\item Hokuyo LIDAR
+		\item VectorNav IMU
+	\end{itemize}
+	\item Software
+	\begin{itemize}
+		\item Robot Operating System (ROS)
+		\item Vicon Tracker (mocap\_vicon)
+		\item Geometric Control: coordinate-free attitude control
+	\end{itemize}
+\end{itemize}
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{.4\linewidth}
+    \centering\includegraphics[height=.5\linewidth]{quad_top.png}
+    \caption*{Quadrotor Top}
+  \end{subfigure}
+  \begin{subfigure}[t]{.4\linewidth}
+    \centering\includegraphics[height=.5\linewidth]{quad_bottom.png}
+    \caption*{Quadrotor Bottom}
+  \end{subfigure}
+\end{figure}
+\end{frame}
+
+% TODO: uncomment!
 %\begin{frame}
-%\frametitle{Introduction}
-%\begin{itemize}
-%        	\item The environment is initially uncertain, so the robot path is initially unknown
-%	\item Simultaneous localization and mapping assumes robotic actions are provided
-%	\item Human interaction is common to guide the robot
-%	\item Goal: develop effective policies to govern robotic motion 
-%\end{itemize}
+%\frametitle{Large Flight Test Experiment Video}
+%
+%
+%	\centering{
+%		\includemedia[
+%  		width=0.720\textwidth,
+%  		height=0.576\textwidth,
+%  		activate=pageopen,
+%  		addresource=nrl_experiment_narrated.mp4,
+%  		flashvars={source=nrl_experiment_narrated.mp4}
+%		]{}{VPlayer.swf}
+%		}
+%
 %
 %\end{frame}
-%
-%
+
+\begin{frame}
+\frametitle{Flying Vehicle in a Complex 3D Environment}
+\begin{itemize}
+        	\item Full 3D mapping and exploration
+	\item Objects are not vertically-uniform
+\end{itemize}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{.4\linewidth}
+    \centering\includegraphics[height=.7\linewidth]{explore3D_space_NW.JPG}
+    \caption*{Northwest View}
+  \end{subfigure}
+  \begin{subfigure}[t]{.4\linewidth}
+    \centering\includegraphics[height=.7\linewidth]{explore3D_space_NE.JPG}
+    \caption*{Northeast View}
+  \end{subfigure}
+\end{figure}
+\end{frame}
+
+% TODO: uncomment! and update!
 %\begin{frame}
-%\frametitle{Problem Definition}
-%\begin{itemize}
-%        	\item Uncertainty-Based Exploration
-%	\begin{itemize}
-%		\item Main idea: choose robot motions that minimize the uncertainty of the occupancy grid map
-%		\item Equivalently maximize map information gain
-%	\end{itemize}
-%\end{itemize}
-%\begin{minipage}[t]{7.0cm}
-%\begin{itemize}
-%	\item Entropy
-%	\begin{itemize}		
-%		\item \emph{Shannon's entropy} serves as an uncertainty measure
-%		\begin{align*}
-%			H(P)=-P\log P-\bar{P}\log \bar{P}%-(1-P)\log(1-P)
-%		\end{align*}
-%		with cell occupancy probability $P$ and complement $\bar{P}=1-P$
-%		\item Entropy $H$ is maximized when $P=0.5$ and minimized as $P\rightarrow0$ or $P\rightarrow1$
-%	\end{itemize}
-%\end{itemize}
-%\end{minipage}
-%\begin{minipage}[t]{3.0cm}
-%\vspace*{0.5cm}
-%\begin{figure}[!htbp]
-%	\centerline{
-%		\hspace*{1.25cm}
-%   		\includegraphics[width=5.0cm]{H_Plotted_square.pdf}%\hspace*{0.1cm}
+%\frametitle{Full 3D Flying Experiment Video}
+%
+%
+%	\centering{
+%		\includemedia[
+%  		width=0.720\textwidth,
+%  		height=0.576\textwidth,
+%  		activate=pageopen,
+%  		addresource=gwu_experiment_narrated.mp4,
+%  		flashvars={source=gwu_experiment_narrated.mp4}
+%		]{}{VPlayer.swf}
 %		}
-%\end{figure}
-%\end{minipage}
+%
 %
 %\end{frame}
-%
-%
-%
-%\begin{frame}
-%\frametitle{Problem Definition}
-%\begin{itemize}
-%        	\item Map Information Gain Maximization
-%	\begin{itemize}
-%		\item Goal: choose a set of optimal future poses to maximize map information gain
-%		\pause
-%		\item Challenge: the complete set cannot be calculated accurately with an initially highly-uncertain environment
-%		\pause
-%		\item Solution: choose short-term optimal pose goals, reevaluating an updated map for the subsequent optimizations
-%		\pause
-%		\item The information gain considering candidate future pose $X_c$:% is
-%		\begin{align*}
-%			\mathcal I(X_c)&=H(P(m|X_{1:t},Z_{1:t}))-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right]
-%		\end{align*}
-%		\pause
-%		\item Optimal pose $X^*_c$ satisfies
-%		\begin{align*}
-%			X_c^*=\argmax_{X_c}{\mathcal I(X_c)}
-%		\end{align*}
-%	\end{itemize}
-%\end{itemize}
-%
-%
-%\end{frame}
-%
-%
-%
-%\begin{frame}
-%\frametitle{Related Work}
-%\begin{itemize}
-%        	\item Frontier-Based Exploration
-%	\begin{itemize}
-%		\item Main idea: robot moves toward boundaries between certain and uncertain space, takes measurements, then pushes back the boundaries
-%		\item Non-entropy-based exploration
-%		\item Heuristic
-%	\end{itemize}
-%	\pause
-%	\item Entropy-Based Exploration
-%	\begin{itemize}
-%		\item Cell occupancy probabilities are inaccurate: inverse sensor model is heuristic or learned, and frequently estimated with Rao-Blackwellized particle filters
-%		\item Inaccurate probabilities imply inaccurate entropy measures
-%		\item Approaches use ``hallucination measurements'' to predict how future measurements \emph{might} affect the map uncertainty,
-%		\begin{align*}
-%			\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right]\approx H(P(m|X_{1:t},Z_{1:t},X_c,\text{E}\left[Z_c\right]))
-%		\end{align*}
-%	\end{itemize}
-%\end{itemize}
-%
-%
-%\end{frame}
-%
-%
-%\begin{frame}
-%\frametitle{Direct Entropy Calculation}
-%\begin{itemize}
-%        	\item Instead of relying on frontiers or making assumptions about measurements, the expected map entropy is \emph{calculated directly} by the law of total expectation:
-%	\begin{align*}
-%		\text{E}[H(P(m|X,z))]=\int_{z_\text{min}}^{z_\text{max}}
-%		H(P(m|X,z))p(z|X)dz,
-%	\end{align*}
-%	\item The robot pose is chosen to maximize map information gain from this expectation
-%	\item Dijkstra's algorithm provides collision-free waypoints for the robot path
-%	\item Smooth trajectory: constrained polynomial least squares 
-%\end{itemize}
-%
-%
-%\end{frame}
-%
-%
-%\begin{frame}
-%\frametitle{Numerical Example}
-%\begin{itemize}
-%        	\item A benchmark example floor plan is explored and mapped using the proposed mapping and exploration algorithms
-%	\item Simulations in Robot Operating System (ROS)
-%\end{itemize}
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=0.4\textwidth]{intel_clean.png}
-%    \caption*{Modified Intel Research Lab floor plan form the SLAM benchmark}
-%\end{figure}
-%
-%\end{frame}
-%
-%%\begin{frame}
-%%\frametitle{Numerical Results}
-%%	\centering{
-%%	\includemovie[poster,autoplay]{0.720\textwidth}{0.576\textwidth}{../../Fig/JIRS16_simulation_speedup10x_then_speedup100x.mov}}
-%%\end{frame}
-%
-%\begin{frame}
-%\frametitle{Numerical Results}
-%
-%\begin{figure}[!ht]
-%    \centering
-%    \begin{subfigure}[t]{0.2\columnwidth}
-%        \centering
-%        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{0min.png}
-%        \caption*{$t=0$ min}
-%        \label{fig:IRL0min}
-%    \end{subfigure}
-%    \begin{subfigure}[t]{0.2\columnwidth}
-%        \centering
-%        \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{5min.png}
-%        \caption*{$t=5$ min}
-%        \label{fig:IRL5min}
-%    \end{subfigure}
-%    \begin{subfigure}[t]{0.2\columnwidth}
-%           \centering
-%           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{10min.png}
-%        \caption*{$t=10$ min}
-%        \label{fig:IRL10min}
-%    \end{subfigure}
-%    \begin{subfigure}[t]{0.2\columnwidth}
-%           \centering
-%           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{15min.png}
-%        \caption*{$t=15$ min}
-%        \label{fig:IRL15min}
-%    \end{subfigure}
-%    \begin{subfigure}[t]{0.2\columnwidth}
-%         \centering
-%         \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{20min.png}
-%        \caption*{$t=20$ min}
-%        \label{fig:IRL20min}
-%    \end{subfigure}
-%    \begin{subfigure}[t]{0.2\columnwidth}
-%           \centering
-%           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{25min.png}
-%        \caption*{$t=25$ min}
-%        \label{fig:IRL25min}
-%    \end{subfigure}
-%    \begin{subfigure}[t]{0.2\columnwidth}
-%           \centering
-%           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{30min.png}
-%        \caption*{$t=30$ min}
-%        \label{fig:IRL30min}
-%    \end{subfigure}
-%    \begin{subfigure}[t]{0.2\columnwidth}
-%           \centering
-%           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{35min.png}
-%        \caption*{$t=35$ min}
-%        \label{fig:IRL35min}
-%    \end{subfigure}
-%%    \begin{subfigure}[t]{0.2\columnwidth}
-%%           \centering
-%%           \includegraphics[trim = {4.6cm 3.8cm 4.6cm 0}, clip, width=\textwidth]{40min.png}
-%%        \caption*{$t=40$ min}
-%%        \label{fig:IRL40min}
-%%    \end{subfigure}
-%    \caption*{A robot (green) measures a room in ROS Stage simulator using modified Intel Research Lab floor plan using the proposed mapping and exploration algorithms.}
-%    \label{fig:IRL}
-%\end{figure}
-%
-%
-%\end{frame}
-%
-%
-%
-%
-%
-%\begin{frame}
-%\frametitle{Experimental Example}
-%\begin{itemize}
-%        	\item A pioneer ground vehicle explores a room composed of Styrofoam walls
-%	\item Experiments use ROS, a Microsoft Kinect depth sensor, and Vicon Tracker for pose information
-%\end{itemize}
-%\begin{figure}
-%	\centering
-%    	\begin{subfigure}[b]{0.28\textwidth}
-%        		\includegraphics[width=\textwidth]{pioneer.png}
-%        		\caption*{Pioneer}
-%    	\end{subfigure}    	
-%	\hspace*{0.04\textwidth}
-%	\begin{subfigure}[b]{0.28\textwidth}
-%        		\includegraphics[width=\textwidth]{test_setup_1.jpg}
-%        		\caption*{Bottom-left view}
-%    	\end{subfigure}
-%	\hspace*{0.03\textwidth}
-%	\begin{subfigure}[b]{0.28\textwidth}
-%        		\includegraphics[width=\textwidth]{test_setup_2.jpg}
-%        		\caption*{Bottom-right view}
-%    	\end{subfigure}
-%\end{figure}
-%
-%\end{frame}
-%
-%%\begin{frame}
-%%\frametitle{Experimental Results}
-%%	\centering{
-%%	\includemovie[poster,autoplay]{0.720\textwidth}{0.576\textwidth}{../../Fig/JIRS16_experiment_SideBySide_speedup8x.mov}}
-%%\end{frame}
-%
-%\begin{frame}
-%\frametitle{Experimental Results}
-%
-%\begin{figure}
-%	\centering{
-%    	\begin{subfigure}[b]{0.19\textwidth}
-%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_start.jpg}
-%        		\caption*{$t=0.0$min}
-%        		\label{fig:Experiment_ogm_t0}
-%    	\end{subfigure}
-%	\begin{subfigure}[b]{0.19\textwidth}
-%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_0p5min.jpg}
-%        		\caption*{$t=0.5$min}
-%        		\label{fig:Experiment_ogm_t0p5}
-%    	\end{subfigure}    
-%	\begin{subfigure}[b]{0.19\textwidth}
-%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1min.jpg}
-%        		\caption*{$t=1.0$min}
-%        		\label{fig:Experiment_ogm_t1}
-%    	\end{subfigure}
-%	\begin{subfigure}[b]{0.19\textwidth}
-%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1p5min.jpg}
-%        		\caption*{$t=1.5$min}
-%        		\label{fig:Experiment_ogm_t1p5}
-%    	\end{subfigure}
-%	\begin{subfigure}[b]{0.19\textwidth}
-%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2min.jpg}
-%        		\caption*{$t=2.0$min}
-%        		\label{fig:Experiment_ogm_t2}
-%    	\end{subfigure}
-%		\vspace*{0.05\textwidth}
-%
-%	}
-%	\vspace*{-1cm}
-%	\centering{
-%    	\begin{subfigure}[b]{0.19\textwidth}
-%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2p5min.jpg}
-%        		\caption*{$t=2.5$min}
-%        		\label{fig:Experiment_ogm_t2p5}
-%    	\end{subfigure}
-%	\begin{subfigure}[b]{0.19\textwidth}
-%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3min.jpg}
-%        		\caption*{$t=3.0$min}
-%        		\label{fig:Experiment_ogm_t3}
-%    	\end{subfigure}    
-%	\begin{subfigure}[b]{0.19\textwidth}
-%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3p5min.jpg}
-%        		\caption*{$t=3.5$min}
-%        		\label{fig:Experiment_ogm_t3p5}
-%    	\end{subfigure}
-%	\begin{subfigure}[b]{0.19\textwidth}
-%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4min.jpg}
-%        		\caption*{$t=4.0$min}
-%        		\label{fig:Experiment_ogm_t4}
-%    	\end{subfigure}
-%	\begin{subfigure}[b]{0.19\textwidth}
-%        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4p5min.jpg}
-%        		\caption*{$t=4.5$min}
-%        		\label{fig:Experiment_ogm_t4p5}
-%    	\end{subfigure}
-%	}
-%\end{figure}
-%
-%\end{frame}
-%
-%
-%
-%\section*{}
-%
-%\begin{frame}
-%\frametitle{}
-%%\center
-%\center{\bf \color{blue} Future Direction:\\Multi-Robot Autonomous Exploration}
-%\end{frame}
-%
-%\subsection*{Future Direction: Multi-Robot Autonomous Exploration}
-%
-%\begin{frame}
-%\frametitle{Proposed Topics}
-%\begin{itemize}
-%        	\item The ray-by-ray exact occupancy grid mapping approach in centralized probabilistic mapping
-%	\item Multiple robots choosing optimal motions considering each other
-%	\item Robots avoiding collisions with each other
-%	\item Coverage optimization of robots from different poses
-%\end{itemize}
-%
-%\end{frame}
-%
-%\begin{frame}
-%\frametitle{Proposed Future Steps}
-%\begin{itemize}
-%        	\item Mapping:
-%	\begin{itemize}
-%		\item Restructure algorithms for centralized map generation
-%		\item Communicate with multiple measurement sources
-%	\end{itemize}
-%	\item Exploration:
-%	\begin{itemize}
-%		\item Simplify enormous computation: bidding-based approach
-%		\item Temporary theoretical map building
-%		\item Collision avoidance between robots
-%	\end{itemize}
-%\end{itemize}
-%
-%\end{frame}
-%
-%
-%
-%\section*{}
-%
-%\begin{frame}
-%\frametitle{}
-%%\center
-%\center{\bf \color{blue} Conclusions and Publications}
-%\end{frame}
-%
-%
-%\subsection*{Conclusions and Publications}
-%
-%\begin{frame}
-%\frametitle{Conclusions}
-%\begin{itemize}
-%        	\item Proposed an occupancy grid mapping technique that uses the \emph{exact probabilistic solution}
-%	\item Computational cost is reduced \emph{substantially} for \emph{real-time implementation} using probabilisitic properties and exploiting mathematical patterns
-%	\item Exact mapping concepts are extended to map uncertainty to accomplish \emph{autonomous exploration}
-%	\item Proposing multiple vehicles can explore an uncertain environment using a \emph{bidding}-based approach
-%\end{itemize}
-%\end{frame}
+
+\begin{frame}
+\frametitle{Experiment Conclusions}
+\begin{itemize}
+        	\item Experimental environments in 2D and 3D are modeled with occupancy grids
+	\item Future entropy prediction and collision-free regions use these occupancy probabilities
+	\item A geometric controller is used during flight testing to follow the collision-free trajectory
+	\item All algorithms are run in real-time without expensive post-processing 
+\end{itemize}
+\end{frame}
+
+\section*{}
+\subsection*{Conclusions and Publications}
+
+\begin{frame}
+\frametitle{Conclusions}
+\begin{itemize}
+        	\item Proposed an occupancy grid mapping technique that uses the \emph{exact probabilistic solution}
+	\item Computational cost is reduced \emph{substantially} for \emph{real-time implementation} using probabilisitic properties and exploiting mathematical patterns
+	\item Exact mapping concepts are extended to map uncertainty to accomplish \emph{autonomous exploration}
+	\item Multiple vehicles can explore and patrol an uncertain environment using a \emph{bidding}-based approach
+	\item These are demonstrated with numerous simulations and experiments
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Future Directions}
+	\begin{itemize}
+		\item Localization: use probabilistic map and other onboard sensors to localize, thereby completing the SLAM problem with the novel mapping approach
+		\item Computational improvements: determine collision-free costs more efficiently than Dijkstra's cost map
+		\item Decentralize the algorithms so processes may be distributed among multiple agents
+	\end{itemize}
+\end{frame}
 

### 2018-07-12 Dissertation/Defense/defense.tex
done except for putting in updated videos (3)
completed with videos, need tlee edits applied still

@@
 
-\begin{minipage}[t]{5.0cm}
-\vspace*{0.5cm}
 \begin{itemize}
 	\item A standard \emph{approximate} inverse sensor model technique is compared with the proposed \emph{exact} method
-\end{itemize}
-\end{minipage}
-\begin{minipage}[t]{5.0cm}
+	\item The exact algorithm yields a more \emph{certain} map from the same measurement set
+\end{itemize}
+
 %\vspace*{0.5cm}
@@
     \centering
-    \begin{subfigure}{0.5\textwidth}
+    \begin{subfigure}{0.2\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{AISM_Image_19.pdf}
+        \caption*{Approx. Map}
+    \end{subfigure}
+%    \hspace*{-0.1\textwidth}
+    \begin{subfigure}{0.2\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{EISM_Image_19.pdf}
+        \caption*{Exact Map}
+    \end{subfigure}
+    \begin{subfigure}{0.2\textwidth}
         \centering
         \includegraphics[width=\textwidth]{AISM_Image_inf_19.pdf}
-        \caption*{Approximate}
-    \end{subfigure}
-    \hspace*{-0.1\textwidth}
-    \begin{subfigure}{0.5\textwidth}
+        \caption*{Approx. $H$}
+    \end{subfigure}
+%    \hspace*{-0.1\textwidth}
+    \begin{subfigure}{0.2\textwidth}
         \centering
         \includegraphics[width=\textwidth]{EISM_Image_inf_19.pdf}
-        \caption*{Exact}
-    \end{subfigure}
-\end{figure}
-\end{minipage}
-
-\begin{itemize}
-\item The exact algorithm yields a more \emph{certain} map from the same measurement set
-\end{itemize}
+        \caption*{Exact $H$}
+    \end{subfigure}
+\end{figure}
 
@@
 
-% TODO: uncomment!
-%\begin{frame}
-%	\frametitle{Numerical Results}
-%	\centering{
-%		\includemedia[
-%  		width=0.720\textwidth,
-%  		height=0.576\textwidth,
-%  		activate=pageopen,
-%  		addresource=IRL.mp4,
-%  		flashvars={source=IRL.mp4}
-%		]{}{VPlayer.swf}
-%		}
-%\end{frame}
-
-%\frametitle{Experimental Results}
-%	\centering{
-%	\includemovie[poster,autoplay]{0.720\textwidth}{0.576\textwidth}{JIRS16_experiment_SideBySide_speedup8x.mov}}
-%\end{frame}
+\begin{frame}
+	\frametitle{2D Autonomous Exploration: Intel Research Lab}
+	\centering{
+		\includemedia[
+  		width=0.720\textwidth,
+  		height=0.576\textwidth,
+  		activate=pageopen,
+  		addresource=IRL.mp4,
+  		flashvars={source=IRL.mp4}
+		]{}{VPlayer.swf}
+		}
+\end{frame}
+
+\begin{frame}
+	\frametitle{3D Autonomous Exploration: Mars}
+	\centering{
+		\includemedia[
+  		width=0.720\textwidth,
+  		height=0.576\textwidth,
+  		activate=pageopen,
+  		addresource=Topological Mars Mapping and Autonomous Exploration in 3D.mp4,
+  		flashvars={source=Topological Mars Mapping and Autonomous Exploration in 3D.mp4}
+		]{}{VPlayer.swf}
+		}
+\end{frame}
 
@@
 
-% TODO: uncomment!
-%\begin{frame}
-%\frametitle{Numerical Example: SEH Second Floor Multi-Vehicle Autonomous Exploration}
-%
-%
-%	\centering{
-%		\includemedia[
-%  		width=0.720\textwidth,
-%  		height=0.576\textwidth,
-%  		activate=pageopen,
-%  		addresource=Multi-Vehicle Mapping and Exploration.mp4,
-%  		flashvars={source=Multi-Vehicle Mapping and Exploration.mp4}
-%		]{}{VPlayer.swf}
-%		}
-%
-%
-%\end{frame}
+\begin{frame}
+\frametitle{Numerical Example: SEH Second Floor Multi-Vehicle Autonomous Exploration}
+
+	\centering{
+		\includemedia[
+  		width=0.720\textwidth,
+  		height=0.576\textwidth,
+  		activate=pageopen,
+  		addresource=Multi-Vehicle Mapping and Exploration.mp4,
+  		flashvars={source=Multi-Vehicle Mapping and Exploration.mp4}
+		]{}{VPlayer.swf}
+		}
+
+\end{frame}
 
@@
 
-% TODO: uncomment!
-%\begin{frame}
-%\frametitle{Numerical Example: SEH Second Floor Multi-Vehicle Autonomous Patrol}
-%
-%
-%	\centering{
-%		\includemedia[
-%  		width=0.720\textwidth,
-%  		height=0.576\textwidth,
-%  		activate=pageopen,
-%  		addresource=Multi-Vehicle Cooperative Patrol.mp4,
-%  		flashvars={source=Multi-Vehicle Cooperative Patrol.mp4}
-%		]{}{VPlayer.swf}
-%		}
-%
-%
-%\end{frame}
+\begin{frame}
+\frametitle{Numerical Example: SEH Second Floor Multi-Vehicle Autonomous Patrol}
+
+	\centering{
+		\includemedia[
+  		width=0.720\textwidth,
+  		height=0.576\textwidth,
+  		activate=pageopen,
+  		addresource=Multi-Vehicle Cooperative Patrol.mp4,
+  		flashvars={source=Multi-Vehicle Cooperative Patrol.mp4}
+		]{}{VPlayer.swf}
+		}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Cell Degradation and Patrol}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.65\textwidth]{entropy_patrol_9minSwitch.pdf}
+\end{figure}
+
+\begin{itemize}
+	\item Entropy of the map increases if patrol stops at $9$ minutes if cell degradation continues
+\end{itemize}
+
+\end{frame}
 
@@
 
-% TODO: uncomment!
-%\begin{frame}
-%\frametitle{Ground Experiment Video}
-%
-%
-%	\centering{
-%		\includemedia[
-%  		width=0.720\textwidth,
-%  		height=0.576\textwidth,
-%  		activate=pageopen,
-%  		addresource=NRL_ground_test.mp4,
-%  		flashvars={source=NRL_ground_test.mp4}
-%		]{}{VPlayer.swf}
-%		}
-%
-%
-%\end{frame}
+\begin{frame}
+\frametitle{Ground Experiment Video}
+
+	\centering{
+		\includemedia[
+  		width=0.720\textwidth,
+  		height=0.576\textwidth,
+  		activate=pageopen,
+  		addresource=NRL_ground_test.mp4,
+  		flashvars={source=NRL_ground_test.mp4}
+		]{}{VPlayer.swf}
+		}
+
+\end{frame}
 
@@
 
-% TODO: uncomment!
-%\begin{frame}
-%\frametitle{Large Flight Test Experiment Video}
-%
-%
-%	\centering{
-%		\includemedia[
-%  		width=0.720\textwidth,
-%  		height=0.576\textwidth,
-%  		activate=pageopen,
-%  		addresource=nrl_experiment_narrated.mp4,
-%  		flashvars={source=nrl_experiment_narrated.mp4}
-%		]{}{VPlayer.swf}
-%		}
-%
-%
-%\end{frame}
+\begin{frame}
+\frametitle{Large Flight Test Experiment Video}
+
+	\centering{
+		\includemedia[
+  		width=0.720\textwidth,
+  		height=0.576\textwidth,
+  		activate=pageopen,
+  		addresource=NRL_3D_large_space.mp4,
+  		flashvars={source=NRL_3D_large_space.mp4}
+		]{}{VPlayer.swf}
+		}
+
+\end{frame}
 
@@
 
-% TODO: uncomment! and update!
-%\begin{frame}
-%\frametitle{Full 3D Flying Experiment Video}
-%
-%
-%	\centering{
-%		\includemedia[
-%  		width=0.720\textwidth,
-%  		height=0.576\textwidth,
-%  		activate=pageopen,
-%  		addresource=gwu_experiment_narrated.mp4,
-%  		flashvars={source=gwu_experiment_narrated.mp4}
-%		]{}{VPlayer.swf}
-%		}
-%
-%
-%\end{frame}
+\begin{frame}
+\frametitle{Full 3D Flying Experiment Video}
+
+	\centering{
+		\includemedia[
+  		width=0.720\textwidth,
+  		height=0.576\textwidth,
+  		activate=pageopen,
+  		addresource=explore3D_SEH.mp4,
+  		flashvars={source=explore3D_SEH.mp4}
+		]{}{VPlayer.swf}
+		}
+
+\end{frame}
 
@@
 \end{frame}
-
-% TODO: update with dissertation version!
 

### 2018-07-13 Dissertation/Defense/defense.tex
Acknowleged the NRL

@@
 \institute{\footnotesize
-{\normalsize Evan Kaufman\\
-\vspace*{0.2cm}
-Dissertation Director: Taeyoung Lee}\\
+{\normalsize Evan Kaufman}\\
 \vspace*{0.3cm}
-{\normalsize Committee Members:\\
- \vspace*{0.2cm}
-Taeyoung Lee, James Hahn, Robert Pless,\\Chung-Hyuk Park, and Zhuming Ai}
-\vspace*{0.4cm}\\
 Mechanical and Aerospace Engineering\\George Washington University
@@
 \begin{itemize}
+	\item Mobile Robotics
+	\begin{itemize}
 		\item Environments can be dangerous and uncertain
 		\item Robots can replace humans for tasks such as surveillance, search-and-rescue, and cleaning
+	\end{itemize}
+	\pause
+	\item Robotic Capabilities
+	\begin{itemize}
 		\item Robot must understand their environment by generating a map, referred to as \emph{mapping}
 		\item Selecting and performing robotic actions to maximize map information, referred to as \emph{autonomous exploration}, is necessary to optimally explore initially-uncertain environments
-	\item Cooperative robots can share information to achieve a collective goal, referred to as \emph{multi-vehicle coordination}
-	\item These task must be performed in 2D and 3D environments, where computational efficiency is paramount to real-time implementation
+		\item Cooperative robots can share information to achieve a collective goal with \emph{multi-vehicle exploration and patrol}
+		\item These tasks must be performed in 2D and 3D environments, where computational efficiency is paramount to real-time implementation
+	\end{itemize}
 \end{itemize}
@@
 \begin{frame}
-\frametitle{High-Level Contributions}
-\begin{itemize}
-	\item Occupancy Grid Mapping: solving the Bayesian solution to occupancy probability, known as the \emph{inverse sensor model}
+\frametitle{Research Contributions}
+\begin{itemize}
+	\item Occupancy Grid Mapping
+		\begin{itemize}
+		\item Computation of the exact occupancy probability
+		\item Incorporation of prior knowledge and new measurements
+		\item Bayesian information fusion
+	\end{itemize}
 	\pause
-	\item Autonomous Exploration: novel approach to predicting future map uncertainty, known as \emph{Shannon's entropy}
+	\item Autonomous Exploration
+	\begin{itemize}
+		\item Novel approach to predicting future map uncertainty
+		\item Optimal planning to maximize information gain
+	\end{itemize}
 	\pause
-	\item Multi-Vehicle Coordination: \emph{bidding}-based framework for assigning tasks while exploring uncertain environments
+	\item Multi-Vehicle Exploration and Patrol
+	\begin{itemize}
+		\item Bidding-based framework for assigning tasks while exploring uncertain environments
+		\item Map degradation to encourage revisiting regions from the past 
+	\end{itemize}
 \end{itemize}
@@
 \frametitle{Outline}
-\begin{itemize}
+\begin{enumerate}[I]
 	\item Occupancy Grid Mapping
@@
 		\item Optimal pose selection
-		\item Numerical Examples
-	\end{itemize}
-\end{itemize}
+		\item Numerical examples
+	\end{itemize}
+\end{enumerate}
 \end{frame}
@@
 \frametitle{Outline}
-\begin{itemize}
-	\item Multi-Vehicle Coordination
+\begin{enumerate}[I]
+\addtocounter{enumi}{2}
+	\item Multi-Vehicle Exploration and Patrol
 	\begin{itemize}
@@
 		\item Ground vehicle 2D mapping and autonomous exploration
-		\item Flying quadrotor in large vertically-uniform space
-		\item Flying quadrotor in complex 3D environment
-	\end{itemize}
-\end{itemize}
+		\item Aerial quadrotor in large vertically-uniform space
+		\item Aerial quadrotor in complex 3D environment
+	\end{itemize}
+\end{enumerate}
 \end{frame}
@@
 \subsection*{Probabilistic Occupancy Grid Mapping}
+
+\begin{frame}
+\frametitle{Probabilistic Occupancy Grid Mapping}
+\begin{enumerate}[I]
+	\item Probabilistic Occupancy Grid Mapping
+\end{enumerate}
+\end{frame}
 
@@
 \begin{frame}
-\frametitle{Inexact Solutions}
+\frametitle{Approximate Solutions}
 
@@
 \begin{frame}
-\frametitle{New Approach: Grouping}
+\frametitle{Proposed Approach: Utilize Invariance}
 
@@
 	\item Solution: $P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})=\eta\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})$
-	\item \emph{Linear Complexity} ($\mathcal{O}(n)$ for \emph{all} $n$ cells in FOV)
+	\item \emph{Linear Complexity}: $\mathcal{O}(n)$ for \emph{all} $n$ cells in FOV compared with $\mathcal O(n\times2^n)$ without using invariance
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Full Measurement Scan}
-
-\begin{itemize}
-	\item Two Methods
-	\begin{itemize}
-		\item \emph{Ray-By-Ray Inverse Sensor Model}: each measurement ray is evaluated individually for updating the map
-		\item \emph{Synergistic Scan Inverse Sensor Model}: all measurement rays are evaluated simultaneously
-	\end{itemize}
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Ray-By-Ray Inverse Sensor Model}
-
-\begin{itemize}
-	\item Main idea: given a scan of measurement rays, update the map based on each measurement ray \emph{individually} and \emph{sequentially}
-	\item Conditional probability:
-	\begin{align*}
-		P(\mathbf{m}_i|X_{1:t}&,Z_{1:t})\nonumber\\&
-		=P((\dots(((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z})
-	\end{align*}
+\frametitle{Scan Combination Approaches}
+
+\begin{itemize}
+	\item \emph{Ray-By-Ray Update}: given a scan of measurement rays, update the map based on each measurement ray \emph{individually} and \emph{sequentially}
+	\begin{itemize}
 		\item Benefit: the measurement rays maintain some dependency as they capture the same map
@@
 	\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Synergistic Scan Inverse Sensor Model}
-
-\begin{itemize}
-	\item Main idea: consider every measurement ray inside a scan \emph{simultaneously} to update all grid cells inside the FOV
-	\item Probability derived from a Bayesian approach:
-	\begin{align*}
-		P(\mathbf{m}_i|X_{1:t},Z_{1:t})
-		&=\xi_i P(\mathbf{m}_i|{X_{1:t-1}},Z_{1:t-1})\nonumber\\&\quad\times
-		\prod_{l\in\mathcal L_i}
-		\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}),
-		\\
-		\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})&
-		\triangleq \frac{\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})}
-	\end{align*}
+%	\item Conditional probability:
+%	\begin{align*}
+%		P(\mathbf{m}_i|X_{1:t}&,Z_{1:t})\nonumber\\&
+%		=P((\dots(((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z})
+%	\end{align*}
+	\item \emph{Synergistic Update}: consider every measurement ray inside a scan \emph{simultaneously} to update all grid cells inside the FOV
+	\begin{itemize}
 		\item Benefit: synergistic update provides a method where all rays are considered simultaneously
@@
 	\end{itemize}
+\end{itemize}
 
@@
 \begin{frame}
-\frametitle{Probabilistic Occupancy Grid Mapping Conclusions}
+\frametitle{Probabilistic Occupancy Grid Mapping Summary}
 
@@
 \begin{frame}
+\frametitle{Entropy-Based Autonomous Exploration}
+\begin{enumerate}[I]
+\addtocounter{enumi}{1}
+	\item Entropy-Based Autonomous Exploration
+\end{enumerate}
+\end{frame}
+
+\begin{frame}
 \frametitle{Autonomous Exploration}
@@
 	\item Goal: determine robotic actions that maximize map information
-	\item Solution: predict future map uncertainty and travel distance to select an optimal pose
+	\item Approach: predict future map uncertainty and travel distance to select an optimal pose
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Autonomous Exploration Conclusions}
+\frametitle{Autonomous Exploration Summary}
 
@@
 \section*{}
-\subsection*{Multi-Vehicle Coordination}
+\subsection*{Multi-Vehicle Exploration and Patrol}
+
+\begin{frame}
+\frametitle{Multi-Vehicle Exploration and Patrol}
+\begin{enumerate}[I]
+\addtocounter{enumi}{2}
+	\item Multi-Vehicle Exploration and Patrol
+\end{enumerate}
+\end{frame}
 
@@
 
+\begin{itemize}
+	\item Multi-Vehicle Autonomous Exploration
 	\begin{itemize}
@@
 		\item Problem: multi-vehicle map information gain optimality can be computationally-intractable
-	\item Solution: bidding-based solution simplifies a complicated optimization into a series of simplified optimizations
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Related Work}
+		\item Approach: bidding-based solution simplifies a complicated optimization into a series of simplified optimizations
+	\end{itemize}
+	\pause
+	\item Multi-Vehicle Autonomous Patrol
+	\begin{itemize}
+		\item Patrolling large environments with humans is expensive and sometimes dangerous
+		\item Several robotic approaches introduce new variables for patrol
+		\item Approach: degrade probabilistic occupancy grid over time to encourage revisiting past regions
+	\end{itemize}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Related Work in Multi-Vehicle Exploration}
 
@@
 \begin{frame}
-\frametitle{Proposed Approach}
+\frametitle{Proposed Approach in Multi-Vehicle Exploration}
 
@@
 \begin{frame}
-\frametitle{Numerical Example: SEH Second Floor Multi-Vehicle Autonomous Exploration}
+\frametitle{Numerical Example: SEH Multi-Vehicle Exploration}
 
@@
 	\end{itemize}
+	\pause
 	\item Proposed Approach
@@
 \begin{frame}
-\frametitle{Numerical Example: SEH Second Floor Multi-Vehicle Autonomous Patrol}
+\frametitle{Numerical Example: SEH Multi-Vehicle Patrol}
 
@@
 \begin{frame}
-\frametitle{Multi-Vehicle Autonomous Exploration and Patrol Conclusions}
+\frametitle{Multi-Vehicle Exploration and Patrol Summary}
 
@@
 \begin{frame}
+\frametitle{Experiments}
+\begin{enumerate}[I]
+\addtocounter{enumi}{3}
+	\item Experiments
+\end{enumerate}
+\end{frame}
+
+\begin{frame}
 \frametitle{Experimental Overview}
@@
 	\item Ground vehicle experiment for 2D occupancy grid mapping and autonomous exploration
-	\item Flying vehicle experiment in large 3D space resembling an office
-	\item Flying vehicle experiment in a complex 3D space
+	\item Aerial vehicle experiment in large 3D space resembling an office
+	\item Aerial vehicle experiment in a complex 3D space
 \end{enumerate}
@@
 \begin{frame}
-\frametitle{Flying Vehicle in a Large Office-Like Environment}
+\frametitle{Aerial Vehicle in a Large Office-Like Environment}
 \begin{itemize}
@@
     \caption*{Quadrotor Bottom}
+  \end{subfigure}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Aerial Vehicle in a Large Office-Like Environment}
+\begin{itemize}
+        	\item Environment Size: $10.8$ m $\times$ $11.8$ m $\times$ $1.65$ m
+\end{itemize}
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{.4\linewidth}
+    \centering\includegraphics[width=\linewidth]{experiment_north.jpg}
+    \caption*{North Region}
+  \end{subfigure}
+  \begin{subfigure}[t]{.4\linewidth}
+    \centering\includegraphics[width=\linewidth]{experiment_south.jpg}
+    \caption*{South Region}
   \end{subfigure}
@@
 \begin{frame}
-\frametitle{Flying Vehicle in a Complex 3D Environment}
+\frametitle{Aerial Vehicle in a Complex 3D Environment}
 \begin{itemize}
@@
 	\item Objects are not vertically-uniform
+	\item Environment Size: $9.0$ m $\times$ $8.5$ m $\times$ $2.25$ m
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Full 3D Flying Experiment Video}
+\frametitle{Full 3D Aerial Experiment Video}
 
@@
 \begin{frame}
-\frametitle{Experiment Conclusions}
+\frametitle{Experiment Summary}
 \begin{itemize}
@@
 \section*{}
-\subsection*{Conclusions and Publications}
+\subsection*{Conclusions}
 
@@
 
+\begin{frame}
+\frametitle{Thanks}
+\begin{itemize}
+        	\item Special thanks to the US Naval Research Lab for support under project ``Intelligent Microflyer''
+	\item Questions
+\end{itemize}
+\end{frame}
+
 \end{document}

### 2018-07-13 Dissertation/Defense/defense.tex
videos full size

@@
 		\includemedia[
-  		width=0.720\textwidth,
-  		height=0.576\textwidth,
+  		width=\textwidth,
+  		height=0.8\textwidth,
   		activate=pageopen,
@@
 		\includemedia[
-  		width=0.720\textwidth,
-  		height=0.576\textwidth,
+    		width=\textwidth,
+  		height=0.8\textwidth,
   		activate=pageopen,
@@
 		\item Decentralize the algorithms so processes may be distributed among multiple agents
+		\vspace*{0.3cm}
+		\item Special thanks to the US Naval Research Lab for support under project ``Intelligent Microflyer''
 	\end{itemize}
@@
 
-\begin{frame}
-\frametitle{Thanks}
-\begin{itemize}
-        	\item Special thanks to the US Naval Research Lab for support under project ``Intelligent Microflyer''
-	\item Questions
-\end{itemize}
-\end{frame}
-
 \end{document}

### 2018-07-19 Dissertation/Thesis/tex/intro.tex
dissertation final edit done through chap 1

@@
 Much like single-vehicle autonomous exploration, the multi-vehicle problem requires a computationally-inexpensive algorithm to handle numerous possible robot motions.
-These nontrivial goals can be handled by extending single-vehicle autonomous exploration with a bidding-based approach.
+These nontrivial goals can be handled similarly to single-vehicle autonomous exploration, but following a bidding-based framework.
 
@@
 distance to a wall).
-This probability density is referred to as the \emph{forward sensor model}, as it relates a depth measurement conditioned on the occupancy of the map.
+This probability density is referred to as the \emph{forward sensor model}, as it relates a depth measurement range to the occupancy of the map.
 
-In contrast, the inverse sensor model relates the occupancy of the map conditioned on the depth measurement.
+In contrast, the inverse sensor model relates the occupancy of the map to the depth measurement range.
 This problem is solved using Bayes' theorem.
@@
 \paragraph{Log-Odds Ratio} Both approaches described above use the log-odds ratio to update grid cell probabilities.
-The log-odds ratio formulation is a popular framework for updating binary random variables with static state within a Bayesian filter.
+The log-odds ratio formulation follows a popular framework for updating binary random variables with static state within a Bayesian filter.
 The main idea is that instead of multiplying terms from prior and current time steps, the properties of logarithms allow these terms to be simply added, while avoiding some truncation issues associated with probabilities close to $0$ or $1$~\cite{ThrBurFox05,Thr03}.
@@
 As such, human teleoperation and monitoring are often essential to guide the vehicles safely through unknown surroundings.
-Therefore, it is desirable that an active motion planning scheme is developed and integrated with SLAM such that vehicles are able to determine their path without human intervention, and explore unknown areas autonomously.
+Therefore, it is desirable that an active motion planning scheme is developed and integrated with SLAM such that vehicles are able to determine their paths without human intervention, and explore unknown areas autonomously.
 The two approaches used to solve this problem are described next.
@@
 The most popular approach to solve the autonomous exploration problem is known as frontier-based exploration, originally-proposed in~\cite{Yam97,Yam98} for 2D applications, and was later extended to 3D with a visibility metric in~\cite{SawKriSri09} and~\cite{ZhuDinLinWu15,SenWan16,KleDor13} using the popular Octomap representation (a variation on occupancy grid mapping).
-The main idea is that robots toward the border between certain and uncertain space, referred to as a frontier.
+The main idea is that robots move toward the border between certain and uncertain space, referred to as a frontier.
 Then the robot takes depth measurement at this frontier, thus pushing back the boundary.
@@
 \paragraph{Entropy-Based Autonomous Exploration}
-Entropy-based approaches address the suboptimal nature of frontier-based approaches by selecting robotic actions designed to decrease minimize map uncertainty~\cite{CarDamKumCas15,StaGriBur05,BurMooStaSch05}.
-These approaches use a measure of uncertainty known as Shannon's entropy, which becomes smaller as cell probabilities approach $0$ or $1$ (becoming more certain).
+Entropy-based approaches address the suboptimal nature of frontier-based approaches by selecting robotic actions designed to decrease map uncertainty~\cite{CarDamKumCas15,StaGriBur05,BurMooStaSch05}.
+These approaches use a measure of probabilistic uncertainty known as Shannon's entropy, which becomes smaller as cell probabilities approach $0$ or $1$ (becoming more certain).
 Existing entropy-based approaches simulate possible future measurements from various locations with so-called ``hallucination measurements,'' which typically corresponds to expected depth measurements from the robot to the closest grid cell that is possibly occupied.
@@
 Instead, an auction-based approach addresses the computational restrictions effectively.
-Bidding-based approaches were introduced with~\cite{SimApfBurFoxMooThrYou00,DiaSte00}, and extended to several other robotic problems~\cite{GerMaj02,ZloSteDiaTha02,SarBal05,ChoBruHow09} exploring decentralization and task allocation with varying robot capabilities.
+Bidding-based approaches were introduced with~\cite{SimApfBurFoxMooThrYou00,DiaSte00}, and were extended to several other robotic problems~\cite{GerMaj02,ZloSteDiaTha02,SarBal05,ChoBruHow09} exploring decentralization and task allocation with varying robot capabilities.
 In~\cite{SimApfBurFoxMooThrYou00}, a central executive assigns tasks to robots for maximizing overall utility of the group.
@@
 Robots submit bids based on expected coverage of uncertain spaces and travel distance.
-The robot winning the auction is awarded the task, and the remaining bids are discounted to prevent robots from covering the same area, thereby coordinating the mapping efforts.
+The robot winning the auction is awarded the task, and the remaining bids for subsequent auctions are discounted to prevent robots from covering the same area, thereby coordinating the mapping efforts.
 Bidding-based approaches simplify multi-vehicle cooperation in a computationally-efficient manner, and integrate effectively with the autonomous mapping and exploration approaches proposed in this dissertation.
@@
 Then, Chapter \ref{chap:ae2D} extends these contributions to autonomous exploration in 2D space, where a computational approach to predict future map uncertainty is presented.
-Next, Chapter \ref{chap:ae2D} applies the proposed occupancy grid mapping approach to 3D environments, and presents a method to simplify autonomous exploration in 3D environments with map projections.
+Next, Chapter \ref{chap:ae3Dsimple} applies the proposed occupancy grid mapping approach to 3D environments, and presents a method to simplify autonomous exploration in 3D environments with map projections.
 Chapter \ref{chap:multivehicle} applies this 3D mapping and exploration to multi-vehicle exploration and patrol scenarios using a bidding-based framework.
@@
 The first contribution is an exact Bayesian solution to occupancy grid mapping.
-The proposed approach considers that occupied spaces occlude measurement behind them, which allows certain mapping outcomes to be systematically grouped together.
+The proposed approach considers that a depth measurement captures the closest occupied space, which allows certain mapping outcomes to be systematically grouped together.
 The exact solution to occupancy grid cell probability, which was previously-believed to have exponential complexity, is now possible to solve in real-time.
@@
 \sloppy The final contribution coordinates the mapping and exploration efforts of multiple robots together.
-The mapping process is multi-threaded such that multiple members update the same occupancy grid.
+The mapping process is multi-threaded such that multiple members update the same occupancy grid simultaneously.
 The exploration algorithm follows a bidding-based structure.
@@
 
-The following is a list of publications. These include research included in this dissertation, as well as contributions in control, data association, and estimation.
+The following is a list of publications. These cover research included in this dissertation, as well as contributions in control, data association, and estimation.
 

### 2018-07-25 Dissertation/Thesis/tex/pogm.tex
dissertation final edit done through chap 2

@@
 
-The occupancy probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$ is based on the forward sensor model $p(z_{t,l}|m,X_{1:t},Z_{1:t-1})$ that describe the distribution of the measurements for the given robot pose and the map.
+The occupancy probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$ is based on the forward sensor model $p(z_{t,l}|m,X_{1:t},Z_{1:t-1})$ that describes the distribution of the measurements for the given robot pose and the map.
 This specifies the stochastic characteristics of the sensor are a known distribution, specific to a particular depth sensor.
@@
 In each case, cells the first three cells are free, the fourth cell is occupied, and the fifth and sixth cells may or may not be occupied.
-All above outcomes correspond to the event $\mathbf{r}_{l,4+}$.
+All above outcomes correspond to the event $\mathbf{r}_{l,4+}$, and therefore share the same forward sensor model.
   \label{fig:show_rkplus}
@@
 \end{align}
-where $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ is chosen for convenience and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the probability density of the measurement when all of cells in the field of view of the $l$-th ray is not occupied.
+where $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ is chosen for convenience and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the probability density of the measurement when all of cells in the field of view of the $l$-th ray are not occupied.
 The normalizer $\eta_{t,l}$ is given by
@@
 For the second ray, namely $z_{t,2}$, this may depend on $z_{t,1}$, but not subsequent rays $z_{t,3:n_z}$.
-By the time $z_{t,n_z}$ is considered, it may depend on all prior rays $z_{t,1:n_z-1}$.
+By the time $z_{t,n_z}$ is considered, it may depend on any combination of prior rays $z_{t,1:n_z-1}$.
 Therefore, the ray-by-ray approach considers rays within a single scan as partially dependent on each other.
@@
 First, we compare the proposed exact solution to the ray-by-ray inverse sensor model summarized in Algorithm \ref{alg:RayByRayISM} with an approximations of \refeqn{ISM_Approx_1}--\refeqn{ISM_Approx_2}.
-Table~\ref{tab:penn} shows the mapping parameters and data specifications for the actual odometry and lidar measurements from University of Pennsylvania through Coursera open course~\cite{coursera}.
+Table~\ref{tab:penn} shows the mapping parameters and data specifications for the actual odometry and lidar measurements from University of Pennsylvania through the Coursera open course~\cite{coursera}.
 Among these parameters, cell resolution and the number of rays have the greatest impact on computation.
@@
 As shown in the figure, both approaches capture the structure of the environment.
-However, the significant differences can be seen, particularly with an enlarged mapping section in Figure~\ref{fig:penn_map_zoom}.
+However, significant differences can be seen, particularly with an enlarged mapping section in Figure~\ref{fig:penn_map_zoom}.
 
@@
 \end{align*}
-which is maximized when the probability of occupancy is $0.5$ for all cells (more uncertain), and it is minimized when they are either $0$ or $1$ (less uncertain). 
+which is maximized when the probability of occupancy is $0.5$ for all cells (more uncertain), and is minimized as probabilities approach either $0$ or $1$ (less uncertain). 
 The entropy differences among the algorithms can be clearly depicted in Figure~\ref{fig:entropy_comp} at frame 2280.
@@
         \includegraphics[trim={1.5cm 3.5cm 1.5cm 1.5cm},clip,width=\textwidth]{map_comparison.png}
-        \caption{Comparison of the total mapped environment based on the approximate (left) and proposed (right) inverse sensor model.}
+        \caption{Approximate (left) and proposed (right) inverse sensor model of the map}
         \label{fig:penn_map_total}
@@
         \centering
-        \includegraphics[trim={1.5cm 3.5cm 1.5cm 1.5cm},clip,width=\textwidth]{map_zoom.png}
-        \caption{Evidence of significant improvement at certain area of the map.}
+        \includegraphics[trim={1.5cm 3.35cm 1.5cm 1.5cm},clip,width=\textwidth]{map_zoom.png}
+        \caption{Approximate (left) and proposed (right) inverse sensor model of an area}
         \label{fig:penn_map_zoom}
@@
         \includegraphics[width=\textwidth]{AISM_Image_inf_19.pdf}
-        \caption{Approximate inverse sensor model}
+        \caption{Approx. inverse sensor model}
         \label{fig:AISM}
@@
 
-In this chapter, we presented the exact solution to the inverse sensor model for a single measurement ray, then extended this result to 2D occupancy grids.
-Measurement scans were analyzed with two techniques, namely ray-by-ray and synergistic combinations.
+In this chapter, we present the exact solution to the inverse sensor model for a single measurement ray, then extend this result to 2D occupancy grids.
+Measurement scans are analyzed with two techniques, namely ray-by-ray and synergistic combinations.
 Then, numerical results show how both proposed solutions outperform existing approximate solutions.

### 2018-07-25 Dissertation/Thesis/tex/ae2D.tex
dissertation final edit done through chap 3

@@
 Shannon's entropy commonly serves as an uncertainty measure~\cite{StaGriBur05}.
-Provided grid cell probabilities of map $m$, Shannon's entropy is defined as
+Provided grid cell probabilities of map $ m$, Shannon's entropy is defined as
 \begin{align}
@@
 \end{prop}
-\begin{proof}% TODO: add ref?
-See Appendix B.
+\begin{proof}
+See Appendix \ref{app:ExpectedRayEntropy}.
 \end{proof}
@@
 \label{eqn:CollisionInequalityConstraint}
-P_\text{collision}(X)=1-\prod_{i\in\mathcal C_X}(P(\bar{\mathbf{m}}_i)\leq\beta,
+P_\text{collision}(X)=1-\prod_{i\in\mathcal C_X}P(\bar{\mathbf{m}}_i)\leq\beta,
 \end{align}
@@
 \end{align}
-is computed by a summation about the $n_d$ rays as %\refeqn{Objective},
+and this is computed by a summation about the $n_d$ rays as %\refeqn{Objective},
 \begin{align}
@@
 Here, a robot autonomously maps and explores a 2D environment composed of two rooms and one hallway.
-The robot models its surroundings with an occupancy grid of $15,000$ cells where grid cell edges have length $\alpha=0.2$m, composing a map with dimensions $30\text{m}\times20\text{m}$.
-The initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) for grid cells covered by the circular robot of radius $0.1$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
+The robot models its surroundings with an occupancy grid of $15,000$ cells where grid cell edges have length $\alpha=0.2$ m, composing a map with dimensions $30\text{ m}\times20\text{ m}$.
+The initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) for grid cells covered by the circular robot of radius $0.1$ m and $P(\mathbf{m}_i)=0.5$ for all other cells.
 At each time step, the robot receives a measurement scan, where the probabilistic properties of the sensor are taken from~\cite{PirRutBisSch11,KhoElb12}.
-Then, the $n_c=8$ evenly-spaced candidate locations about a circle of radius $\delta=0.5$m around the current pose location are considered, where $n_d=32$ measurement rays are evenly-spaced about the candidate location.
+Then, the $n_c=8$ evenly-spaced candidate locations about a circle of radius $\delta=0.5$ m around the current pose location are considered, where $n_d=32$ measurement rays are evenly-spaced about the candidate location.
 When no current candidates yield expected information gains above $\mathcal I_\text{min}=2$, $\lambda=1.25$ is multiplied to $n_c$ and $\delta$.
@@
 Next, the proposed autonomous exploration approach is applied to the floor plan of the Intel Research Lab, illustrated in Figure~\ref{fig:intel}.
-The robot explores its surroundings with an occupancy grid with $90,000$ cells where grid cell edges are $\alpha=0.1$m, composing a map with dimensions $30\text{m}\times30\text{m}$.
-Similar to the prior example, the initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) is chosen for grid cells covered by the circular robot of radius of $0.3$m and $P(\mathbf{m}_i)=0.5$ for all other cells.
-For added safety, cells falling within $0.6$m of the robot are considered inside the possible collision zone $\mathcal{C}_X$, from \refeqn{CollisionInequalityConstraint}, where $\beta=0.1$ is selected.
+The robot explores its surroundings with an occupancy grid with $90,000$ cells where grid cell edges are $\alpha=0.1$ m, composing a map with dimensions $30\text{ m}\times30\text{ m}$.
+Similar to the prior example, the initial probability $P(\mathbf{m}_i)=1\times10^{-10}\approx0$ (minimum value for free space) is chosen for grid cells covered by the circular robot of radius of $0.3$ m and $P(\mathbf{m}_i)=0.5$ for all other cells.
+For added safety, cells falling within $0.6$ m of the robot are considered inside the possible collision zone $\mathcal{C}_X$, from \refeqn{CollisionInequalityConstraint}, where $\beta=0.1$ is selected.
 
@@
 
-The time sequence of the exploration for every 5~minutes are simulated in the ROS Stage 2D environment.
+The large 2D exploration is simulated in the ROS Stage 2D environment.
 The corresponding simulation results are illustrated in Figure \ref{fig:IRL} and a video is available at \href{https://www.youtube.com/watch?v=5VdzKHreB_s}{\WriteBlue{https://www.youtube.com/watch?v=5VdzKHreB\_s}}.

### 2018-07-25 Dissertation/Thesis/tex/ae3Dsimple.tex
updated figs of chap 5
updated references of myself
dissertation final edit done through chap 4

@@
 
-Next, we describe two approaches to projecting the 3D probabilistic map onto a 2D map at the exploration height.
+Next, we describe two approaches to projecting the 3D probabilistic map onto a 2D map at the exploration height~\cite{KauTakAiLee18}.
 Assuming projections are sufficient is consistent with rooms and hallways with non-complex geometries.
@@
 Gazebo is a 3D dynamic simulator~\cite{KaeHow04}, with plugins to simulate an Asus Xtion color depth sensor and a Hokuyo LIDAR.
-Gazebo also provides position and attitude transforms from the world to the quadrotor body, and the ability to set robot model states directly.
+Gazebo also provides position and attitude transformations from the world to the quadrotor body, and the ability to set robot model states directly.
 
@@
 These serve as inputs for ray casting in 3D, which provides the cell probabilities and depths along the ray through geometry.
-Then, the inverse sensor model serves to update the map ray-by-ray.
+Then, the inverse sensor model updates the map ray-by-ray.
 The node is designed to work for any number of sensors, specified by sensor parameters in a yaml configuration file.
@@
 \begin{figure}[!t]
+	\centering
     	\begin{subfigure}[t]{0.24\columnwidth}

### 2018-07-25 Dissertation/Thesis/tex/multivehicle.tex
dissertation final edit done through chap 5

@@
 \end{align}
-where the expected scan $Z$ is given from \refeqn{expectedInfoGainRay} and \refeqn{FindRc}.
+where the expected scan $Z_c$ is given from \refeqn{expectedInfoGainRay} and \refeqn{FindRc}.
 
@@
 The first is that patrol is included in the exploration scheme to incentivize the robots to revisit areas after time passes.
-The second is that the robots must handle a human is walking around the space, unknown to the cooperating robots, requiring the receding horizon framework to avoid collisions with this dynamic obstacle.
+The second is that the robots must handle a human walking around the space, unknown to the cooperating robots, requiring the receding horizon framework to avoid collisions with this dynamic obstacle.
 
@@
 While degrading cells with \refeqn{CellDegradationScalar} is inexpensive, degradation must cycle through all cells, which can be costly with a map containing over $45$ million cells.
-Since all cells are affected by degradation, simply sending the map \emph{changes} from measurement scan updates alone is insufficient for the exploration node to track these changing probabilities.
+Since all cells are affected by degradation, simply sending the map \emph{changes} from measurement scan updates alone is insufficient for the exploration node to track all changing probabilities.
 Thus, the complete map is sent from this mapping thread, and an additional thread in the exploration node subscribes to this message to update the projected map $m_\text{2D}$ slowly.
@@
 
-The robot generated a fairly-complete occupancy grid within $9$ minutes, and proceeded to patrol the space during the remaining time.
+The robot generates a fairly-complete occupancy grid within $9$ minutes, and proceeds to patrol the space during the remaining time.
 After $9$ minutes, the total map entropy no longer increases or decreases significantly; the rate of cell degradation is roughly equal and opposite to improvements to the probabilistic map from the robot team.
@@
 This chapter introduced a bidding-based framework for cooperative autonomous exploration and patrol.
-These contributions exploit the values of the objective functions used in autonomous exploration optimizations for a bids over a series of auctions.
+These contributions exploit the values of the objective functions used in autonomous exploration optimizations for bids over a series of auctions.
 Between auctions, a copy of the map and collision criteria are updated, thereby coordinating the exploration efforts.

### 2018-07-25 Dissertation/Thesis/tex/ae3Dcomplex.tex
dissertation final edit done through chap 6

@@
 The bump functions use $f_\text{max}=1.0$, $f_\text{far}=0.1$, and $\beta=0.1$ to account for travel costs with \refeqn{BumpFun}.
-The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.25$ m/sec and the computation time varies from $1.8$ sec to $2.5$ sec.
+The receding horizon optimal time $d_\text{opt}$ is based on a fixed robot velocity of $0.25$ m/sec and the computation times vary from $1.8$ sec to $2.5$ sec.
 
@@
 
-The simulation was run twice.
-Case 1 was as described in this paper.
-Case 2 was identical to Case 1, except the reduced map $ m_\text{reduced}$ is used when computing \refeqn{Iscan}.
+The simulation is run twice.
+Case 1 is as described in this paper.
+Case 2 is identical to Case 1, except the reduced map $ m_\text{reduced}$ is used when computing \refeqn{Iscan}.
 Case 2 is largely inspired by the promising experimental results covered in Section \ref{sec:QuadrotorNRL}, where a projected map based on the same criteria as \refeqn{Proj3DMapComb} proved effective for level flight.
@@
 \begin{figure}
-	\centerline{
+	\centering
 	\includegraphics[width=0.9\columnwidth]{scitech_entropy_comparison_flat.pdf}
-	}
 	\caption{Entropy Histories During Mars Exploration}
@@
 
-However, the choice of occupancy grid map used for entropy predictions introduced an interesting tradeoff.
-In Case 1, when the full probabilistic map $m$ was used in \refeqn{Iscan}, regions where the grid cells were partially-known were frequently reconsidered until the space was well-known.
-Conversely in Case 2, when $m_\text{reduced}$ was used instead, the robot repeatedly left spaces that were missing a few grid cells to visit new terrain.
-This is because $m_\text{reduced}$ contained some cells with large occupancy probabilities based on \refeqn{Proj3DMapComb}, which allowed some grid cells of $m$ enclosed within a cell of $m_\text{reduced}$ to be uncertain.
-The exploration policy of Case 2 incorrectly assumed these regions were well-known.
-Ironically, this false assumption actually led to greater information gains when the vehicle moved on to unvisited terrain.
-Conversely with Case $1$, which used $m$ for computing expected entropy, the total map entropy decreased more steadily and generated a more-complete 3D occupancy grid of the surface of Mars.
+However, the choice of occupancy grid map used for entropy predictions introduces an interesting tradeoff.
+In Case 1, when the full probabilistic map $m$ is used in \refeqn{Iscan}, regions where the grid cells are partially-known are frequently reconsidered until the space is well-known.
+Conversely in Case 2, when $m_\text{reduced}$ is used instead, the robot repeatedly leaves spaces that are missing a few grid cells to visit new terrain.
+This is because $m_\text{reduced}$ contains some cells with large occupancy probabilities based on \refeqn{Proj3DMapComb}, which allows some grid cells of $m$ enclosed within a cell of $m_\text{reduced}$ to be uncertain.
+The exploration policy of Case 2 incorrectly assumes these regions are well-known.
+Ironically, this false assumption can actually lead to greater information gains when the vehicle moves on to unvisited terrain.
+Conversely with Case $1$, which uses $m$ for computing expected entropy, the total map entropy decreases more steadily and generates a more-complete 3D occupancy grid of the surface of Mars.
 
-In short, the proposed 3D probabilistic occupancy grid mapping and autonomous exploration were simulated successfully in real-time over the surface of Mars using ROS and Gazebo.
-Choosing $m$ for entropy predictions produced a more-complete map, but choosing $m_\text{reduced}$ was sometimes beneficial for exploring new terrain faster, while forgoing map completeness.
+In short, the proposed 3D probabilistic occupancy grid mapping and autonomous exploration are simulated successfully in real-time over the surface of Mars using ROS and Gazebo.
+Choosing $m$ for entropy predictions produces a more-complete map, but choosing $m_\text{reduced}$ is sometimes beneficial for exploring new terrain faster, while forgoing map completeness.
 

### 2018-07-25 Dissertation/Thesis/tex/results.tex
dissertation final edit done through chap 7.3.2

@@
 The depth readings are obtained by a Kinect depth sensor.
-Since the experiment is to map and explore a 2D environment, only a single central row of the 3D Kinect depth scan provide the necessary measurements.
+Since the experiment is to map and explore a 2D environment, only a single central row of the 3D Kinect depth scan provides the necessary measurements.
 The robot captures an experimental environment with walls constructed from Styrofoam.
@@
 Any time delay between the robot pose estimation and the sensor depth readings can cause conflicting information, which may be harmful to the probabilistic map.
-Thus, ROS approximate message filters are applied such that the pose from Vicon Tracker ($100$Hz) and the Microsoft Kinect ($30$Hz) are nearly synchronized, which provides validity of the assumption that the measurement ray positions and directions are known deterministically.
+Thus, ROS Message Filters are applied such that the pose from Vicon Tracker ($100$ Hz) and the Microsoft Kinect ($30$ Hz) are nearly synchronized, which provides validity of the assumption that the measurement ray positions and directions are known deterministically.
 
@@
         		\includegraphics[width=\textwidth]{test_setup_bottom_left_cropped.jpg}
-        		\caption{Bottom-left view}
+        		\caption{Bottom-Left View}
         		\label{fig:Experiment_blv}
@@
         		\includegraphics[width=\textwidth]{test_setup_bottom_right_cropped.jpg}
-        		\caption{Bottom-right view}
+        		\caption{Bottom-Right View}
         		\label{fig:Experiment_brv}
@@
 In some cases, the robot is able to update the cell occupancy probabilities; in other cases, cells fall inside regions occluded from all reachable space, and the robot cannot update these cells.
-Among those grid cells visible from collision-free locations, the vast majority of cells were updated in less than $3$ min while following a very slow ($5$ cm/sec) trajectory.
+Among those grid cells visible from collision-free locations, the vast majority of cells are updated in less than $3$ min while following a very slow ($5$ cm/sec) trajectory.
 
@@
 	The 3D experimental environment is designed to resemble a large room within an office building.
-The walls are build from from cardboard suspended from stanchions and gray sheeting with $80/20$ supports, and miscellaneous objects are placed around the room as well.
+The walls are built from from cardboard suspended from stanchions and gray sheeting with $80/20$ supports, and miscellaneous objects are placed around the room as well.
 	\label{fig:exp3DEnvironment}
@@
 The robot generates a 3D probabilistic map (Figure \ref{fig:exp3DOccMap}) and uses this for entropy prediction and collision-free motion planning with a 3D cost map (Figure \ref{fig:Full3DExplorationCostmaps}).
+A video of this experiment is available at \href{https://www.youtube.com/watch?v=2Q2_-d8kNu0}{\WriteBlue{https://www.youtube.com/watch?v=2Q2\_-d8kNu0}}.
 

### 2018-08-02 Dissertation/Thesis/tex/conclusions.tex
added Final Papers folder for unchanging final papers

@@
 The approach follows a receding horizon, which improves collision-avoidance and substantially reduces unnecessary robotic actions.
-Furthermore, this approach is extended to autonomous patrol, where an exponential map degradation incentivizes robot team members to periodically revisit areas.
+Furthermore, this approach is extended to autonomous patrol, where exponential map degradation incentivizes robot team members to periodically revisit areas.
 These extensions to occupancy grid mapping and autonomous exploration are demonstrated with numerous simulations and experimental results.

### 2018-08-02 Dissertation/Thesis/tex/frontmatter.tex
updated section 7.4 of dissertation to bring figures together with section 7.3, started defense presentation
added Final Papers folder for unchanging final papers

@@
 
-James Hahn\\
-Department of Computer Science\\
-The George Washington University\\
-Committee Chair\\
+\vspace*{0.05\textwidth}
+Taeyoung Lee, Associate Professor of Mechanical and Aerospace Engineering, Dissertation Director\\
 
-Taeyoung Lee\\
-Department of Mechanical and Aerospace Engineering\\
-The George Washington University\\ 
-Committee Member\\ % remember to add a space between committee members
+James Hahn, Professor of Computer Science, Committee Member\\
 
-Robert Pless\\
-Department of Computer Science\\
-The George Washington University\\
-Committee Member\\
+Robert Pless, Professor of Computer Science, Committee Member\\
 
-Chung Hyuk Park\\
-Department of Biomedical Engineering\\
-The George Washington University\\ 
-Committee Member\\
+Chung Hyuk Park, Assistant Professor of Biomedical Engineering, Committee Member\\
 
-Zhuming Ai\\
-Information Management and Decision Architectures\\
-U.S. Naval Research Laboratory\\ 
-Committee Member
+Zhuming Ai, Computer Engineer, U.S. Naval Research Laboratory, Committee Member
 
@@
 
-\phdgrad{July 13, 2018}
+\phdgrad{August 31, 2018}
 \defensedate{July 13, 2018}
@@
 \showlistoffigures
-\showlistoftables
+\hidelistoftables
 \hidenomenclature
@@
 % you can hide/show the glossaries page
-\showglossarieslistofabbreviations
-\showglossarieslistofsymbols
-\showglossariesglossaryofterms
+\hideglossarieslistofabbreviations
+\hideglossarieslistofsymbols
+\hideglossariesglossaryofterms
 
-% acronyms defined in glossaries
-\newabbreviation{lidar}{LIDAR}{Light Detection and Ranging}
-% defining abbreviations like this allows for autocompletion
-\newglossaryentry{filo}{
-    name={FILO},
-    type=\glsxtrabbrvtype,
-    description={first in last out},
-    first={first in last out (FILO)}
-}
-
-% glossary entries
-\newglossaryentry{linux}{
-    name=Linux,
-    description={is a generic term referring to the family of Unix-like computer operating systems that use the Linux kernel},
-    plural=Linuces
-}
-
-\newglossaryentry{matrix}{
-    name={matrix},
-    plural={matrices},
-    description={rectangular array of quantities}
-}
-
-% symbols
-\newglossaryentry{M}{
-    type=symbols,
-    name={\ensuremath{M}},
-    sort=M,
-    description={a \gls{matrix}}
-}
-
-\newglossaryentry{F}{
-    type=symbols,
-    name={\ensuremath{F}},
-    sort=F,
-    description={External Force}
-}
-% Some abstract text
 \abstract{

### 2018-08-28 Dissertation/Defense/phd_research_20min.tex
shortened version for RR (unclassified & open source prior to RR)

--- 2018-07-13 Dissertation/Defense/defense.tex
@@
 \institute{\footnotesize
-{\normalsize Evan Kaufman}\\
+{\\
 \vspace*{0.3cm}
-Mechanical and Aerospace Engineering\\George Washington University
+\normalsize Evan Kaufman}\\
+\vspace*{0.3cm}
+Robotic Research LLC
+\\
+\vspace*{0.3cm}
+September 5, 2018
 }
@@
 	\begin{itemize}
-		\item Environments can be dangerous and uncertain
-		\item Robots can replace humans for tasks such as surveillance, search-and-rescue, and cleaning
+		\item Robots can replace humans dangerous/uncertain tasks such as surveillance, combat, and search-and-rescue
+		\item Key question: when robots explore new spaces, how can they maximize their knowledge about their environment?
 	\end{itemize}
@@
 		\item Robot must understand their environment by generating a map, referred to as \emph{mapping}
-		\item Selecting and performing robotic actions to maximize map information, referred to as \emph{autonomous exploration}, is necessary to optimally explore initially-uncertain environments
+		\item Selecting and performing robotic actions to maximize map information, referred to as \emph{autonomous exploration}
 		\item Cooperative robots can share information to achieve a collective goal with \emph{multi-vehicle exploration and patrol}
@@
 
-\begin{frame}
-\frametitle{Ray Casting}
-
-\begin{itemize}
-	\item Required for 2D and 3D occupancy grid maps
-	\item Main idea: use simple geometry to find the distance from the robot sensor to each cell intersected by the measurement ray
-\end{itemize}
-
-\begin{figure}[!ht]
-    \centering
-    \begin{subfigure}[t]{0.4\columnwidth}
-        \centering
-        \includegraphics[width=\textwidth]{RayCastingIllustration.png}
-%        \label{fig:penn_map_total}
-    \end{subfigure}
-    \hspace*{0.05\columnwidth}
-    \begin{subfigure}[t]{0.4\columnwidth}
-        \centering
-        \includegraphics[width=\textwidth]{RayCastIllustrationReducedMapOnly.png}
-%        \label{fig:penn_map_zoom}
-    \end{subfigure}
-\end{figure}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Scan Combination Approaches}
-
-\begin{itemize}
-	\item \emph{Ray-By-Ray Update}: given a scan of measurement rays, update the map based on each measurement ray \emph{individually} and \emph{sequentially}
-	\begin{itemize}
-		\item Benefit: the measurement rays maintain some dependency as they capture the same map
-		\item Drawback: the best order for evaluating rays is not necessarily chosen
-	\end{itemize}
-%	\item Conditional probability:
-%	\begin{align*}
-%		P(\mathbf{m}_i|X_{1:t}&,Z_{1:t})\nonumber\\&
-%		=P((\dots(((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z})
-%	\end{align*}
-	\item \emph{Synergistic Update}: consider every measurement ray inside a scan \emph{simultaneously} to update all grid cells inside the FOV
-	\begin{itemize}
-		\item Benefit: synergistic update provides a method where all rays are considered simultaneously
-		\item Drawback: required assumption that the measurement rays are independent, though they capture the same map
-	\end{itemize}
-\end{itemize}
-
-\end{frame}
+%\begin{frame}
+%\frametitle{Ray Casting}
+%
+%\begin{itemize}
+%	\item Required for 2D and 3D occupancy grid maps
+%	\item Main idea: use simple geometry to find the distance from the robot sensor to each cell intersected by the measurement ray
+%\end{itemize}
+%
+%\begin{figure}[!ht]
+%    \centering
+%    \begin{subfigure}[t]{0.4\columnwidth}
+%        \centering
+%        \includegraphics[width=\textwidth]{RayCastingIllustration.png}
+%%        \label{fig:penn_map_total}
+%    \end{subfigure}
+%    \hspace*{0.05\columnwidth}
+%    \begin{subfigure}[t]{0.4\columnwidth}
+%        \centering
+%        \includegraphics[width=\textwidth]{RayCastIllustrationReducedMapOnly.png}
+%%        \label{fig:penn_map_zoom}
+%    \end{subfigure}
+%\end{figure}
+%
+%\end{frame}
+
+%\begin{frame}
+%\frametitle{Scan Combination Approaches}
+%
+%\begin{itemize}
+%	\item \emph{Ray-By-Ray Update}: given a scan of measurement rays, update the map based on each measurement ray \emph{individually} and \emph{sequentially}
+%	\begin{itemize}
+%		\item Benefit: the measurement rays maintain some dependency as they capture the same map
+%		\item Drawback: the best order for evaluating rays is not necessarily chosen
+%	\end{itemize}
+%%	\item Conditional probability:
+%%	\begin{align*}
+%%		P(\mathbf{m}_i|X_{1:t}&,Z_{1:t})\nonumber\\&
+%%		=P((\dots(((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z})
+%%	\end{align*}
+%	\item \emph{Synergistic Update}: consider every measurement ray inside a scan \emph{simultaneously} to update all grid cells inside the FOV
+%	\begin{itemize}
+%		\item Benefit: synergistic update provides a method where all rays are considered simultaneously
+%		\item Drawback: required assumption that the measurement rays are independent, though they capture the same map
+%	\end{itemize}
+%\end{itemize}
+%
+%\end{frame}
 
@@
 
-\begin{frame}
-	\frametitle{2D Autonomous Exploration: Intel Research Lab}
-	\centering{
-		\includemedia[
-  		width=\textwidth,
-  		height=0.8\textwidth,
-  		activate=pageopen,
-  		addresource=IRL.mp4,
-  		flashvars={source=IRL.mp4}
-		]{}{VPlayer.swf}
-		}
-\end{frame}
+%\begin{frame}
+%	\frametitle{2D Autonomous Exploration: Intel Research Lab}
+%	\centering{
+%		\includemedia[
+%  		width=\textwidth,
+%  		height=0.8\textwidth,
+%  		activate=pageopen,
+%  		addresource=IRL.mp4,
+%  		flashvars={source=IRL.mp4}
+%		]{}{VPlayer.swf}
+%		}
+%\end{frame}
 

### 2018-09-23 Dissertation/Defense/phd_research_20min.tex
updated & submitted JINT18 paper

@@
 	\item Several sample future measurement rays and Dijkstra's cost map combine in an objective function for optimality
-	\item Numerical simulations show a large 2D environment, and this approach can be extended to 3D environments
+	\item A numerical simulation shows that this approach can be extended to 3D environments
 \end{itemize}
@@
 		\item Decentralize the algorithms so processes may be distributed among multiple agents
-		\vspace*{0.3cm}
-		\item Special thanks to the US Naval Research Lab for support under project ``Intelligent Microflyer''
-	\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Publications on Dissertation Topics}
-{\tiny 
-\begin{itemize}
-	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Multi-Vehicle Cooperative Exploration and Patrol of Uncertain 3D Environments'', in preparation.
-	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Bidding-Based Autonomous Exploration of Cooperative Robots Building a 3D Probabilistic Occupancy Grid Map'', in preparation.
-	\item E.
-Kaufman and T.
-Lee, ``Autonomous Aerial Exploration for Topological Mapping of Mars Environments'', \textit{2019 AIAA Guidance, Navigation, and Control Conference}, San Diego, CA, January 2019, submitted.
-	\item E.
-Kaufman, K.
-Takami, Z.
-Ai, and T.
-Lee, ``Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities,'' \textit{The Second IEEE International Conference on Robotic Computing}, pp.
-49--55, Laguna Hills, CA, January 2018.
-	\item E.
-Kaufman, K.
-Takami, T.
-Lee, and Z.
-Ai, ``Autonomous Exploration with Exact Inverse Sensor Models,'' \textit{Journal of Intelligent }\&\textit{ Robotic Systems}, 2017, doi: 10.1007/s10846-017-0710-7.
-	\item E.
-Kaufman, T.
-Lee, and Z.
-Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, pp.
-246--251, San Francisco, CA, December 2016.
-	\item E.
-Kaufman, T.
-Lee, Z.
-Ai, and I.
-S.
-Moskowitz, ``Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model,'' \textit{Proceedings of the American Control Conference}, pp.
-5709-5715, Boston, MA, July 2016.
-\end{itemize}
-}
-\end{frame}
-
-\begin{frame}
-\frametitle{Publications on Data Association, Estimation, and Control}
-{\tiny 
-\begin{itemize}
-	\item E.
-Kaufman, T.
-A.
-Lovell, and T.
-Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{The Journal of the Astronautical Sciences}, 63(1): pp.
-60-80, 2016, doi: 10.1007/s40295-015-0082-9.
-	\item E.
-Kaufman, T.
-A.
-Lovell, and T.
-Lee, ``Minimum Uncertainty JPDA Filters and Coalescence Avoidance for Multiple Object Tracking,'' \textit{The Journal of the Astronautical Sciences}, 63(4): pp.
-308-334, 2016, doi: 10.1007/s40295-016-0092-2.
-	\item T.
-Wu, E.
-Kaufman, and T.
-Lee, ``Globally Asymptotically Stable Attitude Observer on SO(3),'' \textit{Proceedings of the 54th IEEE Conference on Decision and Control}, pp.
-2164-2168, Osaka, Japan, December 2015.
-	\item E.
-Kaufman, T.
-A.
-Lovell, and T.
-Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting}, AAS 15-451, Williamsburg, VA, January 2015.
-	\item E.
-Kaufman, T.
-A.
-Lovell, and T.
-Lee, ``Minimum Uncertainty JPDA Filter and Coalescence Avoidance Performance Evaluations,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting, Williamsburg}, AAS 15-432, VA, January 2015.
-	\item E.
-Kaufman, K.
-Caldwell, D.
-Lee, and T.
-Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, ASC 14-2527, March 2014, Big Sky, MT.
-	\item E.
-Kaufman, T.
-A.
-Lovell, and T.
-Lee, ``Optimal Joint Probabilistic Data Association Filter Avoiding Coalescence in Close Proximity,'' \textit{Proceedings of the European Control Conference}, pp.
-2709-2714, Strasbourg, June 2014.
-\end{itemize}
-}
+	\end{itemize}
 \end{frame}

### 2018-09-23 Journal/JINT 2018/JINT18.tex
updated & submitted JINT18 paper

@@
 Another popular approach is based on simulated learning trials~\cite{Thr01,ThrBurFox05,SouMaiGon12} to determine an inverse sensor model.
-This approach is sensitive to simulation errors, and is not suitable for the higher dimensional configurations composed of maps, poses, and measurements due to the large number of required samples.
+This approach is complicated and inexact due to sensitivity of simulation errors.
 In short, none of the existing approaches use the sensor properties to directly obtain a Bayesian solution to occupancy probability.
@@
 The approximate nature of the occupancy probability, either by a simplified function or a learned solution, motivates an exact Bayesian solution to the inverse sensor model.
-In this paper, we propose a computationally-efficient algorithm to obtain the exact solution to the inverse sensor model.
-This serves to generate probabilities in a 3D environment following the framework of~\cite{KauLeeAiMos16,KauTakAiLee17}.
-The proposed approach integrates stochastic properties of all sensors and the prior map estimate, following a Bayesian framework, to obtain the exact 3D occupancy probabilities~\cite{KauTakAiLee18}.
+In this paper, we propose a computationally-efficient algorithm to obtain this solution, which serves to generate probabilities in a 3D environment following the framework of~\cite{KauLeeAiMos16,KauTakAiLee17}.
+This approach approach integrates stochastic properties of all sensors and the prior map estimate, following a Bayesian framework, to obtain the exact 3D occupancy probabilities~\cite{KauTakAiLee18}.
 As such, it provides a unified framework for sensor fusion between multiple sensors with various properties, such as sensor accuracy and fields of view.
@@
 \label{subsec:POGM}
-Let map $m$ be an occupancy grid map decomposed into $n_m$ evenly-spaced cubic grid cells with fixed edge length $\alpha$.
-The $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,\ldots,n_m}$, that is defined as $\mathbf{m}_i=1$ when occupied, and $\mathbf{m}_i=0$ when free.
-Let another random variable be defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience, which is the event that the $i$-th cell is free, i.e., $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$.
-The location and size of each grid cell is known, and the random variables composing $m$ are assumed mutually independent, i.e., $P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i)$, so $2^{n_m}$ maps are possible.
-
-Consider a range sensor that repeatedly scans the environment surrounding the robot and returns the distances to the closest occupied spaces.
-Let the measurement scan at the $t$-th time step be $Z_t$, where $z\in Z_t$ is a single measurement ray.
-Let the $t$-th pose be denoted by $X_t=\braces{x_t,R_t}$ with location $x_t\in\Re^3$ and attitude $R_t\in\SO=\braces{R\in\Re^{3\times3}|R^\T R=I,\det{R}=1}$, such that $X_t$ is assumed to be known.
-Given the sensor probabilistic properties $p(z|X_t,m)$, commonly referred to as the \emph{forward sensor model}, the goal is finding the \emph{inverse sensor model}, $P(m|z,X_{1:t},Z_{1:t-1})$.
-The authors of~\cite{KauLeeAiMos16,KauTakAiLee17} proposed a reduced map $r\subset m$ composed of $n_r$ cells along $z$, where cells are indexed by increasing distance from $x_t$.
-Defining $\mathbf{r}_{k+}$ as the event that the $k$-th cell of map $r$ is the closest occupied space, the inverse sensor model for the $k$-th cell is
+An occupancy grid map $m$ is composed of $n_m$ grid cells, where each grid cell is a cube with fixed edge length $\alpha$.
+Occupancy grid cells are considered static binary random variables, i.e., the $i$-th grid cell $\mathbf{m}_i=1$ if occupied or $\mathbf{m}_i=0$ if free (unoccupied).
+For convenience, we define its complement as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ such that $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$.
+Since we assume grid cell occupancies are mutually independent, $2^{n_m}$ map combinations are possible.
+
+The occupancy grid map cell probabilities are repeatedly updated by a range sensor.
+Through the $t$-th time step, we assume the history of poses $X_{1:t}$ and measurement scans $Z_{1:t}$ are given.
+The $t$-th pose $X_t=\braces{x_t,R_t}$ consists of position $x_t\in\Re^3$ and attitude $R_t\in\SO=\braces{R\in\Re^{3\times3}|R^\T R=I,\det{R}=1}$.
+However, measurements are subject to random noise; let individual measurement ray $z\in Z_t$ be characterized by the \emph{forward sensor model}, which is a probability density $p(z|X_t,m)$, known either from the sensor manufacturer or obtained empirically.
+The goal of probabilistic occupancy grid mapping is using the stochastic properties of the forward sensor model to obtain the \emph{inverse sensor model}, $P(m|z,X_{1:t},Z_{1:t-1})$.
+
+\sloppy
+The inverse sensor model problem was solved in~\cite{KauLeeAiMos16,KauTakAiLee17} to provide a real-time solution that uses the forward sensor model directly with map invariance.
+The process is fairly straightforward: define a reduced map $r\subset m$ of $n_r$ grid cells along $z$, indexed by increasing distance.
+Let $\mathbf{r}_{k+}$ correspond to the $k$-th cell being the closest occupied space.
+Then, the inverse sensor model is
 \begin{align}
@@
 \end{align}
-such that the probability summation of the entire space is $1$, and the unnormalized probability of the inverse sensor model is defined as
+such that the summation of any cell probability and its complement is $1$. The unnormalized probability of the inverse sensor model is defined as
 \begin{align}
 \label{eqn:Unnormalized}
-\tilde P(\mathbf{r}_{k}|z,X_{1:t},&Z_{1:t-1})\nonumber\\
+\tilde P(\mathbf{r}&_{k}|z,X_{1:t},Z_{1:t-1})\nonumber\\
 &=\mathbf{P}_k^-
-\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]\nonumber\\
-&\quad + \bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{k+},X_t)\mathbf{P}_k^-,
-\end{align}
-where a priori probability is $\mathbf{P}_k^-=P(\mathbf{r}_{k}|X_{1:t-1},Z_{1:t-1})$, its complement is $\bar{\mathbf{P}}_k^-=1-\mathbf{P}_k^-$, $P(\bar{\mathbf{r}}_{0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience, and $p(z|\mathbf{r}_{(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading.
-The proofs of \refeqn{RayISMAnswer}--\refeqn{Unnormalized} are given in~\cite{KauLeeAiMos16}.
-Since $\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})$ from \refeqn{Unnormalized} uses several repeated terms from the prior $\tilde P(\mathbf{r}_{k-1}|z,X_{1:t},Z_{1:t-1})$, and $\eta$ is easily obtained from these as well, the computational cost of \refeqn{RayISMAnswer} is linear with respect to the number of cells along a measurement ray, amortized to $\mathcal{O}(1)$ for each cell, so the exact inverse sensor model can be applied in real-time.
-The process of updating each cell along a measurement ray is repeated for all rays composing scan $Z_t$.
+\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]+ \bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{k+},X_t)\mathbf{P}_k^-,
+\end{align}
+where probability $\mathbf{P}_k^-=P(\mathbf{r}_{k}|X_{1:t-1},Z_{1:t-1})$ and its complement $\bar{\mathbf{P}}_k^-=1-\mathbf{P}_k^-$ are known a priori, we select $P(\bar{\mathbf{r}}_{0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience, and the forward sensor model maximum reading case is $p(z|\mathbf{r}_{(n_r+1)+},X_t)$.
+The proofs are given in~\cite{KauLeeAiMos16} for \refeqn{RayISMAnswer}--\refeqn{Unnormalized}.
+Since these equations use several repeated terms, \refeqn{RayISMAnswer} exhibits linear complexity with respect to the number of cells along a measurement ray, amortized to $\mathcal{O}(1)$ for each cell, so the exact inverse sensor model can be applied in real-time.
 
@@
 
-Here we present how the probabilistic properties of an occupancy grid map provide key information about the uncertainty of the space for motion planning.
-Shannon's entropy commonly serves as an uncertainty measure~\cite{StaGriBur05}.
-Provided grid cell probabilities of map $m$, Shannon's entropy is defined as
+Here we introduce Shannon's entropy as a measure of probabilistic map uncertainty~\cite{StaGriBur05}, which provides key information for optimal motion planning. Shannon's entropy is defined as
 \begin{align}
@@
 \end{align}
-for an individual cell and the entire map, respectively.
-Thus, entropy is maximized when $P(\mathbf{m}_i)=0.5$, which corresponds to the largest uncertainty; similarly, entropy is minimized as $P(\mathbf{m}_i)$ approaches $0$ or $1$, which corresponds to the smallest uncertainty.
-%; thus Shannon's entropy is a measure of uncertainty.
+for a single grid cell and the entire map, respectively.
+Thus, $P(\mathbf{m}_i)=0.5$ maximizes \refeqn{ShannonsEntropyCell}, corresponding to the largest uncertainty; similarly, \refeqn{ShannonsEntropyCell} is minimized as $P(\mathbf{m}_i)$ approaches $0$ or $1$, corresponding to the smallest uncertainty.
 
 Suppose a future pose candidate and its associated future measurement scan is $X_c=\braces{x_c,R_c}$ and $Z_c$, respectively, where $c\in\mathcal C$ such that $\mathcal C=\braces{1,2,...,n_c}$ accounts for all $n_c$ candidates under consideration for the next robot pose.
-The current pose of the robot is \emph{not} $X_c$ in general, so any change to the probabilistic map from $X_c$ must be predicted.
-Much like the probabilistic mapping, this is achieved ray-by-ray~\cite{KauAiLee16,KauTakAiLee17}.
-Considering that all grid cell probabilities are conditioned on the history of poses $X_{1:t}$ and measurement scans $Z_{1:t}$, these terms are removed from the remaining equations for simplicity.
-For candidate ray $z_c$, the expected entropy is
+The current pose of the robot is \emph{not} $X_c$ in general, so changes to the probabilistic map from $X_c$ can only be predicted.
+This is achieved by considering the expected entropy change of individual measurement rays~\cite{KauAiLee16,KauTakAiLee17}.
+Removing $X_{1:t}$ and $Z_{1:t}$ for simplicity, the expected entropy for candidate ray $z_c$ is
 \begin{align}
@@
 \end{align}
-where $z_{c,k}$ refers to the distance from $x_c$ to the $k$-th grid cell along the measurement ray.
-The first term of the summation of \refeqn{DiscExpEntropyRay}, namely $H(P(m|x_c,z_{c,k}))$, is obtained with entropy definitions \refeqn{ShannonsEntropyCell}, \refeqn{ShannonsEntropyMap} and the inverse sensor model \refeqn{RayISMAnswer}--\refeqn{Unnormalized}.
+where $z_{c,k}$is distance along measurement ray $z_{c}$ from position $x_c$ to the $k$-th grid cell.
+The first term in \refeqn{DiscExpEntropyRay}, namely $H(P(m|x_c,z_{c,k}))$, is found with the inverse sensor model \refeqn{RayISMAnswer}--\refeqn{Unnormalized} and grid cell entropy \refeqn{ShannonsEntropyCell}--\refeqn{ShannonsEntropyMap}.
 The second term is derived from \refeqn{allEta} with
@@
 \end{align}
-Then, \refeqn{DiscExpEntropyRay} is applied repeatedly over several sample measurement rays that compose scan $Z_c$ as shown in~\cite{KauAiLee16}.
+Then, we repeat \refeqn{DiscExpEntropyRay} for several sample measurement rays composing scan $Z_c$~\cite{KauAiLee16}.
 
@@
 This paper introduces a near-optimal bidding-based solution to multi-vehicle exploration and patrol to map a 3D space.
-The proposed approach integrates recent developments in obtaining occupancy grid probabilities and entropies into several auctions to assign tasks to robots.
+This approach integrates recent developments in obtaining occupancy grid probabilities and entropies into several auctions to assign tasks to robots.
 The bidding process relies on expected map information gains and travel distances, and modifies bids between auctions to promote robotic cooperation.

### 2019-01-09 Conference/ACC 2019.2/ACC19.tex
added mars presentation

--- 2018-06-09 Conference/ICRA 2018.1/ICUAS18.tex
@@
-\documentclass[letterpaper, 10pt, conference]{ieeeconf}
-\IEEEoverridecommandlockouts \overrideIEEEmargins
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%2345678901234567890123456789012345678901234567890123456789012345678901234567890
+%        1         2         3         4         5         6         7         8
+
+\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
+
+%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
+
+\IEEEoverridecommandlockouts                              % This command is only needed if 
+                                                          % you want to use the \thanks command
+
+\overrideIEEEmargins                                      % Needed to meet printer requirements.
+
 \usepackage{amsmath,amssymb,url,times,subfigure,graphicx,theorem}
@@
 
-%\usepackage{algorithmicx}
-%\usepackage{algpseudocode}
-
-%\usepackage{algorithm}
-%\usepackage{algpseudocode}
-%\usepackage{pifont}
-
-%\usepackage{amsmath}
-%\usepackage{algorithm}
-%\usepackage[noend]{algpseudocode}
-%
-%\makeatletter
-%\def\BState{\State\hskip-\ALG@thistlm}
-%\makeatother
-
-%\definecolor{red}{rgb}{0,0,0}
 \graphicspath{{../../Fig/}}
@@
 
-\newcommand{\EditEK}[1]{{\color{blue}\protect #1}}
-%\renewcommand{\EditTL}[1]{{\protect #1}}
-
 \newtheorem{definition}{Definition}
@@
 \begin{document}
-\allowdisplaybreaks
 
-\maketitle \thispagestyle{empty} \pagestyle{empty}
+\maketitle
+\thispagestyle{empty}
+\pagestyle{empty}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
+
 This paper presents a bidding-based cooperative robotic exploration strategy to generate accurate occupancy grid maps with multiple vehicles.
@@
 This multi-vehicle exploration approach is demonstrated with a numerical simulation where three robots simultaneously explore a large building floor plan cooperatively.
+
 \end{abstract}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-
-% Multivehicle MAPPING in 3D 
-% what is mapping?
-% Occupancy grid mapping this way (ACC, JIRS, IRC)
-% Designed to work with multiple sensors
-% 
-% Exploration:
-% Frontier, entropy-based
-% Recent papers (SIMPAR, JIRS, IRC)
-% Main idea: maximizing a reward
-% Multivehicle: 
-%
 
@@
 
-%In short, the robot with the largest bid wins the first auction, then discounts all candidate poses within a small neighborhood of the auction-winning robot, and this robot is removed from further auctions.
-This process is repeated, removing a robot from consideration each auction, until just a single robot remains.
-
 The proposed approach uses auctions and bid modifications based on total map information gain and collision-avoidance, promoting exploration of different spaces without explicit consideration of coverage overlaps.
@@
 	\end{figure}
-
-%              C++
-%              if(distToOptimalPose < explore.discountRadius)
-%                proximityDiscount = pow(distToOptimalPose/explore.discountRadius, 2);
-%              else
-%                proximityDiscount = 1.0;
-
-%\begin{align}
-%\label{eqn:bumpFun}
-%\text{bump}(d_\text{cell})= 
-%\begin{cases}
-%    e^{1-\frac1{1-(d_\text{cell}/d_\text{max})^2}},			& \text{if }-d_\text{max}<d_\text{cell}<d_\text{max}\\
-%    0,              				& \text{otherwise}
-%\end{cases},
-%\end{align}
 
@@
 		\centerline{
-			\subfigure[Floor Plan]{\includegraphics[height=0.5\columnwidth]{SEH_cropped.png}}
+			\subfigure[Floor Plan]{\includegraphics[height=0.5\columnwidth]{SEH_cropped.jpeg}}
 			\hspace*{0.025\columnwidth}
@@
 		\centerline{
-			\subfigure[$0$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_0min.jpg}}
+			\subfigure[$0$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_0min_ss.jpg}}
 			\hspace*{0.025\columnwidth}
-			\subfigure[$1$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_1min.jpg}}
+			\subfigure[$1$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_1min_ss.jpg}}
 		}
@@
 		\centerline{
-			\subfigure[$2$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_2min.jpg}}
+			\subfigure[$2$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_2min_ss.jpg}}
 			\hspace*{0.025\columnwidth}
-			\subfigure[$3$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_3min.jpg}}
+			\subfigure[$3$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_3min_ss.jpg}}
 		}
@@
 		\centerline{
-			\subfigure[$4$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_4min.jpg}}
+			\subfigure[$4$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_4min_ss.jpg}}
 			\hspace*{0.025\columnwidth}
-			\subfigure[$5$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_5min.jpg}}
+			\subfigure[$5$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_5min_ss.jpg}}
 		}
@@
 		\centerline{
-			\subfigure[$6$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_6min.jpg}}
+			\subfigure[$6$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_6min_ss.jpg}}
 			\hspace*{0.025\columnwidth}
-			\subfigure[$7$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_7min.jpg}}
+			\subfigure[$7$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_7min_ss.jpg}}
 		}
@@
 		\centerline{
-			\subfigure[$8$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_8min.jpg}}
+			\subfigure[$8$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_8min_ss.jpg}}
 			\hspace*{0.025\columnwidth}
-			\subfigure[$9$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_9min.jpg}}
+			\subfigure[$9$ min]{\includegraphics[trim={22cm 5cm 16cm 3.5cm}, clip, height=0.45\columnwidth]{multi_discount_raw_2D_9min_ss.jpg}}
 		}
@@
 \end{document}
-
-\end{document}

### 2019-01-09 Dissertation/Defense/scitech19_presentation.tex
added mars presentation

--- 2018-09-23 Dissertation/Defense/phd_research_20min.tex
@@
   \leavevmode%
-  \hbox{\begin{beamercolorbox}[wd=.7\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{author in head/foot}%
+  \hbox{\begin{beamercolorbox}[wd=.85\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{author in head/foot}%
     \usebeamerfont{author in head/foot}\insertshorttitle
   \end{beamercolorbox}%
-  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm]{title in head/foot}
+  \begin{beamercolorbox}[wd=.15\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm]{title in head/foot}
     \usebeamerfont{title in head/foot}\mypaper\hfill \insertframenumber/\inserttotalframenumber
-    \usebeamerfont{title in head/foot}\hfill \insertframenumber/\inserttotalframenumber
   \end{beamercolorbox}}%
@@
 
-\title[Multi-Robot Probabilistic Mapping and Exploration]{\large Multi-Robot Probabilistic Mapping and Exploration}
+\title[Autonomous Aerial Exploration for Topological Mapping of Mars Environments]{\large Autonomous Aerial Exploration for Topological Mapping of Mars Environments}
 
@@
 \vspace*{0.3cm}
-\normalsize Evan Kaufman}\\
-\vspace*{0.3cm}
-Robotic Research LLC
+\normalsize Evan Kaufman and Taeyoung Lee}
 \\
 \vspace*{0.3cm}
-September 5, 2018
+January 10, 2019
 }
@@
 \begin{itemize}
-	\item Mobile Robotics
+	\item Topography of Mars
 	\begin{itemize}
-		\item Robots can replace humans dangerous/uncertain tasks such as surveillance, combat, and search-and-rescue
-		\item Key question: when robots explore new spaces, how can they maximize their knowledge about their environment?
+		\item Future mission planning
+		\item Spacecraft landing
+		\item Image distortion correction
+		\item The Mars Orbiting Laser Altimeter (MOLA) uncertainty is about $8$ meters
 	\end{itemize}
 	\pause
-	\item Robotic Capabilities
+	\item Mars Rovers for Mapping Mars
 	\begin{itemize}
-		\item Robot must understand their environment by generating a map, referred to as \emph{mapping}
-		\item Selecting and performing robotic actions to maximize map information, referred to as \emph{autonomous exploration}
-		\item Cooperative robots can share information to achieve a collective goal with \emph{multi-vehicle exploration and patrol}
-		\item These tasks must be performed in 2D and 3D environments, where computational efficiency is paramount to real-time implementation
+		\item Four successful driving missions, Curiosity still functional and communicating
+		\item Driving robots are incapable of traversing steep or dangerous terrain
+		\item Humans cannot control rovers directly due to a time delay from Earth
 	\end{itemize}
@@
 		\begin{itemize}
-		\item Computation of the exact occupancy probability
+		\item Computation of the exact occupancy probability of the initially-uncertain Mars environment
 		\item Incorporation of prior knowledge and new measurements
-		\item Bayesian information fusion
+		\item Bayesian information fusion in 3D
 	\end{itemize}
@@
 		\item Optimal planning to maximize information gain
-	\end{itemize}
-	\pause
-	\item Multi-Vehicle Exploration and Patrol
-	\begin{itemize}
-		\item Bidding-based framework for assigning tasks while exploring uncertain environments
-		\item Map degradation to encourage revisiting regions from the past 
+		\item Complicated 3D Mars environment explicitly considered in pose selection and path planning
 	\end{itemize}
@@
 
-\begin{frame}
-\frametitle{Outline}
-\begin{enumerate}[I]
-	\item Occupancy Grid Mapping
-	\begin{itemize}
-		\item Background and problem definition
-		\item Common approaches and approximations
-		\item Inverse sensor model in 1D
-		\item Ray casting and measurement scans for 2D and 3D environments
-		\item Numerical examples
-	\end{itemize}
-	\pause
-	\item Autonomous Exploration
-	\begin{itemize}
-		\item Existing solutions: frontiers, expected measurements
-		\item Proposed approach: entropy expectation
-		\item Collision-free trajectories
-		\item Optimal pose selection
-		\item Numerical examples
-	\end{itemize}
-\end{enumerate}
-\end{frame}
-
-\begin{frame}
-\frametitle{Outline}
-\begin{enumerate}[I]
-\addtocounter{enumi}{2}
-	\item Multi-Vehicle Exploration and Patrol
-	\begin{itemize}
-		\item Cooperative approaches
-		\item Bidding-based framework
-		\item Receding horizon
-		\item Multi-vehicle autonomous exploration numerical example
-		\item Autonomous patrol
-		\item Multi-vehicle autonomous patrol numerical example
-	\end{itemize}
-	\pause
-	\item Experiments
-	\begin{itemize}
-		\item Ground vehicle 2D mapping and autonomous exploration
-		\item Aerial quadrotor in large vertically-uniform space
-		\item Aerial quadrotor in complex 3D environment
-	\end{itemize}
-\end{enumerate}
-\end{frame}
-
 \section*{}
 \subsection*{Probabilistic Occupancy Grid Mapping}
-
-\begin{frame}
-\frametitle{Probabilistic Occupancy Grid Mapping}
-\begin{enumerate}[I]
-	\item Probabilistic Occupancy Grid Mapping
-\end{enumerate}
-\end{frame}
 
@@
 	\pause
-	\item Useful for collision-avoidance and graph searching
+	\item Useful for collision-avoidance and path-planning
     \end{itemize}
@@
 \begin{frame}
-\frametitle{Approximate Solutions}
-
-\begin{minipage}[t]{7.0cm}
-\begin{itemize}
-	\item Approximate a function for the inverse sensor model based on intuition
-	\begin{itemize}
-		\item Simple to implement, but mathematically inaccurate
-	\end{itemize}
-\end{itemize}
-\end{minipage}
-\begin{minipage}[t]{3.0cm}
-\begin{figure}[!htbp]
-\centerline{
-    \vspace*{-0.5cm}
-%	\hspace*{0.25cm}
-    \includegraphics[width=4.0cm]{Approx_ISM_shortened.png}\hspace*{0.1cm}
-    }
-\centerline{
-    \includegraphics[width=3.5cm]{1D_True_Grid.png}
-%    \hspace*{0.75cm}
-    }
-		\end{figure}
-\end{minipage}
-\vspace*{-0.5cm}
-\vspace*{0.0cm}\pause
-\begin{itemize}
-	\item Find a solution through learning
-	\begin{itemize}
-		\item Simulate maps, poses, and measurements and use learning to obtain an inverse sensor model
-		\item Complicated, unclear how these parameters are chosen
-	\end{itemize}
-	\vspace*{0.0cm}\pause
-	\item Goal: design a \emph{simple} and \emph{accurate} occupancy grid mapping method avoiding log-odds ratio assumptions
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Proposed Approach: Utilize Invariance}
+\frametitle{Inverse Sensor Model with Invariance}
 
@@
 	\end{itemize}
-\end{itemize}
 \setcounter{subfigure}{0}
 % uncomment
+\vspace*{-0.03\textwidth}
 \begin{figure}
@@
 \end{figure}
+	\item Avoids common approximated or learned solutions
+\end{itemize}
 
@@
 
-% Ray Casting
-
-%\begin{frame}
-%\frametitle{Ray Casting}
-%
-%\begin{itemize}
-%	\item Required for 2D and 3D occupancy grid maps
-%	\item Main idea: use simple geometry to find the distance from the robot sensor to each cell intersected by the measurement ray
-%\end{itemize}
-%
-%\begin{figure}[!ht]
-%    \centering
-%    \begin{subfigure}[t]{0.4\columnwidth}
-%        \centering
-%        \includegraphics[width=\textwidth]{RayCastingIllustration.png}
-%%        \label{fig:penn_map_total}
-%    \end{subfigure}
-%    \hspace*{0.05\columnwidth}
-%    \begin{subfigure}[t]{0.4\columnwidth}
-%        \centering
-%        \includegraphics[width=\textwidth]{RayCastIllustrationReducedMapOnly.png}
-%%        \label{fig:penn_map_zoom}
-%    \end{subfigure}
-%\end{figure}
-%
-%\end{frame}
-
-%\begin{frame}
-%\frametitle{Scan Combination Approaches}
-%
-%\begin{itemize}
-%	\item \emph{Ray-By-Ray Update}: given a scan of measurement rays, update the map based on each measurement ray \emph{individually} and \emph{sequentially}
-%	\begin{itemize}
-%		\item Benefit: the measurement rays maintain some dependency as they capture the same map
-%		\item Drawback: the best order for evaluating rays is not necessarily chosen
-%	\end{itemize}
-%%	\item Conditional probability:
-%%	\begin{align*}
-%%		P(\mathbf{m}_i|X_{1:t}&,Z_{1:t})\nonumber\\&
-%%		=P((\dots(((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z})
-%%	\end{align*}
-%	\item \emph{Synergistic Update}: consider every measurement ray inside a scan \emph{simultaneously} to update all grid cells inside the FOV
-%	\begin{itemize}
-%		\item Benefit: synergistic update provides a method where all rays are considered simultaneously
-%		\item Drawback: required assumption that the measurement rays are independent, though they capture the same map
-%	\end{itemize}
-%\end{itemize}
-%
-%\end{frame}
-
-\begin{frame}
-\frametitle{Numerical Example}
-
-\begin{itemize}
-	\item A standard \emph{approximate} inverse sensor model technique is compared with the proposed \emph{exact} method
-	\item The exact algorithm yields a more \emph{certain} map from the same measurement set
-\end{itemize}
-
-%\vspace*{0.5cm}
-\begin{figure}[!htbp]
-    \centering
-    \begin{subfigure}{0.2\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{AISM_Image_19.pdf}
-        \caption*{Approx. Map}
-    \end{subfigure}
-%    \hspace*{-0.1\textwidth}
-    \begin{subfigure}{0.2\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{EISM_Image_19.pdf}
-        \caption*{Exact Map}
-    \end{subfigure}
-    \begin{subfigure}{0.2\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{AISM_Image_inf_19.pdf}
-        \caption*{Approx. $H$}
-    \end{subfigure}
-%    \hspace*{-0.1\textwidth}
-    \begin{subfigure}{0.2\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{EISM_Image_inf_19.pdf}
-        \caption*{Exact $H$}
-    \end{subfigure}
-\end{figure}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Probabilistic Occupancy Grid Mapping Summary}
-
-\begin{itemize}
-	\item The proposed novel Bayesian solution to the inverse sensor model provides the exact solution without approximate functions or learned solutions
-	\item The computational complexity is linear with respect to the number of cells considered, making real-time implementation possible
-	\item 2D and 3D maps are updated with ray casting and ray-by-ray combination
-\end{itemize}
-
-\end{frame}
-
 \section*{}
 \subsection*{Entropy-Based Autonomous Exploration}
-
-\begin{frame}
-\frametitle{Entropy-Based Autonomous Exploration}
-\begin{enumerate}[I]
-\addtocounter{enumi}{1}
-	\item Entropy-Based Autonomous Exploration
-\end{enumerate}
-\end{frame}
 
@@
 \begin{frame}
-\frametitle{Expected Information Gain Computation}
-
-\begin{itemize}
-	\item \emph{Squared Complexity}: $\mathcal O(n_r^2)$ for $n_r$ cells along the ray
-	\vspace*{0.0cm}\pause
-	\item Two Approaches:
-	\begin{itemize}
-		\vspace*{0.0cm}\pause
-		\item Decrease the cells considered by selecting $n_p$ most probable cells: $\mathcal O(n_r^2)$ reduces to $\mathcal O(n_p^2)$
-		\vspace*{0.0cm}\pause
-		\item Use 3D cells in a 2D \emph{projection}
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{.3\linewidth}
-    \centering\includegraphics[height=.75\linewidth]{experiment_ogm3D_2min47sec.jpg}
-    \caption*{\ \ \ \ \ \ \ \ \ \ \ \ \ \ 3D Map}
-  \end{subfigure}
-  \hspace*{0.25\linewidth}
-  \begin{subfigure}[t]{.3\linewidth}
-    \centering\includegraphics[height=.75\linewidth]{experiment_2min47sec.jpg}
-    \caption*{2D Projection}
-  \end{subfigure}
-\end{figure}
-	\end{itemize}
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Pose Selection}
+\frametitle{Pose Information Gain in 2D}
 
@@
 
+\end{frame}
+
+\begin{frame}
+\frametitle{Pose Information Gain in 3D}
+
+\begin{itemize}
+	\item Several sample pitch angles per horizontal direction
+	\item Optimal attitude assuming level flight, i.e., zero roll/pitch
+	\item Consideration of grid cells above and below the vehicle
+\end{itemize}
+\begin{figure}[!t]
+	\centering
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim = {21cm 5cm 20cm 5cm}, clip, height=1.0\textwidth]{mars_general.png}
+        		\caption{General View}
+    	\end{subfigure}
+    	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim = {25cm 6cm 16cm 4cm}, clip, height=1.0\textwidth]{mars_psi.png}
+        		\caption{$\psi$ Rotation}
+    	\end{subfigure}
+	\begin{subfigure}[t]{0.3\columnwidth}
+           	\centering
+          	\includegraphics[trim = {23cm 4cm 17cm 6cm}, clip, height=1.0\textwidth]{mars_theta.png}
+        		\caption{$\theta$ Rotation}
+    	\end{subfigure}
+%\includegraphics[width=2.5in]{myfigure}
+% where an .eps filename suffix will be assumed under latex, 
+% and a .pdf suffix will be assumed for pdflatex; or what has been declared
+% via \DeclareGraphicsExtensions.
+\label{fig:transforms}
+\end{figure}
 \end{frame}
@@
 
-%\begin{frame}
-%	\frametitle{2D Autonomous Exploration: Intel Research Lab}
-%	\centering{
-%		\includemedia[
-%  		width=\textwidth,
-%  		height=0.8\textwidth,
-%  		activate=pageopen,
-%  		addresource=IRL.mp4,
-%  		flashvars={source=IRL.mp4}
-%		]{}{VPlayer.swf}
-%		}
-%\end{frame}
+\begin{frame}
+\frametitle{Optimal Pose Selection in 3D}
+
+\begin{itemize}
+	\item Apply Dijkstra's algorithm about 3D graph of occupancy grid
+	\item Locations and associated distance costs for reachable positions in 3D
+\end{itemize}
+\begin{figure}
+	\centering
+	\begin{subfigure}[t]{0.45\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{mars_closeup.png}
+        		\caption{Mars Simulation Poses}
+    	\end{subfigure}
+	\begin{subfigure}[t]{0.45\columnwidth}
+           	\centering
+          	\includegraphics[height=0.5\textwidth]{disconnected_costmap.png}
+        		\caption{3D Dijkstra's Search}
+    	\end{subfigure}
+\label{fig:marsZoomedIn}
+\end{figure}
+\end{frame}
 
 \begin{frame}
-	\frametitle{3D Autonomous Exploration: Mars}
+	\frametitle{3D Mars Exploration Conclusions}
+	\begin{itemize}
+		\item Expected information gain of several poses in 3D space capturing complex obstacles
+		\item Distance cost accounts for collision-avoidance
+		\item Mapping and exploration are computationally-efficient for real-time implementation
+	\end{itemize}
+		\begin{figure}
+		\centerline{
+			\includegraphics[width=0.8\columnwidth]{node_diagram_3D_mars.png}
+		}
+		\label{fig:NodesDiagram}
+	\end{figure}
+\end{frame}
+
+\section*{}
+\subsection*{Numerical Simulation \& Experiment}
+
+\begin{frame}
+	\frametitle{3D Autonomous Exploration Mars}
+	\begin{itemize}
+		\item A color image of Mars is draped over its height map using the Gazebo simulator
+		\item Mapping and autonomous exploration are run in real-time using ROS nodes
+		\item 3D depth measurements are simulated with Gaussian noise
+		\item The robot is free to move within free space inside a fixed 3D volume
+	\end{itemize}
+	\begin{figure}[!t]
+		\centering
+		\begin{subfigure}[t]{0.3\columnwidth}
+           		\centering
+          		\includegraphics[height=0.9\textwidth]{mars_heightmap.png}
+        			\caption{Height Map}
+    		\end{subfigure}
+    		\begin{subfigure}[t]{0.3\columnwidth}
+			\centering
+			\includegraphics[height=0.9\textwidth]{mars_colormap.png}
+      	  		\caption{Color Map}
+   	 	\end{subfigure}
+  	  	\begin{subfigure}[t]{0.3\columnwidth}
+			\centering{
+          		\includegraphics[height=0.9\textwidth]{mars_gazebo_full_lighting.png}}
+        			\caption{Gazebo}
+    		\end{subfigure}
+	\label{fig:MarsGazebo}
+	\end{figure}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Mars Video}
 	\centering{
@@
 \begin{frame}
-\frametitle{Autonomous Exploration Summary}
-
-\begin{itemize}
-	\item Other approaches use heuristic frontiers or entropy of expected measurements
-	\item The proposed approach solves for the expected value of future map entropy directly
-	\item Several sample future measurement rays and Dijkstra's cost map combine in an objective function for optimality
-	\item A numerical simulation shows that this approach can be extended to 3D environments
-\end{itemize}
-
-\end{frame}
-
-\section*{}
-\subsection*{Multi-Vehicle Exploration and Patrol}
-
-\begin{frame}
-\frametitle{Multi-Vehicle Exploration and Patrol}
-\begin{enumerate}[I]
-\addtocounter{enumi}{2}
-	\item Multi-Vehicle Exploration and Patrol
-\end{enumerate}
-\end{frame}
-
-\begin{frame}
-\frametitle{Motivation and Goal}
-
-\begin{itemize}
-	\item Multi-Vehicle Autonomous Exploration
-	\begin{itemize}
-		\item Multiple agents can accomplish tasks faster and more reliably than a single agent
-		\item Large groups are capable of coverage in large environments
-		\item Problem: multi-vehicle map information gain optimality can be computationally-intractable
-		\item Approach: bidding-based solution simplifies a complicated optimization into a series of simplified optimizations
-	\end{itemize}
-	\pause
-	\item Multi-Vehicle Autonomous Patrol
-	\begin{itemize}
-		\item Patrolling large environments with humans is expensive and sometimes dangerous
-		\item Several robotic approaches introduce new variables for patrol
-		\item Approach: degrade probabilistic occupancy grid over time to encourage revisiting past regions
-	\end{itemize}
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Related Work in Multi-Vehicle Exploration}
-
-\begin{itemize}
-	\item Most existing bidding-based techniques deal with auctioning tasks among a few agents locally in decentralized frameworks
-	\item The other approach is for a central executive to run a series of auctions
-	\item Bids are discounted between auctions to coordinate the efforts based on coverage overlap
-	\item Heuristic, ad hoc
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Proposed Approach in Multi-Vehicle Exploration}
-
-\begin{itemize}
-\item Central executive completes $n$ auctions for $n$ robots
-\begin{enumerate}
-	\item Each robot determines its optimal future pose, and bids the objective function
-	\item The winner is awarded this future pose, and the remaining robots update their bids according to
-	\begin{itemize}
-		\item New expected map changes from the winner
-		\item Collision-avoidance with the winner
-	\end{itemize}
-	\item Repeat these steps with the remaining robots
-\end{enumerate}
-\item The efforts are coordinated by updating bids between auctions
-\item The algorithm follows a receding-horizon framework: future pose optimization repeats as quickly as possible
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Numerical Example: SEH Multi-Vehicle Exploration}
-
-	\centering{
-		\includemedia[
-  		width=\textwidth,
-  		height=0.8\textwidth,
-  		activate=pageopen,
-  		addresource=Multi-Vehicle Mapping and Exploration.mp4,
-  		flashvars={source=Multi-Vehicle Mapping and Exploration.mp4}
-		]{}{VPlayer.swf}
-		}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Cooperative Autonomous Patrol}
-
-\begin{itemize}
-	\item Existing Approaches
-	\begin{itemize}
-		\item Rely on a known map and use Voronoi diagrams to disperse
-		\item Another approach introduces cell periodicity: observe long-term dynamics of objects at the same location on the occupancy grid
-		\item Bio-inspired approach uses digital pheromones 
-		\item Minimizing robot idleness
-	\end{itemize}
-	\pause
-	\item Proposed Approach
-	\begin{itemize}
-		\item Degrade occupancy grid cells toward their initial values with exponential decay
-		\item Only affects existing probability values without new variables to track
-		\item Promotes exploration first, then patrol
-	\end{itemize}
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Continuous-Time Markov Process}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{markov_diagram_continuous.pdf}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.6\textwidth]{DegradeExamples.pdf}
-\end{figure}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Numerical Example: SEH Multi-Vehicle Patrol}
-
-	\centering{
-		\includemedia[
-    		width=\textwidth,
-  		height=0.8\textwidth,
-  		activate=pageopen,
-  		addresource=Multi-Vehicle Cooperative Patrol.mp4,
-  		flashvars={source=Multi-Vehicle Cooperative Patrol.mp4}
-		]{}{VPlayer.swf}
-		}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Cell Degradation and Patrol}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.65\textwidth]{entropy_patrol_9minSwitch.pdf}
-\end{figure}
-
-\begin{itemize}
-	\item Entropy of the map increases if patrol stops at $9$ minutes if cell degradation continues
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Multi-Vehicle Exploration and Patrol Summary}
-
-\begin{itemize}
-	\item Robotic tasks are assigned following a bidding-based framework to avoid a computationally-intractable solution
-	\item Only one robot wins an auction, and bids are updated for the remaining robots in subsequent auctions
-	\item A receding-horizon framework optimizes the time to update the optimal future poses
-	\item Cell degradation provides a simple means for autonomous patrol via a continuous-time Markov process
-	\item Numerical examples of a large experimental space demonstrate both algorithms
-\end{itemize}
-\end{frame}
-
-\section*{}
-\subsection*{Experiments}
-
-\begin{frame}
-\frametitle{Experiments}
-\begin{enumerate}[I]
-\addtocounter{enumi}{3}
-	\item Experiments
-\end{enumerate}
-\end{frame}
-
-\begin{frame}
-\frametitle{Experimental Overview}
-
-\begin{enumerate}
-	\item Ground vehicle experiment for 2D occupancy grid mapping and autonomous exploration
-	\item Aerial vehicle experiment in large 3D space resembling an office
-	\item Aerial vehicle experiment in a complex 3D space
-\end{enumerate}
-\end{frame}
-
-\begin{frame}
-\frametitle{Pioneer Ground Vehicle 2D Example}
-\begin{itemize}
-        	\item A pioneer ground vehicle explores a room composed of Styrofoam walls
-	\item Experiments use ROS, a Microsoft Kinect depth sensor, and Vicon Tracker for pose information
-\end{itemize}
-\begin{figure}
-	\centering
-    	\begin{subfigure}[b]{0.28\textwidth}
-        		\includegraphics[width=\textwidth]{pioneer.png}
-        		\caption*{Pioneer}
-    	\end{subfigure}    	
-	\hspace*{0.04\textwidth}
-	\begin{subfigure}[b]{0.28\textwidth}
-        		\includegraphics[width=\textwidth]{test_setup_1.jpg}
-        		\caption*{Bottom-left view}
-    	\end{subfigure}
-	\hspace*{0.03\textwidth}
-	\begin{subfigure}[b]{0.28\textwidth}
-        		\includegraphics[width=\textwidth]{test_setup_2.jpg}
-        		\caption*{Bottom-right view}
-    	\end{subfigure}
-\end{figure}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Ground Experiment Video}
-
-	\centering{
-		\includemedia[
-  		width=\textwidth,
-  		height=0.8\textwidth,
-  		activate=pageopen,
-  		addresource=NRL_ground_test.mp4,
-  		flashvars={source=NRL_ground_test.mp4}
-		]{}{VPlayer.swf}
-		}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Ground Vehicle Experimental Results}
-
-\begin{figure}
-	\centering{
-    	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_start.jpg}
-        		\caption*{$t=0.0$min}
-        		\label{fig:Experiment_ogm_t0}
-    	\end{subfigure}
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_0p5min.jpg}
-        		\caption*{$t=0.5$min}
-        		\label{fig:Experiment_ogm_t0p5}
-    	\end{subfigure}    
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1min.jpg}
-        		\caption*{$t=1.0$min}
-        		\label{fig:Experiment_ogm_t1}
-    	\end{subfigure}
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1p5min.jpg}
-        		\caption*{$t=1.5$min}
-        		\label{fig:Experiment_ogm_t1p5}
-    	\end{subfigure}
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2min.jpg}
-        		\caption*{$t=2.0$min}
-        		\label{fig:Experiment_ogm_t2}
-    	\end{subfigure}
-		\vspace*{0.05\textwidth}
-
-	}
-	\vspace*{-1cm}
-	\centering{
-    	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2p5min.jpg}
-        		\caption*{$t=2.5$min}
-        		\label{fig:Experiment_ogm_t2p5}
-    	\end{subfigure}
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3min.jpg}
-        		\caption*{$t=3.0$min}
-        		\label{fig:Experiment_ogm_t3}
-    	\end{subfigure}    
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3p5min.jpg}
-        		\caption*{$t=3.5$min}
-        		\label{fig:Experiment_ogm_t3p5}
-    	\end{subfigure}
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4min.jpg}
-        		\caption*{$t=4.0$min}
-        		\label{fig:Experiment_ogm_t4}
-    	\end{subfigure}
-	\begin{subfigure}[b]{0.19\textwidth}
-        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4p5min.jpg}
-        		\caption*{$t=4.5$min}
-        		\label{fig:Experiment_ogm_t4p5}
-    	\end{subfigure}
-	}
-\end{figure}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Aerial Vehicle in a Large Office-Like Environment}
-\begin{itemize}
-        	\item Hardware
-	\begin{itemize}
-		\item NVIDIA Jetson Ubuntu computer-on-module
-		\item Asus Xtion IR color depth sensor
-		\item Hokuyo LIDAR
-		\item VectorNav IMU
-	\end{itemize}
-	\item Software
-	\begin{itemize}
-		\item Robot Operating System (ROS)
-		\item Vicon Tracker (mocap\_vicon)
-		\item Geometric Control: coordinate-free attitude control
-	\end{itemize}
-\end{itemize}
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{.4\linewidth}
-    \centering\includegraphics[height=.5\linewidth]{quad_top.png}
-    \caption*{Quadrotor Top}
-  \end{subfigure}
-  \begin{subfigure}[t]{.4\linewidth}
-    \centering\includegraphics[height=.5\linewidth]{quad_bottom.png}
-    \caption*{Quadrotor Bottom}
-  \end{subfigure}
-\end{figure}
-\end{frame}
-
-\begin{frame}
-\frametitle{Aerial Vehicle in a Large Office-Like Environment}
-\begin{itemize}
-        	\item Environment Size: $10.8$ m $\times$ $11.8$ m $\times$ $1.65$ m
-\end{itemize}
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{.4\linewidth}
-    \centering\includegraphics[width=\linewidth]{experiment_north.jpg}
-    \caption*{North Region}
-  \end{subfigure}
-  \begin{subfigure}[t]{.4\linewidth}
-    \centering\includegraphics[width=\linewidth]{experiment_south.jpg}
-    \caption*{South Region}
-  \end{subfigure}
-\end{figure}
-\end{frame}
-
-\begin{frame}
-\frametitle{Large Flight Test Experiment Video}
-
-	\centering{
-		\includemedia[
-  		width=\textwidth,
-  		height=0.8\textwidth,
-  		activate=pageopen,
-  		addresource=NRL_3D_large_space.mp4,
-  		flashvars={source=NRL_3D_large_space.mp4}
-		]{}{VPlayer.swf}
-		}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Aerial Vehicle in a Complex 3D Environment}
+\frametitle{Experiment: Aerial Vehicle in a Complex 3D Environment}
 \begin{itemize}
@@
 \begin{itemize}
-        	\item Experimental environments in 2D and 3D are modeled with occupancy grids
-	\item Future entropy prediction and collision-free regions use these occupancy probabilities
-	\item A geometric controller is used during flight testing to follow the collision-free trajectory
+        	\item Environments in complex 3D are modeled with occupancy grids in simulation and experimentation
+	\item Predicted entropy of certain and uncertain spaces provide a policy for 3D autonomous exploration
+	\item Dijkstra's search provides a collision-free path and estimated travel costs that account for obstacles
 	\item All algorithms are run in real-time without expensive post-processing 
@@
 
-\begin{frame}
-\frametitle{Future Directions}
-	\begin{itemize}
-		\item Localization: use probabilistic map and other onboard sensors to localize, thereby completing the SLAM problem with the novel mapping approach
-		\item Computational improvements: determine collision-free costs more efficiently than Dijkstra's cost map
-		\item Decentralize the algorithms so processes may be distributed among multiple agents
-	\end{itemize}
-\end{frame}
-
 \end{document}

### 2019-01-09 Journal/JINT 2018/JINT18.tex
added mars presentation

@@
 \author{Evan Kaufman \and Kuya Takami \and \\ Zhuming Ai \and Taeyoung Lee}%
-%\thanks{Evan Kaufman, Taeyoung Lee, Mechanical and Aerospace Engineering, George Washington University, Washington DC 20052 {\tt \{evankaufman,tylee\}@gwu.edu}}
-%\thanks{Zhuming Ai, Ira S. Moskowitz, Information Management \& Decision Architectures, U.S. Naval Research Laboratory,  Washington, DC 20375}
-%}
-
-%\authorrunning{Short form of author list} % if too long for running head
 
@@
 \date{Received: date / Accepted: date}
-% The correct dates will be entered by the editor
 
@@
 
-%Mapping is of fundamental importance for robots to understand their environment, avoid collisions, and efficiently maneuver through their surroundings.
-%Several grid-based map representations have been studied, e.g.,~\cite{WolSuk05,MeyBeiBur12,TanThoWolBus14}, because they provide abundant occupancy knowledge about the environment for collision-free maneuvers and are computationally tractable.
-
 Mapping is an important and well-studied topic in robotics for understanding an environment, avoiding collisions, and efficiently planning robotic motions.
@@
 
-%        1.      sensor fusion in mapping
 The approximate nature of the occupancy probability, either by a simplified function or a learned solution, motivates an exact Bayesian solution to the inverse sensor model.
@@
 
-%        2.
-proposing a method to combine 3D mapping and 2D exploration: Also, we need to motivate why we are not doing full 3D exploration.
-Stress that we are proposing a computationally friendly approach to perform 3D mapping with projection using the fact there is less variation in the floor plan along the vertical direction.
 The 3D occupancy grid map provides key information to produce the policies that govern robotic motion, referred to as autonomous exploration.
@@
 
-%        3.      hardware/software for aerial exploration experiment.
 In this paper, we present a flight experiment for the proposed single-vehicle exploration scheme.
@@
 
-%In short, there are two contributions of this paper.
-First, single-vehicle 2D mapping and autonomous exploration approaches are extended to realistic 3D environments.
-Second, multi-vehicle cooperative patrol is proposed using bidding-based autonomous exploration and map grid cell degradation.
-These two concepts are demonstrated with an experiment and a simulation.
-
 \section{Problem Formulation}
@@
 \begin{align}
-%\label{eqn:objectiveFunOptimized}
-%X_c^*&=\argmax_{X_c}{\ \mathcal I(X_c)},
-%\\
 \label{eqn:expectedInfoGainRay}
@@
 
-% EditEK 1: example of 2 sensor measurements fused, starting with "For example, a robot may be equipped..." until the end of the paragraph.
-
 However, several careful considerations are required for 3D implementation.
@@
 
-%While auctions proceed, the occupancy grid cell probabilities are degrading according to a continuous-time Markov process.
-Cell degradation increases the expected information gain, thereby incentivizing robots to revisit regions after some time passes.
-
-%Here, we formulate a cooperative patrol scheme based on a coordinated autonomous exploration with multiple robots using sequential auctions.
-The first auction is based on the expected information gain presented in the prior section, scaled by travel distance.
-Once the winner of the first auction is determined, the map for bidding is updated based on the expected measurements of the first robot.
-The second auction for the remaining robots is determined by the information gain from the revised map that is further scaled by the travel distance and a penalty for collision-avoidance.
-This process is repeated until all robots are assigned tasks.
-While auctions proceed, the occupancy grid cell probabilities are degrading according to a continuous-time Markov process.
-Cell degradation increases the expected information gain, thereby incentivizing robots to revisit regions after some time passes.
-
 \subsection{Objective Function for the First Auction}
@@
 Second, the waypoints from a candidate pose to the current robot pose are easily obtained along the cost map using steepest descent.
-
-%    if x(k) < d
-%        f(k) = 0.5*fMax*(1-cos(pi*x(k)/d));
-%    else
-%        f(k) = fDiff*exp(-1e-2*(x(k)-d)^2)+fFar;
-%    end
-
-%Dijkstra's search demonstrates that the distances from each robot to $X_c$ are different in general, so travel times to $X_c$ vary.
-This motivates a unified approach to account for travel time, which promotes local travel and discourages robots from traversing across the map when possible.
 
@@
 
-\noindent{\bf Evan Kaufman} received his B.S.
-degree in Mechanical Engineering from Bucknell University in 2012, and is currently a Ph.D.
-candidate at The George Washington University in the Mechanical and Aerospace department.
+\noindent{\bf Evan Kaufman} is a robotics engineer at Robotic Research LLC.
+He received his B.S.
+degree in Mechanical Engineering from Bucknell University in 2012, and his Ph.D.
+in Aerospace Engineering at The George Washington University in 2018.
 His research focusses are in robotic mapping and exploration with range-based sensors, estimation, data association, and flight controls.
@@
 
-\noindent{\bf Kuya Takami} is a postdoctoral research fellow at the Department of Mechanical and Aerospace Engineering at the George Washington University.
+\noindent{\bf Kuya Takami} is a scientific software developer at Enthought and was a postdoctoral research fellow at the Department of Mechanical and Aerospace Engineering at the George Washington University.
 He received his doctoral degree in Mechanical Engineering at Virginia Polytechnic Institute and State University in 2015 specializing in robot perception.

### 2019-01-10 Dissertation/Defense/scitech19_presentation.tex
minor fixes to the e mars presentation... looking good after single readthrough

@@
 		\item Novel approach to predicting future map uncertainty
-		\item Optimal planning to maximize information gain
+		\item Optimal planning to maximize information gain for a \emph{flying} Mars robot
 		\item Complicated 3D Mars environment explicitly considered in pose selection and path planning
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{disconnected_costmap.png}
+          	\includegraphics[height=0.5\textwidth]{connected_costmap.png}
         		\caption{3D Dijkstra's Search}
@@
 	\begin{itemize}
-		\item Expected information gain of several poses in 3D space capturing complex obstacles
+		\item Expected information gain of several poses in 3D space capturing complex obstacles while flying
 		\item Distance cost accounts for collision-avoidance

### 2019-01-10 Dissertation/Defense/scitech19_presentation.tex
minor edits to prepresentation

@@
 		\item Driving robots are incapable of traversing steep or dangerous terrain
-		\item Humans cannot control rovers directly due to a time delay from Earth
+		\item Humans cannot control rovers directly in real-time due to a time delay from Earth
 	\end{itemize}
@@
 \begin{itemize}
-	\item Bayesian Framework
-	\begin{itemize}
-	\item Markov Assumption: latest a priori cell occupancy probabilities capture the information from all prior observations
-	\item Log-Odds Ratio: popular representation due to simple additive update structure and probability truncation avoidance, but requires the \emph{assumption}
-	\begin{align*}
-		P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})\approx P(Z_t|\mathbf{m}_i,X_t)
-	\end{align*}
-	\end{itemize}
+	\item Bayesian Framework with a Markov Assumption: latest a priori cell occupancy probabilities capture the information from all prior observations
 \vspace*{0.0cm}\pause
@@
 \end{align*}
-	\begin{itemize}
 	\item Given $n$ grid cells: $\mathcal O(2^n)$ is \emph{computationally-intractable}, motivating a different solution
-	\end{itemize}
 \end{itemize}
@@
 	\begin{itemize}
-		\item Each trajectory is solved independently as a function of time, i.e. $x(t)$, $y(t)$
+		\item Each trajectory is solved independently as a function of time, i.e. $x(t)$, $y(t)$, and $z(t)$
 		\item Starting and terminal locations are fixed
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{mars_closeup.png}
-        		\caption{Mars Simulation Poses}
+          	\includegraphics[height=0.5\textwidth]{connected_costmap.png}
+        		\caption{3D Dijkstra's Search}
     	\end{subfigure}
@@
            	\centering
-          	\includegraphics[height=0.5\textwidth]{connected_costmap.png}
-        		\caption{3D Dijkstra's Search}
+          	\includegraphics[height=0.5\textwidth]{mars_closeup.png}
+        		\caption{Mars Simulation Poses}
     	\end{subfigure}
@@
 \begin{frame}
-\frametitle{Experiment Summary}
+\frametitle{Simulation and Experiment Summary}
 \begin{itemize}
@@
         	\item Proposed an occupancy grid mapping technique that uses the \emph{exact probabilistic solution}
-	\item Computational cost is reduced \emph{substantially} for \emph{real-time implementation} using probabilisitic properties and exploiting mathematical patterns
-	\item Exact mapping concepts are extended to map uncertainty to accomplish \emph{autonomous exploration}
-	\item Multiple vehicles can explore and patrol an uncertain environment using a \emph{bidding}-based approach
-	\item These are demonstrated with numerous simulations and experiments
+	\item Exact mapping concepts are extended to map uncertainty to accomplish \emph{autonomous exploration} for a flying vehicle above the surface of Mars
+	\item These are demonstrated with a Mars simulation and a flight experiment
 \end{itemize}

### 2019-03-07 Journal/JINT 2018/JINT_ans.tex
added a reviewer response letter that needs to be filled in

--- 2018-06-09 Journal/JINT 2017/JIRS16_ans1.tex
@@
 
-\externaldocument[]{JIRS16_rev1}
+\externaldocument[]{JINT18}
 
@@
 
-\section*{Response to the Reviewers' Comments for JINT-D-16-00659}
+\section*{Response to the Reviewers' Comments for JINT-D-18-00265R2}
 
@@
 \subsection*{Reviewer 1}
-
-\setlength{\leftmargini}{0pt}
-\begin{itemize}\setlength{\itemsep}{2\parsep}
-
-\item {\itshape Reviewer \#1: The paper presents `probabilistic occupancy grid mapping and motion planning approaches such that a robot may build autonomously a map and explore a target area  in real time'.
-The sensor characteristics are explored probabilistically and the approaches are presented and summarized using algorithms.
-An Analysis of  computational efficient  is also presented together with numerical and experimental results.
-The news are described by authors and the experimental results help the readers to understanding the contributions.
-
-However, I have some questions for the authors;}
+\begin{itemize}
+\item {\itshape Reviewer \#1: The paper is well written.
+I suggest making the title more accurate rather than generic.
+The presented methods are specifically applicable for small flying vehicles(computational limitation) in structured indoor environment (as assumed in converting 3D to 2D occupancy grid) and the specific contribution is Bayesian grid mapping and receding horizons based patrol.
+There have been other papers in the literature addressing other grid mapping methods and other horizons for similar application.
+Consider a title such as ``Bayesian grid mapping based exploration and receding horizons based patrol of structured indoor environments using multi aerial vehicles''.}
 \end{itemize}
 
-%\renewcommand{\labelenumi}{\theenumi}
-%\renewcommand{\theenumi}{(\roman{enumi})}%
-
-\begin{enumerate}
-\item {\itshape Reviewer \#1: Figure 1 and the explanation in subsection ``Computationally Efficient Approach'' is not so clear for readers.
-The authors work with bi-dimensional space and present in figure an example in one-dimensional space.
-Also, the explanation in subsection is not clear and the second paragraph of the subsection is concluded with ``we present the following computational algorithm to compute the inverse sensor model'', but only the ``Proposition 1'' is seen in sequence.}
-
-The proposed algorithm to compute the exact inverse sensor model is applied to any of one, two, or three-dimensional depth measurements and mappings.
-The authors decided to use one-dimensional example in Figure 1, only to illustrate the proposed approach based on cell occlusion more explicitly through simpler cases.
-
-To clarify these, a few sentences are added to explain how the 1D concept is extended to 2D maps.
-Additionally, several other minor changes are added for clarity.
-Also, the last statement refers to Section \ref{sec:SA} for a summary of the proposed computational algorithm.
-
-\begin{correction}
-%
-We propose a computational algorithm to evaluate \refeqn{InvSenModWithProbDens} efficiently. 
-Since the cells outside of the sensor field of view (FOV) are not affected, we focus on a reduced map $r_l$ in the FOV of the $l$-th ray.
-This reduced map is chosen such that each cell of $r_l$ corresponds to a grid cell of map $m$ that the $l$-th ray intersects, ordered by increasing distance, which is easily determined from geometry.
-Let $\mathbf{r}_{l,k}$ be the binary random variable representing the occupancy of the $k$-th cell of the $l$-th ray.
-The number of cells in the reduced map is denoted by $n_{r,l}\leq n_m$.
-
-Next, let $\mathbf{r}_{l,k+}$ correspond to the event that the $k$-th cell of the $l$-th ray is occupied, cells with lower index (closer cells) are free, and cells with greater index (farther cells) may or may not be occupied, i.e., event $\mathbf{r}_{l,k+}$ occurs when \\$r\in\braces{r\in\{0,1\}^{{n_{r,l}}}\,|\mathbf{r}_{l,1}=0,\mathbf{r}_{l,2}=0,\ldots,\mathbf{r}_{l,k-1}=0,\mathbf{r}_{l,k}=1}$.
-In other words, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray.
-For example, see Figure \ref{fig:show_rkplus} that illustrates the event of $\mathbf{r}_{l,k+}$ when $l=1$ and $k=4$ for an one-dimensional cell array.
-This concept of grouping map outcomes can be easily extended to 2D using ray casting, where a 1D ray is spanning 2D space.
-The cells considered for the measurement ray are determined with the geometric intersection between the ray and map cell edges.
-Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell.
-Based on this, we present the mathematical expressions for the exact inverse sensor model, as summarized by \refeqn{RayISMAnswer} in Proposition 1.
-Later in Section \ref{sec:SA}, these are also rearranged as a computational algorithm.
-\end{correction}
-
-%The algorithm was written sequentially with LaTeX, but placed awkwardly in the PDF, which is now resolved.
-
-\item {\itshape Reviewer \#1: In page 13, the authors present two algorithms as table.
-Well, algorithms can not be table and seemly the authors are using latex.
-So, the authors must use the  algorithm2e package to present all algorithms into the paper and also to citing them suitably.}
-
-Both algorithms are rewritten using the suggested \texttt{algorithm2e} package and they are referenced properly in the revised manuscript.
-
-\item {\itshape Reviewer \#1: About the computational analysis, the authors talk in introduction of the paper that the approaches are to be used in real time.
-However, discussion about real time questions are not done into paper.
-For example: what information about each cell are stored in the memory in real time?
-For the experiment, what the computational load of the algorithms were?
-what time slice (tick) was?
-how much memories were used in the examples presented in figures 8 and 10?
-Well, I think that the discussion about real time could be improved.}\\
-
-The following discussion on real-time implementation and memory is added to Section 7 of the experimental results.
-
-%\EditTL{It is common to specify the processor and the programming language when presenting computation time. We may also add a remark that those were suitable for our experiments.}
-
-\begin{correction}The resulting map and exploration commands are generated in real time.
-The grid cells are stored as double-float variables in a vector, where the vector index is mapped to a location on the occupancy grid.
-Thus, the memory requirements are proportional to the number of grid cells, which is roughly the mapping area divided by the area of a grid cell.
-Since cell locations need not be saved, memory requirements are reduced.
-On a Lenovo T540p laptop with an Intel Core i7-4900MQ processor (quad-core, 2.8GHz per core) and 16GB of RAM, the mean time for updating the occupancy grid is $0.0115$ seconds, and the mean time to determine an exploration strategy is $0.6820$ seconds.
-Time requirements for mapping and exploration are easily modified by changing grid cell resolution or the number of exploration pose candidates, respectively.
-In short, there is a tradeoff between computational speed and accuracy of mapping or exploration.\end{correction}
-
-\end{enumerate}
-
-%\vspace*{0.1\textwidth}
-
-\clearpage\newpage
+This paper RETURN HERE!
 
@@
 \setlength{\leftmargini}{0pt}
+
 \begin{itemize}\setlength{\itemsep}{2\parsep}
 
-\item {\itshape Reviewer \#2: This paper serves to improve the accuracy occupancy grid mapping and the policy governing robotic motion during autonomous exploration.This paper provides an exact solution to this complicated probability problem using cell occlusion in the forward sensor model, which yields a substantially simpler inverse sensor model, which avoids a potentially harmful Markov assumption that commonly appears in the log-odd ratio formulations.
-There are some minor comments:}
-\end{itemize}
+\item {\itshape Reviewer \#2: The paper introduces a stochastic framework for autonomous exploration and patrol with multiple cooperating robots.
+Frankly, the contributions of the authors sound interesting.
+The paper is mostly well-written and easy to understand.
+However, the following comments have to be addressed:}
 
@@
 
-\item {\itshape Reviewer \#2: ``The robot motion is governed by the information gain-maximizing policy described in this paper.'' What is the advantage to maximize the information gain?
-In the field of motion control, the big control gain is unwanted when designing an algorithm.}
+\item {\itshape Reviewer \#2: A well-written abstract should briefly introduce a problem and suggest a solution for it.
+The authors should briefly introduce the problem that they are attempting to solve before suggesting a solution.
+However,  the abstract should be rewritten to introduce the importance of the authors' contributions without too much details.}
 
-The objective of exploration pursued in this paper is to build an accurate probabilistic map in an optimal fashion.
-As described in \refeqn{ShannonsEntropyDef}, the entropy of a probabilistic map has been used as a measure of uncertainties.
-Therefore, the decrease of the entropy due to measurements is regarded as the information gain to obtain a more accurate map.
-By planning the motion of robots to maximize the information gain, we can construct the complete map in an optimal fashion.
-
-The proposed optimal motion planning scheme to generate the desired path and the pose of the robot should be distinguished from the motion control problem that is applied in the inner loop to physically direct the robot toward the desired path.
-
-In other words, the information gain is separate from control gain.
-This paper serves to determine where the robot should go (optimal pose) based on its available information of a probabilistic map, not the tracking control inputs along a trajectory to the optimal pose, or which type of controller to use.
-To clarify this point, the following is added to the last paragraph of Section 2.2.
-\begin{correction}It is considered that once a collision-free trajectory toward $X_c^*$ is obtained, the robot moves along the trajectory controlled by a motion control system in the inner loop.
-Once the robot has translated to $X_c^*$, the above process is repeated.\end{correction}
-
-\item {\itshape Reviewer \#2: Does the methods proposed in this paper can be applied in controlling mobile robots with visual servoing? for example
-
-[1] Chen H, Wang C, Liang Z, et al.
-Robust practical stabilization of nonholonomic mobile robots based on visual servoing feedback with inputs saturation[J].
-Asian Journal of Control, 2014, 16(3): 692-702.
-
-[2] Chen H, Ding S, Chen X, et al. Global finite-time stabilization for nonholonomic mobile robots based on visual servoing[J]. International Journal of Advanced Robotic Systems, 2014, 11(11): 180.
-
-[3] Chen H, Zhang J, Chen B, et al.
-Global practical stabilization for non-holonomic mobile robots with uncalibrated visual parameters by using a switching controller[J].
-IMA Journal of Mathematical Control and Information, 2013: dns044.}
-
-%In the context of this paper, maximizing information gain is based on how expected measurement ray depths determine some action.
-
-As discussed above, the proposed algorithm generates the collision-free paths to construct a probabilistic map in an optimal fashion.
-Therefore, it can be integrated with any control algorithm, including methods based on visual servoing.
-To clarify these, the last part of Section 5 has been revised as follows.
-
-%In this sense, a geometric interpretation of visual servoing relates to the proposed research, i.e., an image does not directly control the robot degrees of freedom.
-However, the main difference is that this paper provides a desired robot trajectory, whereas visual servoing focusses heavily on robotic control.
-%So, if the proposed research in this paper were combined with control, the research could be classified under visual servoing with a primary focus on map information.
-%The authors chose not to include a particular controller in this paper to maintain generality to the resulting method.
-The following sentence is added to the motion planning description of Section 5.1.
-
-\begin{correction}Then, a controller on the robot tracks this trajectory until the robot falls within acceptable thresholds of the final optimal pose.
-Any control scheme for the motion of the robot can be integrated with the proposed exploration algorithm.
+%The focus of this paper is probabilistic mapping and autonomous exploration, not control system design for stabilization or tracking.
+In fact, the exploration algorithm serves to determine an optimal trajectory for the robot as it gathers knowledge about its environment, and as such, it deals with a guidance problem.
+Designing a specific control system to follow this trajectory may be achieved by a wide range of controllers, and it is out of scope of this paper.
+%
+%To clarify this, the end of Section 5 is modified as follows. 
+%
+%\begin{correction}The starting and ending positions and attitudes may be constrained, and polynomials patched together for long trajectories share a common position and velocity with respect to time.
+Then, the robot tracks this trajectory until the robot falls within acceptable thresholds of the final optimal pose.
+For generality of the proposed exploration algorithm, any position controller yielding robotic motion that follows this desired trajectory may be selected.
 Once the robot completes this motion, the entire process is repeated.\end{correction}
 
+\item {\itshape Reviewer \#2: A reference should be added to support the following sentence in Section 3.2: " In particular, the computational complexity for expected entropy of a single measurement ray intersecting $n_r$ cells is $\mathcal{O}(n_r^2)$''.
+The authors can alternatively show how the computational complexity was calculated.}
+
+\item {\itshape Reviewer \#2: The following claim mentioned in the conclusion should be justified: "This paper introduces a near-optimal bidding-based solution to multi-vehicle exploration and patrol to map a 3D space".
+How did the authors conclude that their proposed solution is a near-optimal one?}
+
+\item {\itshape Reviewer \#2: The authors should discuss the limitations and\/or disadvantages of the proposed framework in the conclusion section.}
+
+\item {\itshape Reviewer \#2: The reference list should be reviewed and all the missing details should be added (e.g., references 9 and 11).}
+
+\item {\itshape Reviewer \#2: The authors may suggest some directions for future work based on the proposed framework.
+}
+
 \end{enumerate}
-
-%\vspace*{0.1\textwidth}
-
-\clearpage\newpage
-\subsection*{Reviewer 3}
-
-\begin{itemize}\setlength{\itemsep}{2\parsep}
-
-\item {\itshape Reviewer \#3: Equation (5), determination of beta, collision threshold, should be clarified through numeric/experimental section as well...}
-
-The collision threshold $\beta$ corresponds to the probability of collision that may occur when the robot follows the desired trajectory.
-As $\beta$ decreases, the chances for collision decreases, but at the time, the planned trajectory may become longer and complicated while keeping a distance from potential obstacles.
-
-%Setting the particular value of $\beta$ depend
-
-%is an important parameter selected to avoid collisions. 
-%It is sensitive to grid cell size, since cell resolution determines the number of grid cells falling inside the possible collision region $\mathcal{C}_X$.
-One practical implication of (5) is that cells with greater occupancy probabilities inside $\mathcal{C}_X$ tend to fall near the perimeter of the region as the robot moves because internal cells are typically well-known to be free assuming the robot covered open space beforehand.
-This point, and
-The following description about collision probability is added to Section 6 (numerical exploration example), and the same properties are held for Section 7 (experiment).
-\begin{correction}For added safety, cells falling within $0.6$m of the robot are considered inside the possible collision zone $\mathcal{C}_X$, from \refeqn{CollisionInequalityConstraint}, where $\beta=0.1$ is selected.\end{correction}
-
-\item {\itshape Reviewer \#3: Just before equation (7) forward sensor model p(...) should be explained more.}
-
-The following is added to the ``Bayesian Framework'' paragraph of Section 3.2.
-
-\begin{correction}The occupancy probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$ is based on the forward sensor model $p(z_{t,l}|m,X_{1:t},Z_{1:t-1})$ that describe the distribution of the measurements for the given robot pose and the map.
-This specifies the stochastic characteristics of the sensor, such as the mean squared error, and it is assumed as a known distribution, specific to a particular depth sensor.
-One could certainly obtain a stochastic model by fitting a probability density to a controlled set of measurements.
-\end{correction}
-
-\item {\itshape Reviewer \#3: Likewise, in the Proposition 1, ``$p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading'' should be explained more rigorously through the Appendix 1 as well.}
-
-The statement in Proposition 1 is revised as follows.
-
-\begin{correction}$p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the probability density of the measurement when all of cells in the field of view of the $l$-th ray is not occupied.\end{correction}
-
-Referring to the maximum reading, the following explanation is added after (34).
-
-\begin{correction}For this special case, probability density can be easily represented with a uniform distribution over a short distance.
-The properties of this distribution depend on the depth sensor fidelity, because a maximum reading may indicate that all cells within the sensor range are empty, but it might also represent a failed reading.\end{correction}
-
-\item {\itshape Reviewer \#3: Uncertainty on position and attitude, cell size adjustment can be discussed at the beginning.}
-
-Accounting for localization is very important, particularly in conjunction with mapping and autonomous exploration, though this is beyond the paper scope.
-The following is added to the Section 1 (introduction).
-
-\begin{correction}Accounting for the stochastic properties of position and attitude is important to the field of robotics, and accounting for localization in conjunction with mapping and exploration are considered as a future step to this paper.\end{correction}
-
-A brief explanation of grid cell size is added to the first paragraph of Section 2.1.
-
-\begin{correction}The location and size of each grid cell is assumed known, where a smaller cell size (greater grid resolution) better represents a space, but increases computation and memory.
-\end{correction}
-
-Further discussion on cell resolution is addressed two comments below.
-
-\item {\itshape Reviewer \#3: Section 5, motion planning should cite the following paper: (grid-based, Dijkstra)}
-
-N. Gasilov, M. Dogan, V. Arici, `Two-stage shortest path algorithm for solving optimal obstacle avoidance problem', IETE Journal of Research, 57(3), 278-285, 2011, doi:10.4103/0377-2063.83650.
-
-The reference is added.
-
-%\EditTL{Do we need a brief comparison with this?}
-
-\item {\itshape Reviewer \#3: Choosing the optimal parameters for each case, e.g. ray numbers, resolution in Table 1, can be explained more to understand the proposed algorithm.}
-
-The following is added in reference to Table 1 in Section 6.1.
-
-\begin{correction}Among these parameters, cell resolution and the number of rays have the greatest impact on computation.
-In general, the finer the grid, the greater the memory and computational requirements are because the computational order of (10) grows linearly with the number of grid cells inside the sensor range limits.
-Since the inverse sensor models are combined sequentially as shown in (13), the number of rays considered are proportional to the computation order as well.\end{correction}
-
-\item {\itshape Reviewer \#3: Convergence rate of algorithm, or speed of robot while achieving its goal, should be discussed, e.g.
-possible extensions or improvements to overcome some unnecessary repetitions and back-and-forth small step size bottlenecks.}
-
-Thanks for the comments. This is an important point, especially as it relates to future goals of the research. The following is added as the last paragraph of Section 7 (experiment).
-
-\begin{correction}Both the numerical and experimental results demonstrated successful exploration strategies, though the algorithm can be improved.
-In particular, the robot sometimes revisits spaces that it has explored before, most frequently when the robot did not take measurements toward a particular direction.
-A few methods to improve this issue are treated as future work.
-One options is adding cost functions (e.g.
-distance based, number of past measurements inside an area) to prevent the robot from leaving an area before it is well-known.
-A second option is partitioning the space into sections, where the robot moves toward sections with large numbers of pose candidates with high expected information gains.
-Another option is reevaluating the future pose optimization while the robot is translating between poses because the occupancy grid map is changing, and so is expected information gain.
-The efficacy of these proposed changes will be evaluated, particularly as they relate to more useful systems, such as 3D map generation and multi-robot exploration.\end{correction}
-
 \end{itemize}

### 2019-03-08 Journal/JINT 2018/JINT18_rev1.tex
mostly done with reviewer response letter & edits

--- 2019-01-09 Journal/JINT 2018/JINT18.tex
@@
 The first contribution extends the authors' prior work in single-robot exact occupancy grid mapping and autonomous exploration in a 2D environment to mapping and exploring in a 3D environment.
-The proposed 3D occupancy grid map is computed efficiently and is scalable to larger and more complex scenarios for real-time mapping.
-Important aspects of this 3D map are projected onto a 2D occupancy grid map, where an autonomous exploration algorithm is designed to select robotic motions that maximize map information gain.
+The proposed 3D occupancy grid map is computed efficiently using an inverse sensor model that accounts for the sensor uncertainty, where we propose several measurement sources may be fused together by considering depth readings individually.
+This approach is scalable to larger and more complex scenarios for real-time mapping.
+Furthermore, this paper shows how important aspects of a 3D map representing a structured environment are projected onto a 2D occupancy grid map, where an autonomous exploration algorithm is designed to select robotic motions that maximize map information gain.
 The mapping and exploration algorithms are demonstrated with an experiment where a quadrotor autonomously maps and explores an initially-uncertain environment.
 The second contribution is a novel approach to multi-vehicle cooperative patrol of environments based on map uncertainty.
-A cooperative autonomous exploration algorithm applies a bidding-based framework to coordinate robotic efforts for improving occupancy grid map information gain.
+We propose a cooperative autonomous exploration algorithm, which applies a bidding-based framework to coordinate robotic efforts for improving occupancy grid map information gain.
 Since these exploration approaches are based on probabilistic knowledge about the map, the 3D occupancy grid map is systematically degraded over time to encourage the robots to revisit regions as time passes, thereby patrolling the environment.
@@
 For these processes, careful consideration must be placed on computation.
-In particular, the computational complexity for expected entropy of a single measurement ray intersecting $n_r$ cells is $\mathcal O(n_r^2)$.
+In particular, the computational complexity for expected entropy of a single measurement ray intersecting $n_r$ cells is $\mathcal O(n_r^2)$~\cite{KauAiLee16}.
 To address this issue, we propose projecting the 3D map onto a 2D occupancy grid map, which is located at the desired height of exploration.
@@
 The proposed approach uses auctions and bid modifications based on total map expected information gain and collision-avoidance, promoting exploration of different spaces without explicit consideration of coverage overlaps.
+For the vehicle that wins the first bid, it moves toward the candidate pose that maximizes its contribution to map information gain with travel cost.
+The subsequent auctions produce an optimal solution except for the actions of prior auction-winning robots, but these actions are explicitly considered.
+In this sense, we achieve a near-optimal coordinated solution without major computational bottlenecks.
 The pseudocode for the bidding process is shown with Algorithm \ref{alg:bidding}.
@@
 
+The research presented in this paper may be enhanced in several ways.
+First, this research assumes fixed map limits.
+In practice, a robot may not be given knowledge of the environment size.
+A mapping and exploration algorithm that is capable of growing as it explores could increase the number of applications for such robots.
+Additionally, localization can be challenging in uncertain environments, can be included in the exploration policy.
+
 %\section*{Appendix}\label{append}

### 2019-03-08 Journal/JINT 2018/JINT_ans.tex
mostly done with reviewer response letter & edits

@@
 
-%The focus of this paper is probabilistic mapping and autonomous exploration, not control system design for stabilization or tracking.
-In fact, the exploration algorithm serves to determine an optimal trajectory for the robot as it gathers knowledge about its environment, and as such, it deals with a guidance problem.
-Designing a specific control system to follow this trajectory may be achieved by a wide range of controllers, and it is out of scope of this paper.
-%
-%To clarify this, the end of Section 5 is modified as follows. 
-%
-%\begin{correction}The starting and ending positions and attitudes may be constrained, and polynomials patched together for long trajectories share a common position and velocity with respect to time.
-Then, the robot tracks this trajectory until the robot falls within acceptable thresholds of the final optimal pose.
-For generality of the proposed exploration algorithm, any position controller yielding robotic motion that follows this desired trajectory may be selected.
-Once the robot completes this motion, the entire process is repeated.\end{correction}
+The abstract is revised to place more focus on the authors' contributions as follows.
 
-\item {\itshape Reviewer \#2: A reference should be added to support the following sentence in Section 3.2: " In particular, the computational complexity for expected entropy of a single measurement ray intersecting $n_r$ cells is $\mathcal{O}(n_r^2)$''.
+\begin{correction}This paper presents a stochastic framework for autonomous exploration and patrol with multiple cooperating robots.
+The first contribution extends the authors' prior work in single-robot exact occupancy grid mapping and autonomous exploration in a 2D environment to mapping and exploring in a 3D environment.
+The proposed 3D occupancy grid map is computed efficiently using an inverse sensor model that accounts for the sensor uncertainty, where we propose several measurement sources may be fused together by considering depth readings individually.
+This approach is scalable to larger and more complex scenarios for real-time mapping.
+Furthermore, this paper shows how important aspects of a 3D map representing a structured environment are projected onto a 2D occupancy grid map, where an autonomous exploration algorithm is designed to select robotic motions that maximize map information gain.
+The mapping and exploration algorithms are demonstrated with an experiment where a quadrotor autonomously maps and explores an initially-uncertain environment.
+The second contribution is a novel approach to multi-vehicle cooperative patrol of environments based on map uncertainty.
+We propose a cooperative autonomous exploration algorithm, which applies a bidding-based framework to coordinate robotic efforts for improving occupancy grid map information gain.
+Since these exploration approaches are based on probabilistic knowledge about the map, the 3D occupancy grid map is systematically degraded over time to encourage the robots to revisit regions as time passes, thereby patrolling the environment.
+Furthermore, using a Bayesian framework and receding horizons, the algorithm is robust to dynamic obstacles within the mapping space.
+The efficacy of the proposed multi-vehicle cooperative patrol is illustrated with a simulation involving three robots patrolling a large floor plan with a non-cooperative person walking around the space.\end{correction}
+
+\item {\itshape Reviewer \#2: A reference should be added to support the following sentence in Section 3.2: ``In particular, the computational complexity for expected entropy of a single measurement ray intersecting $n_r$ cells is $\mathcal{O}(n_r^2)$''.
 The authors can alternatively show how the computational complexity was calculated.}
 
-\item {\itshape Reviewer \#2: The following claim mentioned in the conclusion should be justified: "This paper introduces a near-optimal bidding-based solution to multi-vehicle exploration and patrol to map a 3D space".
+This calculation is shown in reference [9], which now follows the statement in this paper.
+
+\item {\itshape Reviewer \#2: The following claim mentioned in the conclusion should be justified: ``This paper introduces a near-optimal bidding-based solution to multi-vehicle exploration and patrol to map a 3D space''.
 How did the authors conclude that their proposed solution is a near-optimal one?}
 
-\item {\itshape Reviewer \#2: The authors should discuss the limitations and\/or disadvantages of the proposed framework in the conclusion section.}
+The following is added to the end of Section 4.2 to justify this claim.
+
+\begin{correction}For the vehicle that wins the first bid, it moves toward the candidate pose that maximizes its contribution to map information gain with travel cost.
+The subsequent auctions produce an optimal solution except for the actions of prior auction-winning robots, but these actions are explicitly considered.
+In this sense, we achieve a near-optimal coordinated solution without major computational bottlenecks.\end{correction}
+
+\item {\itshape Reviewer \#2: The authors should discuss the limitations and/or disadvantages of the proposed framework in the conclusion section.
+\\
+The authors may suggest some directions for future work based on the proposed framework.
+}
+
+Some of the disadvantages, which motivate future work, relate to implementation challenges. The following is added to the conclusions.
+
+\begin{correction}ADD\end{correction}
 
@@
 
-\item {\itshape Reviewer \#2: The authors may suggest some directions for future work based on the proposed framework.
-}
+Reference [9] and [11] are auto-generated this way because they have an identical author list.
 

### 2019-03-10 Journal/JINT 2018/JINT18_rev1.tex
jint18/19 ready for submission after tlee check

@@
 
-\title{Multi-Vehicle Cooperative Exploration and Patrol of Uncertain 3D Environments}
+\title{Bayesian Mapping-Based Autonomous Exploration and Patrol of 3D Structured Indoor Environments with Multiple Flying Robots}
+%\title{Multi-Vehicle Cooperative Exploration and Patrol of Uncertain 3D Environments}
 
@@
 \begin{abstract}
+Mobile robots are frequently faced with mapping and exploring uncertain environments in surveillance, military, and convenience tasks.
+Often times, human teleoperation is either inconvenient or infeasible for these kinds missions.
+Furthermore, these tasks can be improved by cooperative multi-agent systems, where coordinating robotic efforts can be complicated and computationally-expensive.
 This paper presents a stochastic framework for autonomous exploration and patrol with multiple cooperating robots.
 The first contribution extends the authors' prior work in single-robot exact occupancy grid mapping and autonomous exploration in a 2D environment to mapping and exploring in a 3D environment.
-The proposed 3D occupancy grid map is computed efficiently using an inverse sensor model that accounts for the sensor uncertainty, where we propose several measurement sources may be fused together by considering depth readings individually.
+The proposed 3D occupancy grid map is computed efficiently using an inverse sensor model that accounts for the sensor uncertainty, where we propose how several measurement sources may be fused together by considering depth readings individually.
 This approach is scalable to larger and more complex scenarios for real-time mapping.
@@
 For the vehicle that wins the first bid, it moves toward the candidate pose that maximizes its contribution to map information gain with travel cost.
-The subsequent auctions produce an optimal solution except for the actions of prior auction-winning robots, but these actions are explicitly considered.
+The subsequent auctions produce an optimal solution except for the actions of prior auction-winning robots, but these prior actions are explicitly considered.
 In this sense, we achieve a near-optimal coordinated solution without major computational bottlenecks.
@@
 In practice, a robot may not be given knowledge of the environment size.
-A mapping and exploration algorithm that is capable of growing as it explores could increase the number of applications for such robots.
-Additionally, localization can be challenging in uncertain environments, can be included in the exploration policy.
+A mapping and exploration algorithm that is capable of expanding as it explores could increase the number of applications for such robots.
+Additionally, localization can be challenging in uncertain environments.
+Therefore, localization can be included in the exploration policy.
+The uncertainty of the robot pose estimate could further improve the accuracy of the probabilistic map, expected entropy calculations, and collision-avoidance.
 

### 2019-03-10 Journal/JINT 2018/JINT_ans.tex
jint18/19 ready for submission after tlee check

@@
 \begin{itemize}
-\item {\itshape Reviewer \#1: The paper is well written.
-I suggest making the title more accurate rather than generic.
-The presented methods are specifically applicable for small flying vehicles(computational limitation) in structured indoor environment (as assumed in converting 3D to 2D occupancy grid) and the specific contribution is Bayesian grid mapping and receding horizons based patrol.
+\item {\itshape Reviewer \#1: The paper is well written.}
+
+The authors appreciate the comment.
+
+\item {\itshape Reviewer \#1:  I suggest making the title more accurate rather than generic.
+The presented methods are specifically applicable for small flying vehicles (computational limitation) in structured indoor environment (as assumed in converting 3D to 2D occupancy grid) and the specific contribution is Bayesian grid mapping and receding horizons based patrol.
 There have been other papers in the literature addressing other grid mapping methods and other horizons for similar application.
@@
 
-This paper RETURN HERE!
+This paper is most relevant to structured indoor environments, which should be reflected in the title. The authors have selected the following title.
+
+\begin{correction}Bayesian Mapping-Based Autonomous Exploration and Patrol of 3D Structured Indoor Environments with Multiple Flying Robots\end{correction}
 
@@
 
-The abstract is revised to place more focus on the authors' contributions as follows.
+The abstract is revised to place more focus on the motivation the authors' contributions as follows.
 
-\begin{correction}This paper presents a stochastic framework for autonomous exploration and patrol with multiple cooperating robots.
+\sloppy\begin{correction}Mobile robots are frequently faced with mapping and exploring uncertain environments in surveillance, military, and convenience tasks.
+Often times, human teleoperation is either inconvenient or infeasible for these kinds missions.
+Furthermore, these tasks can be improved by cooperative multi-agent systems, where coordinating robotic efforts can be complicated and computationally-expensive.
+This paper presents a stochastic framework for autonomous exploration and patrol with multiple cooperating robots.
 The first contribution extends the authors' prior work in single-robot exact occupancy grid mapping and autonomous exploration in a 2D environment to mapping and exploring in a 3D environment.
-The proposed 3D occupancy grid map is computed efficiently using an inverse sensor model that accounts for the sensor uncertainty, where we propose several measurement sources may be fused together by considering depth readings individually.
+The proposed 3D occupancy grid map is computed efficiently using an inverse sensor model that accounts for the sensor uncertainty, where we propose how several measurement sources may be fused together by considering depth readings individually.
 This approach is scalable to larger and more complex scenarios for real-time mapping.
@@
 \begin{correction}For the vehicle that wins the first bid, it moves toward the candidate pose that maximizes its contribution to map information gain with travel cost.
-The subsequent auctions produce an optimal solution except for the actions of prior auction-winning robots, but these actions are explicitly considered.
+The subsequent auctions produce an optimal solution except for the actions of prior auction-winning robots, but these prior actions are explicitly considered.
 In this sense, we achieve a near-optimal coordinated solution without major computational bottlenecks.\end{correction}
@@
 
-Some of the disadvantages, which motivate future work, relate to implementation challenges. The following is added to the conclusions.
+Some of the disadvantages, which motivate future work, relate to fixed map limits and localization. The following is added to the conclusions.
 
-\begin{correction}ADD\end{correction}
+\begin{correction}The research presented in this paper may be enhanced in several ways.
+First, this research assumes fixed map limits.
+In practice, a robot may not be given knowledge of the environment size.
+A mapping and exploration algorithm that is capable of expanding as it explores could increase the number of applications for such robots.
+Additionally, localization can be challenging in uncertain environments.
+Therefore, localization can be included in the exploration policy.
+The uncertainty of the robot pose estimate could further improve the accuracy of the probabilistic map, expected entropy calculations, and collision-avoidance.\end{correction}
 
